### 2016-07-27 experiments/e0/writeup/vgn0.Rmd
adding e0, the pilot experiment with subitiable stimlui.

---
title: "Experiment 0 pilot"
author: "Matt Green"
output:
  html_document:
    toc: yes
    mathjax: default
    template: default
    code_folding: hide
    theme: default
---

This document aims to be a complete and up to date record of the analysis for vagueness 0, the pilot experiment}

Start clean.

```{r, startClean}
rm(list=ls())
```

Knitr doc options.

```{r, docOpts}
library(knitr)
knit_hooks$set(crop=hook_pdfcrop)
opts_chunk$set(cache=TRUE, autodep=TRUE)
#opts_chunk$set(size="footnotesize")
#options(width=60)
```

Load libraries.

```{r, loadLibraries}
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(reshape))
suppressPackageStartupMessages(library(lattice))
suppressPackageStartupMessages(library(LMERConvenienceFunctions))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(gridExtra))
library(grid)
suppressPackageStartupMessages(library(lmerTest))
```

Source additional R files with functions for computing confidence intervals.

```{r, sourceFunctions}
source("functions/CIfunctions.R")
```

# Preprocessing
Get the data.

```{r, getData}
system("cat ../experimentCode/output/*.dat > data.txt", wait=TRUE)
d <- read.table("data.txt", header=FALSE)
mynames <- list("gap", "target", "side", "vagueness", "pair", "Version", "Left",
								"Right", "Text", "Condition", "correctAnswer", "subject",
								"actualAnswer", "isCorrect", "rt")
names(d) <- mynames
d$obs <- 1:dim(d)[1]
d$trial=0
for (s in unique(d$subject)){
  d[d$subject==s,"trial"] <- 1:nrow(subset(d,subject==s))
}
# For some reason there is one rt of -6. Set it to NA
d[d$rt==-6,"rt"] <- NA
# make a df
d <- data.frame(
  obs = d$obs, trial = d$trial, 
  subject = d$subject, 
  item = factor(d$pair, label = c("2:4", "3:5", "6:8", "7:9", "2:6", "7:3", "8:4", "5:9")),
  gap = factor(d$gap, levels = c("BigGap", "SmallGap"), labels = c("Big Gap", "Small Gap")), 
  quan = factor(d$target, levels = c("BigTarget", "SmallTarget"), labels = c("Big Target", "Small Target")), 
  vag = factor(d$vagueness, levels = c("Precise", "Vague"), labels = c("Crisp", "Vague")), 
  side = d$side, err = 1 - d$isCorrect, rt = d$rt)
d$subit <- rep("Not subitizable", times = nrow(d))
d$subit[d$item == "2:4"] <- "Subitizable"
d$subit[d$item == "3:5"] <- "Subitizable"
d$subit[d$item == "2:6"] <- "Subitizable"
d$subit[d$item == "7:3"] <- "Subitizable"
d$subit = as.factor(d$subit)
d <- subset(d, select = c("obs", "trial", "subject", "item", "gap", "quan", 
                          "vag", "subit", "side", "err", "rt"))
```

Size of the data is \Sexpr{nrow(d)} trials. 1 was removed for being minus 6 rt.
Remove erroneous response trials. There are \Sexpr{nrow(subset(d,err==1))} error trials.

```{r, removeErrorTrials}
# by setting the RT to NA
d[d$err==1,"rt"] <- NA
```

Do some factors and contrasts.  Set subject and item as factors. Sum code vagueness and subitizability.

```{r, postProcessData}
d$subject <- as.factor(d$subject)
d$item <- as.factor(d$item)
d$vag <- relevel(d$vag, ref="Vague")
contrasts(d$vag) <- contr.sum(2)/2
d$subit <- relevel(d$subit, ref="Subitizable")
contrasts(d$subit) <- contr.sum(2)/2
contrasts(d$quan) <- contr.sum(2)/2
contrasts(d$gap) <- contr.sum(2)/2
```

Check the contrasts.

```{r, checkContrasts}
contrasts(d$vag)
contrasts(d$subit)
contrasts(d$quan)
contrasts(d$gap)
```

Add transformations of rt

```{r, addTransform}
d$logRT <- log(d$rt)
d$recipRT <- -10000/d$rt
```

For the untrimmed data, show effects of transforming rt.

```{r, plotDistributions, fig.width=4, fig.height=2}
xs=melt(d,measure.vars=c("rt","logRT","recipRT"))
densityplot(~value|variable, data=xs, ylab=NULL,xlab=NULL,yaxt=F, scales = list(x = list(relation="free", rot=45), y=list(relation="free",at=NULL)),layout=c(3,1),cex=.3)
```

Trim the data. This function doesn't work well if there are NA values in the response, so explicitly remove them first.

```{r, trimData}
dummy <- subset(d, !is.na(rt))
data.trim <- perSubjectTrim.fnc(dummy, response="rt",subject="subject", trim=2.5)
```

Check the distribution of the trimmed data to see if a transformation is suggested.

```{r, checkTrim, fig.width=4, fig.height=2}
xs=melt(data.trim$data,  measure.vars=c("rt","logRT","recipRT"))
densityplot(~value|variable,data=xs,
            ylab=NULL,xlab=NULL,yaxt=F,
            scales=list(x=list(relation="free",rot=45),y=list(relation="free",at=NULL)),cex=.2)
```

Select either trimmed or untrimmed data by comparing this plot with the untrimmed data. Also use this plot to choose the most normal data (logRT).

```{r, selectTrim}
dat <- data.trim$data
```

# RT modelling
Do a first lmer model. You can't put subit or gap over items because they are between-items.

```{r, lmers}
lmer2 <- lmer(data=dat, logRT~subit*vag + (subit*vag|subject) + (vag|item))
# lmer3 is closer to the original anova than lmer2
lmer3 <- lmer(data=dat, logRT~subit*vag+quan+gap + (subit*vag|subject) + (vag|item))
```

```{r, summaryLMER, results='asis'}
print(xtable(summary(lmer3)$coefficients, digits=3), type="html")
```

## subitizable
```{r, subitRT, results='asis'}
print(xtable(summary(lmer(data=dat, subset=subit=="Subitizable",logRT~vag+quan+gap +(vag|subject)+(vag|item)))$coefficients, digits=3), type='html')
```

## not subitizable
```{r, notsubitrT, results='asis'}
print(xtable(summary(lmer(data=dat, subset=subit=="Not subitizable", logRT~vag+quan+gap +(vag|subject)+(vag|item)))$coefficients, digits=3), type='html')
```

# Error rate analysis
```{r, errorrates, results='asis'}
print(xtable(summary(glmer(data=d, err~vag*subit+quan+gap + (vag|subject) + (1|item), family=binomial, control = glmerControl(optimizer = "bobyqa")))$coefficients, digits=3), type='html')
```

# Plots of significant effects
```{r, plot1}
a <- summarySEwithin(data=dat, measurevar="logRT", withinvars=c("vag","subit","item"), idvar="subject")
as <- subset(a,subit=="Subitizable")
as$item <- as.character(as$item)
as$item[as$item=="7:3"] <- "3:7"
as$item <- factor(as$item)
pd <- position_dodge(.05)
plot1 = ggplot(data=as, aes(y=logRT, x=item, ymin=logRT-ci, ymax=ci+logRT, fill=vag, group=vag)) + 
  geom_line(position=pd) +
  geom_errorbar(width=.1, position=pd)+
  geom_point(pch=21, size=4, position=pd) +
  scale_fill_grey(name=element_blank(),start=0,end=1)+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key=element_blank()) +
  ggtitle("Response time")

```

```{r, plot2}
a <- summarySEwithin(data=d, measurevar="err", withinvars=c("vag","subit","item"), idvar="subject")
as <- subset(a,subit=="Subitizable")
as$item <- as.character(as$item)
as$item[as$item=="7:3"] <- "3:7"
as$item <- factor(as$item)
pd <- position_dodge(.05)
plot2 = ggplot(data=as, aes(y=err, x=item, ymin=err-ci, ymax=ci+err, group=vag, fill=vag))+
  geom_line(position=pd) +
  geom_errorbar(width=.1, position=pd)+
  geom_point(pch=21, size=4, position=pd) +
  scale_fill_grey(name=element_blank(),start=0,end=1)+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key=element_blank()) +
  ggtitle("Error rate")

```

```{r, gridarrange, fig.width=7.5, fig.height=3}
grid.arrange(plot1, plot2, nrow=1, top="Subitizable only")
```

```{r, barchart-rts}
bwtheme <- standard.theme("pdf", color=FALSE)
b1=barchart(data=ddply(dat,.(vag,subit),summarise,rt=mean(logRT)), rt~subit, group=vag, auto.key=list(),par.settings=bwtheme)
```

```{r, barchart-errors}
b2=barchart(data=ddply(d,.(vag,subit),summarise,err=mean(err)), err~subit, group=vag, auto.key=list(),par.settings=bwtheme)
```

```{r, both-barcharts, fig.width=7, fig.height=3}
grid.arrange(b1,b2,ncol=2)
```

Subitizability x Vagueness x Quantity

```{r, plotMeansa}
a <- summarySEwithin(data=dat, measurevar="logRT", withinvars=c("vag","quan","subit"), idvar="subject")
pd <- position_dodge(.1)
meansplot1 <- ggplot(a,aes(y=logRT,x=vag,group=quan,shape=quan,col=quan,ymin=logRT-ci, ymax=ci+logRT))+geom_point(size=4, position=pd)+geom_line(position=pd) +facet_grid(~subit) + geom_errorbar(width=.1, position=pd) + theme_bw()+ theme(panel.grid.major=element_blank()) 
```

```{r, plotMeans, fig.width=7, fig.height=3}
meansplot1
```

```{r, qqplotLMER, fig.width=3, fig.height=3.5}
qqPlot(residuals(lmer2),cex=.5)
```

### 2016-07-27 papers/2014/internal_format/vcr.tex
adding the 2014 LCP submission

\documentclass[man,floatmark]{apa}
% with basic edits to correct typos, by Matt in July 2016.
\usepackage{graphicx}
\usepackage[firstpage]{draftwatermark}
\SetWatermarkText{Internal format 2014}
\SetWatermarkScale{3}
\title{Vagueness as cost reduction}
\author{Matt Green (mjgreen@abdn.ac.uk) and \\Kees van Deemter (k.vdeemter@abdn.ac.uk)}
\affiliation{Department of Computing Science \\ University of Aberdeen\\ AB24 3UE}
\note{corresponding author: Matt Green\\ Rm 244, Computing Science\\Meston Building\\Aberdeen\\AB24 3UE\\email: mjgreen@abdn.ac.uk\\tel: 07964339219}
\abstract{Three experiments investigated the effects of vagueness in referring expressions, in order to test a prediction from game theory models of communication to the effect that crisp communications always have more utility than vague communications, under specified conditions. The experiments used a forced choice task, where participants were required to select an object on screen in response to an instruction in the form of a referring expression. The instruction referred to the number of dots the object contained, and was manipulated such that it referred either vaguely or crisply to the number. In Experiment 1, it was found that when either of two objects contained a subitizable (i.e., very small) number of dots, participants responded more quickly and accurately when the instruction was crisp than when it was vague. In contrast, in Experiment 2, using non-subitizable (i.e., larger) numbers, participants responded more quickly and accurately when the instruction was vague than when it was crisp, with diminishing returns for vagueness as the number of dots involved grew larger. Experiment 3 introduced an additional manipulation on whether the instruction contained a numeral, such that there were both vague and crisp versions of instructions that did and did not contain numerals. It was found that instructions that did not contain numerals were advantageous over those that did, and that within this effect, there was no reliable difference between vague and crisp alternatives.}
\shorttitle{Vagueness as cost reduction}
\rightheader{Vagueness as cost reduction}
\leftheader{Matt Green and Kees van Deemter}
\begin{document}
\maketitle                            
Vagueness pervades the language that we use on a daily basis. In everyday use, language use may be called vague for various reasons.\footnote{See e.g. the entry ``vague" in \cite{penguin}.} In most academic use though, the word `vagueness' has a more particular meaning.  Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to orites paradoxes" \cite[p.\ 4]{keefe1996vagueness} (a similar definition can be found in \citeA{EgreKlinedinst}, among others).  The most crucial of these criteria is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects. By contrast, a word is vague if it is not precise" \cite[p.\ 1]{lipmanvague}. A typical example is the word ``tall", as applied to people from example, because here is no precise, known height which separates those who are tall from those who are not. The crucial point is that ``tall'' admits borderline cases (i.e., people who may or may not count as tall), which are the hallmark of vagueness as we use the term.

Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions, which are used so frequently \cite{Lipman:2000fk, lipmanvague}. By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss" \cite<>[p.~1]{lipmanvague}. Lipman explains the point by means of a scenario in which a speaker describes a person to a hearer, who needs to identify that person in the arrivals hall of of an airport. Lipman argues that, in such a scenario, a precise description of the person's height (e.g., ``The person's height is 187.96 cm") would be more useful than a vague one (``The person is tall''). Lipman uses this scenario to explain why standard game theory models of communication \cite<e.g.,>{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act of communication that communicates the same state of affairs. The relevant conditions are, broadly speaking, that both interlocutors know all the relevant facts (e.g., both know the person's height precisely) and that the setting for the communication is co-operative. These conditions exclude such communicative situations as deliberate deception, and rhetorical situations like political debate and advertising where the intention is to persuade the interlocutor to adopt some point of view, where the persuasion might not be intended to be in the addressee's best interest.

Lipman observed that such the efficiency loss resulting from vague expressions would be unlikely to have arisen unless there are advantages as well as disadvantages associated with vague expressions. Lipman asked, essentially, what these advantages might be, and how they might find a place in a game-theoretical explanation. In this paper, we focus on the first part of Lipman's question.

Several tentative answers to Lipman's question have been offered  \cite<see>{van2009utility, van2010vagueness}. Prominent among these answers is the idea that vague expressions are somehow easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite<e.g.,>[]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}. For example, \citeA<>[p.\ 11]{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze". We shall refer to this characterisation of the utility of vague language as the \emph{cost reduction} hypothesis. The idea of vagueness as cost reduction can take various shapes \cite<>[]{van2009utility}, but the basic hypothesis is that it is easier for people to think in terms of loosely defined categories (such as ``quite a few", or ``many") than in terms of crisply defined ones (such as ``thirteen", or ``237"). This predicts that whenever a vague expression can perform the same communicative task as a precise one, it is rational to choose the vague expression. The corollary for comprehension is that vague expressions should be understood more readily than precise expressions.

Questions concerning optimal language use have many practical applications. Natural Language Generation\footnote{NLG systems take data or formulas as input, and transform them into natural language outputs \cite<>[]{reiter2000building}. The process parallels language production in humans.}  (henceforth NLG) systems must make decisions between different formulations of the same information. For example, if a man's height is 6 foot 2 inches, this could be expressed as ``187.96 metres", ``6 foot 2", or ``tall", among other ways, and the NLG system must decide between these.  The problem is particularly relevant for NLG systems that take numbers as input, as many do. In the context of an NLG system faced with a practical decision of this kind, Lipman's question becomes ``Under what circumstances should vague terms be produced?"  Relevant applications include weather forecasting on the basis of numerical weather data such as temperature and wind speed \cite{goldberg1994using,turner2006generating}, and medical decision support on the basis of clinical measurement such as oxygen saturation, heart rhythm, etc. \cite{Hripcsak01032009, hunter2008summarising, portet2009automatic}. At present, such NLG systems make decisions concerning the level of precision in the utterances that they generate (e.g., ``the temperature will be in the high twenties tomorrow'') on the basis of little more than intuition. A better understanding of the benefits of different precision levels for readers would allow these systems to become more useful.

The cost reduction hypothesis is of direct relevance to psycholinguists interested in language comprehension, and additionally to psycholinguists interested in language production, for example in connection with the question of audience design \cite{Clark1982287}. For, to the extent that speakers and writers choose vague expressions over and above crisp ones because the former are easier to process for hearers than the latter, the cost reduction hypothesis suggests that speakers design their utterance for optimal benefit to their hearers -- out of altruism, so to speak.

The utility of vagueness is the attested aim of a small number of studies, but most of these have focussed on vagueness in a different sense, and focussing on different types of benefits for hearers. Two recent studies can be used to illustrate both issues.

In a study of behaviour modification, \citeA<>{Mishra01042011} manipulated the presentation format of information about quantities. They compared information presented as a range (\emph{between 0.5 and 1.5 g of cocoa}) with information presented as a single value at the midpoint of the range (e.g., \emph{1 g of cocoa}). Information presented as a range was considered to be vague. The authors managed to effect behaviour modification in their participants, in the domains of mental acuity, physical strength, and weight loss, in laboratory and quasi-field studies. Their \emph{vague} conditions improved performance relative to a \emph{precise} condition. They explain the improvement in the range-of-values conditions as resulting from the participants' freedom to distort the information by focussing on one end of the range -- the end that indicates proximity to the goal for motivated participants. The ``illusion of proximity" \cite<>[p.~4]{Mishra01042011} to the goal is argued to allow participants to generate positive expectancies that lead to behaviours that improve performance. In contrast, in the midpoint-value conditions, participants did not have this freedom of interpretation, and could not distort the information to bring about the beneficial \emph{illusion of proximity}. These results are interesting, and of obvious potential practical importance. We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \citeA{keefe1996vagueness}  \citeA{EgreKlinedinst}, since an exact range does not admit borderline cases. (In the terminology of \citeA{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}.) Furthermore, the experiments of \citeA{Mishra01042011} did not explore benefits in terms of processing cost, but in terms of (long-term) behaviour change.

Similar issues arise from the work of \citeA{peters2009bringing}. The authors carried out a series of studies where participants were required to rate hospitals based on various sources of information about quality of care. There was a between-subjects manipulation based on numeracy. The format of the information was manipulated within subjects: either numbers only were presented, or both numbers and evaluative categories were presented (e.g., \emph{Poor}, \emph{Fair}, \emph{Good}, \emph{Excellent}, with visual boundary lines between the categories). Results showed that, for low-numeracy participants, the presence of evaluative categories resulted in a diminished influence of an irrelevant affective state on the ratings. For all participants, the presence of evaluative categories resulted in a greater use of numerical information when making decisions; and greater use of an important indicator that is hard to interpret (i.e., survival rates). Participants' decisions were evaluated with respect to a `gold-standard' set of interpretations derived from experts. It is interesting to consider whether their evaluative categories manipulation can be considered a manipulation of vagueness. Certainly, terms like \emph{Fair} admit the possibility of borderline cases. However, when the visual boundary lines are taken into account, which map the terms to exact ranges,  it becomes doubtful whether any borderline cases could be conceived to arise in fact. For example,  \emph{Fair} was mapped to 60 -- 70 \% for the variable \emph{percentage of heart attack patients given recommended treatment (ACE inhibitor)}.\footnote{This distinction between the potential for vagueness, and the realisation of vagueness in the context of a particular stimulus, is something we will return to when discussing our own experiments.} Accordingly, rather than the vagueness of categories such as \emph{Poor}, Peters et al. emphasise the evaluative content inherent in these categories, and their affective potential.

The experiments reported in the present paper put the cost reduction hypothesis to the test. The question that we are trying to answer is whether vague expressions are processed more easily by readers than crisp ones. Like Lipman, we focus on situations where numerical information is used in order to identify a referent. Reference, in other words, will be the linguistic task on which we focus, partly because of the interest that this topic has recently drawn from the NLG community.  In focussing on benefits for the hearer, we will leave aside the question of audience design, leaving this for later research.

Game theory and the cost reduction hypothesis both make claims about vagueness in terms of its effects on people engaged in communicative linguistic acts. For game theory, when vagueness influences communication, the term \emph{utility} is used to capture this influence: vagueness is said to have less utility than crispness. The cost reduction hypothesis notes that we use vague language frequently, and assumes that the reason for the high frequency of use is that vagueness brings about an \emph{advantage}, for producer or comprehender. In this paper, we set out to measure these effects so that we might have some empirical basis for preferring one account over the other. An obstacle to this effort is that whereas we can obtain, in the laboratory setting, several quantitative measures of the effects of language, the opposing game theory and cost reduction accounts do not specify a metric in which \emph{utility} or \emph{advantage} should be measured. Therefore it is for us as experimenters to suggest a way to capture and measure the effects of vague and crisp expressions such that they can meaningfully be compared. This requires us to choose, and motivate the choice of, both a task that brings about some measurable behaviour (or alternatively, some measurable physiological state) that is consequent upon whether the language used to bring it about was vague or crisp; and also a measure of the behaviour that is sufficiently fine-grained to distinguish the vague case from the crisp case.

A good candidate for the task comes from the \emph{forced choice} paradigm in which a participant is required to make a choice quickly among alternatives, indicating the choice by a physical response such as a button-press. These alternatives can be presented using vague and crisp language in such a way that the participant's choice is at least partly consequent upon whether the presentation was vague or crisp: indeed, in the case where other factors are held constant, it is plausible, and often assumed, that differences in the response are consequent to a great extent upon the mode of presentation. The button-press behaviour offers two metrics which are commonly taken to index cognitive load: response time and response accuracy, where fast accurate responses indicate low cognitive load, and slow or inaccurate responses indicate high cognitive load. In the case that vague presentations lead to lower cognitive load than crisp presentations, or vice versa, (where load is indexed by response time and accuracy), it is plausible to argue that this constitutes an advantage, or greater utility, of one presentation mode over the other.

In the experiments that we present here, we induced people to make a particular choice among alternatives by indicating one alternative using a referring expression.  In the particular situation that we used, participants were presented with a number of objects on screen that constituted the alternatives.  These objects were squares containing a number of dots, where the number in each was different. The referring expression referred to one square by indicating the number of dots it contained, and the participant was required to indicate that square by pressing the appropriate button on the keyboard. The referring expression offered a way to manipulate whether the square was indicated with a crisp or a vague expression of quantity, while holding other factors relatively constant. For example, the participant could be instructed ``Choose the square with many dots" in the vague condition, or ``Choose the square with 20 dots" in the crisp condition.

\section{Experiment 1} 
In our first experiment we set out to compare responses to vague and crisp instructions in a forced choice task using response time and accuracy to measure the responses, and a variety of combinations of numbers. We aimed to identify cases where a difference was elicited such that these cases could form the focus of  subsequent experiments.

\subsection{Method}

\subsubsection{Participants}
Twenty-five students and staff from the University of Aberdeen participated in the study in return for a cash payment of five pounds. Their median age was 23, ranging from 18 to 40. All participants self-reported fluency in English, and had normal (or corrected to normal) vision. Participants were recruited by advertising for participants on a university message board and a university mailing list, offering five pounds to volunteers.

\subsubsection{Apparatus}
A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants. Stimuli were created and presented using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.

\subsubsection{Design}
We presented participants with 64 trials in experiment one in a single block. The basic properties of a trial were that it consisted of two elements: (1) two squares each containing a diffferent number of dots, and (2) a referring expression that referred to one of the squares by indicating the number of dots that it contained, leaving the other as a distractor. We measured two dependent variables: response time, and response accuracy. It was always the case that two squares appeared on the screen, each containing a number of dots: therefore items can be described using the notation $pair_{(p=1... 8)}([n],m)$ to indicate that for the stimulus pair $p$, one of the numbers of dots was $n$ and the other $m$; and that the square-bracketed element $[n]$ was the target. There were 8 such pairs : $pair_{(1)}(2,4)$; $pair_{(2)}(3,5)$; $pair_{(3)}(2,6)$; $pair_{(4)}(3,7)$; $pair_{(5)}(6,8)$; $pair_{(6)}(7,9)$; $pair_{(7)}(4,8)$; $pair_{(8)}(5,9)$. See Table \ref{tablee1} for the properties of the stimuli.

\begin{table}[tbp]
\caption{Properties of stimuli used in Experiment 1, arranged by pair}
\label{tablee1}
\begin{tabular}{lll}
subitizabiity	& gap size  	&	contents\\
\hline
subitizable	&small gap	&	$pair_{(1)}(2,4)$\\
 			&			&	$pair_{(2)}(3,5)$\\
\hline
			&large gap	&	$pair_{(3)}(2,6)$\\
			&			&	$pair_{(4)}(3,7)$\\
\hline
not subitizable	&small gap	&	$pair_{(5)}(6,8)$\\
			&			&	$pair_{(6)}(7,9)$\\
\hline
			&large gap	&	$pair_{(7)}(4,8)$\\
			&			&	$pair_{(8)}(5,9)$\\
\hline
\end{tabular}
\end{table}

We manipulated four independent variables, and explicitly controlled one other, as follows.

The first independent variable was whether the item contained a subitizable number of dots. This was manipulated between items, and divided the items into two sets: one set comprised items that did contain a subitizable number; the other set comprised items that did not contain a subitizable number. There is evidence \cite{trick1994small} that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities. Four of our pairs contained such a subitizable number, and four did not. This allowed us to ask whether vagueness or crispness might exert differential effects depending on whether this subitizing mechanism was, or was not, involved. The four pairs containing a subitizable number were: $pair_{(1)}(2,4)$; $pair_{(2)}(3,5)$; $pair_{(3)}(2,6)$; $pair_{(4)}(3,7)$; and the four pairs not containing a subitizable number were: $pair_{(5)}(6,8)$; $pair_{(6)}(7,9)$; $pair_{(7)}(4,8)$; $pair_{(8)}(5,9)$.

The second independent variable was whether the gap between the smaller and the larger number was small or large. At each level of the subitizable variable, half the items had a small gap, and the other half had a large gap. There is evidence that when the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the numerosity of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene199647}. By systematically manipulating this difference, we were able to ask whether effects of vagueness or crispness might differ according to whether the pair was a more or less discriminable pair.

The third independent variable was vagueness with two levels, vague and crisp: the effects of vague and crisp language were to be elicited by the way in which an instruction referred to an expression of quantity. Each stimulus was presented equally often with a  vague expression of quantity, and a crisp expression of quantity. Taking the item $pair_{(1)}([2],4)$ as an example, where the target was the square with 2 dots, the crisp referring expression was ``Choose the square with two dots" and the vague referring expression was ``Choose the square with few dots".

Because we presented each item with the smaller number as target, and with the larger number as target, we were able to ask whether responses differed as a function of whether the target was large or small (with respect to the distractor). This constituted our fourth independent variable.

We also had to control which side the smaller number appeared on. This was manipulated within items such that each item had a version where the smaller number was on the left and a version where the smaller number was on the right. There is evidence  that when two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}. Therefore we needed to present $pair_{(1)}(2,4)$ once as $pair_{(1)}([2],4)$ and once as $pair_{(1)}(4,[2])$. This accounts for two of the four presentations. The other two presentations were $pair_{(1)}(2,[4])$ and $pair_{(1)}([4],2)$: i.e., with the larger number [4] rather than the smaller number as the target square; again once on the left and once on the right. This accounts for the eight presentations of each $pair_{(p)}(n,m)$.

\subsubsection{Stimuli}
Each stimulus consisted of two elements: (1) two squares containing dots; (2) the text of the referring expression indicating the target. Presentation of the two elements was simultaneous. The manipulation of vagueness was effected in the instruction, which referred to the number of dots in the target square as either \begin{seriate} \item a cardinal numeral in the crisp conditions (i.e., \emph{two, three, four, five, six, seven, eight, nine}), or \item a linguistic quantifier in the vague conditions (e.g., \emph{few, many}) \end{seriate}. The position of the dots in each square was randomised per trial. An example stimulus is given in Fig. (\ref{stime1}).

\begin{figure}[tbp]
\fitfigure{images/stimuluse1}
\caption{Example stimulus from experiment 1. Stimulus consists of two elements (1) the two squares with a number of dots, one of which is the referent of the instruction; (2) the text of the instruction that identifies the target with a referring expression.}
\label{stime1}
\end{figure}

\subsubsection{Procedure}
The experiment was conducted in a small quiet room. On arrival in the room, the participant was told that he or she would be presented with objects on screen, together with an instruction to choose one of them by pressing the appropriate key on the keyboard (which was left cloverleaf for a target on the left, and right cloverleaf for a target on the right). The participant was instructed to respond quickly while avoiding errors.
There were 4 practice trials. After the practice trials the participant was invited to ask any questions they had about the procedure, and then the experimenter left the cubicle before the experimental blocks began.
The order in which trials were presented was randomised per participant. Each trial would time out after 60 seconds if the participant did not respond. No feedback was given on correct trials, but there was feedback on error trials in the form of the word ``\textsc{wrong!!}" which flashed on screen. Response time was measured from the appearance of the stimulus on screen until the button-press indicating the participant's choice. The trial sequence was keypress $\rightarrow$ instruction and stimulus $\rightarrow$ keypress.

\subsection{Results}
Preprocessing was different for the response time analysis and for the analysis of error rates. For the response time analysis, trials with response times measuring fewer than 250 ms were discarded, as were trials with response times of greater than 10,000 ms. Response times from trials with erroneous responses were also discarded.  This led to the loss of 44 out of 1600 trials, which represented 2.75\% of trials. One participant was excluded from the analysis of variance due to making errors in all responses in one condition. This would have led to an unbalanced design for ANOVA had the participant been retained. This participant's exclusion led to the loss of another 59 trials, the other 5 having been excluded by the time and error trimming described earlier in this section. Overall, 6.4\% of trials were excluded from the response time analysis (103 of 1600 trials). For the analysis of error rates however, no trials were discarded.

A 4-way (2 x 2 x 2 x 2) repeated measures ANOVA was carried out on (correct response) RT as the dependent variable, with gap (small:2 or large:4), subitizability (whether the display included a suitizable number or not), quantity (whether the larger or smaller number was the target), and vagueness  (vague instruction or crisp instruction) as independent variables each with two levels.

Grand mean RT was 1637 ms, with standard error 114 ms. The manipulations revealed significant main effects of gap size, target quantity, and subitizability, but no significant main effect of vagueness. There were reliable interactions of vagueness with quantity; and of gap with subitizability, and a reliable three-way interaction between vagueness, quantity, and subitizability. No other interactions were reliable. Details of the reliable effects and interactions follow.

Vague instructions led to numerically slightly longer RT than precise instructions but this difference was not significant (1600 ms vs 1673 ms, $F(1,23)=2.1, p = .161,  \eta^2 =.084$). Participants responded significantly faster when the numerical gap was large at 4 dots than when it was small at 2 dots (1397ms vs 1876 ms, $F(1,23)=70.6, p<.001, \eta^2 = .754$, halfwidth of the 95\% confidence interval for the 479 ms difference was 118 ms). Hereafter halfwidths are referred to simply as CI. When one of the squares contained a subitizable number of dots, participants responded significantly  faster than when neither of the squares contained a subitizable number of dots (1292 ms vs 1981 ms, $F(1,23)=83.9, p<.001, \eta^2 =.785$, 95\% CI of the 689 ms difference was 155 ms).

When participants were instructed to choose the square with the larger number of dots, they responded 165 ms more slowly than when they were instructed to choose the square with the smaller number of dots, for the same pair of numbers (1719 vs 1554 ms, $F(1,23)=15.4, p<.01,  \eta^2 =.401$,95\% CI of the 165 ms difference was 86.5 ms). Vagueness interacted significantly with quantity ($F(1,23)=25.065,p<.001,\eta^2=.521$), such that vagueness was a significant advantage of 116 ms when participants were instructed to choose the square with the larger number of dots (1761ms vs 1645 ms, $t(23)=2.17,p<.05$, 95\% CI of the 116 ms difference was 111 ms) but a significant disadvantage of 275ms when participants were instructed to choose the square with the smaller number of dots (1409 ms vs 1684 ms, $t(23)=-3.88, p < .001$, 95\% CI of the 275 ms difference was 147 ms).

The effect of a subitizable number in one of the squares varied as a function of the size of the gap ($F(1,23)=49.669,p<.001,\eta^2=.683$). In the big gap conditions, the advantage for a subitizable number was smaller than it was in the small gap conditions. In the big gap conditions, the advantage due to a subitizable number was 308 ms (1551 ms vs 1243ms, $t(23=7.5, p<.001$, 95\% CI of the 308 ms difference was 85 ms), whereas in the small gap conditions, the advantage due to a subitizable number was greater at 1059 ms (2398 ms vs 1338 ms, $t(23) = 8.6, p<.001$, 95\% CI of the 1059 difference was 254 ms).

There was a reliable three-way interaction between subitizability, quantity, and vagueness ($F(1,23)=6.653,p<.05,\eta^2=.224$), such that vague instructions were advantageous when subjects were instructed to choose the bigger quantity and there was no subitizable number, but disadvantageous when the participants were instructed to choose the smaller number, regardless whether there was a subitizable number or not. This interaction is plotted in Fig. (\ref{resultse1}).

The grand mean participant error rate was low at $1.3\%$. A generalized linear mixed model \cite{jaeger2008categorical} was fit to the error data, with task, quantity, vagueness and subitizability, as well as all interactions between them, as fixed effects, and with random intercepts for subject and item. None of the main effects or interactions showed reliable effects, apart from the main effect of gap. Mean error rates rose by 2.2\% in the small gap conditions (mean error rate in the big gap conditions: $1.38\%$; mean error rate in the small gap conditions: $3.63\%$, $Wald~Z = 2.323, p<.05 $).

\begin{figure}[tbp]
\fitfigure{images/resultse1.pdf}
\caption{Experiment one results. This figure plots the vagueness advantage in each of the conditions involved in the significant three-way interaction between vagueness, subitizability, and target size, such that a positive number reflects an \emph{advantage} of vagueness in that condition, and a negative number reflects a \emph{disadvantage} of vagueness. The plot shows that there is a vagueness advantage in the  condition ``not-subitizable'  when the referring expression indicates the bigger number as target, but a consistent disadvantage of vagueness in the other conditions.}
\label{resultse1}
\end{figure}

\subsection{Discussion}
Response times did not differ reliably according to whether the quantifier was vague or crisp, all other things being equal, but vagueness exerted effects in interactions with other effects.

\citeA{moyer1967time} observed that participants compared two numbers faster as the numerical distance, or gap, between them grew larger. Our subjects showed a similar effect, responding nearly 500 ms faster (and with fewer than half as many errors) to the instruction to choose a square containing a number of dots when the numerical distance was four than when it was two. This effect did not interact with vagueness for response times or error rates, suggesting that  participants respond similarly in the small and large gap conditions regardless whether the instruction is vague or precise.

Similarly in line with previous findings \cite<e.g.,>[]{mandler1982subitizing},  our participants responded faster when one of the squares contained a subitizable number of dots than when neither did. Vagueness did not interact with this effect, suggesting that participants' rapid identification of very small quantities is not affected by whether they are identified using precise or vague quantifiers. The advantage for a subitizable number was smaller when the gap was big than when the gap was small. This interaction can be explained if a big gap size confers a nearly maximal benefit that is not improved by the presence of a subitizable number. When gap size is small, the task is harder, and the benefit conferred by a subitizable number has an opportunity to manifest itself.

Participants were faster to choose the square with the smaller number of dots than they were to choose the square with the larger number of dots. Vagueness exerted a considerable influence on this effect, acting like a damper to even out response speeds across the two conditions. A vague quantifier sped up the choice of the larger number of dots; and slowed down the choice of the smaller number of dots.

The results showed a three-way interaction between subitizability, quantity and vagueness, which is interpreted here. When participants were instructed to choose the smaller number of dots rather than the larger number, they responded faster to instructions that mentioned the number than to vague instructions that identified the same square, and although the presence of a subitizable number speeded responses, it did so equally for vague and numerical quantifiers. In contrast, when participants were instructed to choose the larger number of dots, participants responded slower to precise instructions than to vague instructions, but only when there was no subitizable number. Cases where vagueness was advantageous include 
$pair_{(6)}(7,9)$ with an instruction to choose the square with many dots;  $pair_{(5)}(6,8)$ with an instruction to choose the square with many dots. Cases where vagueness was disadvantageous include $pair_{(1)}(2,4)$ with an instruction to choose the square with few dots; $pair_{(6)}(7,9)$  with an instruction to choose the square with few dots. This suggests that participants had to count the dots only when there was no subitizable number, or in other words that vague quantifiers confer a benefit over precise quantifiers only when the precise quantifiers identify a large number of dots.

\section{Experiment 2}

Experiment 1 showed that vagueness does not exert much influence on the processing of numbers in the subitizable range. Evidence suggests that such numbers are processed very quickly using a different system than larger numbers \cite<e.g.,>[]{trick1994small}. For experiment two we wanted to avoid the potential confound introduced by the inclusion of numbers in the subitizable range. We expected that vagueness would be positively beneficial when larger numbers of dots were involved. This is because we observed reliable advantages of vagueness in the pilot experiment only for choosing the bigger of two numbers of dots (and then only when there was no subitizable number in the display). We decided to use larger numbers of dots in order to explore this finding more thoroughly. We used more levels of gap in this experiment. This meant that we had a kind of index of discriminability, with small gaps resulting in displays with low discriminability, and bigger gaps resulting in increasingly discriminable displays. Because we always used a baseline number of 25 dots, it was possible to directly compare different gap sizes against the same constant baseline.

We also made a change to the procedure of the experiment. In experiment one the response time that we used as a dependent variable included the reading of the instruction -- the participant pressed the key to initiate the trial and the response time before reading the instruction. This could have conflated two separate processes: comprehending the instruction; and selecting a response. In experiment two we tried to separate these processes. We did this by requiring the participant to read the instruction before pressing the key that started the trial and displayed the squares and the dots. The response time was measured from this key press. In this way the response time for experiment two should index only the selection process.

We hypothesised that under these conditions: \begin{seriate} \item RT will be faster for vague than precise instructions; \item Vagueness will be more advantageous at smaller distances (i.e., low discriminability) than at large distances (i.e, high discriminability). \end{seriate}

\subsection{Method}

\subsubsection{Participants}

Twenty participants were recruited and paid in the same way as in experiment one, and similarly aged between 18 and 45, with a median age of 26. All participants self-reported fluency in English, and had normal, or corrected-to-normal vision.

\subsubsection{Apparatus}

A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants. Stimuli were created and presented using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.

\subsubsection{Design}

We presented participants with 256 trials, arranged in 4 blocks of 64 trials each. It was always the case that two squares appeared on screen. One of these always contained 25 dots, and the other contained a number of dots that varied, and which was always the target. We can use the notation $pair_{(p)}(n,25)$ to represent a pair. The pairs were: $pair_{(1)}(5,25)$; $pair_{(2)}(10,25)$; $pair_{(3)}(15,25)$; $pair_{(4)}(20,25)$; $pair_{(5)}(30,25)$; $pair_{(6)}(35,25)$; $pair_{(7)}(40,25)$; $pair_{(8)}(45,25)$.

We manipulated two independent variables: gap size, with four levels (gap size = 5, 10, 15, or 20), and vagueness, with two levels (crisp and vague). We had to control another two variables: target size (whether the referring expression indicated the larger number of dots or the smaller number of dots); and target side (for a given level of target size, whether the target appeared on the left or the right). We measured, as our dependent variables, response time and response accuracy.

Our manipulation on gap size divided the items into four groups, one for each gap size. Items with gap size 5 were: $pair_{(4)}(20,25)$; and $pair_{(5)}(30,25)$; items with gap size 10 were: $pair_{(3)}(15,25)$; and $pair_{(6)}(35,25)$; items with gap size 15 were $pair_{(2)}(10,25)$ and $pair_{(7)}(40,25)$; and items with gap size 20 were $pair_{(1)}(5,25)$ and $pair_{(8)}(45,25)$.

Our manipulation on vagueness meant generating crisp and precise versions of each referring expression. Each pair was presented 8 times with a vague expression of quantity, and 8 times with a crisp expression of quantity. Taking $pair_{(1)}(5,25)$ as an example, the crisp referring expression was ``Choose the square with 5 dots", and the vague referring expression was ``Choose the square with few dots".

We needed to control which side the target was presented on. Each pair was presented equally often with with the target on the left, as with the target on the right.

We also needed to control whether the target was the smaller or larger number. The items with numbers less than 25 (pairs 1 to 4) formed a group with the smaller number as target and the other items (pairs 5 to 8) formed a balancing group with the larger number as the target number

Table (\ref{tablee2}) shows the organisation of trials in experiment 2.

\begin{table}[tbp]
\caption{Properties of stimuli used in Experiment 2, arranged by pair}
\label{tablee2}
\begin{tabular}{lccccl}
&&&gap&&\\
\hline
target&5&10&15&20\\
\hline
	small&		$pair_{(4)}$(20,25)&$pair_{(3)}$(15,25)&$pair_{(2)}$(10,25)&$pair_{(3)}$(5,25)\\
 	big     &		$pair_{(5)}$(30,25)&$pair_{(6)}$(35,25)&$pair_{(7)}$(40,25)&$pair_{(8)}$(45,25)\\
\end{tabular}
\end{table}

\subsubsection{Stimuli}

Each stimulus consisted of two elements \begin{seriate} \item the text of the referring expression indicating the target; \item two squares containing dots. \end{seriate} Where a number was mentioned in the instruction, it was always in the form of an arabic numeral (e.g., 1,2). This contrasts with experiment one, where numbers were given in natural language form. The position of the dots was randomised per-trial.
Fig (\ref{stimuluse2}) gives an example stimulus.

\begin{figure}[tbp]
\fitfigure{images/stimuluse2}
\caption{An example stimulus from experiment 2. First the referring expression was presented (left panel). Then after a keypress, and a fixation cross (not pictured) the squares and dots were presented without repetition of the referring expression (right panel)}
\label{stimuluse2}
\end{figure}

\subsubsection{Procedure}

The experiment was conducted in a quiet room. On arrival in the room, the participant was told that he or she would be presented with an instruction to choose one of two squares by reference to how many dots it contained. Participants were required to press the space key after reading the instruction. Then there was a central fixation cross for 1000 ms, and a blank screen for 500 ms, followed by the squares and dots (without repetition of the referring expression). Response time was measured as the latency between the presentation of the dots and squares, and the keypress identifying the decision. The display would stay on screen until the participant responded (there was no timing-out). Participants were asked to respond quickly while avoiding errors. There were 8 practice trials, after which the participant was invited to ask any questions about procedure. After answering these the experimenter left the cubicle for the duration of the experiment. The order in which trials were presented was randomised per-participant, and stimuli were presented in 4 blocks of 64 trials each, between which the participant could rest. No feedback was given on correct trials, but there was feedback on error trials in the form of the word ``\textsc{wrong!!}" which flashed on screen.

\subsection{Results}

Because the numbers of dots in this experiment were much greater than in experiment one, no maximum RT cutoff was imposed. The same minimum cutoff as experiment one, 250 ms, was used, but no trial had RT less than 250 ms in this experiment. RTs for trials with erroneous responses were discarded, leading to the loss of 354 trials from 5120, representing 6.9\% of the trials.

Means for response times and error rates are given in Fig. (\ref{resultse2}).

An omnibus 2-way (2 x 4; vagueness x gap size) ANOVA was carried out on participant mean correct RT as the dependent variable, and Vagueness and Gap size as the independent variables manipulated within subjects. This analysis averaged RT over Quantity. Grand mean RT was 1088 ms with standard error 125 ms, 95\% confidence interval from 826 ms to 1350 ms. Vague quantifiers attracted reliably faster mean RT than precise quantifiers (precise: 1246ms, vague: 930 ms, $F(1,19)=12.106, p<.01, \eta^2=.389$). There were reliable differences across RT at the different levels of gap (main effect of gap: $F(1.2, 22.793)=17.839, p<.001, \eta^2= .484$, after Huynh-Feldt correction for sphericity violation). Gap and vagueness interacted reliably ($F(1.711, 32.5 = 7.498, p<.01, \eta^2=.283$, after Huynh-Feldt correction for sphericity violation). RTs reliably grew faster as the gap size increased (overall linear trend: $F(1,19)=20.127, p<.001,\eta^2=.514$). The linear trend was significant at each level of vagueness (linear trend in the precise conditions: $F(1,19)=23.611$; linear trend in the vague conditions: $(F(1,19)=10.946, p<.01)$). This linear increase in speed differed reliably between the precise and vague conditions ($F(1,19)=14.140,p<.01,\eta^2=.427$). In the precise conditions RTs started out much slower than in the vague conditions, at the smallest gap size, but the two conditions converged to very fast times at the largest gap size. The interaction of the linear trend across gap size with vagueness is plotted in Fig. (\ref{resultse2}).

\label{accann}
A generalized linear mixed model \cite{jaeger2008categorical} was fit to the error data, with vagueness and gap and their interaction as fixed effects, and with subject as a random effect. Vagueness exerted a reliable effect on accuracy, with precise instructions reliably inducing more errors  than vague instructions: $Wald~z = 2.326, p<.05$). There was a reliable linear trend in error rates across gap, with error rates falling as gap size increased ($Wald~z = -10.203, p<.001$). This linear trend interacted with vagueness ($Wald~z=-5.366,p<.001$). Error rates were greater in the precise conditions than the vague conditions when gap size was small, and that this difference diminished with increasing gap size until it reversed at the biggest gap size.

\begin{figure}[htbp]
\fitfigure{images/resultse2.pdf}
\caption{Experiment 2 results. The plot shows 3 significant trends in response time:  the linear trends for RT to decline across increasing numerical distance in each of the precise and vague conditions; and the significant interaction between these 2 linear trends}
\label{resultse2}
\end{figure}

\subsection{Discussion}

The experiment provided support for hypothesis 1, that RT would be faster for the vague instruction conditions than for the precise instruction conditions. The experiment also provided support for hypothesis 2, that  vagueness would be more advantageous at smaller than at larger distances.  The findings of experiment two can be summarised thus: vagueness is easier to process than crispness, and returns for vagueness diminish as the stimuli become more easily discriminable. We consider here four explanations of this pattern.

\begin{seriate}
\item The \emph{cost reduction hypothesis} explains the vagueness advantage by claiming that the vague referring expressions place less cognitive load on the comprehender than the crisp referring expressions. It explains the diminishing returns for vagueness in more-discriminable stimuli by claiming that load is low in both conditions for the easily-discriminable stimuli, and that therefore there is no extra benefit to be had from vagueness in the easily-discriminable stimuli. A possible explanation for the pattern of results is cast in terms of the precision of an estimate of quantity that is required to carry out the task. For example, in the display pair$_{(1)}(5, 25)$, with a large distance, vague instructions (\emph{many}, \emph{few}) might be able to  be carried out by a visual comparison that identifies one square as more numerous than the other. Precise instructions (to identify the square with 5 dots) might be able to be carried accurately out with an estimate with a fairly large error of 19. In the display pair$_{(1)}(5, 25)$, which has a small distance, vague instructions might be able to be be carried out with a visual comparison that identifies one square as more numerous than the other. Precise instructions might require an estimate with a maximum error of 4 in this case. The pattern of results can be explained if we assume that the degree to which an estimate with error 4 is harder than an estimate with error 19 is greater than the degree to which a visual comparison between 20 and 25 is harder than a visual comparison between 5 and 25.

\item An account that we will dub the \emph{instruction format hypothesis}, or the number / no-number account, explains the vagueness advantage by placing importance on the observation that the vague conditions did not mention a numeral whereas the crisp conditions did. This could make the difference between a task that taps numerical processing, and one that taps magnitude judgements. The diminishing returns for vagueness can be explained by assuming that the numerical task is particularly challenging when the stimuli are less-discriminable, whereas the magnitude judgement is relatively less difficult in the less-discriminable trials.

\item \citeA{Rayna:1989lr} can account for the pattern in a framework they called the \emph{hierarchy of gist}. They proposed that reasoners encode representations at different levels of precision, and that these representations can be ordered with respect to precision, forming a hierarchy of gist. They further claimed that reasoning operated at the lowest level in this hierarchy that would allow one to accomplish the assigned task. Among the levels of this hierarchy are  (1) ratio representations at the top, which encode numerical information exactly; (2) ordinal representations that capture relative magnitude but do not encode numerical differences, i.e., they establish a rank ordering of quantities by magnitude, e.g., \emph{large}, \emph{larger}, \emph{largest}; (3) nominal or categorical representations that capture only the presence of absence of quantity e.g., \emph{some lives}; \emph{no lives}. \citeA{brainerd1994development} presented evidence that children spontaneously encode relative numerosity, at the ordinal level of the hierarchy mentioned above, when they are presented with stimuli that vary in numerosity. The hierarchy of gist offers an explanation of our vagueness advantage in experiment two that is independent of vagueness in the \citeA{keefe1996vagueness} sense. In experiment two, the participants were presented with the instruction for the trial first, and then with the squares and dots. This was done in order to try to separate the processes involved in processing the instruction from those involved in processing the stimuli. A consequence of this serial presentation though is that participants can infer from the instruction which level of representation is the minimum level necessary to carry out the impending task successfully. We do not suggest that this takes the form of a conscious deliberation -- rather it might be apprehended spontaneously in the same way that the children in \citeA{brainerd1994development} did without conscious deliberation. For example, it follows naturally from the instruction \emph{Please choose the square with fewest dots} that an ordinal representation will be necessary and sufficient for the purposes of the current trial: similarly it follows naturally from the instruction \emph{Please choose the square with 34 dots} that a higher level of the hierarchy will be necessary and suficient for the purposes of this trial - the ratio level that encodes numerosity exactly. We found a response time advantage in experiment two for instructions that used the quantifiers \emph{many} and \emph{few}, when compared with times for quantifiers like \emph{15} and \emph{45}. In the terms of \citeA{Rayna:1989lr}, this can be explained as an advantage for the sufficiency of a lower level of representation -- without reference to vagueness.

\item The \emph{two-systems account} of decision making \cite<e.g.,>{sloman2007two} can offer an explanation of the pattern. Dual process models of decision making propose two systems of processing: the quick and affective System 1 and the deliberative and rule-based System 2. This two-systems account can explain our finding of diminishing returns for vagueness in experiment 2. When the gap is small, participants use System 2 (slow, deliberative) for the precise instructions. When the gap is large, they take a shortcut for the precise conditions and merely establish which square is more (or less) numerous, using System 1 (quick, heuristic). In the vague conditions, they use the heuristic system whether the gap is small or large. This account explains the diminishing returns for vagueness that we observed in terms of a change of strategy in the baseline precise conditions as the gap size grew larger.
\end{seriate}

The potential for vagueness in the vague conditions can be said to have been unrealised in experiment two. In both experiment one and experiment two, the vague instructions were  \emph{choose the square with many dots} and \emph{choose the square with few dots}. \emph{Many} and \emph{few} have the potential for vagueness defined as the existence of borderline cases. Given a square with 15 dots, and asked the question `are there many dots in the square?', we can imagine that some people would be unsure (contrast this with the questions `are there fewer than 20 dots in the square?' to which there is no room for uncertainty). However, this potential for vagueness may be unrealised in the context of a choice between two squares. Given two squares, one with 20 dots, and another with 25 dots, and asked to `choose the square with few dots', it seems that no one could be unsure which square is the intended target. If there is no room for uncertainty, then the potential for \emph{many} and \emph{few} to be vague can be said to be unrealised. On these grounds experiments one and two can be said not to have used vague conditions with true vagueness.

\section{Experiment 3}

In experiments one and two, the participants chose one of two squares. The `vague' quantifiers (e.g., `few') uniquely identified one square. Recall our definition of vague  -- ``a word is precise if it describes a well-defined set of objects. By contrast, a word is vague if it is not precise".  In the first two experiments, the quantifiers in the vague conditions did not really meet this definition. This is because there were no borderline cases of the referent that could make the referent set `not well-defined'. Experiment three used three squares so that the vague quantifiers always had more than one possible referent. To enhance the potential for the vague quantifiers to have true vagueness, we also used indefinite articles in the vague instructions. Thus, instead of the instruction \emph{choose the square with few dots}, we used \emph{choose a square with few dots}.

The core finding in experiment two was that vagueness exerted a beneficial effect compared to precise equivalents in the domain of quantifiers. A problem with this finding is that vagueness in the strict sense that we are interested in was confounded in that experiment with the format of the quantifier. For example, contrast an expression from the vague condition: `the square with few dots' with an expression from the precise condition: `the square with 15 dots'. One difference is that `few' is vague (or at least has the potential for vagueness) and `15' is precise. Another difference is that `few' is a linguistic, or verbal quantifier while `15' is a numerical quantifier, in the sense that a number is mentioned explicitly. Since these two differences were confounded in experiment two, the finding is vulnerable to an alternative interpretation, that our vagueness advantage was in contrast an advantage for the linguistic or verbal form of the quantifier. The present experiment, experiment three, pitted these two alternative interpretations against each other in a factorial design.

The cost reduction hypothesis predicts that there will be a main effect of vagueness such that the vague instructions confer a processing advantage relative to the precise instructions.  In contrast, the instruction format hypothesis predicts that there will be a main effect of instruction format, with verbal format being advantageous compared to numerical format.

\subsection{Method}

\subsubsection{Participants}
Thirty participants were recruited and paid in the same way as experiment 2. They were aged between 18 and 45 with a median age of 28. All participants self-reported fluency in English, and had normal, or corrected-to-normal vision.

\subsubsection{Apparatus}
A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants. Stimuli were created and presented using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.

\subsubsection{Design}
We presented participants with 256 trials, arranged in 4 blocks each with 64 trials.

We used four different triples of numbers: $triple_{(1)}(6,15,24)$; $triple_{(2)}(16,25,34)$; $triple_{(3)}(26,35,44)$; $triple_{(4)}(36,45,54)$. Each triple had the following properties: it comprised three squares (instead of two as in experiments one and two); the central number was always presented in the middle of the three; there were two flanking numbers where one was smaller than the central number and one was bigger. The triples are given in Table \ref{tablee3}.

\begin{table}[tbp]
\caption{Table of stimuli}
\label{tablee3}
\begin{tabular}{rcccc}
triple&1              &2              &3              & 4\\
              \hline
contents              &6:15:24  &16:25:34 &26:35:44&36:45:54 \\
\end{tabular}
\end{table}

We manipulated two independent variables: Vagueness with two levels (vague and crisp); and Format with two levels (numerical, verbal). This yielded four conditions (vague numerical; vague verbal; crisp numerical; crisp verbal). Each condition had a different referring expression, as follows, using $triple_{(1)}(6,15,24)$ as an example. 
The vague numerical condition's referring expression was ``Choose a square with about 10 dots". No square contained 10 dots. 10 is slightly closer to 6 than to 15. Therefore the best referent for this referring expression was the square with 6 dots;  the borderline response was the square with 15 dots; and the poorest referent was the square with 24 dots. The crisp numerical conditions's referring expression was ``Choose the square with 6 dots" on half the presentations in that condition and ``Choose the square with 24 dots" on the other half. One square did contain the exact number mentioned. The square with 15 dots was the borderline response; and the remaining square was the poorest referent. The vague verbal condition's referring expression was ``Choose a square with few dots" on half of the presentations, and ``Choose a square with many dots" on the other half. For this condition, the best referent was the square with 6 dots for `few' and 24 dots for 'many'; the borderline case was the square with 15 dots; and the remaining square was the poorest referent. The vague numerical condition's referring expression was ``Choose the square with the fewest dots" on half of presentations, and ``Choose the square with the most dots" on the other half. For this condition, the square with 6 dots (for few; 24 dots for many) was the best referent; the square with 15 dots was the borderline case, and the remaining square was the poorest referent. Table \ref{instre3} gives examples of an instruction from each condition.

\begin{table}[tbp]
\caption{Table of instructions arranged by condition for the triple (6,15,24)}
\label{instre3}
\begin{tabular}{rll}
&crisp&vague\\
\hline
numerical&Choose the square with 6 dots 		& 	Choose a square with about 10 dots			\\
verbal&Choose the square with the fewest dots			&		Choose a square with few dots\\
\end{tabular}
\end{table}

Each triple was presented in each condition 16 times. 8 of these identified the larger number, and 8 the smaller.

We measured two dependent variables: response time; and the probability of a participant choosing the borderline case.

\subsubsection{Stimuli}
Fig (\ref{stimuluse3}) gives an example stimulus. First, the referring expression that constituted the instruction for that trial was displayed. The participant then pressed a key to indicate that he or she had read the instruction. After 1 second,  the squares and dots were presented, while preserving the text of the referring expression. The position of the dots in the squares was randomised per-trial.

\begin{figure}[tbp]
\fitfigure{images/stimuluse3}
\caption{An example stimulus from experiment 3}
\label{stimuluse3}
\end{figure}

\subsubsection{Procedure}
The experiment was conducted in a small quiet room.
On arrival in the cubicle, the participant was told that he or she would be presented with objects on screen and required to choose one in response to an instruction on screen, by pressing the button corresponding with the object. There were 5 practice trials. After the practice trials the experimenter left the room. There were 4 blocks of 64 trials each. In between blocks the participant had the opportunity to rest before continuing. The response time dependent variable was measured from the presentation of the squares and dots, until the keypress indicating the participant's choice. The trial would timeout after 60 seconds if there was no response. The dependent variable measuring whether the participant chose the borderline case was also recorded at this time. In this experiment, no feedback was given. This was because, in the vague conditions, we did not regard any response as `correct' or `incorrect', but instead as `borderline response', or `not borderline response', and we did not want to draw participants' attention to this distinction explicitly. We simply recorded whether the participant chose the borderline case or not, and how long it took the participant to respond.

\subsection{Results}

No responses were treated as erroneous, because errors were essentially undefined for the vague instructions (e.g., `about ten'). It was noted whether participants chose the borderline square. Trials with RT less than 250 ms, and trials with RT greater than 50,000 ms were discarded from RT analysis. This led to 3 trials being discarded, representing less than one percent of the datapoints (.039\% of the datapoints; 3 of 7680 trials).

A three-way (2 x 2 x 2) repeated measures ANOVA was carried out, with Quantity (2 levels, instruction to choose the biggest number; instruction to choose the smallest number), Task (2 levels, Numerical instruction, Verbal instruction), and Vagueness (2 levels, Crisp and Vague) entered as independent variables manipulated within-subjects in a factorial design. Grand mean RT was 2842 ms, with standard error 331 ms. There was a significant main effect of task, but no other main effects were significant, and no interactions were significant. Figure (\ref{resultse3}) shows response time as a function of stimulus and instruction format. Mean responses to the instructions to choose the biggest square were slightly faster than to the instructions to choose the smallest square, but the difference was not significant (2817 vs 2868 ms, $F(1,29) < 1$). Responses to the numerical  task instructions were reliably much slower than responses to the verbal  task instructions (3773 ms vs 1911 ms, $F(1,29) = 13.369, p<.01, \eta^2 = .316$). Mean responses to the crisp instructions were slightly faster than to the vague responses, but the difference only approached significance at the 5\% level (2751 vs 2933 ms, $F(1,29)=3.470, p=.073$). The quantity x vagueness interaction was not reliable ($F(1,29)=2.913,p=.099$). The task x vagueness interaction was not reliable ($F(1,29)=2.311,p=.139$). The quantity x  task interaction was not reliable ($F(1,29)=1.785,p=.192$). The three-way interaction between quantity, task, and vagueness was not reliable ($F(1,29)=1.753,p=.196$).

Participant grand mean percentage of borderline selections was $16.6\%$. A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with task, vagueness and quantity as fixed effects, and with random intercepts for subject and item. Results are plotted in Figure (\ref{resultse3}). There was a significant main effect of each of task, vagueness, and quantity;  a significant interaction between task and vagueness; and a significant three-way interaction between task, vagueness, and quantity. Participants were significantly more likely to choose the borderline option for vague instructions than for precise instructions (21.9\% vs 11.3\%, $p<.001$), and reliably more likely to choose the borderline response for the smaller quantity than for the larger quantity (12.2\% vs 20.9\%, $p<.001$). Participants were also reliably more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%, $p<.001$). Task and vagueness interacted reliably ($p<.001$) such that the propensity to choose the borderline case in response to numerical format instructions was greater when they were expressed vaguely (e.g., `about 20'). The three-way interaction between task, vagueness, and quantity was due to bigger quantity generally being less likely to result in a borderline response, except in the verbal, vague conditions where instructions to choose the bigger quantity were more likely to result in a borderline response.

\begin{figure}[htbp]
\fitfigure{images/resultse3.pdf}
\caption{Results for experiment 3}
\label{resultse3}
\end{figure}

\subsection{Discussion}

This experiment was designed to test the hypothesis that vagueness would interact with instruction format. If found, such an interaction would show that the benefit for vagueness that was observed in experiment two was additional to any benefit due to the verbal instruction format. There was no such interaction. Instead, instruction format exerted a large effect that was not reliably modulated by vagueness. Thus it appears that the vagueness advantage elicited in experiment two was really a benefit of the verbal instruction format over the numerical format.

A potential confound is that the no-number quantifiers are asymmetrical while the precise quantifiers are symmetrical. For example, given triple$_{(1)}(6,15,24)$, the expressions \emph{about 10} and \emph{10} are symmetric about 10, while the expressions \emph{few} and \emph{fewest} are asymmetric about 10 -- the distribution is skewed towards lower numbers. Support for this confound is that in the skewed conditions, the borderline case was chosen less frequently than in the symmetric conditions. In other words, people were more likely to choose the middle value when the distribution encompassed this value than when it was skewed away from it.

In experiment three, where the potential for vagueness was realised in the form of borderline cases, verbal-vague (the equivalent of the vague condition in experiment two) attracts faster times than numerical-crisp (the equivalent of the crisp condition in experiment two). This is the same pattern that we observed in experiment two. This indicates that the vague conditions in experiment two were not faster as a result of the lack of borderline cases; but rather because they did not specify a number in the instruction.

\section{General Discussion}

Several answers to Lipman's question -- why vagueness is as prevalent as it is in human communication -- have been suggested \cite{van2009utility,van2010vagueness}. In this paper, we have tested one them, namely the cost reduction hypothesis. As far as we know, no other answers to Lipman's question have been subjected to experimental scrutiny yet. Has the cost reduction hypothesis been confirmed by our experiments? In other words, do readers benefit from vagueness, as the cost reduction hypothesis asserts?

At one level, the answer to these questions is affirmative. In all our experiments, we have found consistent and fairly substantial benefits, in terms of hearers' reaction times, for referring expressions involving vague quantifiers like ``many" over crisp ones involving numerical quantifiers, such as ``25". An example of an NLG domain that could benefit from these results is the GIVE challenge \cite{byron2009report} \cite{ koller2010report}. This is an environment where NLG systems issue instructions to human comprehenders who are navigating a virtual environment. A typical situation might involve a panel with ten buttons, for example, that the human should approach in the virtual world in order to press a particular button, when more than one panel with buttons on it is available in the immediate environment. Our results suggest that an instruction like  \emph{approach the panel with many buttons}  that refers to the panel without reference to the particular number might be processed faster by the human player than an instruction that specifies the number like  \emph{approach the panel with ten buttons}, and therefore be a better referring expression for the NLG system to produce. This is a potentially useful result, although it is as yet unclear whether it generalises to situations where the quantifier is not used as part of a reference task.

At a deeper level, the answer must be negative, because our third experiment shows that the benefits reported in the previous paragraph were not a consequence of the vagueness of the words involved (i.e., of the fact that they allow borderline cases), but of the fact that they used a non-numerical expression, that is, the description does not rely on counting. In other words, if the choice for the NLG system mentioned above is broadened to include expressions that are both verbal and crisp, as well as ones that are both numerical and vague, then the only advice that can be given on the basis of our experiments is that the numerical ones should be avoided; we found no clear difference, in terms of benefit for readers, between verbal descriptions that are crisp and those that are vague.

\bibliography{vcr}

\end{document}

### 2016-07-27 presentations/nlg_talk_june_2011/nlg_group_talk_7_june_2011_template.tex
adding presentations, short conference papers, and posters

\documentclass{beamer}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{beamerthemesplit}
\usepackage{graphics,epsfig, subfigure}
\usepackage{apacite}
\usepackage{amsfonts}
\newcommand{\tickYes}{\checkmark}
\usepackage{pifont}
\newcommand{\tickNo}{\hspace{1pt}\ding{55}}
%\usepackage{apalike}
\usepackage{url}

\definecolor{kugreen}{RGB}{50,93,61}
\definecolor{kugreenlys}{RGB}{132,158,139}
\definecolor{kugreenlyslys}{RGB}{173,190,177}
\definecolor{kugreenlyslyslys}{RGB}{214,223,216}
\setbeamercovered{transparent}
\mode<presentation>
{  \usetheme{PaloAlto}
  \usecolortheme[named=kugreen]{structure}
  \useinnertheme{circles}
  \usefonttheme[onlymath]{serif}
  \setbeamercovered{transparent}
  \setbeamertemplate{blocks}[rounded][shadow=true]
}
%\setbeamertemplate{background}{\includegraphics[width=1\textwidth]{natfak_baggrund.pdf}}
\logo{\includegraphics[width=1cm]{../logo/Crest_Colour}}

%\title{Et bachelorprojekt om roterende vand}
%\subtitle{Et heldigt tilf¾lde}
%\author{Thomas R. N. Jansson}
%\institute{Niels Bohr Institute \\ University of Copenhagen}
%\date{9 November 2007}

% \usepackage{beamerthemesplit} // Activate for custom appearance

\title{The utility of vagueness}
\subtitle{an empirical test of the cost reduction hypothesis in a reference task}
\author{Matt Green \& Kees van Deemter}
\institute{NLG group \\ University of Aberdeen}
\date{7th June 2011}

\begin{document}

\frame{\titlepage}

\section[]{}
\frame{\frametitle{Outline}
\begin{enumerate}
\visible<1->{\item Rationale (what, why, how?)}
\visible<2->{\item Experiment one (tests the water)}
\visible<3->{\item Experiment two (extends the findings)}
\visible<4->{\item Interim summary (where do we stand now?)}
\visible<5->{\item Experiment three (addresses problems with E1 \& E2)}
\visible<6->{\item Interpretation (what does it all mean?)}
%\item Take home message
\end{enumerate}
%\tableofcontents}
}
\section[Rationale]{Rationale (what, why, how?)}
\frame{
            \frametitle{A definition of vagueness}
            Vagueness can be defined as the existence of borderline cases 
            \begin{itemize}
            \item e.g., \emph{tall}
            \end{itemize}}

%\section[Backgound]{Why study the effects of vagueness?}
\frame{
            \frametitle{Why study the effects of vagueness?}
            2 main reasons
            \begin{enumerate}
            \item Game theory needs an update
            \begin{itemize}
            \item Game-theory predicts that precision always `wins'
            \item But we use vague language a lot
            \item There must be some utility of vagueness that the models don't account for
            \end{itemize} 
           \visible<2->{ \item NLG systems need help
            \begin{itemize}
            \item Practical NLG systems must decide between generating vague and precise terms \cite<e.g., >[]{portet2009automatic}
            \item These decisions are (almost) guesswork currently
            \item Data from human comprehenders would provide a better basis for these decisions
            \item There isn't much data like this at the moment
            \end{itemize} }
            \end{enumerate}}

%\section[]{Example NLG decision}
\frame{
            \frametitle{Example NLG decision}
            Which of these is better?
            \begin{enumerate}
%            \item ``By 10:29 there had been 2 successive spikes in TcPO2 up to 18.1"
%            \item ``There were a couple of spikes in the TcPO2 trace"
\item `Winds southwest $15$ to $25$ knots \textbf{becoming southeast 15} later this evening'
\item `Winds southwest $15$ to $25$ knots \textbf{diminishing to light} later this evening'
            \end{enumerate} 
            \visible<2->{\begin{itemize}
            \item The answer might depend on the audience: offshore oil rig, mountain resuce, flying doctor, air traffic control \ldots
            \end{itemize}}
            \visible<3>{\begin{itemize}
            \item It depends quite a lot what `better' means: speed, accuracy, recall, comprehension, effect on affect \ldots
            \end{itemize} } }

%\section{Vagueness in reference}
\frame{\frametitle{Vagueness in reference}
\begin{itemize}
	\item Vagueness in reference is highly context-dependent
	\item \underline{Potential} for vagueness is not necessarily \underline{realised}
\end{itemize} }

%\frame{\frametitle{Pinning vagueness down}
%Vagueness, or something else?\\
%\begin{itemize}
%\item Several, separate research comunities
%\begin{itemize} 
%\item Vagueness $\approx$ low specificity
%\item Vagueness $\approx$ large grain size
%\end{itemize}
%\end{itemize}
%}
%

\frame{\frametitle{Pinning utility down}
\begin{itemize}
\item Utility in \underline{co-operative} situations is our focus
\item Non-co-operative situations where vagueness has utility
	\begin{itemize}
	\item Sales, marketing, advertising \tickNo
	\item Politics, campaigning \tickNo
	\item Persuasion, behaviour modification \tickNo
	\end{itemize}
\item Co-operative situations where vagueness may have utility
	\begin{itemize}
	\item Local vagueness \tickNo
	\item Information reduction \tickNo
	\item Reducing cognitive load on the comprehender \tickYes
	\end{itemize}
\end{itemize}
}

\frame{\frametitle{The cost reduction hypothesis} 
``For the listener, information which is too specific may require more effort to analyze" \cite{lipmanvague}
}

\section[Test the water]{Experiment one (tests the water)}

\frame{\frametitle{Experiment one}
%            \frametitle{Stimulus}
            \begin{figure}
            \includegraphics{/Users/mjgmac562/Documents/vagueness-folder/writing/figures/drawingexp1b}
            \end{figure}}

\frame{\frametitle{Experiment one}
\begin{itemize}
\item Pilot experiment
\item $H_{0}$: \emph{No difference between vague and precise}
\item $H_{1}$: \emph{Some difference between vague and precise\\  ~~~~~ \ldots V better than P?}
\item Method: forced choice
\item DVs: RT, errors
\item IVs: Vagueness (Vague, Precise)
\item Number combinations: \{$2,4$\}, \{$2,6$\}, \{$3,5$\}, \{$5,9$\}, \{$6,8$\}, \{$7,3$\}, \{$7,9$\}, \{$8,4$\}
\end{itemize}
}

\frame{
	\frametitle{Results, experiment one}
	\begin{figure}
            \includegraphics[width=\textwidth]{/Users/mjgmac562/Documents/vagueness-folder/writing/figures/e1resultscolourpresentation}
            \end{figure}
	}

\section[Extend findings]{Experiment two (extends findings)}

\frame{\frametitle{Experiment two}
\begin{itemize}
\item Try larger numbers (not subitizable)
\item $H_{0}$: \emph{No difference between vague and precise}
\item $H_{1}$: \emph{Vagueness helps when the numbers are bigger, and when the difference between the numbers is greater}
\item Number combinations: \{$5,25$\} \{$10,25$\} \{$15,25$\} \{$20,25$\} \{$30,25$\} \{$35,25$\} \{$40,25$\}
\end{itemize}
}

\frame{\frametitle{Instruction first}
\begin{figure}
 \includegraphics[width=.8\textwidth]{/Users/mjgmac562/Documents/vagueness-folder/writing/figures/drawingexp2instruction}
\end{figure}
}

\frame{\frametitle{Stimulus next}
\begin{figure}
 \includegraphics[width=.8\textwidth]{/Users/mjgmac562/Documents/vagueness-folder/writing/figures/drawingexp2stimulusonly}
\end{figure}
}

\frame{
            \frametitle{Results, experiment two}
            \begin{figure}
            \includegraphics[width=\textwidth]{/Users/mjgmac562/Documents/vagueness-folder/writing/figures/e2resultscolourpresentation}
            \end{figure}
            }

\section[Where do we stand?]{Interim summary (where do we stand?)}
\frame{\frametitle{Interim summary (where do we stand?)}
\begin{itemize}
\item Vagueness is worse for very small numbers (expt 1)
\visible<2->{\item Vagueness is better for larger numbers (expt 2)}
\visible<3->{\item Diminishing returns for vagueness as gap size grows very large (expt 2)}
\end{itemize} }

\frame{\frametitle{Issues to address}
\begin{itemize}
\visible<1->{\item Potential for vagueness not realised? (expts 1 \& 2)}
\visible<2->{\begin{itemize}\item Definite articles uniquely identify}
\visible<3->{\item Two squares means no borderline case}
\visible<4->{\item Solution: use indefinite articles; use $>2$ squares\end{itemize}}
\visible<5->{\item Vagueness confounded with absence of a number in the instructions?}
\visible<6->{\begin{itemize} \item V: \{few, many\}; P: \{5, twenty\} }
\visible<7->{\item Solution: factorially manipulate instruction format 2 x 2}
\visible<8->{ \begin{table}[htdp]
%\caption{default}
\begin{center}
\begin{tabular}{c|c|c}
                     & Vague                              &  Precise \\ 
                     \hline
Numerical  &  \{about 20, about 30\}  &  \{16, 34\} \\ 
\hline
Verbal         & \{few, many\}                    &  \{fewest, most\}  \\
\hline
\end{tabular}
\end{center}
\label{default}
\end{table}%
}\end{itemize} \end{itemize} }
%; Precise-numerical: \{16, 34\}, Precise-verbal \{fewest, most\}, Vague-verbal: \{few, many\}; Vague-numerical \{about 20, about 40\}  }\end{itemize} \end{itemize} }

\section[Address problems]{Experiment three (addresses problems)}

\frame{
            \frametitle{Stimulus, experiment three}
            \begin{figure}
            \includegraphics[width=.7\textwidth]{/Users/mjgmac562/Documents/vagueness-folder/writing/figures/drawingexp3}
            \end{figure}
            }

\frame{
            \frametitle{Operationalising borderline cases}
            \begin{figure}
            \includegraphics[width=.7\textwidth]{/Users/mjgmac562/Documents/vagueness-folder/writing/figures/borderline}
            \end{figure}
            Number combinations: \{$6,15,24$\}, \{$16,25,34$\}, \{$26,35,44$\}, \{$36,45,54$\}
\begin{itemize} \item            Numerical-precise: Choose the square with 16 dots
           \item Numerical-vague: Choose a square with about 20 dots
           \item  Verbal-precise: Choose the square with fewest dots
            \item Verbal-vague: Choose a square with few dots\end{itemize}
            }

\frame{
            \frametitle{Experiment three (more squares \& use indefinites)}
            \begin{figure}
            \includegraphics[width=1\textwidth]{/Users/mjgmac562/Documents/vagueness-folder/e3/figures/borderlineresponseplottaskvgncolourpresentationfirstslide}
            \end{figure}
            }

\frame{
            \frametitle{Experiment three (more squares \& use indefinites)}
            \begin{figure}
            \includegraphics[width=1\textwidth]{/Users/mjgmac562/Documents/vagueness-folder/e3/figures/borderlineresponseplottaskvgncolourpresentationsecondslide}
            \end{figure}
            }

\frame{
            \frametitle{Experiment three (more squares \& use indefinites)}
            \begin{figure}
            \includegraphics[width=1\textwidth]{/Users/mjgmac562/Documents/vagueness-folder/e3/figures/borderlineresponseplottaskvgncolourpresentationthirdslide}
            \end{figure}
            }

\frame{
            \frametitle{Experiment three (more squares \& use indefinites)}
            \begin{figure}
            \includegraphics[width=1\textwidth]{/Users/mjgmac562/Documents/vagueness-folder/e3/figures/borderlineresponseplottaskvgncolourpresentationfourthslide}
            \end{figure}
            }

\section[What does it all mean?]{Interpretation (what does it all mean?)}
\frame{
            \frametitle{Interpretation and Conclusions}
            \visible<1->{Summary of findings}
           \visible<2->{ \begin{itemize} 
            \item We did find some benefits for using some vague terms \ldots}
            \visible<3->{\item \dots but verbal format is largely responsible}
            \end{itemize}
            \visible<4->{Summary of explanations}
            \visible<5->{\begin{itemize} 
            \item Hierarchy of gist account}
           \visible<6->{ \item Heuristic / Systematic account
            \end{itemize} }
            \visible<7->{Lessons for NLG}
            \begin{itemize}
            \visible<8->{\item Data from human comprehenders can help NLG systems make better decisions}
            \visible<9->{\item Vagueness is really hard to pin down experimentally}
            \visible<10->{\item Using vague terms can make a big difference for cognitive load on comprehenders}
            \end{itemize} 
            }
%\frame{\titlepage}
%
%\section[Outline]{}
%\frame{\tableofcontents}
%
%\section{Introduction}
%\subsection{Overview of the Beamer Class}
%\frame
%{
%  \frametitle{Features of the Beamer Class}
%
%  \begin{itemize}
%  \item<1-> Normal LaTeX class.
%  \item<2-> Easy overlays.
%  \item<3-> No external programs needed.      
%  \end{itemize}
%}

%\renewcommand*{\refname}{}
\section[]{Take home message}
\frame{
\frametitle{Take home message}
\visible<2->{Vagueness does make things easier in some situations, but really it is non-numerical format that does the heavy lifting.}
\begin{center}
\visible<1->{\includegraphics[width=.4\textwidth]{/Users/mjgmac562/Documents/vagueness-folder/writing/figures/barometer}}
\includegraphics[width=.4\textwidth]{/Users/mjgmac562/Documents/vagueness-folder/writing/figures/digitalbarometer}\\
~~\\
\visible<3->{Thank you for listening. 5 mins for Q \& A}
\end{center}
}

\frame{
            \frametitle{Next}
            \begin{figure}
            \includegraphics[width=1\textwidth]{/Users/mjgmac562/Documents/vagueness-folder/writing/figures/lifesavinggraph}
            \end{figure}
            }

\renewcommand*{\refname}{}
\section*{}
\frame{
\frametitle{References}
\bibliographystyle{apacite}
\tiny{\bibliography{/Users/mjgmac562/Documents/vagueness-folder/writing/bibliography/vag.bib} }
}

\end{document}

### 2016-07-27 papers/2015/internal_response_to_reviews/Kees_response_to_reviews.txt
adding papers

Hi Matt,

As promised, a short summary of our plans regarding our LCP paper. Perhaps you can take a quick look to see if you think this is where we stand.

===========================================================================================================================

Summary of discussion on our LCP submission, 17 May 2012.

Timing: Your thesis has higher priority, but we should try to keep the momentum, so we can resubmit not too long after our thesis submission (in early October?)

Issues from the reviews:

1. We want to keep the focus on vagueness (contra reviewer 3) but (as suggested by reviewer 1) we need to flag at an early stage that vagueness is difficult to isolate for a number of reasons:

(a) the difference between potential/actual vagueness
(b) the difference between being vague and not containing a number
(c) the difference between vagueness and non-specificity (as in Mishra's work for instance)

2. We need to be clearer perhaps about the research question that we're testing. The suggestion from reviewer 2, to state alternative hypotheses before each experiment, might be good.

3. We need to read all the suggested literature: Dehaene 97, Feigenson et al 2004, Dehaene 2008, Sauerland & Stateva 2007. The latter contends that "about x" is not vague. Maybe a quick check of this paper will suffice to dismiss it? (There could be a link with the issue 1(c) above.)

4. Consider whether we want to measure reading times (as suggested by reviewer 1). If so, what method should we use? Also, should this influence the choice of materials (e.g., should we try harder to find expressions of similar length?) It's hard enough to find suitable materials as it is, so we should probably avoid further complications ...

5. In a new version of Experiment 3, we should probably make the gaps between numbers proportional to the size of the numbers.

6. The big issue: finding out how much evaluation procedures contribute to the "cost" of a description. Concretely, reviewer 1 worries that the difference that we describe to presence/absence of Number might actually be caused by the possibility/impossibility of doing the task on the basis of a comparison strategy (i.e., "choose the largest/smallest of the two").

-- The first suggestion made by he reviewer might be useful: he suggests to compare "find the square with many dots" vs. "find the square with more than x dots". Both of these can be solved by a comparison strategy, so they're alike in that regard. Our hypothesis would be that the first one is easier. This sounds fine, but can't we argue that we have done almost precisely this as part of experiment 3, when we compared "few" and "fewest"?

-- One could think of an analogous experiment where none of the two cases can be resolved using comparison. For example, "with about x dots" vs "with x dots". On the other hand, we did exactly this as part of experiment 3 ... The question that this raises is whether perhaps we can address this reviewer's main criticism without doing a new experiment.

-- A possible new version of experiment 3, to address the comparison strategy confound. Ideally we'd look at all 8 combinations of YesNumber/NoNumber, YesVague/NoVague, YesComparison/NoComparison. "YesComparison" means that the task can be solved by a simple comparison (as opposed to laborious counting). Here's an imperfect attempt:

YN, YV, YC: far fewer than 10
YN, YV, NC: about 10  -- I do find this an intriguing contrast. We could do this as a separate experiment (analogous to the suggestion of reviewer 1).

YN, NV, YC: fewer than 10
YN, NV, NC: 10

NN, YV, YC: few
NN, YV, NC: few but not very few (?) -- Wouldn't subjects simply choose the one in the middle? (What if we used more than 3 numbers?)

NN, NV, YC: fewest
NN, NV, NC: fewest except one (?) -- Problem: this is solvable through a slightly more complex kind of comparison. Can we find a better way to fill this slot?

Can we perhaps argue that the NN part of this table is impossible to complete, so the best we can do is an experiment focussing on YN (the first four items in the list)? This strikes me as an approach that might satisfy reviewer 1.

========================================================================================================================

k.

### 2016-07-27 papers/2015/cover_letter/lcn-cover-letter.tex
adding papers

%basic cover letter template
\documentclass{letter}
\usepackage{amssymb,amsmath}
\usepackage{graphicx}

\oddsidemargin=.2in
\evensidemargin=.2in
\textwidth=5.9in
\topmargin=-.5in
\textheight=9in

\address{Matthew Green and Kees van Deemter\\Department of Computing Sciences\\University of Aberdeen\\Aberdeen, AB24 3UE\\tel: 07896310566\\ email: mjgreen@abdn.ac.uk}

\newcommand {\qed}{\mbox{$\Box$}}
\renewcommand {\iff}{\Longleftrightarrow}
\newcommand {\R}{\mathbb{R}}
\newcommand {\N}{\mathbb{N}}
\newcommand {\Q}{\mathbb{Q}}
\newcommand {\Z}{\mathbb{Z}}

\newcommand {\sub}{\mbox{SB}}

\begin{document}
\begin{letter}{Editor\\
Language, Cognition and Neuroscience}

\opening{Dear Editor,}

We are pleased to submit our manuscript ``Vagueness as cost reduction" for consideration for publication in Language, Cognition and Neuroscience. The manuscript is co-authored by Matthew J. Green and Kees van Deemter (Department of Computing Science, University of Aberdeen). The manuscript is not currently under review with any other journal.

A previous version of the manuscript was submitted to what was then Language and Cognitive Processes on 01-Nov-2011, with Manuscript ID PLCP-2011-OP-9118. We received helpful reviews in response to that submission and we will refer to those reviews below. The original submission contained experiments 1 to 3 of the current submission, which adds experiments 4 and 5.

Vagueness is pervasive in everyday language use, but its pervasiveness  is not well understood. From some perspectives, including that of some game theorists, a precise, crisp term should always be preferable to a vague equivalent. Vague terms create potential misunderstandings, and can be considered to impose an efficiency loss. However, this efficiency loss is unlikely to have arisen unless vague terms confer advantages as well as disadvantages. At present there is little published work that investigates the utility of vagueness from an empirical perspective. The present manuscript seeks to make a contribution to empirical work on the utility of vagueness.

Reviews of the original submission all mentioned that the contrast we established between vague and crisp terms was also consistent with a difference between `comparison' and `matching' verification procedures (or selection algorithms). To spell this out, and explain how we addressed this concern, it will be necessary to describe our experimental setup briefly. Participants were presented with a set of dot arrays, and required to identify one set in response to an instruction of the form: ``Choose a square with $n$ dots" where $n$ was either a numerical quantifier (e.g., 5) or a verbal quantifier (e.g., few).  The concern of the reviewers was that an instruction like ``Choose a square with 5 dots", which we had cast as the crisp condition, invited a `matching' task where the participant had to match the quantity in the instruction to a quantity in the stimulus, whereas an instruction like ``Choose a square with few dots", which we had cast as the vague condition, invited a `comparison' task where the participant had to compare the dot arrays with each other to satisfy the instruction. In other words our crisp / vague dimension was confounded with the matching / comparison verification method.

One reviewer focused on another potential confound - that the difference which we had described as crisp / vague was also a difference between numerals in the crisp cases and verbal quantifiers in the vague cases.  This reviewer suggested that numerals invited counting where verbal quantifiers invited estimation, and that this difference could explain what we had described as effects of vagueness.  We have added material to the Introduction addressing these confounds, as suggested by reviewers.

Reviewers suggested that we should carry out further experiments to address these confounds. We did this in experiments 4 and 5 of the current submission.  Experiment 4 used numerals in each condition - we devised crisp and vague forms of the instructions, and comparison and matching conditions for each of the crisp and vague forms. In Experiment 5 we did the same for verbal quantifiers. In this way we crossed vagueness with verification method at each level of the numeral / verbal quantifier dimension.

The findings of the new experiments 4 and 5 are that neither the verification method nor the numerical / verbal format of the instruction can fully explain the effects of vagueness that we identified: at each level of the numerical / verbal format dimension vagueness exerted benefits in the comparison conditions but exerted disadvantages in the matching conditions.

In closing we would like to raise the possibility of inviting the original reviewers to review the current submission, in order to capitalise on the work already done in examining a previous version of the manuscript, but of course we understand that this is the editor's decision.

\closing{Sincerely,\\
\includegraphics[width=3cm]{mysignature.pdf}
}
Matthew Green\\
Kees van Deemter
\end{letter}

\end{document}

### 2016-07-27 papers/2015/resubmission_drafts/2014_resubmission_draft/K-lcn-aug-2014.tex
adding papers

% changing to a version that only reports experiments 2 to 5
\documentclass[doc,apacite]{apa6}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{color}
\usepackage{booktabs}
\usepackage{layout}
\usepackage{apacite}
\usepackage{datetime}

%\usepackage[draft]{changes}
\usepackage[final]{changes}
\setdeletedmarkup{\textcolor{red}{\sout{#1}}}
%\usepackage[right=2 in]{geometry}
\setlength{\marginparwidth}{72pt}

\title{Vagueness as cost reduction}
\date{\today}
\authornote{\today ~\currenttime}
\shorttitle{Vagueness as cost reduction}
\twoauthors{Matt Green}{Kees van Deemter}
\twoaffiliations{Computing Science\\ University of Aberdeen}{Computing Science\\ University of Aberdeen}

% KvD: Throughout the paper I've tried to use "instruction format" consistently, instead of sometimes "instruction format", sometimes "quantifier format".

\abstract{Much of everyday language is vague, yet standard game-theoretic models make it difficult to see how vague expressions can have benefits over crisp ones in co-operative situations. We report a series of experiments that aims to separate the utility of vagueness from the utility of other factors that come together with vagueness. We argue that the results support a view of vagueness where the benefits that vague terms exert are due to other influences that vagueness brings with it rather than to \deleted{influences of} vagueness itself.}

\begin{document}
\maketitle               
\section{Introduction}             
Vagueness pervades the language that we use on a daily basis. In everyday use, language use may be called vague for various reasons.\footnote{See e.g. the entry ``vague'' in \cite{penguin}.} In most academic use though, the word `vagueness' has a more particular meaning.  Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1996vagueness} (a similar definition can be found in \citeA{EgreKlinedinst}, among many others).  The most crucial of these criteria is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects. By contrast, a word is vague if it is not precise'' \cite[p.\ 1]{lipmanvague}. A typical example is the word ``tall'', as applied to people for example, because here is no precise, known height which separates those who are tall from those who are not. The crucial point is that ``tall'' admits borderline cases (i.e., people who may or may not count as tall), which are the hallmark of vagueness as we use the term.

Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions, which are used so frequently \cite{Lipman:2000fk, lipmanvague}. By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss'' \cite<>[p.~1]{lipmanvague}. Lipman explains the point by means of a scenario in which a speaker describes a person to a hearer, who needs to identify that person in the arrivals hall of of an airport. Lipman argues that, in such a scenario, a precise description of the person's height (e.g., ``The person's height is 187.96 cm'') would be more useful than a vague one (``The person is tall''). Lipman uses this scenario to explain why standard game theory models of communication \cite<e.g.,>{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act \deleted{of communication} that communicates the same state of affairs. The relevant conditions are, broadly speaking, that both interlocutors know all the relevant facts (e.g., both know the person's height precisely) and that the setting for the communication is co-operative. These conditions exclude deliberate deception, and rhetorical situations like political debate and advertising where the intention is to persuade the interlocutor to adopt some point of view, where the persuasion might not be intended to be in the addressee's best interest.

Lipman \replaced{argued}{observed} that such the efficiency loss resulting from vague expressions would be unlikely to have arisen unless there are advantages as well as disadvantages associated with vague expressions. Lipman asked, essentially, what these advantages might be, and how they might find a place in a game-theoretical explanation. In this paper, we focus on the first part of Lipman's question.

Several tentative answers to Lipman's question have been offered  \cite<see>{van2009utility, van2010vagueness}. Prominent among these answers is the idea that vague expressions are somehow easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite<e.g.,>[]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}. For example, \citeA<>[p.\ 11]{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze''. We shall refer to this characterisation of the utility of vague language as the \emph{cost reduction} hypothesis. The idea of vagueness as cost reduction can take various shapes~\deleted{\protect\cite<>[]{van2009utility}}, but the basic hypothesis is that it is easier for people to think in terms of loosely defined categories (such as ``quite a few'', or ``many'') than in terms of crisply defined ones (such as ``thirteen'', or ``237''). This predicts that whenever a vague expression can perform the same communicative task as a precise one, it is rational to choose the vague expression. The corollary for comprehension is that vague expressions should be understood more readily than precise expressions.

Questions concerning optimal language use have many practical applications. Natural Language Generation\footnote{NLG systems take data or formulas as input, and transform them into natural language outputs \cite<>[]{reiter2000building}. The process parallels language production in humans.}  (henceforth NLG) systems must make decisions between different formulations of the same information. For example, if a man's height is 6 foot 2 inches, this could be expressed as ``187.96 metres'', ``6 foot 2'', or ``tall'', among other ways, and the NLG system must decide between these.  The problem is particularly relevant for NLG systems that take numbers as input, as many do. In the context of an NLG system faced with a practical decision of this kind, Lipman's question becomes ``Under what circumstances should vague terms be produced?''  Relevant applications include weather forecasting on the basis of numerical weather data such as temperature and wind speed \cite{goldberg1994using,turner2006generating}, and medical decision support on the basis of clinical measurement such as oxygen saturation, heart rhythm, etc. \cite{Hripcsak01032009, hunter2008summarising, portet2009automatic}. At present, such NLG systems \replaced{are often forced to make}{make} decisions concerning the level of precision in the utterances that they generate (e.g., ``the temperature will be in the high twenties tomorrow'') on the basis of little more than intuition. A better understanding of the benefits of different precision levels for readers would allow these systems to become more useful.

The cost reduction hypothesis is of direct relevance to psycholinguists interested in language comprehension, and additionally to psycholinguists interested in language production, for example in connection with the question of audience design \cite{Clark1982287}. For, to the extent that speakers and writers choose vague expressions over and above crisp ones because the former are easier to process for hearers than the latter, the cost reduction hypothesis suggests that speakers design their utterance for optimal benefit to their hearers -- out of altruism, so to speak.

The utility of vagueness is the attested aim of a small number of studies, but most of these have focussed on vagueness in a different sense, and focussing on different types of benefits for hearers. Two recent studies can illustrate both issues.

In a \added{series of} studies of behaviour modification, \citeA<>{Mishra01042011} manipulated the presentation format of information about quantities \added{in the domains of mental acuity, physical strength, and weight loss}. \added{In the weight loss study, participants were told that the study was designed to test the validity of a new (actually fictitious) health index, the HHI (Holistic Health Index). They were told that an ideal HHI score lies in the range of 45 to 55. In a longitudinal study, participants submitted their height, weight, hydration level, gender, and age to a computer each week. Participants were told that two algorithms would be used to compute their HHI, and that it was possible that the two algorithms might give different values initially, but would converge over the course of the study to a single value. They were also told that if the two algorithms did give different values, then the true score lay between the two values. In one condition, which the authors called the precise condition, the two algorithms gave the same score. In the other condition, which the authors called the vague condition, one algorithm added 3\% to the score while the other algorithm subtracted 3\% from the score, yielding a range of values whose midpoint was the same as the two values given in the precise condition.

One group of participants was given HHI scores in the ideal range: for this group their weight loss did not differ depending on whether they were given vague or precise HHI values. However for the other group, who were given HHI scores outside the ideal range, their weight loss was significantly greater if they were given vague HHI scores than if they were given precise HHI scores. The authors explain the improvement in the vague condition for this group as resulting from the participants' freedom to think of themselves a positioned on one end of the range - the end closest to the ideal HHI scores.} 
%
This ``illusion of proximity'' \cite<>[p.~4]{Mishra01042011} to the goal is argued to allow participants to generate positive expectancies that lead to behaviours that improve performance. In contrast, in the precise conditions, participants did not have this freedom of interpretation, and could not distort the information to bring about the beneficial \emph{illusion of proximity}. These results are interesting, and of obvious potential practical importance. We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \cite{keefe1996vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases. In the terminology of \citeA{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}. Furthermore, the experiments of \citeA{Mishra01042011} did not explore benefits in terms of processing cost, but in terms of long-term behaviour change.

Similar issues arise from the work of \citeA{peters2009bringing}. The authors carried out a series of studies where participants were required to rate hospitals based on various sources of information about quality of care. There was a between-subjects manipulation based on numeracy. The format of the information was manipulated within subjects: either numbers only were presented, or both numbers and evaluative categories were presented (e.g., \emph{Poor}, \emph{Fair}, \emph{Good}, \emph{Excellent}, with \added{crisp} visual boundary lines between the categories). Results showed that, for low-numeracy participants, the presence of evaluative categories resulted in a diminished influence of an irrelevant affective state on the ratings. For all participants, the presence of evaluative categories resulted in better decisions and in a greater use of the most important and reliable types of information, such as survival rates. 
%Participants' decisions were evaluated with respect to a gold-standard set of interpretations derived from experts. 
% KvD The last sentence is worth checking.

It is, however, questionable whether the ``evaluative categories" manipulation in this study can be considered a manipulation of vagueness. Certainly, terms like \emph{Fair} admit the possibility of borderline cases. However, \replaced{given that the boundaries between the categories were marked crisply, and that therefore the categories mapped crisply to numerical values} {when the visual boundary lines are taken into account, which map the terms to exact ranges}, it becomes doubtful whether any borderline cases could be conceived to arise in fact. For example, \emph{Fair} was mapped to 60\% -- 70\% for the variable \emph{percentage of heart attack patients given recommended treatment (ACE inhibitor)}. {\deleted\footnote{This distinction between the potential for vagueness, and the realisation of vagueness in the context of a particular stimulus, is something we will return to when discussing our own experiments.}} Accordingly, rather than the vagueness of categories such as \emph{Poor}, Peters et al. emphasise the evaluative content inherent in these categories, and \replaced{the affective potential of the evaluative content rather than the vagueness of the terms like {\em Fair}}{their affective potential}.\\[2ex]
%
The experiments reported in the present paper put the cost reduction hypothesis to the test. The question that we are trying to answer is whether vague expressions are processed more easily by readers than crisp ones. Like Lipman, we focus on situations where numerical information is used in order to identify a referent. Reference, in other words, will be the linguistic task on which we focus, partly because of the interest that this topic has recently drawn from the NLG community.  In focussing on benefits for the hearer, we will leave aside the question of audience design, leaving this for later research.

\added{In using references to quantities to test the cost reduction hypothesis we are only testing one aspect of vagueness in a particular context. This limits the applicability of our results. It also has the advantage that we can explore the costs and benefits of vagueness more thoroughly in that context. Since one prevalent view of vagueness is that a vague expression is never preferable to a crisp equivalent, a demonstration of a benefit for vagueness in any context would advance the discussion.}

\added{In our experiments we present participants with an instruction to select a referent from among alternatives, using vague and crisp referring expressions to identify the referent. In this context, speed and accuracy of selecting the referent are the key dimensions on which benefits or costs for vagueness should be apparent. We used dot arrays as the possible referents of the instructions, and we varied how the instructions identified a particular dot array, using vague and crisp versions of the instructions. For example, the participant could be instructed `Choose the square with many dots' in the vague condition, or `Choose the square with 20 dots' in the crisp condition.}

\deleted{Game theory and the cost reduction hypothesis both make claims about vagueness in terms of its effects on people engaged in communicative linguistic acts. For game theory, when vagueness influences communication, the term \emph{utility} is used to capture this influence: vagueness is said to have less utility than crispness. The cost reduction hypothesis notes that we use vague language frequently, and assumes that the reason for the high frequency of use is that vagueness brings about an \emph{advantage}, for producer or comprehender. }

\deleted{In this paper, we set out to measure these effects so that we might have some empirical basis for preferring one account over the other. An obstacle to this effort is that whereas we can obtain, in the laboratory setting, several quantitative measures of the effects of language, neither the game theory account nor the cost reduction account specifies a metric in which \emph{utility} or \emph{advantage} should be measured. Therefore it is for us as experimenters to suggest a way to capture and measure the effects of vague and crisp expressions such that they can meaningfully be compared. This requires us to choose, and motivate the choice of, both a task that brings about some measurable behaviour (or alternatively, some measurable physiological state) that is consequent upon whether the language used to bring it about was vague or crisp; and also a measure of the behaviour that is sufficiently fine-grained to distinguish the vague case from the crisp case.}

\deleted{
A good candidate for the task comes from the \emph{forced choice} paradigm in which a participant is required to make a choice quickly among alternatives, indicating the choice by a physical response such as a button-press. These alternatives can be presented using vague and crisp language in such a way that the participant's choice is at least partly dependent upon whether the presentation was vague or crisp: indeed, in the case where other factors are held constant, it is plausible, and often assumed, that differences in the response are consequent to a great extent upon the mode of presentation. The button-press behaviour offers two metrics which are commonly taken to index cognitive load: response time and response accuracy, where fast accurate responses indicate low cognitive load, and slow or inaccurate responses indicate high cognitive load. If vague presentations lead to lower cognitive load than crisp presentations, or vice versa, (where load is indexed by response time and accuracy), it is plausible to argue that this constitutes an advantage, or greater utility, of one presentation mode over the other.}

\deleted{
In the experiments that we present here, we induced people to make a particular choice among alternatives by indicating one alternative using a referring expression.  In the particular situation that we used, participants were presented with a number of objects on screen that constituted the alternatives.  These objects were squares containing a number of dots, where the number in each was different. The referring expression referred to one square by indicating the number of dots it contained, and the participant was required to indicate that square by pressing the appropriate button on the keyboard. The referring expression offered a way to manipulate whether the square was indicated with a crisp or a vague expression of quantity, while holding other factors relatively constant. For example, the participant could be instructed ``Choose the square with many dots'' in the vague condition, or ``Choose the square with 20 dots'' in the crisp condition.}

% Inserting introduction material that mentions the other factors that go along with vagueness in this setup. This is in response to reviewer comments to mention the confounds earlier - MG

In practice manipulating vagueness independently of other factors in these forced choice tasks with numeric expressions is very difficult.
%
\deleted{For example when comparing few/many against small numbers like 2,3,4 (subitizable numbers) it is expected that the specialised routines that the brain has for processing small numbers will come into play: these could overwhelm the influence of vagueness. Experiment One addresses the role of vagueness for subitizable and not-subitizable numbers.} 
%
For example when comparing few/many against \deleted{non-subitizable} numbers like 10, 20, although few/many are vague and 10,20 are crisp, they also differ according to numerical and verbal format. This difference between numerical and verbal format could explain any differences that are found between the vague and crisp conditions for these instructions. \deleted{Experiment Two demonstrates that there are differences for these instructions that can be interpreted in line with the cost reduction hypothesis, while acknowledging that numerical / verbal format could be the underlying driver of the effect.} It is possible to create numerical and verbal versions of both vague and crisp instructions in order to test whether numerical/verbal format drives \replaced{any differences observed between}{the processing advantage for} few/many versus 10/20. \deleted{demonstrated in Experiment Two.} \added{For example, the numerical crisp conditions (10, 20) have numerical vague versions (about 10, about 20); and the verbal vague conditions (few, many) have verbal crisp versions (the fewest, the most).}

\deleted{
Experiment Three takes the existing numerical crisp (10/20) and verbal vague (few, many) conditions and adds instructions in  numerical vague (about 10 / about 20), and verbal crisp (the fewest / the most) expressions. We increase the number of squares to three for this experiment, so that expressions like few and many have a borderline case.
}

Another non-vagueness  difference between few/many and 10/20 is that the process of identifying the square (the selection algorithm for making the decision) might differ. For example, in the vague condition, identifying the square with few dots might be done by a comparison algorithm, identifying that one square is less numerous than the others without establishing the numerosity of either square; whereas, in the crisp condition, establishing that one of the squares has about 10 dots would require an estimate of the numerosity of the squares (a matching algorithm, matching the number mentioned in the instruction to estimates of the numerosity of the squares). Thus if participants respond more quickly when instructed to choose the square with few dots, this could be due to vagueness or to comparison being easier than matching.
%
%\deleted{There is some evidence that the size of the array can influence the selection algorithm used to identify the referent of an instruction. \protect\citeA{Feigenson2004307} discusss two systems for representing number. They write that ``large arrays [\ldots] activate a system for representing sets and comparing their approximate cardinal values'' whereas``small arrays [\ldots] activate a system for representing and tracking numerically distinct individuals, which allows for computations [\ldots] of the number of individuals in the array'' \protect\cite[p.~311]{Feigenson2004307}.}
% KvDTentatively deleted this.

\added{ Experiments 4 and 5 take up this issue and the numeric/verbal format issue, by comparing comparison and matching versions of both vague and crisp instructions, when number-use is held constant: numbers are always mentioned in Experiment 4, while numbers are avoided throughout Experiment 5.
% KvD: I've tentatively reinstated this, because I think we cannot simply state the problem without announcing our approach to it. If you know a better way to do this then that's fine. 
}

%\section{\deleted{Experiment 1}}
% \deleted{
%In our first experiment we set out to compare responses to vague and crisp instructions in a forced choice task using response time and accuracy to measure the responses, and a variety of combinations of numbers. We aimed to identify cases where a difference was elicited such that these cases could form the focus of  subsequent experiments.
%}
%\subsection{\deleted{Method}}
%
%\subsubsection{\deleted{Participants}}
%\deleted{
%Twenty-five students and staff from the University of Aberdeen participated in the study in return for a cash payment of five pounds. Their median age was 23, ranging from 18 to 40. All participants self-reported fluency in English, and had normal (or corrected to normal) vision. Participants were recruited by advertising for participants on a university message board and a university mailing list, offering five pounds to volunteers. 
%}
%
%\subsubsection{\deleted{Apparatus}}\deleted{
%A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants. Stimuli were created and presented using the language GNU Octave \protect\cite{eaton:2002} and the Psychophysics Toolbox extensions \protect \cite{ptbx1, ptbx2}.
%}
%
%\subsubsection{\deleted{Design}}
%\deleted{
%We presented participants with 64 trials in experiment one in a single block. The basic properties of a trial were that it consisted of two elements: (1) two squares each containing a diffferent number of dots, and (2) a referring expression that referred to one of the squares by indicating the number of dots that it contained, leaving the other as a distractor. We measured two dependent variables: response time, and response accuracy. It was always the case that two squares appeared on the screen, each containing a number of dots: therefore items can be described using the notation $pair_{(p=1... 8)}([n],m)$ to indicate that for the stimulus pair $p$, one of the numbers of dots was $n$ and the other $m$; and that the square-bracketed element $[n]$ was the target. There were 8 such pairs : $pair_{(1)}(2,4)$; $pair_{(2)}(3,5)$; $pair_{(3)}(2,6)$; $pair_{(4)}(3,7)$; $pair_{(5)}(6,8)$; $pair_{(6)}(7,9)$; $pair_{(7)}(4,8)$; $pair_{(8)}(5,9)$. See Table \ref{tablee1} for the properties of the stimuli.
%}
%
%
%\begin{table}[tbp]
%\caption{\deleted{Properties of stimuli used in Experiment 1, arranged by pair}}
%\label{tablee1}
%\begin{tabular}{lll}
%\deleted{subitizabiity}	& \deleted{gap size}  	&	\deleted{contents}\\
%\hline
%\deleted{subitizable}	&\deleted{small gap}	&	\deleted{$pair_{(1)}(2,4)$}\\
% 			&			&	\deleted{$pair_{(2)}(3,5)$}\\
%\hline
%			&\deleted{large gap}	&	\deleted{$pair_{(3)}(2,6)$}\\
%			&			&	\deleted{$pair_{(4)}(3,7)$}\\
%\hline
%\deleted{not subitizable}	&\deleted{small gap}	&	\deleted{$pair_{(5)}(6,8)$}\\
%			&			&	\deleted{$pair_{(6)}(7,9)$}\\
%\hline
%			&\deleted{large gap}	&	\deleted{$pair_{(7)}(4,8)$}\\
%			&			&	\deleted{$pair_{(8)}(5,9)$}\\
%\hline
%\end{tabular}
%\end{table}
%
%
%
%\begin{table}[htbp]
%\caption{\deleted{Table of instructions for the pair (2,4). Experiment One}}
%\label{instructionse1}
%\begin{tabular}{rl}
%\toprule
%\deleted{vagueness}&\deleted{example}\\
%\midrule
%\deleted{crisp} 	& 	\deleted{Choose the square with two dots }\\
%\deleted{vague}	&	\deleted{Choose the square with few dots}\\
%\bottomrule
%\end{tabular}
%\end{table}
%
%
%
%\deleted{We manipulated four independent variables, and explicitly controlled one other, as follows. }
%
%\deleted{
%The first independent variable was whether the item contained a subitizable number of dots (we use \emph{subitizable} to include 2 and 3: 4 and above are considered not subitizable). This was manipulated between items, and divided the items into two sets: one set comprised items that did contain a subitizable number; the other set comprised items that did not contain a subitizable number. There is evidence \protect\cite{trick1994small} that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities. Four of our pairs contained such a subitizable number, and four did not. This allowed us to ask whether vagueness or crispness might exert differential effects depending on whether this subitizing mechanism was, or was not, involved. The four pairs containing a subitizable number were: $pair_{(1)}(2,4)$; $pair_{(2)}(3,5)$; $pair_{(3)}(2,6)$; $pair_{(4)}(3,7)$; and the four pairs not containing a subitizable number were: $pair_{(5)}(6,8)$; $pair_{(6)}(7,9)$; $pair_{(7)}(4,8)$; $pair_{(8)}(5,9)$.
%}
%
%
%\deleted{The second independent variable was whether the gap between the smaller and the larger number was small or large. At each level of the subitizable variable, half the items had a small gap, and the other half had a large gap. }
%\deleted{There is evidence that when the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the numerosity of two sets of dots \protect\cite{van123} and for processing Arabic numerals and number words \protect \cite{Dehaene199647}.} 
%\deleted{By systematically manipulating this difference, we were able to ask whether effects of vagueness or crispness might differ according to whether the pair was a more or less discriminable pair. }
%
%\deleted{
%The third independent variable was vagueness with two levels, vague and crisp: the effects of vague and crisp language were to be elicited by the way in which an instruction referred to an expression of quantity. Each stimulus was presented equally often with a  vague expression of quantity, and a crisp expression of quantity. Taking the item $pair_{(1)}([2],4)$ as an example, where the target was the square with 2 dots, the crisp referring expression was ``Choose the square with two dots" and the vague referring expression was ``Choose the square with few dots".}
%
%\deleted{Because we presented each item with the smaller number as target, and with the larger number as target, we were able to ask whether responses differed as a function of whether the target was large or small (with respect to the distractor). This constituted our fourth independent variable.}
%
%\deleted{We also had to control which side the smaller number appeared on. This was manipulated within items such that each item had a version where the smaller number was on the left and a version where the smaller number was on the right. There is evidence  that when two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \protect\cite{dehaene1993mental, gevers2006automatic}. Therefore we needed to present $pair_{(1)}(2,4)$ once as $pair_{(1)}([2],4)$ and once as $pair_{(1)}(4,[2])$. This accounts for two of the four presentations. The other two presentations were $pair_{(1)}(2,[4])$ and $pair_{(1)}([4],2)$: i.e., with the larger number [4] rather than the smaller number as the target square; again once on the left and once on the right. This accounts for the eight presentations of each $pair_{(p)}(n,m)$.}
%
%\subsubsection{\deleted{Stimuli}}
%\deleted{Each stimulus consisted of two elements: (1) two squares containing dots; (2) the text of the referring expression indicating the target. Presentation of the two elements was simultaneous. The manipulation of vagueness was effected in the instruction, which referred to the number of dots in the target square as either (a) a cardinal numeral in the crisp conditions (i.e., \emph{two, three, four, five, six, seven, eight, nine}), or (b) a linguistic quantifier in the vague conditions (e.g., \emph{few, many}) . The position of the dots in each square was randomised per trial. An example stimulus is given in Fig. (\protect\ref{stime1}).}
%
%\begin{figure}[tbp]
%\centering
%\includegraphics[draft,width=.5\textwidth]{images/stimuluse1}
%\caption{\deleted{Example stimulus from experiment 1. Stimulus consists of two elements (1) the two squares with a number of dots, one of which is the referent of the instruction; (2) the text of the instruction that identifies the target with a referring expression.}}
%\label{stime1}
%\end{figure}
%
%\subsubsection{\deleted{Procedure}}
%\deleted{The experiment was conducted in a small quiet room. On arrival in the room, the participant was told that he or she would be presented with objects on screen, together with an instruction to choose one of them by pressing the appropriate key on the keyboard (which was left cloverleaf for a target on the left, and right cloverleaf for a target on the right). The participant was instructed to respond quickly while avoiding errors.
%There were 4 practice trials. After the practice trials the participant was invited to ask any questions they had about the procedure, and then the experimenter left the cubicle before the experimental blocks began.
%The order in which trials were presented was randomised per participant. Each trial would time out after 60 seconds if the participant did not respond. No feedback was given on correct trials, but there was feedback on error trials in the form of the word ``\textsc{wrong!!}" which flashed on screen. Response time was measured from the appearance of the stimulus on screen until the button-press indicating the participant's choice. The trial sequence was keypress $\rightarrow$ instruction and stimulus $\rightarrow$ keypress.}
%
%\subsection{\deleted{Results}}
%
%\deleted{For the response time analysis 40 erroneous trials were removed, and then the data were trimmed at 2.5 s.d. separately for each participant, removing another 60 cases. In all, 6.3\% of the data were removed.}
%
%\deleted{A linear mixed effects regression model was built of the log response time data. The model had vagueness and subitizability as the fixed effects of main interest, and control terms for quantity and gap size, with random slopes for vagueness over items and for the vagueness x subitizability interaction over subjects.}
%
%\deleted{The main effect of subitizability was significant: subitizable stimuli resulted in faster responses than non-subitizable stimuli ($\beta=-.303, se=.062, t=-4.867, p=.002$). The main effect of vagueness was not significant ($\beta=.017, se=.021, t=.809, p=.435$). However vagueness did exert an interaction effect with subitizability ($\beta=.172, se=.037, t=4.640, p<.01$), such that vagueness increased response times when the analysis was restricted to only subitizable stimuli ($\beta=.103, se=.035, t=2.930, p<.05$). but reduced response times when the analysis was restricted to non-subitizable stimuli ($\beta=-.066, se=.023, t=-2.826, p<.05$). }
%
%\deleted{Error rates were low at 2.5\% overall. A generalized logit mixed model \protect\cite{jaeger2008categorical} was fit to the error data. Fixed effects were vagueness, subitizability and their interaction, with quantity and gap size as control terms. The random effects structure was random slopes for vagueness and random intercepts for item (models with more complex structure did not converge). A large gap led to fewer errors ($\beta=-.976, se=.443, z=-2.201, p<.05$). There was a reliable vagueness x subitizability interaction ($\beta=2.559, se=.880, z=2.906, p <.01$) with the same shape as that for response times (see Fig \protect\ref{fige1}).}
%
%\begin{figure}
%\centering
%\includegraphics[draft,width=.8\textwidth]{images/resultse1}
%\caption{\deleted{Results for experiment one}\label{fige1}}
%\end{figure}
%
%\subsection{\deleted{Discussion}}
%\deleted{As expected, response times were faster for stimuli that contained a subitizable number of dots. Our main question was whether vagueness would interact with this effect. Response times did not differ reliably according to whether the quantifier was vague or crisp but vagueness did affect response times differently for subitizable and non-subitizable stimuli. When the stimulus contained a subitizable number of dots vagueness slowed responses and led to more errors. When neither square contained a subitizable number of dots, vagueness speeded response times and led to fewer errors.}
%
%% added Feigenson ref - MG
%
%\deleted{A possible explanation of this pattern of results is that when subitizable numbers are present, vagueness hinders the fast automatic system that deals with small numbers, whereas when these specialised routines are not available because there is no subitizable number, vagueness helps participants to identify the right square by allowing them to use estimates of cardinality instead of counting. }
%\deleted{This explanation has some parallels with the two systems for representing number discussed in \protect\citeA{Feigenson2004307}. They write that ``large arrays [\ldots] activate a system for representing sets and comparing their approximate cardinal values'' whereas``small arrays [\ldots] activate a system for representing and tracking numerically distinct individuals, which allows for computations [\ldots] of the number of individuals in the array'' \protect\cite[p.~311]{Feigenson2004307}.}

\section{Experiment Two}

\subsection{Introduction}
We used a forced choice task to compare choices made in response to vague instructions against choices made in response to crisp instructions. 
The participant was presented with an instruction like ``Choose the square with many dots in the vague conditions", or ``Choose the square with 20 dots in the crisp conditions."
Then two dot arrays were presented in the form of squares containing a number of dots.
The participant was required to identify the square that corresponded with the instruction, by pressing the appropriate key.
Response time and accuracy were recorded for analysis.
\added{Fig \ref{stimuluse2} shows an example stimulus.}

Our main manipulation was of the vagueness of the instruction, with two levels, vague and crisp. Table \ref{instructionse2} shows examples from each condition. 
We also manipulated how discriminable the dot arrays were. One array always contained 25 dots: the other varied between 5, 10, 15, 20, 30, 35, 40, 45 dots. This led to numerical differences of 5, 10, 15, and 20, with lower differences representing less discriminable arrays and larger differences representing more discriminable arrays. There is evidence that when the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the numerosity of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene199647}.
Where a number was mentioned in the instruction, it was always in the form of an arabic numeral.

The instructions indicated the larger of the two dot arrays equally often as they indicated the smaller of the dot arrays, so that participants could not systematically choose the larger or smaller array as a response strategy.
There is evidence that when two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}. We controlled which side the smaller number appeared on to avoid systematic influences of this effect.

\begin{table}[htbp]
\caption{Table of instructions for the pair (5,25). Experiment Two}
\label{instructionse2}
\begin{tabular}{rl}
\toprule
vagueness&example\\
\midrule
crisp 	& 	Choose the square with 5 dots \\
vague	&	Choose the square with few dots\\
\bottomrule
\end{tabular}
\end{table}

\subsection{Hypotheses}
%We hypothesised that under these conditions: (1) RT will be faster for vague than precise instructions; (2) Vagueness will be more advantageous at smaller gap sizes (i.e., low discriminability) than at large gap sizes (i.e, high discriminability).

The {\bf cost reduction} hypothesis predicts that responses will be faster and more accurate for the vague conditions than for the crisp condition because the vague conditions impose a lower cognitive load than the crisp conditions.  The {\bf instruction format} hypothesis predicts that, for non-subitisable numbers, responses will be faster and more accurate in the verbal conditions than in the numerical ones because numbers are harder to process than verbal references to quantities. There is abundant evidence (e.g., \protect\cite{trick1994small}) that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities. We performed a pilot experiment in which we were able to confirm this finding in the experimental settings on which we are focussing in this paper \cite{our-earlier-paper}. In particular, we found that, when participants were confronted with an array of two boxes, instructions of the form ``Choose the square with $n$ dots" led to consistently lower Response Times than instructions of the form ``Choose the square with many/few dots" when $2 \leq n \leq 5$. The converse was true for $n>5$. % KvD: Can you check this addition about subitisable numbers, Matt? I'm not sure this stuff has found its right place in the paper yet. Where it is at the moment, it seems to break the story. Is it better after all to make this Experiment 1 (announced as a replication)?
The {\bf selection algorithm} account predicts that responses will be faster and more accurate in the comparison conditions than in the matching conditions because it is easier to carry out comparison than matching. Thus all three accounts predict a main effect advantage of vagueness, but for different reasons.

Regardless of which of the above hypotheses one adopts, one would predict that responses will be faster and more accurate when the numerical distance between the dot arrays is greater than when it is smaller because greater distances result in more easily discriminable arrays. Thus all three accounts predict a main effect of numerical distance, or gap size. For the same reason, one would predict that any advantage for the vague instructions should be smaller than when the arrays are least discriminable. This constitutes a prediction for an interaction between vagueness and numerical distance. % I've combined the old points (2) and (3) and removed the numbering. Maybe you can polish this further, finding a better place for these observations?

% KvD I've tentatively deleted the following. I think it's probably superfluous.
%The present experiment will demonstrate an advantage for our vague conditions, but not that the advantage is due to the vagueness of the instruction, because the vague conditions were also (a) verbal not numeric and (b) answerable with a comparison strategy not a matching strategy.  Our subsequent experiments will discriminate between the alternative accounts.

% KvD I'm tentatively starting with a general explanation of the methodology of this group of experiments.

{\bf General.} All our experiments shared essentially the same methodology: Participants aged between ..., self-reported fluent speakers of English, were recruited by .... and were paid 10 ponds for participating. All had normal, or corrected-to-normal vision. A MacBook ... Stimuli were created ... In all except the last experiment (Experiment 5), items were presented as follows: (...) (IF WE DO ADD THIS GENERAL DESCRIPTION THEN REMOVE THE RELEVANT INFO FROM THE DESCRIPTION OF EACH OF THE INDIVIDUAL EXPERIMENTS.)

\subsection{Method}
%\subsubsection{Participants}
Twenty participants were recruited by mailing list and paid ten pounds for participating. Participants were aged between 18 and 45, with a median age of 26. All participants self-reported fluency in English, and had normal, or corrected-to-normal vision.

%\subsubsection{Apparatus}
A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants. Stimuli were created and presented using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.

%\subsubsection{\deleted{Design}}
\deleted{We presented participants with 256 trials, arranged in 4 blocks of 64 trials each. }

\deleted{It was always the case that two squares appeared on screen. One of these always contained 25 dots, and the other contained a number of dots that varied, and which was always the target. }

\deleted{We can use the notation $pair_{(p)}(n,25)$ to represent a pair. The pairs were: $pair_{(1)}(5,25)$; $pair_{(2)}(10,25)$; $pair_{(3)}(15,25)$; $pair_{(4)}(20,25)$; $pair_{(5)}(30,25)$; $pair_{(6)}(35,25)$; $pair_{(7)}(40,25)$; $pair_{(8)}(45,25)$.}
%
\deleted{We manipulated two independent variables: gap size, with four levels (gap size = 5, 10, 15, or 20), and vagueness, with two levels (crisp and vague). }
\deleted{We had to control another two variables: target size (whether the referring expression indicated the larger number of dots or the smaller number of dots); and target side (for a given level of target size, whether the target appeared on the left or the right). }
\deleted{We measured, as our dependent variables, response time and response accuracy.}
%
\deleted{Our manipulation on gap size divided the items into four groups, one for each gap size. Items with gap size 5 were: $pair_{(4)}(20,25)$; and $pair_{(5)}(30,25)$; items with gap size 10 were: $pair_{(3)}(15,25)$; and $pair_{(6)}(35,25)$; items with gap size 15 were $pair_{(2)}(10,25)$ and $pair_{(7)}(40,25)$; and items with gap size 20 were $pair_{(1)}(5,25)$ and $pair_{(8)}(45,25)$.}
%
\deleted{Our manipulation on vagueness meant generating vague and crisp versions of each referring expression. Each pair was presented 8 times with a vague expression of quantity, and 8 times with a crisp expression of quantity. Taking $pair_{(1)}(5,25)$ as an example, the crisp referring expression was ``Choose the square with 5 dots'', and the vague referring expression was ``Choose the square with few dots''.
%
We needed to control which side the target was presented on. Each pair was presented equally often with with the target on the left, as with the target on the right.
%
We also needed to control whether the target was the smaller or larger number. The items with numbers less than 25 (pairs 1 to 4) formed a group with the smaller number as target and the other items (pairs 5 to 8) formed a balancing group with the larger number as the target number.}

%\subsubsection{\deleted{Stimuli}}

\deleted{Each stimulus consisted of two elements (a) the text of the referring expression indicating the target; (b) two squares containing dots. }
\deleted{Where a number was mentioned in the instruction, it was always in the form of an arabic numeral (e.g., 1,2). }
\deleted{This contrasts with experiment one, where numbers were given in natural language form.}
\deleted{The position of the dots was randomised per-trial.}
\deleted{Fig (\ref{stimuluse2}) gives an example stimulus.}

\begin{figure}[tbp]
\fitfigure{images/stimuluse2}
\caption{An example stimulus from Experiment Two. First the referring expression was presented (left panel). Then after a keypress, and a fixation cross (not pictured) the squares and dots were presented without repetition of the referring expression (right panel)}
\label{stimuluse2}
\end{figure}

%\subsubsection{Procedure}

The experiment was conducted in a quiet room. On arrival in the room, the participant was told that he or she would be presented with an instruction to choose one of two squares by reference to how many dots it contained. Participants were required to press the space key after reading the instruction. Then there was a central fixation cross for 1000 ms, and a blank screen for 500 ms, followed by the squares and dots (without repetition of the referring expression).  The position of the dots was randomised per-trial. Response time was measured as the latency between the presentation of the dots and squares, and the keypress identifying the decision; in this way, the Response Times can be separated from readings times, which is important since we are only interested in the latter. % KvD OK?

The display would stay on screen until the participant responded (there was no timing-out). Participants were asked to respond quickly while avoiding errors. There were 8 practice trials, after which the participant was invited to ask any questions about procedure. After answering these the experimenter left the cubicle for the duration of the experiment. The order in which trials were presented was randomised per-participant. There were 256 trials, presented in 4 blocks of 64 trials each, between which the participant could rest. No feedback was given on correct trials, but there was feedback on error trials in the form of the word ``\textsc{wrong!!}'' which flashed on screen.

\subsection{Results}

A response was counted as erroneous if the square with the wrong number of dots was chosen (when the instruction contained a number); if the square with the larger number of dots was selected (when the instruction was ``Choose the square with few dots"); or if the square with the smaller number of dots was selected (when the instruction was ``Choose the square with many dots".) % KvD Is this addition defining "erroneous" OK?
RTs for trials with erroneous responses were discarded, leading to the loss of 354 trials from 5120, representing 6.9\% of the trials. The correct response RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of 160 trials, 3.4\% of the correct responses. Means for response times and error rates are given in Fig. (\ref{resultse2}).
A linear mixed model of RT was built using as independent variables vagueness and gap size and their interaction, with random slopes for vagueness and gap size over participants. Vagueness was sum coded: vague $= -.5$, crisp $= .5$; gap size was Helmert coded. Helmert contrasts compare each level against the mean of the previous levels. Level one of this contrast is gap size 5 compared with gap size 10; level two is the mean of gap size 5 and gap size 10 versus gap size 15; and level three is the mean of gap sizes 5, 10 and 15 versus gap size 20. 
$p$ values were calculated using the R package \emph{lmerTest} \cite{lmerTest}.

RTs were faster for vague instructions ($\beta=.109, se=.022, t=4.9, p<.001$). 
%
RT grew faster as gap size increased: level one $(\beta=-.116, se=.012, t=-9.3, p<.05)$, level 2 ($\beta=-.103, se=.009, t=-11.1, p<.001$) and level three ($\beta=-.082, se=.007, t=-11.0, p<.05$). Since discriminability of the dot arrays is easier for larger gap sizes, discriminability probably underlies this effect. Gap size and vagueness interacted significantly for larger gap sizes when modelling RT. % KvD Is it worth saying how they interacted?
The interactions at the different levels of gap size were: level one: ($\beta=.001, se=.016, t=.03, p=.974$); level two ($\beta=-.039, se=.009, t=-4.460, p<.001$); level three ($\beta=-.043, se=.006,t=-7.0, p<.001$). In the crisp conditions RTs started out much slower than in the vague conditions, at the smallest gap size, but the two conditions converged to very fast times at the largest gap size. There were diminishing returns for vagueness as gap size increased.

\label{accann}
Error rate data were analysed using a generalized logit mixed model \cite{jaeger2008categorical}, with vagueness and gap size and their interaction as independent variables, and with random slopes for vagueness and gap size over participants. 
%
The effect of vagueness on error rates approached significance, with the vague instructions leading to fewer errors ($\beta=.307, se=.173, t=1.8, p=.077$).
%
Error rates decreased as gap size increased: level one ($\beta=-.585, se=.092, ,z=-6.4, p<.001$), level 2 ($\beta=-.434, z=-5.3,p<.001$) and level three ($\beta=-.250,z=-4.1, p<.001$). 
%
Error rates were greater in the crisp conditions than the vague conditions when gap size was small, and this difference diminished with increasing gap size until it reversed at the biggest gap size.
\begin{figure}[htbp]
\fitfigure{images/plotMeansBoth.pdf}
\caption{Experiment Two results. }
\label{resultse2}
\end{figure}

\subsection{Discussion}

The experiment provided support for hypothesis (1), that responses would be faster and more accurate for the vague instruction conditions than for the crisp instruction conditions. 
We also found that responses grew faster and more accurate as the numerical distance between the dot arrays grew larger, in line with hypothesis (2).
The experiment also provided support for hypothesis (3) that vagueness would be more advantageous at smaller than at larger gap sizes.  
The findings of Experiment Two might be summarised thus: vague instructions are easier to process than crisp instructions, and returns for vagueness diminish as the stimuli become more easily discriminable.

The cost reduction hypothesis explains the vagueness advantage by claiming that the vague referring expressions place less cognitive load on the comprehender than the crisp referring expressions. It explains the diminishing returns for vagueness in more-discriminable stimuli by claiming that load is low in both conditions for the easily-discriminable stimuli, and that therefore there is no extra benefit to be had from vagueness in the easily-discriminable stimuli.

The instruction format hypothesis explains the main effect advantage for the vague instructions by observing that the vague instructions used verbal quantifiers whereas the crisp conditions used numerical quantifiers. Under this account it is avoiding numbers that explains the vagueness advantage main effect. The main effect of numerical distance is explained by assuming that larger distances result in more easily discriminable arrays. The vagueness x numerical distance interaction can be explained by assuming that the numerical quantifiers make the task particularly challenging when the stimuli are less discriminable.

The selection algorithm account explains the main effect advantage of the vague conditions as due to the vague conditions allowing a comparison strategy rather than a matching strategy in the crisp conditions. The main effect of numerical distance is explained by claiming that comparison is easier for more discriminable arrays. The diminishing returns for vagueness as numerical distance grows are explained by claiming that the advantage of being able to use comparison is greater for less discriminable arrays than for more discriminable arrays.

%A possible explanation for the pattern of results is cast in terms of the precision of an estimate of quantity that is required to carry out the task. For example, in the display pair$_{(1)}(5, 25)$, with a large gap size, vague instructions (\emph{many}, \emph{few}) might be able to  be carried out by a visual comparison that identifies one square as more numerous than the other. Precise instructions (to identify the square with 5 dots) might be able to be carried accurately out with an estimate with a fairly large error of 19. In the display pair$_{(8)}(20, 25)$, which has a small gap size, vague instructions might be able to be be carried out with a visual comparison that identifies one square as more numerous than the other. Precise instructions might require an estimate with a maximum error of 4 in this case. The pattern of results can be explained if we assume that the degree to which an estimate with error 4 is harder than an estimate with error 19 is greater than the degree to which a visual comparison between 20 and 25 is harder than a visual comparison between 5 and 25.

%\item An account that we will dub the \emph{instruction format hypothesis}, or the number / no-number account, explains the vagueness advantage by placing importance on the observation that the vague conditions did not mention a numeral whereas the crisp conditions did. This could make the difference between a task that taps numerical processing, and one that taps magnitude judgements. The diminishing returns for vagueness can be explained by assuming that the numerical task is particularly challenging when the stimuli are less-discriminable, whereas the magnitude judgement is relatively less difficult in the less-discriminable trials.

%\item \citeA{Rayna:1989lr} can account for the pattern in a framework they called the \emph{hierarchy of gist}. They proposed that reasoners encode representations at different levels of precision, and that these representations can be ordered with respect to precision, forming a hierarchy of gist. They further claimed that reasoning operated at the lowest level in this hierarchy that would allow one to accomplish the assigned task. Among the levels of this hierarchy are  (1) ratio representations at the top, which encode numerical information exactly; (2) ordinal representations that capture relative magnitude but do not encode numerical differences, i.e., they establish a rank ordering of quantities by magnitude, e.g., \emph{large}, \emph{larger}, \emph{largest}; (3) nominal or categorical representations that capture only the presence of absence of quantity e.g., \emph{some lives}; \emph{no lives}. \citeA{brainerd1994development} presented evidence that children spontaneously encode relative numerosity, at the ordinal level of the hierarchy mentioned above, when they are presented with stimuli that vary in numerosity. The hierarchy of gist offers an explanation of our vagueness advantage in experiment two that is independent of vagueness in the \citeA{keefe1996vagueness} sense. In experiment two, the participants were presented with the instruction for the trial first, and then with the squares and dots. This was done in order to try to separate the processes involved in processing the instruction from those involved in processing the stimuli. A consequence of this serial presentation though is that participants can infer from the instruction which level of representation is the minimum level necessary to carry out the impending task successfully. We do not suggest that this takes the form of a conscious deliberation -- rather it might be apprehended spontaneously in the same way that the children in \citeA{brainerd1994development} did without conscious deliberation. For example, it follows naturally from the instruction \emph{Please choose the square with fewest dots} that an ordinal representation will be necessary and sufficient for the purposes of the current trial: similarly it follows naturally from the instruction \emph{Please choose the square with 34 dots} that a higher level of the hierarchy will be necessary and sufficient for the purposes of this trial - the ratio level that encodes numerosity exactly. We found a response time advantage in experiment two for instructions that used the quantifiers \emph{many} and \emph{few}, when compared with times for quantifiers like \emph{15} and \emph{45}. In the terms of \citeA{Rayna:1989lr}, this can be explained as an advantage for the sufficiency of a lower level of representation -- without reference to vagueness.
%
%\item The \emph{two-systems account} of decision making \cite<e.g.,>{sloman2007two} can offer an explanation of the pattern. Dual process models of decision making propose two systems of processing: the quick and affective System 1 and the deliberative and rule-based System 2. This two-systems account can explain our finding of diminishing returns for vagueness in experiment 2. When the gap size is small, participants use System 2 (slow, deliberative) for the precise instructions. When the gap size is large, they take a shortcut for the crisp conditions and merely establish which square is more (or less) numerous, using System 1 (quick, heuristic). In the vague conditions, they use the heuristic system whether the gap size is small or large. This account explains the diminishing returns for vagueness that we observed in terms of a change of strategy in the baseline precise conditions as the gap size grew larger.
%\end{seriate}

%The potential for vagueness in the vague conditions can be said to have been unrealised in Experiment Two. The vague instructions were  \emph{choose the square with many dots} and \emph{choose the square with few dots}. \emph{Many} and \emph{few} have the potential for vagueness defined as the existence of borderline cases. Given a square with 15 dots, and asked the question `are there many dots in the square?', we can imagine that some people would be unsure (contrast this with the questions `are there fewer than 20 dots in the square?' to which there is no room for uncertainty). However, this potential for vagueness may be unrealised in the context of a choice between two squares. Given two squares, one with 20 dots, and another with 25 dots, and asked to `choose the square with few dots', it seems that no one could be unsure which square is the intended target. If there is no room for uncertainty, then the potential for \emph{many} and \emph{few} to be vague can be said to be unrealised.

\section{Experiment Three}

\subsection{Introduction}

The main result from experiment two was that responses were faster and more accurate for vague instructions than for crisp instructions. This finding can be interpreted in line with three different hypotheses.%: the cost reduction hypothesis, the instruction format account, and the selection algorithm account. 
In experiment three we set out to distinguish between the cost reduction hypothesis and the instruction format account, leaving the selection algorithm account for experiments four and five. Contrast an expression from the vague condition: `the square with few dots' with an expression from the crisp condition: `the square with 15 dots'. One difference is that `few' is vague (or at least has the potential for vagueness) and `15' is crisp. Another difference is that `few' is a linguistic, or verbal quantifier while `15' is a numerical quantifier, in the sense that a number is mentioned explicitly. Since these two differences were confounded in Experiment Two, the vagueness advantage finding is vulnerable to an alternative interpretation, that what we saw as a vagueness advantage was in contrast an advantage for the linguistic or verbal form of the quantifier. In the present experiment three we pitted these alternative interpretations against each other in a factual design.

In Experiment Two, the participants chose one of two squares. The `vague' quantifiers (e.g., `few') uniquely identified one square. Recall our definition of vague  -- ``a word is precise if it describes a well-defined set of objects. By contrast, a word is vague if it is not precise''.  In the first experiment, the quantifiers in the vague conditions did not really meet this definition. This is because there were no borderline cases of the referent that could make the referent set `not well-defined'. Experiment three used three squares so that the vague quantifiers always had more than one possible referent. To enhance the potential for the vague quantifiers to have true vagueness in Experiment Three, we also used indefinite articles in the vague instructions. Thus, instead of the instruction \emph{Choose the square with few dots}, we used \emph{Choose a square with few dots}.

In experiment three, an item was a referring expression followed by a trio of numbers, representing the number of squares in the left, middle, and right squares. We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54). Each triple had the following properties: it comprised three squares (instead of two as in Experiment Two); the central number was always presented in the middle of the three; there were two flanking numbers where one was smaller than the central number and one was bigger.

There was a numerical and a verbal version of each of the vague and crisp referring expressions. See Table \ref{instructionseB} for examples.
The vague numerical condition's referring expression was ``Choose a square with about 10 dots''. None of the squares displayed contained 10 dots. %10 is slightly closer to 6 than to 15. Therefore the best referent for this referring expression was the square with 6 dots;  the borderline response was the square with 15 dots; and the poorest referent was the square with 24 dots. 
% KvD: tentatively removed the above.
%
% KvD: In what follows, I've made a number of fairly trivial replacements and reorderings. Please check!
The {\em crisp numerical} conditions's referring expression was ``Choose the square with 6 dots'' on half the presentations in that condition and ``Choose the square with 24 dots'' on the other half. One square did contain the exact number mentioned.  In the {\em crisp verbal} condition, we used the referring expression ``Choose a square with fewer than $20$ (more than $10$) dots".  The {\em vague numerical} condition's referring expression was ``Choose the square with far fewer than $20$ dots'' on half of presentations, making the squares with 6 and $15$ dots possible referents; iin the other case, the referring expression was ``Choose the square with far more than $10$ dots'', which made $15$ and $24$ possible referents.The {\em vague verbal} condition's referring expression was ``Choose a square with few dots'' on half of the presentations, and ``Choose a square with many dots'' on the other half. For this condition, the best referent was the square with 6 dots for `few' and 24 dots for `many'; the borderline case was the square with 15 dots; and the remaining square was the poorest referent. 
% KvD: OK, Matt? This needs checking once again.
%
% An indication that the manipulation of vagueness was successful is that participants chose the borderline case square on 16\% of trials.
% KvD This last sentence needs to be reformulated, talking about {6,15} both.

\begin{table}[htbp]
\caption{Table of instructions arranged by condition for the triple (6,15,24). Experiment Three}
\label{instructionseB}
\begin{tabular}{rlll}
\toprule
vagueness&instruction format&selection task&example\\
\midrule
crisp 	& 	numeric	& matching	&Choose the square with 6 dots \\
vague	&	numeric 	& matching	&Choose a square with about 10 dots\\
crisp		&	verbal	& comparison	&Choose the square with the fewest dots\\
vague	&	verbal	& comparison	&Choose a square with few dots\\
\bottomrule
\end{tabular}
\end{table}

\subsection{Hypotheses}
The instruction format account predicts a main effect of instruction format such that the numeric instruction conditions will attract longer response times, but has no prediction of any vagueness effects or interaction effects with vagueness.

The selection task account predicts the same main effect but treats it as an effect of the task mandated by the instruction, with the matching task conditions predicted to take longer than the comparison task conditions.

The cost reduction account predicts that there will be a main effect of vagueness such that the vague instruction conditions attract faster responses than the crisp instruction conditions. The cost reduction account also predicts an interaction with the instruction format (or selection task) variable: at each level of instruction format (or selection task) the vague condition should attract faster responses than the crisp condition.

\subsection{Method}

Thirty participants were recruited and paid in the same way as Experiment Two. They were aged between 18 and 45 with a median age of 28. All participants self-reported fluency in English, and had normal, or corrected-to-normal vision.

The same apparatus was used as for Experiment Two.

We presented participants with 256 trials, arranged in 4 blocks each with 64 trials. Each triple was presented in each condition 16 times. 8 of these identified the larger number, and 8 the smaller. We manipulated two independent variables: vagueness with two levels (vague and crisp); and instruction format with two levels (numeric, verbal). In this experiment the selection task variable mapped onto the instruction format variable: the numeric instruction conditions both mandated a matching selection task and the verbal instruction conditions both mandated a comparison task. This yielded four conditions (vague numeric; vague verbal; crisp numeric; crisp verbal). Each condition had a different referring expression, as follows, using the triple $(6,15,24)$ as an example: Choose a square with about 10 dots; Choose a square with few dots; Choose the square with 6 dots; Choose the square with the fewest dots. We measured two dependent variables: response time; and the probability of a participant choosing the borderline case.

Fig (\ref{stimuluse3}) gives an example stimulus. First, the referring expression that constituted the instruction for that trial was displayed. The participant then pressed a key to indicate that he or she had read the instruction. After 1 second,  the squares and dots were presented, while preserving the text of the referring expression. The position of the dots in the squares was randomised per-trial.

\begin{figure}[tbp]
\centering
\includegraphics[width=.4\textwidth]{images/stimuluse3}
\caption{An example stimulus from Experiment Three}
\label{stimuluse3}
\end{figure}

The experiment was conducted in a small quiet room.
On arrival in the cubicle, the participant was told that he or she would be presented with objects on screen and required to choose one in response to an instruction on screen, by pressing the button corresponding with the object. There were 5 practice trials. After the practice trials the experimenter left the room. There were 4 blocks of 64 trials each. In between blocks the participant had the opportunity to rest before continuing. The response time dependent variable was measured from the presentation of the squares and dots, until the keypress indicating the participant's choice. The trial would timeout after 60 seconds if there was no response. The dependent variable measuring whether the participant chose the borderline case was also recorded at this time. In this experiment, no feedback was given. This was because, in the vague conditions, we did not regard any response as `correct' or `incorrect', but instead as `borderline response', or `not borderline response', and we did not want to draw participants' attention to this distinction explicitly. We simply recorded whether the participant chose the borderline case or not, and how long it took the participant to respond.

\subsection{Results}

This time around, no responses were treated as erroneous, because errors were essentially undefined for the vague instructions (e.g., `about ten'). It was noted whether participants chose the borderline square. 
%
The RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data. Means for response times and error rates are given in Fig. (\ref{resultse3}).

A linear mixed model was constructed for the response times, with selection task (or instruction format), vagueness, and item and their interactions as independent variables, and random slopes over participant for task and vagueness (and their interaction) and for item. Task and vagueness were sum coded and item was centred. There was a significant effect of task with numerical conditions attracting longer responses than the verbal conditions ($\beta=.37, se=.07, t=5.1, p<.001$). The effect of vagueness was to slow responses down  ($\beta=.058, se=.013, t=4.8, p<.001$). There was an interaction effect between vagueness and task, such that the disadvantage for vagueness was greater in the numerical than in the verbal instruction conditions.  There were differences in response time as a function of item ($\beta=.06, se=.008, t=7.0, p<.001$) but there was no clear trend across items.

Participant grand mean percentage of borderline selections was $16.6\%$. A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with task, vagueness and item as fixed effects, and with random slopes for task and vagueness and item over participants.  Participants were significantly more likely to choose the borderline option for vague instructions than for precise instructions (21.9\% vs 11.3\%, $\beta=.79, se=.25, z=3.2, p<.01$). Participants were significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%, $\beta=3.57, se=.26, z=13.6, p<.001$).

\begin{figure}[htbp]
\centering
\includegraphics[width=.65\textwidth]{images/resultse3.pdf}
\caption{Results for Experiment Three}
\label{resultse3}
\end{figure}

\subsection{Discussion}

This experiment tested to see whether when borderline cases are present, vague instructions would speed responses as they did in Experiment Two when there were no borderline squares. We actually found a \emph{disadvantage} of vague instructions: vague instructions slowed people down by 112 ms on average. We also found that the effect of instruction format was significant, with numerical format slowing responses by 689 ms on average. The disadvantage of numerical format overwhelms the contribution of vagueness. The verbal vague condition was still responded to faster than the numerical crisp condition, so the pattern from Experiment Two is reproduced, but in the light of the evidence from Experiment Three in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like either an advantage of verbal instruction format, or alternatively as an advantage of the comparison task, according to the selection task account.

% KvD: I've tentatively added the following:
Having effectively separated the {\bf cost reduction} hypothesis from the {\bf instruction format} hypothesis, it is important to observe that, in Experiment 3, instruction format went hand in hand with {\bf selection algorithm}: as shown by the Table of instructions \ref{instructionseB}, the instructions that used a verbal instruction format allowed a comparison algorithm, whereas the instructions that used a numeric format allowed on a matching algorithm. Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison algorithm (which is known to be faster than matching). % KvD I assume we need to say more about this, citing some appropriate papers to support this idea.

In the next two experiments we pitted the comparison and matching selection tasks against each other while controlling vagueness and instruction format. In Experiment Four we would restrict all the instructions to numeric quantifiers while factorially manipulating vagueness and the comparison / matching selection tasks. In Experiment Five we would ensure that all instructions used verbal quantifiers, while also factorially manipulating vagueness and the comparison / matching selection tasks. This allows us to distinguish between the predictions of the selection task account and the instruction format account.

\section{Experiment Four}

\subsection{Introduction}

The main aim of experiment four was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies. The main changes from experiment three were that the selection task was explicitly controlled, and that all conditions were constrained to mention a number. We used the same stimuli as in experiment three. Table \ref{Instructions for e4} shows the instructions. % KvD I think we should motivate our choice of materials (especially "far ..er")

\begin{table}[htbp]
\centering
\caption{Experiment Four: Instructions, assuming $6,15,24$ dots as the Item, and showing \emph{fewer} instead of \emph{more}}
\label{Instructions for e4}
\begin{tabular}{llll}
\hline
vagueness&instruction format&selection task&instruction\\
\hline
crisp & numeric&matching & Choose a square with 6 dots \\ 
crisp & numeric&comparison & Choose a square with fewer than 20 dots \\
vague & numeric&matching & Choose a square with about 10 dots \\ 
vague & numeric&comparison & Choose a square with far fewer than 20 dots \\ 
\hline
\end{tabular}
\end{table}%

\subsection{Hypotheses}
% (1) Vague instructions are easier for the reader than crisp alternatives (main effect of vagueness)
%(2) Comparison is easier for the reader than matching (main effect of selection)
%(3) Effects of vagueness are different depending on whether selection is matching or comparison (interaction effect selection x vagueness).

The instruction format account predicts no differences between the conditions, since all conditions used numeric quantifiers.

The selection task account predicts a main effect of selection task such that the comparison conditions would attract faster responses than the matching conditions.

The cost reduction account predicts that there will be a main effect of vagueness such that the vague instruction conditions attract faster responses than the crisp instruction conditions and particularly that at each level of selection task the vague condition should attract faster responses than the crisp condition.

\subsection{Method}

38 volunteers were recruited via internal messaging at University of Aberdeen, with self-reported fluency in English. They were paid ten pounds each for participating.
The apparatus was the same as Experiment Two.
The design was a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e4}).
Each stimulus was an instruction followed by a triple of dots.
First a referring expression instruction was presented. Participants pressed a key to dismiss the instruction and proceed to the squares with dots in them.

\subsection{Results}
A linear mixed effects regression model was built for log response times. The structure of the model was as follows: fixed effects were vagueness, selection task (both sum coded) and centred item and their interactions: random effects were vagueness, selection task, and item (but not their interactions - the model failed to converge when these interactions were included). The means are plotted in Figure \ref{resultse4}.

The results showed that vagueness was beneficial for comparison but detrimental for matching. There was no significant main effect of vagueness ($\beta =.003, se=.014, t=.202, p=.841$). There was a main effect of task type, with the comparison task speeding responses compared to the matching task ($\beta=-.165, se=.027, t=-6.218, p<.001$). Vagueness exerted effects in different directions for the comparison task and for the matching task. Separate analyses were conducted at each level of the selection task to see whether within each task type there were significant effects of vagueness. There were: in the comparison task vagueness significantly speeded response times compared with crisp controls ($\beta=-.07, se=.02, t=3.52, p<.01$). In the matching task vagueness significantly slowed response times compared with crisp controls ($\beta=-.07, se=.02, t=-2.89, p<.05$).

None of the accounts set out in the Hypotheses section emerge well from the results. The instruction format account wrongly predicts no differences between the conditions. The selection task correctly predicted the main effect of selection task, but has no coverage of the interaction with vagueness. The cost reduction account is wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.

\begin{figure}[htbp]
\includegraphics[width=.5\textwidth]{images/resultse4.pdf}
\caption{Results for Experiment Four}
\label{resultse4}
\end{figure}

\section{Experiment Five}
\subsection{Introduction}
This experiment investigated response times for instructions that did not use a number. 
We manipulated vagueness and the selection task (comparison and matching). 
In order to implement the experiment without mentioning numbers,  we used a prime to visually show the numbers of dots that we wanted to refer to as {\em the target} in the instructions. 
\added{This presentation of a prime before the main trial shares some features with Experiment 2 in \protect \citeA{Izard20081221}, although in that experiment participants were told the numerosity of the prime - called an \emph{inducer} in that paper - in our experiment we did not tell participants the numerosity of the prime array.}
An item was thus a combination of a visual prime, a numeric triple, and a referring expression.
The referring expressions were constrained to never mention a numeral, as in Table \ref{Instructions for e5}.

\begin{table*}[htbp]
\centering
\caption{Experiment Five: Instructions}
\label{Instructions for e5}
\begin{tabular}{lllp{7cm}}
\hline
vagueness&instruction format& selection task&instruction\\
\hline
crisp & verbal&matching & Choose a square with the same number of dots as the target \\ 
crisp & verbal&comparison& Choose a square with fewer dots than the target \\
vague & verbal&matching & Choose a square with about the same number of dots as the target \\ 
vague & verbal&compaison& Choose a square with far fewer dots than the target \\ 
\hline
\end{tabular}
\end{table*}%

\subsection{Hypotheses}
The instruction format account predicts no differences between the conditions, since all conditions used verbal quantifiers.

The selection task account predicts a main effect of selection task such that the comparison conditions would attract faster responses than the matching conditions.

The cost reduction account predicts that there will be a main effect of vagueness such that the vague instruction conditions attract faster responses than the crisp instruction conditions and particularly that at each level of selection task the vague condition should attract faster responses than the crisp condition.

\subsection{Method}

40 volunteers recruited via internal messaging at University of Aberdeen, with self-reported fluency in English. They were paid ten pounds for participating. 
The apparatus was the same as Experiment Two.
The design was a 2 x 2 factorial manipulation of vagueness and selection task.
Each stimulus was a sequence of prime, instruction, and squares.
The procedure for this experiment was different from the others, to accommodate the requirement not to use numbers in the instruction. We had to have a different way to indicate a numerosity in the instruction, which we did by adding a visual `prime', a square that contained the number of dots that we wanted to refer to.

\subsection{Results}
The results showed that vagueness was beneficial for comparison but detrimental for matching (the same as Experiment Four) even when no numbers were allowed in the instructions. Figure \ref{resultse5} shows the means by condition. There was no main effect of vagueness ($\beta=.01, se=.01, t=1.51, p=.14$). There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=-.18, se=.02, t=-10.38, p<.01$). This effect was in the same direction as Experiment Four. Vagueness did  exert different effects depending on the selection task ($\beta=.12, se=.03, t=4.32, p<.05$). Separate analyses were conducted for the comparison task and for the matching task. In the comparison task, vagueness resulted in faster response times ($\beta=-0.08, se=.02, t=4.30, p<.05$). In the matching task vagueness slowed response times ($\beta=.05, se=.01, t=3.72, p<.05$). These results are in the same direction as Experiment Four.

\begin{figure}[htbp]
\includegraphics[width=.5\textwidth]{images/resultse5.pdf}
\caption{Results for Experiment Five}
\label{resultse5}
\end{figure}

Again none of the accounts set out in the Hypotheses section emerge well from the results. The instruction format account wrongly predicts no differences between the conditions. The selection task correctly predicted the main effect of selection task, but has no coverage of the interaction with vagueness. The cost reduction account is wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.

\subsection{Discussion of experiments 4 and 5}

The main aim of these two experiments was to test whether vagueness confers any cognitive benefits over and above those due to differences in the selection task according to whether the instruction mandates a \emph{comparison} selection task or a \emph{matching} selection task, when number-use is held constant.  The main effect of selection task showed that the assumption that the \emph{comparison} task is easier than the \emph{matching} task is well-founded. In both experiments people were reliably faster at responding in the \emph{comparison} task.

Vagueness, which was the phenomenon  on which our investigation focussed, did not exert a main effect in response time. However when the comparison and selection tasks were analysed separately, there was small reliable speedup in RT from the crisp to the vague \emph{comparison} tasks, but a small reliable slowdown in RT from the crisp to the vague \emph{matching} tasks.

\section{General Discussion}

Experiment 2 showed us that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment could not distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format; and mandated a comparison strategy rather than a matching strategy.  Experiment 3 showed us that number-avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional explanatory power in either the verbal format instructions or the numerical format instructions when we generated verbal and numerical versions of both crisp and vague instructions. However the experiment could not distinguish benefits of number-avoidance in the verbal instructions from benefits of the comparison selection task: the verbal instructions also mandated a comparison strategy rather than a matching strategy. In experiments 4 and 5 we manipulated vagueness and the selection task independently of numerical format. We found that there are effects of the selection task mandated by the instruction, with the comparison task instructions attracting faster response times than the matching instructions, and that vagueness exerts benefits when the selection task is \emph{comparison}, but not when the task is \emph{matching}.

The benefits of vagueness in the \emph{comparison} task in experiments 4 and 5 could be explained as differences in the number of valid targets for the expression, as follows. Taking as an example the stimulus with (6,15,24) dots, it could be argued that the vague comparison instruction (e.g., \emph{a square with far fewer than 20 dots}) has one valid target, the square with 6 dots, while the crisp comparison instruction (e.g., \emph{a square with fewer than 20 dots}) has two valid targets, the squares with 6 and 15 dots. In both experiments 4 and 5 we found that people were quicker to identify a square when the instruction only had one valid target. This leads us to speculate that the benefit for vagueness here could be due to the vague expression foregrounding a particular valid target while the crisp expression carries with it the additional task of distinguishing between two alternative valid targets, something we propose to call a ``range-reduction'' benefit.

What is one entitled to conclude? Given that we were able to identify a class of situations -- namely: situations in which a comparison strategy suffices to identify the intended referent -- in which vague expressions led to faster response times than crisp ones, would it be valid to conclude that we have finally discovered an advantage for vagueness that cannot be ascribed to some other factor? We believe the answer to this question is negative. To see why, consider figures \ref{resultse4} and \ref{resultse5}. Both figures depict four conditions, depending on whether the expression was crisp or vague, and depending on whether the referent could be identified using a comparison strategy or not. Two of the resulting four conditions result in an expression that can denote either of two referents; the other two conditions result in an expression that can only denote one referent, with the other possible referent being a marginal candidate at best:

\begin{table}[htbp]
\caption{Vagueness as range reduction}
\begin{center}
\begin{tabular}{lll}
\hline
vagueness & selection task & candidates\\
\hline
crisp & matching & 1 candidate\\
crisp & comparison & 2 candidates\\
vague & matching & 2 candidates\\
vague & comparison & 1 candidate\\
\hline
\end{tabular}
\end{center}
\label{default}
\end{table}

To see why vagueness thus has opposite effects, depending on whether it is used in matching or comparison situations, compare an instruction like `Choose a square with 6 dots' with its vague counterpart `Choose a square with about 10 dots': by adding the word `about', we broaden the range of squares that the expression might be referring to. On the other hand, compare `Choose a square with fewer than 20 dots' with its vague counterpart `Choose a square with far fewer than 20 dots': by adding the word `far', we did not broaden the range of squares denotable by the expression: we narrow it down, because only some of the squares that have fewer dots may have {\em far} fewer dots.

The observation that conditions with 1 candidate lead to shorter response times than conditions with 2 candidates is consistent with the range reduction hypothesis, but not with the idea that vagueness has a beneficial effect. It appears, in other words, that range reduction causes shorter response times, suggesting that shorter response times will only result from a vague expression if this expression leads to range reduction. Once again, it seems, it is not vagueness itself that has advantages but a phenomenon (namely range reduction) that is an automatic concomitant of vagueness in some types of situations.

The findings from our experiments show that when vague expressions are compared with crisp alternatives in our forced choice task, vague expressions appear to yield benefits in some situations, but that the observed benefits may be due to factors other than vagueness itself that the vague forms bring along with them: factors like avoiding numbers; permitting comparison tasks; and range reduction. The picture that is starting to emerge, in other words, is subtle: on the one hand, in the situations that we have been studying -- where cooperative speakers refer to an object (e.g., a square) by means of some quantity associated with the object (e.g., the number of dots contained in the square) -- vagueness is not intrinsically beneficial. On the other hand, vague expressions often have other features that {\em are} beneficial, and these are what give us the incorrect impression that vagueness itself is beneficial. Vagueness may thus have acquired a reputation that it does not deserve.

A comparison may clarify the logic of the situation. In recent years a number of studies, focussing typically on red wine, have suggested that alcohol, consumed in low doses, may have health benefits. An alternative explanation, however, asserts that it is not the alcohol in the wine that is beneficial, but the antioxidants from grapes. If this alternative explanation is correct then alcohol may not be as beneficial as some of us would have it. % KvD Good hey?

\bibliography{lcn-aug-2014}
\end{document}

### 2016-07-27 papers/2015/resubmission_drafts/2015_resubmission_draft/vagueness-dec-2015.tex
adding papers

--- 2016-07-27 papers/2015/resubmission_drafts/2014_resubmission_draft/K-lcn-aug-2014.tex
@@
-% changing to a version that only reports experiments 2 to 5
 \documentclass[doc,apacite]{apa6}
@@
 \usepackage{datetime}
-
-%\usepackage[draft]{changes}
-\usepackage[final]{changes}
-\setdeletedmarkup{\textcolor{red}{\sout{#1}}}
-%\usepackage[right=2 in]{geometry}
-\setlength{\marginparwidth}{72pt}
-
+\usepackage{rotating}
 \title{Vagueness as cost reduction}
@@
 
-% KvD: Throughout the paper I've tried to use "instruction format" consistently, instead of sometimes "instruction format", sometimes "quantifier format".
-
 \abstract{Much of everyday language is vague, yet standard game-theoretic models make it difficult to see how vague expressions can have benefits over crisp ones in co-operative situations.
 We report a series of experiments that aims to separate the utility of vagueness from the utility of other factors that come together with vagueness.
-We argue that the results support a view of vagueness where the benefits that vague terms exert are due to other influences that vagueness brings with it rather than to \deleted{influences of} vagueness itself.}
+We argue that the results support a view of vagueness where the benefits that vague terms exert are due to other influences that vagueness brings with it rather than to vagueness itself.}
 
@@
 Lipman argues that, in such a scenario, a precise description of the person's height (e.g., ``The person's height is 187.96 cm'') would be more useful than a vague one (``The person is tall'').
-Lipman uses this scenario to explain why standard game theory models of communication \cite<e.g.,>{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act \deleted{of communication} that communicates the same state of affairs.
+Lipman uses this scenario to explain why standard game theory models of communication \cite<e.g.,>{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs.
 The relevant conditions are, broadly speaking, that both interlocutors know all the relevant facts (e.g., both know the person's height precisely) and that the setting for the communication is co-operative.
@@
 
-Lipman \replaced{argued}{observed} that such the efficiency loss resulting from vague expressions would be unlikely to have arisen unless there are advantages as well as disadvantages associated with vague expressions.
+Lipman argued that such the efficiency loss resulting from vague expressions would be unlikely to have arisen unless there are advantages as well as disadvantages associated with vague expressions.
 Lipman asked, essentially, what these advantages might be, and how they might find a place in a game-theoretical explanation.
@@
 We shall refer to this characterisation of the utility of vague language as the \emph{cost reduction} hypothesis.
-The idea of vagueness as cost reduction can take various shapes~\deleted{\protect\cite<>[]{van2009utility}}, but the basic hypothesis is that it is easier for people to think in terms of loosely defined categories (such as ``quite a few'', or ``many'') than in terms of crisply defined ones (such as ``thirteen'', or ``237'').
+The idea of vagueness as cost reduction can take various shapes, but the basic hypothesis is that it is easier for people to think in terms of loosely defined categories (such as ``quite a few'', or ``many'') than in terms of crisply defined ones (such as ``thirteen'', or ``237'').
 This predicts that whenever a vague expression can perform the same communicative task as a precise one, it is rational to choose the vague expression.
@@
 \cite{Hripcsak01032009, hunter2008summarising, portet2009automatic}.
-At present, such NLG systems \replaced{are often forced to make}{make} decisions concerning the level of precision in the utterances that they generate (e.g., ``the temperature will be in the high twenties tomorrow'') on the basis of little more than intuition.
+At present, such NLG systems are often forced to make decisions concerning the level of precision in the utterances that they generate (e.g., ``the temperature will be in the high twenties tomorrow'') on the basis of little more than intuition.
 A better understanding of the benefits of different precision levels for readers would allow these systems to become more useful.
@@
 
-In a \added{series of} studies of behaviour modification, \citeA<>{Mishra01042011} manipulated the presentation format of information about quantities \added{in the domains of mental acuity, physical strength, and weight loss}.
-\added{In the weight loss study, participants were told that the study was designed to test the validity of a new (actually fictitious) health index, the HHI (Holistic Health Index).
+In a series of studies of behaviour modification, \citeA<>{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss.
+In the weight loss study, participants were told that the study was designed to test the validity of a new (actually fictitious) health index, the HHI (Holistic Health Index).
 They were told that an ideal HHI score lies in the range of 45 to 55.
@@
 However for the other group, who were given HHI scores outside the ideal range, their weight loss was significantly greater if they were given vague HHI scores than if they were given precise HHI scores.
-The authors explain the improvement in the vague condition for this group as resulting from the participants' freedom to think of themselves a positioned on one end of the range - the end closest to the ideal HHI scores.}
-%
+The authors explain the improvement in the vague condition for this group as resulting from the participants' freedom to think of themselves a positioned on one end of the range - the end closest to the ideal HHI scores.
 This ``illusion of proximity'' \cite<>[p.~4]{Mishra01042011} to the goal is argued to allow participants to generate positive expectancies that lead to behaviours that improve performance.
@@
 There was a between-subjects manipulation based on numeracy.
-The format of the information was manipulated within subjects: either numbers only were presented, or both numbers and evaluative categories were presented (e.g., \emph{Poor}, \emph{Fair}, \emph{Good}, \emph{Excellent}, with \added{crisp} visual boundary lines between the categories).
+The format of the information was manipulated within subjects: either numbers only were presented, or both numbers and evaluative categories were presented (e.g., \emph{Poor}, \emph{Fair}, \emph{Good}, \emph{Excellent}, with crisp visual boundary lines between the categories).
 Results showed that, for low-numeracy participants, the presence of evaluative categories resulted in a diminished influence of an irrelevant affective state on the ratings.
 For all participants, the presence of evaluative categories resulted in better decisions and in a greater use of the most important and reliable types of information, such as survival rates.
-%Participants' decisions were evaluated with respect to a gold-standard set of interpretations derived from experts. 
-% KvD The last sentence is worth checking.
 
@@
 Certainly, terms like \emph{Fair} admit the possibility of borderline cases.
-However, \replaced{given that the boundaries between the categories were marked crisply, and that therefore the categories mapped crisply to numerical values} {when the visual boundary lines are taken into account, which map the terms to exact ranges}, it becomes doubtful whether any borderline cases could be conceived to arise in fact.
+However, given that the boundaries between the categories were marked crisply, and that therefore the categories mapped crisply to numerical values, it becomes doubtful whether any borderline cases could be conceived to arise in fact.
 For example, \emph{Fair} was mapped to 60\% -- 70\% for the variable \emph{percentage of heart attack patients given recommended treatment (ACE inhibitor)}.
-{\deleted\footnote{This distinction between the potential for vagueness, and the realisation of vagueness in the context of a particular stimulus, is something we will return to when discussing our own experiments.}} Accordingly, rather than the vagueness of categories such as \emph{Poor}, Peters et al.
-emphasise the evaluative content inherent in these categories, and \replaced{the affective potential of the evaluative content rather than the vagueness of the terms like {\em Fair}}{their affective potential}.\\[2ex]
-%
+Accordingly, rather than the vagueness of categories such as \emph{Poor}, Peters et al.
+emphasise the evaluative content inherent in these categories, and the affective potential of the evaluative content rather than the vagueness of the terms like {\em Fair}.
+%\\[2ex]
+\subsection{}
 The experiments reported in the present paper put the cost reduction hypothesis to the test.
@@
 
-\added{In using references to quantities to test the cost reduction hypothesis we are only testing one aspect of vagueness in a particular context.
+In using references to quantities to test the cost reduction hypothesis we are only testing one aspect of vagueness in a particular context.
 This limits the applicability of our results.
 It also has the advantage that we can explore the costs and benefits of vagueness more thoroughly in that context.
-Since one prevalent view of vagueness is that a vague expression is never preferable to a crisp equivalent, a demonstration of a benefit for vagueness in any context would advance the discussion.}
+Since one prevalent view of vagueness is that a vague expression is never preferable to a crisp equivalent, a demonstration of a benefit for vagueness in any context would advance the discussion.
 
-\added{In our experiments we present participants with an instruction to select a referent from among alternatives, using vague and crisp referring expressions to identify the referent.
-In this context, speed and accuracy of selecting the referent are the key dimensions on which benefits or costs for vagueness should be apparent.
-We used dot arrays as the possible referents of the instructions, and we varied how the instructions identified a particular dot array, using vague and crisp versions of the instructions.
-For example, the participant could be instructed `Choose the square with many dots' in the vague condition, or `Choose the square with 20 dots' in the crisp condition.}
+In our experiments we used a speeded forced choice task to compare the processing costs of different references to quantities.
+In this context, speed and accuracy of responses are the key dimensions on which the different references can be compared.
+The stimuli in the experiments were sets of dot arrays containing various numbers of dots.
+The forced choice was to identify one dot array given a reference to a given quantity of dots.
+We manipulated the references in several ways across a series of four experiments.
 
-\deleted{Game theory and the cost reduction hypothesis both make claims about vagueness in terms of its effects on people engaged in communicative linguistic acts.
-For game theory, when vagueness influences communication, the term \emph{utility} is used to capture this influence: vagueness is said to have less utility than crispness.
-The cost reduction hypothesis notes that we use vague language frequently, and assumes that the reason for the high frequency of use is that vagueness brings about an \emph{advantage}, for producer or comprehender.
-}
+Our main manipulation was always of vagueness: we constructed crisp and vague versions of references to the same dot array.
+For example, in experiment 1 the instruction presented to the participant could be \emph{Choose the square with many dots.} (vague condition), or \emph{Choose the square with 20 dots.} (crisp condition), identifying the same dot array.
 
-\deleted{In this paper, we set out to measure these effects so that we might have some empirical basis for preferring one account over the other.
-An obstacle to this effort is that whereas we can obtain, in the laboratory setting, several quantitative measures of the effects of language, neither the game theory account nor the cost reduction account specifies a metric in which \emph{utility} or \emph{advantage} should be measured.
-Therefore it is for us as experimenters to suggest a way to capture and measure the effects of vague and crisp expressions such that they can meaningfully be compared.
-This requires us to choose, and motivate the choice of, both a task that brings about some measurable behaviour (or alternatively, some measurable physiological state) that is consequent upon whether the language used to bring it about was vague or crisp; and also a measure of the behaviour that is sufficiently fine-grained to distinguish the vague case from the crisp case.}
+Ideally, we would manipulate vagueness independently of other variables.
+However, in practice, and given the constraints imposed by the use of natural language, any manipulation of vagueness introduces variance along other dimensions as well.
+Therefore we carried out further experiments to try to address and control variance along other dimensions.
 
-\deleted{
-A good candidate for the task comes from the \emph{forced choice} paradigm in which a participant is required to make a choice quickly among alternatives, indicating the choice by a physical response such as a button-press.
-These alternatives can be presented using vague and crisp language in such a way that the participant's choice is at least partly dependent upon whether the presentation was vague or crisp: indeed, in the case where other factors are held constant, it is plausible, and often assumed, that differences in the response are consequent to a great extent upon the mode of presentation.
-The button-press behaviour offers two metrics which are commonly taken to index cognitive load: response time and response accuracy, where fast accurate responses indicate low cognitive load, and slow or inaccurate responses indicate high cognitive load.
-If vague presentations lead to lower cognitive load than crisp presentations, or vice versa, (where load is indexed by response time and accuracy), it is plausible to argue that this constitutes an advantage, or greater utility, of one presentation mode over the other.}
+One non-vagueness source of variance was the difference between numerical and verbal format in the references.
+For example, in the instructions for experiment 1 given above, \emph{many} is in verbal format (as well as vague) and \emph{20} is in numerical format (as well as crisp).
+Therefore any difference observed between responses to the \emph{many} instruction and the \emph{20} instruction could be due either to vagueness or to {\bf instruction format}.
+It is possible to create vague and crisp references in each of these instruction formats: for example in numerical instruction format we can have \emph{20} (crisp) and \emph{about 20} (vague), and in verbal instruction format we can have \emph{the most} (crisp) and \emph{many} (vague).
+Experiment 2 varied vagueness and instruction format factorially in a similar way to this, to try to tease apart these two sources of variance.
 
-\deleted{
-In the experiments that we present here, we induced people to make a particular choice among alternatives by indicating one alternative using a referring expression.
-In the particular situation that we used, participants were presented with a number of objects on screen that constituted the alternatives.
-These objects were squares containing a number of dots, where the number in each was different.
-The referring expression referred to one square by indicating the number of dots it contained, and the participant was required to indicate that square by pressing the appropriate button on the keyboard.
-The referring expression offered a way to manipulate whether the square was indicated with a crisp or a vague expression of quantity, while holding other factors relatively constant.
-For example, the participant could be instructed ``Choose the square with many dots'' in the vague condition, or ``Choose the square with 20 dots'' in the crisp condition.}
+Another potential non-vagueness source of variance between responses to the \emph{many} (verbal vague) references and the \emph{20} (numeric crisp) references lies in the method used to identify a referent (we call this the {\bf selection algorithm} source of variance).
+Identifying the array with \emph{many} dots might be done by a \emph{comparison} algorithm, identifying that one array is more numerous than the others without establishing the numerosity of any arrays, whereas identifying the array with \emph{20} dots, which might be done by a \emph{matching} algorithm would require an estimate of the numerosity of the arrays.
+It seems reasonable that comparison would be faster than matching because it does not require estimates of cardinality.
+Furthermore the problem persists when using references like \emph{the most} (verbal crisp) and \emph{about 20} (numeric vague), since they too vary not only along the vague/crisp dimension and the numeric/verbal dimension but also along the comparison/matching dimension.
+Experiments 3 and 4 addressed the \emph{selection algorithm} source of variance - experiment 3 crossed vague/crisp and comparison/matching factorially using only numerical references, and experiment 4 crossed vague/crisp and comparison/matching factorially using only verbal references.
 
-% Inserting introduction material that mentions the other factors that go along with vagueness in this setup. This is in response to reviewer comments to mention the confounds earlier - MG
-
-In practice manipulating vagueness independently of other factors in these forced choice tasks with numeric expressions is very difficult.
-%
-\deleted{For example when comparing few/many against small numbers like 2,3,4 (subitizable numbers) it is expected that the specialised routines that the brain has for processing small numbers will come into play: these could overwhelm the influence of vagueness.
-Experiment One addresses the role of vagueness for subitizable and not-subitizable numbers.}
-%
-For example when comparing few/many against \deleted{non-subitizable} numbers like 10, 20, although few/many are vague and 10,20 are crisp, they also differ according to numerical and verbal format.
-This difference between numerical and verbal format could explain any differences that are found between the vague and crisp conditions for these instructions.
-\deleted{Experiment Two demonstrates that there are differences for these instructions that can be interpreted in line with the cost reduction hypothesis, while acknowledging that numerical / verbal format could be the underlying driver of the effect.} It is possible to create numerical and verbal versions of both vague and crisp instructions in order to test whether numerical/verbal format drives \replaced{any differences observed between}{the processing advantage for} few/many versus 10/20.
-\deleted{demonstrated in Experiment Two.} \added{For example, the numerical crisp conditions (10, 20) have numerical vague versions (about 10, about 20); and the verbal vague conditions (few, many) have verbal crisp versions (the fewest, the most).}
-
-\deleted{
-Experiment Three takes the existing numerical crisp (10/20) and verbal vague (few, many) conditions and adds instructions in  numerical vague (about 10 / about 20), and verbal crisp (the fewest / the most) expressions.
-We increase the number of squares to three for this experiment, so that expressions like few and many have a borderline case.
-}
-
-Another non-vagueness  difference between few/many and 10/20 is that the process of identifying the square (the selection algorithm for making the decision) might differ.
-For example, in the vague condition, identifying the square with few dots might be done by a comparison algorithm, identifying that one square is less numerous than the others without establishing the numerosity of either square; whereas, in the crisp condition, establishing that one of the squares has about 10 dots would require an estimate of the numerosity of the squares (a matching algorithm, matching the number mentioned in the instruction to estimates of the numerosity of the squares).
-Thus if participants respond more quickly when instructed to choose the square with few dots, this could be due to vagueness or to comparison being easier than matching.
-%
-%\deleted{There is some evidence that the size of the array can influence the selection algorithm used to identify the referent of an instruction.
-\protect\citeA{Feigenson2004307} discusss two systems for representing number.
-They write that ``large arrays [\ldots] activate a system for representing sets and comparing their approximate cardinal values'' whereas``small arrays [\ldots] activate a system for representing and tracking numerically distinct individuals, which allows for computations [\ldots] of the number of individuals in the array'' \protect\cite[p.~311]{Feigenson2004307}.}
-% KvDTentatively deleted this.
-
-\added{ Experiments 4 and 5 take up this issue and the numeric/verbal format issue, by comparing comparison and matching versions of both vague and crisp instructions, when number-use is held constant: numbers are always mentioned in Experiment 4, while numbers are avoided throughout Experiment 5.
-% KvD: I've tentatively reinstated this, because I think we cannot simply state the problem without announcing our approach to it. If you know a better way to do this then that's fine. 
-}
-
-%\section{\deleted{Experiment 1}}
-% \deleted{
-%In our first experiment we set out to compare responses to vague and crisp instructions in a forced choice task using response time and accuracy to measure the responses, and a variety of combinations of numbers.
-We aimed to identify cases where a difference was elicited such that these cases could form the focus of  subsequent experiments.
-%}
-%\subsection{\deleted{Method}}
-%
-%\subsubsection{\deleted{Participants}}
-%\deleted{
-%Twenty-five students and staff from the University of Aberdeen participated in the study in return for a cash payment of five pounds.
-Their median age was 23, ranging from 18 to 40.
-All participants self-reported fluency in English, and had normal (or corrected to normal) vision.
-Participants were recruited by advertising for participants on a university message board and a university mailing list, offering five pounds to volunteers.
-%}
-%
-%\subsubsection{\deleted{Apparatus}}\deleted{
-%A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants.
-Stimuli were created and presented using the language GNU Octave \protect\cite{eaton:2002} and the Psychophysics Toolbox extensions \protect \cite{ptbx1, ptbx2}.
-%}
-%
-%\subsubsection{\deleted{Design}}
-%\deleted{
-%We presented participants with 64 trials in experiment one in a single block.
-The basic properties of a trial were that it consisted of two elements: (1) two squares each containing a diffferent number of dots, and (2) a referring expression that referred to one of the squares by indicating the number of dots that it contained, leaving the other as a distractor.
-We measured two dependent variables: response time, and response accuracy.
-It was always the case that two squares appeared on the screen, each containing a number of dots: therefore items can be described using the notation $pair_{(p=1...
-8)}([n],m)$ to indicate that for the stimulus pair $p$, one of the numbers of dots was $n$ and the other $m$; and that the square-bracketed element $[n]$ was the target.
-There were 8 such pairs : $pair_{(1)}(2,4)$; $pair_{(2)}(3,5)$; $pair_{(3)}(2,6)$; $pair_{(4)}(3,7)$; $pair_{(5)}(6,8)$; $pair_{(6)}(7,9)$; $pair_{(7)}(4,8)$; $pair_{(8)}(5,9)$.
-See Table \ref{tablee1} for the properties of the stimuli.
-%}
-%
-%
-%\begin{table}[tbp]
-%\caption{\deleted{Properties of stimuli used in Experiment 1, arranged by pair}}
-%\label{tablee1}
-%\begin{tabular}{lll}
-%\deleted{subitizabiity}	& \deleted{gap size}  	&	\deleted{contents}\\
-%\hline
-%\deleted{subitizable}	&\deleted{small gap}	&	\deleted{$pair_{(1)}(2,4)$}\\
-% 			&			&	\deleted{$pair_{(2)}(3,5)$}\\
-%\hline
-%			&\deleted{large gap}	&	\deleted{$pair_{(3)}(2,6)$}\\
-%			&			&	\deleted{$pair_{(4)}(3,7)$}\\
-%\hline
-%\deleted{not subitizable}	&\deleted{small gap}	&	\deleted{$pair_{(5)}(6,8)$}\\
-%			&			&	\deleted{$pair_{(6)}(7,9)$}\\
-%\hline
-%			&\deleted{large gap}	&	\deleted{$pair_{(7)}(4,8)$}\\
-%			&			&	\deleted{$pair_{(8)}(5,9)$}\\
-%\hline
-%\end{tabular}
-%\end{table}
-%
-%
-%
-%\begin{table}[htbp]
-%\caption{\deleted{Table of instructions for the pair (2,4). Experiment One}}
-%\label{instructionse1}
-%\begin{tabular}{rl}
-%\toprule
-%\deleted{vagueness}&\deleted{example}\\
-%\midrule
-%\deleted{crisp} 	& 	\deleted{Choose the square with two dots }\\
-%\deleted{vague}	&	\deleted{Choose the square with few dots}\\
-%\bottomrule
-%\end{tabular}
-%\end{table}
-%
-%
-%
-%\deleted{We manipulated four independent variables, and explicitly controlled one other, as follows. }
-%
-%\deleted{
-%The first independent variable was whether the item contained a subitizable number of dots (we use \emph{subitizable} to include 2 and 3: 4 and above are considered not subitizable).
-This was manipulated between items, and divided the items into two sets: one set comprised items that did contain a subitizable number; the other set comprised items that did not contain a subitizable number.
-There is evidence \protect\cite{trick1994small} that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities.
-Four of our pairs contained such a subitizable number, and four did not.
-This allowed us to ask whether vagueness or crispness might exert differential effects depending on whether this subitizing mechanism was, or was not, involved.
-The four pairs containing a subitizable number were: $pair_{(1)}(2,4)$; $pair_{(2)}(3,5)$; $pair_{(3)}(2,6)$; $pair_{(4)}(3,7)$; and the four pairs not containing a subitizable number were: $pair_{(5)}(6,8)$; $pair_{(6)}(7,9)$; $pair_{(7)}(4,8)$; $pair_{(8)}(5,9)$.
-%}
-%
-%
-%\deleted{The second independent variable was whether the gap between the smaller and the larger number was small or large.
-At each level of the subitizable variable, half the items had a small gap, and the other half had a large gap.
-}
-%\deleted{There is evidence that when the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the numerosity of two sets of dots \protect\cite{van123} and for processing Arabic numerals and number words \protect \cite{Dehaene199647}.}
-%\deleted{By systematically manipulating this difference, we were able to ask whether effects of vagueness or crispness might differ according to whether the pair was a more or less discriminable pair.
-}
-%
-%\deleted{
-%The third independent variable was vagueness with two levels, vague and crisp: the effects of vague and crisp language were to be elicited by the way in which an instruction referred to an expression of quantity.
-Each stimulus was presented equally often with a  vague expression of quantity, and a crisp expression of quantity.
-Taking the item $pair_{(1)}([2],4)$ as an example, where the target was the square with 2 dots, the crisp referring expression was ``Choose the square with two dots" and the vague referring expression was ``Choose the square with few dots".}
-%
-%\deleted{Because we presented each item with the smaller number as target, and with the larger number as target, we were able to ask whether responses differed as a function of whether the target was large or small (with respect to the distractor).
-This constituted our fourth independent variable.}
-%
-%\deleted{We also had to control which side the smaller number appeared on.
-This was manipulated within items such that each item had a version where the smaller number was on the left and a version where the smaller number was on the right.
-There is evidence  that when two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \protect\cite{dehaene1993mental, gevers2006automatic}.
-Therefore we needed to present $pair_{(1)}(2,4)$ once as $pair_{(1)}([2],4)$ and once as $pair_{(1)}(4,[2])$.
-This accounts for two of the four presentations.
-The other two presentations were $pair_{(1)}(2,[4])$ and $pair_{(1)}([4],2)$: i.e., with the larger number [4] rather than the smaller number as the target square; again once on the left and once on the right.
-This accounts for the eight presentations of each $pair_{(p)}(n,m)$.}
-%
-%\subsubsection{\deleted{Stimuli}}
-%\deleted{Each stimulus consisted of two elements: (1) two squares containing dots; (2) the text of the referring expression indicating the target.
-Presentation of the two elements was simultaneous.
-The manipulation of vagueness was effected in the instruction, which referred to the number of dots in the target square as either (a) a cardinal numeral in the crisp conditions (i.e., \emph{two, three, four, five, six, seven, eight, nine}), or (b) a linguistic quantifier in the vague conditions (e.g., \emph{few, many}) .
-The position of the dots in each square was randomised per trial.
-An example stimulus is given in Fig.
-(\protect\ref{stime1}).}
-%
-%\begin{figure}[tbp]
-%\centering
-%\includegraphics[draft,width=.5\textwidth]{images/stimuluse1}
-%\caption{\deleted{Example stimulus from experiment 1.
-Stimulus consists of two elements (1) the two squares with a number of dots, one of which is the referent of the instruction; (2) the text of the instruction that identifies the target with a referring expression.}}
-%\label{stime1}
-%\end{figure}
-%
-%\subsubsection{\deleted{Procedure}}
-%\deleted{The experiment was conducted in a small quiet room.
-On arrival in the room, the participant was told that he or she would be presented with objects on screen, together with an instruction to choose one of them by pressing the appropriate key on the keyboard (which was left cloverleaf for a target on the left, and right cloverleaf for a target on the right).
-The participant was instructed to respond quickly while avoiding errors.
-%There were 4 practice trials.
-After the practice trials the participant was invited to ask any questions they had about the procedure, and then the experimenter left the cubicle before the experimental blocks began.
-%The order in which trials were presented was randomised per participant.
-Each trial would time out after 60 seconds if the participant did not respond.
-No feedback was given on correct trials, but there was feedback on error trials in the form of the word ``\textsc{wrong!!}" which flashed on screen.
-Response time was measured from the appearance of the stimulus on screen until the button-press indicating the participant's choice.
-The trial sequence was keypress $\rightarrow$ instruction and stimulus $\rightarrow$ keypress.}
-%
-%\subsection{\deleted{Results}}
-%
-%\deleted{For the response time analysis 40 erroneous trials were removed, and then the data were trimmed at 2.5 s.d.
-separately for each participant, removing another 60 cases.
-In all, 6.3\% of the data were removed.}
-%
-%\deleted{A linear mixed effects regression model was built of the log response time data.
-The model had vagueness and subitizability as the fixed effects of main interest, and control terms for quantity and gap size, with random slopes for vagueness over items and for the vagueness x subitizability interaction over subjects.}
-%
-%\deleted{The main effect of subitizability was significant: subitizable stimuli resulted in faster responses than non-subitizable stimuli ($\beta=-.303, se=.062, t=-4.867, p=.002$).
-The main effect of vagueness was not significant ($\beta=.017, se=.021, t=.809, p=.435$).
-However vagueness did exert an interaction effect with subitizability ($\beta=.172, se=.037, t=4.640, p<.01$), such that vagueness increased response times when the analysis was restricted to only subitizable stimuli ($\beta=.103, se=.035, t=2.930, p<.05$).
-but reduced response times when the analysis was restricted to non-subitizable stimuli ($\beta=-.066, se=.023, t=-2.826, p<.05$).
-}
-%
-%\deleted{Error rates were low at 2.5\% overall.
-A generalized logit mixed model \protect\cite{jaeger2008categorical} was fit to the error data.
-Fixed effects were vagueness, subitizability and their interaction, with quantity and gap size as control terms.
-The random effects structure was random slopes for vagueness and random intercepts for item (models with more complex structure did not converge).
-A large gap led to fewer errors ($\beta=-.976, se=.443, z=-2.201, p<.05$).
-There was a reliable vagueness x subitizability interaction ($\beta=2.559, se=.880, z=2.906, p <.01$) with the same shape as that for response times (see Fig \protect\ref{fige1}).}
-%
-%\begin{figure}
-%\centering
-%\includegraphics[draft,width=.8\textwidth]{images/resultse1}
-%\caption{\deleted{Results for experiment one}\label{fige1}}
-%\end{figure}
-%
-%\subsection{\deleted{Discussion}}
-%\deleted{As expected, response times were faster for stimuli that contained a subitizable number of dots.
-Our main question was whether vagueness would interact with this effect.
-Response times did not differ reliably according to whether the quantifier was vague or crisp but vagueness did affect response times differently for subitizable and non-subitizable stimuli.
-When the stimulus contained a subitizable number of dots vagueness slowed responses and led to more errors.
-When neither square contained a subitizable number of dots, vagueness speeded response times and led to fewer errors.}
-%
-%% added Feigenson ref - MG
-%
-%\deleted{A possible explanation of this pattern of results is that when subitizable numbers are present, vagueness hinders the fast automatic system that deals with small numbers, whereas when these specialised routines are not available because there is no subitizable number, vagueness helps participants to identify the right square by allowing them to use estimates of cardinality instead of counting.
-}
-%\deleted{This explanation has some parallels with the two systems for representing number discussed in \protect\citeA{Feigenson2004307}.
-They write that ``large arrays [\ldots] activate a system for representing sets and comparing their approximate cardinal values'' whereas``small arrays [\ldots] activate a system for representing and tracking numerically distinct individuals, which allows for computations [\ldots] of the number of individuals in the array'' \protect\cite[p.~311]{Feigenson2004307}.}
-
-\section{Experiment Two}
+\section{Experiment One}
 
@@
 We used a forced choice task to compare choices made in response to vague instructions against choices made in response to crisp instructions.
-The participant was presented with an instruction like ``Choose the square with many dots in the vague conditions", or ``Choose the square with 20 dots in the crisp conditions."
+The participant was presented with an instruction like ``Choose the square with many dots'' in the vague conditions, or ``Choose the square with 20 dots'' in the crisp conditions.
 Then two dot arrays were presented in the form of squares containing a number of dots.
+Fig \ref{stimuluse1} shows an example stimulus.
 The participant was required to identify the square that corresponded with the instruction, by pressing the appropriate key.
 Response time and accuracy were recorded for analysis.
-\added{Fig \ref{stimuluse2} shows an example stimulus.}
 
-Our main manipulation was of the vagueness of the instruction, with two levels, vague and crisp. Table \ref{instructionse2} shows examples from each condition. 
+Our main manipulation was of the vagueness of the instruction, with two levels, vague and crisp.
+Table \ref{instructionse1} shows examples from each condition.
 We also manipulated how discriminable the dot arrays were.
-One array always contained 25 dots: the other varied between 5, 10, 15, 20, 30, 35, 40, 45 dots.
-This led to numerical differences of 5, 10, 15, and 20, with lower differences representing less discriminable arrays and larger differences representing more discriminable arrays.
+One array always contained 25 dots: the other contained either 5, 10, 15, 20, 30, 35, 40, or 45 dots.
+This led to numerical differences of 5, 10, 15, and 20, with lower differences resulting in less discriminable arrays and larger differences representing more discriminable arrays.
 There is evidence that when the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the numerosity of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene199647}.
-Where a number was mentioned in the instruction, it was always in the form of an arabic numeral.
+Where a number was mentioned in the instruction, it was always in the form of an Arabic numeral.
 
-The instructions indicated the larger of the two dot arrays equally often as they indicated the smaller of the dot arrays, so that participants could not systematically choose the larger or smaller array as a response strategy.
+The instructions indicated the larger of the two dot arrays equally often as they indicated the smaller of the dot arrays, so that participants could not systematically choose the larger or smaller array as a successful response strategy.
 There is evidence that when two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}.
@@
 \begin{table}[htbp]
-\caption{Table of instructions for the pair (5,25). Experiment Two}
-\label{instructionse2}
+\caption{Table of instructions for the pair (5,25). Experiment 1}
+\label{instructionse1}
 \begin{tabular}{rl}
@@
 \subsection{Hypotheses}
-%We hypothesised that under these conditions: (1) RT will be faster for vague than precise instructions; (2) Vagueness will be more advantageous at smaller gap sizes (i.e., low discriminability) than at large gap sizes (i.e, high discriminability).
+The {\bf cost reduction} hypothesis predicts that responses will be faster and more accurate for the vague conditions than for the crisp conditions because the vague conditions impose a lower cognitive load than the crisp conditions.
 
-The {\bf cost reduction} hypothesis predicts that responses will be faster and more accurate for the vague conditions than for the crisp condition because the vague conditions impose a lower cognitive load than the crisp conditions.
-The {\bf instruction format} hypothesis predicts that, for non-subitisable numbers, responses will be faster and more accurate in the verbal conditions than in the numerical ones because numbers are harder to process than verbal references to quantities.
-There is abundant evidence (e.g., \protect\cite{trick1994small}) that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities.
-We performed a pilot experiment in which we were able to confirm this finding in the experimental settings on which we are focussing in this paper \cite{our-earlier-paper}.
-In particular, we found that, when participants were confronted with an array of two boxes, instructions of the form ``Choose the square with $n$ dots" led to consistently lower Response Times than instructions of the form ``Choose the square with many/few dots" when $2 \leq n \leq 5$.
-The converse was true for $n>5$.
-% KvD: Can you check this addition about subitisable numbers, Matt?
-I'm not sure this stuff has found its right place in the paper yet.
-Where it is at the moment, it seems to break the story.
-Is it better after all to make this Experiment 1 (announced as a replication)?
-The {\bf selection algorithm} account predicts that responses will be faster and more accurate in the comparison conditions than in the matching conditions because it is easier to carry out comparison than matching.
+The {\bf instruction format} hypothesis predicts that responses will be faster and more accurate in the vague (i.e., verbal) conditions than in the crisp (i.e., numerical) conditions because numbers are harder to process than verbal references to quantities.
+
+The {\bf selection algorithm} account predicts that responses will be faster and more accurate in the vague (i.e., comparison) conditions than in the crisp (i.e., matching) conditions because it is easier to carry out comparison than matching.
+
 Thus all three accounts predict a main effect advantage of vagueness, but for different reasons.
-
-Regardless of which of the above hypotheses one adopts, one would predict that responses will be faster and more accurate when the numerical distance between the dot arrays is greater than when it is smaller because greater distances result in more easily discriminable arrays.
-Thus all three accounts predict a main effect of numerical distance, or gap size.
-For the same reason, one would predict that any advantage for the vague instructions should be smaller than when the arrays are least discriminable.
-This constitutes a prediction for an interaction between vagueness and numerical distance.
-% I've combined the old points (2) and (3) and removed the numbering.
-Maybe you can polish this further, finding a better place for these observations?
-
-% KvD I've tentatively deleted the following. I think it's probably superfluous.
-%The present experiment will demonstrate an advantage for our vague conditions, but not that the advantage is due to the vagueness of the instruction, because the vague conditions were also (a) verbal not numeric and (b) answerable with a comparison strategy not a matching strategy.
-Our subsequent experiments will discriminate between the alternative accounts.
+All three accounts also predict a main effect advantage for greater numerical distance, because greater numerical distances lead to more easily discriminable arrays.
+The accounts also share the prediction of an interaction effect whereby the main effect advantage of vagueness should diminish as numerical distance increases.
 
@@
 
-{\bf General.} All our experiments shared essentially the same methodology: Participants aged between ..., self-reported fluent speakers of English, were recruited by ....
+%{\bf General.} All our experiments shared essentially the same methodology: Participants aged between ..., self-reported fluent speakers of English, were recruited by ....
 and were paid 10 ponds for participating.
@@
 \subsection{Method}
-%\subsubsection{Participants}
 Twenty participants were recruited by mailing list and paid ten pounds for participating.
@@
 All participants self-reported fluency in English, and had normal, or corrected-to-normal vision.
-
-%\subsubsection{Apparatus}
 A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants.
 Stimuli were created and presented using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.
-
-%\subsubsection{\deleted{Design}}
-\deleted{We presented participants with 256 trials, arranged in 4 blocks of 64 trials each. }
-
-\deleted{It was always the case that two squares appeared on screen. One of these always contained 25 dots, and the other contained a number of dots that varied, and which was always the target. }
-
-\deleted{We can use the notation $pair_{(p)}(n,25)$ to represent a pair.
-The pairs were: $pair_{(1)}(5,25)$; $pair_{(2)}(10,25)$; $pair_{(3)}(15,25)$; $pair_{(4)}(20,25)$; $pair_{(5)}(30,25)$; $pair_{(6)}(35,25)$; $pair_{(7)}(40,25)$; $pair_{(8)}(45,25)$.}
-%
-\deleted{We manipulated two independent variables: gap size, with four levels (gap size = 5, 10, 15, or 20), and vagueness, with two levels (crisp and vague). }
-\deleted{We had to control another two variables: target size (whether the referring expression indicated the larger number of dots or the smaller number of dots); and target side (for a given level of target size, whether the target appeared on the left or the right).
-}
-\deleted{We measured, as our dependent variables, response time and response accuracy.}
-%
-\deleted{Our manipulation on gap size divided the items into four groups, one for each gap size.
-Items with gap size 5 were: $pair_{(4)}(20,25)$; and $pair_{(5)}(30,25)$; items with gap size 10 were: $pair_{(3)}(15,25)$; and $pair_{(6)}(35,25)$; items with gap size 15 were $pair_{(2)}(10,25)$ and $pair_{(7)}(40,25)$; and items with gap size 20 were $pair_{(1)}(5,25)$ and $pair_{(8)}(45,25)$.}
-%
-\deleted{Our manipulation on vagueness meant generating vague and crisp versions of each referring expression.
-Each pair was presented 8 times with a vague expression of quantity, and 8 times with a crisp expression of quantity.
-Taking $pair_{(1)}(5,25)$ as an example, the crisp referring expression was ``Choose the square with 5 dots'', and the vague referring expression was ``Choose the square with few dots''.
-%
-We needed to control which side the target was presented on. Each pair was presented equally often with with the target on the left, as with the target on the right.
-%
+We needed to control which side the target was presented on.
+Each pair was presented equally often with with the target on the left, as with the target on the right.
 We also needed to control whether the target was the smaller or larger number.
-The items with numbers less than 25 (pairs 1 to 4) formed a group with the smaller number as target and the other items (pairs 5 to 8) formed a balancing group with the larger number as the target number.}
-
-%\subsubsection{\deleted{Stimuli}}
-
-\deleted{Each stimulus consisted of two elements (a) the text of the referring expression indicating the target; (b) two squares containing dots. }
-\deleted{Where a number was mentioned in the instruction, it was always in the form of an arabic numeral (e.g., 1,2). }
-\deleted{This contrasts with experiment one, where numbers were given in natural language form.}
-\deleted{The position of the dots was randomised per-trial.}
-\deleted{Fig (\ref{stimuluse2}) gives an example stimulus.}
-
-\begin{figure}[tbp]
-\fitfigure{images/stimuluse2}
-\caption{An example stimulus from Experiment Two.
-First the referring expression was presented (left panel).
-Then after a keypress, and a fixation cross (not pictured) the squares and dots were presented without repetition of the referring expression (right panel)}
-\label{stimuluse2}
-\end{figure}
-
-%\subsubsection{Procedure}
+The items with numbers less than 25 (pairs 1 to 4) formed a group with the smaller number as target and the other items (pairs 5 to 8) formed a balancing group with the larger number as the target number.
 
@@
 The position of the dots was randomised per-trial.
-Response time was measured as the latency between the presentation of the dots and squares, and the keypress identifying the decision; in this way, the Response Times can be separated from readings times, which is important since we are only interested in the latter.
-% KvD OK?
+Response time was measured as the latency between the presentation of the dots and squares, and the keypress identifying the decision; in this way, the response times can be separated from time spent reading the instructions, which is important since we are only interested in the former.
 
@@
 No feedback was given on correct trials, but there was feedback on error trials in the form of the word ``\textsc{wrong!!}'' which flashed on screen.
+
+\begin{figure}[tbp]
+\fitfigure{images/stimuluse1}
+\caption{An example stimulus from Experiment 1.
+First the referring expression was presented (left panel).
+Then after a keypress, and a fixation cross (not pictured) the squares and dots were presented without repetition of the referring expression (right panel)}
+\label{stimuluse1}
+\end{figure}
 
@@
 \begin{figure}[htbp]
-\fitfigure{images/plotMeansBoth.pdf}
-\caption{Experiment Two results. }
-\label{resultse2}
+\fitfigure{images/resultse1.pdf}
+\caption{Experiment 1 results. }
+\label{resultse1}
 \end{figure}
@@
 
-The experiment provided support for hypothesis (1), that responses would be faster and more accurate for the vague instruction conditions than for the crisp instruction conditions. 
-We also found that responses grew faster and more accurate as the numerical distance between the dot arrays grew larger, in line with hypothesis (2).
-The experiment also provided support for hypothesis (3) that vagueness would be more advantageous at smaller than at larger gap sizes.  
-The findings of Experiment Two might be summarised thus: vague instructions are easier to process than crisp instructions, and returns for vagueness diminish as the stimuli become more easily discriminable.
+Results from experiment 1 were in line with the three main accounts.
+Responses were faster and more accurate when the instructions were in the vague (or verbal, or comparison) conditions than when they were in the crisp (or numerical, or selection) conditions.
+The speed and accuracy diminished as the arrays became more easily discriminable, and the advantage for vague (or verbal, or comparison) instructions tailed off as the arrays became more easily discriminable.
 
@@
 
-%A possible explanation for the pattern of results is cast in terms of the precision of an estimate of quantity that is required to carry out the task.
-For example, in the display pair$_{(1)}(5, 25)$, with a large gap size, vague instructions (\emph{many}, \emph{few}) might be able to  be carried out by a visual comparison that identifies one square as more numerous than the other.
-Precise instructions (to identify the square with 5 dots) might be able to be carried accurately out with an estimate with a fairly large error of 19.
-In the display pair$_{(8)}(20, 25)$, which has a small gap size, vague instructions might be able to be be carried out with a visual comparison that identifies one square as more numerous than the other.
-Precise instructions might require an estimate with a maximum error of 4 in this case.
-The pattern of results can be explained if we assume that the degree to which an estimate with error 4 is harder than an estimate with error 19 is greater than the degree to which a visual comparison between 20 and 25 is harder than a visual comparison between 5 and 25.
-
-%\item An account that we will dub the \emph{instruction format hypothesis}, or the number / no-number account, explains the vagueness advantage by placing importance on the observation that the vague conditions did not mention a numeral whereas the crisp conditions did.
-This could make the difference between a task that taps numerical processing, and one that taps magnitude judgements.
-The diminishing returns for vagueness can be explained by assuming that the numerical task is particularly challenging when the stimuli are less-discriminable, whereas the magnitude judgement is relatively less difficult in the less-discriminable trials.
-
-%\item \citeA{Rayna:1989lr} can account for the pattern in a framework they called the \emph{hierarchy of gist}.
-They proposed that reasoners encode representations at different levels of precision, and that these representations can be ordered with respect to precision, forming a hierarchy of gist.
-They further claimed that reasoning operated at the lowest level in this hierarchy that would allow one to accomplish the assigned task.
-Among the levels of this hierarchy are  (1) ratio representations at the top, which encode numerical information exactly; (2) ordinal representations that capture relative magnitude but do not encode numerical differences, i.e., they establish a rank ordering of quantities by magnitude, e.g., \emph{large}, \emph{larger}, \emph{largest}; (3) nominal or categorical representations that capture only the presence of absence of quantity e.g., \emph{some lives}; \emph{no lives}.
-\citeA{brainerd1994development} presented evidence that children spontaneously encode relative numerosity, at the ordinal level of the hierarchy mentioned above, when they are presented with stimuli that vary in numerosity.
-The hierarchy of gist offers an explanation of our vagueness advantage in experiment two that is independent of vagueness in the \citeA{keefe1996vagueness} sense.
-In experiment two, the participants were presented with the instruction for the trial first, and then with the squares and dots.
-This was done in order to try to separate the processes involved in processing the instruction from those involved in processing the stimuli.
-A consequence of this serial presentation though is that participants can infer from the instruction which level of representation is the minimum level necessary to carry out the impending task successfully.
-We do not suggest that this takes the form of a conscious deliberation -- rather it might be apprehended spontaneously in the same way that the children in \citeA{brainerd1994development} did without conscious deliberation.
-For example, it follows naturally from the instruction \emph{Please choose the square with fewest dots} that an ordinal representation will be necessary and sufficient for the purposes of the current trial: similarly it follows naturally from the instruction \emph{Please choose the square with 34 dots} that a higher level of the hierarchy will be necessary and sufficient for the purposes of this trial - the ratio level that encodes numerosity exactly.
-We found a response time advantage in experiment two for instructions that used the quantifiers \emph{many} and \emph{few}, when compared with times for quantifiers like \emph{15} and \emph{45}.
-In the terms of \citeA{Rayna:1989lr}, this can be explained as an advantage for the sufficiency of a lower level of representation -- without reference to vagueness.
-%
-%\item The \emph{two-systems account} of decision making \cite<e.g.,>{sloman2007two} can offer an explanation of the pattern.
-Dual process models of decision making propose two systems of processing: the quick and affective System 1 and the deliberative and rule-based System 2.
-This two-systems account can explain our finding of diminishing returns for vagueness in experiment 2.
-When the gap size is small, participants use System 2 (slow, deliberative) for the precise instructions.
-When the gap size is large, they take a shortcut for the crisp conditions and merely establish which square is more (or less) numerous, using System 1 (quick, heuristic).
-In the vague conditions, they use the heuristic system whether the gap size is small or large.
-This account explains the diminishing returns for vagueness that we observed in terms of a change of strategy in the baseline precise conditions as the gap size grew larger.
-%\end{seriate}
-
-%The potential for vagueness in the vague conditions can be said to have been unrealised in Experiment Two.
-The vague instructions were  \emph{choose the square with many dots} and \emph{choose the square with few dots}.
-\emph{Many} and \emph{few} have the potential for vagueness defined as the existence of borderline cases.
-Given a square with 15 dots, and asked the question `are there many dots in the square?', we can imagine that some people would be unsure (contrast this with the questions `are there fewer than 20 dots in the square?' to which there is no room for uncertainty).
-However, this potential for vagueness may be unrealised in the context of a choice between two squares.
-Given two squares, one with 20 dots, and another with 25 dots, and asked to `choose the square with few dots', it seems that no one could be unsure which square is the intended target.
-If there is no room for uncertainty, then the potential for \emph{many} and \emph{few} to be vague can be said to be unrealised.
-
-\section{Experiment Three}
+\section{Experiment 2}
 
@@
 
-The main result from experiment two was that responses were faster and more accurate for vague instructions than for crisp instructions.
-This finding can be interpreted in line with three different hypotheses.%: the cost reduction hypothesis, the instruction format account, and the selection algorithm account.
-In experiment three we set out to distinguish between the cost reduction hypothesis and the instruction format account, leaving the selection algorithm account for experiments four and five.
+The main result from experiment 1 was that responses were faster and more accurate for vague instructions than for crisp instructions.
+This finding can be interpreted in line with three different accounts: the cost reduction account, the instruction format account, and the selection algorithm account.
+
+In experiment 2 we set out to distinguish between the cost reduction account and the instruction format account, leaving the selection algorithm account for experiments 3 and 4.
 Contrast an expression from the vague condition: `the square with few dots' with an expression from the crisp condition: `the square with 15 dots'.
@@
 Another difference is that `few' is a linguistic, or verbal quantifier while `15' is a numerical quantifier, in the sense that a number is mentioned explicitly.
-Since these two differences were confounded in Experiment Two, the vagueness advantage finding is vulnerable to an alternative interpretation, that what we saw as a vagueness advantage was in contrast an advantage for the linguistic or verbal form of the quantifier.
-In the present experiment three we pitted these alternative interpretations against each other in a factual design.
+Since these two differences were confounded in Experiment 1, the vagueness advantage finding is vulnerable to an alternative interpretation, that what we saw as a vagueness advantage was in contrast an advantage for the verbal form of the quantifier.
+In the present experiment 2 we pitted these alternative interpretations against each other in a factorial design.
 
-In Experiment Two, the participants chose one of two squares.
+In Experiment 1, the participants chose one of two squares.
 The `vague' quantifiers (e.g., `few') uniquely identified one square.
@@
 This is because there were no borderline cases of the referent that could make the referent set `not well-defined'.
-Experiment three used three squares so that the vague quantifiers always had more than one possible referent.
-To enhance the potential for the vague quantifiers to have true vagueness in Experiment Three, we also used indefinite articles in the vague instructions.
+Experiment 2 used three squares so that the vague quantifiers always had more than one possible referent.
+To enhance the potential for the vague quantifiers to have true vagueness in Experiment 2, we also used indefinite articles in the vague instructions.
 Thus, instead of the instruction \emph{Choose the square with few dots}, we used \emph{Choose a square with few dots}.
 
-In experiment three, an item was a referring expression followed by a trio of numbers, representing the number of squares in the left, middle, and right squares.
+In experiment 2, an item was a referring expression followed by a triple of numbers, representing the number of squares in the left, middle, and right squares.
 We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54).
-Each triple had the following properties: it comprised three squares (instead of two as in Experiment Two); the central number was always presented in the middle of the three; there were two flanking numbers where one was smaller than the central number and one was bigger.
+Each triple had the following properties: it comprised three squares (instead of two as in Experiment 1); the central number was always presented in the middle of the three; there were two flanking numbers where one was smaller than the central number and one was bigger.
 
-There was a numerical and a verbal version of each of the vague and crisp referring expressions. See Table \ref{instructionseB} for examples.
-The vague numerical condition's referring expression was ``Choose a square with about 10 dots''.
+There was a numerical and a verbal version of each of the vague and crisp referring expressions. See Table \ref{instructionse2} for examples.
+The {\em vague numerical condition}'s referring expression was, on half of the presentations in that condition, ``Choose a square with about 10 dots''.
 None of the squares displayed contained 10 dots.
-%10 is slightly closer to 6 than to 15.
+10 is slightly closer to 6 than to 15.
 Therefore the best referent for this referring expression was the square with 6 dots;  the borderline response was the square with 15 dots; and the poorest referent was the square with 24 dots.
-% KvD: tentatively removed the above.
-%
-% KvD: In what follows, I've made a number of fairly trivial replacements and reorderings. Please check!
+On the other half of the presentations in that condition the referring expression was ``Choose a square with about 20 dots''.
 The {\em crisp numerical} conditions's referring expression was ``Choose the square with 6 dots'' on half the presentations in that condition and ``Choose the square with 24 dots'' on the other half.
@@
 In the {\em crisp verbal} condition, we used the referring expression ``Choose a square with fewer than $20$ (more than $10$) dots".
-The {\em vague numerical} condition's referring expression was ``Choose the square with far fewer than $20$ dots'' on half of presentations, making the squares with 6 and $15$ dots possible referents; iin the other case, the referring expression was ``Choose the square with far more than $10$ dots'', which made $15$ and $24$ possible referents.The {\em vague verbal} condition's referring expression was ``Choose a square with few dots'' on half of the presentations, and ``Choose a square with many dots'' on the other half.
+The {\em vague numerical} condition's referring expression was ``Choose the square with far fewer than $20$ dots'' on half of presentations, making the squares with 6 and $15$ dots possible referents; in the other case, the referring expression was ``Choose the square with far more than $10$ dots'', which made $15$ and $24$ possible referents.The {\em vague verbal} condition's referring expression was ``Choose a square with few dots'' on half of the presentations, and ``Choose a square with many dots'' on the other half.
 For this condition, the best referent was the square with 6 dots for `few' and 24 dots for `many'; the borderline case was the square with 15 dots; and the remaining square was the poorest referent.
-% KvD: OK, Matt? This needs checking once again.
-%
-% An indication that the manipulation of vagueness was successful is that participants chose the borderline case square on 16\% of trials.
-% KvD This last sentence needs to be reformulated, talking about {6,15} both.
+
+An indication that the manipulation of vagueness was successful is that participants chose the borderline case square on 16\% of trials.
 
 \begin{table}[htbp]
-\caption{Table of instructions arranged by condition for the triple (6,15,24). Experiment Three}
-\label{instructionseB}
+\caption{Table of instructions arranged by condition for the triple (6,15,24). Experiment 2}
+\label{instructionse2}
 \begin{tabular}{rlll}
@@
 \subsection{Hypotheses}
+The cost reduction account predicts a main effect of vagueness such that the vague instruction conditions attract faster responses than the crisp instruction conditions.
+The cost reduction account also predicts an interaction with the instruction format (or selection algorithm) variable: at each level of instruction format (or selection algorithm) the vague condition should attract faster responses than the crisp condition.%why is this interaction predicted?
+
 The instruction format account predicts a main effect of instruction format such that the numeric instruction conditions will attract longer response times, but has no prediction of any vagueness effects or interaction effects with vagueness.
@@
 
-The cost reduction account predicts that there will be a main effect of vagueness such that the vague instruction conditions attract faster responses than the crisp instruction conditions.
-The cost reduction account also predicts an interaction with the instruction format (or selection task) variable: at each level of instruction format (or selection task) the vague condition should attract faster responses than the crisp condition.
-
 \subsection{Method}
 
-Thirty participants were recruited and paid in the same way as Experiment Two.
+Thirty participants were recruited and paid in the same way as Experiment 1.
 They were aged between 18 and 45 with a median age of 28.
@@
 
-The same apparatus was used as for Experiment Two.
+The same apparatus was used as for Experiment 1.
 
@@
 
-Fig (\ref{stimuluse3}) gives an example stimulus.
+Fig (\ref{stimuluse2}) gives an example stimulus.
 First, the referring expression that constituted the instruction for that trial was displayed.
@@
 \centering
-\includegraphics[width=.4\textwidth]{images/stimuluse3}
-\caption{An example stimulus from Experiment Three}
-\label{stimuluse3}
+\includegraphics[width=.4\textwidth]{images/stimuluse2}
+\caption{An example stimulus from Experiment 2}
+\label{stimuluse2}
 \end{figure}
@@
 
-This time around, no responses were treated as erroneous, because errors were essentially undefined for the vague instructions (e.g., `about ten').
+This time around, no responses were treated as erroneous, because errors were essentially undefined for the vague instructions (e.g., `about 10').
 It was noted whether participants chose the borderline square.
 %
-The RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data. Means for response times and error rates are given in Fig. (\ref{resultse3}).
+The RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data. Means for response times are given in Fig. (\ref{resultse2}).
 
-A linear mixed model was constructed for the response times, with selection task (or instruction format), vagueness, and item and their interactions as independent variables, and random slopes over participant for task and vagueness (and their interaction) and for item.
-Task and vagueness were sum coded and item was centred.
-There was a significant effect of task with numerical conditions attracting longer responses than the verbal conditions ($\beta=.37, se=.07, t=5.1, p<.001$).
-The effect of vagueness was to slow responses down  ($\beta=.058, se=.013, t=4.8, p<.001$).
-There was an interaction effect between vagueness and task, such that the disadvantage for vagueness was greater in the numerical than in the verbal instruction conditions.
-There were differences in response time as a function of item ($\beta=.06, se=.008, t=7.0, p<.001$) but there was no clear trend across items.
+A linear mixed model was constructed for the response times.
+%
+Response times were logged; \emph{selection task / instruction format} and \emph{vagueness} were sum-coded and \emph{Item} was centred. 
+% 
+The fixed effects in the model were \emph{selection task / instruction format} and \emph{vagueness} and their interaction, and \emph{item}.
+% 
+The random effects in the model were \emph{participant}, and slopes over \emph{participant} for \emph{selection task / instruction format} and \emph{vagueness} and their interaction, and for \emph{item}.
+
+There was a significant effect of \emph{selection task /instruction format} with numerical conditions attracting longer responses than the verbal conditions (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; $\beta=.37, se=.07, t=5.1, p<.001$).
+% 
+The effect of \emph{vagueness} was to slow responses down (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06, se=.01, t=4.6, p<.001$). 
+%
+There was a main effect of \emph{item} ($\beta=.06, se=.008, t=7.0, p<.001$) indicating that response times differed in some way according to which item was presented.
+However further analysis revealed that there was no consistent smooth trend across items - this effect seems likely to be due to the very fast responses for the smallest item.
+This item had the largest ratio difference so may have been particularly discriminable for participants.
+% Reviews asked for an explanation of the difference for numeric crisp over items.
+Also reviews mention ratio differences as something that we did not control.
+%
+There was an interaction effect between vagueness and task, such that the disadvantage for vagueness was greater in the numerical than in the verbal instruction conditions ($\beta=-0.13, se= 0.02, t= -6.6, p=0.000$).
+[IS THIS CONSISTENT WITH THE ALL-NUMBERS EXPERIMENT 5?]
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.65\textwidth]{images/response-time-sep-1.pdf}
+\caption{Response time results for Experiment 2}
+\label{resultse2}
+\end{figure}
 
@@
 A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with task, vagueness and item as fixed effects, and with random slopes for task and vagueness and item over participants.
+%
+The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{borderline-response-distribution-e2}.
+%
 Participants were significantly more likely to choose the borderline option for vague instructions than for precise instructions (21.9\% vs 11.3\%, $\beta=.79, se=.25, z=3.2, p<.01$).
@@
 \centering
-\includegraphics[width=.65\textwidth]{images/resultse3.pdf}
-\caption{Results for Experiment Three}
-\label{resultse3}
+\includegraphics[]{images/histbd4-ggplot-version-1.pdf}
+\caption{Distribution of Nearest match, borderline case, and furthest match responses}
+\label{borderline-response-distribution-e2}
 \end{figure}
@@
 The disadvantage of numerical format overwhelms the contribution of vagueness.
-The verbal vague condition was still responded to faster than the numerical crisp condition, so the pattern from Experiment Two is reproduced, but in the light of the evidence from Experiment Three in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like either an advantage of verbal instruction format, or alternatively as an advantage of the comparison task, according to the selection task account.
+The verbal vague condition was still responded to faster than the numerical crisp condition, so the pattern from Experiment 1 is reproduced, but in the light of the evidence from Experiment 2 in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like either an advantage of verbal instruction format, or alternatively as an advantage of the comparison task, according to the selection task account.
 
 % KvD: I've tentatively added the following:
-Having effectively separated the {\bf cost reduction} hypothesis from the {\bf instruction format} hypothesis, it is important to observe that, in Experiment 3, instruction format went hand in hand with {\bf selection algorithm}: as shown by the Table of instructions \ref{instructionseB}, the instructions that used a verbal instruction format allowed a comparison algorithm, whereas the instructions that used a numeric format allowed on a matching algorithm.
+Having effectively separated the {\bf cost reduction} hypothesis from the {\bf instruction format} hypothesis, it is important to observe that, in Experiment 2, instruction format went hand in hand with {\bf selection algorithm}: as shown by the Table of instructions \ref{instructionse2}, the instructions that used a verbal instruction format allowed a comparison algorithm, whereas the instructions that used a numeric format allowed on a matching algorithm.
 Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison algorithm (which is known to be faster than matching).
@@
 In the next two experiments we pitted the comparison and matching selection tasks against each other while controlling vagueness and instruction format.
-In Experiment Four we would restrict all the instructions to numeric quantifiers while factorially manipulating vagueness and the comparison / matching selection tasks.
-In Experiment Five we would ensure that all instructions used verbal quantifiers, while also factorially manipulating vagueness and the comparison / matching selection tasks.
+In Experiment 3 we would restrict all the instructions to numeric quantifiers while factorially manipulating vagueness and the comparison / matching selection tasks.
+In Experiment 4 we would ensure that all instructions used verbal quantifiers, while also factorially manipulating vagueness and the comparison / matching selection tasks.
 This allows us to distinguish between the predictions of the selection task account and the instruction format account.
 
-\section{Experiment Four}
+\section{Experiment 3}
 
@@
 
-The main aim of experiment four was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies.
-The main changes from experiment three were that the selection task was explicitly controlled, and that all conditions were constrained to mention a number.
-We used the same stimuli as in experiment three.
-Table \ref{Instructions for e4} shows the instructions.
+The main aim of experiment 3 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies.
+The main changes from experiment 2 were that the selection task was explicitly controlled, and that all conditions were constrained to mention a number.
+We used the same stimuli as in experiment 2.
+Table \ref{Instructions for e3} shows the instructions.
 % KvD I think we should motivate our choice of materials (especially "far ..er")
@@
 \centering
-\caption{Experiment Four: Instructions, assuming $6,15,24$ dots as the Item, and showing \emph{fewer} instead of \emph{more}}
-\label{Instructions for e4}
+\caption{Experiment 3: Instructions, assuming $6,15,24$ dots as the Item, and showing \emph{fewer} instead of \emph{more}}
+\label{Instructions for e3}
 \begin{tabular}{llll}
@@
 38 volunteers were recruited via internal messaging at University of Aberdeen, with self-reported fluency in English. They were paid ten pounds each for participating.
-The apparatus was the same as Experiment Two.
-The design was a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e4}).
+The apparatus was the same as Experiment 1.
+The design was a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e3}).
 Each stimulus was an instruction followed by a triple of dots.
@@
 The structure of the model was as follows: fixed effects were vagueness, selection task (both sum coded) and centred item and their interactions: random effects were vagueness, selection task, and item (but not their interactions - the model failed to converge when these interactions were included).
-The means are plotted in Figure \ref{resultse4}.
+The means are plotted in Figure \ref{resultse3}.
 
@@
 \begin{figure}[htbp]
-\includegraphics[width=.5\textwidth]{images/resultse4.pdf}
-\caption{Results for Experiment Four}
-\label{resultse4}
+\centering
+\includegraphics[width=.5\textwidth]{images/response-time-two-panels-1.pdf}
+\caption{Results for Experiment 3}
+\label{resultse3}
 \end{figure}
 
-\section{Experiment Five}
+\section{Experiment 4}
 \subsection{Introduction}
@@
 In order to implement the experiment without mentioning numbers,  we used a prime to visually show the numbers of dots that we wanted to refer to as {\em the target} in the instructions. 
-\added{This presentation of a prime before the main trial shares some features with Experiment 2 in \protect \citeA{Izard20081221}, although in that experiment participants were told the numerosity of the prime - called an \emph{inducer} in that paper - in our experiment we did not tell participants the numerosity of the prime array.}
+This presentation of a prime before the main trial shares some features with Experiment 2 in \protect \citeA{Izard20081221}, although in that experiment participants were told the numerosity of the prime - called an \emph{inducer} in that paper - in our experiment we did not tell participants the numerosity of the prime array.
 An item was thus a combination of a visual prime, a numeric triple, and a referring expression.
-The referring expressions were constrained to never mention a numeral, as in Table \ref{Instructions for e5}.
+The referring expressions were constrained to never mention a numeral, as in Table \ref{Instructions for e4}.
 
@@
 \centering
-\caption{Experiment Five: Instructions}
-\label{Instructions for e5}
+\caption{Experiment 4: Instructions}
+\label{Instructions for e4}
 \begin{tabular}{lllp{7cm}}
@@
 40 volunteers recruited via internal messaging at University of Aberdeen, with self-reported fluency in English. They were paid ten pounds for participating. 
-The apparatus was the same as Experiment Two.
+The apparatus was the same as Experiment 1.
 The design was a 2 x 2 factorial manipulation of vagueness and selection task.
@@
 \subsection{Results}
-The results showed that vagueness was beneficial for comparison but detrimental for matching (the same as Experiment Four) even when no numbers were allowed in the instructions.
-Figure \ref{resultse5} shows the means by condition.
+The results showed that vagueness was beneficial for comparison but detrimental for matching (the same as Experiment 3) even when no numbers were allowed in the instructions.
+Figure \ref{resultse4} shows the means by condition.
 There was no main effect of vagueness ($\beta=.01, se=.01, t=1.51, p=.14$).
@@
 In the matching task vagueness slowed response times ($\beta=.05, se=.01, t=3.72, p<.05$).
-These results are in the same direction as Experiment Four.
+These results are in the same direction as Experiment 3.
 
 \begin{figure}[htbp]
-\includegraphics[width=.5\textwidth]{images/resultse5.pdf}
-\caption{Results for Experiment Five}
-\label{resultse5}
+\centering
+\includegraphics[width=.5\textwidth]{images/response-time-two-panels-1-e4.pdf}
+\caption{Results for Experiment 4}
+\label{resultse4}
 \end{figure}
@@
 
-\subsection{Discussion of experiments 4 and 5}
+\subsection{Discussion of experiments 3 and 4}
 
@@
 
-Experiment 2 showed us that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment could not distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format; and mandated a comparison strategy rather than a matching strategy.
-Experiment 3 showed us that number-avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional explanatory power in either the verbal format instructions or the numerical format instructions when we generated verbal and numerical versions of both crisp and vague instructions.
+Experiment 1 showed us that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment could not distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format; and mandated a comparison strategy rather than a matching strategy.
+Experiment 2 showed us that number-avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional explanatory power in either the verbal format instructions or the numerical format instructions when we generated verbal and numerical versions of both crisp and vague instructions.
 However the experiment could not distinguish benefits of number-avoidance in the verbal instructions from benefits of the comparison selection task: the verbal instructions also mandated a comparison strategy rather than a matching strategy.
-In experiments 4 and 5 we manipulated vagueness and the selection task independently of numerical format.
+In experiments 3 and 4 we manipulated vagueness and the selection task independently of numerical format.
 We found that there are effects of the selection task mandated by the instruction, with the comparison task instructions attracting faster response times than the matching instructions, and that vagueness exerts benefits when the selection task is \emph{comparison}, but not when the task is \emph{matching}.
 
-The benefits of vagueness in the \emph{comparison} task in experiments 4 and 5 could be explained as differences in the number of valid targets for the expression, as follows.
+The benefits of vagueness in the \emph{comparison} task in experiments 3 and 4 could be explained as differences in the number of valid targets for the expression, as follows.
 Taking as an example the stimulus with (6,15,24) dots, it could be argued that the vague comparison instruction (e.g., \emph{a square with far fewer than 20 dots}) has one valid target, the square with 6 dots, while the crisp comparison instruction (e.g., \emph{a square with fewer than 20 dots}) has two valid targets, the squares with 6 and 15 dots.
-In both experiments 4 and 5 we found that people were quicker to identify a square when the instruction only had one valid target.
+In both experiments 3 and 4 we found that people were quicker to identify a square when the instruction only had one valid target.
 This leads us to speculate that the benefit for vagueness here could be due to the vague expression foregrounding a particular valid target while the crisp expression carries with it the additional task of distinguishing between two alternative valid targets, something we propose to call a ``range-reduction'' benefit.
@@
 We believe the answer to this question is negative.
-To see why, consider figures \ref{resultse4} and \ref{resultse5}.
+To see why, consider figures \ref{resultse3} and \ref{resultse4}.
 Both figures depict four conditions, depending on whether the expression was crisp or vague, and depending on whether the referent could be identified using a comparison strategy or not.
@@
 
-\bibliography{lcn-aug-2014}
+\bibliography{vagueness-dec-2015}
+\clearpage
+\appendix
+
+\begin{sidewaystable}
+\centering
+\caption{All Instructions}
+\label{All Instructions}
+\begin{tabular}{lp{5cm}llllp{2cm}l}
+\midrule
+Experiment&Example instruction&vagueness&quantifier type&symmetry&selection&number of valid targets&definiteness\\
+\midrule
+1&Choose the square with 5 dots&crisp&numeric&symmetric&matching&1&definite\\
+1&Choose the square with few dots&vague&linguistic&asymmetric&comparison&1&definite\\
+\midrule
+2&Choose the square with 6 dots&crisp&numeric&symmetric&matching&1&definite\\
+2&Choose a square with about 10 dots&vague&numeric&symmetric&matching&2&indefinite\\
+2&Choose the square with the fewest dots&crisp&linguistic&asymmetric&comparison&1&definite\\
+2&Choose a square with few dots&vague&linguistic&asymmetric&comparison&2&indefinite\\
+\midrule
+3&Choose a square with 6 dots&crisp&numeric&symmetric&matching&1&indefinite\\
+3&Choose a square with fewer than 20 dots&crisp&numeric&asymmetric&comparison&2&indefinite\\
+3&Choose a square with about 10 dots&vague&numeric&symmetric&matching&2&indefinite\\
+3&Choose a square with far fewer than 20 dots&vague&numeric&asymmetric&comparison&1&indefinite\\
+\midrule
+4&Choose a square with the same number of dots as the target&crisp&linguistic&symmetric&matching&1&indefinite\\
+4&Choose a square with fewer dots than the target&crisp&linguistic&asymmetric&comparison&2&indefinite\\
+4&Choose a square with about the same number of dots as the target&vague&linguistic&symmetric&matching&2&indefinite\\
+4&Choose a square with far fewer dots than the target&vague&linguistic&asymmetric&comparison&1&indefinite\\
+\midrule
+\end{tabular}
+\end{sidewaystable}
 \end{document}

### 2016-07-27 experiments/e1/writeup/e1.Rmd
adding e1 of 4

---
title: "Experiment 1 of 4"
output:
  html_document:
    toc: yes
    mathjax: default
    template: default
    code_folding: hide
    theme: default
---

<style type="text/css">
.table {width: 50%;}
</style>

vagueness |instruction
----------|-------------------------------------------------------------------
crisp     |Choose the the square with many dots
vague     |Choose the square with 20 dots

items| a
-----|-----
1    |5,25
2    |10,25
3    |15,25
4    |20,25
-----|-----
5    |30,25
6    |35,25
7    |40,25
8    |45,25

In this experiment it was possible to identify errors, so error trials were removed. This is because there were only 2 squares and an istruction uniquely identified one of them.

# Prep data

```{r prepData}
rm(list=ls())
suppressPackageStartupMessages(library(knitr))
#opts_chunk$set(cache=TRUE, autodep=TRUE)
#opts_chunk$set(size="footnotesize")
#options(width=60)
```

```{r loadLibs}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(lattice))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(LMERConvenienceFunctions))
suppressPackageStartupMessages(library(lmerTest))
```

```{r sourceFunctions}
source("functions/CIfunctions.R")
```

## get data

```{r manipulateDataFrame, echo=FALSE}
system("cat ../experimentCode/output/*.dat > ../data/data.txt", wait=TRUE)
dat <- read.delim("../data/data.txt", header=FALSE)
names(dat) <- c("condition", "leftN", "rightN", "vgn", "side", "string", "subject", "rt", "error")
dat$subject <- as.factor(paste("sub",sep="",sprintf("%02d",dat$subject,sep="")))
dat$error <- 1-dat$error
dat$Ord <- ifelse(dat$leftN < dat$rightN, 'LtoR', 'RtoL')
dat$c_Ord <- ifelse(dat$Ord=='LtoR', -.5, .5)
dat$bigside <- ifelse(dat$leftN > dat$rightN, 'L', 'R')
dat$Qty <- ifelse(dat$side==dat$bigside, "Large", "Small")
dat$bigside=NULL
dat$c_Qty <- ifelse(dat$Qty=='Small', -.5, .5)
dat$gap <- as.character(paste("gap", sep=" ", sprintf("%02d",abs(dat$leftN-dat$rightN))))
dat$vgn <- as.factor(ifelse(dat$vgn=="P", "Crisp", "Vague"))
dat$c_Vgn <- ifelse(dat$vgn=='Crisp', -.5, .5)
dat$logrt <- log(dat$rt)
# item_ratio
# NOT DONE YET
write.table(x=dat, "all_trials.txt")
dat$subject <- as.factor(dat$subject)
dat$vgn <- ifelse(dat$vgn=='Crisp', -.5, .5)
dat$gap[dat$gap=="gap 05"] <- -1.5
dat$gap[dat$gap=="gap 10"] <- -.5
dat$gap[dat$gap=="gap 15"] <- .5
dat$gap[dat$gap=="gap 20"] <- 1.5
dat$gap <- as.numeric(dat$gap)
# Separate off correct trials
alltrials <- dat
correct <- subset(dat, error==0)
errors <-  subset(dat, error==1)
```

# RT model

```{r tableRTmodel, results='asis'}
print(
	xtable(
		digits=3, 
		summary(
			rt.lmer <- lmer(data=correct, logrt~gap*vgn + (vgn|subject))
			)$coefficients,
		caption="Most maximal response time model", 
		caption.placement="top"), type='html')
```

## RT means

For correct rts df, show means. Vagueness appears to be beneficial.

```{r plotCorrectRTMeans, fig.width=3, fig.height=3}
a <- summarySEwithin(correct,  measurevar="logrt",  withinvars=c("vgn"),  idvar="subject")
ggplot(data=a, aes(y=logrt_norm, x=vgn, group=1, ymin=logrt_norm-ci, ymax=ci+logrt_norm)) + geom_line() + geom_point() + geom_errorbar(width=.05) + xlab("") + ylab("log rt") + theme_bw() + theme(aspect.ratio=1) 
```

# Error rates model

```{r tableErrorModel, results='asis'}
print(xtable(digits=3, summary(ac.lmer <- 
															 	glmer(data=alltrials, error~gap*vgn + (gap+vgn|subject), 
															 				family="binomial", 
															 				control = glmerControl(optimizer = "bobyqa")))$coefficients,
						 caption="Most maximal accuracy model",
						 caption.placement="top"),type='html')
```

```{r qq, fig.width=5, fig.height=3}
op <- par(mfrow=c(1,2))
qqPlot(residuals(rt.lmer), cex=.2)
qqPlot(residuals(ac.lmer), cex=.2)
```

```{r plotMeansBoth, fig.width=10, fig.height=4, message=FALSE}
rts <- summarySEwithin(correct, measurevar="logrt", withinvars=c("vgn","gap"))
err <- summarySEwithin(alltrials, measurevar="error", withinvars=c("vgn","gap"))
dodge=.06

plot1 =
  ggplot(rts, aes(y=logrt_norm, x=gap, group=vgn, fill=vgn, shape=vgn, ymin=logrt_norm-ci, ymax=logrt_norm+ci)) + 
  geom_line(position=position_dodge(dodge)) + 
  geom_errorbar(width=.3, lty=1, position=position_dodge(dodge)) + 
  geom_point(pch=21, size=3, position=position_dodge(dodge)) +
  ggtitle("Response time") + 
  ylab("log RT") + 
  xlab("Gap Size") +
  scale_fill_grey(name=element_blank(), start=0,end=1)+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank() ) +
  theme(aspect.ratio=1)+xlab("")

plot2 = 
  ggplot(err, aes(y=error_norm, x=gap, group=vgn, fill=vgn, shape=vgn, ymin=error_norm-ci, ymax=error_norm+ci)) + 
  geom_line(position=position_dodge(dodge)) + 
  geom_errorbar(width=.3, lty=1, position=position_dodge(dodge)) + 
  geom_point(pch=21, size=3, position=position_dodge(dodge)) +
  ggtitle("Error rate") + 
  ylab("error rate") + 
  xlab("Gap Size") +
  scale_fill_grey(name=element_blank(), start=0,end=1)+
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank() )+
  theme(aspect.ratio=1)+xlab("")

grid.arrange(plot1,plot2,ncol=2)
```

### 2016-07-27 experiments/e1/experimentCode/experiment.m

@@ -1,0 +1,690 @@
+function experiment(SUBJECT)
+% This second study has the following structure: practice block followed by:4 blocks with a rest in
+% between. Each block has 2 presentations of 32 trials. The 32 trial are
+% composed of 8 pairs of quantities presented once with vague instructions
+% and once with precise instructions, with then element of the pair that is
+% being picked out presented on the left and on the right (8 pairs * 2
+% sides * 2 instructions types * sides = 32 trials, presented twice per
+% block = 64 trials; each block presented 4 times = 256 trials).
+%
+% A trial consists of the instruction on screen while participants read it
+% for as long as they like; a button press on SPACEBAR to dismiss the
+% instruction; a blank screen of 100ms duration; a fixation cross of 100ms
+% duration; a blank of 100ms duration; a screen with the pair presented
+% (currently for unlimited duration terminated by a key press from a
+% restricted set (left cmd, right cmd) ); 1000ms ISI.
+%
+% Participant id > 100 is used for testing and can be overwritten
+
+howLongToWaitForAResponse = 60; % seconds. Set low for test runs
+testing = 0; % 0 for real runs
+
+% Clear Matlab/Octave window:
+clc;
+
+% check for Opengl compatibility, abort otherwise:
+AssertOpenGL;
+
+% Reseed the random-number generator for each expt.
+rand('state',sum(100*clock));
+
+% Make sure keyboard mapping is the same on all supported operating systems
+% Apple MacOS/X, MS-Windows and GNU/Linux:
+KbName('UnifyKeyNames');
+
+%     % Turn off the annoying echo to the command line
+%     ListenChar(2);
+
+% Define allowable keys
+leftresp=KbName('LeftGUI'); %227
+rightresp=KbName('RightGUI'); %231
+spacebar=KbName('space'); %44
+%    RestrictKeysForKbCheck([leftresp,rightresp,spacebar]);
+
+%%%%%%%%%%%%%%%%%%%%%%
+% file handling
+%%%%%%%%%%%%%%%%%%%%%%
+
+%practice stimuli to read from
+practiceStimuliFilename = 'practiceStimuli.txt';
+
+%practice data output file to write to
+practicedatafilename = strcat('output/subject',...
+    num2str(SUBJECT),'Practice','.dat');
+
+% experimental stimuli to read from
+experimentalStimuliFilename = 'experimentalStimuli.txt';
+
[…]

### 2016-07-27 experiments/e1/experimentCode/experimentpresentation.m

@@ -1,0 +1,690 @@
+function experimentpresentation(SUBJECT)
+% This second study has the following structure: practice block followed by:4 blocks with a rest in
+% between. Each block has 2 presentations of 32 trials. The 32 trial are
+% composed of 8 pairs of quantities presented once with vague instructions
+% and once with precise instructions, with then element of the pair that is
+% being picked out presented on the left and on the right (8 pairs * 2
+% sides * 2 instructions types * sides = 32 trials, presented twice per
+% block = 64 trials; each block presented 4 times = 256 trials).
+%
+% A trial consists of the instruction on screen while participants read it
+% for as long as they like; a button press on SPACEBAR to dismiss the
+% instruction; a blank screen of 100ms duration; a fixation cross of 100ms
+% duration; a blank of 100ms duration; a screen with the pair presented
+% (currently for unlimited duration terminated by a key press from a
+% restricted set (left cmd, right cmd) ); 1000ms ISI.
+%
+% Participant id > 100 is used for testing and can be overwritten
+
+howLongToWaitForAResponse = 60; % seconds. Set low for test runs
+testing = 0; % 0 for real runs
+
+% Clear Matlab/Octave window:
+clc;
+
+% check for Opengl compatibility, abort otherwise:
+AssertOpenGL;
+
+% Reseed the random-number generator for each expt.
+rand('state',sum(100*clock));
+
+% Make sure keyboard mapping is the same on all supported operating systems
+% Apple MacOS/X, MS-Windows and GNU/Linux:
+KbName('UnifyKeyNames');
+
+%     % Turn off the annoying echo to the command line
+%     ListenChar(2);
+
+% Define allowable keys
+leftresp=KbName('q'); %227
+rightresp=KbName('p'); %231
+spacebar=KbName('space'); %44
+%    RestrictKeysForKbCheck([leftresp,rightresp,spacebar]);
+
+%%%%%%%%%%%%%%%%%%%%%%
+% file handling
+%%%%%%%%%%%%%%%%%%%%%%
+
+%practice stimuli to read from
+practiceStimuliFilename = 'practiceStimuli.txt';
+
+%practice data output file to write to
+practicedatafilename = strcat('dummy/subject',...
+    num2str(SUBJECT),'Practice','.dat');
+
+% experimental stimuli to read from
+experimentalStimuliFilename = 'experimentalStimuli.txt';
+
[…]

### 2016-07-27 experiments/e1/writeup/functions/CIfunctions.R

@@ -1,0 +1,140 @@
+## Summarizes data.
+## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
+##   data: a data frame.
+##   measurevar: the name of a column that contains the variable to be summariezed
+##   groupvars: a vector containing names of columns that contain grouping variables
+##   na.rm: a boolean that indicates whether to ignore NA's
+##   conf.interval: the percent range of the confidence interval (default is 95%)
+summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
+                      conf.interval=.95, .drop=TRUE) {
+  require(plyr)
+  
+  # New version of length which can handle NA's: if na.rm==T, don't count them
+  length2 <- function (x, na.rm=FALSE) {
+    if (na.rm) sum(!is.na(x))
+    else       length(x)
+  }
+  
+  # This is does the summary; it's not easy to understand...
+  datac <- ddply(data, groupvars, .drop=.drop,
+                 .fun= function(xx, col, na.rm) {
+                   c( N    = length2(xx[,col], na.rm=na.rm),
+                      mean = mean   (xx[,col], na.rm=na.rm),
+                      sd   = sd     (xx[,col], na.rm=na.rm)
+                   )
+                 },
+                 measurevar,
+                 na.rm
+  )
+  
+  # Rename the "mean" column    
+  datac <- rename(datac, c("mean"=measurevar))
+  
+  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
+  
+  # Confidence interval multiplier for standard error
+  # Calculate t-statistic for confidence interval: 
+  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
+  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
+  datac$ci <- datac$se * ciMult
+  
+  return(datac)
+}
+
+
+## Summarizes data, handling within-subjects variables by removing inter-subject variability.
+## It will still work if there are no within-S variables.
+## Gives count, un-normed mean, normed mean (with same between-group mean),
+##   standard deviation, standard error of the mean, and confidence interval.
+## If there are within-subject variables, calculate adjusted values using method from Morey (2008).
+##   data: a data frame.
+##   measurevar: the name of a column that contains the variable to be summariezed
[…]

### 2016-07-28 experiments/e2/writeup/main.Rmd
adding e2 with renaming to e2

---
title: "Experiment 2 of 4: main"
output:
  html_document:
    toc: true
    <!--toc_float: true-->
    mathjax: default
    template: default
    code_folding: hide
    theme: default
---

```{r child="print_info_on_design.Rmd"}
```

# Distributional stuff, including transformations.

```{r child = 'transformations_table.Rmd'}
```

```{r child = 'libraries.Rmd'}
```

```{r child = 'data_wrangling.Rmd'}
```

```{r child = 'plot_transformations_against_normal_equivalents.Rmd'}
```

# Now working with a data set without impossible trials.

```{r child = 'plot_boxplot_subjects_and_items.Rmd'}
```

```{r child = 'plot_fast_and_slow_subjects_and_items.Rmd'}
```

```{r child = 'plot_fast_and_slow_subjects_and_items_RecSqt.Rmd'}
```

```{r child = 'plot_practice_and_fatigue_effects.Rmd'}
```

# Plot main effects and interactions

```{r child = 'plot_interaction_effects.Rmd'}
```

# Show compression of item ratios

```{r child = 'show_compression_of_ratios.Rmd'}
```

# Model ols()

```{r child = "model_ols_1_with_baayen4plot_model_criticism.Rmd"}
```

# Model lmer()

```{r child = "model_lmer_0.Rmd"}
```

### 2016-07-28 experiments/e2/writeup/model_code_not_used.Rmd
adding e2 with renaming to e2

models not used

## ols model v1, Rec

```{r, fig.width=11, fig.height=3}
# do ols model

# define datadist
dd.dd = datadist(dd)
options(datadist = "dd.dd")
# build model
v1 <- ols(data=dd, RT_Rec ~ c_Vag + c_Num + c_Itm + c_Qty + c_Ord + s_Trl, x=T, y=T)
cat("R2 = ", v1$stats[4])
# show model
#v1
# show model summary
#summary(v1)
# do anova 
an1 = anova(v1)
# show anova
#an1
# generate predictions
p1=Predict(v1)
# plot 'partial effects'
plot(p1, anova=an1, pval=T, layout=c(6,1), aspect=1, main="ols model v1")
```

```{r, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

par(mfrow=c(1,4), pty='s')
# create scaled residuals
v1$rstand = as.vector(scale(resid(v1)))
# plot scaled residuals density
plot(density(v1$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(v1$rstand, cex=.5)
qqline(v1$rstand)
# plot standardised residuals versus fitted values
plot(v1$rstand ~ fitted(v1), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))
# create dffits
dffits=abs(resid(v1, "dffits"))
# plot dffits
plot(dffits, type='h')
```

## ols model v2, Rec, with polynomials

```{r, fig.width=11, fig.height=3}
# do ols model

# define datadist
dd.dd = datadist(dd)
options(datadist = "dd.dd")
# build model
v2 <- ols(data=dd, RT_Rec ~ c_Vag * c_Num + pol(c_Itm,2) + c_Qty + c_Ord + pol(s_Trl,6) + c_Vag:c_Qty, x=T, y=T)
cat("R2 = ", v2$stats[4])
# show model
#v2
# show model summary
#summary(v2)
# do anova 
an1 = anova(v2)
# show anova
#an1
# generate predictions
p1=Predict(v2)
# plot 'partial effects'
plot(p1, anova=an1, pval=T, layout=c(6,1), aspect=1, main="ols model v2")
```

```{r, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

par(mfrow=c(1,4), pty='s')
# create scaled residuals
v2$rstand = as.vector(scale(resid(v2)))
# plot scaled residuals density
plot(density(v2$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(v2$rstand, cex=.5)
qqline(v2$rstand)
# plot standardised residuals versus fitted values
plot(v2$rstand ~ fitted(v2), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))
# create dffits
dffits=abs(resid(v2, "dffits"))
# plot dffits
plot(dffits, type='h')
```

## ols model v3, RecSqt, with polynomials

```{r, fig.width=11, fig.height=3}
# do ols model

# define datadist
dd.dd = datadist(dd)
options(datadist = "dd.dd")
# build model
v3 <- ols(data=dd, RT_RecSqt ~ c_Vag * c_Num + pol(c_Itm,2) + c_Qty + c_Ord + pol(s_Trl,6) + c_Vag:c_Qty, x=T, y=T)
cat("R2 = ", v3$stats[4])
# show model
#v3
# show model summary
#summary(v3)
# do anova 
an1 = anova(v3)
# show anova
#an1
# generate predictions
p1=Predict(v3)
# plot 'partial effects'
plot(p1, anova=an1, pval=T, layout=c(6,1), aspect=1, main="ols model v3")
```

```{r, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

par(mfrow=c(1,4), pty='s')
# create scaled residuals
v3$rstand = as.vector(scale(resid(v3)))
# plot scaled residuals density
plot(density(v3$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(v3$rstand, cex=.5)
qqline(v3$rstand)
# plot standardised residuals versus fitted values
plot(v3$rstand ~ fitted(v3), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))
# create dffits
dffits=abs(resid(v3, "dffits"))
# plot dffits
plot(dffits, type='h')
```

### 2016-07-28 experiments/e2/writeup/plot_interaction_effects.Rmd
adding e2 with renaming to e2

---
title: "Experiment 2 of 4: interaction_effects"
output:
  html_document:
    template: default
    code_folding: hide
    theme: default
---

```{r, message=FALSE, results='hide', echo=FALSE}
load("dd.Rda")
source('functions/CIfunctions.R')
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(plyr))
if(require(reshape2)){}
```

Main effects

Plot main effects on several transformations

```{r, fig.width=16, fig.height=6, warning=FALSE}
# plot main effects on observed data

# allow reference to transformation
temp <- melt(dd,
             id.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
             measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log",  "RT_raw"),
             variable.name="transformation",
             value.name="score")

# allow reference to effect
temp2 <- melt(temp, 
              id.vars=c("transformation", "score"),
              measure.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
              variable.name="effect",
              value.name="level")

# do the plot
ggplot(temp2, aes(y=score, x=level, group=1, col=effect)) +
  facet_wrap(~transformation + effect, scales="free", shrink=TRUE, ncol=5) +
  stat_summary(fun.data=mean_cl_normal, geom="pointrange") +
  stat_summary(fun.y=mean, geom="line") +
  theme(aspect.ratio=1, axis.text.y=element_blank())
```

Main effects

```{r}
# prep main effects rts data frame
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness")));names(vrts)[3]<-"Levels"
nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number")));names(nrts)[3]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Quantity")));names(qrts)[3]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Order")));names(orts)[3]<-"Levels"
irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Item")));names(irts)[3]<-"Levels"
rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
```

```{r, fig.width=9, fig.height=3}
# plot main effects
ggplot(rts, aes(y=RT_RecSqt, x=Levels, group=1, ymin=mins, ymax=maxs)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Effect, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) 
```

Main effects over items

```{r, message=FALSE, fig.width=10, fig.height=3}
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Vagueness")));names(vrts)[4]<-"Levels"
nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Number")));names(nrts)[4]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Quantity")));names(qrts)[4]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Order")));names(orts)[4]<-"Levels"

rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci

rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))

ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, group=Levels, ymin=mins, ymax=maxs, shape=Levels, fill=Levels, col=Levels)) +
  scale_shape_manual(values=c(21,21,22,22,23,23,24,24)) +
  scale_fill_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
  scale_color_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Effect) + 
  theme_bw() +
  theme(aspect.ratio=1, 
        panel.grid=element_blank()) +
  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))

```

2 way interactions not involving items.

```{r}
# prep 2 ways rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Number")); names(rts1)[1] <- "F1"; names(rts1)[2] <- "F2"
rts1$Effect <- rep("Vagueness:Number",4)
rts2 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Quantity")); names(rts2)[1] <- "F1"; names(rts2)[2] <- "F2"
rts2$Effect <- rep("Vagueness:Quantity",4)
rts3 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Order")); names(rts3)[1] <- "F1"; names(rts3)[2] <- "F2"
rts3$Effect <- rep("Vagueness:Order",4)
rts5 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Quantity")); names(rts5)[1] <- "F1"; names(rts5)[2] <- "F2"
rts5$Effect <- rep("Number:Quantity",4)
rts6 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Order")); names(rts6)[1] <- "F1"; names(rts6)[2] <- "F2"
rts6$Effect <- rep("Number:Order",4)
rts8 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Quantity","Order")); names(rts8)[1] <- "F1"; names(rts8)[2] <- "F2"
rts8$Effect <- rep("Quantity:Order",4)
rts <- rbind(rts1,rts2,rts3,rts5,rts6,rts8)
rts <- subset(rts, select=c(Effect, F1, F2, RT_RecSqt, RT_RecSqt_norm, sd, se, ci))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
```

```{r,fig.width=10, fig.height=3}
# plot 2 way interactions
ggplot(rts, aes(y=RT_RecSqt, x=F2, group=F1, ymin=mins, ymax=maxs, pch=F1)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_wrap(~Effect, scale="free_x", nrow=1) + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank()) +
  xlab("Levels")
```

Plot 2 way interactions over items

```{r, message=FALSE}
# prep 2 ways over items rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Number")); names(rts1)[2] <- "F1"; names(rts1)[3] <- "F2"
rts1$E1 <- "Vagueness"
rts1$E2 <- "Number"
rts2 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Quantity")); names(rts2)[2] <- "F1"; names(rts2)[3] <- "F2"
rts2$E1 <- "Vagueness"
rts2$E2 <- "Quantity"
rts3 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Order")); names(rts3)[2] <- "F1"; names(rts3)[3] <- "F2"
rts3$E1 <- "Vagueness"
rts3$E2 <- "Order"
rts4 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Quantity", "Number")); names(rts4)[2] <- "F1"; names(rts4)[3] <- "F2"
rts4$E1 <- "Quantity"
rts4$E2 <- "Number"
rts5 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Order","Number")); names(rts5)[2] <- "F1"; names(rts5)[3] <- "F2"
rts5$E1 <- "Order"
rts5$E2 <- "Number"
rts6 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Quantity","Order")); names(rts6)[2] <- "F1"; names(rts6)[3] <- "F2"
rts6$E1 <- "Quantity"
rts6$E2 <- "Order"

rts <- rbind(rts1,rts2,rts3,rts4,rts5,rts6)
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci

rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
rts$F3 <- factor(paste(rts$F1,rts$F2))

rts$E1 <- factor(rts$E1, levels=c("Number","Order","Quantity","Vagueness"))
rts$E2 <- factor(rts$E2, levels=c("Number","Order","Quantity","Vagueness"))

# The palette with black:
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, group=F3, col=F2, pch=F1)) + 
  facet_grid(E2~E1, drop=F) + 
  geom_point(cex=3) + geom_line() + theme_bw() + 
  theme(aspect.ratio=1, panel.grid=element_blank()) +
  scale_color_manual(name="colour",values=cbbPalette) +
  scale_shape_discrete(name="shape") +
  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))

```

Plot vagueness by number interaction over items

```{r, fig.width=7, fig.height=3.5, eval=TRUE, cache=FALSE}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
dodge = position_dodge(width=0.2)
ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
                group=condition, shape=condition, fill=condition)) +
  geom_line(position=dodge) + 
  geom_errorbar(width=.2, position=dodge) + 
  geom_point(size=4, position=dodge) + 
  scale_shape_manual("",values = c(22, 22, 21, 21)) + 
  scale_fill_manual("",values=c("black","white","black","white")) +
  ggtitle("Response time") + 
  ylab("RT_RecSqt") + 
  xlab("") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(~Number)
```

3 way interactions

```{r}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Quantity"))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
p1=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Vagueness, ymin=mins, ymax=maxs, pch=Vagueness)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Quantity, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank(), legend.position="top")
```

```{r}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Order"))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
p2=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Vagueness, ymin=mins, ymax=maxs, pch=Vagueness)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Order, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank(), legend.position="top")
```

```{r}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Quantity","Order"))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
p3=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Quantity, ymin=mins, ymax=maxs, pch=Quantity)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Order, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank(), legend.position="top")
```

```{r}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Quantity","Order"))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
p4=ggplot(rts, aes(y=RT_RecSqt, x=Vagueness, group=Quantity, ymin=mins, ymax=maxs, pch=Quantity)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Order, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank(), legend.position="top")
```

```{r, fig.width=10,fig.height=5}
grid.arrange(nrow=2, p1,p2,p3,p4)
```

Plot vagueness by number by quantity over items.

```{r, fig.width=10, fig.height=4, eval=TRUE}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item", "Quantity"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
dodge = position_dodge(width=0.2)
ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
                group=Vagueness, 
                fill=Vagueness)) +
  geom_line(position=dodge) + 
  geom_errorbar(width=.2, position=dodge) + 
  geom_point(pch=21, size=4, position=dodge) + 
  scale_fill_manual("",values=c("black","white")) +
  ggtitle("Response time") + 
  ylab("RT_Rec_Sqt") + 
  xlab("") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background=element_rect(fill="white",color=1),
        legend.key = element_blank(), 
        aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(~Number+Quantity)
```

4 way interaction

```{r}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness",  "Quantity", "Order"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
ggplot(rts, aes(y=RT_RecSqt, x=Vagueness, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, group=Quantity, pch=Quantity)) +
  facet_grid(~Order+Number, scale="free_x") +
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank())
```

4 way interaction split over items

```{r, fig.width=8, fig.height=8, eval=TRUE, cache=FALSE, echo=TRUE}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness", "Item", "Quantity", "Order"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
dodge = position_dodge(width=0.2)
ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
                group=Vagueness,fill=Vagueness)) +
  geom_line(position=dodge) + 
  geom_errorbar(width=.2, position=dodge) + 
  geom_point(pch=21, size=4, position=dodge) + 
  scale_fill_manual("",values=c("black","white","black","white")) +
  ggtitle("Number * Vagueness * Quantity * Order") + 
  ylab("RT_RecSqt") + 
  xlab("") +
  theme_bw() +
  theme( legend.key = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), axis.text.x = element_text(angle = 45, hjust = 1), aspect.ratio=1) +
  facet_grid(Vagueness+Order~Number+Quantity)
```

4 way interaction split over items showing ratio of item dots.

```{r, fig.width=8, fig.height=8, eval=TRUE, cache=FALSE, echo=TRUE, message=FALSE, warning=FALSE}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness", "item_mean_ratio", "Quantity", "Order"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
dodge = position_dodge(width=0.2)
ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
                group=Vagueness,fill=Vagueness)) +
  geom_line(position=dodge) + 
  geom_errorbar(width=.2, position=dodge) + 
  geom_point(pch=21, size=4, position=dodge) + 
  scale_fill_manual("",values=c("black","white","black","white")) +
  ggtitle("Number * Vagueness * Quantity * Order") + 
  ylab("RT_RecSqt") + 
  xlab("") +
  theme_bw() +
  theme( legend.key = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), axis.text.x = element_text(angle = 45, hjust = 1), aspect.ratio=1) +
  facet_grid(Vagueness+Order~Number+Quantity)
```

### 2016-07-28 experiments/e2/writeup/print_info_on_design.Rmd
adding e2 with renaming to e2

```{r}
# print info on design, number of cells
cat(paste('number of subjects = ',30,'\n',
          '---','\n',
          'number of trials per subject = ',256,'\n', 
          '---','\n',          
          'number of unique cells in the design = Item(4) * Number(2) * Vagueness(2) * Order(2) * Quantity(2) = ',4*2*2*2*2,'\n',
          'number of times a subject saw a unique cell in the design = 256 / 64 = ', 256/64,'\n',
          '---','\n',
          'number of higher-level cells in the design (abstracting over Order and Quantity) = Item(4) * Number(2) * Vagueness(2) = ',4*2*2,'\n',
          'number of times a subject saw a higher level cell in the design =  = 256 / 16 = ', 4*2*2, '\n',
          sep=''))
```

### 2016-07-28 experiments/e2/writeup/show_compression_of_ratios.Rmd
adding e2 with renaming to e2

```{r, fig.width=10, fig.height=7, echo=F}
#  Show compression of ratios of numbers of dots in the items
item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54)
item_range_ratio_equally_spaced = seq(from = 6 / 24, to = 36 / 54, length.out = 4)
item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54)))
item_mean_ratio_equally_spaced = seq(from = mean(c(6 / 15, 15 / 24)), to = mean(c(36 / 45, 45 / 54)), length.out = 4)

plot(y=c(2,2,2,2), x=item_range_ratio, xlim=c(0,1), ylim=c(0,3), ylab="", yaxt='n', col="blue", pch=21, bg='blue', type='o', xaxt='n', xlab="ratio: small is easily discriminable", main="Discriminability of each item, expressed as ratios between array magnitudes")
text(y=c(2.15,2.15,2.15,2.15), x=item_range_ratio, labels=round(item_range_ratio,2), col="blue")
text(y=2.15, x=0.1, labels="ratio min/max", col='blue')
points(y=c(2.3,2.3,2.3,2.3), x=item_range_ratio_equally_spaced, col="lightblue", pch=21, bg='lightblue', type='o')

points(y=c(1,1,1,1), x=item_mean_ratio, pch=21, col='red', bg='red', type='o')
text(y=c(1.15,1.15,1.15,1.15), x=item_mean_ratio, labels=round(item_mean_ratio,2), col="red")
text(y=1.15, x=0.1, labels="mean (ratio min/med, ratio med/max)", col='red')
points(y=c(1.3,1.3,1.3,1.3), x=item_mean_ratio_equally_spaced, col="pink", pch=21, bg='pink', type='o')
axis(1, at=seq(0,1,0.1))

```

### 2016-07-28 experiments/e1/experimentCode/exp1screenshots.m

@@ -1,0 +1,430 @@
+function exp2screenshots(SUBJECT)
+% This second study has the following structure: practice block followed by:4 blocks with a rest in
+% between. Each block has 2 presentations of 32 trials. The 32 trial are
+% composed of 8 pairs of quantities presented once with vague instructions
+% and once with precise instructions, with then element of the pair that is
+% being picked out presented on the left and on the right (8 pairs * 2
+% sides * 2 instructions types * sides = 32 trials, presented twice per
+% block = 64 trials; each block presented 4 times = 256 trials).
+%
+% A trial consists of the instruction on screen while participants read it
+% for as long as they like; a button press on SPACEBAR to dismiss the
+% instruction; a blank screen of 100ms duration; a fixation cross of 100ms
+% duration; a blank of 100ms duration; a screen with the pair presented
+% (currently for unlimited duration terminated by a key press from a
+% restricted set (left cmd, right cmd) ); 1000ms ISI.
+%
+% Participant id > 100 is used for testing and can be overwritten
+
+SUBJECT=1000;
+
+howLongToWaitForAResponse = 60; % seconds. Set low for test runs
+testing = 1; % 0 for real runs
+
+% Clear Matlab/Octave window:
+clc;
+
+% check for Opengl compatibility, abort otherwise:
+AssertOpenGL;
+
+% Reseed the random-number generator for each expt.
+rand('state',sum(100*clock));
+
+% Make sure keyboard mapping is the same on all supported operating systems
+% Apple MacOS/X, MS-Windows and GNU/Linux:
+KbName('UnifyKeyNames');
+
+%     % Turn off the annoying echo to the command line
+%     ListenChar(2);
+
+% Define allowable keys
+leftresp=KbName('LeftGUI'); %227
+rightresp=KbName('RightGUI'); %231
+spacebar=KbName('space'); %44
+
+
+%%%%%%%%%%%%%%%%%%%%%%
+% file handling
+%%%%%%%%%%%%%%%%%%%%%%
+
+
+%practice data output file to write to
+practicedatafilename = strcat('output/subject',...
+    num2str(SUBJECT),'Practice','.dat');
+
+% experimental stimuli to read from
+experimentalStimuliFilename = 'experimentalStimuli.txt';
+
+% experimental data output file to write to
[…]

### 2016-07-28 experiments/e1/experimentCode/exp2screenshots.m

@@ -1,430 +1,0 @@
-function exp2screenshots(SUBJECT)
-% This second study has the following structure: practice block followed by:4 blocks with a rest in
-% between. Each block has 2 presentations of 32 trials. The 32 trial are
-% composed of 8 pairs of quantities presented once with vague instructions
-% and once with precise instructions, with then element of the pair that is
-% being picked out presented on the left and on the right (8 pairs * 2
-% sides * 2 instructions types * sides = 32 trials, presented twice per
-% block = 64 trials; each block presented 4 times = 256 trials).
-%
-% A trial consists of the instruction on screen while participants read it
-% for as long as they like; a button press on SPACEBAR to dismiss the
-% instruction; a blank screen of 100ms duration; a fixation cross of 100ms
-% duration; a blank of 100ms duration; a screen with the pair presented
-% (currently for unlimited duration terminated by a key press from a
-% restricted set (left cmd, right cmd) ); 1000ms ISI.
-%
-% Participant id > 100 is used for testing and can be overwritten
-
-SUBJECT=1000;
-
-howLongToWaitForAResponse = 60; % seconds. Set low for test runs
-testing = 1; % 0 for real runs
-
-% Clear Matlab/Octave window:
-clc;
-
-% check for Opengl compatibility, abort otherwise:
-AssertOpenGL;
-
-% Reseed the random-number generator for each expt.
-rand('state',sum(100*clock));
-
-% Make sure keyboard mapping is the same on all supported operating systems
-% Apple MacOS/X, MS-Windows and GNU/Linux:
-KbName('UnifyKeyNames');
-
-%     % Turn off the annoying echo to the command line
-%     ListenChar(2);
-
-% Define allowable keys
-leftresp=KbName('LeftGUI'); %227
-rightresp=KbName('RightGUI'); %231
-spacebar=KbName('space'); %44
-
-
-%%%%%%%%%%%%%%%%%%%%%%
-% file handling
-%%%%%%%%%%%%%%%%%%%%%%
-
-
-%practice data output file to write to
-practicedatafilename = strcat('output/subject',...
-    num2str(SUBJECT),'Practice','.dat');
-
-% experimental stimuli to read from
-experimentalStimuliFilename = 'experimentalStimuli.txt';
-
-% experimental data output file to write to
[…]

### 2016-07-28 experiments/e2/experimentCode/exp2.m

@@ -1,0 +1,778 @@
+function exp3(SUBJECT)
+% This third study has the following structure:
+% Practice, Rest, Block1 Block 2 Block 3, Block 4
+% There are 256 trials
+% There are 64 unique trials
+% There are 4 Blocks
+% Each trial consists of 3 squares each with a number of dots.
+%
+% A trial consists of:
+% the instruction on screen while participants read it for as long as they like;
+% a button press on SPACEBAR to bring up
+% a screen with the instruction and the dots presented
+%
+% Participant id > 100 is used for testing and can be overwritten
+
+howLongToWaitForAResponse = 60; % seconds. Set low for test runs
+testing = 1; % 0 for real runs, 1 for testing
+takescreenshot=1;
+SUBJECT=1001;
+%takescreenshot=0;
+
+% Clear Matlab/Octave window:
+clc;
+
+% check for Opengl compatibility, abort otherwise:
+AssertOpenGL;
+
+% Reseed the random-number generator for each expt.
+rand('state',sum(100*clock));
+
+% Make sure keyboard mapping is the same on all supported operating systems
+% Apple MacOS/X, MS-Windows and GNU/Linux:
+KbName('UnifyKeyNames');
+
+% Define keys by ID
+leftresp=KbName('c'); % is 6
+middleresp=KbName('b'); % is 5
+rightresp=KbName('m'); % is 16
+escapekey=KbName('ESCAPE'); % is 41
+spacebar=KbName('space'); % is 44
+returnkey=KbName('return'); % is 40
+
+%%%%%%%%%%%%%%%%%%%%%%
+% define stimuli
+%%%%%%%%%%%%%%%%%%%%%%
+
+% order in headers gives the answer to "Which square had the small quantity"? and does not give the correct response
+
+%Condition 1 is numerical vague
+%Condition 2 is numerical precise
+%Condition 3 is visual vague
+%Condition 4 is visual precise
+%
+%Order 1 is small number on left
+%Order 2 is small number on right
+%
+%Quantity 1 is small
+%Quantity 2 is big
+
+
+exp3headers={
+	"Item","Condition","Order","Quantity","Left","Mid","Right","Instruction"};
+
+exp3practicestimuli = {
+	0,0,0,0,6,15,24,"Choose a square with about 10 dots";
+	0,0,0,0,34,25,16,"Choose the square with 16 dots";
[…]

### 2016-07-28 experiments/e2/experimentCode/exp2dummy.m

@@ -1,0 +1,586 @@
+function exp3dummy(SUBJECT)
+% This third study has the following structure:
+% Practice, Rest, Block1 Block 2 Block 3, Block 4
+% There are 256 trials
+% There are 64 unique trials
+% There are 4 Blocks
+% Each trial consists of 3 squares each with a number of dots.
+%
+% A trial consists of:
+% the instruction on screen while participants read it for as long as they like;
+% a button press on SPACEBAR to bring up
+% a screen with the instruction and the dots presented
+%
+% Participant id > 100 is used for testing and can be overwritten
+
+howLongToWaitForAResponse = 60; % seconds. Set low for test runs
+testing = 1; % 0 for real runs, 1 for testing
+takescreenshot=1;
+SUBJECT=1001;
+
+
+% Clear Matlab/Octave window:
+clc;
+
+% check for Opengl compatibility, abort otherwise:
+AssertOpenGL;
+
+% Reseed the random-number generator for each expt.
+rand('state',sum(100*clock));
+
+% Make sure keyboard mapping is the same on all supported operating systems
+% Apple MacOS/X, MS-Windows and GNU/Linux:
+KbName('UnifyKeyNames');
+
+% Define keys by ID
+leftresp=KbName('c'); % is 6
+middleresp=KbName('b'); % is 5
+rightresp=KbName('m'); % is 16
+escapekey=KbName('ESCAPE'); % is 41
+spacebar=KbName('space'); % is 44
+returnkey=KbName('return'); % is 40
+
+%%%%%%%%%%%%%%%%%%%%%%
+% define stimuli
+%%%%%%%%%%%%%%%%%%%%%%
+
+% order in headers gives the answer to "Which square had the small quantity"? and does not give the correct response
+
+%Condition 1 is numerical vague
+%Condition 2 is numerical precise
+%Condition 3 is visual vague
+%Condition 4 is visual precise
+%
+%Order 1 is small number on left
+%Order 2 is small number on right
+%
+%Quantity 1 is small
+%Quantity 2 is big
+
+
+exp3headers={
+	"Item","Condition","Order","Quantity","Left","Mid","Right","Instruction"};
+
+exp3practicestimuli = {
+	0,0,0,0,6,15,24,"Choose a square with about 10 dots";
+	0,0,0,0,34,25,16,"Choose the square with 16 dots";
[…]

### 2016-07-28 experiments/e2/experimentCode/exp2fakescreenshots.m

@@ -1,0 +1,816 @@
+function exp3fakescreenshots(SUBJECT)
+% This third study has the following structure:
+% Practice, Rest, Block1 Block 2 Block 3, Block 4
+% There are 256 trials
+% There are 64 unique trials
+% There are 4 Blocks
+% Each trial consists of 3 squares each with a number of dots.
+%
+% A trial consists of:
+% the instruction on screen while participants read it for as long as they like;
+% a button press on SPACEBAR to bring up
+% a screen with the instruction and the dots presented
+%
+% Participant id > 100 is used for testing and can be overwritten
+
+howLongToWaitForAResponse = 60; % seconds. Set low for test runs
+testing = 1; % 0 for real runs, 1 for testing
+takescreenshot=1;
+SUBJECT=1001;
+
+
+% Clear Matlab/Octave window:
+clc;
+
+% check for Opengl compatibility, abort otherwise:
+AssertOpenGL;
+
+% Reseed the random-number generator for each expt.
+rand('state',sum(100*clock));
+
+% Make sure keyboard mapping is the same on all supported operating systems
+% Apple MacOS/X, MS-Windows and GNU/Linux:
+KbName('UnifyKeyNames');
+
+% Define keys by ID
+leftresp=KbName('c'); % is 6
+middleresp=KbName('b'); % is 5
+rightresp=KbName('m'); % is 16
+escapekey=KbName('ESCAPE'); % is 41
+spacebar=KbName('space'); % is 44
+returnkey=KbName('return'); % is 40
+
+%%%%%%%%%%%%%%%%%%%%%%
+% define stimuli
+%%%%%%%%%%%%%%%%%%%%%%
+
+% order in headers gives the answer to "Which square had the small quantity"? and does not give the correct response
+
+%Condition 1 is numerical vague
+%Condition 2 is numerical precise
+%Condition 3 is visual vague
+%Condition 4 is visual precise
+%
+%Order 1 is small number on left
+%Order 2 is small number on right
+%
+%Quantity 1 is small
+%Quantity 2 is big
+
+
+exp3headers={
+	"Item","Condition","Order","Quantity","Left","Mid","Right","Instruction"};
+
+exp3practicestimuli = {
+	0,0,0,0,6,15,24,"Choose a square with about 10 dots";
+	0,0,0,0,34,25,16,"Choose the square with 16 dots";
[…]

### 2016-07-28 experiments/e2/experimentCode/exp2test.m

@@ -1,0 +1,53 @@
+function exp3test(SUBJECT)
+SUBJECT=1001;
+
+% Clear Matlab/Octave window:
+clc;
+
+% check for Opengl compatibility, abort otherwise:
+AssertOpenGL;
+
+screenNumber = 0;
+
+% Returns as default the mean gray value of screen:
+gray=GrayIndex(screenNumber);
+
+% [windowPtr,rect]=SCREEN(windowPtrOrScreenNumber,'OpenWindow',[color],[rect],[pixelSize]);
+% [w, rect] = Screen(screenNumber, 'OpenWindow', gray, [1,1,801,601],[]);
+[w, rect] = Screen(screenNumber, 'OpenWindow', gray, [],[]);
+    
+% Set text size
+Screen('TextSize', w, 32);
+
+try
+
+instructionText = sprintf(strcat...
+         ('You will see 3 squares with dots in them, and an instruction\n',...
+         'to choose one of them, as in the picture above.\n\n',...
+         'First you see the instruction, then press space to see the squares.\n\n',...
+         'Use the keys C, B, or M to choose the square that you think matches\n',...
+         'the instruction best. There will be some practice trials so you\n',...
+         'can get used to the routine.'));
+         
+instructionY = 200;
+instructionX = 100;
+
+% Text of introductory instructions to back buffer
+DrawFormattedText(w, instructionText, instructionX, instructionY, WhiteIndex(w));
+
+% Flip introductory instructions to screen
+Screen('Flip', w);
+
+% Wait for key press
+KbWait();
+
+% Go to grey screen
+Screen(w,'FillRect',gray);
+Screen('Flip', w);
+Screen('CloseAll');
+
+catch
+
+Screen('CloseAll');
+
+end_try_catch

### 2016-07-28 experiments/e2/writeup/functions/ci-functions.R

@@ -1,0 +1,304 @@
+library(gmodels)
+library(mnormt)
+
+lm.ci <- function(data.frame, conf.level = 0.95, difference = FALSE) {
+  #loftus-masson within-subject CIs
+  k = ncol(data.frame)
+  n <- nrow(data.frame)
+  df.stack <- stack(data.frame)
+  require(nlme)
+  parts <- rep(1:n, k)
+  root.ms.error <- lme(values ~ 0 + ind, random = ~1 | parts, cbind(parts, 
+                                                                    df.stack))[[6]]
+  detach(package:nlme)
+  mean.mat <- matrix(, k, 1)
+  ci.mat <- matrix(, k, 2)
+  if (difference == TRUE) 
+    diff.factor = 2^0.5/2
+  else diff.factor = 1
+  moe <- root.ms.error/n^0.5 * qt(1 - (1 - conf.level)/2, (n - 1) * (k - 
+                                                                       1)) * diff.factor
+  for (i in 1:k) mean.mat[i, ] <- mean(data.frame[,i])
+  for (i in 1:k) {
+    ci.mat[i, 1] <- mean.mat[i] - moe
+    ci.mat[i, 2] <- mean.mat[i] + moe
+  }
+  dimnames(ci.mat) <- list(names(data.frame), c("lower", "upper"))
+  ci.mat
+}
+
+bs.ci <- function(data.frame, conf.level = 0.95, difference = FALSE) {
+  # between-subject CIs
+  k = ncol(data.frame)
+  n <- nrow(data.frame)
+  df.stack <- stack(data.frame)
+  group.means <- mean(as.matrix(data.frame), na.rm = TRUE)
+  if (difference == TRUE) 
+    ci.mat <- (confint(lm(values ~ 0 + ind, df.stack)) - group.means) * 
+    2^0.5/2 + group.means
+  else ci.mat <- confint(lm(values ~ 0 + ind, df.stack))
+  dimnames(ci.mat) <- list(names(data.frame), c("lower", "upper"))
+  ci.mat
+}
+
+cm.ci.u <- function(data.frame, conf.level = 0.95) {
+  #cousineau uncorrected within-subject CIs
+  k = ncol(data.frame)
+  n <- nrow(data.frame)
+  df.stack <- stack(data.frame)
+  index <- rep(1:n, k)
+  p.means <- tapply(df.stack$values, index, mean)
+  norm.df <- data.frame - p.means + (sum(data.frame)/(n * k))
+  output <- matrix(, k, 2)
[…]

### 2016-07-28 experiments/e2/writeup/functions/classifyResponse.R

@@ -1,0 +1,188 @@
+classifyResponse = function(dat) {
+  
+  # what were they expected to respond?
+  
+  dat$crossed = as.factor(paste('Con', dat$Condition, ':Quan', dat$Quantity, ':Item', dat$Item, sep=''))
+  
+  dat[dat$crossed=='Con1:Quan1:Item1', 'Exp_Num'] <-  6 
+  dat[dat$crossed=='Con1:Quan1:Item1', 'Bline_Num']     <- 15  
+  dat[dat$crossed=='Con1:Quan1:Item1', 'Extr_Num']      <- 24
+  
+  dat[dat$crossed=='Con1:Quan1:Item2', 'Exp_Num'] <- 16
+  dat[dat$crossed=='Con1:Quan1:Item2', 'Bline_Num']     <- 25  
+  dat[dat$crossed=='Con1:Quan1:Item2', 'Extr_Num']      <- 34
+  
+  dat[dat$crossed=='Con1:Quan1:Item3', 'Exp_Num'] <- 26
+  dat[dat$crossed=='Con1:Quan1:Item3', 'Bline_Num']     <- 35
+  dat[dat$crossed=='Con1:Quan1:Item3', 'Extr_Num']      <- 44
+  
+  dat[dat$crossed=='Con1:Quan1:Item4', 'Exp_Num'] <- 36
+  dat[dat$crossed=='Con1:Quan1:Item4', 'Bline_Num']     <- 45
+  dat[dat$crossed=='Con1:Quan1:Item4', 'Extr_Num']      <- 54
+  
+  dat[dat$crossed=='Con1:Quan2:Item1', 'Exp_Num'] <- 24
+  dat[dat$crossed=='Con1:Quan2:Item1', 'Bline_Num']     <- 15 
+  dat[dat$crossed=='Con1:Quan2:Item1', 'Extr_Num']      <-  6
+  
+  dat[dat$crossed=='Con1:Quan2:Item2', 'Exp_Num'] <- 34
+  dat[dat$crossed=='Con1:Quan2:Item2', 'Bline_Num']     <- 25  
+  dat[dat$crossed=='Con1:Quan2:Item2', 'Extr_Num']      <- 16
+  
+  dat[dat$crossed=='Con1:Quan2:Item3', 'Exp_Num'] <- 44
+  dat[dat$crossed=='Con1:Quan2:Item3', 'Bline_Num']     <- 35
+  dat[dat$crossed=='Con1:Quan2:Item3', 'Extr_Num']      <- 26
+  
[…]

### 2016-07-28 experiments/e2/writeup/functions/gatherData.R

@@ -1,0 +1,24 @@
+# R version of gatherData.sh
+
+gatherData = function(number_of_valid_subjects) {
+  # message("gathering data ... ")
+  # number_of_valid_subjects <- 30
+  data_dir <- '../experimentCode/output/'
+  
+  column.headers.df <-  head( read.table(
+    paste(data_dir,'subject1.data',sep=''),
+    header=TRUE),0)
+  
+  gathered.data <- column.headers.df
+  
+  for (subject in 1:number_of_valid_subjects) {
+    current.filename <- paste(data_dir, 'subject', subject, '.data', sep='')
+    current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = FALSE)
+    gathered.data <- rbind(gathered.data, current.data)
+  }
+  
+  # message("returning gathered data")
+  
+  return(gathered.data)
+  
+}

### 2016-07-28 experiments/e2/writeup/functions/merFunctions.R

@@ -1,0 +1,1 @@
+r. <- function (formula, ...) rstandard(lm(formula, ...))

### 2016-07-28 experiments/e2/writeup/which subjects violate normality.R

@@ -1,0 +1,12 @@
+# which subjects violate normality
+f = function(dfr) return(shapiro.test(dfr$logRT)$p.value)
+p = as.vector(by(dat, dat$Subject, f))
+names(p) = levels(dat$Subject)
+names(p[p < 0.05])
+length(names(p[p < 0.05]))
+qqmath(~logRT | Subject, data = dat,
+       prepanel = prepanel.qqmathline,
+       panel = function(x, ...) {
+         panel.qqmathline(x, ...)
+         panel.qqmath(x, ...)
+       })

### 2016-07-28 experiments/e3/writeup/experiment3.Rmd
added e3, but lots of its files and contents need renaming from exp4 to e3 to be consistent with the 2016 numbering

---
title: "Experiment 3 of 4"
output:
  html_document:
    toc: yes
---  
# Experiment 3

```{r, startClean}
#Start clean.
rm(list=ls())
```

```{r,docOpts}
#Knitr doc options.
library(knitr)
#knit_hooks$set(crop=hook_pdfcrop)
#opts_chunk$set(cache=TRUE, autodep=TRUE)
#options(tidy=TRUE)
#options(replace.assign=TRUE, width=40)
```

```{r, loadLibraries}
# Load libraries.
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(reshape))
suppressPackageStartupMessages(library(lattice))
suppressPackageStartupMessages(library(LMERConvenienceFunctions))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(lmerTest))
```

# Experiment Structure
Each subject saw 192 trials, being 3 cycles of 64 trials. There were 38 participants. There were 7296 rows in the full data set.

# Trial Procedure
The procedure for each trial was as follows. First an instruction that was a referring expression was presented until participants dismissed it with a key press. Then the squares with dots were presented until the participant responded with a key press, with a 60 second timeout if they did not respond.

# Prepare the data
Source additional R files with functions for computing confidence intervals.

```{r,sourceFunctions}
source("functions/CIfunctions.R")
```

Get data.

```{r,readdata}
dat <- read.delim("data/allresults.txt", stringsAsFactors = FALSE)

# declare local variables
number_of_valid_subjects <- 38
number_of_rows <- 7296
number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # 192

# sort out borderline responses into expected near and far
# How many squares in the square they chose? 
dat$RESPONSE <- as.character(dat$RESPONSE)
for (row in 1 : nrow(dat) ) {
  switch(dat[row,'RESPONSE'],
         'LEFT' = {dat[row, 'choice'] <- dat[row, 'Left']},
         'MIDDLE' = {dat[row, 'choice'] <- dat[row, 'Mid']},
         'RIGHT' = {dat[row, 'choice'] <- dat[row, 'Right']}         
  )
}
dat$crossed=paste('Con',dat$Condition,':Quan',dat$Quantity,':Item',dat$Item,sep="")

dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseExpected'] <-  6 
dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseNear']     <- 15  
dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseFar']      <- 24

dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseExpected'] <- 16
dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseFar']      <- 34

dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseExpected'] <- 26
dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseFar']      <- 44

dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseExpected'] <- 36
dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseFar']      <- 54

dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseExpected'] <- 24
dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseNear']     <- 15 
dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseFar']      <-  6

dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseExpected'] <- 34
dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseFar']      <- 16

dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseExpected'] <- 44
dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseFar']      <- 26

dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseExpected'] <- 54
dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseFar']      <- 36

dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseExpected'] <-  6 
dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseNear']     <- 15  
dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseFar']      <- 24

dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseExpected'] <- 16
dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseFar']      <- 34

dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseExpected'] <- 26
dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseFar']      <- 44

dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseExpected'] <- 36
dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseFar']      <- 54

dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseExpected'] <- 24
dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseNear']     <- 15 
dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseFar']      <-  6

dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseExpected'] <- 34
dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseFar']      <- 16

dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseExpected'] <- 44
dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseFar']      <- 26

dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseExpected'] <- 54
dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseFar']      <- 36

dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseExpected'] <-  6 
dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseNear']     <- 15  
dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseFar']      <- 24

dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseExpected'] <- 16
dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseFar']      <- 34

dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseExpected'] <- 26
dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseFar']      <- 44

dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseExpected'] <- 36
dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseFar']      <- 54

dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseExpected'] <- 24
dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseNear']     <- 15 
dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseFar']      <-  6

dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseExpected'] <- 34
dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseFar']      <- 16

dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseExpected'] <- 44
dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseFar']      <- 26

dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseExpected'] <- 54
dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseFar']      <- 36

dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseExpected'] <-  6 
dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseNear']     <- 15  
dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseFar']      <- 24

dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseExpected'] <- 16
dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseFar']      <- 34

dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseExpected'] <- 26
dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseFar']      <- 44

dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseExpected'] <- 36
dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseFar']      <- 54

dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseExpected'] <- 24
dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseNear']     <- 15 
dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseFar']      <-  6

dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseExpected'] <- 34
dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseFar']      <- 16

dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseExpected'] <- 44
dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseFar']      <- 26

dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseExpected'] <- 54
dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseFar']      <- 36

dat$isResponseExpected <- dat$choice == dat$ResponseExpected
dat$isResponseNear <- dat$choice == dat$ResponseNear
dat$isResponseFar <- dat$choice == dat$ResponseFar

for ( row in 1:nrow(dat) ) {
  dat[row, 'switchResponse'] <- 
    ifelse(dat[row, 'isResponseExpected']==TRUE, 'Expected', 
           ifelse(dat[row, 'isResponseNear']==TRUE, 'Near', 'Far') ) 
}

## treat variables

# SUBJECT
# ensure Subject is a factor 
dat$Subject=factor(paste("s",sprintf("%02d",dat$Subject),sep=""))

# TRIAL
# Trial for subject, 1 to 192
dat$Trial = rep(x = 1:number_of_trials_per_subject, times = number_of_valid_subjects)

# make a centred Trial for modeling
dat$c_Trl <-dat$Trial - mean(dat$Trial)

# make a scaled Trial for modelling
dat$s_Trl <- as.numeric(scale(dat$Trial))

# ID
# id is a unique identifier for the 7680 row data
dat$id <- factor(paste(paste(dat$Subject), 
                       paste("t", sprintf("%03d", dat$Trial), sep="") , sep=":"))

# ITEM
# create a centred numeric item variable for modeling
dat$c_Itm <- ifelse(dat$Item==1, -.75, 
                    ifelse(dat$Item==2, -.25,
                           ifelse(dat$Item==3, .25, .75)))

# make Item be a factor and assign labels
dat$Item <- factor(dat$Item, levels=c(1,2,3,4), labels=c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))

# add ratios for Item
item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54) 
# 0.2500000 0.4705882 0.5909091 0.6666667
item_range_ratio_scaled = as.vector(scale(c(6 / 24, 16 / 34, 26 / 44, 36 / 54)))
# -1.3441995 -0.1316642  0.5297187  0.9461450
item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))) 
# 0.5125000 0.6876471 0.7691558 0.8166667
item_mean_ratio_scaled = as.vector(scale(c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))))) 
# -1.37582241 -0.06614191  0.54334858 0.89861574

dat[dat$Item == "06:15:24", "item_range_ratio"] <-  0.2500000
dat[dat$Item == "16:25:34", "item_range_ratio"] <-  0.4705882
dat[dat$Item == "26:35:44", "item_range_ratio"] <-  0.5909091
dat[dat$Item == "36:45:54", "item_range_ratio"] <-  0.6666667

dat[dat$Item == "06:15:24", "item_range_ratio_scaled"] <-  -1.3441995
dat[dat$Item == "16:25:34", "item_range_ratio_scaled"] <-  -0.1316642
dat[dat$Item == "26:35:44", "item_range_ratio_scaled"] <-   0.5297187  
dat[dat$Item == "36:45:54", "item_range_ratio_scaled"] <-   0.9461450

dat[dat$Item == "06:15:24", "item_mean_ratio"] <-  0.5125000 
dat[dat$Item == "16:25:34", "item_mean_ratio"] <-  0.6876471 
dat[dat$Item == "26:35:44", "item_mean_ratio"] <-  0.7691558
dat[dat$Item == "36:45:54", "item_mean_ratio"] <-  0.8166667

dat[dat$Item == "06:15:24", "item_mean_ratio_scaled"] <-   -1.37582241 
dat[dat$Item == "16:25:34", "item_mean_ratio_scaled"] <-   -0.06614191 
dat[dat$Item == "26:35:44", "item_mean_ratio_scaled"] <-    0.54334858 
dat[dat$Item == "36:45:54", "item_mean_ratio_scaled"] <-    0.89861574

# VAGUENESS
# Create a factor coding for Vagueness
dat[ dat$Condition==1 , 'Vagueness'] <- 'Vague'
dat[ dat$Condition==2 , 'Vagueness'] <- 'Crisp'
dat[ dat$Condition==3 , 'Vagueness'] <- 'Vague'
dat[ dat$Condition==4 , 'Vagueness'] <- 'Crisp'
dat$Vagueness <- as.factor(dat$Vagueness)

# manually center Vagueness
dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)

dat$Selection = ""
dat[dat$Condition %in% c(1,2), "Selection"] <- "matching"
dat[dat$Condition %in% c(3,4), "Selection"] <- "comparison"
dat$Selection.sum <- ifelse(dat$Selection=="matching", .5, -.5)
dat$Selection <- as.factor(dat$Selection)

# remove one zero RT by replacing with NA
dat$RT[dat$RT==0]<-NA

# ORDER
# give the levels of Order meaningful names
dat$Order <- factor(dat$Order, levels=c(1,2), labels=c('LtoR','RtoL'))

# make a manually centred Order
dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)

# QUANTITY
# give the levels of Quantity meaningful names
dat$Quantity <- factor(dat$Quantity, levels=c(1,2), labels=c('Small','Large'))

# make a manually centred Quantity
dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)

# INSTRUCTION
# add number of characters in the instruction # 29 30 34 36 38
dat$nchar_instr = nchar(as.character(dat$Instruction))
dat$nchar_instr_scaled = as.vector(scale(nchar(as.character(dat$Instruction)), scale=TRUE))

# make Instruction be a factor (17 levels)
dat$Instruction <- as.factor(dat$Instruction)

# RT
# add transformations of RT
dat$RT_RecSqd    <- 1/dat$RT^2
dat$RT_Rec       <- 1/dat$RT
dat$RT_RecSqt    <- 1/sqrt(dat$RT) 
dat$RT_log       <- log(dat$RT)
dat$RT_sqt       <- sqrt(dat$RT)
dat$RT_raw       <- dat$RT
dat$RT_sqd       <- dat$RT^2

# # print to file a table with information about the design
# design_info <- unique(subset(dat, select=c(Item, Condition, Vagueness, Selection, Quantity, Order, 
#                                            Left, Mid, Right, Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side,Instruction)))
# design.info <- design_info[order(design_info$Item, design_info$Condition, design_info$Quantity, design_info$Order),]
# row.names(design_info) <- NULL
# capture.output(  print.data.frame(design_info, row.names=F, print.gap=3, quote=F, right=F), file="design_info.table")

# # put dat in better column order
# dat <- subset(dat, 
#               select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Selection, 
#                          Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,
#                          c_Trl, s_Trl, c_Itm, c_Vag, Selection.sum, f_Cnd, c_Ord, c_Qty,
#                          RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
#                          Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
#                          Instruction, nchar_instr
#               ))

# save dat as dat.Rda
# This data set contains *all* trials including impossible trials and is mainly for graphs comparing different removals
save(dat, file="dat.Rda")

 # lose impossible trials
  dd <- dat
  dd$RT[dd$RT == 1 ] <- NA
  dd$RT[dd$RT > 40000 ] <- NA
  dd <- dd[complete.cases(dd),]
  row.names(dd) <- NULL

# add preceding RT: because we removed impossible trials, the value for preceding RT for a trial following an impossible trial is the value of the trial that preceded the impossible trial.
dd$RTprev <- NA
for (s in levels(dd$Subject)) {
  nrows = nrow(dd[dd$Subject==s,])
  for (i in 1:nrows) {
    if (i==1) {
      dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]
    }
    else
      dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1]
  }
}

# add transformations of previous RT
dd$RTprev_RecSqd    <- 1/dd$RTprev^2
dd$RTprev_Rec       <- 1/dd$RTprev
dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev) 
dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
dd$RTprev_log       <- log(dd$RTprev)
dd$RTprev_sqt       <- sqrt(dd$RTprev)
dd$RTprev_raw       <- dd$RTprev
dd$RTprev_sqd       <- dd$RTprev^2

save(dd, file="dd.Rda")

```

Check the distribution of the trimmed data to see if a transformation is suggested. Log transformation seems to be best so we adopt that for the rest of the analyses.

```{r}
# boxplot subjects and items

load("dd.Rda")
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(ggplot2))

# allow reference to transformation
subdata = melt(dd,
               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
               variable.name="transformation",
               value.name="value")

# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ variable, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
  xlab("") + ylab("")
```

```{r}
load("dd.Rda")
suppressPackageStartupMessages(require(lme4))
suppressPackageStartupMessages(require(lmerTest))
suppressPackageStartupMessages(require(languageR))
```

```{r}
v5 <- lmer(data=dd, 
           RT_RecSqt ~ 
             c_Vag + Selection.sum + c_Qty + c_Ord + 
             Selection.sum:c_Vag:c_Qty +
             item_mean_ratio + 
             s_Trl +
             RTprev_RecSqt +
             nchar_instr +
             (1+c_Vag + Selection.sum + c_Qty + c_Ord|Subject)
)
```

```{r}
summary(v5)
```

```{r}
cat("R^2")
cor(fitted(v5), dd$RT_RecSqt)^2
```

```{r}
par(mfrow=c(2,5))
plotLMER.fnc(v5)
```

Plot model coefficients and ci's

```{r}
par(las=2, mar=c(14,6,1,1))
y=as.data.frame(summary(v5)$coef[-1,1]) # estimates, without intercept
q=confint(v5)[-c(1,2,3),]
y=cbind(y,q)
plot(y[,1], ylim=range(y[,1],y[,2],y[,3]), xaxt="n", xlab="", ylab="estimate\n",  pch='-', cex=2)
segments(x0=1:nrow(y), x1=1:nrow(y), y[,2], y[,3])
abline(h=0)
axis(1, labels=row.names(y), at=1:nrow(y), srt=45)
grid()
```

Model criticism baayen plots

```{r, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

par(mfrow=c(1,4), pty='s')
# create scaled residuals
v5$rstand = as.vector(scale(resid(v5)))
# plot scaled residuals density
plot(density(v5$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(v5$rstand, cex=.5)
qqline(v5$rstand)
# plot standardised residuals versus fitted values
plot(v5$rstand ~ fitted(v5), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))
# create dffits
dffits=abs(resid(v5, "dffits"))
# plot dffits
plot(dffits, type='h')
```

### 2016-07-28 experiments/e3/experimentCode/e4strings.m

@@ -1,0 +1,11 @@
+    %%%%%%%%%%%%%%%%%%%%%%%
+    % define text strings %
+    %%%%%%%%%%%%%%%%%%%%%%%
+
+    instructionText = sprintf(strcat...
+         ('Press Enter to start the practice trials.'));
+
+    oopsText = sprintf(strcat(...
+        'Ooops... Something went wrong.\n\n',...
+        'Please tell the experimenter.'));
+

### 2016-07-28 experiments/e3/writeup/acquire/acquire.R

@@ -1,0 +1,25 @@
+rm(list=ls())
+x <- read.delim("allresults.txt")
+x[x$Condition %in% c(1,3),"vagueness"] <- "vague"
+x[x$Condition %in% c(2,4),"vagueness"] <- "crisp"
+x$vagueness <- factor(x$vagueness)
+contrasts(x$vagueness) <- contr.sum(2)
+x[x$Condition %in% c(1,2),"selection"] <- "matching"
+x[x$Condition %in% c(3,4),"selection"] <- "comparison"
+x$selection <- factor(x$selection)
+contrasts(x$selection) <- contr.sum(2)
+x <- subset(x,subset=RT>0&RT<49000)
+x$cRT <- scale(x$RT, scale=F)
+x$logRT <- log(x$RT)
+x$cLogRT <- scale(x$logRT,scale=F)
+x$Item <- as.factor(x$Item)
+x$Subject <- as.factor(x$Subject)
+x$Effect <- x$selection:x$vagueness
+x$TARGET<-"TARGET"
+x[x$Order==1&x$Quantity==1,"TARGET"] <- "LEFT"
+x[x$Order==1&x$Quantity==2,"TARGET"] <- "RIGHT"
+x[x$Order==2&x$Quantity==1,"TARGET"] <- "RIGHT"
+x[x$Order==2&x$Quantity==2,"TARGET"] <- "LEFT"
+x$EXTREME <- ifelse(x$TARGET==x$RESPONSE,1,0)
+x <- subset(x, select=c(Subject,Item,vagueness,selection, Effect,RT,logRT,EXTREME))
+save(x, file="e4vagueness.RData")

### 2016-07-28 experiments/e3/archive/scratch3.Rnw
tidying

<<getData, echo=TRUE>>=
  x=read.table("../data/data.txt")
  x$rt <- x$RT; x$RT <- NULL # for consistency
  x$logRT <- log(x$rt)
  x$recipRT <- -1000/x$rt
  # remove rt of 0
  (before <- nrow(x))
  x <- subset(x, rt>0)
  (after <- nrow(x))
  before - after
  @

    Set factors etc

  <<setFactors>>=
    x$vagueness <- factor(x$vagueness)
  contrasts(x$vagueness) <- contr.sum(2)/2
  x$selection <- factor(x$selection)
  contrasts(x$selection) <- contr.sum(2)/2
  x$cItm <- scale(x$Item)
  x$Item <- as.factor(x$Item)
  x$Subject <- as.factor(x$Subject)
  @

    Check contrasts

  <<checkContrasts>>=
    contrasts(x$vagueness)
  contrasts(x$selection)
  unique(x$cItm)
  @

    For the untrimmed data, show effects of transforming rt.

  \begin{marginfigure}
  <<plotDistributions, fig.width=4, fig.height=2, out.width="\\linewidth", echo=FALSE, fig.pos="Hhtbp", crop=TRUE>>=
    xs=melt(x,measure.vars=c("rt","logRT","recipRT"))
  densityplot(~value|variable,data=xs, ylab=NULL, xlab=NULL, yaxt=F, scales=list(x=list(relation="free",rot=45), y=list(relation="free",at=NULL)), cex=.2)
  @
    \caption{Untrimmed data}
  \end{marginfigure}

  Trim the data.

  <<trimData>>=
    data.trim <- perSubjectTrim.fnc(x, response="rt", subject="Subject", trim=2.5)
  @

    Check the distribution of the trimmed data to see if a transformation is suggested.

  \begin{marginfigure}
  <<checkTrim, fig.width=4, fig.height=2, out.width="\\linewidth", fig.pos="Hhtbp", echo=FALSE, crop=TRUE>>=
    xs=melt(data.trim$data,  measure.vars=c("rt","logRT","recipRT"))
  densityplot(~value|variable, data=xs, ylab=NULL, xlab=NULL, yaxt=F, scales=list(x=list(relation="free", rot=45), y=list(relation="free", at=NULL)), cex=.2)
  @
    \caption{Trimmed data}
  \end{marginfigure}

  Select either trimmed or untrimmed data.

  <<selectTrim>>=
    dat <- data.trim$data
  @

    Show means.

  \begin{figure}[htbp]
  <<plotMeans, echo=FALSE, fig.width=5, fig.height=3, out.width='.5\\linewidth', crop=TRUE>>=
    a=summarySEwithin(dat, measurevar="logRT", withinvars=c("vagueness","selection"), idvar="Subject", na.rm=TRUE)
  a$mygroup=paste(a$selection,a$vagueness)
  pd <- position_dodge(.05)
  ggplot(a, aes(x=vagueness, y=logRT, group=selection, col=mygroup, shape=mygroup, fill=mygroup, ymin=logRT-ci, ymax=logRT+ci)) + geom_errorbar(width=.1, position=pd) +  geom_line(position=pd) +  geom_point(position=pd,size=5) +  scale_colour_manual(values = c("black","black", "black", "black")) +   scale_shape_manual(values = c(21, 23, 21, 23)) + scale_fill_manual(values = c("black", "black", "white", "white")) + xlab("Vagueness") + theme(legend.title=element_blank())
  @
    \caption{Means for vagueness and selection}
  \end{figure}

  Edinburgh slides plot.

  \begin{figure}[htbp]
  <<plotMeansEdin, echo=FALSE, fig.width=9, fig.height=5>>=
    a=summarySEwithin(dat, measurevar="logRT", withinvars=c("vagueness","selection"), idvar="Subject", na.rm=TRUE)
  a$vagueness <- relevel(a$vagueness,ref="crisp")
  a$selection <- relevel(a$selection,ref="matching")
  a$mygroup=a$selection:a$vagueness
  a$mygroup=a$selection
  ggplot(a, aes(x=vagueness, y=logRT, group=selection, col=selection, ymin=logRT-ci, ymax=logRT+ci)) + 
    geom_errorbar(width=.1,lwd=1.1) +  
    geom_line(lwd=1.1) +  
    geom_point(size=5) +  
    scale_colour_manual(values = c("red","blue","red","blue")) +  
    #   scale_shape_manual(values = c(21, 23, 21, 23)) + 
    scale_fill_manual(values = c("blue", "blue", "white", "white")) +
    theme(legend.title=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_text(size=18,angle=0),
          axis.text.x=element_text(size=18),
          axis.text.y=element_text(size=18),
          panel.background=element_rect(),
          legend.text=element_text(size=15) ,
          legend.key.size=unit(1.5,"cm")) +
    scale_y_continuous(limits = c(7.05,7.505))
  @
    \caption{Means for vagueness and selection}
  \end{figure}

  Break the means down over items in response to significant 3-way interaction in the lm.

  \begin{figure}[htbp]
  <<PlotOfConditionMeansBrokenOverItems, fig.width=9, fig.height=2.9, out.width='\\linewidth',crop=TRUE, echo=FALSE>>=
    d2 <- summarySEwithin(dat, measurevar="logRT",withinvars=c("vagueness","selection", "Item"),idvar="Subject",na.rm=TRUE)
  d2$mygroup=paste(d2$selection,d2$vagueness)
  d2$Item <- paste("Item:",d2$Item,sep="")
  pd <- position_dodge(.1)
  ggplot(d2, aes(x=vagueness, y=logRT_norm, group=selection, col=selection, shape=selection, fill=selection,  ymin=logRT_norm-ci, ymax=logRT_norm+ci)) + geom_errorbar(width=.1, position=pd) + facet_grid(~Item) +  geom_line(position=pd) +  geom_point(position=pd, size=3) +  scale_colour_manual(values = c("black","black")) +   scale_shape_manual(values = c(21, 21)) + scale_fill_manual(values = c("black", "white")) + guides(col = guide_legend(reverse = TRUE, title="Task"), fill=guide_legend(reverse = TRUE, title="Task"), shape=guide_legend(reverse = TRUE, title="Task"))
  @
    \caption{Condition means broken down over items}
  \end{figure}

  Specify a lm model.

  <<lm1>>=
    lm1 <- lm(data=dat, logRT~vagueness*selection*cItm)
  @

    <<lm1-table, results='asis', echo=FALSE>>=
    print(xtable(summary(lm1),caption="lm model"), table.placement="htbp")
  @

    Specify a random-intercepts-only model.

  <<rimodel>>=
    lmer.rt.ri <- lmer(data=dat, formula=logRT~vagueness*selection*cItm+ (1|Subject)) 
  @

    Specify a maximal model.

  <<mxmodel, cache=TRUE>>=
    # this doesn't converge
    lmer.rt.mx <- lmer(data=dat, formula=logRT~vagueness*selection*cItm+ (vagueness*selection*cItm|Subject)) 
  @

    Drop highest-order interaction random slope for effects involving cItm.

  <<mmmodel0, cache=TRUE>>=
    # this doesn't converge
    lmer.rt.mm0 <- lmer(data=dat, formula=logRT~vagueness*selection*cItm+ (vagueness*selection+cItm|Subject)) 
  @

    Drop highest-order interaction random slope for effects involving selection.

  <<mmmodel1, cache=TRUE>>=
    # this converges
    lmer.rt.mm1 <- lmer(data=dat, formula=logRT~vagueness*selection*cItm+ (vagueness+selection+cItm|Subject)) 
  @

    <<modelTables, results='asis', echo=FALSE>>=
    print(xtable(summary(lmer.rt.mm1)$coefficients, digits=3))
  @

    Check the residuals of the most maximal model that converges.

  \begin{marginfigure}
  <<qq, fig.width=5, fig.height=5, crop=TRUE, fig.pos='htbp', out.width="\\linewidth", echo=FALSE>>=
    qqPlot(residuals(lmer.rt.mm1), cex=.2)
  @
    \caption{Model residuals for the most maximal model that converges}
  \end{marginfigure}

  Vagueness appears to be beneficial in the comparison condition. Do a focussed comparison.

  <<datc>>=
    datc.lmer <- lmer(data=dat, subset=selection=="comparison", formula=logRT~vagueness*cItm+ (vagueness+cItm|Subject))
  @

    <<datc-table, results='asis', echo=FALSE>>=
    print(xtable(summary(datc.lmer)$coef,caption="Restricted to comparison condition"))
  @

    Vagueness appears to be detrimental in the matching condition. Do a focussed comparison.

  <<datm>>=
    datm.lmer <- lmer(data=dat, subset=selection=="matching", formula=logRT~vagueness*cItm+ (vagueness+cItm|Subject))
  @

    <<datm-table, results='asis', echo=FALSE>>=
    print(xtable(summary(datm.lmer)$coef,caption="Restricted to matching condition"))
  @

### 2016-07-28 experiments/e3/writeup/experiment3.Rmd
tidying

@@
   html_document:
-    toc: yes
+    toc: true
+    <!--toc_float: true-->
+    mathjax: default
+    template: default
+    code_folding: hide
+    theme: default
 ---  
@@
 
-```{r, startClean}
-#Start clean.
-rm(list=ls())
+```{r docOpts}
+library(knitr)
 ```
 
-```{r,docOpts}
-#Knitr doc options.
-library(knitr)
-#knit_hooks$set(crop=hook_pdfcrop)
-#opts_chunk$set(cache=TRUE, autodep=TRUE)
-#options(tidy=TRUE)
-#options(replace.assign=TRUE, width=40)
-```
-
-```{r, loadLibraries}
-# Load libraries.
+```{r loadLibraries}
 suppressPackageStartupMessages(library(MASS))
-suppressPackageStartupMessages(library(knitr))
 suppressPackageStartupMessages(library(ggplot2))
@@
 suppressPackageStartupMessages(library(lmerTest))
+suppressPackageStartupMessages(library(reshape2))
+suppressPackageStartupMessages(library(ggplot2))
+suppressPackageStartupMessages(require(languageR))
 ```
@@
 
-```{r,sourceFunctions}
+```{r sourceFunctions}
 source("functions/CIfunctions.R")
@@
 
-```{r,readdata}
+```{r readdata}
 dat <- read.delim("data/allresults.txt", stringsAsFactors = FALSE)
@@
 
-# # print to file a table with information about the design
-# design_info <- unique(subset(dat, select=c(Item, Condition, Vagueness, Selection, Quantity, Order, 
-#                                            Left, Mid, Right, Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side,Instruction)))
-# design.info <- design_info[order(design_info$Item, design_info$Condition, design_info$Quantity, design_info$Order),]
-# row.names(design_info) <- NULL
-# capture.output(  print.data.frame(design_info, row.names=F, print.gap=3, quote=F, right=F), file="design_info.table")
-
-# # put dat in better column order
-# dat <- subset(dat, 
-#               select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Selection, 
-#                          Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,
-#                          c_Trl, s_Trl, c_Itm, c_Vag, Selection.sum, f_Cnd, c_Ord, c_Qty,
-#                          RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
-#                          Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
-#                          Instruction, nchar_instr
-#               ))
-
 # save dat as dat.Rda
@@
 
+# save dd as dd.Rda
+# This data set non-impossible trials and renamed/recomputed variables
 save(dd, file="dd.Rda")
-
 ```
@@
 
-```{r}
-# boxplot subjects and items
-
+```{r boxplot subjects and items}
 load("dd.Rda")
-suppressPackageStartupMessages(library(reshape2))
-suppressPackageStartupMessages(library(ggplot2))
-
 # allow reference to transformation
@@
                value.name="value") 
-
 # boxplots for subjects and items separately for each dependent variable
@@
 
-```{r}
+```{r load_dd}
 load("dd.Rda")
-suppressPackageStartupMessages(require(lme4))
-suppressPackageStartupMessages(require(lmerTest))
-suppressPackageStartupMessages(require(languageR))
 ```
 
-```{r}
+```{r do_lmer}
 v5 <- lmer(data=dd, 
@@
 
-```{r}
+```{r summary_lmer}
 summary(v5)
@@
 
-```{r}
+```{r R2}
 cat("R^2")
@@
 
-```{r}
+```{r plotLMER}
 par(mfrow=c(2,5))
@@
 
-```{r}
+```{r coeff}
 par(las=2, mar=c(14,6,1,1))
@@
 
-```{r, fig.width=9, fig.height=3}
+```{r baayenplots, fig.width=9, fig.height=3}
 # Baayen 4-plot model criticism

### 2016-07-28 experiments/e0/writeup/vgn0.Rmd
tidying up.

@@
 
-This document aims to be a complete and up to date record of the analysis for vagueness 0, the pilot experiment}
-
-Start clean.
-
-```{r, startClean}
-rm(list=ls())
-```
+This document aims to be a complete and up to date record of the analysis for vagueness 0, the pilot experiment.
 
@@
 
-```{r, docOpts}
+```{r docOpts}
 library(knitr)
 knit_hooks$set(crop=hook_pdfcrop)
-opts_chunk$set(cache=TRUE, autodep=TRUE)
-#opts_chunk$set(size="footnotesize")
-#options(width=60)
 ```
@@
 
-```{r, loadLibraries}
+```{r loadLibraries}
 suppressPackageStartupMessages(library(MASS))
@@
 suppressPackageStartupMessages(library(gridExtra))
-library(grid)
+suppressPackageStartupMessages(library(grid))
 suppressPackageStartupMessages(library(lmerTest))
@@
 
-```{r, sourceFunctions}
-source("functions/CIfunctions.R")
+```{r sourceCIFunctions}
+source("CIfunctions.R")
 ```
@@
 # Preprocessing
+
 Get the data.
 
-```{r, getData}
+```{r getData}
+if( file.exists('data.txt') ) 
+	{
+	d <- read.table("data.txt", header=FALSE) 
+} else {
 	system("cat ../experimentCode/output/*.dat > data.txt", wait=TRUE)
-d <- read.table("data.txt", header=FALSE)
+}
 mynames <- list("gap", "target", "side", "vagueness", "pair", "Version", "Left",
@@
 names(d) <- mynames
-d$obs <- 1:dim(d)[1]
-d$trial=0
+d$obs <- 1:nrow(d)
+d$trial <- 0
 for (s in unique(d$subject)){
-  d[d$subject==s,"trial"] <- 1:nrow(subset(d,subject==s))
+  d[d$subject==s, "trial"] <- 1:nrow(subset(d,subject==s))
 }
@@
 d <- data.frame(
-  obs = d$obs, trial = d$trial, 
+  obs = d$obs, 
+  trial = d$trial, 
   subject = d$subject, 
@@
   vag = factor(d$vagueness, levels = c("Precise", "Vague"), labels = c("Crisp", "Vague")), 
-  side = d$side, err = 1 - d$isCorrect, rt = d$rt)
-d$subit <- rep("Not subitizable", times = nrow(d))
+  side = d$side, 
+  err = 1 - d$isCorrect, 
+  rt = d$rt)
+d$subit <- "Not subitizable"
 d$subit[d$item == "2:4"] <- "Subitizable"
@@
 
-```{r, removeErrorTrials}
+```{r removeErrorTrials}
 # by setting the RT to NA
@@
 
-```{r, postProcessData}
+```{r postProcessData}
 d$subject <- as.factor(d$subject)
@@
 
-```{r, checkContrasts}
+```{r checkContrasts}
 contrasts(d$vag)
@@
 
-```{r, addTransform}
+```{r addTransform}
 d$logRT <- log(d$rt)
@@
 
-```{r, plotDistributions, fig.width=4, fig.height=2}
-xs=melt(d,measure.vars=c("rt","logRT","recipRT"))
-densityplot(~value|variable, data=xs, ylab=NULL,xlab=NULL,yaxt=F, scales = list(x = list(relation="free", rot=45), y=list(relation="free",at=NULL)),layout=c(3,1),cex=.3)
+```{r plotDistributions, fig.width=4, fig.height=2}
+xs = melt(d,  
+					measure.vars = c("rt","logRT","recipRT"))
+densityplot(
+	~value|variable,
+	data=xs,
+	xlab=NULL,
+	yaxt=F,
+  scales=list(x=list(relation="free", rot=45), 
+   						y=list(relation="free", at=NULL)),
+	cex=.2,
+	plot.points=FALSE,
+	layout=c(3,1)
+	)
 ```
@@
 
-```{r, trimData}
+```{r trimData}
 dummy <- subset(d, !is.na(rt))
-data.trim <- perSubjectTrim.fnc(dummy, response="rt",subject="subject", trim=2.5)
+data.trim <- perSubjectTrim.fnc(dummy, response="rt",subject="subject", trim=2.5)$data
 ```
@@
 
-```{r, checkTrim, fig.width=4, fig.height=2}
-xs=melt(data.trim$data,  measure.vars=c("rt","logRT","recipRT"))
-densityplot(~value|variable,data=xs,
-            ylab=NULL,xlab=NULL,yaxt=F,
-            scales=list(x=list(relation="free",rot=45),y=list(relation="free",at=NULL)),cex=.2)
+```{r checkTrim, fig.width=4, fig.height=2}
+xs = melt(data.trim,  
+					measure.vars = c("rt","logRT","recipRT"))
+densityplot(
+	~value|variable,
+	data=xs,
+	xlab=NULL,
+	yaxt=F,
+  scales=list(x=list(relation="free", rot=45), 
+   						y=list(relation="free", at=NULL)),
+	cex=.2,
+	plot.points=FALSE,
+	layout=c(3,1)
+	)
 ```
@@
 
-```{r, selectTrim}
-dat <- data.trim$data
+```{r selectTrim}
+dat <- data.trim
 ```
@@
 # RT modelling
+
 Do a first lmer model. You can't put subit or gap over items because they are between-items.
@@
 ## subitizable
+
 ```{r, subitRT, results='asis'}
@@
 ## not subitizable
+
 ```{r, notsubitrT, results='asis'}
@@
 # Error rate analysis
-```{r, errorrates, results='asis'}
-print(xtable(summary(glmer(data=d, err~vag*subit+quan+gap + (vag|subject) + (1|item), family=binomial, control = glmerControl(optimizer = "bobyqa")))$coefficients, digits=3), type='html')
+
+```{r errorrates, results='asis'}
+print(xtable(summary(
+	glmer(data=d, err~vag*subit+quan+gap + (vag|subject) + (1|item), 
+				family=binomial, 
+				control = glmerControl(optimizer = "bobyqa")
+				)
+	)$coefficients,
+	digits=3), 
+	type='html')
 ```
@@
 # Plots of significant effects
-```{r, plot1}
+
+```{r plot_sig, fig.width=7.5, fig.height=3}
 a <- summarySEwithin(data=dat, measurevar="logRT", withinvars=c("vag","subit","item"), idvar="subject")
@@
   geom_line(position=pd) +
-  geom_errorbar(width=.1, position=pd)+
+  geom_errorbar(width=.2, position=pd)+
   geom_point(pch=21, size=4, position=pd) +
@@
 
-```
-
-```{r, plot2}
 a <- summarySEwithin(data=d, measurevar="err", withinvars=c("vag","subit","item"), idvar="subject")
@@
 
-```
-
-```{r, gridarrange, fig.width=7.5, fig.height=3}
 grid.arrange(plot1, plot2, nrow=1, top="Subitizable only")
@@
 
-```{r, barchart-rts}
+```{r barcharts, fig.width=9, fig.height=3}
 bwtheme <- standard.theme("pdf", color=FALSE)
-b1=barchart(data=ddply(dat,.(vag,subit),summarise,rt=mean(logRT)), rt~subit, group=vag, auto.key=list(),par.settings=bwtheme)
-```
-
-```{r, barchart-errors}
-b2=barchart(data=ddply(d,.(vag,subit),summarise,err=mean(err)), err~subit, group=vag, auto.key=list(),par.settings=bwtheme)
-```
-
-```{r, both-barcharts, fig.width=7, fig.height=3}
+b1=barchart(data=ddply(dat,.(vag,subit),summarise, logRT=mean(logRT)), 
+						logRT~subit, 
+						group=vag, auto.key=list(space='right', cex=.7), par.settings=bwtheme)
+b2=barchart(data=ddply(d,.(vag,subit),summarise, Error_Rate=mean(err)), 
+						Error_Rate~subit, 
+						group=vag, auto.key=list(space='right', cex=.7), par.settings=bwtheme)
 grid.arrange(b1,b2,ncol=2)
@@
 
-```{r, plotMeansa}
+```{r plotMeansa, fig.width=7, fig.height=3}
 a <- summarySEwithin(data=dat, measurevar="logRT", withinvars=c("vag","quan","subit"), idvar="subject")
 pd <- position_dodge(.1)
-meansplot1 <- ggplot(a,aes(y=logRT,x=vag,group=quan,shape=quan,col=quan,ymin=logRT-ci, ymax=ci+logRT))+geom_point(size=4, position=pd)+geom_line(position=pd) +facet_grid(~subit) + geom_errorbar(width=.1, position=pd) + theme_bw()+ theme(panel.grid.major=element_blank())
+	ggplot(
+		a, 
+		aes(y=logRT, x=vag, group=quan, shape=quan, col=quan, ymin=logRT-ci, ymax=ci+logRT)
+	) + 
+	geom_point(size=4, position=pd) + 
+	geom_line(position=pd) + 
+	facet_grid(~subit) + 
+	geom_errorbar(width=.1, position=pd) + 
+	theme_bw() + 
+	theme(panel.grid.major=element_blank()) 
 ```
 
-```{r, plotMeans, fig.width=7, fig.height=3}
-meansplot1
-```
-
-```{r, qqplotLMER, fig.width=3, fig.height=3.5}
+```{r qqplotLMER, fig.width=3, fig.height=3.5}
 qqPlot(residuals(lmer2),cex=.5)

### 2016-07-28 experiments/e0/writeup/acquire/gatherData.sh

@@ -1,11 +1,0 @@
-FIRST=
-for FILE in ../experimentCode/output/*.dat
-do
-    exec 5<"$FILE" # Open file
-    read LINE <&5 # Read first line
-    [ -z "$FIRST" ] && echo "$LINE" # Print it only from first file
-    FIRST="no"
-    cat <&5 # Print the rest directly to standard output
-    exec 5<&- # Close file
-    # Redirect stdout for this section into file.out
-done > ../data/data.txt

### 2016-07-28 experiments/e0/writeup/e0.Rmd
tidying up

--- 2016-07-28 experiments/e0/writeup/vgn0.Rmd
@@
     toc: yes
+    <!--toc_float: true-->
     mathjax: default

### 2016-07-28 experiments/e1/writeup/e1.Rmd
tidying up

@@
 title: "Experiment 1 of 4"
+author: "Matt Green"
 output:
@@
 ```{r sourceFunctions}
-source("functions/CIfunctions.R")
+source("CIfunctions.R")
 ```

### 2016-07-28 experiments/e2/writeup/main.Rmd
tidying up

@@
 title: "Experiment 2 of 4: main"
+author: "Matt Green"
 output:

### 2016-07-28 experiments/e3/writeup/e3.Rmd
tidying up

--- 2016-07-28 experiments/e3/writeup/experiment3.Rmd
@@
 title: "Experiment 3 of 4"
+author: "Matt Green"
 output:
@@
 ---
+
 # Experiment 3
@@
 ```{r readdata}
-dat <- read.delim("data/allresults.txt", stringsAsFactors = FALSE)
+dat <- read.delim("allresults.txt", stringsAsFactors = FALSE)
 
@@
 
-# save dat as dat.Rda
-# This data set contains *all* trials including impossible trials and is mainly for graphs comparing different removals
-save(dat, file="dat.Rda")
+# The dat data set contains *all* trials including impossible trials 
+# and is mainly for graphs comparing different removals
 
@@
 
-# save dd as dd.Rda
+# save dd as data.Rda
 # This data set non-impossible trials and renamed/recomputed variables
-save(dd, file="dd.Rda")
+save(dd, file="data.Rda")
 ```
@@
 ```{r boxplot subjects and items}
-load("dd.Rda")
+load("data.Rda")
 # allow reference to transformation
@@
   xlab("") + ylab("")
-```
-
-```{r load_dd}
-load("dd.Rda")
 ```
@@
 ```{r coeff}
-par(las=2, mar=c(14,6,1,1))
-y=as.data.frame(summary(v5)$coef[-1,1]) # estimates, without intercept
-q=confint(v5)[-c(1,2,3),]
-y=cbind(y,q)
+v5 <- lme4::lmer(data=dd, 
+           RT_RecSqt ~ 
+             c_Vag + Selection.sum + c_Qty + c_Ord + 
+             Selection.sum:c_Vag:c_Qty +
+             item_mean_ratio + 
+             s_Trl +
+             RTprev_RecSqt +
+             nchar_instr +
+             (1+c_Vag + Selection.sum + c_Qty + c_Ord|Subject)
+)
+par(las=2, mar=c(6,6,1,1))
+y=data.frame(coef=summary(v5)$coef[-1,1]) # estimates, without intercept
+q=confint(v5, method="Wald")[-c(1,2,3),]
+y=cbind(y,q[15:23,])
 plot(y[,1], ylim=range(y[,1],y[,2],y[,3]), xaxt="n", xlab="", ylab="estimate\n",  pch='-', cex=2)
@@
 abline(h=0)
-axis(1, labels=row.names(y), at=1:nrow(y), srt=45)
+axis(1, labels = abbreviate(row.names(y), minlength=10), , at=1:nrow(y), las=2)
 grid()
@@
 # create scaled residuals
-v5$rstand = as.vector(scale(resid(v5)))
+dd$rstand = as.vector(scale(resid(v5)))
 # plot scaled residuals density
-plot(density(v5$rstand))
+plot(density(dd$rstand))
 # plot sample quantiles versus theoretical quantiles
-qqnorm(v5$rstand, cex=.5)
-qqline(v5$rstand)
+qqnorm(dd$rstand, cex=.5)
+qqline(dd$rstand)
 # plot standardised residuals versus fitted values
-plot(v5$rstand ~ fitted(v5), pch='.')
+plot(dd$rstand ~ fitted(v5), pch='.')
 # absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
@@
 # create dffits
-dffits=abs(resid(v5, "dffits"))
+#dffits=abs(resid(v5, "dffits"))
 # plot dffits
-plot(dffits, type='h')
+#plot(dffits, type='h')
 ```

### 2016-07-28 experiments/e4/writeup/e4.Rmd
adding e4

---
title: "Experiment 4 of 4"
author: "Matt Green"
output:
  html_document:
    toc: true
    <!--toc_float: true-->
    mathjax: default
    template: default
    code_folding: hide
    theme: default
---  
# Experiment 4

Each subject saw 192 trials, being 3 cycles of 64 trials. There were 40 subjects. There were 7680 rows in the full data.

# Trial Procedure
First an instruction was presented until subjects dismissed it with a key press. Then a grey screen was presented for 1 second. Then a prime was presented for 2 seconds. Then a grey screen was presented for 1 second. Then the squares with dots were presented until the subject responded, with a 10 second timeout if they did not respond.

<figure>
<img src="./flow-of-experiment4.pdf" alt="Mountain View" style="width:480px;height:300px;">
<figcaption>Flow of experiment 4.</figcaption>
</figure>

# Describe the data

Here is a description of each column in the raw data.

* **Sub** identifier for subject. There are 40 subjects.
* **Trl** identifier for the trial. There are 64 Trials for each subject.
* **Itm** identifier for item. There are 4 items. An item is a set of Selections for the dots in the squares, e.g., Item 1 is (6,15,24) in any order. This should be included as a (possibly ordered) factor because the task will probably take longer for more dots. Include this in the final model if it does increase rt. Table below shows the dots for each item.
* **Cnd** just a single vector for the combination of vagueness and selection. Useful for some plotting routines that don't accept pairs of grouping variables. 
* **Ord** whether the order was for example (6,15,24) or (24,14,6). This matters because when the smaller quantity is on the left people are quicker than when the smaller quantity is on the right (i.e, SNARC effect). It is a control variable rather than a direct predictor. It could be left out of the main model assuming it is non-significant.
* **Qty** whether the instruction identifies bigger or smaller quantities. This is a factor with 2 levels sum coded. An example of bigger is _"Choose a square with more dots than the target"_. An example of smaller is _"Choose a square with fewer dots than the target"_.
* **Prm** how many dots there were in the prime (prime is _target_ in the instructions to the participants). Primes were 6 10 16 20 24 26 30 34 36 40 44 50 54.
* **Lft** how many dots there were in the left square
* **Mid** how many dots there were in the middle square
* **Rgt** how many dots there were in the right-hand side square
* **key** which key the subject pressed as a response on a trial.
* **rt** raw response time in milliseconds
* **Ins** a character array for the instruction that was presented to the subject on that trial. Table below shows an example of an instruction in each condition.
* **V** factor for vagueness with 2 levels sum coded
* **S** factor for selection task with 2 levels sum coded

Here is a table of instructions.

<style type="text/css">
.table {width: 50%;}
</style>

vagueness | selection | instruction
----------|-----------|--------------------------------------------------------------------
crisp     | match     | Choose a square with the same number of dots as the target  
crisp     | compare   | Choose a square with fewer dots than the target 
vague     | match     | Choose a square with about the same number of dots as the target 
vague     | compare   | Choose a square with far fewer dots than the target

Item   | dots
-------|-----------
Item 1 | (6,15,24)
Item 2 | (16,25,34)
Item 3 | (26,35,44)
Item 4 | (36,45,54)

```{r, startClean, echo=FALSE}
# start clean
rm(list=ls())
```

```{r, docOpts, echo=FALSE, cache=FALSE}
# Knitr doc options.
library(knitr)
knit_hooks$set(crop=hook_pdfcrop)
#options(tidy=TRUE)
#options(replace.assign=TRUE, width=50)
```

```{r, loadLibraries, echo=FALSE}
# Load libraries.
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(lattice))
suppressPackageStartupMessages(library(LMERConvenienceFunctions))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(lmerTest))
```

Source additional R files with functions for computing confidence intervals.

```{r, sourceFunctions}
source("functions/CIfunctions.R")
```

Get data from the raw per-subject files.

```{r, getData}
#if( file.exists('dd.Rda') ) load('dd.Rda') else {
#system(wait=TRUE, paste("head -n 1 ", "../experimentCode","/output/", "subject01.data ", "> ", "../data/data.txt", sep=""))
#system(wait=TRUE, paste("tail -q -n +2 ", "../experimentCode", "/output/", "*.data ",  ">> ", "../data/data.txt", sep=""))
dat <- read.delim(file="../data/data.txt", header=T, stringsAsFactors=FALSE)

# declare local variables
Selection_of_valid_subjects <- 40
Selection_of_rows <- 7680
Selection_of_trials_per_subject <- Selection_of_rows / Selection_of_valid_subjects # 192

# sort out borderline responses into expected near and far
# How many squares in the square they chose? 
dat$RESPONSE <- as.character(dat$key)
for (row in 1 : nrow(dat) ) {
  switch(dat[row,'RESPONSE'],
         'L' = {dat[row, 'choice'] <- dat[row, 'Lft']},
         'M' = {dat[row, 'choice'] <- dat[row, 'Mid']},
         'R' = {dat[row, 'choice'] <- dat[row, 'Rgt']}         
  )
}
  dat$Item <- dat$Itm
  dat$Itm <- NULL
dat$crossed=paste('Con',dat$Cnd,':Quan',dat$Qty,':Item',dat$Item,sep="")

dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseExpected'] <-  6 
dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseNear']     <- 15  
dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseFar']      <- 24

dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseExpected'] <- 16
dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseFar']      <- 34

dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseExpected'] <- 26
dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseFar']      <- 44

dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseExpected'] <- 36
dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseFar']      <- 54

dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseExpected'] <- 24
dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseNear']     <- 15 
dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseFar']      <-  6

dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseExpected'] <- 34
dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseFar']      <- 16

dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseExpected'] <- 44
dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseFar']      <- 26

dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseExpected'] <- 54
dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseFar']      <- 36

dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseExpected'] <-  6 
dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseNear']     <- 15  
dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseFar']      <- 24

dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseExpected'] <- 16
dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseFar']      <- 34

dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseExpected'] <- 26
dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseFar']      <- 44

dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseExpected'] <- 36
dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseFar']      <- 54

dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseExpected'] <- 24
dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseNear']     <- 15 
dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseFar']      <-  6

dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseExpected'] <- 34
dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseFar']      <- 16

dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseExpected'] <- 44
dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseFar']      <- 26

dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseExpected'] <- 54
dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseFar']      <- 36

dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseExpected'] <-  6 
dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseNear']     <- 15  
dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseFar']      <- 24

dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseExpected'] <- 16
dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseFar']      <- 34

dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseExpected'] <- 26
dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseFar']      <- 44

dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseExpected'] <- 36
dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseFar']      <- 54

dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseExpected'] <- 24
dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseNear']     <- 15 
dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseFar']      <-  6

dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseExpected'] <- 34
dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseFar']      <- 16

dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseExpected'] <- 44
dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseFar']      <- 26

dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseExpected'] <- 54
dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseFar']      <- 36

dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseExpected'] <-  6 
dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseNear']     <- 15  
dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseFar']      <- 24

dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseExpected'] <- 16
dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseFar']      <- 34

dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseExpected'] <- 26
dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseFar']      <- 44

dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseExpected'] <- 36
dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseFar']      <- 54

dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseExpected'] <- 24
dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseNear']     <- 15 
dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseFar']      <-  6

dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseExpected'] <- 34
dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseNear']     <- 25  
dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseFar']      <- 16

dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseExpected'] <- 44
dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseNear']     <- 35
dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseFar']      <- 26

dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseExpected'] <- 54
dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseNear']     <- 45
dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseFar']      <- 36

dat$isResponseExpected <- dat$choice == dat$ResponseExpected
dat$isResponseNear <- dat$choice == dat$ResponseNear
dat$isResponseFar <- dat$choice == dat$ResponseFar

for ( row in 1:nrow(dat) ) {
  dat[row, 'switchResponse'] <- 
    ifelse(dat[row, 'isResponseExpected']==TRUE, 'Expected', 
           ifelse(dat[row, 'isResponseNear']==TRUE, 'Near', 'Far') ) 
}

# SUBJECT
# ensure Subject is a factor 
dat$Subject=factor(paste("s",sprintf("%02d",dat$Sub),sep=""))

# TRIAL
# Trial for subject, 1 to 192
dat$Trial = rep(x = 1:Selection_of_trials_per_subject, times = Selection_of_valid_subjects)

# make a centred Trial for modeling
dat$c_Trl <-dat$Trial - mean(dat$Trial)

# make a scaled Trial for modelling
dat$s_Trl <- as.numeric(scale(dat$Trial))

# ID
# id is a unique identifier for the 7680 row data
dat$id <- factor(paste(paste(dat$Subject), 
                       paste("t", sprintf("%03d", dat$Trial), sep="") , sep=":"))

# ITEM
# create a centred numeric item variable for modeling
dat$c_Itm <- ifelse(dat$Item==1, -.75, 
                    ifelse(dat$Item==2, -.25,
                           ifelse(dat$Item==3, .25, .75)))

# make Item be a factor and assign labels
dat$Item <- factor(dat$Item, levels=c(1,2,3,4), labels=c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))

# add ratios for Item
item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54) 
# 0.2500000 0.4705882 0.5909091 0.6666667
item_range_ratio_scaled = as.vector(scale(c(6 / 24, 16 / 34, 26 / 44, 36 / 54)))
# -1.3441995 -0.1316642  0.5297187  0.9461450
item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))) 
# 0.5125000 0.6876471 0.7691558 0.8166667
item_mean_ratio_scaled = as.vector(scale(c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))))) 
# -1.37582241 -0.06614191  0.54334858 0.89861574

dat[dat$Item == "06:15:24", "item_range_ratio"] <-  0.2500000
dat[dat$Item == "16:25:34", "item_range_ratio"] <-  0.4705882
dat[dat$Item == "26:35:44", "item_range_ratio"] <-  0.5909091
dat[dat$Item == "36:45:54", "item_range_ratio"] <-  0.6666667

dat[dat$Item == "06:15:24", "item_range_ratio_scaled"] <-  -1.3441995
dat[dat$Item == "16:25:34", "item_range_ratio_scaled"] <-  -0.1316642
dat[dat$Item == "26:35:44", "item_range_ratio_scaled"] <-   0.5297187  
dat[dat$Item == "36:45:54", "item_range_ratio_scaled"] <-   0.9461450

dat[dat$Item == "06:15:24", "item_mean_ratio"] <-  0.5125000 
dat[dat$Item == "16:25:34", "item_mean_ratio"] <-  0.6876471 
dat[dat$Item == "26:35:44", "item_mean_ratio"] <-  0.7691558
dat[dat$Item == "36:45:54", "item_mean_ratio"] <-  0.8166667

dat[dat$Item == "06:15:24", "item_mean_ratio_scaled"] <-   -1.37582241 
dat[dat$Item == "16:25:34", "item_mean_ratio_scaled"] <-   -0.06614191 
dat[dat$Item == "26:35:44", "item_mean_ratio_scaled"] <-    0.54334858 
dat[dat$Item == "36:45:54", "item_mean_ratio_scaled"] <-    0.89861574

# VAGUENESS
# Create a factor coding for Vagueness
dat[ dat$Cnd==1 , 'Vagueness'] <- 'Vague'
dat[ dat$Cnd==2 , 'Vagueness'] <- 'Crisp'
dat[ dat$Cnd==3 , 'Vagueness'] <- 'Vague'
dat[ dat$Cnd==4 , 'Vagueness'] <- 'Crisp'
dat$Vagueness <- as.factor(dat$Vagueness)

# manually center Vagueness
dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)

dat$Selection = ""
dat[dat$Cnd %in% c(1,2), "Selection"] <- "matching"
dat[dat$Cnd %in% c(3,4), "Selection"] <- "comparison"
dat$Selection.sum <- ifelse(dat$Selection=="matching", .5, -.5)
dat$Selection <- as.factor(dat$Selection)

# remove redundant confusing columns
dat$Sub <- NULL
dat$Trl <- NULL

# ORDER
# give the levels of Order meaningful names
dat$Order <- factor(dat$Ord, levels=c(1,2), labels=c('LtoR','RtoL'))

# make a manually centred Order
dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)

# QUANTITY
# give the levels of Quantity meaningful names
dat$Quantity <- factor(dat$Qty, levels=c(1,2), labels=c('Small','Large'))

# make a manually centred Quantity
dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)

# INSTRUCTION
# add Selection of characters in the instruction # 29 30 34 36 38
dat$nchar_instr = nchar(as.character(dat$Ins))
dat$nchar_instr_scaled = as.vector(scale(nchar(as.character(dat$Ins)), scale=TRUE))

# make Instruction be a factor (17 levels)
dat$Ins <- as.factor(dat$Ins)

# RT
# add transformations of RT
dat$RT <- dat$rt; dat$rt <-  NULL
dat$RT_RecSqd    <- 1/dat$RT^2
dat$RT_Rec       <- 1/dat$RT
dat$RT_RecSqt    <- 1/sqrt(dat$RT) 
dat$RT_log       <- log(dat$RT)
dat$RT_sqt       <- sqrt(dat$RT)
dat$RT_raw       <- dat$RT
dat$RT_sqd       <- dat$RT^2

save(dat, file="dat.Rda")

dd <- dat

 # lose impossible trials
  dd <- dat
  dd$RT[dd$RT > 9000 ] <- NA
  dd <- dd[complete.cases(dd),]
  row.names(dd) <- NULL

# add preceding RT: because we removed impossible trials, the value for preceding RT for a trial following an impossible trial is the value of the trial that preceded the impossible trial.
dd$RTprev <- NA
for (s in levels(dd$Subject)) {
  nrows = nrow(dd[dd$Subject==s,])
  for (i in 1:nrows) {
    if (i==1) {
      dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]
    }
    else
      dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1]
  }
}

# add transformations of previous RT
dd$RTprev_RecSqd    <- 1/dd$RTprev^2
dd$RTprev_Rec       <- 1/dd$RTprev
dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev) 
dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
dd$RTprev_log       <- log(dd$RTprev)
dd$RTprev_sqt       <- sqrt(dd$RTprev)
dd$RTprev_raw       <- dd$RTprev
dd$RTprev_sqd       <- dd$RTprev^2
  # save dd
  save(dd, file="dd.Rda")
#}
```

Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions.

```{r,fig.width=20, fig.height=12, message=FALSE}
## plot the comparison of the transformations against their normal equivalents to see distributional properties using full data

suppressPackageStartupMessages(library("reshape2"))
suppressPackageStartupMessages(library("ggplot2"))
load("dat.Rda")

# first create a data frame specially for transformations
# this won't be used for analysis, just for this plot
dat_transforms <- dat

# dfSubset says whether any data points were removed
dat_transforms$dfSubset <- "original"

# RTtype says whether the row contains data for an observed RT or 
# a sample from a normal distribution with the same mean and sd
dat_transforms$RTtype <- "observed"

# create extra rows for dfs with removed (outlier) data points
dat1 <- dat_transforms                                       # nrow 7680
dat2 <- subset(dat_transforms, RT<9000)              # nrow 7677
dat2$dfSubset <- "RT<9000"

# bring the dfs back into main df
dat_transforms <- rbind(dat1,dat2)
dat_transforms$dfSubset <- factor(dat_transforms$dfSubset, levels=c("original", "RT<9000"))

# create extra rows for normally distributed transformed variables
dato = dat_transforms
datn = dat_transforms; datn$RTtype <- "normal"
dat_transforms <- rbind(dato,datn)
dat_transforms$RTtype <- factor(dat_transforms$RTtype, levels=c("observed","normal"))

# for safety
dat_transforms[dat_transforms$RTtype=="normal", c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd")] <- NA

# Create the normally distributed transformed variables, replacing the NAs declared above
for (k in levels(dat_transforms$dfSubset)){
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_RecSqd"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_RecSqd"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_RecSqd"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_RecSqd"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_Rec"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_Rec"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_Rec"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_Rec"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_RecSqt"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_RecSqt"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_RecSqt"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_RecSqt"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_log"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_log"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_log"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_log"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_sqt"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_sqt"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_sqt"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_sqt"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_raw"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_raw"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_raw"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_raw"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_sqd"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_sqd"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_sqd"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_sqd"]))
}

temp <- melt(data=dat_transforms, 
             id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Selection", "Quantity", "Order"),
             measure.vars=c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd"),
             variable.name="transformation",
             value.name="score")

ggplot(temp, aes(score, colour=RTtype)) + 
  geom_freqpoly() + 
  facet_wrap(dfSubset~transformation, scales="free", nrow=2) +
  theme_bw() + theme(aspect.ratio=1) + scale_color_manual(values=c("blue","red"))
```

Show how transformations of RT affect distribution of times, and how they affect which times are outliers.

```{r, fig.width=20, fig.height=5}
# boxplot subjects and items

load("dd.Rda")
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(ggplot2))

# allow reference to transformation
subdata = melt(dd,
               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
               variable.name="transformation",
               value.name="value")

# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ transformation, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
  xlab("") + ylab("")
```

Show how mean times for individual subjects and items vary with respect to the grand mean RT.

```{r, fig.width=10, fig.height=5}
# Show how mean times for individual subjects and items vary with respect to the grand mean RT.

suppressMessages(library(plyr))
suppressMessages(library(lattice))
suppressMessages(library(latticeExtra))

subs=ddply(dd, .(level=Subject), summarise, effect="subject", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
itms=ddply(dd, .(level=Item), summarise, effect="item", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
si=rbind(subs,itms)
si$effect <- factor(si$effect, levels=c("subject", "item"))
si$lsign <- factor(si$lsign, levels=c("slow","fast"))

my1 <- xyplot(data=si, lmean~level | effect, 
              aspect=1,
              layout=c(2,1),
              ylim = extendrange(si$lmean, f = 0.10),
              par.settings=list(trellis.par.set(superpose.symbol = list(col = c("blue","red"), pch=19), superpose.line=list(col = c("blue","red")))),
              groups=si$lsign,
              xlab=NULL,
              ylab="RT (in ms)",
              key=list(text=list(levels(si$lsign)), lines=list(lty=c(1,1), col=trellis.par.get('superpose.line')$col, pch=c(19,19), type='o'), divide=1),
              mean=unique(si$emean),
              scales=list(x=list(relation="free", at=list(15.5, 32.5), labels=list("subject number", "item identifier"), cex=1 ), 
                          y=list(at=c(0, 2000, unique(si$emean), 4000, 6000, 8000,  10000), labels=c(0, 2000,  "Mean", 4000,  6000, 8000, 10000), cex=.9) ),
              panel=function(x,y,type,subscripts,groups=groups,...){
                panel.xyplot(x,y,type='p',groups=groups, subscripts=subscripts,...)
                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, groups=groups, subscripts=subscripts, code=3, length=0, col=trellis.par.get('superpose.line')$col[groups][subscripts], ...)
                panel.abline(h=unique(si$emean), col="lightgrey",lty=1)
                panel.text(x, y=(ifelse(y<unique(si$emean),y-200,y+200)), as.numeric(si$level)[subscripts][panel.number()==1], cex=.75)
                panel.text(x, y=(ifelse(y<unique(si$emean),y-200,y+200)), as.character(si$level)[subscripts][panel.number()==2], cex=.75)
              })

my2 <- xyplot(data=si, ldiff~level | effect, 
              aspect=1, type='n',
              ylab="Deviation from mean RT (in ms)",
              ylim=extendrange(si$ldiff,f=0.10))

print(doubleYScale(my1, my2, add.axis=TRUE, add.ylab2=TRUE, use.style=FALSE))
```

Show how practice and fatigue effects vary over subjects

```{r, message=F, results='hide'}
if(require(ggplot2)){}
load("dd.Rda")
```

```{r, eval=TRUE, fig.width=20, fig.height=10, cache=FALSE, echo=TRUE}
# Show how practice/fatigue effects vary over subjects
ggplot(dd, aes(y=RT_log, x=Trial)) + facet_wrap(~Subject,nrow=3) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial", breaks=NULL) + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1) 
```

Show practice and fatigue effects vary over items.

```{r, eval=TRUE, fig.width=9, fig.height=4, cache=FALSE, echo=TRUE}
# Show how practice/fatigue effects vary over items
ggplot(dd, aes(y=RT_log, x=Trial)) + facet_grid(~Item) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial") + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1) 
```

# Show means.

Main effects

```{r}
# prep main effects rts data frame
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness")));names(vrts)[3]<-"Levels"
nrts <- cbind(Effect="Selection",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection")));names(nrts)[3]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Quantity")));names(qrts)[3]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Order")));names(orts)[3]<-"Levels"
irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Item")));names(irts)[3]<-"Levels"
rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_log-rts$ci
rts$maxs=rts$RT_log+rts$ci

# plot main effects
ggplot(rts, aes(y=RT_log, x=Levels, group=1, ymin=mins, ymax=maxs)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Effect, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) 
```

Main effects over items

```{r, message=FALSE, fig.width=10, fig.height=3}
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Vagueness")));names(vrts)[4]<-"Levels"
nrts <- cbind(Effect="Selection",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Selection")));names(nrts)[4]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Quantity")));names(qrts)[4]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Order")));names(orts)[4]<-"Levels"

rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_log-rts$ci
rts$maxs=rts$RT_log+rts$ci

rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))

ggplot(rts, aes(y=RT_log, x=item_mean_ratio, group=Levels, ymin=mins, ymax=maxs, shape=Levels, fill=Levels, col=Levels)) +
  scale_shape_manual(values=c(21,21,22,22,23,23,24,24)) +
  scale_fill_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
  scale_color_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
  geom_line() +
  geom_errorbar(width=.01) +
  geom_point(cex=3) + 
  facet_grid(~Effect) + 
  theme_bw() +
  theme(aspect.ratio=1, 
        panel.grid=element_blank()) +
  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))

```

2 way interactions ignoring item

```{r}
# prep 2 ways rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Selection")); names(rts1)[1] <- "F1"; names(rts1)[2] <- "F2"
rts1$Effect <- rep("Vagueness:Selection",4)
rts2 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Quantity")); names(rts2)[1] <- "F1"; names(rts2)[2] <- "F2"
rts2$Effect <- rep("Vagueness:Quantity",4)
rts3 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Order")); names(rts3)[1] <- "F1"; names(rts3)[2] <- "F2"
rts3$Effect <- rep("Vagueness:Order",4)
rts5 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection","Quantity")); names(rts5)[1] <- "F1"; names(rts5)[2] <- "F2"
rts5$Effect <- rep("Selection:Quantity",4)
rts6 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection","Order")); names(rts6)[1] <- "F1"; names(rts6)[2] <- "F2"
rts6$Effect <- rep("Selection:Order",4)
rts8 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Quantity","Order")); names(rts8)[1] <- "F1"; names(rts8)[2] <- "F2"
rts8$Effect <- rep("Quantity:Order",4)
rts <- rbind(rts1,rts2,rts3,rts5,rts6,rts8)
rts <- subset(rts, select=c(Effect, F1, F2, RT_log, RT_log_norm, sd, se, ci))
rts$mins=rts$RT_log-rts$ci
rts$maxs=rts$RT_log+rts$ci
```

```{r,fig.width=10, fig.height=3}
# plot 2 way interactions
ggplot(rts, aes(y=RT_log, x=F2, group=F1, ymin=mins, ymax=maxs, pch=F1)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_wrap(~Effect, scale="free_x", nrow=1) + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank()) +
  xlab("Levels")
```

Plot 2 way interactions over items

```{r, message=FALSE}
# prep 2 ways over items rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Vagueness","Selection")); names(rts1)[2] <- "F1"; names(rts1)[3] <- "F2"
rts1$E1 <- "Vagueness"
rts1$E2 <- "Selection"
rts2 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Vagueness","Quantity")); names(rts2)[2] <- "F1"; names(rts2)[3] <- "F2"
rts2$E1 <- "Vagueness"
rts2$E2 <- "Quantity"
rts3 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Vagueness","Order")); names(rts3)[2] <- "F1"; names(rts3)[3] <- "F2"
rts3$E1 <- "Vagueness"
rts3$E2 <- "Order"
rts4 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Quantity", "Selection")); names(rts4)[2] <- "F1"; names(rts4)[3] <- "F2"
rts4$E1 <- "Quantity"
rts4$E2 <- "Selection"
rts5 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Order","Selection")); names(rts5)[2] <- "F1"; names(rts5)[3] <- "F2"
rts5$E1 <- "Order"
rts5$E2 <- "Selection"
rts6 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Quantity","Order")); names(rts6)[2] <- "F1"; names(rts6)[3] <- "F2"
rts6$E1 <- "Quantity"
rts6$E2 <- "Order"

rts <- rbind(rts1,rts2,rts3,rts4,rts5,rts6)
rts$mins=rts$RT_log-rts$ci
rts$maxs=rts$RT_log+rts$ci

rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
rts$F3 <- factor(paste(rts$F1,rts$F2))

rts$E1 <- factor(rts$E1, levels=c("Selection","Order","Quantity","Vagueness"))
rts$E2 <- factor(rts$E2, levels=c("Selection","Order","Quantity","Vagueness"))

# The palette with black:
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

ggplot(rts, aes(y=RT_log, x=item_mean_ratio, group=F3, col=F2, pch=F1)) + 
  facet_grid(E2~E1, drop=F) + 
  geom_point(cex=3) + geom_line() + theme_bw() + 
  theme(aspect.ratio=1, panel.grid=element_blank()) +
  scale_color_manual(name="colour",values=cbbPalette) +
  scale_shape_discrete(name="shape") +
  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))

```

# RT model all items

```{r, rtmodel-allitems}
# backing off A LOT from maximal model to avoid warning about maxfun
rtm <- lmer(
  data=dd,
  RT_log ~ 
    c_Vag * Selection.sum + item_mean_ratio + c_Qty + c_Ord + s_Trl +
    RTprev_log  + 
  ( c_Vag * Selection.sum | Subject)
)
rtm.summary <- summary(rtm)
rtm.confint <- confint(rtm)
```

Print a table of coefficients.

```{r, rtmodeltable, results='asis', echo=FALSE}
print(xtable(
  digits=c(0,2,2,1,1,3), 
  rtm.summary$coefficients), type='html')
```

Plot coefficients

```{r}
par(las=2, mar=c(14,6,1,1))
y=coef(rtm.summary)[-1,1] # omit intercept
n=length(y)
plot(y, xaxt="n", xlab="", ylab="RT_log\n", pch=19, ylim=extendrange(y, f=.10), main="Coefficients")
abline(h=0)
axis(1, labels=names(y) , at=1:n)
grid()
segments(x0=1:n, x1=1:n, y0, y1, lwd=2)
```

Plot partial effects

```{r}
par(mfrow=c(2,4))
plotLMER.fnc(rtm)
```

```{r}
suppressPackageStartupMessages(require(languageR))
cat("k is:",
collin.fnc(dd[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])$cnumber)
```

```{r, fig.width=6, fig.height=5}
suppressPackageStartupMessages(require(rms))
par(mar=c(1.1,3.1,1.1,1.1), pty='s')
plot(varclus(as.matrix(dd[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])))
```

Model criticism baayen plots

```{r, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

par(mfrow=c(1,4), pty='s')
# create scaled residuals
dd$rstand = as.vector(scale(resid(rtm)))
# plot scaled residuals density
plot(density(dd$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(dd$rstand, cex=.5)
qqline(dd$rstand)
# plot standardised residuals versus fitted values
plot(dd$rstand ~ fitted(rtm), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))

```

Refit model with outliers removed (3.5 sd)

```{r}
dd2 <- subset(dd, abs(rstand)<3.5 )
```

```{r}
# backing off A LOT from maximal model to avoid warning about maxfun
rtm <- lmer(
  data=dd2,
  RT_log ~ 
    c_Vag * Selection.sum + item_mean_ratio + c_Qty + c_Ord + s_Trl +
    RTprev_log  + 
  ( c_Vag * Selection.sum | Subject)
)
rtm.summary <- summary(rtm)
rtm.confint <- confint(rtm)
```

Print a table of coefficients.

```{r,  results='asis', echo=FALSE}
print(xtable(
  digits=c(0,2,2,1,1,3), 
  rtm.summary$coefficients), type='html')
```

Plot coefficients

```{r}
par(las=2, mar=c(14,6,1,1))
y=coef(rtm.summary)[-1,1] # omit intercept
n=length(y)
plot(y, xaxt="n", xlab="", ylab="RT_log\n", pch=19, ylim=extendrange(y, f=.10), main="Coefficients")
abline(h=0)
axis(1, labels=names(y) , at=1:n)
grid()
segments(x0=1:n, x1=1:n, y0, y1, lwd=2)
```

Plot partial effects

```{r}
par(mfrow=c(2,4))
plotLMER.fnc(rtm)
```

```{r}
suppressPackageStartupMessages(require(languageR))
cat("k is:",
collin.fnc(dd2[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])$cnumber
)

```

```{r, fig.width=6, fig.height=5}
suppressPackageStartupMessages(require(rms))
par(mar=c(1.1,3.1,1.1,1.1), pty='s')
plot(varclus(as.matrix(dd2[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])))
```

Model criticism baayen plots

```{r, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

par(mfrow=c(1,4), pty='s')
# create scaled residuals
dd2$rstand = as.vector(scale(resid(rtm)))
# plot scaled residuals density
plot(density(dd2$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(dd2$rstand, cex=.5)
qqline(dd2$rstand)
# plot standardised residuals versus fitted values
plot(dd2$rstand ~ fitted(rtm), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))

```

Separate analyses for vague vs crisp at each level of selection (matching, and comparison)

not done yet.

```{r,  results='asis', echo=FALSE}
print(xtable(
  digits=c(0,2,2,1,1,3), 
  rtm.summary$coefficients), type='latex')
```

# Borderline analysis

### 2016-07-28 experiments/e4/experimentCode/e4.m

@@ -1,0 +1,387 @@
+function e5(SUBJECT)
+
+diary on;
+more off;
+newLevel=3; % adjust verbosity
+oldLevel = Screen('Preference', 'Verbosity', [newLevel]);
+ListenChar(2); % disable keyboard input to octave while running experiment (no echo to terminal)
+% If we are in testing mode then do skip synch tests, If we are in real mode don't skip synch tests
+skipTestValue=0
+if SUBJECT >99
+	skipTestValue=1
+end
+
+
+Screen('Preference', 'SkipSyncTests', skipTestValue); % 0 = don't skip sync tests; 1 = do skip run sync tests
+
+% VARIABLES
+howLongToWaitForAResponse = 10; % in seconds. Set low for test runs
+timeOutTime = 2; % in seconds.
+maskDuration = 1; % in seconds
+interTrialDelay = 1; % in seconds
+leftBox  = [170 300 370 500];
+midBox   = [540 300 740 500];
+rightBox = [910 300 1110 500];
+leftCenter  = [270 400];
+midCenter   = [640 400];
+rightCenter = [1010 400];
+messagey = 150; 
+instructionY = 400;
+instructionX = 100;
+gray=127.5;
+white=225;
+black=0;
+startrt=0;   % initialise start rt variable
+endrt=0;     % initialise end rt variable
+instruction_on_time=0;
+instruction_off_time=0;
+numberOfBlocks = 4; % 1 is practice, 2 is block1, 3 is block2, 4 is block3
+numberOfColumns = 9; % there are always 9 columns in the input
+rand('state',sum(100*clock));
+
+% COMMANDS
+% Clear Matlab/Octave window:
+clc;
+% check for Opengl compatibility, abort otherwise:
+AssertOpenGL;
+% Make sure keyboard mapping is the same on all supported operating systems
+KbName('UnifyKeyNames');
+% Define keys by ID
+leftresp=KbName('c'); 
+middleresp=KbName('b'); 
+rightresp=KbName('m'); 
+escapekey=KbName('ESCAPE'); 
+spacebar=KbName('space'); 
+returnkey=KbName('return'); 
+% read separate files containing dotlocations, stimuli, and strings 
+source('e5stimuli.m');
+source('e5strings.m');
+source('e5dotlocations.m');
+
+% set up output files
+% practice data output file to write to
+practicedatafilename = strcat('output/practice/subject',...
+    num2str(SUBJECT),'Practice','.data');
+% experimental data output file to write to
[…]

### 2016-07-28 experiments/e4/experimentCode/e4strings.m

@@ -1,0 +1,10 @@
+    %%%%%%%%%%%%%%%%%%%%%%%
+    % define text strings %
+    %%%%%%%%%%%%%%%%%%%%%%%
+
+    instructionText = sprintf(strcat('Press Enter to start the practice trials.'));
+
+    oopsText = sprintf(strcat(...
+        'Ooops... Something went wrong.\n\n',...
+        'Please tell the experimenter.'));
+

### 2016-07-28 experiments/e4/experimentCode/touchpadoff.sh

@@ -1,0 +1,2 @@
+synclient TouchpadOff=1
+

### 2016-07-28 experiments/e4/writeup/acquire/gatherData.sh

@@ -1,0 +1,14 @@
+FIRST=
+
+for FILE in ../code/output/*.data
+do
+        exec 5<"$FILE" # Open file
+        read LINE <&5 # Read first line
+        [ -z "$FIRST" ] && echo "$LINE" # Print it only from first file
+        FIRST="no"
+
+        cat <&5 # Print the rest directly to standard output
+        exec 5<&- # Close file
+        # Redirect stdout for this section into file.out
+
+done > allresults.txt

### 2016-07-28 experiments/e4/writeup/e4.Rmd
e4 edits

@@
 
-```{r, startClean, echo=FALSE}
-# start clean
-rm(list=ls())
-```
-
 ```{r, docOpts, echo=FALSE, cache=FALSE}
-# Knitr doc options.
 library(knitr)
 knit_hooks$set(crop=hook_pdfcrop)
-#options(tidy=TRUE)
-#options(replace.assign=TRUE, width=50)
 ```
@@
 suppressPackageStartupMessages(library(lattice))
+suppressPackageStartupMessages(library(latticeExtra))
 suppressPackageStartupMessages(library(LMERConvenienceFunctions))
@@
 ```{r, sourceFunctions}
-source("functions/CIfunctions.R")
+source("CIfunctions.R")
 ```
 
-Get data from the raw per-subject files.
+Get data from the raw per-subject files if necessary.
 
 ```{r, getData}
-#if( file.exists('dd.Rda') ) load('dd.Rda') else {
-#system(wait=TRUE, paste("head -n 1 ", "../experimentCode","/output/", "subject01.data ", "> ", "../data/data.txt", sep=""))
-#system(wait=TRUE, paste("tail -q -n +2 ", "../experimentCode", "/output/", "*.data ",  ">> ", "../data/data.txt", sep=""))
-dat <- read.delim(file="../data/data.txt", header=T, stringsAsFactors=FALSE)
+if (file.exists('data.txt')) {
+	dat <- read.delim(file="data.txt", header=T, stringsAsFactors=FALSE)
+} else {
+	system(wait=TRUE, paste("head -n 1 ", "../experimentCode","/output/", "subject01.data ", "> ", "data.txt", sep=""))
+	system(wait=TRUE, paste("tail -q -n +2 ", "../experimentCode", "/output/", "*.data ",  ">> ", "data.txt", sep=""))
+}
 
@@
 
-save(dat, file="dat.Rda")
-
-dd <- dat
-
 # lose impossible trials
@@
 dd$RTprev_sqd       <- dd$RTprev^2
-  # save dd
-  save(dd, file="dd.Rda")
-#}
+
+save(dd, file="data.Rda")
 ```
@@
 
-```{r,fig.width=20, fig.height=12, message=FALSE}
+```{r foo,fig.width=20, fig.height=12, message=FALSE}
 ## plot the comparison of the transformations against their normal equivalents to see distributional properties using full data
-
-suppressPackageStartupMessages(library("reshape2"))
-suppressPackageStartupMessages(library("ggplot2"))
-load("dat.Rda")
 
@@
 
-temp <- melt(data=dat_transforms, 
+temp <- reshape2::melt(data=dat_transforms, 
              id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Selection", "Quantity", "Order"),
-             measure.vars=c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd"),
-             variable.name="transformation",
-             value.name="score")
+             measure.vars=c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd"))
 
-ggplot(temp, aes(score, colour=RTtype)) + 
+ggplot(temp, 
+			 aes(value, colour=RTtype)) + 
 	geom_freqpoly() + 
-  facet_wrap(dfSubset~transformation, scales="free", nrow=2) +
+  facet_wrap(dfSubset~variable, scales="free", nrow=2) +
   theme_bw() + theme(aspect.ratio=1) + scale_color_manual(values=c("blue","red"))
@@
 
-```{r, fig.width=20, fig.height=5}
+```{r bar, fig.width=20, fig.height=5}
 # boxplot subjects and items
-
-load("dd.Rda")
-suppressPackageStartupMessages(library(reshape2))
-suppressPackageStartupMessages(library(ggplot2))
-
+load("data.Rda")
 # allow reference to transformation
 subdata = melt(dd,
-               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
-               variable.name="transformation",
-               value.name="value")
-
+               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw")) 
 # boxplots for subjects and items separately for each dependent variable
 ggplot(subdata) + 
-  facet_wrap(~ transformation, scales="free", ncol=4) +
+  facet_wrap(~ variable, scales="free", ncol=4) +
   geom_boxplot(aes(y=value, x=Item, col="Items")) + 
@@
 
-```{r, fig.width=10, fig.height=5}
+```{r, bar2, fig.width=10, fig.height=5}
 # Show how mean times for individual subjects and items vary with respect to the grand mean RT.
-
-suppressMessages(library(plyr))
-suppressMessages(library(lattice))
-suppressMessages(library(latticeExtra))
-
 subs=ddply(dd, .(level=Subject), summarise, effect="subject", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
@@
 
-```{r, message=F, results='hide'}
-if(require(ggplot2)){}
-load("dd.Rda")
-```
-
-```{r, eval=TRUE, fig.width=20, fig.height=10, cache=FALSE, echo=TRUE}
+```{r barr3, eval=TRUE, fig.width=20, fig.height=10, cache=FALSE, echo=TRUE}
 # Show how practice/fatigue effects vary over subjects
@@
 
-```{r, eval=TRUE, fig.width=9, fig.height=4, cache=FALSE, echo=TRUE}
+```{r, var3, eval=TRUE, fig.width=9, fig.height=4, cache=FALSE, echo=TRUE}
 # Show how practice/fatigue effects vary over items
@@
 
-```{r}
+```{r, bar4}
 # prep main effects rts data frame
@@
 
-```{r, message=FALSE, fig.width=10, fig.height=3}
+```{r, bar5, message=FALSE, fig.width=10, fig.height=3}
 vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Vagueness")));names(vrts)[4]<-"Levels"
@@
   scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1)) 
-
 ```
@@
 
-```{r}
+```{r, bar6}
 # prep 2 ways rts data frame
@@
 
-```{r,fig.width=10, fig.height=3}
+```{r,bar7,fig.width=10, fig.height=3}
 # plot 2 way interactions
@@
 
-```{r, message=FALSE}
+```{r foo2, message=FALSE}
 # prep 2 ways over items rts data frame
@@
   scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))
-
 ```
@@
 
-```{r, rtmodel-allitems}
+```{r, foo3, rtmodel-allitems}
 # backing off A LOT from maximal model to avoid warning about maxfun
-rtm <- lmer(
+rtm <- lme4::lmer(
   data=dd,
@@
 rtm.summary <- summary(rtm)
-rtm.confint <- confint(rtm)
+rtm.confint <- confint(rtm, method='Wald')[13:20,]
 ```
@@
 
-```{r, rtmodeltable, results='asis', echo=FALSE}
+```{r, foo4, rtmodeltable, results='asis', echo=FALSE}
 print(xtable(
-  digits=c(0,2,2,1,1,3), 
+  digits=c(0,2,2,1), 
   rtm.summary$coefficients), type='html')
@@
 
-```{r}
+```{r, foo5}
 par(las=2, mar=c(14,6,1,1))
@@
 n=length(y)
-plot(y, xaxt="n", xlab="", ylab="RT_log\n", pch=19, ylim=extendrange(y, f=.10), main="Coefficients")
+plot(y, xaxt="n", xlab="", ylab="RT_log\n", pch=19, ylim=extendrange(y, f=.20), main="Coefficients")
 abline(h=0)
@@
 grid()
-segments(x0=1:n, x1=1:n, y0, y1, lwd=2)
+segments(x0=1:n, x1=1:n, y0=rtm.confint[,1], y1=rtm.confint[,2], lwd=2)
 ```
@@
 
-```{r}
+```{r foo6}
 par(mfrow=c(2,4))
@@
 
-```{r}
+```{r foo7}
 suppressPackageStartupMessages(require(languageR))
@@
 
-```{r, fig.width=6, fig.height=5}
+```{r foo8, fig.width=6, fig.height=5}
 suppressPackageStartupMessages(require(rms))
@@
 
-```{r, fig.width=9, fig.height=3}
+```{r daz, fig.width=9, fig.height=3}
 # Baayen 4-plot model criticism
@@
 
-```{r}
+```{r daz2}
 dd2 <- subset(dd, abs(rstand)<3.5 )
@@
 
-```{r}
+```{r daz3}
 # backing off A LOT from maximal model to avoid warning about maxfun
@@
 rtm.summary <- summary(rtm)
-rtm.confint <- confint(rtm)
+rtm.confint <- confint(rtm, method='Wald')
 ```
@@
 
-```{r,  results='asis', echo=FALSE}
+```{r daz4,  results='asis', echo=FALSE}
 print(xtable(
@@
 
-```{r}
+```{r daz5}
 par(las=2, mar=c(14,6,1,1))
@@
 grid()
-segments(x0=1:n, x1=1:n, y0, y1, lwd=2)
+segments(x0=1:n, x1=1:n, y0=rtm.confint[13:20,1], y1=rtm.confint[13:20,2], lwd=2)
 ```
@@
 
-```{r}
+```{r daz6}
 par(mfrow=c(2,4))
@@
 
-```{r}
+```{r daz7}
 suppressPackageStartupMessages(require(languageR))
@@
 )
-
 ```
 
-```{r, fig.width=6, fig.height=5}
+```{r daz9, fig.width=6, fig.height=5}
 suppressPackageStartupMessages(require(rms))
@@
 
-```{r, fig.width=9, fig.height=3}
+```{r mc1, fig.width=9, fig.height=3}
 # Baayen 4-plot model criticism
@@
 
-```{r,  results='asis', echo=FALSE}
+```{r sep1,  results='asis', echo=FALSE}
 print(xtable(
   digits=c(0,2,2,1,1,3), 
-  rtm.summary$coefficients), type='latex')
+  rtm.summary$coefficients), type='html')
 ```
@@
 # Borderline analysis
+
+not done yet.

### 2016-07-29 experiments/e2/archive/lmer_code_temp.Rmd
archiving old e2 Rmd files

## lm model in logRT

```{r, results='asis', cache=FALSE, eval=TRUE}
lm.logRT.0 <- lm (logRT ~  c_Trl + expos.c + expos.q +
                    c_Num + c_Vag + Quantity + Order +
                    c_Num : c_Vag , dat)
co <- as.data.frame(summary(lm.logRT.0)$coef)
# do stars
x=co[,"Pr(>|t|)"]
s=rep(NA, times=length(x))
for (i in 1:length(x)) {s[i]=ifelse(x[i]<.001, "$***$", ifelse(x[i]<.01,"$**$", ifelse(x[i]<.05,"$*$",  ifelse(x[i]<.1, "§", " "))))}
# cut p value into minimum <.001
p=rep(NA, times=length(x))
for (i in 1:length(x)) {p[i] <- as.character(ifelse(x[i]<.001, "<0.001", formatC(round(x[i],3), format='f', digits=3)))}
# bind coefs with p and s
dd=data.frame(beta=rownames(co), Estimate=co[,1], Std_Error=co[,2], t_value = co[,3], p_value = p, signif=s, row.names=NULL)
print(  xtable (dd, caption="Signif. codes: $***$ < 0.001; $**$ < 0.01; $*$ < 0.05; § < 0.1", align=rep("r",times=length(dd)+1)), type='html', include.rownames=FALSE)

```

## lmer models in logRT

```{r, eval=TRUE, cache=TRUE, warning=TRUE, message=TRUE}
# baseline model should be without experimental manipulations because rt is typically mostly detrmined by mean RT and practice/fatigue
lmer(data=dat, logRT ~ (1 | Subject) + (1 | f_Itm) ) -> m0
lmer(data=dat, logRT ~ s_Trl + (1 | Subject) + (1 | f_Itm) ) -> m1
lmer(data=dat, logRT ~ s_Trl + (1 + s_Trl | Subject) + (1 | f_Itm) ) -> m2
lmer(data=dat, logRT ~ s_Trl + (1 + s_Trl | Subject) + (1 + s_Trl | f_Itm) ) -> m3
# add in experimental manipulations
lmer(data=dat, logRT ~ s_Trl + c_Num + (1 + s_Trl | Subject) + (1 + s_Trl | f_Itm) ) -> m4
lmer(data=dat, logRT ~ s_Trl + c_Num + (1 + s_Trl + c_Num | Subject) + (1 + s_Trl | f_Itm) ) -> m5
lmer(data=dat, logRT ~ s_Trl + c_Num + (1 + s_Trl + c_Num | Subject) + (1 + s_Trl + c_Num | f_Itm) ) -> m6
lmer(data=dat, logRT ~ s_Trl + c_Num + c_Vag + (1 + s_Trl + c_Num | Subject) + (1 + s_Trl + c_Num | f_Itm) ) -> m7
lmer(data=dat, logRT ~ s_Trl + c_Num + c_Vag + (1 + s_Trl + c_Num + c_Vag | Subject) + (1 + s_Trl + c_Num | f_Itm) ) -> m8
lmer(data=dat, logRT ~ s_Trl + c_Num + c_Vag + (1 + s_Trl + c_Num + c_Vag | Subject) + (1 + s_Trl + c_Num + c_Vag | f_Itm) ) -> m9
```

```{r, eval=TRUE, cache=TRUE, warning=TRUE, message=TRUE}
# interactions of number and vagueness
lmer(data=dat, logRT ~ s_Trl + c_Num * c_Vag + (1 + s_Trl + c_Num + c_Vag | Subject) + (1 + s_Trl + c_Num + c_Vag | f_Itm) ) -> m10 # no interaction in the randoms
lmer(data=dat, logRT ~ s_Trl + c_Num * c_Vag + (1 + s_Trl + c_Num * c_Vag | Subject) + (1 + s_Trl + c_Num + c_Vag | f_Itm) ) -> m11 # interaction in the subjects only * vs m10
# no converge below
lmer(data=dat, logRT ~ s_Trl + c_Num * c_Vag + (1 + s_Trl + c_Num + c_Vag | Subject) + (1 + s_Trl + c_Num * c_Vag | f_Itm) ) -> m12 # interaction in the items only no converge
# no converge below
lmer(data=dat, logRT ~ s_Trl + c_Num * c_Vag + (1 + s_Trl + c_Num * c_Vag | Subject) + (1 + s_Trl + c_Num * c_Vag | f_Itm) ) -> m13 # interaction in both randoms no converge
# no converge below
lmer(data=dat, logRT ~ s_Trl + c_Num * c_Vag + c_Qty + (1 + s_Trl + c_Num * c_Vag | Subject) + (1 + s_Trl + c_Num * c_Vag | f_Itm) ) -> m14 # interaction in both randoms, add Q, no converge
```

```{r, eval=TRUE, cache=TRUE, warning=TRUE, message=TRUE}
lmer(data=dat, logRT ~ s_Trl + c_Num * c_Vag + c_Qty + (1 + s_Trl + c_Num + c_Vag | Subject) + (1 + s_Trl + c_Num + c_Vag | f_Itm) ) -> m15 # no interaction in the randoms, add Q. *** vs m10
lmer(data=dat, logRT ~ s_Trl + c_Num * c_Vag * c_Qty + (1 + s_Trl + c_Num + c_Vag | Subject) + (1 + s_Trl + c_Num + c_Vag | f_Itm) ) -> m17 # no interaction in the randoms, add interaction Q ** vs m15
lmer(data=dat, logRT ~ s_Trl + c_Num * c_Vag * c_Qty + (1 + s_Trl + c_Num * c_Vag | Subject) + (1 + s_Trl + c_Num + c_Vag | f_Itm) ) -> m18 # interaction in the subjects only, keep interaction Q
lmer(data=dat, logRT ~ s_Trl + c_Num * c_Vag * c_Qty + (1 + s_Trl + c_Num * c_Vag | Subject) + (1 + s_Trl + c_Num * c_Vag | f_Itm) ) -> m19 # interaction in both randoms, keep interaction Q
# no converge below
lmer(data=dat, logRT ~ s_Trl + c_Num * c_Vag * c_Qty + c_Ord + (1 + s_Trl + c_Num * c_Vag | Subject) + (1 + s_Trl + c_Num * c_Vag | f_Itm) ) -> m20 # interaction in both randoms, keep interaction Q, add Ord
```

## invlam section

```{r, eval=TRUE, cache=TRUE, warning=TRUE, message=TRUE}
# no converge below
lmer(data=dat, invlam ~ s_Trl + c_Num * c_Vag * c_Qty + (1 + s_Trl + c_Num * c_Vag | Subject) + (1 + s_Trl + c_Num * c_Vag | f_Itm) ) -> m19.i # no converge
lmer(data=dat, invlam ~ s_Trl + c_Num * c_Vag * c_Qty + c_Ord + (1 + s_Trl + c_Num * c_Vag | Subject) + (1 + s_Trl + c_Num * c_Vag | f_Itm) ) -> m20.i # interaction in both randoms, keep interaction Q, add Ord
```

## Borderline response

```{r, eval=TRUE, cache=FALSE}
round(table(dat$Response) / nrow(dat) * 100, 0)
```

```{r, eval=TRUE, cache=FALSE}
round(table(dat$Vagueness, dat$Response) / nrow(dat) * 100, 0)
```

```{r, eval=TRUE, cache=FALSE}
# make dat be long format with respect to Response 
# now put it long so there are three rows for each Trial.s
b = melt(dat, measure.vars=c("Expected", "Near", "Far"))  
```

```{r, eval=TRUE, cache=FALSE, fig.width=4, fig.height=4}
# show overall distribution of responses
b2 <- summarySEwithin(b, measurevar='value', withinvars = 'variable')
b2$value=b2$value*100
b2$ci <- b2$ci * 100
ggplot(b2, aes(y=value, x=variable, fill=variable, ymin=value-ci, ymax=value+ci)) + 
  geom_bar(stat='identity', position='dodge') +
  scale_fill_grey("") + theme_bw() + theme(legend.key=element_blank(), aspect.ratio=1) + xlab("") + ylab("percent") +
  geom_errorbar(position=position_dodge(width=0.9), width=.2)
```

```{r, eval=TRUE, cache=FALSE}
# show distribution of responses broken down over number
b2 = summarySEwithin(b, measurevar="value", withinvars=c("Number","variable"))
b2$value=b2$value*100
b2$ci <- b2$ci * 100
ggplot(b2, aes(y=value, x=variable, fill=variable, ymin=value-ci, ymax=value+ci)) + 
  geom_bar(stat='identity', position='dodge') +
  scale_fill_grey("") + theme_bw() + theme(legend.key=element_blank(), aspect.ratio=1) + xlab("") + ylab("percent") +
  geom_errorbar(position=position_dodge(width=0.9), width=.2) +
  facet_grid(~Number)
```

```{r, eval=TRUE, cache=FALSE}
# show distribution of responses broken down over vagueness
b2 = summarySEwithin(b, measurevar="value", withinvars=c("Vagueness","variable"))
b2$value=b2$value*100
b2$ci <- b2$ci * 100
ggplot(b2, aes(y=value, x=variable, fill=variable, ymin=value-ci, ymax=value+ci)) + 
  geom_bar(stat='identity', position='dodge') +
  scale_fill_grey("") + theme_bw() + theme(legend.key=element_blank(), aspect.ratio=1) + xlab("") + ylab("percent") + 
  geom_errorbar(position=position_dodge(width=0.9), width=.2) +
  facet_grid(~Vagueness)
```

```{r, eval=TRUE, cache=FALSE}
# show distribution of responses broken down over number and vagueness
b2 = summarySEwithin(b, measurevar="value", withinvars=c("Number","Vagueness","variable"))
b2$value=b2$value*100
b2$ci <- b2$ci * 100
ggplot(b2, aes(y=value, x=variable, fill=variable, ymin=value-ci, ymax=value+ci)) + 
  geom_bar(stat='identity', position='dodge') +
  scale_fill_grey("") + theme_bw() + theme(legend.key=element_blank(), aspect.ratio=1) + xlab("") + ylab("percent") + 
  geom_errorbar(position=position_dodge(width=0.9), width=.2) +
  facet_grid(~Number+Vagueness)
```

```{r, eval=TRUE, cache=FALSE}
b2 = summarySEwithin(b, measurevar="value", withinvars=c("Number","Vagueness","f_Itm","variable"))
b2$value=b2$value*100
b2$ci <- b2$ci * 100
ggplot(b2, aes(y=value, x=variable, fill=variable, ymin=value-ci, ymax=value+ci)) + 
  geom_bar(stat='identity', position='dodge') +
  scale_fill_grey("") + theme_bw() + theme(legend.key=element_blank(), aspect.ratio=.75) + xlab("") + ylab("percent") +
  geom_errorbar(position=position_dodge(width=0.9), width=.2) +
  facet_grid(~Number+Vagueness~f_Itm)
```

```{r, eval=TRUE, cache=TRUE, warning=TRUE, message=TRUE}
lmer.bd.mm0 <- glmer(data=dat, 
                     Near ~ 
                       c_Num * c_Vag * c_Itm 
                     + (c_Num * c_Vag + c_Itm | Subject), 
                     family=binomial, 
                     control=glmerControl(optimizer="bobyqa"))
lmer.bd.summary <- summary(lmer.bd.mm0)
```

```{r, results='asis', eval=TRUE, cache=FALSE}
print(xtable(digits=3,lmer.bd.summary$coefficients,
             caption="The most maximal model of bd that converges"),caption.placement="top", table.placement='htbp', type='html')
```

```{r, fig.width=4, fig.height=4, eval=TRUE, cache=FALSE}
qqPlot(residuals(lmer.bd.mm0))
```

```{r, eval=TRUE, cache=FALSE, echo=TRUE}
beep("fanfare")
```

### 2016-07-29 experiments/e2/archive/model_ols_1_with_baayen4plot_model_criticism.Rmd
archiving old e2 Rmd files

---
output:
  html_document:
    toc: true
    <!--toc_float: true-->
    mathjax: default
    template: default
    code_folding: hide
    <!--theme: default-->
---

## ols model v4, RecSqt, in progress

```{r}
load("dd.Rda")
suppressPackageStartupMessages(require(rms))
suppressPackageStartupMessages(require(languageR))
options(width=140)
```

Pairs plot

```{r, fig.width=7, fig.height=7}
# pairs plot
pairs(dd[,c("RT_RecSqt", "Number", "Vagueness", "RTprev_RecSqt")], pch = 19, cex=.075)
```

Build the model

```{r}
## do ols model
# define datadist
dd.dd = datadist(dd)
options(datadist = "dd.dd")
# build model
v4 <- ols(data=dd, 
          RT_RecSqt ~ 
            c_Vag + c_Num + c_Qty + c_Ord + 
            c_Num:c_Vag:c_Qty+
            item_mean_ratio + 
            #c_Vag:c_Num + 
            #c_Vag:c_Num:item_mean_ratio + 
            # pol(s_Trl, 3) +
            s_Trl+
            # pol(RTprev_RecSqt,3) +
            RTprev_RecSqt +
            nchar_instr
            # cell+
            # measurement,
            ,
          x=T, y=T )
```

```{r}
# print out p value for likelihood ratio with df
cat("lik.ratio p value, whether the model as a whole is explanatory: p =", 1-pchisq(v4$stats[2],v4$stats[3]))
# print out R2 for the model
cat("R^2 = ", v4$stats[4])
```

Show the model tables

```{r}
# show model
v4
# show model summary
summary(v4)
# do anova 
an1 = anova(v4)
# show anova
an1
```

plot 'partial effects'

```{r, fig.width=12, fig.height=5}
## plot partial effects
# Compute Predicted Values and Confidence Limits
p1=Predict(v4)
# plot 'partial effects'
plot(p1, anova=an1, pval=T, aspect=1, main="ols model v4")
```

```{r}
# add predicted RT_RecSqt to dd
dd$RT_Predicted <- predict (v4)
```

Plot model coefficients and ci's

```{r}
par(las=2, mar=c(14,6,1,1))
y=coef(v4) # with intercept
n=length(y)
y0=confint(v4)[1:n,1]
y1=confint(v4)[1:n,2]

y=coef(v4)[-1] # omit intercept
n=length(y)
y0=confint(v4)[-1,1]
y1=confint(v4)[-1,2]

plot(y, xaxt="n", xlab="", ylab="RT_RecSqt\n", pch=19, ylim=extendrange(y, f=.10), main="Speed")
abline(h=0)
axis(1, labels=names(y) , at=1:n)
grid()
segments(x0=1:n, x1=1:n, y0, y1, lwd=2)
```

Collinearity is assessed by the condition number $k$.\
The greater the collinearity, the closer the matrix of predictors is is to becoming _singular_.\
$k$ estimates the extent to which a matrix is $\text{singular}$\
$0\ldots6$ no collinearity\
$\text{around} 15$ medium collinearity\
$>30$ indicates potentially harmful collinearity\

```{r}
cat("k is:",
collin.fnc(dd[,c("s_Trl", "c_Num", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_RecSqt", "nchar_instr", "measurement")])$cnumber
)
```

```{r, fig.width=6, fig.height=5}
par(mar=c(1.1,3.1,1.1,1.1), pty='s')
plot(varclus(as.matrix(dd[,c("s_Trl", "c_Num", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_RecSqt", "nchar_instr", "measurement")])))
```

Model criticism baayen plots

```{r, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

par(mfrow=c(1,4), pty='s')
# create scaled residuals
v4$rstand = as.vector(scale(resid(v4)))
# plot scaled residuals density
plot(density(v4$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(v4$rstand, cex=.5)
qqline(v4$rstand)
# plot standardised residuals versus fitted values
plot(v4$rstand ~ fitted(v4), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))
# create dffits
dffits=abs(resid(v4, "dffits"))
# plot dffits
plot(dffits, type='h')
```

```{r}
# here no overly influential
w = which.influence(fit = v4, cutoff = 0.4)
w
```

Model validation

Validation tests overfitting using bootstrap.

Bootstrapping draws as a bootstrap sample the same number of samples as the original data, at random with replacement, from the original data.

Then fit the model to the data in the bootstrap sample, and use this model to predict the reaction times for the original full data (which contains many samples that are new to the bootstrap model).

Then compare the goodness of fit of the bootstrap model with the goodness of fit of the original model.

Averaged over a large number of bootstrap models these comparisons of goodness of fit  reveal to what extent the original model overfits the data.

Function validate() in rms does this _re-sampling validation_.

```{r, val_1, echo=TRUE}
# argument B is the number of bootstrap runs
# argument pr is whether to print the results of each run
v_1 <- validate(v4, bw = T, B = 200, pr=FALSE, estimates=TRUE)
```

```{r}
# in print.validate, B is the number of re-samples to show outocmes for each resample
print(v_1, B=0)
```

```{r}
# show how often each number of variables kept was observed across the 200 runs
xtabs(~rowSums(attr(v_1,"kept")))
```

### 2016-07-29 experiments/e2/archive/plot_boxplot_subjects_and_items.Rmd
archiving old e2 Rmd files

Show how transformations of RT affect distribution of times, and how they affect which times are outliers.

```{r, fig.width=20, fig.height=5}
# boxplot subjects and items

load("dd.Rda")
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(ggplot2))

# allow reference to transformation
subdata = melt(dd,
               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
               variable.name="transformation",
               value.name="value")

# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ transformation, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
  xlab("") + ylab("")
```

### 2016-07-29 experiments/e2/archive/plot_fast_and_slow_subjects_and_items_RecSqt.Rmd
archiving old e2 Rmd files

Show how mean times for individual subjects and items vary with respect to the grand mean RT_RecSqt.

```{r, fig.width=10, fig.height=5}
# Show how mean times for individual subjects and items vary with respect to the grand mean RT_RecSqt.

suppressMessages(library(plyr))
suppressMessages(library(lattice))
suppressMessages(library(latticeExtra))

subs <- ddply(dd, .(level=Subject), summarise, 
              effect="subject", emean=mean(dd$RT_RecSqt), lmean=mean(RT_RecSqt), ldiff=mean(RT_RecSqt)-mean(dd$RT_RecSqt),
              lsign=ifelse(mean(RT_RecSqt)-mean(dd$RT_RecSqt)<0,"slow","fast"))
itms <- ddply(dd, .(level=Item), summarise, effect="item", emean=mean(dd$RT_RecSqt), lmean=mean(RT_RecSqt),
              ldiff=mean(RT_RecSqt)-mean(dd$RT_RecSqt), lsign=ifelse(mean(RT_RecSqt)-mean(dd$RT_RecSqt)<0,"slow","fast"))
si <- rbind(subs,itms)
si$effect <- factor(si$effect, levels=c("subject", "item"))
si$lsign <- factor(si$lsign, levels=c("slow","fast"))

my1 <- xyplot(data=si, lmean~level | effect, 
              aspect=1,
              layout=c(2,1),
              ylim = extendrange(si$lmean, f = 0.10),
              par.settings = list(trellis.par.set(superpose.symbol = list(col = c("blue","red"), pch=19), superpose.line=list(col = c("blue","red")))),
              groups = si$lsign,
              xlab = NULL,
              ylab = "RT_RecSqt: reciprocal of square root of RT in ms\n",
              key=list(text=list(levels(si$lsign)), lines=list(lty=c(1,1), col=trellis.par.get('superpose.line')$col, pch=c(19,19), type='o'), divide=1),
              mean = unique(si$emean),
              scales = list(x=list(relation="free", at=list(15.5, 32.5), labels=list("subject number", "item identifier"), cex=1 ) ),
              panel=function(x,y,type,subscripts,groups=groups,...){
                panel.xyplot(x,y,type='p',groups=groups, subscripts=subscripts,...)
                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, groups=groups, subscripts=subscripts, code=3, length=0, col=trellis.par.get('superpose.line')$col[groups][subscripts], ...)
                panel.abline(h=unique(si$emean), col="lightgrey",lty=1)
                panel.text(x, y=(ifelse(y<unique(si$emean),y-.001,y+.001)), as.numeric(si$level)[subscripts][panel.number()==1], cex=.75)
                panel.text(x, y=(ifelse(y<unique(si$emean),y-.001,y+.001)), as.character(si$level)[subscripts][panel.number()==2], cex=.75)
              })

my2 <- xyplot(data=si, ldiff~level | effect, 
              aspect=1, type='n',
              ylab="\nDeviation from mean RT_RecSqt (in ms)",
              ylim=extendrange(si$ldiff,f=0.10))

print(doubleYScale(my1, my2, add.axis=TRUE, add.ylab2=TRUE, use.style=FALSE))
```

### 2016-07-29 experiments/e2/archive/plot_practice_and_fatigue_effects.Rmd
archiving old e2 Rmd files

Show how practice and fatigue effects vary over subjects

```{r, message=F, results='hide'}
if(require(ggplot2)){}
load("dd.Rda")
```

```{r, eval=TRUE, fig.width=20, fig.height=10, cache=FALSE, echo=TRUE}
# Show how practice/fatigue effects vary over subjects
ggplot(dd, aes(y=RT_RecSqt, x=Trial)) + facet_wrap(~Subject,nrow=3) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial", breaks=NULL) + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1) 
```

Show practice and fatigue effects vary over items.

```{r, eval=TRUE, fig.width=9, fig.height=4, cache=FALSE, echo=TRUE}
# Show how practice/fatigue effects vary over items
ggplot(dd, aes(y=RT_RecSqt, x=Trial)) + facet_grid(~Item) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial") + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1) 
```

### 2016-07-29 experiments/e2/archive/plot_transformations_against_normal_equivalents.Rmd
archiving old e2 Rmd files

--- 2016-07-28 experiments/e4/writeup/e4.Rmd
@@
----
-title: "Experiment 4 of 4"
-author: "Matt Green"
-output:
-  html_document:
-    toc: true
-    <!--toc_float: true-->
-    mathjax: default
-    template: default
-    code_folding: hide
-    theme: default
----  
-# Experiment 4
-
-Each subject saw 192 trials, being 3 cycles of 64 trials. There were 40 subjects. There were 7680 rows in the full data.
-
-# Trial Procedure
-First an instruction was presented until subjects dismissed it with a key press.
-Then a grey screen was presented for 1 second.
-Then a prime was presented for 2 seconds.
-Then a grey screen was presented for 1 second.
-Then the squares with dots were presented until the subject responded, with a 10 second timeout if they did not respond.
-
-<figure>
-<img src="./flow-of-experiment4.pdf" alt="Mountain View" style="width:480px;height:300px;">
-<figcaption>Flow of experiment 4.</figcaption>
-</figure>
-
-# Describe the data
-
-Here is a description of each column in the raw data.
-
-* **Sub** identifier for subject. There are 40 subjects.
-* **Trl** identifier for the trial. There are 64 Trials for each subject.
-* **Itm** identifier for item.
-There are 4 items.
-An item is a set of Selections for the dots in the squares, e.g., Item 1 is (6,15,24) in any order.
-This should be included as a (possibly ordered) factor because the task will probably take longer for more dots.
-Include this in the final model if it does increase rt.
-Table below shows the dots for each item.
-* **Cnd** just a single vector for the combination of vagueness and selection. Useful for some plotting routines that don't accept pairs of grouping variables. 
-* **Ord** whether the order was for example (6,15,24) or (24,14,6).
-This matters because when the smaller quantity is on the left people are quicker than when the smaller quantity is on the right (i.e, SNARC effect).
-It is a control variable rather than a direct predictor.
-It could be left out of the main model assuming it is non-significant.
-* **Qty** whether the instruction identifies bigger or smaller quantities.
-This is a factor with 2 levels sum coded.
-An example of bigger is _"Choose a square with more dots than the target"_.
-An example of smaller is _"Choose a square with fewer dots than the target"_.
-* **Prm** how many dots there were in the prime (prime is _target_ in the instructions to the participants). Primes were 6 10 16 20 24 26 30 34 36 40 44 50 54.
-* **Lft** how many dots there were in the left square
-* **Mid** how many dots there were in the middle square
-* **Rgt** how many dots there were in the right-hand side square
-* **key** which key the subject pressed as a response on a trial.
-* **rt** raw response time in milliseconds
-* **Ins** a character array for the instruction that was presented to the subject on that trial. Table below shows an example of an instruction in each condition.
-* **V** factor for vagueness with 2 levels sum coded
-* **S** factor for selection task with 2 levels sum coded
-
-Here is a table of instructions.
-
-<style type="text/css">
-.table {width: 50%;}
-</style>
-
-vagueness | selection | instruction
-----------|-----------|--------------------------------------------------------------------
-crisp     | match     | Choose a square with the same number of dots as the target  
-crisp     | compare   | Choose a square with fewer dots than the target 
-vague     | match     | Choose a square with about the same number of dots as the target 
-vague     | compare   | Choose a square with far fewer dots than the target
-
-Item   | dots
--------|-----------
-Item 1 | (6,15,24)
-Item 2 | (16,25,34)
-Item 3 | (26,35,44)
-Item 4 | (36,45,54)
-
-```{r, startClean, echo=FALSE}
-# start clean
-rm(list=ls())
-```
-
-```{r, docOpts, echo=FALSE, cache=FALSE}
-# Knitr doc options.
-library(knitr)
-knit_hooks$set(crop=hook_pdfcrop)
-#options(tidy=TRUE)
-#options(replace.assign=TRUE, width=50)
-```
-
-```{r, loadLibraries, echo=FALSE}
-# Load libraries.
-suppressPackageStartupMessages(library(MASS))
-suppressPackageStartupMessages(library(knitr))
-suppressPackageStartupMessages(library(ggplot2))
-suppressPackageStartupMessages(library(lme4))
-suppressPackageStartupMessages(library(xtable))
-suppressPackageStartupMessages(library(plyr))
-suppressPackageStartupMessages(library(reshape2))
-suppressPackageStartupMessages(library(lattice))
-suppressPackageStartupMessages(library(LMERConvenienceFunctions))
-suppressPackageStartupMessages(library(car))
-suppressPackageStartupMessages(library(grid))
-suppressPackageStartupMessages(library(lmerTest))
-```
-
-Source additional R files with functions for computing confidence intervals.
-
-```{r, sourceFunctions}
-source("functions/CIfunctions.R")
-```
-
-Get data from the raw per-subject files.
-
-```{r, getData}
-#if( file.exists('dd.Rda') ) load('dd.Rda') else {
-#system(wait=TRUE, paste("head -n 1 ", "../experimentCode","/output/", "subject01.data ", "> ", "../data/data.txt", sep=""))
-#system(wait=TRUE, paste("tail -q -n +2 ", "../experimentCode", "/output/", "*.data ",  ">> ", "../data/data.txt", sep=""))
-dat <- read.delim(file="../data/data.txt", header=T, stringsAsFactors=FALSE)
-
-# declare local variables
-Selection_of_valid_subjects <- 40
-Selection_of_rows <- 7680
-Selection_of_trials_per_subject <- Selection_of_rows / Selection_of_valid_subjects # 192
-
-# sort out borderline responses into expected near and far
-# How many squares in the square they chose? 
-dat$RESPONSE <- as.character(dat$key)
-for (row in 1 : nrow(dat) ) {
-  switch(dat[row,'RESPONSE'],
-         'L' = {dat[row, 'choice'] <- dat[row, 'Lft']},
-         'M' = {dat[row, 'choice'] <- dat[row, 'Mid']},
-         'R' = {dat[row, 'choice'] <- dat[row, 'Rgt']}         
-  )
-}
-  dat$Item <- dat$Itm
-  dat$Itm <- NULL
-dat$crossed=paste('Con',dat$Cnd,':Quan',dat$Qty,':Item',dat$Item,sep="")
-
-dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseExpected'] <-  6 
-dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseNear']     <- 15  
-dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseFar']      <- 24
-
-dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseExpected'] <- 16
-dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseFar']      <- 34
-
-dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseExpected'] <- 26
-dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseFar']      <- 44
-
-dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseExpected'] <- 36
-dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseFar']      <- 54
-
-dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseExpected'] <- 24
-dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseNear']     <- 15 
-dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseFar']      <-  6
-
-dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseExpected'] <- 34
-dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseFar']      <- 16
-
-dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseExpected'] <- 44
-dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseFar']      <- 26
-
-dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseExpected'] <- 54
-dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseFar']      <- 36
-
-dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseExpected'] <-  6 
-dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseNear']     <- 15  
-dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseFar']      <- 24
-
-dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseExpected'] <- 16
-dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseFar']      <- 34
-
-dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseExpected'] <- 26
-dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseFar']      <- 44
-
-dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseExpected'] <- 36
-dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseFar']      <- 54
-
-dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseExpected'] <- 24
-dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseNear']     <- 15 
-dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseFar']      <-  6
-
-dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseExpected'] <- 34
-dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseFar']      <- 16
-
-dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseExpected'] <- 44
-dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseFar']      <- 26
-
-dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseExpected'] <- 54
-dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseFar']      <- 36
-
-dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseExpected'] <-  6 
-dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseNear']     <- 15  
-dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseFar']      <- 24
-
-dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseExpected'] <- 16
-dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseFar']      <- 34
-
-dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseExpected'] <- 26
-dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseFar']      <- 44
-
-dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseExpected'] <- 36
-dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseFar']      <- 54
-
-dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseExpected'] <- 24
-dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseNear']     <- 15 
-dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseFar']      <-  6
-
-dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseExpected'] <- 34
-dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseFar']      <- 16
-
-dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseExpected'] <- 44
-dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseFar']      <- 26
-
-dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseExpected'] <- 54
-dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseFar']      <- 36
-
-dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseExpected'] <-  6 
-dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseNear']     <- 15  
-dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseFar']      <- 24
-
-dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseExpected'] <- 16
-dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseFar']      <- 34
-
-dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseExpected'] <- 26
-dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseFar']      <- 44
-
-dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseExpected'] <- 36
-dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseFar']      <- 54
-
-dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseExpected'] <- 24
-dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseNear']     <- 15 
-dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseFar']      <-  6
-
-dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseExpected'] <- 34
-dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseFar']      <- 16
-
-dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseExpected'] <- 44
-dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseFar']      <- 26
-
-dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseExpected'] <- 54
-dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseFar']      <- 36
-
-dat$isResponseExpected <- dat$choice == dat$ResponseExpected
-dat$isResponseNear <- dat$choice == dat$ResponseNear
-dat$isResponseFar <- dat$choice == dat$ResponseFar
-
-for ( row in 1:nrow(dat) ) {
-  dat[row, 'switchResponse'] <- 
-    ifelse(dat[row, 'isResponseExpected']==TRUE, 'Expected', 
-           ifelse(dat[row, 'isResponseNear']==TRUE, 'Near', 'Far') ) 
-}
-
-# SUBJECT
-# ensure Subject is a factor 
-dat$Subject=factor(paste("s",sprintf("%02d",dat$Sub),sep=""))
-
-# TRIAL
-# Trial for subject, 1 to 192
-dat$Trial = rep(x = 1:Selection_of_trials_per_subject, times = Selection_of_valid_subjects)
-
-# make a centred Trial for modeling
-dat$c_Trl <-dat$Trial - mean(dat$Trial)
-
-# make a scaled Trial for modelling
-dat$s_Trl <- as.numeric(scale(dat$Trial))
-
-# ID
-# id is a unique identifier for the 7680 row data
-dat$id <- factor(paste(paste(dat$Subject), 
-                       paste("t", sprintf("%03d", dat$Trial), sep="") , sep=":"))
-
-# ITEM
-# create a centred numeric item variable for modeling
-dat$c_Itm <- ifelse(dat$Item==1, -.75, 
-                    ifelse(dat$Item==2, -.25,
-                           ifelse(dat$Item==3, .25, .75)))
-
-# make Item be a factor and assign labels
-dat$Item <- factor(dat$Item, levels=c(1,2,3,4), labels=c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
-
-# add ratios for Item
-item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54) 
-# 0.2500000 0.4705882 0.5909091 0.6666667
-item_range_ratio_scaled = as.vector(scale(c(6 / 24, 16 / 34, 26 / 44, 36 / 54)))
-# -1.3441995 -0.1316642  0.5297187  0.9461450
-item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))) 
-# 0.5125000 0.6876471 0.7691558 0.8166667
-item_mean_ratio_scaled = as.vector(scale(c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))))) 
-# -1.37582241 -0.06614191  0.54334858 0.89861574
-
-dat[dat$Item == "06:15:24", "item_range_ratio"] <-  0.2500000
-dat[dat$Item == "16:25:34", "item_range_ratio"] <-  0.4705882
-dat[dat$Item == "26:35:44", "item_range_ratio"] <-  0.5909091
-dat[dat$Item == "36:45:54", "item_range_ratio"] <-  0.6666667
-
-dat[dat$Item == "06:15:24", "item_range_ratio_scaled"] <-  -1.3441995
-dat[dat$Item == "16:25:34", "item_range_ratio_scaled"] <-  -0.1316642
-dat[dat$Item == "26:35:44", "item_range_ratio_scaled"] <-   0.5297187  
-dat[dat$Item == "36:45:54", "item_range_ratio_scaled"] <-   0.9461450
-
-dat[dat$Item == "06:15:24", "item_mean_ratio"] <-  0.5125000 
-dat[dat$Item == "16:25:34", "item_mean_ratio"] <-  0.6876471 
-dat[dat$Item == "26:35:44", "item_mean_ratio"] <-  0.7691558
-dat[dat$Item == "36:45:54", "item_mean_ratio"] <-  0.8166667
-
-dat[dat$Item == "06:15:24", "item_mean_ratio_scaled"] <-   -1.37582241 
-dat[dat$Item == "16:25:34", "item_mean_ratio_scaled"] <-   -0.06614191 
-dat[dat$Item == "26:35:44", "item_mean_ratio_scaled"] <-    0.54334858 
-dat[dat$Item == "36:45:54", "item_mean_ratio_scaled"] <-    0.89861574
-
-# VAGUENESS
-# Create a factor coding for Vagueness
-dat[ dat$Cnd==1 , 'Vagueness'] <- 'Vague'
-dat[ dat$Cnd==2 , 'Vagueness'] <- 'Crisp'
-dat[ dat$Cnd==3 , 'Vagueness'] <- 'Vague'
-dat[ dat$Cnd==4 , 'Vagueness'] <- 'Crisp'
-dat$Vagueness <- as.factor(dat$Vagueness)
-
-# manually center Vagueness
-dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)
-
-dat$Selection = ""
-dat[dat$Cnd %in% c(1,2), "Selection"] <- "matching"
-dat[dat$Cnd %in% c(3,4), "Selection"] <- "comparison"
-dat$Selection.sum <- ifelse(dat$Selection=="matching", .5, -.5)
-dat$Selection <- as.factor(dat$Selection)
-
-# remove redundant confusing columns
-dat$Sub <- NULL
-dat$Trl <- NULL
-
-# ORDER
-# give the levels of Order meaningful names
-dat$Order <- factor(dat$Ord, levels=c(1,2), labels=c('LtoR','RtoL'))
-
-# make a manually centred Order
-dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)
-
-# QUANTITY
-# give the levels of Quantity meaningful names
-dat$Quantity <- factor(dat$Qty, levels=c(1,2), labels=c('Small','Large'))
-
-# make a manually centred Quantity
-dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)
-
-# INSTRUCTION
-# add Selection of characters in the instruction # 29 30 34 36 38
-dat$nchar_instr = nchar(as.character(dat$Ins))
-dat$nchar_instr_scaled = as.vector(scale(nchar(as.character(dat$Ins)), scale=TRUE))
-
-# make Instruction be a factor (17 levels)
-dat$Ins <- as.factor(dat$Ins)
-
-# RT
-# add transformations of RT
-dat$RT <- dat$rt; dat$rt <-  NULL
-dat$RT_RecSqd    <- 1/dat$RT^2
-dat$RT_Rec       <- 1/dat$RT
-dat$RT_RecSqt    <- 1/sqrt(dat$RT) 
-dat$RT_log       <- log(dat$RT)
-dat$RT_sqt       <- sqrt(dat$RT)
-dat$RT_raw       <- dat$RT
-dat$RT_sqd       <- dat$RT^2
-
-save(dat, file="dat.Rda")
-
-dd <- dat
-
- # lose impossible trials
-  dd <- dat
-  dd$RT[dd$RT > 9000 ] <- NA
-  dd <- dd[complete.cases(dd),]
-  row.names(dd) <- NULL
-
-# add preceding RT: because we removed impossible trials, the value for preceding RT for a trial following an impossible trial is the value of the trial that preceded the impossible trial.
-dd$RTprev <- NA
-for (s in levels(dd$Subject)) {
-  nrows = nrow(dd[dd$Subject==s,])
-  for (i in 1:nrows) {
-    if (i==1) {
-      dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]
-    }
-    else
-      dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1]
-  }
-}
-
-# add transformations of previous RT
-dd$RTprev_RecSqd    <- 1/dd$RTprev^2
-dd$RTprev_Rec       <- 1/dd$RTprev
-dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev) 
-dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
-dd$RTprev_log       <- log(dd$RTprev)
-dd$RTprev_sqt       <- sqrt(dd$RTprev)
-dd$RTprev_raw       <- dd$RTprev
-dd$RTprev_sqd       <- dd$RTprev^2
-  # save dd
-  save(dd, file="dd.Rda")
-#}
-```
-
 Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions.
@@
 dat1 <- dat_transforms                                       # nrow 7680
-dat2 <- subset(dat_transforms, RT<9000)              # nrow 7677
-dat2$dfSubset <- "RT<9000"
+dat2 <- subset(dat_transforms, RT>1)                         # nrow 7678
+dat2$dfSubset <- "RT>1"
+dat3 <- subset(dat_transforms, RT>1 & RT<59999)              # nrow 7677
+dat3$dfSubset <- "RT>1&RT<59999"
+dat4 <- subset(dat_transforms, RT>1 & RT<59999 & RT<30001)   # nrow 7658
+dat4$dfSubset <- "RT>1&RT<30001"
 
 # bring the dfs back into main df
-dat_transforms <- rbind(dat1,dat2)
-dat_transforms$dfSubset <- factor(dat_transforms$dfSubset, levels=c("original", "RT<9000"))
+dat_transforms <- rbind(dat1, dat2, dat3, dat4)
+dat_transforms$dfSubset <- factor(dat_transforms$dfSubset, levels=c("original", "RT>1", "RT>1&RT<59999", "RT>1&RT<30001"))
 
@@
 
-temp <- melt(data=dat_transforms, 
-             id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Selection", "Quantity", "Order"),
+temp <- melt(dat_transforms, 
+             id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Number", "Quantity", "Order"),
              measure.vars=c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd"),
@@
   geom_freqpoly() + 
-  facet_wrap(dfSubset~transformation, scales="free", nrow=2) +
+  facet_wrap(dfSubset~transformation, scales="free", nrow=4) +
   theme_bw() + theme(aspect.ratio=1) + scale_color_manual(values=c("blue","red"))
 ```
-
-Show how transformations of RT affect distribution of times, and how they affect which times are outliers.
-
-```{r, fig.width=20, fig.height=5}
-# boxplot subjects and items
-
-load("dd.Rda")
-suppressPackageStartupMessages(library(reshape2))
-suppressPackageStartupMessages(library(ggplot2))
-
-# allow reference to transformation
-subdata = melt(dd,
-               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
-               variable.name="transformation",
-               value.name="value")
-
-# boxplots for subjects and items separately for each dependent variable
-ggplot(subdata) + 
-  facet_wrap(~ transformation, scales="free", ncol=4) +
-  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
-  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
-  theme_bw() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
-  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
-  xlab("") + ylab("")
-```
-
-Show how mean times for individual subjects and items vary with respect to the grand mean RT.
-
-```{r, fig.width=10, fig.height=5}
-# Show how mean times for individual subjects and items vary with respect to the grand mean RT.
-
-suppressMessages(library(plyr))
-suppressMessages(library(lattice))
-suppressMessages(library(latticeExtra))
-
-subs=ddply(dd, .(level=Subject), summarise, effect="subject", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
-itms=ddply(dd, .(level=Item), summarise, effect="item", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
-si=rbind(subs,itms)
-si$effect <- factor(si$effect, levels=c("subject", "item"))
-si$lsign <- factor(si$lsign, levels=c("slow","fast"))
-
-my1 <- xyplot(data=si, lmean~level | effect, 
-              aspect=1,
-              layout=c(2,1),
-              ylim = extendrange(si$lmean, f = 0.10),
-              par.settings=list(trellis.par.set(superpose.symbol = list(col = c("blue","red"), pch=19), superpose.line=list(col = c("blue","red")))),
-              groups=si$lsign,
-              xlab=NULL,
-              ylab="RT (in ms)",
-              key=list(text=list(levels(si$lsign)), lines=list(lty=c(1,1), col=trellis.par.get('superpose.line')$col, pch=c(19,19), type='o'), divide=1),
-              mean=unique(si$emean),
-              scales=list(x=list(relation="free", at=list(15.5, 32.5), labels=list("subject number", "item identifier"), cex=1 ), 
-                          y=list(at=c(0, 2000, unique(si$emean), 4000, 6000, 8000,  10000), labels=c(0, 2000,  "Mean", 4000,  6000, 8000, 10000), cex=.9) ),
-              panel=function(x,y,type,subscripts,groups=groups,...){
-                panel.xyplot(x,y,type='p',groups=groups, subscripts=subscripts,...)
-                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, groups=groups, subscripts=subscripts, code=3, length=0, col=trellis.par.get('superpose.line')$col[groups][subscripts], ...)
-                panel.abline(h=unique(si$emean), col="lightgrey",lty=1)
-                panel.text(x, y=(ifelse(y<unique(si$emean),y-200,y+200)), as.numeric(si$level)[subscripts][panel.number()==1], cex=.75)
-                panel.text(x, y=(ifelse(y<unique(si$emean),y-200,y+200)), as.character(si$level)[subscripts][panel.number()==2], cex=.75)
-              })
-
-my2 <- xyplot(data=si, ldiff~level | effect, 
-              aspect=1, type='n',
-              ylab="Deviation from mean RT (in ms)",
-              ylim=extendrange(si$ldiff,f=0.10))
-
-print(doubleYScale(my1, my2, add.axis=TRUE, add.ylab2=TRUE, use.style=FALSE))
-```
-
-Show how practice and fatigue effects vary over subjects
-
-```{r, message=F, results='hide'}
-if(require(ggplot2)){}
-load("dd.Rda")
-```
-
-```{r, eval=TRUE, fig.width=20, fig.height=10, cache=FALSE, echo=TRUE}
-# Show how practice/fatigue effects vary over subjects
-ggplot(dd, aes(y=RT_log, x=Trial)) + facet_wrap(~Subject,nrow=3) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial", breaks=NULL) + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1)
-```
-
-Show practice and fatigue effects vary over items.
-
-```{r, eval=TRUE, fig.width=9, fig.height=4, cache=FALSE, echo=TRUE}
-# Show how practice/fatigue effects vary over items
-ggplot(dd, aes(y=RT_log, x=Trial)) + facet_grid(~Item) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial") + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1)
-```
-
-# Show means.
-
-Main effects
-
-```{r}
-# prep main effects rts data frame
-vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness")));names(vrts)[3]<-"Levels"
-nrts <- cbind(Effect="Selection",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection")));names(nrts)[3]<-"Levels"
-qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Quantity")));names(qrts)[3]<-"Levels"
-orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Order")));names(orts)[3]<-"Levels"
-irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Item")));names(irts)[3]<-"Levels"
-rts=rbind(vrts,nrts,qrts,orts)
-rts$mins=rts$RT_log-rts$ci
-rts$maxs=rts$RT_log+rts$ci
-
-# plot main effects
-ggplot(rts, aes(y=RT_log, x=Levels, group=1, ymin=mins, ymax=maxs)) + 
-  geom_line() +
-  geom_errorbar(width=.1) +
-  geom_point(cex=3) + 
-  facet_grid(~Effect, scale="free_x") + 
-  theme_bw() +
-  theme(aspect.ratio=1) 
-```
-
-Main effects over items
-
-```{r, message=FALSE, fig.width=10, fig.height=3}
-vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Vagueness")));names(vrts)[4]<-"Levels"
-nrts <- cbind(Effect="Selection",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Selection")));names(nrts)[4]<-"Levels"
-qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Quantity")));names(qrts)[4]<-"Levels"
-orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Order")));names(orts)[4]<-"Levels"
-
-rts=rbind(vrts,nrts,qrts,orts)
-rts$mins=rts$RT_log-rts$ci
-rts$maxs=rts$RT_log+rts$ci
-
-rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
-
-ggplot(rts, aes(y=RT_log, x=item_mean_ratio, group=Levels, ymin=mins, ymax=maxs, shape=Levels, fill=Levels, col=Levels)) +
-  scale_shape_manual(values=c(21,21,22,22,23,23,24,24)) +
-  scale_fill_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
-  scale_color_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
-  geom_line() +
-  geom_errorbar(width=.01) +
-  geom_point(cex=3) + 
-  facet_grid(~Effect) + 
-  theme_bw() +
-  theme(aspect.ratio=1, 
-        panel.grid=element_blank()) +
-  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))
-
-```
-
-2 way interactions ignoring item
-
-```{r}
-# prep 2 ways rts data frame
-rts1 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Selection")); names(rts1)[1] <- "F1"; names(rts1)[2] <- "F2"
-rts1$Effect <- rep("Vagueness:Selection",4)
-rts2 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Quantity")); names(rts2)[1] <- "F1"; names(rts2)[2] <- "F2"
-rts2$Effect <- rep("Vagueness:Quantity",4)
-rts3 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Order")); names(rts3)[1] <- "F1"; names(rts3)[2] <- "F2"
-rts3$Effect <- rep("Vagueness:Order",4)
-rts5 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection","Quantity")); names(rts5)[1] <- "F1"; names(rts5)[2] <- "F2"
-rts5$Effect <- rep("Selection:Quantity",4)
-rts6 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection","Order")); names(rts6)[1] <- "F1"; names(rts6)[2] <- "F2"
-rts6$Effect <- rep("Selection:Order",4)
-rts8 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Quantity","Order")); names(rts8)[1] <- "F1"; names(rts8)[2] <- "F2"
-rts8$Effect <- rep("Quantity:Order",4)
-rts <- rbind(rts1,rts2,rts3,rts5,rts6,rts8)
-rts <- subset(rts, select=c(Effect, F1, F2, RT_log, RT_log_norm, sd, se, ci))
-rts$mins=rts$RT_log-rts$ci
-rts$maxs=rts$RT_log+rts$ci
-```
-
-```{r,fig.width=10, fig.height=3}
-# plot 2 way interactions
-ggplot(rts, aes(y=RT_log, x=F2, group=F1, ymin=mins, ymax=maxs, pch=F1)) + 
-  geom_line() +
-  geom_errorbar(width=.1) +
-  geom_point(cex=3) + 
-  facet_wrap(~Effect, scale="free_x", nrow=1) + 
-  theme_bw() +
-  theme(aspect.ratio=1) +
-  theme(legend.title=element_blank()) +
-  xlab("Levels")
-```
-
-Plot 2 way interactions over items
-
-```{r, message=FALSE}
-# prep 2 ways over items rts data frame
-rts1 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Vagueness","Selection")); names(rts1)[2] <- "F1"; names(rts1)[3] <- "F2"
-rts1$E1 <- "Vagueness"
-rts1$E2 <- "Selection"
-rts2 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Vagueness","Quantity")); names(rts2)[2] <- "F1"; names(rts2)[3] <- "F2"
-rts2$E1 <- "Vagueness"
-rts2$E2 <- "Quantity"
-rts3 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Vagueness","Order")); names(rts3)[2] <- "F1"; names(rts3)[3] <- "F2"
-rts3$E1 <- "Vagueness"
-rts3$E2 <- "Order"
-rts4 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Quantity", "Selection")); names(rts4)[2] <- "F1"; names(rts4)[3] <- "F2"
-rts4$E1 <- "Quantity"
-rts4$E2 <- "Selection"
-rts5 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Order","Selection")); names(rts5)[2] <- "F1"; names(rts5)[3] <- "F2"
-rts5$E1 <- "Order"
-rts5$E2 <- "Selection"
-rts6 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Quantity","Order")); names(rts6)[2] <- "F1"; names(rts6)[3] <- "F2"
-rts6$E1 <- "Quantity"
-rts6$E2 <- "Order"
-
-rts <- rbind(rts1,rts2,rts3,rts4,rts5,rts6)
-rts$mins=rts$RT_log-rts$ci
-rts$maxs=rts$RT_log+rts$ci
-
-rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
-rts$F3 <- factor(paste(rts$F1,rts$F2))
-
-rts$E1 <- factor(rts$E1, levels=c("Selection","Order","Quantity","Vagueness"))
-rts$E2 <- factor(rts$E2, levels=c("Selection","Order","Quantity","Vagueness"))
-
-# The palette with black:
-cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-
-ggplot(rts, aes(y=RT_log, x=item_mean_ratio, group=F3, col=F2, pch=F1)) + 
-  facet_grid(E2~E1, drop=F) + 
-  geom_point(cex=3) + geom_line() + theme_bw() + 
-  theme(aspect.ratio=1, panel.grid=element_blank()) +
-  scale_color_manual(name="colour",values=cbbPalette) +
-  scale_shape_discrete(name="shape") +
-  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))
-
-```
-
-# RT model all items
-
-```{r, rtmodel-allitems}
-# backing off A LOT from maximal model to avoid warning about maxfun
-rtm <- lmer(
-  data=dd,
-  RT_log ~ 
-    c_Vag * Selection.sum + item_mean_ratio + c_Qty + c_Ord + s_Trl +
-    RTprev_log  + 
-  ( c_Vag * Selection.sum | Subject)
-)
-rtm.summary <- summary(rtm)
-rtm.confint <- confint(rtm)
-```
-
-Print a table of coefficients.
-
-```{r, rtmodeltable, results='asis', echo=FALSE}
-print(xtable(
-  digits=c(0,2,2,1,1,3), 
-  rtm.summary$coefficients), type='html')
-```
-
-Plot coefficients
-
-```{r}
-par(las=2, mar=c(14,6,1,1))
-y=coef(rtm.summary)[-1,1] # omit intercept
-n=length(y)
-plot(y, xaxt="n", xlab="", ylab="RT_log\n", pch=19, ylim=extendrange(y, f=.10), main="Coefficients")
-abline(h=0)
-axis(1, labels=names(y) , at=1:n)
-grid()
-segments(x0=1:n, x1=1:n, y0, y1, lwd=2)
-```
-
-Plot partial effects
-
-```{r}
-par(mfrow=c(2,4))
-plotLMER.fnc(rtm)
-```
-
-```{r}
-suppressPackageStartupMessages(require(languageR))
-cat("k is:",
-collin.fnc(dd[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])$cnumber)
-```
-
-```{r, fig.width=6, fig.height=5}
-suppressPackageStartupMessages(require(rms))
-par(mar=c(1.1,3.1,1.1,1.1), pty='s')
-plot(varclus(as.matrix(dd[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])))
-```
-
-Model criticism baayen plots
-
-```{r, fig.width=9, fig.height=3}
-# Baayen 4-plot model criticism
-
-par(mfrow=c(1,4), pty='s')
-# create scaled residuals
-dd$rstand = as.vector(scale(resid(rtm)))
-# plot scaled residuals density
-plot(density(dd$rstand))
-# plot sample quantiles versus theoretical quantiles
-qqnorm(dd$rstand, cex=.5)
-qqline(dd$rstand)
-# plot standardised residuals versus fitted values
-plot(dd$rstand ~ fitted(rtm), pch='.')
-# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
-abline(h=c(-2.5,2.5))
-
-```
-
-Refit model with outliers removed (3.5 sd)
-
-```{r}
-dd2 <- subset(dd, abs(rstand)<3.5 )
-```
-
-```{r}
-# backing off A LOT from maximal model to avoid warning about maxfun
-rtm <- lmer(
-  data=dd2,
-  RT_log ~ 
-    c_Vag * Selection.sum + item_mean_ratio + c_Qty + c_Ord + s_Trl +
-    RTprev_log  + 
-  ( c_Vag * Selection.sum | Subject)
-)
-rtm.summary <- summary(rtm)
-rtm.confint <- confint(rtm)
-```
-
-Print a table of coefficients.
-
-```{r,  results='asis', echo=FALSE}
-print(xtable(
-  digits=c(0,2,2,1,1,3), 
-  rtm.summary$coefficients), type='html')
-```
-
-Plot coefficients
-
-```{r}
-par(las=2, mar=c(14,6,1,1))
-y=coef(rtm.summary)[-1,1] # omit intercept
-n=length(y)
-plot(y, xaxt="n", xlab="", ylab="RT_log\n", pch=19, ylim=extendrange(y, f=.10), main="Coefficients")
-abline(h=0)
-axis(1, labels=names(y) , at=1:n)
-grid()
-segments(x0=1:n, x1=1:n, y0, y1, lwd=2)
-```
-
-Plot partial effects
-
-```{r}
-par(mfrow=c(2,4))
-plotLMER.fnc(rtm)
-```
-
-```{r}
-suppressPackageStartupMessages(require(languageR))
-cat("k is:",
-collin.fnc(dd2[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])$cnumber
-)
-
-```
-
-```{r, fig.width=6, fig.height=5}
-suppressPackageStartupMessages(require(rms))
-par(mar=c(1.1,3.1,1.1,1.1), pty='s')
-plot(varclus(as.matrix(dd2[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])))
-```
-
-Model criticism baayen plots
-
-```{r, fig.width=9, fig.height=3}
-# Baayen 4-plot model criticism
-
-par(mfrow=c(1,4), pty='s')
-# create scaled residuals
-dd2$rstand = as.vector(scale(resid(rtm)))
-# plot scaled residuals density
-plot(density(dd2$rstand))
-# plot sample quantiles versus theoretical quantiles
-qqnorm(dd2$rstand, cex=.5)
-qqline(dd2$rstand)
-# plot standardised residuals versus fitted values
-plot(dd2$rstand ~ fitted(rtm), pch='.')
-# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
-abline(h=c(-2.5,2.5))
-
-```
-
-Separate analyses for vague vs crisp at each level of selection (matching, and comparison)
-
-not done yet.
-
-```{r,  results='asis', echo=FALSE}
-print(xtable(
-  digits=c(0,2,2,1,1,3), 
-  rtm.summary$coefficients), type='latex')
-```
-
-# Borderline analysis

### 2016-07-29 experiments/e2/archive/table_raw_mean_RT_tables.Rmd
archiving old e2 Rmd files

## Raw mean RT tables

```{r, results='asis', eval=TRUE, cache=FALSE}
var1 = ddply(dat, .(Vagueness), summarise, RT=mean(RT, na.rm=TRUE) )
print(xtable(var1, digits=0),type='html', include.rownames=F)
```

```{r, results='asis', eval=TRUE, cache=FALSE}
var2=ddply(dat, .(Number), summarise, RT=mean(RT,na.rm=TRUE))
print(xtable(var2, digits=0),type='html', include.rownames=F)
```

```{r, results='asis', eval=TRUE, cache=FALSE}
var3=ddply(dat, .(Vagueness,Number), summarise, RT=mean(RT, na.rm=TRUE))
print(xtable(var3, digits=0),type='html', include.rownames=F)
```

```{r, results='asis', eval=TRUE, cache=FALSE}
var4=ddply(dat, .(f_Itm), summarise, RT=mean(RT, na.rm=TRUE))
print(xtable(var3, digits=0),type='html', include.rownames=F)
```

### 2016-07-29 experiments/e2/writeup/e2.Rmd
tidying up

---
title: "Experiment 2 (of 4)"
author: "Matt Green"
output: 
  html_document:
    toc: true
    <!--toc_float: true-->
    mathjax: default
    template: default
    code_folding: hide
    theme: default
---

```{r print_info_on_design}
# print info on design, number of cells
cat(paste('number of subjects = ',30,'\n',
          '---','\n',
          'number of trials per subject = ',256,'\n', 
          '---','\n',          
          'number of unique cells in the design = Item(4) * Number(2) * Vagueness(2) * Order(2) * Quantity(2) = ',4*2*2*2*2,'\n',
          'number of times a subject saw a unique cell in the design = 256 / 64 = ', 256/64,'\n',
          '---','\n',
          'number of higher-level cells in the design (abstracting over Order and Quantity) = Item(4) * Number(2) * Vagueness(2) = ',4*2*2,'\n',
          'number of times a subject saw a higher level cell in the design =  = 256 / 16 = ', 4*2*2, '\n',
          sep=''))
```

# Distributional stuff, including transformations.

```{r transformations_table, results='asis'}
require(xtable, quietly=TRUE)
x=445
lambda = c(-2,-1,-0.5, 0,0.5,1,2)

xd=data.frame(
  lambda = c(
    lambda[1],
    lambda[2],
    lambda[3],
    lambda[4],
    lambda[5],
    lambda[6],
    lambda[7]
  ),
  x=c(    
    paste("x = ", sep="", x),
    paste("x = ", sep="", x),
    paste("x = ", sep="", x),
    paste("x = ", sep="", x),
    paste("x = ", sep="", x),
    paste("x = ", sep="", x),
    paste("x = ", sep="", x)
     ),
  conventional=c(
    '$$y=\\frac{1}{x^2}$$', 
    '$$y=\\frac{1}{x}$$', 
    '$$y=\\frac{1}{\\sqrt{x}}$$', 
    '$$y=log_e(x)$$', 
    '$$y=\\sqrt{x}$$', 
    '$$y=x$$', 
    '$$y=x^2$$'
    ),
#   y1=c(    
#     paste("y = ", sep="", 1/(x^2)),
#     paste("y = ", sep="", 1/x),
#     paste("y = ", sep="", 1/sqrt(x)),
#     paste("y = ", sep="", round(log(x),2)),
#     paste("y = ", sep="", sqrt(x)),
#     paste("y = ", sep="", x),
#     paste("y = ", sep="", x^2)
#   ),
  power=c(
    '$$y=x^{-2}$$', 
    '$$y=x^{-1}$$', 
    '$$y=x^{-0.5}$$', 
    '$$y=log_e(x)$$', 
    '$$y=x^{0.5}$$', 
    '$$y=x^{1}$$', 
    '$$y=x^{2}$$'
    ),
  y=c(     
    paste("y = ", sep="", format(round(x^lambda[1],6), sc=F) ),
    paste("y = ", sep="", round(x^lambda[2],3) ),
    paste("y = ", sep="", round(x^lambda[3],3) ),
    paste("y = ", sep="", round(log(x),3) ),
    paste("y = ", sep="", round(x^lambda[5],2) ),
    paste("y = ", sep="", round(x^lambda[6],6) ),
    paste("y = ", sep="", round(x^lambda[7],6) )
  ),
  back = c(
    '$$x=y^{\\frac{1}{-2}}$$',
    '$$x=y^{\\frac{1}{-1}}$$', 
    '$$x=y^{\\frac{1}{-0.5}}$$', 
    '$$x=exp(y)$$',  
    '$$x=y^{\\frac{1}{0.5}}$$', 
    '$$x=y^{\\frac{1}{1}}$$',  
    '$$x=y^{\\frac{1}{2}}$$'
  ),
  back_lambda=c(
    '$$x=y^{1/\\lambda}$$',
    '$$x=y^{1/\\lambda}$$',
    '$$x=y^{1/\\lambda}$$',
    '$$x=exp(y)$$',  
    '$$x=y^{1/\\lambda}$$',
    '$$x=y^{1/\\lambda}$$',
    '$$x=y^{1/\\lambda}$$'
  ),
  y_back=c(  
    paste( 'x = ', (x^lambda[1]) ^ {1/lambda[1]} ,sep=""),
    paste( 'x = ', (x^lambda[2]) ^ {1/lambda[2]} ,sep=""),
    paste( 'x = ', (x^lambda[3]) ^ {1/lambda[3]} ,sep=""),
    paste( 'x = ', exp(log(x)), sep=""),
    paste( 'x = ', (x^lambda[5]) ^ {1/lambda[5]} ,sep=""),
    paste( 'x = ', (x^lambda[6]) ^ {1/lambda[6]} ,sep=""),
    paste( 'x = ', (x^lambda[7]) ^ {1/lambda[7]} ,sep="")
  )
)

print(xtable(xd, digits=1, align=c(rep('l', times=ncol(xd)+1))), type='html', include.rownames=F, digits=1)
```

```{r libs, message=FALSE, results='hide'}
if(require(MASS)){}
if(require(ggplot2)){}
if(require(lmerTest)){}
if(require(xtable)){}
if(require(plyr)){}
if(require(reshape2)){}
if(require(lattice)){}
if(require(LMERConvenienceFunctions)){} # for perSubjectTrim
if(require(car)){}
if(require(rms)){} # for varclus
if(require(gridExtra)){}
if(require(grid)){}
if(require(latticeExtra)){}
if(require(beepr)){}
suppressPackageStartupMessages(require(lme4))
suppressPackageStartupMessages(require(lmerTest))
suppressPackageStartupMessages(require(languageR))
suppressPackageStartupMessages(library("reshape2"))
suppressPackageStartupMessages(library("ggplot2"))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(ggplot2))
if(require(ggplot2)){}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(gridExtra))
suppressPackageStartupMessages(library(plyr))
if(require(reshape2)){}
suppressPackageStartupMessages(require(rms))
suppressPackageStartupMessages(require(languageR))
```

Get the data if file dat.Rda does not already exist.

```{r, get_the_data}
# if the output file dat.Rda already exists then load it, else do data wrangling
if( file.exists('dat.Rda') ) load('dat.Rda') else {

  # source required data wrangling functions
  source('functions/classifyResponse.R')
  source('functions/gatherData.R')

  # declare local variables
  number_of_valid_subjects <- 30
  number_of_rows <- 7680 # 7680
  number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # 256

    ## call user-defined functions

  # collect all individual subject files into one data frame
  dat <- gatherData(number_of_valid_subjects)

  # classify the response as expected, near, or far
  dat <- classifyResponse(dat)

   ## treat variables

  # SUBJECT
  # ensure Subject is a factor 
  dat$Subject=factor(paste("s",sprintf("%02d",dat$Subject),sep=""))

  # TRIAL
  # Trial for subject, 1 to 256
  dat$Trial = rep(x = 1:number_of_trials_per_subject, times = number_of_valid_subjects)

  # make a centred Trial for modeling
  dat$c_Trl <-dat$Trial - mean(dat$Trial)

  # make a scaled Trial for modelling
  dat$s_Trl <- as.numeric(scale(dat$Trial))

  # ID
  # id is a unique identifier for the 7680 row data
  dat$id <- factor(paste(paste(dat$Subject), 
                         paste("t", sprintf("%03d", dat$Trial), sep="") , sep=":"))

  # ITEM
  # create a centred numeric item variable for modeling
  dat$c_Itm <- ifelse(dat$Item==1, -.75, 
                      ifelse(dat$Item==2, -.25,
                             ifelse(dat$Item==3, .25, .75)))

  # make Item be a factor and assign labels
  dat$Item <- factor(dat$Item, levels=c(1,2,3,4), labels=c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))

  # add ratios for Item
  item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54) 
          # 0.2500000 0.4705882 0.5909091 0.6666667
  item_range_ratio_scaled = as.vector(scale(c(6 / 24, 16 / 34, 26 / 44, 36 / 54)))
          # -1.3441995 -0.1316642  0.5297187  0.9461450
  item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))) 
          # 0.5125000 0.6876471 0.7691558 0.8166667
  item_mean_ratio_scaled = as.vector(scale(c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))))) 
          # -1.37582241 -0.06614191  0.54334858 0.89861574

  dat[dat$Item == "06:15:24", "item_range_ratio"] <-  0.2500000
  dat[dat$Item == "16:25:34", "item_range_ratio"] <-  0.4705882
  dat[dat$Item == "26:35:44", "item_range_ratio"] <-  0.5909091
  dat[dat$Item == "36:45:54", "item_range_ratio"] <-  0.6666667

  dat[dat$Item == "06:15:24", "item_range_ratio_scaled"] <-  -1.3441995
  dat[dat$Item == "16:25:34", "item_range_ratio_scaled"] <-  -0.1316642
  dat[dat$Item == "26:35:44", "item_range_ratio_scaled"] <-   0.5297187  
  dat[dat$Item == "36:45:54", "item_range_ratio_scaled"] <-   0.9461450

  dat[dat$Item == "06:15:24", "item_mean_ratio"] <-  0.5125000 
  dat[dat$Item == "16:25:34", "item_mean_ratio"] <-  0.6876471 
  dat[dat$Item == "26:35:44", "item_mean_ratio"] <-  0.7691558
  dat[dat$Item == "36:45:54", "item_mean_ratio"] <-  0.8166667

  dat[dat$Item == "06:15:24", "item_mean_ratio_scaled"] <-   -1.37582241 
  dat[dat$Item == "16:25:34", "item_mean_ratio_scaled"] <-   -0.06614191 
  dat[dat$Item == "26:35:44", "item_mean_ratio_scaled"] <-    0.54334858 
  dat[dat$Item == "36:45:54", "item_mean_ratio_scaled"] <-    0.89861574

  # VAGUENESS
  # Create a factor coding for Vagueness
  dat[ dat$Condition==1 , 'Vagueness'] <- 'Vague'
  dat[ dat$Condition==2 , 'Vagueness'] <- 'Crisp'
  dat[ dat$Condition==3 , 'Vagueness'] <- 'Vague'
  dat[ dat$Condition==4 , 'Vagueness'] <- 'Crisp'
  dat$Vagueness <- as.factor(dat$Vagueness)

  # manually center Vagueness
  dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)

  # NUMBER
  # Create a factor coding for Number use
  dat[ dat$Condition==1 , 'Number'] <- 'Numeric'
  dat[ dat$Condition==2 , 'Number'] <- 'Numeric'
  dat[ dat$Condition==3 , 'Number'] <- 'Verbal'
  dat[ dat$Condition==4 , 'Number'] <- 'Verbal'
  dat$Number <- as.factor(dat$Number)

  # manually center Number
  dat$c_Num <- ifelse(dat$Number=='Numeric', -.5, .5)

  # CONDITION
  # make a factor out of Condition, as f_Cnd
  dat$f_Cnd <- factor(dat$Condition, levels=c(1,2,3,4), labels=c('Vg:Nm', 'Cr:Nm', 'Vg:Vb', 'Cr:Vb'))

  # ORDER
  # give the levels of Order meaningful names
  dat$Order <- factor(dat$Order, levels=c(1,2), labels=c('LtoR','RtoL'))

  # make a manually centred Order
  dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)

  # QUANTITY
  # give the levels of Quantity meaningful names
  dat$Quantity <- factor(dat$Quantity, levels=c(1,2), labels=c('Small','Large'))

  # make a manually centred Quantity
  dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)

  # INSTRUCTION
  # add number of characters in the instruction # 29 30 34 36 38
  dat$nchar_instr = nchar(dat$Instruction)
  dat$nchar_instr_scaled = as.vector(scale(nchar(dat$Instruction), scale=TRUE))

  # make Instruction be a factor (17 levels)
  dat$Instruction <- as.factor(dat$Instruction)

  # RT
  # add transformations of RT
  dat$RT_RecSqd    <- 1/dat$RT^2
  dat$RT_Rec       <- 1/dat$RT
  dat$RT_RecSqt    <- 1/sqrt(dat$RT) 
  dat$RT_log       <- log(dat$RT)
  dat$RT_sqt       <- sqrt(dat$RT)
  dat$RT_raw       <- dat$RT
  dat$RT_sqd       <- dat$RT^2

  # print to file a table with information about the design
  design_info <- unique(subset(dat, select=c(Item, Condition, Vagueness, Number, Quantity, Order, 
                                             Left, Mid, Right, Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side,Instruction)))
  design.info <- design_info[order(design_info$Item, design_info$Condition, design_info$Quantity, design_info$Order),]
  row.names(design_info) <- NULL
  capture.output(  print.data.frame(design_info, row.names=F, print.gap=3, quote=F, right=F), file="design_info.table")

  # put dat in better column order
  dat <- subset(dat, 
                select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number, 
                           Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,
                           c_Trl, s_Trl, c_Itm, c_Vag, c_Num, f_Cnd, c_Ord, c_Qty,
                           RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
                           Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
                           Instruction, nchar_instr
                ))

  # save dat as dat.Rda
  # This data set contains *all* trials 7680 including impossible trials and is mainly for graphs comparing different removals
  save(dat, file="dat.Rda")

  ####################################################
  # remove impossible trials from dat in new data dd #
  ####################################################

  # Throw out RT = 1 and RT = 59998, and RTprev = 1 and RTprev = 59998 
  # i.e., throw out sticky fingers and timeouts, 
  # and the trials that followed sticky fingers and timeouts 
  # since they were likely affected by unusual previous trials.

  # lose impossible trials
  dd <- dat
  dd$RT[dd$RT == 1 ] <- NA
  dd$RT[dd$RT == 59998 ] <- NA
  dd <- dd[complete.cases(dd),]
  row.names(dd) <- NULL

  # add preceding RT: because we removed impossible trials, the value for preceding RT for a trial following an impossible trial is the value of the trial that preceded the impossible trial.
  dd$RTprev <- NA
  for (s in levels(dd$Subject)) {
    nrows = nrow(dd[dd$Subject==s,])
    for (i in 1:nrows) {
      if (i==1) {
        dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]
      }
      else
        dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1]
    }
  }

  # add transformations of previous RT
  dd$RTprev_RecSqd    <- 1/dd$RTprev^2
  dd$RTprev_Rec       <- 1/dd$RTprev
  dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev) 
  dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
  dd$RTprev_log       <- log(dd$RTprev)
  dd$RTprev_sqt       <- sqrt(dd$RTprev)
  dd$RTprev_raw       <- dd$RTprev
  dd$RTprev_sqd       <- dd$RTprev^2

  # number of higher-level cells in the design (abstracting over Order) = Item(4) * Number(2) * Vagueness(2) * Quantity(2) = 32
  # number of measurements in eaach cell = 8
  # 332 * 8 = 256
  # 256 * 30 = 7680

  dd$measurement=0
  dd$cell=0
  for (s in levels(dd$Subject)) {
    cellcount=0
    for (i in levels(dd$Item)) {
      for (n in levels(dd$Number)) {
        for (v in levels(dd$Vagueness)) {
          for (q in levels(dd$Quantity)) {
            measurementcount=0
            cellcount=cellcount+1
            for (t in dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q, "Trial"]) {
              dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "cell"] <- cellcount
              measurementcount=measurementcount+1
              dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "measurement"] = measurementcount
            }
          }
        }
      }
    }    
  }

  # put dd in better column order
  dd <- subset(dd, 
               select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number, 
                          Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,
                          c_Trl, s_Trl, c_Itm, c_Vag, c_Num, f_Cnd, c_Ord, c_Qty,
                          RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
                          RTprev, RTprev_RecSqd, RTprev_Rec, RTprev_RecSqt, RTprev_log, RTprev_sqt, RTprev_raw, RTprev_sqd,
                          Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
                          Instruction, nchar_instr,
                          cell, measurement
               ))

  # save dd
  save(dd, file="dd.Rda")
}
```

Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions.

```{r plotDists, fig.width=20, fig.height=12, message=FALSE}
## plot the comparison of the transformations against their normal equivalents to see distributional properties using full data

load("dat.Rda")

# first create a data frame specially for transformations
# this won't be used for analysis, just for this plot
dat_transforms <- dat

# dfSubset says whether any data points were removed
dat_transforms$dfSubset <- "original"

# RTtype says whether the row contains data for an observed RT or 
# a sample from a normal distribution with the same mean and sd
dat_transforms$RTtype <- "observed"

# create extra rows for dfs with removed (outlier) data points
dat1 <- dat_transforms                                       # nrow 7680
dat2 <- subset(dat_transforms, RT>1)                         # nrow 7678
dat2$dfSubset <- "RT>1"
dat3 <- subset(dat_transforms, RT>1 & RT<59999)              # nrow 7677
dat3$dfSubset <- "RT>1&RT<59999"
dat4 <- subset(dat_transforms, RT>1 & RT<59999 & RT<30001)   # nrow 7658
dat4$dfSubset <- "RT>1&RT<30001"

# bring the dfs back into main df
dat_transforms <- rbind(dat1, dat2, dat3, dat4)
dat_transforms$dfSubset <- factor(dat_transforms$dfSubset, levels=c("original", "RT>1", "RT>1&RT<59999", "RT>1&RT<30001"))

# create extra rows for normally distributed transformed variables
dato = dat_transforms
datn = dat_transforms; datn$RTtype <- "normal"
dat_transforms <- rbind(dato,datn)
dat_transforms$RTtype <- factor(dat_transforms$RTtype, levels=c("observed","normal"))

# for safety
dat_transforms[dat_transforms$RTtype=="normal", c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd")] <- NA

# Create the normally distributed transformed variables, replacing the NAs declared above
for (k in levels(dat_transforms$dfSubset)){
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_RecSqd"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_RecSqd"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_RecSqd"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_RecSqd"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_Rec"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_Rec"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_Rec"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_Rec"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_RecSqt"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_RecSqt"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_RecSqt"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_RecSqt"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_log"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_log"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_log"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_log"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_sqt"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_sqt"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_sqt"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_sqt"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_raw"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_raw"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_raw"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_raw"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_sqd"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_sqd"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_sqd"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_sqd"]))
}

temp <- melt(dat_transforms, 
             id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Number", "Quantity", "Order"),
             measure.vars=c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd"),
             variable.name="transformation",
             value.name="score")

ggplot(temp, aes(score, colour=RTtype)) + 
  geom_freqpoly() + 
  facet_wrap(dfSubset~transformation, scales="free", nrow=4) +
  theme_bw() + theme(aspect.ratio=1) + scale_color_manual(values=c("blue","red"))
```

# Now working with a data set without impossible trials.

Show how transformations of RT affect distribution of times, and how they affect which times are outliers.

```{r boxPlots, fig.width=20, fig.height=5}
# boxplot subjects and items

load("dd.Rda")

# allow reference to transformation
subdata = melt(dd,
               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
               variable.name="transformation",
               value.name="value")

# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ transformation, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
  xlab("") + ylab("")
```

Show how mean times for individual subjects and items vary with respect to the grand mean RT.

```{r fastSlowRT, fig.width=10, fig.height=5}
# Show how mean times for individual subjects and items vary with respect to the grand mean RT.

suppressMessages(library(plyr))
suppressMessages(library(lattice))
suppressMessages(library(latticeExtra))

subs=ddply(dd, .(level=Subject), summarise, effect="subject", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
itms=ddply(dd, .(level=Item), summarise, effect="item", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
si=rbind(subs,itms)
si$effect <- factor(si$effect, levels=c("subject", "item"))
si$lsign <- factor(si$lsign, levels=c("slow","fast"))

my1 <- xyplot(data=si, lmean~level | effect, 
              aspect=1,
              layout=c(2,1),
              ylim = extendrange(si$lmean, f = 0.10),
              par.settings=list(trellis.par.set(superpose.symbol = list(col = c("blue","red"), pch=19), superpose.line=list(col = c("blue","red")))),
              groups=si$lsign,
              xlab=NULL,
              ylab="RT (in ms)",
              key=list(text=list(levels(si$lsign)), lines=list(lty=c(1,1), col=trellis.par.get('superpose.line')$col, pch=c(19,19), type='o'), divide=1),
              mean=unique(si$emean),
              scales=list(x=list(relation="free", at=list(15.5, 32.5), labels=list("subject number", "item identifier"), cex=1 ), 
                          y=list(at=c(0, 2000, unique(si$emean), 4000, 6000, 8000,  10000), labels=c(0, 2000,  "Mean", 4000,  6000, 8000, 10000), cex=.9) ),
              panel=function(x,y,type,subscripts,groups=groups,...){
                panel.xyplot(x,y,type='p',groups=groups, subscripts=subscripts,...)
                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, groups=groups, subscripts=subscripts, code=3, length=0, col=trellis.par.get('superpose.line')$col[groups][subscripts], ...)
                panel.abline(h=unique(si$emean), col="lightgrey",lty=1)
                panel.text(x, y=(ifelse(y<unique(si$emean),y-300,y+300)), as.numeric(si$level)[subscripts][panel.number()==1], cex=.75)
                panel.text(x, y=(ifelse(y<unique(si$emean),y-300,y+300)), as.character(si$level)[subscripts][panel.number()==2], cex=.75)
              })

my2 <- xyplot(data=si, ldiff~level | effect, 
              aspect=1, type='n',
              ylab="Deviation from mean RT (in ms)",
              ylim=extendrange(si$ldiff,f=0.10))

print(doubleYScale(my1, my2, add.axis=TRUE, add.ylab2=TRUE, use.style=FALSE))
```

Show how mean times for individual subjects and items vary with respect to the grand mean RT_RecSqt.

```{r fastSlowRecSqt, fig.width=10, fig.height=5}
# Show how mean times for individual subjects and items vary with respect to the grand mean RT_RecSqt.

suppressMessages(library(plyr))
suppressMessages(library(lattice))
suppressMessages(library(latticeExtra))

subs <- ddply(dd, .(level=Subject), summarise, 
              effect="subject", emean=mean(dd$RT_RecSqt), lmean=mean(RT_RecSqt), ldiff=mean(RT_RecSqt)-mean(dd$RT_RecSqt),
              lsign=ifelse(mean(RT_RecSqt)-mean(dd$RT_RecSqt)<0,"slow","fast"))
itms <- ddply(dd, .(level=Item), summarise, effect="item", emean=mean(dd$RT_RecSqt), lmean=mean(RT_RecSqt),
              ldiff=mean(RT_RecSqt)-mean(dd$RT_RecSqt), lsign=ifelse(mean(RT_RecSqt)-mean(dd$RT_RecSqt)<0,"slow","fast"))
si <- rbind(subs,itms)
si$effect <- factor(si$effect, levels=c("subject", "item"))
si$lsign <- factor(si$lsign, levels=c("slow","fast"))

my1 <- xyplot(data=si, lmean~level | effect, 
              aspect=1,
              layout=c(2,1),
              ylim = extendrange(si$lmean, f = 0.10),
              par.settings = list(trellis.par.set(superpose.symbol = list(col = c("blue","red"), pch=19), superpose.line=list(col = c("blue","red")))),
              groups = si$lsign,
              xlab = NULL,
              ylab = "RT_RecSqt: reciprocal of square root of RT in ms\n",
              key=list(text=list(levels(si$lsign)), lines=list(lty=c(1,1), col=trellis.par.get('superpose.line')$col, pch=c(19,19), type='o'), divide=1),
              mean = unique(si$emean),
              scales = list(x=list(relation="free", at=list(15.5, 32.5), labels=list("subject number", "item identifier"), cex=1 ) ),
              panel=function(x,y,type,subscripts,groups=groups,...){
                panel.xyplot(x,y,type='p',groups=groups, subscripts=subscripts,...)
                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, groups=groups, subscripts=subscripts, code=3, length=0, col=trellis.par.get('superpose.line')$col[groups][subscripts], ...)
                panel.abline(h=unique(si$emean), col="lightgrey",lty=1)
                panel.text(x, y=(ifelse(y<unique(si$emean),y-.001,y+.001)), as.numeric(si$level)[subscripts][panel.number()==1], cex=.75)
                panel.text(x, y=(ifelse(y<unique(si$emean),y-.001,y+.001)), as.character(si$level)[subscripts][panel.number()==2], cex=.75)
              })

my2 <- xyplot(data=si, ldiff~level | effect, 
              aspect=1, type='n',
              ylab="\nDeviation from mean RT_RecSqt (in ms)",
              ylim=extendrange(si$ldiff,f=0.10))

print(doubleYScale(my1, my2, add.axis=TRUE, add.ylab2=TRUE, use.style=FALSE))
```

Show how practice and fatigue effects vary over subjects

```{r, message=F, results='hide'}

load("dd.Rda")
```

```{r, eval=TRUE, fig.width=20, fig.height=10, cache=FALSE, echo=TRUE}
# Show how practice/fatigue effects vary over subjects
ggplot(dd, aes(y=RT_RecSqt, x=Trial)) + facet_wrap(~Subject,nrow=3) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial", breaks=NULL) + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1) 
```

Show practice and fatigue effects vary over items.

```{r, eval=TRUE, fig.width=9, fig.height=4, cache=FALSE, echo=TRUE}
# Show how practice/fatigue effects vary over items
ggplot(dd, aes(y=RT_RecSqt, x=Trial)) + facet_grid(~Item) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial") + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1) 
```

# Plot main effects and interactions

```{r interactionEffects, message=FALSE, results='hide', echo=FALSE}
load("dd.Rda")
source('functions/CIfunctions.R')

```

Main effects

Plot main effects on several transformations

```{r plorMainsa, fig.width=16, fig.height=6, warning=FALSE}
# plot main effects on observed data

# allow reference to transformation
temp <- melt(dd,
             id.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
             measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log",  "RT_raw"),
             variable.name="transformation",
             value.name="score")

# allow reference to effect
temp2 <- melt(temp, 
              id.vars=c("transformation", "score"),
              measure.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
              variable.name="effect",
              value.name="level")

# do the plot
ggplot(temp2, aes(y=score, x=level, group=1, col=effect)) +
  facet_wrap(~transformation + effect, scales="free", shrink=TRUE, ncol=5) +
  stat_summary(fun.data=mean_cl_normal, geom="pointrange") +
  stat_summary(fun.y=mean, geom="line") +
  theme(aspect.ratio=1, axis.text.y=element_blank())
```

Main effects

```{r mainB}
# prep main effects rts data frame
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness")));names(vrts)[3]<-"Levels"
nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number")));names(nrts)[3]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Quantity")));names(qrts)[3]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Order")));names(orts)[3]<-"Levels"
irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Item")));names(irts)[3]<-"Levels"
rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
```

```{r mainC, fig.width=9, fig.height=3}
# plot main effects
ggplot(rts, aes(y=RT_RecSqt, x=Levels, group=1, ymin=mins, ymax=maxs)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Effect, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) 
```

Main effects over items

```{r mainD, message=FALSE, fig.width=10, fig.height=3}
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Vagueness")));names(vrts)[4]<-"Levels"
nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Number")));names(nrts)[4]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Quantity")));names(qrts)[4]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Order")));names(orts)[4]<-"Levels"

rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci

rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))

ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, group=Levels, ymin=mins, ymax=maxs, shape=Levels, fill=Levels, col=Levels)) +
  scale_shape_manual(values=c(21,21,22,22,23,23,24,24)) +
  scale_fill_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
  scale_color_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Effect) + 
  theme_bw() +
  theme(aspect.ratio=1, 
        panel.grid=element_blank()) +
  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))

```

2 way interactions not involving items.

```{r 2way}
# prep 2 ways rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Number")); names(rts1)[1] <- "F1"; names(rts1)[2] <- "F2"
rts1$Effect <- rep("Vagueness:Number",4)
rts2 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Quantity")); names(rts2)[1] <- "F1"; names(rts2)[2] <- "F2"
rts2$Effect <- rep("Vagueness:Quantity",4)
rts3 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Order")); names(rts3)[1] <- "F1"; names(rts3)[2] <- "F2"
rts3$Effect <- rep("Vagueness:Order",4)
rts5 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Quantity")); names(rts5)[1] <- "F1"; names(rts5)[2] <- "F2"
rts5$Effect <- rep("Number:Quantity",4)
rts6 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Order")); names(rts6)[1] <- "F1"; names(rts6)[2] <- "F2"
rts6$Effect <- rep("Number:Order",4)
rts8 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Quantity","Order")); names(rts8)[1] <- "F1"; names(rts8)[2] <- "F2"
rts8$Effect <- rep("Quantity:Order",4)
rts <- rbind(rts1,rts2,rts3,rts5,rts6,rts8)
rts <- subset(rts, select=c(Effect, F1, F2, RT_RecSqt, RT_RecSqt_norm, sd, se, ci))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
```

```{r foo,fig.width=10, fig.height=3}
# plot 2 way interactions
ggplot(rts, aes(y=RT_RecSqt, x=F2, group=F1, ymin=mins, ymax=maxs, pch=F1)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_wrap(~Effect, scale="free_x", nrow=1) + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank()) +
  xlab("Levels")
```

Plot 2 way interactions over items

```{r 22way, message=FALSE}
# prep 2 ways over items rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Number")); names(rts1)[2] <- "F1"; names(rts1)[3] <- "F2"
rts1$E1 <- "Vagueness"
rts1$E2 <- "Number"
rts2 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Quantity")); names(rts2)[2] <- "F1"; names(rts2)[3] <- "F2"
rts2$E1 <- "Vagueness"
rts2$E2 <- "Quantity"
rts3 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Order")); names(rts3)[2] <- "F1"; names(rts3)[3] <- "F2"
rts3$E1 <- "Vagueness"
rts3$E2 <- "Order"
rts4 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Quantity", "Number")); names(rts4)[2] <- "F1"; names(rts4)[3] <- "F2"
rts4$E1 <- "Quantity"
rts4$E2 <- "Number"
rts5 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Order","Number")); names(rts5)[2] <- "F1"; names(rts5)[3] <- "F2"
rts5$E1 <- "Order"
rts5$E2 <- "Number"
rts6 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Quantity","Order")); names(rts6)[2] <- "F1"; names(rts6)[3] <- "F2"
rts6$E1 <- "Quantity"
rts6$E2 <- "Order"

rts <- rbind(rts1,rts2,rts3,rts4,rts5,rts6)
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci

rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
rts$F3 <- factor(paste(rts$F1,rts$F2))

rts$E1 <- factor(rts$E1, levels=c("Number","Order","Quantity","Vagueness"))
rts$E2 <- factor(rts$E2, levels=c("Number","Order","Quantity","Vagueness"))

# The palette with black:
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, group=F3, col=F2, pch=F1)) + 
  facet_grid(E2~E1, drop=F) + 
  geom_point(cex=3) + geom_line() + theme_bw() + 
  theme(aspect.ratio=1, panel.grid=element_blank()) +
  scale_color_manual(name="colour",values=cbbPalette) +
  scale_shape_discrete(name="shape") +
  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))

```

Plot vagueness by number interaction over items

```{r plotInt1, fig.width=7, fig.height=3.5, eval=TRUE, cache=FALSE}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
dodge = position_dodge(width=0.2)
ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
                group=condition, shape=condition, fill=condition)) +
  geom_line(position=dodge) + 
  geom_errorbar(width=.2, position=dodge) + 
  geom_point(size=4, position=dodge) + 
  scale_shape_manual("",values = c(22, 22, 21, 21)) + 
  scale_fill_manual("",values=c("black","white","black","white")) +
  ggtitle("Response time") + 
  ylab("RT_RecSqt") + 
  xlab("") +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(~Number)
```

3 way interactions

```{r plot3way}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Quantity"))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
p1=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Vagueness, ymin=mins, ymax=maxs, pch=Vagueness)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Quantity, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank(), legend.position="top")
```

```{r bar}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Order"))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
p2=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Vagueness, ymin=mins, ymax=maxs, pch=Vagueness)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Order, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank(), legend.position="top")
```

```{r daz}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Quantity","Order"))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
p3=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Quantity, ymin=mins, ymax=maxs, pch=Quantity)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Order, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank(), legend.position="top")
```

```{r foo2}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Quantity","Order"))
rts$mins=rts$RT_RecSqt-rts$ci
rts$maxs=rts$RT_RecSqt+rts$ci
p4=ggplot(rts, aes(y=RT_RecSqt, x=Vagueness, group=Quantity, ymin=mins, ymax=maxs, pch=Quantity)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Order, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank(), legend.position="top")
```

```{r foo3, fig.width=10,fig.height=5}
grid.arrange(nrow=2, p1,p2,p3,p4)
```

Plot vagueness by number by quantity over items.

```{r vnitems, fig.width=10, fig.height=4, eval=TRUE}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item", "Quantity"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
dodge = position_dodge(width=0.2)
ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
                group=Vagueness, 
                fill=Vagueness)) +
  geom_line(position=dodge) + 
  geom_errorbar(width=.2, position=dodge) + 
  geom_point(pch=21, size=4, position=dodge) + 
  scale_fill_manual("",values=c("black","white")) +
  ggtitle("Response time") + 
  ylab("RT_Rec_Sqt") + 
  xlab("") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        panel.background=element_rect(fill="white",color=1),
        legend.key = element_blank(), 
        aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_grid(~Number+Quantity)
```

4 way interaction

```{r 4way}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness",  "Quantity", "Order"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
ggplot(rts, aes(y=RT_RecSqt, x=Vagueness, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, group=Quantity, pch=Quantity)) +
  facet_grid(~Order+Number, scale="free_x") +
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank())
```

4 way interaction split over items

```{r 4waysplit, fig.width=8, fig.height=8, eval=TRUE, cache=FALSE, echo=TRUE}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness", "Item", "Quantity", "Order"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
                group=Vagueness,fill=Vagueness)) +
  geom_line() + 
  geom_errorbar(width=.12) + 
  geom_point(pch=21, size=2) + 
  scale_fill_manual("",values=c("black","white","black","white")) +
  ggtitle("Number * Vagueness * Quantity * Order") + 
  ylab("RT_RecSqt") + 
  xlab("") +
  theme_bw() +
  theme( legend.key = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), axis.text.x = element_text(angle = 45, hjust = 1), aspect.ratio=1) +
  facet_grid(Vagueness+Order~Number+Quantity)
```

4 way interaction split over items showing ratio of item dots.

```{r 4waySplitRatio, fig.width=8, fig.height=8, eval=TRUE, cache=FALSE, echo=TRUE, message=FALSE, warning=FALSE}
rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness", "item_mean_ratio", "Quantity", "Order"))
rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
#dodge = position_dodge(width=0.2)
ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
                group=Vagueness,fill=Vagueness)) +
  # geom_line(position=dodge) + 
	geom_line() +
  geom_errorbar(width=.02) + 
  geom_point(pch=21, size=2) + 
  scale_fill_manual("",values=c("black","white","black","white")) +
  ggtitle("Number * Vagueness * Quantity * Order") + 
  ylab("RT_RecSqt") + 
  xlab("") +
  theme_bw() +
  theme( legend.key = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), axis.text.x = element_text(angle = 45, hjust = 1), aspect.ratio=1) +
  facet_grid(Vagueness+Order~Number+Quantity)
```

# Show compression of item ratios

```{r showCOmpression, fig.width=10, fig.height=7, echo=F}
#  Show compression of ratios of numbers of dots in the items
item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54)
item_range_ratio_equally_spaced = seq(from = 6 / 24, to = 36 / 54, length.out = 4)
item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54)))
item_mean_ratio_equally_spaced = seq(from = mean(c(6 / 15, 15 / 24)), to = mean(c(36 / 45, 45 / 54)), length.out = 4)

plot(y=c(2,2,2,2), x=item_range_ratio, xlim=c(0,1), ylim=c(0,3), ylab="", yaxt='n', col="blue", pch=21, bg='blue', type='o', xaxt='n', xlab="ratio: small is easily discriminable", main="Discriminability of each item, expressed as ratios between array magnitudes")
text(y=c(2.15,2.15,2.15,2.15), x=item_range_ratio, labels=round(item_range_ratio,2), col="blue")
text(y=2.15, x=0.1, labels="ratio min/max", col='blue')
points(y=c(2.3,2.3,2.3,2.3), x=item_range_ratio_equally_spaced, col="lightblue", pch=21, bg='lightblue', type='o')

points(y=c(1,1,1,1), x=item_mean_ratio, pch=21, col='red', bg='red', type='o')
text(y=c(1.15,1.15,1.15,1.15), x=item_mean_ratio, labels=round(item_mean_ratio,2), col="red")
text(y=1.15, x=0.2, labels="mean (ratio min/med,\n            ratio med/max)", col='red')
points(y=c(1.3,1.3,1.3,1.3), x=item_mean_ratio_equally_spaced, col="pink", pch=21, bg='pink', type='o')
axis(1, at=seq(0,1,0.1))

```

## ols model v4, RecSqt, in progress

```{r ols}
load("dd.Rda")

options(width=140)
```

Pairs plot

```{r pairs, fig.width=7, fig.height=7}
# pairs plot
pairs(dd[,c("RT_RecSqt", "Number", "Vagueness", "RTprev_RecSqt")], pch = 19, cex=.075)
```

Build the model

```{r doModelols}
## do ols model
# define datadist
dd.dd = datadist(dd)
options(datadist = "dd.dd")
# build model
v4 <- ols(data=dd, 
          RT_RecSqt ~ 
            c_Vag + c_Num + c_Qty + c_Ord + 
            c_Num:c_Vag:c_Qty+
            item_mean_ratio + 
            #c_Vag:c_Num + 
            #c_Vag:c_Num:item_mean_ratio + 
            # pol(s_Trl, 3) +
            s_Trl+
            # pol(RTprev_RecSqt,3) +
            RTprev_RecSqt +
            nchar_instr
            # cell+
            # measurement,
            ,
          x=T, y=T )
```

```{r likRat}
# print out p value for likelihood ratio with df
cat("lik.ratio p value, whether the model as a whole is explanatory: p =", 1-pchisq(v4$stats[2],v4$stats[3]))
# print out R2 for the model
cat("R^2 = ", v4$stats[4])
```

Show the model tables

```{r modTabs}
# show model
v4
# show model summary
summary(v4)
# do anova 
an1 = anova(v4)
# show anova
an1
```

plot 'partial effects'

```{r partial, fig.width=12, fig.height=5}
## plot partial effects
# Compute Predicted Values and Confidence Limits
p1=Predict(v4)
# plot 'partial effects'
plot(p1, anova=an1, pval=T, aspect=1, main="ols model v4")
```

```{r daz99}
# add predicted RT_RecSqt to dd
dd$RT_Predicted <- predict (v4)
```

Plot model coefficients and ci's

```{r coefsAndCiS}
par(las=2, mar=c(14,6,1,1))
y=coef(v4) # with intercept
n=length(y)
y0=confint(v4)[1:n,1]
y1=confint(v4)[1:n,2]

y=coef(v4)[-1] # omit intercept
n=length(y)
y0=confint(v4)[-1,1]
y1=confint(v4)[-1,2]

plot(y, xaxt="n", xlab="", ylab="RT_RecSqt\n", pch=19, ylim=extendrange(y, f=.10), main="Speed")
abline(h=0)
axis(1, labels=names(y) , at=1:n)
grid()
segments(x0=1:n, x1=1:n, y0, y1, lwd=2)
```

Collinearity is assessed by the condition number $k$.\
The greater the collinearity, the closer the matrix of predictors is is to becoming _singular_.\
$k$ estimates the extent to which a matrix is $\text{singular}$\
$0\ldots6$ no collinearity\
$\text{around} 15$ medium collinearity\
$>30$ indicates potentially harmful collinearity\

```{r kis}
cat("k is:",
collin.fnc(dd[,c("s_Trl", "c_Num", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_RecSqt", "nchar_instr", "measurement")])$cnumber
)
```

```{r plotg, fig.width=6, fig.height=5}
par(mar=c(1.1,3.1,1.1,1.1), pty='s')
plot(varclus(as.matrix(dd[,c("s_Trl", "c_Num", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_RecSqt", "nchar_instr", "measurement")])))
```

Model criticism baayen plots

```{r baayneMplots, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

par(mfrow=c(1,4), pty='s')
# create scaled residuals
v4$rstand = as.vector(scale(resid(v4)))
# plot scaled residuals density
plot(density(v4$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(v4$rstand, cex=.5)
qqline(v4$rstand)
# plot standardised residuals versus fitted values
plot(v4$rstand ~ fitted(v4), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))
# create dffits
dffits=abs(resid(v4, "dffits"))
# plot dffits
plot(dffits, type='h')
```

```{r which}
# here no overly influential
w = which.influence(fit = v4, cutoff = 0.4)
w
```

Model validation

Validation tests overfitting using bootstrap.

Bootstrapping draws as a bootstrap sample the same number of samples as the original data, at random with replacement, from the original data.

Then fit the model to the data in the bootstrap sample, and use this model to predict the reaction times for the original full data (which contains many samples that are new to the bootstrap model).

Then compare the goodness of fit of the bootstrap model with the goodness of fit of the original model.

Averaged over a large number of bootstrap models these comparisons of goodness of fit  reveal to what extent the original model overfits the data.

Function validate() in rms does this _re-sampling validation_.

```{r val_1, echo=TRUE}
# argument B is the number of bootstrap runs
# argument pr is whether to print the results of each run
v_1 <- validate(v4, bw = T, B = 200, pr=FALSE, estimates=TRUE)
```

```{r prval}
# in print.validate, B is the number of re-samples to show outocmes for each resample
print(v_1, B=0)
```

```{r kept}
# show how often each number of variables kept was observed across the 200 runs
xtabs(~rowSums(attr(v_1,"kept")))
```

# model lmer

```{r rmme}
load("dd.Rda")
```

```{r lmer5}
v5 <- lmer(data=dd, 
           RT_RecSqt ~ 
             c_Vag + c_Num + c_Qty + c_Ord + 
             c_Num:c_Vag:c_Qty +
             item_mean_ratio + 
             s_Trl +
             RTprev_RecSqt +
             nchar_instr +
             (1+c_Vag + c_Num + c_Qty + c_Ord|Subject)
)
```

```{r v51}
summary(v5)
```

```{r v52}
cat("R^2")
cor(fitted(v5), dd$RT_RecSqt)^2
```

```{r v53}
par(mfrow=c(2,5))
plotLMER.fnc(v5)
```

Plot model coefficients and ci's

```{r v54}
e=data.frame(coef=summary(v5)$coef[-1,1]) # estimates, without intercept
q=as.data.frame(confint(v5, method='Wald')[18:26,1:2])
eq=cbind(e,q)
y=eq[rev(rownames(eq)),]

par(mar=c(4,10,1,1))
plot(y=1:nrow(y), x=y$coef, xlim=range(y$"2.5 %", y$"97.5 %"), type='n', axes=F, xlab="", ylab="")
segments(y0=1:nrow(y), y1=1:nrow(y), x0=y$"2.5 %", x1=y$"97.5 %", lwd=.55)
points(y=1:nrow(y), x=y$coef, pch=20, cex=.75)
abline(v=0, lty=3, col='grey')
axis(2, labels=row.names(y), at=1:nrow(y), las=1)
axis(1, las=1)
mtext("estimate", side=1, line=2.5, cex.lab=1, las=1)
box(col='grey')
```

Model criticism baayen plots

```{r baayenPLots99, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

par(mfrow=c(1,4), pty='s')
# create scaled residuals
dd$rstand = as.vector(scale(resid(v5)))
# plot scaled residuals density
plot(density(dd$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(dd$rstand, cex=.5)
qqline(dd$rstand)
# plot standardised residuals versus fitted values
plot(dd$rstand ~ fitted(v5), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))
```

### 2016-07-29 experiments/e2/writeup/blups_temp.R

@@ -1,23 +1,0 @@
-par(mfrow=c(3,2))
-plot(ranef(m2)$Subject,type="n",main="m2 subj BLUPS")
-text(ranef(m2)$Subject, labels=rownames(ranef(m2)$Subject), cex=.7)
-abline(h=0,lty=3);abline(v=0,lty=3)
-plot(coef(m2)$Subject,type="n",main="m2 subj coefficients")
-text(coef(m2)$Subject, labels=rownames(coef(m2)$Subject), cex=.7)
-abline(h=0,lty=3)
-
-
-plot(ranef(m3)$Subject,type="n",main="m3 subj BLUPS")
-text(ranef(m3)$Subject, labels=rownames(ranef(m3)$Subject), cex=.7)
-abline(h=0,lty=3);abline(v=0,lty=3)
-plot(coef(m3)$Subject,type="n",main="m3 subj coefficients")
-text(coef(m3)$Subject, labels=rownames(coef(m3)$Subject), cex=.7)
-abline(h=0,lty=3)
-
-
-plot(ranef(m3)$f_Itm,type="n",main="m3 item BLUPS")
-text(ranef(m3)$f_Itm, labels=rownames(ranef(m3)$f_Itm), cex=.7)
-abline(h=0,lty=3);abline(v=0,lty=3)
-plot(coef(m3)$f_Itm,type="n",main="m3 item coefficients")
-text(coef(m3)$f_Itm, labels=rownames(coef(m3)$f_Itm), cex=.7)
-abline(h=0,lty=3)

### 2016-07-29 experiments/e2/writeup/e2.Rmd
simplifying the directory by adding function code into the main .Rmd file

@@
 
+```{r loadKnitr}
+suppressPackageStartupMessages(library(knitr))
+```
+
+```{r loadLibs}
+suppressPackageStartupMessages(library(MASS))
+suppressPackageStartupMessages(library(xtable))
+suppressPackageStartupMessages(library(lattice))
+suppressPackageStartupMessages(library(latticeExtra))
+suppressPackageStartupMessages(library(car))
+suppressPackageStartupMessages(library(beepr))
+suppressPackageStartupMessages(library(lme4))
+suppressPackageStartupMessages(library(lmerTest))
+suppressPackageStartupMessages(library(LMERConvenienceFunctions)) # for perSubjectTrim
+suppressPackageStartupMessages(library(languageR))
+suppressPackageStartupMessages(library(reshape2))
+suppressPackageStartupMessages(library(ggplot2))
+suppressPackageStartupMessages(library(grid))
+suppressPackageStartupMessages(library(gridExtra))
+suppressPackageStartupMessages(library(plyr))
+suppressPackageStartupMessages(library(rms))                      # for varclus
+suppressPackageStartupMessages(library(languageR))
+```
+
+```{r CIfunctions}
+## Summarizes data.
+## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
+##   data: a data frame.
+##   measurevar: the name of a column that contains the variable to be summariezed
+##   groupvars: a vector containing names of columns that contain grouping variables
+##   na.rm: a boolean that indicates whether to ignore NA's
+##   conf.interval: the percent range of the confidence interval (default is 95%)
+summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
+                      conf.interval=.95, .drop=TRUE) {
+  require(plyr)
+  # New version of length which can handle NA's: if na.rm==T, don't count them
+  length2 <- function (x, na.rm=FALSE) {
+    if (na.rm) sum(!is.na(x))
+    else       length(x)
+  }
+  # This is does the summary; it's not easy to understand...
+  datac <- ddply(data, groupvars, .drop=.drop,
+                 .fun= function(xx, col, na.rm) {
+                   c( N    = length2(xx[,col], na.rm=na.rm),
+                      mean = mean   (xx[,col], na.rm=na.rm),
+                      sd   = sd     (xx[,col], na.rm=na.rm)
+                   )
+                 },
+                 measurevar,
+                 na.rm
+  )
+  # Rename the "mean" column    
+  datac <- rename(datac, c("mean"=measurevar))
+  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
+  # Confidence interval multiplier for standard error
+  # Calculate t-statistic for confidence interval: 
+  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
+  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
+  datac$ci <- datac$se * ciMult
+  return(datac)
+}
+## Summarizes data, handling within-subjects variables by removing inter-subject variability.
+## It will still work if there are no within-S variables.
+## Gives count, un-normed mean, normed mean (with same between-group mean),
+##   standard deviation, standard error of the mean, and confidence interval.
+## If there are within-subject variables, calculate adjusted values using method from Morey (2008).
+##   data: a data frame.
+##   measurevar: the name of a column that contains the variable to be summariezed
+##   betweenvars: a vector containing names of columns that are between-subjects variables
+##   withinvars: a vector containing names of columns that are within-subjects variables
+##   idvar: the name of a column that identifies each subject (or matched subjects)
+##   na.rm: a boolean that indicates whether to ignore NA's
+##   conf.interval: the percent range of the confidence interval (default is 95%)
+summarySEwithin <- function(data=NULL, measurevar, betweenvars=NULL, withinvars=NULL,
+                            idvar=NULL, na.rm=FALSE, conf.interval=.95, .drop=TRUE) {
+  # Ensure that the betweenvars and withinvars are factors
+  factorvars <- vapply(data[, c(betweenvars, withinvars), drop=FALSE],
+                       FUN=is.factor, FUN.VALUE=logical(1))
+  if (!all(factorvars)) {
+    nonfactorvars <- names(factorvars)[!factorvars]
+    message("Automatically converting the following non-factors to factors: ",
+            paste(nonfactorvars, collapse = ", "))
+    data[nonfactorvars] <- lapply(data[nonfactorvars], factor)
+  }
+  # Get the means from the un-normed data
+  datac <- summarySE(data, measurevar, groupvars=c(betweenvars, withinvars),
+                     na.rm=na.rm, conf.interval=conf.interval, .drop=.drop)
+  # Drop all the unused columns (these will be calculated with normed data)
+  datac$sd <- NULL
+  datac$se <- NULL
+  datac$ci <- NULL
+  # Norm each subject's data
+  ndata <- normDataWithin(data, idvar, measurevar, betweenvars, na.rm, .drop=.drop)
+  # This is the name of the new column
+  measurevar_n <- paste(measurevar, "_norm", sep="")
+  # Collapse the normed data - now we can treat between and within vars the same
+  ndatac <- summarySE(ndata, measurevar_n, groupvars=c(betweenvars, withinvars),
+                      na.rm=na.rm, conf.interval=conf.interval, .drop=.drop)
+  # Apply correction from Morey (2008) to the standard error and confidence interval
+  #  Get the product of the number of conditions of within-S variables
+  nWithinGroups    <- prod(vapply(ndatac[,withinvars, drop=FALSE], FUN=nlevels,
+                                  FUN.VALUE=numeric(1)))
+  correctionFactor <- sqrt( nWithinGroups / (nWithinGroups-1) )
+  # Apply the correction factor
+  ndatac$sd <- ndatac$sd * correctionFactor
+  ndatac$se <- ndatac$se * correctionFactor
+  ndatac$ci <- ndatac$ci * correctionFactor
+  # Combine the un-normed means with the normed results
+  merge(datac, ndatac)
+}
+## Norms the data within specified groups in a data frame; it normalizes each
+## subject (identified by idvar) so that they have the same mean, within each group
+## specified by betweenvars.
+##   data: a data frame.
+##   idvar: the name of a column that identifies each subject (or matched subjects)
+##   measurevar: the name of a column that contains the variable to be summariezed
+##   betweenvars: a vector containing names of columns that are between-subjects variables
+##   na.rm: a boolean that indicates whether to ignore NA's
+normDataWithin <- function(data=NULL, idvar, measurevar, betweenvars=NULL,
+                           na.rm=FALSE, .drop=TRUE) {
+  require(plyr)
+  # Measure var on left, idvar + between vars on right of formula.
+  data.subjMean <- ddply(data, c(idvar, betweenvars), .drop=.drop,
+                         .fun = function(xx, col, na.rm) {
+                           c(subjMean = mean(xx[,col], na.rm=na.rm))
+                         },
+                         measurevar,
+                         na.rm
+  )
+  # Put the subject means with original data
+  data <- merge(data, data.subjMean)
+  # Get the normalized data in a new column
+  measureNormedVar <- paste(measurevar, "_norm", sep="")
+  data[,measureNormedVar] <- data[,measurevar] - data[,"subjMean"] +
+    mean(data[,measurevar], na.rm=na.rm)
+  # Remove this subject mean column
+  data$subjMean <- NULL
+  return(data)
+}
+```
+
 ```{r print_info_on_design}
-# print info on design, number of cells
 cat(paste('number of subjects = ',30,'\n',
@@
 ```{r transformations_table, results='asis'}
-require(xtable, quietly=TRUE)
 x=445
@@
     ),
-#   y1=c(    
-#     paste("y = ", sep="", 1/(x^2)),
-#     paste("y = ", sep="", 1/x),
-#     paste("y = ", sep="", 1/sqrt(x)),
-#     paste("y = ", sep="", round(log(x),2)),
-#     paste("y = ", sep="", sqrt(x)),
-#     paste("y = ", sep="", x),
-#     paste("y = ", sep="", x^2)
-#   ),
   power=c(
@@
 
-```{r libs, message=FALSE, results='hide'}
-if(require(MASS)){}
-if(require(ggplot2)){}
-if(require(lmerTest)){}
-if(require(xtable)){}
-if(require(plyr)){}
-if(require(reshape2)){}
-if(require(lattice)){}
-if(require(LMERConvenienceFunctions)){} # for perSubjectTrim
-if(require(car)){}
-if(require(rms)){} # for varclus
-if(require(gridExtra)){}
-if(require(grid)){}
-if(require(latticeExtra)){}
-if(require(beepr)){}
-suppressPackageStartupMessages(require(lme4))
-suppressPackageStartupMessages(require(lmerTest))
-suppressPackageStartupMessages(require(languageR))
-suppressPackageStartupMessages(library("reshape2"))
-suppressPackageStartupMessages(library("ggplot2"))
-suppressPackageStartupMessages(library(reshape2))
-suppressPackageStartupMessages(library(ggplot2))
-if(require(ggplot2)){}
-suppressPackageStartupMessages(library(ggplot2))
-suppressPackageStartupMessages(library(gridExtra))
-suppressPackageStartupMessages(library(plyr))
-if(require(reshape2)){}
-suppressPackageStartupMessages(require(rms))
-suppressPackageStartupMessages(require(languageR))
-```
-
-Get the data if file dat.Rda does not already exist.
+Load the file dat.Rda if it exists, else Get the data if file dat.Rda does not already exist.
 
@@
 # if the output file dat.Rda already exists then load it, else do data wrangling
-if( file.exists('dat.Rda') ) load('dat.Rda') else {
-
-  # source required data wrangling functions
-  source('functions/classifyResponse.R')
-  source('functions/gatherData.R')
+if( file.exists('dat.Rda') && file.exists('dd.Rda') ) {
+	load('dat.Rda'); load('dd.Rda')
+	} else {
+	gatherData = function(number_of_valid_subjects) {
+		data_dir <- '../experimentCode/output/'
+		column.headers.df <-  head( read.table(
+			paste(data_dir,'subject1.data',sep=''),
+			header=TRUE),0)
+		gathered.data <- column.headers.df
+		for (subject in 1:number_of_valid_subjects) {
+			current.filename <- paste(data_dir, 'subject', subject, '.data', sep='')
+			current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = FALSE)
+			gathered.data <- rbind(gathered.data, current.data)
+		}
+		return(gathered.data)
+	} # end of gatherData function
+	classifyResponse = function(dat) {
+		# what were they expected to respond?
+		dat$crossed = as.factor(paste('Con', dat$Condition, ':Quan', dat$Quantity, ':Item', dat$Item, sep=''))
+		dat[dat$crossed=='Con1:Quan1:Item1', 'Exp_Num'] <-  6
+		dat[dat$crossed=='Con1:Quan1:Item1', 'Bline_Num']     <- 15
+		dat[dat$crossed=='Con1:Quan1:Item1', 'Extr_Num']      <- 24
+		dat[dat$crossed=='Con1:Quan1:Item2', 'Exp_Num'] <- 16
+		dat[dat$crossed=='Con1:Quan1:Item2', 'Bline_Num']     <- 25
+		dat[dat$crossed=='Con1:Quan1:Item2', 'Extr_Num']      <- 34
+		dat[dat$crossed=='Con1:Quan1:Item3', 'Exp_Num'] <- 26
+		dat[dat$crossed=='Con1:Quan1:Item3', 'Bline_Num']     <- 35
+		dat[dat$crossed=='Con1:Quan1:Item3', 'Extr_Num']      <- 44
+		dat[dat$crossed=='Con1:Quan1:Item4', 'Exp_Num'] <- 36
+		dat[dat$crossed=='Con1:Quan1:Item4', 'Bline_Num']     <- 45
+		dat[dat$crossed=='Con1:Quan1:Item4', 'Extr_Num']      <- 54
+		dat[dat$crossed=='Con1:Quan2:Item1', 'Exp_Num'] <- 24
+		dat[dat$crossed=='Con1:Quan2:Item1', 'Bline_Num']     <- 15
+		dat[dat$crossed=='Con1:Quan2:Item1', 'Extr_Num']      <-  6
+		dat[dat$crossed=='Con1:Quan2:Item2', 'Exp_Num'] <- 34
+		dat[dat$crossed=='Con1:Quan2:Item2', 'Bline_Num']     <- 25
+		dat[dat$crossed=='Con1:Quan2:Item2', 'Extr_Num']      <- 16
+		dat[dat$crossed=='Con1:Quan2:Item3', 'Exp_Num'] <- 44
+		dat[dat$crossed=='Con1:Quan2:Item3', 'Bline_Num']     <- 35
+		dat[dat$crossed=='Con1:Quan2:Item3', 'Extr_Num']      <- 26
+		dat[dat$crossed=='Con1:Quan2:Item4', 'Exp_Num'] <- 54
+		dat[dat$crossed=='Con1:Quan2:Item4', 'Bline_Num']     <- 45
+		dat[dat$crossed=='Con1:Quan2:Item4', 'Extr_Num']      <- 36
+		dat[dat$crossed=='Con2:Quan1:Item1', 'Exp_Num'] <-  6
+		dat[dat$crossed=='Con2:Quan1:Item1', 'Bline_Num']     <- 15
+		dat[dat$crossed=='Con2:Quan1:Item1', 'Extr_Num']      <- 24
+		dat[dat$crossed=='Con2:Quan1:Item2', 'Exp_Num'] <- 16
+		dat[dat$crossed=='Con2:Quan1:Item2', 'Bline_Num']     <- 25
+		dat[dat$crossed=='Con2:Quan1:Item2', 'Extr_Num']      <- 34
+		dat[dat$crossed=='Con2:Quan1:Item3', 'Exp_Num'] <- 26
+		dat[dat$crossed=='Con2:Quan1:Item3', 'Bline_Num']     <- 35
+		dat[dat$crossed=='Con2:Quan1:Item3', 'Extr_Num']      <- 44
+		dat[dat$crossed=='Con2:Quan1:Item4', 'Exp_Num'] <- 36
+		dat[dat$crossed=='Con2:Quan1:Item4', 'Bline_Num']     <- 45
+		dat[dat$crossed=='Con2:Quan1:Item4', 'Extr_Num']      <- 54
+		dat[dat$crossed=='Con2:Quan2:Item1', 'Exp_Num'] <- 24
+		dat[dat$crossed=='Con2:Quan2:Item1', 'Bline_Num']     <- 15
+		dat[dat$crossed=='Con2:Quan2:Item1', 'Extr_Num']      <-  6
+		dat[dat$crossed=='Con2:Quan2:Item2', 'Exp_Num'] <- 34
+		dat[dat$crossed=='Con2:Quan2:Item2', 'Bline_Num']     <- 25
+		dat[dat$crossed=='Con2:Quan2:Item2', 'Extr_Num']      <- 16
+		dat[dat$crossed=='Con2:Quan2:Item3', 'Exp_Num'] <- 44
+		dat[dat$crossed=='Con2:Quan2:Item3', 'Bline_Num']     <- 35
+		dat[dat$crossed=='Con2:Quan2:Item3', 'Extr_Num']      <- 26
+		dat[dat$crossed=='Con2:Quan2:Item4', 'Exp_Num'] <- 54
+		dat[dat$crossed=='Con2:Quan2:Item4', 'Bline_Num']     <- 45
+		dat[dat$crossed=='Con2:Quan2:Item4', 'Extr_Num']      <- 36
+		dat[dat$crossed=='Con3:Quan1:Item1', 'Exp_Num'] <-  6
+		dat[dat$crossed=='Con3:Quan1:Item1', 'Bline_Num']     <- 15
+		dat[dat$crossed=='Con3:Quan1:Item1', 'Extr_Num']      <- 24
+		dat[dat$crossed=='Con3:Quan1:Item2', 'Exp_Num'] <- 16
+		dat[dat$crossed=='Con3:Quan1:Item2', 'Bline_Num']     <- 25
+		dat[dat$crossed=='Con3:Quan1:Item2', 'Extr_Num']      <- 34
+		dat[dat$crossed=='Con3:Quan1:Item3', 'Exp_Num'] <- 26
+		dat[dat$crossed=='Con3:Quan1:Item3', 'Bline_Num']     <- 35
+		dat[dat$crossed=='Con3:Quan1:Item3', 'Extr_Num']      <- 44
+		dat[dat$crossed=='Con3:Quan1:Item4', 'Exp_Num'] <- 36
+		dat[dat$crossed=='Con3:Quan1:Item4', 'Bline_Num']     <- 45
+		dat[dat$crossed=='Con3:Quan1:Item4', 'Extr_Num']      <- 54
+		dat[dat$crossed=='Con3:Quan2:Item1', 'Exp_Num'] <- 24
+		dat[dat$crossed=='Con3:Quan2:Item1', 'Bline_Num']     <- 15
+		dat[dat$crossed=='Con3:Quan2:Item1', 'Extr_Num']      <-  6
+		dat[dat$crossed=='Con3:Quan2:Item2', 'Exp_Num'] <- 34
+		dat[dat$crossed=='Con3:Quan2:Item2', 'Bline_Num']     <- 25
+		dat[dat$crossed=='Con3:Quan2:Item2', 'Extr_Num']      <- 16
+		dat[dat$crossed=='Con3:Quan2:Item3', 'Exp_Num'] <- 44
+		dat[dat$crossed=='Con3:Quan2:Item3', 'Bline_Num']     <- 35
+		dat[dat$crossed=='Con3:Quan2:Item3', 'Extr_Num']      <- 26
+		dat[dat$crossed=='Con3:Quan2:Item4', 'Exp_Num'] <- 54
+		dat[dat$crossed=='Con3:Quan2:Item4', 'Bline_Num']     <- 45
+		dat[dat$crossed=='Con3:Quan2:Item4', 'Extr_Num']      <- 36
+		dat[dat$crossed=='Con4:Quan1:Item1', 'Exp_Num'] <-  6
+		dat[dat$crossed=='Con4:Quan1:Item1', 'Bline_Num']     <- 15
+		dat[dat$crossed=='Con4:Quan1:Item1', 'Extr_Num']      <- 24
+		dat[dat$crossed=='Con4:Quan1:Item2', 'Exp_Num'] <- 16
+		dat[dat$crossed=='Con4:Quan1:Item2', 'Bline_Num']     <- 25
+		dat[dat$crossed=='Con4:Quan1:Item2', 'Extr_Num']      <- 34
+		dat[dat$crossed=='Con4:Quan1:Item3', 'Exp_Num'] <- 26
+		dat[dat$crossed=='Con4:Quan1:Item3', 'Bline_Num']     <- 35
+		dat[dat$crossed=='Con4:Quan1:Item3', 'Extr_Num']      <- 44
+		dat[dat$crossed=='Con4:Quan1:Item4', 'Exp_Num'] <- 36
+		dat[dat$crossed=='Con4:Quan1:Item4', 'Bline_Num']     <- 45
+		dat[dat$crossed=='Con4:Quan1:Item4', 'Extr_Num']      <- 54
+		dat[dat$crossed=='Con4:Quan2:Item1', 'Exp_Num'] <- 24
+		dat[dat$crossed=='Con4:Quan2:Item1', 'Bline_Num']     <- 15
+		dat[dat$crossed=='Con4:Quan2:Item1', 'Extr_Num']      <-  6
+		dat[dat$crossed=='Con4:Quan2:Item2', 'Exp_Num'] <- 34
+		dat[dat$crossed=='Con4:Quan2:Item2', 'Bline_Num']     <- 25
+		dat[dat$crossed=='Con4:Quan2:Item2', 'Extr_Num']      <- 16
+		dat[dat$crossed=='Con4:Quan2:Item3', 'Exp_Num'] <- 44
+		dat[dat$crossed=='Con4:Quan2:Item3', 'Bline_Num']     <- 35
+		dat[dat$crossed=='Con4:Quan2:Item3', 'Extr_Num']      <- 26
+		dat[dat$crossed=='Con4:Quan2:Item4', 'Exp_Num'] <- 54
+		dat[dat$crossed=='Con4:Quan2:Item4', 'Bline_Num']     <- 45
+		dat[dat$crossed=='Con4:Quan2:Item4', 'Extr_Num']      <- 36
+		dat$crossed <- NULL
+		# what side LEFT, MIDDLE, RIGHT corresponds with Expected, Borderline, Extreme?
+		for (row in 1:nrow(dat)) {
+			if (dat[row, 'Exp_Num']   == dat[row, 'Left'])  {dat[row, 'Exp_side']   <- 'left'}
+			if (dat[row, 'Exp_Num']   == dat[row, 'Mid'])   {dat[row, 'Exp_side']   <- 'mid'}
+			if (dat[row, 'Exp_Num']   == dat[row, 'Right']) {dat[row, 'Exp_side']   <- 'right'}
+			if (dat[row, 'Bline_Num'] == dat[row, 'Left'])  {dat[row, 'Bline_side'] <- 'left'}
+			if (dat[row, 'Bline_Num'] == dat[row, 'Mid'])   {dat[row, 'Bline_side'] <- 'mid'}
+			if (dat[row, 'Bline_Num'] == dat[row, 'Right']) {dat[row, 'Bline_side'] <- 'right'}
+			if (dat[row, 'Extr_Num']  == dat[row, 'Left'])  {dat[row, 'Extr_side']  <- 'left'}
+			if (dat[row, 'Extr_Num']  == dat[row, 'Mid'])   {dat[row, 'Extr_side']  <- 'mid'}
+			if (dat[row, 'Extr_Num']  == dat[row, 'Right']) {dat[row, 'Extr_side']  <- 'right'}
+		}
+		# what button press did the subject actually make? LEFT, MIDDLE, RIGHT, NOANSWER?
+		dat$RESPONSE <- as.factor(dat$RESPONSE)
+		# what number of dots corresponds with the subject's button press?
+		for (row in 1 : nrow(dat) ) {
+			switch(as.character(dat[row,'RESPONSE']),
+						 'LEFT' = {dat[row, 'response_num'] <- dat[row, 'Left']},
+						 'MIDDLE' = {dat[row, 'response_num'] <- dat[row, 'Mid']},
+						 'RIGHT' = {dat[row, 'response_num'] <- dat[row, 'Right']},
+						 'NOANSWER' = {dat[row, 'response_num'] <- NA}
+			)
+		}
+		# what side was the subject's button-press? Left, mid right?
+		dat$response_side <- tolower(dat$RESPONSE)
+		dat$response_side[dat$response_side == "middle"] <- 'mid'
+		dat$response_side <- factor(dat$response_side, exclude="noanswer")
+		# what category was the subject's response? Expected, Borderline, Extreme
+		dat$response_category <- "nocat"
+		for (row in row.names(na.omit(dat))) {
+			if (dat[row, 'response_num'] == dat[row, 'Exp_Num']) {dat[row, 'response_category'] <- 'expected'}
+			if (dat[row, 'response_num'] == dat[row, 'Bline_Num']) {dat[row, 'response_category'] <- 'borderline'}
+			if (dat[row, 'response_num'] == dat[row, 'Extr_Num']) {dat[row, 'response_category'] <- 'extreme'}
+		}
+		dat$response_category <- factor(dat$response_category, exclude="nocat")
+		dat$RESPONSE <- NULL
+		return(dat)
+	} # end of classifyResponse function
 
 	# declare local variables
-  number_of_valid_subjects <- 30
+	number_of_valid_subjects <- 30 # = 30
 	number_of_rows <- 7680 # 7680
-  number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # 256
-
-    ## call user-defined functions
-
-  # collect all individual subject files into one data frame
-  dat <- gatherData(number_of_valid_subjects)
-
-  # classify the response as expected, near, or far
-  dat <- classifyResponse(dat)
-
-   ## treat variables
+	number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # = 256
+	dat <- gatherData(number_of_valid_subjects) # = 30
+	dat <- classifyResponse(dat) # classify the response as expected, near, or far
 
 	# SUBJECT
-  # ensure Subject is a factor 
 	dat$Subject=factor(paste("s",sprintf("%02d",dat$Subject),sep=""))
-
 	# TRIAL
-  # Trial for subject, 1 to 256
 	dat$Trial = rep(x = 1:number_of_trials_per_subject, times = number_of_valid_subjects)
-
 	# make a centred Trial for modeling
 	dat$c_Trl <-dat$Trial - mean(dat$Trial)
-
 	# make a scaled Trial for modelling
 	dat$s_Trl <- as.numeric(scale(dat$Trial))
-
 	# ID
@@
 												 paste("t", sprintf("%03d", dat$Trial), sep="") , sep=":"))
-
 	# ITEM
@@
 														 ifelse(dat$Item==3, .25, .75)))
-
 	# make Item be a factor and assign labels
 	dat$Item <- factor(dat$Item, levels=c(1,2,3,4), labels=c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
-
 	# add ratios for Item
@@
 	# -1.37582241 -0.06614191  0.54334858 0.89861574
-
 	dat[dat$Item == "06:15:24", "item_range_ratio"] <-  0.2500000
@@
 	dat[dat$Item == "36:45:54", "item_range_ratio"] <-  0.6666667
-
 	dat[dat$Item == "06:15:24", "item_range_ratio_scaled"] <-  -1.3441995
@@
 	dat[dat$Item == "36:45:54", "item_range_ratio_scaled"] <-   0.9461450
-
 	dat[dat$Item == "06:15:24", "item_mean_ratio"] <-  0.5125000
@@
 	dat[dat$Item == "36:45:54", "item_mean_ratio"] <-  0.8166667
-
 	dat[dat$Item == "06:15:24", "item_mean_ratio_scaled"] <-   -1.37582241
@@
 	dat[dat$Item == "36:45:54", "item_mean_ratio_scaled"] <-    0.89861574
-
 	# VAGUENESS
@@
 	dat$Vagueness <- as.factor(dat$Vagueness)
-
 	# manually center Vagueness
 	dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)
-
 	# NUMBER
@@
 	dat$Number <- as.factor(dat$Number)
-
 	# manually center Number
 	dat$c_Num <- ifelse(dat$Number=='Numeric', -.5, .5)
-
 	# CONDITION
@@
 	dat$f_Cnd <- factor(dat$Condition, levels=c(1,2,3,4), labels=c('Vg:Nm', 'Cr:Nm', 'Vg:Vb', 'Cr:Vb'))
-
 	# ORDER
@@
 	dat$Order <- factor(dat$Order, levels=c(1,2), labels=c('LtoR','RtoL'))
-
 	# make a manually centred Order
 	dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)
-
   # QUANTITY
@@
 	dat$Quantity <- factor(dat$Quantity, levels=c(1,2), labels=c('Small','Large'))
-
 	# make a manually centred Quantity
 	dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)
-
 	# INSTRUCTION
@@
 	dat$nchar_instr_scaled = as.vector(scale(nchar(dat$Instruction), scale=TRUE))
-
 	# make Instruction be a factor (17 levels)
 	dat$Instruction <- as.factor(dat$Instruction)
-
 	# RT
@@
 	dat$RT_sqd       <- dat$RT^2
-
   # print to file a table with information about the design
@@
 	row.names(design_info) <- NULL
-  capture.output(  print.data.frame(design_info, row.names=F, print.gap=3, quote=F, right=F), file="design_info.table")
-
+	capture.output(print.data.frame(design_info, row.names=F, print.gap=3, quote=F, right=F),
+								 file="design_info.table")
 	# put dat in better column order
@@
 								))
-
 	# save dat as dat.Rda
-  # This data set contains *all* trials 7680 including impossible trials and is mainly for graphs comparing different removals
+	# This data set contains *all* trials 7680 including impossible trials and is
+	# mainly for graphs comparing different removals
 	save(dat, file="dat.Rda")
-
-  ####################################################
 	# remove impossible trials from dat in new data dd #
-  ####################################################
-
 	# Throw out RT = 1 and RT = 59998, and RTprev = 1 and RTprev = 59998
@@
 	# since they were likely affected by unusual previous trials.
-
 	# lose impossible trials
@@
 	row.names(dd) <- NULL
-
 	# add preceding RT: because we removed impossible trials, the value for preceding RT for a trial following an impossible trial is the value of the trial that preceded the impossible trial.
@@
 	}
-
 	# add transformations of previous RT
@@
 	dd$RTprev_sqd       <- dd$RTprev^2
-
-  # number of higher-level cells in the design (abstracting over Order) = Item(4) * Number(2) * Vagueness(2) * Quantity(2) = 32
+	# number of higher-level cells in the design (abstracting over Order) =
+	# Item(4) * Number(2) * Vagueness(2) * Quantity(2) = 32
 	# number of measurements in eaach cell = 8
@@
 	# 256 * 30 = 7680
-
 	dd$measurement=0
@@
 	}
-
 	# put dd in better column order
@@
 							 ))
-
-  # save dd
 	save(dd, file="dd.Rda")
-}
+} # end of the else clause, which is run if both of the files dat.Rda and dd.Rda do not exist.
 ```
@@
 ## plot the comparison of the transformations against their normal equivalents to see distributional properties using full data
-
-load("dat.Rda")
-
 # first create a data frame specially for transformations
@@
 dat_transforms <- dat
-
 # dfSubset says whether any data points were removed
 dat_transforms$dfSubset <- "original"
-
 # RTtype says whether the row contains data for an observed RT or 
@@
 dat_transforms$RTtype <- "observed"
-
 # create extra rows for dfs with removed (outlier) data points
@@
 dat4$dfSubset <- "RT>1&RT<30001"
-
 # bring the dfs back into main df
@@
 dat_transforms$dfSubset <- factor(dat_transforms$dfSubset, levels=c("original", "RT>1", "RT>1&RT<59999", "RT>1&RT<30001"))
-
 # create extra rows for normally distributed transformed variables
@@
 dat_transforms$RTtype <- factor(dat_transforms$RTtype, levels=c("observed","normal"))
-
 # for safety
 dat_transforms[dat_transforms$RTtype=="normal", c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd")] <- NA
-
 # Create the normally distributed transformed variables, replacing the NAs declared above
@@
           sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_sqd"]))
-}
-
-temp <- melt(dat_transforms, 
+} # end of create normally distributed equivalents for the data transforms.
+temp <- reshape2::melt(dat_transforms, 
              id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Number", "Quantity", "Order"),
@@
              value.name="score")
-
 ggplot(temp, aes(score, colour=RTtype)) + 
@@
   facet_wrap(dfSubset~transformation, scales="free", nrow=4) +
-  theme_bw() + theme(aspect.ratio=1) + scale_color_manual(values=c("blue","red"))
+  theme_bw() + 
+	theme(aspect.ratio=1) + 
+	scale_color_manual(values=c("blue","red"))
 ```
@@
 ```{r boxPlots, fig.width=20, fig.height=5}
-# boxplot subjects and items
-
-load("dd.Rda")
-
 # allow reference to transformation
-subdata = melt(dd,
+subdata = reshape2::melt(dd,
                measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
@@
 # Show how mean times for individual subjects and items vary with respect to the grand mean RT.
-
-suppressMessages(library(plyr))
-suppressMessages(library(lattice))
-suppressMessages(library(latticeExtra))
-
-subs=ddply(dd, .(level=Subject), summarise, effect="subject", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
-itms=ddply(dd, .(level=Item), summarise, effect="item", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
+subs=ddply(dd, .(level=Subject), 
+					 summarise, effect="subject", emean=mean(dd$RT), lmean=mean(RT),
+					 ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0, "slow", "fast"))
+itms=ddply(dd, .(level=Item), 
+					 summarise, effect="item", emean=mean(dd$RT), lmean=mean(RT), 
+					 ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0, "slow", "fast"))
 si=rbind(subs,itms)
@@
 si$lsign <- factor(si$lsign, levels=c("slow","fast"))
-
 my1 <- xyplot(data=si, lmean~level | effect, 
@@
               ylim = extendrange(si$lmean, f = 0.10),
-              par.settings=list(trellis.par.set(superpose.symbol = list(col = c("blue","red"), pch=19), superpose.line=list(col = c("blue","red")))),
+              par.settings=list(trellis.par.set(superpose.symbol = list(col = c("blue","red"), pch=19), 
+              																	superpose.line=list(col = c("blue","red")))),
               groups=si$lsign,
@@
               ylab="RT (in ms)",
-              key=list(text=list(levels(si$lsign)), lines=list(lty=c(1,1), col=trellis.par.get('superpose.line')$col, pch=c(19,19), type='o'), divide=1),
+              key=list(text=list(levels(si$lsign)), 
+              				 lines=list(lty=c(1,1), col=trellis.par.get('superpose.line')$col, 
+              				 					 pch=c(19,19), type='o'), divide=1),
               mean=unique(si$emean),
-              scales=list(x=list(relation="free", at=list(15.5, 32.5), labels=list("subject number", "item identifier"), cex=1 ), 
-                          y=list(at=c(0, 2000, unique(si$emean), 4000, 6000, 8000,  10000), labels=c(0, 2000,  "Mean", 4000,  6000, 8000, 10000), cex=.9) ),
+              scales=list(x=list(relation="free", at=list(15.5, 32.5), 
+              									 labels=list("subject number", "item identifier"), cex=1 ),
+                          y=list(at=c(0, 2000, unique(si$emean), 4000, 6000, 8000,  10000), 
+                          			 labels=c(0, 2000,  "Mean", 4000,  6000, 8000, 10000), cex=.9) ),
               panel=function(x,y,type,subscripts,groups=groups,...){
                 panel.xyplot(x,y,type='p',groups=groups, subscripts=subscripts,...)
-                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, groups=groups, subscripts=subscripts, code=3, length=0, col=trellis.par.get('superpose.line')$col[groups][subscripts], ...)
+                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, groups=groups, subscripts=subscripts, 
+                						 code=3, length=0, col=trellis.par.get('superpose.line')$col[groups][subscripts], ...)
                 panel.abline(h=unique(si$emean), col="lightgrey",lty=1)
-                panel.text(x, y=(ifelse(y<unique(si$emean),y-300,y+300)), as.numeric(si$level)[subscripts][panel.number()==1], cex=.75)
-                panel.text(x, y=(ifelse(y<unique(si$emean),y-300,y+300)), as.character(si$level)[subscripts][panel.number()==2], cex=.75)
-              })
-
+                panel.text(x, y=(ifelse(y<unique(si$emean),y-300,y+300)), 
+                					 as.numeric(si$level)[subscripts][panel.number()==1], cex=.75)
+                panel.text(x, y=(ifelse(y<unique(si$emean),y-300,y+300)), 
+                					 as.character(si$level)[subscripts][panel.number()==2], cex=.75)
+              }) # end of my1
 my2 <- xyplot(data=si, ldiff~level | effect, 
@@
               ylab="Deviation from mean RT (in ms)",
-              ylim=extendrange(si$ldiff,f=0.10))
-
-print(doubleYScale(my1, my2, add.axis=TRUE, add.ylab2=TRUE, use.style=FALSE))
+              ylim=extendrange(si$ldiff,f=0.10)
+							) # creates a second y-axis scale
+print(doubleYScale(my1, my2, 
+									 add.axis=TRUE, 
+									 add.ylab2=TRUE, 
+									 use.style=FALSE))
 ```
@@
 # Show how mean times for individual subjects and items vary with respect to the grand mean RT_RecSqt.
-
-suppressMessages(library(plyr))
-suppressMessages(library(lattice))
-suppressMessages(library(latticeExtra))
-
 subs <- ddply(dd, .(level=Subject), summarise, 
-              effect="subject", emean=mean(dd$RT_RecSqt), lmean=mean(RT_RecSqt), ldiff=mean(RT_RecSqt)-mean(dd$RT_RecSqt),
+              effect="subject", emean=mean(dd$RT_RecSqt), lmean=mean(RT_RecSqt), 
+							ldiff=mean(RT_RecSqt)-mean(dd$RT_RecSqt),
               lsign=ifelse(mean(RT_RecSqt)-mean(dd$RT_RecSqt)<0,"slow","fast"))
 itms <- ddply(dd, .(level=Item), summarise, effect="item", emean=mean(dd$RT_RecSqt), lmean=mean(RT_RecSqt),
-              ldiff=mean(RT_RecSqt)-mean(dd$RT_RecSqt), lsign=ifelse(mean(RT_RecSqt)-mean(dd$RT_RecSqt)<0,"slow","fast"))
+              ldiff=mean(RT_RecSqt)-mean(dd$RT_RecSqt), 
+							lsign=ifelse(mean(RT_RecSqt)-mean(dd$RT_RecSqt)<0,"slow","fast"))
 si <- rbind(subs,itms)
@@
               ylab = "RT_RecSqt: reciprocal of square root of RT in ms\n",
-              key=list(text=list(levels(si$lsign)), lines=list(lty=c(1,1), col=trellis.par.get('superpose.line')$col, pch=c(19,19), type='o'), divide=1),
+              key=list(text=list(levels(si$lsign)), 
+              				 lines=list(lty=c(1,1), 
+              				 					 col=trellis.par.get('superpose.line')$col,
+              				 					 pch=c(19,19), type='o'),
+              				 divide=1),
               mean = unique(si$emean),
-              scales = list(x=list(relation="free", at=list(15.5, 32.5), labels=list("subject number", "item identifier"), cex=1 ) ),
+              scales = list(x=list(relation="free", at=list(15.5, 32.5), 
+              										 labels=list("subject number", "item identifier"), cex=1 ) ),
               panel=function(x,y,type,subscripts,groups=groups,...){
                 panel.xyplot(x,y,type='p',groups=groups, subscripts=subscripts,...)
-                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, groups=groups, subscripts=subscripts, code=3, length=0, col=trellis.par.get('superpose.line')$col[groups][subscripts], ...)
+                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, 
+                						 groups=groups, subscripts=subscripts, code=3, length=0, col=trellis.par.get('superpose.line')$col[groups][subscripts], ...)
                 panel.abline(h=unique(si$emean), col="lightgrey",lty=1)
-                panel.text(x, y=(ifelse(y<unique(si$emean),y-.001,y+.001)), as.numeric(si$level)[subscripts][panel.number()==1], cex=.75)
-                panel.text(x, y=(ifelse(y<unique(si$emean),y-.001,y+.001)), as.character(si$level)[subscripts][panel.number()==2], cex=.75)
+                panel.text(x, y=(ifelse(y<unique(si$emean),y-.001,y+.001)), 
+                					 as.numeric(si$level)[subscripts][panel.number()==1], cex=.75)
+                panel.text(x, y=(ifelse(y<unique(si$emean),y-.001,y+.001)), 
+                					 as.character(si$level)[subscripts][panel.number()==2], cex=.75)
               })
-
 my2 <- xyplot(data=si, ldiff~level | effect, 
@@
               ylim=extendrange(si$ldiff,f=0.10))
-
 print(doubleYScale(my1, my2, add.axis=TRUE, add.ylab2=TRUE, use.style=FALSE))
@@
 
-```{r, message=F, results='hide'}
-
-load("dd.Rda")
-```
-
-```{r, eval=TRUE, fig.width=20, fig.height=10, cache=FALSE, echo=TRUE}
+```{r pracFatigueSubjs, eval=TRUE, fig.width=20, fig.height=10, cache=FALSE, echo=TRUE}
 # Show how practice/fatigue effects vary over subjects
-ggplot(dd, aes(y=RT_RecSqt, x=Trial)) + facet_wrap(~Subject,nrow=3) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial", breaks=NULL) + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1)
+ggplot(dd, aes(y=RT_RecSqt, x=Trial)) + 
+	facet_wrap(~Subject,nrow=3) + 
+	theme(panel.grid=element_blank(), 
+				panel.margin = unit(0, "lines"), 
+				panel.background = element_rect(fill = 'white', colour='black' ), 
+				legend.position="top") + 
+	scale_x_continuous("Trial", breaks=NULL) + 
+	geom_point(pch=19, cex=.25, alpha=.5) + 
+	stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + 
+	theme(aspect.ratio = 1) 
 ```
@@
 
-```{r, eval=TRUE, fig.width=9, fig.height=4, cache=FALSE, echo=TRUE}
+```{r pracFatigueItms, eval=TRUE, fig.width=9, fig.height=4, cache=FALSE, echo=TRUE}
 # Show how practice/fatigue effects vary over items
-ggplot(dd, aes(y=RT_RecSqt, x=Trial)) + facet_grid(~Item) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial") + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1)
+ggplot(dd, aes(y=RT_RecSqt, x=Trial)) + 
+	facet_grid(~Item) + 
+	theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), 
+				panel.background = element_rect(fill = 'white', colour='black' ), 
+				legend.position="top") + 
+	scale_x_continuous("Trial") + 
+	geom_point(pch=19, cex=.25, alpha=.5) + 
+	stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + 
+	theme(aspect.ratio = 1) 
 ```
@@
 
-```{r interactionEffects, message=FALSE, results='hide', echo=FALSE}
-load("dd.Rda")
-source('functions/CIfunctions.R')
-
-```
-
 Main effects
@@
 
-```{r plorMainsa, fig.width=16, fig.height=6, warning=FALSE}
+```{r plotMainsa, fig.width=16, fig.height=6, warning=FALSE}
 # plot main effects on observed data
-
 # allow reference to transformation
-temp <- melt(dd,
+temp <- reshape2::melt(dd,
              id.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
@@
              value.name="score")
-
 # allow reference to effect
-temp2 <- melt(temp, 
+temp2 <- reshape2::melt(temp, 
               id.vars=c("transformation", "score"),
@@
               value.name="level")
-
 # do the plot
@@
 
-```{r mainB}
+```{r mainB, fig.width=9, fig.height=3}
 # prep main effects rts data frame
@@
 rts$maxs=rts$RT_RecSqt+rts$ci
-```
-
-```{r mainC, fig.width=9, fig.height=3}
 # plot main effects
@@
 orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Order")));names(orts)[4]<-"Levels"
-
 rts=rbind(vrts,nrts,qrts,orts)
@@
 rts$maxs=rts$RT_RecSqt+rts$ci
-
 rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
-
-ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, group=Levels, ymin=mins, ymax=maxs, shape=Levels, fill=Levels, col=Levels)) +
+ggplot(rts, 
+			 aes(y=RT_RecSqt, x=item_mean_ratio, group=Levels, ymin=mins, 
+			 		ymax=maxs, shape=Levels, fill=Levels, col=Levels)) +
   scale_shape_manual(values=c(21,21,22,22,23,23,24,24)) +
-  scale_fill_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
-  scale_color_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
+  scale_fill_manual(values=c("black", "gray", "black", "gray", "black", "gray", "black", "gray")) +
+  scale_color_manual(values=c("black", "gray", "black", "gray", "black", "gray", "black", "gray")) +
   geom_line() +
@@
 
-```{r 2way}
+```{r 2way, fig.width=10, fig.height=3}
 # prep 2 ways rts data frame
@@
 rts$maxs=rts$RT_RecSqt+rts$ci
-```
-
-```{r foo,fig.width=10, fig.height=3}
 # plot 2 way interactions
@@
 rts6$E2 <- "Order"
-
 rts <- rbind(rts1,rts2,rts3,rts4,rts5,rts6)
@@
 rts$maxs=rts$RT_RecSqt+rts$ci
-
 rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
 rts$F3 <- factor(paste(rts$F1,rts$F2))
-
 rts$E1 <- factor(rts$E1, levels=c("Number","Order","Quantity","Vagueness"))
 rts$E2 <- factor(rts$E2, levels=c("Number","Order","Quantity","Vagueness"))
-
 # The palette with black:
 cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-
 ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, group=F3, col=F2, pch=F1)) + 
   facet_grid(E2~E1, drop=F) + 
-  geom_point(cex=3) + geom_line() + theme_bw() + 
+  geom_point(cex=3) + 
+	geom_line() + 
+	theme_bw() + 
   theme(aspect.ratio=1, panel.grid=element_blank()) +
@@
   theme_bw() +
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust = 1)) +
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
+  			legend.key = element_blank(), aspect.ratio=1, 
+  			axis.text.x = element_text(angle = 45, hjust = 1)) +
   facet_grid(~Number)
@@
 
-```{r plot3way}
+```{r plot3way, fig.width=10,fig.height=5}
 rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Quantity"))
@@
   theme(legend.title=element_blank(), legend.position="top")
-```
-
-```{r bar}
 rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Order"))
@@
   theme(legend.title=element_blank(), legend.position="top")
-```
-
-```{r daz}
 rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Quantity","Order"))
@@
   theme(legend.title=element_blank(), legend.position="top")
-```
-
-```{r foo2}
 rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Quantity","Order"))
@@
   theme(legend.title=element_blank(), legend.position="top")
-```
-
-```{r foo3, fig.width=10,fig.height=5}
 grid.arrange(nrow=2, p1,p2,p3,p4)
@@
   theme_bw() +
-  theme( legend.key = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), axis.text.x = element_text(angle = 45, hjust = 1), aspect.ratio=1) +
+  theme( legend.key = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1),
+  			 axis.text.x = element_text(angle = 45, hjust = 1), aspect.ratio=1) +
   facet_grid(Vagueness+Order~Number+Quantity)
@@
   theme_bw() +
-  theme( legend.key = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), axis.text.x = element_text(angle = 45, hjust = 1), aspect.ratio=1) +
+  theme( legend.key = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), 
+  			 axis.text.x = element_text(angle = 45, hjust = 1), aspect.ratio=1) +
   facet_grid(Vagueness+Order~Number+Quantity)
@@
 
-```{r showCOmpression, fig.width=10, fig.height=7, echo=F}
+```{r showCompression, fig.width=10, fig.height=7, echo=F}
 #  Show compression of ratios of numbers of dots in the items
@@
 item_mean_ratio_equally_spaced = seq(from = mean(c(6 / 15, 15 / 24)), to = mean(c(36 / 45, 45 / 54)), length.out = 4)
-
 plot(y=c(2,2,2,2), x=item_range_ratio, xlim=c(0,1), ylim=c(0,3), ylab="", yaxt='n', col="blue", pch=21, bg='blue', type='o', xaxt='n', xlab="ratio: small is easily discriminable", main="Discriminability of each item, expressed as ratios between array magnitudes")
@@
 points(y=c(2.3,2.3,2.3,2.3), x=item_range_ratio_equally_spaced, col="lightblue", pch=21, bg='lightblue', type='o')
-
 points(y=c(1,1,1,1), x=item_mean_ratio, pch=21, col='red', bg='red', type='o')
@@
 axis(1, at=seq(0,1,0.1))
-
 ```
@@
 
-```{r ols}
-load("dd.Rda")
-
-options(width=140)
-```
-
 Pairs plot
@@
 ```{r pairs, fig.width=7, fig.height=7}
-# pairs plot
 pairs(dd[,c("RT_RecSqt", "Number", "Vagueness", "RTprev_RecSqt")], pch = 19, cex=.075)
@@
 n=length(y)
-y0=confint(v4)[1:n,1]
-y1=confint(v4)[1:n,2]
-
+y0=confint(v4)[,1]
+y1=confint(v4)[,2]
 y=coef(v4)[-1] # omit intercept
@@
 y1=confint(v4)[-1,2]
-
 plot(y, xaxt="n", xlab="", ylab="RT_RecSqt\n", pch=19, ylim=extendrange(y, f=.10), main="Speed")
@@
 cat("k is:",
-collin.fnc(dd[,c("s_Trl", "c_Num", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_RecSqt", "nchar_instr", "measurement")])$cnumber
+collin.fnc(dd[,	c("s_Trl", "c_Num", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", 
+									"RTprev_RecSqt", "nchar_instr", "measurement")
+							])$cnumber
 )
@@
 par(mar=c(1.1,3.1,1.1,1.1), pty='s')
-plot(varclus(as.matrix(dd[,c("s_Trl", "c_Num", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_RecSqt", "nchar_instr", "measurement")])))
+plot(varclus(
+	as.matrix(dd[,c("s_Trl", "c_Num", "c_Vag", "c_Qty", "c_Ord",
+									"item_mean_ratio", "RTprev_RecSqt", "nchar_instr", "measurement")])))
 ```
@@
 
-```{r baayneMplots, fig.width=9, fig.height=3}
+```{r baayenMplots, fig.width=9, fig.height=3}
 # Baayen 4-plot model criticism
-
 par(mfrow=c(1,4), pty='s')
@@
 # plot scaled residuals density
-plot(density(v4$rstand))
+plot(density(v4$rstand), main='density of\nscaled residuals')
 # plot sample quantiles versus theoretical quantiles
@@
 # plot standardised residuals versus fitted values
-plot(v4$rstand ~ fitted(v4), pch='.')
-# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
+plot(v4$rstand ~ fitted(v4), pch='.', main='std resid\nv fitted')
+# absolute standardised residuals greater than 2.5 are candidates 
+# for being outliers, the abline identifies them on the plot
 abline(h=c(-2.5,2.5))
@@
 # plot dffits
-plot(dffits, type='h')
-```
-
-```{r which}
+plot(dffits, type='h', main='dffits')
 # here no overly influential
 w = which.influence(fit = v4, cutoff = 0.4)
-w
+if (length(w)==0) {
+	print('There are no overly influential data points')
+} else {
+	print('There are some overly influential variables:');
+	print(w)
+}
 ```
@@
 ```{r lmer5}
-v5 <- lmer(data=dd, 
+v5 <- lmerTest::lmer(data=dd, 
            RT_RecSqt ~ 
@@
 y=eq[rev(rownames(eq)),]
-
 par(mar=c(4,10,1,1))
@@
 # Baayen 4-plot model criticism
-
 par(mfrow=c(1,4), pty='s')

### 2016-07-29 experiments/e2/writeup/functions/gatherData.R

@@ -1,24 +1,0 @@
-# R version of gatherData.sh
-
-gatherData = function(number_of_valid_subjects) {
-  # message("gathering data ... ")
-  # number_of_valid_subjects <- 30
-  data_dir <- '../experimentCode/output/'
-  
-  column.headers.df <-  head( read.table(
-    paste(data_dir,'subject1.data',sep=''),
-    header=TRUE),0)
-  
-  gathered.data <- column.headers.df
-  
-  for (subject in 1:number_of_valid_subjects) {
-    current.filename <- paste(data_dir, 'subject', subject, '.data', sep='')
-    current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = FALSE)
-    gathered.data <- rbind(gathered.data, current.data)
-  }
-  
-  # message("returning gathered data")
-  
-  return(gathered.data)
-  
-}

### 2016-07-30 papers/2016/vcr.tex
got a working version of a 2016 draft in apa6 format

--- 2016-07-27 papers/2015/resubmission_drafts/2015_resubmission_draft/vagueness-dec-2015.tex
@@
-\documentclass[doc,apacite]{apa6}
-\usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{color}
-\usepackage{booktabs}
-\usepackage{layout}
-\usepackage{apacite}
+\documentclass[
+a4paper 
+%, man
+, doc
+%, jou
+, draftfirst 
+, longtable
+]{apa6} 
+\usepackage{rotating} 
+                          % use this in jou and doc modes only, with \sideways. 
+                          % In man mode use the documentclass option longtable and do not load rotating.
+\usepackage{csquotes}  
+\usepackage[american]{babel}                     
+\usepackage[style=apa, sortcites=true, sorting=nyt, backend=biber, doi=false, isbn=false]{biblatex}
+    \DeclareLanguageMapping{american}{american-apa}
+    \addbibresource{vcr.bib}
 \usepackage{datetime}
-\usepackage{rotating}
-\title{Vagueness as cost reduction}
-\date{\today}
-\authornote{\today ~\currenttime}
-\shorttitle{Vagueness as cost reduction}
+\usepackage{url}
+
+                         %% OTHER DOCUMENTCLASS OPTIONS 
+                         % floatsintext (in man mode do not flush floats, better for internal reading, 
+                         %                    but do not use for submission to journals)
+                         % mask (masks references that are marked as self-references, for blind peer review)
+                         % longtable (in man mode only)
+                         % notxfonts (in jou mode suppresses txfonts in case pslatex or times is preferred)
+                         % notimes (in jou mode use Computer Modern instead of loading txfonts or pslatex or times)
+                         % notab (in jou mode suppress auto stretching of tabular environments to fill their float)
+                         % helv (in man mode use Helvetica instead of Computer Modern)
+                         % tt (in man mode use typewriter font)
+                         % draftfirst or draftall (places the word DRAFT on first page or all pages)
+                         % draft (is passed to, and handled by, the article documentclass)
+
+\SetWatermarkAngle{0} % Angle at which the watermark text is drawn
+\SetWatermarkColor{red} % Color of the watermark
+%\SetWatermarkLightness{?real?} Lightness of the watermark text (1=white, 0=black)
+%\SetWatermarkFontSize{?length?} Font size of the watermark text
+\SetWatermarkScale{0.5} % Scaling of the watermark text 
+%\SetWatermarkHorCenter{?length?} Horizontal center of watermark text 
+\SetWatermarkVerCenter{2cm} % Vertical center of watermark text
+\SetWatermarkText{Internal Draft} %Watermark text
+
+\title{Vagueness as Cost Reduction}
+\shorttitle{Vagueness as Cost Reduction}
 \twoauthors{Matt Green}{Kees van Deemter}
+\leftheader{Green, van Deemter} % for even-page headers in jou mode
 \twoaffiliations{Computing Science\\ University of Aberdeen}{Computing Science\\ University of Aberdeen}
-
 \abstract{Much of everyday language is vague, yet standard game-theoretic models make it difficult to see how vague expressions can have benefits over crisp ones in co-operative situations.
@@
 We argue that the results support a view of vagueness where the benefits that vague terms exert are due to other influences that vagueness brings with it rather than to vagueness itself.}
+\keywords{keyword 1, keyword2, keyword3, keyword4}
+\authornote{Contact author is Matt Green mjglab@gmail.com} % The Author Note, containing contact information, acknowledgements, etc.
+\note{\date{\today~\currenttime}} % Notation of manuscript date or other information de- sired beneath affiliation line
+\journal{Aimed at Language, Cognition, and Neuroscience} % typeset in the top left header of page 1 (jou and doc modes only)
+\volume{volume info} % Volume, number, pages; typeset in the top left header in jou and doc modes, underneath the content of \journal
+\ccoppy{Copyright notice etc} % Copyright notice, etc.; typeset in the top right header of page 1 (jou and doc modes only)
+\copnum{Additional text} % Any additional text needed;  typeset in the top right header in jou and doc modes, underneath the content of \ccoppy
 
 \begin{document}
-\maketitle               
-\section{Introduction}
-Vagueness pervades the language that we use on a daily basis.
-In everyday use, language use may be called vague for various reasons.\footnote{See e.g.
-the entry ``vague'' in \cite{penguin}.} In most academic use though, the word `vagueness' has a more particular meaning.
-Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1996vagueness} (a similar definition can be found in \citeA{EgreKlinedinst}, among many others).
+\maketitle %
+%\section{Introduction} % APA does not use Introduction header
+Vagueness pervades the language that we use on a daily basis. In everyday use, language use may be called vague for various reasons (see e.g. the entry ``vague'' in \cite{penguin}.) 
+In most academic use though, the word `vagueness' has a more particular meaning.
+Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1996vagueness} (a similar definition can be found in \textcite{EgreKlinedinst}, among many others).
 The most crucial of these criteria is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects.
@@
 Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions, which are used so frequently \cite{Lipman:2000fk, lipmanvague}.
-By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss'' \cite<>[p.~1]{lipmanvague}.
+By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss'' \cite[][p.~1]{lipmanvague}.
 Lipman explains the point by means of a scenario in which a speaker describes a person to a hearer, who needs to identify that person in the arrivals hall of of an airport.
 Lipman argues that, in such a scenario, a precise description of the person's height (e.g., ``The person's height is 187.96 cm'') would be more useful than a vague one (``The person is tall'').
-Lipman uses this scenario to explain why standard game theory models of communication \cite<e.g.,>{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs.
+Lipman uses this scenario to explain why standard game theory models of communication \cite[e.g.,]{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs.
 The relevant conditions are, broadly speaking, that both interlocutors know all the relevant facts (e.g., both know the person's height precisely) and that the setting for the communication is co-operative.
@@
 
-Several tentative answers to Lipman's question have been offered  \cite<see>{van2009utility, van2010vagueness}.
-Prominent among these answers is the idea that vague expressions are somehow easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite<e.g.,>[]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
-For example, \citeA<>[p.\ 11]{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze''.
+Several tentative answers to Lipman's question have been offered  \parencite[see][]{van2009utility, van2010vagueness}.
+Prominent among these answers is the idea that vague expressions are somehow easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite[e.g.,][]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
+For example, \textcite[][p.\ 11]{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze''.
 We shall refer to this characterisation of the utility of vague language as the \emph{cost reduction} hypothesis.
@@
 Questions concerning optimal language use have many practical applications.
-Natural Language Generation\footnote{NLG systems take data or formulas as input, and transform them into natural language outputs \cite<>[]{reiter2000building}.
+Natural Language Generation\footnote{NLG systems take data or formulas as input, and transform them into natural language outputs \cite[][]{reiter2000building}.
 The process parallels language production in humans.}  (henceforth NLG) systems must make decisions between different formulations of the same information.
@@
 
-In a series of studies of behaviour modification, \citeA<>{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss.
+In a series of studies of behaviour modification, \textcite[]{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss.
 In the weight loss study, participants were told that the study was designed to test the validity of a new (actually fictitious) health index, the HHI (Holistic Health Index).
@@
 The authors explain the improvement in the vague condition for this group as resulting from the participants' freedom to think of themselves a positioned on one end of the range - the end closest to the ideal HHI scores.
-This ``illusion of proximity'' \cite<>[p.~4]{Mishra01042011} to the goal is argued to allow participants to generate positive expectancies that lead to behaviours that improve performance.
+This ``illusion of proximity'' \cite[][p.~4]{Mishra01042011} to the goal is argued to allow participants to generate positive expectancies that lead to behaviours that improve performance.
 In contrast, in the precise conditions, participants did not have this freedom of interpretation, and could not distort the information to bring about the beneficial \emph{illusion of proximity}.
@@
 We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \cite{keefe1996vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases.
-In the terminology of \citeA{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}.
-Furthermore, the experiments of \citeA{Mishra01042011} did not explore benefits in terms of processing cost, but in terms of long-term behaviour change.
+In the terminology of \textcite{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}.
+Furthermore, the experiments of \textcite{Mishra01042011} did not explore benefits in terms of processing cost, but in terms of long-term behaviour change.
 
-Similar issues arise from the work of \citeA{peters2009bringing}.
+Similar issues arise from the work of \textcite{peters2009bringing}.
 The authors carried out a series of studies where participants were required to rate hospitals based on various sources of information about quality of care.
@@
 Error rates were greater in the crisp conditions than the vague conditions when gap size was small, and this difference diminished with increasing gap size until it reversed at the biggest gap size.
-\begin{figure}[htbp]
+\begin{figure*}[htbp]
 \fitfigure{images/resultse1.pdf}
@@
 \label{resultse1}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{table}[htbp]
+\begin{table*}[htbp]
+\begin{center}
 \caption{Table of instructions arranged by condition for the triple (6,15,24). Experiment 2}
@@
 \end{tabular}
-\end{table}
+\end{center}
+\end{table*}
 
@@
 
-\begin{figure}[htbp]
+\begin{figure*}[htbp]
 \centering
@@
 \label{resultse2}
-\end{figure}
+\end{figure*}
 
@@
 \label{borderline-response-distribution-e2}
-\end{figure}
+\end{figure*}
 
@@
 We used the same stimuli as in experiment 2.
-Table \ref{Instructions for e3} shows the instructions.
+Table \ref{instructionse3} shows the instructions.
 % KvD I think we should motivate our choice of materials (especially "far ..er")
 
-\begin{table}[htbp]
-\centering
-\caption{Experiment 3: Instructions, assuming $6,15,24$ dots as the Item, and showing \emph{fewer} instead of \emph{more}}
-\label{Instructions for e3}
+\begin{table*}[htbp]
+\begin{center}
+\caption{Experiment 3 Instructions assuming 6,15,24 dots as the Item, and showing fewer instead of more}
+\label{instructionse3}
 \begin{tabular}{llll}
-\hline
+\toprule
 vagueness&instruction format&selection task&instruction\\
-\hline
+\midrule
 crisp & numeric&matching & Choose a square with 6 dots \\ 
@@
 vague & numeric&comparison & Choose a square with far fewer than 20 dots \\ 
-\hline
+\bottomrule
 \end{tabular}
-\end{table}%
+\end{center}
+\end{table*}
 
@@
 The apparatus was the same as Experiment 1.
-The design was a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e3}).
+The design was a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{instructionse3}).
 Each stimulus was an instruction followed by a triple of dots.
@@
 
-\begin{figure}[htbp]
+\begin{figure*}[htbp]
 \centering
-\includegraphics[width=.5\textwidth]{images/response-time-two-panels-1.pdf}
+\includegraphics[width=.75\textwidth]{images/response-time-two-panels-1.pdf}
 \caption{Results for Experiment 3}
 \label{resultse3}
-\end{figure}
+\end{figure*}
 
@@
 In order to implement the experiment without mentioning numbers,  we used a prime to visually show the numbers of dots that we wanted to refer to as {\em the target} in the instructions. 
-This presentation of a prime before the main trial shares some features with Experiment 2 in \protect \citeA{Izard20081221}, although in that experiment participants were told the numerosity of the prime - called an \emph{inducer} in that paper - in our experiment we did not tell participants the numerosity of the prime array.
+This presentation of a prime before the main trial shares some features with Experiment 2 in \protect \textcite{Izard20081221}, although in that experiment participants were told the numerosity of the prime - called an \emph{inducer} in that paper - in our experiment we did not tell participants the numerosity of the prime array.
 An item was thus a combination of a visual prime, a numeric triple, and a referring expression.
@@
 \end{tabular}
-\end{table*}%
+\end{table*}
 
@@
 
-\begin{figure}[htbp]
+\begin{figure*}[htbp]
 \centering
-\includegraphics[width=.5\textwidth]{images/response-time-two-panels-1-e4.pdf}
+\includegraphics[width=.75\textwidth]{images/response-time-two-panels-1-e4.pdf}
 \caption{Results for Experiment 4}
 \label{resultse4}
-\end{figure}
+\end{figure*}
 
@@
 
-\bibliography{vagueness-dec-2015}
-\clearpage
+\printbibliography
+
 \appendix
+\section{}
+
+Table \ref{AllInstructions} shows the relevant properties of all the experiments in a single view.
+
+\newgeometry{bottom=2cm}
+
+\newcommand\T{\rule{0pt}{2.6ex}}       % Top strut
+\newcommand\B{\rule[-2.2ex]{0pt}{0pt}} % Bottom strut
 
 \begin{sidewaystable}
-\centering
 \caption{All Instructions}
-\label{All Instructions}
-\begin{tabular}{lp{5cm}llllp{2cm}l}
-\midrule
-Experiment&Example instruction&vagueness&quantifier type&symmetry&selection&number of valid targets&definiteness\\
-\midrule
-1&Choose the square with 5 dots&crisp&numeric&symmetric&matching&1&definite\\
-1&Choose the square with few dots&vague&linguistic&asymmetric&comparison&1&definite\\
-\midrule
-2&Choose the square with 6 dots&crisp&numeric&symmetric&matching&1&definite\\
-2&Choose a square with about 10 dots&vague&numeric&symmetric&matching&2&indefinite\\
-2&Choose the square with the fewest dots&crisp&linguistic&asymmetric&comparison&1&definite\\
-2&Choose a square with few dots&vague&linguistic&asymmetric&comparison&2&indefinite\\
-\midrule
-3&Choose a square with 6 dots&crisp&numeric&symmetric&matching&1&indefinite\\
-3&Choose a square with fewer than 20 dots&crisp&numeric&asymmetric&comparison&2&indefinite\\
-3&Choose a square with about 10 dots&vague&numeric&symmetric&matching&2&indefinite\\
-3&Choose a square with far fewer than 20 dots&vague&numeric&asymmetric&comparison&1&indefinite\\
-\midrule
-4&Choose a square with the same number of dots as the target&crisp&linguistic&symmetric&matching&1&indefinite\\
-4&Choose a square with fewer dots than the target&crisp&linguistic&asymmetric&comparison&2&indefinite\\
-4&Choose a square with about the same number of dots as the target&vague&linguistic&symmetric&matching&2&indefinite\\
-4&Choose a square with far fewer dots than the target&vague&linguistic&asymmetric&comparison&1&indefinite\\
-\midrule
+\label{AllInstructions}
+
+    \resizebox{\linewidth}{!}{%
+    \begin{tabular}{clcccccc}
+%    \midrule
+    $E$&Example instruction&$V$&$Q$&Symmetry&Algorithm&$N$&Def\\
+       \toprule
+    1&Choose the square with 5 dots&Crisp&Number&Symmetric&Matching&1&Definite\\
+    1&Choose the square with few dots&Vague&Word&Asymmetric&Comparison&1&Definite\B\\
+%    \midrule
+%\rule{0pt}{0cm} \\
+    2&Choose the square with 6 dots&Crisp&Number&Symmetric&Matching&1&Definite\\
+    2&Choose a square with about 10 dots&Vague&Number&Symmetric&Matching&2&Indefinite\\
+    2&Choose the square with the fewest dots&Crisp&Word&Asymmetric&Comparison&1&Definite\\
+    2&Choose a square with few dots&Vague&Word&Asymmetric&Comparison&2&Indefinite\B\\
+%    \midrule
+    3&Choose a square with 6 dots&Crisp&Number&Symmetric&Matching&1&Indefinite\\
+    3&Choose a square with fewer than 20 dots&Crisp&Number&Asymmetric&Comparison&2&Indefinite\\
+    3&Choose a square with about 10 dots&Vague&Number&Symmetric&Matching&2&Indefinite\\
+    3&Choose a square with far fewer than 20 dots&Vague&Number&Asymmetric&Comparison&1&Indefinite\B\\
+%    \midrule
+    4&Choose a square with the same number of dots as the target&Crisp&Word&Symmetric&Matching&1&Indefinite\\
+    4&Choose a square with fewer dots than the target&Crisp&Word&Asymmetric&Comparison&2&Indefinite\\
+    4&Choose a square with about the same number of dots as the target&Vague&Word&Symmetric&Matching&2&Indefinite\\
+    4&Choose a square with far fewer dots than the target&Vague&Word&Asymmetric&Comparison&1&Indefinite\\
+    \bottomrule
     \end{tabular} 
+    }
+    ~\\
+    E:~Experiment number\\
+    $V$:~Vagueness\\
+    $Q$:~Quantifier used\\
+%    Symmetry:~\\
+%    Algorithm:~\\
+     $N$:~Number of valid targets for the instruction\\
+     Def:~Definiteness of the article in the instruction: a/the
+
 \end{sidewaystable}
+
 \end{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{}
+
+\end{document}
+
+%\begin{figure*}
+\includegraphics[angle=90,origin=c, scale=.5]{images/appendix_table}
+%\end{figure*}

### 2016-08-02 experiments/e0/writeup/e0.Rmd
no reshape only reshape2

@@
 suppressPackageStartupMessages(library(plyr))
-suppressPackageStartupMessages(library(reshape))
+suppressPackageStartupMessages(library(reshape2))
 suppressPackageStartupMessages(library(lattice))

### 2016-08-02 experiments/e3/writeup/e3.Rmd
no reshape only reshape2

@@
 suppressPackageStartupMessages(library(plyr))
-suppressPackageStartupMessages(library(reshape))
 suppressPackageStartupMessages(library(lattice))
@@
 
-```{r boxplot subjects and items}
+```{r boxplot subjects and items, width-12, height=3}
 load("data.Rda")
@@
 ggplot(subdata) + 
-  facet_wrap(~ variable, scales="free", ncol=4) +
+  facet_wrap(~ transformation, scales="free", ncol=4) +
   geom_boxplot(aes(y=value, x=Item, col="Items")) + 

### 2016-08-03 papers/2016/vcr.tex
no longer tracking contribution

@@
 %, jou
-, draftfirst 
+%, draftfirst 
 , longtable
@@
 
-\SetWatermarkAngle{0} % Angle at which the watermark text is drawn
-\SetWatermarkColor{red} % Color of the watermark
-%\SetWatermarkLightness{?real?} Lightness of the watermark text (1=white, 0=black)
-%\SetWatermarkFontSize{?length?} Font size of the watermark text
-\SetWatermarkScale{0.5} % Scaling of the watermark text 
-%\SetWatermarkHorCenter{?length?} Horizontal center of watermark text 
-\SetWatermarkVerCenter{2cm} % Vertical center of watermark text
-\SetWatermarkText{Internal Draft} %Watermark text
+%\SetWatermarkAngle{0} % Angle at which the watermark text is drawn
+%\SetWatermarkColor{red} % Color of the watermark
+%\SetWatermarkLightness{0.7} %Lightness of the watermark text (1=white, 0=black)
+%%\SetWatermarkFontSize{?length?} Font size of the watermark text
+%\SetWatermarkScale{0.5} % Scaling of the watermark text 
+%%\SetWatermarkHorCenter{?length?} Horizontal center of watermark text 
+%\SetWatermarkVerCenter{2cm} % Vertical center of watermark text
+%\SetWatermarkText{Internal Draft} %Watermark text
 
@@
 \keywords{keyword 1, keyword2, keyword3, keyword4}
-\authornote{Contact author is Matt Green mjglab@gmail.com} % The Author Note, containing contact information, acknowledgements, etc.
-\note{\date{\today~\currenttime}} % Notation of manuscript date or other information de- sired beneath affiliation line
-\journal{Aimed at Language, Cognition, and Neuroscience} % typeset in the top left header of page 1 (jou and doc modes only)
-\volume{volume info} % Volume, number, pages; typeset in the top left header in jou and doc modes, underneath the content of \journal
-\ccoppy{Copyright notice etc} % Copyright notice, etc.; typeset in the top right header of page 1 (jou and doc modes only)
-\copnum{Additional text} % Any additional text needed;  typeset in the top right header in jou and doc modes, underneath the content of \ccoppy
-
+\authornote{Contact author is Matt Green: \textsf{mjglab@gmail.com}} 
+% The Author Note, containing contact information, acknowledgements, etc.
+\note{This version compiled at: \textsf{\currenttime~\today}} 
+% Notation of manuscript date or other information desired beneath affiliation line
+\journal{Aimed at:}
+% typeset in the top left header of page 1 (jou and doc modes only)
+\volume{\textsc{Language, Cognition, and Neuroscience}} 
+% Volume, number, pages; typeset in the top left header in jou and doc modes, underneath the content of \journal
+\ccoppy{Draft date:}
+% Copyright notice, etc.; typeset in the top right header of page 1 (jou and doc modes only)
+\copnum{\today}
+% Any additional text needed;  typeset in the top right header in jou and doc modes, underneath the content of \ccoppy
 \begin{document}
-\maketitle %
+\longdate
+\maketitle % no blank line
 %\section{Introduction} % APA does not use Introduction header
@@
 In most academic use though, the word `vagueness' has a more particular meaning.
-Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1996vagueness} (a similar definition can be found in \textcite{EgreKlinedinst}, among many others).
+Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1997vagueness} (a similar definition can be found in \textcite{EgreKlinedinst}, among many others).
 The most crucial of these criteria is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects.
@@
 These results are interesting, and of obvious potential practical importance.
-We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \cite{keefe1996vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases.
+We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \cite{keefe1997vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases.
 In the terminology of \textcite{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}.

### 2016-08-03 papers/springer-chapter/chapter.tex
adding springer-chapter to the main repo. Some other deletions

%!TEX root = book.tex 
% springer volume on vagueness & rationality
%%%%%%%%%%%%%%%%%%%%% chapter.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample chapter
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag %%%%%%%%%%%%%%%%%%%%%%%%%%
\motto{This is a \textsf{motto}:  Use the template \emph{chapter.tex} to style the various elements of your chapter content.}
%% \bibliographystyle{spbasic}
%% spbasic info
% This is an author-year citation style bibliography. As such, it is non-standard LaTeX, and requires a special package file to function properly. Such a package is natbib.sty by Patrick W. Daly
% For Springer medical, life sciences, chemistry, geology, engineering and computer science publications.   
% For use with the natbib package (see below). Default is author-year citations. When citations are numbered, please use \usepackage[numbers]{natbib}.
% The \cite command functions as follows:
 %   \citet{key} ==>>                Jones et al. (1990)
 %   \citet*{key} ==>>               Jones, Baker, and Smith (1990)
 %   \citep{key} ==>>                (Jones et al., 1990)
 %   \citep*{key} ==>>               (Jones, Baker, and Smith, 1990)
 %   \citep[chap. 2]{key} ==>>       (Jones et al., 1990, chap. 2)
 %   \citep[e.g.][]{key} ==>>        (e.g. Jones et al., 1990)
 %   \citep[e.g.][p. 32]{key} ==>>   (e.g. Jones et al., p. 32)
 %   \citeauthor{key} ==>>           Jones et al.
 %   \citeauthor*{key} ==>>          Jones, Baker, and Smith
 %   \citeyear{key} ==>>             1990
\chapter{The Elusive Benefits of Vagueness: the Evidence So Far}
\label{intro} % Always give a unique label
% use \chaptermark{}
% to alter or adjust the chapter heading in the running head

\abstract{Much of everyday language is vague, yet the benefits of vagueness for hearers and readers are proving to be elusive. We discuss a range of earlier controlled experiments with human participants, and we report on a new series of experiments that we conducted in recent years. These new experiments, which focus on vague expressions that are part of referential noun phrases, aim to separate the utility of vagueness (as defined by the existence of borderline cases, as is common in theoretical work on vagueness) from the utility of other factors that tend to co-occur with vagueness. Having presented the evidence, we argue that the evidence so far supports a view of vagueness where the benefits that vague terms exert are due to other influences, rather than to vagueness itself. These factors include: low granularity; the use of evaluative words; the avoidance of overtly numerical words; the existence of comparison strategies; and, lastly and more tentatively, a phenomenon that we call range reduction. Although it is possible that other types of vague expressions (i.e., vague words outside referential noun phrases) behave differently, our work suggests that vagueness itself may not increase the utility of an expression.}

\section{Introduction}
\label{sec:1}

Vagueness pervades the language that we use on a daily basis. In everyday use, language use may be called vague for various reasons -- see e.g. the entry for \emph{vague} in \cite{penguin}.  In most academic use though, the word \emph{vagueness} has a more particular meaning. \citeauthor{keefe1997vagueness}, for example, state \begin{quotation}vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes \cite[p.\ 4]{keefe1997vagueness}\end{quotation} (a similar definition can be found in \cite{EgreKlinedinst}, among many others).  The most crucial of these criteria is the existence of borderline cases: \begin{quotation}a word is precise if it describes a well-defined set of objects. By contrast, a word is vague if it is not precise \cite[p.\ 1]{lipmanvague}\end{quotation} A typical example is the word \emph{tall}, as applied to people for example, because here is no precise, known height which separates those who are tall from those who are not. The crucial point is that \emph{tall} admits borderline cases (i.e., people who may or may not count as tall), which are the hallmark of vagueness as we use the term.

Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions, which are used so frequently \cite{Lipman:2000fk, lipmanvague}. By introducing borderline cases, these expressions create potential misunderstandings, thereby creating \begin{quotation}a worldwide several-thousand year efficiency loss \cite[p.\ 1]{lipmanvague}\end{quotation} \citeauthor{lipmanvague} explains the point by means of a scenario in which a speaker describes a person to a hearer, who needs to identify that person in the arrivals hall of of an airport. Lipman argues that, in such a scenario, a precise description of the person's height, e.g., \textsl{The person's height is 187.96~cm} would be more useful than a vague one, e.g., \textsl{The person is tall}. Lipman uses this scenario to explain why standard game theory models of communication, \citep[e.g.][]{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs. The relevant conditions are, broadly speaking, that both interlocutors know all the relevant facts (e.g., both know the person's height precisely) and that the setting for the communication is co-operative. These conditions exclude deliberate deception, and rhetorical situations like political debate and advertising where the intention is to persuade the interlocutor to adopt some point of view, where the persuasion might not be intended to be in the addressee's best interest.

Lipman argued that such the efficiency loss resulting from vague expressions would be unlikely to have arisen unless there are advantages as well as disadvantages associated with vague expressions. Lipman asked, essentially, what these advantages might be, and how they might find a place in a game-theoretical explanation. In this paper, we focus on the first part of Lipman's question.

Several tentative answers to Lipman's question have been offered -- see \citet{van2009utility, van2010vagueness}. Prominent among these answers is the idea that vague expressions are somehow easier to process, by a speaker and/or a hearer, than expressions that are not \emph{vague}, i.e., that are \emph{crisp} \citep[e.g.,][]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}. For example, Lipman writes: \begin{quotation}For the listener, information which is too specific may require more effort to analyze \citep[][p.\ 11]{lipmanvague}  \end{quotation} We shall refer to this characterisation of the utility of vague language as the \emph{cost reduction} hypothesis. The idea of vagueness as cost reduction can take various shapes, but the basic hypothesis is that it is easier for people to think in terms of loosely defined categories, such as \emph{quite a few}, or \emph{many}, than in terms of crisply defined ones, such as \emph{thirteen}, or \emph{237}. This predicts that whenever a vague expression can perform the same communicative task as a crisp one, it is rational to choose the vague expression. The corollary for comprehension is that vague expressions should be understood more readily than precise expressions.

Questions concerning optimal language use have many practical applications. Natural Language Generation\footnote{\textsc{nlg} systems take data or formulas as input, and transform them into natural language outputs \cite[][]{reiter2000building}. The process parallels language production in humans.}  (henceforth \textsc{nlg}) systems must make decisions between different formulations of the same information. For example, if a man's height is 6 foot 2 inches, this could be expressed as \emph{187.96 metres}, \emph{6 foot 2}, or \emph{tall}, among other ways, and the \textsc{nlg} system must decide between these.  The problem is particularly relevant for \textsc{nlg} systems that take numbers as input, as many do. In the context of an \textsc{nlg} system faced with a practical decision of this kind, Lipman's question becomes \emph{Under what circumstances should vague terms be produced?}  Relevant applications include weather forecasting on the basis of numerical weather data such as temperature and wind speed \citep{goldberg1994using, turner2006generating}, and medical decision support on the basis of clinical measurement such as oxygen saturation, heart rhythm, etc. \citep{Hripcsak01032009, hunter2008summarising, portet2009automatic}. At present, such \textsc{nlg} systems are often forced to make decisions concerning the level of precision in the utterances that they generate (e.g., \emph{the temperature will be in the high twenties tomorrow}) on the basis of little more than intuition. A better understanding of the benefits of different precision levels for readers would allow these systems to become more useful.

The cost reduction hypothesis is of direct relevance to psycholinguists interested in language comprehension, and additionally to psycholinguists interested in language production, for example in connection with the question of audience design \citep{Clark1982287}. For, to the extent that speakers and writers choose vague expressions over and above crisp ones because the former are easier to process for hearers than the latter, the cost reduction hypothesis suggests that speakers design their utterance for optimal benefit to their hearers -- out of altruism, so to speak.

The utility of vagueness is the attested aim of a small number of studies, but most of these have focussed on vagueness in a different sense, and focussing on different types of benefits for hearers. Two recent studies can illustrate both issues.

\paragraph{Behaviour modification experiments} 
In a series of studies of behaviour modification, \citet[]{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss. In the weight loss study, participants were told that the study was designed to test the validity of a new (actually fictitious) health index, the HHI (Holistic Health Index). They were told that an ideal HHI score lies in the range of 45 to 55. In a longitudinal study, participants submitted their height, weight, hydration level, gender, and age to a computer each week. Participants were told that two algorithms would be used to compute their HHI, and that it was possible that the two algorithms might give different values initially, but would converge over the course of the study to a single value. They were also told that if the two algorithms did give different values, then the true score lay between the two values. In one condition, which the authors called the precise condition, the two algorithms gave the same score. In the other condition, which the authors called the vague condition, one algorithm added 3\% to the score while the other algorithm subtracted 3\% from the score, yielding a range of values whose midpoint was the same as the two values given in the precise condition.

\paragraph{Ratings experiments} 
Similar issues arise from the work of \citet{peters2009bringing}. The authors carried out a series of studies where participants were required to rate hospitals based on various sources of information about quality of care. There was a between-subjects manipulation based on numeracy. The format of the information was manipulated within subjects: either numbers only were presented, or both numbers and evaluative categories were presented (e.g., \emph{Poor}, \emph{Fair}, \emph{Good}, \emph{Excellent}, with crisp visual boundary lines between the categories). Results showed that, for low-numeracy participants, the presence of evaluative categories resulted in a diminished influence of an irrelevant affective state on the ratings. For all participants, the presence of evaluative categories resulted in better decisions and in a greater use of the most important and reliable types of information, such as survival rates.

It is, however, questionable whether the \emph{evaluative categories} manipulation in this study can be considered a manipulation of vagueness. Certainly, terms like \emph{Fair} admit the possibility of borderline cases. However, given that the boundaries between the categories were marked crisply, and that therefore the categories mapped crisply to numerical values, it becomes doubtful whether any borderline cases could be conceived to arise in fact. For example, \emph{Fair} was mapped to 60\% -- 70\% for the variable \emph{percentage of heart attack patients given recommended treatment (ACE inhibitor)}. Accordingly, rather than the vagueness of categories such as \emph{Poor}, Peters et al. emphasise the evaluative content inherent in these categories, and the affective potential of the evaluative content rather than the vagueness of the terms like \emph{Fair}.

\paragraph{Our own experiments}
The experiments reported in the present paper put the cost reduction hypothesis to the test. The question that we are trying to answer is whether vague expressions are processed more easily by readers than crisp ones. Like Lipman, we focus on situations where numerical information is used in order to identify a referent. Reference, in other words, will be the linguistic task on which we focus, partly because of the interest that this topic has recently drawn from the NLG community.  In focussing on benefits for the hearer, we will leave aside the question of audience design, leaving this for later research.

In using references to quantities to test the cost reduction hypothesis we are only testing one aspect of vagueness in a particular context. This limits the applicability of our results. It also has the advantage that we can explore the costs and benefits of vagueness more thoroughly in that context. Since one prevalent view of vagueness is that a vague expression is never preferable to a crisp equivalent, a demonstration of a benefit for vagueness in any context would advance the discussion.

In our experiments we used a speeded forced choice task to compare the processing costs of different references to quantities. In this context, speed and accuracy of responses are the key dimensions on which the different references can be compared. The stimuli in the experiments were sets of dot arrays containing various numbers of dots. The forced choice was to identify one dot array given a reference to a given quantity of dots. We manipulated the references in several ways across a series of four experiments.

Our main manipulation was always of vagueness: we constructed crisp and vague versions of references to the same dot array. For example, in experiment 1 the instruction presented to the participant could be \emph{Choose the square with many dots.} (vague condition), or \emph{Choose the square with 20 dots.} (crisp condition), identifying the same dot array.

Ideally, we would manipulate vagueness independently of other variables. However, in practice, and given the constraints imposed by the use of natural language, any manipulation of vagueness introduces variance along other dimensions as well. Therefore we carried out further experiments to try to address and control variance along other dimensions.

One non-vagueness source of variance was the difference between numerical and verbal format in the references. For example, in the instructions for experiment 1 given above, \emph{many} is in verbal format (as well as vague) and \emph{20} is in numerical format (as well as crisp). Therefore any difference observed between responses to the \emph{many} instruction and the \emph{20} instruction could be due either to vagueness or to {\bf instruction format}. It is possible to create vague and crisp references in each of these instruction formats: for example in numerical instruction format we can have \emph{20} (crisp) and \emph{about 20} (vague), and in verbal instruction format we can have \emph{the most} (crisp) and \emph{many} (vague). Experiment 2 varied vagueness and instruction format factorially in a similar way to this, to try to tease apart these two sources of variance.

Another potential non-vagueness source of variance between responses to the \emph{many} (verbal vague) references and the \emph{20} (numeric crisp) references lies in the method used to identify a referent (we call this the {\bf selection algorithm} source of variance). Identifying the array with \emph{many} dots might be done by a \emph{comparison} algorithm, identifying that one array is more numerous than the others without establishing the numerosity of any arrays, whereas identifying the array with \emph{20} dots, which might be done by a \emph{matching} algorithm would require an estimate of the numerosity of the arrays. It seems reasonable that comparison would be faster than matching because it does not require estimates of cardinality. Furthermore the problem persists when using references like \emph{the most} (verbal crisp) and \emph{about 20} (numeric vague), since they too vary not only along the vague/crisp dimension and the numeric/verbal dimension but also along the comparison/matching dimension. Experiments 3 and 4 addressed the \emph{selection algorithm} source of variance - experiment 3 crossed vague/crisp and comparison/matching factorially using only numerical references, and experiment 4 crossed vague/crisp and comparison/matching factorially using only verbal references.

\section{Experiment 1}
\label{sec:2}
% Always give a unique label
% and use \ref{<label>} for cross-references
% and \cite{<label>} for bibliographic references
% use \sectionmark{}
% to alter or adjust the section heading in the running head
Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Further on please use the \LaTeX\ automatism for all your cross-references and citations.

\subsection{Introduction}%\label{}
We used a forced choice task to compare choices made in response to vague instructions against choices made in response to crisp instructions. The participant was presented with an instruction like ``Choose the square with many dots'' in the vague conditions, or ``Choose the square with 20 dots'' in the crisp conditions. Then two dot arrays were presented in the form of squares containing a number of dots. Fig \ref{stimuluse1} shows an example stimulus. The participant was required to identify the square that corresponded with the instruction, by pressing the appropriate key.
Response time and accuracy were recorded for analysis.

Our main manipulation was of the vagueness of the instruction, with two levels, vague and crisp. Table \ref{instructionse1} shows examples from each condition.  We also manipulated how discriminable the dot arrays were. One array always contained 25 dots: the other contained either 5, 10, 15, 20, 30, 35, 40, or 45 dots. This led to numerical differences of 5, 10, 15, and 20, with lower differences resulting in less discriminable arrays and larger differences representing more discriminable arrays. There is evidence that when the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the numerosity of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene199647}. Where a number was mentioned in the instruction, it was always in the form of an Arabic numeral.

The instructions indicated the larger of the two dot arrays equally often as they indicated the smaller of the dot arrays, so that participants could not systematically choose the larger or smaller array as a successful response strategy. There is evidence that when two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}. We controlled which side the smaller number appeared on to avoid systematic influences of this effect.

\begin{table}
\caption{Table of instructions for the pair (5,25). Experiment 1}
\label{instructionse1}       % Give a unique label
\begin{tabular}{| l | l |}
\hline\noalign{\smallskip}
vagueness&example\\
\noalign{\smallskip}\svhline\noalign{\smallskip}
crisp 	& 	Choose the square with 5 dots \\
vague	&	Choose the square with few dots\\
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
$^a$ Table foot note (with superscript)\\
\end{table}

\section{Experiment 2}
\label{sec:3}
% Always give a unique label
% and use \ref{<label>} for cross-references
% and \cite{<label>} for bibliographic references
% use \sectionmark{}
% to alter or adjust the section heading in the running head
Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Further on please use the \LaTeX\ automatism for all your cross-references and citations.

\section{Experiment 3}
\label{sec:4}
% Always give a unique label
% and use \ref{<label>} for cross-references
% and \cite{<label>} for bibliographic references
% use \sectionmark{}
% to alter or adjust the section heading in the running head
Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Further on please use the \LaTeX\ automatism for all your cross-references and citations.

\section{Experiment 4}
\label{sec:4}
% Always give a unique label
% and use \ref{<label>} for cross-references
% and \cite{<label>} for bibliographic references
% use \sectionmark{}
% to alter or adjust the section heading in the running head
Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Further on please use the \LaTeX\ automatism for all your cross-references and citations.

\section{Dummy Template Section}
\label{sec:5}
% Always give a unique label
% and use \ref{<label>} for cross-references
% and \cite{<label>} for bibliographic references
% use \sectionmark{}
% to alter or adjust the section heading in the running head
Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Further on please use the \LaTeX\ automatism for all your cross-references and citations.

Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.

Use the standard \verb|equation| environment to typeset your equations, e.g.
%
\begin{equation}
a \times b = c\;,
\end{equation}
%
however, for multiline equations we recommend to use the \verb|eqnarray|
environment\footnote{In physics texts please activate the class option \texttt{vecphys} to depict your vectors in \textbf{\itshape boldface-italic} type - as is customary for a wide range of physical subjects.}.
\begin{eqnarray}
a \times b = c \nonumber\\
\vec{a} \cdot \vec{b}=\vec{c}
\label{eq:01}
\end{eqnarray}

\subsection{Subsection Heading}
\label{subsec:2}
Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Furtheron please use the \LaTeX\ automatism for all your cross-references\index{cross-references} and citations\index{citations} as has already been described in Sect.~\ref{sec:2}.

\begin{quotation}
Please do not use quotation marks when quoting texts! Simply use the \verb|quotation| environment -- it will automatically render Springer's preferred layout.
\end{quotation}

\subsubsection{Subsubsection Heading}
Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Furtheron please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{subsec:2}, see also Fig.~\ref{fig:1}\footnote{If you copy text passages, figures, or tables from other works, you must obtain \textit{permission} from the copyright holder (usually the original publisher). Please enclose the signed permission with the manucript. The sources\index{permission to print} must be acknowledged either in the captions, as footnotes or in a separate section of the book.}

Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.

% For figures use
%
\begin{figure}[b]
\sidecaption
% Use the relevant command for your figure-insertion program
% to insert the figure file.
% For example, with the option graphics use
\includegraphics[scale=.65]{figure}
%
% If not, use
%\picplace{5cm}{2cm} % Give the correct figure height and width in cm
%
\caption{If the width of the figure is less than 7.8 cm use the \texttt{sidecapion} command to flush the caption on the left side of the page. If the figure is positioned at the top of the page, align the sidecaption with the top of the figure -- to achieve this you simply need to use the optional argument \texttt{[t]} with the \texttt{sidecaption} command}
\label{fig:1}       % Give a unique label
\end{figure}

\paragraph{Paragraph Heading} %
Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Furtheron please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.

Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.

For typesetting numbered lists we recommend to use the \verb|enumerate| environment -- it will automatically render Springer's preferred layout.

\begin{enumerate}
\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
\begin{enumerate}
\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
\end{enumerate}
\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
\end{enumerate}

\subparagraph{Subparagraph Heading} In order to avoid simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}, see also Fig.~\ref{fig:2}.

Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.

For unnumbered list we recommend to use the \verb|itemize| environment -- it will automatically render Springer's preferred layout.

\begin{itemize}
\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development, cf. Table~\ref{tab:1}.}
\begin{itemize}
\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
\end{itemize}
\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
\end{itemize}

\begin{figure}[t]
\sidecaption[t]
% Use the relevant command for your figure-insertion program
% to insert the figure file.
% For example, with the option graphics use
\includegraphics[scale=.65]{figure}
%
% If not, use
%\picplace{5cm}{2cm} % Give the correct figure height and width in cm
%
\caption{Please write your figure caption here}
\label{fig:2}       % Give a unique label
\end{figure}

\runinhead{Run-in Heading Boldface Version} Use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.

\subruninhead{Run-in Heading Italic Version} Use the \LaTeX\ automatism for all your cross-refer\-ences and citations as has already been described in Sect.~\ref{sec:2}\index{paragraph}.
% Use the \index{} command to code your index words
%
% For tables use
%
\begin{table}
\caption{Please write your table caption here}
\label{tab:1}       % Give a unique label
%
% For LaTeX tables use
%
\begin{tabular}{p{2cm}p{2.4cm}p{2cm}p{4.9cm}}
\hline\noalign{\smallskip}
Classes & Subclass & Length & Action Mechanism  \\
\noalign{\smallskip}\svhline\noalign{\smallskip}
Translation & mRNA$^a$  & 22 (19--25) & Translation repression, mRNA cleavage\\
Translation & mRNA cleavage & 21 & mRNA cleavage\\
Translation & mRNA  & 21--22 & mRNA cleavage\\
Translation & mRNA  & 24--26 & Histone and DNA Modification\\
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
$^a$ Table foot note (with superscript)
\end{table}
%
%\section{Section Heading}
%\label{sec:3}
% Always give a unique label
% and use \ref{<label>} for cross-references
% and \cite{<label>} for bibliographic references
% use \sectionmark{}
% to alter or adjust the section heading in the running head
Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Furtheron please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.

Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.

If you want to list definitions or the like we recommend to use the Springer-enhanced \verb|description| environment -- it will automatically render Springer's preferred layout.

\begin{description}[Type 1]
\item[Type 1]{That addresses central themes pertainng to migration, health, and disease. In Sect.~\ref{sec:1}, Wilson discusses the role of human migration in infectious disease distributions and patterns.}
\item[Type 2]{That addresses central themes pertainng to migration, health, and disease. In Sect.~\ref{subsec:2}, Wilson discusses the role of human migration in infectious disease distributions and patterns.}
\end{description}

\subsection{Subsection Heading} %
In order to avoid simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Use the \LaTeX\ automatism for all your cross-references and citations citations as has already been described in Sect.~\ref{sec:2}.

Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.

\begin{svgraybox}
If you want to emphasize complete paragraphs of texts we recommend to use the newly defined Springer class option \verb|graybox| and the newly defined environment \verb|svgraybox|. This will produce a 15 percent screened box 'behind' your text.

If you want to emphasize complete paragraphs of texts we recommend to use the newly defined Springer class option and environment \verb|svgraybox|. This will produce a 15 percent screened box 'behind' your text.
\end{svgraybox}

\subsubsection{Subsubsection Heading}
Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Furtheron please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.

Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.

\begin{theorem}
Theorem text goes here.
\end{theorem}
%
% or
%
\begin{definition}
Definition text goes here.
\end{definition}

\begin{proof}
%\smartqed
Proof text goes here.
\qed
\end{proof}

\paragraph{Paragraph Heading} %
Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text. Furtheron please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.

Note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.
%
% For built-in environments use
%
\begin{theorem}
Theorem text goes here.
\end{theorem}
%
\begin{definition}
Definition text goes here.
\end{definition}
%
\begin{proof}
\smartqed
Proof text goes here.
\qed
\end{proof}
%
\begin{acknowledgement}
If you want to include acknowledgments of assistance and the like at the end of an individual chapter please use the \verb|acknowledgement| environment -- it will automatically render Springer's preferred layout.
\end{acknowledgement}
%
\section*{Appendix}
\addcontentsline{toc}{section}{Appendix}
%
When placed at the end of a chapter or contribution (as opposed to at the end of the book), the numbering of tables, figures, and equations in the appendix section continues on from that in the main text. Hence please \textit{do not} use the \verb|appendix| command when writing an appendix at the end of your chapter or contribution. If there is only one the appendix is designated ``Appendix'', or ``Appendix 1'', or ``Appendix 2'', etc. if there is more than one.

\begin{equation}
a \times b = c
\end{equation}
% Problems or Exercises should be sorted chapterwise
%\section*{Problems}
%\addcontentsline{toc}{section}{Problems}
%
% Use the following environment.
% Don't forget to label each problem;
%% the label is needed for the solutions' environment
%\begin{prob}
%\label{prob1}
%A given problem or Excercise is described here. The
%problem is described here. The problem is described here.
%\end{prob}
%
%\begin{prob}
%\label{prob2}
%\textbf{Problem Heading}\\
%(a) The first part of the problem is described here.\\
%(b) The second part of the problem is described here.
%\end{prob}

%\input{referenc}
% BibTeX users please use
\bibliographystyle{spbasic} %plain, harvard, apalike, chicago, astron, authordate, natbib.
\bibliography{bibliography}

### 2016-08-03 papers/springer-chapter/chapter_template.tex
adding springer-chapter to the main repo. Some other deletions

--- 2016-08-03 papers/springer-chapter/chapter.tex
@@
-%!TEX root = book.tex 
-% springer volume on vagueness & rationality
 %%%%%%%%%%%%%%%%%%%%% chapter.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag %%%%%%%%%%%%%%%%%%%%%%%%%%
-\motto{This is a \textsf{motto}:  Use the template \emph{chapter.tex} to style the various elements of your chapter content.}
-%% \bibliographystyle{spbasic}
-%% spbasic info
-% This is an author-year citation style bibliography. As such, it is non-standard LaTeX, and requires a special package file to function properly. Such a package is natbib.sty by Patrick W. Daly
-% For Springer medical, life sciences, chemistry, geology, engineering and computer science publications.   
-% For use with the natbib package (see below). Default is author-year citations. When citations are numbered, please use \usepackage[numbers]{natbib}.
-% The \cite command functions as follows:
- %   \citet{key} ==>>                Jones et al. (1990)
- %   \citet*{key} ==>>               Jones, Baker, and Smith (1990)
- %   \citep{key} ==>>                (Jones et al., 1990)
- %   \citep*{key} ==>>               (Jones, Baker, and Smith, 1990)
- %   \citep[chap. 2]{key} ==>>       (Jones et al., 1990, chap. 2)
- %   \citep[e.g.][]{key} ==>>        (e.g. Jones et al., 1990)
- %   \citep[e.g.][p. 32]{key} ==>>   (e.g. Jones et al., p. 32)
- %   \citeauthor{key} ==>>           Jones et al.
- %   \citeauthor*{key} ==>>          Jones, Baker, and Smith
- %   \citeyear{key} ==>>             1990
-\chapter{The Elusive Benefits of Vagueness: the Evidence So Far}
+%\motto{Use the template \emph{chapter.tex} to style the various elements of your chapter content.}
+\chapter{Chapter Heading}
 \label{intro} % Always give a unique label
@@
 
-\abstract{Much of everyday language is vague, yet the benefits of vagueness for hearers and readers are proving to be elusive.
-We discuss a range of earlier controlled experiments with human participants, and we report on a new series of experiments that we conducted in recent years.
-These new experiments, which focus on vague expressions that are part of referential noun phrases, aim to separate the utility of vagueness (as defined by the existence of borderline cases, as is common in theoretical work on vagueness) from the utility of other factors that tend to co-occur with vagueness.
-Having presented the evidence, we argue that the evidence so far supports a view of vagueness where the benefits that vague terms exert are due to other influences, rather than to vagueness itself.
-These factors include: low granularity; the use of evaluative words; the avoidance of overtly numerical words; the existence of comparison strategies; and, lastly and more tentatively, a phenomenon that we call range reduction.
-Although it is possible that other types of vague expressions (i.e., vague words outside referential noun phrases) behave differently, our work suggests that vagueness itself may not increase the utility of an expression.}
+\abstract*{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
+The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
+This allows unregistered users to read the abstract as a teaser for the complete chapter.
+As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.
+Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
+source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
+Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
 
-\section{Introduction}
+\abstract{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
+The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
+This allows unregistered users to read the abstract as a teaser for the complete chapter.
+As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.\newline\indent
+Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
+source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
+Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
+
+\section{Section Heading}
 \label{sec:1}
+Use the template \emph{chapter.tex} together with the Springer document class SVMono (monograph-type books) or SVMult (edited books) to style the various elements of your chapter content in the Springer layout.
 
-Vagueness pervades the language that we use on a daily basis.
-In everyday use, language use may be called vague for various reasons -- see e.g.
-the entry for \emph{vague} in \cite{penguin}.
-In most academic use though, the word \emph{vagueness} has a more particular meaning.
-\citeauthor{keefe1997vagueness}, for example, state \begin{quotation}vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes \cite[p.\ 4]{keefe1997vagueness}\end{quotation} (a similar definition can be found in \cite{EgreKlinedinst}, among many others).
-The most crucial of these criteria is the existence of borderline cases: \begin{quotation}a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise \cite[p.\ 1]{lipmanvague}\end{quotation} A typical example is the word \emph{tall}, as applied to people for example, because here is no precise, known height which separates those who are tall from those who are not.
-The crucial point is that \emph{tall} admits borderline cases (i.e., people who may or may not count as tall), which are the hallmark of vagueness as we use the term.
-
-Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions, which are used so frequently \cite{Lipman:2000fk, lipmanvague}.
-By introducing borderline cases, these expressions create potential misunderstandings, thereby creating \begin{quotation}a worldwide several-thousand year efficiency loss \cite[p.\ 1]{lipmanvague}\end{quotation} \citeauthor{lipmanvague} explains the point by means of a scenario in which a speaker describes a person to a hearer, who needs to identify that person in the arrivals hall of of an airport.
-Lipman argues that, in such a scenario, a precise description of the person's height, e.g., \textsl{The person's height is 187.96~cm} would be more useful than a vague one, e.g., \textsl{The person is tall}.
-Lipman uses this scenario to explain why standard game theory models of communication, \citep[e.g.][]{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs.
-The relevant conditions are, broadly speaking, that both interlocutors know all the relevant facts (e.g., both know the person's height precisely) and that the setting for the communication is co-operative.
-These conditions exclude deliberate deception, and rhetorical situations like political debate and advertising where the intention is to persuade the interlocutor to adopt some point of view, where the persuasion might not be intended to be in the addressee's best interest.
-
-Lipman argued that such the efficiency loss resulting from vague expressions would be unlikely to have arisen unless there are advantages as well as disadvantages associated with vague expressions.
-Lipman asked, essentially, what these advantages might be, and how they might find a place in a game-theoretical explanation.
-In this paper, we focus on the first part of Lipman's question.
-
-Several tentative answers to Lipman's question have been offered -- see \citet{van2009utility, van2010vagueness}.
-Prominent among these answers is the idea that vague expressions are somehow easier to process, by a speaker and/or a hearer, than expressions that are not \emph{vague}, i.e., that are \emph{crisp} \citep[e.g.,][]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
-For example, Lipman writes: \begin{quotation}For the listener, information which is too specific may require more effort to analyze \citep[][p.\ 11]{lipmanvague}  \end{quotation} We shall refer to this characterisation of the utility of vague language as the \emph{cost reduction} hypothesis.
-The idea of vagueness as cost reduction can take various shapes, but the basic hypothesis is that it is easier for people to think in terms of loosely defined categories, such as \emph{quite a few}, or \emph{many}, than in terms of crisply defined ones, such as \emph{thirteen}, or \emph{237}.
-This predicts that whenever a vague expression can perform the same communicative task as a crisp one, it is rational to choose the vague expression.
-The corollary for comprehension is that vague expressions should be understood more readily than precise expressions.
-
-Questions concerning optimal language use have many practical applications.
-Natural Language Generation\footnote{\textsc{nlg} systems take data or formulas as input, and transform them into natural language outputs \cite[][]{reiter2000building}.
-The process parallels language production in humans.}  (henceforth \textsc{nlg}) systems must make decisions between different formulations of the same information.
-For example, if a man's height is 6 foot 2 inches, this could be expressed as \emph{187.96 metres}, \emph{6 foot 2}, or \emph{tall}, among other ways, and the \textsc{nlg} system must decide between these.
-The problem is particularly relevant for \textsc{nlg} systems that take numbers as input, as many do.
-In the context of an \textsc{nlg} system faced with a practical decision of this kind, Lipman's question becomes \emph{Under what circumstances should vague terms be produced?}  Relevant applications include weather forecasting on the basis of numerical weather data such as temperature and wind speed \citep{goldberg1994using, turner2006generating}, and medical decision support on the basis of clinical measurement such as oxygen saturation, heart rhythm, etc.
-\citep{Hripcsak01032009, hunter2008summarising, portet2009automatic}.
-At present, such \textsc{nlg} systems are often forced to make decisions concerning the level of precision in the utterances that they generate (e.g., \emph{the temperature will be in the high twenties tomorrow}) on the basis of little more than intuition.
-A better understanding of the benefits of different precision levels for readers would allow these systems to become more useful.
-
-The cost reduction hypothesis is of direct relevance to psycholinguists interested in language comprehension, and additionally to psycholinguists interested in language production, for example in connection with the question of audience design \citep{Clark1982287}.
-For, to the extent that speakers and writers choose vague expressions over and above crisp ones because the former are easier to process for hearers than the latter, the cost reduction hypothesis suggests that speakers design their utterance for optimal benefit to their hearers -- out of altruism, so to speak.
-
-The utility of vagueness is the attested aim of a small number of studies, but most of these have focussed on vagueness in a different sense, and focussing on different types of benefits for hearers.
-Two recent studies can illustrate both issues.
-
-\paragraph{Behaviour modification experiments} 
-In a series of studies of behaviour modification, \citet[]{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss.
-In the weight loss study, participants were told that the study was designed to test the validity of a new (actually fictitious) health index, the HHI (Holistic Health Index).
-They were told that an ideal HHI score lies in the range of 45 to 55.
-In a longitudinal study, participants submitted their height, weight, hydration level, gender, and age to a computer each week.
-Participants were told that two algorithms would be used to compute their HHI, and that it was possible that the two algorithms might give different values initially, but would converge over the course of the study to a single value.
-They were also told that if the two algorithms did give different values, then the true score lay between the two values.
-In one condition, which the authors called the precise condition, the two algorithms gave the same score.
-In the other condition, which the authors called the vague condition, one algorithm added 3\% to the score while the other algorithm subtracted 3\% from the score, yielding a range of values whose midpoint was the same as the two values given in the precise condition.
-
-\paragraph{Ratings experiments} 
-Similar issues arise from the work of \citet{peters2009bringing}.
-The authors carried out a series of studies where participants were required to rate hospitals based on various sources of information about quality of care.
-There was a between-subjects manipulation based on numeracy.
-The format of the information was manipulated within subjects: either numbers only were presented, or both numbers and evaluative categories were presented (e.g., \emph{Poor}, \emph{Fair}, \emph{Good}, \emph{Excellent}, with crisp visual boundary lines between the categories).
-Results showed that, for low-numeracy participants, the presence of evaluative categories resulted in a diminished influence of an irrelevant affective state on the ratings.
-For all participants, the presence of evaluative categories resulted in better decisions and in a greater use of the most important and reliable types of information, such as survival rates.
-
-It is, however, questionable whether the \emph{evaluative categories} manipulation in this study can be considered a manipulation of vagueness.
-Certainly, terms like \emph{Fair} admit the possibility of borderline cases.
-However, given that the boundaries between the categories were marked crisply, and that therefore the categories mapped crisply to numerical values, it becomes doubtful whether any borderline cases could be conceived to arise in fact.
-For example, \emph{Fair} was mapped to 60\% -- 70\% for the variable \emph{percentage of heart attack patients given recommended treatment (ACE inhibitor)}.
-Accordingly, rather than the vagueness of categories such as \emph{Poor}, Peters et al.
-emphasise the evaluative content inherent in these categories, and the affective potential of the evaluative content rather than the vagueness of the terms like \emph{Fair}.
-
-\paragraph{Our own experiments}
-The experiments reported in the present paper put the cost reduction hypothesis to the test.
-The question that we are trying to answer is whether vague expressions are processed more easily by readers than crisp ones.
-Like Lipman, we focus on situations where numerical information is used in order to identify a referent.
-Reference, in other words, will be the linguistic task on which we focus, partly because of the interest that this topic has recently drawn from the NLG community.
-In focussing on benefits for the hearer, we will leave aside the question of audience design, leaving this for later research.
-
-In using references to quantities to test the cost reduction hypothesis we are only testing one aspect of vagueness in a particular context.
-This limits the applicability of our results.
-It also has the advantage that we can explore the costs and benefits of vagueness more thoroughly in that context.
-Since one prevalent view of vagueness is that a vague expression is never preferable to a crisp equivalent, a demonstration of a benefit for vagueness in any context would advance the discussion.
-
-In our experiments we used a speeded forced choice task to compare the processing costs of different references to quantities.
-In this context, speed and accuracy of responses are the key dimensions on which the different references can be compared.
-The stimuli in the experiments were sets of dot arrays containing various numbers of dots.
-The forced choice was to identify one dot array given a reference to a given quantity of dots.
-We manipulated the references in several ways across a series of four experiments.
-
-Our main manipulation was always of vagueness: we constructed crisp and vague versions of references to the same dot array.
-For example, in experiment 1 the instruction presented to the participant could be \emph{Choose the square with many dots.} (vague condition), or \emph{Choose the square with 20 dots.} (crisp condition), identifying the same dot array.
-
-Ideally, we would manipulate vagueness independently of other variables.
-However, in practice, and given the constraints imposed by the use of natural language, any manipulation of vagueness introduces variance along other dimensions as well.
-Therefore we carried out further experiments to try to address and control variance along other dimensions.
-
-One non-vagueness source of variance was the difference between numerical and verbal format in the references.
-For example, in the instructions for experiment 1 given above, \emph{many} is in verbal format (as well as vague) and \emph{20} is in numerical format (as well as crisp).
-Therefore any difference observed between responses to the \emph{many} instruction and the \emph{20} instruction could be due either to vagueness or to {\bf instruction format}.
-It is possible to create vague and crisp references in each of these instruction formats: for example in numerical instruction format we can have \emph{20} (crisp) and \emph{about 20} (vague), and in verbal instruction format we can have \emph{the most} (crisp) and \emph{many} (vague).
-Experiment 2 varied vagueness and instruction format factorially in a similar way to this, to try to tease apart these two sources of variance.
-
-Another potential non-vagueness source of variance between responses to the \emph{many} (verbal vague) references and the \emph{20} (numeric crisp) references lies in the method used to identify a referent (we call this the {\bf selection algorithm} source of variance).
-Identifying the array with \emph{many} dots might be done by a \emph{comparison} algorithm, identifying that one array is more numerous than the others without establishing the numerosity of any arrays, whereas identifying the array with \emph{20} dots, which might be done by a \emph{matching} algorithm would require an estimate of the numerosity of the arrays.
-It seems reasonable that comparison would be faster than matching because it does not require estimates of cardinality.
-Furthermore the problem persists when using references like \emph{the most} (verbal crisp) and \emph{about 20} (numeric vague), since they too vary not only along the vague/crisp dimension and the numeric/verbal dimension but also along the comparison/matching dimension.
-Experiments 3 and 4 addressed the \emph{selection algorithm} source of variance - experiment 3 crossed vague/crisp and comparison/matching factorially using only numerical references, and experiment 4 crossed vague/crisp and comparison/matching factorially using only verbal references.
-
-\section{Experiment 1}
+\section{Section Heading}
 \label{sec:2}
@@
 Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Further on please use the \LaTeX\ automatism for all your cross-references and citations.
-
-\subsection{Introduction}%\label{}
-We used a forced choice task to compare choices made in response to vague instructions against choices made in response to crisp instructions.
-The participant was presented with an instruction like ``Choose the square with many dots'' in the vague conditions, or ``Choose the square with 20 dots'' in the crisp conditions.
-Then two dot arrays were presented in the form of squares containing a number of dots.
-Fig \ref{stimuluse1} shows an example stimulus.
-The participant was required to identify the square that corresponded with the instruction, by pressing the appropriate key.
-Response time and accuracy were recorded for analysis.
-
-Our main manipulation was of the vagueness of the instruction, with two levels, vague and crisp.
-Table \ref{instructionse1} shows examples from each condition.
-We also manipulated how discriminable the dot arrays were.
-One array always contained 25 dots: the other contained either 5, 10, 15, 20, 30, 35, 40, or 45 dots.
-This led to numerical differences of 5, 10, 15, and 20, with lower differences resulting in less discriminable arrays and larger differences representing more discriminable arrays.
-There is evidence that when the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the numerosity of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene199647}.
-Where a number was mentioned in the instruction, it was always in the form of an Arabic numeral.
-
-The instructions indicated the larger of the two dot arrays equally often as they indicated the smaller of the dot arrays, so that participants could not systematically choose the larger or smaller array as a successful response strategy.
-There is evidence that when two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}.
-We controlled which side the smaller number appeared on to avoid systematic influences of this effect.
-
-\begin{table}
-\caption{Table of instructions for the pair (5,25). Experiment 1}
-\label{instructionse1}       % Give a unique label
-\begin{tabular}{| l | l |}
-\hline\noalign{\smallskip}
-vagueness&example\\
-\noalign{\smallskip}\svhline\noalign{\smallskip}
-crisp 	& 	Choose the square with 5 dots \\
-vague	&	Choose the square with few dots\\
-\noalign{\smallskip}\hline\noalign{\smallskip}
-\end{tabular}
-$^a$ Table foot note (with superscript)\\
-\end{table}
-
-\section{Experiment 2}
-\label{sec:3}
-% Always give a unique label
-% and use \ref{<label>} for cross-references
-% and \cite{<label>} for bibliographic references
-% use \sectionmark{}
-% to alter or adjust the section heading in the running head
-Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Further on please use the \LaTeX\ automatism for all your cross-references and citations.
-
-\section{Experiment 3}
-\label{sec:4}
-% Always give a unique label
-% and use \ref{<label>} for cross-references
-% and \cite{<label>} for bibliographic references
-% use \sectionmark{}
-% to alter or adjust the section heading in the running head
-Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Further on please use the \LaTeX\ automatism for all your cross-references and citations.
-
-\section{Experiment 4}
-\label{sec:4}
-% Always give a unique label
-% and use \ref{<label>} for cross-references
-% and \cite{<label>} for bibliographic references
-% use \sectionmark{}
-% to alter or adjust the section heading in the running head
-Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Further on please use the \LaTeX\ automatism for all your cross-references and citations.
-
-\section{Dummy Template Section}
-\label{sec:5}
-% Always give a unique label
-% and use \ref{<label>} for cross-references
-% and \cite{<label>} for bibliographic references
-% use \sectionmark{}
-% to alter or adjust the section heading in the running head
-Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Further on please use the \LaTeX\ automatism for all your cross-references and citations.
+Furtheron please use the \LaTeX\ automatism for all your cross-references and citations.
 
@@
 %
-%\section{Section Heading}
-%\label{sec:3}
+\section{Section Heading}
+\label{sec:3}
 % Always give a unique label
@@
 % Problems or Exercises should be sorted chapterwise
-%\section*{Problems}
-%\addcontentsline{toc}{section}{Problems}
+\section*{Problems}
+\addcontentsline{toc}{section}{Problems}
 %
@@
 % Don't forget to label each problem;
-%% the label is needed for the solutions' environment
-%\begin{prob}
-%\label{prob1}
-%A given problem or Excercise is described here. The
-%problem is described here. The problem is described here.
-%\end{prob}
-%
-%\begin{prob}
-%\label{prob2}
-%\textbf{Problem Heading}\\
-%(a) The first part of the problem is described here.\\
-%(b) The second part of the problem is described here.
-%\end{prob}
+% the label is needed for the solutions' environment
+\begin{prob}
+\label{prob1}
+A given problem or Excercise is described here. The
+problem is described here. The problem is described here.
+\end{prob}
+
+\begin{prob}
+\label{prob2}
+\textbf{Problem Heading}\\
+(a) The first part of the problem is described here.\\
+(b) The second part of the problem is described here.
+\end{prob}
 
 %\input{referenc}
-% BibTeX users please use
-\bibliographystyle{spbasic} %plain, harvard, apalike, chicago, astron, authordate, natbib.
-\bibliography{bibliography}

### 2016-08-03 papers/springer-chapter/MACROS.tex
images

--- 2016-08-03 papers/springer-chapter/chapter_template.tex
@@
-%%%%%%%%%%%%%%%%%%%%% chapter.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% sample chapter
-%
-% Use this file as a template for your own input.
-%
-%%%%%%%%%%%%%%%%%%%%%%%% Springer-Verlag %%%%%%%%%%%%%%%%%%%%%%%%%%
-%\motto{Use the template \emph{chapter.tex} to style the various elements of your chapter content.}
-\chapter{Chapter Heading}
-\label{intro} % Always give a unique label
-% use \chaptermark{}
-% to alter or adjust the chapter heading in the running head
-
-\abstract*{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
-The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
-This allows unregistered users to read the abstract as a teaser for the complete chapter.
-As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.
-Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
-source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
-Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
-
-\abstract{Each chapter should be preceded by an abstract (10--15 lines long) that summarizes the content.
-The abstract will appear \textit{online} at \url{www.SpringerLink.com} and be available with unrestricted access.
-This allows unregistered users to read the abstract as a teaser for the complete chapter.
-As a general rule the abstracts will not appear in the printed version of your book unless it is the style of your particular book or that of the series to which your book belongs.\newline\indent
-Please use the 'starred' version of the new Springer \texttt{abstract} command for typesetting the text of the online abstracts (cf.
-source file of this chapter template \texttt{abstract}) and include them with the source files of your manuscript.
-Use the plain \texttt{abstract} command if the abstract is also to appear in the printed version of the book.}
-
-\section{Section Heading}
-\label{sec:1}
-Use the template \emph{chapter.tex} together with the Springer document class SVMono (monograph-type books) or SVMult (edited books) to style the various elements of your chapter content in the Springer layout.
-
-\section{Section Heading}
-\label{sec:2}
-% Always give a unique label
-% and use \ref{<label>} for cross-references
-% and \cite{<label>} for bibliographic references
-% use \sectionmark{}
-% to alter or adjust the section heading in the running head
-Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Furtheron please use the \LaTeX\ automatism for all your cross-references and citations.
-
-Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.
-
-Use the standard \verb|equation| environment to typeset your equations, e.g.
-%
-\begin{equation}
-a \times b = c\;,
-\end{equation}
-%
-however, for multiline equations we recommend to use the \verb|eqnarray|
-environment\footnote{In physics texts please activate the class option \texttt{vecphys} to depict your vectors in \textbf{\itshape boldface-italic} type - as is customary for a wide range of physical subjects.}.
-\begin{eqnarray}
-a \times b = c \nonumber\\
-\vec{a} \cdot \vec{b}=\vec{c}
-\label{eq:01}
-\end{eqnarray}
-
-\subsection{Subsection Heading}
-\label{subsec:2}
-Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Furtheron please use the \LaTeX\ automatism for all your cross-references\index{cross-references} and citations\index{citations} as has already been described in Sect.~\ref{sec:2}.
-
-\begin{quotation}
-Please do not use quotation marks when quoting texts! Simply use the \verb|quotation| environment -- it will automatically render Springer's preferred layout.
-\end{quotation}
-
-\subsubsection{Subsubsection Heading}
-Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Furtheron please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{subsec:2}, see also Fig.~\ref{fig:1}\footnote{If you copy text passages, figures, or tables from other works, you must obtain \textit{permission} from the copyright holder (usually the original publisher).
-Please enclose the signed permission with the manucript.
-The sources\index{permission to print} must be acknowledged either in the captions, as footnotes or in a separate section of the book.}
-
-Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.
-
-% For figures use
-%
 \begin{figure}[b]
 \sidecaption
-% Use the relevant command for your figure-insertion program
-% to insert the figure file.
-% For example, with the option graphics use
-\includegraphics[scale=.65]{figure}
-%
-% If not, use
-%\picplace{5cm}{2cm} % Give the correct figure height and width in cm
-%
-\caption{If the width of the figure is less than 7.8 cm use the \texttt{sidecapion} command to flush the caption on the left side of the page.
+\includegraphics[scale=.5]{images/resultse1}
+\caption{Experiment 1 resultsIf the width of the figure is less than 7.8 cm use the \texttt{sidecapion} command to flush the caption on the left side of the page.
 If the figure is positioned at the top of the page, align the sidecaption with the top of the figure -- to achieve this you simply need to use the optional argument \texttt{[t]} with the \texttt{sidecaption} command}
-\label{fig:1}       % Give a unique label
+\label{resultse1}
 \end{figure}
 
-\paragraph{Paragraph Heading} %
-Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Furtheron please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.
-
-Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.
-
-For typesetting numbered lists we recommend to use the \verb|enumerate| environment -- it will automatically render Springer's preferred layout.
-
-\begin{enumerate}
-\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
-\begin{enumerate}
-\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
-\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
-\end{enumerate}
-\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
-\end{enumerate}
-
-\subparagraph{Subparagraph Heading} In order to avoid simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}, see also Fig.~\ref{fig:2}.
-
-Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.
-
-For unnumbered list we recommend to use the \verb|itemize| environment -- it will automatically render Springer's preferred layout.
-
-\begin{itemize}
-\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development, cf. Table~\ref{tab:1}.}
-\begin{itemize}
-\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
-\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
-\end{itemize}
-\item{Livelihood and survival mobility are oftentimes coutcomes of uneven socioeconomic development.}
-\end{itemize}
-
-\begin{figure}[t]
-\sidecaption[t]
-% Use the relevant command for your figure-insertion program
-% to insert the figure file.
-% For example, with the option graphics use
-\includegraphics[scale=.65]{figure}
-%
-% If not, use
-%\picplace{5cm}{2cm} % Give the correct figure height and width in cm
-%
-\caption{Please write your figure caption here}
-\label{fig:2}       % Give a unique label
-\end{figure}
-
-\runinhead{Run-in Heading Boldface Version} Use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.
-
-\subruninhead{Run-in Heading Italic Version} Use the \LaTeX\ automatism for all your cross-refer\-ences and citations as has already been described in Sect.~\ref{sec:2}\index{paragraph}.
-% Use the \index{} command to code your index words
-%
-% For tables use
-%
 \begin{table}
-\caption{Please write your table caption here}
-\label{tab:1}       % Give a unique label
-%
-% For LaTeX tables use
-%
-\begin{tabular}{p{2cm}p{2.4cm}p{2cm}p{4.9cm}}
+\caption{Table of instructions for the pair (5,25). Experiment 1}
+\label{instructionse1} 
+\begin{tabular}{| l | l |}
 \hline\noalign{\smallskip}
-Classes & Subclass & Length & Action Mechanism  \\
+vagueness&example\\
 \noalign{\smallskip}\svhline\noalign{\smallskip}
-Translation & mRNA$^a$  & 22 (19--25) & Translation repression, mRNA cleavage\\
-Translation & mRNA cleavage & 21 & mRNA cleavage\\
-Translation & mRNA  & 21--22 & mRNA cleavage\\
-Translation & mRNA  & 24--26 & Histone and DNA Modification\\
+crisp 	& 	Choose the square with 5 dots \\
+vague	&	Choose the square with few dots\\
 \noalign{\smallskip}\hline\noalign{\smallskip}
 \end{tabular}
-$^a$ Table foot note (with superscript)
+$^a$ Table foot note (with superscript)\\
 \end{table}
-%
-\section{Section Heading}
-\label{sec:3}
-% Always give a unique label
-% and use \ref{<label>} for cross-references
-% and \cite{<label>} for bibliographic references
-% use \sectionmark{}
-% to alter or adjust the section heading in the running head
-Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Furtheron please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.
 
-Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.
+\begin{table}
+\caption{}
+\label{} 
+\begin{tabular}{}
+\hline\noalign{\smallskip}
 
-If you want to list definitions or the like we recommend to use the Springer-enhanced \verb|description| environment -- it will automatically render Springer's preferred layout.
+\noalign{\smallskip}\svhline\noalign{\smallskip}
 
-\begin{description}[Type 1]
-\item[Type 1]{That addresses central themes pertainng to migration, health, and disease.
-In Sect.~\ref{sec:1}, Wilson discusses the role of human migration in infectious disease distributions and patterns.}
-\item[Type 2]{That addresses central themes pertainng to migration, health, and disease.
-In Sect.~\ref{subsec:2}, Wilson discusses the role of human migration in infectious disease distributions and patterns.}
-\end{description}
-
-\subsection{Subsection Heading} %
-In order to avoid simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Use the \LaTeX\ automatism for all your cross-references and citations citations as has already been described in Sect.~\ref{sec:2}.
-
-Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.
-
-\begin{svgraybox}
-If you want to emphasize complete paragraphs of texts we recommend to use the newly defined Springer class option \verb|graybox| and the newly defined environment \verb|svgraybox|.
-This will produce a 15 percent screened box 'behind' your text.
-
-If you want to emphasize complete paragraphs of texts we recommend to use the newly defined Springer class option and environment \verb|svgraybox|.
-This will produce a 15 percent screened box 'behind' your text.
-\end{svgraybox}
-
-\subsubsection{Subsubsection Heading}
-Instead of simply listing headings of different levels we recommend to let every heading be followed by at least a short passage of text.
-Furtheron please use the \LaTeX\ automatism for all your cross-references and citations as has already been described in Sect.~\ref{sec:2}.
-
-Please note that the first line of text that follows a heading is not indented, whereas the first lines of all subsequent paragraphs are.
+\noalign{\smallskip}\hline\noalign{\smallskip}
+\end{tabular}
+$^a$ Table foot note (with superscript)\\
+\end{table}
 
@@
 %
-\begin{acknowledgement}
-If you want to include acknowledgments of assistance and the like at the end of an individual chapter please use the \verb|acknowledgement| environment -- it will automatically render Springer's preferred layout.
-\end{acknowledgement}
-%
-\section*{Appendix}
-\addcontentsline{toc}{section}{Appendix}
-%
-When placed at the end of a chapter or contribution (as opposed to at the end of the book), the numbering of tables, figures, and equations in the appendix section continues on from that in the main text.
-Hence please \textit{do not} use the \verb|appendix| command when writing an appendix at the end of your chapter or contribution.
-If there is only one the appendix is designated ``Appendix'', or ``Appendix 1'', or ``Appendix 2'', etc.
-if there is more than one.
 
-\begin{equation}
-a \times b = c
-\end{equation}
-% Problems or Exercises should be sorted chapterwise
+ Problems or Exercises should be sorted chapterwise
 \section*{Problems}
 \addcontentsline{toc}{section}{Problems}
-%
-% Use the following environment.
-% Don't forget to label each problem;
+
+ Use the following environment.
+ Don't forget to label each problem;
 % the label is needed for the solutions' environment
@@
 
-%\input{referenc}
+\begin{equation}
+a \times b = c
+\end{equation}

### 2016-08-07 experiments/e2/writeup/e2.Rmd
small edit
compiled
tidied up e2 considerably, switched to pdf output

@@
 output: 
-  html_document:
+  pdf_document:
     toc: true
-    <!--toc_float: true-->
-    mathjax: default
-    template: default
-    code_folding: hide
-    theme: default
 ---
 
-```{r loadKnitr}
+```{r loadKnitr, echo=FALSE}
 suppressPackageStartupMessages(library(knitr))
@@
 
-```{r loadLibs}
-suppressPackageStartupMessages(library(MASS))
-suppressPackageStartupMessages(library(xtable))
-suppressPackageStartupMessages(library(lattice))
-suppressPackageStartupMessages(library(latticeExtra))
-suppressPackageStartupMessages(library(car))
-suppressPackageStartupMessages(library(beepr))
-suppressPackageStartupMessages(library(lme4))
-suppressPackageStartupMessages(library(lmerTest))
-suppressPackageStartupMessages(library(LMERConvenienceFunctions)) # for perSubjectTrim
-suppressPackageStartupMessages(library(languageR))
-suppressPackageStartupMessages(library(reshape2))
-suppressPackageStartupMessages(library(ggplot2))
-suppressPackageStartupMessages(library(grid))
-suppressPackageStartupMessages(library(gridExtra))
-suppressPackageStartupMessages(library(plyr))
-suppressPackageStartupMessages(library(rms))                      # for varclus
-suppressPackageStartupMessages(library(languageR))
-```
-
-```{r CIfunctions}
-## Summarizes data.
-## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
-##   data: a data frame.
-##   measurevar: the name of a column that contains the variable to be summariezed
-##   groupvars: a vector containing names of columns that contain grouping variables
-##   na.rm: a boolean that indicates whether to ignore NA's
-##   conf.interval: the percent range of the confidence interval (default is 95%)
-summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
-                      conf.interval=.95, .drop=TRUE) {
-  require(plyr)
-  # New version of length which can handle NA's: if na.rm==T, don't count them
-  length2 <- function (x, na.rm=FALSE) {
-    if (na.rm) sum(!is.na(x))
-    else       length(x)
-  }
-  # This is does the summary; it's not easy to understand...
-  datac <- ddply(data, groupvars, .drop=.drop,
-                 .fun= function(xx, col, na.rm) {
-                   c( N    = length2(xx[,col], na.rm=na.rm),
-                      mean = mean   (xx[,col], na.rm=na.rm),
-                      sd   = sd     (xx[,col], na.rm=na.rm)
-                   )
-                 },
-                 measurevar,
-                 na.rm
-  )
-  # Rename the "mean" column    
-  datac <- rename(datac, c("mean"=measurevar))
-  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
-  # Confidence interval multiplier for standard error
-  # Calculate t-statistic for confidence interval: 
-  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
-  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
-  datac$ci <- datac$se * ciMult
-  return(datac)
-}
-## Summarizes data, handling within-subjects variables by removing inter-subject variability.
-## It will still work if there are no within-S variables.
-## Gives count, un-normed mean, normed mean (with same between-group mean),
-##   standard deviation, standard error of the mean, and confidence interval.
-## If there are within-subject variables, calculate adjusted values using method from Morey (2008).
-##   data: a data frame.
-##   measurevar: the name of a column that contains the variable to be summariezed
-##   betweenvars: a vector containing names of columns that are between-subjects variables
-##   withinvars: a vector containing names of columns that are within-subjects variables
-##   idvar: the name of a column that identifies each subject (or matched subjects)
-##   na.rm: a boolean that indicates whether to ignore NA's
-##   conf.interval: the percent range of the confidence interval (default is 95%)
-summarySEwithin <- function(data=NULL, measurevar, betweenvars=NULL, withinvars=NULL,
-                            idvar=NULL, na.rm=FALSE, conf.interval=.95, .drop=TRUE) {
-  # Ensure that the betweenvars and withinvars are factors
-  factorvars <- vapply(data[, c(betweenvars, withinvars), drop=FALSE],
-                       FUN=is.factor, FUN.VALUE=logical(1))
-  if (!all(factorvars)) {
-    nonfactorvars <- names(factorvars)[!factorvars]
-    message("Automatically converting the following non-factors to factors: ",
-            paste(nonfactorvars, collapse = ", "))
-    data[nonfactorvars] <- lapply(data[nonfactorvars], factor)
-  }
-  # Get the means from the un-normed data
-  datac <- summarySE(data, measurevar, groupvars=c(betweenvars, withinvars),
-                     na.rm=na.rm, conf.interval=conf.interval, .drop=.drop)
-  # Drop all the unused columns (these will be calculated with normed data)
-  datac$sd <- NULL
-  datac$se <- NULL
-  datac$ci <- NULL
-  # Norm each subject's data
-  ndata <- normDataWithin(data, idvar, measurevar, betweenvars, na.rm, .drop=.drop)
-  # This is the name of the new column
-  measurevar_n <- paste(measurevar, "_norm", sep="")
-  # Collapse the normed data - now we can treat between and within vars the same
-  ndatac <- summarySE(ndata, measurevar_n, groupvars=c(betweenvars, withinvars),
-                      na.rm=na.rm, conf.interval=conf.interval, .drop=.drop)
-  # Apply correction from Morey (2008) to the standard error and confidence interval
-  #  Get the product of the number of conditions of within-S variables
-  nWithinGroups    <- prod(vapply(ndatac[,withinvars, drop=FALSE], FUN=nlevels,
-                                  FUN.VALUE=numeric(1)))
-  correctionFactor <- sqrt( nWithinGroups / (nWithinGroups-1) )
-  # Apply the correction factor
-  ndatac$sd <- ndatac$sd * correctionFactor
-  ndatac$se <- ndatac$se * correctionFactor
-  ndatac$ci <- ndatac$ci * correctionFactor
-  # Combine the un-normed means with the normed results
-  merge(datac, ndatac)
-}
-## Norms the data within specified groups in a data frame; it normalizes each
-## subject (identified by idvar) so that they have the same mean, within each group
-## specified by betweenvars.
-##   data: a data frame.
-##   idvar: the name of a column that identifies each subject (or matched subjects)
-##   measurevar: the name of a column that contains the variable to be summariezed
-##   betweenvars: a vector containing names of columns that are between-subjects variables
-##   na.rm: a boolean that indicates whether to ignore NA's
-normDataWithin <- function(data=NULL, idvar, measurevar, betweenvars=NULL,
-                           na.rm=FALSE, .drop=TRUE) {
-  require(plyr)
-  # Measure var on left, idvar + between vars on right of formula.
-  data.subjMean <- ddply(data, c(idvar, betweenvars), .drop=.drop,
-                         .fun = function(xx, col, na.rm) {
-                           c(subjMean = mean(xx[,col], na.rm=na.rm))
-                         },
-                         measurevar,
-                         na.rm
-  )
-  # Put the subject means with original data
-  data <- merge(data, data.subjMean)
-  # Get the normalized data in a new column
-  measureNormedVar <- paste(measurevar, "_norm", sep="")
-  data[,measureNormedVar] <- data[,measurevar] - data[,"subjMean"] +
-    mean(data[,measurevar], na.rm=na.rm)
-  # Remove this subject mean column
-  data$subjMean <- NULL
-  return(data)
-}
-```
-
-```{r print_info_on_design}
-cat(paste('number of subjects = ',30,'\n',
-          '---','\n',
-          'number of trials per subject = ',256,'\n', 
-          '---','\n',          
-          'number of unique cells in the design = Item(4) * Number(2) * Vagueness(2) * Order(2) * Quantity(2) = ',4*2*2*2*2,'\n',
-          'number of times a subject saw a unique cell in the design = 256 / 64 = ', 256/64,'\n',
-          '---','\n',
-          'number of higher-level cells in the design (abstracting over Order and Quantity) = Item(4) * Number(2) * Vagueness(2) = ',4*2*2,'\n',
-          'number of times a subject saw a higher level cell in the design =  = 256 / 16 = ', 4*2*2, '\n',
-          sep=''))
-```
-
-# Distributional stuff, including transformations.
-
-```{r transformations_table, results='asis'}
-x=445
-lambda = c(-2,-1,-0.5, 0,0.5,1,2)
-
-xd=data.frame(
-  lambda = c(
-    lambda[1],
-    lambda[2],
-    lambda[3],
-    lambda[4],
-    lambda[5],
-    lambda[6],
-    lambda[7]
-  ),
-  x=c(    
-    paste("x = ", sep="", x),
-    paste("x = ", sep="", x),
-    paste("x = ", sep="", x),
-    paste("x = ", sep="", x),
-    paste("x = ", sep="", x),
-    paste("x = ", sep="", x),
-    paste("x = ", sep="", x)
-     ),
-  conventional=c(
-    '$$y=\\frac{1}{x^2}$$', 
-    '$$y=\\frac{1}{x}$$', 
-    '$$y=\\frac{1}{\\sqrt{x}}$$', 
-    '$$y=log_e(x)$$', 
-    '$$y=\\sqrt{x}$$', 
-    '$$y=x$$', 
-    '$$y=x^2$$'
-    ),
-  power=c(
-    '$$y=x^{-2}$$', 
-    '$$y=x^{-1}$$', 
-    '$$y=x^{-0.5}$$', 
-    '$$y=log_e(x)$$', 
-    '$$y=x^{0.5}$$', 
-    '$$y=x^{1}$$', 
-    '$$y=x^{2}$$'
-    ),
-  y=c(     
-    paste("y = ", sep="", format(round(x^lambda[1],6), sc=F) ),
-    paste("y = ", sep="", round(x^lambda[2],3) ),
-    paste("y = ", sep="", round(x^lambda[3],3) ),
-    paste("y = ", sep="", round(log(x),3) ),
-    paste("y = ", sep="", round(x^lambda[5],2) ),
-    paste("y = ", sep="", round(x^lambda[6],6) ),
-    paste("y = ", sep="", round(x^lambda[7],6) )
-  ),
-  back = c(
-    '$$x=y^{\\frac{1}{-2}}$$',
-    '$$x=y^{\\frac{1}{-1}}$$', 
-    '$$x=y^{\\frac{1}{-0.5}}$$', 
-    '$$x=exp(y)$$',  
-    '$$x=y^{\\frac{1}{0.5}}$$', 
-    '$$x=y^{\\frac{1}{1}}$$',  
-    '$$x=y^{\\frac{1}{2}}$$'
-  ),
-  back_lambda=c(
-    '$$x=y^{1/\\lambda}$$',
-    '$$x=y^{1/\\lambda}$$',
-    '$$x=y^{1/\\lambda}$$',
-    '$$x=exp(y)$$',  
-    '$$x=y^{1/\\lambda}$$',
-    '$$x=y^{1/\\lambda}$$',
-    '$$x=y^{1/\\lambda}$$'
-  ),
-  y_back=c(  
-    paste( 'x = ', (x^lambda[1]) ^ {1/lambda[1]} ,sep=""),
-    paste( 'x = ', (x^lambda[2]) ^ {1/lambda[2]} ,sep=""),
-    paste( 'x = ', (x^lambda[3]) ^ {1/lambda[3]} ,sep=""),
-    paste( 'x = ', exp(log(x)), sep=""),
-    paste( 'x = ', (x^lambda[5]) ^ {1/lambda[5]} ,sep=""),
-    paste( 'x = ', (x^lambda[6]) ^ {1/lambda[6]} ,sep=""),
-    paste( 'x = ', (x^lambda[7]) ^ {1/lambda[7]} ,sep="")
-  )
-)
-
-print(xtable(xd, digits=1, align=c(rep('l', times=ncol(xd)+1))), type='html', include.rownames=F, digits=1)
-```
-
-Load the file dat.Rda if it exists, else Get the data if file dat.Rda does not already exist.
-
-```{r, get_the_data}
-# if the output file dat.Rda already exists then load it, else do data wrangling
-if( file.exists('dat.Rda') && file.exists('dd.Rda') ) {
-	load('dat.Rda'); load('dd.Rda')
+```{r loadLibs, echo=FALSE, message=FALSE}
+library(beepr)                     #
+library(car)                       #
+library(ggplot2)                   #
+library(grid)                      #
+library(gridExtra)                 #
+library(languageR)                 #
+library(lattice)                   #
+library(latticeExtra)              # for doubleYScale
+library(lme4)                      #
+library(lmerTest)                  #
+library(LMERConvenienceFunctions)  # for perSubjectTrim
+library(MASS)                      #
+library(plyr)                      #
+library(reshape2)                  #
+library(rms)                       # for varclus
+library(Rmisc)                     # for summarySEwithin and friends
+library(xtable)                    #
+```
+
+```{r get_the_data, echo=FALSE}
+# if the file data.Rda already exists then load it, else do data wrangling
+if( file.exists('data_processed.Rda') ) {
+	load('data_processed.Rda')
 	} else {
@@
 		column.headers.df <-  head( read.table(
-			paste(data_dir,'subject1.data',sep=''),
+			paste(data_dir,'subject01.data',sep=''),
 			header=TRUE),0)
@@
 		for (subject in 1:number_of_valid_subjects) {
-			current.filename <- paste(data_dir, 'subject', subject, '.data', sep='')
+			current.filename <- paste(data_dir, 'subject', sprintf("%02d",subject), '.data', sep='')
 			current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = FALSE)
@@
 								))
-	# save dat as dat.Rda
-	# This data set contains *all* trials 7680 including impossible trials and is
+	# This data set (dat) contains *all* trials 7680 including impossible trials and is
 	# mainly for graphs comparing different removals
-	save(dat, file="dat.Rda")
+	save(dat, file='data_raw.Rda')
 	# remove impossible trials from dat in new data dd #
@@
 							 ))
-	save(dd, file="dd.Rda")
-} # end of the else clause, which is run if both of the files dat.Rda and dd.Rda do not exist.
+	save(dd, file="data_processed.Rda")
+} # end of the else clause, which is run if the file data.Rda does not exist.
 ```
@@
 
-```{r plotDists, fig.width=20, fig.height=12, message=FALSE}
-## plot the comparison of the transformations against their normal equivalents to see distributional properties using full data
+```{r plotDists, fig.width=5, fig.height=5, message=FALSE, echo=FALSE}
 # first create a data frame specially for transformations
 # this won't be used for analysis, just for this plot
+load('data_raw.Rda') # dat
 dat_transforms <- dat
@@
 # create extra rows for dfs with removed (outlier) data points
-dat1 <- dat_transforms                                       # nrow 7680
-dat2 <- subset(dat_transforms, RT>1)                         # nrow 7678
-dat2$dfSubset <- "RT>1"
 dat3 <- subset(dat_transforms, RT>1 & RT<59999)              # nrow 7677
@@
 # bring the dfs back into main df
-dat_transforms <- rbind(dat1, dat2, dat3, dat4)
-dat_transforms$dfSubset <- factor(dat_transforms$dfSubset, levels=c("original", "RT>1", "RT>1&RT<59999", "RT>1&RT<30001"))
+dat_transforms <- rbind(dat3, dat4)
+dat_transforms$dfSubset <- factor(dat_transforms$dfSubset, levels=c("RT>1&RT<59999", "RT>1&RT<30001"))
 # create extra rows for normally distributed transformed variables
@@
 # for safety
-dat_transforms[dat_transforms$RTtype=="normal", c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd")] <- NA
+dat_transforms[dat_transforms$RTtype=="normal", c("RT_log", "RT_raw")] <- NA
 # Create the normally distributed transformed variables, replacing the NAs declared above
 for (k in levels(dat_transforms$dfSubset)){
-  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_RecSqd"]  <- 
-    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_RecSqd"]), 
-          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_RecSqd"]), 
-          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_RecSqd"]))
-  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_Rec"]  <- 
-    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_Rec"]), 
-          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_Rec"]), 
-          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_Rec"]))
-  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_RecSqt"]  <- 
-    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_RecSqt"]), 
-          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_RecSqt"]), 
-          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_RecSqt"]))
     dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_log"]  <- 
@@
           sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_log"]))
-  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_sqt"]  <- 
-    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_sqt"]), 
-          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_sqt"]), 
-          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_sqt"]))
     dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_raw"]  <- 
@@
           sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_raw"]))
-  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_sqd"]  <- 
-    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_sqd"]), 
-          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_sqd"]), 
-          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_sqd"]))
-} # end of create normally distributed equivalents for the data transforms.
+  }# end of create normally distributed equivalents for the data transforms.
 temp <- reshape2::melt(dat_transforms, 
              id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Number", "Quantity", "Order"),
-             measure.vars=c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd"),
+             measure.vars=c("RT_log", "RT_raw"),
              variable.name="transformation",
-             value.name="score")
-ggplot(temp, aes(score, colour=RTtype)) + 
-  geom_freqpoly() + 
-  facet_wrap(dfSubset~transformation, scales="free", nrow=4) +
-  theme_bw() + 
-	theme(aspect.ratio=1) + 
-	scale_color_manual(values=c("blue","red"))
+             value.name="time")
+ggplot(temp, aes(time, colour=RTtype)) + geom_freqpoly() + facet_wrap(transformation~dfSubset, scales='free') + scale_color_manual(values=c("blue","red")) + theme_bw() + theme(aspect.ratio=1)
 ```
@@
 
-```{r boxPlots, fig.width=20, fig.height=5}
+```{r boxPlots, fig.width=7, fig.height=3, echo=FALSE}
+load('data_processed.Rda')
 # allow reference to transformation
 subdata = reshape2::melt(dd,
-               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
+               measure.vars=c("RT_log", "RT_raw"),
                variable.name="transformation",
@@
   facet_wrap(~ transformation, scales="free", ncol=4) +
-  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
-  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
-  theme_bw() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
-  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
+  geom_boxplot(aes(y=value, x=Item, col="Items"), outlier.colour = "black", outlier.shape = 21, outlier.size=0.1) + 
+  geom_boxplot(aes(y=value, x=Subject, col="Subjects"), outlier.colour = "black", outlier.shape = 21, outlier.size=0.1) + 
+  theme_bw() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=5), aspect.ratio=1, panel.grid=element_blank()) + 
+  scale_colour_manual("",values=c("Items"="orange","Subjects"="blue")) +
   xlab("") + ylab("")
@@
 
-Show how mean times for individual subjects and items vary with respect to the grand mean RT.
-
-```{r fastSlowRT, fig.width=10, fig.height=5}
+Show how mean times for individual subjects and items vary with respect to the grand mean Log RT.
+
+```{r fastSlowRT, fig.width=7, fig.height=3, message=FALSE, echo=FALSE}
 # Show how mean times for individual subjects and items vary with respect to the grand mean RT.
 subs=ddply(dd, .(level=Subject), 
-					 summarise, effect="subject", emean=mean(dd$RT), lmean=mean(RT),
-					 ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0, "slow", "fast"))
+					 summarise, effect="subject", emean=mean(dd$RT_log), lmean=mean(RT_log),
+					 ldiff=mean(RT_log)-mean(dd$RT_log))
 itms=ddply(dd, .(level=Item), 
-					 summarise, effect="item", emean=mean(dd$RT), lmean=mean(RT), 
-					 ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0, "slow", "fast"))
+					 summarise, effect="item", emean=mean(dd$RT_log), lmean=mean(RT_log), 
+					 ldiff=mean(RT_log)-mean(dd$RT_log))
 si=rbind(subs,itms)
 si$effect <- factor(si$effect, levels=c("subject", "item"))
-si$lsign <- factor(si$lsign, levels=c("slow","fast"))
 my1 <- xyplot(data=si, lmean~level | effect, 
@@
               ylim = extendrange(si$lmean, f = 0.10),
-              par.settings=list(trellis.par.set(superpose.symbol = list(col = c("blue","red"), pch=19), 
-              																	superpose.line=list(col = c("blue","red")))),
-              groups=si$lsign,
+              par.settings=list(trellis.par.set(superpose.symbol = list(col='black'), 
+              																	superpose.line=list(col='black'))),
               xlab=NULL,
-              ylab="RT (in ms)",
-              key=list(text=list(levels(si$lsign)), 
-              				 lines=list(lty=c(1,1), col=trellis.par.get('superpose.line')$col, 
-              				 					 pch=c(19,19), type='o'), divide=1),
+              ylab=NULL,
+              main='How mean times for individual subjects and items\nvary with respect to the grand mean Log RT',
               mean=unique(si$emean),
-              scales=list(x=list(relation="free", at=list(15.5, 32.5), 
+              scales=list(x=list(relation="free", at=NULL, 
               									 labels=list("subject number", "item identifier"), cex=1 ),
-                          y=list(at=c(0, 2000, unique(si$emean), 4000, 6000, 8000,  10000), 
-                          			 labels=c(0, 2000,  "Mean", 4000,  6000, 8000, 10000), cex=.9) ),
-              panel=function(x,y,type,subscripts,groups=groups,...){
-                panel.xyplot(x,y,type='p',groups=groups, subscripts=subscripts,...)
-                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, groups=groups, subscripts=subscripts, 
-                						 code=3, length=0, col=trellis.par.get('superpose.line')$col[groups][subscripts], ...)
-                panel.abline(h=unique(si$emean), col="lightgrey",lty=1)
-                panel.text(x, y=(ifelse(y<unique(si$emean),y-300,y+300)), 
-                					 as.numeric(si$level)[subscripts][panel.number()==1], cex=.75)
-                panel.text(x, y=(ifelse(y<unique(si$emean),y-300,y+300)), 
-                					 as.character(si$level)[subscripts][panel.number()==2], cex=.75)
+                          y=list(at=    c(0,7,unique(si$emean),8,9),
+                          			 labels=c("0","7","Mean Log RT","8","9"))),
+              panel=function(x,y,type,subscripts,...){
+                panel.abline(h=unique(si$emean), col="lightgrey",lty=1,...)
+                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, subscripts=subscripts, code=3, length=0, ...)
+                panel.xyplot(x,y,type='p', pch=21, cex=.5, col='black', fill='black', subscripts=subscripts,...)
               }) # end of my1
@@
               aspect=1, type='n',
-              ylab="Deviation from mean RT (in ms)",
+              ylab="Deviation from\nmean Log RT",
               ylim=extendrange(si$ldiff,f=0.10)
@@
 
-Show how mean times for individual subjects and items vary with respect to the grand mean RT_RecSqt.
-
-```{r fastSlowRecSqt, fig.width=10, fig.height=5}
-# Show how mean times for individual subjects and items vary with respect to the grand mean RT_RecSqt.
-subs <- ddply(dd, .(level=Subject), summarise, 
-              effect="subject", emean=mean(dd$RT_RecSqt), lmean=mean(RT_RecSqt), 
-							ldiff=mean(RT_RecSqt)-mean(dd$RT_RecSqt),
-              lsign=ifelse(mean(RT_RecSqt)-mean(dd$RT_RecSqt)<0,"slow","fast"))
-itms <- ddply(dd, .(level=Item), summarise, effect="item", emean=mean(dd$RT_RecSqt), lmean=mean(RT_RecSqt),
-              ldiff=mean(RT_RecSqt)-mean(dd$RT_RecSqt), 
-							lsign=ifelse(mean(RT_RecSqt)-mean(dd$RT_RecSqt)<0,"slow","fast"))
-si <- rbind(subs,itms)
-si$effect <- factor(si$effect, levels=c("subject", "item"))
-si$lsign <- factor(si$lsign, levels=c("slow","fast"))
-my1 <- xyplot(data=si, lmean~level | effect, 
-              aspect=1,
-              layout=c(2,1),
-              ylim = extendrange(si$lmean, f = 0.10),
-              par.settings = list(trellis.par.set(superpose.symbol = list(col = c("blue","red"), pch=19), superpose.line=list(col = c("blue","red")))),
-              groups = si$lsign,
-              xlab = NULL,
-              ylab = "RT_RecSqt: reciprocal of square root of RT in ms\n",
-              key=list(text=list(levels(si$lsign)), 
-              				 lines=list(lty=c(1,1), 
-              				 					 col=trellis.par.get('superpose.line')$col,
-              				 					 pch=c(19,19), type='o'),
-              				 divide=1),
-              mean = unique(si$emean),
-              scales = list(x=list(relation="free", at=list(15.5, 32.5), 
-              										 labels=list("subject number", "item identifier"), cex=1 ) ),
-              panel=function(x,y,type,subscripts,groups=groups,...){
-                panel.xyplot(x,y,type='p',groups=groups, subscripts=subscripts,...)
-                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, 
-                						 groups=groups, subscripts=subscripts, code=3, length=0, col=trellis.par.get('superpose.line')$col[groups][subscripts], ...)
-                panel.abline(h=unique(si$emean), col="lightgrey",lty=1)
-                panel.text(x, y=(ifelse(y<unique(si$emean),y-.001,y+.001)), 
-                					 as.numeric(si$level)[subscripts][panel.number()==1], cex=.75)
-                panel.text(x, y=(ifelse(y<unique(si$emean),y-.001,y+.001)), 
-                					 as.character(si$level)[subscripts][panel.number()==2], cex=.75)
-              })
-my2 <- xyplot(data=si, ldiff~level | effect, 
-              aspect=1, type='n',
-              ylab="\nDeviation from mean RT_RecSqt (in ms)",
-              ylim=extendrange(si$ldiff,f=0.10))
-print(doubleYScale(my1, my2, add.axis=TRUE, add.ylab2=TRUE, use.style=FALSE))
-```
-
-Show how practice and fatigue effects vary over subjects
-
-```{r pracFatigueSubjs, eval=TRUE, fig.width=20, fig.height=10, cache=FALSE, echo=TRUE}
-# Show how practice/fatigue effects vary over subjects
-ggplot(dd, aes(y=RT_RecSqt, x=Trial)) + 
-	facet_wrap(~Subject,nrow=3) + 
-	theme(panel.grid=element_blank(), 
-				panel.margin = unit(0, "lines"), 
-				panel.background = element_rect(fill = 'white', colour='black' ), 
-				legend.position="top") + 
-	scale_x_continuous("Trial", breaks=NULL) + 
-	geom_point(pch=19, cex=.25, alpha=.5) + 
-	stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + 
-	theme(aspect.ratio = 1) 
-```
-
-Show practice and fatigue effects vary over items.
-
-```{r pracFatigueItms, eval=TRUE, fig.width=9, fig.height=4, cache=FALSE, echo=TRUE}
-# Show how practice/fatigue effects vary over items
-ggplot(dd, aes(y=RT_RecSqt, x=Trial)) + 
-	facet_grid(~Item) + 
-	theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), 
-				panel.background = element_rect(fill = 'white', colour='black' ), 
-				legend.position="top") + 
-	scale_x_continuous("Trial") + 
-	geom_point(pch=19, cex=.25, alpha=.5) + 
-	stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + 
-	theme(aspect.ratio = 1) 
-```
-
 # Plot main effects and interactions
@@
 
-```{r plotMainsa, fig.width=16, fig.height=6, warning=FALSE}
+```{r plotMainsa, fig.width=6, fig.height=6, warning=FALSE, echo=FALSE}
 # plot main effects on observed data
@@
 
-```{r mainB, fig.width=9, fig.height=3}
+```{r mainB, fig.width=6, fig.height=3, echo=FALSE}
 # prep main effects rts data frame
@@
 
-```{r mainD, message=FALSE, fig.width=10, fig.height=3}
+```{r mainD, message=FALSE, fig.width=6, fig.height=3, echo=FALSE}
 vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Vagueness")));names(vrts)[4]<-"Levels"
@@
 
-```{r 2way, fig.width=10, fig.height=3}
+```{r 2way, fig.width=6, fig.height=3, echo=FALSE}
 # prep 2 ways rts data frame
@@
 rts <- rbind(rts1,rts2,rts3,rts5,rts6,rts8)
-rts <- subset(rts, select=c(Effect, F1, F2, RT_RecSqt, RT_RecSqt_norm, sd, se, ci))
+rts <- subset(rts, select=c(Effect, F1, F2, RT_RecSqt, sd, se, ci))
 rts$mins=rts$RT_RecSqt-rts$ci
@@
 
-```{r 22way, message=FALSE}
+```{r 22way, message=FALSE, fig.width=6, fig.height=3, echo=FALSE}
 # prep 2 ways over items rts data frame
@@
 
-```{r plotInt1, fig.width=7, fig.height=3.5, eval=TRUE, cache=FALSE}
+```{r plotInt1, fig.width=7, fig.height=3.5, echo=FALSE}
 rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item"))
@@
 
-```{r plot3way, fig.width=10,fig.height=5}
+```{r plot3way, fig.width=7, fig.height=5, echo=FALSE}
 rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Quantity"))
@@
 
-```{r vnitems, fig.width=10, fig.height=4, eval=TRUE}
+```{r vnitems, fig.width=7, fig.height=4, echo=FALSE}
 rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item", "Quantity"))
@@
 
-```{r 4way}
+```{r 4way, fig.width=6, fig.height=3, echo=FALSE}
 rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness",  "Quantity", "Order"))
@@
 
-```{r 4waysplit, fig.width=8, fig.height=8, eval=TRUE, cache=FALSE, echo=TRUE}
+```{r 4waysplit, fig.width=7, fig.height=7, echo=FALSE}
 rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness", "Item", "Quantity", "Order"))
@@
 
-```{r 4waySplitRatio, fig.width=8, fig.height=8, eval=TRUE, cache=FALSE, echo=TRUE, message=FALSE, warning=FALSE}
+```{r 4waySplitRatio, fig.width=7, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}
 rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness", "item_mean_ratio", "Quantity", "Order"))
@@
 
-```{r showCompression, fig.width=10, fig.height=7, echo=F}
+```{r showCompression, fig.width=7, fig.height=7, echo=FALSE}
 #  Show compression of ratios of numbers of dots in the items
@@
 
-## ols model v4, RecSqt, in progress
-
-Pairs plot
-
-```{r pairs, fig.width=7, fig.height=7}
-pairs(dd[,c("RT_RecSqt", "Number", "Vagueness", "RTprev_RecSqt")], pch = 19, cex=.075)
-```
-
-Build the model
-
-```{r doModelols}
-## do ols model
-# define datadist
-dd.dd = datadist(dd)
-options(datadist = "dd.dd")
-# build model
-v4 <- ols(data=dd, 
-          RT_RecSqt ~ 
-            c_Vag + c_Num + c_Qty + c_Ord + 
-            c_Num:c_Vag:c_Qty+
-            item_mean_ratio + 
-            #c_Vag:c_Num + 
-            #c_Vag:c_Num:item_mean_ratio + 
-            # pol(s_Trl, 3) +
-            s_Trl+
-            # pol(RTprev_RecSqt,3) +
-            RTprev_RecSqt +
-            nchar_instr
-            # cell+
-            # measurement,
-            ,
-          x=T, y=T )
-```
-
-```{r likRat}
-# print out p value for likelihood ratio with df
-cat("lik.ratio p value, whether the model as a whole is explanatory: p =", 1-pchisq(v4$stats[2],v4$stats[3]))
-# print out R2 for the model
-cat("R^2 = ", v4$stats[4])
-```
-
-Show the model tables
-
-```{r modTabs}
-# show model
-v4
-# show model summary
-summary(v4)
-# do anova 
-an1 = anova(v4)
-# show anova
-an1
-```
-
-plot 'partial effects'
-
-```{r partial, fig.width=12, fig.height=5}
-## plot partial effects
-# Compute Predicted Values and Confidence Limits
-p1=Predict(v4)
-# plot 'partial effects'
-plot(p1, anova=an1, pval=T, aspect=1, main="ols model v4")
-```
-
-```{r daz99}
-# add predicted RT_RecSqt to dd
-dd$RT_Predicted <- predict (v4)
-```
-
-Plot model coefficients and ci's
-
-```{r coefsAndCiS}
-par(las=2, mar=c(14,6,1,1))
-y=coef(v4) # with intercept
-n=length(y)
-y0=confint(v4)[,1]
-y1=confint(v4)[,2]
-y=coef(v4)[-1] # omit intercept
-n=length(y)
-y0=confint(v4)[-1,1]
-y1=confint(v4)[-1,2]
-plot(y, xaxt="n", xlab="", ylab="RT_RecSqt\n", pch=19, ylim=extendrange(y, f=.10), main="Speed")
-abline(h=0)
-axis(1, labels=names(y) , at=1:n)
-grid()
-segments(x0=1:n, x1=1:n, y0, y1, lwd=2)
-```
-
-Collinearity is assessed by the condition number $k$.\
-The greater the collinearity, the closer the matrix of predictors is is to becoming _singular_.\
-$k$ estimates the extent to which a matrix is $\text{singular}$\
-$0\ldots6$ no collinearity\
-$\text{around} 15$ medium collinearity\
-$>30$ indicates potentially harmful collinearity\
-
-```{r kis}
-cat("k is:",
-collin.fnc(dd[,	c("s_Trl", "c_Num", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", 
-									"RTprev_RecSqt", "nchar_instr", "measurement")
-							])$cnumber
-)
-```
-
-```{r plotg, fig.width=6, fig.height=5}
-par(mar=c(1.1,3.1,1.1,1.1), pty='s')
-plot(varclus(
-	as.matrix(dd[,c("s_Trl", "c_Num", "c_Vag", "c_Qty", "c_Ord",
-									"item_mean_ratio", "RTprev_RecSqt", "nchar_instr", "measurement")])))
-```
-
-Model criticism baayen plots
-
-```{r baayenMplots, fig.width=9, fig.height=3}
-# Baayen 4-plot model criticism
-par(mfrow=c(1,4), pty='s')
-# create scaled residuals
-v4$rstand = as.vector(scale(resid(v4)))
-# plot scaled residuals density
-plot(density(v4$rstand), main='density of\nscaled residuals')
-# plot sample quantiles versus theoretical quantiles
-qqnorm(v4$rstand, cex=.5)
-qqline(v4$rstand)
-# plot standardised residuals versus fitted values
-plot(v4$rstand ~ fitted(v4), pch='.', main='std resid\nv fitted')
-# absolute standardised residuals greater than 2.5 are candidates 
-# for being outliers, the abline identifies them on the plot
-abline(h=c(-2.5,2.5))
-# create dffits
-dffits=abs(resid(v4, "dffits"))
-# plot dffits
-plot(dffits, type='h', main='dffits')
-# here no overly influential
-w = which.influence(fit = v4, cutoff = 0.4)
-if (length(w)==0) {
-	print('There are no overly influential data points')
-} else {
-	print('There are some overly influential variables:');
-	print(w)
-}
-```
-
-Model validation
-
-Validation tests overfitting using bootstrap.
-
-Bootstrapping draws as a bootstrap sample the same number of samples as the original data, at random with replacement, from the original data.
-
-Then fit the model to the data in the bootstrap sample, and use this model to predict the reaction times for the original full data (which contains many samples that are new to the bootstrap model).
-
-Then compare the goodness of fit of the bootstrap model with the goodness of fit of the original model.
-
-Averaged over a large number of bootstrap models these comparisons of goodness of fit  reveal to what extent the original model overfits the data.
-
-Function validate() in rms does this _re-sampling validation_.
-
-```{r val_1, echo=TRUE}
-# argument B is the number of bootstrap runs
-# argument pr is whether to print the results of each run
-v_1 <- validate(v4, bw = T, B = 200, pr=FALSE, estimates=TRUE)
-```
-
-```{r prval}
-# in print.validate, B is the number of re-samples to show outocmes for each resample
-print(v_1, B=0)
-```
-
-```{r kept}
-# show how often each number of variables kept was observed across the 200 runs
-xtabs(~rowSums(attr(v_1,"kept")))
-```
-
-# model lmer
+# Lmer model
 
 ```{r rmme}
-load("dd.Rda")
+load("data_processed.Rda")
 ```
@@
 v5 <- lmerTest::lmer(data=dd, 
-           RT_RecSqt ~ 
+           RT_log ~ 
              c_Vag + c_Num + c_Qty + c_Ord + 
@@
              s_Trl +
-             RTprev_RecSqt +
+             RTprev_log +
              nchar_instr +
-             (1+c_Vag + c_Num + c_Qty + c_Ord|Subject)
-)
+             (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
+v5b <- lme4::lmer(data=dd, 
+           RT_log ~ 
+             c_Vag + c_Num + c_Qty + c_Ord + 
+             c_Num:c_Vag:c_Qty +
+             item_mean_ratio + 
+             s_Trl +
+             RTprev_log +
+             nchar_instr +
+             (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
 ```
@@
 summary(v5)
+```
+
+```{r v511, results='asis'}
+print(xtable(coef(summary(v5b))))
 ```
@@
 
-```{r v54}
+```{r v54, fig.width=7, fig.height=4}
 e=data.frame(coef=summary(v5)$coef[-1,1]) # estimates, without intercept

### 2016-08-07 experiments/e2/experimentCode/exp2dummy.m

@@ -1,586 +1,0 @@
-function exp3dummy(SUBJECT)
-% This third study has the following structure:
-% Practice, Rest, Block1 Block 2 Block 3, Block 4
-% There are 256 trials
-% There are 64 unique trials
-% There are 4 Blocks
-% Each trial consists of 3 squares each with a number of dots.
-%
-% A trial consists of:
-% the instruction on screen while participants read it for as long as they like;
-% a button press on SPACEBAR to bring up
-% a screen with the instruction and the dots presented
-%
-% Participant id > 100 is used for testing and can be overwritten
-
-howLongToWaitForAResponse = 60; % seconds. Set low for test runs
-testing = 1; % 0 for real runs, 1 for testing
-takescreenshot=1;
-SUBJECT=1001;
-
-
-% Clear Matlab/Octave window:
-clc;
-
-% check for Opengl compatibility, abort otherwise:
-AssertOpenGL;
-
-% Reseed the random-number generator for each expt.
-rand('state',sum(100*clock));
-
-% Make sure keyboard mapping is the same on all supported operating systems
-% Apple MacOS/X, MS-Windows and GNU/Linux:
-KbName('UnifyKeyNames');
-
-% Define keys by ID
-leftresp=KbName('c'); % is 6
-middleresp=KbName('b'); % is 5
-rightresp=KbName('m'); % is 16
-escapekey=KbName('ESCAPE'); % is 41
-spacebar=KbName('space'); % is 44
-returnkey=KbName('return'); % is 40
-
-%%%%%%%%%%%%%%%%%%%%%%
-% define stimuli
-%%%%%%%%%%%%%%%%%%%%%%
-
-% order in headers gives the answer to "Which square had the small quantity"? and does not give the correct response
-
-%Condition 1 is numerical vague
-%Condition 2 is numerical precise
-%Condition 3 is visual vague
-%Condition 4 is visual precise
-%
-%Order 1 is small number on left
-%Order 2 is small number on right
-%
-%Quantity 1 is small
-%Quantity 2 is big
-
-
-exp3headers={
-	"Item","Condition","Order","Quantity","Left","Mid","Right","Instruction"};
-
-exp3practicestimuli = {
-	0,0,0,0,6,15,24,"Choose a square with about 10 dots";
-	0,0,0,0,34,25,16,"Choose the square with 16 dots";
[…]

### 2016-08-07 experiments/e2/experimentCode/exp2fakescreenshots.m

@@ -1,816 +1,0 @@
-function exp3fakescreenshots(SUBJECT)
-% This third study has the following structure:
-% Practice, Rest, Block1 Block 2 Block 3, Block 4
-% There are 256 trials
-% There are 64 unique trials
-% There are 4 Blocks
-% Each trial consists of 3 squares each with a number of dots.
-%
-% A trial consists of:
-% the instruction on screen while participants read it for as long as they like;
-% a button press on SPACEBAR to bring up
-% a screen with the instruction and the dots presented
-%
-% Participant id > 100 is used for testing and can be overwritten
-
-howLongToWaitForAResponse = 60; % seconds. Set low for test runs
-testing = 1; % 0 for real runs, 1 for testing
-takescreenshot=1;
-SUBJECT=1001;
-
-
-% Clear Matlab/Octave window:
-clc;
-
-% check for Opengl compatibility, abort otherwise:
-AssertOpenGL;
-
-% Reseed the random-number generator for each expt.
-rand('state',sum(100*clock));
-
-% Make sure keyboard mapping is the same on all supported operating systems
-% Apple MacOS/X, MS-Windows and GNU/Linux:
-KbName('UnifyKeyNames');
-
-% Define keys by ID
-leftresp=KbName('c'); % is 6
-middleresp=KbName('b'); % is 5
-rightresp=KbName('m'); % is 16
-escapekey=KbName('ESCAPE'); % is 41
-spacebar=KbName('space'); % is 44
-returnkey=KbName('return'); % is 40
-
-%%%%%%%%%%%%%%%%%%%%%%
-% define stimuli
-%%%%%%%%%%%%%%%%%%%%%%
-
-% order in headers gives the answer to "Which square had the small quantity"? and does not give the correct response
-
-%Condition 1 is numerical vague
-%Condition 2 is numerical precise
-%Condition 3 is visual vague
-%Condition 4 is visual precise
-%
-%Order 1 is small number on left
-%Order 2 is small number on right
-%
-%Quantity 1 is small
-%Quantity 2 is big
-
-
-exp3headers={
-	"Item","Condition","Order","Quantity","Left","Mid","Right","Instruction"};
-
-exp3practicestimuli = {
-	0,0,0,0,6,15,24,"Choose a square with about 10 dots";
-	0,0,0,0,34,25,16,"Choose the square with 16 dots";
[…]

### 2016-08-07 experiments/e2/experimentCode/exp2screenshots.m

@@ -1,814 +1,0 @@
-function exp3screenshots(SUBJECT)
-% This third study has the following structure:
-% Practice, Rest, Block1 Block 2 Block 3, Block 4
-% There are 256 trials
-% There are 64 unique trials
-% There are 4 Blocks
-% Each trial consists of 3 squares each with a number of dots.
-%
-% A trial consists of:
-% the instruction on screen while participants read it for as long as they like;
-% a button press on SPACEBAR to bring up
-% a screen with the instruction and the dots presented
-%
-% Participant id > 100 is used for testing and can be overwritten
-
-howLongToWaitForAResponse = 60; % seconds. Set low for test runs
-testing = 1; % 0 for real runs, 1 for testing
-takescreenshot=1;
-SUBJECT=1001;
-
-
-% Clear Matlab/Octave window:
-clc;
-
-% check for Opengl compatibility, abort otherwise:
-AssertOpenGL;
-
-% Reseed the random-number generator for each expt.
-rand('state',sum(100*clock));
-
-% Make sure keyboard mapping is the same on all supported operating systems
-% Apple MacOS/X, MS-Windows and GNU/Linux:
-KbName('UnifyKeyNames');
-
-% Define keys by ID
-leftresp=KbName('c'); % is 6
-middleresp=KbName('b'); % is 5
-rightresp=KbName('m'); % is 16
-escapekey=KbName('ESCAPE'); % is 41
-spacebar=KbName('space'); % is 44
-returnkey=KbName('return'); % is 40
-
-%%%%%%%%%%%%%%%%%%%%%%
-% define stimuli
-%%%%%%%%%%%%%%%%%%%%%%
-
-% order in headers gives the answer to "Which square had the small quantity"? and does not give the correct response
-
-%Condition 1 is numerical vague
-%Condition 2 is numerical precise
-%Condition 3 is visual vague
-%Condition 4 is visual precise
-%
-%Order 1 is small number on left
-%Order 2 is small number on right
-%
-%Quantity 1 is small
-%Quantity 2 is big
-
-
-exp3headers={
-	"Item","Condition","Order","Quantity","Left","Mid","Right","Instruction"};
-
-exp3practicestimuli = {
-	0,0,0,0,6,15,24,"Choose a square with about 10 dots";
-	0,0,0,0,34,25,16,"Choose the square with 16 dots";
[…]

### 2016-08-07 experiments/e2/experimentCode/exp2test.m

@@ -1,53 +1,0 @@
-function exp3test(SUBJECT)
-SUBJECT=1001;
-
-% Clear Matlab/Octave window:
-clc;
-
-% check for Opengl compatibility, abort otherwise:
-AssertOpenGL;
-
-screenNumber = 0;
-
-% Returns as default the mean gray value of screen:
-gray=GrayIndex(screenNumber);
-
-% [windowPtr,rect]=SCREEN(windowPtrOrScreenNumber,'OpenWindow',[color],[rect],[pixelSize]);
-% [w, rect] = Screen(screenNumber, 'OpenWindow', gray, [1,1,801,601],[]);
-[w, rect] = Screen(screenNumber, 'OpenWindow', gray, [],[]);
-    
-% Set text size
-Screen('TextSize', w, 32);
-
-try
-
-instructionText = sprintf(strcat...
-         ('You will see 3 squares with dots in them, and an instruction\n',...
-         'to choose one of them, as in the picture above.\n\n',...
-         'First you see the instruction, then press space to see the squares.\n\n',...
-         'Use the keys C, B, or M to choose the square that you think matches\n',...
-         'the instruction best. There will be some practice trials so you\n',...
-         'can get used to the routine.'));
-         
-instructionY = 200;
-instructionX = 100;
-
-% Text of introductory instructions to back buffer
-DrawFormattedText(w, instructionText, instructionX, instructionY, WhiteIndex(w));
-
-% Flip introductory instructions to screen
-Screen('Flip', w);
-
-% Wait for key press
-KbWait();
-
-% Go to grey screen
-Screen(w,'FillRect',gray);
-Screen('Flip', w);
-Screen('CloseAll');
-
-catch
-
-Screen('CloseAll');
-
-end_try_catch

### 2016-08-07 experiments/e2/writeup/scratch.R

@@ -1,0 +1,294 @@
+<!--html_document:-->
+  <!--toc: true-->
+  <!--toc_float: true-->
+  <!--mathjax: default-->
+  <!--template: default-->
+  <!--code_folding: hide-->
+  <!--theme: default-->
+  
+  ```{r print_info_on_design}
+cat(paste('number of subjects = ',30,'\n',
+          '---','\n',
+          'number of trials per subject = ',256,'\n', 
+          '---','\n',          
+          'number of unique cells in the design = Item(4) * Number(2) * Vagueness(2) * Order(2) * Quantity(2) = ',4*2*2*2*2,'\n',
+          'number of times a subject saw a unique cell in the design = 256 / 64 = ', 256/64,'\n',
+          '---','\n',
+          'number of higher-level cells in the design (abstracting over Order and Quantity) = Item(4) * Number(2) * Vagueness(2) = ',4*2*2,'\n',
+          'number of times a subject saw a higher level cell in the design =  = 256 / 16 = ', 4*2*2, '\n',
+          sep=''))
+```
+
+
+# Distributional stuff, including transformations.
+
+```{r transformations_table, results='asis'}
+x=445
+lambda = c(-2,-1,-0.5, 0,0.5,1,2)
+
+xd=data.frame(
+  lambda = c(
+    lambda[1],
+    lambda[2],
+    lambda[3],
+    lambda[4],
+    lambda[5],
+    lambda[6],
+    lambda[7]
+  ),
+  x=c(    
+    paste("x = ", sep="", x),
+    paste("x = ", sep="", x),
+    paste("x = ", sep="", x),
+    paste("x = ", sep="", x),
+    paste("x = ", sep="", x),
+    paste("x = ", sep="", x),
+    paste("x = ", sep="", x)
+  ),
+  conventional=c(
+    '$$y=\\frac{1}{x^2}$$', 
+    '$$y=\\frac{1}{x}$$', 
+    '$$y=\\frac{1}{\\sqrt{x}}$$', 
+    '$$y=log_e(x)$$', 
+    '$$y=\\sqrt{x}$$', 
+    '$$y=x$$', 
+    '$$y=x^2$$'
+  ),
+  power=c(
+    '$$y=x^{-2}$$', 
+    '$$y=x^{-1}$$', 
+    '$$y=x^{-0.5}$$', 
+    '$$y=log_e(x)$$', 
+    '$$y=x^{0.5}$$', 
+    '$$y=x^{1}$$', 
+    '$$y=x^{2}$$'
[…]

### 2016-08-10 experiments/e2/writeup/e2-b-knitr.Rnw
Switched to knitr latex Rnw

--- 2016-08-07 experiments/e2/writeup/e2.Rmd
@@
----
-title: "Experiment 2 (of 4)"
-author: "Matt Green"
-output: 
-  pdf_document:
-    toc: true
----
+\documentclass[a4paper]{article}
+\usepackage{geometry}
+\usepackage{placeins}
+\begin{document}
 
-```{r loadKnitr, echo=FALSE}
-suppressPackageStartupMessages(library(knitr))
-```
+<<loadKnitr, echo=FALSE, message=FALSE>>=
+library(knitr)
+opts_chunk$set(cache=FALSE, autodep = TRUE)
+@
 
-```{r loadLibs, echo=FALSE, message=FALSE}
+<<defineFonts>>=
+# .define.fonts <- function () {
+#     quartzFonts(avenir =    c('Avenir Book', 'Avenir Black', 'Avenir Book Oblique', 'Avenir Black Oblique'),
+#                 helvetica = c('Helvetica New Light', 'Helvetica New Bold', 'Helvetica New Light Italic', 'Helvetica New Bold Italic'),
+#                 cmodern = c('CMU Serif', 'CMU Serif Extra','CMU Classical Serif','CMU Serif Extra'))
+# }
+# .define.fonts()
+@
+
+<<loadFonts>>=
+library(extrafont)
+loadfonts()
+@
+
+<<loadLibs, echo=FALSE, message=FALSE>>=
 library(beepr)                     #
@@
 library(xtable)                    #
-```
+@
 
-```{r get_the_data, echo=FALSE}
+\section{Get the data}
+
+<<getTheData, echo=FALSE>>=
 # if the file data.Rda already exists then load it, else do data wrangling
@@
 } # end of the else clause, which is run if the file data.Rda does not exist.
-```
+@
 
-Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions.
+\section{Plots}
 
-```{r plotDists, fig.width=5, fig.height=5, message=FALSE, echo=FALSE}
+<<prepPlotData, echo=FALSE>>=
 # first create a data frame specially for transformations
@@
              value.name="time")
-ggplot(temp, aes(time, colour=RTtype)) + geom_freqpoly() + facet_wrap(transformation~dfSubset, scales='free') + scale_color_manual(values=c("blue","red")) + theme_bw() + theme(aspect.ratio=1)
-```
+@
 
-# Now working with a data set without impossible trials.
+<<plotDists, fig.width=5, fig.height=5, fig.cap='Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions', fig.pos='hbtp', fig.align='right', message=FALSE, warning=FALSE, echo=FALSE>>=
+ggplot(temp, aes(time, colour=RTtype)) + 
+  geom_freqpoly() + 
+  facet_wrap(transformation~dfSubset, scales='free') + 
+  scale_color_manual(values=c("blue","red")) + 
+  theme_bw() + 
+  theme(aspect.ratio=1, text=element_text(family='CMU Serif', size=12, face='italic'))
+@
 
-Show how transformations of RT affect distribution of times, and how they affect which times are outliers.
+<<exampleOfFonts, fig.width=3, fig.height=3, fig.cap='font example', echo=FALSE>>=
+p <- ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point() +
+ggtitle("Fuel Efficiency of 32 Cars") +
+xlab("Weight (x1000 lb)") + ylab("Miles per Gallon") +
+theme_bw() +
+theme(text=element_text(family="CMU Serif", size=14, face='plain'))
+p
+ggsave("ggplot_CMU_Serif.pdf", p, width=3.5, height=3.5)
+@
 
-```{r boxPlots, fig.width=7, fig.height=3, echo=FALSE}
+<<boxPlots, fig.width=7, fig.height=3, fig.cap='Show how transformations of RT affect distribution of times, and how they affect which times are outliers.', echo=FALSE>>=
 load('data_processed.Rda')
@@
   xlab("") + ylab("")
-```
-
-Show how mean times for individual subjects and items vary with respect to the grand mean Log RT.
-
-```{r fastSlowRT, fig.width=7, fig.height=3, message=FALSE, echo=FALSE}
+  @
+  <<fastSlowRT, fig.width=7, fig.height=3, fig.cap='Show how mean times for individual subjects and items vary with respect to the grand mean Log RT.', message=FALSE, echo=FALSE>>=
 # Show how mean times for individual subjects and items vary with respect to the grand mean RT.
@@
 use.style=FALSE))
-```
+@
 
-# Plot main effects and interactions
-
-Main effects
-
-Plot main effects on several transformations
-
-```{r plotMainsa, fig.width=6, fig.height=6, warning=FALSE, echo=FALSE}
+<<plotMainsa, fig.width=6, fig.height=6, fig.cap='Plot main effects on several transformations', message=FALSE, echo=FALSE, warning=FALSE>>=
 # plot main effects on observed data
@@
 theme(aspect.ratio=1, axis.text.y=element_blank())
-```
+@
 
-Main effects
-
-```{r mainB, fig.width=6, fig.height=3, echo=FALSE}
+<<mainB, fig.width=6, fig.height=3, fig.cap='Just the main effects', message=FALSE, echo=FALSE>>=
 # prep main effects rts data frame
-vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness")));names(vrts)[3]<-"Levels"
-nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number")));names(nrts)[3]<-"Levels"
-qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Quantity")));names(qrts)[3]<-"Levels"
-orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Order")));names(orts)[3]<-"Levels"
-irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Item")));names(irts)[3]<-"Levels"
+vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness")));names(vrts)[3]<-"Levels"
+nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number")));names(nrts)[3]<-"Levels"
+qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Quantity")));names(qrts)[3]<-"Levels"
+orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Order")));names(orts)[3]<-"Levels"
+irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Item")));names(irts)[3]<-"Levels"
 rts=rbind(vrts,nrts,qrts,orts)
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
+rts$mins=rts$RT_log-rts$ci
+rts$maxs=rts$RT_log+rts$ci
 # plot main effects
-ggplot(rts, aes(y=RT_RecSqt, x=Levels, group=1, ymin=mins, ymax=maxs)) + 
+ggplot(rts, aes(y=RT_log, x=Levels, group=1, ymin=mins, ymax=maxs)) +
 geom_line() +
@@
 theme(aspect.ratio=1)
-```
+@
 
-Main effects over items
+<<mainD, fig.width=6, fig.height=3.5, fig.cap='Main effects over item ratios', message=FALSE, echo=FALSE>>=
+nrts <- cbind(Effect="Number",
+xlev=factor(c(1,2)), 
+myshape='a',
+summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Number")))
+names(nrts)[5]<-"Levels"
 
-```{r mainD, message=FALSE, fig.width=6, fig.height=3, echo=FALSE}
-vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Vagueness")));names(vrts)[4]<-"Levels"
-nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Number")));names(nrts)[4]<-"Levels"
-qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Quantity")));names(qrts)[4]<-"Levels"
-orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Order")));names(orts)[4]<-"Levels"
-rts=rbind(vrts,nrts,qrts,orts)
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
+orts <- cbind(Effect="Order",
+xlev=factor(c(1,2)), 
+myshape='b',
+summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Order")))
+names(orts)[5]<-"Levels"
+
+qrts <- cbind(Effect="Quantity",
+xlev=factor(c(1,2)), 
+myshape='d',
+summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Quantity")))
+names(qrts)[5]<-"Levels"
+
+vrts <- cbind(Effect="Vagueness",
+xlev=factor(c(1,2)),
+myshape='e',
+summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Vagueness")))
+names(vrts)[5]<-"Levels"
+
+rts=rbind(nrts,orts,qrts,vrts)
+rts$mins=rts$RT_log-rts$ci
+rts$maxs=rts$RT_log+rts$ci
 rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
-ggplot(rts, 
-			 aes(y=RT_RecSqt, x=item_mean_ratio, group=Levels, ymin=mins, 
-			 		ymax=maxs, shape=Levels, fill=Levels, col=Levels)) +
-  scale_shape_manual(values=c(21,21,22,22,23,23,24,24)) +
-  scale_fill_manual(values=c("black", "gray", "black", "gray", "black", "gray", "black", "gray")) +
-  scale_color_manual(values=c("black", "gray", "black", "gray", "black", "gray", "black", "gray")) +
-  geom_line() +
-  geom_errorbar(width=.1) +
-  geom_point(cex=3) + 
+#rts$myshape=rep(c(21,22,23,24), each=8)
+
+p1=
+ggplot(data=rts, aes(x=item_mean_ratio, y=RT_log, group=Levels)) +
+geom_errorbar(data=droplevels(subset(rts, Effect=='Number')),width=.01, aes(ymin=mins, ymax=maxs)) +
+geom_line(data=droplevels(subset(rts, Effect=='Number'))) +
+geom_point(data=droplevels(subset(rts, Effect=='Number')),
+aes(shape=Levels, fill=Levels), size=3 ) +
+geom_errorbar(data=droplevels(subset(rts, Effect=='Order')), width=.01, aes(ymin=mins, ymax=maxs)) +
+geom_line(data=droplevels(subset(rts, Effect=='Order'))) +
+geom_point(data=droplevels(subset(rts, Effect=='Order')),
+aes(shape=Levels, fill=Levels), size=3  ) +
+geom_errorbar(data=droplevels(subset(rts, Effect=='Quantity')),width=.01, aes(ymin=mins, ymax=maxs)) +
+geom_line(data=droplevels(subset(rts, Effect=='Quantity'))) +
+geom_point(data=droplevels(subset(rts, Effect=='Quantity')),
+aes(shape=Levels, fill=Levels), size=3  ) +
+geom_errorbar(data=droplevels(subset(rts, Effect=='Vagueness')),width=.01, aes(ymin=mins, ymax=maxs)) +
+geom_line(data=droplevels(subset(rts, Effect=='Vagueness'))) +
+geom_point(data=droplevels(subset(rts, Effect=='Vagueness')),
+aes(shape=Levels, fill=Levels), size=3  ) +
 facet_grid(~Effect) +
-  theme_bw() +
+scale_shape_manual(values = c(
+"Numeric"=21,
+"Verbal"=21,
+"LtoR"=22,
+"RtoL"=22,
+"Small"=23,
+"Large"=23,
+"Crisp"=24,
+"Vague"=24),
+breaks=c('Numeric','Verbal','LtoR','RtoL','Small','Large','Crisp','Vague'))  +
+scale_fill_manual(values = c(
+"Numeric"="black",
+"Verbal"="white",
+"LtoR"="black",
+"RtoL"="white",
+"Small"="black",
+"Large"="white",
+"Crisp"="black",
+"Vague"="white"),
+breaks=c('Numeric','Verbal','LtoR','RtoL','Small','Large','Crisp','Vague')) +
+scale_x_continuous(minor_breaks=NULL,
+breaks=c(0.5125000, 0.6876471, 0.7691558, 0.8166667),
+labels=abbreviate(c(0.5125000, 0.6876471, 0.7691558, 0.8166667))) +
 theme(aspect.ratio=1,
-        panel.grid=element_blank()) +
-  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1)) 
-```
+legend.position='top',
+axis.text.x=element_text(size=7),
+legend.key=element_rect(fill = 'white'),
+panel.background=element_rect(fill = 'grey96'))
+p1
+@
 
-2 way interactions not involving items.
-
-```{r 2way, fig.width=6, fig.height=3, echo=FALSE}
-# prep 2 ways rts data frame
-rts1 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Number")); names(rts1)[1] <- "F1"; names(rts1)[2] <- "F2"
-rts1$Effect <- rep("Vagueness:Number",4)
-rts2 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Quantity")); names(rts2)[1] <- "F1"; names(rts2)[2] <- "F2"
-rts2$Effect <- rep("Vagueness:Quantity",4)
-rts3 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Order")); names(rts3)[1] <- "F1"; names(rts3)[2] <- "F2"
-rts3$Effect <- rep("Vagueness:Order",4)
-rts5 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Quantity")); names(rts5)[1] <- "F1"; names(rts5)[2] <- "F2"
-rts5$Effect <- rep("Number:Quantity",4)
-rts6 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Order")); names(rts6)[1] <- "F1"; names(rts6)[2] <- "F2"
-rts6$Effect <- rep("Number:Order",4)
-rts8 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Quantity","Order")); names(rts8)[1] <- "F1"; names(rts8)[2] <- "F2"
-rts8$Effect <- rep("Quantity:Order",4)
-rts <- rbind(rts1,rts2,rts3,rts5,rts6,rts8)
-rts <- subset(rts, select=c(Effect, F1, F2, RT_RecSqt, sd, se, ci))
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-# plot 2 way interactions
-ggplot(rts, aes(y=RT_RecSqt, x=F2, group=F1, ymin=mins, ymax=maxs, pch=F1)) + 
-  geom_line() +
-  geom_errorbar(width=.1) +
-  geom_point(cex=3) + 
-  facet_wrap(~Effect, scale="free_x", nrow=1) + 
-  theme_bw() +
-  theme(aspect.ratio=1) +
-  theme(legend.title=element_blank()) +
-  xlab("Levels")
-```
-
-Plot 2 way interactions over items
-
-```{r 22way, message=FALSE, fig.width=6, fig.height=3, echo=FALSE}
+<<22way, fig.width=6, fig.height=3, fig.cap='2-way interactions over items', message=FALSE, echo=FALSE>>=
 # prep 2 ways over items rts data frame
@@
 scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))
-```
+@
 
-Plot vagueness by number interaction over items
-
-```{r plotInt1, fig.width=7, fig.height=3.5, echo=FALSE}
+<<plotInt1, fig.width=7, fig.height=3.5, fig.cap='Plot vagueness by number interaction over items', echo=FALSE>>=
 rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item"))
@@
 facet_grid(~Number)
-```
+@
 
-3 way interactions
-
-```{r plot3way, fig.width=7, fig.height=5, echo=FALSE}
+<<plot3way, fig.width=7, fig.height=5, fig.cap='3 way interactions',echo=FALSE>>=
 rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Quantity"))
@@
 grid.arrange(nrow=2, p1,p2,p3,p4)
-```
+@
 
-Plot vagueness by number by quantity over items.
-
-```{r vnitems, fig.width=7, fig.height=4, echo=FALSE}
+<<vnitems, fig.width=7, fig.height=4, fig.cap='Plot vagueness by number by quantity over items.',echo=FALSE>>=
 rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item", "Quantity"))
@@
 facet_grid(~Number+Quantity)
-```
+@
 
-4 way interaction
-
-```{r 4way, fig.width=6, fig.height=3, echo=FALSE}
+<<4way, fig.width=6, fig.height=3, fig.cap='4 way interaction', echo=FALSE>>=
 rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness",  "Quantity", "Order"))
@@
 theme(legend.title=element_blank())
-```
+@
 
-4 way interaction split over items
-
-```{r 4waysplit, fig.width=7, fig.height=7, echo=FALSE}
+<<4waysplit, fig.width=7, fig.height=7, fig.cap='4 way interaction split over items',echo=FALSE>>=
 rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness", "Item", "Quantity", "Order"))
@@
 facet_grid(Vagueness+Order~Number+Quantity)
-```
+@
 
-4 way interaction split over items showing ratio of item dots.
-
-```{r 4waySplitRatio, fig.width=7, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}
+<<4waySplitRatio, fig.width=7, fig.height=7, fig.cap='4 way interaction split over items showing ratio of item dots.',echo=FALSE, message=FALSE, warning=FALSE>>=
 rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness", "item_mean_ratio", "Quantity", "Order"))
@@
 facet_grid(Vagueness+Order~Number+Quantity)
-```
+@
 
-# Show compression of item ratios
-
-```{r showCompression, fig.width=7, fig.height=7, echo=FALSE}
+<<showCompression, fig.width=7, fig.height=7, fig.cap='Show compression of item ratios', echo=FALSE>>=
 #  Show compression of ratios of numbers of dots in the items
@@
 item_mean_ratio_equally_spaced = seq(from = mean(c(6 / 15, 15 / 24)), to = mean(c(36 / 45, 45 / 54)), length.out = 4)
-plot(y=c(2,2,2,2), x=item_range_ratio, xlim=c(0,1), ylim=c(0,3), ylab="", yaxt='n', col="blue", pch=21, bg='blue', type='o', xaxt='n', xlab="ratio: small is easily discriminable", main="Discriminability of each item, expressed as ratios between array magnitudes")
+plot(y=c(2,2,2,2), x=item_range_ratio, xlim=c(0,1), ylim=c(0,3), ylab="", yaxt='n', col="blue", pch=21, bg='blue', type='o', xaxt='n', xlab="ratio: small is easily discriminable", main="Discriminability of each item,\n expressed as ratios between array magnitudes")
 text(y=c(2.15,2.15,2.15,2.15), x=item_range_ratio, labels=round(item_range_ratio,2), col="blue")
@@
 axis(1, at=seq(0,1,0.1))
-```
+@
 
-# Lmer model
+\FloatBarrier
+\section{Lmer model}
 
-```{r rmme}
+<<rmme>>=
 load("data_processed.Rda")
-```
+@
 
-```{r lmer5}
+<<lmer5, cache=TRUE>>=
 v5 <- lmerTest::lmer(data=dd,
@@
 (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
-```
+@
 
-```{r v51}
+<<v51, cache=TRUE>>=
 summary(v5)
-```
+@
 
-```{r v511, results='asis'}
+<<v511, results='asis', cache=TRUE>>=
 print(xtable(coef(summary(v5b))))
-```
+@
 
-```{r v52}
+<<v52>>=
 cat("R^2")
 cor(fitted(v5), dd$RT_RecSqt)^2
-```
+@
 
-```{r v53}
-par(mfrow=c(2,5))
+<<plotLMERfnc, fig.width=7, fig.height=3.5, fig.cap='caption'>>=
+par(mfrow=c(3,4), family='cmodern')
 plotLMER.fnc(v5)
-```
+@
 
@@
 
-```{r v54, fig.width=7, fig.height=4}
+<<v54, fig.width=7, fig.height=4, fig.cap='Estimates plot'>>=
 e=data.frame(coef=summary(v5)$coef[-1,1]) # estimates, without intercept
@@
 box(col='grey')
-```
+@
 
@@
 
-```{r baayenPLots99, fig.width=9, fig.height=3}
+<<baayenPLots99, fig.width=9, fig.height=3, fig.cap='caption'>>=
 # Baayen 4-plot model criticism
@@
 abline(h=c(-2.5,2.5))
-```
+@
+
+\end{document}

### 2016-08-10 experiments/e2/writeup/e2-b-knitr.tex
Switched to knitr latex Rnw

\documentclass[a4paper]{article}\usepackage[]{graphicx}\usepackage[]{color}
%% maxwidth is the original width if it is less than linewidth
%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\usepackage{geometry}
\usepackage{placeins}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# .define.fonts <- function () \{}
\hlcom{#     quartzFonts(avenir =    c('Avenir Book', 'Avenir Black', 'Avenir Book Oblique', 'Avenir Black Oblique'),}
\hlcom{#                 helvetica = c('Helvetica New Light', 'Helvetica New Bold', 'Helvetica New Light Italic', 'Helvetica New Bold Italic'),}
\hlcom{#                 cmodern = c('CMU Serif', 'CMU Serif Extra','CMU Classical Serif','CMU Serif Extra'))}
\hlcom{# \}}
\hlcom{# .define.fonts()}
\end{alltt}
\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{library}\hlstd{(extrafont)}
\end{alltt}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Registering fonts with R}}\begin{alltt}
\hlkwd{loadfonts}\hlstd{()}
\end{alltt}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# .Keyboard already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Andale Mono already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# More than one version of regular/bold/italic found for Apple Braille. Skipping setup for this font.}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# AppleMyungjo already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Arial Black already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Arial already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Arial Narrow already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Arial Rounded MT Bold already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Arial Unicode MS already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Bodoni Ornaments already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Bodoni 72 Smallcaps already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# No regular (non-bold, non-italic) version of . Skipping setup for this font.}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# No regular (non-bold, non-italic) version of Brush Script MT. Skipping setup for this font.}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Comic Sans MS already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Courier New already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# No regular (non-bold, non-italic) version of DIN Alternate. Skipping setup for this font.}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# No regular (non-bold, non-italic) version of DIN Condensed. Skipping setup for this font.}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Georgia already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Impact already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Khmer Sangam MN already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Lao Sangam MN already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Luminari already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Microsoft Sans Serif already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Tahoma already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Times New Roman already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Trattatello already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Trebuchet MS already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Verdana already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Webdings already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Wingdings already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Wingdings 2 already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Wingdings 3 already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# More than one version of regular/bold/italic found for CMU Bright. Skipping setup for this font.}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# No regular (non-bold, non-italic) version of CMU Classical Serif. Skipping setup for this font.}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# CMU Concrete already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# CMU Sans Serif already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# CMU Sans Serif Demi Condensed already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# CMU Serif already registered with pdfFonts().}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# No regular (non-bold, non-italic) version of CMU Serif Extra. Skipping setup for this font.}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# No regular (non-bold, non-italic) version of CMU Serif Upright Italic. Skipping setup for this font.}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# More than one version of regular/bold/italic found for CMU Typewriter Text. Skipping setup for this font.}}

{\ttfamily\noindent\itshape\color{messagecolor}{\#\# CMU Typewriter Text Variable Width already registered with pdfFonts().}}\end{kframe}
\end{knitrout}

\section{Get the data}

\section{Plots}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]

\hfill{}\includegraphics[width=\maxwidth]{figure/plotDists-1}

\caption[Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions]{Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions}\label{fig:plotDists}
\end{figure}

\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/exampleOfFonts-1} \caption[font example]{font example}\label{fig:exampleOfFonts}
\end{figure}

\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/boxPlots-1} \caption[Show how transformations of RT affect distribution of times, and how they affect which times are outliers]{Show how transformations of RT affect distribution of times, and how they affect which times are outliers.}\label{fig:boxPlots}
\end{figure}

\end{knitrout}
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/fastSlowRT-1} \caption[Show how mean times for individual subjects and items vary with respect to the grand mean Log RT]{Show how mean times for individual subjects and items vary with respect to the grand mean Log RT.}\label{fig:fastSlowRT}
\end{figure}

\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/plotMainsa-1} \caption[Plot main effects on several transformations]{Plot main effects on several transformations}\label{fig:plotMainsa}
\end{figure}

\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/mainB-1} \caption[Just the main effects]{Just the main effects}\label{fig:mainB}
\end{figure}

\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/mainD-1} \caption[Main effects over item ratios]{Main effects over item ratios}\label{fig:mainD}
\end{figure}

\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/22way-1} \caption[2-way interactions over items]{2-way interactions over items}\label{fig:22way}
\end{figure}

\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/plotInt1-1} \caption[Plot vagueness by number interaction over items]{Plot vagueness by number interaction over items}\label{fig:plotInt1}
\end{figure}

\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/plot3way-1} \caption[3 way interactions]{3 way interactions}\label{fig:plot3way}
\end{figure}

\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/vnitems-1} \caption[Plot vagueness by number by quantity over items]{Plot vagueness by number by quantity over items.}\label{fig:vnitems}
\end{figure}

\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/4way-1} \caption[4 way interaction]{4 way interaction}\label{fig:4way}
\end{figure}

\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/4waysplit-1} \caption[4 way interaction split over items]{4 way interaction split over items}\label{fig:4waysplit}
\end{figure}

\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/4waySplitRatio-1} \caption[4 way interaction split over items showing ratio of item dots]{4 way interaction split over items showing ratio of item dots.}\label{fig:4waySplitRatio}
\end{figure}

\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/showCompression-1} \caption[Show compression of item ratios]{Show compression of item ratios}\label{fig:showCompression}
\end{figure}

\end{knitrout}

\FloatBarrier
\section{Lmer model}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{load}\hlstd{(}\hlstr{"data_processed.Rda"}\hlstd{)}
\end{alltt}
\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{v5} \hlkwb{<-} \hlstd{lmerTest}\hlopt{::}\hlkwd{lmer}\hlstd{(}\hlkwc{data}\hlstd{=dd,}
\hlstd{RT_log} \hlopt{~}
\hlstd{c_Vag} \hlopt{+} \hlstd{c_Num} \hlopt{+} \hlstd{c_Qty} \hlopt{+} \hlstd{c_Ord} \hlopt{+}
\hlstd{c_Num}\hlopt{:}\hlstd{c_Vag}\hlopt{:}\hlstd{c_Qty} \hlopt{+}
\hlstd{item_mean_ratio} \hlopt{+}
\hlstd{s_Trl} \hlopt{+}
\hlstd{RTprev_log} \hlopt{+}
\hlstd{nchar_instr} \hlopt{+}
\hlstd{(}\hlnum{1}\hlopt{+}\hlstd{c_Vag} \hlopt{+} \hlstd{c_Num} \hlopt{+} \hlstd{c_Qty} \hlopt{+} \hlstd{c_Ord}\hlopt{|}\hlstd{Subject))}
\hlstd{v5b} \hlkwb{<-} \hlstd{lme4}\hlopt{::}\hlkwd{lmer}\hlstd{(}\hlkwc{data}\hlstd{=dd,}
\hlstd{RT_log} \hlopt{~}
\hlstd{c_Vag} \hlopt{+} \hlstd{c_Num} \hlopt{+} \hlstd{c_Qty} \hlopt{+} \hlstd{c_Ord} \hlopt{+}
\hlstd{c_Num}\hlopt{:}\hlstd{c_Vag}\hlopt{:}\hlstd{c_Qty} \hlopt{+}
\hlstd{item_mean_ratio} \hlopt{+}
\hlstd{s_Trl} \hlopt{+}
\hlstd{RTprev_log} \hlopt{+}
\hlstd{nchar_instr} \hlopt{+}
\hlstd{(}\hlnum{1}\hlopt{+}\hlstd{c_Vag} \hlopt{+} \hlstd{c_Num} \hlopt{+} \hlstd{c_Qty} \hlopt{+} \hlstd{c_Ord}\hlopt{|}\hlstd{Subject))}
\end{alltt}
\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{summary}\hlstd{(v5)}
\end{alltt}
\begin{verbatim}
## Linear mixed model fit by REML t-tests use Satterthwaite approximations
##   to degrees of freedom [lmerMod]
## Formula: RT_log ~ c_Vag + c_Num + c_Qty + c_Ord + c_Num:c_Vag:c_Qty +  
##     item_mean_ratio + s_Trl + RTprev_log + nchar_instr + (1 +  
##     c_Vag + c_Num + c_Qty + c_Ord | Subject)
##    Data: dd
## 
## REML criterion at convergence: 11474.8
## 
## Scaled residuals: 
##     Min      1Q  Median      3Q     Max 
## -4.5470 -0.6351 -0.0955  0.5372  5.0914 
## 
## Random effects:
##  Groups   Name        Variance Std.Dev. Corr                   
##  Subject  (Intercept) 0.153949 0.39236                         
##           c_Vag       0.001546 0.03932   0.69                  
##           c_Num       0.165314 0.40659  -0.67 -0.64            
##           c_Qty       0.008148 0.09027   0.16  0.26 -0.34      
##           c_Ord       0.001559 0.03949  -0.13  0.02 -0.39 -0.52
##  Residual             0.249734 0.49973                         
## Number of obs: 7677, groups:  Subject, 30
## 
## Fixed effects:
##                     Estimate Std. Error         df t value Pr(>|t|)    
## (Intercept)        6.400e+00  1.252e-01  2.600e+02  51.112  < 2e-16 ***
## c_Vag              6.094e-02  1.387e-02  3.300e+01   4.393 0.000112 ***
## c_Num             -4.336e-01  7.515e-02  2.900e+01  -5.770 2.97e-06 ***
## c_Qty             -6.907e-02  2.005e-02  2.900e+01  -3.445 0.001743 ** 
## c_Ord              1.796e-02  1.350e-02  5.100e+01   1.331 0.189164    
## item_mean_ratio    7.714e-01  4.927e-02  7.551e+03  15.658  < 2e-16 ***
## s_Trl             -1.070e-01  5.807e-03  7.558e+03 -18.421  < 2e-16 ***
## RTprev_log         6.047e-02  9.692e-03  7.594e+03   6.239 4.63e-10 ***
## nchar_instr        6.086e-03  1.944e-03  7.551e+03   3.131 0.001749 ** 
## c_Vag:c_Num:c_Qty  1.049e-01  4.607e-02  7.551e+03   2.278 0.022742 *  
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## 
## Correlation of Fixed Effects:
##             (Intr) c_Vag  c_Num  c_Qty  c_Ord  itm_m_ s_Trl  RTprv_ nchr_n
## c_Vag        0.083                                                        
## c_Num       -0.360 -0.337                                                 
## c_Qty        0.067  0.117 -0.278                                          
## c_Ord       -0.051  0.007 -0.206 -0.228                                   
## item_men_rt -0.257 -0.004  0.001  0.000 -0.001                            
## s_Trl       -0.104  0.003 -0.001 -0.002  0.006 -0.017                     
## RTprev_log  -0.587  0.006 -0.004 -0.003  0.019 -0.015  0.185              
## nchar_instr -0.505  0.237 -0.034  0.018  0.000 -0.017  0.000  0.006       
## c_Vg:c_N:_Q  0.070 -0.033  0.005 -0.003  0.000  0.002 -0.001 -0.002 -0.138
\end{verbatim}
\end{kframe}
\end{knitrout}

\begin{kframe}
\begin{alltt}
\hlkwd{print}\hlstd{(}\hlkwd{xtable}\hlstd{(}\hlkwd{coef}\hlstd{(}\hlkwd{summary}\hlstd{(v5b))))}
\end{alltt}
\end{kframe}% latex table generated in R 3.3.1 by xtable 1.8-2 package
% Wed Aug 10 22:41:39 2016
\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
  \hline
 & Estimate & Std. Error & t value \\ 
  \hline
(Intercept) & 6.40 & 0.13 & 51.11 \\ 
  c\_Vag & 0.06 & 0.01 & 4.39 \\ 
  c\_Num & -0.43 & 0.08 & -5.77 \\ 
  c\_Qty & -0.07 & 0.02 & -3.45 \\ 
  c\_Ord & 0.02 & 0.01 & 1.33 \\ 
  item\_mean\_ratio & 0.77 & 0.05 & 15.66 \\ 
  s\_Trl & -0.11 & 0.01 & -18.42 \\ 
  RTprev\_log & 0.06 & 0.01 & 6.24 \\ 
  nchar\_instr & 0.01 & 0.00 & 3.13 \\ 
  c\_Vag:c\_Num:c\_Qty & 0.10 & 0.05 & 2.28 \\ 
   \hline
\end{tabular}
\end{table}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{cat}\hlstd{(}\hlstr{"R^2"}\hlstd{)}
\end{alltt}
\begin{verbatim}
## R^2
\end{verbatim}
\begin{alltt}
\hlkwd{cor}\hlstd{(}\hlkwd{fitted}\hlstd{(v5), dd}\hlopt{$}\hlstd{RT_RecSqt)}\hlopt{^}\hlnum{2}
\end{alltt}
\begin{verbatim}
## [1] 0.5095834
\end{verbatim}
\end{kframe}
\end{knitrout}

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlkwd{par}\hlstd{(}\hlkwc{mfrow}\hlstd{=}\hlkwd{c}\hlstd{(}\hlnum{3}\hlstd{,}\hlnum{4}\hlstd{),} \hlkwc{family}\hlstd{=}\hlstr{'cmodern'}\hlstd{)}
\hlkwd{plotLMER.fnc}\hlstd{(v5)}
\end{alltt}
\begin{verbatim}
## effect size (range) for  c_Vag is  0.03470056 
## effect size (range) for  c_Num is  0.4073765 
## effect size (range) for  c_Qty is  0.09530422 
## effect size (range) for  c_Ord is  0.0179595 
## effect size (range) for  item_mean_ratio is  0.2346348 
## effect size (range) for  s_Trl is  0.369093 
## effect size (range) for  RTprev_log is  0.2759498 
## effect size (range) for  nchar_instr is  0.05477539
\end{verbatim}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot.new(): figure margins too large}}\end{kframe}
\end{knitrout}

Plot model coefficients and ci's

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{e}\hlkwb{=}\hlkwd{data.frame}\hlstd{(}\hlkwc{coef}\hlstd{=}\hlkwd{summary}\hlstd{(v5)}\hlopt{$}\hlstd{coef[}\hlopt{-}\hlnum{1}\hlstd{,}\hlnum{1}\hlstd{])} \hlcom{# estimates, without intercept}
\hlstd{q}\hlkwb{=}\hlkwd{as.data.frame}\hlstd{(}\hlkwd{confint}\hlstd{(v5,} \hlkwc{method}\hlstd{=}\hlstr{'Wald'}\hlstd{)[}\hlnum{18}\hlopt{:}\hlnum{26}\hlstd{,}\hlnum{1}\hlopt{:}\hlnum{2}\hlstd{])}
\hlstd{eq}\hlkwb{=}\hlkwd{cbind}\hlstd{(e,q)}
\hlstd{y}\hlkwb{=}\hlstd{eq[}\hlkwd{rev}\hlstd{(}\hlkwd{rownames}\hlstd{(eq)),]}
\hlkwd{par}\hlstd{(}\hlkwc{mar}\hlstd{=}\hlkwd{c}\hlstd{(}\hlnum{4}\hlstd{,}\hlnum{10}\hlstd{,}\hlnum{1}\hlstd{,}\hlnum{1}\hlstd{))}
\hlkwd{plot}\hlstd{(}\hlkwc{y}\hlstd{=}\hlnum{1}\hlopt{:}\hlkwd{nrow}\hlstd{(y),} \hlkwc{x}\hlstd{=y}\hlopt{$}\hlstd{coef,} \hlkwc{xlim}\hlstd{=}\hlkwd{range}\hlstd{(y}\hlopt{$}\hlstr{"2.5 %"}\hlstd{, y}\hlopt{$}\hlstr{"97.5 %"}\hlstd{),} \hlkwc{type}\hlstd{=}\hlstr{'n'}\hlstd{,} \hlkwc{axes}\hlstd{=F,} \hlkwc{xlab}\hlstd{=}\hlstr{""}\hlstd{,} \hlkwc{ylab}\hlstd{=}\hlstr{""}\hlstd{)}
\hlkwd{segments}\hlstd{(}\hlkwc{y0}\hlstd{=}\hlnum{1}\hlopt{:}\hlkwd{nrow}\hlstd{(y),} \hlkwc{y1}\hlstd{=}\hlnum{1}\hlopt{:}\hlkwd{nrow}\hlstd{(y),} \hlkwc{x0}\hlstd{=y}\hlopt{$}\hlstr{"2.5 %"}\hlstd{,} \hlkwc{x1}\hlstd{=y}\hlopt{$}\hlstr{"97.5 %"}\hlstd{,} \hlkwc{lwd}\hlstd{=}\hlnum{.55}\hlstd{)}
\hlkwd{points}\hlstd{(}\hlkwc{y}\hlstd{=}\hlnum{1}\hlopt{:}\hlkwd{nrow}\hlstd{(y),} \hlkwc{x}\hlstd{=y}\hlopt{$}\hlstd{coef,} \hlkwc{pch}\hlstd{=}\hlnum{20}\hlstd{,} \hlkwc{cex}\hlstd{=}\hlnum{.75}\hlstd{)}
\hlkwd{abline}\hlstd{(}\hlkwc{v}\hlstd{=}\hlnum{0}\hlstd{,} \hlkwc{lty}\hlstd{=}\hlnum{3}\hlstd{,} \hlkwc{col}\hlstd{=}\hlstr{'grey'}\hlstd{)}
\hlkwd{axis}\hlstd{(}\hlnum{2}\hlstd{,} \hlkwc{labels}\hlstd{=}\hlkwd{row.names}\hlstd{(y),} \hlkwc{at}\hlstd{=}\hlnum{1}\hlopt{:}\hlkwd{nrow}\hlstd{(y),} \hlkwc{las}\hlstd{=}\hlnum{1}\hlstd{)}
\hlkwd{axis}\hlstd{(}\hlnum{1}\hlstd{,} \hlkwc{las}\hlstd{=}\hlnum{1}\hlstd{)}
\hlkwd{mtext}\hlstd{(}\hlstr{"estimate"}\hlstd{,} \hlkwc{side}\hlstd{=}\hlnum{1}\hlstd{,} \hlkwc{line}\hlstd{=}\hlnum{2.5}\hlstd{,} \hlkwc{cex.lab}\hlstd{=}\hlnum{1}\hlstd{,} \hlkwc{las}\hlstd{=}\hlnum{1}\hlstd{)}
\hlkwd{box}\hlstd{(}\hlkwc{col}\hlstd{=}\hlstr{'grey'}\hlstd{)}
\end{alltt}
\end{kframe}\begin{figure}
\includegraphics[width=\maxwidth]{figure/v54-1} \caption[Estimates plot]{Estimates plot}\label{fig:v54}
\end{figure}

\end{knitrout}

Model criticism baayen plots

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# Baayen 4-plot model criticism}
\hlkwd{par}\hlstd{(}\hlkwc{mfrow}\hlstd{=}\hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,}\hlnum{4}\hlstd{),} \hlkwc{pty}\hlstd{=}\hlstr{'s'}\hlstd{)}
\hlcom{# create scaled residuals}
\hlstd{dd}\hlopt{$}\hlstd{rstand} \hlkwb{=} \hlkwd{as.vector}\hlstd{(}\hlkwd{scale}\hlstd{(}\hlkwd{resid}\hlstd{(v5)))}
\hlcom{# plot scaled residuals density}
\hlkwd{plot}\hlstd{(}\hlkwd{density}\hlstd{(dd}\hlopt{$}\hlstd{rstand))}
\hlcom{# plot sample quantiles versus theoretical quantiles}
\hlkwd{qqnorm}\hlstd{(dd}\hlopt{$}\hlstd{rstand,} \hlkwc{cex}\hlstd{=}\hlnum{.5}\hlstd{)}
\hlkwd{qqline}\hlstd{(dd}\hlopt{$}\hlstd{rstand)}
\hlcom{# plot standardised residuals versus fitted values}
\hlkwd{plot}\hlstd{(dd}\hlopt{$}\hlstd{rstand} \hlopt{~} \hlkwd{fitted}\hlstd{(v5),} \hlkwc{pch}\hlstd{=}\hlstr{'.'}\hlstd{)}
\hlcom{# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot}
\hlkwd{abline}\hlstd{(}\hlkwc{h}\hlstd{=}\hlkwd{c}\hlstd{(}\hlopt{-}\hlnum{2.5}\hlstd{,}\hlnum{2.5}\hlstd{))}
\end{alltt}
\end{kframe}\begin{figure}
\includegraphics[width=\maxwidth]{figure/baayenPLots99-1} \caption[caption]{caption}\label{fig:baayenPLots99}
\end{figure}

\end{knitrout}

\end{document}

### 2016-08-10 experiments/e2/writeup/e2-b.Rnw
Switched to knitr latex Rnw

--- 2016-08-10 experiments/e2/writeup/e2-b-knitr.Rnw
@@
-\documentclass[a4paper]{article}
-\usepackage{geometry}
+\documentclass{article}
+
 \usepackage{placeins}
@@
 
-<<loadKnitr, echo=FALSE, message=FALSE>>=
-library(knitr)
-opts_chunk$set(cache=FALSE, autodep = TRUE)
-@
-
-<<defineFonts>>=
-# .define.fonts <- function () {
-#     quartzFonts(avenir =    c('Avenir Book', 'Avenir Black', 'Avenir Book Oblique', 'Avenir Black Oblique'),
-#                 helvetica = c('Helvetica New Light', 'Helvetica New Bold', 'Helvetica New Light Italic', 'Helvetica New Bold Italic'),
-#                 cmodern = c('CMU Serif', 'CMU Serif Extra','CMU Classical Serif','CMU Serif Extra'))
-# }
-# .define.fonts()
-@
-
-<<loadFonts>>=
-library(extrafont)
-loadfonts()
+<<loadKnitr, echo=FALSE>>=
+suppressPackageStartupMessages(library(knitr))
 @
@@
 
-<<getTheData, echo=FALSE>>=
+<<getTheData, echo=1:4>>=
 # if the file data.Rda already exists then load it, else do data wrangling
@@
 \section{Plots}
-
-<<prepPlotData, echo=FALSE>>=
+\FloatBarrier
+\begin{figure}[hbtp]
+<<plotDists, fig=TRUE, message=FALSE, echo=FALSE, cache=TRUE>>=
 # first create a data frame specially for transformations
@@
              value.name="time")
+ggplot(temp, aes(time, colour=RTtype)) + geom_freqpoly() + facet_wrap(transformation~dfSubset, scales='free') + scale_color_manual(values=c("blue","red")) + theme_bw() + theme(aspect.ratio=1)
 @
-
-<<plotDists, fig.width=5, fig.height=5, fig.cap='Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions', fig.pos='hbtp', fig.align='right', message=FALSE, warning=FALSE, echo=FALSE>>=
-ggplot(temp, aes(time, colour=RTtype)) + 
-  geom_freqpoly() + 
-  facet_wrap(transformation~dfSubset, scales='free') + 
-  scale_color_manual(values=c("blue","red")) + 
-  theme_bw() + 
-  theme(aspect.ratio=1, text=element_text(family='CMU Serif', size=12, face='italic'))
-@
-
-<<exampleOfFonts, fig.width=3, fig.height=3, fig.cap='font example', echo=FALSE>>=
-p <- ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point() +
-ggtitle("Fuel Efficiency of 32 Cars") +
-xlab("Weight (x1000 lb)") + ylab("Miles per Gallon") +
-theme_bw() +
-theme(text=element_text(family="CMU Serif", size=14, face='plain'))
-p
-ggsave("ggplot_CMU_Serif.pdf", p, width=3.5, height=3.5)
-@
-
-<<boxPlots, fig.width=7, fig.height=3, fig.cap='Show how transformations of RT affect distribution of times, and how they affect which times are outliers.', echo=FALSE>>=
-load('data_processed.Rda')
-# allow reference to transformation
-subdata = reshape2::melt(dd,
-                         measure.vars=c("RT_log", "RT_raw"),
-                         variable.name="transformation",
-                         value.name="value") 
-# boxplots for subjects and items separately for each dependent variable
-ggplot(subdata) + 
-  facet_wrap(~ transformation, scales="free", ncol=4) +
-  geom_boxplot(aes(y=value, x=Item, col="Items"), outlier.colour = "black", outlier.shape = 21, outlier.size=0.1) + 
-  geom_boxplot(aes(y=value, x=Subject, col="Subjects"), outlier.colour = "black", outlier.shape = 21, outlier.size=0.1) + 
-  theme_bw() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=5), aspect.ratio=1, panel.grid=element_blank()) + 
-  scale_colour_manual("",values=c("Items"="orange","Subjects"="blue")) +
-  xlab("") + ylab("")
-  @
-  <<fastSlowRT, fig.width=7, fig.height=3, fig.cap='Show how mean times for individual subjects and items vary with respect to the grand mean Log RT.', message=FALSE, echo=FALSE>>=
-# Show how mean times for individual subjects and items vary with respect to the grand mean RT.
-subs=ddply(dd, .(level=Subject), 
-summarise, effect="subject", emean=mean(dd$RT_log), lmean=mean(RT_log),
-ldiff=mean(RT_log)-mean(dd$RT_log))
-itms=ddply(dd, .(level=Item), 
-summarise, effect="item", emean=mean(dd$RT_log), lmean=mean(RT_log), 
-ldiff=mean(RT_log)-mean(dd$RT_log))
-si=rbind(subs,itms)
-si$effect <- factor(si$effect, levels=c("subject", "item"))
-my1 <- xyplot(data=si, lmean~level | effect, 
-aspect=1,
-layout=c(2,1),
-ylim = extendrange(si$lmean, f = 0.10),
-par.settings=list(trellis.par.set(superpose.symbol = list(col='black'), 
-superpose.line=list(col='black'))),
-xlab=NULL,
-ylab=NULL,
-main='How mean times for individual subjects and items\nvary with respect to the grand mean Log RT',
-mean=unique(si$emean),
-scales=list(x=list(relation="free", at=NULL, 
-labels=list("subject number", "item identifier"), cex=1 ),
-y=list(at=    c(0,7,unique(si$emean),8,9),
-labels=c("0","7","Mean Log RT","8","9"))),
-panel=function(x,y,type,subscripts,...){
-panel.abline(h=unique(si$emean), col="lightgrey",lty=1,...)
-panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, subscripts=subscripts, code=3, length=0, ...)
-panel.xyplot(x,y,type='p', pch=21, cex=.5, col='black', fill='black', subscripts=subscripts,...)
-}) # end of my1
-my2 <- xyplot(data=si, ldiff~level | effect, 
-aspect=1, type='n',
-ylab="Deviation from\nmean Log RT",
-ylim=extendrange(si$ldiff,f=0.10)
-) # creates a second y-axis scale
-print(doubleYScale(my1, my2, 
-add.axis=TRUE, 
-add.ylab2=TRUE, 
-use.style=FALSE))
-@
-
-<<plotMainsa, fig.width=6, fig.height=6, fig.cap='Plot main effects on several transformations', message=FALSE, echo=FALSE, warning=FALSE>>=
-# plot main effects on observed data
-# allow reference to transformation
-temp <- reshape2::melt(dd,
-id.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
-measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log",  "RT_raw"),
-variable.name="transformation",
-value.name="score")
-# allow reference to effect
-temp2 <- reshape2::melt(temp,
-id.vars=c("transformation", "score"),
-measure.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
-variable.name="effect",
-value.name="level")
-# do the plot
-ggplot(temp2, aes(y=score, x=level, group=1, col=effect)) +
-facet_wrap(~transformation + effect, scales="free", shrink=TRUE, ncol=5) +
-stat_summary(fun.data=mean_cl_normal, geom="pointrange") +
-stat_summary(fun.y=mean, geom="line") +
-theme(aspect.ratio=1, axis.text.y=element_blank())
-@
-
-<<mainB, fig.width=6, fig.height=3, fig.cap='Just the main effects', message=FALSE, echo=FALSE>>=
-# prep main effects rts data frame
-vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness")));names(vrts)[3]<-"Levels"
-nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number")));names(nrts)[3]<-"Levels"
-qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Quantity")));names(qrts)[3]<-"Levels"
-orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Order")));names(orts)[3]<-"Levels"
-irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Item")));names(irts)[3]<-"Levels"
-rts=rbind(vrts,nrts,qrts,orts)
-rts$mins=rts$RT_log-rts$ci
-rts$maxs=rts$RT_log+rts$ci
-# plot main effects
-ggplot(rts, aes(y=RT_log, x=Levels, group=1, ymin=mins, ymax=maxs)) +
-geom_line() +
-geom_errorbar(width=.1) +
-geom_point(cex=3) +
-facet_grid(~Effect, scale="free_x") +
-theme_bw() +
-theme(aspect.ratio=1)
-@
-
-<<mainD, fig.width=6, fig.height=3.5, fig.cap='Main effects over item ratios', message=FALSE, echo=FALSE>>=
-nrts <- cbind(Effect="Number",
-xlev=factor(c(1,2)), 
-myshape='a',
-summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Number")))
-names(nrts)[5]<-"Levels"
-
-orts <- cbind(Effect="Order",
-xlev=factor(c(1,2)), 
-myshape='b',
-summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Order")))
-names(orts)[5]<-"Levels"
-
-qrts <- cbind(Effect="Quantity",
-xlev=factor(c(1,2)), 
-myshape='d',
-summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Quantity")))
-names(qrts)[5]<-"Levels"
-
-vrts <- cbind(Effect="Vagueness",
-xlev=factor(c(1,2)),
-myshape='e',
-summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Vagueness")))
-names(vrts)[5]<-"Levels"
-
-rts=rbind(nrts,orts,qrts,vrts)
-rts$mins=rts$RT_log-rts$ci
-rts$maxs=rts$RT_log+rts$ci
-rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
-#rts$myshape=rep(c(21,22,23,24), each=8)
-
-p1=
-ggplot(data=rts, aes(x=item_mean_ratio, y=RT_log, group=Levels)) +
-geom_errorbar(data=droplevels(subset(rts, Effect=='Number')),width=.01, aes(ymin=mins, ymax=maxs)) +
-geom_line(data=droplevels(subset(rts, Effect=='Number'))) +
-geom_point(data=droplevels(subset(rts, Effect=='Number')),
-aes(shape=Levels, fill=Levels), size=3 ) +
-geom_errorbar(data=droplevels(subset(rts, Effect=='Order')), width=.01, aes(ymin=mins, ymax=maxs)) +
-geom_line(data=droplevels(subset(rts, Effect=='Order'))) +
-geom_point(data=droplevels(subset(rts, Effect=='Order')),
-aes(shape=Levels, fill=Levels), size=3  ) +
-geom_errorbar(data=droplevels(subset(rts, Effect=='Quantity')),width=.01, aes(ymin=mins, ymax=maxs)) +
-geom_line(data=droplevels(subset(rts, Effect=='Quantity'))) +
-geom_point(data=droplevels(subset(rts, Effect=='Quantity')),
-aes(shape=Levels, fill=Levels), size=3  ) +
-geom_errorbar(data=droplevels(subset(rts, Effect=='Vagueness')),width=.01, aes(ymin=mins, ymax=maxs)) +
-geom_line(data=droplevels(subset(rts, Effect=='Vagueness'))) +
-geom_point(data=droplevels(subset(rts, Effect=='Vagueness')),
-aes(shape=Levels, fill=Levels), size=3  ) +
-facet_grid(~Effect) +
-scale_shape_manual(values = c(
-"Numeric"=21,
-"Verbal"=21,
-"LtoR"=22,
-"RtoL"=22,
-"Small"=23,
-"Large"=23,
-"Crisp"=24,
-"Vague"=24),
-breaks=c('Numeric','Verbal','LtoR','RtoL','Small','Large','Crisp','Vague'))  +
-scale_fill_manual(values = c(
-"Numeric"="black",
-"Verbal"="white",
-"LtoR"="black",
-"RtoL"="white",
-"Small"="black",
-"Large"="white",
-"Crisp"="black",
-"Vague"="white"),
-breaks=c('Numeric','Verbal','LtoR','RtoL','Small','Large','Crisp','Vague')) +
-scale_x_continuous(minor_breaks=NULL,
-breaks=c(0.5125000, 0.6876471, 0.7691558, 0.8166667),
-labels=abbreviate(c(0.5125000, 0.6876471, 0.7691558, 0.8166667))) +
-theme(aspect.ratio=1,
-legend.position='top',
-axis.text.x=element_text(size=7),
-legend.key=element_rect(fill = 'white'),
-panel.background=element_rect(fill = 'grey96'))
-p1
-@
-
-<<22way, fig.width=6, fig.height=3, fig.cap='2-way interactions over items', message=FALSE, echo=FALSE>>=
-# prep 2 ways over items rts data frame
-rts1 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Number")); names(rts1)[2] <- "F1"; names(rts1)[3] <- "F2"
-rts1$E1 <- "Vagueness"
-rts1$E2 <- "Number"
-rts2 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Quantity")); names(rts2)[2] <- "F1"; names(rts2)[3] <- "F2"
-rts2$E1 <- "Vagueness"
-rts2$E2 <- "Quantity"
-rts3 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Order")); names(rts3)[2] <- "F1"; names(rts3)[3] <- "F2"
-rts3$E1 <- "Vagueness"
-rts3$E2 <- "Order"
-rts4 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Quantity", "Number")); names(rts4)[2] <- "F1"; names(rts4)[3] <- "F2"
-rts4$E1 <- "Quantity"
-rts4$E2 <- "Number"
-rts5 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Order","Number")); names(rts5)[2] <- "F1"; names(rts5)[3] <- "F2"
-rts5$E1 <- "Order"
-rts5$E2 <- "Number"
-rts6 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Quantity","Order")); names(rts6)[2] <- "F1"; names(rts6)[3] <- "F2"
-rts6$E1 <- "Quantity"
-rts6$E2 <- "Order"
-rts <- rbind(rts1,rts2,rts3,rts4,rts5,rts6)
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
-rts$F3 <- factor(paste(rts$F1,rts$F2))
-rts$E1 <- factor(rts$E1, levels=c("Number","Order","Quantity","Vagueness"))
-rts$E2 <- factor(rts$E2, levels=c("Number","Order","Quantity","Vagueness"))
-# The palette with black:
-cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, group=F3, col=F2, pch=F1)) +
-facet_grid(E2~E1, drop=F) +
-geom_point(cex=3) +
-geom_line() +
-theme_bw() +
-theme(aspect.ratio=1, panel.grid=element_blank()) +
-scale_color_manual(name="colour",values=cbbPalette) +
-scale_shape_discrete(name="shape") +
-scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))
-@
-
-<<plotInt1, fig.width=7, fig.height=3.5, fig.cap='Plot vagueness by number interaction over items', echo=FALSE>>=
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item"))
-rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
-dodge = position_dodge(width=0.2)
-ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci,
-group=condition, shape=condition, fill=condition)) +
-geom_line(position=dodge) +
-geom_errorbar(width=.2, position=dodge) +
-geom_point(size=4, position=dodge) +
-scale_shape_manual("",values = c(22, 22, 21, 21)) +
-scale_fill_manual("",values=c("black","white","black","white")) +
-ggtitle("Response time") +
-ylab("RT_RecSqt") +
-xlab("") +
-theme_bw() +
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-legend.key = element_blank(), aspect.ratio=1,
-axis.text.x = element_text(angle = 45, hjust = 1)) +
-facet_grid(~Number)
-@
-
-<<plot3way, fig.width=7, fig.height=5, fig.cap='3 way interactions',echo=FALSE>>=
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Quantity"))
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-p1=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Vagueness, ymin=mins, ymax=maxs, pch=Vagueness)) +
-geom_line() +
-geom_errorbar(width=.1) +
-geom_point(cex=3) +
-facet_grid(~Quantity, scale="free_x") +
-theme_bw() +
-theme(aspect.ratio=1) +
-theme(legend.title=element_blank(), legend.position="top")
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Order"))
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-p2=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Vagueness, ymin=mins, ymax=maxs, pch=Vagueness)) +
-geom_line() +
-geom_errorbar(width=.1) +
-geom_point(cex=3) +
-facet_grid(~Order, scale="free_x") +
-theme_bw() +
-theme(aspect.ratio=1) +
-theme(legend.title=element_blank(), legend.position="top")
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Quantity","Order"))
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-p3=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Quantity, ymin=mins, ymax=maxs, pch=Quantity)) +
-geom_line() +
-geom_errorbar(width=.1) +
-geom_point(cex=3) +
-facet_grid(~Order, scale="free_x") +
-theme_bw() +
-theme(aspect.ratio=1) +
-theme(legend.title=element_blank(), legend.position="top")
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Quantity","Order"))
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-p4=ggplot(rts, aes(y=RT_RecSqt, x=Vagueness, group=Quantity, ymin=mins, ymax=maxs, pch=Quantity)) +
-geom_line() +
-geom_errorbar(width=.1) +
-geom_point(cex=3) +
-facet_grid(~Order, scale="free_x") +
-theme_bw() +
-theme(aspect.ratio=1) +
-theme(legend.title=element_blank(), legend.position="top")
-grid.arrange(nrow=2, p1,p2,p3,p4)
-@
-
-<<vnitems, fig.width=7, fig.height=4, fig.cap='Plot vagueness by number by quantity over items.',echo=FALSE>>=
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item", "Quantity"))
-rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
-dodge = position_dodge(width=0.2)
-ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci,
-group=Vagueness,
-fill=Vagueness)) +
-geom_line(position=dodge) +
-geom_errorbar(width=.2, position=dodge) +
-geom_point(pch=21, size=4, position=dodge) +
-scale_fill_manual("",values=c("black","white")) +
-ggtitle("Response time") +
-ylab("RT_Rec_Sqt") +
-xlab("") +
-theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-panel.background=element_rect(fill="white",color=1),
-legend.key = element_blank(),
-aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust = 1)) +
-facet_grid(~Number+Quantity)
-@
-
-<<4way, fig.width=6, fig.height=3, fig.cap='4 way interaction', echo=FALSE>>=
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness",  "Quantity", "Order"))
-rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
-ggplot(rts, aes(y=RT_RecSqt, x=Vagueness, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, group=Quantity, pch=Quantity)) +
-facet_grid(~Order+Number, scale="free_x") +
-geom_line() +
-geom_errorbar(width=.1) +
-geom_point(cex=3) +
-theme_bw() +
-theme(aspect.ratio=1) +
-theme(legend.title=element_blank())
-@
-
-<<4waysplit, fig.width=7, fig.height=7, fig.cap='4 way interaction split over items',echo=FALSE>>=
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness", "Item", "Quantity", "Order"))
-rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
-ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci,
-group=Vagueness,fill=Vagueness)) +
-geom_line() +
-geom_errorbar(width=.12) +
-geom_point(pch=21, size=2) +
-scale_fill_manual("",values=c("black","white","black","white")) +
-ggtitle("Number * Vagueness * Quantity * Order") +
-ylab("RT_RecSqt") +
-xlab("") +
-theme_bw() +
-theme( legend.key = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1),
-axis.text.x = element_text(angle = 45, hjust = 1), aspect.ratio=1) +
-facet_grid(Vagueness+Order~Number+Quantity)
-@
-
-<<4waySplitRatio, fig.width=7, fig.height=7, fig.cap='4 way interaction split over items showing ratio of item dots.',echo=FALSE, message=FALSE, warning=FALSE>>=
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness", "item_mean_ratio", "Quantity", "Order"))
-rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
-rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
-#dodge = position_dodge(width=0.2)
-ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci,
-group=Vagueness,fill=Vagueness)) +
-# geom_line(position=dodge) +
-geom_line() +
-geom_errorbar(width=.02) +
-geom_point(pch=21, size=2) +
-scale_fill_manual("",values=c("black","white","black","white")) +
-ggtitle("Number * Vagueness * Quantity * Order") +
-ylab("RT_RecSqt") +
-xlab("") +
-theme_bw() +
-theme( legend.key = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1),
-axis.text.x = element_text(angle = 45, hjust = 1), aspect.ratio=1) +
-facet_grid(Vagueness+Order~Number+Quantity)
-@
-
-<<showCompression, fig.width=7, fig.height=7, fig.cap='Show compression of item ratios', echo=FALSE>>=
-#  Show compression of ratios of numbers of dots in the items
-item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54)
-item_range_ratio_equally_spaced = seq(from = 6 / 24, to = 36 / 54, length.out = 4)
-item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54)))
-item_mean_ratio_equally_spaced = seq(from = mean(c(6 / 15, 15 / 24)), to = mean(c(36 / 45, 45 / 54)), length.out = 4)
-plot(y=c(2,2,2,2), x=item_range_ratio, xlim=c(0,1), ylim=c(0,3), ylab="", yaxt='n', col="blue", pch=21, bg='blue', type='o', xaxt='n', xlab="ratio: small is easily discriminable", main="Discriminability of each item,\n expressed as ratios between array magnitudes")
-text(y=c(2.15,2.15,2.15,2.15), x=item_range_ratio, labels=round(item_range_ratio,2), col="blue")
-text(y=2.15, x=0.1, labels="ratio min/max", col='blue')
-points(y=c(2.3,2.3,2.3,2.3), x=item_range_ratio_equally_spaced, col="lightblue", pch=21, bg='lightblue', type='o')
-points(y=c(1,1,1,1), x=item_mean_ratio, pch=21, col='red', bg='red', type='o')
-text(y=c(1.15,1.15,1.15,1.15), x=item_mean_ratio, labels=round(item_mean_ratio,2), col="red")
-text(y=1.15, x=0.2, labels="mean (ratio min/med,\n            ratio med/max)", col='red')
-points(y=c(1.3,1.3,1.3,1.3), x=item_mean_ratio_equally_spaced, col="pink", pch=21, bg='pink', type='o')
-axis(1, at=seq(0,1,0.1))
-@
-
-\FloatBarrier
-\section{Lmer model}
-
-<<rmme>>=
-load("data_processed.Rda")
-@
-
-<<lmer5, cache=TRUE>>=
-v5 <- lmerTest::lmer(data=dd,
-RT_log ~
-c_Vag + c_Num + c_Qty + c_Ord +
-c_Num:c_Vag:c_Qty +
-item_mean_ratio +
-s_Trl +
-RTprev_log +
-nchar_instr +
-(1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
-v5b <- lme4::lmer(data=dd,
-RT_log ~
-c_Vag + c_Num + c_Qty + c_Ord +
-c_Num:c_Vag:c_Qty +
-item_mean_ratio +
-s_Trl +
-RTprev_log +
-nchar_instr +
-(1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
-@
-
-<<v51, cache=TRUE>>=
-summary(v5)
-@
-
-<<v511, results='asis', cache=TRUE>>=
-print(xtable(coef(summary(v5b))))
-@
-
-<<v52>>=
-cat("R^2")
-cor(fitted(v5), dd$RT_RecSqt)^2
-@
-
-<<plotLMERfnc, fig.width=7, fig.height=3.5, fig.cap='caption'>>=
-par(mfrow=c(3,4), family='cmodern')
-plotLMER.fnc(v5)
-@
-
-Plot model coefficients and ci's
-
-<<v54, fig.width=7, fig.height=4, fig.cap='Estimates plot'>>=
-e=data.frame(coef=summary(v5)$coef[-1,1]) # estimates, without intercept
-q=as.data.frame(confint(v5, method='Wald')[18:26,1:2])
-eq=cbind(e,q)
-y=eq[rev(rownames(eq)),]
-par(mar=c(4,10,1,1))
-plot(y=1:nrow(y), x=y$coef, xlim=range(y$"2.5 %", y$"97.5 %"), type='n', axes=F, xlab="", ylab="")
-segments(y0=1:nrow(y), y1=1:nrow(y), x0=y$"2.5 %", x1=y$"97.5 %", lwd=.55)
-points(y=1:nrow(y), x=y$coef, pch=20, cex=.75)
-abline(v=0, lty=3, col='grey')
-axis(2, labels=row.names(y), at=1:nrow(y), las=1)
-axis(1, las=1)
-mtext("estimate", side=1, line=2.5, cex.lab=1, las=1)
-box(col='grey')
-@
-
-Model criticism baayen plots
-
-<<baayenPLots99, fig.width=9, fig.height=3, fig.cap='caption'>>=
-# Baayen 4-plot model criticism
-par(mfrow=c(1,4), pty='s')
-# create scaled residuals
-dd$rstand = as.vector(scale(resid(v5)))
-# plot scaled residuals density
-plot(density(dd$rstand))
-# plot sample quantiles versus theoretical quantiles
-qqnorm(dd$rstand, cex=.5)
-qqline(dd$rstand)
-# plot standardised residuals versus fitted values
-plot(dd$rstand ~ fitted(v5), pch='.')
-# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
-abline(h=c(-2.5,2.5))
-@
+\caption{Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions.}
+\end{figure}
 

### 2016-08-10 experiments/e2/writeup/scratch.R

@@ -292,3 +292,6 @@ print(v_1, B=0)
 xtabs(~rowSums(attr(v_1,"kept")))
 ```
 
+
+
+  

### 2016-08-12 experiments/e2/writeup/e2-b-knitr.Rnw
more consistency, better plots, using discriminability instead of item ratio

@@
-\documentclass[a4paper]{article}
-\usepackage{geometry}
+\documentclass[a4paper,12pt]{article}
+\usepackage[margin=2cm]{geometry}
 \usepackage{placeins}
+
+\title{Exp 2 of 4}
 \begin{document}
 
-<<loadKnitr, echo=FALSE, message=FALSE>>=
+<<setup, include=FALSE, cache=FALSE>>=
 library(knitr)
-opts_chunk$set(cache=FALSE, autodep = TRUE)
+options(formatR.arrow=TRUE, 
+        width=100)
+opts_chunk$set(fig.path='figure/graphics-', 
+               cache.path='cache/graphics-', 
+               fig.align='center', 
+               dev='pdf', 
+               fig.width=5, 
+               fig.height=5, 
+               fig.show='hold', 
+               cache=FALSE, 
+               par=TRUE,
+               size='scriptsize',
+               tidy=TRUE,
+               comment = NA,
+               tidy.opts = list(blank=FALSE, 
+                                width.cutoff=100))
+knit_hooks$set(crop=hook_pdfcrop)
 @
 
-<<defineFonts>>=
-# .define.fonts <- function () {
-#     quartzFonts(avenir =    c('Avenir Book', 'Avenir Black', 'Avenir Book Oblique', 'Avenir Black Oblique'),
-#                 helvetica = c('Helvetica New Light', 'Helvetica New Bold', 'Helvetica New Light Italic', 'Helvetica New Bold Italic'),
-#                 cmodern = c('CMU Serif', 'CMU Serif Extra','CMU Classical Serif','CMU Serif Extra'))
-# }
-# .define.fonts()
+\maketitle
+\tableofcontents
+\clearpage
+
+<<loadFonts, message=FALSE, echo=FALSE>>=
+library(fontcm)
+library(extrafont)
+#font_import()
+#loadfonts()
 @
 
-<<loadFonts>>=
-library(extrafont)
-loadfonts()
-@
-
-<<loadLibs, echo=FALSE, message=FALSE>>=
+<<loadLibs, message=FALSE, echo=FALSE>>=
 library(beepr)                     #
@@
 
-\section{Get the data}
-
-<<getTheData, echo=FALSE>>=
-# if the file data.Rda already exists then load it, else do data wrangling
-if( file.exists('data_processed.Rda') ) {
-	load('data_processed.Rda')
-	} else {
+<<gatherDataFunction, echo=FALSE>>=
 gatherData = function(number_of_valid_subjects) {
@@
   }
+  message('Returning gathered data')
   return(gathered.data)
 } # end of gatherData function
-	classifyResponse = function(dat) {
+@
+
+<<classifyResponseFuntion, echo=FALSE>>=
+classifyResponses = function(dat) {
   # what were they expected to respond?
@@
            'RIGHT' = {dat[row, 'response_num'] <- dat[row, 'Right']},
-						 'NOANSWER' = {dat[row, 'response_num'] <- NA}
-			)
-		}
+           'NOANSWER' = {dat[row, 'response_num'] <- NA})}
   # what side was the subject's button-press? Left, mid right?
@@
   dat$RESPONSE <- NULL
+  message('Returning classified data')
   return(dat)
 } # end of classifyResponse function
+@
 
-	# declare local variables
-	number_of_valid_subjects <- 30 # = 30
-	number_of_rows <- 7680 # 7680
-	number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # = 256
-	dat <- gatherData(number_of_valid_subjects) # = 30
-	dat <- classifyResponse(dat) # classify the response as expected, near, or far
-
+<<processDataFunction, echo=FALSE>>=
+processData = function(dat) {
   # SUBJECT
@@
   # id is a unique identifier for the 7680 row data
-	dat$id <- factor(paste(paste(dat$Subject),
-												 paste("t", sprintf("%03d", dat$Trial), sep="") , sep=":"))
+  dat$id <- factor(paste(paste(dat$Subject), paste("t", sprintf("%03d", dat$Trial), sep="") , sep=":"))
   # ITEM
   # create a centred numeric item variable for modeling
-	dat$c_Itm <- ifelse(dat$Item==1, -.75,
-											ifelse(dat$Item==2, -.25,
-														 ifelse(dat$Item==3, .25, .75)))
+  dat$c_Itm <- ifelse(dat$Item==1, -.75, ifelse(dat$Item==2, -.25, ifelse(dat$Item==3, .25, .75)))
   # make Item be a factor and assign labels
   dat$Item <- factor(dat$Item, levels=c(1,2,3,4), labels=c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
-	# add ratios for Item
-	item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54)
-	# 0.2500000 0.4705882 0.5909091 0.6666667
-	item_range_ratio_scaled = as.vector(scale(c(6 / 24, 16 / 34, 26 / 44, 36 / 54)))
-	# -1.3441995 -0.1316642  0.5297187  0.9461450
-	item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54)))
-	# 0.5125000 0.6876471 0.7691558 0.8166667
-	item_mean_ratio_scaled = as.vector(scale(c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54)))))
-	# -1.37582241 -0.06614191  0.54334858 0.89861574
-	dat[dat$Item == "06:15:24", "item_range_ratio"] <-  0.2500000
-	dat[dat$Item == "16:25:34", "item_range_ratio"] <-  0.4705882
-	dat[dat$Item == "26:35:44", "item_range_ratio"] <-  0.5909091
-	dat[dat$Item == "36:45:54", "item_range_ratio"] <-  0.6666667
-	dat[dat$Item == "06:15:24", "item_range_ratio_scaled"] <-  -1.3441995
-	dat[dat$Item == "16:25:34", "item_range_ratio_scaled"] <-  -0.1316642
-	dat[dat$Item == "26:35:44", "item_range_ratio_scaled"] <-   0.5297187
-	dat[dat$Item == "36:45:54", "item_range_ratio_scaled"] <-   0.9461450
-	dat[dat$Item == "06:15:24", "item_mean_ratio"] <-  0.5125000
-	dat[dat$Item == "16:25:34", "item_mean_ratio"] <-  0.6876471
-	dat[dat$Item == "26:35:44", "item_mean_ratio"] <-  0.7691558
-	dat[dat$Item == "36:45:54", "item_mean_ratio"] <-  0.8166667
-	dat[dat$Item == "06:15:24", "item_mean_ratio_scaled"] <-   -1.37582241
-	dat[dat$Item == "16:25:34", "item_mean_ratio_scaled"] <-   -0.06614191
-	dat[dat$Item == "26:35:44", "item_mean_ratio_scaled"] <-    0.54334858
-	dat[dat$Item == "36:45:54", "item_mean_ratio_scaled"] <-    0.89861574
   # VAGUENESS
@@
   # add transformations of RT
-	dat$RT_RecSqd    <- 1/dat$RT^2
-	dat$RT_Rec       <- 1/dat$RT
-	dat$RT_RecSqt    <- 1/sqrt(dat$RT)
   dat$RT_log       <- log(dat$RT)
-	dat$RT_sqt       <- sqrt(dat$RT)
   dat$RT_raw       <- dat$RT
-	dat$RT_sqd       <- dat$RT^2
   # print to file a table with information about the design
-	design_info <- unique(subset(dat, select=c(Item, Condition, Vagueness, Number, Quantity, Order,
-																						 Left, Mid, Right, Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side,Instruction)))
+  design_info <- unique(subset(dat, select=c(Item, Condition, Vagueness, Number, Quantity, Order, Left, Mid, Right, Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, Instruction)))
   design.info <- design_info[order(design_info$Item, design_info$Condition, design_info$Quantity, design_info$Order),]
   row.names(design_info) <- NULL
-	capture.output(print.data.frame(design_info, row.names=F, print.gap=3, quote=F, right=F),
-								 file="design_info.table")
+  capture.output(print.data.frame(design_info, row.names=F, print.gap=3, quote=F, right=F), file="design_info-table.txt")
+  # Add discriminability metric with reference to item
+  discriminability_range = c(0.7500000, 0.5294118, 0.4090909, 0.3333333)
+  discriminability_range_scaled = c(1.3441995, 0.1316642, -0.5297187, -0.9461450)  
+  discriminability = c(0.4875000, 0.3123529, 0.2308442, 0.1833333)
+  discriminability_scaled = c(1.37582241, 0.06614191, -0.54334858, -0.89861574)
+  dat[dat$Item == "06:15:24", "discriminability"] <- 0.4875000
+  dat[dat$Item == "16:25:34", "discriminability"] <- 0.3123529
+  dat[dat$Item == "26:35:44", "discriminability"] <- 0.2308442
+  dat[dat$Item == "36:45:54", "discriminability"] <- 0.1833333
   # put dat in better column order
-	dat <- subset(dat,
-								select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number,
-													 Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,
+  dat <- subset(dat, select = c(id, Subject, Trial, Condition, Order, 
+                           Quantity, Vagueness, Number, Item, discriminability, 
                            c_Trl, s_Trl, c_Itm, c_Vag, c_Num, f_Cnd, c_Ord, c_Qty,
-													 RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
-													 Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
-													 Instruction, nchar_instr
-								))
-	# This data set (dat) contains *all* trials 7680 including impossible trials and is
-	# mainly for graphs comparing different removals
+                           RT, RT_log, RT_raw, Exp_Num, Bline_Num, 
+                           Extr_Num, Exp_side, Bline_side, Extr_side, 
+                           response_num, response_side, response_category, 
+                           Left, Mid, Right, Instruction, nchar_instr))
+  # This data set (dat) contains *all* trials 7680 including impossible trials and is mainly for graphs comparing different removals
   save(dat, file='data_raw.Rda')
-	# remove impossible trials from dat in new data dd #
-	# Throw out RT = 1 and RT = 59998, and RTprev = 1 and RTprev = 59998
-	# i.e., throw out sticky fingers and timeouts,
-	# and the trials that followed sticky fingers and timeouts
-	# since they were likely affected by unusual previous trials.
-	# lose impossible trials
+  message('Returning processed data')
+  return(dat)
+} # end of function processData
+@
+
+<<postProcessDataFunction, echo=FALSE>>=
+postProcessData = function(dat) {
+  # dd removes impossible trials from dat
+  # Throw out RT = 1 and RT = 59998, and RTprev = 1 and RTprev = 59998 i.e., throw out sticky fingers and timeouts, and the trials that followed sticky fingers and timeouts since they were likely affected by unusual previous trials.
+Also lose impossible trials
   dd <- dat
@@
     for (i in 1:nrows) {
-			if (i==1) {
-				dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]
-			}
+      if (i==1) {dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]}
       else
-				dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1]
-		}
-	}
+        dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1] }}
   # add transformations of previous RT
-	dd$RTprev_RecSqd    <- 1/dd$RTprev^2
-	dd$RTprev_Rec       <- 1/dd$RTprev
-	dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev)
-	dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
   dd$RTprev_log       <- log(dd$RTprev)
-	dd$RTprev_sqt       <- sqrt(dd$RTprev)
   dd$RTprev_raw       <- dd$RTprev
-	dd$RTprev_sqd       <- dd$RTprev^2
-	# number of higher-level cells in the design (abstracting over Order) =
-	# Item(4) * Number(2) * Vagueness(2) * Quantity(2) = 32
-	# number of measurements in eaach cell = 8
-	# 332 * 8 = 256
-	# 256 * 30 = 7680
-	dd$measurement=0
-	dd$cell=0
-	for (s in levels(dd$Subject)) {
-		cellcount=0
-		for (i in levels(dd$Item)) {
-			for (n in levels(dd$Number)) {
-				for (v in levels(dd$Vagueness)) {
-					for (q in levels(dd$Quantity)) {
-						measurementcount=0
-						cellcount=cellcount+1
-						for (t in dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q, "Trial"]) {
-							dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "cell"] <- cellcount
-							measurementcount=measurementcount+1
-							dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "measurement"] = measurementcount
-						}
-					}
-				}
-			}
-		}
-	}
   # put dd in better column order
-	dd <- subset(dd,
-							 select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number,
-							 					 Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,
-							 					 c_Trl, s_Trl, c_Itm, c_Vag, c_Num, f_Cnd, c_Ord, c_Qty,
-							 					 RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
-							 					 RTprev, RTprev_RecSqd, RTprev_Rec, RTprev_RecSqt, RTprev_log, RTprev_sqt, RTprev_raw, RTprev_sqd,
-							 					 Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
-							 					 Instruction, nchar_instr,
-							 					 cell, measurement
-							 ))
+  dd <- subset(dd, select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number, Item, discriminability, c_Trl, s_Trl, c_Itm, c_Vag, c_Num, f_Cnd, c_Ord, c_Qty, RT, RT_log, RT_raw, RTprev, RTprev_log, RTprev_raw, Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right, Instruction, nchar_instr ))
   save(dd, file="data_processed.Rda")
-} # end of the else clause, which is run if the file data.Rda does not exist.
+  message('Returning post processed data')
+  return(dd)
+} # end of function postProcessResponses
 @
 
+\clearpage
+\section{Get the data}
+
+<<getTheData>>=
+# if the file data_processed.Rda already exists then load it, else do data wrangling
+if( file.exists('data_processed.Rda') ) {
+  load('data_processed.Rda')
+} else {
+  # declare local variables
+  number_of_valid_subjects <- 30 # = 30
+  number_of_rows <- 7680 # 7680
+  number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # = 256
+  # call functions
+  dat <- gatherData(number_of_valid_subjects) # = 30
+  dat <- classifyResponses(dat) # classify the response as expected, near, or far
+  dat <- processData(dat) # remove impossible trials and re-do previous rt measures
+  dd  <- postProcessData(dat)
+} # end else do data wrangling
+@
+
+<<showDataNames>>=
+names(dd)
+@
+
+<<showDataHead>>=
+head(dd)
+@
+
+<<showDataSummary>>=
+summary(dd)
+@
+
+<<showDataStructure>>=
+str(dd)
+@
+
+\clearpage
 \section{Plots}
+\clearpage
+\subsection{Discriminability}
+
+<<showCompression, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='Ratios for different numbers of dots in the arrays: smaller values are more discriminable.
+Blue is for the ratio between the smallest number in the array and the largest number in an array.
+Red is for the mean of two ratios, one for the smallest number to the middle number, the other for the middle number to the largest number in the array', echo=FALSE>>=
+#  Show compression of ratios of numbers of dots in the items
+discriminability = sort(unique(as.numeric(dd$discriminability)))
+plot(y=c(1.5,1.5,1.5,1.5), 
+     x=discriminability, 
+     ylab='',
+     xlab='',
+     yaxt='n',
+     type='n', 
+     xaxt='n',
+     axes=F)
+lines(y=c(1.5,1.5,1.5,1.5), x=discriminability, type='o', pch=19)
+text(y=c(1.15,1.15,1.15,1.15), x=discriminability, labels=round(discriminability,2),  cex=.7)
+mtext('mean (ratio min/med,\nratio med/max)', side = 2, line = 1)
+mtext('Discriminability index', side=1, line=3)
+mtext('Discriminability for each item', side=3, line=1, cex=1.5)
+axis(1)
+box(col='grey')
+#text(y=1.00, x=0.07, labels="mean\n(ratio min/med,\nratio med/max)", col='red')
+# axis(1, at=seq(-.1, 1,0.1)[4:11])
+# axis(1)
+@
+
+\clearpage
 
@@
 }# end of create normally distributed equivalents for the data transforms.
+@
+
+<<plotDistFreqPolys, fig.width=6, fig.height=4, fig.cap='Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions', fig.pos='hbtp', fig.align='center', message=FALSE, warning=FALSE, echo=FALSE>>=
 temp <- reshape2::melt(dat_transforms, 
@@
                        value.name="time")
+ggplot(temp, aes(time, colour=RTtype)) + 
+  geom_freqpoly() + 
+  facet_wrap(transformation~dfSubset, scales='free', ncol=4, nrow=1) + 
+  scale_color_manual('RT Type: ', values=c("blue","red")) + 
+  theme_bw() + 
+  theme(aspect.ratio=1, 
+        text=element_text(family='CMU Serif', size=10, face='plain'),
+        axis.title.x=element_blank(),
+        axis.text.x=element_text(angle=25),
+        panel.grid=element_blank(),
+        legend.position='top',
+        legend.key=element_rect(color = 'white'))
 @
 
-<<plotDists, fig.width=5, fig.height=5, fig.cap='Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions', fig.pos='hbtp', fig.align='right', message=FALSE, warning=FALSE, echo=FALSE>>=
-ggplot(temp, aes(time, colour=RTtype)) + 
-  geom_freqpoly() + 
-  facet_wrap(transformation~dfSubset, scales='free') + 
-  scale_color_manual(values=c("blue","red")) + 
-  theme_bw() + 
-  theme(aspect.ratio=1, text=element_text(family='CMU Serif', size=12, face='italic'))
-@
+\clearpage
 
-<<exampleOfFonts, fig.width=3, fig.height=3, fig.cap='font example', echo=FALSE>>=
-p <- ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point() +
-ggtitle("Fuel Efficiency of 32 Cars") +
-xlab("Weight (x1000 lb)") + ylab("Miles per Gallon") +
-theme_bw() +
-theme(text=element_text(family="CMU Serif", size=14, face='plain'))
-p
-ggsave("ggplot_CMU_Serif.pdf", p, width=3.5, height=3.5)
-@
-
-<<boxPlots, fig.width=7, fig.height=3, fig.cap='Show how transformations of RT affect distribution of times, and how they affect which times are outliers.', echo=FALSE>>=
+<<plotBoxPlots, fig.width=6, fig.height=4, fig.cap='Show how transformations of RT affect distribution of times, and how they affect which times are outliers.', fig.pos='hbtp', echo=FALSE>>=
 load('data_processed.Rda')
@@
 ggplot(subdata) + 
-  facet_wrap(~ transformation, scales="free", ncol=4) +
+  facet_wrap(~ transformation, scales="free", ncol=2, nrow=1) +
   geom_boxplot(aes(y=value, x=Item, col="Items"), outlier.colour = "black", outlier.shape = 21, outlier.size=0.1) + 
@@
   theme_bw() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=5), aspect.ratio=1, panel.grid=element_blank()) + 
+  theme(aspect.ratio=1, 
+        text=element_text(family='CMU Serif', size=10, face='plain'),
+        panel.grid=element_blank(),
+        axis.ticks.x=element_blank(),
+        axis.text.x=element_blank(),
+        legend.position='top',
+        legend.key=element_rect(color = 'white')) + 
   scale_colour_manual("",values=c("Items"="orange","Subjects"="blue")) +
@@
 @
-  <<fastSlowRT, fig.width=7, fig.height=3, fig.cap='Show how mean times for individual subjects and items vary with respect to the grand mean Log RT.', message=FALSE, echo=FALSE>>=
+
+\clearpage
+
+<<plotFastSlowRTs, fig.width=6, fig.height=4, fig.cap='Show how mean times for individual subjects and items vary with respect to the grand mean Log RT.', fig.pos='hbtp', message=FALSE, echo=FALSE>>=
 # Show how mean times for individual subjects and items vary with respect to the grand mean RT.
@@
            ldiff=mean(RT_log)-mean(dd$RT_log))
-itms=ddply(dd, .(level=Item), 
-summarise, effect="item", emean=mean(dd$RT_log), lmean=mean(RT_log), 
+subs$level=1:30
+subs$effect <- factor(subs$effect, levels=c("subject", "discriminability"))
+itms=ddply(dd, .(level=discriminability), 
+           summarise, effect="discriminability", emean=mean(dd$RT_log), lmean=mean(RT_log), 
            ldiff=mean(RT_log)-mean(dd$RT_log))
+itms$effect <- factor(itms$effect, levels=c("subject", "discriminability"))
 si=rbind(subs,itms)
-si$effect <- factor(si$effect, levels=c("subject", "item"))
+si$effect <- factor(si$effect, levels=c("subject", "discriminability"))
 my1 <- xyplot(data=si, lmean~level | effect, 
@@
               xlab=NULL,
-ylab=NULL,
-main='How mean times for individual subjects and items\nvary with respect to the grand mean Log RT',
+              ylab='Log RT',
+              #main='How mean times for individual subjects and items\nvary with respect to the grand mean Log RT',
               mean=unique(si$emean),
@@
                                  labels=list("subject number", "item identifier"), cex=1 ),
-y=list(at=    c(0,7,unique(si$emean),8,9),
-labels=c("0","7","Mean Log RT","8","9"))),
+                          y=list(at = c(0, 7, unique(si$emean), 8, 9),
+                                 labels=c("0", "7", "Mean", "8", "9"))),
               panel=function(x,y,type,subscripts,...){
@@
                 panel.xyplot(x,y,type='p', pch=21, cex=.5, col='black', fill='black', subscripts=subscripts,...)
-}) # end of my1
+              }
+) # end of my1
 my2 <- xyplot(data=si, ldiff~level | effect, 
@@
 
-<<plotMainsa, fig.width=6, fig.height=6, fig.cap='Plot main effects on several transformations', message=FALSE, echo=FALSE, warning=FALSE>>=
+\clearpage
+
+<<plotMainsa, fig.width=6, fig.height=4, fig.cap='Plot main effects on several transformations', fig.pos='hbtp', message=FALSE, echo=FALSE, warning=FALSE>>=
 # plot main effects on observed data
@@
                        id.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
-measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log",  "RT_raw"),
+                       measure.vars=c("RT_log",  "RT_raw"),
                        variable.name="transformation",
@@
 ggplot(temp2, aes(y=score, x=level, group=1, col=effect)) +
-facet_wrap(~transformation + effect, scales="free", shrink=TRUE, ncol=5) +
+  facet_wrap(~transformation + effect, scales="free", ncol=5) +
   stat_summary(fun.data=mean_cl_normal, geom="pointrange") +
   stat_summary(fun.y=mean, geom="line") +
-theme(aspect.ratio=1, axis.text.y=element_blank())
+  theme_bw() +
+  theme(aspect.ratio=1, 
+        axis.text.y=element_blank(),
+        axis.ticks.y=element_blank(),
+        axis.text.x = element_text(angle = 45, hjust = 1, size=8),
+        legend.position="none") +
+  ylab('Response latency\n') + xlab('')
 @
 
-<<mainB, fig.width=6, fig.height=3, fig.cap='Just the main effects', message=FALSE, echo=FALSE>>=
+\clearpage
+
+<<mainB, fig.width=6, fig.height=4, fig.cap='Just the main effects', fig.pos='hbtp', message=FALSE, echo=FALSE>>=
 # prep main effects rts data frame
@@
   geom_line() +
-geom_errorbar(width=.1) +
-geom_point(cex=3) +
+  geom_errorbar(width=.05) +
+  geom_point(pch=19) +
   facet_grid(~Effect, scale="free_x") +
   theme_bw() +
-theme(aspect.ratio=1)
+  theme(aspect.ratio=1,
+        axis.text.x = element_text(size=8),
+        axis.text.y = element_text(size=8, angle=15),
+        panel.grid=element_blank()) +
+  ylab("RT (log)") + xlab('')
 @
 
-<<mainD, fig.width=6, fig.height=3.5, fig.cap='Main effects over item ratios', message=FALSE, echo=FALSE>>=
+\clearpage
+
+<<mainD, fig.width=6, fig.height=5, fig.cap='Main effects over item ratios', fig.pos='hbtp', message=FALSE, echo=FALSE>>=
 nrts <- cbind(Effect="Number",
@@
               myshape='a',
-summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Number")))
+              summarySEwithin(dd, measurevar="RT_log", withinvars=c("discriminability","Number")))
 names(nrts)[5]<-"Levels"
@@
               myshape='b',
-summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Order")))
+              summarySEwithin(dd, measurevar="RT_log", withinvars=c("discriminability","Order")))
 names(orts)[5]<-"Levels"
@@
               myshape='d',
-summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Quantity")))
+              summarySEwithin(dd, measurevar="RT_log", withinvars=c("discriminability","Quantity")))
 names(qrts)[5]<-"Levels"
@@
               myshape='e',
-summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Vagueness")))
+              summarySEwithin(dd, measurevar="RT_log", withinvars=c("discriminability","Vagueness")))
 names(vrts)[5]<-"Levels"
@@
 rts$maxs=rts$RT_log+rts$ci
-rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
+rts$discriminability <- as.numeric(as.character(rts$discriminability))
 #rts$myshape=rep(c(21,22,23,24), each=8)
@@
 p1=
-ggplot(data=rts, aes(x=item_mean_ratio, y=RT_log, group=Levels)) +
+  ggplot(data=rts, aes(x=discriminability, y=RT_log, group=Levels)) +
   geom_errorbar(data=droplevels(subset(rts, Effect=='Number')),width=.01, aes(ymin=mins, ymax=maxs)) +
@@
   scale_x_continuous(minor_breaks=NULL,
-breaks=c(0.5125000, 0.6876471, 0.7691558, 0.8166667),
-labels=abbreviate(c(0.5125000, 0.6876471, 0.7691558, 0.8166667))) +
+                     breaks=1-c(0.5125000, 0.6876471, 0.7691558, 0.8166667),
+                     labels=abbreviate(1-c(0.5125000, 0.6876471, 0.7691558, 0.8166667))) +
+  theme_bw() +
   theme(aspect.ratio=1,
         legend.position='top',
-axis.text.x=element_text(size=7),
-legend.key=element_rect(fill = 'white'),
-panel.background=element_rect(fill = 'grey96'))
+        axis.text.x=element_text(size=8, angle=35),
+        axis.text.y=element_text(size=8, angle=15),
+        legend.key=element_rect(color = 'white')
+        #,        panel.grid=element_blank()
+        ) +
+  ylab("RT (log)") + xlab("Item discriminability measure")
 p1
@@
 
-<<22way, fig.width=6, fig.height=3, fig.cap='2-way interactions over items', message=FALSE, echo=FALSE>>=
+\clearpage
+
+<<22way, fig.width=7, fig.height=7, fig.cap='2-way interactions over item ratios', fig.pos='hbtp', message=FALSE, echo=FALSE>>=
 # prep 2 ways over items rts data frame
-rts1 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Number")); names(rts1)[2] <- "F1"; names(rts1)[3] <- "F2"
+rts1 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("discriminability", "Vagueness","Number")); names(rts1)[2] <- "F1"; names(rts1)[3] <- "F2"
 rts1$E1 <- "Vagueness"
 rts1$E2 <- "Number"
-rts2 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Quantity")); names(rts2)[2] <- "F1"; names(rts2)[3] <- "F2"
+rts2 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("discriminability", "Vagueness","Quantity")); names(rts2)[2] <- "F1"; names(rts2)[3] <- "F2"
 rts2$E1 <- "Vagueness"
 rts2$E2 <- "Quantity"
-rts3 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Order")); names(rts3)[2] <- "F1"; names(rts3)[3] <- "F2"
+rts3 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("discriminability", "Vagueness","Order")); names(rts3)[2] <- "F1"; names(rts3)[3] <- "F2"
 rts3$E1 <- "Vagueness"
 rts3$E2 <- "Order"
-rts4 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Quantity", "Number")); names(rts4)[2] <- "F1"; names(rts4)[3] <- "F2"
+rts4 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("discriminability", "Quantity", "Number")); names(rts4)[2] <- "F1"; names(rts4)[3] <- "F2"
 rts4$E1 <- "Quantity"
 rts4$E2 <- "Number"
-rts5 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Order","Number")); names(rts5)[2] <- "F1"; names(rts5)[3] <- "F2"
+rts5 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("discriminability", "Order","Number")); names(rts5)[2] <- "F1"; names(rts5)[3] <- "F2"
 rts5$E1 <- "Order"
 rts5$E2 <- "Number"
-rts6 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Quantity","Order")); names(rts6)[2] <- "F1"; names(rts6)[3] <- "F2"
+rts6 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("discriminability", "Quantity","Order")); names(rts6)[2] <- "F1"; names(rts6)[3] <- "F2"
 rts6$E1 <- "Quantity"
@@
 rts <- rbind(rts1,rts2,rts3,rts4,rts5,rts6)
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
+rts$mins=rts$RT_log-rts$ci
+rts$maxs=rts$RT_log+rts$ci
+rts$discriminability <- as.numeric(as.character(rts$discriminability))
 rts$F3 <- factor(paste(rts$F1,rts$F2))
@@
 cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, group=F3, col=F2, pch=F1)) +
+ggplot(rts, aes(y=RT_log, x=discriminability, group=F3, col=F2, pch=F1)) +
   facet_grid(E2~E1, drop=F) +
-geom_point(cex=3) +
+  geom_point() +
   geom_line() +
   theme_bw() +
-theme(aspect.ratio=1, panel.grid=element_blank()) +
+  theme(
+    aspect.ratio=1
+    #,  panel.grid=element_blank()
+    , axis.text.x=element_text(size=7, angle=45,hjust=1)
+    , axis.text.y=element_text(size=7)
+    ) +
   scale_color_manual(name="colour",values=cbbPalette) +
   scale_shape_discrete(name="shape") +
-scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))
+  #scale_x_continuous(breaks = round(sort(unique(rts$discriminability)),1)) +
+  scale_x_continuous(minor_breaks=NULL,
+                     breaks=1-c(0.5125000, 0.6876471, 0.7691558, 0.8166667),
+                     labels=abbreviate(1-c(0.5125000, 0.6876471, 0.7691558, 0.8166667))) 
 @
 
-<<plotInt1, fig.width=7, fig.height=3.5, fig.cap='Plot vagueness by number interaction over items', echo=FALSE>>=
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item"))
+\clearpage
+
+<<plotInt1, fig.width=6, fig.height=4, fig.cap='vagueness by number interaction over items', fig.pos='hbtp', message=FALSE, echo=FALSE>>=
+rts <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number","Vagueness","Item"))
 rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
 dodge = position_dodge(width=0.2)
-ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci,
+ggplot(rts, aes(y=RT_log, x=Item, ymin=RT_log-ci, ymax=RT_log+ci,
                 group=condition, shape=condition, fill=condition)) +
@@
   ggtitle("Response time") +
-ylab("RT_RecSqt") +
+  ylab("RT_log") +
   xlab("") +
@@
 
-<<plot3way, fig.width=7, fig.height=5, fig.cap='3 way interactions',echo=FALSE>>=
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Quantity"))
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-p1=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Vagueness, ymin=mins, ymax=maxs, pch=Vagueness)) +
+\clearpage
+
+<<plotInt2, fig.width=6, fig.height=4, fig.cap='vagueness by number interaction over item ratios', fig.pos='hbtp', message=FALSE, echo=FALSE>>=
+rts <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number","Vagueness","discriminability"))
+rts$discriminability <- as.numeric(as.character(rts$discriminability))
+rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
+
+ggplot(rts, aes(y=RT_log, x=discriminability, ymin=RT_log-ci, ymax=RT_log+ci,
+                group=condition, shape=condition, fill=condition)) +
+  geom_line(position = position_dodge(width=0.005)) +
+  geom_errorbar(position = position_dodge(width=0.005)) +
+  geom_point(position = position_dodge(width=0.005), cex=3) +
+  scale_shape_manual("",values = c(22, 22, 21, 21)) +
+  scale_fill_manual("",values=c("black","white","black","white")) +
+  ggtitle("Response time") +
+  ylab("RT_log") +
+  xlab("") +
+  theme_bw() +
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        legend.key = element_blank(), aspect.ratio=1,
+        axis.text.x = element_text(angle = 45, hjust = 1)) +
+  facet_grid(~Number) +
+  scale_x_continuous(minor_breaks=NULL,
+                     breaks=1-c(0.5125000, 0.6876471, 0.7691558, 0.8166667),
+                     labels=abbreviate(1-c(0.5125000, 0.6876471, 0.7691558, 0.8166667)))
+@
+
+\clearpage
+
+<<plot3way, fig.width=7, fig.height=7, fig.pos='hbtp', fig.cap='3 way interactions',echo=FALSE>>=
+rts <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number","Vagueness","Quantity"))
+rts$mins=rts$RT_log-rts$ci
+rts$maxs=rts$RT_log+rts$ci
+p1=ggplot(rts, aes(y=RT_log, x=Number, group=Vagueness, ymin=mins, ymax=maxs, pch=Vagueness)) +
   geom_line() +
@@
   theme(legend.title=element_blank(), legend.position="top")
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Order"))
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-p2=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Vagueness, ymin=mins, ymax=maxs, pch=Vagueness)) +
+rts <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number","Vagueness","Order"))
+rts$mins=rts$RT_log-rts$ci
+rts$maxs=rts$RT_log+rts$ci
+p2=ggplot(rts, aes(y=RT_log, x=Number, group=Vagueness, ymin=mins, ymax=maxs, pch=Vagueness)) +
   geom_line() +
@@
   theme(legend.title=element_blank(), legend.position="top")
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Quantity","Order"))
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-p3=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Quantity, ymin=mins, ymax=maxs, pch=Quantity)) +
+rts <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number","Quantity","Order"))
+rts$mins=rts$RT_log-rts$ci
+rts$maxs=rts$RT_log+rts$ci
+p3=ggplot(rts, aes(y=RT_log, x=Number, group=Quantity, ymin=mins, ymax=maxs, pch=Quantity)) +
   geom_line() +
@@
   theme(legend.title=element_blank(), legend.position="top")
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Quantity","Order"))
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-p4=ggplot(rts, aes(y=RT_RecSqt, x=Vagueness, group=Quantity, ymin=mins, ymax=maxs, pch=Quantity)) +
+rts <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Quantity","Order"))
+rts$mins=rts$RT_log-rts$ci
+rts$maxs=rts$RT_log+rts$ci
+p4=ggplot(rts, aes(y=RT_log, x=Vagueness, group=Quantity, ymin=mins, ymax=maxs, pch=Quantity)) +
   geom_line() +
@@
 
-<<vnitems, fig.width=7, fig.height=4, fig.cap='Plot vagueness by number by quantity over items.',echo=FALSE>>=
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item", "Quantity"))
+\clearpage
+
+<<vnitems, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='Plot vagueness by number by quantity over item ratios.',message = FALSE, echo=FALSE>>=
+rts <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number","Vagueness","discriminability", "Quantity"))
 rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
-dodge = position_dodge(width=0.2)
-ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci,
+rts$discriminability <- as.numeric(as.character(rts$discriminability))
+ggplot(rts, aes(y=RT_log, 
+                x=discriminability, 
+                ymin=RT_log-ci, 
+                ymax=RT_log+ci,
                 group=Vagueness,
-fill=Vagueness)) +
-geom_line(position=dodge) +
-geom_errorbar(width=.2, position=dodge) +
-geom_point(pch=21, size=4, position=dodge) +
+                fill=Vagueness,
+                shape=Vagueness)) +
+  geom_line(position=position_dodge(width=.025)) +
+  geom_errorbar(position=position_dodge(width=.025), width=.02) +
+  geom_point(position=position_dodge(width=.025), cex=3) +
+  scale_shape_manual("",values = c(22, 21)) +
   scale_fill_manual("",values=c("black","white")) +
@@
   ylab("RT_Rec_Sqt") +
-xlab("") +
+  xlab("Item discriminability") +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
@@
         legend.key = element_blank(),
-aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust = 1)) +
-facet_grid(~Number+Quantity)
+        aspect.ratio=1, 
+        axis.text.x = element_text(angle = 45, hjust = 1)) +
+  facet_grid(~Number+Quantity) +
+    scale_x_continuous(minor_breaks=NULL,
+                     breaks=1-c(0.5125000, 0.6876471, 0.7691558, 0.8166667),
+                     labels=abbreviate(1-c(0.5125000, 0.6876471, 0.7691558, 0.8166667)))
 @
 
-<<4way, fig.width=6, fig.height=3, fig.cap='4 way interaction', echo=FALSE>>=
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness",  "Quantity", "Order"))
+\clearpage
+
+<<4way, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='4 way interaction', echo=FALSE>>=
+rts <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number", "Vagueness",  "Quantity", "Order"))
 rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
-ggplot(rts, aes(y=RT_RecSqt, x=Vagueness, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, group=Quantity, pch=Quantity)) +
+ggplot(rts, aes(y=RT_log, x=Vagueness, ymin=RT_log-ci, ymax=RT_log+ci, group=Quantity, pch=Quantity)) +
   facet_grid(~Order+Number, scale="free_x") +
-geom_line() +
-geom_errorbar(width=.1) +
-geom_point(cex=3) +
+  geom_line(position = position_dodge(width = 0.2)) +
+  geom_errorbar(position = position_dodge(width = 0.2), width=.1) +
+  geom_point(position = position_dodge(width = 0.2), cex=3) +
   theme_bw() + 
   theme(aspect.ratio=1) +
-theme(legend.title=element_blank())
+  theme(legend.title=element_blank(),
+        legend.position='right',
+        legend.key=element_rect(color = 'white'))  
 @
 
-<<4waysplit, fig.width=7, fig.height=7, fig.cap='4 way interaction split over items',echo=FALSE>>=
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness", "Item", "Quantity", "Order"))
+\clearpage
+
+<<4waysplit, fig.width=7, fig.height=7, fig.cap='4 way interaction split over item discriminability', fig.pos='hbtp', message=FALSE, echo=FALSE>>=
+rts <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number", "Vagueness", "discriminability", "Quantity", "Order"))
 rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
-ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci,
-group=Vagueness,fill=Vagueness)) +
-geom_line() +
-geom_errorbar(width=.12) +
-geom_point(pch=21, size=2) +
-scale_fill_manual("",values=c("black","white","black","white")) +
-ggtitle("Number * Vagueness * Quantity * Order") +
-ylab("RT_RecSqt") +
-xlab("") +
-theme_bw() +
-theme( legend.key = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1),
-axis.text.x = element_text(angle = 45, hjust = 1), aspect.ratio=1) +
-facet_grid(Vagueness+Order~Number+Quantity)
-@
-
-<<4waySplitRatio, fig.width=7, fig.height=7, fig.cap='4 way interaction split over items showing ratio of item dots.',echo=FALSE, message=FALSE, warning=FALSE>>=
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness", "item_mean_ratio", "Quantity", "Order"))
-rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
-rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
-#dodge = position_dodge(width=0.2)
-ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci,
-group=Vagueness,fill=Vagueness)) +
-# geom_line(position=dodge) +
+rts$discriminability <- as.numeric(as.character(rts$discriminability))
+ggplot(rts, 
+       aes(y=RT_log, x=discriminability, ymin=RT_log-ci, ymax=RT_log+ci, group=Vagueness, fill=Vagueness)
+       ) +
+  facet_grid(Vagueness+Order~Number+Quantity) +
   geom_line() +
@@
   ggtitle("Number * Vagueness * Quantity * Order") +
-ylab("RT_RecSqt") +
+  ylab("RT_log") +
   xlab("") +
   theme_bw() +
-theme( legend.key = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1),
-axis.text.x = element_text(angle = 45, hjust = 1), aspect.ratio=1) +
-facet_grid(Vagueness+Order~Number+Quantity)
+  theme( legend.key = element_blank(), 
+         axis.text.x = element_text(angle = 45, hjust = 1, size=8), 
+         axis.text.y = element_text(angle = 45, hjust = 1, size=6),
+         aspect.ratio=1) + 
+  scale_x_continuous(minor_breaks=NULL,
+                     breaks=1-c(0.5125000, 0.6876471, 0.7691558, 0.8166667),
+                     labels=abbreviate(1-c(0.5125000, 0.6876471, 0.7691558, 0.8166667)))
 @
 
-<<showCompression, fig.width=7, fig.height=7, fig.cap='Show compression of item ratios', echo=FALSE>>=
-#  Show compression of ratios of numbers of dots in the items
-item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54)
-item_range_ratio_equally_spaced = seq(from = 6 / 24, to = 36 / 54, length.out = 4)
-item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54)))
-item_mean_ratio_equally_spaced = seq(from = mean(c(6 / 15, 15 / 24)), to = mean(c(36 / 45, 45 / 54)), length.out = 4)
-plot(y=c(2,2,2,2), x=item_range_ratio, xlim=c(0,1), ylim=c(0,3), ylab="", yaxt='n', col="blue", pch=21, bg='blue', type='o', xaxt='n', xlab="ratio: small is easily discriminable", main="Discriminability of each item,\n expressed as ratios between array magnitudes")
-text(y=c(2.15,2.15,2.15,2.15), x=item_range_ratio, labels=round(item_range_ratio,2), col="blue")
-text(y=2.15, x=0.1, labels="ratio min/max", col='blue')
-points(y=c(2.3,2.3,2.3,2.3), x=item_range_ratio_equally_spaced, col="lightblue", pch=21, bg='lightblue', type='o')
-points(y=c(1,1,1,1), x=item_mean_ratio, pch=21, col='red', bg='red', type='o')
-text(y=c(1.15,1.15,1.15,1.15), x=item_mean_ratio, labels=round(item_mean_ratio,2), col="red")
-text(y=1.15, x=0.2, labels="mean (ratio min/med,\n            ratio med/max)", col='red')
-points(y=c(1.3,1.3,1.3,1.3), x=item_mean_ratio_equally_spaced, col="pink", pch=21, bg='pink', type='o')
-axis(1, at=seq(0,1,0.1))
-@
-
-\FloatBarrier
+\clearpage
 \section{Lmer model}
@@
 
-<<lmer5, cache=TRUE>>=
-v5 <- lmerTest::lmer(data=dd,
+<<lmer5, tidy=FALSE, cache=TRUE>>=
+v5 <- lme4::lmer(data=dd,
                  RT_log ~
@@
                    c_Num:c_Vag:c_Qty +
-item_mean_ratio +
-s_Trl +
-RTprev_log +
-nchar_instr +
-(1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
-v5b <- lme4::lmer(data=dd,
-RT_log ~
-c_Vag + c_Num + c_Qty + c_Ord +
-c_Num:c_Vag:c_Qty +
-item_mean_ratio +
+                   discriminability +
                    s_Trl +
@@
 
-<<v51, cache=TRUE>>=
-summary(v5)
+<<printXtableV5, echo=FALSE, results='asis'>>=
+print(
+  xtable(
+    coef(summary(v5)
+         ),
+  caption = 'xtable v5'
+  ),
+  table.placement='htbp',
+  latex.environments='',
+  size='small'
+)
 @
 
-<<v511, results='asis', cache=TRUE>>=
-print(xtable(coef(summary(v5b))))
+<<rSquaredInfo, echo=FALSE>>=
+cat("R^2")
+cor(fitted(v5), dd$RT_log)^2
 @
 
-<<v52>>=
-cat("R^2")
-cor(fitted(v5), dd$RT_RecSqt)^2
+\clearpage
+
+<<plotModelCoefsAndCis, fig.width=6, fig.height=4, fig.cap='Coefficient estimates and their (Wald) 95 per cent confidence intervals', fig.pos='hbtp', echo=FALSE>>=
+# Plot model coefficients and ci's
+dfr <- data.frame(coef = summary(v5)$coef[-1, 1], 
+                     ci2.5  = confint(v5, method='Wald')[18:26, 1],
+                     ci97.5 = confint(v5, method='Wald')[18:26, 2] )
+dfr <- dfr[rev(rownames(dfr)),]
+par(mar=c(5,10,1,1))
+plot(y=1:nrow(dfr), 
+     x=dfr$coef, 
+     xlim=range(dfr$ci2.5, dfr$ci97.5), 
+     type='n', axes=F, xlab="", ylab="")
+abline(v=0, lty=3, col='grey')
+abline(h=c(1,2,3,4,5,6,7,8,9), col='grey', lty=3)
+arrows(y0=1:nrow(dfr), y1=1:nrow(dfr), 
+       x0=dfr$ci2.5, x1=dfr$ci97.5, 
+       length=0.035,
+       col='black',
+       lwd=2, code=3, angle=90)
+points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white')
+axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE)
+axis(1)
+mtext("Coefficient estimates, in units of log(RT),\nwith 95% confidence intervals", side=1, line=3.5, cex.lab=1, las=1)
+mtext("Model terms", side=2, line=8)
+box(col='grey')
 @
 
-<<plotLMERfnc, fig.width=7, fig.height=3.5, fig.cap='caption'>>=
-par(mfrow=c(3,4), family='cmodern')
+\clearpage
+
+<<plotLMERfnc, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='plotMLERfnc', echo=TRUE>>=
+par(mfrow=c(2,4))
 plotLMER.fnc(v5)
@@
 
-Plot model coefficients and ci's
+\clearpage
 
-<<v54, fig.width=7, fig.height=4, fig.cap='Estimates plot'>>=
-e=data.frame(coef=summary(v5)$coef[-1,1]) # estimates, without intercept
-q=as.data.frame(confint(v5, method='Wald')[18:26,1:2])
-eq=cbind(e,q)
-y=eq[rev(rownames(eq)),]
-par(mar=c(4,10,1,1))
-plot(y=1:nrow(y), x=y$coef, xlim=range(y$"2.5 %", y$"97.5 %"), type='n', axes=F, xlab="", ylab="")
-segments(y0=1:nrow(y), y1=1:nrow(y), x0=y$"2.5 %", x1=y$"97.5 %", lwd=.55)
-points(y=1:nrow(y), x=y$coef, pch=20, cex=.75)
-abline(v=0, lty=3, col='grey')
-axis(2, labels=row.names(y), at=1:nrow(y), las=1)
-axis(1, las=1)
-mtext("estimate", side=1, line=2.5, cex.lab=1, las=1)
-box(col='grey')
-@
-
-Model criticism baayen plots
-
-<<baayenPLots99, fig.width=9, fig.height=3, fig.cap='caption'>>=
+<<baayenPLots99, fig.width=6, fig.height=4, fig.cap='Baayen Model Criticism Plots', fig.pos='hbtp', echo=FALSE>>=
 # Baayen 4-plot model criticism
-par(mfrow=c(1,4), pty='s')
+par(mfrow=c(1,3), pty='s')
 # create scaled residuals
@@
 
+\clearpage
+
+\section{lmerTest Version}
+
+<<lmerTestVersionOfModel, cache=TRUE>>=
+v6 <- lmerTest::lmer(data=dd,
+RT_log ~
+c_Vag + c_Num + c_Qty + c_Ord +
+c_Num:c_Vag:c_Qty +
+discriminability +
+s_Trl +
+RTprev_log +
+nchar_instr +
+(1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
+@
+
+<<summaryLmerTest, cache=TRUE>>=
+summary(v6)
+@
+
+\clearpage
+\section{Borderline responses}
+
+<<barplotBorderline, fig.width=6, fig.height=4, echo=FALSE>>=
+dd$response_category <- relevel(dd$response_category, ref = 'expected')
+#barplot(table(dd$response_category,dd$Vagueness), beside=TRUE, legend=TRUE)
+barplot(table(dd$Vagueness,dd$response_category), beside=TRUE, legend=TRUE, space=c(0,.5))
+@
+
+<<tabBdl, echo=FALSE, results='asis'>>=
+print(xtable(table(dd$response_category,dd$Vagueness),
+             caption='Borderline cases counts'),
+      table.placement='htbp',
+      latex.environments='center',
+      size='small')
+@
+
 \end{document}

### 2016-08-12 experiments/e2/writeup/e2-b-knitr.tex
more consistency, better plots, using discriminability instead of item ratio

@@
-\documentclass[a4paper]{article}\usepackage[]{graphicx}\usepackage[]{color}
+\documentclass[a4paper,12pt]{article}\usepackage[]{graphicx}\usepackage[]{color}
 %% maxwidth is the original width if it is less than linewidth
@@
 \usepackage{alltt}
-\usepackage{geometry}
+\usepackage[margin=2cm]{geometry}
 \usepackage{placeins}
+
+\title{Exp 2 of 4}
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
@@
 
-\begin{knitrout}
+\maketitle
+\tableofcontents
+\clearpage
+
+\clearpage
+\section{Get the data}
+
+\begin{knitrout}\scriptsize
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 \begin{alltt}
-\hlcom{# .define.fonts <- function () \{}
-\hlcom{#     quartzFonts(avenir =    c('Avenir Book', 'Avenir Black', 'Avenir Book Oblique', 'Avenir Black Oblique'),}
-\hlcom{#                 helvetica = c('Helvetica New Light', 'Helvetica New Bold', 'Helvetica New Light Italic', 'Helvetica New Bold Italic'),}
-\hlcom{#                 cmodern = c('CMU Serif', 'CMU Serif Extra','CMU Classical Serif','CMU Serif Extra'))}
-\hlcom{# \}}
-\hlcom{# .define.fonts()}
+\hlcom{# if the file data_processed.Rda already exists then load it, else do data wrangling}
+\hlkwa{if} \hlstd{(}\hlkwd{file.exists}\hlstd{(}\hlstr{"data_processed.Rda"}\hlstd{)) \{}
+    \hlkwd{load}\hlstd{(}\hlstr{"data_processed.Rda"}\hlstd{)}
+\hlstd{\}} \hlkwa{else} \hlstd{\{}
+    \hlcom{# declare local variables}
+    \hlstd{number_of_valid_subjects} \hlkwb{<-} \hlnum{30}  \hlcom{# = 30}
+    \hlstd{number_of_rows} \hlkwb{<-} \hlnum{7680}  \hlcom{# 7680}
+    \hlstd{number_of_trials_per_subject} \hlkwb{<-} \hlstd{number_of_rows}\hlopt{/}\hlstd{number_of_valid_subjects}  \hlcom{# = 256}
+    \hlcom{# call functions}
+    \hlstd{dat} \hlkwb{<-} \hlkwd{gatherData}\hlstd{(number_of_valid_subjects)}  \hlcom{# = 30}
+    \hlstd{dat} \hlkwb{<-} \hlkwd{classifyResponses}\hlstd{(dat)}  \hlcom{# classify the response as expected, near, or far}
+    \hlstd{dat} \hlkwb{<-} \hlkwd{processData}\hlstd{(dat)}  \hlcom{# remove impossible trials and re-do previous rt measures}
+    \hlstd{dd} \hlkwb{<-} \hlkwd{postProcessData}\hlstd{(dat)}
+\hlstd{\}}  \hlcom{# end else do data wrangling}
 \end{alltt}
@@
 
-\begin{knitrout}
+\begin{knitrout}\scriptsize
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 \begin{alltt}
-\hlkwd{library}\hlstd{(extrafont)}
+\hlkwd{names}\hlstd{(dd)}
 \end{alltt}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Registering fonts with R}}\begin{alltt}
-\hlkwd{loadfonts}\hlstd{()}
-\end{alltt}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# .Keyboard already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Andale Mono already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# More than one version of regular/bold/italic found for Apple Braille. Skipping setup for this font.}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# AppleMyungjo already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Arial Black already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Arial already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Arial Narrow already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Arial Rounded MT Bold already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Arial Unicode MS already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Bodoni Ornaments already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Bodoni 72 Smallcaps already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# No regular (non-bold, non-italic) version of . Skipping setup for this font.}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# No regular (non-bold, non-italic) version of Brush Script MT. Skipping setup for this font.}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Comic Sans MS already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Courier New already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# No regular (non-bold, non-italic) version of DIN Alternate. Skipping setup for this font.}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# No regular (non-bold, non-italic) version of DIN Condensed. Skipping setup for this font.}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Georgia already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Impact already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Khmer Sangam MN already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Lao Sangam MN already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Luminari already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Microsoft Sans Serif already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Tahoma already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Times New Roman already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Trattatello already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Trebuchet MS already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Verdana already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Webdings already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Wingdings already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Wingdings 2 already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Wingdings 3 already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# More than one version of regular/bold/italic found for CMU Bright. Skipping setup for this font.}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# No regular (non-bold, non-italic) version of CMU Classical Serif. Skipping setup for this font.}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# CMU Concrete already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# CMU Sans Serif already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# CMU Sans Serif Demi Condensed already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# CMU Serif already registered with pdfFonts().}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# No regular (non-bold, non-italic) version of CMU Serif Extra. Skipping setup for this font.}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# No regular (non-bold, non-italic) version of CMU Serif Upright Italic. Skipping setup for this font.}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# More than one version of regular/bold/italic found for CMU Typewriter Text. Skipping setup for this font.}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# CMU Typewriter Text Variable Width already registered with pdfFonts().}}\end{kframe}
+\begin{verbatim}
+ [1] "id"                "Subject"           "Trial"             "Condition"        
+ [5] "Order"             "Quantity"          "Vagueness"         "Number"           
+ [9] "Item"              "discriminability"  "c_Trl"             "s_Trl"            
+[13] "c_Itm"             "c_Vag"             "c_Num"             "f_Cnd"            
+[17] "c_Ord"             "c_Qty"             "RT"                "RT_log"           
+[21] "RT_raw"            "RTprev"            "RTprev_log"        "RTprev_raw"       
+[25] "Exp_Num"           "Bline_Num"         "Extr_Num"          "Exp_side"         
+[29] "Bline_side"        "Extr_side"         "response_num"      "response_side"    
+[33] "response_category" "Left"              "Mid"               "Right"            
+[37] "Instruction"       "nchar_instr"      
+\end{verbatim}
+\end{kframe}
 \end{knitrout}
 
-\section{Get the data}
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{head}\hlstd{(dd)}
+\end{alltt}
+\begin{verbatim}
+        id Subject Trial Condition Order Quantity Vagueness  Number     Item discriminability
+1 s01:t001     s01     1         2  RtoL    Small     Crisp Numeric 06:15:24        0.4875000
+2 s01:t002     s01     2         1  LtoR    Large     Vague Numeric 16:25:34        0.3123529
+3 s01:t003     s01     3         4  LtoR    Small     Crisp  Verbal 26:35:44        0.2308442
+4 s01:t004     s01     4         1  RtoL    Large     Vague Numeric 36:45:54        0.1833333
+5 s01:t005     s01     5         1  RtoL    Small     Vague Numeric 06:15:24        0.4875000
+6 s01:t006     s01     6         3  LtoR    Large     Vague  Verbal 16:25:34        0.3123529
+   c_Trl     s_Trl c_Itm c_Vag c_Num f_Cnd c_Ord c_Qty   RT   RT_log RT_raw RTprev RTprev_log
+1 -127.5 -1.725186 -0.75  -0.5  -0.5 Cr:Nm   0.5  -0.5 1517 7.324490   1517   1517   7.324490
+2 -126.5 -1.711655 -0.25   0.5  -0.5 Vg:Nm  -0.5   0.5 1920 7.560080   1920   1517   7.324490
+3 -125.5 -1.698124  0.25  -0.5   0.5 Cr:Vb  -0.5  -0.5 2346 7.760467   2346   1920   7.560080
+4 -124.5 -1.684593  0.75   0.5  -0.5 Vg:Nm   0.5   0.5 1773 7.480428   1773   2346   7.760467
+5 -123.5 -1.671062 -0.75   0.5  -0.5 Vg:Nm   0.5  -0.5 2556 7.846199   2556   1773   7.480428
+6 -122.5 -1.657531 -0.25   0.5   0.5 Vg:Vb  -0.5   0.5 2043 7.622175   2043   2556   7.846199
+  RTprev_raw Exp_Num Bline_Num Extr_Num Exp_side Bline_side Extr_side response_num response_side
+1       1517       6        15       24    right        mid      left            6         right
+2       1517      34        25       16    right        mid      left           25           mid
+3       1920      26        35       44     left        mid     right           26          left
+4       2346      54        45       36     left        mid     right           45           mid
+5       1773       6        15       24    right        mid      left            6         right
+6       2556      34        25       16    right        mid      left           34         right
+  response_category Left Mid Right                            Instruction nchar_instr
+1          expected   24  15     6          Choose the square with 6 dots          29
+2        borderline   16  25    34     Choose a square with about 30 dots          34
+3          expected   26  35    44 Choose the square with the fewest dots          38
+4        borderline   54  45    36     Choose a square with about 50 dots          34
+5          expected   24  15     6     Choose a square with about 10 dots          34
+6          expected   16  25    34         Choose a square with many dots          30
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
 
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{summary}\hlstd{(dd)}
+\end{alltt}
+\begin{verbatim}
+        id          Subject         Trial         Condition    Order       Quantity    Vagueness   
+ s01:t001:   1   s01    : 256   Min.   :  1.0   Min.   :1.0   LtoR:3838   Small:3840   Crisp:3840  
+ s01:t002:   1   s02    : 256   1st Qu.: 65.0   1st Qu.:2.0   RtoL:3839   Large:3837   Vague:3837  
+ s01:t003:   1   s03    : 256   Median :129.0   Median :3.0                                        
+ s01:t004:   1   s04    : 256   Mean   :128.5   Mean   :2.5                                        
+ s01:t005:   1   s05    : 256   3rd Qu.:193.0   3rd Qu.:4.0                                        
+ s01:t006:   1   s06    : 256   Max.   :256.0   Max.   :4.0                                        
+ (Other) :7671   (Other):6141                                                                      
+     Number           Item      discriminability     c_Trl                s_Trl           
+ Numeric:3838   06:15:24:1919   Min.   :0.1833   Min.   :-127.50000   Min.   :-1.7251858  
+ Verbal :3839   16:25:34:1919   1st Qu.:0.2308   1st Qu.: -63.50000   1st Qu.:-0.8592102  
+                26:35:44:1920   Median :0.2308   Median :   0.50000   Median : 0.0067654  
+                36:45:54:1919   Mean   :0.3035   Mean   :   0.02377   Mean   : 0.0003217  
+                                3rd Qu.:0.3124   3rd Qu.:  64.50000   3rd Qu.: 0.8727411  
+                                Max.   :0.4875   Max.   : 127.50000   Max.   : 1.7251858
+
+     c_Itm               c_Vag                c_Num             f_Cnd          c_Ord          
+ Min.   :-7.50e-01   Min.   :-0.5000000   Min.   :-5.00e-01   Vg:Nm:1918   Min.   :-5.00e-01  
+ 1st Qu.:-2.50e-01   1st Qu.:-0.5000000   1st Qu.:-5.00e-01   Cr:Nm:1920   1st Qu.:-5.00e-01  
+ Median : 2.50e-01   Median :-0.5000000   Median : 5.00e-01   Vg:Vb:1919   Median : 5.00e-01  
+ Mean   : 3.26e-05   Mean   :-0.0001954   Mean   : 6.51e-05   Cr:Vb:1920   Mean   : 6.51e-05  
+ 3rd Qu.: 2.50e-01   3rd Qu.: 0.5000000   3rd Qu.: 5.00e-01                3rd Qu.: 5.00e-01  
+ Max.   : 7.50e-01   Max.   : 0.5000000   Max.   : 5.00e-01                Max.   : 5.00e-01
+
+     c_Qty                  RT            RT_log           RT_raw          RTprev     
+ Min.   :-0.5000000   Min.   :  445   Min.   : 6.098   Min.   :  445   Min.   :  445  
+ 1st Qu.:-0.5000000   1st Qu.: 1240   1st Qu.: 7.123   1st Qu.: 1240   1st Qu.: 1240  
+ Median :-0.5000000   Median : 1727   Median : 7.454   Median : 1727   Median : 1727  
+ Mean   :-0.0001954   Mean   : 2840   Mean   : 7.595   Mean   : 2840   Mean   : 2835  
+ 3rd Qu.: 0.5000000   3rd Qu.: 2699   3rd Qu.: 7.901   3rd Qu.: 2699   3rd Qu.: 2697  
+ Max.   : 0.5000000   Max.   :42685   Max.   :10.662   Max.   :42685   Max.   :42685
+
+   RTprev_log       RTprev_raw       Exp_Num     Bline_Num     Extr_Num    Exp_side        
+ Min.   : 6.098   Min.   :  445   Min.   : 6   Min.   :15   Min.   : 6   Length:7677       
+ 1st Qu.: 7.123   1st Qu.: 1240   1st Qu.:16   1st Qu.:25   1st Qu.:24   Class :character  
+ Median : 7.454   Median : 1727   Median :26   Median :35   Median :34   Mode  :character  
+ Mean   : 7.594   Mean   : 2835   Mean   :30   Mean   :30   Mean   :30                     
+ 3rd Qu.: 7.900   3rd Qu.: 2697   3rd Qu.:36   3rd Qu.:35   3rd Qu.:44                     
+ Max.   :10.662   Max.   :42685   Max.   :54   Max.   :45   Max.   :54
+
+  Bline_side         Extr_side          response_num   response_side  response_category
+ Length:7677        Length:7677        Min.   : 6.00   left :3215    borderline:1274   
+ Class :character   Class :character   1st Qu.:24.00   mid  :1274    expected  :6108   
+ Mode  :character   Mode  :character   Median :34.00   right:3188    extreme   : 295   
+                                       Mean   :30.87                                   
+                                       3rd Qu.:44.00                                   
+                                       Max.   :54.00
+
+      Left         Mid         Right                                    Instruction  
+ Min.   : 6   Min.   :15   Min.   : 6   Choose a square with few dots         : 960  
+ 1st Qu.:24   1st Qu.:25   1st Qu.:24   Choose the square with the fewest dots: 960  
+ Median :34   Median :35   Median :26   Choose the square with the most dots  : 960  
+ Mean   :30   Mean   :30   Mean   :30   Choose a square with many dots        : 959  
+ 3rd Qu.:44   3rd Qu.:35   3rd Qu.:36   Choose a square with about 30 dots    : 480  
+ Max.   :54   Max.   :45   Max.   :54   Choose a square with about 40 dots    : 480  
+                                        (Other)                               :2878  
+  nchar_instr   
+ Min.   :29.00  
+ 1st Qu.:30.00  
+ Median :30.00  
+ Mean   :32.59  
+ 3rd Qu.:36.00  
+ Max.   :38.00
+
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{str}\hlstd{(dd)}
+\end{alltt}
+\begin{verbatim}
+'data.frame':	7677 obs. of  38 variables:
+ $ id               : Factor w/ 7680 levels "s01:t001","s01:t002",..: 1 2 3 4 5 6 7 8 9 10 ...
+ $ Subject          : Factor w/ 30 levels "s01","s02","s03",..: 1 1 1 1 1 1 1 1 1 1 ...
+ $ Trial            : int  1 2 3 4 5 6 7 8 9 10 ...
+ $ Condition        : int  2 1 4 1 1 3 3 4 4 2 ...
+ $ Order            : Factor w/ 2 levels "LtoR","RtoL": 2 1 1 2 2 1 2 2 1 2 ...
+ $ Quantity         : Factor w/ 2 levels "Small","Large": 1 2 1 2 1 2 1 1 2 2 ...
+ $ Vagueness        : Factor w/ 2 levels "Crisp","Vague": 1 2 1 2 2 2 2 1 1 1 ...
+ $ Number           : Factor w/ 2 levels "Numeric","Verbal": 1 1 2 1 1 2 2 2 2 1 ...
+ $ Item             : Factor w/ 4 levels "06:15:24","16:25:34",..: 1 2 3 4 1 2 3 4 1 2 ...
+ $ discriminability : num  0.487 0.312 0.231 0.183 0.487 ...
+ $ c_Trl            : num  -128 -126 -126 -124 -124 ...
+ $ s_Trl            : num  -1.73 -1.71 -1.7 -1.68 -1.67 ...
+ $ c_Itm            : num  -0.75 -0.25 0.25 0.75 -0.75 -0.25 0.25 0.75 -0.75 -0.25 ...
+ $ c_Vag            : num  -0.5 0.5 -0.5 0.5 0.5 0.5 0.5 -0.5 -0.5 -0.5 ...
+ $ c_Num            : num  -0.5 -0.5 0.5 -0.5 -0.5 0.5 0.5 0.5 0.5 -0.5 ...
+ $ f_Cnd            : Factor w/ 4 levels "Vg:Nm","Cr:Nm",..: 2 1 4 1 1 3 3 4 4 2 ...
+ $ c_Ord            : num  0.5 -0.5 -0.5 0.5 0.5 -0.5 0.5 0.5 -0.5 0.5 ...
+ $ c_Qty            : num  -0.5 0.5 -0.5 0.5 -0.5 0.5 -0.5 -0.5 0.5 0.5 ...
+ $ RT               : int  1517 1920 2346 1773 2556 2043 2384 3078 1760 2218 ...
+ $ RT_log           : num  7.32 7.56 7.76 7.48 7.85 ...
+ $ RT_raw           : int  1517 1920 2346 1773 2556 2043 2384 3078 1760 2218 ...
+ $ RTprev           : int  1517 1517 1920 2346 1773 2556 2043 2384 3078 1760 ...
+ $ RTprev_log       : num  7.32 7.32 7.56 7.76 7.48 ...
+ $ RTprev_raw       : int  1517 1517 1920 2346 1773 2556 2043 2384 3078 1760 ...
+ $ Exp_Num          : num  6 34 26 54 6 34 26 36 24 34 ...
+ $ Bline_Num        : num  15 25 35 45 15 25 35 45 15 25 ...
+ $ Extr_Num         : num  24 16 44 36 24 16 44 54 6 16 ...
+ $ Exp_side         : chr  "right" "right" "left" "left" ...
+ $ Bline_side       : chr  "mid" "mid" "mid" "mid" ...
+ $ Extr_side        : chr  "left" "left" "right" "right" ...
+ $ response_num     : int  6 25 26 45 6 34 26 36 24 25 ...
+ $ response_side    : Factor w/ 3 levels "left","mid","right": 3 2 1 2 3 3 3 3 3 2 ...
+ $ response_category: Factor w/ 3 levels "borderline","expected",..: 2 1 2 1 2 2 2 2 2 1 ...
+ $ Left             : int  24 16 26 54 24 16 44 54 6 34 ...
+ $ Mid              : int  15 25 35 45 15 25 35 45 15 25 ...
+ $ Right            : int  6 34 44 36 6 34 26 36 24 16 ...
+ $ Instruction      : Factor w/ 17 levels "Choose a square with about 10 dots",..: 15 3 16 5 1 7 6 16 17 11 ...
+ $ nchar_instr      : int  29 34 38 34 34 30 29 38 36 30 ...
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+\clearpage
 \section{Plots}
+\clearpage
+\subsection{Discriminability}
 
-\begin{knitrout}
+\begin{knitrout}\scriptsize
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
 
-\hfill{}\includegraphics[width=\maxwidth]{figure/plotDists-1}
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-showCompression-1}
 
-\caption[Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions]{Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions}\label{fig:plotDists}
+}
+
+\caption[Ratios for different numbers of dots in the arrays]{Ratios for different numbers of dots in the arrays: smaller values are more discriminable.
+Blue is for the ratio between the smallest number in the array and the largest number in an array.
+Red is for the mean of two ratios, one for the smallest number to the middle number, the other for the middle number to the largest number in the array}\label{fig:showCompression}
 \end{figure}
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/exampleOfFonts-1} \caption[font example]{font example}\label{fig:exampleOfFonts}
+\clearpage
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-plotDistFreqPolys-1}
+
+}
+
+\caption[Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions]{Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions}\label{fig:plotDistFreqPolys}
 \end{figure}
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/boxPlots-1} \caption[Show how transformations of RT affect distribution of times, and how they affect which times are outliers]{Show how transformations of RT affect distribution of times, and how they affect which times are outliers.}\label{fig:boxPlots}
-\end{figure}
+\clearpage
 
-\end{knitrout}
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/fastSlowRT-1} \caption[Show how mean times for individual subjects and items vary with respect to the grand mean Log RT]{Show how mean times for individual subjects and items vary with respect to the grand mean Log RT.}\label{fig:fastSlowRT}
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-plotBoxPlots-1}
+
+}
+
+\caption[Show how transformations of RT affect distribution of times, and how they affect which times are outliers]{Show how transformations of RT affect distribution of times, and how they affect which times are outliers.}\label{fig:plotBoxPlots}
 \end{figure}
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/plotMainsa-1} \caption[Plot main effects on several transformations]{Plot main effects on several transformations}\label{fig:plotMainsa}
+\clearpage
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-plotFastSlowRTs-1}
+
+}
+
+\caption[Show how mean times for individual subjects and items vary with respect to the grand mean Log RT]{Show how mean times for individual subjects and items vary with respect to the grand mean Log RT.}\label{fig:plotFastSlowRTs}
 \end{figure}
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/mainB-1} \caption[Just the main effects]{Just the main effects}\label{fig:mainB}
+\clearpage
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-plotMainsa-1}
+
+}
+
+\caption[Plot main effects on several transformations]{Plot main effects on several transformations}\label{fig:plotMainsa}
 \end{figure}
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/mainD-1} \caption[Main effects over item ratios]{Main effects over item ratios}\label{fig:mainD}
+\clearpage
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-mainB-1}
+
+}
+
+\caption[Just the main effects]{Just the main effects}\label{fig:mainB}
 \end{figure}
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/22way-1} \caption[2-way interactions over items]{2-way interactions over items}\label{fig:22way}
+\clearpage
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-mainD-1}
+
+}
+
+\caption[Main effects over item ratios]{Main effects over item ratios}\label{fig:mainD}
 \end{figure}
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/plotInt1-1} \caption[Plot vagueness by number interaction over items]{Plot vagueness by number interaction over items}\label{fig:plotInt1}
+\clearpage
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-22way-1}
+
+}
+
+\caption[2-way interactions over item ratios]{2-way interactions over item ratios}\label{fig:22way}
 \end{figure}
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/plot3way-1} \caption[3 way interactions]{3 way interactions}\label{fig:plot3way}
+\clearpage
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-plotInt1-1}
+
+}
+
+\caption[vagueness by number interaction over items]{vagueness by number interaction over items}\label{fig:plotInt1}
 \end{figure}
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/vnitems-1} \caption[Plot vagueness by number by quantity over items]{Plot vagueness by number by quantity over items.}\label{fig:vnitems}
+\clearpage
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-plotInt2-1}
+
+}
+
+\caption[vagueness by number interaction over item ratios]{vagueness by number interaction over item ratios}\label{fig:plotInt2}
 \end{figure}
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/4way-1} \caption[4 way interaction]{4 way interaction}\label{fig:4way}
+\clearpage
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-plot3way-1}
+
+}
+
+\caption[3 way interactions]{3 way interactions}\label{fig:plot3way}
 \end{figure}
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/4waysplit-1} \caption[4 way interaction split over items]{4 way interaction split over items}\label{fig:4waysplit}
+\clearpage
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-vnitems-1}
+
+}
+
+\caption[Plot vagueness by number by quantity over item ratios]{Plot vagueness by number by quantity over item ratios.}\label{fig:vnitems}
 \end{figure}
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/4waySplitRatio-1} \caption[4 way interaction split over items showing ratio of item dots]{4 way interaction split over items showing ratio of item dots.}\label{fig:4waySplitRatio}
+\clearpage
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-4way-1}
+
+}
+
+\caption[4 way interaction]{4 way interaction}\label{fig:4way}
 \end{figure}
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/showCompression-1} \caption[Show compression of item ratios]{Show compression of item ratios}\label{fig:showCompression}
+\clearpage
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-4waysplit-1}
+
+}
+
+\caption[4 way interaction split over item discriminability]{4 way interaction split over item discriminability}\label{fig:4waysplit}
 \end{figure}
@@
 
-\FloatBarrier
+\clearpage
 \section{Lmer model}
 
-\begin{knitrout}
+\begin{knitrout}\scriptsize
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
@@
 
-\begin{knitrout}
+\begin{knitrout}\scriptsize
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 \begin{alltt}
-\hlstd{v5} \hlkwb{<-} \hlstd{lmerTest}\hlopt{::}\hlkwd{lmer}\hlstd{(}\hlkwc{data}\hlstd{=dd,}
+\hlstd{v5} \hlkwb{<-} \hlstd{lme4}\hlopt{::}\hlkwd{lmer}\hlstd{(}\hlkwc{data}\hlstd{=dd,}
                  \hlstd{RT_log} \hlopt{~}
@@
                    \hlstd{c_Num}\hlopt{:}\hlstd{c_Vag}\hlopt{:}\hlstd{c_Qty} \hlopt{+}
-\hlstd{item_mean_ratio} \hlopt{+}
-\hlstd{s_Trl} \hlopt{+}
-\hlstd{RTprev_log} \hlopt{+}
-\hlstd{nchar_instr} \hlopt{+}
-\hlstd{(}\hlnum{1}\hlopt{+}\hlstd{c_Vag} \hlopt{+} \hlstd{c_Num} \hlopt{+} \hlstd{c_Qty} \hlopt{+} \hlstd{c_Ord}\hlopt{|}\hlstd{Subject))}
-\hlstd{v5b} \hlkwb{<-} \hlstd{lme4}\hlopt{::}\hlkwd{lmer}\hlstd{(}\hlkwc{data}\hlstd{=dd,}
-\hlstd{RT_log} \hlopt{~}
-\hlstd{c_Vag} \hlopt{+} \hlstd{c_Num} \hlopt{+} \hlstd{c_Qty} \hlopt{+} \hlstd{c_Ord} \hlopt{+}
-\hlstd{c_Num}\hlopt{:}\hlstd{c_Vag}\hlopt{:}\hlstd{c_Qty} \hlopt{+}
-\hlstd{item_mean_ratio} \hlopt{+}
+                   \hlstd{discriminability} \hlopt{+}
                    \hlstd{s_Trl} \hlopt{+}
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{summary}\hlstd{(v5)}
-\end{alltt}
-\begin{verbatim}
-## Linear mixed model fit by REML t-tests use Satterthwaite approximations
-##   to degrees of freedom [lmerMod]
-## Formula: RT_log ~ c_Vag + c_Num + c_Qty + c_Ord + c_Num:c_Vag:c_Qty +  
-##     item_mean_ratio + s_Trl + RTprev_log + nchar_instr + (1 +  
-##     c_Vag + c_Num + c_Qty + c_Ord | Subject)
-##    Data: dd
-## 
-## REML criterion at convergence: 11474.8
-## 
-## Scaled residuals: 
-##     Min      1Q  Median      3Q     Max 
-## -4.5470 -0.6351 -0.0955  0.5372  5.0914 
-## 
-## Random effects:
-##  Groups   Name        Variance Std.Dev. Corr                   
-##  Subject  (Intercept) 0.153949 0.39236                         
-##           c_Vag       0.001546 0.03932   0.69                  
-##           c_Num       0.165314 0.40659  -0.67 -0.64            
-##           c_Qty       0.008148 0.09027   0.16  0.26 -0.34      
-##           c_Ord       0.001559 0.03949  -0.13  0.02 -0.39 -0.52
-##  Residual             0.249734 0.49973                         
-## Number of obs: 7677, groups:  Subject, 30
-## 
-## Fixed effects:
-##                     Estimate Std. Error         df t value Pr(>|t|)    
-## (Intercept)        6.400e+00  1.252e-01  2.600e+02  51.112  < 2e-16 ***
-## c_Vag              6.094e-02  1.387e-02  3.300e+01   4.393 0.000112 ***
-## c_Num             -4.336e-01  7.515e-02  2.900e+01  -5.770 2.97e-06 ***
-## c_Qty             -6.907e-02  2.005e-02  2.900e+01  -3.445 0.001743 ** 
-## c_Ord              1.796e-02  1.350e-02  5.100e+01   1.331 0.189164    
-## item_mean_ratio    7.714e-01  4.927e-02  7.551e+03  15.658  < 2e-16 ***
-## s_Trl             -1.070e-01  5.807e-03  7.558e+03 -18.421  < 2e-16 ***
-## RTprev_log         6.047e-02  9.692e-03  7.594e+03   6.239 4.63e-10 ***
-## nchar_instr        6.086e-03  1.944e-03  7.551e+03   3.131 0.001749 ** 
-## c_Vag:c_Num:c_Qty  1.049e-01  4.607e-02  7.551e+03   2.278 0.022742 *  
-## ---
-## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
-## 
-## Correlation of Fixed Effects:
-##             (Intr) c_Vag  c_Num  c_Qty  c_Ord  itm_m_ s_Trl  RTprv_ nchr_n
-## c_Vag        0.083                                                        
-## c_Num       -0.360 -0.337                                                 
-## c_Qty        0.067  0.117 -0.278                                          
-## c_Ord       -0.051  0.007 -0.206 -0.228                                   
-## item_men_rt -0.257 -0.004  0.001  0.000 -0.001                            
-## s_Trl       -0.104  0.003 -0.001 -0.002  0.006 -0.017                     
-## RTprev_log  -0.587  0.006 -0.004 -0.003  0.019 -0.015  0.185              
-## nchar_instr -0.505  0.237 -0.034  0.018  0.000 -0.017  0.000  0.006       
-## c_Vg:c_N:_Q  0.070 -0.033  0.005 -0.003  0.000  0.002 -0.001 -0.002 -0.138
-\end{verbatim}
-\end{kframe}
-\end{knitrout}
-
-\begin{kframe}
-\begin{alltt}
-\hlkwd{print}\hlstd{(}\hlkwd{xtable}\hlstd{(}\hlkwd{coef}\hlstd{(}\hlkwd{summary}\hlstd{(v5b))))}
-\end{alltt}
-\end{kframe}% latex table generated in R 3.3.1 by xtable 1.8-2 package
-% Wed Aug 10 22:41:39 2016
-\begin{table}[ht]
-\centering
+% latex table generated in R 3.3.1 by xtable 1.8-2 package
+% Fri Aug 12 22:28:46 2016
+\begin{table}[htbp]
+\begingroup\small
 \begin{tabular}{rrrr}
@@
   \hline
-(Intercept) & 6.40 & 0.13 & 51.11 \\ 
+(Intercept) & 7.17 & 0.12 & 58.67 \\ 
   c\_Vag & 0.06 & 0.01 & 4.39 \\ 
@@
   c\_Ord & 0.02 & 0.01 & 1.33 \\ 
-  item\_mean\_ratio & 0.77 & 0.05 & 15.66 \\ 
+  discriminability & -0.77 & 0.05 & -15.66 \\ 
   s\_Trl & -0.11 & 0.01 & -18.42 \\ 
@@
 \end{tabular}
+\endgroup
+\caption{xtable v5} 
 \end{table}
 
-\begin{knitrout}
+\begin{knitrout}\scriptsize
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{cat}\hlstd{(}\hlstr{"R^2"}\hlstd{)}
-\end{alltt}
 \begin{verbatim}
-## R^2
-\end{verbatim}
-\begin{alltt}
-\hlkwd{cor}\hlstd{(}\hlkwd{fitted}\hlstd{(v5), dd}\hlopt{$}\hlstd{RT_RecSqt)}\hlopt{^}\hlnum{2}
-\end{alltt}
-\begin{verbatim}
-## [1] 0.5095834
+R^2
+[1] 0.533722
 \end{verbatim}
@@
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlkwd{par}\hlstd{(}\hlkwc{mfrow}\hlstd{=}\hlkwd{c}\hlstd{(}\hlnum{3}\hlstd{,}\hlnum{4}\hlstd{),} \hlkwc{family}\hlstd{=}\hlstr{'cmodern'}\hlstd{)}
-\hlkwd{plotLMER.fnc}\hlstd{(v5)}
-\end{alltt}
-\begin{verbatim}
-## effect size (range) for  c_Vag is  0.03470056 
-## effect size (range) for  c_Num is  0.4073765 
-## effect size (range) for  c_Qty is  0.09530422 
-## effect size (range) for  c_Ord is  0.0179595 
-## effect size (range) for  item_mean_ratio is  0.2346348 
-## effect size (range) for  s_Trl is  0.369093 
-## effect size (range) for  RTprev_log is  0.2759498 
-## effect size (range) for  nchar_instr is  0.05477539
-\end{verbatim}
+\clearpage
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot.new(): figure margins too large}}\end{kframe}
-\end{knitrout}
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
 
-Plot model coefficients and ci's
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-plotModelCoefsAndCis-1}
 
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
-\begin{alltt}
-\hlstd{e}\hlkwb{=}\hlkwd{data.frame}\hlstd{(}\hlkwc{coef}\hlstd{=}\hlkwd{summary}\hlstd{(v5)}\hlopt{$}\hlstd{coef[}\hlopt{-}\hlnum{1}\hlstd{,}\hlnum{1}\hlstd{])} \hlcom{# estimates, without intercept}
-\hlstd{q}\hlkwb{=}\hlkwd{as.data.frame}\hlstd{(}\hlkwd{confint}\hlstd{(v5,} \hlkwc{method}\hlstd{=}\hlstr{'Wald'}\hlstd{)[}\hlnum{18}\hlopt{:}\hlnum{26}\hlstd{,}\hlnum{1}\hlopt{:}\hlnum{2}\hlstd{])}
-\hlstd{eq}\hlkwb{=}\hlkwd{cbind}\hlstd{(e,q)}
-\hlstd{y}\hlkwb{=}\hlstd{eq[}\hlkwd{rev}\hlstd{(}\hlkwd{rownames}\hlstd{(eq)),]}
-\hlkwd{par}\hlstd{(}\hlkwc{mar}\hlstd{=}\hlkwd{c}\hlstd{(}\hlnum{4}\hlstd{,}\hlnum{10}\hlstd{,}\hlnum{1}\hlstd{,}\hlnum{1}\hlstd{))}
-\hlkwd{plot}\hlstd{(}\hlkwc{y}\hlstd{=}\hlnum{1}\hlopt{:}\hlkwd{nrow}\hlstd{(y),} \hlkwc{x}\hlstd{=y}\hlopt{$}\hlstd{coef,} \hlkwc{xlim}\hlstd{=}\hlkwd{range}\hlstd{(y}\hlopt{$}\hlstr{"2.5 %"}\hlstd{, y}\hlopt{$}\hlstr{"97.5 %"}\hlstd{),} \hlkwc{type}\hlstd{=}\hlstr{'n'}\hlstd{,} \hlkwc{axes}\hlstd{=F,} \hlkwc{xlab}\hlstd{=}\hlstr{""}\hlstd{,} \hlkwc{ylab}\hlstd{=}\hlstr{""}\hlstd{)}
-\hlkwd{segments}\hlstd{(}\hlkwc{y0}\hlstd{=}\hlnum{1}\hlopt{:}\hlkwd{nrow}\hlstd{(y),} \hlkwc{y1}\hlstd{=}\hlnum{1}\hlopt{:}\hlkwd{nrow}\hlstd{(y),} \hlkwc{x0}\hlstd{=y}\hlopt{$}\hlstr{"2.5 %"}\hlstd{,} \hlkwc{x1}\hlstd{=y}\hlopt{$}\hlstr{"97.5 %"}\hlstd{,} \hlkwc{lwd}\hlstd{=}\hlnum{.55}\hlstd{)}
-\hlkwd{points}\hlstd{(}\hlkwc{y}\hlstd{=}\hlnum{1}\hlopt{:}\hlkwd{nrow}\hlstd{(y),} \hlkwc{x}\hlstd{=y}\hlopt{$}\hlstd{coef,} \hlkwc{pch}\hlstd{=}\hlnum{20}\hlstd{,} \hlkwc{cex}\hlstd{=}\hlnum{.75}\hlstd{)}
-\hlkwd{abline}\hlstd{(}\hlkwc{v}\hlstd{=}\hlnum{0}\hlstd{,} \hlkwc{lty}\hlstd{=}\hlnum{3}\hlstd{,} \hlkwc{col}\hlstd{=}\hlstr{'grey'}\hlstd{)}
-\hlkwd{axis}\hlstd{(}\hlnum{2}\hlstd{,} \hlkwc{labels}\hlstd{=}\hlkwd{row.names}\hlstd{(y),} \hlkwc{at}\hlstd{=}\hlnum{1}\hlopt{:}\hlkwd{nrow}\hlstd{(y),} \hlkwc{las}\hlstd{=}\hlnum{1}\hlstd{)}
-\hlkwd{axis}\hlstd{(}\hlnum{1}\hlstd{,} \hlkwc{las}\hlstd{=}\hlnum{1}\hlstd{)}
-\hlkwd{mtext}\hlstd{(}\hlstr{"estimate"}\hlstd{,} \hlkwc{side}\hlstd{=}\hlnum{1}\hlstd{,} \hlkwc{line}\hlstd{=}\hlnum{2.5}\hlstd{,} \hlkwc{cex.lab}\hlstd{=}\hlnum{1}\hlstd{,} \hlkwc{las}\hlstd{=}\hlnum{1}\hlstd{)}
-\hlkwd{box}\hlstd{(}\hlkwc{col}\hlstd{=}\hlstr{'grey'}\hlstd{)}
-\end{alltt}
-\end{kframe}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/v54-1} \caption[Estimates plot]{Estimates plot}\label{fig:v54}
+}
+
+\caption[Coefficient estimates and their (Wald) 95 per cent confidence intervals]{Coefficient estimates and their (Wald) 95 per cent confidence intervals}\label{fig:plotModelCoefsAndCis}
 \end{figure}
@@
 
-Model criticism baayen plots
+\clearpage
 
-\begin{knitrout}
+\begin{knitrout}\scriptsize
 \definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 \begin{alltt}
-\hlcom{# Baayen 4-plot model criticism}
-\hlkwd{par}\hlstd{(}\hlkwc{mfrow}\hlstd{=}\hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,}\hlnum{4}\hlstd{),} \hlkwc{pty}\hlstd{=}\hlstr{'s'}\hlstd{)}
-\hlcom{# create scaled residuals}
-\hlstd{dd}\hlopt{$}\hlstd{rstand} \hlkwb{=} \hlkwd{as.vector}\hlstd{(}\hlkwd{scale}\hlstd{(}\hlkwd{resid}\hlstd{(v5)))}
-\hlcom{# plot scaled residuals density}
-\hlkwd{plot}\hlstd{(}\hlkwd{density}\hlstd{(dd}\hlopt{$}\hlstd{rstand))}
-\hlcom{# plot sample quantiles versus theoretical quantiles}
-\hlkwd{qqnorm}\hlstd{(dd}\hlopt{$}\hlstd{rstand,} \hlkwc{cex}\hlstd{=}\hlnum{.5}\hlstd{)}
-\hlkwd{qqline}\hlstd{(dd}\hlopt{$}\hlstd{rstand)}
-\hlcom{# plot standardised residuals versus fitted values}
-\hlkwd{plot}\hlstd{(dd}\hlopt{$}\hlstd{rstand} \hlopt{~} \hlkwd{fitted}\hlstd{(v5),} \hlkwc{pch}\hlstd{=}\hlstr{'.'}\hlstd{)}
-\hlcom{# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot}
-\hlkwd{abline}\hlstd{(}\hlkwc{h}\hlstd{=}\hlkwd{c}\hlstd{(}\hlopt{-}\hlnum{2.5}\hlstd{,}\hlnum{2.5}\hlstd{))}
+\hlkwd{par}\hlstd{(}\hlkwc{mfrow} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{2}\hlstd{,} \hlnum{4}\hlstd{))}
+\hlkwd{plotLMER.fnc}\hlstd{(v5)}
 \end{alltt}
-\end{kframe}\begin{figure}
-\includegraphics[width=\maxwidth]{figure/baayenPLots99-1} \caption[caption]{caption}\label{fig:baayenPLots99}
+\begin{verbatim}
+effect size (range) for  c_Vag is  0.03470056 
+effect size (range) for  c_Num is  0.4073765 
+effect size (range) for  c_Qty is  0.09530422 
+effect size (range) for  c_Ord is  0.0179595 
+effect size (range) for  discriminability is  0.2346348 
+effect size (range) for  s_Trl is  0.369093 
+effect size (range) for  RTprev_log is  0.2759499 
+effect size (range) for  nchar_instr is  0.05477539 
+\end{verbatim}
+\end{kframe}\begin{figure}[hbtp]
+
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-plotLMERfnc-1}
+
+}
+
+\caption[plotMLERfnc]{plotMLERfnc}\label{fig:plotLMERfnc}
 \end{figure}
@@
 
+\clearpage
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-baayenPLots99-1}
+
+}
+
+\caption[Baayen Model Criticism Plots]{Baayen Model Criticism Plots}\label{fig:baayenPLots99}
+\end{figure}
+
+\end{knitrout}
+
+\clearpage
+
+\section{lmerTest Version}
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{v6} \hlkwb{<-} \hlstd{lmerTest}\hlopt{::}\hlkwd{lmer}\hlstd{(}\hlkwc{data} \hlstd{= dd, RT_log} \hlopt{~} \hlstd{c_Vag} \hlopt{+} \hlstd{c_Num} \hlopt{+} \hlstd{c_Qty} \hlopt{+} \hlstd{c_Ord} \hlopt{+} \hlstd{c_Num}\hlopt{:}\hlstd{c_Vag}\hlopt{:}\hlstd{c_Qty} \hlopt{+} \hlstd{discriminability} \hlopt{+}
+    \hlstd{s_Trl} \hlopt{+} \hlstd{RTprev_log} \hlopt{+} \hlstd{nchar_instr} \hlopt{+} \hlstd{(}\hlnum{1} \hlopt{+} \hlstd{c_Vag} \hlopt{+} \hlstd{c_Num} \hlopt{+} \hlstd{c_Qty} \hlopt{+} \hlstd{c_Ord} \hlopt{|} \hlstd{Subject))}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{summary}\hlstd{(v6)}
+\end{alltt}
+\begin{verbatim}
+Linear mixed model fit by REML t-tests use Satterthwaite approximations to degrees of freedom [
+lmerMod]
+Formula: RT_log ~ c_Vag + c_Num + c_Qty + c_Ord + c_Num:c_Vag:c_Qty +  
+    discriminability + s_Trl + RTprev_log + nchar_instr + (1 +  
+    c_Vag + c_Num + c_Qty + c_Ord | Subject)
+   Data: dd
+
+REML criterion at convergence: 11474.8
+
+Scaled residuals: 
+    Min      1Q  Median      3Q     Max 
+-4.5470 -0.6351 -0.0955  0.5372  5.0914
+
+Random effects:
+ Groups   Name        Variance Std.Dev. Corr                   
+ Subject  (Intercept) 0.153949 0.39236                         
+          c_Vag       0.001546 0.03932   0.69                  
+          c_Num       0.165314 0.40659  -0.67 -0.64            
+          c_Qty       0.008148 0.09027   0.16  0.26 -0.34      
+          c_Ord       0.001559 0.03949  -0.13  0.02 -0.39 -0.52
+ Residual             0.249734 0.49973                         
+Number of obs: 7677, groups:  Subject, 30
+
+Fixed effects:
+                    Estimate Std. Error         df t value Pr(>|t|)    
+(Intercept)        7.172e+00  1.222e-01  2.370e+02  58.671  < 2e-16 ***
+c_Vag              6.094e-02  1.387e-02  3.300e+01   4.393 0.000112 ***
+c_Num             -4.336e-01  7.515e-02  2.900e+01  -5.770 2.97e-06 ***
+c_Qty             -6.907e-02  2.005e-02  2.900e+01  -3.445 0.001743 ** 
+c_Ord              1.796e-02  1.350e-02  5.100e+01   1.331 0.189164    
+discriminability  -7.714e-01  4.927e-02  7.551e+03 -15.658  < 2e-16 ***
+s_Trl             -1.070e-01  5.807e-03  7.558e+03 -18.421  < 2e-16 ***
+RTprev_log         6.047e-02  9.692e-03  7.594e+03   6.239 4.63e-10 ***
+nchar_instr        6.086e-03  1.944e-03  7.551e+03   3.131 0.001749 ** 
+c_Vag:c_Num:c_Qty  1.049e-01  4.607e-02  7.551e+03   2.278 0.022742 *  
+---
+Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
+
+Correlation of Fixed Effects:
+            (Intr) c_Vag  c_Num  c_Qty  c_Ord  dscrmn s_Trl  RTprv_ nchr_n
+c_Vag        0.083                                                        
+c_Num       -0.369 -0.337                                                 
+c_Qty        0.069  0.117 -0.278                                          
+c_Ord       -0.052  0.007 -0.206 -0.228                                   
+discrmnblty -0.140  0.004 -0.001  0.000  0.001                            
+s_Trl       -0.114  0.003 -0.001 -0.002  0.006  0.017                     
+RTprev_log  -0.607  0.006 -0.004 -0.003  0.019  0.015  0.185              
+nchar_instr -0.524  0.237 -0.034  0.018  0.000  0.017  0.000  0.006       
+c_Vg:c_N:_Q  0.073 -0.033  0.005 -0.003  0.000 -0.002 -0.001 -0.002 -0.138
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+\clearpage
+\section{Borderline responses}
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-barplotBorderline-1}
+
+}
+
+\end{knitrout}
+
+% latex table generated in R 3.3.1 by xtable 1.8-2 package
+% Fri Aug 12 22:28:47 2016
+\begin{table}[htbp]
+\centering
+\begingroup\small
+\begin{tabular}{rrr}
+  \hline
+ & Crisp & Vague \\ 
+  \hline
+expected & 3332 & 2776 \\ 
+  borderline & 433 & 841 \\ 
+  extreme &  75 & 220 \\ 
+   \hline
+\end{tabular}
+\endgroup
+\caption{Borderline cases counts} 
+\end{table}
+
 \end{document}

### 2016-08-12 experiments/e2/writeup/scratch.Rnw
more consistency, better plots, using discriminability instead of item ratio

<!--html_document:-->
  <!--toc: true-->
  <!--toc_float: true-->
  <!--mathjax: default-->
  <!--template: default-->
  <!--code_folding: hide-->
  <!--theme: default-->

  ```{r print_info_on_design}
cat(paste('number of subjects = ',30,'\n',
          '---','\n',
          'number of trials per subject = ',256,'\n', 
          '---','\n',          
          'number of unique cells in the design = Item(4) * Number(2) * Vagueness(2) * Order(2) * Quantity(2) = ',4*2*2*2*2,'\n',
          'number of times a subject saw a unique cell in the design = 256 / 64 = ', 256/64,'\n',
          '---','\n',
          'number of higher-level cells in the design (abstracting over Order and Quantity) = Item(4) * Number(2) * Vagueness(2) = ',4*2*2,'\n',
          'number of times a subject saw a higher level cell in the design =  = 256 / 16 = ', 4*2*2, '\n',
          sep=''))
```

# Distributional stuff, including transformations.

```{r transformations_table, results='asis'}
x=445
lambda = c(-2,-1,-0.5, 0,0.5,1,2)

xd=data.frame(
  lambda = c(
    lambda[1],
    lambda[2],
    lambda[3],
    lambda[4],
    lambda[5],
    lambda[6],
    lambda[7]
  ),
  x=c(    
    paste("x = ", sep="", x),
    paste("x = ", sep="", x),
    paste("x = ", sep="", x),
    paste("x = ", sep="", x),
    paste("x = ", sep="", x),
    paste("x = ", sep="", x),
    paste("x = ", sep="", x)
  ),
  conventional=c(
    '$$y=\\frac{1}{x^2}$$', 
    '$$y=\\frac{1}{x}$$', 
    '$$y=\\frac{1}{\\sqrt{x}}$$', 
    '$$y=log_e(x)$$', 
    '$$y=\\sqrt{x}$$', 
    '$$y=x$$', 
    '$$y=x^2$$'
  ),
  power=c(
    '$$y=x^{-2}$$', 
    '$$y=x^{-1}$$', 
    '$$y=x^{-0.5}$$', 
    '$$y=log_e(x)$$', 
    '$$y=x^{0.5}$$', 
    '$$y=x^{1}$$', 
    '$$y=x^{2}$$'
  ),
  y=c(     
    paste("y = ", sep="", format(round(x^lambda[1],6), sc=F) ),
    paste("y = ", sep="", round(x^lambda[2],3) ),
    paste("y = ", sep="", round(x^lambda[3],3) ),
    paste("y = ", sep="", round(log(x),3) ),
    paste("y = ", sep="", round(x^lambda[5],2) ),
    paste("y = ", sep="", round(x^lambda[6],6) ),
    paste("y = ", sep="", round(x^lambda[7],6) )
  ),
  back = c(
    '$$x=y^{\\frac{1}{-2}}$$',
    '$$x=y^{\\frac{1}{-1}}$$', 
    '$$x=y^{\\frac{1}{-0.5}}$$', 
    '$$x=exp(y)$$',  
    '$$x=y^{\\frac{1}{0.5}}$$', 
    '$$x=y^{\\frac{1}{1}}$$',  
    '$$x=y^{\\frac{1}{2}}$$'
  ),
  back_lambda=c(
    '$$x=y^{1/\\lambda}$$',
    '$$x=y^{1/\\lambda}$$',
    '$$x=y^{1/\\lambda}$$',
    '$$x=exp(y)$$',  
    '$$x=y^{1/\\lambda}$$',
    '$$x=y^{1/\\lambda}$$',
    '$$x=y^{1/\\lambda}$$'
  ),
  y_back=c(  
    paste( 'x = ', (x^lambda[1]) ^ {1/lambda[1]} ,sep=""),
    paste( 'x = ', (x^lambda[2]) ^ {1/lambda[2]} ,sep=""),
    paste( 'x = ', (x^lambda[3]) ^ {1/lambda[3]} ,sep=""),
    paste( 'x = ', exp(log(x)), sep=""),
    paste( 'x = ', (x^lambda[5]) ^ {1/lambda[5]} ,sep=""),
    paste( 'x = ', (x^lambda[6]) ^ {1/lambda[6]} ,sep=""),
    paste( 'x = ', (x^lambda[7]) ^ {1/lambda[7]} ,sep="")
  )
)

print(xtable(xd, digits=1, align=c(rep('l', times=ncol(xd)+1))), type='html', include.rownames=F, digits=1)
```

Show practice and fatigue effects vary over items.

```{r pracFatigueItms, eval=TRUE, fig.width=7, fig.height=3.5, echo=FALSE}
# Show how practice/fatigue effects vary over items
ggplot(dd, aes(y=RT_RecSqt, x=Trial)) + 
  facet_grid(~Item) + 
  theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), 
        panel.background = element_rect(fill = 'white', colour='black' ), 
        legend.position="top") + 
  scale_x_continuous("Trial") + 
  geom_point(pch=19, cex=.25, alpha=.5) + 
  stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + 
  theme(aspect.ratio = 1) 
```

## ols model v4, RecSqt, in progress

Pairs plot

```{r pairs, fig.width=7, fig.height=7}
pairs(dd[,c("RT_RecSqt", "Number", "Vagueness", "RTprev_RecSqt")], pch = 19, cex=.075)
```

Build the model

```{r doModelols}
## do ols model
# define datadist
dd.dd = datadist(dd)
options(datadist = "dd.dd")
# build model
v4 <- ols(data=dd, 
          RT_RecSqt ~ 
            c_Vag + c_Num + c_Qty + c_Ord + 
            c_Num:c_Vag:c_Qty+
            item_mean_ratio + 
            #c_Vag:c_Num + 
            #c_Vag:c_Num:item_mean_ratio + 
            # pol(s_Trl, 3) +
            s_Trl+
            # pol(RTprev_RecSqt,3) +
            RTprev_RecSqt +
            nchar_instr
          # cell+
          # measurement,
          ,
          x=T, y=T )
```

```{r likRat}
# print out p value for likelihood ratio with df
cat("lik.ratio p value, whether the model as a whole is explanatory: p =", 1-pchisq(v4$stats[2],v4$stats[3]))
# print out R2 for the model
cat("R^2 = ", v4$stats[4])
```

Show the model tables

```{r modTabs}
# show model
v4
# show model summary
summary(v4)
# do anova 
an1 = anova(v4)
# show anova
an1
```

plot 'partial effects'

```{r partial, fig.width=12, fig.height=5}
## plot partial effects
# Compute Predicted Values and Confidence Limits
p1=Predict(v4)
# plot 'partial effects'
plot(p1, anova=an1, pval=T, aspect=1, main="ols model v4")
```

```{r daz99}
# add predicted RT_RecSqt to dd
dd$RT_Predicted <- predict (v4)
```

Plot model coefficients and ci's

```{r coefsAndCiS}
par(las=2, mar=c(14,6,1,1))
y=coef(v4) # with intercept
n=length(y)
y0=confint(v4)[,1]
y1=confint(v4)[,2]
y=coef(v4)[-1] # omit intercept
n=length(y)
y0=confint(v4)[-1,1]
y1=confint(v4)[-1,2]
plot(y, xaxt="n", xlab="", ylab="RT_RecSqt\n", pch=19, ylim=extendrange(y, f=.10), main="Speed")
abline(h=0)
axis(1, labels=names(y) , at=1:n)
grid()
segments(x0=1:n, x1=1:n, y0, y1, lwd=2)
```

Collinearity is assessed by the condition number $k$.\
The greater the collinearity, the closer the matrix of predictors is is to becoming _singular_.\
$k$ estimates the extent to which a matrix is $\text{singular}$\
$0\ldots6$ no collinearity\
$\text{around} 15$ medium collinearity\
$>30$ indicates potentially harmful collinearity\

```{r kis}
cat("k is:",
collin.fnc(dd[,	c("s_Trl", "c_Num", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", 
"RTprev_RecSqt", "nchar_instr", "measurement")
])$cnumber
)
```

```{r plotg, fig.width=6, fig.height=5}
par(mar=c(1.1,3.1,1.1,1.1), pty='s')
plot(varclus(
as.matrix(dd[,c("s_Trl", "c_Num", "c_Vag", "c_Qty", "c_Ord",
"item_mean_ratio", "RTprev_RecSqt", "nchar_instr", "measurement")])))
```

Model criticism baayen plots

```{r baayenMplots, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism
par(mfrow=c(1,4), pty='s')
# create scaled residuals
v4$rstand = as.vector(scale(resid(v4)))
# plot scaled residuals density
plot(density(v4$rstand), main='density of\nscaled residuals')
# plot sample quantiles versus theoretical quantiles
qqnorm(v4$rstand, cex=.5)
qqline(v4$rstand)
# plot standardised residuals versus fitted values
plot(v4$rstand ~ fitted(v4), pch='.', main='std resid\nv fitted')
# absolute standardised residuals greater than 2.5 are candidates 
# for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))
# create dffits
dffits=abs(resid(v4, "dffits"))
# plot dffits
plot(dffits, type='h', main='dffits')
# here no overly influential
w = which.influence(fit = v4, cutoff = 0.4)
if (length(w)==0) {
print('There are no overly influential data points')
} else {
print('There are some overly influential variables:');
print(w)
}
```

Model validation

Validation tests overfitting using bootstrap.

Bootstrapping draws as a bootstrap sample the same number of samples as the original data, at random with replacement, from the original data.

Then fit the model to the data in the bootstrap sample, and use this model to predict the reaction times for the original full data (which contains many samples that are new to the bootstrap model).

Then compare the goodness of fit of the bootstrap model with the goodness of fit of the original model.

Averaged over a large number of bootstrap models these comparisons of goodness of fit  reveal to what extent the original model overfits the data.

Function validate() in rms does this _re-sampling validation_.

```{r val_1, echo=TRUE}
# argument B is the number of bootstrap runs
# argument pr is whether to print the results of each run
v_1 <- validate(v4, bw = T, B = 200, pr=FALSE, estimates=TRUE)
```

```{r prval}
# in print.validate, B is the number of re-samples to show outocmes for each resample
print(v_1, B=0)
```

```{r kept}
# show how often each number of variables kept was observed across the 200 runs
xtabs(~rowSums(attr(v_1,"kept")))
```

<<exampleOfFonts, fig.width=3, fig.height=3, fig.cap='font example', echo=FALSE, eval=FALSE>>=
p <- ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point() +
ggtitle("Fuel Efficiency of 32 Cars") +
xlab("Weight (x1000 lb)") + ylab("Miles per Gallon") +
theme_bw() +
theme(text=element_text(family="CMU Serif", size=14, face='plain'))
p
ggsave("ggplot_CMU_Serif.pdf", p, width=3.5, height=3.5)
@

<<defineFonts>>=
# .define.fonts <- function () {
#     quartzFonts(avenir =    c('Avenir Book', 'Avenir Black', 'Avenir Book Oblique', 'Avenir Black Oblique'),
#                 helvetica = c('Helvetica New Light', 'Helvetica New Bold', 'Helvetica New Light Italic', 'Helvetica New Bold Italic'),
#                 cmodern = c('CMU Serif', 'CMU Serif Extra','CMU Classical Serif','CMU Serif Extra'))
# }
# .define.fonts()
@

<<fig.cap='base graphic special font'>>=
#base graphic special font
pdf("plot_cm.pdf", family="CM Roman", width=5.5, height=5)

curve(dnorm, from=-3, to=3, main="Normal Distribution")
text(x=0, y=0.1, cex=1.5, expression(italic(y == frac(1, sqrt(2 * pi)) *
e ^ {-frac(x^2, 2)} )))

dev.off()
embed_fonts("plot_cm.pdf", outfile="plot_cm_embed.pdf")
@

### 2016-08-12 experiments/e2/writeup/Untitled.R

@@ -1,0 +1,9 @@
+# add discriminability with reference to item
+discriminability_range = c(0.7500000, 0.5294118, 0.4090909, 0.3333333)
+discriminability_range_scaled = c(1.3441995, 0.1316642, -0.5297187, -0.9461450)  
+discriminability = c(0.4875000, 0.3123529, 0.2308442, 0.1833333)
+discriminability_scaled = c(1.37582241, 0.06614191, -0.54334858, -0.89861574)
+dat[dat$Item == "06:15:24", "discriminability"] <- 0.4875000
+dat[dat$Item == "16:25:34", "discriminability"] <- 0.3123529
+dat[dat$Item == "26:35:44", "discriminability"] <- 0.2308442
+dat[dat$Item == "36:45:54", "discriminability"] <- 0.1833333

### 2016-08-12 experiments/e2/writeup/measurement.R

@@ -1,0 +1,25 @@
+# number of higher-level cells in the design (abstracting over Order) =
+# Item(4) * Number(2) * Vagueness(2) * Quantity(2) = 32
+# number of measurements in eaach cell = 8
+# 332 * 8 = 256
+# 256 * 30 = 7680
+dd$measurement=0
+dd$cell=0
+for (s in levels(dd$Subject)) {
+  cellcount=0
+  for (i in levels(dd$Item)) {
+    for (n in levels(dd$Number)) {
+      for (v in levels(dd$Vagueness)) {
+        for (q in levels(dd$Quantity)) {
+          measurementcount=0
+          cellcount=cellcount+1
+          for (t in dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q, "Trial"]) {
+            dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "cell"] <- cellcount
+            measurementcount=measurementcount+1
+            dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "measurement"] = measurementcount
+          }
+        }
+      }
+    }
+  }
+}

### 2016-08-13 experiments/e2/writeup/e2-b-knitr.Rnw
tidying up

@@
-\documentclass[a4paper,12pt]{article}
+\documentclass[a4paper,12pt,twoside]{article}
 \usepackage[margin=2cm]{geometry}
 \usepackage{placeins}
-
+\usepackage{graphicx}
+\usepackage{color}
+\usepackage{hyperref}
 \title{Exp 2 of 4}
@@
                                 width.cutoff=100))
-knit_hooks$set(crop=hook_pdfcrop)
+#knit_hooks$set(crop=hook_pdfcrop)
 @
@@
 
+\clearpage
+
 <<showDataNames>>=
@@
 
+\clearpage
+
 <<showDataHead>>=
@@
 
+\clearpage
+
 <<showDataSummary>>=
@@
 @
+
+\clearpage
 
@@
 axis(1)
-box(col='grey')
+box(which='plot', col='grey')
+
 #text(y=1.00, x=0.07, labels="mean\n(ratio min/med,\nratio med/max)", col='red')
@@
 \clearpage
+\subsection{Consider using log RT}
 
@@
         legend.position='top',
-        legend.key=element_rect(color = 'white'))
-@
-
-\clearpage
+        legend.key=element_rect(color = 'white'),
+        plot.background=element_rect(fill=NA, color='white'))
+@
+
+\clearpage
+\subsection{How logging RT affects the distribution}
 
@@
         legend.position='top',
-        legend.key=element_rect(color = 'white')) + 
+        legend.key=element_rect(color = 'white'),
+        plot.background=element_rect(fill=NA, color='white')) + 
   scale_colour_manual("",values=c("Items"="orange","Subjects"="blue")) +
@@
 \clearpage
+\subsection{Identify fast and slow subjects and items}
 
@@
               xlab=NULL,
-              ylab='Log RT',
+              ylab='Log RT\n',
               #main='How mean times for individual subjects and items\nvary with respect to the grand mean Log RT',
@@
 \clearpage
-
-<<plotMainsa, fig.width=6, fig.height=4, fig.cap='Plot main effects on several transformations', fig.pos='hbtp', message=FALSE, echo=FALSE, warning=FALSE>>=
+\subsection{Plot main effects in both transformations}
+
+<<prepplotsMainEffects, message=FALSE, echo=FALSE, warning=FALSE>>=
 # plot main effects on observed data
 # allow reference to transformation
-temp <- reshape2::melt(dd,
+# raw
+tempraw <- reshape2::melt(dd,
                           id.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
-                       measure.vars=c("RT_log",  "RT_raw"),
+                          measure.vars=c("RT_raw"),
                           variable.name="transformation",
@@
 # allow reference to effect
-temp2 <- reshape2::melt(temp,
+tempraw2 <- reshape2::melt(tempraw,
                            id.vars=c("transformation", "score"),
@@
                            value.name="level")
-# do the plot
-ggplot(temp2, aes(y=score, x=level, group=1, col=effect)) +
-  facet_wrap(~transformation + effect, scales="free", ncol=5) +
+# log
+templog <- reshape2::melt(dd,
+                          id.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
+                          measure.vars=c("RT_log"),
+                          variable.name="transformation",
+                          value.name="score")
+# allow reference to effect
+templog2 <- reshape2::melt(templog,
+                           id.vars=c("transformation", "score"),
+                           measure.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
+                           variable.name="effect",
+                           value.name="level")
+@
+
+<<plotMainEffectsRaw, fig.width=6, fig.height=3, fig.cap='Plot main effects in raw RT', fig.pos='hbtp', echo=FALSE>>=
+# raw
+ggplot(tempraw2, aes(y=score, x=level, group=1, col=effect)) +
+  facet_wrap(~effect, scales="free", ncol=5) +
   stat_summary(fun.data=mean_cl_normal, geom="pointrange") +
@@
   theme(aspect.ratio=1, 
-        axis.text.y=element_blank(),
-        axis.ticks.y=element_blank(),
+        # axis.text.y=element_blank(),
+        # axis.ticks.y=element_blank(),
+        panel.grid=element_blank(),
         axis.text.x = element_text(angle = 45, hjust = 1, size=8),
-        legend.position="none") +
-  ylab('Response latency\n') + xlab('')
-@
-
-\clearpage
-
-<<mainB, fig.width=6, fig.height=4, fig.cap='Just the main effects', fig.pos='hbtp', message=FALSE, echo=FALSE>>=
-# prep main effects rts data frame
-vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness")));names(vrts)[3]<-"Levels"
-nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number")));names(nrts)[3]<-"Levels"
-qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Quantity")));names(qrts)[3]<-"Levels"
-orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Order")));names(orts)[3]<-"Levels"
-irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Item")));names(irts)[3]<-"Levels"
-rts=rbind(vrts,nrts,qrts,orts)
-rts$mins=rts$RT_log-rts$ci
-rts$maxs=rts$RT_log+rts$ci
-# plot main effects
-ggplot(rts, aes(y=RT_log, x=Levels, group=1, ymin=mins, ymax=maxs)) +
-  geom_line() +
-  geom_errorbar(width=.05) +
-  geom_point(pch=19) +
-  facet_grid(~Effect, scale="free_x") +
+        legend.position="none",
+        plot.background=element_rect(fill=NA, color='white')) +
+  xlab('')
+@
+
+<<plotMainEffectsLog, fig.width=6, fig.height=3, fig.cap='Plot main effects in log RT', fig.pos='htbp', echo=FALSE>>=
+# log
+ggplot(templog2, aes(y=score, x=level, group=1, col=effect)) +
+  facet_wrap(~effect, scales="free", ncol=5) +
+  stat_summary(fun.data=mean_cl_normal, geom="pointrange") +
+  stat_summary(fun.y=mean, geom="line") +
   theme_bw() +
   theme(aspect.ratio=1, 
-        axis.text.x = element_text(size=8),
-        axis.text.y = element_text(size=8, angle=15),
-        panel.grid=element_blank()) +
-  ylab("RT (log)") + xlab('')
-@
-
-\clearpage
-
-<<mainD, fig.width=6, fig.height=5, fig.cap='Main effects over item ratios', fig.pos='hbtp', message=FALSE, echo=FALSE>>=
+        # axis.text.y=element_blank(),
+        # axis.ticks.y=element_blank(),
+        panel.grid=element_blank(),
+        axis.text.x = element_text(angle = 45, hjust = 1, size=8),
+        legend.position="none",
+        plot.background=element_rect(fill=NA, color='white')) +
+  xlab('')
+@
+
+\clearpage
+\subsection{Main effects in log RT over discriminability}
+
+<<mainD, fig.width=6, fig.height=5, fig.cap='Main effects in log RT over discriminability', fig.pos='hbtp', message=FALSE, echo=FALSE>>=
 nrts <- cbind(Effect="Number",
@@
         axis.text.y=element_text(size=8, angle=15),
-        legend.key=element_rect(color = 'white')
+        legend.key=element_rect(color = 'white'),
+        plot.background=element_rect(fill=NA, color='white')
         #,        panel.grid=element_blank()
         ) +
-  ylab("RT (log)") + xlab("Item discriminability measure")
+  xlab("Item discriminability measure")
 p1
@@
 \clearpage
-
-<<22way, fig.width=7, fig.height=7, fig.cap='2-way interactions over item ratios', fig.pos='hbtp', message=FALSE, echo=FALSE>>=
+\subsection{2-Way interactions over discriminability}
+
+<<22way, fig.width=7, fig.height=7, fig.cap='2-way interactions over discriminability', fig.pos='hbtp', message=FALSE, echo=FALSE>>=
 # prep 2 ways over items rts data frame
@@
     , axis.text.x=element_text(size=7, angle=45,hjust=1)
-    , axis.text.y=element_text(size=7)
+    , axis.text.y=element_text(size=7),
+    plot.background=element_rect(fill=NA, color='white')
     ) +
@@
 \clearpage
+\subsection{Vagueness by number interaction over items}
 
@@
         legend.key = element_blank(), aspect.ratio=1,
-        axis.text.x = element_text(angle = 45, hjust = 1)) +
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        plot.background=element_rect(fill=NA, color='white')) +
   facet_grid(~Number) 
@@
 \clearpage
-
-<<plotInt2, fig.width=6, fig.height=4, fig.cap='vagueness by number interaction over item ratios', fig.pos='hbtp', message=FALSE, echo=FALSE>>=
+\subsection{Vagueness by number interaction over discriminability}
+
+<<plotInt2, fig.width=6, fig.height=4, fig.cap='vagueness by number interaction over discriminability', fig.pos='hbtp', message=FALSE, echo=FALSE>>=
 rts <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number","Vagueness","discriminability"))
@@
         legend.key = element_blank(), aspect.ratio=1,
-        axis.text.x = element_text(angle = 45, hjust = 1)) +
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        plot.background=element_rect(fill=NA, color='white')) +
   facet_grid(~Number) +
@@
 \clearpage
-
-<<plot3way, fig.width=7, fig.height=7, fig.pos='hbtp', fig.cap='3 way interactions',echo=FALSE>>=
+\subsection{3-Way interactions}
+
+<<plot3way, fig.width=7, fig.height=7, fig.pos='hbtp', fig.cap='3-Way interactions',echo=FALSE>>=
 rts <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number","Vagueness","Quantity"))
@@
   theme(aspect.ratio=1) +
-  theme(legend.title=element_blank(), legend.position="top")
+  theme(legend.title=element_blank(), legend.position="top", plot.background=element_rect(fill=NA, color='white'))
 rts <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number","Vagueness","Order"))
@@
   theme(aspect.ratio=1) +
-  theme(legend.title=element_blank(), legend.position="top")
+  theme(legend.title=element_blank(), legend.position="top", plot.background=element_rect(fill=NA, color='white'))
 rts <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number","Quantity","Order"))
@@
   theme(aspect.ratio=1) +
-  theme(legend.title=element_blank(), legend.position="top")
+  theme(legend.title=element_blank(), legend.position="top", plot.background=element_rect(fill=NA, color='white'))
 rts <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Quantity","Order"))
@@
   theme(aspect.ratio=1) +
-  theme(legend.title=element_blank(), legend.position="top")
+  theme(legend.title=element_blank(), legend.position="top", plot.background=element_rect(fill=NA, color='white'))
 grid.arrange(nrow=2, p1,p2,p3,p4)
@@
 \clearpage
-
-<<vnitems, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='Plot vagueness by number by quantity over item ratios.',message = FALSE, echo=FALSE>>=
+\subsection{Vagueness by number by quantity over discriminability}
+
+<<vnitems, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='Vagueness by number by quantity over discriminability',message = FALSE, echo=FALSE>>=
 rts <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number","Vagueness","discriminability", "Quantity"))
@@
   ggtitle("Response time") +
-  ylab("RT_Rec_Sqt") +
+  #ylab("RT_Rec_Sqt") +
   xlab("Item discriminability") +
@@
         aspect.ratio=1, 
-        axis.text.x = element_text(angle = 45, hjust = 1)) +
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        plot.background=element_rect(fill=NA, color='white')) +
   facet_grid(~Number+Quantity) +
@@
 \clearpage
-
-<<4way, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='4 way interaction', echo=FALSE>>=
+\subsection{4-Way interaction}
+
+<<4way, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='4-Way interaction', echo=FALSE>>=
 rts <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number", "Vagueness",  "Quantity", "Order"))
@@
         legend.position='right',
-        legend.key=element_rect(color = 'white'))  
-@
-
-\clearpage
-
-<<4waysplit, fig.width=7, fig.height=7, fig.cap='4 way interaction split over item discriminability', fig.pos='hbtp', message=FALSE, echo=FALSE>>=
+        legend.key=element_rect(color = 'white'),
+        plot.background=element_rect(fill=NA, color='white'))  
+@
+
+\clearpage
+\subsection{4-Way interaction split over discriminability}
+
+<<4waysplit, fig.width=7, fig.height=7, fig.cap='4-Way interaction split over discriminability', fig.pos='hbtp', message=FALSE, echo=FALSE>>=
 rts <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Number", "Vagueness", "discriminability", "Quantity", "Order"))
@@
          axis.text.y = element_text(angle = 45, hjust = 1, size=6),
-         aspect.ratio=1) + 
+         aspect.ratio=1,
+         plot.background=element_rect(fill=NA, color='white')) + 
   scale_x_continuous(minor_breaks=NULL,
@@
 \clearpage
-\section{Lmer model}
+\section{Lmer model: before outlier removal}
 
@@
 
+<<summaryV5, echo=FALSE>>=
+summary(v5)
+@
+
+\clearpage
+
 <<printXtableV5, echo=FALSE, results='asis'>>=
@@
   table.placement='htbp',
-  latex.environments='',
+  latex.environments='center',
   size='small'
@@
 
-\clearpage
-
-<<plotModelCoefsAndCis, fig.width=6, fig.height=4, fig.cap='Coefficient estimates and their (Wald) 95 per cent confidence intervals', fig.pos='hbtp', echo=FALSE>>=
+%\clearpage
+
+<<plotModelCoefsAndCis, fig.width=5.5, fig.height=4, fig.cap='Coefficient estimates and their (Wald) 95 per cent confidence intervals', fig.pos='hbtp', echo=FALSE>>=
 # Plot model coefficients and ci's
@@
 mtext("Model terms", side=2, line=8)
-box(col='grey')
+box(which='plot', col='grey')
+
 @
@@
 plotLMER.fnc(v5)
+
 @
@@
 abline(h=c(-2.5,2.5))
+
 @
@@
 
-<<lmerTestVersionOfModel, cache=TRUE>>=
+<<lmerTestVersionOfModel, tidy=FALSE, cache=TRUE>>=
 v6 <- lmerTest::lmer(data=dd,
@@
 \clearpage
+\section{Lmer model: after outlier removal}
+
+\clearpage
 \section{Borderline responses}
@@
 dd$response_category <- relevel(dd$response_category, ref = 'expected')
-#barplot(table(dd$response_category,dd$Vagueness), beside=TRUE, legend=TRUE)
 barplot(table(dd$Vagueness,dd$response_category), beside=TRUE, legend=TRUE, space=c(0,.5))
+
 @
@@
 
+\clearpage
+\appendix
+\section{Functions listing}
+\clearpage
+\subsection{Gather Data}
+
+<<showgFunction, tidy=TRUE, size='scriptsize', echo=TRUE>>=
+<<gatherDataFunction>>
+@
+
+\clearpage
+\subsection{Classify Response}
+
+<<showcFunction, tidy=TRUE, size='scriptsize', echo=TRUE>>=
+<<classifyResponseFuntion>>
+@
+
+\clearpage
+\subsection{Processing}
+
+<<showpFunction, tidy=TRUE, size='scriptsize', echo=TRUE>>=
+<<processDataFunction>>
+@
+
+\clearpage
+\subsection{Post-processing}
+
+<<showppFunction, tidy=TRUE, size='scriptsize', echo=TRUE>>=
+<<postProcessDataFunction>>
+@
+
 \end{document}

### 2016-08-13 experiments/e2/writeup/e2-b-knitr.tex
tidying up

@@
-\documentclass[a4paper,12pt]{article}\usepackage[]{graphicx}\usepackage[]{color}
+\documentclass[a4paper,12pt,twoside]{article}\usepackage[]{graphicx}\usepackage[]{color}
 %% maxwidth is the original width if it is less than linewidth
@@
 \usepackage{placeins}
-
+\usepackage{graphicx}
+\usepackage{color}
+\usepackage{hyperref}
 \title{Exp 2 of 4}
@@
 
+\clearpage
+
 \begin{knitrout}\scriptsize
@@
 \end{knitrout}
+
+\clearpage
 
@@
 \clearpage
+\subsection{How logging RT affects the distribution}
 
@@
 \clearpage
+\subsection{Plot main effects in both transformations}
 
@@
 
-{\centering \includegraphics[width=\maxwidth]{figure/graphics-plotMainsa-1}
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-plotMainEffectsRaw-1}
 
@@
 
-\caption[Plot main effects on several transformations]{Plot main effects on several transformations}\label{fig:plotMainsa}
+\caption[Plot main effects in raw RT]{Plot main effects in raw RT}\label{fig:plotMainEffectsRaw}
+\end{figure}
+
+\end{knitrout}
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[htbp]
+
+{\centering \includegraphics[width=\maxwidth]{figure/graphics-plotMainEffectsLog-1}
+
+}
+
+\caption[Plot main effects in log RT]{Plot main effects in log RT}\label{fig:plotMainEffectsLog}
 \end{figure}
@@
 \clearpage
-
-\begin{knitrout}\scriptsize
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
-
-{\centering \includegraphics[width=\maxwidth]{figure/graphics-mainB-1}
-
-}
-
-\caption[Just the main effects]{Just the main effects}\label{fig:mainB}
-\end{figure}
-
-\end{knitrout}
-
-\clearpage
+\subsection{Main effects in log RT over discriminability}
 
@@
 
-\caption[Main effects over item ratios]{Main effects over item ratios}\label{fig:mainD}
+\caption[Main effects in log RT over discriminability]{Main effects in log RT over discriminability}\label{fig:mainD}
 \end{figure}
@@
 \clearpage
+\subsection{2-Way interactions over discriminability}
 
@@
 
-\caption[2-way interactions over item ratios]{2-way interactions over item ratios}\label{fig:22way}
+\caption[2-way interactions over discriminability]{2-way interactions over discriminability}\label{fig:22way}
 \end{figure}
@@
 \clearpage
+\subsection{Vagueness by number interaction over discriminability}
 
@@
 
-\caption[vagueness by number interaction over item ratios]{vagueness by number interaction over item ratios}\label{fig:plotInt2}
+\caption[vagueness by number interaction over discriminability]{vagueness by number interaction over discriminability}\label{fig:plotInt2}
 \end{figure}
@@
 \clearpage
+\subsection{3-Way interactions}
 
@@
 
-\caption[3 way interactions]{3 way interactions}\label{fig:plot3way}
+\caption[3-Way interactions]{3-Way interactions}\label{fig:plot3way}
 \end{figure}
@@
 \clearpage
+\subsection{Vagueness by number by quantity over discriminability}
 
@@
 
-\caption[Plot vagueness by number by quantity over item ratios]{Plot vagueness by number by quantity over item ratios.}\label{fig:vnitems}
+\caption[Vagueness by number by quantity over discriminability]{Vagueness by number by quantity over discriminability}\label{fig:vnitems}
 \end{figure}
@@
 \clearpage
+\subsection{4-Way interaction}
 
@@
 
-\caption[4 way interaction]{4 way interaction}\label{fig:4way}
+\caption[4-Way interaction]{4-Way interaction}\label{fig:4way}
 \end{figure}
@@
 \clearpage
+\subsection{4-Way interaction split over discriminability}
 
@@
 
-\caption[4 way interaction split over item discriminability]{4 way interaction split over item discriminability}\label{fig:4waysplit}
+\caption[4-Way interaction split over discriminability]{4-Way interaction split over discriminability}\label{fig:4waysplit}
 \end{figure}
@@
 
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+Linear mixed model fit by REML ['lmerMod']
+Formula: RT_log ~ c_Vag + c_Num + c_Qty + c_Ord + c_Num:c_Vag:c_Qty +  
+    discriminability + s_Trl + RTprev_log + nchar_instr + (1 +  
+    c_Vag + c_Num + c_Qty + c_Ord | Subject)
+   Data: dd
+
+REML criterion at convergence: 11474.8
+
+Scaled residuals: 
+    Min      1Q  Median      3Q     Max 
+-4.5470 -0.6351 -0.0955  0.5372  5.0914
+
+Random effects:
+ Groups   Name        Variance Std.Dev. Corr                   
+ Subject  (Intercept) 0.153949 0.39236                         
+          c_Vag       0.001546 0.03932   0.69                  
+          c_Num       0.165314 0.40659  -0.67 -0.64            
+          c_Qty       0.008148 0.09027   0.16  0.26 -0.34      
+          c_Ord       0.001559 0.03949  -0.13  0.02 -0.39 -0.52
+ Residual             0.249734 0.49973                         
+Number of obs: 7677, groups:  Subject, 30
+
+Fixed effects:
+                   Estimate Std. Error t value
+(Intercept)        7.171685   0.122236   58.67
+c_Vag              0.060938   0.013873    4.39
+c_Num             -0.433614   0.075148   -5.77
+c_Qty             -0.069067   0.020047   -3.45
+c_Ord              0.017959   0.013497    1.33
+discriminability  -0.771402   0.049267  -15.66
+s_Trl             -0.106972   0.005807  -18.42
+RTprev_log         0.060469   0.009692    6.24
+nchar_instr        0.006086   0.001944    3.13
+c_Vag:c_Num:c_Qty  0.104949   0.046066    2.28
+
+Correlation of Fixed Effects:
+            (Intr) c_Vag  c_Num  c_Qty  c_Ord  dscrmn s_Trl  RTprv_ nchr_n
+c_Vag        0.083                                                        
+c_Num       -0.369 -0.337                                                 
+c_Qty        0.069  0.117 -0.278                                          
+c_Ord       -0.052  0.007 -0.206 -0.228                                   
+discrmnblty -0.140  0.004 -0.001  0.000  0.001                            
+s_Trl       -0.114  0.003 -0.001 -0.002  0.006  0.017                     
+RTprev_log  -0.607  0.006 -0.004 -0.003  0.019  0.015  0.185              
+nchar_instr -0.524  0.237 -0.034  0.018  0.000  0.017  0.000  0.006       
+c_Vg:c_N:_Q  0.073 -0.033  0.005 -0.003  0.000 -0.002 -0.001 -0.002 -0.138
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+\clearpage
+
 % latex table generated in R 3.3.1 by xtable 1.8-2 package
-% Fri Aug 12 22:28:46 2016
+% Sat Aug 13 09:11:20 2016
 \begin{table}[htbp]
+\centering
 \begingroup\small
@@
 
-\clearpage
+%\clearpage
 
@@
 \begin{alltt}
-\hlstd{v6} \hlkwb{<-} \hlstd{lmerTest}\hlopt{::}\hlkwd{lmer}\hlstd{(}\hlkwc{data} \hlstd{= dd, RT_log} \hlopt{~} \hlstd{c_Vag} \hlopt{+} \hlstd{c_Num} \hlopt{+} \hlstd{c_Qty} \hlopt{+} \hlstd{c_Ord} \hlopt{+} \hlstd{c_Num}\hlopt{:}\hlstd{c_Vag}\hlopt{:}\hlstd{c_Qty} \hlopt{+} \hlstd{discriminability} \hlopt{+}
-    \hlstd{s_Trl} \hlopt{+} \hlstd{RTprev_log} \hlopt{+} \hlstd{nchar_instr} \hlopt{+} \hlstd{(}\hlnum{1} \hlopt{+} \hlstd{c_Vag} \hlopt{+} \hlstd{c_Num} \hlopt{+} \hlstd{c_Qty} \hlopt{+} \hlstd{c_Ord} \hlopt{|} \hlstd{Subject))}
+\hlstd{v6} \hlkwb{<-} \hlstd{lmerTest}\hlopt{::}\hlkwd{lmer}\hlstd{(}\hlkwc{data}\hlstd{=dd,}
+                     \hlstd{RT_log} \hlopt{~}
+                       \hlstd{c_Vag} \hlopt{+} \hlstd{c_Num} \hlopt{+} \hlstd{c_Qty} \hlopt{+} \hlstd{c_Ord} \hlopt{+}
+                       \hlstd{c_Num}\hlopt{:}\hlstd{c_Vag}\hlopt{:}\hlstd{c_Qty} \hlopt{+}
+                       \hlstd{discriminability} \hlopt{+}
+                       \hlstd{s_Trl} \hlopt{+}
+                       \hlstd{RTprev_log} \hlopt{+}
+                       \hlstd{nchar_instr} \hlopt{+}
+                       \hlstd{(}\hlnum{1}\hlopt{+}\hlstd{c_Vag} \hlopt{+} \hlstd{c_Num} \hlopt{+} \hlstd{c_Qty} \hlopt{+} \hlstd{c_Ord}\hlopt{|}\hlstd{Subject))}
 \end{alltt}
@@
 % latex table generated in R 3.3.1 by xtable 1.8-2 package
-% Fri Aug 12 22:28:47 2016
+% Sat Aug 13 09:11:21 2016
 \begin{table}[htbp]
@@
 
+\clearpage
+\appendix
+\section{Functions listing}
+\clearpage
+\subsection{Gather Data}
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{gatherData} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{number_of_valid_subjects}\hlstd{) \{}
+    \hlstd{data_dir} \hlkwb{<-} \hlstr{"../experimentCode/output/"}
+    \hlstd{column.headers.df} \hlkwb{<-} \hlkwd{head}\hlstd{(}\hlkwd{read.table}\hlstd{(}\hlkwd{paste}\hlstd{(data_dir,} \hlstr{"subject01.data"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{""}\hlstd{),} \hlkwc{header} \hlstd{=} \hlnum{TRUE}\hlstd{),}
+        \hlnum{0}\hlstd{)}
+    \hlstd{gathered.data} \hlkwb{<-} \hlstd{column.headers.df}
+    \hlkwa{for} \hlstd{(subject} \hlkwa{in} \hlnum{1}\hlopt{:}\hlstd{number_of_valid_subjects) \{}
+        \hlstd{current.filename} \hlkwb{<-} \hlkwd{paste}\hlstd{(data_dir,} \hlstr{"subject"}\hlstd{,} \hlkwd{sprintf}\hlstd{(}\hlstr{"%02d"}\hlstd{, subject),} \hlstr{".data"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{""}\hlstd{)}
+        \hlstd{current.data} \hlkwb{<-} \hlkwd{read.table}\hlstd{(}\hlkwc{file} \hlstd{= current.filename,} \hlkwc{header} \hlstd{=} \hlnum{TRUE}\hlstd{,} \hlkwc{stringsAsFactors} \hlstd{=} \hlnum{FALSE}\hlstd{)}
+        \hlstd{gathered.data} \hlkwb{<-} \hlkwd{rbind}\hlstd{(gathered.data, current.data)}
+    \hlstd{\}}
+    \hlkwd{message}\hlstd{(}\hlstr{"Returning gathered data"}\hlstd{)}
+    \hlkwd{return}\hlstd{(gathered.data)}
+\hlstd{\}}  \hlcom{# end of gatherData function}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\clearpage
+\subsection{Classify Response}
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{classifyResponses} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{dat}\hlstd{) \{}
+    \hlcom{# what were they expected to respond?}
+    \hlstd{dat}\hlopt{$}\hlstd{crossed} \hlkwb{<-} \hlkwd{as.factor}\hlstd{(}\hlkwd{paste}\hlstd{(}\hlstr{"Con"}\hlstd{, dat}\hlopt{$}\hlstd{Condition,} \hlstr{":Quan"}\hlstd{, dat}\hlopt{$}\hlstd{Quantity,} \hlstr{":Item"}\hlstd{, dat}\hlopt{$}\hlstd{Item,} \hlkwc{sep} \hlstd{=} \hlstr{""}\hlstd{))}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan1:Item1"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{6}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan1:Item1"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{15}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan1:Item1"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{24}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan1:Item2"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{16}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan1:Item2"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{25}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan1:Item2"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{34}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan1:Item3"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{26}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan1:Item3"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{35}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan1:Item3"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{44}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan1:Item4"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{36}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan1:Item4"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{45}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan1:Item4"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{54}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan2:Item1"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{24}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan2:Item1"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{15}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan2:Item1"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{6}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan2:Item2"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{34}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan2:Item2"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{25}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan2:Item2"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{16}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan2:Item3"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{44}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan2:Item3"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{35}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan2:Item3"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{26}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan2:Item4"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{54}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan2:Item4"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{45}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con1:Quan2:Item4"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{36}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan1:Item1"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{6}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan1:Item1"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{15}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan1:Item1"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{24}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan1:Item2"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{16}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan1:Item2"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{25}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan1:Item2"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{34}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan1:Item3"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{26}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan1:Item3"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{35}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan1:Item3"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{44}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan1:Item4"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{36}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan1:Item4"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{45}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan1:Item4"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{54}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan2:Item1"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{24}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan2:Item1"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{15}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan2:Item1"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{6}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan2:Item2"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{34}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan2:Item2"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{25}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan2:Item2"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{16}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan2:Item3"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{44}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan2:Item3"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{35}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan2:Item3"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{26}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan2:Item4"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{54}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan2:Item4"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{45}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con2:Quan2:Item4"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{36}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan1:Item1"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{6}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan1:Item1"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{15}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan1:Item1"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{24}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan1:Item2"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{16}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan1:Item2"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{25}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan1:Item2"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{34}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan1:Item3"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{26}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan1:Item3"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{35}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan1:Item3"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{44}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan1:Item4"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{36}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan1:Item4"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{45}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan1:Item4"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{54}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan2:Item1"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{24}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan2:Item1"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{15}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan2:Item1"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{6}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan2:Item2"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{34}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan2:Item2"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{25}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan2:Item2"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{16}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan2:Item3"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{44}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan2:Item3"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{35}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan2:Item3"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{26}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan2:Item4"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{54}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan2:Item4"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{45}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con3:Quan2:Item4"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{36}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan1:Item1"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{6}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan1:Item1"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{15}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan1:Item1"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{24}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan1:Item2"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{16}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan1:Item2"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{25}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan1:Item2"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{34}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan1:Item3"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{26}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan1:Item3"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{35}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan1:Item3"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{44}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan1:Item4"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{36}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan1:Item4"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{45}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan1:Item4"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{54}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan2:Item1"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{24}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan2:Item1"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{15}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan2:Item1"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{6}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan2:Item2"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{34}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan2:Item2"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{25}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan2:Item2"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{16}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan2:Item3"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{44}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan2:Item3"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{35}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan2:Item3"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{26}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan2:Item4"}\hlstd{,} \hlstr{"Exp_Num"}\hlstd{]} \hlkwb{<-} \hlnum{54}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan2:Item4"}\hlstd{,} \hlstr{"Bline_Num"}\hlstd{]} \hlkwb{<-} \hlnum{45}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{crossed} \hlopt{==} \hlstr{"Con4:Quan2:Item4"}\hlstd{,} \hlstr{"Extr_Num"}\hlstd{]} \hlkwb{<-} \hlnum{36}
+    \hlstd{dat}\hlopt{$}\hlstd{crossed} \hlkwb{<-} \hlkwa{NULL}
+    \hlcom{# what side LEFT, MIDDLE, RIGHT corresponds with Expected, Borderline, Extreme?}
+    \hlkwa{for} \hlstd{(row} \hlkwa{in} \hlnum{1}\hlopt{:}\hlkwd{nrow}\hlstd{(dat)) \{}
+        \hlkwa{if} \hlstd{(dat[row,} \hlstr{"Exp_Num"}\hlstd{]} \hlopt{==} \hlstd{dat[row,} \hlstr{"Left"}\hlstd{]) \{}
+            \hlstd{dat[row,} \hlstr{"Exp_side"}\hlstd{]} \hlkwb{<-} \hlstr{"left"}
+        \hlstd{\}}
+        \hlkwa{if} \hlstd{(dat[row,} \hlstr{"Exp_Num"}\hlstd{]} \hlopt{==} \hlstd{dat[row,} \hlstr{"Mid"}\hlstd{]) \{}
+            \hlstd{dat[row,} \hlstr{"Exp_side"}\hlstd{]} \hlkwb{<-} \hlstr{"mid"}
+        \hlstd{\}}
+        \hlkwa{if} \hlstd{(dat[row,} \hlstr{"Exp_Num"}\hlstd{]} \hlopt{==} \hlstd{dat[row,} \hlstr{"Right"}\hlstd{]) \{}
+            \hlstd{dat[row,} \hlstr{"Exp_side"}\hlstd{]} \hlkwb{<-} \hlstr{"right"}
+        \hlstd{\}}
+        \hlkwa{if} \hlstd{(dat[row,} \hlstr{"Bline_Num"}\hlstd{]} \hlopt{==} \hlstd{dat[row,} \hlstr{"Left"}\hlstd{]) \{}
+            \hlstd{dat[row,} \hlstr{"Bline_side"}\hlstd{]} \hlkwb{<-} \hlstr{"left"}
+        \hlstd{\}}
+        \hlkwa{if} \hlstd{(dat[row,} \hlstr{"Bline_Num"}\hlstd{]} \hlopt{==} \hlstd{dat[row,} \hlstr{"Mid"}\hlstd{]) \{}
+            \hlstd{dat[row,} \hlstr{"Bline_side"}\hlstd{]} \hlkwb{<-} \hlstr{"mid"}
+        \hlstd{\}}
+        \hlkwa{if} \hlstd{(dat[row,} \hlstr{"Bline_Num"}\hlstd{]} \hlopt{==} \hlstd{dat[row,} \hlstr{"Right"}\hlstd{]) \{}
+            \hlstd{dat[row,} \hlstr{"Bline_side"}\hlstd{]} \hlkwb{<-} \hlstr{"right"}
+        \hlstd{\}}
+        \hlkwa{if} \hlstd{(dat[row,} \hlstr{"Extr_Num"}\hlstd{]} \hlopt{==} \hlstd{dat[row,} \hlstr{"Left"}\hlstd{]) \{}
+            \hlstd{dat[row,} \hlstr{"Extr_side"}\hlstd{]} \hlkwb{<-} \hlstr{"left"}
+        \hlstd{\}}
+        \hlkwa{if} \hlstd{(dat[row,} \hlstr{"Extr_Num"}\hlstd{]} \hlopt{==} \hlstd{dat[row,} \hlstr{"Mid"}\hlstd{]) \{}
+            \hlstd{dat[row,} \hlstr{"Extr_side"}\hlstd{]} \hlkwb{<-} \hlstr{"mid"}
+        \hlstd{\}}
+        \hlkwa{if} \hlstd{(dat[row,} \hlstr{"Extr_Num"}\hlstd{]} \hlopt{==} \hlstd{dat[row,} \hlstr{"Right"}\hlstd{]) \{}
+            \hlstd{dat[row,} \hlstr{"Extr_side"}\hlstd{]} \hlkwb{<-} \hlstr{"right"}
+        \hlstd{\}}
+    \hlstd{\}}
+    \hlcom{# what button press did the subject actually make? LEFT, MIDDLE, RIGHT, NOANSWER?}
+    \hlstd{dat}\hlopt{$}\hlstd{RESPONSE} \hlkwb{<-} \hlkwd{as.factor}\hlstd{(dat}\hlopt{$}\hlstd{RESPONSE)}
+    \hlcom{# what number of dots corresponds with the subject's button press?}
+    \hlkwa{for} \hlstd{(row} \hlkwa{in} \hlnum{1}\hlopt{:}\hlkwd{nrow}\hlstd{(dat)) \{}
+        \hlkwd{switch}\hlstd{(}\hlkwd{as.character}\hlstd{(dat[row,} \hlstr{"RESPONSE"}\hlstd{]),} \hlkwc{LEFT} \hlstd{= \{}
+            \hlstd{dat[row,} \hlstr{"response_num"}\hlstd{]} \hlkwb{<-} \hlstd{dat[row,} \hlstr{"Left"}\hlstd{]}
+        \hlstd{\},} \hlkwc{MIDDLE} \hlstd{= \{}
+            \hlstd{dat[row,} \hlstr{"response_num"}\hlstd{]} \hlkwb{<-} \hlstd{dat[row,} \hlstr{"Mid"}\hlstd{]}
+        \hlstd{\},} \hlkwc{RIGHT} \hlstd{= \{}
+            \hlstd{dat[row,} \hlstr{"response_num"}\hlstd{]} \hlkwb{<-} \hlstd{dat[row,} \hlstr{"Right"}\hlstd{]}
+        \hlstd{\},} \hlkwc{NOANSWER} \hlstd{= \{}
+            \hlstd{dat[row,} \hlstr{"response_num"}\hlstd{]} \hlkwb{<-} \hlnum{NA}
+        \hlstd{\})}
+    \hlstd{\}}
+    \hlcom{# what side was the subject's button-press? Left, mid right?}
+    \hlstd{dat}\hlopt{$}\hlstd{response_side} \hlkwb{<-} \hlkwd{tolower}\hlstd{(dat}\hlopt{$}\hlstd{RESPONSE)}
+    \hlstd{dat}\hlopt{$}\hlstd{response_side[dat}\hlopt{$}\hlstd{response_side} \hlopt{==} \hlstr{"middle"}\hlstd{]} \hlkwb{<-} \hlstr{"mid"}
+    \hlstd{dat}\hlopt{$}\hlstd{response_side} \hlkwb{<-} \hlkwd{factor}\hlstd{(dat}\hlopt{$}\hlstd{response_side,} \hlkwc{exclude} \hlstd{=} \hlstr{"noanswer"}\hlstd{)}
+    \hlcom{# what category was the subject's response? Expected, Borderline, Extreme}
+    \hlstd{dat}\hlopt{$}\hlstd{response_category} \hlkwb{<-} \hlstr{"nocat"}
+    \hlkwa{for} \hlstd{(row} \hlkwa{in} \hlkwd{row.names}\hlstd{(}\hlkwd{na.omit}\hlstd{(dat))) \{}
+        \hlkwa{if} \hlstd{(dat[row,} \hlstr{"response_num"}\hlstd{]} \hlopt{==} \hlstd{dat[row,} \hlstr{"Exp_Num"}\hlstd{]) \{}
+            \hlstd{dat[row,} \hlstr{"response_category"}\hlstd{]} \hlkwb{<-} \hlstr{"expected"}
+        \hlstd{\}}
+        \hlkwa{if} \hlstd{(dat[row,} \hlstr{"response_num"}\hlstd{]} \hlopt{==} \hlstd{dat[row,} \hlstr{"Bline_Num"}\hlstd{]) \{}
+            \hlstd{dat[row,} \hlstr{"response_category"}\hlstd{]} \hlkwb{<-} \hlstr{"borderline"}
+        \hlstd{\}}
+        \hlkwa{if} \hlstd{(dat[row,} \hlstr{"response_num"}\hlstd{]} \hlopt{==} \hlstd{dat[row,} \hlstr{"Extr_Num"}\hlstd{]) \{}
+            \hlstd{dat[row,} \hlstr{"response_category"}\hlstd{]} \hlkwb{<-} \hlstr{"extreme"}
+        \hlstd{\}}
+    \hlstd{\}}
+    \hlstd{dat}\hlopt{$}\hlstd{response_category} \hlkwb{<-} \hlkwd{factor}\hlstd{(dat}\hlopt{$}\hlstd{response_category,} \hlkwc{exclude} \hlstd{=} \hlstr{"nocat"}\hlstd{)}
+    \hlstd{dat}\hlopt{$}\hlstd{RESPONSE} \hlkwb{<-} \hlkwa{NULL}
+    \hlkwd{message}\hlstd{(}\hlstr{"Returning classified data"}\hlstd{)}
+    \hlkwd{return}\hlstd{(dat)}
+\hlstd{\}}  \hlcom{# end of classifyResponse function}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\clearpage
+\subsection{Processing}
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{processData} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{dat}\hlstd{) \{}
+    \hlcom{# SUBJECT}
+    \hlstd{dat}\hlopt{$}\hlstd{Subject} \hlkwb{<-} \hlkwd{factor}\hlstd{(}\hlkwd{paste}\hlstd{(}\hlstr{"s"}\hlstd{,} \hlkwd{sprintf}\hlstd{(}\hlstr{"%02d"}\hlstd{, dat}\hlopt{$}\hlstd{Subject),} \hlkwc{sep} \hlstd{=} \hlstr{""}\hlstd{))}
+    \hlcom{# TRIAL}
+    \hlstd{dat}\hlopt{$}\hlstd{Trial} \hlkwb{<-} \hlkwd{rep}\hlstd{(}\hlkwc{x} \hlstd{=} \hlnum{1}\hlopt{:}\hlstd{number_of_trials_per_subject,} \hlkwc{times} \hlstd{= number_of_valid_subjects)}
+    \hlcom{# make a centred Trial for modeling}
+    \hlstd{dat}\hlopt{$}\hlstd{c_Trl} \hlkwb{<-} \hlstd{dat}\hlopt{$}\hlstd{Trial} \hlopt{-} \hlkwd{mean}\hlstd{(dat}\hlopt{$}\hlstd{Trial)}
+    \hlcom{# make a scaled Trial for modelling}
+    \hlstd{dat}\hlopt{$}\hlstd{s_Trl} \hlkwb{<-} \hlkwd{as.numeric}\hlstd{(}\hlkwd{scale}\hlstd{(dat}\hlopt{$}\hlstd{Trial))}
+    \hlcom{# ID id is a unique identifier for the 7680 row data}
+    \hlstd{dat}\hlopt{$}\hlstd{id} \hlkwb{<-} \hlkwd{factor}\hlstd{(}\hlkwd{paste}\hlstd{(}\hlkwd{paste}\hlstd{(dat}\hlopt{$}\hlstd{Subject),} \hlkwd{paste}\hlstd{(}\hlstr{"t"}\hlstd{,} \hlkwd{sprintf}\hlstd{(}\hlstr{"%03d"}\hlstd{, dat}\hlopt{$}\hlstd{Trial),} \hlkwc{sep} \hlstd{=} \hlstr{""}\hlstd{),} \hlkwc{sep} \hlstd{=} \hlstr{":"}\hlstd{))}
+    \hlcom{# ITEM create a centred numeric item variable for modeling}
+    \hlstd{dat}\hlopt{$}\hlstd{c_Itm} \hlkwb{<-} \hlkwd{ifelse}\hlstd{(dat}\hlopt{$}\hlstd{Item} \hlopt{==} \hlnum{1}\hlstd{,} \hlopt{-}\hlnum{0.75}\hlstd{,} \hlkwd{ifelse}\hlstd{(dat}\hlopt{$}\hlstd{Item} \hlopt{==} \hlnum{2}\hlstd{,} \hlopt{-}\hlnum{0.25}\hlstd{,} \hlkwd{ifelse}\hlstd{(dat}\hlopt{$}\hlstd{Item} \hlopt{==} \hlnum{3}\hlstd{,} \hlnum{0.25}\hlstd{,}
+        \hlnum{0.75}\hlstd{)))}
+    \hlcom{# make Item be a factor and assign labels}
+    \hlstd{dat}\hlopt{$}\hlstd{Item} \hlkwb{<-} \hlkwd{factor}\hlstd{(dat}\hlopt{$}\hlstd{Item,} \hlkwc{levels} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{,} \hlnum{3}\hlstd{,} \hlnum{4}\hlstd{),} \hlkwc{labels} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"06:15:24"}\hlstd{,} \hlstr{"16:25:34"}\hlstd{,} \hlstr{"26:35:44"}\hlstd{,}
+        \hlstr{"36:45:54"}\hlstd{))}
+    \hlcom{# VAGUENESS Create a factor coding for Vagueness}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{Condition} \hlopt{==} \hlnum{1}\hlstd{,} \hlstr{"Vagueness"}\hlstd{]} \hlkwb{<-} \hlstr{"Vague"}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{Condition} \hlopt{==} \hlnum{2}\hlstd{,} \hlstr{"Vagueness"}\hlstd{]} \hlkwb{<-} \hlstr{"Crisp"}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{Condition} \hlopt{==} \hlnum{3}\hlstd{,} \hlstr{"Vagueness"}\hlstd{]} \hlkwb{<-} \hlstr{"Vague"}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{Condition} \hlopt{==} \hlnum{4}\hlstd{,} \hlstr{"Vagueness"}\hlstd{]} \hlkwb{<-} \hlstr{"Crisp"}
+    \hlstd{dat}\hlopt{$}\hlstd{Vagueness} \hlkwb{<-} \hlkwd{as.factor}\hlstd{(dat}\hlopt{$}\hlstd{Vagueness)}
+    \hlcom{# manually center Vagueness}
+    \hlstd{dat}\hlopt{$}\hlstd{c_Vag} \hlkwb{<-} \hlkwd{ifelse}\hlstd{(dat}\hlopt{$}\hlstd{Vagueness} \hlopt{==} \hlstr{"Crisp"}\hlstd{,} \hlopt{-}\hlnum{0.5}\hlstd{,} \hlnum{0.5}\hlstd{)}
+    \hlcom{# NUMBER Create a factor coding for Number use}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{Condition} \hlopt{==} \hlnum{1}\hlstd{,} \hlstr{"Number"}\hlstd{]} \hlkwb{<-} \hlstr{"Numeric"}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{Condition} \hlopt{==} \hlnum{2}\hlstd{,} \hlstr{"Number"}\hlstd{]} \hlkwb{<-} \hlstr{"Numeric"}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{Condition} \hlopt{==} \hlnum{3}\hlstd{,} \hlstr{"Number"}\hlstd{]} \hlkwb{<-} \hlstr{"Verbal"}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{Condition} \hlopt{==} \hlnum{4}\hlstd{,} \hlstr{"Number"}\hlstd{]} \hlkwb{<-} \hlstr{"Verbal"}
+    \hlstd{dat}\hlopt{$}\hlstd{Number} \hlkwb{<-} \hlkwd{as.factor}\hlstd{(dat}\hlopt{$}\hlstd{Number)}
+    \hlcom{# manually center Number}
+    \hlstd{dat}\hlopt{$}\hlstd{c_Num} \hlkwb{<-} \hlkwd{ifelse}\hlstd{(dat}\hlopt{$}\hlstd{Number} \hlopt{==} \hlstr{"Numeric"}\hlstd{,} \hlopt{-}\hlnum{0.5}\hlstd{,} \hlnum{0.5}\hlstd{)}
+    \hlcom{# CONDITION make a factor out of Condition, as f_Cnd}
+    \hlstd{dat}\hlopt{$}\hlstd{f_Cnd} \hlkwb{<-} \hlkwd{factor}\hlstd{(dat}\hlopt{$}\hlstd{Condition,} \hlkwc{levels} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{,} \hlnum{3}\hlstd{,} \hlnum{4}\hlstd{),} \hlkwc{labels} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"Vg:Nm"}\hlstd{,} \hlstr{"Cr:Nm"}\hlstd{,} \hlstr{"Vg:Vb"}\hlstd{,}
+        \hlstr{"Cr:Vb"}\hlstd{))}
+    \hlcom{# ORDER give the levels of Order meaningful names}
+    \hlstd{dat}\hlopt{$}\hlstd{Order} \hlkwb{<-} \hlkwd{factor}\hlstd{(dat}\hlopt{$}\hlstd{Order,} \hlkwc{levels} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{),} \hlkwc{labels} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"LtoR"}\hlstd{,} \hlstr{"RtoL"}\hlstd{))}
+    \hlcom{# make a manually centred Order}
+    \hlstd{dat}\hlopt{$}\hlstd{c_Ord} \hlkwb{<-} \hlkwd{ifelse}\hlstd{(dat}\hlopt{$}\hlstd{Order} \hlopt{==} \hlstr{"LtoR"}\hlstd{,} \hlopt{-}\hlnum{0.5}\hlstd{,} \hlnum{0.5}\hlstd{)}
+    \hlcom{# QUANTITY give the levels of Quantity meaningful names}
+    \hlstd{dat}\hlopt{$}\hlstd{Quantity} \hlkwb{<-} \hlkwd{factor}\hlstd{(dat}\hlopt{$}\hlstd{Quantity,} \hlkwc{levels} \hlstd{=} \hlkwd{c}\hlstd{(}\hlnum{1}\hlstd{,} \hlnum{2}\hlstd{),} \hlkwc{labels} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{"Small"}\hlstd{,} \hlstr{"Large"}\hlstd{))}
+    \hlcom{# make a manually centred Quantity}
+    \hlstd{dat}\hlopt{$}\hlstd{c_Qty} \hlkwb{<-} \hlkwd{ifelse}\hlstd{(dat}\hlopt{$}\hlstd{Quantity} \hlopt{==} \hlstr{"Small"}\hlstd{,} \hlopt{-}\hlnum{0.5}\hlstd{,} \hlnum{0.5}\hlstd{)}
+    \hlcom{# INSTRUCTION add number of characters in the instruction # 29 30 34 36 38}
+    \hlstd{dat}\hlopt{$}\hlstd{nchar_instr} \hlkwb{<-} \hlkwd{nchar}\hlstd{(dat}\hlopt{$}\hlstd{Instruction)}
+    \hlstd{dat}\hlopt{$}\hlstd{nchar_instr_scaled} \hlkwb{<-} \hlkwd{as.vector}\hlstd{(}\hlkwd{scale}\hlstd{(}\hlkwd{nchar}\hlstd{(dat}\hlopt{$}\hlstd{Instruction),} \hlkwc{scale} \hlstd{=} \hlnum{TRUE}\hlstd{))}
+    \hlcom{# make Instruction be a factor (17 levels)}
+    \hlstd{dat}\hlopt{$}\hlstd{Instruction} \hlkwb{<-} \hlkwd{as.factor}\hlstd{(dat}\hlopt{$}\hlstd{Instruction)}
+    \hlcom{# RT add transformations of RT}
+    \hlstd{dat}\hlopt{$}\hlstd{RT_log} \hlkwb{<-} \hlkwd{log}\hlstd{(dat}\hlopt{$}\hlstd{RT)}
+    \hlstd{dat}\hlopt{$}\hlstd{RT_raw} \hlkwb{<-} \hlstd{dat}\hlopt{$}\hlstd{RT}
+    \hlcom{# print to file a table with information about the design}
+    \hlstd{design_info} \hlkwb{<-} \hlkwd{unique}\hlstd{(}\hlkwd{subset}\hlstd{(dat,} \hlkwc{select} \hlstd{=} \hlkwd{c}\hlstd{(Item, Condition, Vagueness, Number, Quantity, Order,}
+        \hlstd{Left, Mid, Right, Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, Instruction)))}
+    \hlstd{design.info} \hlkwb{<-} \hlstd{design_info[}\hlkwd{order}\hlstd{(design_info}\hlopt{$}\hlstd{Item, design_info}\hlopt{$}\hlstd{Condition, design_info}\hlopt{$}\hlstd{Quantity, design_info}\hlopt{$}\hlstd{Order),}
+        \hlstd{]}
+    \hlkwd{row.names}\hlstd{(design_info)} \hlkwb{<-} \hlkwa{NULL}
+    \hlkwd{capture.output}\hlstd{(}\hlkwd{print.data.frame}\hlstd{(design_info,} \hlkwc{row.names} \hlstd{= F,} \hlkwc{print.gap} \hlstd{=} \hlnum{3}\hlstd{,} \hlkwc{quote} \hlstd{= F,} \hlkwc{right} \hlstd{= F),}
+        \hlkwc{file} \hlstd{=} \hlstr{"design_info-table.txt"}\hlstd{)}
+    \hlcom{# Add discriminability metric with reference to item}
+    \hlstd{discriminability_range} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlnum{0.75}\hlstd{,} \hlnum{0.5294118}\hlstd{,} \hlnum{0.4090909}\hlstd{,} \hlnum{0.3333333}\hlstd{)}
+    \hlstd{discriminability_range_scaled} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlnum{1.3441995}\hlstd{,} \hlnum{0.1316642}\hlstd{,} \hlopt{-}\hlnum{0.5297187}\hlstd{,} \hlopt{-}\hlnum{0.946145}\hlstd{)}
+    \hlstd{discriminability} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlnum{0.4875}\hlstd{,} \hlnum{0.3123529}\hlstd{,} \hlnum{0.2308442}\hlstd{,} \hlnum{0.1833333}\hlstd{)}
+    \hlstd{discriminability_scaled} \hlkwb{<-} \hlkwd{c}\hlstd{(}\hlnum{1.37582241}\hlstd{,} \hlnum{0.06614191}\hlstd{,} \hlopt{-}\hlnum{0.54334858}\hlstd{,} \hlopt{-}\hlnum{0.89861574}\hlstd{)}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{Item} \hlopt{==} \hlstr{"06:15:24"}\hlstd{,} \hlstr{"discriminability"}\hlstd{]} \hlkwb{<-} \hlnum{0.4875}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{Item} \hlopt{==} \hlstr{"16:25:34"}\hlstd{,} \hlstr{"discriminability"}\hlstd{]} \hlkwb{<-} \hlnum{0.3123529}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{Item} \hlopt{==} \hlstr{"26:35:44"}\hlstd{,} \hlstr{"discriminability"}\hlstd{]} \hlkwb{<-} \hlnum{0.2308442}
+    \hlstd{dat[dat}\hlopt{$}\hlstd{Item} \hlopt{==} \hlstr{"36:45:54"}\hlstd{,} \hlstr{"discriminability"}\hlstd{]} \hlkwb{<-} \hlnum{0.1833333}
+    \hlcom{# put dat in better column order}
+    \hlstd{dat} \hlkwb{<-} \hlkwd{subset}\hlstd{(dat,} \hlkwc{select} \hlstd{=} \hlkwd{c}\hlstd{(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number,}
+        \hlstd{Item, discriminability, c_Trl, s_Trl, c_Itm, c_Vag, c_Num, f_Cnd, c_Ord, c_Qty, RT, RT_log, RT_raw,}
+        \hlstd{Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category,}
+        \hlstd{Left, Mid, Right, Instruction, nchar_instr))}
+    \hlcom{# This data set (dat) contains *all* trials 7680 including impossible trials and is mainly for graphs}
+    \hlcom{# comparing different removals}
+    \hlkwd{save}\hlstd{(dat,} \hlkwc{file} \hlstd{=} \hlstr{"data_raw.Rda"}\hlstd{)}
+    \hlkwd{message}\hlstd{(}\hlstr{"Returning processed data"}\hlstd{)}
+    \hlkwd{return}\hlstd{(dat)}
+\hlstd{\}}  \hlcom{# end of function processData}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
+\clearpage
+\subsection{Post-processing}
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlstd{postProcessData} \hlkwb{<-} \hlkwa{function}\hlstd{(}\hlkwc{dat}\hlstd{) \{}
+    \hlcom{# dd removes impossible trials from dat Throw out RT = 1 and RT = 59998, and RTprev = 1 and RTprev =}
+    \hlcom{# 59998 i.e., throw out sticky fingers and timeouts, and the trials that followed sticky fingers and}
+    \hlcom{# timeouts since they were likely affected by unusual previous trials. Also lose impossible trials}
+    \hlstd{dd} \hlkwb{<-} \hlstd{dat}
+    \hlstd{dd}\hlopt{$}\hlstd{RT[dd}\hlopt{$}\hlstd{RT} \hlopt{==} \hlnum{1}\hlstd{]} \hlkwb{<-} \hlnum{NA}
+    \hlstd{dd}\hlopt{$}\hlstd{RT[dd}\hlopt{$}\hlstd{RT} \hlopt{==} \hlnum{59998}\hlstd{]} \hlkwb{<-} \hlnum{NA}
+    \hlstd{dd} \hlkwb{<-} \hlstd{dd[}\hlkwd{complete.cases}\hlstd{(dd), ]}
+    \hlkwd{row.names}\hlstd{(dd)} \hlkwb{<-} \hlkwa{NULL}
+    \hlcom{# add preceding RT: because we removed impossible trials, the value for preceding RT for a trial}
+    \hlcom{# following an impossible trial is the value of the trial that preceded the impossible trial.}
+    \hlstd{dd}\hlopt{$}\hlstd{RTprev} \hlkwb{<-} \hlnum{NA}
+    \hlkwa{for} \hlstd{(s} \hlkwa{in} \hlkwd{levels}\hlstd{(dd}\hlopt{$}\hlstd{Subject)) \{}
+        \hlstd{nrows} \hlkwb{<-} \hlkwd{nrow}\hlstd{(dd[dd}\hlopt{$}\hlstd{Subject} \hlopt{==} \hlstd{s, ])}
+        \hlkwa{for} \hlstd{(i} \hlkwa{in} \hlnum{1}\hlopt{:}\hlstd{nrows) \{}
+            \hlkwa{if} \hlstd{(i} \hlopt{==} \hlnum{1}\hlstd{) \{}
+                \hlstd{dd[dd}\hlopt{$}\hlstd{Subject} \hlopt{==} \hlstd{s,} \hlstr{"RTprev"}\hlstd{][i]} \hlkwb{<-} \hlstd{dd[dd}\hlopt{$}\hlstd{Subject} \hlopt{==} \hlstd{s,} \hlstr{"RT"}\hlstd{][i]}
+            \hlstd{\}} \hlkwa{else} \hlstd{dd[dd}\hlopt{$}\hlstd{Subject} \hlopt{==} \hlstd{s,} \hlstr{"RTprev"}\hlstd{][i]} \hlkwb{<-} \hlstd{dd[dd}\hlopt{$}\hlstd{Subject} \hlopt{==} \hlstd{s,} \hlstr{"RT"}\hlstd{][i} \hlopt{-} \hlnum{1}\hlstd{]}
+        \hlstd{\}}
+    \hlstd{\}}
+    \hlcom{# add transformations of previous RT}
+    \hlstd{dd}\hlopt{$}\hlstd{RTprev_log} \hlkwb{<-} \hlkwd{log}\hlstd{(dd}\hlopt{$}\hlstd{RTprev)}
+    \hlstd{dd}\hlopt{$}\hlstd{RTprev_raw} \hlkwb{<-} \hlstd{dd}\hlopt{$}\hlstd{RTprev}
+    \hlcom{# put dd in better column order}
+    \hlstd{dd} \hlkwb{<-} \hlkwd{subset}\hlstd{(dd,} \hlkwc{select} \hlstd{=} \hlkwd{c}\hlstd{(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number, Item,}
+        \hlstd{discriminability, c_Trl, s_Trl, c_Itm, c_Vag, c_Num, f_Cnd, c_Ord, c_Qty, RT, RT_log, RT_raw,}
+        \hlstd{RTprev, RTprev_log, RTprev_raw, Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side,}
+        \hlstd{response_num, response_side, response_category, Left, Mid, Right, Instruction, nchar_instr))}
+    \hlkwd{save}\hlstd{(dd,} \hlkwc{file} \hlstd{=} \hlstr{"data_processed.Rda"}\hlstd{)}
+    \hlkwd{message}\hlstd{(}\hlstr{"Returning post processed data"}\hlstd{)}
+    \hlkwd{return}\hlstd{(dd)}
+\hlstd{\}}  \hlcom{# end of function postProcessResponses}
+\end{alltt}
+\end{kframe}
+\end{knitrout}
+
 \end{document}

### 2016-08-13 experiments/e2/writeup/Untitled.R

@@ -1,9 +1,0 @@
-# add discriminability with reference to item
-discriminability_range = c(0.7500000, 0.5294118, 0.4090909, 0.3333333)
-discriminability_range_scaled = c(1.3441995, 0.1316642, -0.5297187, -0.9461450)  
-discriminability = c(0.4875000, 0.3123529, 0.2308442, 0.1833333)
-discriminability_scaled = c(1.37582241, 0.06614191, -0.54334858, -0.89861574)
-dat[dat$Item == "06:15:24", "discriminability"] <- 0.4875000
-dat[dat$Item == "16:25:34", "discriminability"] <- 0.3123529
-dat[dat$Item == "26:35:44", "discriminability"] <- 0.2308442
-dat[dat$Item == "36:45:54", "discriminability"] <- 0.1833333

### 2016-08-13 experiments/e2/writeup/measurement.R

@@ -1,25 +1,0 @@
-# number of higher-level cells in the design (abstracting over Order) =
-# Item(4) * Number(2) * Vagueness(2) * Quantity(2) = 32
-# number of measurements in eaach cell = 8
-# 332 * 8 = 256
-# 256 * 30 = 7680
-dd$measurement=0
-dd$cell=0
-for (s in levels(dd$Subject)) {
-  cellcount=0
-  for (i in levels(dd$Item)) {
-    for (n in levels(dd$Number)) {
-      for (v in levels(dd$Vagueness)) {
-        for (q in levels(dd$Quantity)) {
-          measurementcount=0
-          cellcount=cellcount+1
-          for (t in dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q, "Trial"]) {
-            dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "cell"] <- cellcount
-            measurementcount=measurementcount+1
-            dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "measurement"] = measurementcount
-          }
-        }
-      }
-    }
-  }
-}

### 2016-09-14 experiments/e1/writeup/e1.Rmd
renamed the springer chapter template for easier bash-completion
e1 stuff

@@
 
-In this experiment it was possible to identify errors, so error trials were removed. This is because there were only 2 squares and an istruction uniquely identified one of them.
+In this experiment it was possible to identify errors, so error trials were removed. This is because there were only 2 squares and an instruction uniquely identified one of them.
 
@@
 # item_ratio
-# NOT DONE YET
+for (c in unique(dat$condition)) {
+  x=subset(dat, condition==c)
+  x$item_ratio=unique(min(x$leftN,x$rightN)/max(x$leftN,x$rightN))
+  dat[dat$condition==c, 'item_ratio']=x$item_ratio
+}
 write.table(x=dat, "all_trials.txt")
@@
 ```
+
+```{r plotMeansBothV2, fig.width=10, fig.height=4, message=FALSE}
+rts <- summarySEwithin(correct, measurevar="logrt", withinvars=c("vgn","item_ratio"))
+err <- summarySEwithin(alltrials, measurevar="error", withinvars=c("vgn","item_ratio"))
+dodge=0
+
+plot1 =
+  ggplot(rts, aes(y=logrt_norm, x=1-as.numeric(abbreviate(item_ratio)), group=vgn, fill=vgn, shape=vgn, ymin=logrt_norm-ci, ymax=logrt_norm+ci)) + 
+  geom_line(position=position_dodge(dodge)) + 
+  geom_errorbar(width=.01, lty=1, position=position_dodge(dodge)) + 
+  geom_point(pch=21, size=3, position=position_dodge(dodge)) +
+  ggtitle("Response time") + 
+  ylab("log RT") + 
+  xlab("item_ratio Size") +
+  scale_x_continuous(breaks = 1-as.numeric(abbreviate(sort(unique(dat$item_ratio))))) +
+  scale_fill_grey(name=element_blank(), start=0,end=1)+
+  theme_bw()+
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank() ) +
+  theme(aspect.ratio=1)+xlab("")
+
+plot2 = 
+  ggplot(err, aes(y=error_norm, x=1-as.numeric(abbreviate(item_ratio)), group=vgn, fill=vgn, shape=vgn, ymin=error_norm-ci, ymax=error_norm+ci)) + 
+  geom_line(position=position_dodge(dodge)) + 
+  geom_errorbar(width=.01, lty=1, position=position_dodge(dodge)) + 
+  geom_point(pch=21, size=3, position=position_dodge(dodge)) +
+  ggtitle("Error rate") + 
+  ylab("error rate") + 
+  xlab("item_ratio Size") +
+  scale_x_continuous(breaks = 1-as.numeric(abbreviate(sort(unique(dat$item_ratio))))) +
+  scale_fill_grey(name=element_blank(), start=0,end=1)+
+  theme_bw()+
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank() )+
+  theme(aspect.ratio=1)+xlab("")
+
+grid.arrange(plot1,plot2,ncol=2)
+```

### 2016-09-17 papers/2016-Special-Issue/v-special-issue-2016.tex
new structure

--- 2016-07-27 papers/2015/resubmission_drafts/2015_resubmission_draft/vagueness-dec-2015.tex
@@
 \usepackage{rotating}
-\title{Vagueness as cost reduction}
+\title{Is Vagueness Beneficial for Hearers?\\Evidence from Experiments}
 \date{\today}
 \authornote{\today ~\currenttime}
-\shorttitle{Vagueness as cost reduction}
+\shorttitle{Is Vagueness Beneficial for Hearers?}
 \twoauthors{Matt Green}{Kees van Deemter}
@@
 
-\abstract{Much of everyday language is vague, yet standard game-theoretic models make it difficult to see how vague expressions can have benefits over crisp ones in co-operative situations.
-We report a series of experiments that aims to separate the utility of vagueness from the utility of other factors that come together with vagueness.
-We argue that the results support a view of vagueness where the benefits that vague terms exert are due to other influences that vagueness brings with it rather than to vagueness itself.}
+\abstract{Much of everyday language is vague, but the causes of this phenomenon are not well understood.
+Consequently, it is difficult for the designers of a Natural Language Generation (NLG) system to know when and how to let the system generate vague expressions.
+The present paper is an attempt to find out what benefits vagueness might have for readers.% , by means of a series of carefully controlled psycholinguistic experiments that focus on the use of vagueness in referring expressions.
+The work is motivated by NLG but has implications for linguistics, philosophical logic and game theory.
+%
+The paper reports on a series of experiments that aim to separate the utility of vagueness (as defined by the existence of borderline cases) from the utility of other factors that tend to co-occur with vagueness.
+We argue that the evidence so far supports a view of vagueness where the benefits that vague terms exert are due to other influences, rather than to vagueness itself.
+These factors include: low granularity; the use of evaluative words; the avoidance of overtly numerical words; the existence of comparison strategies; and, lastly and more tentatively, a phenomenon that we call range reduction.
+Although it is possible that other types of vague expressions (i.e., vague words outside referential noun phrases) behave differently, our work suggests that vagueness itself may not increase the utility of an expression.
+The paper concludes with a brief discussion of the implications of our work for practical NLG.}
 
@@
 \maketitle               
-\section{Introduction}
-Vagueness pervades the language that we use on a daily basis.
-In everyday use, language use may be called vague for various reasons.\footnote{See e.g.
-the entry ``vague'' in \cite{penguin}.} In most academic use though, the word `vagueness' has a more particular meaning.
-Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1996vagueness} (a similar definition can be found in \citeA{EgreKlinedinst}, among many others).
-The most crucial of these criteria is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects.
+\section{Introduction}\label{introduction}
+Vagueness pervades the language that we use on a daily basis, and the challenge of understanding vague language has been a prominent concern in many areas of logic and linguistics, involving both theoretical and applied work, including the area known as Natural Language Generation (NLG).
+
+NLG systems transform data and formulas into language \cite<e.g.,>[]{reiter2000building}.
+NLG systems routinely make decisions between different formulations of the same information.
+For example, if the temperature is 27.2 degrees Celsius, this could be expressed as ``27.2 degrees'', ``approximately 27 degrees', ``above 25 degrees", or ``warm'', and the system must decide between these.
+The problem is especially important for NLG systems that take numbers as input, for example in the generation of textual weather reports from numerical weather data such as temperature and wind speed \cite{goldberg1994using,turner2006generating}, and medical decision support on the basis of clinical measurement such as oxygen saturation, heart rhythm, etc.
+\cite{Hripcsak01032009, hunter2008summarising, portet2009automatic}.
+Such systems are often forced to make decisions concerning the level of precision in the utterances that they generate on the basis of little more than intuition.
+Even when NLG systems are designed to mimic human language use (e.g., Konstas and Lapata 2013),
+% Ioannis Konstas and Mirella Lapata. 2013. A global model for concept-to-text generation. J. Artif. Intell. Res. (JAIR), 48:305Ð346. 
+there is no guarantee that these decisions taken by these systems benefit readers.
+A better understanding of the benefits (for readers and hearers) of different precision levels would allow these systems to become more useful.
+The present article investigates the benefits,  or de-benefits, of vagueness.
+
+%.
+In the context of an NLG system faced with a practical decision of this kind, Lipman's question becomes ``Under what circumstances should vague terms be produced?''  Relevant applications include weather forecasting
+
+Language use may be called vague for various reasons.\footnote{See e.g.
+the entry ``vague'' in \cite{penguin}.} In most academic use though, the word `vagueness' has a specific meaning.
+Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1996vagueness}, also \citeA{EgreKlinedinst}).
+The crucial criterion is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects.
 By contrast, a word is vague if it is not precise'' \cite[p.\ 1]{lipmanvague}.
@@
 
-Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions, which are used so frequently \cite{Lipman:2000fk, lipmanvague}.
+Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions \cite{Lipman:2000fk, lipmanvague}.
 By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss'' \cite<>[p.~1]{lipmanvague}.
 Lipman explains the point by means of a scenario in which a speaker describes a person to a hearer, who needs to identify that person in the arrivals hall of of an airport.
-Lipman argues that, in such a scenario, a precise description of the person's height (e.g., ``The person's height is 187.96 cm'') would be more useful than a vague one (``The person is tall'').
+In such a scenario, a precise description of the person's height (e.g., ``The person's height is 187.96 cm'') would be more useful than a vague one (``The person is tall'').
 Lipman uses this scenario to explain why standard game theory models of communication \cite<e.g.,>{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs.
-The relevant conditions are, broadly speaking, that both interlocutors know all the relevant facts (e.g., both know the person's height precisely) and that the setting for the communication is co-operative.
+%The relevant conditions are, broadly speaking, that both interlocutors know all the relevant facts (e.g., both know the person's height precisely) and that the setting for the communication is co-operative.
 These conditions exclude deliberate deception, and rhetorical situations like political debate and advertising where the intention is to persuade the interlocutor to adopt some point of view, where the persuasion might not be intended to be in the addressee's best interest.
 
-Lipman argued that such the efficiency loss resulting from vague expressions would be unlikely to have arisen unless there are advantages as well as disadvantages associated with vague expressions.
-Lipman asked, essentially, what these advantages might be, and how they might find a place in a game-theoretical explanation.
+Lipman argued that the efficiency loss resulting from vague expressions would be unlikely to have arisen unless there are advantages as well as disadvantages associated with vague expressions.
+Lipman asked, essentially, what these advantages might be.
+%, and how they might find a place in a game-theoretical explanation.
 In this paper, we focus on the first part of Lipman's question.
-
 Several tentative answers to Lipman's question have been offered  \cite<see>{van2009utility, van2010vagueness}.
-Prominent among these answers is the idea that vague expressions are somehow easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite<e.g.,>[]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
+Prominent among these answers is the idea that vague expressions are easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite<e.g.,>[]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
 For example, \citeA<>[p.\ 11]{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze''.
-We shall refer to this characterisation of the utility of vague language as the \emph{cost reduction} hypothesis.
-The idea of vagueness as cost reduction can take various shapes, but the basic hypothesis is that it is easier for people to think in terms of loosely defined categories (such as ``quite a few'', or ``many'') than in terms of crisply defined ones (such as ``thirteen'', or ``237'').
-This predicts that whenever a vague expression can perform the same communicative task as a precise one, it is rational to choose the vague expression.
+We shall refer to this as the \emph{cost reduction} hypothesis.
+%The idea of vagueness as cost reduction can take various shapes, but the basic hypothesis is that it is easier for people to think in terms of loosely defined categories (such as ``quite a few'', or ``many'') than in terms of crisply defined ones (such as ``thirteen'', or ``237'').
+%This hypothesis predicts that when a vague expression can perform the same communicative task as a precise one, then, the things being equal, it is rational to choose the vague expression.
 The corollary for comprehension is that vague expressions should be understood more readily than precise expressions.
 
-Questions concerning optimal language use have many practical applications.
-Natural Language Generation\footnote{NLG systems take data or formulas as input, and transform them into natural language outputs \cite<>[]{reiter2000building}.
-The process parallels language production in humans.}  (henceforth NLG) systems must make decisions between different formulations of the same information.
-For example, if a man's height is 6 foot 2 inches, this could be expressed as ``187.96 metres'', ``6 foot 2'', or ``tall'', among other ways, and the NLG system must decide between these.
-The problem is particularly relevant for NLG systems that take numbers as input, as many do.
-In the context of an NLG system faced with a practical decision of this kind, Lipman's question becomes ``Under what circumstances should vague terms be produced?''  Relevant applications include weather forecasting on the basis of numerical weather data such as temperature and wind speed \cite{goldberg1994using,turner2006generating}, and medical decision support on the basis of clinical measurement such as oxygen saturation, heart rhythm, etc.
-\cite{Hripcsak01032009, hunter2008summarising, portet2009automatic}.
-At present, such NLG systems are often forced to make decisions concerning the level of precision in the utterances that they generate (e.g., ``the temperature will be in the high twenties tomorrow'') on the basis of little more than intuition.
-A better understanding of the benefits of different precision levels for readers would allow these systems to become more useful.
-
-The cost reduction hypothesis is of direct relevance to psycholinguists interested in language comprehension, and additionally to psycholinguists interested in language production, for example in connection with the question of audience design \cite{Clark1982287}.
+%The cost reduction hypothesis is of direct relevance to psycholinguists interested in language comprehension, and additionally to psycholinguists interested in language production, for example in connection with the question of audience design \cite{Clark1982287}.
 For, to the extent that speakers and writers choose vague expressions over and above crisp ones because the former are easier to process for hearers than the latter, the cost reduction hypothesis suggests that speakers design their utterance for optimal benefit to their hearers -- out of altruism, so to speak.
 
-The utility of vagueness is the attested aim of a small number of studies, but most of these have focussed on vagueness in a different sense, and focussing on different types of benefits for hearers.
+Charting the utility of vagueness is the attested aim of a small number of studies, but most of these have focussed on vagueness in a different sense, and focussing on different types of benefits for hearers.
 Two recent studies can illustrate both issues.
@@
 This limits the applicability of our results.
-It also has the advantage that we can explore the costs and benefits of vagueness more thoroughly in that context.
+However, it has the advantage that it enables us to explore the costs and benefits of vagueness more thoroughly.
 Since one prevalent view of vagueness is that a vague expression is never preferable to a crisp equivalent, a demonstration of a benefit for vagueness in any context would advance the discussion.
@@
 Experiment 1 showed us that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment could not distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format; and mandated a comparison strategy rather than a matching strategy.
-Experiment 2 showed us that number-avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional explanatory power in either the verbal format instructions or the numerical format instructions when we generated verbal and numerical versions of both crisp and vague instructions.
+Experiment 2 showed us that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional explanatory power in either the verbal format instructions or the numerical format instructions when we generated verbal and numerical versions of both crisp and vague instructions.
 However the experiment could not distinguish benefits of number-avoidance in the verbal instructions from benefits of the comparison selection task: the verbal instructions also mandated a comparison strategy rather than a matching strategy.
@@
 In recent years a number of studies, focussing typically on red wine, have suggested that alcohol, consumed in low doses, may have health benefits.
-An alternative explanation, however, asserts that it is not the alcohol in the wine that is beneficial, but the antioxidants from grapes.
-If this alternative explanation is correct then alcohol may not be as beneficial as some of us would have it.
-% KvD Good hey?
+An alternative explanation, however, asserts that it is not the alcohol in the wine that was beneficial, but antioxidants from grapes.
+If this alternative explanation is correct then alcohol may not be as beneficial as some would like to think.\\[2ex]
+%
+{\bf Implications for practical NLG systems.} Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
+The question of how a particular piece of quantitative information is best conveyed through language is certainly an important one for practical NLG.
+Years of research on the logic of vagueness in natural language have primed the research community to expect that the utility of vagueness is an important part of the answer, but our findings call this into question.
+%It would appear that this is true for most vague expressions, and not only the ones investigated in our experiments, where a speaker is referring to a square with dots.
 
-\bibliography{vagueness-dec-2015}
+Consider, once again, a number of ways in which a given temperature can be reported. Given a temperature measurement (or prediction) as input, an NLG system might justifiably say that it was
+%
+\begin{quote}
+(a) {\em 27.2 degrees}, or\\
+(b) {\em approximately 27 degrees}, or\\
+(c) {\em above 25 degrees}, or\\
+(d) {\em warm}, 
+\end{quote}
+%
+among other candidate expressions.
+If standard accounts are to be believed, then options (a) and (c) convey crisp information, whereas (b) and (d) are vague (i.e., they permit borderline cases).
+But, to the best of our knowledge, there is no experimental evidence that the fact in itself that an expression is vague confers a benefit (or a de-benefit) on it for hearers.
+Rather than asking whether a candidate expression is vague, other questions might shed more light on the NLG system's choice, similar (though not necessarily identical) to the ones identified in our empirical studies.
+For a start, both (c) and (d) designate an interval that is half-bounded, whereas both (a) and (b) express closed intervals, and this might affect their comprehension by a hearer (cf., our finding about comparison versus matching strategies).
+Furthermore, the expressions (c) and (d) -- one of which is vague while the other is crisp -- appear to convey very similar amounts of information, saying that the temperature is higher than some (fairly high) standard.
+Perhaps most importantly, "warm" is shorter than the other three expressions and avoids the use of numbers, and our experiments suggest that this may make ``warm" more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of the word ``warm" (see section 1 and \citeA{peters2009bringing}).
+
+One way to see why vagueness may not matter as much for NLG -- and for human communication more generally -- as is often thought is the following thought experiment.
+Many speakers understand the word ``warm" as vague.
+Now suppose a group of users agreed to give it a precise definition; according to this precisification, ``warm" means $>25$ degrees (as in (c)).
+It seems unlikely that this change, from a vague meaning to a crisp one, would change the utility of the word.
+
+\bibliography{v-special-issue-2016}
 \clearpage

### 2016-09-17 papers/2016-Springer-Chapter/v-chapter-tables.tex
new structure

\begin{table}
\caption{Table of instructions for the pair (5,25). Experiment 1}
\label{instructionse1} 
\begin{tabular}{lll}
\hline\noalign{\smallskip}
vagueness&example\\
\noalign{\smallskip}\svhline\noalign{\smallskip}
crisp 	& 	Choose the square with 5 dots \\
vague	&	Choose the square with few dots\\
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
$^a$ Table foot note (with superscript)\\
\end{table}

\begin{table}
\caption{Table of instructions arranged by condition for the triple (6,15,24). Experiment 2}
\label{instructionse2} 
\begin{tabular}{llll}
\hline\noalign{\smallskip}
vagueness&instruction format&selection task&example\\
\noalign{\smallskip}\svhline\noalign{\smallskip}
crisp 	& 	numeric	& matching	&	Choose the square with 6 dots \\
vague	&	numeric 	& matching	&	Choose a square with about 10 dots\\
crisp		&	verbal	& comparison	&	Choose the square with the fewest dots\\
vague	&	verbal	& comparison	&	Choose a square with few dots\\
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
$^a$ Table foot note (with superscript)\\
\end{table}

\begin{table}
\caption{Experiment 3 Instructions assuming 6,15,24 dots as the Item, and showing fewer instead of more}
\label{instructionse3} 
\begin{tabular}{llll}
\hline\noalign{\smallskip}
vagueness&instruction format&selection task&instruction\\
\noalign{\smallskip}\svhline\noalign{\smallskip}
crisp 	& numeric		&matching 	& Choose a square with 6 dots \\ 
crisp 	& numeric		&comparison 	& Choose a square with fewer than 20 dots \\
vague 	& numeric		&matching 	& Choose a square with about 10 dots \\ 
vague 	& numeric		&comparison 	& Choose a square with far fewer than 20 dots \\ 
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
$^a$ Table foot note (with superscript)\\
\end{table}

\begin{table}
\caption{Experiment 4: Instructions}
\label{instructionse4}
\begin{tabular}{llll}
\hline\noalign{\smallskip}
vagueness&instruction format& selection task&instruction\\
\noalign{\smallskip}\svhline\noalign{\smallskip}
crisp 	& verbal	&matching 	& Choose a square with the same number of dots as the target \\ 
crisp 	& verbal	&comparison	& Choose a square with fewer dots than the target \\
vague 	& verbal	&matching 	& Choose a square with about the same number of dots as the target \\ 
vague 	& verbal	&comparison	& Choose a square with far fewer dots than the target \\ 
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
$^a$ Table foot note (with superscript)\\
\end{table}

\begin{table}
\caption{Vagueness as range reduction}
\label{vaguenessasrangereduction} 
\begin{tabular}{lll}
\hline\noalign{\smallskip}
vagueness & selection task & candidates\\
\noalign{\smallskip}\svhline\noalign{\smallskip}
crisp 	& matching 	& 1 candidate\\
crisp 	& comparison 	& 2 candidates\\
vague 	& matching 	& 2 candidates\\
vague 	& comparison 	& 1 candidate\\
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
$^a$ Table foot note (with superscript)\\
\end{table}

### 2016-09-17 papers/2016-Special-Issue/v-special-issue-2016.tex
holding point

@@
 emphasise the evaluative content inherent in these categories, and the affective potential of the evaluative content rather than the vagueness of the terms like {\em Fair}.
+\clearpage
 %\\[2ex]
-\subsection{}
+\subsection{General Methodology}
 The experiments reported in the present paper put the cost reduction hypothesis to the test. 
@@
 In this context, speed and accuracy of responses are the key dimensions on which the different references can be compared.
-The stimuli in the experiments were sets of dot arrays containing various numbers of dots.
-The forced choice was to identify one dot array given a reference to a given quantity of dots.
-We manipulated the references in several ways across a series of four experiments.
 
-Our main manipulation was always of vagueness: we constructed crisp and vague versions of references to the same dot array.
-For example, in experiment 1 the instruction presented to the participant could be \emph{Choose the square with many dots.} (vague condition), or \emph{Choose the square with 20 dots.} (crisp condition), identifying the same dot array.
+Each stimulus in the experiments was a set of dot arrays containing various number of dots, together with a preceding instruction (in the form of a referring expression) to choose one of the arrays with respect to its number of dots, as quickly as possible while avoiding errors.
+We manipulated the stimuli in several ways across a series of four experiments.
 
-Ideally, we would manipulate vagueness independently of other variables.
-However, in practice, and given the constraints imposed by the use of natural language, any manipulation of vagueness introduces variance along other dimensions as well.
-Therefore we carried out further experiments to try to address and control variance along other dimensions.
+% added/moved out of expt 1 by matt 17 sept
+There is evidence that when the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the numerosity of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene199647}.
+Where a number was mentioned in the instruction, it was always in the form of an Arabic numeral.
 
-One non-vagueness source of variance was the difference between numerical and verbal format in the references.
-For example, in the instructions for experiment 1 given above, \emph{many} is in verbal format (as well as vague) and \emph{20} is in numerical format (as well as crisp).
-Therefore any difference observed between responses to the \emph{many} instruction and the \emph{20} instruction could be due either to vagueness or to {\bf instruction format}.
-It is possible to create vague and crisp references in each of these instruction formats: for example in numerical instruction format we can have \emph{20} (crisp) and \emph{about 20} (vague), and in verbal instruction format we can have \emph{the most} (crisp) and \emph{many} (vague).
-Experiment 2 varied vagueness and instruction format factorially in a similar way to this, to try to tease apart these two sources of variance.
+% added/moved out of expt 1 by matt 17 sept
+There is evidence that when two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}.
+We controlled which side the smaller number appeared on to avoid systematic influences of this effect.
 
-Another potential non-vagueness source of variance between responses to the \emph{many} (verbal vague) references and the \emph{20} (numeric crisp) references lies in the method used to identify a referent (we call this the {\bf selection algorithm} source of variance).
-Identifying the array with \emph{many} dots might be done by a \emph{comparison} algorithm, identifying that one array is more numerous than the others without establishing the numerosity of any arrays, whereas identifying the array with \emph{20} dots, which might be done by a \emph{matching} algorithm would require an estimate of the numerosity of the arrays.
-It seems reasonable that comparison would be faster than matching because it does not require estimates of cardinality.
-Furthermore the problem persists when using references like \emph{the most} (verbal crisp) and \emph{about 20} (numeric vague), since they too vary not only along the vague/crisp dimension and the numeric/verbal dimension but also along the comparison/matching dimension.
-Experiments 3 and 4 addressed the \emph{selection algorithm} source of variance - experiment 3 crossed vague/crisp and comparison/matching factorially using only numerical references, and experiment 4 crossed vague/crisp and comparison/matching factorially using only verbal references.
+% General methodology moved to here
+All the experiments shared the following properties: 
+The experiment was conducted in a quiet room.
+All participants self-reported fluency in English, and had normal, or corrected-to-normal vision.
+Participants were asked to respond quickly while avoiding errors.
+Stimuli were created and presented using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.
+There was a block of practice trials after which participants could ask any questions, following which the experimenter left the room.
+The position of the dots was randomised per-trial.
+The order in which trials were presented was randomised per-participant.
+A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants.
+
+\clearpage
 
@@
 The participant was presented with an instruction like ``Choose the square with many dots'' in the vague conditions, or ``Choose the square with 20 dots'' in the crisp conditions. 
-Then two dot arrays were presented in the form of squares containing a number of dots.
-Fig \ref{stimuluse1} shows an example stimulus.
-The participant was required to identify the square that corresponded with the instruction, by pressing the appropriate key.
+Then two dot arrays were presented in the form of squares containing a number of dots. %Fig \ref{stimuluse1} shows an example stimulus. 
+The participant was required to identify the array that corresponded with the instruction, by pressing the appropriate key.
 Response time and accuracy were recorded for analysis.
 
-Our main manipulation was of the vagueness of the instruction, with two levels, vague and crisp.
-Table \ref{instructionse1} shows examples from each condition.
+Our main manipulation was of the vagueness of the instruction, with two levels, vague and crisp. %Table \ref{instructionse1} shows examples from each condition.  
 We also manipulated how discriminable the dot arrays were. 
@@
 This led to numerical differences of 5, 10, 15, and 20, with lower differences resulting in less discriminable arrays and larger differences representing more discriminable arrays.
-There is evidence that when the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the numerosity of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene199647}.
-Where a number was mentioned in the instruction, it was always in the form of an Arabic numeral.
-
-The instructions indicated the larger of the two dot arrays equally often as they indicated the smaller of the dot arrays, so that participants could not systematically choose the larger or smaller array as a successful response strategy.
-There is evidence that when two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}.
-We controlled which side the smaller number appeared on to avoid systematic influences of this effect.
-
-\begin{table}[htbp]
-\caption{Table of instructions for the pair (5,25). Experiment 1}
-\label{instructionse1}
-\begin{tabular}{rl}
-\toprule
-vagueness&example\\
-\midrule
-crisp 	& 	Choose the square with 5 dots \\
-vague	&	Choose the square with few dots\\
-\bottomrule
-\end{tabular}
-\end{table}
 
 \subsection{Hypotheses}
-The {\bf cost reduction} hypothesis predicts that responses will be faster and more accurate for the vague conditions than for the crisp conditions because the vague conditions impose a lower cognitive load than the crisp conditions.
+(1) A main effect of vagueness: vague instructions impose a lower cognitive load for the comprehender than crisp alternatives.
 
-The {\bf instruction format} hypothesis predicts that responses will be faster and more accurate in the vague (i.e., verbal) conditions than in the crisp (i.e., numerical) conditions because numbers are harder to process than verbal references to quantities.
+(2) A main effect advantage of increasing numerical distance: the task will become easier as the numerical distance increases, because the two arrays are more discriminable.
 
-The {\bf selection algorithm} account predicts that responses will be faster and more accurate in the vague (i.e., comparison) conditions than in the crisp (i.e., matching) conditions because it is easier to carry out comparison than matching.
-
-Thus all three accounts predict a main effect advantage of vagueness, but for different reasons.
-All three accounts also predict a main effect advantage for greater numerical distance, because greater numerical distances lead to more easily discriminable arrays.
-The accounts also share the prediction of an interaction effect whereby the main effect advantage of vagueness should diminish as numerical distance increases.
-
-% KvD I'm tentatively starting with a general explanation of the methodology of this group of experiments.
-
-%{\bf General.} All our experiments shared essentially the same methodology: Participants aged between ..., self-reported fluent speakers of English, were recruited by ....
-and were paid 10 ponds for participating.
-All had normal, or corrected-to-normal vision.
-A MacBook ...
-Stimuli were created ...
-In all except the last experiment (Experiment 5), items were presented as follows: (...) (IF WE DO ADD THIS GENERAL DESCRIPTION THEN REMOVE THE RELEVANT INFO FROM THE DESCRIPTION OF EACH OF THE INDIVIDUAL EXPERIMENTS.)
+(3) An interaction between vagueness and numerical distance: i.e., any facilitation for vagueness should be greater at smaller numerical distances.
 
 \subsection{Method}
-Twenty participants were recruited by mailing list and paid ten pounds for participating.
-Participants were aged between 18 and 45, with a median age of 26.
-All participants self-reported fluency in English, and had normal, or corrected-to-normal vision.
-A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants.
-Stimuli were created and presented using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.
-We needed to control which side the target was presented on.
-Each pair was presented equally often with with the target on the left, as with the target on the right.
-We also needed to control whether the target was the smaller or larger number.
-The items with numbers less than 25 (pairs 1 to 4) formed a group with the smaller number as target and the other items (pairs 5 to 8) formed a balancing group with the larger number as the target number.
-
-The experiment was conducted in a quiet room.
-On arrival in the room, the participant was told that he or she would be presented with an instruction to choose one of two squares by reference to how many dots it contained.
+Twenty participants were recruited by mailing list; aged between 18 and 45 with a median age of 26; and paid ten pounds for participating. 
+On arrival in the room, each participant was told that on each trial he or she would be presented with an instruction to choose one of two squares on screen by reference to how many dots the square contained.
 Participants were required to press the space key after reading the instruction. 
-Then there was a central fixation cross for 1000 ms, and a blank screen for 500 ms, followed by the squares and dots (without repetition of the referring expression).
-The position of the dots was randomised per-trial.
-Response time was measured as the latency between the presentation of the dots and squares, and the keypress identifying the decision; in this way, the response times can be separated from time spent reading the instructions, which is important since we are only interested in the former.
-
-The display would stay on screen until the participant responded (there was no timing-out).
-Participants were asked to respond quickly while avoiding errors.
-There were 8 practice trials, after which the participant was invited to ask any questions about procedure.
-After answering these the experimenter left the cubicle for the duration of the experiment.
-The order in which trials were presented was randomised per-participant.
+Then there was a central fixation cross for 1000 ms, and a blank screen for 500 ms, followed by the array (without repetition of the referring expression). 
+The arrays would stay on screen until the participant responded (there was no timing-out). 
+Response time was measured as the latency between the presentation of the arrays, and the keypress identifying the choice: in this way, the response time was separated from time spent reading the instructions, which is important since we are only interested in the former.
 There were 256 trials, presented in 4 blocks of 64 trials each, between which the participant could rest. 
@@
 
-\begin{figure}[tbp]
-\fitfigure{images/stimuluse1}
-\caption{An example stimulus from Experiment 1.
-First the referring expression was presented (left panel).
-Then after a keypress, and a fixation cross (not pictured) the squares and dots were presented without repetition of the referring expression (right panel)}
-\label{stimuluse1}
-\end{figure}
-
 \subsection{Results}
-
 A response was counted as erroneous if the square with the wrong number of dots was chosen (when the instruction contained a number); if the square with the larger number of dots was selected (when the instruction was ``Choose the square with few dots"); or if the square with the smaller number of dots was selected (when the instruction was ``Choose the square with many dots".) % KvD Is this addition defining "erroneous" OK?
 RTs for trials with erroneous responses were discarded, leading to the loss of 354 trials from 5120, representing 6.9\% of the trials. 
-The correct response RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of 160 trials, 3.4\% of the correct responses.
-Means for response times and error rates are given in Fig.
-(\ref{resultse2}).
-A linear mixed model of RT was built using as independent variables vagueness and gap size and their interaction, with random slopes for vagueness and gap size over participants.
-Vagueness was sum coded: vague $= -.5$, crisp $= .5$; gap size was Helmert coded.
+The correct response RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of a further 160 trials, or 3.4\% of the remaining correct responses. 
+Means for response times and error rates are given in Fig. (\ref{resultse2}).
+A linear mixed model of RT was built using as independent variables vagueness and numerical distance and their interaction, with random slopes for vagueness and numerical distance over participants. 
+Vagueness was sum coded: vague $= -.5$, crisp $= .5$; numerical distance was Helmert coded. 
 Helmert contrasts compare each level against the mean of the previous levels. 
-Level one of this contrast is gap size 5 compared with gap size 10; level two is the mean of gap size 5 and gap size 10 versus gap size 15; and level three is the mean of gap sizes 5, 10 and 15 versus gap size 20.
-$p$ values were calculated using the R package \emph{lmerTest} \cite{lmerTest}.
+Level one of this contrast is numerical distance 5 compared with numerical distance 10; level two is the mean of numerical distance 5 and numerical distance 10 versus numerical distance 15; and level three is the mean of numerical distances 5, 10 and 15 versus numerical distance 20.
+All $p$ values were calculated using the R package \emph{lmerTest} \cite{lmerTest}.
 
-RTs were faster for vague instructions ($\beta=.109, se=.022, t=4.9, p<.001$). 
-%
-RT grew faster as gap size increased: level one $(\beta=-.116, se=.012, t=-9.3, p<.05)$, level 2 ($\beta=-.103, se=.009, t=-11.1, p<.001$) and level three ($\beta=-.082, se=.007, t=-11.0, p<.05$).
-Since discriminability of the dot arrays is easier for larger gap sizes, discriminability probably underlies this effect.
-Gap size and vagueness interacted significantly for larger gap sizes when modelling RT.
-% KvD Is it worth saying how they interacted?
-The interactions at the different levels of gap size were: level one: ($\beta=.001, se=.016, t=.03, p=.974$); level two ($\beta=-.039, se=.009, t=-4.460, p<.001$); level three ($\beta=-.043, se=.006,t=-7.0, p<.001$).
-In the crisp conditions RTs started out much slower than in the vague conditions, at the smallest gap size, but the two conditions converged to very fast times at the largest gap size.
-There were diminishing returns for vagueness as gap size increased.
+(1) RTs were faster for vague instructions than for crisp instructions ($\beta=.109, se=.022, t=4.9, p<.001$). 
+(2) RTs grew faster as numerical distance increased: level one $(\beta=-.116, se=.012, t=-9.3, p<.05)$, level 2 ($\beta=-.103, se=.009, t=-11.1, p<.001$) and level three ($\beta=-.082, se=.007, t=-11.0, p<.05$).
+Since discriminability of the arrays is easier for larger numerical distances, discriminability probably underlies this effect. 
+(3) Numerical distance and vagueness interacted significantly for larger numerical distances when modelling RT: essentially there were diminishing returns for vagueness as numerical distance increased.% KvD Is it worth saying how they interacted?
+% MG I think this is answered now?
+The interactions at the different levels of numerical distance were: level one: ($\beta=.001, se=.016, t=.03, p=.974$); level two ($\beta=-.039, se=.009, t=-4.460, p<.001$); level three ($\beta=-.043, se=.006,t=-7.0, p<.001$).
+In the crisp conditions RTs started out much slower than in the vague conditions, at the smallest numerical distance, but the two conditions converged to very fast times at the largest numerical distance.
 
-\label{accann}
-Error rate data were analysed using a generalized logit mixed model \cite{jaeger2008categorical}, with vagueness and gap size and their interaction as independent variables, and with random slopes for vagueness and gap size over participants.
-%
+Error rate data were analysed using a generalized logit mixed model \cite{jaeger2008categorical}, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
 The effect of vagueness on error rates approached significance, with the vague instructions leading to fewer errors ($\beta=.307, se=.173, t=1.8, p=.077$).
-%
-Error rates decreased as gap size increased: level one ($\beta=-.585, se=.092, ,z=-6.4, p<.001$), level 2 ($\beta=-.434, z=-5.3,p<.001$) and level three ($\beta=-.250,z=-4.1, p<.001$). 
-%
-Error rates were greater in the crisp conditions than the vague conditions when gap size was small, and this difference diminished with increasing gap size until it reversed at the biggest gap size.
+Error rates decreased as numerical distance increased: level one ($\beta=-.585, se=.092, ,z=-6.4, p<.001$), level 2 ($\beta=-.434, z=-5.3,p<.001$) and level three ($\beta=-.250,z=-4.1, p<.001$). 
+Error rates were greater in the crisp conditions than the vague conditions when numerical distance was small, and this difference diminished with increasing numerical distance until it reversed at the biggest numerical distance.
+
 \begin{figure}[htbp]
 \fitfigure{images/resultse1.pdf}
-\caption{Experiment 1 results. }
+\caption{Experiment 1 results: response times and error rates}
 \label{resultse1}
@@
 \subsection{Discussion}
-
-Results from experiment 1 were in line with the three main accounts.
-Responses were faster and more accurate when the instructions were in the vague (or verbal, or comparison) conditions than when they were in the crisp (or numerical, or selection) conditions.
-The speed and accuracy diminished as the arrays became more easily discriminable, and the advantage for vague (or verbal, or comparison) instructions tailed off as the arrays became more easily discriminable.
+Responses were faster and more accurate when the instructions were in the vague conditions than when they were in the crisp conditions. 
+The speed and accuracy diminished as the arrays became more easily discriminable, and the advantage for vague instructions tailed off as the arrays became more easily discriminable.
 
 The cost reduction hypothesis explains the vagueness advantage by claiming that the vague referring expressions place less cognitive load on the comprehender than the crisp referring expressions. 
-It explains the diminishing returns for vagueness in more-discriminable stimuli by claiming that load is low in both conditions for the easily-discriminable stimuli, and that therefore there is no extra benefit to be had from vagueness in the easily-discriminable stimuli.
+It explains the diminishing returns for vagueness in more-discriminable stimuli (i.e., the vagueness by numerical distance interaction) by claiming that load is low in both conditions for the easily-discriminable stimuli, and that therefore there is no extra benefit to be had from vagueness in the easily-discriminable stimuli.
 
-The instruction format hypothesis explains the main effect advantage for the vague instructions by observing that the vague instructions used verbal quantifiers whereas the crisp conditions used numerical quantifiers.
-Under this account it is avoiding numbers that explains the vagueness advantage main effect.
-The main effect of numerical distance is explained by assuming that larger distances result in more easily discriminable arrays.
-The vagueness x numerical distance interaction can be explained by assuming that the numerical quantifiers make the task particularly challenging when the stimuli are less discriminable.
-
-The selection algorithm account explains the main effect advantage of the vague conditions as due to the vague conditions allowing a comparison strategy rather than a matching strategy in the crisp conditions.
-The main effect of numerical distance is explained by claiming that comparison is easier for more discriminable arrays.
-The diminishing returns for vagueness as numerical distance grows are explained by claiming that the advantage of being able to use comparison is greater for less discriminable arrays than for more discriminable arrays.
-
+\clearpage
 \section{Experiment 2}
@@
 The main result from experiment 1 was that responses were faster and more accurate for vague instructions than for crisp instructions.
-This finding can be interpreted in line with three different accounts: the cost reduction account, the instruction format account, and the selection algorithm account.
 
-In experiment 2 we set out to distinguish between the cost reduction account and the instruction format account, leaving the selection algorithm account for experiments 3 and 4.
-Contrast an expression from the vague condition: `the square with few dots' with an expression from the crisp condition: `the square with 15 dots'.
-One difference is that `few' is vague (or at least has the potential for vagueness) and `15' is crisp.
-Another difference is that `few' is a linguistic, or verbal quantifier while `15' is a numerical quantifier, in the sense that a number is mentioned explicitly.
-Since these two differences were confounded in Experiment 1, the vagueness advantage finding is vulnerable to an alternative interpretation, that what we saw as a vagueness advantage was in contrast an advantage for the verbal form of the quantifier.
-In the present experiment 2 we pitted these alternative interpretations against each other in a factorial design.
+Contrast, from Experiment 1, an expression from the vague condition: `the square with few dots' with an expression from the crisp condition: `the square with 5 dots'. 
+One difference is that `few' is vague (or at least has the potential for vagueness) and `5' is crisp. 
+Another difference is that `few' is a linguistic, or verbal quantifier while `5' is a numerical quantifier, in the sense that a number is mentioned explicitly. 
+Since these two differences could not be separated in Experiment 1, the vagueness advantage finding is vulnerable to an alternative interpretation, that what we saw as a vagueness advantage was in contrast an advantage for the verbal form of the quantifier.
+In Experiment 2 we created verbal and numeric versions of each of the vague and crisp instructions so that we could compare vague and crisp conditions while taking account of verbal/numeric format.%In the present experiment 2 we pitted these alternative interpretations against each other in a factorial design.
 
-In Experiment 1, the participants chose one of two squares.
-The `vague' quantifiers (e.g., `few') uniquely identified one square.
-Recall our definition of vague  -- ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise''.
-In the first experiment, the quantifiers in the vague conditions did not really meet this definition.
-This is because there were no borderline cases of the referent that could make the referent set `not well-defined'.
-Experiment 2 used three squares so that the vague quantifiers always had more than one possible referent.
-To enhance the potential for the vague quantifiers to have true vagueness in Experiment 2, we also used indefinite articles in the vague instructions.
-Thus, instead of the instruction \emph{Choose the square with few dots}, we used \emph{Choose a square with few dots}.
+In Experiment 1, the participants chose one of two squares: therefore the `vague' quantifiers (e.g., `few') uniquely identified one square. 
+Recall our definition of vague  -- ``a word is precise if it describes a well-defined set of objects. By contrast, a word is vague if it is not precise''.  
+In Experiment 1, it might be said that the quantifiers in the vague conditions did not realise their potential for vagueness. 
+This is because there were no borderline cases of the referent that could make the referent set `not well-defined', and perhaps because using definite articles in the instructions implied that only one option was correct.
+Using error feedback in experiment 1 could have exacerbated this.
+Experiment 2 used three squares so that the vague quantifiers always had more than one possible referent, and used indefinite articles in the vague instructions to avoid the impression that only one response counted as correct, and was carried out without error feedback.
+An indication that the potential for vagueness was realised in Experiment 2 is that the borderline response was chosen fairly often --16\% of the time.
 
-In experiment 2, an item was a referring expression followed by a triple of numbers, representing the number of squares in the left, middle, and right squares.
+In Experiment 2, an item was a referring expression followed by a triple of numbers, representing the number of squares in the left, middle, and right squares. 
 We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54). 
 Each triple had the following properties: it comprised three squares (instead of two as in Experiment 1); the central number was always presented in the middle of the three; there were two flanking numbers where one was smaller than the central number and one was bigger.
-
-There was a numerical and a verbal version of each of the vague and crisp referring expressions. See Table \ref{instructionse2} for examples.
-The {\em vague numerical condition}'s referring expression was, on half of the presentations in that condition, ``Choose a square with about 10 dots''.
+Examples of crisp and vague versions of the numerical and verbal instructions follow: the examples assume the triple (6,15,24) and reference to the smaller numbers of dots, such that 6 was classified as the expected response; 15 was classified as the borderline response; and 24 was classified as the incorrect response.
+In the {\em vague numerical} condition we used ``Choose a square with about 10 dots''.
 None of the squares displayed contained 10 dots.
-10 is slightly closer to 6 than to 15.
-Therefore the best referent for this referring expression was the square with 6 dots;  the borderline response was the square with 15 dots; and the poorest referent was the square with 24 dots.
-On the other half of the presentations in that condition the referring expression was ``Choose a square with about 20 dots''.
-The {\em crisp numerical} conditions's referring expression was ``Choose the square with 6 dots'' on half the presentations in that condition and ``Choose the square with 24 dots'' on the other half.
-One square did contain the exact number mentioned.
-In the {\em crisp verbal} condition, we used the referring expression ``Choose a square with fewer than $20$ (more than $10$) dots".
-The {\em vague numerical} condition's referring expression was ``Choose the square with far fewer than $20$ dots'' on half of presentations, making the squares with 6 and $15$ dots possible referents; in the other case, the referring expression was ``Choose the square with far more than $10$ dots'', which made $15$ and $24$ possible referents.The {\em vague verbal} condition's referring expression was ``Choose a square with few dots'' on half of the presentations, and ``Choose a square with many dots'' on the other half.
-For this condition, the best referent was the square with 6 dots for `few' and 24 dots for `many'; the borderline case was the square with 15 dots; and the remaining square was the poorest referent.
-
-An indication that the manipulation of vagueness was successful is that participants chose the borderline case square on 16\% of trials.
-
-\begin{table}[htbp]
-\caption{Table of instructions arranged by condition for the triple (6,15,24). Experiment 2}
-\label{instructionse2}
-\begin{tabular}{rlll}
-\toprule
-vagueness&instruction format&selection task&example\\
-\midrule
-crisp 	& 	numeric	& matching	&Choose the square with 6 dots \\
-vague	&	numeric & matching	&Choose a square with about 10 dots\\
-crisp	&	verbal	& comparison	&Choose the square with the fewest dots\\
-vague	&	verbal	& comparison	&Choose a square with few dots\\
-\bottomrule
-\end{tabular}
-\end{table}
+10 is slightly closer to 6 than to 15, justifying 6 as the best response and 15 as the borderline response.
+In the {\em vague verbal} condition we used ``Choose a square with few dots''.
+In the {\em crisp numerical} condition we used ``Choose the square with 6 dots''. One square did contain the exact number mentioned.
+In the {\em crisp verbal} condition, we used ``Choose the square with the fewest dots".
 
 \subsection{Hypotheses}
-The cost reduction account predicts a main effect of vagueness such that the vague instruction conditions attract faster responses than the crisp instruction conditions.
-The cost reduction account also predicts an interaction with the instruction format (or selection algorithm) variable: at each level of instruction format (or selection algorithm) the vague condition should attract faster responses than the crisp condition.%why is this interaction predicted?
+(1) A main effect advantage of vagueness.
 
-The instruction format account predicts a main effect of instruction format such that the numeric instruction conditions will attract longer response times, but has no prediction of any vagueness effects or interaction effects with vagueness.
+(2) An advantage of vagueness both in \emph{numeric} and in \emph{verbal} instructions.
 
-The selection task account predicts the same main effect but treats it as an effect of the task mandated by the instruction, with the matching task conditions predicted to take longer than the comparison task conditions.
+(3) No large main effect of instruction format.
+%MG If we are right that vagueness was responsible for the vagueness advantage in e 1, that could also have been a number avoidance advantage, then we would not expect that number avoidance, or, instruction format, plays a role.
+
+(4) On the basis of Experiment 1, we would expect an effect of array, with faster responses to more discriminable arrays
 
@@
 
-Thirty participants were recruited and paid in the same way as Experiment 1.
-They were aged between 18 and 45 with a median age of 28.
-All participants self-reported fluency in English, and had normal, or corrected-to-normal vision.
-
-The same apparatus was used as for Experiment 1.
-
+Participants were aged between 18 and 45 with a median age of 28. 
 We presented participants with 256 trials, arranged in 4 blocks each with 64 trials. 
-Each triple was presented in each condition 16 times.
-8 of these identified the larger number, and 8 the smaller.
-We manipulated two independent variables: vagueness with two levels (vague and crisp); and instruction format with two levels (numeric, verbal).
-In this experiment the selection task variable mapped onto the instruction format variable: the numeric instruction conditions both mandated a matching selection task and the verbal instruction conditions both mandated a comparison task.
-This yielded four conditions (vague numeric; vague verbal; crisp numeric; crisp verbal).
-Each condition had a different referring expression, as follows, using the triple $(6,15,24)$ as an example: Choose a square with about 10 dots; Choose a square with few dots; Choose the square with 6 dots; Choose the square with the fewest dots.
+We manipulated as independent variables vagueness and instruction format, yielding four conditions, \emph{vague numeric}; \emph{vague verbal}; \emph{crisp numeric}; \emph{crisp verbal}.
 We measured two dependent variables: response time; and the probability of a participant choosing the borderline case.
-
-Fig (\ref{stimuluse2}) gives an example stimulus.
-First, the referring expression that constituted the instruction for that trial was displayed.
+On each trial, first the referring expression that constituted the instruction for that trial was displayed. 
 The participant then pressed a key to indicate that he or she had read the instruction. 
 After 1 second, the squares and dots were presented, while preserving the text of the referring expression. 
-The position of the dots in the squares was randomised per-trial.
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=.4\textwidth]{images/stimuluse2}
-\caption{An example stimulus from Experiment 2}
-\label{stimuluse2}
-\end{figure}
-
-The experiment was conducted in a small quiet room.
-On arrival in the cubicle, the participant was told that he or she would be presented with objects on screen and required to choose one in response to an instruction on screen, by pressing the button corresponding with the object.
-There were 5 practice trials.
-After the practice trials the experimenter left the room.
-There were 4 blocks of 64 trials each.
-In between blocks the participant had the opportunity to rest before continuing.
 The response time dependent variable was measured from the presentation of the squares and dots, until the keypress indicating the participant's choice. 
+The dependent variable measuring whether the participant chose the borderline case was also recorded at this time. 
 The trial would timeout after 60 seconds if there was no response.
-The dependent variable measuring whether the participant chose the borderline case was also recorded at this time.
 In this experiment, no feedback was given. 
 This was because, in the vague conditions, we did not regard any response as `correct' or `incorrect', but instead as `borderline response', or `not borderline response', and we did not want to draw participants' attention to this distinction explicitly.
-We simply recorded whether the participant chose the borderline case or not, and how long it took the participant to respond.
+We simply recorded whether the participant chose the best referent, the borderline case or the poorest referent, and how long it took the participant to respond.
 
 \subsection{Results}
+Means for response times and proportion of borderline responses are given in Fig. (\ref{resultse2}). 
+Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data. 
+A linear mixed model was constructed for the (logged) response times.
+The \emph{instruction format} and \emph{vagueness} variables were sum-coded and \emph{Item} was centred.
+The fixed effects in the model were \emph{instruction format} and \emph{vagueness} and their interaction, and \emph{item}.
+The random effects in the model were \emph{participant}, and slopes over \emph{participant} for \emph{instruction format}, \emph{vagueness}, and their interaction
 
-This time around, no responses were treated as erroneous, because errors were essentially undefined for the vague instructions (e.g., `about 10').
-It was noted whether participants chose the borderline square.
-%
-The RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data. Means for response times are given in Fig. (\ref{resultse2}).
-
-A linear mixed model was constructed for the response times.
-%
-Response times were logged; \emph{selection task / instruction format} and \emph{vagueness} were sum-coded and \emph{Item} was centred. 
-% 
-The fixed effects in the model were \emph{selection task / instruction format} and \emph{vagueness} and their interaction, and \emph{item}.
-% 
-The random effects in the model were \emph{participant}, and slopes over \emph{participant} for \emph{selection task / instruction format} and \emph{vagueness} and their interaction, and for \emph{item}.
-
-There was a significant effect of \emph{selection task /instruction format} with numerical conditions attracting longer responses than the verbal conditions (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; $\beta=.37, se=.07, t=5.1, p<.001$).
-% 
-The effect of \emph{vagueness} was to slow responses down (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06, se=.01, t=4.6, p<.001$). 
-%
-There was a main effect of \emph{item} ($\beta=.06, se=.008, t=7.0, p<.001$) indicating that response times differed in some way according to which item was presented.
-However further analysis revealed that there was no consistent smooth trend across items - this effect seems likely to be due to the very fast responses for the smallest item.
+(1) The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06, se=.01, t=4.6, p<.001$).
+(2) This disadvantage for vagueness persisted at each level of instruction format, offering evidence against hypothesis 2.
+[GIVE VALUES OF A FOCUSSED T TEST] There was an interaction effect between vagueness and instruction format, such that the disadvantage for vagueness was greater in the numerical than in the verbal conditions ($\beta=-0.13, se= 0.02, t= -6.6, p=0.000$).
+(3) There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; $\beta=.37, se=.07, t=5.1, p<.001$).
+(4) There was a significant main effect of \emph{item} ($\beta=.06, se=.008, t=7.0, p<.001$) indicating that response times differed in some way according to which item was presented.
+However further analysis revealed that there was no consistent smooth trend across items.
+This effect seems likely to be due to the very fast responses for the smallest item.
 This item had the largest ratio difference so may have been particularly discriminable for participants.
-% Reviews asked for an explanation of the difference for numeric crisp over items.
+% MG reviews asked for an explanation of the difference for numeric crisp over items.
 Also reviews mention ratio differences as something that we did not control.
-%
-There was an interaction effect between vagueness and task, such that the disadvantage for vagueness was greater in the numerical than in the verbal instruction conditions ($\beta=-0.13, se= 0.02, t= -6.6, p=0.000$).
-[IS THIS CONSISTENT WITH THE ALL-NUMBERS EXPERIMENT 5?]
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.65\textwidth]{images/response-time-sep-1.pdf}
-\caption{Response time results for Experiment 2}
-\label{resultse2}
-\end{figure}
 
@@
 A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with task, vagueness and item as fixed effects, and with random slopes for task and vagueness and item over participants.
-%
-The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{borderline-response-distribution-e2}.
-%
+The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{resultse2}.
 Participants were significantly more likely to choose the borderline option for vague instructions than for precise instructions (21.9\% vs 11.3\%, $\beta=.79, se=.25, z=3.2, p<.01$). 
@@
 \centering
-\includegraphics[]{images/histbd4-ggplot-version-1.pdf}
-\caption{Distribution of Nearest match, borderline case, and furthest match responses}
-\label{borderline-response-distribution-e2}
+\includegraphics[width=.49\textwidth]{images/response-time-sep-1.pdf}
+\includegraphics[width=.49\textwidth]{images/histbd4-ggplot-version-1.pdf}
+\label{resultse2}
+\caption{Experiment 2 results: response times and proportion of borderline cases}
 \end{figure}
@@
 
-This experiment tested to see whether when borderline cases are present, vague instructions would speed responses as they did in Experiment Two when there were no borderline squares.
+Experiment 2 tested to see whether when borderline cases are present, vague instructions would speed responses as they did in Experiment 1 when there were no borderline squares. 
 We actually found a \emph{disadvantage} of vague instructions: vague instructions slowed people down by 112 ms on average. 
@@
 The disadvantage of numerical format overwhelms the contribution of vagueness. 
-The verbal vague condition was still responded to faster than the numerical crisp condition, so the pattern from Experiment 1 is reproduced, but in the light of the evidence from Experiment 2 in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like either an advantage of verbal instruction format, or alternatively as an advantage of the comparison task, according to the selection task account.
+The \emph{verbal vague} condition was still responded to faster than the \emph{numerical crisp} condition, so the pattern from Experiment 1 is reproduced, but in the light of the evidence from Experiment 2 in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like either an advantage of verbal instruction format.
 
 % KvD: I've tentatively added the following:
-Having effectively separated the {\bf cost reduction} hypothesis from the {\bf instruction format} hypothesis, it is important to observe that, in Experiment 2, instruction format went hand in hand with {\bf selection algorithm}: as shown by the Table of instructions \ref{instructionse2}, the instructions that used a verbal instruction format allowed a comparison algorithm, whereas the instructions that used a numeric format allowed on a matching algorithm.
-Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison algorithm (which is known to be faster than matching).
-% KvD I assume we need to say more about this, citing some appropriate papers to support this idea.
+Having effectively separated the {\bf cost reduction} hypothesis from the {\bf instruction format} hypothesis, it is important to observe that, in Experiment 2, instruction format went hand in hand with {\bf selection algorithm}:
+ as shown by the Table of instructions \ref{instructionse2}, the instructions that used a verbal instruction format allowed a comparison algorithm, whereas the instructions that used a numeric format allowed a matching algorithm.
+Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison algorithm
+(which is known to be faster than matching). % KvD I assume we need to say more about this, citing some appropriate papers to support this idea.
 
-In the next two experiments we pitted the comparison and matching selection tasks against each other while controlling vagueness and instruction format.
-In Experiment 3 we would restrict all the instructions to numeric quantifiers while factorially manipulating vagueness and the comparison / matching selection tasks.
-In Experiment 4 we would ensure that all instructions used verbal quantifiers, while also factorially manipulating vagueness and the comparison / matching selection tasks.
+In the next two experiments we pitted the comparison algorithm and matching algorithm selection tasks against each other while controlling vagueness and instruction format. 
+In Experiment 3 we would restrict all the instructions to numeric quantifiers while factorially manipulating vagueness and selection task. 
+In Experiment 4 we would ensure that all instructions used verbal quantifiers, while also factorially manipulating vagueness and selection task. 
 This allows us to distinguish between the predictions of the selection task account and the instruction format account.
 
+\clearpage
 \section{Experiment 3}
@@
 The main changes from experiment 2 were that the selection task was explicitly controlled, and that all conditions were constrained to mention a number. 
-We used the same stimuli as in experiment 2.
-Table \ref{Instructions for e3} shows the instructions.
-% KvD I think we should motivate our choice of materials (especially "far ..er")
+We used the same stimuli as in experiment 2. Table \ref{Instructions for e3} shows the instructions. % KvD I think we should motivate our choice of materials (especially "far ..er")
 
-\begin{table}[htbp]
-\centering
-\caption{Experiment 3: Instructions, assuming $6,15,24$ dots as the Item, and showing \emph{fewer} instead of \emph{more}}
-\label{Instructions for e3}
-\begin{tabular}{llll}
-\hline
-vagueness&instruction format&selection task&instruction\\
-\hline
 crisp & numeric&matching & Choose a square with 6 dots \\ 
@@
 vague & numeric&comparison & Choose a square with far fewer than 20 dots \\
-\hline
-\end{tabular}
-\end{table}%
 
 \subsection{Hypotheses}
-% (1) Vague instructions are easier for the reader than crisp alternatives (main effect of vagueness)
-%(2) Comparison is easier for the reader than matching (main effect of selection)
-%(3) Effects of vagueness are different depending on whether selection is matching or comparison (interaction effect selection x vagueness).
+(1) Vague instructions are easier for the reader than crisp alternatives (main effect of vagueness)
 
-The instruction format account predicts no differences between the conditions, since all conditions used numeric quantifiers.
+(2) Comparison is easier for the reader than matching (main effect of selection)
 
-The selection task account predicts a main effect of selection task such that the comparison conditions would attract faster responses than the matching conditions.
+(3) Effects of vagueness are different depending on whether selection is matching or comparison (interaction effect selection x vagueness).
 
-The cost reduction account predicts that there will be a main effect of vagueness such that the vague instruction conditions attract faster responses than the crisp instruction conditions and particularly that at each level of selection task the vague condition should attract faster responses than the crisp condition.
+%The instruction format account predicts no differences between the conditions, since all conditions used numeric quantifiers.
+%
+%The selection task account predicts a main effect of selection task such that the comparison conditions would attract faster responses than the matching conditions.
+%
+%The cost reduction account predicts that there will be a main effect of vagueness such that the vague instruction conditions attract faster responses than the crisp instruction conditions and particularly that at each level of selection task the vague condition should attract faster responses than the crisp condition.
 
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.5\textwidth]{images/response-time-two-panels-1.pdf}
-\caption{Results for Experiment 3}
-\label{resultse3}
-\end{figure}
-
+\clearpage
 \section{Experiment 4}
@@
 The referring expressions were constrained to never mention a numeral, as in Table \ref{Instructions for e4}.
-
-\begin{table*}[htbp]
-\centering
-\caption{Experiment 4: Instructions}
-\label{Instructions for e4}
-\begin{tabular}{lllp{7cm}}
-\hline
-vagueness&instruction format& selection task&instruction\\
-\hline
-crisp & verbal&matching & Choose a square with the same number of dots as the target \\ 
-crisp & verbal&comparison& Choose a square with fewer dots than the target \\
-vague & verbal&matching & Choose a square with about the same number of dots as the target \\ 
-vague & verbal&compaison& Choose a square with far fewer dots than the target \\ 
-\hline
-\end{tabular}
-\end{table*}%
 
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.5\textwidth]{images/response-time-two-panels-1-e4.pdf}
-\caption{Results for Experiment 4}
-\label{resultse4}
-\end{figure}
-
 Again none of the accounts set out in the Hypotheses section emerge well from the results.
@@
 
+\clearpage
 \subsection{Discussion of experiments 3 and 4}
@@
 
+\clearpage
 \section{General Discussion}
@@
 \end{sidewaystable}
+
+\clearpage
+
+\begin{table}[htbp]
+\caption{Table of instructions for the pair (5,25). Experiment 1}
+\label{instructionse1}
+\begin{tabular}{rl}
+\toprule
+vagueness&example\\
+\midrule
+crisp 	& 	Choose the square with 5 dots \\
+vague	&	Choose the square with few dots\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{figure}[tbp]
+\fitfigure{images/stimuluse1}
+\caption{An example stimulus from Experiment 1.
+First the referring expression was presented (left panel).
+Then after a keypress, and a fixation cross (not pictured) the squares and dots were presented without repetition of the referring expression (right panel)}
+\label{stimuluse1}
+\end{figure}
+
+\begin{figure}[htbp]
+\fitfigure{images/resultse1.pdf}
+\caption{Experiment 1 results. }
+\label{resultse1}
+\end{figure}
+
+\begin{table}[htbp]
+\caption{Table of instructions arranged by condition for the triple (6,15,24). Experiment 2}
+\label{instructionse2}
+\begin{tabular}{rlll}
+\toprule
+vagueness&instruction format&selection task&example\\
+\midrule
+vague	&	numeric 	& matching	&	Choose a square with about 10 dots\\
+vague	&	verbal	& comparison	&	Choose a square with few dots\\
+crisp 	& 	numeric	& matching	&	Choose the square with 6 dots \\
+crisp		&	verbal	& comparison	&	Choose the square with the fewest dots\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{figure}[tbp]
+\centering
+\includegraphics[width=.4\textwidth]{images/stimuluse2}
+\caption{An example stimulus from Experiment 2}
+\label{stimuluse2}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.65\textwidth]{images/response-time-sep-1.pdf}
+\caption{Response time results for Experiment 2}
+\label{resultse2}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[]{images/histbd4-ggplot-version-1.pdf}
+\caption{Expt 2: Distribution of Nearest match, borderline case, and furthest match responses}
+\label{borderline-response-distribution-e2}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.5\textwidth]{images/response-time-two-panels-1.pdf}
+\caption{Results for Experiment 3}
+\label{resultse3}
+\end{figure}
+
+\begin{table}[htbp]
+\centering
+\caption{Experiment 3: Instructions, assuming $6,15,24$ dots as the Item, and showing \emph{fewer} instead of \emph{more}}
+\label{Instructions for e3}
+\begin{tabular}{llll}
+\hline
+vagueness&instruction format&selection task&instruction\\
+\hline
+crisp & numeric&matching & Choose a square with 6 dots \\ 
+crisp & numeric&comparison & Choose a square with fewer than 20 dots \\
+vague & numeric&matching & Choose a square with about 10 dots \\ 
+vague & numeric&comparison & Choose a square with far fewer than 20 dots \\ 
+\hline
+\end{tabular}
+\end{table}%
+
+\begin{table*}[htbp]
+\centering
+\caption{Experiment 4: Instructions}
+\label{Instructions for e4}
+\begin{tabular}{lllp{7cm}}
+\hline
+vagueness&instruction format& selection task&instruction\\
+\hline
+crisp & verbal&matching & Choose a square with the same number of dots as the target \\ 
+crisp & verbal&comparison& Choose a square with fewer dots than the target \\
+vague & verbal&matching & Choose a square with about the same number of dots as the target \\ 
+vague & verbal&compaison& Choose a square with far fewer dots than the target \\ 
+\hline
+\end{tabular}
+\end{table*}%
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.5\textwidth]{images/response-time-two-panels-1-e4.pdf}
+\caption{Results for Experiment 4}
+\label{resultse4}
+\end{figure}
+
 \end{document}

### 2016-09-17 papers/2016-Special-Issue/v-special-issue-2016.tex
line 313, got up to e3 and e4, make hypotheses match results with numbers,

@@
 \clearpage
+
 \section{Experiment 2}
@@
 One difference is that `few' is vague (or at least has the potential for vagueness) and `5' is crisp. 
-Another difference is that `few' is a linguistic, or verbal quantifier while `5' is a numerical quantifier, in the sense that a number is mentioned explicitly. 
+Another difference is that `few' is a verbal quantifier while `5' is a numerical quantifier, in the sense that a number is mentioned explicitly. 
 Since these two differences could not be separated in Experiment 1, the vagueness advantage finding is vulnerable to an alternative interpretation, that what we saw as a vagueness advantage was in contrast an advantage for the verbal form of the quantifier.
-In Experiment 2 we created verbal and numeric versions of each of the vague and crisp instructions so that we could compare vague and crisp conditions while taking account of verbal/numeric format.%In the present experiment 2 we pitted these alternative interpretations against each other in a factorial design.
+In Experiment 2 we created verbal and numeric versions of each of the vague and crisp instructions so that we could compare vague and crisp conditions while taking account of verbal / numeric format.
+%In the present experiment 2 we pitted these alternative interpretations against each other in a factorial design.
 
@@
 Using error feedback in experiment 1 could have exacerbated this.
-Experiment 2 used three squares so that the vague quantifiers always had more than one possible referent, and used indefinite articles in the vague instructions to avoid the impression that only one response counted as correct, and was carried out without error feedback.
+Experiment 2 used three arrays so that the vague quantifiers always had more than one possible referent, and used indefinite articles in the vague instructions to avoid the impression that only one response counted as correct, and was carried out without error feedback.
 An indication that the potential for vagueness was realised in Experiment 2 is that the borderline response was chosen fairly often --16\% of the time.
 
-In Experiment 2, an item was a referring expression followed by a triple of numbers, representing the number of squares in the left, middle, and right squares. 
+In Experiment 2, an item was a referring expression instruction followed by a set of dot arrays defined by triple of numbers, representing the number of dots in the left, middle, and right arrays. 
 We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54). 
-Each triple had the following properties: it comprised three squares (instead of two as in Experiment 1); the central number was always presented in the middle of the three; there were two flanking numbers where one was smaller than the central number and one was bigger.
-Examples of crisp and vague versions of the numerical and verbal instructions follow: the examples assume the triple (6,15,24) and reference to the smaller numbers of dots, such that 6 was classified as the expected response; 15 was classified as the borderline response; and 24 was classified as the incorrect response.
+Each set of arrays had the following properties: it comprised three arrays (instead of two as in Experiment 1); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had fewer.
+Examples of crisp and vague versions of the numerical and verbal instructions follow: the examples assume the array (6,15,24) and reference to the smaller number of dots, such that 6 was classified as the expected response; 15 was classified as the borderline response; and 24 was classified as the incorrect response.
 In the {\em vague numerical} condition we used ``Choose a square with about 10 dots''.
@@
 
+(5) Participants should make more borderline case choices for vague than crisp instructions.
+
 \subsection{Method}
-
 Participants were aged between 18 and 45 with a median age of 28. 
@@
 
-Participant grand mean percentage of borderline selections was $16.6\%$.
+(4) Participant grand mean percentage of borderline selections was $16.6\%$.
 A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with task, vagueness and item as fixed effects, and with random slopes for task and vagueness and item over participants.
@@
 We actually found a \emph{disadvantage} of vague instructions: vague instructions slowed people down by 112 ms on average. 
-We also found that the effect of instruction format was significant, with numerical format slowing responses by 689 ms on average. 
-The disadvantage of numerical format overwhelms the contribution of vagueness. 
-The \emph{verbal vague} condition was still responded to faster than the \emph{numerical crisp} condition, so the pattern from Experiment 1 is reproduced, but in the light of the evidence from Experiment 2 in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like either an advantage of verbal instruction format.
+We also found that the effect of instruction format was significant, with numerical format slowing responses by 689 ms on average, such that the disadvantage of numerical format overwhelmed the contribution of vagueness.
+The \emph{verbal vague} condition was still responded to faster than the \emph{numerical crisp} condition, so the pattern from Experiment 1 was reproduced, but in the light of the evidence from Experiment 2, in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like an advantage of verbal instruction format.
 
@@
 In Experiment 4 we would ensure that all instructions used verbal quantifiers, while also factorially manipulating vagueness and selection task. 
-This allows us to distinguish between the predictions of the selection task account and the instruction format account.
+This allowed us to distinguish between the predictions of the selection task account and the instruction format account.
 
 \clearpage
+
 \section{Experiment 3}
@@
 \subsection{Introduction}
-
 The main aim of experiment 3 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies.
 The main changes from experiment 2 were that the selection task was explicitly controlled, and that all conditions were constrained to mention a number. 
-We used the same stimuli as in experiment 2. Table \ref{Instructions for e3} shows the instructions. % KvD I think we should motivate our choice of materials (especially "far ..er")
+We used the same arrays as in experiment 2. 
+Table \ref{Instructions for e3} shows the instructions. 
+% KvD I think we should motivate our choice of materials (especially "far ..er")
 
-crisp & numeric&matching & Choose a square with 6 dots \\ 
-crisp & numeric&comparison & Choose a square with fewer than 20 dots \\
-vague & numeric&matching & Choose a square with about 10 dots \\ 
-vague & numeric&comparison & Choose a square with far fewer than 20 dots \\
+crisp  numeric matching  Choose a square with 6 dots
+
+crisp  numeric comparison  Choose a square with fewer than 20 dots
+
+vague  numeric matching  Choose a square with about 10 dots
+
+vague  numeric comparison  Choose a square with far fewer than 20 dots
 
@@
 
-The results showed that vagueness was beneficial for comparison but detrimental for matching.
-There was no significant main effect of vagueness ($\beta =.003, se=.014, t=.202, p=.841$).
+The results showed that vagueness was beneficial for comparison but detrimental for matching. There was no significant main effect of vagueness ($\beta =.003, se=.014, t=.202, p=.841$). 
 There was a main effect of task type, with the comparison task speeding responses compared to the matching task ($\beta=-.165, se=.027, t=-6.218, p<.001$). 
@@
 
+crisp  verbal matching  Choose a square with the same number of dots as the target
+
+crisp  verbal comparison Choose a square with fewer dots than the target
+
+vague  verbal matching  Choose a square with about the same number of dots as the target
+
+vague  verbal comparison Choose a square with far fewer dots than the target
+
 \subsection{Hypotheses}
-The instruction format account predicts no differences between the conditions, since all conditions used verbal quantifiers.
+%The instruction format account predicts no differences between the conditions, since all conditions used verbal quantifiers.
+%
+%The selection task account predicts a main effect of selection task such that the comparison conditions would attract faster responses than the matching conditions.
+%
+%The cost reduction account predicts that there will be a main effect of vagueness such that the vague instruction conditions attract faster responses than the crisp instruction conditions and particularly that at each level of selection task the vague condition should attract faster responses than the crisp condition.
 
-The selection task account predicts a main effect of selection task such that the comparison conditions would attract faster responses than the matching conditions.
+(1) Vague instructions are easier for the reader than crisp alternatives (main effect of vagueness)
 
-The cost reduction account predicts that there will be a main effect of vagueness such that the vague instruction conditions attract faster responses than the crisp instruction conditions and particularly that at each level of selection task the vague condition should attract faster responses than the crisp condition.
+(2) Comparison is easier for the reader than matching (main effect of selection)
+
+(3) Effects of vagueness are different depending on whether selection is matching or comparison (interaction effect selection x vagueness).
 
@@
 vague & verbal&matching & Choose a square with about the same number of dots as the target \\ 
-vague & verbal&compaison& Choose a square with far fewer dots than the target \\ 
+vague & verbal&comparison& Choose a square with far fewer dots than the target \\ 
 \hline

### 2016-09-18 papers/2016-Special-Issue/v-special-issue-2016.tex
At this point I am sending back to Kees. I have tried to resinstate the naive version. I have left out a couple of tables that seemed better in the main body.

@@
-\documentclass[doc,apacite]{apa6}
+\documentclass[doc,floatsintext,apacite]{apa6}
 \usepackage{graphicx}
@@
 \usepackage{datetime}
-\usepackage{rotating}
-\title{Is Vagueness Beneficial for Hearers?\\Evidence from Experiments}
+\usepackage{rotating} % comment out for man mode
+\title{Is Vagueness Beneficial for Hearers? \protect \\ Evidence from Experiments}
 \date{\today}
@@
 \maketitle
+
 \section{Introduction}\label{introduction}
@@
 emphasise the evaluative content inherent in these categories, and the affective potential of the evaluative content rather than the vagueness of the terms like {\em Fair}.
-\clearpage
-%\\[2ex]
-\subsection{General Methodology}
+
+\subsubsection{General Methodology}
 The experiments reported in the present paper put the cost reduction hypothesis to the test. 
@@
 In this context, speed and accuracy of responses are the key dimensions on which the different references can be compared. 
+Each stimulus in the experiments was a set of dot arrays containing various number of dots, together with a preceding instruction (in the form of a referring expression) to choose one of the arrays with respect to its cardinality.
+The participant was asked to respond as quickly as possible while avoiding errors. 
+We manipulated the instructions and the arrays in several ways across a series of four experiments.
 
-Each stimulus in the experiments was a set of dot arrays containing various number of dots, together with a preceding instruction (in the form of a referring expression) to choose one of the arrays with respect to its number of dots, as quickly as possible while avoiding errors.
-We manipulated the stimuli in several ways across a series of four experiments.
+% MG added/moved out of expt 1
+There is evidence that when the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the cardinality of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene199647}.
+We manipulated the number of dots in each array such that some sets of arrays had smaller numerical distances and others had larger numerical distances. 
+Where a number was mentioned in the instructions, it was always in the form of an Arabic numeral.
 
-% added/moved out of expt 1 by matt 17 sept
-There is evidence that when the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the numerosity of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene199647}.
-Where a number was mentioned in the instruction, it was always in the form of an Arabic numeral.
-
-% added/moved out of expt 1 by matt 17 sept
+% MG added/moved out of expt 1
 There is evidence that when two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}.
@@
 
-% General methodology moved to here
+% MG General methodology moved to here: per-experiment method sections only talk about deviations from this general method
 All the experiments shared the following properties: 
+Stimuli were created using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.
+The position of the dots in each array was randomised per-trial.
+The order in which trials were presented was randomised per-participant. 
+There were 256 trials, presented in 4 blocks of 64 trials each, between which the participant could rest. 
+A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants and recorded responses. 
+Participants were recruited using email lists at the University of Aberdeen, and paid ten pounds for participating.
+All participants self-reported fluency in English, and had normal, or corrected-to-normal vision.
 The experiment was conducted in a quiet room. 
-All participants self-reported fluency in English, and had normal, or corrected-to-normal vision.
-Participants were asked to respond quickly while avoiding errors.
-Stimuli were created and presented using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.
+Participants were asked to respond as quickly as possible while avoiding errors.
 There was a block of practice trials after which participants could ask any questions, following which the experimenter left the room.
-The position of the dots was randomised per-trial.
-The order in which trials were presented was randomised per-participant.
-A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants.
-
-\clearpage
 
@@
 \subsection{Introduction}
-We used a forced choice task to compare choices made in response to vague instructions against choices made in response to crisp instructions. 
-The participant was presented with an instruction like ``Choose the square with many dots'' in the vague conditions, or ``Choose the square with 20 dots'' in the crisp conditions. 
-Then two dot arrays were presented in the form of squares containing a number of dots. %Fig \ref{stimuluse1} shows an example stimulus. 
-The participant was required to identify the array that corresponded with the instruction, by pressing the appropriate key.
+We used a forced choice task to elicit responses to crisp or vague forms of instructions that required the participant to choose one of two dot arrays by referring to its cardinality.  
+The participant was presented with an instruction in the form \emph{Choose the square with \ldots dots}. 
+Then a set of two two dot arrays was presented, each in the form of a square containing some number of dots. 
+ %Fig \ref{stimuluse1} shows an example stimulus. 
+The participant was required to identify the array that corresponded with the instruction, by pressing the appropriate key, as quickly as possible while avoiding errors. 
 Response time and accuracy were recorded for analysis.
 
-Our main manipulation was of the vagueness of the instruction, with two levels, vague and crisp. %Table \ref{instructionse1} shows examples from each condition.  
-We also manipulated how discriminable the dot arrays were. 
+We manipulated how discriminable the dot arrays were by varying the numerical distance between them. 
 One array always contained 25 dots: the other contained either 5, 10, 15, 20, 30, 35, 40, or 45 dots. 
-This led to numerical differences of 5, 10, 15, and 20, with lower differences resulting in less discriminable arrays and larger differences representing more discriminable arrays.
+This gave us numerical distances of 5, 10, 15, and 20, with smaller numerical distances resulting in less discriminable arrays and larger distances resulting in more discriminable arrays.
+
+Our main manipulation was of the vagueness of the instruction, with two levels, \emph{crisp} and \emph{vague}.
+Assuming the dot array [5, 25], and the instruction referring to the smaller cardinality, the \emph{crisp} instruction was \emph{Choose the square with 5 dots} and its \emph{vague} counterpart was \emph{Choose the square with few dots}.
+%Table \ref{instructionse1} shows examples from each condition.
 
 \subsection{Hypotheses}
-(1) A main effect of vagueness: vague instructions impose a lower cognitive load for the comprehender than crisp alternatives.
+(1) A main effect advantage for vagueness: vague instructions impose a lower cognitive load for the comprehender than crisp alternatives.
 
-(2) A main effect advantage of increasing numerical distance: the task will become easier as the numerical distance increases, because the two arrays are more discriminable.
+(2) A main effect advantage of increasing numerical distance: the task will become easier as the numerical distance increases, because the two arrays are then more discriminable.
 
-(3) An interaction between vagueness and numerical distance: i.e., any facilitation for vagueness should be greater at smaller numerical distances.
+(3) An interaction between vagueness and numerical distance: i.e., any facilitation for vagueness should be greater at smaller numerical distances than at larger numerical distances.
 
 \subsection{Method}
-Twenty participants were recruited by mailing list; aged between 18 and 45 with a median age of 26; and paid ten pounds for participating. 
-On arrival in the room, each participant was told that on each trial he or she would be presented with an instruction to choose one of two squares on screen by reference to how many dots the square contained.
-Participants were required to press the space key after reading the instruction. 
-Then there was a central fixation cross for 1000 ms, and a blank screen for 500 ms, followed by the array (without repetition of the referring expression). 
+Twenty participants were recruited, aged between 18 and 45 with a median age of 26.
+On each trial a participant was presented with an instruction to choose one of two dot arrays on screen by reference to its cardinality. 
+Following a keypress to indicate that the participant had read the instruction there was a central fixation cross for 1000 ms, and a blank screen for 500 ms, followed by the array (without repetition of the referring expression).
 The arrays would stay on screen until the participant responded (there was no timing-out). 
 Response time was measured as the latency between the presentation of the arrays, and the keypress identifying the choice: in this way, the response time was separated from time spent reading the instructions, which is important since we are only interested in the former.
-There were 256 trials, presented in 4 blocks of 64 trials each, between which the participant could rest. 
+A response was counted as erroneous if the square with the wrong number of dots was chosen (when the instruction contained a number); 
+if the square with the larger number of dots was selected (when the instruction was \emph{Choose the square with few dots}); 
+or if the square with the smaller number of dots was selected (when the instruction was \emph{Choose the square with many dots}).
+% KvD Is this addition defining "erroneous" OK?
+MG Yes, I moved it though
 No feedback was given on correct trials, but there was feedback on error trials in the form of the word ``\textsc{wrong!!}'' which flashed on screen.
@@
 \subsection{Results}
-A response was counted as erroneous if the square with the wrong number of dots was chosen (when the instruction contained a number); if the square with the larger number of dots was selected (when the instruction was ``Choose the square with few dots"); or if the square with the smaller number of dots was selected (when the instruction was ``Choose the square with many dots".) % KvD Is this addition defining "erroneous" OK?
-RTs for trials with erroneous responses were discarded, leading to the loss of 354 trials from 5120, representing 6.9\% of the trials. 
+\subsubsection{Response times}
+Response times (RTs) for trials with erroneous responses were discarded, leading to the loss of 354 trials from 5120, representing 6.9\% of the trials. 
 The correct response RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of a further 160 trials, or 3.4\% of the remaining correct responses. 
-Means for response times and error rates are given in Fig. (\ref{resultse2}).
-A linear mixed model of RT was built using as independent variables vagueness and numerical distance and their interaction, with random slopes for vagueness and numerical distance over participants. 
-Vagueness was sum coded: vague $= -.5$, crisp $= .5$; numerical distance was Helmert coded. 
-Helmert contrasts compare each level against the mean of the previous levels. 
-Level one of this contrast is numerical distance 5 compared with numerical distance 10; level two is the mean of numerical distance 5 and numerical distance 10 versus numerical distance 15; and level three is the mean of numerical distances 5, 10 and 15 versus numerical distance 20.
+Means for response times and error rates are given in Fig. (\ref{resultse1}).
+A linear mixed model of RT was built, using as independent variables \emph{vagueness} and \emph{numerical distance} and their interaction, with random slopes for \emph{vagueness} and \emph{numerical distance} over participants.
+\emph{Vagueness} was sum coded: \emph{vague} $= -.5$, \emph{crisp} $= .5$; \emph{numerical distance} was Helmert coded, i.e., the variable is ordered and each level is compared against the mean of all the previous levels to that point.
+% MG a reviewer said he was not familiar with Helmert contrasts
+%Level one of this contrast is numerical distance 5 compared with numerical distance 10; level two is the mean of numerical distance 5 and numerical distance 10 versus numerical distance 15; and level three is the mean of numerical distances 5, 10 and 15 versus numerical distance 20.
 All $p$ values were calculated using the R package \emph{lmerTest} \cite{lmerTest}.
+Numbering below corresponds to the numbering of the hypotheses.
 
-(1) RTs were faster for vague instructions than for crisp instructions ($\beta=.109, se=.022, t=4.9, p<.001$). 
-(2) RTs grew faster as numerical distance increased: level one $(\beta=-.116, se=.012, t=-9.3, p<.05)$, level 2 ($\beta=-.103, se=.009, t=-11.1, p<.001$) and level three ($\beta=-.082, se=.007, t=-11.0, p<.05$).
+(1) RTs were faster for vague instructions than for crisp instructions ($\beta=.109$, $se=.022$, $t=4.9$, $p<.001$).
+
+(2) RTs grew faster as numerical distance increased: level one ($\beta=-.116$, $se=.012$, $t=-9.3$, $p<.05$), level 2 ($\beta=-.103$, $se=.009$, $t=-11.1$, $p<.001$) and level three ($\beta=-.082$, $se=.007$, $t=-11.0$, $p<.05$).
 Since discriminability of the arrays is easier for larger numerical distances, discriminability probably underlies this effect.
+
 (3) Numerical distance and vagueness interacted significantly for larger numerical distances when modelling RT: essentially there were diminishing returns for vagueness as numerical distance increased.% KvD Is it worth saying how they interacted?
 % MG I think this is answered now?
-The interactions at the different levels of numerical distance were: level one: ($\beta=.001, se=.016, t=.03, p=.974$); level two ($\beta=-.039, se=.009, t=-4.460, p<.001$); level three ($\beta=-.043, se=.006,t=-7.0, p<.001$).
+The interactions at the different levels of numerical distance were: level one: ($\beta=.001$, $se=.016$, $t=.03$, $p=.974$); level two ($\beta=-.039$, $se=.009$, $t=-4.460$, $p<.001$); level three ($\beta=-.043$, $se=.006$, $t=-7.0$, $p<.001$).
 In the crisp conditions RTs started out much slower than in the vague conditions, at the smallest numerical distance, but the two conditions converged to very fast times at the largest numerical distance.
 
+\subsubsection{Error rates}
 Error rate data were analysed using a generalized logit mixed model \cite{jaeger2008categorical}, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
-The effect of vagueness on error rates approached significance, with the vague instructions leading to fewer errors ($\beta=.307, se=.173, t=1.8, p=.077$).
-Error rates decreased as numerical distance increased: level one ($\beta=-.585, se=.092, ,z=-6.4, p<.001$), level 2 ($\beta=-.434, z=-5.3,p<.001$) and level three ($\beta=-.250,z=-4.1, p<.001$). 
-Error rates were greater in the crisp conditions than the vague conditions when numerical distance was small, and this difference diminished with increasing numerical distance until it reversed at the biggest numerical distance.
+
+(1) The effect of vagueness on error rates approached significance, with the vague instructions leading to fewer errors ($\beta=.307, se=.173, t=1.8, p=.077$).
+
+(2) Error rates decreased as numerical distance increased: level one ($\beta=-.585, se=.092, ,z=-6.4, p<.001$), level 2 ($\beta=-.434, z=-5.3,p<.001$) and level three ($\beta=-.250,z=-4.1, p<.001$).
+
+(3) Error rates were greater in the crisp conditions than the vague conditions when numerical distance was small, and this difference diminished with increasing numerical distance until it reversed at the biggest numerical distance.
+% MG NEED TO INSERT STATS FOR THIS INTERACTION
 
@@
 \subsection{Discussion}
-Responses were faster and more accurate when the instructions were in the vague conditions than when they were in the crisp conditions. 
-The speed and accuracy diminished as the arrays became more easily discriminable, and the advantage for vague instructions tailed off as the arrays became more easily discriminable.
+Responses were faster and more accurate for vague than crisp instructions. 
+Response times and error rates improved as the arrays became more easily discriminable. 
+The advantage for vague instructions tailed off as the arrays became more easily discriminable.
 
@@
 It explains the diminishing returns for vagueness in more-discriminable stimuli (i.e., the vagueness by numerical distance interaction) by claiming that load is low in both conditions for the easily-discriminable stimuli, and that therefore there is no extra benefit to be had from vagueness in the easily-discriminable stimuli.
-
-\clearpage
 
@@
 Experiment 2 used three arrays so that the vague quantifiers always had more than one possible referent, and used indefinite articles in the vague instructions to avoid the impression that only one response counted as correct, and was carried out without error feedback.
-An indication that the potential for vagueness was realised in Experiment 2 is that the borderline response was chosen fairly often --16\% of the time.
+An indication that the potential for vagueness was realised in Experiment 2 is that the borderline response was chosen fairly often: 16\% of the time.
 
@@
 We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54). 
-Each set of arrays had the following properties: it comprised three arrays (instead of two as in Experiment 1); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had fewer.
+Each set of arrays had the following properties: it comprised three arrays (instead of two as in Experiment 1); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had more.
+
 Examples of crisp and vague versions of the numerical and verbal instructions follow: the examples assume the array (6,15,24) and reference to the smaller number of dots, such that 6 was classified as the expected response; 15 was classified as the borderline response; and 24 was classified as the incorrect response.
-In the {\em vague numerical} condition we used ``Choose a square with about 10 dots''.
+In the {\em vague numerical} condition we used \emph{Choose a square with about 10 dots}.
 None of the squares displayed contained 10 dots.
 10 is slightly closer to 6 than to 15, justifying 6 as the best response and 15 as the borderline response.
-In the {\em vague verbal} condition we used ``Choose a square with few dots''.
-In the {\em crisp numerical} condition we used ``Choose the square with 6 dots''. One square did contain the exact number mentioned.
-In the {\em crisp verbal} condition, we used ``Choose the square with the fewest dots".
+In the {\em vague verbal} condition we used \emph{Choose a square with few dots}.
+In the {\em crisp numerical} condition we used \emph{Choose the square with 6 dots} and one square always did contain the number mentioned.
+In the {\em crisp verbal} condition, we used \emph{Choose the square with the fewest dots}.
 
 \subsection{Hypotheses}
-(1) A main effect advantage of vagueness.
+(1) A main effect RT advantage for vagueness.
 
-(2) An advantage of vagueness both in \emph{numeric} and in \emph{verbal} instructions.
+(2) An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
 
 (3) No large main effect of instruction format.
-%MG If we are right that vagueness was responsible for the vagueness advantage in e 1, that could also have been a number avoidance advantage, then we would not expect that number avoidance, or, instruction format, plays a role.
+%MG If we are right that vagueness was responsible for the vagueness advantage in e 1, that could also have been a number avoidance advantage, then we would not expect that number avoidance, or, instruction format, plays a role (?)
 
-(4) On the basis of Experiment 1, we would expect an effect of array, with faster responses to more discriminable arrays
+(4) On the basis of Experiment 1, we would expect an effect of triple, with faster responses to more discriminable triples.
 
@@
 Participants were aged between 18 and 45 with a median age of 28. 
-We presented participants with 256 trials, arranged in 4 blocks each with 64 trials. 
 We manipulated as independent variables vagueness and instruction format, yielding four conditions, \emph{vague numeric}; \emph{vague verbal}; \emph{crisp numeric}; \emph{crisp verbal}.
@@
 On each trial, first the referring expression that constituted the instruction for that trial was displayed. 
-The participant then pressed a key to indicate that he or she had read the instruction. 
-After 1 second, the squares and dots were presented, while preserving the text of the referring expression. 
-The response time dependent variable was measured from the presentation of the squares and dots, until the keypress indicating the participant's choice. 
-The dependent variable measuring whether the participant chose the borderline case was also recorded at this time. 
+Participants then pressed a key to indicate that they had read the instruction. 
+After 1000 ms, the arrays were presented, while preserving the text of the referring expression. 
+The response time dependent variable was measured from the presentation of the arrays, until the keypress indicating the participant's choice, which was also recorded. 
 The trial would timeout after 60 seconds if there was no response.
@@
 \subsection{Results}
+\subsubsection{Response times}
 Means for response times and proportion of borderline responses are given in Fig. (\ref{resultse2}). 
@@
 (1) The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06, se=.01, t=4.6, p<.001$).
-(2) This disadvantage for vagueness persisted at each level of instruction format, offering evidence against hypothesis 2.
-[GIVE VALUES OF A FOCUSSED T TEST] There was an interaction effect between vagueness and instruction format, such that the disadvantage for vagueness was greater in the numerical than in the verbal conditions ($\beta=-0.13, se= 0.02, t= -6.6, p=0.000$).
+
+(2) Vagueness was disadvantageous in both the \emph{numeric} and \emph{verbal} conditions, offering evidence against hypothesis 2. %MG [? give values of a focussed t test]
+The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to an interaction effect between vagueness and instruction format ($\beta=-0.13, se= 0.02, t= -6.6, p=0.000$).
+
 (3) There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; $\beta=.37, se=.07, t=5.1, p<.001$).
-(4) There was a significant main effect of \emph{item} ($\beta=.06, se=.008, t=7.0, p<.001$) indicating that response times differed in some way according to which item was presented.
-However further analysis revealed that there was no consistent smooth trend across items.
-This effect seems likely to be due to the very fast responses for the smallest item.
-This item had the largest ratio difference so may have been particularly discriminable for participants.
-% MG reviews asked for an explanation of the difference for numeric crisp over items.
-Also reviews mention ratio differences as something that we did not control.
 
-(4) Participant grand mean percentage of borderline selections was $16.6\%$.
+(4) There was a significant main effect of \emph{triple} ($\beta=.06, se=.008, t=7.0, p<.001$) indicating that response times differed in some way according to which triple was presented. 
+However further analysis revealed that there was no consistent smooth trend across different triples. 
+This effect seems likely to be due to the very fast responses for the smallest triple, which had the largest ratio difference and so may have been particularly discriminable for participants. 
+% MG reviews asked for an explanation of the difference for numeric crisp over items. Also reviews mention ratio differences as something that we did not control.
+
+\subsubsection{Borderline responses}
+(5) Participant grand mean percentage of borderline selections was $16.6\%$.
 A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with task, vagueness and item as fixed effects, and with random slopes for task and vagueness and item over participants.
@@
 \includegraphics[width=.49\textwidth]{images/histbd4-ggplot-version-1.pdf}
+\caption{Experiment 2 results: response times and proportion of borderline cases}
 \label{resultse2}
-\caption{Experiment 2 results: response times and proportion of borderline cases}
 \end{figure}
@@
 Having effectively separated the {\bf cost reduction} hypothesis from the {\bf instruction format} hypothesis, it is important to observe that, in Experiment 2, instruction format went hand in hand with {\bf selection algorithm}:
- as shown by the Table of instructions \ref{instructionse2}, the instructions that used a verbal instruction format allowed a comparison algorithm, whereas the instructions that used a numeric format allowed a matching algorithm.
+ %as shown by the Table of instructions \ref{instructionse2}, 
+the instructions that used a verbal instruction format allowed a comparison algorithm, whereas the instructions that used a numeric format allowed a matching algorithm. 
 Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison algorithm
@@
 In the next two experiments we pitted the comparison algorithm and matching algorithm selection tasks against each other while controlling vagueness and instruction format. 
-In Experiment 3 we would restrict all the instructions to numeric quantifiers while factorially manipulating vagueness and selection task. 
-In Experiment 4 we would ensure that all instructions used verbal quantifiers, while also factorially manipulating vagueness and selection task. 
+In Experiment 3 we restricted all the instructions to \emph{numeric} quantifiers while factorially manipulating vagueness and selection task. 
+In Experiment 4 we ensured that all instructions used \emph{verbal} quantifiers, while also factorially manipulating vagueness and selection task. 
 This allowed us to distinguish between the predictions of the selection task account and the instruction format account.
-
-\clearpage
 
@@
 We used the same arrays as in experiment 2. 
-Table \ref{Instructions for e3} shows the instructions. 
+Table \ref{Instructions for e3} shows the instructions for each condition.
+
 % KvD I think we should motivate our choice of materials (especially "far ..er")
 
-crisp  numeric matching  Choose a square with 6 dots
-
-crisp  numeric comparison  Choose a square with fewer than 20 dots
-
-vague  numeric matching  Choose a square with about 10 dots
-
-vague  numeric comparison  Choose a square with far fewer than 20 dots
+\begin{table}[htbp]
+\centering
+\caption{Experiment 3: Instructions, assuming $6,15,24$ dots as the Item, and showing \emph{fewer} instead of \emph{more}}
+\label{Instructions for e3}
+\begin{tabular}{llll}
+\hline
+vagueness&instruction format&selection task&instruction\\
+\hline
+crisp & numeric&matching & Choose a square with 6 dots \\ 
+crisp & numeric&comparison & Choose a square with fewer than 20 dots \\
+vague & numeric&matching & Choose a square with about 10 dots \\ 
+vague & numeric&comparison & Choose a square with far fewer than 20 dots \\ 
+\hline
+\end{tabular}
+\end{table}%
 
@@
 
-(2) Comparison is easier for the reader than matching (main effect of selection)
+(2) Comparison is easier for the reader than matching (main effect of selection task)
 
@@
 %
-%The selection task account predicts a main effect of selection task such that the comparison conditions would attract faster responses than the matching conditions.
-%
 %The cost reduction account predicts that there will be a main effect of vagueness such that the vague instruction conditions attract faster responses than the crisp instruction conditions and particularly that at each level of selection task the vague condition should attract faster responses than the crisp condition.
@@
 
-38 volunteers were recruited via internal messaging at University of Aberdeen, with self-reported fluency in English. They were paid ten pounds each for participating.
-The apparatus was the same as Experiment 1.
+38 participants were recruited. 
 The design was a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e3}).
-Each stimulus was an instruction followed by a triple of dots.
-First a referring expression instruction was presented. Participants pressed a key to dismiss the instruction and proceed to the squares with dots in them.
+On each trial a referring expression instruction was presented: participants pressed a key to dismiss the instruction, when the dot arrays were presented until the participant responded, and the response time and choice were recorded.
 
@@
 
-The results showed that vagueness was beneficial for comparison but detrimental for matching. There was no significant main effect of vagueness ($\beta =.003, se=.014, t=.202, p=.841$). 
-There was a main effect of task type, with the comparison task speeding responses compared to the matching task ($\beta=-.165, se=.027, t=-6.218, p<.001$). 
-Vagueness exerted effects in different directions for the comparison task and for the matching task. 
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.75\textwidth]{images/response-time-two-panels-1.pdf}
+\caption{Results for Experiment 3}
+\label{resultse3}
+\end{figure}
+
+(1) There was no significant main effect of vagueness ($\beta =.003, se=.014, t=.202, p=.841$). 
+The results showed that vagueness was beneficial for comparison but detrimental for matching. 
+(2) There was a main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=-.165, se=.027, t=-6.218, p<.001$). 
+(3) Vagueness exerted effects in different directions for the comparison task and for the matching task. 
 Separate analyses were conducted at each level of the selection task to see whether within each task type there were significant effects of vagueness. 
@@
 
-\clearpage
 \section{Experiment 4}
@@
 We manipulated vagueness and the selection task (comparison and matching). 
-In order to implement the experiment without mentioning numbers,  we used a prime to visually show the numbers of dots that we wanted to refer to as {\em the target} in the instructions. 
-This presentation of a prime before the main trial shares some features with Experiment 2 in \protect \citeA{Izard20081221}, although in that experiment participants were told the numerosity of the prime - called an \emph{inducer} in that paper - in our experiment we did not tell participants the numerosity of the prime array.
-An item was thus a combination of a visual prime, a numeric triple, and a referring expression.
-The referring expressions were constrained to never mention a numeral, as in Table \ref{Instructions for e4}.
+In order to implement the experiment without mentioning numbers in the instructions, we changed the sequence of each trial to include a 'target' (i.e., a dot array of a particular cardinality) before the instruction, so that we could then refer back to the target's cardinality in the instruction using expressions like \emph{the same number of dots as the target}; \emph{fewer dots than the target}.
+%?MG WHAT WERE THE VALUES OF THE TARGETS?
+This presentation of a target before the main body of the trial shares some features with \citeA[Experiment 2]{Izard20081221}, although in that experiment participants were told the cardinality of the target (called an \emph{inducer} in that paper) whereas in our experiment we did not tell participants the cardinality of the prime array.
+An item was thus a combination of a target dot array, an instruction that did not explicitly contain a number, and a set of dot arrays taking their cardinalities from the same triples used in experiment 2.
+Table \ref{Instructions for e4} spells out how the instructions were constrained not to mention a numeral.
 
-crisp  verbal matching  Choose a square with the same number of dots as the target
-
-crisp  verbal comparison Choose a square with fewer dots than the target
-
-vague  verbal matching  Choose a square with about the same number of dots as the target
-
-vague  verbal comparison Choose a square with far fewer dots than the target
+\begin{table}[htbp]
+\centering
+\caption{Experiment 4: Instructions}
+\label{Instructions for e4}
+\begin{tabular}{lllp{7cm}}
+\hline
+vagueness&instruction format& selection task&instruction\\
+\hline
+crisp 	& verbal	&	matching 		& Choose a square with the same number of dots as the target \\ 
+crisp 	& verbal	&	comparison	& Choose a square with fewer dots than the target \\
+vague 	& verbal	&	matching 		& Choose a square with about the same number of dots as the target \\ 
+vague 	& verbal	&	comparison	& Choose a square with far fewer dots than the target \\ 
+\hline
+\end{tabular}
+\end{table}%
 
@@
 
-\subsection{Method}
+%\subsection{Method}
 
-40 volunteers recruited via internal messaging at University of Aberdeen, with self-reported fluency in English. They were paid ten pounds for participating. 
-The apparatus was the same as Experiment 1.
-The design was a 2 x 2 factorial manipulation of vagueness and selection task.
-Each stimulus was a sequence of prime, instruction, and squares.
-The procedure for this experiment was different from the others, to accommodate the requirement not to use numbers in the instruction.
-We had to have a different way to indicate a numerosity in the instruction, which we did by adding a visual `prime', a square that contained the number of dots that we wanted to refer to.
+%40 participants were recruited. % via internal messaging at University of Aberdeen, with self-reported fluency in English. They were paid ten pounds for participating. 
+%The apparatus was the same as Experiment 1.
+%The design was a 2 x 2 factorial manipulation of vagueness and selection task.
+%Each stimulus was a sequence of prime, instruction, and squares.
+%The procedure for this experiment was different from the others, to accommodate the requirement not to use numbers in the instruction. 
+%We had to have a different way to indicate a cardinality in the instruction, which we did by adding a visual `target', a square that contained the number of dots that we wanted to refer to.
 
 \subsection{Results}
+40 volunteers participated.
 The results showed that vagueness was beneficial for comparison but detrimental for matching (the same as Experiment 3) even when no numbers were allowed in the instructions. 
 Figure \ref{resultse4} shows the means by condition.
-There was no main effect of vagueness ($\beta=.01, se=.01, t=1.51, p=.14$).
-There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=-.18, se=.02, t=-10.38, p<.01$).
-This effect was in the same direction as Experiment Four.
-Vagueness did  exert different effects depending on the selection task ($\beta=.12, se=.03, t=4.32, p<.05$).
-Separate analyses were conducted for the comparison task and for the matching task.
-In the comparison task, vagueness resulted in faster response times ($\beta=-0.08, se=.02, t=4.30, p<.05$).
-In the matching task vagueness slowed response times ($\beta=.05, se=.01, t=3.72, p<.05$).
-These results are in the same direction as Experiment 3.
 
-Again none of the accounts set out in the Hypotheses section emerge well from the results.
+(1) There was no main effect of vagueness ($\beta=.01, se=.01, t=1.51, p=.14$).
+
+(2) There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=-.18, se=.02, t=-10.38, p<.01$).
+This effect was in the same direction as Experiment 3.
+
+(3) Vagueness did exert different effects depending on the selection task ($\beta=.12, se=.03, t=4.32, p<.05$). 
+Separate analyses were conducted for the comparison task and for the matching task. In the comparison task, vagueness resulted in faster response times ($\beta=-0.08, se=.02, t=4.30, p<.05$). 
+In the matching task vagueness slowed response times ($\beta=.05, se=.01, t=3.72, p<.05$). These results are in the same direction as Experiment 3.
+
+The cost reduction account is wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
 The instruction format account wrongly predicts no differences between the conditions. 
-The selection task correctly predicted the main effect of selection task, but has no coverage of the interaction with vagueness.
-The cost reduction account is wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
+The selection task account correctly predicted the main effect of selection task, but has no coverage of the interaction of selection task with vagueness.
 
-\clearpage
-\subsection{Discussion of experiments 3 and 4}
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.75\textwidth]{images/response-time-two-panels-1-e4.pdf}
+\caption{Results for Experiment 4}
+\label{resultse4}
+\end{figure}
+
+\section{Discussion of experiments 3 and 4}
 
@@
 The main effect of selection task showed that the assumption that the \emph{comparison} task is easier than the \emph{matching} task is well-founded. 
-In both experiments people were reliably faster at responding in the \emph{comparison} task.
+In both experiments people were reliably quicker to respond in the \emph{comparison} task.
 
 Vagueness, which was the phenomenon on which our investigation focussed, did not exert a main effect in response time. 
-However when the comparison and selection tasks were analysed separately, there was small reliable speedup in RT from the crisp to the vague \emph{comparison} tasks, but a small reliable slowdown in RT from the crisp to the vague \emph{matching} tasks.
+However when the comparison and selection tasks were analysed separately, there was small reliable advantage for vagueness in the \emph{comparison} tasks, but a small reliable disadvantage for vagueness in the \emph{matching} tasks.
 
-\clearpage
 \section{General Discussion}
@@
 \bibliography{v-special-issue-2016}
+\end{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \clearpage
@@
 \end{figure}
+crisp  numeric matching  Choose a square with 6 dots
 
-\end{document}
+crisp  numeric comparison  Choose a square with fewer than 20 dots
+
+vague  numeric matching  Choose a square with about 10 dots
+
+vague  numeric comparison  Choose a square with far fewer than 20 dots
+
+crisp  verbal matching  Choose a square with the same number of dots as the target
+
+crisp  verbal comparison Choose a square with fewer dots than the target
+
+vague  verbal matching  Choose a square with about the same number of dots as the target
+
+vague  verbal comparison Choose a square with far fewer dots than the target

### 2016-09-21 papers/2016-Special-Issue/v-special-issue-2016.tex
Kees's email said: The changes you’ve made look to me to be on the right track. Just in case you’ve got time to look at this again before tomorrow, I’ve revised the closing paragraphs of the paper a bit (where applications to NLG are discussed).

@@
 the entry ``vague'' in \cite{penguin}.} In most academic use though, the word `vagueness' has a specific meaning.
-Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1996vagueness}, also \citeA{EgreKlinedinst}).
+Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1997vagueness}, also \citeA{EgreKlinedinst}).
 The crucial criterion is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects.
@@
 These results are interesting, and of obvious potential practical importance.
-We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \cite{keefe1996vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases.
+We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \cite{keefe1997vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases.
 In the terminology of \citeA{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}.
@@
 {\bf Implications for practical NLG systems.} Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
-The question of how a particular piece of quantitative information is best conveyed through language is certainly an important one for practical NLG.
-Years of research on the logic of vagueness in natural language have primed the research community to expect that the utility of vagueness is an important part of the answer, but our findings call this into question.
-%It would appear that this is true for most vague expressions, and not only the ones investigated in our experiments, where a speaker is referring to a square with dots.
+The question of how a particular piece of quantitative information is best conveyed through language is an important one for practical NLG.
+Years of research on the logic of vagueness in natural language have primed the research community to expect that the utility of vagueness is an important part of the answer, but our findings call this expectation into question.
+Although our own studies have focussed on vagueness in descriptive Noun Phrases, it seems plausible that vagueness plays a similar role in other linguistic constructs.
 
-Consider, once again, a number of ways in which a given temperature can be reported. Given a temperature measurement (or prediction) as input, an NLG system might justifiably say that it was
+For example, consider air temperature, once again. Given a temperature measurement (or prediction) as input, an NLG system that addresses a non-expert audience might say that it was (or will be)
 %
@@
 If standard accounts are to be believed, then options (a) and (c) convey crisp information, whereas (b) and (d) are vague (i.e., they permit borderline cases).
-But, to the best of our knowledge, there is no experimental evidence that the fact in itself that an expression is vague confers a benefit (or a de-benefit) on it for hearers.
-Rather than asking whether a candidate expression is vague, other questions might shed more light on the NLG system's choice, similar (though not necessarily identical) to the ones identified in our empirical studies.
-For a start, both (c) and (d) designate an interval that is half-bounded, whereas both (a) and (b) express closed intervals, and this might affect their comprehension by a hearer (cf., our finding about comparison versus matching strategies).
-Furthermore, the expressions (c) and (d) -- one of which is vague while the other is crisp -- appear to convey very similar amounts of information, saying that the temperature is higher than some (fairly high) standard.
-Perhaps most importantly, "warm" is shorter than the other three expressions and avoids the use of numbers, and our experiments suggest that this may make ``warm" more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of the word ``warm" (see section 1 and \citeA{peters2009bringing}).
+From our own experiments or the ones we found in the literature, there is no evidence that the fact in itself that an expression is vague should confer a benefit on it for hearers.
+Rather than asking whether a candidate expression is vague, other questions might shed more light on the NLG system's choice, similar (though not necessarily identical) to the ones identified in our studies.
+These questions might focus on the amount of information that a given expression conveys (i.e., on granularity), on the avoidance of numbers, and on the use of evaluative terms.
+Let's see how this might pan out for the examples at hand.
 
-One way to see why vagueness may not matter as much for NLG -- and for human communication more generally -- as is often thought is the following thought experiment.
-Many speakers understand the word ``warm" as vague.
-Now suppose a group of users agreed to give it a precise definition; according to this precisification, ``warm" means $>25$ degrees (as in (c)).
-It seems unlikely that this change, from a vague meaning to a crisp one, would change the utility of the word.
+The expressions (c) and (d) -- one of which is vague while the other is crisp -- convey similar amounts of information, saying that the temperature is higher than some standard (i.e., designating a half-open temperature interval, as opposed to the closed intervals designated by (a) and (b)).
+Yet (d) (``warm") is shorter than the other three expressions and avoids the use of numbers, and our experiments suggest that this may make ``warm" more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm" (see section 1).
+Results of this kind can be applied in NLG, though obvious care would need to be exercised to take the purpose of the communication into account.\footnote{In the experiments reported in this paper, the purpose of the utterances investigated was always clear (e.g., instructing people on a weight loss program \cite{Mishra01042011}; informing a patient's choice of hospital \cite{peters2009bringing}; or selecting a dot array (in our own studies).}
+
+One way to see why vagueness may not matter as much for NLG -- and for human communication -- as is often thought, is the following thought experiment.
+Suppose a group of speakers understand the word ``warm" as vague, agreeing that temperatures above 26 count as warm, and temperatures below 24 do not count as warm, but considering temperatures between 24 and 26 as borderline cases.
+Now suppose these speakers agreed to ``precisify" their definition deciding that, henceforth, ``warm" means ``$>25$ degrees" (as in (c) above): this decision resolves the borderline cases, while everything else remains the same.
+It seems unlikely that this change in language use, from a vague meaning to a crisp one (i.e., one that has no borderline cases), would lower the utility of the word.
+Our experimental findings are consistent with this idea.
+
+%\bibliography{bibliography}
 

### 2016-09-21 papers/2016-Special-Issue/v-special-issue-2016.tex
small change to general discussion, moved table too

@@
 Experiment 2 showed us that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional explanatory power in either the verbal format instructions or the numerical format instructions when we generated verbal and numerical versions of both crisp and vague instructions.
-However the experiment could not distinguish benefits of number-avoidance in the verbal instructions from benefits of the comparison selection task: the verbal instructions also mandated a comparison strategy rather than a matching strategy.
-In experiments 3 and 4 we manipulated vagueness and the selection task independently of numerical format.
-We found that there are effects of the selection task mandated by the instruction, with the comparison task instructions attracting faster response times than the matching instructions, and that vagueness exerts benefits when the selection task is \emph{comparison}, but not when the task is \emph{matching}.
+% MJG edit
+However the experiment could not distinguish benefits of number avoidance in the verbal instructions from benefits of 
+the comparison selection task in the verbal instructions, since the verbal instructions mandated a comparison strategy, and the numeric instructions mandated a matching strategy.
+In experiments 3 and 4 we manipulated vagueness and the selection task separately at each level of numerical format.
+Across the two experiments, we found that there are effects of the selection task mandated by the instruction, with the comparison task instructions attracting faster response times than the matching instructions, and within the experiments we found that vagueness exerts benefits when the selection task is \emph{comparison}, but not when the task is \emph{matching}.
 
@@
 This leads us to speculate that the benefit for vagueness here could be due to the vague expression foregrounding a particular valid target while the crisp expression carries with it the additional task of distinguishing between two alternative valid targets, something we propose to call a ``range-reduction'' benefit.
+
+\begin{table}[htbp]
+\caption{Vagueness as range reduction: a summary of Experiments 3 and 4}
+\label{Vagueness as range reduction}
+\centering
+\begin{tabular}{cccc}
+selection task 				& vagueness		& candidates	& effect of vageueness	\\
+\toprule
+\multirow{ 2}{*}{comparison} 	& crisp 			& 2			& \multirow{ 2}{*}{vagueness advantage}  		\\
+  						& vague			& 1			&                                               				\\
+\midrule
+\multirow{ 2}{*}{matching} 		& crisp 			& 1 			&\multirow{ 2}{*}{vagueness disadvantage}	\\
+ 						& vague			& 2			&								 	\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 Two of the resulting four conditions result in an expression that can denote either of two referents; the other two conditions result in an expression that can only denote one referent, with the other possible referent being a marginal candidate at best:
-
-\begin{table}[htbp]
-\caption{Vagueness as range reduction}
-\begin{center}
-\begin{tabular}{lll}
-\hline
-vagueness & selection task & candidates\\
-\hline
-crisp & matching & 1 candidate\\
-crisp & comparison & 2 candidates\\
-vague & matching & 2 candidates\\
-vague & comparison & 1 candidate\\
-\hline
-\end{tabular}
-\end{center}
-\label{default}
-\end{table}
 

### 2016-09-24 experiments/e1/writeup/e1.Rmd
e1 is now looking ok as a template for the other experiments.

--- 2016-07-27 experiments/e1/writeup/e1.Rmd
@@
     toc: yes
+    <!--toc_float: true-->
     mathjax: default
     template: default
-    code_folding: hide
+    <!--code_folding: show-->
     theme: default
 ---  
-
-<style type="text/css">
-.table {width: 50%;}
+<style>
+table, th, td {
+    border: 1px grey;
+    border-collapse: collapse;
+    padding:2px 5px 2px 5px
+}  
 </style>
 
-vagueness |instruction
-----------|-------------------------------------------------------------------
-crisp     |Choose the the square with many dots
-vague     |Choose the square with 20 dots
-
-items| a
------|-----
-1    |5,25
-2    |10,25
-3    |15,25
-4    |20,25
------|-----
-5    |30,25
-6    |35,25
-7    |40,25
-8    |45,25
-
-In this experiment it was possible to identify errors, so error trials were removed. This is because there were only 2 squares and an istruction uniquely identified one of them.
-
-# Prep data
-
-```{r prepData}
-rm(list=ls())
-suppressPackageStartupMessages(library(knitr))
-#opts_chunk$set(cache=TRUE, autodep=TRUE)
-#opts_chunk$set(size="footnotesize")
-#options(width=60)
+```{r pander, echo=FALSE}
+library(pander)
 ```
 
-```{r loadLibs}
-suppressPackageStartupMessages(library(ggplot2))
-suppressPackageStartupMessages(library(xtable))
-suppressPackageStartupMessages(library(plyr))
-suppressPackageStartupMessages(library(reshape2))
-suppressPackageStartupMessages(library(lattice))
-suppressPackageStartupMessages(library(car))
-suppressPackageStartupMessages(library(gridExtra))
-suppressPackageStartupMessages(library(lme4))
-suppressPackageStartupMessages(library(LMERConvenienceFunctions))
-suppressPackageStartupMessages(library(lmerTest))
+```{r showInstructions, echo = FALSE}
+v = c("crisp","vague")
+e <- c("Choose the the square with many dots", "Choose the square with 20 dots")
+i = data.frame(v, e)
+names(i) = c("vagueness", "example instruction")
+pander(i, justify=c('center', 'left'))
 ```
 
-```{r sourceFunctions}
-source("functions/CIfunctions.R")
+```{r showItems, echo=FALSE}
+items = c(1,2,3,4,5,6,7,8)
+i = data.frame(Gap = c("05","10","15","20"),
+               x = c("25:20","25:15","25:10","25:05"),
+               z = c("25:30","25:35","25:40","25:45") )
+names(i) = c('Gap Size', 'Small Target', 'Large Target')
+pander(i, justify=c('center', 'center', 'center'))
 ```
 
-## get data
-
-```{r manipulateDataFrame, echo=FALSE}
-system("cat ../experimentCode/output/*.dat > ../data/data.txt", wait=TRUE)
-dat <- read.delim("../data/data.txt", header=FALSE)
-names(dat) <- c("condition", "leftN", "rightN", "vgn", "side", "string", "subject", "rt", "error")
-dat$subject <- as.factor(paste("sub",sep="",sprintf("%02d",dat$subject,sep="")))
-dat$error <- 1-dat$error
-dat$Ord <- ifelse(dat$leftN < dat$rightN, 'LtoR', 'RtoL')
-dat$c_Ord <- ifelse(dat$Ord=='LtoR', -.5, .5)
-dat$bigside <- ifelse(dat$leftN > dat$rightN, 'L', 'R')
-dat$Qty <- ifelse(dat$side==dat$bigside, "Large", "Small")
-dat$bigside=NULL
-dat$c_Qty <- ifelse(dat$Qty=='Small', -.5, .5)
-dat$gap <- as.character(paste("gap", sep=" ", sprintf("%02d",abs(dat$leftN-dat$rightN))))
-dat$vgn <- as.factor(ifelse(dat$vgn=="P", "Crisp", "Vague"))
-dat$c_Vgn <- ifelse(dat$vgn=='Crisp', -.5, .5)
-dat$logrt <- log(dat$rt)
-# item_ratio
-# NOT DONE YET
-write.table(x=dat, "all_trials.txt")
-dat$subject <- as.factor(dat$subject)
-dat$vgn <- ifelse(dat$vgn=='Crisp', -.5, .5)
-dat$gap[dat$gap=="gap 05"] <- -1.5
-dat$gap[dat$gap=="gap 10"] <- -.5
-dat$gap[dat$gap=="gap 15"] <- .5
-dat$gap[dat$gap=="gap 20"] <- 1.5
-dat$gap <- as.numeric(dat$gap)
-# Separate off correct trials
-alltrials <- dat
-correct <- subset(dat, error==0)
-errors <-  subset(dat, error==1)
+```{r loadLibraries, message=FALSE, echo = FALSE}
+library(knitr); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); source('summarySEwithin2.R')
 ```
 
-# RT model
+In this experiment it was possible to identify errors.
+This is because there were only 2 squares and an instruction uniquely identified one of them.
+Therefore error trials were removed from analysis of response times:
 
-```{r tableRTmodel, results='asis'}
-print(
-	xtable(
-		digits=3, 
-		summary(
-			rt.lmer <- lmer(data=correct, logrt~gap*vgn + (vgn|subject))
-			)$coefficients,
-		caption="Most maximal response time model", 
-		caption.placement="top"), type='html')
+```{r separateDataFramesForRTandErrorAnalyses, echo=T}
+rawdata <- read.table('data.txt', sep=' ')
+rtdata <- subset(rawdata, error==FALSE)
+erdata <- rawdata
 ```
 
-## RT means
+This code creates the condition means with confidence intervals used for plotting:
 
-For correct rts df, show means. Vagueness appears to be beneficial.
-
-```{r plotCorrectRTMeans, fig.width=3, fig.height=3}
-a <- summarySEwithin(correct,  measurevar="logrt",  withinvars=c("vgn"),  idvar="subject")
-ggplot(data=a, aes(y=logrt_norm, x=vgn, group=1, ymin=logrt_norm-ci, ymax=ci+logrt_norm)) + geom_line() + geom_point() + geom_errorbar(width=.05) + xlab("") + ylab("log rt") + theme_bw() + theme(aspect.ratio=1)
+```{r computeSummariesForGraphs, cache=TRUE}
+rtdataplot1 = summarySEwithin2(rtdata,  measurevar="logrt",  withinvars=c("Vgn","Gap"), idvar="subject")
+erdataplot1 <- summarySEwithin2(erdata, measurevar="error", withinvars=c("Vgn","Gap"), idvar="subject")
 ```
 
-# Error rates model
-
-```{r tableErrorModel, results='asis'}
-print(xtable(digits=3, summary(ac.lmer <- 
-															 	glmer(data=alltrials, error~gap*vgn + (gap+vgn|subject), 
-															 				family="binomial", 
-															 				control = glmerControl(optimizer = "bobyqa")))$coefficients,
-						 caption="Most maximal accuracy model",
-						 caption.placement="top"),type='html')
-```
-
-```{r qq, fig.width=5, fig.height=3}
-op <- par(mfrow=c(1,2))
-qqPlot(residuals(rt.lmer), cex=.2)
-qqPlot(residuals(ac.lmer), cex=.2)
-```
-
-```{r plotMeansBoth, fig.width=10, fig.height=4, message=FALSE}
-rts <- summarySEwithin(correct, measurevar="logrt", withinvars=c("vgn","gap"))
-err <- summarySEwithin(alltrials, measurevar="error", withinvars=c("vgn","gap"))
+```{r rtplot, echo = FALSE}
 dodge=.06
-
-plot1 =
-  ggplot(rts, aes(y=logrt_norm, x=gap, group=vgn, fill=vgn, shape=vgn, ymin=logrt_norm-ci, ymax=logrt_norm+ci)) + 
+rtplot = ggplot(rtdataplot1, aes(y=logrtNormed, x=Gap, group=Vgn, fill=Vgn, shape=Vgn, ymin=logrtNormed-ci, ymax=logrtNormed+ci)) + 
   geom_line(position=position_dodge(dodge)) + 
@@
   xlab("Gap Size") +
-  scale_fill_grey(name=element_blank(), start=0,end=1)+
-  theme_bw()+
+  scale_fill_grey(name=element_blank(), start=0, end=1) +
+  theme_bw() +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank() ) +
-  theme(aspect.ratio=1)+xlab("")
+  theme(aspect.ratio=1) + xlab("")
+```
 
-plot2 = 
-  ggplot(err, aes(y=error_norm, x=gap, group=vgn, fill=vgn, shape=vgn, ymin=error_norm-ci, ymax=error_norm+ci)) + 
+```{r erplot, echo=FALSE}
+dodge=.06
+erplot = ggplot(erdataplot1, aes(y=errorNormed, x=Gap, group=Vgn, fill=Vgn, shape=Vgn, ymin=errorNormed-ci, ymax=errorNormed+ci)) + 
   geom_line(position=position_dodge(dodge)) + 
@@
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank() )+
-  theme(aspect.ratio=1)+xlab("")
+  theme(aspect.ratio=1) + xlab("")
+```
 
-grid.arrange(plot1,plot2,ncol=2)
+```{r plotRTandErrorRates, fig.width=10, fig.height=4, message=FALSE, echo=FALSE}
+grid.arrange(rtplot, erplot, ncol=2)
 ```
+
+This code computes the models:
+
+```{r showVersions, echo=FALSE, comment=NA}
+cat (paste(sep="", 
+           c("lmerTest version "), c(packageDescription("lmerTest", fields = "Version")),
+           "; ", 
+           c("lme4 version "), c(packageDescription("lme4", fields = "Version"))) ) 
+```
+
+```{r computeLmerModels, cache = TRUE, tidy=FALSE}
+rt.lmer.1 <- lmerTest::lmer(
+  data=rtdata, logrt ~ c_Vgn * c_Gap + (c_Vgn * c_Gap|subject))
+rt.lmer.2 <- lmerTest::lmer(
+  data=rtdata, logrt ~ c_Vgn * c_Gap + c_Ord + c_Qty + (c_Vgn * c_Gap + c_Ord + c_Qty | subject))
+ac.lmer.1 <- lme4::glmer(
+  data = erdata, error ~ c_Vgn * c_Gap + (c_Vgn * c_Gap | subject), 
+  family="binomial", control = glmerControl(optimizer = "bobyqa"))
+ac.lmer.2 <- lme4::glmer(
+  data = erdata, error ~ c_Vgn * c_Gap + c_Ord + c_Qty + (c_Vgn * c_Gap + c_Ord + c_Qty | subject), 
+  family="binomial", control = glmerControl(optimizer = "bobyqa"))
+```
+
+# Response times
+
+```{r, results='asis', echo=F}
+print(
+  xtable(rtdataplot1,
+         caption='Condition means'),
+  include.rownames=FALSE,
+  caption.placement='top',
+  type='html')
+```
+
+```{r tableRTmodelSimple, results='asis', echo=FALSE, cache=TRUE}
+x = as.data.frame(format(summary(rt.lmer.1)$coef, scientific=FALSE), stringsAsFactors = FALSE)
+
+for (i in 1:(ncol(x)-1)) {
+  for (j in 1:nrow(x)) {
+    x[j,i] <-  substr(x[j,i], start=1, stop=7)
+  }
+}
+
+for (i in 1:(ncol(x)-1)) {
+  x[,i] <- as.numeric(x[,i])
+}
+
+for (i in ncol(x)) {
+  for (j in 1:nrow(x)) {
+    x[j,i] <-  substr(x[j,i], start=1, stop=7)
+  }
+}
+print(
+  xtable(
+    align='lrrrrr', 
+    digits=c(0,  2,2,2,2,3),
+    x,
+    caption="Simple response time model"), 
+    caption.placement="top", 
+  type='html')
+```
+
+```{r tableRTmodelMaximal, results='asis', echo=FALSE, cache=TRUE}
+x = as.data.frame(format(summary(rt.lmer.2)$coef, scientific=FALSE), stringsAsFactors = FALSE)
+
+for (i in 1:(ncol(x)-1)) {
+  for (j in 1:nrow(x)) {
+    x[j,i] <-  substr(x[j,i], start=1, stop=7)
+  }
+}
+
+for (i in 1:(ncol(x)-1)) {
+  x[,i] <- as.numeric(x[,i])
+}
+
+for (i in ncol(x)) {
+  for (j in 1:nrow(x)) {
+    x[j,i] <-  substr(x[j,i], start=1, stop=7)
+  }
+}
+print(
+  xtable(
+    align='lrrrrr', 
+    digits=c(0,  2,2,2,2,3),
+    x,
+    caption="Maximal response time model"), 
+    caption.placement="top", 
+  type='html')
+```
+
+# Error rates
+
+```{r erdataplot1, results='asis', echo=F}
+print(
+  xtable(erdataplot1,
+         caption='Condition means'),
+  include.rownames=FALSE,
+  caption.placement='top',
+  type='html')
+```
+
+```{r tableErrorModelsimple, results='asis', cache=TRUE, echo=FALSE}
+print(
+  xtable(summary(
+    ac.lmer.1)$coefficients,
+         digits=3, 
+         caption="Simple error rates model"),
+         caption.placement="top", type='html')
+```
+
+```{r tableErrorModelMaximal, results='asis', cache=TRUE, echo=FALSE}
+print(
+  xtable(
+    digits=3, summary(
+      ac.lmer.2)$coefficients,
+             caption="Maximal error rates model"),
+             caption.placement="top", type='html')
+```

### 2016-09-24 experiments/e1/writeup/e1.Rmd
I dropped the maximal models into commented out code

@@
   data=rtdata, logrt ~ c_Vgn * c_Gap + (c_Vgn * c_Gap|subject))
-rt.lmer.2 <- lmerTest::lmer(
-  data=rtdata, logrt ~ c_Vgn * c_Gap + c_Ord + c_Qty + (c_Vgn * c_Gap + c_Ord + c_Qty | subject))
+
 ac.lmer.1 <- lme4::glmer(
   data = erdata, error ~ c_Vgn * c_Gap + (c_Vgn * c_Gap | subject), 
-  family="binomial", control = glmerControl(optimizer = "bobyqa"))
-ac.lmer.2 <- lme4::glmer(
-  data = erdata, error ~ c_Vgn * c_Gap + c_Ord + c_Qty + (c_Vgn * c_Gap + c_Ord + c_Qty | subject), 
   family="binomial", control = glmerControl(optimizer = "bobyqa"))
@@
     x,
-    caption="Simple response time model"), 
-    caption.placement="top", 
-  type='html')
-```
-
-```{r tableRTmodelMaximal, results='asis', echo=FALSE, cache=TRUE}
-x = as.data.frame(format(summary(rt.lmer.2)$coef, scientific=FALSE), stringsAsFactors = FALSE)
-
-for (i in 1:(ncol(x)-1)) {
-  for (j in 1:nrow(x)) {
-    x[j,i] <-  substr(x[j,i], start=1, stop=7)
-  }
-}
-
-for (i in 1:(ncol(x)-1)) {
-  x[,i] <- as.numeric(x[,i])
-}
-
-for (i in ncol(x)) {
-  for (j in 1:nrow(x)) {
-    x[j,i] <-  substr(x[j,i], start=1, stop=7)
-  }
-}
-print(
-  xtable(
-    align='lrrrrr', 
-    digits=c(0,  2,2,2,2,3),
-    x,
-    caption="Maximal response time model"), 
+    caption="Response time model"), 
     caption.placement="top", 
@@
          digits=3, 
-         caption="Simple error rates model"),
+         caption="Error rates model"),
          caption.placement="top", type='html')
@@
 
-```{r tableErrorModelMaximal, results='asis', cache=TRUE, echo=FALSE}
-print(
-  xtable(
-    digits=3, summary(
-      ac.lmer.2)$coefficients,
-             caption="Maximal error rates model"),
-             caption.placement="top", type='html')
-```
+<!-- ```{r} -->
+<!-- rt.lmer.2 <- lmerTest::lmer( -->
+<!--   data=rtdata, logrt ~ c_Vgn * c_Gap + c_Ord + c_Qty + (c_Vgn * c_Gap + c_Ord + c_Qty | subject)) -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- ac.lmer.2 <- lme4::glmer( -->
+<!--   data = erdata, error ~ c_Vgn * c_Gap + c_Ord + c_Qty + (c_Vgn * c_Gap + c_Ord + c_Qty | subject),  -->
+<!--   family="binomial", control = glmerControl(optimizer = "bobyqa")) -->
+<!-- ``` -->
+
+<!-- ```{r tableRTmodelMaximal, results='asis', echo=FALSE, cache=TRUE} -->
+<!-- x = as.data.frame(format(summary(rt.lmer.2)$coef, scientific=FALSE), stringsAsFactors = FALSE) -->
+
+<!-- for (i in 1:(ncol(x)-1)) { -->
+<!--   for (j in 1:nrow(x)) { -->
+<!--     x[j,i] <-  substr(x[j,i], start=1, stop=7) -->
+<!--   } -->
+<!-- } -->
+
+<!-- for (i in 1:(ncol(x)-1)) { -->
+<!--   x[,i] <- as.numeric(x[,i]) -->
+<!-- } -->
+
+<!-- for (i in ncol(x)) { -->
+<!--   for (j in 1:nrow(x)) { -->
+<!--     x[j,i] <-  substr(x[j,i], start=1, stop=7) -->
+<!--   } -->
+<!-- } -->
+<!-- print( -->
+<!--   xtable( -->
+<!--     align='lrrrrr',  -->
+<!--     digits=c(0,  2,2,2,2,3), -->
+<!--     x, -->
+<!--     caption="Maximal response time model"),  -->
+<!--     caption.placement="top",  -->
+<!--   type='html') -->
+<!-- ``` -->
+
+<!-- ```{r tableErrorModelMaximal, results='asis', cache=TRUE, echo=FALSE} -->
+<!-- print( -->
+<!--   xtable( -->
+<!--     digits=3, summary( -->
+<!--       ac.lmer.2)$coefficients, -->
+<!--              caption="Maximal error rates model"), -->
+<!--              caption.placement="top", type='html') -->
+<!-- ``` -->

### 2016-09-24 experiments/e0/writeup/e0.Rmd
practicing for html upload

@@
 ```{r getData}
-if( file.exists('data.txt') ) 
-	{
-	d <- read.table("data.txt", header=FALSE) 
-} else {
-	system("cat ../experimentCode/output/*.dat > data.txt", wait=TRUE)
+if( !file.exists('data.txt') ) {
+	system("cat ../experimentCode/output/*.dat > data.txt", wait=TRUE) # concatenate the individual subject data files into one file 
 } 
+```
+
+```{r}
+# # This bit is for generating the PURE file
+d <- read.table("data.txt", header=FALSE) # read in the data for all subjects
+mynames <- list("gap_size", 
+                "target_size", 
+                "correct_side", 
+                "vagueness",
+                "pair", 
+                "Version", 
+                "Left",
+                "Right", 
+                "Text", 
+                "Condition", 
+                "correctAnswer",
+                "subject",
+                "actualAnswer",
+                "isCorrect",
+                "rt")
+names(d) <- mynames
+# For some reason there is one rt of -6. Set it to NA
+d[d$rt==-6,"rt"] <- NA
+d$observation <- 1:nrow(d)
+d$trial <- 0
+for (s in unique(d$subject)){
+  d[d$subject==s, "trial"] <- 1:nrow(subset(d,subject==s))
+}
+d$item = ordered(d$pair, label = c("2:4", "3:5", "6:8", "7:9", "2:6", "7:3", "8:4", "5:9"))
+d$subit <- "Not subitizable"
+d$subit[d$item == "2:4"] <- "Subitizable"
+d$subit[d$item == "3:5"] <- "Subitizable"
+d$subit[d$item == "2:6"] <- "Subitizable"
+d$subit[d$item == "7:3"] <- "Subitizable"
+d$dotpair_id=d$item; d$item=NULL
+d$subit = as.factor(d$subit)
+d$gap_size = factor(d$gap_size, levels = c("BigGap", "SmallGap"), labels = c("Big", "Small"))
+d$target_size = factor(d$target_size, levels = c("BigTarget", "SmallTarget"), labels = c("Big", "Small"))
+d$vagueness = factor(d$vagueness, levels = c("Precise", "Vague"), labels = c("Crisp", "Vague")) 
+d$is_error = 1 - d$isCorrect
+d$dots_right=d$Right; d$Right=NULL
+d$dots_left=d$Left; d$Left=NULL
+dd <- subset(d, select=c(
+  observation, subject, trial, gap_size, target_size, correct_side, dots_left, dots_right, dotpair_id, vagueness, Text, is_error, rt))
+write.table(dd, file='pilot_experiment_data.txt', col.names=T, row.names=F, quote=F, sep='\t')
+```
+
+```{r}
+# This bit is for carrying forward
+d <- read.table("data.txt", header=FALSE) # read in the data for all subjects
 mynames <- list("gap", "target", "side", "vagueness", "pair", "Version", "Left",

### 2016-09-24 experiments/e0/writeup/scratch.R

@@ -1,0 +1,45 @@
+rm(d)
+d <- read.table("data.txt", header=FALSE) # read in the data for all subjects
+mynames <- list("gap_size", 
+                "target_size", 
+                "correct_side", 
+                "vagueness",
+                "pair", 
+                "Version", 
+                "Left",
+                "Right", 
+                "Text", 
+                "Condition", 
+                "correctAnswer",
+                "subject",
+                "actualAnswer",
+                "isCorrect",
+                "rt")
+names(d) <- mynames
+# For some reason there is one rt of -6. Set it to NA
+d[d$rt==-6,"rt"] <- NA
+d$observation <- 1:nrow(d)
+d$trial <- 0
+for (s in unique(d$subject)){
+  d[d$subject==s, "trial"] <- 1:nrow(subset(d,subject==s))
+}
+d$item = ordered(d$pair, label = c("2:4", "3:5", "6:8", "7:9", "2:6", "7:3", "8:4", "5:9"))
+d$subit <- "Not subitizable"
+d$subit[d$item == "2:4"] <- "Subitizable"
+d$subit[d$item == "3:5"] <- "Subitizable"
+d$subit[d$item == "2:6"] <- "Subitizable"
+d$subit[d$item == "7:3"] <- "Subitizable"
+d$dotpair_id=d$item); d$item=NULL
+d$subit = as.factor(d$subit)
+d$gap_size = factor(d$gap_size, levels = c("BigGap", "SmallGap"), labels = c("Big", "Small"))
+d$target_size = factor(d$target_size, levels = c("BigTarget", "SmallTarget"), labels = c("Big", "Small"))
+d$vagueness = factor(d$vagueness, levels = c("Precise", "Vague"), labels = c("Crisp", "Vague")) 
+d$is_error = 1 - d$isCorrect
+d$dots_right=d$Right; d$Right=NULL
+d$dots_left=d$Left; d$Left=NULL
+dd <- subset(d, select=c(
+  observation, subject, trial, gap_size, target_size, correct_side, dots_left, dots_right, dotpair_id, vagueness, Text, is_error, rt))
+
+head(dd,30)
+
+summary(dd)

### 2016-09-24 experiments/e1/e1.R

@@ -1,0 +1,165 @@
+## ----pander, echo=FALSE--------------------------------------------------
+library(pander)
+
+## ----showInstructions, echo = FALSE--------------------------------------
+v = c("crisp","vague")
+e <- c("Choose the the square with many dots", "Choose the square with 20 dots")
+i = data.frame(v, e)
+names(i) = c("vagueness", "example instruction")
+pander(i, justify=c('center', 'left'))
+
+## ----showItems, echo=FALSE-----------------------------------------------
+items = c(1,2,3,4,5,6,7,8)
+i = data.frame(Gap = c("05","10","15","20"),
+               x = c("25:20","25:15","25:10","25:05"),
+               z = c("25:30","25:35","25:40","25:45") )
+names(i) = c('Gap Size', 'Small Target', 'Large Target')
+pander(i, justify=c('center', 'center', 'center'))
+
+## ----loadLibraries, message=FALSE, echo = FALSE--------------------------
+library(knitr); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); source('summarySEwithin2.R') 
+
+## ----separateDataFramesForRTandErrorAnalyses, echo=T---------------------
+rawdata <- read.table('data.txt', sep=' ')
+rtdata <- subset(rawdata, error==FALSE)
+erdata <- rawdata
+
+## ----computeSummariesForGraphs, cache=TRUE-------------------------------
+rtdataplot1 = summarySEwithin2(rtdata,  measurevar="logrt",  withinvars=c("Vgn","Gap"), idvar="subject")
+erdataplot1 <- summarySEwithin2(erdata, measurevar="error", withinvars=c("Vgn","Gap"), idvar="subject")
+
+## ----rtplot, echo = FALSE------------------------------------------------
+dodge=.06
+rtplot = ggplot(rtdataplot1, aes(y=logrtNormed, x=Gap, group=Vgn, fill=Vgn, shape=Vgn, ymin=logrtNormed-ci, ymax=logrtNormed+ci)) + 
+  geom_line(position=position_dodge(dodge)) + 
+  geom_errorbar(width=.3, lty=1, position=position_dodge(dodge)) + 
[…]

### 2016-09-24 experiments/e1/trashme/e1.R

@@ -1,0 +1,97 @@
+## ----loadLibs------------------------------------------------------------
+suppressPackageStartupMessages(library(knitr))
+suppressPackageStartupMessages(library(ggplot2))
+suppressPackageStartupMessages(library(xtable))
+suppressPackageStartupMessages(library(gridExtra))
+suppressPackageStartupMessages(library(lme4))
+suppressPackageStartupMessages(library(lmerTest)) 
+suppressPackageStartupMessages(library(data.table)) 
+suppressPackageStartupMessages(library(dplyr))
+suppressPackageStartupMessages(library(dtplyr))  
+source('summarySEwithin2.R') 
+
+## ----separateDataFramesForRTandErrorAnalyses, echo=T---------------------
+dat <- read.table('data.txt', sep=' ')
+str(dat)
+erdata <- dat
+rtdata <- subset(dat, error==FALSE)
+
+## ----rtplot, message=FALSE-----------------------------------------------
+(rtdataplot1 = summarySEwithin2(rtdata,  measurevar="logrt",  withinvars=c("Vgn","Gap"), idvar="subject"))
+dodge=.06
+rtplot = ggplot(rtdataplot1, aes(y=logrtNormed, x=Gap, group=Vgn, fill=Vgn, shape=Vgn, ymin=logrtNormed-ci, ymax=logrtNormed+ci)) + 
+  geom_line(position=position_dodge(dodge)) + 
+  geom_errorbar(width=.3, lty=1, position=position_dodge(dodge)) + 
+  geom_point(pch=21, size=3, position=position_dodge(dodge)) +
+  ggtitle("Response time") + 
+  ylab("log RT") + 
+  xlab("Gap Size") +
+  scale_fill_grey(name=element_blank(), start=0, end=1) +
+  theme_bw() +
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank() ) +
+  theme(aspect.ratio=1) + xlab("")
+rtplot
+
+## ----tableRTmodelSimple, results='asis'----------------------------------
+print(
+  xtable(
+    digits=3, 
+    summary(
+      rt.lmer <- lmer(data=rtdata, logrt ~ c_Vgn * c_Gap + (c_Vgn * c_Gap|subject))
+    )$coefficients,
+    caption="a response time model", 
+    caption.placement="top"), type='html')
+
+## ----tableRTmodelMaximal, results='asis'---------------------------------
+print(
+  xtable(
+    digits=3, 
+    summary(
[…]

### 2016-09-25 papers/2016-Special-Issue/v-special-issue-2016.tex
pasted kees changes

@@
 \usepackage{rotating} % comment out for man mode
+
+\renewcommand{\baselinestretch}{1.7} % added 21 Sept 2016, copied from CRM book
+
 \title{Is Vagueness Beneficial for Hearers? \protect \\ Evidence from Experiments}
@@
 
-\abstract{Much of everyday language is vague, but the causes of this phenomenon are not well understood.
-Consequently, it is difficult for the designers of a Natural Language Generation (NLG) system to know when and how to let the system generate vague expressions.
+\abstract{Much of everyday language use is vague, but the causes of this phenomenon are not well understood.
+Consequently, it is difficult for the designers of a Natural Language Generation (NLG) system to know when to let the system generate vague expressions.
 The present paper is an attempt to find out what benefits vagueness might have for readers.% , by means of a series of carefully controlled psycholinguistic experiments that focus on the use of vagueness in referring expressions.
@@
 %
-The paper reports on a series of experiments that aim to separate the utility of vagueness (as defined by the existence of borderline cases) from the utility of other factors that tend to co-occur with vagueness.
-We argue that the evidence so far supports a view of vagueness where the benefits that vague terms exert are due to other influences, rather than to vagueness itself.
+This article reports on a series of experiments that aim to separate the utility of vagueness (as defined by the existence of borderline cases) from the utility of other factors that tend to co-occur with vagueness.
+We argue that the evidence so far supports a view where the benefits that vague terms exert are due to other influences, rather than to vagueness itself.
 These factors include: low granularity; the use of evaluative words; the avoidance of overtly numerical words; the existence of comparison strategies; and, lastly and more tentatively, a phenomenon that we call range reduction.
-Although it is possible that other types of vague expressions (i.e., vague words outside referential noun phrases) behave differently, our work suggests that vagueness itself may not increase the utility of an expression.
-The paper concludes with a brief discussion of the implications of our work for practical NLG.}
+The paper concludes with a brief discussion of the implications for practical NLG.}
 
@@
 They were told that an ideal HHI score lies in the range of 45 to 55.
-In a longitudinal study, participants submitted their height, weight, hydration level, gender, and age to a computer each week.
-Participants were told that two algorithms would be used to compute their HHI, and that it was possible that the two algorithms might give different values initially, but would converge over the course of the study to a single value.
-They were also told that if the two algorithms did give different values, then the true score lay between the two values.
+In a longitudinal study, participants submitted their weight to a computer each week.
+Participants were told that two algorithms would be used to compute their HHI, and that the two might give different values initially, in which case the true score lay between the two values.
 In one condition, which the authors called the precise condition, the two algorithms gave the same score.
@@
 
-\subsubsection{General Methodology}
+\subsubsection{General Methodology}\label{method}
 The experiments reported in the present paper put the cost reduction hypothesis to the test. 
@@
 
-% MG added/moved out of expt 1
-There is evidence that when the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the cardinality of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene199647}.
-We manipulated the number of dots in each array such that some sets of arrays had smaller numerical distances and others had larger numerical distances. 
-Where a number was mentioned in the instructions, it was always in the form of an Arabic numeral.
-
-% MG added/moved out of expt 1
-There is evidence that when two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}.
-We controlled which side the smaller number appeared on to avoid systematic influences of this effect.
-
 % MG General methodology moved to here: per-experiment method sections only talk about deviations from this general method
-All the experiments shared the following properties: 
+%All the experiments shared the following properties: 
 Stimuli were created using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.
-The position of the dots in each array was randomised per-trial.
+The position of the dots was randomised per-trial.
 The order in which trials were presented was randomised per-participant. 
-There were 256 trials, presented in 4 blocks of 64 trials each, between which the participant could rest. 
+There were 256 trials, presented in 4 blocks of 64 each, between which the participant could rest. 
 A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants and recorded responses. 
@@
 
+% KvD I've moved the following two paragraphs towards the end of the section.
+% MG added/moved out of expt 1
+
+When the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the cardinality of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene199647}.
+We manipulated the number of dots in each array such that some sets of arrays had smaller numerical distances and others had larger numerical distances.
+Where a number was mentioned in the instructions, it was always in the form of an Arabic numeral.
+% MG added/moved out of expt 1
+When two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}.
+We controlled which side the smaller number appeared on to avoid systematic influences of this effect.
+
+% KvD added the following from a previous version of the paper:
+There is abundant evidence (e.g., \protect\cite{trick1994small}) that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities.
+We performed a pilot experiment in which we were able to confirm this finding in the experimental settings on which we are focussing in this paper \cite{our-earlier-paper}.
+As one might expect, we found that, when participants were confronted with an array of two boxes, instructions of the form ``Choose the square with $n$ dots" led to consistently lower Response Times than instructions of the form ``Choose the square with many/few dots" when $2 \leq n \leq 5$; the converse was true for $n>5$.
+We henceforth focussed our studies on non-subitizable numbers, because it is there that vagueness is expected to have benefits.
+
 \section{Experiment One}
@@
 \subsection{Introduction}
+
 We used a forced choice task to elicit responses to crisp or vague forms of instructions that required the participant to choose one of two dot arrays by referring to its cardinality.
 The participant was presented with an instruction in the form \emph{Choose the square with \ldots dots}.
-Then a set of two two dot arrays was presented, each in the form of a square containing some number of dots. 
+Then a set of two dot arrays was presented, each in the form of a square containing some number of dots.
  %Fig \ref{stimuluse1} shows an example stimulus. 
@@
 
-\subsection{Hypotheses}
-(1) A main effect advantage for vagueness: vague instructions impose a lower cognitive load for the comprehender than crisp alternatives.
-
-(2) A main effect advantage of increasing numerical distance: the task will become easier as the numerical distance increases, because the two arrays are then more discriminable.
-
-(3) An interaction between vagueness and numerical distance: i.e., any facilitation for vagueness should be greater at smaller numerical distances than at larger numerical distances.
-
+\subsection{Hypothesis}
+(H1) A main effect advantage for vagueness: vague instructions impose a lower cognitive load for the comprehender than crisp alternatives.
 \subsection{Method}
@@
 No feedback was given on correct trials, but there was feedback on error trials in the form of the word ``\textsc{wrong!!}'' which flashed on screen.
+Hypothesis H1 was tested looking both at response times and at error rates.
 
 \subsection{Results}
+
 \subsubsection{Response times}
@@
 All $p$ values were calculated using the R package \emph{lmerTest} \cite{lmerTest}.
-Numbering below corresponds to the numbering of the hypotheses.
 
-(1) RTs were faster for vague instructions than for crisp instructions ($\beta=.109$, $se=.022$, $t=4.9$, $p<.001$).
-
-(2) RTs grew faster as numerical distance increased: level one ($\beta=-.116$, $se=.012$, $t=-9.3$, $p<.05$), level 2 ($\beta=-.103$, $se=.009$, $t=-11.1$, $p<.001$) and level three ($\beta=-.082$, $se=.007$, $t=-11.0$, $p<.05$).
-Since discriminability of the arrays is easier for larger numerical distances, discriminability probably underlies this effect.
-
-(3) Numerical distance and vagueness interacted significantly for larger numerical distances when modelling RT: essentially there were diminishing returns for vagueness as numerical distance increased.% KvD Is it worth saying how they interacted?
-% MG I think this is answered now?
-The interactions at the different levels of numerical distance were: level one: ($\beta=.001$, $se=.016$, $t=.03$, $p=.974$); level two ($\beta=-.039$, $se=.009$, $t=-4.460$, $p<.001$); level three ($\beta=-.043$, $se=.006$, $t=-7.0$, $p<.001$).
-In the crisp conditions RTs started out much slower than in the vague conditions, at the smallest numerical distance, but the two conditions converged to very fast times at the largest numerical distance.
+(H1$_{rt}$) RTs were faster for vague instructions than for crisp instructions ($\beta=.109$, $se=.022$, $t=4.9$, $p<.001$).
 
@@
 
-(1) The effect of vagueness on error rates approached significance, with the vague instructions leading to fewer errors ($\beta=.307, se=.173, t=1.8, p=.077$).
-
-(2) Error rates decreased as numerical distance increased: level one ($\beta=-.585, se=.092, ,z=-6.4, p<.001$), level 2 ($\beta=-.434, z=-5.3,p<.001$) and level three ($\beta=-.250,z=-4.1, p<.001$).
-
-(3) Error rates were greater in the crisp conditions than the vague conditions when numerical distance was small, and this difference diminished with increasing numerical distance until it reversed at the biggest numerical distance.
-% MG NEED TO INSERT STATS FOR THIS INTERACTION
+(H1$_{er}$) The effect of vagueness on error rates approached significance, with the vague instructions leading to fewer errors ($\beta=.307, se=.173, t=1.8, p=.077$).
 
@@
 \subsection{Discussion}
-Responses were faster and more accurate for vague than crisp instructions. 
-Response times and error rates improved as the arrays became more easily discriminable. 
-The advantage for vague instructions tailed off as the arrays became more easily discriminable.
 
-The cost reduction hypothesis explains the vagueness advantage by claiming that the vague referring expressions place less cognitive load on the comprehender than the crisp referring expressions. 
+As expected, responses were faster and more accurate for vague than crisp instructions. Post hoc, we formulated some further hypotheses:
+
+(H2) A main effect advantage of increasing numerical distance: the task will become easier as the numerical distance increases, because the two arrays are then more discriminable.
+
+(H3) An interaction between vagueness and numerical distance: i.e., any facilitation for vagueness should be greater at smaller numerical distances than at larger numerical distances.
+
+Looking at Response Times, these two post-hoc hypotheses were confirmed:
+
+(H2$_{rt}$) RTs increased with numerical distance level one ($\beta=-.116$, $se=.012$, $t=-9.3$, $p<.05$), level 2 ($\beta=-.103$, $se=.009$, $t=-11.1$, $p<.001$) and level three ($\beta=-.082$, $se=.007$, $t=-11.0$, $p<.05$).
+Since discriminability of the arrays is easier for larger numerical distances, discriminability probably underlies this effect.
+
+(H3$_{rt}$) Numerical distance and vagueness interacted significantly for larger numerical distances when modelling RT: essentially there were diminishing returns for vagueness as numerical distance increased.% KvD Is it worth saying how they interacted?
+% MG I think this is answered now?
+The interactions at the different levels of numerical distance were: level one: ($\beta=.001$, $se=.016$, $t=.03$, $p=.974$); level two ($\beta=-.039$, $se=.009$, $t=-4.460$, $p<.001$); level three ($\beta=-.043$, $se=.006$, $t=-7.0$, $p<.001$).
+In the crisp conditions RTs started out much slower than in the vague conditions, at the smallest numerical distance, but the two conditions converged to very fast times at the largest numerical distance.
+
+Looking at Error Rates, the two post-hoc hypotheses were confirmed as well:
+
+(H2$_{er}$) Error rates decreased as numerical distance increased: level one ($\beta=-.585, se=.092, ,z=-6.4, p<.001$), level 2 ($\beta=-.434, z=-5.3,p<.001$) and level three ($\beta=-.250,z=-4.1, p<.001$).
+
+(H3$_{er}$) Error rates were greater in the crisp conditions than the vague conditions when numerical distance was small, and this difference diminished with increasing numerical distance until it reversed at the biggest numerical distance.
+% MG NEED TO INSERT STATS FOR THIS INTERACTION -- KvD Can you add this please?
+If it's a problem, we can omit H3.
+
+So far, these results are in line with the idea of cost reduction.
+Cost reduction explains the vagueness advantage by claiming that the vague referring expressions place less cognitive load on the comprehender than the crisp referring expressions.
 It explains the diminishing returns for vagueness in more-discriminable stimuli (i.e., the vagueness by numerical distance interaction) by claiming that load is low in both conditions for the easily-discriminable stimuli, and that therefore there is no extra benefit to be had from vagueness in the easily-discriminable stimuli.
@@
 
-The main result from experiment 1 was that responses were faster and more accurate for vague instructions than for crisp instructions.
+% KvD: I've added some syntactic sugar, in a number of places, to clarify the relation between Ex 1 and Ex 2.
 
+However, the picture painted by these findings might be misleading.
+First of all, there is a possibly confounding factor.
 Contrast, from Experiment 1, an expression from the vague condition: `the square with few dots' with an expression from the crisp condition: `the square with 5 dots'.
-One difference is that `few' is vague (or at least has the potential for vagueness) and `5' is crisp. 
-Another difference is that `few' is a verbal quantifier while `5' is a numerical quantifier, in the sense that a number is mentioned explicitly. 
+One difference is that `few' has the potential for vagueness, whereas `5' is crisp. But
+another difference is that `few' is verbal while `5' is numerical, in the sense that a number is mentioned explicitly. 
 Since these two differences could not be separated in Experiment 1, the vagueness advantage finding is vulnerable to an alternative interpretation, that what we saw as a vagueness advantage was in contrast an advantage for the verbal form of the quantifier.
-In Experiment 2 we created verbal and numeric versions of each of the vague and crisp instructions so that we could compare vague and crisp conditions while taking account of verbal / numeric format.
+In Experiment 2 we therefore created verbal and numeric versions of each of the vague and crisp instructions so that we could compare vague and crisp conditions while taking account of verbal / numeric format.
 %In the present experiment 2 we pitted these alternative interpretations against each other in a factorial design.
 
-In Experiment 1, the participants chose one of two squares: therefore the `vague' quantifiers (e.g., `few') uniquely identified one square. 
-Recall our definition of vague  -- ``a word is precise if it describes a well-defined set of objects. By contrast, a word is vague if it is not precise''.  
-In Experiment 1, it might be said that the quantifiers in the vague conditions did not realise their potential for vagueness. 
+Another potential problem with Experiment 1 is the following.
+Participants chose one of two squares: therefore the `vague' quantifiers (e.g., `few') uniquely identified one square.
+Recall our definition of vague  -- ``a word is precise if it describes a well-defined set of objects.
+By contrast, a word is vague if it is not precise''.
+In Experiment 1, the quantifiers in the vague conditions did not realise their potential for vagueness.
 This is because there were no borderline cases of the referent that could make the referent set `not well-defined', and perhaps because using definite articles in the instructions implied that only one option was correct.
-Using error feedback in experiment 1 could have exacerbated this.
+Using error feedback in Experiment 1 could have exacerbated this.
+
+To find out what happens when words are used in a context where their potential for vagueness comes to the fore,
 Experiment 2 used three arrays so that the vague quantifiers always had more than one possible referent, and used indefinite articles in the vague instructions to avoid the impression that only one response counted as correct, and was carried out without error feedback.
@@
 
-In Experiment 2, an item was a referring expression instruction followed by a set of dot arrays defined by triple of numbers, representing the number of dots in the left, middle, and right arrays. 
+In Experiment 2, an item was a referring expression instruction followed by a set of dot arrays defined by a triple of numbers, representing the number of dots in the left, middle, and right arrays.
 We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54).
@@
 In the {\em vague numerical} condition we used \emph{Choose a square with about 10 dots}.
-None of the squares displayed contained 10 dots.
+None of the squares contained 10 dots.
 10 is slightly closer to 6 than to 15, justifying 6 as the best response and 15 as the borderline response.
 In the {\em vague verbal} condition we used \emph{Choose a square with few dots}.
-In the {\em crisp numerical} condition we used \emph{Choose the square with 6 dots} and one square always did contain the number mentioned.
-In the {\em crisp verbal} condition, we used \emph{Choose the square with the fewest dots}.
+In the {\em crisp numerical} condition we used \emph{Choose the square with 6 dots}, and one square always did contain the number mentioned.
+For {\em crisp verbal}, we used \emph{Choose the square with the fewest dots}.
+
+% KvD I've re-instated the table below, because this makes it easier to understand the instruction format/selection algorithm confound.
+
+\begin{table}[tbp]
+\caption{Table of instructions arranged by condition for the triple (6,15,24)}
+\label{instre3}
+\begin{tabular}{rll}
+&crisp&vague\\
+\hline
+numerical	&		Choose the square with 6 dots 			& 	Choose a square with about 10 dots			\\
+verbal	&		Choose the square with the fewest dots	&	Choose a square with few dots\\
+\end{tabular}
+\end{table}
 
 \subsection{Hypotheses}
-(1) A main effect RT advantage for vagueness.
+(H1) A main effect RT advantage for vagueness.
 
-(2) An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
+(H2) An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
 
-(3) No large main effect of instruction format.
+(H3) No large main effect of instruction format (numeric versus verbal). 
 %MG If we are right that vagueness was responsible for the vagueness advantage in e 1, that could also have been a number avoidance advantage, then we would not expect that number avoidance, or, instruction format, plays a role (?)
 
-(4) On the basis of Experiment 1, we would expect an effect of triple, with faster responses to more discriminable triples.
+(H4) On the basis of Experiment 1, we would expect faster responses to more discriminable triples (i.e., an effect of triple).
 
-(5) Participants should make more borderline case choices for vague than crisp instructions.
+(H5) Participants should make more borderline case choices for vague than crisp instructions.
 
@@
 
-(1) The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06, se=.01, t=4.6, p<.001$).
+(H1) The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06, se=.01, t=4.6, p<.001$).
 
-(2) Vagueness was disadvantageous in both the \emph{numeric} and \emph{verbal} conditions, offering evidence against hypothesis 2. %MG [? give values of a focussed t test]
+(H2) Vagueness was disadvantageous in both the \emph{numeric} and \emph{verbal} conditions, offering evidence against hypothesis 2. %MG [? give values of a focussed t test]
 The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to an interaction effect between vagueness and instruction format ($\beta=-0.13, se= 0.02, t= -6.6, p=0.000$).
 
-(3) There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; $\beta=.37, se=.07, t=5.1, p<.001$).
+(H3) There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; $\beta=.37, se=.07, t=5.1, p<.001$).
 
-(4) There was a significant main effect of \emph{triple} ($\beta=.06, se=.008, t=7.0, p<.001$) indicating that response times differed in some way according to which triple was presented. 
+(H4) There was a significant main effect of \emph{triple} ($\beta=.06, se=.008, t=7.0, p<.001$) indicating that response times differed in some way according to which triple was presented. 
 However further analysis revealed that there was no consistent smooth trend across different triples. 
@@
 \subsubsection{Borderline responses}
-(5) Participant grand mean percentage of borderline selections was $16.6\%$.
+(H5) Participant grand mean percentage of borderline selections was $16.6\%$.
 A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with task, vagueness and item as fixed effects, and with random slopes for task and vagueness and item over participants.
@@
 
-% KvD: I've tentatively added the following:
-Having effectively separated the {\bf cost reduction} hypothesis from the {\bf instruction format} hypothesis, it is important to observe that, in Experiment 2, instruction format went hand in hand with {\bf selection algorithm}:
+% KvD I've expanded the following:
+However, once again there is a possibly confounding factor.
+Observe that, in Experiment 2, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) {\bf selection algorithm}:
  %as shown by the Table of instructions \ref{instructionse2}, 
-the instructions that used a verbal instruction format allowed a comparison algorithm, whereas the instructions that used a numeric format allowed a matching algorithm. 
-Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison algorithm
-(which is known to be faster than matching). % KvD I assume we need to say more about this, citing some appropriate papers to support this idea.
+%the instructions that used a verbal instruction format allowed a comparison algorithm, whereas the instructions that used a numeric format allowed a matching algorithm. 
+To see this, consider the task of selecting the dot array that contains ``few dots": to do this, it suffices to {\em compare} the three arrays and select the one that contains the fewest elements.
+Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison ``algorithm" (which is known to be faster than matching).
+% KvD I think we need to say more about this, citing some appropriate papers to support this idea.
 
@@
 \centering
-\caption{Experiment 3: Instructions, assuming $6,15,24$ dots as the Item, and showing \emph{fewer} instead of \emph{more}}
+\caption{Experiment 3: Instructions, assuming $6,15,24$ dots as the Item} %, and showing \emph{fewer} instead of \emph{more}}
 \label{Instructions for e3}
@@
 \hline
-crisp & numeric&matching & Choose a square with 6 dots \\ 
-crisp & numeric&comparison & Choose a square with fewer than 20 dots \\
-vague & numeric&matching & Choose a square with about 10 dots \\ 
-vague & numeric&comparison & Choose a square with far fewer than 20 dots \\ 
+crisp 	& numeric		&	matching		& Choose a square with 6 dots \\ 
+crisp 	& numeric		&	comparison 	& Choose a square with fewer than 20 dots \\
+vague	& numeric		&	matching 		& Choose a square with about 10 dots \\ 
+vague 	& numeric		&	comparison 	& Choose a square with far fewer than 20 dots \\ 
 \hline
@@
 \subsection{Hypotheses}
-(1) Vague instructions are easier for the reader than crisp alternatives (main effect of vagueness)
+(H1) Vague instructions are easier for the reader than crisp alternatives (main effect of vagueness)
 
-(2) Comparison is easier for the reader than matching (main effect of selection task)
+(H2) Comparison is easier for the reader than matching (main effect of selection task)
 
-(3) Effects of vagueness are different depending on whether selection is matching or comparison (interaction effect selection x vagueness).
+(H3) Effects of vagueness are different depending on whether selection is matching or comparison (interaction effect selection x vagueness).
 
@@
 
-(1) There was no significant main effect of vagueness ($\beta =.003, se=.014, t=.202, p=.841$). 
+(H1) There was no significant main effect of vagueness ($\beta =.003, se=.014, t=.202, p=.841$). 
 The results showed that vagueness was beneficial for comparison but detrimental for matching. 
-(2) There was a main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=-.165, se=.027, t=-6.218, p<.001$). 
-(3) Vagueness exerted effects in different directions for the comparison task and for the matching task. 
+(H2) There was a main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=-.165, se=.027, t=-6.218, p<.001$). 
+(H3) Vagueness exerted effects in different directions for the comparison task and for the matching task. 
 Separate analyses were conducted at each level of the selection task to see whether within each task type there were significant effects of vagueness. 
@@
 
-None of the accounts set out in the Hypotheses section emerge well from the results. 
-The instruction format account wrongly predicts no differences between the conditions. 
-The selection task correctly predicted the main effect of selection task, but has no coverage of the interaction with vagueness. 
-The cost reduction account is wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
+% KvD: deleted the following:
+%None of the accounts set out in the Hypotheses section emerge well from the results. 
+%The instruction format account wrongly predicts no differences between the conditions. 
+%The selection task correctly predicted the main effect of selection task, but has no coverage of the interaction with vagueness.
+
+The cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
 
@@
 This presentation of a target before the main body of the trial shares some features with \citeA[Experiment 2]{Izard20081221}, although in that experiment participants were told the cardinality of the target (called an \emph{inducer} in that paper) whereas in our experiment we did not tell participants the cardinality of the prime array.
-An item was thus a combination of a target dot array, an instruction that did not explicitly contain a number, and a set of dot arrays taking their cardinalities from the same triples used in experiment 2.
+An item was thus a combination of a target dot array, an instruction that did not contain a number, and a set of dot arrays taking their cardinalities from the same triples used in Experiment 2.
 Table \ref{Instructions for e4} spells out how the instructions were constrained not to mention a numeral.
@@
 
-(1) Vague instructions are easier for the reader than crisp alternatives (main effect of vagueness)
+(H1) Vague instructions are easier for the reader than crisp alternatives (main effect of vagueness)
 
-(2) Comparison is easier for the reader than matching (main effect of selection)
+(H2) Comparison is easier for the reader than matching (main effect of selection)
 
-(3) Effects of vagueness are different depending on whether selection is matching or comparison (interaction effect selection x vagueness).
+(H3) Effects of vagueness are different depending on whether selection is matching or comparison (interaction effect selection x vagueness).
 
@@
 
-(1) There was no main effect of vagueness ($\beta=.01, se=.01, t=1.51, p=.14$).
+(H1) There was no main effect of vagueness ($\beta=.01, se=.01, t=1.51, p=.14$).
 
-(2) There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=-.18, se=.02, t=-10.38, p<.01$).
+(H2) There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=-.18, se=.02, t=-10.38, p<.01$).
 This effect was in the same direction as Experiment 3.
 
-(3) Vagueness did exert different effects depending on the selection task ($\beta=.12, se=.03, t=4.32, p<.05$). 
+(H3) Vagueness did exert different effects depending on the selection task ($\beta=.12, se=.03, t=4.32, p<.05$). 
 Separate analyses were conducted for the comparison task and for the matching task. In the comparison task, vagueness resulted in faster response times ($\beta=-0.08, se=.02, t=4.30, p<.05$). 
@@
 
-The cost reduction account is wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
-The instruction format account wrongly predicts no differences between the conditions. 
-The selection task account correctly predicted the main effect of selection task, but has no coverage of the interaction of selection task with vagueness.
+The cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
+
+% KvD: deleted:
+%The instruction format account wrongly predicts no differences between the conditions. 
+%The selection task account correctly predicted the main effect of selection task, but has no coverage of the interaction of selection task with vagueness.
 
@@
 
-Experiment 1 showed us that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment could not distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format; and mandated a comparison strategy rather than a matching strategy.
-Experiment 2 showed us that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional explanatory power in either the verbal format instructions or the numerical format instructions when we generated verbal and numerical versions of both crisp and vague instructions.
+Experiment 1 showed hat responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment could not distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format; and mandated a comparison strategy rather than a matching strategy.
+Experiment 2 showed that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional explanatory power in either the verbal format instructions or the numerical format instructions.
+% when we generated verbal and numerical versions of both crisp and vague instructions.
 % MJG edit
-However the experiment could not distinguish benefits of number avoidance in the verbal instructions from benefits of 
-the comparison selection task in the verbal instructions, since the verbal instructions mandated a comparison strategy, and the numeric instructions mandated a matching strategy.
-In experiments 3 and 4 we manipulated vagueness and the selection task separately at each level of numerical format.
-Across the two experiments, we found that there are effects of the selection task mandated by the instruction, with the comparison task instructions attracting faster response times than the matching instructions, and within the experiments we found that vagueness exerts benefits when the selection task is \emph{comparison}, but not when the task is \emph{matching}.
+
+However, Experiment 2 could not distinguish benefits of number avoidance from benefits of 
+the comparison selection task. %, since the verbal instructions mandated a comparison strategy, and the numeric instructions mandated a matching strategy. 
+In Experiments 3 and 4 we manipulated vagueness and the selection task separately at each level of numerical format.
+Across the two experiments, we found that %there are effects of the selection task mandated by the instruction, with
+the comparison task instructions attracted faster response times than the matching instructions.
+Within the two experiments we found that vagueness exerts benefits when the selection task is \emph{comparison}, but not when the task is \emph{matching}.
 
@@
 We believe the answer to this question is negative.
-To see why, consider figures \ref{resultse3} and \ref{resultse4}.
+To see why, consider Figures \ref{resultse3} and \ref{resultse4}.
 Both figures depict four conditions, depending on whether the expression was crisp or vague, and depending on whether the referent could be identified using a comparison strategy or not.
-Two of the resulting four conditions result in an expression that can denote either of two referents; the other two conditions result in an expression that can only denote one referent, with the other possible referent being a marginal candidate at best:
+Two of the resulting four conditions result in an expression that can denote either of two referents; the other two conditions result in an expression that can only denote one referent, with the other possible referent being a marginal candidate at best.
+
+% KvD Something missing here?
 
@@
 
-The observation that conditions with 1 candidate lead to shorter response times than conditions with 2 candidates is consistent with the range reduction hypothesis, but not with the idea that vagueness has a beneficial effect.
-It appears, in other words, that range reduction causes shorter response times, suggesting that shorter response times will only result from a vague expression if this expression leads to range reduction.
-Once again, it seems, it is not vagueness itself that has advantages but a phenomenon (namely range reduction) that is an automatic concomitant of vagueness in some types of situations.
+The observation that conditions with 1 candidate lead to shorter response times than conditions with 2 candidates is consistent with the range reduction hypothesis, but not with the idea that vagueness is beneficial.
+It appears, in other words, that %range reduction causes shorter response times, suggesting that
+shorter response times will only result from a vague expression if this expression leads to range reduction.
+Once again, it is not vagueness itself that has advantages but a phenomenon (namely range reduction) that is an automatic concomitant of vagueness in some types of situations.
 
-The findings from our experiments show that when vague expressions are compared with crisp alternatives in our forced choice task, vague expressions appear to yield benefits in some situations, but that the observed benefits may be due to factors other than vagueness itself that the vague forms bring along with them: factors like avoiding numbers; permitting comparison tasks; and range reduction.
-The picture that is starting to emerge, in other words, is subtle: on the one hand, in the situations that we have been studying -- where cooperative speakers refer to an object (e.g., a square) by means of some quantity associated with the object (e.g., the number of dots contained in the square) -- vagueness is not intrinsically beneficial.
-On the other hand, vague expressions often have other features that {\em are} beneficial, and these are what give us the incorrect impression that vagueness itself is beneficial.
+Our findings suggest that the observed benefits of vague expressions in certain situations may be due to factors other than vagueness itself: factors like avoiding numbers; permitting comparison tasks; and range reduction.
+The picture that is starting to emerge is subtle: on the one hand, in the situations that we have been studying -- where cooperative speakers refer to an object (e.g., a square) by means of some quantity associated with the object -- vagueness is not intrinsically beneficial.
+On the other hand, vague expressions frequently possess other features that {\em are} beneficial, and these are what give us the incorrect impression that vagueness itself is beneficial.
 Vagueness may thus have acquired a reputation that it does not deserve.
+To answer to Lipman's question, of why vagueness permeates human language (see our Introduction), may lie in a different direction after all, possibly relating to a combination of pure necessity and benefits for the speaker rather than the hearer (see \cite{van2010vagueness}, chapter 11 for discussion of some possible directions).
 
 A comparison may clarify the logic of the situation.
-In recent years a number of studies, focussing typically on red wine, have suggested that alcohol, consumed in low doses, may have health benefits.
+In recent years a number of studies, focussing on red wine, have suggested that alcohol, consumed in low doses, may have health benefits.
 An alternative explanation, however, asserts that it is not the alcohol in the wine that was beneficial, but antioxidants from grapes.
-If this alternative explanation is correct then alcohol may not be as beneficial as some would like to think.\\[2ex]
+If this alternative explanation is correct, then alcohol may not be healthy after all.\\[2ex]
 %
@@
 The question of how a particular piece of quantitative information is best conveyed through language is an important one for practical NLG.
-Years of research on the logic of vagueness in natural language have primed the research community to expect that the utility of vagueness is an important part of the answer, but our findings call this expectation into question.
-Although our own studies have focussed on vagueness in descriptive Noun Phrases, it seems plausible that vagueness plays a similar role in other linguistic constructs.
+Years of research on the logic of vagueness in natural language -- which has given rise to such logical techniques as Partial Logic (e.g, \cite{Fine}, Probabilistic Logic \cite{Edgington}, and Fuzzy Logic \cite{Zadeh} -- have primed the research community to expect that the utility of vagueness is an important part of the answer, but our findings call this expectation into question.
+Although our own studies in this article have focussed on vagueness in descriptive Noun Phrases, it seems plausible that vagueness plays a similar role in other linguistic constructs.
+
+% Missing bibitems:
+% Fine 1975. Kit Fine. Vagueness, Truth and Logic. {\em Synthese} {\bf 30}, pp.265-300
+% Edgington 1996. Dorothy Edgington. Vagueness by Degrees. In Keefe and Smith 1997.
+% Zadeh 1965. Lotfi Zadeh. Fuzzy Sets. {\em Information and Control} 8 (3) 1965, 338-353.
 
@@
 \begin{quote}
-(a) {\em 27.2 degrees}, or\\
+(a) {\em 27.2 degrees Celsius}, or\\
 (b) {\em approximately 27 degrees}, or\\
@@
 among other candidate expressions.
-If standard accounts are to be believed, then options (a) and (c) convey crisp information, whereas (b) and (d) are vague (i.e., they permit borderline cases).
-From our own experiments or the ones we found in the literature, there is no evidence that the fact in itself that an expression is vague should confer a benefit on it for hearers.
-Rather than asking whether a candidate expression is vague, other questions might shed more light on the NLG system's choice, similar (though not necessarily identical) to the ones identified in our studies.
+If the linguistic literature is to be believed, then options (a) and (c) convey crisp information, whereas (b) and (d) are vague (i.e., they permit borderline cases).
+From our own and previous experiments, there is no evidence that the fact in itself that an expression is vague should confer a benefit on it for hearers.
+Rather than asking whether a candidate expression is vague, other questions might shed more light on the NLG system's choice, similar to the ones identified in our studies.
 These questions might focus on the amount of information that a given expression conveys (i.e., on granularity), on the avoidance of numbers, and on the use of evaluative terms.
@@
 
-The expressions (c) and (d) -- one of which is vague while the other is crisp -- convey similar amounts of information, saying that the temperature is higher than some standard (i.e., designating a half-open temperature interval, as opposed to the closed intervals designated by (a) and (b)).
-Yet (d) (``warm") is shorter than the other three expressions and avoids the use of numbers, and our experiments suggest that this may make ``warm" more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm" (see section 1).
-Results of this kind can be applied in NLG, though obvious care would need to be exercised to take the purpose of the communication into account.\footnote{In the experiments reported in this paper, the purpose of the utterances investigated was always clear (e.g., instructing people on a weight loss program \cite{Mishra01042011}; informing a patient's choice of hospital \cite{peters2009bringing}; or selecting a dot array (in our own studies).}
+First, the experiments by Mishra et al.
+suggest that it is important how much information is conveyed by an expression, and their findings are echoed by our own thoughts about range reduction.
+In the case of (a)--(d) above, it appears that (a) conveys the most detailed information (designating the smallest segment of the temperature scale), followed by (b).
+Expression (b)  appears to be followed by (d), and (d) by (c) (e.g., 40 degrees is above 25, but at 40 Celsius the word ``warm" may no longer be applicable, giving way to words like ``hot"):
+%
+\begin{quote}
+$a < b < d < c$
+\end{quote}
+%
+If these hunches are correct then an {\em expert} may  to prefer expression (a), because it gives her the most detailed information as a basis for her decisions.
+On the other hand, expression (d) (``warm") is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make ``warm" more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm" (see section 1 and  \cite{peters2009bringing}); the factor of evaluation is especially important if the hearer is unfamiliar with the metric used (e.g., because, being American, they are more used to Fahrenheit), in which case ``warm" is much clearer.
+These considerations suggest that {\em non-experts} might prefer expression (d) over all others.
+If results of this kind were to be confirmed by experiments, they would be applicable in practical NLG, though care would need to be exercised to take the purpose of the communication into account.\footnote{One reason why ``warm" might be less preferred than these considerations suggest is that, in the weather context, some very frequently occurring numbers, such as 20 (degrees), could be relatively easy to process, analogous to the subitizable numbers of \protect\cite{trick1994small}, see our General Methodology section.} In the experiments reported in this paper, the purpose of the utterances investigated was always clear (e.g., instructing people on a weight loss program \cite{Mishra01042011}; informing a patient's choice of hospital \cite{peters2009bringing}; or selecting a dot array (in our own studies).
+% KvD See new footnote above.
 
-One way to see why vagueness may not matter as much for NLG -- and for human communication -- as is often thought, is the following thought experiment.
+One way to see why vagueness (as defined in our Introduction) may not matter as much for NLG -- and for human communication -- as is often thought, is the following thought experiment.
 Suppose a group of speakers understand the word ``warm" as vague, agreeing that temperatures above 26 count as warm, and temperatures below 24 do not count as warm, but considering temperatures between 24 and 26 as borderline cases.
-Now suppose these speakers agreed to ``precisify" their definition deciding that, henceforth, ``warm" means ``$>25$ degrees" (as in (c) above): this decision resolves the borderline cases, while everything else remains the same.
-It seems unlikely that this change in language use, from a vague meaning to a crisp one (i.e., one that has no borderline cases), would lower the utility of the word.
+Now one day these speakers agree to sharpen up their definition deciding that, henceforth, ``warm" means ``$>25$ degrees" (as in (c) above): this decision resolves the borderline cases, while everything else remains the same.
+It seems unlikely that this change in language use, from a vague meaning to a crisp one (i.e., one that has no borderline cases anymore), would lower the utility of the word.
 Our experimental findings are consistent with this idea.

### 2016-09-25 experiments/e1/writeup/e1.Rmd
made graph smaller so that text is relatively larger

@@
   geom_errorbar(width=.3, lty=1, position=position_dodge(dodge)) + 
-  geom_point(pch=21, size=3, position=position_dodge(dodge)) +
+  geom_point(pch=21, size=2, position=position_dodge(dodge)) +
   ggtitle("Response time") + 
-  ylab("log RT") + 
+  ylab("Respone time (log (ms))") + 
   xlab("Gap Size") +
-  scale_fill_grey(name=element_blank(), start=0, end=1) +
+  scale_fill_grey(name=element_blank(), start=0, end=1, guide=FALSE) +
   theme_bw() +
@@
   geom_errorbar(width=.3, lty=1, position=position_dodge(dodge)) + 
-  geom_point(pch=21, size=3, position=position_dodge(dodge)) +
+  geom_point(pch=21, size=2, position=position_dodge(dodge)) +
   ggtitle("Error rate") + 
-  ylab("error rate") + 
+  ylab("Error rate (proportion of errors)") + 
   xlab("Gap Size") +
-  scale_fill_grey(name=element_blank(), start=0,end=1)+
+  scale_fill_grey(name=element_blank(), start=0, end=1)+
   theme_bw()+
@@
 
-```{r plotRTandErrorRates, fig.width=10, fig.height=4, message=FALSE, echo=FALSE}
+```{r plotRTandErrorRates, fig.width=7.5, fig.height=3, message=FALSE, echo=FALSE}
 grid.arrange(rtplot, erplot, ncol=2)

### 2016-09-25 papers/2016-Special-Issue/v-special-issue-2016.tex
e1 results graph has more readable text

@@
-\documentclass[doc,floatsintext,apacite]{apa6}
+\documentclass[%
+doc,% saying man here makes it doublespaced; saying doc here instead makes it singlespaced again as supported by the documentclass
+floatsintext,%
+apacite%
+]{apa6}
 \usepackage{graphicx}
@@
 \usepackage{datetime}
-\usepackage{rotating} % comment out for man mode
 
-\renewcommand{\baselinestretch}{1.7} % added 21 Sept 2016, copied from CRM book
+%\renewcommand{\baselinestretch}{1.7} % added 21 Sept 2016, copied from CRM book
 
@@
 \begin{figure}[htbp]
-\fitfigure{images/resultse1.pdf}
+\fitfigure{images/plotRTandErrorRates-1}
 \caption{Experiment 1 results: response times and error rates}
@@
 \begin{table}[tbp]
-\caption{Table of instructions arranged by condition for the triple (6,15,24)}
+\caption{Experiment 1: instructions arranged by condition for the dot triple $(6,15,24)$ and instructions indicating the smaller numbers}
 \label{instre3}
@@
 &crisp&vague\\
-\hline
+\midrule
 numerical	&		Choose the square with 6 dots 			& 	Choose a square with about 10 dots			\\
 verbal	&		Choose the square with the fewest dots	&	Choose a square with few dots\\
+\midrule
 \end{tabular}
@@
 
-\subsection{Hypotheses}
-(H1) A main effect RT advantage for vagueness.
-
-(H2) An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
-
-(H3) No large main effect of instruction format (numeric versus verbal). 
-%MG If we are right that vagueness was responsible for the vagueness advantage in e 1, that could also have been a number avoidance advantage, then we would not expect that number avoidance, or, instruction format, plays a role (?)
-
-(H4) On the basis of Experiment 1, we would expect faster responses to more discriminable triples (i.e., an effect of triple).
-
-(H5) Participants should make more borderline case choices for vague than crisp instructions.
+\begin{APAenumerate}
+	\item [(H1)] A main effect RT advantage for vagueness.
+	\item [(H2)] An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
+	\item [(H3)] No large main effect of instruction format (numeric versus verbal).
+	\item [(H4)] On the basis of Experiment 1, we would expect faster responses to more discriminable triples (i.e., an effect of triple).
+	\item [(H5)] Participants should make more borderline case choices for vague than crisp instructions.
+\end{APAenumerate}
 
@@
 \centering
-\caption{Experiment 3: Instructions, assuming $6,15,24$ dots as the Item} %, and showing \emph{fewer} instead of \emph{more}}
+\caption{Experiment 3: Instructions arranged by condition for the dot triple $(6,15,24)$and instructions indicating the smaller numbers} %, and showing \emph{fewer} instead of \emph{more}}
 \label{Instructions for e3}
 \begin{tabular}{llll}
-\hline
-vagueness&instruction format&selection task&instruction\\
-\hline
-crisp 	& numeric		&	matching		& Choose a square with 6 dots \\ 
-crisp 	& numeric		&	comparison 	& Choose a square with fewer than 20 dots \\
-vague	& numeric		&	matching 		& Choose a square with about 10 dots \\ 
-vague 	& numeric		&	comparison 	& Choose a square with far fewer than 20 dots \\ 
-\hline
+vagueness	&	instruction		&	selection	&	instruction\\
+			&	format		&	task		&			\\
+\toprule
+\multirow{ 2}{*}{crisp} 	& \multirow{ 2}{*}{numeric}	&	matching		& Choose a square with 6 dots \\ 
+					& 						&	comparison 	& Choose a square with fewer than 20 dots \\
+					\midrule
+\multirow{ 2}{*}{vague} 	& \multirow{ 2}{*}{numeric} 	&	matching 		& Choose a square with about 10 dots \\ 
+ 					& 						&	comparison 	& Choose a square with far fewer than 20 dots \\ 
+\bottomrule
 \end{tabular}
@@
 \label{Instructions for e4}
-\begin{tabular}{lllp{7cm}}
-\hline
-vagueness&instruction format& selection task&instruction\\
-\hline
-crisp 	& verbal	&	matching 		& Choose a square with the same number of dots as the target \\ 
-crisp 	& verbal	&	comparison	& Choose a square with fewer dots than the target \\
-vague 	& verbal	&	matching 		& Choose a square with about the same number of dots as the target \\ 
-vague 	& verbal	&	comparison	& Choose a square with far fewer dots than the target \\ 
-\hline
+\begin{tabular}{llll}
+vagueness	&	instruction		&	selection							&	instruction: Choose a square with \ldots\\
+			&	format		&	task		&			\\
+\toprule
+\multirow{ 2}{*}{crisp} 	 	& \multirow{ 2}{*}{verbal} 	&	matching 		& the same number of dots as the target \\ 
+	& 	&	comparison										& fewer dots than the target \\
+	\midrule
+\multirow{ 2}{*}{vague} 	& \multirow{ 2}{*}{verbal} 	&	matching 			& about the same number of dots as the target \\ 
+	& 	&	comparison										& far fewer dots than the target \\ 
+\bottomrule
 \end{tabular}
@@
 vague  verbal comparison Choose a square with far fewer dots than the target
+
+%(H1) A main effect RT advantage for vagueness.
+%
+%(H2) An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
+%
+%(H3) No large main effect of instruction format (numeric versus verbal). 
+%%MG If we are right that vagueness was responsible for the vagueness advantage in e 1, that could also have been a number avoidance advantage, then we would not expect that number avoidance, or, instruction format, plays a role (?)
+%
+%(H4) On the basis of Experiment 1, we would expect faster responses to more discriminable triples (i.e., an effect of triple).
+%
+%(H5) Participants should make more borderline case choices for vague than crisp instructions.

### 2016-09-25 papers/2016-Special-Issue/v-special-issue-2016.tex
some updated stats

@@
 doc,% saying man here makes it doublespaced; saying doc here instead makes it singlespaced again as supported by the documentclass
+%man,
 floatsintext,%
@@
 Such systems are often forced to make decisions concerning the level of precision in the utterances that they generate on the basis of little more than intuition.
-Even when NLG systems are designed to mimic human language use (e.g., Konstas and Lapata 2013),
-% Ioannis Konstas and Mirella Lapata. 2013. A global model for concept-to-text generation. J. Artif. Intell. Res. (JAIR), 48:305Ð346. 
-there is no guarantee that these decisions taken by these systems benefit readers.
+Even when NLG systems are designed to mimic human language use \cite<e.g.,>{konstas2013global} there is no guarantee that these decisions taken by these systems benefit readers.
 A better understanding of the benefits (for readers and hearers) of different precision levels would allow these systems to become more useful.
@@
 % MG General methodology moved to here: per-experiment method sections only talk about deviations from this general method
-%All the experiments shared the following properties: 
+All the experiments shared the following properties: 
 Stimuli were created using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.
@@
 There was a block of practice trials after which participants could ask any questions, following which the experimenter left the room.
+All $p$ values reported for linear models were calculated using the R package \emph{lmerTest} \cite{lmerTest}.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%
+%                            E 1                                     %
+%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Experiment One}
@@
 Then a set of two dot arrays was presented, each in the form of a square containing some number of dots.
- %Fig \ref{stimuluse1} shows an example stimulus. 
 The participant was required to identify the array that corresponded with the instruction, by pressing the appropriate key, as quickly as possible while avoiding errors. 
@@
 Assuming the dot array [5, 25], and the instruction referring to the smaller cardinality, the \emph{crisp} instruction was \emph{Choose the square with 5 dots} and its \emph{vague} counterpart was \emph{Choose the square with few dots}.
-%Table \ref{instructionse1} shows examples from each condition.
 
 \subsection{Hypothesis}
-(H1) A main effect advantage for vagueness: vague instructions impose a lower cognitive load for the comprehender than crisp alternatives.
+\begin{APAenumerate}
+	\item [(H1)] A main effect advantage for vagueness: vague instructions impose a lower cognitive load for the comprehender than crisp alternatives.
+\end{APAenumerate}
+
 \subsection{Method}
-Twenty participants were recruited, aged between 18 and 45 with a median age of 26.
+
+%Twenty participants were recruited, aged between 18 and 45 with a median age of 26.
 On each trial a participant was presented with an instruction to choose one of two dot arrays on screen by reference to its cardinality. 
@@
 or if the square with the smaller number of dots was selected (when the instruction was \emph{Choose the square with many dots}). 
-% KvD Is this addition defining "erroneous" OK?
-MG Yes, I moved it though
 No feedback was given on correct trials, but there was feedback on error trials in the form of the word ``\textsc{wrong!!}'' which flashed on screen.
@@
 \subsection{Results}
-
-\subsubsection{Response times}
-Response times (RTs) for trials with erroneous responses were discarded, leading to the loss of 354 trials from 5120, representing 6.9\% of the trials. 
-The correct response RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of a further 160 trials, or 3.4\% of the remaining correct responses. 
-Means for response times and error rates are given in Fig. (\ref{resultse1}).
-A linear mixed model of RT was built, using as independent variables \emph{vagueness} and \emph{numerical distance} and their interaction, with random slopes for \emph{vagueness} and \emph{numerical distance} over participants.
-\emph{Vagueness} was sum coded: \emph{vague} $= -.5$, \emph{crisp} $= .5$; \emph{numerical distance} was Helmert coded, i.e., the variable is ordered and each level is compared against the mean of all the previous levels to that point.
-% MG a reviewer said he was not familiar with Helmert contrasts
-%Level one of this contrast is numerical distance 5 compared with numerical distance 10; level two is the mean of numerical distance 5 and numerical distance 10 versus numerical distance 15; and level three is the mean of numerical distances 5, 10 and 15 versus numerical distance 20.
-All $p$ values were calculated using the R package \emph{lmerTest} \cite{lmerTest}.
-
-(H1$_{rt}$) RTs were faster for vague instructions than for crisp instructions ($\beta=.109$, $se=.022$, $t=4.9$, $p<.001$).
-
-\subsubsection{Error rates}
-Error rate data were analysed using a generalized logit mixed model \cite{jaeger2008categorical}, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
-
-(H1$_{er}$) The effect of vagueness on error rates approached significance, with the vague instructions leading to fewer errors ($\beta=.307, se=.173, t=1.8, p=.077$).
 
@@
 
+\subsubsection{Response times}
+Response times (RTs) for trials with erroneous responses were discarded, leading to the loss of 354 trials from 5120, representing 6.9\% of the trials. 
+The correct response RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of a further 160 trials, or 3.4\% of the remaining correct responses. 
+Means for response times and error rates are given in Fig. (\ref{resultse1}).
+A linear mixed model of RT was built, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
+
+\begin{APAenumerate}
+	\item [(H1$_{rt}$)] RTs were faster for vague instructions than for crisp instructions 
+	($\beta=-0.11$, $se=.02$, $t=4.7$, $p<.001$).  % MG added the values for these stats on 25 Sept 2016
+\end{APAenumerate}
+
+\subsubsection{Error rates}
+Error rate data were analysed using a generalized logit mixed model \cite{jaeger2008categorical}, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
+
+\begin{APAenumerate}
+	\item [(H1$_{er}$)] There was a statistically non-significant effect of vagueness on error rates: the vague instructions tended to have a lower error rate than the than the crisp instructions 
+	($\beta=-.22$, $se=.18$, $t=-1.2$, $p=.230$).  % MG added the values for these stats on 25 Sept 2016
+\end{APAenumerate}
+
 \subsection{Discussion}
@@
 
-(H2) A main effect advantage of increasing numerical distance: the task will become easier as the numerical distance increases, because the two arrays are then more discriminable.
+\begin{APAenumerate}
+	\item [(H2)] A main effect advantage of increasing numerical distance: the task will become easier as the numerical distance increases, because the two arrays are then more discriminable.
+	\item [(H3)] An interaction between vagueness and numerical distance: i.e., any facilitation for vagueness should be greater at smaller numerical distances than at larger numerical distances.
+\end{APAenumerate}
 
-(H3) An interaction between vagueness and numerical distance: i.e., any facilitation for vagueness should be greater at smaller numerical distances than at larger numerical distances.
+Looking at Response Times, these two post-hoc hypotheses were confirmed, at Bonferroni-corrected significance levels:
 
-Looking at Response Times, these two post-hoc hypotheses were confirmed:
+\begin{APAenumerate}
+	\item [(H2$_{rt}$)] RTs increased with numerical distance 
+	($\beta=-0.18$, $se=0.01$, $t=-11.6$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
+	Since discriminability of the arrays is easier for larger numerical distances, discriminability probably underlies this effect. 
+	\item [(H3$_{rt}$)] Numerical distance and vagueness interacted significantly: essentially there were diminishing returns for vagueness as numerical distance increased: 
+	($\beta=0.07$, $se=0.01$, $t=5.9$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
+\end{APAenumerate}
 
-(H2$_{rt}$) RTs increased with numerical distance level one ($\beta=-.116$, $se=.012$, $t=-9.3$, $p<.05$), level 2 ($\beta=-.103$, $se=.009$, $t=-11.1$, $p<.001$) and level three ($\beta=-.082$, $se=.007$, $t=-11.0$, $p<.05$).
-Since discriminability of the arrays is easier for larger numerical distances, discriminability probably underlies this effect.
+Looking at Error Rates, the two post-hoc hypotheses were confirmed as well, at Bonferroni-corrected significance levels:
 
-(H3$_{rt}$) Numerical distance and vagueness interacted significantly for larger numerical distances when modelling RT: essentially there were diminishing returns for vagueness as numerical distance increased.% KvD Is it worth saying how they interacted?
-% MG I think this is answered now?
-The interactions at the different levels of numerical distance were: level one: ($\beta=.001$, $se=.016$, $t=.03$, $p=.974$); level two ($\beta=-.039$, $se=.009$, $t=-4.460$, $p<.001$); level three ($\beta=-.043$, $se=.006$, $t=-7.0$, $p<.001$).
-In the crisp conditions RTs started out much slower than in the vague conditions, at the smallest numerical distance, but the two conditions converged to very fast times at the largest numerical distance.
-
-Looking at Error Rates, the two post-hoc hypotheses were confirmed as well:
-
-(H2$_{er}$) Error rates decreased as numerical distance increased: level one ($\beta=-.585, se=.092, ,z=-6.4, p<.001$), level 2 ($\beta=-.434, z=-5.3,p<.001$) and level three ($\beta=-.250,z=-4.1, p<.001$).
-
-(H3$_{er}$) Error rates were greater in the crisp conditions than the vague conditions when numerical distance was small, and this difference diminished with increasing numerical distance until it reversed at the biggest numerical distance.
-% MG NEED TO INSERT STATS FOR THIS INTERACTION -- KvD Can you add this please?
-If it's a problem, we can omit H3.
+\begin{APAenumerate}
+	\item [(H2$_{er}$)] Error rates decreased as numerical distance increased: 
+	($\beta=-0.78$, $se=0.08$, $z=-9.6$, $p<0.0005$).% MG added the values for these stats on 25 Sept 2016
+	\item [(H3$_{er}$)] Numerical distance and vagueness interacted significantly: essentially there were diminishing returns for vagueness as numerical distance increased until the biggest numerical distance when vagueness was disadvantageous:
+	($\beta=0.80$, $se=0.15$, $z=5.27$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
+\end{APAenumerate}
 
@@
 
-\subsection{Introduction}
+\subsection{Introduction} % E2
 
@@
 In Experiment 2 we therefore created verbal and numeric versions of each of the vague and crisp instructions so that we could compare vague and crisp conditions while taking account of verbal / numeric format.
-%In the present experiment 2 we pitted these alternative interpretations against each other in a factorial design.
 
@@
 
-\begin{table}[tbp]
+\begin{table}[htbp]
 \caption{Experiment 1: instructions arranged by condition for the dot triple $(6,15,24)$ and instructions indicating the smaller numbers}
@@
 \midrule
-numerical	&		Choose the square with 6 dots 			& 	Choose a square with about 10 dots			\\
+numerical	&		Choose the square with 6 dots 			& 	Choose a square with about 10 dots\\
 verbal	&		Choose the square with the fewest dots	&	Choose a square with few dots\\
@@
 
+\subsection{Hypotheses} % E2
 \begin{APAenumerate}
@@
 
-\subsection{Method}
-Participants were aged between 18 and 45 with a median age of 28. 
+%%%%%%%%%%%%%%%%%%%%%%%%
+%                            E 2                                     %
+%%%%%%%%%%%%%%%%%%%%%%%%
+
+\subsection{Method} % E2
+% Participants were aged between 18 and 45 with a median age of 28. 
 We manipulated as independent variables vagueness and instruction format, yielding four conditions, \emph{vague numeric}; \emph{vague verbal}; \emph{crisp numeric}; \emph{crisp verbal}.
@@
 
-\subsection{Results}
-\subsubsection{Response times}
+\subsection{Results} % E2
+
+\subsubsection{Response times} % E2
 Means for response times and proportion of borderline responses are given in Fig. (\ref{resultse2}). 
@@
 The fixed effects in the model were \emph{instruction format} and \emph{vagueness} and their interaction, and \emph{item}.
-The random effects in the model were \emph{participant}, and slopes over \emph{participant} for \emph{instruction format}, \emph{vagueness}, and their interaction
+The random effects in the model were \emph{participant}, and slopes over \emph{participant} for \emph{instruction format}, \emph{vagueness}, and their interaction. 
+\begin{APAenumerate}
+	\item [(H1)] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms;
+	($\beta=??$, $se=??$, $t=??$, $p<??$) % MG ADD STATS
+	% $\beta=.06$, $se=.01$, $t=4.6$, $p<.001$).
+	\item [(H2)] Vagueness was disadvantageous in both the \emph{numeric} 
+	($\beta=??$, $se=??$, $t=??$, $p<??$) % MG ADD STATS
+	and \emph{verbal} 
+	($\beta=??$, $se=??$, $t=??$, $p<??$) % MG ADD STATS
+	conditions, offering evidence against hypothesis 2. 
+The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to an interaction effect between vagueness and instruction format 
+	($\beta=??$, $se=??$, $t=??$, $p<??$) % MG ADD STATS
+	% ($\beta=-0.13, se= 0.02, t= -6.6, p=0.000$).
+	\item [(H3)] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms;
+	($\beta=??$, $se=??$, $t=??$, $p<??$) % MG ADD STATS
+	% ($\beta=.37, se=.07, t=5.1, p<.001$)
+	\item [(H4)] There was a significant main effect of which triple of numbers were used in the stimulus 
+	($\beta=$, $se=$, $t=$, $p<$). 
+	%($\beta=.06, se=.008, t=7.0, p<.001$) 
+This effect seems likely to be due to the very fast responses for stimuli 
+using arrangements of the smallest numbers of dots (6,15,24), 
+which also had the largest difference in ratio of smallest number to largest number in the stimulus, 
+suggesting that stimuli using this arrangement of numbers of dots may have been particularly discriminable for participants.
+\end{APAenumerate}
 
-(H1) The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06, se=.01, t=4.6, p<.001$).
-
-(H2) Vagueness was disadvantageous in both the \emph{numeric} and \emph{verbal} conditions, offering evidence against hypothesis 2. %MG [? give values of a focussed t test]
-The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to an interaction effect between vagueness and instruction format ($\beta=-0.13, se= 0.02, t= -6.6, p=0.000$).
-
-(H3) There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; $\beta=.37, se=.07, t=5.1, p<.001$).
-
-(H4) There was a significant main effect of \emph{triple} ($\beta=.06, se=.008, t=7.0, p<.001$) indicating that response times differed in some way according to which triple was presented. 
-However further analysis revealed that there was no consistent smooth trend across different triples. 
-This effect seems likely to be due to the very fast responses for the smallest triple, which had the largest ratio difference and so may have been particularly discriminable for participants. 
-% MG reviews asked for an explanation of the difference for numeric crisp over items. Also reviews mention ratio differences as something that we did not control.
-
-\subsubsection{Borderline responses}
-(H5) Participant grand mean percentage of borderline selections was $16.6\%$.
+\subsubsection{Borderline responses} % E2
 A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with task, vagueness and item as fixed effects, and with random slopes for task and vagueness and item over participants.
 The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{resultse2}. 
-Participants were significantly more likely to choose the borderline option for vague instructions than for precise instructions (21.9\% vs 11.3\%, $\beta=.79, se=.25, z=3.2, p<.01$). 
-Participants were significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%, $\beta=3.57, se=.26, z=13.6, p<.001$).
+Participant grand mean percentage of borderline selections was $16.6\%$.  
+\begin{APAenumerate}
+	\item [(H5)] Participants were significantly more likely to choose the borderline option for vague instructions than for precise instructions (21.9\% vs 11.3\%, 
+	($\beta=??$, $se=??$, $z=??$, $p<??$) % MG ADD STATS
+	% $\beta=.79, se=.25, z=3.2, p<.01$).  
+\end{APAenumerate}
+Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%, 
+$\beta=??$, $se=??$, $z=??$, $p<??$) % MG ADD STATS
+% $\beta=3.57, se=.26, z=13.6, p<.001$).
 
@@
 
-\subsection{Discussion}
+\subsection{Discussion} % E2
 
@@
 This allowed us to distinguish between the predictions of the selection task account and the instruction format account.
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+%                            E 3                                     %
+%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%
+%                            E 4                                     %
+%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Experiment 4}
+
 \subsection{Introduction}
@@
 Vagueness may thus have acquired a reputation that it does not deserve.
-To answer to Lipman's question, of why vagueness permeates human language (see our Introduction), may lie in a different direction after all, possibly relating to a combination of pure necessity and benefits for the speaker rather than the hearer (see \cite{van2010vagueness}, chapter 11 for discussion of some possible directions).
+To answer to Lipman's question, of why vagueness permeates human language (see our Introduction), may lie in a different direction after all, possibly relating to a combination of pure necessity and benefits for the speaker rather than the hearer (\citeA<see>[chapter 11]{van2010vagueness}, for discussion of some possible directions).
 
@@
 
-% Missing bibitems:
-% Fine 1975. Kit Fine. Vagueness, Truth and Logic. {\em Synthese} {\bf 30}, pp.265-300
-% Edgington 1996. Dorothy Edgington. Vagueness by Degrees. In Keefe and Smith 1997.
-% Zadeh 1965. Lotfi Zadeh. Fuzzy Sets. {\em Information and Control} 8 (3) 1965, 338-353.
+\paragraph{Implications for practical NLG systems}
+Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
+The question of how a particular piece of quantitative information is best conveyed through language is an important one for practical NLG.
+Years of research on the logic of vagueness in natural language -- which has given rise to such logical techniques as Partial Logic (e.g, \cite{Fine}, Probabilistic Logic \cite{Edgington}, and Fuzzy Logic \cite{Zadeh} -- have primed the research community to expect that the utility of vagueness is an important part of the answer, but our findings call this expectation into question.
+Although our own studies in this article have focussed on vagueness in descriptive Noun Phrases, it seems plausible that vagueness plays a similar role in other linguistic constructs.
 
@@
 %(H5) Participants should make more borderline case choices for vague than crisp instructions.
+
+%(H3$_{rt}$) Numerical distance and vagueness interacted significantly for larger numerical distances when modelling RT: essentially there were diminishing returns for vagueness as numerical distance increased.% KvD Is it worth saying how they interacted?
+% MG I think this is answered now?
+%The interactions at the different levels of numerical distance were: level one: ($\beta=.001$, $se=.016$, $t=.03$, $p=.974$); level two ($\beta=-.039$, $se=.009$, $t=-4.460$, $p<.001$); level three ($\beta=-.043$, $se=.006$, $t=-7.0$, $p<.001$).
+%In the crisp conditions RTs started out much slower than in the vague conditions, at the smallest numerical distance, but the two conditions converged to very fast times at the largest numerical distance.
+
+\begin{APAenumerate}
+	\item 
+	\item 
+\end{APAenumerate}
+%
+%\emph{Vagueness} was sum coded: \emph{vague} $= -.5$, \emph{crisp} $= .5$; \emph{numerical distance} was Helmert coded, i.e., the variable is ordered and each level is compared against the mean of all the previous levels to that point.
+% MG a reviewer said he was not familiar with Helmert contrasts
+%Level one of this contrast is numerical distance 5 compared with numerical distance 10; level two is the mean of numerical distance 5 and numerical distance 10 versus numerical distance 15; and level three is the mean of numerical distances 5, 10 and 15 versus numerical distance 20.
+%
+%(H1$_{rt}$) RTs were faster for vague instructions than for crisp instructions ($\beta=.109$, $se=.022$, $t=4.9$, $p<.001$).
+
+%(H1$_{er}$) The effect of vagueness on error rates approached significance, with the vague instructions leading to fewer errors ($\beta=.307, se=.173, t=1.8, p=.077$).
+
+%In the present experiment 2 we pitted these alternative interpretations against each other in a factorial design.
+
+ %Fig \ref{stimuluse1} shows an example stimulus.
+
+ %Table \ref{instructionse1} shows examples from each condition.
+
+ % KvD Is this addition defining "erroneous" OK? MG Yes, I moved it though
+
+ % MG reviews asked for an explanation of the difference for numeric crisp over items. Also reviews mention ratio differences as something that we did not control.
+
+ %indicating that response times differed in some way according to which triple was presented. % However further analysis revealed that there was no consistent smooth trend across different triples.
+
+ % Missing bibitems: now added
+% Fine 1975. Kit Fine. Vagueness, Truth and Logic. {\em Synthese} {\bf 30}, pp.265-300
+% Edgington 1996. Dorothy Edgington. Vagueness by Degrees. In Keefe and Smith 1997.
+% Zadeh 1965. Lotfi Zadeh. Fuzzy Sets. {\em Information and Control} 8 (3) 1965, 338-353.
+% Ioannis Konstas and Mirella Lapata. 2013. A global model for concept-to-text generation. J. Artif. Intell. Res. (JAIR), 48:305Ð346.

### 2016-09-25 experiments/e1/writeup/e1.Rmd
changed number of digits in some tables

@@
 
-```{r pander, echo=FALSE}
+```{r pander, echo=FALSE, cache=FALSE}
 library(pander)
+library(knitr)
 ```
 
-```{r showInstructions, echo = FALSE}
+```{r showInstructions, echo=FALSE, cache=TRUE}
 v = c("crisp","vague")
@@
 
-```{r showItems, echo=FALSE}
+```{r showItems, echo=FALSE, cache=TRUE}
 items = c(1,2,3,4,5,6,7,8)
@@
 
-```{r loadLibraries, message=FALSE, echo = FALSE}
-library(knitr); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); source('summarySEwithin2.R')
+```{r loadLibraries, message=FALSE, echo=FALSE, cache=FALSE}
+library(knitr); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions); source('summarySEwithin2.R')
 ```
@@
 
-```{r separateDataFramesForRTandErrorAnalyses, echo=T}
+```{r separateDataFramesForRTandErrorAnalyses, echo=TRUE, cache=FALSE}
 rawdata <- read.table('data.txt', sep=' ')
@@
 erdata <- rawdata
+```
+
+Correct responses were trimmed at 2.5 sd separately for each subject:
+
+```{r trimData, echo=TRUE, cache=FALSE}
+tempdata <- perSubjectTrim.fnc(rtdata, response='rt', subject='subject', trim = 2.5)$data
+rtdata <- tempdata
 ```
@@
     align='lrrrrr', 
-    digits=c(0,  2,2,2,2,3),
+    digits=c(0, 2, 2, 1, 1, 3),
     x,
@@
 print(
-  xtable(summary(
-    ac.lmer.1)$coefficients,
-         digits=3, 
+  xtable(
+    summary(
+    ac.lmer.1
+    )$coefficients,
+    digits=c(0, 2, 2, 2, 5), 
     caption="Error rates model"),

### 2016-09-25 experiments/e1/writeup/e1.Rmd
recompiled unsaved digit changes

@@
     )$coefficients,
-    digits=c(0, 2, 2, 2, 5), 
+    digits=c(0, 2, 2, 2, 7), 
     caption="Error rates model"),

### 2016-09-25 experiments/e2/writeup/e2.Rmd
e2 in Rmd fomat for special issue - NEED TO DO STATS AND ADD TO TEX

@@
 ---
-title: "Experiment 2 (of 4)"
+title: "Experiment 2 of 4"
 author: "Matt Green"
 output:
-  pdf_document:
-    toc: true
+  html_document:
+    toc: yes
+    <!--toc_float: true-->
+    mathjax: default
+    template: default
+    <!--code_folding: show-->
+    theme: default
 ---  
+<style>
+table, th, td {
+    border: 1px grey;
+    border-collapse: collapse;
+    padding:2px 5px 2px 5px
+}  
+</style>
 
-```{r loadKnitr, echo=FALSE}
-suppressPackageStartupMessages(library(knitr))
+```{r setup, include=FALSE, cache=FALSE}
+library(knitr)
+knitr::opts_chunk$set(echo = TRUE)
+library(pander)
 ```
 
-```{r loadLibs, echo=FALSE, message=FALSE}
-library(beepr)                     #
-library(car)                       #
-library(ggplot2)                   #
-library(grid)                      #
-library(gridExtra)                 #
-library(languageR)                 #
-library(lattice)                   #
-library(latticeExtra)              # for doubleYScale
-library(lme4)                      #
-library(lmerTest)                  #
-library(LMERConvenienceFunctions)  # for perSubjectTrim
-library(MASS)                      #
-library(plyr)                      #
-library(reshape2)                  #
-library(rms)                       # for varclus
-library(Rmisc)                     # for summarySEwithin and friends
-library(xtable)                    #
+```{r showInstructions, echo=FALSE, cache=TRUE}
+v = c("crisp", "crisp", "vague", "vague")
+n = c('number', 'word', 'number', 'word')
+e <- c("Choose the square with 6 dots", "Choose the square with the fewest dots", "Choose a square with about 10 dots", "Choose a square with few dots")
+i = data.frame(v, n, e)
+names(i) = c("vagueness", "number", "example instruction")
+pander(i, justify=c('center', 'left', 'left'))
 ```
 
-```{r get_the_data, echo=FALSE}
-# if the file data.Rda already exists then load it, else do data wrangling
-if( file.exists('data_processed.Rda') ) {
-	load('data_processed.Rda')
-	} else {
-	gatherData = function(number_of_valid_subjects) {
-		data_dir <- '../experimentCode/output/'
-		column.headers.df <-  head( read.table(
-			paste(data_dir,'subject01.data',sep=''),
-			header=TRUE),0)
-		gathered.data <- column.headers.df
-		for (subject in 1:number_of_valid_subjects) {
-			current.filename <- paste(data_dir, 'subject', sprintf("%02d",subject), '.data', sep='')
-			current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = FALSE)
-			gathered.data <- rbind(gathered.data, current.data)
-		}
-		return(gathered.data)
-	} # end of gatherData function
-	classifyResponse = function(dat) {
-		# what were they expected to respond?
-		dat$crossed = as.factor(paste('Con', dat$Condition, ':Quan', dat$Quantity, ':Item', dat$Item, sep=''))
-		dat[dat$crossed=='Con1:Quan1:Item1', 'Exp_Num'] <-  6
-		dat[dat$crossed=='Con1:Quan1:Item1', 'Bline_Num']     <- 15
-		dat[dat$crossed=='Con1:Quan1:Item1', 'Extr_Num']      <- 24
-		dat[dat$crossed=='Con1:Quan1:Item2', 'Exp_Num'] <- 16
-		dat[dat$crossed=='Con1:Quan1:Item2', 'Bline_Num']     <- 25
-		dat[dat$crossed=='Con1:Quan1:Item2', 'Extr_Num']      <- 34
-		dat[dat$crossed=='Con1:Quan1:Item3', 'Exp_Num'] <- 26
-		dat[dat$crossed=='Con1:Quan1:Item3', 'Bline_Num']     <- 35
-		dat[dat$crossed=='Con1:Quan1:Item3', 'Extr_Num']      <- 44
-		dat[dat$crossed=='Con1:Quan1:Item4', 'Exp_Num'] <- 36
-		dat[dat$crossed=='Con1:Quan1:Item4', 'Bline_Num']     <- 45
-		dat[dat$crossed=='Con1:Quan1:Item4', 'Extr_Num']      <- 54
-		dat[dat$crossed=='Con1:Quan2:Item1', 'Exp_Num'] <- 24
-		dat[dat$crossed=='Con1:Quan2:Item1', 'Bline_Num']     <- 15
-		dat[dat$crossed=='Con1:Quan2:Item1', 'Extr_Num']      <-  6
-		dat[dat$crossed=='Con1:Quan2:Item2', 'Exp_Num'] <- 34
-		dat[dat$crossed=='Con1:Quan2:Item2', 'Bline_Num']     <- 25
-		dat[dat$crossed=='Con1:Quan2:Item2', 'Extr_Num']      <- 16
-		dat[dat$crossed=='Con1:Quan2:Item3', 'Exp_Num'] <- 44
-		dat[dat$crossed=='Con1:Quan2:Item3', 'Bline_Num']     <- 35
-		dat[dat$crossed=='Con1:Quan2:Item3', 'Extr_Num']      <- 26
-		dat[dat$crossed=='Con1:Quan2:Item4', 'Exp_Num'] <- 54
-		dat[dat$crossed=='Con1:Quan2:Item4', 'Bline_Num']     <- 45
-		dat[dat$crossed=='Con1:Quan2:Item4', 'Extr_Num']      <- 36
-		dat[dat$crossed=='Con2:Quan1:Item1', 'Exp_Num'] <-  6
-		dat[dat$crossed=='Con2:Quan1:Item1', 'Bline_Num']     <- 15
-		dat[dat$crossed=='Con2:Quan1:Item1', 'Extr_Num']      <- 24
-		dat[dat$crossed=='Con2:Quan1:Item2', 'Exp_Num'] <- 16
-		dat[dat$crossed=='Con2:Quan1:Item2', 'Bline_Num']     <- 25
-		dat[dat$crossed=='Con2:Quan1:Item2', 'Extr_Num']      <- 34
-		dat[dat$crossed=='Con2:Quan1:Item3', 'Exp_Num'] <- 26
-		dat[dat$crossed=='Con2:Quan1:Item3', 'Bline_Num']     <- 35
-		dat[dat$crossed=='Con2:Quan1:Item3', 'Extr_Num']      <- 44
-		dat[dat$crossed=='Con2:Quan1:Item4', 'Exp_Num'] <- 36
-		dat[dat$crossed=='Con2:Quan1:Item4', 'Bline_Num']     <- 45
-		dat[dat$crossed=='Con2:Quan1:Item4', 'Extr_Num']      <- 54
-		dat[dat$crossed=='Con2:Quan2:Item1', 'Exp_Num'] <- 24
-		dat[dat$crossed=='Con2:Quan2:Item1', 'Bline_Num']     <- 15
-		dat[dat$crossed=='Con2:Quan2:Item1', 'Extr_Num']      <-  6
-		dat[dat$crossed=='Con2:Quan2:Item2', 'Exp_Num'] <- 34
-		dat[dat$crossed=='Con2:Quan2:Item2', 'Bline_Num']     <- 25
-		dat[dat$crossed=='Con2:Quan2:Item2', 'Extr_Num']      <- 16
-		dat[dat$crossed=='Con2:Quan2:Item3', 'Exp_Num'] <- 44
-		dat[dat$crossed=='Con2:Quan2:Item3', 'Bline_Num']     <- 35
-		dat[dat$crossed=='Con2:Quan2:Item3', 'Extr_Num']      <- 26
-		dat[dat$crossed=='Con2:Quan2:Item4', 'Exp_Num'] <- 54
-		dat[dat$crossed=='Con2:Quan2:Item4', 'Bline_Num']     <- 45
-		dat[dat$crossed=='Con2:Quan2:Item4', 'Extr_Num']      <- 36
-		dat[dat$crossed=='Con3:Quan1:Item1', 'Exp_Num'] <-  6
-		dat[dat$crossed=='Con3:Quan1:Item1', 'Bline_Num']     <- 15
-		dat[dat$crossed=='Con3:Quan1:Item1', 'Extr_Num']      <- 24
-		dat[dat$crossed=='Con3:Quan1:Item2', 'Exp_Num'] <- 16
-		dat[dat$crossed=='Con3:Quan1:Item2', 'Bline_Num']     <- 25
-		dat[dat$crossed=='Con3:Quan1:Item2', 'Extr_Num']      <- 34
-		dat[dat$crossed=='Con3:Quan1:Item3', 'Exp_Num'] <- 26
-		dat[dat$crossed=='Con3:Quan1:Item3', 'Bline_Num']     <- 35
-		dat[dat$crossed=='Con3:Quan1:Item3', 'Extr_Num']      <- 44
-		dat[dat$crossed=='Con3:Quan1:Item4', 'Exp_Num'] <- 36
-		dat[dat$crossed=='Con3:Quan1:Item4', 'Bline_Num']     <- 45
-		dat[dat$crossed=='Con3:Quan1:Item4', 'Extr_Num']      <- 54
-		dat[dat$crossed=='Con3:Quan2:Item1', 'Exp_Num'] <- 24
-		dat[dat$crossed=='Con3:Quan2:Item1', 'Bline_Num']     <- 15
-		dat[dat$crossed=='Con3:Quan2:Item1', 'Extr_Num']      <-  6
-		dat[dat$crossed=='Con3:Quan2:Item2', 'Exp_Num'] <- 34
-		dat[dat$crossed=='Con3:Quan2:Item2', 'Bline_Num']     <- 25
-		dat[dat$crossed=='Con3:Quan2:Item2', 'Extr_Num']      <- 16
-		dat[dat$crossed=='Con3:Quan2:Item3', 'Exp_Num'] <- 44
-		dat[dat$crossed=='Con3:Quan2:Item3', 'Bline_Num']     <- 35
-		dat[dat$crossed=='Con3:Quan2:Item3', 'Extr_Num']      <- 26
-		dat[dat$crossed=='Con3:Quan2:Item4', 'Exp_Num'] <- 54
-		dat[dat$crossed=='Con3:Quan2:Item4', 'Bline_Num']     <- 45
-		dat[dat$crossed=='Con3:Quan2:Item4', 'Extr_Num']      <- 36
-		dat[dat$crossed=='Con4:Quan1:Item1', 'Exp_Num'] <-  6
-		dat[dat$crossed=='Con4:Quan1:Item1', 'Bline_Num']     <- 15
-		dat[dat$crossed=='Con4:Quan1:Item1', 'Extr_Num']      <- 24
-		dat[dat$crossed=='Con4:Quan1:Item2', 'Exp_Num'] <- 16
-		dat[dat$crossed=='Con4:Quan1:Item2', 'Bline_Num']     <- 25
-		dat[dat$crossed=='Con4:Quan1:Item2', 'Extr_Num']      <- 34
-		dat[dat$crossed=='Con4:Quan1:Item3', 'Exp_Num'] <- 26
-		dat[dat$crossed=='Con4:Quan1:Item3', 'Bline_Num']     <- 35
-		dat[dat$crossed=='Con4:Quan1:Item3', 'Extr_Num']      <- 44
-		dat[dat$crossed=='Con4:Quan1:Item4', 'Exp_Num'] <- 36
-		dat[dat$crossed=='Con4:Quan1:Item4', 'Bline_Num']     <- 45
-		dat[dat$crossed=='Con4:Quan1:Item4', 'Extr_Num']      <- 54
-		dat[dat$crossed=='Con4:Quan2:Item1', 'Exp_Num'] <- 24
-		dat[dat$crossed=='Con4:Quan2:Item1', 'Bline_Num']     <- 15
-		dat[dat$crossed=='Con4:Quan2:Item1', 'Extr_Num']      <-  6
-		dat[dat$crossed=='Con4:Quan2:Item2', 'Exp_Num'] <- 34
-		dat[dat$crossed=='Con4:Quan2:Item2', 'Bline_Num']     <- 25
-		dat[dat$crossed=='Con4:Quan2:Item2', 'Extr_Num']      <- 16
-		dat[dat$crossed=='Con4:Quan2:Item3', 'Exp_Num'] <- 44
-		dat[dat$crossed=='Con4:Quan2:Item3', 'Bline_Num']     <- 35
-		dat[dat$crossed=='Con4:Quan2:Item3', 'Extr_Num']      <- 26
-		dat[dat$crossed=='Con4:Quan2:Item4', 'Exp_Num'] <- 54
-		dat[dat$crossed=='Con4:Quan2:Item4', 'Bline_Num']     <- 45
-		dat[dat$crossed=='Con4:Quan2:Item4', 'Extr_Num']      <- 36
-		dat$crossed <- NULL
-		# what side LEFT, MIDDLE, RIGHT corresponds with Expected, Borderline, Extreme?
-		for (row in 1:nrow(dat)) {
-			if (dat[row, 'Exp_Num']   == dat[row, 'Left'])  {dat[row, 'Exp_side']   <- 'left'}
-			if (dat[row, 'Exp_Num']   == dat[row, 'Mid'])   {dat[row, 'Exp_side']   <- 'mid'}
-			if (dat[row, 'Exp_Num']   == dat[row, 'Right']) {dat[row, 'Exp_side']   <- 'right'}
-			if (dat[row, 'Bline_Num'] == dat[row, 'Left'])  {dat[row, 'Bline_side'] <- 'left'}
-			if (dat[row, 'Bline_Num'] == dat[row, 'Mid'])   {dat[row, 'Bline_side'] <- 'mid'}
-			if (dat[row, 'Bline_Num'] == dat[row, 'Right']) {dat[row, 'Bline_side'] <- 'right'}
-			if (dat[row, 'Extr_Num']  == dat[row, 'Left'])  {dat[row, 'Extr_side']  <- 'left'}
-			if (dat[row, 'Extr_Num']  == dat[row, 'Mid'])   {dat[row, 'Extr_side']  <- 'mid'}
-			if (dat[row, 'Extr_Num']  == dat[row, 'Right']) {dat[row, 'Extr_side']  <- 'right'}
-		}
-		# what button press did the subject actually make? LEFT, MIDDLE, RIGHT, NOANSWER?
-		dat$RESPONSE <- as.factor(dat$RESPONSE)
-		# what number of dots corresponds with the subject's button press?
-		for (row in 1 : nrow(dat) ) {
-			switch(as.character(dat[row,'RESPONSE']),
-						 'LEFT' = {dat[row, 'response_num'] <- dat[row, 'Left']},
-						 'MIDDLE' = {dat[row, 'response_num'] <- dat[row, 'Mid']},
-						 'RIGHT' = {dat[row, 'response_num'] <- dat[row, 'Right']},
-						 'NOANSWER' = {dat[row, 'response_num'] <- NA}
-			)
-		}
-		# what side was the subject's button-press? Left, mid right?
-		dat$response_side <- tolower(dat$RESPONSE)
-		dat$response_side[dat$response_side == "middle"] <- 'mid'
-		dat$response_side <- factor(dat$response_side, exclude="noanswer")
-		# what category was the subject's response? Expected, Borderline, Extreme
-		dat$response_category <- "nocat"
-		for (row in row.names(na.omit(dat))) {
-			if (dat[row, 'response_num'] == dat[row, 'Exp_Num']) {dat[row, 'response_category'] <- 'expected'}
-			if (dat[row, 'response_num'] == dat[row, 'Bline_Num']) {dat[row, 'response_category'] <- 'borderline'}
-			if (dat[row, 'response_num'] == dat[row, 'Extr_Num']) {dat[row, 'response_category'] <- 'extreme'}
-		}
-		dat$response_category <- factor(dat$response_category, exclude="nocat")
-		dat$RESPONSE <- NULL
-		return(dat)
-	} # end of classifyResponse function
-
-	# declare local variables
-	number_of_valid_subjects <- 30 # = 30
-	number_of_rows <- 7680 # 7680
-	number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # = 256
-	dat <- gatherData(number_of_valid_subjects) # = 30
-	dat <- classifyResponse(dat) # classify the response as expected, near, or far
-
-	# SUBJECT
-	dat$Subject=factor(paste("s",sprintf("%02d",dat$Subject),sep=""))
-	# TRIAL
-	dat$Trial = rep(x = 1:number_of_trials_per_subject, times = number_of_valid_subjects)
-	# make a centred Trial for modeling
-	dat$c_Trl <-dat$Trial - mean(dat$Trial)
-	# make a scaled Trial for modelling
-	dat$s_Trl <- as.numeric(scale(dat$Trial))
-	# ID
-	# id is a unique identifier for the 7680 row data
-	dat$id <- factor(paste(paste(dat$Subject),
-												 paste("t", sprintf("%03d", dat$Trial), sep="") , sep=":"))
-	# ITEM
-	# create a centred numeric item variable for modeling
-	dat$c_Itm <- ifelse(dat$Item==1, -.75,
-											ifelse(dat$Item==2, -.25,
-														 ifelse(dat$Item==3, .25, .75)))
-	# make Item be a factor and assign labels
-	dat$Item <- factor(dat$Item, levels=c(1,2,3,4), labels=c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
-	# add ratios for Item
-	item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54)
-	# 0.2500000 0.4705882 0.5909091 0.6666667
-	item_range_ratio_scaled = as.vector(scale(c(6 / 24, 16 / 34, 26 / 44, 36 / 54)))
-	# -1.3441995 -0.1316642  0.5297187  0.9461450
-	item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54)))
-	# 0.5125000 0.6876471 0.7691558 0.8166667
-	item_mean_ratio_scaled = as.vector(scale(c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54)))))
-	# -1.37582241 -0.06614191  0.54334858 0.89861574
-	dat[dat$Item == "06:15:24", "item_range_ratio"] <-  0.2500000
-	dat[dat$Item == "16:25:34", "item_range_ratio"] <-  0.4705882
-	dat[dat$Item == "26:35:44", "item_range_ratio"] <-  0.5909091
-	dat[dat$Item == "36:45:54", "item_range_ratio"] <-  0.6666667
-	dat[dat$Item == "06:15:24", "item_range_ratio_scaled"] <-  -1.3441995
-	dat[dat$Item == "16:25:34", "item_range_ratio_scaled"] <-  -0.1316642
-	dat[dat$Item == "26:35:44", "item_range_ratio_scaled"] <-   0.5297187
-	dat[dat$Item == "36:45:54", "item_range_ratio_scaled"] <-   0.9461450
-	dat[dat$Item == "06:15:24", "item_mean_ratio"] <-  0.5125000
-	dat[dat$Item == "16:25:34", "item_mean_ratio"] <-  0.6876471
-	dat[dat$Item == "26:35:44", "item_mean_ratio"] <-  0.7691558
-	dat[dat$Item == "36:45:54", "item_mean_ratio"] <-  0.8166667
-	dat[dat$Item == "06:15:24", "item_mean_ratio_scaled"] <-   -1.37582241
-	dat[dat$Item == "16:25:34", "item_mean_ratio_scaled"] <-   -0.06614191
-	dat[dat$Item == "26:35:44", "item_mean_ratio_scaled"] <-    0.54334858
-	dat[dat$Item == "36:45:54", "item_mean_ratio_scaled"] <-    0.89861574
-	# VAGUENESS
-	# Create a factor coding for Vagueness
-	dat[ dat$Condition==1 , 'Vagueness'] <- 'Vague'
-	dat[ dat$Condition==2 , 'Vagueness'] <- 'Crisp'
-	dat[ dat$Condition==3 , 'Vagueness'] <- 'Vague'
-	dat[ dat$Condition==4 , 'Vagueness'] <- 'Crisp'
-	dat$Vagueness <- as.factor(dat$Vagueness)
-	# manually center Vagueness
-	dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)
-	# NUMBER
-	# Create a factor coding for Number use
-	dat[ dat$Condition==1 , 'Number'] <- 'Numeric'
-	dat[ dat$Condition==2 , 'Number'] <- 'Numeric'
-	dat[ dat$Condition==3 , 'Number'] <- 'Verbal'
-	dat[ dat$Condition==4 , 'Number'] <- 'Verbal'
-	dat$Number <- as.factor(dat$Number)
-	# manually center Number
-	dat$c_Num <- ifelse(dat$Number=='Numeric', -.5, .5)
-	# CONDITION
-	# make a factor out of Condition, as f_Cnd
-	dat$f_Cnd <- factor(dat$Condition, levels=c(1,2,3,4), labels=c('Vg:Nm', 'Cr:Nm', 'Vg:Vb', 'Cr:Vb'))
-	# ORDER
-	# give the levels of Order meaningful names
-	dat$Order <- factor(dat$Order, levels=c(1,2), labels=c('LtoR','RtoL'))
-	# make a manually centred Order
-	dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)
-  # QUANTITY
-	# give the levels of Quantity meaningful names
-	dat$Quantity <- factor(dat$Quantity, levels=c(1,2), labels=c('Small','Large'))
-	# make a manually centred Quantity
-	dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)
-	# INSTRUCTION
-	# add number of characters in the instruction # 29 30 34 36 38
-	dat$nchar_instr = nchar(dat$Instruction)
-	dat$nchar_instr_scaled = as.vector(scale(nchar(dat$Instruction), scale=TRUE))
-	# make Instruction be a factor (17 levels)
-	dat$Instruction <- as.factor(dat$Instruction)
-	# RT
-	# add transformations of RT
-	dat$RT_RecSqd    <- 1/dat$RT^2
-	dat$RT_Rec       <- 1/dat$RT
-	dat$RT_RecSqt    <- 1/sqrt(dat$RT)
-	dat$RT_log       <- log(dat$RT)
-	dat$RT_sqt       <- sqrt(dat$RT)
-	dat$RT_raw       <- dat$RT
-	dat$RT_sqd       <- dat$RT^2
-  # print to file a table with information about the design
-	design_info <- unique(subset(dat, select=c(Item, Condition, Vagueness, Number, Quantity, Order,
-																						 Left, Mid, Right, Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side,Instruction)))
-	design.info <- design_info[order(design_info$Item, design_info$Condition, design_info$Quantity, design_info$Order),]
-	row.names(design_info) <- NULL
-	capture.output(print.data.frame(design_info, row.names=F, print.gap=3, quote=F, right=F),
-								 file="design_info.table")
-	# put dat in better column order
-	dat <- subset(dat,
-								select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number,
-													 Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,
-													 c_Trl, s_Trl, c_Itm, c_Vag, c_Num, f_Cnd, c_Ord, c_Qty,
-													 RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
-													 Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
-													 Instruction, nchar_instr
-								))
-	# This data set (dat) contains *all* trials 7680 including impossible trials and is
-	# mainly for graphs comparing different removals
-	save(dat, file='data_raw.Rda')
-	# remove impossible trials from dat in new data dd #
-	# Throw out RT = 1 and RT = 59998, and RTprev = 1 and RTprev = 59998
-	# i.e., throw out sticky fingers and timeouts,
-	# and the trials that followed sticky fingers and timeouts
-	# since they were likely affected by unusual previous trials.
-	# lose impossible trials
-	dd <- dat
-	dd$RT[dd$RT == 1 ] <- NA
-	dd$RT[dd$RT == 59998 ] <- NA
-	dd <- dd[complete.cases(dd),]
-	row.names(dd) <- NULL
-	# add preceding RT: because we removed impossible trials, the value for preceding RT for a trial following an impossible trial is the value of the trial that preceded the impossible trial.
-	dd$RTprev <- NA
-	for (s in levels(dd$Subject)) {
-		nrows = nrow(dd[dd$Subject==s,])
-		for (i in 1:nrows) {
-			if (i==1) {
-				dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]
-			}
-			else
-				dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1]
-		}
-	}
-	# add transformations of previous RT
-	dd$RTprev_RecSqd    <- 1/dd$RTprev^2
-	dd$RTprev_Rec       <- 1/dd$RTprev
-	dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev)
-	dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
-	dd$RTprev_log       <- log(dd$RTprev)
-	dd$RTprev_sqt       <- sqrt(dd$RTprev)
-	dd$RTprev_raw       <- dd$RTprev
-	dd$RTprev_sqd       <- dd$RTprev^2
-	# number of higher-level cells in the design (abstracting over Order) =
-	# Item(4) * Number(2) * Vagueness(2) * Quantity(2) = 32
-	# number of measurements in eaach cell = 8
-	# 332 * 8 = 256
-	# 256 * 30 = 7680
-	dd$measurement=0
-	dd$cell=0
-	for (s in levels(dd$Subject)) {
-		cellcount=0
-		for (i in levels(dd$Item)) {
-			for (n in levels(dd$Number)) {
-				for (v in levels(dd$Vagueness)) {
-					for (q in levels(dd$Quantity)) {
-						measurementcount=0
-						cellcount=cellcount+1
-						for (t in dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q, "Trial"]) {
-							dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "cell"] <- cellcount
-							measurementcount=measurementcount+1
-							dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "measurement"] = measurementcount
-						}
-					}
-				}
-			}
-		}
-	}
-	# put dd in better column order
-	dd <- subset(dd,
-							 select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number,
-							 					 Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,
-							 					 c_Trl, s_Trl, c_Itm, c_Vag, c_Num, f_Cnd, c_Ord, c_Qty,
-							 					 RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
-							 					 RTprev, RTprev_RecSqd, RTprev_Rec, RTprev_RecSqt, RTprev_log, RTprev_sqt, RTprev_raw, RTprev_sqd,
-							 					 Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
-							 					 Instruction, nchar_instr,
-							 					 cell, measurement
-							 ))
-	save(dd, file="data_processed.Rda")
-} # end of the else clause, which is run if the file data.Rda does not exist.
+```{r showItems, echo=FALSE, cache=TRUE}
+items = c(1,2,3,4,5,6,7,8)
+i = data.frame(Item = c("1","2","3","4"),
+               x = c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
+names(i) = c('Item', 'Numbers of dots')
+pander(i, justify=c('center', 'center'))
 ```
 
-Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions.
-
-```{r plotDists, fig.width=5, fig.height=5, message=FALSE, echo=FALSE}
-# first create a data frame specially for transformations
-# this won't be used for analysis, just for this plot
-load('data_raw.Rda') # dat
-dat_transforms <- dat
-# dfSubset says whether any data points were removed
-dat_transforms$dfSubset <- "original"
-# RTtype says whether the row contains data for an observed RT or 
-# a sample from a normal distribution with the same mean and sd
-dat_transforms$RTtype <- "observed"
-# create extra rows for dfs with removed (outlier) data points
-dat3 <- subset(dat_transforms, RT>1 & RT<59999)              # nrow 7677
-dat3$dfSubset <- "RT>1&RT<59999"
-dat4 <- subset(dat_transforms, RT>1 & RT<59999 & RT<30001)   # nrow 7658
-dat4$dfSubset <- "RT>1&RT<30001"
-# bring the dfs back into main df
-dat_transforms <- rbind(dat3, dat4)
-dat_transforms$dfSubset <- factor(dat_transforms$dfSubset, levels=c("RT>1&RT<59999", "RT>1&RT<30001"))
-# create extra rows for normally distributed transformed variables
-dato = dat_transforms
-datn = dat_transforms; datn$RTtype <- "normal"
-dat_transforms <- rbind(dato,datn)
-dat_transforms$RTtype <- factor(dat_transforms$RTtype, levels=c("observed","normal"))
-# for safety
-dat_transforms[dat_transforms$RTtype=="normal", c("RT_log", "RT_raw")] <- NA
-# Create the normally distributed transformed variables, replacing the NAs declared above
-for (k in levels(dat_transforms$dfSubset)){
-    dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_log"]  <- 
-    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_log"]), 
-          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_log"]), 
-          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_log"]))
-    dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_raw"]  <- 
-    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k  & dat_transforms$RTtype=="observed", "RT_raw"]), 
-          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_raw"]), 
-          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_raw"]))
-  }# end of create normally distributed equivalents for the data transforms.
-temp <- reshape2::melt(dat_transforms, 
-             id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Number", "Quantity", "Order"),
-             measure.vars=c("RT_log", "RT_raw"),
-             variable.name="transformation",
-             value.name="time")
-ggplot(temp, aes(time, colour=RTtype)) + geom_freqpoly() + facet_wrap(transformation~dfSubset, scales='free') + scale_color_manual(values=c("blue","red")) + theme_bw() + theme(aspect.ratio=1)
+```{r loadLibraries, message=FALSE, echo=FALSE, cache=FALSE}
+library(knitr); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions); source('summarySEwithin2.R')
 ```
 
-# Now working with a data set without impossible trials.
+In this experiment we did not count any response as an error, instead we coded them as expected, borderline, and extreme.
 
-Show how transformations of RT affect distribution of times, and how they affect which times are outliers.
-
-```{r boxPlots, fig.width=7, fig.height=3, echo=FALSE}
-load('data_processed.Rda')
-# allow reference to transformation
-subdata = reshape2::melt(dd,
-               measure.vars=c("RT_log", "RT_raw"),
-               variable.name="transformation",
-               value.name="value") 
-# boxplots for subjects and items separately for each dependent variable
-ggplot(subdata) + 
-  facet_wrap(~ transformation, scales="free", ncol=4) +
-  geom_boxplot(aes(y=value, x=Item, col="Items"), outlier.colour = "black", outlier.shape = 21, outlier.size=0.1) + 
-  geom_boxplot(aes(y=value, x=Subject, col="Subjects"), outlier.colour = "black", outlier.shape = 21, outlier.size=0.1) + 
-  theme_bw() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=5), aspect.ratio=1, panel.grid=element_blank()) + 
-  scale_colour_manual("",values=c("Items"="orange","Subjects"="blue")) +
-  xlab("") + ylab("")
+```{r getData, echo=T}
+rawdata <- read.table('data.txt', sep=' ')
 ```
 
-Show how mean times for individual subjects and items vary with respect to the grand mean Log RT.
-
-```{r fastSlowRT, fig.width=7, fig.height=3, message=FALSE, echo=FALSE}
-# Show how mean times for individual subjects and items vary with respect to the grand mean RT.
-subs=ddply(dd, .(level=Subject), 
-					 summarise, effect="subject", emean=mean(dd$RT_log), lmean=mean(RT_log),
-					 ldiff=mean(RT_log)-mean(dd$RT_log))
-itms=ddply(dd, .(level=Item), 
-					 summarise, effect="item", emean=mean(dd$RT_log), lmean=mean(RT_log), 
-					 ldiff=mean(RT_log)-mean(dd$RT_log))
-si=rbind(subs,itms)
-si$effect <- factor(si$effect, levels=c("subject", "item"))
-my1 <- xyplot(data=si, lmean~level | effect, 
-              aspect=1,
-              layout=c(2,1),
-              ylim = extendrange(si$lmean, f = 0.10),
-              par.settings=list(trellis.par.set(superpose.symbol = list(col='black'), 
-              																	superpose.line=list(col='black'))),
-              xlab=NULL,
-              ylab=NULL,
-              main='How mean times for individual subjects and items\nvary with respect to the grand mean Log RT',
-              mean=unique(si$emean),
-              scales=list(x=list(relation="free", at=NULL, 
-              									 labels=list("subject number", "item identifier"), cex=1 ),
-                          y=list(at=    c(0,7,unique(si$emean),8,9),
-                          			 labels=c("0","7","Mean Log RT","8","9"))),
-              panel=function(x,y,type,subscripts,...){
-                panel.abline(h=unique(si$emean), col="lightgrey",lty=1,...)
-                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, subscripts=subscripts, code=3, length=0, ...)
-                panel.xyplot(x,y,type='p', pch=21, cex=.5, col='black', fill='black', subscripts=subscripts,...)
-              }) # end of my1
-my2 <- xyplot(data=si, ldiff~level | effect, 
-              aspect=1, type='n',
-              ylab="Deviation from\nmean Log RT",
-              ylim=extendrange(si$ldiff,f=0.10)
-							) # creates a second y-axis scale
-print(doubleYScale(my1, my2, 
-									 add.axis=TRUE, 
-									 add.ylab2=TRUE, 
-									 use.style=FALSE))
+```{r}
+tempdata <- perSubjectTrim.fnc(rawdata, response='RT', subject='Subject', trim = 2.5)$data
+rtdata <- tempdata
+bldata <- tempdata
 ```
 
-# Plot main effects and interactions
+This code creates the condition means with confidence intervals used for plotting:
 
-Main effects
-
-Plot main effects on several transformations
-
-```{r plotMainsa, fig.width=6, fig.height=6, warning=FALSE, echo=FALSE}
-# plot main effects on observed data
-# allow reference to transformation
-temp <- reshape2::melt(dd,
-             id.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
-             measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log",  "RT_raw"),
-             variable.name="transformation",
-             value.name="score")
-# allow reference to effect
-temp2 <- reshape2::melt(temp, 
-              id.vars=c("transformation", "score"),
-              measure.vars=c("Item", "Vagueness", "Number", "Quantity", "Order"),
-              variable.name="effect",
-              value.name="level")
-# do the plot
-ggplot(temp2, aes(y=score, x=level, group=1, col=effect)) +
-  facet_wrap(~transformation + effect, scales="free", shrink=TRUE, ncol=5) +
-  stat_summary(fun.data=mean_cl_normal, geom="pointrange") +
-  stat_summary(fun.y=mean, geom="line") +
-  theme(aspect.ratio=1, axis.text.y=element_blank())
+```{r computeSummariesForGraphs, cache=TRUE, message=FALSE}
+rtdataplot1 <- summarySEwithin2(rtdata, measurevar="RT_log",        withinvars=c("Condition", "Vagueness", "Number", "Item"), idvar="Subject")
 ```
 
-Main effects
-
-```{r mainB, fig.width=6, fig.height=3, echo=FALSE}
-# prep main effects rts data frame
-vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness")));names(vrts)[3]<-"Levels"
-nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number")));names(nrts)[3]<-"Levels"
-qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Quantity")));names(qrts)[3]<-"Levels"
-orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Order")));names(orts)[3]<-"Levels"
-irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Item")));names(irts)[3]<-"Levels"
-rts=rbind(vrts,nrts,qrts,orts)
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-# plot main effects
-ggplot(rts, aes(y=RT_RecSqt, x=Levels, group=1, ymin=mins, ymax=maxs)) + 
-  geom_line() +
-  geom_errorbar(width=.1) +
-  geom_point(cex=3) + 
-  facet_grid(~Effect, scale="free_x") + 
-  theme_bw() +
-  theme(aspect.ratio=1) 
-```
-
-Main effects over items
-
-```{r mainD, message=FALSE, fig.width=6, fig.height=3, echo=FALSE}
-vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Vagueness")));names(vrts)[4]<-"Levels"
-nrts <- cbind(Effect="Number",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Number")));names(nrts)[4]<-"Levels"
-qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Quantity")));names(qrts)[4]<-"Levels"
-orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio","Order")));names(orts)[4]<-"Levels"
-rts=rbind(vrts,nrts,qrts,orts)
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
-ggplot(rts, 
-			 aes(y=RT_RecSqt, x=item_mean_ratio, group=Levels, ymin=mins, 
-			 		ymax=maxs, shape=Levels, fill=Levels, col=Levels)) +
-  scale_shape_manual(values=c(21,21,22,22,23,23,24,24)) +
-  scale_fill_manual(values=c("black", "gray", "black", "gray", "black", "gray", "black", "gray")) +
-  scale_color_manual(values=c("black", "gray", "black", "gray", "black", "gray", "black", "gray")) +
-  geom_line() +
-  geom_errorbar(width=.1) +
-  geom_point(cex=3) + 
-  facet_grid(~Effect) + 
-  theme_bw() +
-  theme(aspect.ratio=1, 
-        panel.grid=element_blank()) +
-  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1)) 
-```
-
-2 way interactions not involving items.
-
-```{r 2way, fig.width=6, fig.height=3, echo=FALSE}
-# prep 2 ways rts data frame
-rts1 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Number")); names(rts1)[1] <- "F1"; names(rts1)[2] <- "F2"
-rts1$Effect <- rep("Vagueness:Number",4)
-rts2 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Quantity")); names(rts2)[1] <- "F1"; names(rts2)[2] <- "F2"
-rts2$Effect <- rep("Vagueness:Quantity",4)
-rts3 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Order")); names(rts3)[1] <- "F1"; names(rts3)[2] <- "F2"
-rts3$Effect <- rep("Vagueness:Order",4)
-rts5 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Quantity")); names(rts5)[1] <- "F1"; names(rts5)[2] <- "F2"
-rts5$Effect <- rep("Number:Quantity",4)
-rts6 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Order")); names(rts6)[1] <- "F1"; names(rts6)[2] <- "F2"
-rts6$Effect <- rep("Number:Order",4)
-rts8 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Quantity","Order")); names(rts8)[1] <- "F1"; names(rts8)[2] <- "F2"
-rts8$Effect <- rep("Quantity:Order",4)
-rts <- rbind(rts1,rts2,rts3,rts5,rts6,rts8)
-rts <- subset(rts, select=c(Effect, F1, F2, RT_RecSqt, sd, se, ci))
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-# plot 2 way interactions
-ggplot(rts, aes(y=RT_RecSqt, x=F2, group=F1, ymin=mins, ymax=maxs, pch=F1)) + 
-  geom_line() +
-  geom_errorbar(width=.1) +
-  geom_point(cex=3) + 
-  facet_wrap(~Effect, scale="free_x", nrow=1) + 
-  theme_bw() +
-  theme(aspect.ratio=1) +
-  theme(legend.title=element_blank()) +
-  xlab("Levels")
-```
-
-Plot 2 way interactions over items
-
-```{r 22way, message=FALSE, fig.width=6, fig.height=3, echo=FALSE}
-# prep 2 ways over items rts data frame
-rts1 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Number")); names(rts1)[2] <- "F1"; names(rts1)[3] <- "F2"
-rts1$E1 <- "Vagueness"
-rts1$E2 <- "Number"
-rts2 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Quantity")); names(rts2)[2] <- "F1"; names(rts2)[3] <- "F2"
-rts2$E1 <- "Vagueness"
-rts2$E2 <- "Quantity"
-rts3 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Vagueness","Order")); names(rts3)[2] <- "F1"; names(rts3)[3] <- "F2"
-rts3$E1 <- "Vagueness"
-rts3$E2 <- "Order"
-rts4 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Quantity", "Number")); names(rts4)[2] <- "F1"; names(rts4)[3] <- "F2"
-rts4$E1 <- "Quantity"
-rts4$E2 <- "Number"
-rts5 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Order","Number")); names(rts5)[2] <- "F1"; names(rts5)[3] <- "F2"
-rts5$E1 <- "Order"
-rts5$E2 <- "Number"
-rts6 <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("item_mean_ratio", "Quantity","Order")); names(rts6)[2] <- "F1"; names(rts6)[3] <- "F2"
-rts6$E1 <- "Quantity"
-rts6$E2 <- "Order"
-rts <- rbind(rts1,rts2,rts3,rts4,rts5,rts6)
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
-rts$F3 <- factor(paste(rts$F1,rts$F2))
-rts$E1 <- factor(rts$E1, levels=c("Number","Order","Quantity","Vagueness"))
-rts$E2 <- factor(rts$E2, levels=c("Number","Order","Quantity","Vagueness"))
-# The palette with black:
-cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, group=F3, col=F2, pch=F1)) + 
-  facet_grid(E2~E1, drop=F) + 
-  geom_point(cex=3) + 
-	geom_line() + 
-	theme_bw() + 
-  theme(aspect.ratio=1, panel.grid=element_blank()) +
-  scale_color_manual(name="colour",values=cbbPalette) +
-  scale_shape_discrete(name="shape") +
-  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))
-```
-
-Plot vagueness by number interaction over items
-
-```{r plotInt1, fig.width=7, fig.height=3.5, echo=FALSE}
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item"))
-rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
+```{r rtplot, fig.width=5, fig.height=3}
 dodge = position_dodge(width=0.2)
-ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
-                group=condition, shape=condition, fill=condition)) +
+rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness))
+rtplot=ggplot(rtdataplot1, aes(y=RT_log, x=Item, ymin=RT_log-ci, ymax=RT_log+ci,
+                group=Condition, shape=Condition, fill=Condition)) +
   geom_line(position=dodge) +
   geom_errorbar(width=.2, position=dodge) +
-  geom_point(size=4, position=dodge) + 
+  geom_point(size=3, position=dodge) +
   scale_shape_manual("",values = c(22, 22, 21, 21)) +
@@
   ggtitle("Response time") +
-  ylab("RT_RecSqt") + 
+  ylab("Respone time (log (ms))") + 
   xlab("") +
@@
         legend.key = element_blank(), aspect.ratio=1,
-  			axis.text.x = element_text(angle = 45, hjust = 1)) +
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        plot.background=element_rect(fill=NA, color='white')) +
   facet_grid(~Number) 
+rtplot
 ```
 
-3 way interactions
-
-```{r plot3way, fig.width=7, fig.height=5, echo=FALSE}
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Quantity"))
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-p1=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Vagueness, ymin=mins, ymax=maxs, pch=Vagueness)) + 
-  geom_line() +
-  geom_errorbar(width=.1) +
-  geom_point(cex=3) + 
-  facet_grid(~Quantity, scale="free_x") + 
+```{r blplot, fig.width=5, fig.height=2.5}
+bldata$Condition <- as.factor(paste(sep=':', bldata$Vagueness, bldata$Number))
+bldata$response_category <- relevel(bldata$response_category, ref = 'expected')
+blplot=ggplot(bldata) + 
+  geom_bar(aes(Vagueness, fill=response_category), position = 'dodge') +
+  #geom_errorbar() +
+  facet_wrap(nrow=1,~Number) +
+  scale_fill_grey(name="Response Type") +
+  xlab(" ") +
+  ggtitle("Response Type") +
   theme_bw() + 
-  theme(aspect.ratio=1) +
-  theme(legend.title=element_blank(), legend.position="top")
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Order"))
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-p2=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Vagueness, ymin=mins, ymax=maxs, pch=Vagueness)) + 
-  geom_line() +
-  geom_errorbar(width=.1) +
-  geom_point(cex=3) + 
-  facet_grid(~Order, scale="free_x") + 
-  theme_bw() +
-  theme(aspect.ratio=1) +
-  theme(legend.title=element_blank(), legend.position="top")
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Quantity","Order"))
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-p3=ggplot(rts, aes(y=RT_RecSqt, x=Number, group=Quantity, ymin=mins, ymax=maxs, pch=Quantity)) + 
-  geom_line() +
-  geom_errorbar(width=.1) +
-  geom_point(cex=3) + 
-  facet_grid(~Order, scale="free_x") + 
-  theme_bw() +
-  theme(aspect.ratio=1) +
-  theme(legend.title=element_blank(), legend.position="top")
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Vagueness","Quantity","Order"))
-rts$mins=rts$RT_RecSqt-rts$ci
-rts$maxs=rts$RT_RecSqt+rts$ci
-p4=ggplot(rts, aes(y=RT_RecSqt, x=Vagueness, group=Quantity, ymin=mins, ymax=maxs, pch=Quantity)) + 
-  geom_line() +
-  geom_errorbar(width=.1) +
-  geom_point(cex=3) + 
-  facet_grid(~Order, scale="free_x") + 
-  theme_bw() +
-  theme(aspect.ratio=1) +
-  theme(legend.title=element_blank(), legend.position="top")
-grid.arrange(nrow=2, p1,p2,p3,p4)
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        legend.key = element_blank(), aspect.ratio=1,
+        plot.background=element_rect(fill=NA, color='white'))
+blplot
 ```
 
-Plot vagueness by number by quantity over items.
+<!-- ```{r} -->
+<!-- dd$response_category <- relevel(dd$response_category, ref = 'expected') -->
+<!-- barplot(table(dd$Vagueness,dd$response_category), beside=TRUE, legend=TRUE, space=c(0,.5)) -->
+<!-- ``` -->
 
-```{r vnitems, fig.width=7, fig.height=4, echo=FALSE}
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number","Vagueness","Item", "Quantity"))
-rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
-dodge = position_dodge(width=0.2)
-ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
-                group=Vagueness, 
-                fill=Vagueness)) +
-  geom_line(position=dodge) + 
-  geom_errorbar(width=.2, position=dodge) + 
-  geom_point(pch=21, size=4, position=dodge) + 
-  scale_fill_manual("",values=c("black","white")) +
-  ggtitle("Response time") + 
-  ylab("RT_Rec_Sqt") + 
-  xlab("") +
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
-        panel.background=element_rect(fill="white",color=1),
-        legend.key = element_blank(), 
-        aspect.ratio=1, axis.text.x = element_text(angle = 45, hjust = 1)) +
-  facet_grid(~Number+Quantity)
-```
-
-4 way interaction
-
-```{r 4way, fig.width=6, fig.height=3, echo=FALSE}
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness",  "Quantity", "Order"))
-rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
-ggplot(rts, aes(y=RT_RecSqt, x=Vagueness, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, group=Quantity, pch=Quantity)) +
-  facet_grid(~Order+Number, scale="free_x") +
-  geom_line() +
-  geom_errorbar(width=.1) +
-  geom_point(cex=3) + 
-  theme_bw() +
-  theme(aspect.ratio=1) +
-  theme(legend.title=element_blank())
-```
-
-4 way interaction split over items
-
-```{r 4waysplit, fig.width=7, fig.height=7, echo=FALSE}
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness", "Item", "Quantity", "Order"))
-rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
-ggplot(rts, aes(y=RT_RecSqt, x=Item, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
-                group=Vagueness,fill=Vagueness)) +
-  geom_line() + 
-  geom_errorbar(width=.12) + 
-  geom_point(pch=21, size=2) + 
-  scale_fill_manual("",values=c("black","white","black","white")) +
-  ggtitle("Number * Vagueness * Quantity * Order") + 
-  ylab("RT_RecSqt") + 
-  xlab("") +
-  theme_bw() +
-  theme( legend.key = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1),
-  			 axis.text.x = element_text(angle = 45, hjust = 1), aspect.ratio=1) +
-  facet_grid(Vagueness+Order~Number+Quantity)
-```
-
-4 way interaction split over items showing ratio of item dots.
-
-```{r 4waySplitRatio, fig.width=7, fig.height=7, echo=FALSE, message=FALSE, warning=FALSE}
-rts <- summarySEwithin(dd, measurevar="RT_RecSqt", withinvars=c("Number", "Vagueness", "item_mean_ratio", "Quantity", "Order"))
-rts$condition=paste(rts$Number,rts$Vagueness, sep= ' ')
-rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
-#dodge = position_dodge(width=0.2)
-ggplot(rts, aes(y=RT_RecSqt, x=item_mean_ratio, ymin=RT_RecSqt-ci, ymax=RT_RecSqt+ci, 
-                group=Vagueness,fill=Vagueness)) +
-  # geom_line(position=dodge) + 
-	geom_line() +
-  geom_errorbar(width=.02) + 
-  geom_point(pch=21, size=2) + 
-  scale_fill_manual("",values=c("black","white","black","white")) +
-  ggtitle("Number * Vagueness * Quantity * Order") + 
-  ylab("RT_RecSqt") + 
-  xlab("") +
-  theme_bw() +
-  theme( legend.key = element_blank(), axis.text.x = element_text(angle = 45, hjust = 1), 
-  			 axis.text.x = element_text(angle = 45, hjust = 1), aspect.ratio=1) +
-  facet_grid(Vagueness+Order~Number+Quantity)
-```
-
-# Show compression of item ratios
-
-```{r showCompression, fig.width=7, fig.height=7, echo=FALSE}
-#  Show compression of ratios of numbers of dots in the items
-item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54)
-item_range_ratio_equally_spaced = seq(from = 6 / 24, to = 36 / 54, length.out = 4)
-item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54)))
-item_mean_ratio_equally_spaced = seq(from = mean(c(6 / 15, 15 / 24)), to = mean(c(36 / 45, 45 / 54)), length.out = 4)
-plot(y=c(2,2,2,2), x=item_range_ratio, xlim=c(0,1), ylim=c(0,3), ylab="", yaxt='n', col="blue", pch=21, bg='blue', type='o', xaxt='n', xlab="ratio: small is easily discriminable", main="Discriminability of each item, expressed as ratios between array magnitudes")
-text(y=c(2.15,2.15,2.15,2.15), x=item_range_ratio, labels=round(item_range_ratio,2), col="blue")
-text(y=2.15, x=0.1, labels="ratio min/max", col='blue')
-points(y=c(2.3,2.3,2.3,2.3), x=item_range_ratio_equally_spaced, col="lightblue", pch=21, bg='lightblue', type='o')
-points(y=c(1,1,1,1), x=item_mean_ratio, pch=21, col='red', bg='red', type='o')
-text(y=c(1.15,1.15,1.15,1.15), x=item_mean_ratio, labels=round(item_mean_ratio,2), col="red")
-text(y=1.15, x=0.2, labels="mean (ratio min/med,\n            ratio med/max)", col='red')
-points(y=c(1.3,1.3,1.3,1.3), x=item_mean_ratio_equally_spaced, col="pink", pch=21, bg='pink', type='o')
-axis(1, at=seq(0,1,0.1))
-```
-
-# Lmer model
-
-```{r rmme}
-load("data_processed.Rda")
-```
-
-```{r lmer5}
-v5 <- lmerTest::lmer(data=dd, 
-           RT_log ~ 
-             c_Vag + c_Num + c_Qty + c_Ord + 
-             c_Num:c_Vag:c_Qty +
-             item_mean_ratio + 
-             s_Trl +
-             RTprev_log +
-             nchar_instr +
-             (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
-v5b <- lme4::lmer(data=dd, 
-           RT_log ~ 
-             c_Vag + c_Num + c_Qty + c_Ord + 
-             c_Num:c_Vag:c_Qty +
-             item_mean_ratio + 
-             s_Trl +
-             RTprev_log +
-             nchar_instr +
-             (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
-```
-
-```{r v51}
-summary(v5)
-```
-
-```{r v511, results='asis'}
-print(xtable(coef(summary(v5b))))
-```
-
-```{r v52}
-cat("R^2")
-cor(fitted(v5), dd$RT_RecSqt)^2
-```
-
-```{r v53}
-par(mfrow=c(2,5))
-plotLMER.fnc(v5)
-```
-
-Plot model coefficients and ci's
-
-```{r v54, fig.width=7, fig.height=4}
-e=data.frame(coef=summary(v5)$coef[-1,1]) # estimates, without intercept
-q=as.data.frame(confint(v5, method='Wald')[18:26,1:2])
-eq=cbind(e,q)
-y=eq[rev(rownames(eq)),]
-par(mar=c(4,10,1,1))
-plot(y=1:nrow(y), x=y$coef, xlim=range(y$"2.5 %", y$"97.5 %"), type='n', axes=F, xlab="", ylab="")
-segments(y0=1:nrow(y), y1=1:nrow(y), x0=y$"2.5 %", x1=y$"97.5 %", lwd=.55)
-points(y=1:nrow(y), x=y$coef, pch=20, cex=.75)
-abline(v=0, lty=3, col='grey')
-axis(2, labels=row.names(y), at=1:nrow(y), las=1)
-axis(1, las=1)
-mtext("estimate", side=1, line=2.5, cex.lab=1, las=1)
-box(col='grey')
-```
-
-Model criticism baayen plots
-
-```{r baayenPLots99, fig.width=9, fig.height=3}
-# Baayen 4-plot model criticism
-par(mfrow=c(1,4), pty='s')
-# create scaled residuals
-dd$rstand = as.vector(scale(resid(v5)))
-# plot scaled residuals density
-plot(density(dd$rstand))
-# plot sample quantiles versus theoretical quantiles
-qqnorm(dd$rstand, cex=.5)
-qqline(dd$rstand)
-# plot standardised residuals versus fitted values
-plot(dd$rstand ~ fitted(v5), pch='.')
-# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
-abline(h=c(-2.5,2.5))
-```
+<!-- , ymin=isBorderlineNormed-ci, ymax=isBorderlineNormed+ci,  group=Condition, fill=Condition)) + -->

### 2016-09-25 experiments/e2/writeup/e2-b-knitr.R

@@ -1,0 +1,1220 @@
+## ----setup, include=FALSE, cache=FALSE-----------------------------------
+library(knitr)
+options(formatR.arrow=TRUE, 
+        width=100)
+opts_chunk$set(fig.path='figure/graphics-', 
+               cache.path='cache/graphics-', 
+               fig.align='center', 
+               dev='pdf', 
+               fig.width=5, 
+               fig.height=5, 
+               fig.show='hold', 
+               cache=FALSE, 
+               par=TRUE,
+               size='scriptsize',
+               tidy=TRUE,
+               comment = NA,
+               tidy.opts = list(blank=FALSE, 
+                                width.cutoff=100))
+#knit_hooks$set(crop=hook_pdfcrop)
+
+## ----loadFonts, message=FALSE, echo=FALSE--------------------------------
+library(fontcm)
+library(extrafont)
+#font_import()
+#loadfonts()
+
+## ----loadLibs, message=FALSE, echo=FALSE---------------------------------
+library(beepr)                     #
+library(car)                       #
+library(ggplot2)                   #
+library(grid)                      #
+library(gridExtra)                 #
+library(languageR)                 #
+library(lattice)                   #
+library(latticeExtra)              # for doubleYScale
+library(lme4)                      #
+library(lmerTest)                  #
+library(LMERConvenienceFunctions)  # for perSubjectTrim
+library(MASS)                      #
+library(plyr)                      #
+library(reshape2)                  #
+library(rms)                       # for varclus
+library(Rmisc)                     # for summarySEwithin and friends
+library(xtable)                    #
+
[…]

### 2016-09-25 experiments/e2/writeup/preprocessing/e2-preprocessing.R

@@ -1,0 +1,296 @@
+## ----gatherDataFunction, echo=FALSE--------------------------------------
+gatherData = function(number_of_valid_subjects) {
+  data_dir <- '../../experimentCode/output/'
+  column.headers.df <-  head( read.table(
+    paste(data_dir,'subject01.data',sep=''),
+    header=TRUE),0)
+  gathered.data <- column.headers.df
+  for (subject in 1:number_of_valid_subjects) {
+    current.filename <- paste(data_dir, 'subject', sprintf("%02d",subject), '.data', sep='')
+    current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = FALSE)
+    gathered.data <- rbind(gathered.data, current.data)
+  }
+  message('Returning gathered data')
+  return(gathered.data)
+} # end of gatherData function
+
+## ----classifyResponseFuntion, echo=FALSE---------------------------------
+classifyResponses = function(dat) {
+  # what were they expected to respond?
+  dat$crossed = as.factor(paste('Con', dat$Condition, ':Quan', dat$Quantity, ':Item', dat$Item, sep=''))
+  dat[dat$crossed=='Con1:Quan1:Item1', 'Exp_Num']       <-  6
+  dat[dat$crossed=='Con1:Quan1:Item1', 'Bline_Num']     <- 15
+  dat[dat$crossed=='Con1:Quan1:Item1', 'Extr_Num']      <- 24
+  dat[dat$crossed=='Con1:Quan1:Item2', 'Exp_Num'] <- 16
+  dat[dat$crossed=='Con1:Quan1:Item2', 'Bline_Num']     <- 25
+  dat[dat$crossed=='Con1:Quan1:Item2', 'Extr_Num']      <- 34
+  dat[dat$crossed=='Con1:Quan1:Item3', 'Exp_Num'] <- 26
+  dat[dat$crossed=='Con1:Quan1:Item3', 'Bline_Num']     <- 35
+  dat[dat$crossed=='Con1:Quan1:Item3', 'Extr_Num']      <- 44
+  dat[dat$crossed=='Con1:Quan1:Item4', 'Exp_Num'] <- 36
[…]

### 2016-09-25 papers/2016-Special-Issue/v-special-issue-2016.tex
got up to e2 graph, BUT NOT STATS

@@
 
-As expected, responses were faster and more accurate for vague than crisp instructions. Post hoc, we formulated some further hypotheses:
+As expected, responses were faster and more accurate for vague than crisp instructions.
+Post hoc, we formulated some further hypotheses, which we tested using Bonferroni's correction for multiple comparisons:
 
@@
 
-Looking at Response Times, these two post-hoc hypotheses were confirmed, at Bonferroni-corrected significance levels:
+Looking at Response Times, these two post-hoc hypotheses were confirmed, at Bonferroni-corrected significance thresholds:
 
@@
 
-Looking at Error Rates, the two post-hoc hypotheses were confirmed as well, at Bonferroni-corrected significance levels:
+Looking at Error Rates, the two post-hoc hypotheses were confirmed as well, at Bonferroni-corrected significance thresholds:
 
@@
 It explains the diminishing returns for vagueness in more-discriminable stimuli (i.e., the vagueness by numerical distance interaction) by claiming that load is low in both conditions for the easily-discriminable stimuli, and that therefore there is no extra benefit to be had from vagueness in the easily-discriminable stimuli.
+
+%%%%%%%%%%%%%%%%%%%%%%%%
+%                            E 2                                     %
+%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 2                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
 \subsection{Method} % E2
@@
 Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%, 
-$\beta=??$, $se=??$, $z=??$, $p<??$) % MG ADD STATS
+$\beta=??$, $se=??$, $z=??$, $p<??$). % MG ADD STATS
 % $\beta=3.57, se=.26, z=13.6, p<.001$).
@@
 \centering
-\includegraphics[width=.49\textwidth]{images/response-time-sep-1.pdf}
-\includegraphics[width=.49\textwidth]{images/histbd4-ggplot-version-1.pdf}
+\includegraphics[width=.49\textwidth]{images/rtplote2.pdf}
+\includegraphics[trim = 0mm -30mm 0mm 0mm, clip, width=.49\textwidth]{images/blplote2.pdf}
 \caption{Experiment 2 results: response times and proportion of borderline cases}

### 2016-09-26 papers/2016-Special-Issue/v-special-issue-2016.tex
mainly bib edits but also e2 missing stats in red

@@
 \documentclass[%
-doc,% saying man here makes it doublespaced; saying doc here instead makes it singlespaced again as supported by the documentclass
-%man,
+%			% choose man or doc immediately below
+doc,			% doc is single spaced
+%man,		% man is double spaced
 floatsintext,%
@@
 
-%\renewcommand{\baselinestretch}{1.7} % added 21 Sept 2016, copied from CRM book
+%\renewcommand{\baselinestretch}{1.7} % added 21 Sept 2016, copied from CRM book, but see documentclass option 'man' for "manuscript'' above which makes it double spaced
 
@@
 
-When the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the cardinality of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene199647}.
+When the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the cardinality of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene1996}.
 We manipulated the number of dots in each array such that some sets of arrays had smaller numerical distances and others had larger numerical distances.
@@
 \fitfigure{images/plotRTandErrorRates-1}
-\caption{Experiment 1 results: response times and error rates}
+\caption{Experiment 1 results: response times and error rates.}
 \label{resultse1}
@@
 \midrule
-numerical	&		Choose the square with 6 dots 			& 	Choose a square with about 10 dots\\
-verbal	&		Choose the square with the fewest dots	&	Choose a square with few dots\\
+numerical	&	Choose the square with 6 dots 			& 	Choose a square with about 10 dots	\\
+verbal	&	Choose the square with the fewest dots	&	Choose a square with few dots		\\
 \midrule
@@
 	\item [(H1)] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms;
-	($\beta=??$, $se=??$, $t=??$, $p<??$) % MG ADD STATS
+	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}) % MG ADD STATS
 	% $\beta=.06$, $se=.01$, $t=4.6$, $p<.001$).
 	\item [(H2)] Vagueness was disadvantageous in both the \emph{numeric} 
-	($\beta=??$, $se=??$, $t=??$, $p<??$) % MG ADD STATS
+	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}) % MG ADD STATS
 	and \emph{verbal} 
-	($\beta=??$, $se=??$, $t=??$, $p<??$) % MG ADD STATS
+	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}) % MG ADD STATS
 	conditions, offering evidence against hypothesis 2. 
 The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to an interaction effect between vagueness and instruction format 
-	($\beta=??$, $se=??$, $t=??$, $p<??$) % MG ADD STATS
+	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}) % MG ADD STATS
 	% ($\beta=-0.13, se= 0.02, t= -6.6, p=0.000$).
 	\item [(H3)] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms;
-	($\beta=??$, $se=??$, $t=??$, $p<??$) % MG ADD STATS
+	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}) % MG ADD STATS
 	% ($\beta=.37, se=.07, t=5.1, p<.001$)
 	\item [(H4)] There was a significant main effect of which triple of numbers were used in the stimulus 
-	($\beta=$, $se=$, $t=$, $p<$). 
+	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}). 
 	%($\beta=.06, se=.008, t=7.0, p<.001$) 
@@
 The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{resultse2}. 
-Participant grand mean percentage of borderline selections was $16.6\%$.  
+Participants chose the borderline square on $16.6\%$ of trials overall.  
 \begin{APAenumerate}
 	\item [(H5)] Participants were significantly more likely to choose the borderline option for vague instructions than for precise instructions (21.9\% vs 11.3\%, 
-	($\beta=??$, $se=??$, $z=??$, $p<??$) % MG ADD STATS
+	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}) % MG ADD STATS
 	% $\beta=.79, se=.25, z=3.2, p<.01$).  
@@
 Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%, 
-$\beta=??$, $se=??$, $z=??$, $p<??$). % MG ADD STATS
+\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}). % MG ADD STATS
 % $\beta=3.57, se=.26, z=13.6, p<.001$).
@@
 \includegraphics[trim = 0mm -30mm 0mm 0mm, clip, width=.49\textwidth]{images/blplote2.pdf}
-\caption{Experiment 2 results: response times and proportion of borderline cases}
+\caption{Experiment 2 results: mean response times by condition and item, and counts of borderline case responses by condition.}
 \label{resultse2}
@@
 % KvD I think we need to say more about this, citing some appropriate papers to support this idea.
+\textcolor{red}{insert citation supporting this}
 
@@
 % Ioannis Konstas and Mirella Lapata. 2013. A global model for concept-to-text generation. J. Artif. Intell. Res. (JAIR), 48:305Ð346.
+
+\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}
+$\beta=??$, $se=??$, $t=??$, $p<??$

### 2016-09-26 experiments/e2/writeup/e2.Rmd
finished updated stats for e2

@@
 
-```{r rtplot, fig.width=5, fig.height=3}
+```{r rtplot, fig.width=5, fig.height=3, echo=F}
 dodge = position_dodge(width=0.2)
@@
 
-```{r blplot, fig.width=5, fig.height=2.5}
+```{r blplot, fig.width=5, fig.height=2.5, echo=F}
 bldata$Condition <- as.factor(paste(sep=':', bldata$Vagueness, bldata$Number))
@@
 
-<!-- ```{r} -->
-<!-- dd$response_category <- relevel(dd$response_category, ref = 'expected') -->
-<!-- barplot(table(dd$Vagueness,dd$response_category), beside=TRUE, legend=TRUE, space=c(0,.5)) -->
-<!-- ``` -->
+```{r rtlmer, cache=TRUE}
+rtlmer = lmerTest::lmer(data=rtdata,
+                        RT_log ~ c_Vag * c_Num + c_Itm +
+                          (1 + c_Vag * c_Num + c_Itm | Subject))
+```
 
-<!-- , ymin=isBorderlineNormed-ci, ymax=isBorderlineNormed+ci,  group=Condition, fill=Condition)) + -->
+```{r rtlmersummaryxtable, results='asis', cache=TRUE}
+print(xtable(summary(rtlmer)$coef, digits=c(0,2,2,1,1,4)),type='html')
+```
+
+```{r seplmersfornum, results='asis', cache=TRUE}
+s1 <- subset(rtdata, subset=Number=='Numeric')
+s1lm <- lmerTest::lmer(data=s1, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
+print(xtable(summary(s1lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
+```
+
+```{r seplmersforvag, results='asis', cache=TRUE}
+s2 <- subset(rtdata, subset=Number=='Verbal')
+s2lm <- lmerTest::lmer(data=s2, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
+print(xtable(summary(s2lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
+```
+
+```{r bllmer, cache=TRUE}
+bllmer = lme4::glmer(data=bldata,
+                        isBorderline ~ c_Vag * c_Num + c_Itm +
+                          (1 + c_Vag * c_Num + c_Itm | Subject),
+                        family="binomial", control = glmerControl(optimizer = "bobyqa"))
+```
+
+```{r bllmersummaryxtable, results='asis', cache=TRUE}
+print(xtable(summary(bllmer)$coef, digits=c(0,2,2,1,4)),type='html')
+```

### 2016-09-26 papers/2016-Special-Issue/v-special-issue-2016.tex
finished updated stats for e2 in tex

@@
 Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data. 
-A linear mixed model was constructed for the (logged) response times.
-The \emph{instruction format} and \emph{vagueness} variables were sum-coded and \emph{Item} was centred.
-The fixed effects in the model were \emph{instruction format} and \emph{vagueness} and their interaction, and \emph{item}.
-The random effects in the model were \emph{participant}, and slopes over \emph{participant} for \emph{instruction format}, \emph{vagueness}, and their interaction. 
+A linear mixed model was constructed for the (logged) response times, 
+with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
 \begin{APAenumerate}
-	\item [(H1)] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms;
-	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}) % MG ADD STATS
-	% $\beta=.06$, $se=.01$, $t=4.6$, $p<.001$).
-	\item [(H2)] Vagueness was disadvantageous in both the \emph{numeric} 
-	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}) % MG ADD STATS
-	and \emph{verbal} 
-	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}) % MG ADD STATS
-	conditions, offering evidence against hypothesis 2. 
-The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to an interaction effect between vagueness and instruction format 
-	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}) % MG ADD STATS
-	% ($\beta=-0.13, se= 0.02, t= -6.6, p=0.000$).
-	\item [(H3)] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms;
-	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}) % MG ADD STATS
-	% ($\beta=.37, se=.07, t=5.1, p<.001$)
-	\item [(H4)] There was a significant main effect of which triple of numbers were used in the stimulus 
-	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}). 
-	%($\beta=.06, se=.008, t=7.0, p<.001$) 
-This effect seems likely to be due to the very fast responses for stimuli 
-using arrangements of the smallest numbers of dots (6,15,24), 
-which also had the largest difference in ratio of smallest number to largest number in the stimulus, 
-suggesting that stimuli using this arrangement of numbers of dots may have been particularly discriminable for participants.
+	\item [(H1)] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06$, $se=.01$, $t=4.6$, $p<0.0001$) % MG added the values for these stats on 26 Sept 2016
+	\item [(H2)] In focussed comparisons, vagueness was significantly disadvantageous in the \emph{numeric} conditions ($\beta=0.09$, $se=0.02$, $t=4.5$, $p<0.0001$)% MG added the values for these stats on 26 Sept 2016
+				, and non-significantly disadvantageous in the \emph{verbal} conditions ($\beta=0.02$, $se=0.02$, $t=1.5$, $p=0.1550$)% MG added the values for these stats on 26 Sept 2016
+				, offering partial evidence against hypothesis 2.
+The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to a significant overall interaction effect between vagueness and instruction format ($\beta=-0.13$, $se=0.02$, $t=-6.6$, $p=0.0169$) % MG added the values for these stats on 26 Sept 2016
+	\item [(H3)] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; ($\beta=-0.36$, $se=0.07$, $t=-5.1$, $p<0.0001$) % MG added the values for these stats on 26 Sept 2016
+	\item [(H4)] There was a significant main effect of item (i.e., of which triple of numbers of dots were used in the stimulus): ($\beta=0.12$, $se=0.02$, $t=7.1$, $p<0.0001$).
+% MG added the values for these stats on 26 Sept 2016
+				This effect seems likely to be due to the very fast responses for stimuli using arrangements of the smallest numbers of dots (6,15,24), which also had the largest difference in ratio of smallest number to largest number in the stimulus, suggesting that stimuli using this arrangement of numbers of dots may have been particularly discriminable for participants.
 \end{APAenumerate}
@@
 \subsubsection{Borderline responses} % E2
-A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with task, vagueness and item as fixed effects, and with random slopes for task and vagueness and item over participants.
+A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
 The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{resultse2}. 
@@
 \begin{APAenumerate}
-	\item [(H5)] Participants were significantly more likely to choose the borderline option for vague instructions than for precise instructions (21.9\% vs 11.3\%, 
-	(\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}) % MG ADD STATS
-	% $\beta=.79, se=.25, z=3.2, p<.01$).  
+	\item [(H5)] Participants were significantly more likely to choose the borderline option for vague instructions than for crisp instructions (21.9\% vs 11.3\%: $\beta=0.62$, $se=0.22$, $z=2.8$, $p=0.0059$).
+% MG added the values for these stats on 26 Sept 2016
+	Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%: $\beta=-3.35$, $se=0.23$, $z=-14.6$, $p<0.0001$).
+% MG added the values for these stats on 26 Sept 2016
 \end{APAenumerate}
-Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%, 
-\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}). % MG ADD STATS
-% $\beta=3.57, se=.26, z=13.6, p<.001$).
 
@@
 
+% MG has not updated these stats yet at 26 Sept 2016
 (H1) There was no significant main effect of vagueness ($\beta =.003, se=.014, t=.202, p=.841$). 
@@
 In order to implement the experiment without mentioning numbers in the instructions, we changed the sequence of each trial to include a 'target' (i.e., a dot array of a particular cardinality) before the instruction, so that we could then refer back to the target's cardinality in the instruction using expressions like \emph{the same number of dots as the target}; \emph{fewer dots than the target}.
-%?MG WHAT WERE THE VALUES OF THE TARGETS?
+% MG WHAT WERE THE VALUES OF THE TARGETS?
+\textcolor{red}{MG insert the values of the targets}
 This presentation of a target before the main body of the trial shares some features with \citeA[Experiment 2]{Izard20081221}, although in that experiment participants were told the cardinality of the target (called an \emph{inducer} in that paper) whereas in our experiment we did not tell participants the cardinality of the prime array.
@@
 
+% MG has not updated these stats yet at 26 Sept 2016
 (H1) There was no main effect of vagueness ($\beta=.01, se=.01, t=1.51, p=.14$).
@@
 $\beta=??$, $se=??$, $t=??$, $p<??$
+
+%The \emph{instruction format} and \emph{vagueness} variables were sum-coded and \emph{Item} was centred.
+%The fixed effects in the model were \emph{instruction format} and \emph{vagueness} and their interaction, and \emph{item}.
+%The random effects in the model were \emph{participant}, and slopes over \emph{participant} for \emph{instruction format}, \emph{vagueness}, and their interaction.

### 2016-09-26 papers/2016-Special-Issue/v-special-issue-2016.tex
numbered sections

@@
 %			% choose man or doc immediately below
-doc,			% doc is single spaced
-%man,		% man is double spaced
+%doc,	% doc is single spaced
+man,		% man is double spaced
 floatsintext,%
@@
 \date{\today}
-\authornote{\today ~\currenttime}
+\authornote{latex compiled \today ~\currenttime}
 \shorttitle{Is Vagueness Beneficial for Hearers?}
@@
 \maketitle   
-
+\setcounter{secnumdepth}{3}
 \section{Introduction}\label{introduction}
@@
 
-\subsubsection{General Methodology}\label{method}
+\subsection{General Methodology}\label{method}
 The experiments reported in the present paper put the cost reduction hypothesis to the test. 
@@
 % MG has not updated these stats yet at 26 Sept 2016
-(H1) There was no significant main effect of vagueness ($\beta =.003, se=.014, t=.202, p=.841$). 
+(H1) There was no significant main effect of vagueness ($\beta =.003$, $se=.014$, $t=.202$, $p=.841$). 
 The results showed that vagueness was beneficial for comparison but detrimental for matching. 
@@
 \toprule
-\multirow{ 2}{*}{crisp} 	 	& \multirow{ 2}{*}{verbal} 	&	matching 		& the same number of dots as the target \\ 
-	& 	&	comparison										& fewer dots than the target \\
+\multirow{ 2}{*}{crisp} 	 	& \multirow{ 2}{*}{verbal} 	&	matching 		& \ldots the same number of dots as the target \\ 
+	& 	&	comparison										& \ldots fewer dots than the target \\
 	\midrule
-\multirow{ 2}{*}{vague} 	& \multirow{ 2}{*}{verbal} 	&	matching 			& about the same number of dots as the target \\ 
-	& 	&	comparison										& far fewer dots than the target \\ 
+\multirow{ 2}{*}{vague} 	& \multirow{ 2}{*}{verbal} 	&	matching 			& \ldots about the same number of dots as the target \\ 
+	& 	&	comparison										& \ldots far fewer dots than the target \\ 
 \bottomrule
@@
 % KvD Something missing here?
+\textcolor{red}{KvD Something missing here?}
 
@@
 {\bf Implications for practical NLG systems.} Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
-The question of how a particular piece of quantitative information is best conveyed through language is an important one for practical NLG.
-Years of research on the logic of vagueness in natural language -- which has given rise to such logical techniques as Partial Logic (e.g, \cite{Fine}, Probabilistic Logic \cite{Edgington}, and Fuzzy Logic \cite{Zadeh} -- have primed the research community to expect that the utility of vagueness is an important part of the answer, but our findings call this expectation into question.
-Although our own studies in this article have focussed on vagueness in descriptive Noun Phrases, it seems plausible that vagueness plays a similar role in other linguistic constructs.
-
-\paragraph{Implications for practical NLG systems}
-Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
 The question of how a particular piece of quantitative information is best conveyed through language is an important one for practical NLG.
@@
 
-%\bibliography{bibliography}
-
+\setcounter{secnumdepth}{0}
 \bibliography{v-special-issue-2016}

### 2016-09-27 experiments/e3/writeup/e3.Rmd
e3 writeup less verbose

--- 2016-07-28 experiments/e3/writeup/experiment3.Rmd
@@
     template: default
-    code_folding: hide
+    <!--code_folding: show-->
     theme: default
 ---  
+<style>
+table, th, td {
+    border: 1px grey;
+    border-collapse: collapse;
+    padding:2px 5px 2px 5px
+}  
+</style>
+
 # Experiment 3
 
-```{r docOpts}
-library(knitr)
+```{r loadLibraries, message=FALSE}
+library(knitr); library(MASS); library(ggplot2); library(lme4); library(xtable); library(plyr); library(lattice); library(LMERConvenienceFunctions); library(car); library(grid); library(lmerTest); library(reshape2); library(ggplot2); library(languageR); library(pander); source('summarySEwithin2.R')
 ```
 
-```{r loadLibraries}
-suppressPackageStartupMessages(library(MASS))
-suppressPackageStartupMessages(library(ggplot2))
-suppressPackageStartupMessages(library(lme4))
-suppressPackageStartupMessages(library(xtable))
-suppressPackageStartupMessages(library(plyr))
-suppressPackageStartupMessages(library(reshape))
-suppressPackageStartupMessages(library(lattice))
-suppressPackageStartupMessages(library(LMERConvenienceFunctions))
-suppressPackageStartupMessages(library(car))
-suppressPackageStartupMessages(library(grid))
-suppressPackageStartupMessages(library(lmerTest))
-suppressPackageStartupMessages(library(reshape2))
-suppressPackageStartupMessages(library(ggplot2))
-suppressPackageStartupMessages(require(languageR))
-```
-
-# Experiment Structure
+Experiment Structure: 
 Each subject saw 192 trials, being 3 cycles of 64 trials. There were 38 participants. There were 7296 rows in the full data set.
 
-# Trial Procedure
-The procedure for each trial was as follows.
+Trial Procedure: 
 First an instruction that was a referring expression was presented until participants dismissed it with a key press.
@@
 
-# Prepare the data
-Source additional R files with functions for computing confidence intervals.
-
-```{r sourceFunctions}
-source("functions/CIfunctions.R")
+```{r showInstructions, echo=FALSE, cache=TRUE}
+v = c("crisp", "crisp", "vague", "vague")
+n = c('number', 'number', 'number', 'number')
+t = c("matching", "comparison", "matching", "comparison")
+e = c("Choose a square with 6 dots", "Choose a square with fewer than 20 dots", "Choose a square with about 10 dots", "Choose a square with far fewer than 20 dots")
+i = data.frame(v, n, t, e)
+names(i) = c("vagueness", "instruction format", "selection task", "example instruction")
+pander(i, justify=c('center', 'left', 'left', 'left'))
 ```
 
-Get data.
-
-```{r readdata}
-dat <- read.delim("data/allresults.txt", stringsAsFactors = FALSE)
-
-# declare local variables
-number_of_valid_subjects <- 38
-number_of_rows <- 7296
-number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # 192
-
-# sort out borderline responses into expected near and far
-# How many squares in the square they chose? 
-dat$RESPONSE <- as.character(dat$RESPONSE)
-for (row in 1 : nrow(dat) ) {
-  switch(dat[row,'RESPONSE'],
-         'LEFT' = {dat[row, 'choice'] <- dat[row, 'Left']},
-         'MIDDLE' = {dat[row, 'choice'] <- dat[row, 'Mid']},
-         'RIGHT' = {dat[row, 'choice'] <- dat[row, 'Right']}         
-  )
-}
-dat$crossed=paste('Con',dat$Condition,':Quan',dat$Quantity,':Item',dat$Item,sep="")
-
-dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseExpected'] <-  6 
-dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseNear']     <- 15  
-dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseFar']      <- 24
-
-dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseExpected'] <- 16
-dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseFar']      <- 34
-
-dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseExpected'] <- 26
-dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseFar']      <- 44
-
-dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseExpected'] <- 36
-dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseFar']      <- 54
-
-dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseExpected'] <- 24
-dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseNear']     <- 15 
-dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseFar']      <-  6
-
-dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseExpected'] <- 34
-dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseFar']      <- 16
-
-dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseExpected'] <- 44
-dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseFar']      <- 26
-
-dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseExpected'] <- 54
-dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseFar']      <- 36
-
-dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseExpected'] <-  6 
-dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseNear']     <- 15  
-dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseFar']      <- 24
-
-dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseExpected'] <- 16
-dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseFar']      <- 34
-
-dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseExpected'] <- 26
-dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseFar']      <- 44
-
-dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseExpected'] <- 36
-dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseFar']      <- 54
-
-dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseExpected'] <- 24
-dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseNear']     <- 15 
-dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseFar']      <-  6
-
-dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseExpected'] <- 34
-dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseFar']      <- 16
-
-dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseExpected'] <- 44
-dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseFar']      <- 26
-
-dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseExpected'] <- 54
-dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseFar']      <- 36
-
-dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseExpected'] <-  6 
-dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseNear']     <- 15  
-dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseFar']      <- 24
-
-dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseExpected'] <- 16
-dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseFar']      <- 34
-
-dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseExpected'] <- 26
-dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseFar']      <- 44
-
-dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseExpected'] <- 36
-dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseFar']      <- 54
-
-dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseExpected'] <- 24
-dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseNear']     <- 15 
-dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseFar']      <-  6
-
-dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseExpected'] <- 34
-dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseFar']      <- 16
-
-dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseExpected'] <- 44
-dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseFar']      <- 26
-
-dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseExpected'] <- 54
-dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseFar']      <- 36
-
-dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseExpected'] <-  6 
-dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseNear']     <- 15  
-dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseFar']      <- 24
-
-dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseExpected'] <- 16
-dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseFar']      <- 34
-
-dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseExpected'] <- 26
-dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseFar']      <- 44
-
-dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseExpected'] <- 36
-dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseFar']      <- 54
-
-dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseExpected'] <- 24
-dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseNear']     <- 15 
-dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseFar']      <-  6
-
-dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseExpected'] <- 34
-dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseFar']      <- 16
-
-dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseExpected'] <- 44
-dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseFar']      <- 26
-
-dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseExpected'] <- 54
-dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseFar']      <- 36
-
-dat$isResponseExpected <- dat$choice == dat$ResponseExpected
-dat$isResponseNear <- dat$choice == dat$ResponseNear
-dat$isResponseFar <- dat$choice == dat$ResponseFar
-
-for ( row in 1:nrow(dat) ) {
-  dat[row, 'switchResponse'] <- 
-    ifelse(dat[row, 'isResponseExpected']==TRUE, 'Expected', 
-           ifelse(dat[row, 'isResponseNear']==TRUE, 'Near', 'Far') ) 
-}
-
-## treat variables
-
-# SUBJECT
-# ensure Subject is a factor 
-dat$Subject=factor(paste("s",sprintf("%02d",dat$Subject),sep=""))
-
-# TRIAL
-# Trial for subject, 1 to 192
-dat$Trial = rep(x = 1:number_of_trials_per_subject, times = number_of_valid_subjects)
-
-# make a centred Trial for modeling
-dat$c_Trl <-dat$Trial - mean(dat$Trial)
-
-# make a scaled Trial for modelling
-dat$s_Trl <- as.numeric(scale(dat$Trial))
-
-# ID
-# id is a unique identifier for the 7680 row data
-dat$id <- factor(paste(paste(dat$Subject), 
-                       paste("t", sprintf("%03d", dat$Trial), sep="") , sep=":"))
-
-# ITEM
-# create a centred numeric item variable for modeling
-dat$c_Itm <- ifelse(dat$Item==1, -.75, 
-                    ifelse(dat$Item==2, -.25,
-                           ifelse(dat$Item==3, .25, .75)))
-
-# make Item be a factor and assign labels
-dat$Item <- factor(dat$Item, levels=c(1,2,3,4), labels=c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
-
-# add ratios for Item
-item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54) 
-# 0.2500000 0.4705882 0.5909091 0.6666667
-item_range_ratio_scaled = as.vector(scale(c(6 / 24, 16 / 34, 26 / 44, 36 / 54)))
-# -1.3441995 -0.1316642  0.5297187  0.9461450
-item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))) 
-# 0.5125000 0.6876471 0.7691558 0.8166667
-item_mean_ratio_scaled = as.vector(scale(c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))))) 
-# -1.37582241 -0.06614191  0.54334858 0.89861574
-
-dat[dat$Item == "06:15:24", "item_range_ratio"] <-  0.2500000
-dat[dat$Item == "16:25:34", "item_range_ratio"] <-  0.4705882
-dat[dat$Item == "26:35:44", "item_range_ratio"] <-  0.5909091
-dat[dat$Item == "36:45:54", "item_range_ratio"] <-  0.6666667
-
-dat[dat$Item == "06:15:24", "item_range_ratio_scaled"] <-  -1.3441995
-dat[dat$Item == "16:25:34", "item_range_ratio_scaled"] <-  -0.1316642
-dat[dat$Item == "26:35:44", "item_range_ratio_scaled"] <-   0.5297187  
-dat[dat$Item == "36:45:54", "item_range_ratio_scaled"] <-   0.9461450
-
-dat[dat$Item == "06:15:24", "item_mean_ratio"] <-  0.5125000 
-dat[dat$Item == "16:25:34", "item_mean_ratio"] <-  0.6876471 
-dat[dat$Item == "26:35:44", "item_mean_ratio"] <-  0.7691558
-dat[dat$Item == "36:45:54", "item_mean_ratio"] <-  0.8166667
-
-dat[dat$Item == "06:15:24", "item_mean_ratio_scaled"] <-   -1.37582241 
-dat[dat$Item == "16:25:34", "item_mean_ratio_scaled"] <-   -0.06614191 
-dat[dat$Item == "26:35:44", "item_mean_ratio_scaled"] <-    0.54334858 
-dat[dat$Item == "36:45:54", "item_mean_ratio_scaled"] <-    0.89861574
-
-# VAGUENESS
-# Create a factor coding for Vagueness
-dat[ dat$Condition==1 , 'Vagueness'] <- 'Vague'
-dat[ dat$Condition==2 , 'Vagueness'] <- 'Crisp'
-dat[ dat$Condition==3 , 'Vagueness'] <- 'Vague'
-dat[ dat$Condition==4 , 'Vagueness'] <- 'Crisp'
-dat$Vagueness <- as.factor(dat$Vagueness)
-
-# manually center Vagueness
-dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)
-
-dat$Selection = ""
-dat[dat$Condition %in% c(1,2), "Selection"] <- "matching"
-dat[dat$Condition %in% c(3,4), "Selection"] <- "comparison"
-dat$Selection.sum <- ifelse(dat$Selection=="matching", .5, -.5)
-dat$Selection <- as.factor(dat$Selection)
-
-# remove one zero RT by replacing with NA
-dat$RT[dat$RT==0]<-NA
-
-# ORDER
-# give the levels of Order meaningful names
-dat$Order <- factor(dat$Order, levels=c(1,2), labels=c('LtoR','RtoL'))
-
-# make a manually centred Order
-dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)
-
-# QUANTITY
-# give the levels of Quantity meaningful names
-dat$Quantity <- factor(dat$Quantity, levels=c(1,2), labels=c('Small','Large'))
-
-# make a manually centred Quantity
-dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)
-
-# INSTRUCTION
-# add number of characters in the instruction # 29 30 34 36 38
-dat$nchar_instr = nchar(as.character(dat$Instruction))
-dat$nchar_instr_scaled = as.vector(scale(nchar(as.character(dat$Instruction)), scale=TRUE))
-
-# make Instruction be a factor (17 levels)
-dat$Instruction <- as.factor(dat$Instruction)
-
-# RT
-# add transformations of RT
-dat$RT_RecSqd    <- 1/dat$RT^2
-dat$RT_Rec       <- 1/dat$RT
-dat$RT_RecSqt    <- 1/sqrt(dat$RT) 
-dat$RT_log       <- log(dat$RT)
-dat$RT_sqt       <- sqrt(dat$RT)
-dat$RT_raw       <- dat$RT
-dat$RT_sqd       <- dat$RT^2
-
-# save dat as dat.Rda
-# This data set contains *all* trials including impossible trials and is mainly for graphs comparing different removals
-save(dat, file="dat.Rda")
-
- # lose impossible trials
-  dd <- dat
-  dd$RT[dd$RT == 1 ] <- NA
-  dd$RT[dd$RT > 40000 ] <- NA
-  dd <- dd[complete.cases(dd),]
-  row.names(dd) <- NULL
-
-# add preceding RT: because we removed impossible trials, the value for preceding RT for a trial following an impossible trial is the value of the trial that preceded the impossible trial.
-dd$RTprev <- NA
-for (s in levels(dd$Subject)) {
-  nrows = nrow(dd[dd$Subject==s,])
-  for (i in 1:nrows) {
-    if (i==1) {
-      dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]
-    }
-    else
-      dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1]
-  }
-}
-
-# add transformations of previous RT
-dd$RTprev_RecSqd    <- 1/dd$RTprev^2
-dd$RTprev_Rec       <- 1/dd$RTprev
-dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev) 
-dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
-dd$RTprev_log       <- log(dd$RTprev)
-dd$RTprev_sqt       <- sqrt(dd$RTprev)
-dd$RTprev_raw       <- dd$RTprev
-dd$RTprev_sqd       <- dd$RTprev^2
-
-# save dd as dd.Rda
-# This data set non-impossible trials and renamed/recomputed variables
-save(dd, file="dd.Rda")
+```{r showItems, echo=FALSE, cache=TRUE}
+items = c(1,2,3,4,5,6,7,8)
+i = data.frame(Item = c("1","2","3","4"),
+               x = c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
+names(i) = c('Item', 'Numbers of dots')
+pander(i, justify=c('center', 'center'))
 ```
 
-Check the distribution of the trimmed data to see if a transformation is suggested. Log transformation seems to be best so we adopt that for the rest of the analyses.
-
-```{r boxplot subjects and items}
-load("dd.Rda")
-# allow reference to transformation
-subdata = melt(dd,
-               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
-               variable.name="transformation",
-               value.name="value") 
-# boxplots for subjects and items separately for each dependent variable
-ggplot(subdata) + 
-  facet_wrap(~ variable, scales="free", ncol=4) +
-  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
-  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
-  theme_bw() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
-  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
-  xlab("") + ylab("")
+```{r getData, echo=T}
+rawdata <- read.table('data.txt', sep=' ')
 ```
 
-```{r load_dd}
-load("dd.Rda")
+```{r trimat25sd}
+trimdata <- perSubjectTrim.fnc(rawdata, response='RT', subject='Subject', trim = 2.5)$data
+rtdata <- trimdata
+```
+```{r computeSummariesForGraphs, cache=TRUE, message=FALSE}
+rtdataplot1 <- summarySEwithin2(rtdata, measurevar="RT_log", withinvars=c("Condition", "Vagueness", "Selection", "Item"), idvar="Subject")
 ```
 
-```{r do_lmer}
-v5 <- lmer(data=dd, 
-           RT_RecSqt ~ 
-             c_Vag + Selection.sum + c_Qty + c_Ord + 
-             Selection.sum:c_Vag:c_Qty +
-             item_mean_ratio + 
-             s_Trl +
-             RTprev_RecSqt +
-             nchar_instr +
-             (1+c_Vag + Selection.sum + c_Qty + c_Ord|Subject)
-)
+```{r rtplot, fig.width=7, fig.height=3, echo=F}
+dodge = position_dodge(width=0.2)
+rtdataplot1$Condition <- as.factor(paste(sep=' ', rtdataplot1$Selection, rtdataplot1$Vagueness))
+rtplot=ggplot(rtdataplot1, aes(y=RT_logNormed, x=Item, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci,
+                group=Condition, shape=Condition, fill=Condition)) +
+  geom_line(position=dodge) +
+  geom_errorbar(width=.2, position=dodge) +
+  geom_point(size=2, position=dodge) +
+  scale_shape_manual("",values = c(22, 22, 21, 21)) +
+  scale_fill_manual("",values=c("black","white","black","white")) +
+  ggtitle("Response time") +
+  ylab("Respone time (log (ms))") + 
+  xlab("") +
+  theme_bw() +
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        legend.key = element_blank(), aspect.ratio=1,
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        plot.background=element_rect(fill=NA, color='white')) +
+  facet_grid(~Selection) 
+rtplot
 ```
 
-```{r summary_lmer}
-summary(v5)
+```{r rtlmer, cache=TRUE}
+rtlmer = lmerTest::lmer(data=rtdata,
+                        RT_log ~ c_Vag * c_Sel + c_Itm +
+                          (1 + c_Vag * c_Sel + c_Itm | Subject))
 ```
 
-```{r R2}
-cat("R^2")
-cor(fitted(v5), dd$RT_RecSqt)^2
+```{r rtlmersummaryxtable, results='asis', cache=TRUE, echo=F}
+print(xtable(summary(rtlmer)$coef, digits=c(0,2,2,1,1,4)),type='html')
 ```
 
-```{r plotLMER}
-par(mfrow=c(2,5))
-plotLMER.fnc(v5)
+```{r seplmersforcomparison, results='asis', cache=TRUE, echo=-3}
+s1 <- subset(rtdata, subset=Selection=='Comparison')
+s1lm <- lmerTest::lmer(data=s1, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
+print(xtable(summary(s1lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
 ```
 
-Plot model coefficients and ci's
-
-```{r coeff}
-par(las=2, mar=c(14,6,1,1))
-y=as.data.frame(summary(v5)$coef[-1,1]) # estimates, without intercept
-q=confint(v5)[-c(1,2,3),]
-y=cbind(y,q)
-plot(y[,1], ylim=range(y[,1],y[,2],y[,3]), xaxt="n", xlab="", ylab="estimate\n",  pch='-', cex=2)
-segments(x0=1:nrow(y), x1=1:nrow(y), y[,2], y[,3])
-abline(h=0)
-axis(1, labels=row.names(y), at=1:nrow(y), srt=45)
-grid()
+```{r seplmersformatching, results='asis', cache=TRUE, echo=-3}
+s2 <- subset(rtdata, subset=Selection=='Matching')
+s2lm <- lmerTest::lmer(data=s2, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
+print(xtable(summary(s2lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
 ```
-
-Model criticism baayen plots
-
-```{r baayenplots, fig.width=9, fig.height=3}
-# Baayen 4-plot model criticism
-
-par(mfrow=c(1,4), pty='s')
-# create scaled residuals
-v5$rstand = as.vector(scale(resid(v5)))
-# plot scaled residuals density
-plot(density(v5$rstand))
-# plot sample quantiles versus theoretical quantiles
-qqnorm(v5$rstand, cex=.5)
-qqline(v5$rstand)
-# plot standardised residuals versus fitted values
-plot(v5$rstand ~ fitted(v5), pch='.')
-# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
-abline(h=c(-2.5,2.5))
-# create dffits
-dffits=abs(resid(v5, "dffits"))
-# plot dffits
-plot(dffits, type='h')
-```

### 2016-09-27 experiments/e3/writeup/preprocessing/preprocessing.Rmd
e3 writeup less verbose

--- 2016-07-28 experiments/e3/writeup/experiment3.Rmd
@@
----
-title: "Experiment 3 of 4"
-output:
-  html_document:
-    toc: true
-    <!--toc_float: true-->
-    mathjax: default
-    template: default
-    code_folding: hide
-    theme: default
----  
-# Experiment 3
-
-```{r docOpts}
-library(knitr)
-```
-
-```{r loadLibraries}
-suppressPackageStartupMessages(library(MASS))
-suppressPackageStartupMessages(library(ggplot2))
-suppressPackageStartupMessages(library(lme4))
-suppressPackageStartupMessages(library(xtable))
-suppressPackageStartupMessages(library(plyr))
-suppressPackageStartupMessages(library(reshape))
-suppressPackageStartupMessages(library(lattice))
-suppressPackageStartupMessages(library(LMERConvenienceFunctions))
-suppressPackageStartupMessages(library(car))
-suppressPackageStartupMessages(library(grid))
-suppressPackageStartupMessages(library(lmerTest))
-suppressPackageStartupMessages(library(reshape2))
-suppressPackageStartupMessages(library(ggplot2))
-suppressPackageStartupMessages(require(languageR))
-```
-
-# Experiment Structure
-Each subject saw 192 trials, being 3 cycles of 64 trials. There were 38 participants. There were 7296 rows in the full data set.
-
-# Trial Procedure
-The procedure for each trial was as follows.
-First an instruction that was a referring expression was presented until participants dismissed it with a key press.
-Then the squares with dots were presented until the participant responded with a key press, with a 60 second timeout if they did not respond.
-
-# Prepare the data
-Source additional R files with functions for computing confidence intervals.
-
-```{r sourceFunctions}
-source("functions/CIfunctions.R")
-```
-
-Get data.
-
 ```{r readdata}
-dat <- read.delim("data/allresults.txt", stringsAsFactors = FALSE)
+dat <- read.delim("../rawdata.txt", stringsAsFactors = FALSE)
 
@@
 # sort out borderline responses into expected near and far
-# How many squares in the square they chose? 
 dat$RESPONSE <- as.character(dat$RESPONSE)
@@
 dat$Selection = ""
-dat[dat$Condition %in% c(1,2), "Selection"] <- "matching"
-dat[dat$Condition %in% c(3,4), "Selection"] <- "comparison"
-dat$Selection.sum <- ifelse(dat$Selection=="matching", .5, -.5)
+dat[dat$Condition %in% c(3,4), "Selection"] <- "Comparison"
+dat[dat$Condition %in% c(1,2), "Selection"] <- "Matching"
+dat$c_Sel <- ifelse(dat$Selection=="Comparison", -.5, .5)
 dat$Selection <- as.factor(dat$Selection)
@@
 
-# save dd as dd.Rda
+# save dd as data.Rda
 # This data set non-impossible trials and renamed/recomputed variables
-save(dd, file="dd.Rda")
+write.table(dd, file="../data.txt")
 ```
-
-Check the distribution of the trimmed data to see if a transformation is suggested. Log transformation seems to be best so we adopt that for the rest of the analyses.
-
-```{r boxplot subjects and items}
-load("dd.Rda")
-# allow reference to transformation
-subdata = melt(dd,
-               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
-               variable.name="transformation",
-               value.name="value") 
-# boxplots for subjects and items separately for each dependent variable
-ggplot(subdata) + 
-  facet_wrap(~ variable, scales="free", ncol=4) +
-  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
-  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
-  theme_bw() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
-  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
-  xlab("") + ylab("")
-```
-
-```{r load_dd}
-load("dd.Rda")
-```
-
-```{r do_lmer}
-v5 <- lmer(data=dd, 
-           RT_RecSqt ~ 
-             c_Vag + Selection.sum + c_Qty + c_Ord + 
-             Selection.sum:c_Vag:c_Qty +
-             item_mean_ratio + 
-             s_Trl +
-             RTprev_RecSqt +
-             nchar_instr +
-             (1+c_Vag + Selection.sum + c_Qty + c_Ord|Subject)
-)
-```
-
-```{r summary_lmer}
-summary(v5)
-```
-
-```{r R2}
-cat("R^2")
-cor(fitted(v5), dd$RT_RecSqt)^2
-```
-
-```{r plotLMER}
-par(mfrow=c(2,5))
-plotLMER.fnc(v5)
-```
-
-Plot model coefficients and ci's
-
-```{r coeff}
-par(las=2, mar=c(14,6,1,1))
-y=as.data.frame(summary(v5)$coef[-1,1]) # estimates, without intercept
-q=confint(v5)[-c(1,2,3),]
-y=cbind(y,q)
-plot(y[,1], ylim=range(y[,1],y[,2],y[,3]), xaxt="n", xlab="", ylab="estimate\n",  pch='-', cex=2)
-segments(x0=1:nrow(y), x1=1:nrow(y), y[,2], y[,3])
-abline(h=0)
-axis(1, labels=row.names(y), at=1:nrow(y), srt=45)
-grid()
-```
-
-Model criticism baayen plots
-
-```{r baayenplots, fig.width=9, fig.height=3}
-# Baayen 4-plot model criticism
-
-par(mfrow=c(1,4), pty='s')
-# create scaled residuals
-v5$rstand = as.vector(scale(resid(v5)))
-# plot scaled residuals density
-plot(density(v5$rstand))
-# plot sample quantiles versus theoretical quantiles
-qqnorm(v5$rstand, cex=.5)
-qqline(v5$rstand)
-# plot standardised residuals versus fitted values
-plot(v5$rstand ~ fitted(v5), pch='.')
-# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
-abline(h=c(-2.5,2.5))
-# create dffits
-dffits=abs(resid(v5, "dffits"))
-# plot dffits
-plot(dffits, type='h')
-```

### 2016-09-27 experiments/e3/writeup/scratch.Rmd
e3 writeup less verbose

```{r boxplotsubjectsanditems}
dd = read.table("data.txt")
# allow reference to transformation
subdata = melt(dd,
               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
               variable.name="transformation",
               value.name="value") 
# boxplots for subjects and items separately for each dependent variable
foo=ggplot(subdata) + 
  facet_wrap(~ transformation, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
  xlab("") + ylab("")
```

```{r do_lmer}
v5 <- lmer(data=dd, 
           RT_RecSqt ~ 
             c_Vag + Selection.sum + c_Qty + c_Ord + 
             Selection.sum:c_Vag:c_Qty +
             item_mean_ratio + 
             s_Trl +
             RTprev_RecSqt +
             nchar_instr +
             (1+c_Vag + Selection.sum + c_Qty + c_Ord|Subject)
)
```

```{r summary_lmer}
summary(v5)
```

```{r R2}
cat("R^2")
cor(fitted(v5), dd$RT_RecSqt)^2
```

```{r plotLMER}
par(mfrow=c(2,5))
plotLMER.fnc(v5)
```

Plot model coefficients and ci's

```{r coeff}
v5 <- lme4::lmer(data=dd, 
RT_RecSqt ~ 
c_Vag + Selection.sum + c_Qty + c_Ord + 
Selection.sum:c_Vag:c_Qty +
item_mean_ratio + 
s_Trl +
RTprev_RecSqt +
nchar_instr +
(1+c_Vag + Selection.sum + c_Qty + c_Ord|Subject)
)
par(las=2, mar=c(6,6,1,1))
y=data.frame(coef=summary(v5)$coef[-1,1]) # estimates, without intercept
q=confint(v5, method="Wald")[-c(1,2,3),]
y=cbind(y,q[15:23,])
plot(y[,1], ylim=range(y[,1],y[,2],y[,3]), xaxt="n", xlab="", ylab="estimate\n",  pch='-', cex=2)
segments(x0=1:nrow(y), x1=1:nrow(y), y[,2], y[,3])
abline(h=0)
axis(1, labels = abbreviate(row.names(y), minlength=10), , at=1:nrow(y), las=2)
grid()
```

Model criticism baayen plots

```{r baayenplots, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

par(mfrow=c(1,4), pty='s')
# create scaled residuals
dd$rstand = as.vector(scale(resid(v5)))
# plot scaled residuals density
plot(density(dd$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(dd$rstand, cex=.5)
qqline(dd$rstand)
# plot standardised residuals versus fitted values
plot(dd$rstand ~ fitted(v5), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))
# create dffits
#dffits=abs(resid(v5, "dffits"))
# plot dffits
#plot(dffits, type='h')
```

### 2016-09-27 papers/2016-Special-Issue/v-special-issue-2016.tex
finished stats for e3

@@
 \subsection{Hypotheses}
-(H1) Vague instructions are easier for the reader than crisp alternatives (main effect of vagueness)
+\begin{APAenumerate}
+	\item [(H1)] Vague instructions are easier for the reader than crisp alternatives (main effect of vagueness)
+	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection task)
+	\item [(H3)] Effects of vagueness are different depending on whether selection is matching or comparison (interaction effect selection x vagueness).
+\end{APAenumerate}
 
-(H2) Comparison is easier for the reader than matching (main effect of selection task)
-
-(H3) Effects of vagueness are different depending on whether selection is matching or comparison (interaction effect selection x vagueness).
-
-%The instruction format account predicts no differences between the conditions, since all conditions used numeric quantifiers.
-%
-%The cost reduction account predicts that there will be a main effect of vagueness such that the vague instruction conditions attract faster responses than the crisp instruction conditions and particularly that at each level of selection task the vague condition should attract faster responses than the crisp condition.
-
-\subsection{Method}
+\subsection{Method} % E 3
 
@@
 
-\subsection{Results}
-A linear mixed effects regression model was built for log response times. 
-The structure of the model was as follows: fixed effects were vagueness, selection task (both sum coded) and centred item and their interactions: random effects were vagueness, selection task, and item (but not their interactions - the model failed to converge when these interactions were included).
-The means are plotted in Figure \ref{resultse3}.
+\subsection{Results} % E 3
+RTs were trimmed at 2.5 sd separately for each subject, leading to the loss of 204 trials (2.8\% of the trials).
+Condition means for the remaining (logged) RTs are plotted in Figure \ref{resultse3}.
+A linear mixed model was constructed for the logged response times, 
+with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
 
@@
 \centering
-\includegraphics[width=.75\textwidth]{images/response-time-two-panels-1.pdf}
-\caption{Results for Experiment 3}
+\includegraphics[trim = 0mm 30mm 0mm 10mm, clip, width=.75\textwidth]{images/rtplote3}
+\caption{Mean response times by condition for Experiment 3 where all instructions were numeric}
 \label{resultse3}
@@
 
-% MG has not updated these stats yet at 26 Sept 2016
-(H1) There was no significant main effect of vagueness ($\beta =.003$, $se=.014$, $t=.202$, $p=.841$). 
-The results showed that vagueness was beneficial for comparison but detrimental for matching. 
-(H2) There was a main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=-.165, se=.027, t=-6.218, p<.001$). 
-(H3) Vagueness exerted effects in different directions for the comparison task and for the matching task. 
-Separate analyses were conducted at each level of the selection task to see whether within each task type there were significant effects of vagueness. 
-There were: in the comparison task vagueness significantly speeded response times compared with crisp controls ($\beta=-.07, se=.02, t=3.52, p<.01$). 
-In the matching task vagueness significantly slowed response times compared with crisp controls ($\beta=-.07, se=.02, t=-2.89, p<.05$).
+% E3
+\begin{APAenumerate}
+	\item [(H1)] Vague instructions were non-significantly easier for the reader than crisp instructions when vagueness was considered as a main effect ($\beta=-0.01$, $se=0.01$, $t=-0.4$, $p=0.7158$).
+% MG added the values for these stats on 26 Sept 2016
+	\item [(H2)] There was a statistically significant main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=0.16$, $se=0.03$, $t=6.2$, $p<0.0001$).
+% MG added the values for these stats on 26 Sept 2016
+	\item [(H3)] Vagueness exerted effects in different directions for the comparison task and for the matching task: there was a significant interaction between vagueness and selection task ($\beta=0.13$, $se=0.03$, $t=4.2$, $p<0.0001$).
+% MG added the values for these stats on 26 Sept 2016
+%				
+Separate analyses were conducted testing for effects of vagueness at each level of the selection task, using Bonferroni significance thresholds to correct for multiple comparisons. 
+Within the comparison task vagueness significantly speeded response times compared with crisp controls ($\beta=-0.07$, $se=0.02$, $t=-3.5$, $p<0.0012$).
+% MG added the values for these stats on 26 Sept 2016
+Within the matching task vagueness significantly \emph{slowed} response times compared with crisp controls ($\beta=0.06$, $se=0.02$, $t=2.9$, $p<0.0061$).
+% MG added the values for these stats on 26 Sept 2016
+\end{APAenumerate}
 
-% KvD: deleted the following:
-%None of the accounts set out in the Hypotheses section emerge well from the results. 
-%The instruction format account wrongly predicts no differences between the conditions. 
-%The selection task correctly predicted the main effect of selection task, but has no coverage of the interaction with vagueness.
-
-The cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
+% MG edited this 27 Sept 2016:
+%The cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
+The cost reduction account was wrong to predict significant main effect advantages for vagueness (although there was a non-significant trend in the \emph{direction} predicted by the cost reduction account), and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was significantly advantageous in the comparison task.
 
@@
 
-\clearpage
-\appendix
+\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}
 
-\begin{sidewaystable}
-\centering
-\caption{All Instructions}
-\label{All Instructions}
-\begin{tabular}{lp{5cm}llllp{2cm}l}
-\midrule
-Experiment&Example instruction&vagueness&quantifier type&symmetry&selection&number of valid targets&definiteness\\
-\midrule
-1&Choose the square with 5 dots&crisp&numeric&symmetric&matching&1&definite\\
-1&Choose the square with few dots&vague&linguistic&asymmetric&comparison&1&definite\\
-\midrule
-2&Choose the square with 6 dots&crisp&numeric&symmetric&matching&1&definite\\
-2&Choose a square with about 10 dots&vague&numeric&symmetric&matching&2&indefinite\\
-2&Choose the square with the fewest dots&crisp&linguistic&asymmetric&comparison&1&definite\\
-2&Choose a square with few dots&vague&linguistic&asymmetric&comparison&2&indefinite\\
-\midrule
-3&Choose a square with 6 dots&crisp&numeric&symmetric&matching&1&indefinite\\
-3&Choose a square with fewer than 20 dots&crisp&numeric&asymmetric&comparison&2&indefinite\\
-3&Choose a square with about 10 dots&vague&numeric&symmetric&matching&2&indefinite\\
-3&Choose a square with far fewer than 20 dots&vague&numeric&asymmetric&comparison&1&indefinite\\
-\midrule
-4&Choose a square with the same number of dots as the target&crisp&linguistic&symmetric&matching&1&indefinite\\
-4&Choose a square with fewer dots than the target&crisp&linguistic&asymmetric&comparison&2&indefinite\\
-4&Choose a square with about the same number of dots as the target&vague&linguistic&symmetric&matching&2&indefinite\\
-4&Choose a square with far fewer dots than the target&vague&linguistic&asymmetric&comparison&1&indefinite\\
-\midrule
-\end{tabular}
-\end{sidewaystable}
+\begin{APAenumerate}
+	\item 
+	\item 
+\end{APAenumerate}
 
-\clearpage
-
-\begin{table}[htbp]
-\caption{Table of instructions for the pair (5,25). Experiment 1}
-\label{instructionse1}
-\begin{tabular}{rl}
-\toprule
-vagueness&example\\
-\midrule
-crisp 	& 	Choose the square with 5 dots \\
-vague	&	Choose the square with few dots\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{figure}[tbp]
-\fitfigure{images/stimuluse1}
-\caption{An example stimulus from Experiment 1.
+%\clearpage
+%\appendix
+%
+%\begin{sidewaystable}
+%\centering
+%\caption{All Instructions}
+%\label{All Instructions}
+%\begin{tabular}{lp{5cm}llllp{2cm}l}
+%\midrule
+%Experiment&Example instruction&vagueness&quantifier type&symmetry&selection&number of valid targets&definiteness\\
+%\midrule
+%1&Choose the square with 5 dots&crisp&numeric&symmetric&matching&1&definite\\
+%1&Choose the square with few dots&vague&linguistic&asymmetric&comparison&1&definite\\
+%\midrule
+%2&Choose the square with 6 dots&crisp&numeric&symmetric&matching&1&definite\\
+%2&Choose a square with about 10 dots&vague&numeric&symmetric&matching&2&indefinite\\
+%2&Choose the square with the fewest dots&crisp&linguistic&asymmetric&comparison&1&definite\\
+%2&Choose a square with few dots&vague&linguistic&asymmetric&comparison&2&indefinite\\
+%\midrule
+%3&Choose a square with 6 dots&crisp&numeric&symmetric&matching&1&indefinite\\
+%3&Choose a square with fewer than 20 dots&crisp&numeric&asymmetric&comparison&2&indefinite\\
+%3&Choose a square with about 10 dots&vague&numeric&symmetric&matching&2&indefinite\\
+%3&Choose a square with far fewer than 20 dots&vague&numeric&asymmetric&comparison&1&indefinite\\
+%\midrule
+%4&Choose a square with the same number of dots as the target&crisp&linguistic&symmetric&matching&1&indefinite\\
+%4&Choose a square with fewer dots than the target&crisp&linguistic&asymmetric&comparison&2&indefinite\\
+%4&Choose a square with about the same number of dots as the target&vague&linguistic&symmetric&matching&2&indefinite\\
+%4&Choose a square with far fewer dots than the target&vague&linguistic&asymmetric&comparison&1&indefinite\\
+%\midrule
+%\end{tabular}
+%\end{sidewaystable}
+%
+%\clearpage
+%
+%\begin{table}[htbp]
+%\caption{Table of instructions for the pair (5,25). Experiment 1}
+%\label{instructionse1}
+%\begin{tabular}{rl}
+%\toprule
+%vagueness&example\\
+%\midrule
+%crisp 	& 	Choose the square with 5 dots \\
+%vague	&	Choose the square with few dots\\
+%\bottomrule
+%\end{tabular}
+%\end{table}
+%
+%\begin{figure}[tbp]
+%\fitfigure{images/stimuluse1}
+%\caption{An example stimulus from Experiment 1.
 First the referring expression was presented (left panel).
 Then after a keypress, and a fixation cross (not pictured) the squares and dots were presented without repetition of the referring expression (right panel)}
-\label{stimuluse1}
-\end{figure}
-
-\begin{figure}[htbp]
-\fitfigure{images/resultse1.pdf}
-\caption{Experiment 1 results. }
-\label{resultse1}
-\end{figure}
-
-\begin{table}[htbp]
-\caption{Table of instructions arranged by condition for the triple (6,15,24). Experiment 2}
-\label{instructionse2}
-\begin{tabular}{rlll}
-\toprule
-vagueness&instruction format&selection task&example\\
-\midrule
-vague	&	numeric 	& matching	&	Choose a square with about 10 dots\\
-vague	&	verbal	& comparison	&	Choose a square with few dots\\
-crisp 	& 	numeric	& matching	&	Choose the square with 6 dots \\
-crisp		&	verbal	& comparison	&	Choose the square with the fewest dots\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{figure}[tbp]
-\centering
-\includegraphics[width=.4\textwidth]{images/stimuluse2}
-\caption{An example stimulus from Experiment 2}
-\label{stimuluse2}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.65\textwidth]{images/response-time-sep-1.pdf}
-\caption{Response time results for Experiment 2}
-\label{resultse2}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[]{images/histbd4-ggplot-version-1.pdf}
-\caption{Expt 2: Distribution of Nearest match, borderline case, and furthest match responses}
-\label{borderline-response-distribution-e2}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.5\textwidth]{images/response-time-two-panels-1.pdf}
-\caption{Results for Experiment 3}
-\label{resultse3}
-\end{figure}
-
-\begin{table}[htbp]
-\centering
-\caption{Experiment 3: Instructions, assuming $6,15,24$ dots as the Item, and showing \emph{fewer} instead of \emph{more}}
-\label{Instructions for e3}
-\begin{tabular}{llll}
-\hline
-vagueness&instruction format&selection task&instruction\\
-\hline
-crisp & numeric&matching & Choose a square with 6 dots \\ 
-crisp & numeric&comparison & Choose a square with fewer than 20 dots \\
-vague & numeric&matching & Choose a square with about 10 dots \\ 
-vague & numeric&comparison & Choose a square with far fewer than 20 dots \\ 
-\hline
-\end{tabular}
-\end{table}%
-
-\begin{table*}[htbp]
-\centering
-\caption{Experiment 4: Instructions}
-\label{Instructions for e4}
-\begin{tabular}{lllp{7cm}}
-\hline
-vagueness&instruction format& selection task&instruction\\
-\hline
-crisp & verbal&matching & Choose a square with the same number of dots as the target \\ 
-crisp & verbal&comparison& Choose a square with fewer dots than the target \\
-vague & verbal&matching & Choose a square with about the same number of dots as the target \\ 
-vague & verbal&comparison& Choose a square with far fewer dots than the target \\ 
-\hline
-\end{tabular}
-\end{table*}%
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.5\textwidth]{images/response-time-two-panels-1-e4.pdf}
-\caption{Results for Experiment 4}
-\label{resultse4}
-\end{figure}
-crisp  numeric matching  Choose a square with 6 dots
-
-crisp  numeric comparison  Choose a square with fewer than 20 dots
-
-vague  numeric matching  Choose a square with about 10 dots
-
-vague  numeric comparison  Choose a square with far fewer than 20 dots
-
-crisp  verbal matching  Choose a square with the same number of dots as the target
-
-crisp  verbal comparison Choose a square with fewer dots than the target
-
-vague  verbal matching  Choose a square with about the same number of dots as the target
-
-vague  verbal comparison Choose a square with far fewer dots than the target
+%\label{stimuluse1}
+%\end{figure}
+%
+%\begin{figure}[htbp]
+%\fitfigure{images/resultse1.pdf}
+%\caption{Experiment 1 results. }
+%\label{resultse1}
+%\end{figure}
+%
+%\begin{table}[htbp]
+%\caption{Table of instructions arranged by condition for the triple (6,15,24). Experiment 2}
+%\label{instructionse2}
+%\begin{tabular}{rlll}
+%\toprule
+%vagueness&instruction format&selection task&example\\
+%\midrule
+%vague	&	numeric 	& matching	&	Choose a square with about 10 dots\\
+%vague	&	verbal	& comparison	&	Choose a square with few dots\\
+%crisp 	& 	numeric	& matching	&	Choose the square with 6 dots \\
+%crisp		&	verbal	& comparison	&	Choose the square with the fewest dots\\
+%\bottomrule
+%\end{tabular}
+%\end{table}
+%
+%\begin{figure}[tbp]
+%\centering
+%\includegraphics[width=.4\textwidth]{images/stimuluse2}
+%\caption{An example stimulus from Experiment 2}
+%\label{stimuluse2}
+%\end{figure}
+%
+%
+%\begin{figure}[htbp]
+%\centering
+%\includegraphics[width=.65\textwidth]{images/response-time-sep-1.pdf}
+%\caption{Response time results for Experiment 2}
+%\label{resultse2}
+%\end{figure}
+%
+%\begin{figure}[htbp]
+%\centering
+%\includegraphics[]{images/histbd4-ggplot-version-1.pdf}
+%\caption{Expt 2: Distribution of Nearest match, borderline case, and furthest match responses}
+%\label{borderline-response-distribution-e2}
+%\end{figure}
+%
+%\begin{figure}[htbp]
+%\centering
+%\includegraphics[width=.5\textwidth]{images/response-time-two-panels-1.pdf}
+%\caption{Results for Experiment 3}
+%\label{resultse3}
+%\end{figure}
+%
+%\begin{table}[htbp]
+%\centering
+%\caption{Experiment 3: Instructions, assuming $6,15,24$ dots as the Item, and showing \emph{fewer} instead of \emph{more}}
+%\label{Instructions for e3}
+%\begin{tabular}{llll}
+%\hline
+%vagueness&instruction format&selection task&instruction\\
+%\hline
+%crisp & numeric&matching & Choose a square with 6 dots \\ 
+%crisp & numeric&comparison & Choose a square with fewer than 20 dots \\
+%vague & numeric&matching & Choose a square with about 10 dots \\ 
+%vague & numeric&comparison & Choose a square with far fewer than 20 dots \\ 
+%\hline
+%\end{tabular}
+%\end{table}%
+%
+%
+%\begin{table*}[htbp]
+%\centering
+%\caption{Experiment 4: Instructions}
+%\label{Instructions for e4}
+%\begin{tabular}{lllp{7cm}}
+%\hline
+%vagueness&instruction format& selection task&instruction\\
+%\hline
+%crisp & verbal&matching & Choose a square with the same number of dots as the target \\ 
+%crisp & verbal&comparison& Choose a square with fewer dots than the target \\
+%vague & verbal&matching & Choose a square with about the same number of dots as the target \\ 
+%vague & verbal&comparison& Choose a square with far fewer dots than the target \\ 
+%\hline
+%\end{tabular}
+%\end{table*}%
+%
+%\begin{figure}[htbp]
+%\centering
+%\includegraphics[width=.5\textwidth]{images/response-time-two-panels-1-e4.pdf}
+%\caption{Results for Experiment 4}
+%\label{resultse4}
+%\end{figure}
+%crisp  numeric matching  Choose a square with 6 dots 
+%
+%crisp  numeric comparison  Choose a square with fewer than 20 dots 
+%
+%vague  numeric matching  Choose a square with about 10 dots 
+%
+%vague  numeric comparison  Choose a square with far fewer than 20 dots 
+%
+%crisp  verbal matching  Choose a square with the same number of dots as the target 
+%
+%crisp  verbal comparison Choose a square with fewer dots than the target 
+%
+%vague  verbal matching  Choose a square with about the same number of dots as the target 
+%
+%vague  verbal comparison Choose a square with far fewer dots than the target
 
@@
 
-\begin{APAenumerate}
-	\item 
-	\item 
-\end{APAenumerate}
 %
@@
 
-\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}
-$\beta=??$, $se=??$, $t=??$, $p<??$
-
 %The \emph{instruction format} and \emph{vagueness} variables were sum-coded and \emph{Item} was centred.

### 2016-09-27 papers/2016-Special-Issue/v-special-issue-2016.tex
removed some dud images

@@
 \centering
-\includegraphics[width=.49\textwidth]{images/rtplote2.pdf}
-\includegraphics[trim = 0mm -30mm 0mm 0mm, clip, width=.49\textwidth]{images/blplote2.pdf}
+\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=.49\textwidth]{images/e2rt.pdf}
+\includegraphics[trim = 0mm -23mm 0mm 0mm, clip, width=.49\textwidth]{images/e2bl.pdf}
 \caption{Experiment 2 results: mean response times by condition and item, and counts of borderline case responses by condition.}
@@
 \begin{tabular}{llll}
-vagueness	&	instruction		&	selection	&	instruction\\
-			&	format		&	task		&			\\
+instruction				&vagueness			&	selection		&	instruction							\\
+format				&					&	task			&									\\
 \toprule
-\multirow{ 2}{*}{crisp} 	& \multirow{ 2}{*}{numeric}	&	matching		& Choose a square with 6 dots \\ 
+\multirow{4}{*}{numeric}	&\multirow{2}{*}{crisp} 	&	matching		& Choose a square with 6 dots 				\\ 
 					&					&	comparison 	& Choose a square with fewer than 20 dots 	\\
-					\midrule
-\multirow{ 2}{*}{vague} 	& \multirow{ 2}{*}{numeric} 	&	matching 		& Choose a square with about 10 dots \\ 
+					\cline{2-4}
+					&\multirow{2}{*}{vague} 	&	matching 		& Choose a square with about 10 dots 		\\ 
  					&					&	comparison 	& Choose a square with far fewer than 20 dots 	\\ 
@@
 \end{table}%
+%&	
+%
+%	instruction				&
+%	
+%	
+%	format				&
+%
 
@@
 
-\section{Experiment 4}
+\section{Experiment 4} % E 4
 
@@
 We manipulated vagueness and the selection task (comparison and matching). 
-In order to implement the experiment without mentioning numbers in the instructions, we changed the sequence of each trial to include a 'target' (i.e., a dot array of a particular cardinality) before the instruction, so that we could then refer back to the target's cardinality in the instruction using expressions like \emph{the same number of dots as the target}; \emph{fewer dots than the target}.
-% MG WHAT WERE THE VALUES OF THE TARGETS?
-\textcolor{red}{MG insert the values of the targets}
+In order to implement the experiment without mentioning numbers in the instructions, we changed the sequence of each trial to include a `target' (i.e., a dot array of a particular cardinality) before the instruction, so that we could then refer back to the target's cardinality in the instruction using expressions like \emph{the same number of dots as the target}; \emph{fewer dots than the target}.
+%
 This presentation of a target before the main body of the trial shares some features with \citeA[Experiment 2]{Izard20081221}, although in that experiment participants were told the cardinality of the target (called an \emph{inducer} in that paper) whereas in our experiment we did not tell participants the cardinality of the prime array.
 An item was thus a combination of a target dot array, an instruction that did not contain a number, and a set of dot arrays taking their cardinalities from the same triples used in Experiment 2.
-Table \ref{Instructions for e4} spells out how the instructions were constrained not to mention a numeral.
+Table \ref{Instructions for e4} spells out how the instructions were constrained not to mention a numeral and gives examples of targets.
 
-\begin{table}[htbp]
-\centering
-\caption{Experiment 4: Instructions}
-\label{Instructions for e4}
-\begin{tabular}{llll}
-vagueness	&	instruction		&	selection							&	instruction: Choose a square with \ldots\\
-			&	format		&	task		&			\\
+\begin{table}[htp]
+\caption{Instructions and targets by condition for experiment 4}
+\begin{center}
+\begin{tabular}{llcrl}
+Item					&Selection				&Vagueness	&Target	&Choose a square with\ldots						\\
 \toprule
-\multirow{ 2}{*}{crisp} 	 	& \multirow{ 2}{*}{verbal} 	&	matching 		& \ldots the same number of dots as the target \\ 
-	& 	&	comparison										& \ldots fewer dots than the target \\
-	\midrule
-\multirow{ 2}{*}{vague} 	& \multirow{ 2}{*}{verbal} 	&	matching 			& \ldots about the same number of dots as the target \\ 
-	& 	&	comparison										& \ldots far fewer dots than the target \\ 
+\multirow{4}{*}{06:15:24} 	&\multirow{2}{*}{Comparison} 	&Crisp		&20		&\ldots fewer dots than the target					\\
+\cline{3-5}
+					&						&Vague		&20		&\ldots far fewer dots than the target					\\
+					\cline{2-5}
+					&\multirow{2}{*}{Matching}		& Crisp		&6		&\ldots the same number of dots as the target			\\
+					\cline{3-5}
+					& 		 				&Vague		&10 		&\ldots about the same number of dots as the target		\\
 					\bottomrule
 \end{tabular}
+\end{center}
+\label{Instructions for e4}
 \end{table}%
 
-\subsection{Hypotheses}
-%The instruction format account predicts no differences between the conditions, since all conditions used verbal quantifiers.
-%
-%The selection task account predicts a main effect of selection task such that the comparison conditions would attract faster responses than the matching conditions.
-%
-%The cost reduction account predicts that there will be a main effect of vagueness such that the vague instruction conditions attract faster responses than the crisp instruction conditions and particularly that at each level of selection task the vague condition should attract faster responses than the crisp condition.
+\subsection{Hypotheses}% E 4
+\begin{APAenumerate}
+	\item [(H1)] Vague instructions are easier for the reader than crisp alternatives (main effect of vagueness)
+	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection)
+	\item [(H3)] Effects of vagueness are different depending on whether selection is matching or comparison (interaction effect selection x vagueness, and focussed comparisons at each level of selection).
+\end{APAenumerate}
 
-(H1) Vague instructions are easier for the reader than crisp alternatives (main effect of vagueness)
-
-(H2) Comparison is easier for the reader than matching (main effect of selection)
-
-(H3) Effects of vagueness are different depending on whether selection is matching or comparison (interaction effect selection x vagueness).
-
-%\subsection{Method}
-
-%40 participants were recruited. % via internal messaging at University of Aberdeen, with self-reported fluency in English. They were paid ten pounds for participating. 
-%The apparatus was the same as Experiment 1.
-%The design was a 2 x 2 factorial manipulation of vagueness and selection task.
-%Each stimulus was a sequence of prime, instruction, and squares.
-%The procedure for this experiment was different from the others, to accommodate the requirement not to use numbers in the instruction. 
-%We had to have a different way to indicate a cardinality in the instruction, which we did by adding a visual `target', a square that contained the number of dots that we wanted to refer to.
-
-\subsection{Results}
+\subsection{Results}% E 4
 40 volunteers participated.
@@
 
-% MG has not updated these stats yet at 26 Sept 2016
-(H1) There was no main effect of vagueness ($\beta=.01, se=.01, t=1.51, p=.14$).
-
-(H2) There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=-.18, se=.02, t=-10.38, p<.01$).
+\begin{APAenumerate}
+	\item [(H1)] There was no significant main effect of vagueness ($\beta=-0.02$, $se=0.01$, $t=-1.5$, $p=0.1296$). % MG added the values for these stats on 27 Sept 2016
+	\item [(H2)] There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=0.18$, $se=0.02$, $t=10.4$, $p<0.0001$).
+% MG added the values for these stats on 27 Sept 2016
 	This effect was in the same direction as Experiment 3. 
-
-(H3) Vagueness did exert different effects depending on the selection task ($\beta=.12, se=.03, t=4.32, p<.05$). 
-Separate analyses were conducted for the comparison task and for the matching task. In the comparison task, vagueness resulted in faster response times ($\beta=-0.08, se=.02, t=4.30, p<.05$). 
-In the matching task vagueness slowed response times ($\beta=.05, se=.01, t=3.72, p<.05$). These results are in the same direction as Experiment 3.
+	\item[(H3)] Vagueness did exert different effects depending on the selection task (main interaction effect of vagueness by selection $\beta=0.12$, $se=0.02$, $t=5.1$, $p<0.0001$).
+% MG added the values for these stats on 27 Sept 2016
+Separate analyses of the effect of vagueness were conducted for the comparison task and for the matching task using Bonferroni-adjusted significance thresholds. 
+In the comparison task, vagueness resulted in faster response times ($\beta=-0.08$, $se=0.02$, $t=-4.3$, $p<0.0001$). % MG added the values for these stats on 27 Sept 2016
+In the matching task vagueness slowed response times ($\beta=0.05$, $se=0.01$, $t=3.7$, $p=0.0004$). % MG added the values for these stats on 27 Sept 2016
+These results are in the same direction as Experiment 3.
+\end{APAenumerate}
 
@@
 
-% KvD: deleted:
-%The instruction format account wrongly predicts no differences between the conditions. 
-%The selection task account correctly predicted the main effect of selection task, but has no coverage of the interaction of selection task with vagueness.
-
 \begin{figure}[htbp]
 \centering
-\includegraphics[width=.75\textwidth]{images/response-time-two-panels-1-e4.pdf}
-\caption{Results for Experiment 4}
+\includegraphics[trim = 0mm 30mm 0mm 10mm, clip, width=.75\textwidth]{images/rtplote4.pdf}
+\caption{Mean response times by condition for Experiment 4}
 \label{resultse4}
@@
 \multirow{ 2}{*}{comparison} 	& crisp 			& 2			& \multirow{ 2}{*}{vagueness advantage}  		\\
+\cline{2-3}
   						& vague			& 1			&                                               				\\
@@
 \multirow{ 2}{*}{matching} 		& crisp 			& 1 			&\multirow{ 2}{*}{vagueness disadvantage}	\\
+\cline{2-3}
  						& vague			& 2			&								 	\\
@@
 %The random effects in the model were \emph{participant}, and slopes over \emph{participant} for \emph{instruction format}, \emph{vagueness}, and their interaction.
+
+\begin{table}[htbp]
+\centering
+\caption{Experiment 4: Instructions and targets}
+\label{Instructions for e4}
+\begin{tabular}{llll}
+vagueness	&	instruction		&	selection							&	instruction: Choose a square with \ldots\\
+			&	format		&	task		&			\\
+\toprule
+\multirow{ 2}{*}{crisp} 	 	& \multirow{ 2}{*}{verbal} 	&	matching 		& \ldots the same number of dots as the target \\ 
+	& 	&	comparison										& \ldots fewer dots than the target \\
+	\midrule
+\multirow{ 2}{*}{vague} 	& \multirow{ 2}{*}{verbal} 	&	matching 			& \ldots about the same number of dots as the target \\ 
+	& 	&	comparison										& \ldots far fewer dots than the target \\ 
+\bottomrule
+\end{tabular}
+\end{table}%

### 2016-09-27 experiments/e4/writeup/e4.Rmd
less verbose e4 writeup

@@
     template: default
-    code_folding: hide
+    <!--code_folding: hide-->
     theme: default
 ---  
+<style>
+table, th, td {
+    border: 1px grey;
+    border-collapse: collapse;
+    padding:2px 5px 2px 5px
+}  
+</style>
+
+```{r libs, message=FALSE}
+library(knitr); library(MASS); library(ggplot2); library(lme4); library(xtable); library(plyr); library(lattice); library(LMERConvenienceFunctions); library(car); library(grid); library(lmerTest); library(reshape2); library(ggplot2); library(languageR); library(pander); source('summarySEwithin2.R')
+```
+
 # Experiment 4
@@
 
-<style type="text/css">
-.table {width: 50%;}
-</style>
+<!-- <style type="text/css"> -->
+<!-- .table {width: 50%;} -->
+<!-- </style> -->
 
@@
 
-```{r, docOpts, echo=FALSE, cache=FALSE}
-library(knitr)
-knit_hooks$set(crop=hook_pdfcrop)
+```{r getdata}
+rawdata <- read.table('data.txt')
 ```
 
-```{r, loadLibraries, echo=FALSE}
-# Load libraries.
-suppressPackageStartupMessages(library(MASS))
-suppressPackageStartupMessages(library(knitr))
-suppressPackageStartupMessages(library(ggplot2))
-suppressPackageStartupMessages(library(lme4))
-suppressPackageStartupMessages(library(xtable))
-suppressPackageStartupMessages(library(plyr))
-suppressPackageStartupMessages(library(reshape2))
-suppressPackageStartupMessages(library(lattice))
-suppressPackageStartupMessages(library(latticeExtra))
-suppressPackageStartupMessages(library(LMERConvenienceFunctions))
-suppressPackageStartupMessages(library(car))
-suppressPackageStartupMessages(library(grid))
-suppressPackageStartupMessages(library(lmerTest))
+```{r targets}
+target <- subset(rawdata, subset= Item=="06:15:24" & Quantity=="Small")
+target2 <- subset(target, select=c(Item, Condition, Ins, Target))
+target3 <- unique(target2[order(target$Condition),] )
+row.names(target3) <- NULL
+pander(target3)
 ```
 
-Source additional R files with functions for computing confidence intervals.
-
-```{r, sourceFunctions}
-source("CIfunctions.R")
+```{r trimat25sd}
+trimdata <- perSubjectTrim.fnc(rawdata, response='RT', subject='Subject', trim = 2.5)$data
+rtdata <- trimdata
 ```
 
-Get data from the raw per-subject files if necessary.
-
-```{r, getData}
-if (file.exists('data.txt')) {
-	dat <- read.delim(file="data.txt", header=T, stringsAsFactors=FALSE)
-} else {
-	system(wait=TRUE, paste("head -n 1 ", "../experimentCode","/output/", "subject01.data ", "> ", "data.txt", sep=""))
-	system(wait=TRUE, paste("tail -q -n +2 ", "../experimentCode", "/output/", "*.data ",  ">> ", "data.txt", sep=""))
-}
-
-# declare local variables
-Selection_of_valid_subjects <- 40
-Selection_of_rows <- 7680
-Selection_of_trials_per_subject <- Selection_of_rows / Selection_of_valid_subjects # 192
-
-# sort out borderline responses into expected near and far
-# How many squares in the square they chose? 
-dat$RESPONSE <- as.character(dat$key)
-for (row in 1 : nrow(dat) ) {
-  switch(dat[row,'RESPONSE'],
-         'L' = {dat[row, 'choice'] <- dat[row, 'Lft']},
-         'M' = {dat[row, 'choice'] <- dat[row, 'Mid']},
-         'R' = {dat[row, 'choice'] <- dat[row, 'Rgt']}         
-  )
-}
-dat$Item <- dat$Itm
-dat$Itm <- NULL
-dat$crossed=paste('Con',dat$Cnd,':Quan',dat$Qty,':Item',dat$Item,sep="")
-
-dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseExpected'] <-  6 
-dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseNear']     <- 15  
-dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseFar']      <- 24
-
-dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseExpected'] <- 16
-dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseFar']      <- 34
-
-dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseExpected'] <- 26
-dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseFar']      <- 44
-
-dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseExpected'] <- 36
-dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseFar']      <- 54
-
-dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseExpected'] <- 24
-dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseNear']     <- 15 
-dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseFar']      <-  6
-
-dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseExpected'] <- 34
-dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseFar']      <- 16
-
-dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseExpected'] <- 44
-dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseFar']      <- 26
-
-dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseExpected'] <- 54
-dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseFar']      <- 36
-
-dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseExpected'] <-  6 
-dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseNear']     <- 15  
-dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseFar']      <- 24
-
-dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseExpected'] <- 16
-dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseFar']      <- 34
-
-dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseExpected'] <- 26
-dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseFar']      <- 44
-
-dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseExpected'] <- 36
-dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseFar']      <- 54
-
-dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseExpected'] <- 24
-dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseNear']     <- 15 
-dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseFar']      <-  6
-
-dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseExpected'] <- 34
-dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseFar']      <- 16
-
-dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseExpected'] <- 44
-dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseFar']      <- 26
-
-dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseExpected'] <- 54
-dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseFar']      <- 36
-
-dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseExpected'] <-  6 
-dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseNear']     <- 15  
-dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseFar']      <- 24
-
-dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseExpected'] <- 16
-dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseFar']      <- 34
-
-dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseExpected'] <- 26
-dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseFar']      <- 44
-
-dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseExpected'] <- 36
-dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseFar']      <- 54
-
-dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseExpected'] <- 24
-dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseNear']     <- 15 
-dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseFar']      <-  6
-
-dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseExpected'] <- 34
-dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseFar']      <- 16
-
-dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseExpected'] <- 44
-dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseFar']      <- 26
-
-dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseExpected'] <- 54
-dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseFar']      <- 36
-
-dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseExpected'] <-  6 
-dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseNear']     <- 15  
-dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseFar']      <- 24
-
-dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseExpected'] <- 16
-dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseFar']      <- 34
-
-dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseExpected'] <- 26
-dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseFar']      <- 44
-
-dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseExpected'] <- 36
-dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseFar']      <- 54
-
-dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseExpected'] <- 24
-dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseNear']     <- 15 
-dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseFar']      <-  6
-
-dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseExpected'] <- 34
-dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseFar']      <- 16
-
-dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseExpected'] <- 44
-dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseFar']      <- 26
-
-dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseExpected'] <- 54
-dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseFar']      <- 36
-
-dat$isResponseExpected <- dat$choice == dat$ResponseExpected
-dat$isResponseNear <- dat$choice == dat$ResponseNear
-dat$isResponseFar <- dat$choice == dat$ResponseFar
-
-for ( row in 1:nrow(dat) ) {
-  dat[row, 'switchResponse'] <- 
-    ifelse(dat[row, 'isResponseExpected']==TRUE, 'Expected', 
-           ifelse(dat[row, 'isResponseNear']==TRUE, 'Near', 'Far') ) 
-}
-
-# SUBJECT
-# ensure Subject is a factor 
-dat$Subject=factor(paste("s",sprintf("%02d",dat$Sub),sep=""))
-
-# TRIAL
-# Trial for subject, 1 to 192
-dat$Trial = rep(x = 1:Selection_of_trials_per_subject, times = Selection_of_valid_subjects)
-
-# make a centred Trial for modeling
-dat$c_Trl <-dat$Trial - mean(dat$Trial)
-
-# make a scaled Trial for modelling
-dat$s_Trl <- as.numeric(scale(dat$Trial))
-
-# ID
-# id is a unique identifier for the 7680 row data
-dat$id <- factor(paste(paste(dat$Subject), 
-                       paste("t", sprintf("%03d", dat$Trial), sep="") , sep=":"))
-
-# ITEM
-# create a centred numeric item variable for modeling
-dat$c_Itm <- ifelse(dat$Item==1, -.75, 
-                    ifelse(dat$Item==2, -.25,
-                           ifelse(dat$Item==3, .25, .75)))
-
-# make Item be a factor and assign labels
-dat$Item <- factor(dat$Item, levels=c(1,2,3,4), labels=c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
-
-# add ratios for Item
-item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54) 
-# 0.2500000 0.4705882 0.5909091 0.6666667
-item_range_ratio_scaled = as.vector(scale(c(6 / 24, 16 / 34, 26 / 44, 36 / 54)))
-# -1.3441995 -0.1316642  0.5297187  0.9461450
-item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))) 
-# 0.5125000 0.6876471 0.7691558 0.8166667
-item_mean_ratio_scaled = as.vector(scale(c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))))) 
-# -1.37582241 -0.06614191  0.54334858 0.89861574
-
-dat[dat$Item == "06:15:24", "item_range_ratio"] <-  0.2500000
-dat[dat$Item == "16:25:34", "item_range_ratio"] <-  0.4705882
-dat[dat$Item == "26:35:44", "item_range_ratio"] <-  0.5909091
-dat[dat$Item == "36:45:54", "item_range_ratio"] <-  0.6666667
-
-dat[dat$Item == "06:15:24", "item_range_ratio_scaled"] <-  -1.3441995
-dat[dat$Item == "16:25:34", "item_range_ratio_scaled"] <-  -0.1316642
-dat[dat$Item == "26:35:44", "item_range_ratio_scaled"] <-   0.5297187  
-dat[dat$Item == "36:45:54", "item_range_ratio_scaled"] <-   0.9461450
-
-dat[dat$Item == "06:15:24", "item_mean_ratio"] <-  0.5125000 
-dat[dat$Item == "16:25:34", "item_mean_ratio"] <-  0.6876471 
-dat[dat$Item == "26:35:44", "item_mean_ratio"] <-  0.7691558
-dat[dat$Item == "36:45:54", "item_mean_ratio"] <-  0.8166667
-
-dat[dat$Item == "06:15:24", "item_mean_ratio_scaled"] <-   -1.37582241 
-dat[dat$Item == "16:25:34", "item_mean_ratio_scaled"] <-   -0.06614191 
-dat[dat$Item == "26:35:44", "item_mean_ratio_scaled"] <-    0.54334858 
-dat[dat$Item == "36:45:54", "item_mean_ratio_scaled"] <-    0.89861574
-
-# VAGUENESS
-# Create a factor coding for Vagueness
-dat[ dat$Cnd==1 , 'Vagueness'] <- 'Vague'
-dat[ dat$Cnd==2 , 'Vagueness'] <- 'Crisp'
-dat[ dat$Cnd==3 , 'Vagueness'] <- 'Vague'
-dat[ dat$Cnd==4 , 'Vagueness'] <- 'Crisp'
-dat$Vagueness <- as.factor(dat$Vagueness)
-
-# manually center Vagueness
-dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)
-
-dat$Selection = ""
-dat[dat$Cnd %in% c(1,2), "Selection"] <- "matching"
-dat[dat$Cnd %in% c(3,4), "Selection"] <- "comparison"
-dat$Selection.sum <- ifelse(dat$Selection=="matching", .5, -.5)
-dat$Selection <- as.factor(dat$Selection)
-
-# remove redundant confusing columns
-dat$Sub <- NULL
-dat$Trl <- NULL
-
-# ORDER
-# give the levels of Order meaningful names
-dat$Order <- factor(dat$Ord, levels=c(1,2), labels=c('LtoR','RtoL'))
-
-# make a manually centred Order
-dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)
-
-# QUANTITY
-# give the levels of Quantity meaningful names
-dat$Quantity <- factor(dat$Qty, levels=c(1,2), labels=c('Small','Large'))
-
-# make a manually centred Quantity
-dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)
-
-# INSTRUCTION
-# add Selection of characters in the instruction # 29 30 34 36 38
-dat$nchar_instr = nchar(as.character(dat$Ins))
-dat$nchar_instr_scaled = as.vector(scale(nchar(as.character(dat$Ins)), scale=TRUE))
-
-# make Instruction be a factor (17 levels)
-dat$Ins <- as.factor(dat$Ins)
-
-# RT
-# add transformations of RT
-dat$RT <- dat$rt; dat$rt <-  NULL
-dat$RT_RecSqd    <- 1/dat$RT^2
-dat$RT_Rec       <- 1/dat$RT
-dat$RT_RecSqt    <- 1/sqrt(dat$RT) 
-dat$RT_log       <- log(dat$RT)
-dat$RT_sqt       <- sqrt(dat$RT)
-dat$RT_raw       <- dat$RT
-dat$RT_sqd       <- dat$RT^2
-
-# lose impossible trials
-dd <- dat
-dd$RT[dd$RT > 9000 ] <- NA
-dd <- dd[complete.cases(dd),]
-row.names(dd) <- NULL
-
-# add preceding RT: because we removed impossible trials, the value for preceding RT for a trial following an impossible trial is the value of the trial that preceded the impossible trial.
-dd$RTprev <- NA
-for (s in levels(dd$Subject)) {
-  nrows = nrow(dd[dd$Subject==s,])
-  for (i in 1:nrows) {
-    if (i==1) {
-      dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]
-    }
-    else
-      dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1]
-  }
-}
-
-# add transformations of previous RT
-dd$RTprev_RecSqd    <- 1/dd$RTprev^2
-dd$RTprev_Rec       <- 1/dd$RTprev
-dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev) 
-dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
-dd$RTprev_log       <- log(dd$RTprev)
-dd$RTprev_sqt       <- sqrt(dd$RTprev)
-dd$RTprev_raw       <- dd$RTprev
-dd$RTprev_sqd       <- dd$RTprev^2
-
-save(dd, file="data.Rda")
+```{r computeSummariesForGraphs, cache=TRUE, message=FALSE}
+rtdataplot1 <- summarySEwithin2(rtdata, measurevar="RT_log", withinvars=c("Condition", "Vagueness", "Selection", "Item"), idvar="Subject")
 ```
 
-Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions.
-
-```{r foo,fig.width=20, fig.height=12, message=FALSE}
-## plot the comparison of the transformations against their normal equivalents to see distributional properties using full data
-
-# first create a data frame specially for transformations
-# this won't be used for analysis, just for this plot
-dat_transforms <- dat
-
-# dfSubset says whether any data points were removed
-dat_transforms$dfSubset <- "original"
-
-# RTtype says whether the row contains data for an observed RT or 
-# a sample from a normal distribution with the same mean and sd
-dat_transforms$RTtype <- "observed"
-
-# create extra rows for dfs with removed (outlier) data points
-dat1 <- dat_transforms                                       # nrow 7680
-dat2 <- subset(dat_transforms, RT<9000)              # nrow 7677
-dat2$dfSubset <- "RT<9000"
-
-# bring the dfs back into main df
-dat_transforms <- rbind(dat1,dat2)
-dat_transforms$dfSubset <- factor(dat_transforms$dfSubset, levels=c("original", "RT<9000"))
-
-# create extra rows for normally distributed transformed variables
-dato = dat_transforms
-datn = dat_transforms; datn$RTtype <- "normal"
-dat_transforms <- rbind(dato,datn)
-dat_transforms$RTtype <- factor(dat_transforms$RTtype, levels=c("observed","normal"))
-
-# for safety
-dat_transforms[dat_transforms$RTtype=="normal", c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd")] <- NA
-
-# Create the normally distributed transformed variables, replacing the NAs declared above
-for (k in levels(dat_transforms$dfSubset)){
-  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_RecSqd"]  <- 
-    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_RecSqd"]), 
-          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_RecSqd"]), 
-          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_RecSqd"]))
-  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_Rec"]  <- 
-    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_Rec"]), 
-          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_Rec"]), 
-          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_Rec"]))
-  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_RecSqt"]  <- 
-    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_RecSqt"]), 
-          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_RecSqt"]), 
-          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_RecSqt"]))
-  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_log"]  <- 
-    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_log"]), 
-          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_log"]), 
-          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_log"]))
-  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_sqt"]  <- 
-    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_sqt"]), 
-          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_sqt"]), 
-          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_sqt"]))
-  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_raw"]  <- 
-    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_raw"]), 
-          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_raw"]), 
-          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_raw"]))
-  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_sqd"]  <- 
-    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_sqd"]), 
-          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_sqd"]), 
-          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_sqd"]))
-}
-
-temp <- reshape2::melt(data=dat_transforms, 
-             id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Selection", "Quantity", "Order"),
-             measure.vars=c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd"))
-
-ggplot(temp, 
-			 aes(value, colour=RTtype)) + 
-	geom_freqpoly() + 
-  facet_wrap(dfSubset~variable, scales="free", nrow=2) +
-  theme_bw() + theme(aspect.ratio=1) + scale_color_manual(values=c("blue","red"))
+```{r rtplot, fig.width=7, fig.height=3, echo=F}
+dodge = position_dodge(width=0.2)
+rtdataplot1$Condition <- as.factor(paste(sep=' ', rtdataplot1$Selection, rtdataplot1$Vagueness))
+rtplot=ggplot(rtdataplot1, aes(y=RT_logNormed, x=Item, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci,
+                group=Condition, shape=Condition, fill=Condition)) +
+  geom_line(position=dodge) +
+  geom_errorbar(width=.2, position=dodge) +
+  geom_point(size=2, position=dodge) +
+  scale_shape_manual("",values = c(22, 22, 21, 21)) +
+  scale_fill_manual("",values=c("black","white","black","white")) +
+  ggtitle("Response time") +
+  ylab("Respone time (log (ms))") + 
+  xlab("") +
+  theme_bw() +
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        legend.key = element_blank(), aspect.ratio=1,
+        axis.text.x = element_text(angle = 45, hjust = 1),
+        plot.background=element_rect(fill=NA, color='white')) +
+  facet_grid(~Selection) 
+rtplot
 ```
 
-Show how transformations of RT affect distribution of times, and how they affect which times are outliers.
-
-```{r bar, fig.width=20, fig.height=5}
-# boxplot subjects and items
-load("data.Rda")
-# allow reference to transformation
-subdata = melt(dd,
-               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw")) 
-# boxplots for subjects and items separately for each dependent variable
-ggplot(subdata) + 
-  facet_wrap(~ variable, scales="free", ncol=4) +
-  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
-  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
-  theme_bw() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
-  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
-  xlab("") + ylab("")
+```{r rtlmer, cache=TRUE}
+rtlmer = lmerTest::lmer(data=rtdata,
+                        RT_log ~ c_Vag * c_Sel + c_Itm +
+                          (1 + c_Vag * c_Sel + c_Itm | Subject))
 ```
 
-Show how mean times for individual subjects and items vary with respect to the grand mean RT.
-
-```{r, bar2, fig.width=10, fig.height=5}
-# Show how mean times for individual subjects and items vary with respect to the grand mean RT.
-subs=ddply(dd, .(level=Subject), summarise, effect="subject", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
-itms=ddply(dd, .(level=Item), summarise, effect="item", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
-si=rbind(subs,itms)
-si$effect <- factor(si$effect, levels=c("subject", "item"))
-si$lsign <- factor(si$lsign, levels=c("slow","fast"))
-
-my1 <- xyplot(data=si, lmean~level | effect, 
-              aspect=1,
-              layout=c(2,1),
-              ylim = extendrange(si$lmean, f = 0.10),
-              par.settings=list(trellis.par.set(superpose.symbol = list(col = c("blue","red"), pch=19), superpose.line=list(col = c("blue","red")))),
-              groups=si$lsign,
-              xlab=NULL,
-              ylab="RT (in ms)",
-              key=list(text=list(levels(si$lsign)), lines=list(lty=c(1,1), col=trellis.par.get('superpose.line')$col, pch=c(19,19), type='o'), divide=1),
-              mean=unique(si$emean),
-              scales=list(x=list(relation="free", at=list(15.5, 32.5), labels=list("subject number", "item identifier"), cex=1 ), 
-                          y=list(at=c(0, 2000, unique(si$emean), 4000, 6000, 8000,  10000), labels=c(0, 2000,  "Mean", 4000,  6000, 8000, 10000), cex=.9) ),
-              panel=function(x,y,type,subscripts,groups=groups,...){
-                panel.xyplot(x,y,type='p',groups=groups, subscripts=subscripts,...)
-                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, groups=groups, subscripts=subscripts, code=3, length=0, col=trellis.par.get('superpose.line')$col[groups][subscripts], ...)
-                panel.abline(h=unique(si$emean), col="lightgrey",lty=1)
-                panel.text(x, y=(ifelse(y<unique(si$emean),y-200,y+200)), as.numeric(si$level)[subscripts][panel.number()==1], cex=.75)
-                panel.text(x, y=(ifelse(y<unique(si$emean),y-200,y+200)), as.character(si$level)[subscripts][panel.number()==2], cex=.75)
-              })
-
-my2 <- xyplot(data=si, ldiff~level | effect, 
-              aspect=1, type='n',
-              ylab="Deviation from mean RT (in ms)",
-              ylim=extendrange(si$ldiff,f=0.10))
-
-print(doubleYScale(my1, my2, add.axis=TRUE, add.ylab2=TRUE, use.style=FALSE))
+```{r rtlmersummaryxtable, results='asis', cache=TRUE, echo=F}
+print(xtable(summary(rtlmer)$coef, digits=c(0,2,2,1,1,4)),type='html')
 ```
 
-Show how practice and fatigue effects vary over subjects
-
-```{r barr3, eval=TRUE, fig.width=20, fig.height=10, cache=FALSE, echo=TRUE}
-# Show how practice/fatigue effects vary over subjects
-ggplot(dd, aes(y=RT_log, x=Trial)) + facet_wrap(~Subject,nrow=3) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial", breaks=NULL) + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1)
+```{r seplmersforcomparison, results='asis', cache=TRUE, echo=-3}
+s1 <- subset(rtdata, subset=Selection=='Comparison')
+s1lm <- lmerTest::lmer(data=s1, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
+print(xtable(summary(s1lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
 ```
 
-Show practice and fatigue effects vary over items.
-
-```{r, var3, eval=TRUE, fig.width=9, fig.height=4, cache=FALSE, echo=TRUE}
-# Show how practice/fatigue effects vary over items
-ggplot(dd, aes(y=RT_log, x=Trial)) + facet_grid(~Item) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial") + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1)
+```{r seplmersformatching, results='asis', cache=TRUE, echo=-3}
+s2 <- subset(rtdata, subset=Selection=='Matching')
+s2lm <- lmerTest::lmer(data=s2, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
+print(xtable(summary(s2lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
 ```
-
-# Show means.
-
-Main effects
-
-```{r, bar4}
-# prep main effects rts data frame
-vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness")));names(vrts)[3]<-"Levels"
-nrts <- cbind(Effect="Selection",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection")));names(nrts)[3]<-"Levels"
-qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Quantity")));names(qrts)[3]<-"Levels"
-orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Order")));names(orts)[3]<-"Levels"
-irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("Item")));names(irts)[3]<-"Levels"
-rts=rbind(vrts,nrts,qrts,orts)
-rts$mins=rts$RT_log-rts$ci
-rts$maxs=rts$RT_log+rts$ci
-
-# plot main effects
-ggplot(rts, aes(y=RT_log, x=Levels, group=1, ymin=mins, ymax=maxs)) + 
-  geom_line() +
-  geom_errorbar(width=.1) +
-  geom_point(cex=3) + 
-  facet_grid(~Effect, scale="free_x") + 
-  theme_bw() +
-  theme(aspect.ratio=1) 
-```
-
-Main effects over items
-
-```{r, bar5, message=FALSE, fig.width=10, fig.height=3}
-vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Vagueness")));names(vrts)[4]<-"Levels"
-nrts <- cbind(Effect="Selection",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Selection")));names(nrts)[4]<-"Levels"
-qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Quantity")));names(qrts)[4]<-"Levels"
-orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Order")));names(orts)[4]<-"Levels"
-
-rts=rbind(vrts,nrts,qrts,orts)
-rts$mins=rts$RT_log-rts$ci
-rts$maxs=rts$RT_log+rts$ci
-
-rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
-
-ggplot(rts, aes(y=RT_log, x=item_mean_ratio, group=Levels, ymin=mins, ymax=maxs, shape=Levels, fill=Levels, col=Levels)) +
-  scale_shape_manual(values=c(21,21,22,22,23,23,24,24)) +
-  scale_fill_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
-  scale_color_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
-  geom_line() +
-  geom_errorbar(width=.01) +
-  geom_point(cex=3) + 
-  facet_grid(~Effect) + 
-  theme_bw() +
-  theme(aspect.ratio=1, 
-        panel.grid=element_blank()) +
-  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1)) 
-```
-
-2 way interactions ignoring item
-
-```{r, bar6}
-# prep 2 ways rts data frame
-rts1 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Selection")); names(rts1)[1] <- "F1"; names(rts1)[2] <- "F2"
-rts1$Effect <- rep("Vagueness:Selection",4)
-rts2 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Quantity")); names(rts2)[1] <- "F1"; names(rts2)[2] <- "F2"
-rts2$Effect <- rep("Vagueness:Quantity",4)
-rts3 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Order")); names(rts3)[1] <- "F1"; names(rts3)[2] <- "F2"
-rts3$Effect <- rep("Vagueness:Order",4)
-rts5 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection","Quantity")); names(rts5)[1] <- "F1"; names(rts5)[2] <- "F2"
-rts5$Effect <- rep("Selection:Quantity",4)
-rts6 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection","Order")); names(rts6)[1] <- "F1"; names(rts6)[2] <- "F2"
-rts6$Effect <- rep("Selection:Order",4)
-rts8 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Quantity","Order")); names(rts8)[1] <- "F1"; names(rts8)[2] <- "F2"
-rts8$Effect <- rep("Quantity:Order",4)
-rts <- rbind(rts1,rts2,rts3,rts5,rts6,rts8)
-rts <- subset(rts, select=c(Effect, F1, F2, RT_log, RT_log_norm, sd, se, ci))
-rts$mins=rts$RT_log-rts$ci
-rts$maxs=rts$RT_log+rts$ci
-```
-
-```{r,bar7,fig.width=10, fig.height=3}
-# plot 2 way interactions
-ggplot(rts, aes(y=RT_log, x=F2, group=F1, ymin=mins, ymax=maxs, pch=F1)) + 
-  geom_line() +
-  geom_errorbar(width=.1) +
-  geom_point(cex=3) + 
-  facet_wrap(~Effect, scale="free_x", nrow=1) + 
-  theme_bw() +
-  theme(aspect.ratio=1) +
-  theme(legend.title=element_blank()) +
-  xlab("Levels")
-```
-
-Plot 2 way interactions over items
-
-```{r foo2, message=FALSE}
-# prep 2 ways over items rts data frame
-rts1 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Vagueness","Selection")); names(rts1)[2] <- "F1"; names(rts1)[3] <- "F2"
-rts1$E1 <- "Vagueness"
-rts1$E2 <- "Selection"
-rts2 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Vagueness","Quantity")); names(rts2)[2] <- "F1"; names(rts2)[3] <- "F2"
-rts2$E1 <- "Vagueness"
-rts2$E2 <- "Quantity"
-rts3 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Vagueness","Order")); names(rts3)[2] <- "F1"; names(rts3)[3] <- "F2"
-rts3$E1 <- "Vagueness"
-rts3$E2 <- "Order"
-rts4 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Quantity", "Selection")); names(rts4)[2] <- "F1"; names(rts4)[3] <- "F2"
-rts4$E1 <- "Quantity"
-rts4$E2 <- "Selection"
-rts5 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Order","Selection")); names(rts5)[2] <- "F1"; names(rts5)[3] <- "F2"
-rts5$E1 <- "Order"
-rts5$E2 <- "Selection"
-rts6 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Quantity","Order")); names(rts6)[2] <- "F1"; names(rts6)[3] <- "F2"
-rts6$E1 <- "Quantity"
-rts6$E2 <- "Order"
-
-rts <- rbind(rts1,rts2,rts3,rts4,rts5,rts6)
-rts$mins=rts$RT_log-rts$ci
-rts$maxs=rts$RT_log+rts$ci
-
-rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
-rts$F3 <- factor(paste(rts$F1,rts$F2))
-
-rts$E1 <- factor(rts$E1, levels=c("Selection","Order","Quantity","Vagueness"))
-rts$E2 <- factor(rts$E2, levels=c("Selection","Order","Quantity","Vagueness"))
-
-# The palette with black:
-cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-
-ggplot(rts, aes(y=RT_log, x=item_mean_ratio, group=F3, col=F2, pch=F1)) + 
-  facet_grid(E2~E1, drop=F) + 
-  geom_point(cex=3) + geom_line() + theme_bw() + 
-  theme(aspect.ratio=1, panel.grid=element_blank()) +
-  scale_color_manual(name="colour",values=cbbPalette) +
-  scale_shape_discrete(name="shape") +
-  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))
-```
-
-# RT model all items
-
-```{r, foo3, rtmodel-allitems}
-# backing off A LOT from maximal model to avoid warning about maxfun
-rtm <- lme4::lmer(
-  data=dd,
-  RT_log ~ 
-    c_Vag * Selection.sum + item_mean_ratio + c_Qty + c_Ord + s_Trl +
-    RTprev_log  + 
-  ( c_Vag * Selection.sum | Subject)
-)
-rtm.summary <- summary(rtm)
-rtm.confint <- confint(rtm, method='Wald')[13:20,]
-```
-
-Print a table of coefficients.
-
-```{r, foo4, rtmodeltable, results='asis', echo=FALSE}
-print(xtable(
-  digits=c(0,2,2,1), 
-  rtm.summary$coefficients), type='html')
-```
-
-Plot coefficients
-
-```{r, foo5}
-par(las=2, mar=c(14,6,1,1))
-y=coef(rtm.summary)[-1,1] # omit intercept
-n=length(y)
-plot(y, xaxt="n", xlab="", ylab="RT_log\n", pch=19, ylim=extendrange(y, f=.20), main="Coefficients")
-abline(h=0)
-axis(1, labels=names(y) , at=1:n)
-grid()
-segments(x0=1:n, x1=1:n, y0=rtm.confint[,1], y1=rtm.confint[,2], lwd=2)
-```
-
-Plot partial effects
-
-```{r foo6}
-par(mfrow=c(2,4))
-plotLMER.fnc(rtm)
-```
-
-```{r foo7}
-suppressPackageStartupMessages(require(languageR))
-cat("k is:",
-collin.fnc(dd[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])$cnumber)
-```
-
-```{r foo8, fig.width=6, fig.height=5}
-suppressPackageStartupMessages(require(rms))
-par(mar=c(1.1,3.1,1.1,1.1), pty='s')
-plot(varclus(as.matrix(dd[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])))
-```
-
-Model criticism baayen plots
-
-```{r daz, fig.width=9, fig.height=3}
-# Baayen 4-plot model criticism
-
-par(mfrow=c(1,4), pty='s')
-# create scaled residuals
-dd$rstand = as.vector(scale(resid(rtm)))
-# plot scaled residuals density
-plot(density(dd$rstand))
-# plot sample quantiles versus theoretical quantiles
-qqnorm(dd$rstand, cex=.5)
-qqline(dd$rstand)
-# plot standardised residuals versus fitted values
-plot(dd$rstand ~ fitted(rtm), pch='.')
-# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
-abline(h=c(-2.5,2.5))
-
-```
-
-Refit model with outliers removed (3.5 sd)
-
-```{r daz2}
-dd2 <- subset(dd, abs(rstand)<3.5 )
-```
-
-```{r daz3}
-# backing off A LOT from maximal model to avoid warning about maxfun
-rtm <- lmer(
-  data=dd2,
-  RT_log ~ 
-    c_Vag * Selection.sum + item_mean_ratio + c_Qty + c_Ord + s_Trl +
-    RTprev_log  + 
-  ( c_Vag * Selection.sum | Subject)
-)
-rtm.summary <- summary(rtm)
-rtm.confint <- confint(rtm, method='Wald')
-```
-
-Print a table of coefficients.
-
-```{r daz4,  results='asis', echo=FALSE}
-print(xtable(
-  digits=c(0,2,2,1,1,3), 
-  rtm.summary$coefficients), type='html')
-```
-
-Plot coefficients
-
-```{r daz5}
-par(las=2, mar=c(14,6,1,1))
-y=coef(rtm.summary)[-1,1] # omit intercept
-n=length(y)
-plot(y, xaxt="n", xlab="", ylab="RT_log\n", pch=19, ylim=extendrange(y, f=.20), main="Coefficients")
-abline(h=0)
-axis(1, labels=names(y) , at=1:n)
-grid()
-segments(x0=1:n, x1=1:n, y0=rtm.confint[13:20,1], y1=rtm.confint[13:20,2], lwd=2)
-```
-
-Plot partial effects
-
-```{r daz6}
-par(mfrow=c(2,4))
-plotLMER.fnc(rtm)
-```
-
-```{r daz7}
-suppressPackageStartupMessages(require(languageR))
-cat("k is:",
-collin.fnc(dd2[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])$cnumber
-)
-```
-
-```{r daz9, fig.width=6, fig.height=5}
-suppressPackageStartupMessages(require(rms))
-par(mar=c(1.1,3.1,1.1,1.1), pty='s')
-plot(varclus(as.matrix(dd2[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])))
-```
-
-Model criticism baayen plots
-
-```{r mc1, fig.width=9, fig.height=3}
-# Baayen 4-plot model criticism
-
-par(mfrow=c(1,4), pty='s')
-# create scaled residuals
-dd2$rstand = as.vector(scale(resid(rtm)))
-# plot scaled residuals density
-plot(density(dd2$rstand))
-# plot sample quantiles versus theoretical quantiles
-qqnorm(dd2$rstand, cex=.5)
-qqline(dd2$rstand)
-# plot standardised residuals versus fitted values
-plot(dd2$rstand ~ fitted(rtm), pch='.')
-# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
-abline(h=c(-2.5,2.5))
-
-```
-
-Separate analyses for vague vs crisp at each level of selection (matching, and comparison)
-
-not done yet.
-
-```{r sep1,  results='asis', echo=FALSE}
-print(xtable(
-  digits=c(0,2,2,1,1,3), 
-  rtm.summary$coefficients), type='html')
-```
-
-# Borderline analysis
-
-not done yet.

### 2016-09-27 experiments/e4/writeup/preprocessing/preprocessing.Rmd
less verbose e4 writeup

--- 2016-07-28 experiments/e3/writeup/experiment3.Rmd
@@
----
-title: "Experiment 3 of 4"
-output:
-  html_document:
-    toc: true
-    <!--toc_float: true-->
-    mathjax: default
-    template: default
-    code_folding: hide
-    theme: default
----  
-# Experiment 3
+Get data from the raw per-subject files if necessary.
 
-```{r docOpts}
-library(knitr)
-```
-
-```{r loadLibraries}
-suppressPackageStartupMessages(library(MASS))
-suppressPackageStartupMessages(library(ggplot2))
-suppressPackageStartupMessages(library(lme4))
-suppressPackageStartupMessages(library(xtable))
-suppressPackageStartupMessages(library(plyr))
-suppressPackageStartupMessages(library(reshape))
-suppressPackageStartupMessages(library(lattice))
-suppressPackageStartupMessages(library(LMERConvenienceFunctions))
-suppressPackageStartupMessages(library(car))
-suppressPackageStartupMessages(library(grid))
-suppressPackageStartupMessages(library(lmerTest))
-suppressPackageStartupMessages(library(reshape2))
-suppressPackageStartupMessages(library(ggplot2))
-suppressPackageStartupMessages(require(languageR))
-```
-
-# Experiment Structure
-Each subject saw 192 trials, being 3 cycles of 64 trials. There were 38 participants. There were 7296 rows in the full data set.
-
-# Trial Procedure
-The procedure for each trial was as follows.
-First an instruction that was a referring expression was presented until participants dismissed it with a key press.
-Then the squares with dots were presented until the participant responded with a key press, with a 60 second timeout if they did not respond.
-
-# Prepare the data
-Source additional R files with functions for computing confidence intervals.
-
-```{r sourceFunctions}
-source("functions/CIfunctions.R")
-```
-
-Get data.
-
-```{r readdata}
-dat <- read.delim("data/allresults.txt", stringsAsFactors = FALSE)
+```{r, getData}
+if (file.exists('rawdata.txt')) {
+  dat <- read.delim(file="rawdata.txt", header=T, stringsAsFactors=FALSE)
+} else {
+  system(wait=TRUE, paste("head -n 1 ", "../../experimentCode","/output/", "subject01.data ", "> ", "rawdata.txt", sep=""))
+  system(wait=TRUE, paste("tail -q -n +2 ", "../../experimentCode", "/output/", "*.data ",  ">> ", "rawdata.txt", sep=""))
+  dat <- read.delim(file="rawdata.txt", header=T, stringsAsFactors=FALSE)
+}
 
 # declare local variables
-number_of_valid_subjects <- 38
-number_of_rows <- 7296
-number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # 192
+Selection_of_valid_subjects <- 40
+Selection_of_rows <- 7680
+Selection_of_trials_per_subject <- Selection_of_rows / Selection_of_valid_subjects # 192
 
@@
 # How many squares in the square they chose? 
-dat$RESPONSE <- as.character(dat$RESPONSE)
+dat$RESPONSE <- as.character(dat$key)
 for (row in 1 : nrow(dat) ) {
   switch(dat[row,'RESPONSE'],
-         'LEFT' = {dat[row, 'choice'] <- dat[row, 'Left']},
-         'MIDDLE' = {dat[row, 'choice'] <- dat[row, 'Mid']},
-         'RIGHT' = {dat[row, 'choice'] <- dat[row, 'Right']}         
+         'L' = {dat[row, 'choice'] <- dat[row, 'Lft']},
+         'M' = {dat[row, 'choice'] <- dat[row, 'Mid']},
+         'R' = {dat[row, 'choice'] <- dat[row, 'Rgt']}         
   )
 }
-dat$crossed=paste('Con',dat$Condition,':Quan',dat$Quantity,':Item',dat$Item,sep="")
+dat$Item <- dat$Itm
+dat$Itm <- NULL
+dat$crossed=paste('Con',dat$Cnd,':Quan',dat$Qty,':Item',dat$Item,sep="")
 
@@
 
-## treat variables
-
 # SUBJECT
 # ensure Subject is a factor 
-dat$Subject=factor(paste("s",sprintf("%02d",dat$Subject),sep=""))
+dat$Subject=factor(paste("s",sprintf("%02d",dat$Sub),sep=""))
 
@@
 # Trial for subject, 1 to 192
-dat$Trial = rep(x = 1:number_of_trials_per_subject, times = number_of_valid_subjects)
+dat$Trial = rep(x = 1:Selection_of_trials_per_subject, times = Selection_of_valid_subjects)
 
@@
 # Create a factor coding for Vagueness
-dat[ dat$Condition==1 , 'Vagueness'] <- 'Vague'
-dat[ dat$Condition==2 , 'Vagueness'] <- 'Crisp'
-dat[ dat$Condition==3 , 'Vagueness'] <- 'Vague'
-dat[ dat$Condition==4 , 'Vagueness'] <- 'Crisp'
+dat[ dat$Cnd==1 , 'Vagueness'] <- 'Vague'
+dat[ dat$Cnd==2 , 'Vagueness'] <- 'Crisp'
+dat[ dat$Cnd==3 , 'Vagueness'] <- 'Vague'
+dat[ dat$Cnd==4 , 'Vagueness'] <- 'Crisp'
 dat$Vagueness <- as.factor(dat$Vagueness)
@@
 dat$Selection = ""
-dat[dat$Condition %in% c(1,2), "Selection"] <- "matching"
-dat[dat$Condition %in% c(3,4), "Selection"] <- "comparison"
-dat$Selection.sum <- ifelse(dat$Selection=="matching", .5, -.5)
+dat[dat$Cnd %in% c(1,2), "Selection"] <- "Matching"
+dat[dat$Cnd %in% c(3,4), "Selection"] <- "Comparison"
+dat$c_Sel <- ifelse(dat$Selection=="Matching", .5, -.5)
 dat$Selection <- as.factor(dat$Selection)
 
-# remove one zero RT by replacing with NA
-dat$RT[dat$RT==0]<-NA
+# remove redundant confusing columns
+dat$Sub <- NULL
+dat$Trl <- NULL
 
@@
 # give the levels of Order meaningful names
-dat$Order <- factor(dat$Order, levels=c(1,2), labels=c('LtoR','RtoL'))
+dat$Order <- factor(dat$Ord, levels=c(1,2), labels=c('LtoR','RtoL'))
 
@@
 # give the levels of Quantity meaningful names
-dat$Quantity <- factor(dat$Quantity, levels=c(1,2), labels=c('Small','Large'))
+dat$Quantity <- factor(dat$Qty, levels=c(1,2), labels=c('Small','Large'))
 
@@
 # INSTRUCTION
-# add number of characters in the instruction # 29 30 34 36 38
-dat$nchar_instr = nchar(as.character(dat$Instruction))
-dat$nchar_instr_scaled = as.vector(scale(nchar(as.character(dat$Instruction)), scale=TRUE))
+# add Selection of characters in the instruction # 29 30 34 36 38
+dat$nchar_instr = nchar(as.character(dat$Ins))
+dat$nchar_instr_scaled = as.vector(scale(nchar(as.character(dat$Ins)), scale=TRUE))
 
 # make Instruction be a factor (17 levels)
-dat$Instruction <- as.factor(dat$Instruction)
+dat$Ins <- as.factor(dat$Ins)
 
@@
 # add transformations of RT
+dat$RT <- dat$rt; dat$rt <-  NULL
 dat$RT_RecSqd    <- 1/dat$RT^2
@@
 
-# save dat as dat.Rda
-# This data set contains *all* trials including impossible trials and is mainly for graphs comparing different removals
-save(dat, file="dat.Rda")
-
 # lose impossible trials
 dd <- dat
-  dd$RT[dd$RT == 1 ] <- NA
-  dd$RT[dd$RT > 40000 ] <- NA
+dd$RT[dd$RT > 9000 ] <- NA
 dd <- dd[complete.cases(dd),]
@@
 
-# add preceding RT: because we removed impossible trials, the value for preceding RT for a trial following an impossible trial is the value of the trial that preceded the impossible trial.
-dd$RTprev <- NA
-for (s in levels(dd$Subject)) {
-  nrows = nrow(dd[dd$Subject==s,])
-  for (i in 1:nrows) {
-    if (i==1) {
-      dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]
-    }
-    else
-      dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1]
-  }
-}
+# # add preceding RT: because we removed impossible trials, the value for preceding RT for a trial following an impossible trial is the value of the trial that preceded the impossible trial.
+# dd$RTprev <- NA
+# for (s in levels(dd$Subject)) {
+#   nrows = nrow(dd[dd$Subject==s,])
+#   for (i in 1:nrows) {
+#     if (i==1) {
+#       dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]
+#     }
+#     else
+#       dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1]
+#   }
+# }
+# 
+# # add transformations of previous RT
+# dd$RTprev_RecSqd    <- 1/dd$RTprev^2
+# dd$RTprev_Rec       <- 1/dd$RTprev
+# dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev) 
+# dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
+# dd$RTprev_log       <- log(dd$RTprev)
+# dd$RTprev_sqt       <- sqrt(dd$RTprev)
+# dd$RTprev_raw       <- dd$RTprev
+# dd$RTprev_sqd       <- dd$RTprev^2
 
-# add transformations of previous RT
-dd$RTprev_RecSqd    <- 1/dd$RTprev^2
-dd$RTprev_Rec       <- 1/dd$RTprev
-dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev) 
-dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
-dd$RTprev_log       <- log(dd$RTprev)
-dd$RTprev_sqt       <- sqrt(dd$RTprev)
-dd$RTprev_raw       <- dd$RTprev
-dd$RTprev_sqd       <- dd$RTprev^2
-
-# save dd as dd.Rda
-# This data set non-impossible trials and renamed/recomputed variables
-save(dd, file="dd.Rda")
+dd$Condition <- as.factor(paste(dd$Selection,dd$Vagueness, sep=' '))
+dd$Target <- dd$Prm
+foo <- subset(dd, select=c(
+  "Subject", "Trial", "Condition", "Order", "Quantity", "Vagueness", "c_Vag", "Selection", "c_Sel", "Item", "c_Itm", "Target", "Ins", "switchResponse", "RT", "RT_log"
+))
+write.table(foo, file='../data.txt')
 ```
-
-Check the distribution of the trimmed data to see if a transformation is suggested. Log transformation seems to be best so we adopt that for the rest of the analyses.
-
-```{r boxplot subjects and items}
-load("dd.Rda")
-# allow reference to transformation
-subdata = melt(dd,
-               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
-               variable.name="transformation",
-               value.name="value") 
-# boxplots for subjects and items separately for each dependent variable
-ggplot(subdata) + 
-  facet_wrap(~ variable, scales="free", ncol=4) +
-  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
-  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
-  theme_bw() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
-  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
-  xlab("") + ylab("")
-```
-
-```{r load_dd}
-load("dd.Rda")
-```
-
-```{r do_lmer}
-v5 <- lmer(data=dd, 
-           RT_RecSqt ~ 
-             c_Vag + Selection.sum + c_Qty + c_Ord + 
-             Selection.sum:c_Vag:c_Qty +
-             item_mean_ratio + 
-             s_Trl +
-             RTprev_RecSqt +
-             nchar_instr +
-             (1+c_Vag + Selection.sum + c_Qty + c_Ord|Subject)
-)
-```
-
-```{r summary_lmer}
-summary(v5)
-```
-
-```{r R2}
-cat("R^2")
-cor(fitted(v5), dd$RT_RecSqt)^2
-```
-
-```{r plotLMER}
-par(mfrow=c(2,5))
-plotLMER.fnc(v5)
-```
-
-Plot model coefficients and ci's
-
-```{r coeff}
-par(las=2, mar=c(14,6,1,1))
-y=as.data.frame(summary(v5)$coef[-1,1]) # estimates, without intercept
-q=confint(v5)[-c(1,2,3),]
-y=cbind(y,q)
-plot(y[,1], ylim=range(y[,1],y[,2],y[,3]), xaxt="n", xlab="", ylab="estimate\n",  pch='-', cex=2)
-segments(x0=1:nrow(y), x1=1:nrow(y), y[,2], y[,3])
-abline(h=0)
-axis(1, labels=row.names(y), at=1:nrow(y), srt=45)
-grid()
-```
-
-Model criticism baayen plots
-
-```{r baayenplots, fig.width=9, fig.height=3}
-# Baayen 4-plot model criticism
-
-par(mfrow=c(1,4), pty='s')
-# create scaled residuals
-v5$rstand = as.vector(scale(resid(v5)))
-# plot scaled residuals density
-plot(density(v5$rstand))
-# plot sample quantiles versus theoretical quantiles
-qqnorm(v5$rstand, cex=.5)
-qqline(v5$rstand)
-# plot standardised residuals versus fitted values
-plot(v5$rstand ~ fitted(v5), pch='.')
-# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
-abline(h=c(-2.5,2.5))
-# create dffits
-dffits=abs(resid(v5, "dffits"))
-# plot dffits
-plot(dffits, type='h')
-```

### 2016-09-27 experiments/e2/writeup/e2.Rmd
tried to resize plots better for latex special issue

@@
 rtplot=ggplot(rtdataplot1, aes(y=RT_log, x=Item, ymin=RT_log-ci, ymax=RT_log+ci,
-                group=Condition, shape=Condition, fill=Condition)) +
+                group=Condition, fill=Vagueness)) +
   geom_line(position=dodge) +
   geom_errorbar(width=.2, position=dodge) +
-  geom_point(size=3, position=dodge) +
-  scale_shape_manual("",values = c(22, 22, 21, 21)) +
-  scale_fill_manual("",values=c("black","white","black","white")) +
+  geom_point(pch=21, size=3, position=dodge) +
+  #scale_shape_manual("",values = c(22, 22)) +
+  scale_fill_manual("",values=c("black","white")) +
   ggtitle("Response time") +
@@
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        legend.key = element_blank(), aspect.ratio=1,
+        legend.key = element_blank(), legend.position="top",
+        aspect.ratio=1,
         axis.text.x = element_text(angle = 45, hjust = 1),
@@
   facet_wrap(nrow=1,~Number) +
-  scale_fill_grey(name="Response Type") +
+  scale_fill_grey(name="") +
   xlab(" ") +
@@
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        legend.key = element_blank(), aspect.ratio=1,
+        legend.key = element_blank(), legend.position="top", aspect.ratio=1,
         plot.background=element_rect(fill=NA, color='white'))

### 2016-09-27 papers/2016-Special-Issue/v-special-issue-2016.tex
added stats for e4 to tex

@@
 % KvD added the following from a previous version of the paper:
-There is abundant evidence (e.g., \protect\cite{trick1994small}) that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities.
-We performed a pilot experiment in which we were able to confirm this finding in the experimental settings on which we are focussing in this paper \cite{our-earlier-paper}.
-As one might expect, we found that, when participants were confronted with an array of two boxes, instructions of the form ``Choose the square with $n$ dots" led to consistently lower Response Times than instructions of the form ``Choose the square with many/few dots" when $2 \leq n \leq 5$; the converse was true for $n>5$.
+There is abundant evidence \cite<e.g.,>{trick1994small} that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities.
+We performed a pilot experiment \cite{our-earlier-paper} in which we were able to confirm this finding in the experimental settings on which we are focussing in this paper.
+As one might expect, we found that, when participants were confronted with % MG changed the next line and added the footnote to remove the doubts we had about whether a dot array must have regular layout, or can have irregular layout - I'm still not sure this is a good description though
+a stimulus consisting of two squares containing different numbers of dots\footnote{such a stimulus is referred to hereafter as a stimulus consisting of a set of some number of dot arrays, where the number of dots in each dot array varies and is referred to as its cardinality, and the physical arrangement of dots in each dot array is irregular},
+instructions of the form \emph{Choose the square with \emph{n} dots} led to consistently faster response times than instructions of the form \emph{Choose the square with many/few dots} when $2 \leq n \leq 5$; the converse was true for $n>5$.
 We henceforth focussed our studies on non-subitizable numbers, because it is there that vagueness is expected to have benefits.
@@
 Then a set of two dot arrays was presented, each in the form of a square containing some number of dots.
-The participant was required to identify the array that corresponded with the instruction, by pressing the appropriate key, as quickly as possible while avoiding errors. 
+The participant was required to identify the dot array that corresponded with the instruction, by pressing the appropriate key, as quickly as possible while avoiding errors. 
 Response time and accuracy were recorded for analysis.
@@
 
-As expected, responses were faster and more accurate for vague than crisp instructions.
-Post hoc, we formulated some further hypotheses, which we tested using Bonferroni's correction for multiple comparisons:
+As expected, responses were faster and more accurate for vague than crisp instructions. Post hoc, we formulated some further hypotheses:
 
@@
 
-Looking at Response Times, these two post-hoc hypotheses were confirmed, at Bonferroni-corrected significance thresholds:
+Looking at response times, these two post-hoc hypotheses were supported:
 
@@
 
-Looking at Error Rates, the two post-hoc hypotheses were confirmed as well, at Bonferroni-corrected significance thresholds:
+Looking at error rates, the two post-hoc hypotheses were supported:
 
@@
 
-In Experiment 2, an item was a referring expression instruction followed by a set of dot arrays defined by a triple of numbers, representing the number of dots in the left, middle, and right arrays.
+In Experiment 2, an item was a referring expression instruction followed by a set of three dot arrays defined by a triple of numbers, representing the number of dots in the left, middle, and right arrays.
 We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54).
@@
 	\item [(H2)] An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
-	\item [(H3)] No large main effect of instruction format (numeric versus verbal).
-	\item [(H4)] On the basis of Experiment 1, we would expect faster responses to more discriminable triples (i.e., an effect of triple).
+	\item [(H3)] No large main effect of instruction format (of numeric versus verbal) since we hypothesised that vagueness rather than instruction format drove the effect in experiment 1.
+	\item [(H4)] On the basis of Experiment 1, we would expect faster responses for stimuli with more discriminable arrays.(i.e., an effect of item).
 	\item [(H5)] Participants should make more borderline case choices for vague than crisp instructions.
@@
 \subsubsection{Response times} % E2
+% MG added the values for these stats on 26 Sept 2016
 Means for response times and proportion of borderline responses are given in Fig. (\ref{resultse2}). 
@@
 \begin{APAenumerate}
-	\item [(H1)] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06$, $se=.01$, $t=4.6$, $p<0.0001$) % MG added the values for these stats on 26 Sept 2016
-	\item [(H2)] In focussed comparisons, vagueness was significantly disadvantageous in the \emph{numeric} conditions ($\beta=0.09$, $se=0.02$, $t=4.5$, $p<0.0001$)% MG added the values for these stats on 26 Sept 2016
-				, and non-significantly disadvantageous in the \emph{verbal} conditions ($\beta=0.02$, $se=0.02$, $t=1.5$, $p=0.1550$)% MG added the values for these stats on 26 Sept 2016
-				, offering partial evidence against hypothesis 2.
-The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to a significant overall interaction effect between vagueness and instruction format ($\beta=-0.13$, $se=0.02$, $t=-6.6$, $p=0.0169$) % MG added the values for these stats on 26 Sept 2016
-	\item [(H3)] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; ($\beta=-0.36$, $se=0.07$, $t=-5.1$, $p<0.0001$) % MG added the values for these stats on 26 Sept 2016
+	\item [(H1)] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06$, $se=.01$, $t=4.6$, $p<0.0001$)
+	\item [(H2)] In focussed comparisons, vagueness was significantly disadvantageous in the \emph{numeric} conditions ($\beta=0.09$, $se=0.02$, $t=4.5$, $p<0.0001$), and non-significantly disadvantageous in the \emph{verbal} conditions ($\beta=0.02$, $se=0.02$, $t=1.5$, $p=0.1550$), offering partial evidence against hypothesis 2.
+The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to a significant overall interaction effect between vagueness and instruction format ($\beta=-0.13$, $se=0.02$, $t=-6.6$, $p=0.0169$).
+	\item [(H3)] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1, but suggesting that instruction format rather than vagueness drove the effect we observed in experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; ($\beta=-0.36$, $se=0.07$, $t=-5.1$, $p<0.0001$).
 	\item [(H4)] There was a significant main effect of item (i.e., of which triple of numbers of dots were used in the stimulus): ($\beta=0.12$, $se=0.02$, $t=7.1$, $p<0.0001$).
-% MG added the values for these stats on 26 Sept 2016
-				This effect seems likely to be due to the very fast responses for stimuli using arrangements of the smallest numbers of dots (6,15,24), which also had the largest difference in ratio of smallest number to largest number in the stimulus, suggesting that stimuli using this arrangement of numbers of dots may have been particularly discriminable for participants.
+This effect seems likely to be due to the very fast responses for stimuli using arrangements of the smallest numbers of dots (6,15,24), which also had the largest difference in ratio of smallest number to largest number in the stimulus, suggesting that stimuli using these numbers of dots may have been particularly discriminable for participants.
 \end{APAenumerate}
@@
 \subsubsection{Borderline responses} % E2
+% MG added the values for these stats on 26 Sept 2016
 A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
@@
 	\item [(H5)] Participants were significantly more likely to choose the borderline option for vague instructions than for crisp instructions (21.9\% vs 11.3\%: $\beta=0.62$, $se=0.22$, $z=2.8$, $p=0.0059$).
-% MG added the values for these stats on 26 Sept 2016
 Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%: $\beta=-3.35$, $se=0.23$, $z=-14.6$, $p<0.0001$).
-% MG added the values for these stats on 26 Sept 2016
 \end{APAenumerate}
@@
 However, once again there is a possibly confounding factor.
-Observe that, in Experiment 2, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) {\bf selection algorithm}:
- %as shown by the Table of instructions \ref{instructionse2}, 
-%the instructions that used a verbal instruction format allowed a comparison algorithm, whereas the instructions that used a numeric format allowed a matching algorithm. 
-To see this, consider the task of selecting the dot array that contains ``few dots": to do this, it suffices to {\em compare} the three arrays and select the one that contains the fewest elements.
+Observe that, in Experiment 2, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) {\bf selection algorithm}: To see this, consider the task of selecting the dot array that contains ``few dots": to do this, it suffices to {\em compare} the three arrays and select the one that contains the fewest elements.
 Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison ``algorithm" (which is known to be faster than matching).
 % KvD I think we need to say more about this, citing some appropriate papers to support this idea.
-\textcolor{red}{insert citation supporting this}
+\textcolor{red}{{\bf--Insert citation}}
 
@@
 
-% KvD I think we should motivate our choice of materials (especially "far ..er")
+%KvD I think we should motivate our choice of materials (especially "far ..er")
+\textcolor{red}{{\bf KvD I think we should motivate our choice of materials (especially "far ..er")}}
 
@@
 \centering
-\caption{Experiment 3: Instructions arranged by condition for the dot triple $(6,15,24)$and instructions indicating the smaller numbers} %, and showing \emph{fewer} instead of \emph{more}}
+\caption{Experiment 3: Instructions arranged by condition for stimuli with $(6,15,24)$ dots and instructions indicating the smaller numbers} 
 \label{Instructions for e3}
@@
 \end{table}%
-%&	
-%
-%	instruction				&
-%	
-%	
-%	format				&
-%
 
-\subsection{Hypotheses}
+\subsection{Hypotheses} % E 3
 \begin{APAenumerate}
@@
 \subsection{Results} % E 3
-RTs were trimmed at 2.5 sd separately for each subject, leading to the loss of 204 trials (2.8\% of the trials).
+
+Response times were trimmed at 2.5 SD separately for each subject, leading to the loss of 204 trials (2.8\% of the trials).
 Condition means for the remaining (logged) RTs are plotted in Figure \ref{resultse3}.
@@
 
-% E3
+% MG added the values for these stats on 26 Sept 2016
 \begin{APAenumerate}
 	\item [(H1)] Vague instructions were non-significantly easier for the reader than crisp instructions when vagueness was considered as a main effect ($\beta=-0.01$, $se=0.01$, $t=-0.4$, $p=0.7158$). 
-% MG added the values for these stats on 26 Sept 2016
 	\item [(H2)] There was a statistically significant main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=0.16$, $se=0.03$, $t=6.2$, $p<0.0001$).
-% MG added the values for these stats on 26 Sept 2016
 	\item [(H3)] Vagueness exerted effects in different directions for the comparison task and for the matching task: there was a significant interaction between vagueness and selection task ($\beta=0.13$, $se=0.03$, $t=4.2$, $p<0.0001$).
-% MG added the values for these stats on 26 Sept 2016
-%				
-Separate analyses were conducted testing for effects of vagueness at each level of the selection task, using Bonferroni significance thresholds to correct for multiple comparisons. 
+Separate analyses were conducted testing for effects of vagueness at each level of the selection task.
 Within the comparison task vagueness significantly speeded response times compared with crisp controls ($\beta=-0.07$, $se=0.02$, $t=-3.5$, $p<0.0012$). 
-% MG added the values for these stats on 26 Sept 2016
 Within the matching task vagueness significantly \emph{slowed} response times compared with crisp controls ($\beta=0.06$, $se=0.02$, $t=2.9$, $p<0.0061$). 
-% MG added the values for these stats on 26 Sept 2016
 \end{APAenumerate}
@@
 
+% MG added the values for these stats on 27 Sept 2016
 \begin{APAenumerate}
-	\item [(H1)] There was no significant main effect of vagueness ($\beta=-0.02$, $se=0.01$, $t=-1.5$, $p=0.1296$). % MG added the values for these stats on 27 Sept 2016
+	\item [(H1)] There was no significant main effect of vagueness ($\beta=-0.02$, $se=0.01$, $t=-1.5$, $p=0.1296$). 
 	\item [(H2)] There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=0.18$, $se=0.02$, $t=10.4$, $p<0.0001$).
-% MG added the values for these stats on 27 Sept 2016
 This effect was in the same direction as Experiment 3.
 	\item[(H3)] Vagueness did exert different effects depending on the selection task (main interaction effect of vagueness by selection $\beta=0.12$, $se=0.02$, $t=5.1$, $p<0.0001$). 
-% MG added the values for these stats on 27 Sept 2016
 Separate analyses of the effect of vagueness were conducted for the comparison task and for the matching task using Bonferroni-adjusted significance thresholds. 
-In the comparison task, vagueness resulted in faster response times ($\beta=-0.08$, $se=0.02$, $t=-4.3$, $p<0.0001$). % MG added the values for these stats on 27 Sept 2016
-In the matching task vagueness slowed response times ($\beta=0.05$, $se=0.01$, $t=3.7$, $p=0.0004$). % MG added the values for these stats on 27 Sept 2016
+In the comparison task, vagueness resulted in faster response times ($\beta=-0.08$, $se=0.02$, $t=-4.3$, $p<0.0001$). 
+In the matching task vagueness slowed response times ($\beta=0.05$, $se=0.01$, $t=3.7$, $p=0.0004$). 
 These results are in the same direction as Experiment 3.
@@
 
-Vagueness, which was the phenomenon on which our investigation focussed, did not exert a main effect in response time. 
-However when the comparison and selection tasks were analysed separately, there was small reliable advantage for vagueness in the \emph{comparison} tasks, but a small reliable disadvantage for vagueness in the \emph{matching} tasks.
+Vagueness, which was the phenomenon on which our investigation focussed, did not exert a significant main effect in response time.
+However when the comparison and selection tasks were analysed separately, there was small significant advantage for vagueness in the \emph{comparison} tasks, but a small significant disadvantage for vagueness in the \emph{matching} tasks.
 
@@
 
-Experiment 1 showed hat responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment could not distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format; and mandated a comparison strategy rather than a matching strategy.
-Experiment 2 showed that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional explanatory power in either the verbal format instructions or the numerical format instructions.
-% when we generated verbal and numerical versions of both crisp and vague instructions.
-% MJG edit
+Experiment 1 showed that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment could not distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format; and mandated a comparison strategy rather than a matching strategy.
+Experiment 2 showed that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional benefit in either the verbal format instructions or the numerical format instructions.
 
-However, Experiment 2 could not distinguish benefits of number avoidance from benefits of 
-the comparison selection task. %, since the verbal instructions mandated a comparison strategy, and the numeric instructions mandated a matching strategy. 
+However, Experiment 2 could not distinguish benefits of number avoidance from benefits of the comparison selection task.
 In Experiments 3 and 4 we manipulated vagueness and the selection task separately at each level of numerical format.
-Across the two experiments, we found that %there are effects of the selection task mandated by the instruction, with
-the comparison task instructions attracted faster response times than the matching instructions.
+Across the two experiments, we found that the comparison-task instructions attracted faster response times than the matching-task instructions.
 Within the two experiments we found that vagueness exerts benefits when the selection task is \emph{comparison}, but not when the task is \emph{matching}.
@@
 Vagueness may thus have acquired a reputation that it does not deserve.
-To answer to Lipman's question, of why vagueness permeates human language (see our Introduction), may lie in a different direction after all, possibly relating to a combination of pure necessity and benefits for the speaker rather than the hearer (\citeA<see>[chapter 11]{van2010vagueness}, for discussion of some possible directions).
+To answer to Lipman's question, of why vagueness permeates human language (see our Introduction), may lie in a different direction after all, possibly relating to a combination of pure necessity and benefits for the speaker rather than the hearer \cite<see>[chapter 11, for discussion of some possible directions]{van2010vagueness}.
 
@@
 \end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\textcolor{red}{$\beta=??$, $se=??$, $t=??$, $p<??$}
-
-\begin{APAenumerate}
-	\item 
-	\item 
-\end{APAenumerate}
-
-%\clearpage
-%\appendix
-%
-%\begin{sidewaystable}
-%\centering
-%\caption{All Instructions}
-%\label{All Instructions}
-%\begin{tabular}{lp{5cm}llllp{2cm}l}
-%\midrule
-%Experiment&Example instruction&vagueness&quantifier type&symmetry&selection&number of valid targets&definiteness\\
-%\midrule
-%1&Choose the square with 5 dots&crisp&numeric&symmetric&matching&1&definite\\
-%1&Choose the square with few dots&vague&linguistic&asymmetric&comparison&1&definite\\
-%\midrule
-%2&Choose the square with 6 dots&crisp&numeric&symmetric&matching&1&definite\\
-%2&Choose a square with about 10 dots&vague&numeric&symmetric&matching&2&indefinite\\
-%2&Choose the square with the fewest dots&crisp&linguistic&asymmetric&comparison&1&definite\\
-%2&Choose a square with few dots&vague&linguistic&asymmetric&comparison&2&indefinite\\
-%\midrule
-%3&Choose a square with 6 dots&crisp&numeric&symmetric&matching&1&indefinite\\
-%3&Choose a square with fewer than 20 dots&crisp&numeric&asymmetric&comparison&2&indefinite\\
-%3&Choose a square with about 10 dots&vague&numeric&symmetric&matching&2&indefinite\\
-%3&Choose a square with far fewer than 20 dots&vague&numeric&asymmetric&comparison&1&indefinite\\
-%\midrule
-%4&Choose a square with the same number of dots as the target&crisp&linguistic&symmetric&matching&1&indefinite\\
-%4&Choose a square with fewer dots than the target&crisp&linguistic&asymmetric&comparison&2&indefinite\\
-%4&Choose a square with about the same number of dots as the target&vague&linguistic&symmetric&matching&2&indefinite\\
-%4&Choose a square with far fewer dots than the target&vague&linguistic&asymmetric&comparison&1&indefinite\\
-%\midrule
-%\end{tabular}
-%\end{sidewaystable}
-%
-%\clearpage
-%
-%\begin{table}[htbp]
-%\caption{Table of instructions for the pair (5,25). Experiment 1}
-%\label{instructionse1}
-%\begin{tabular}{rl}
-%\toprule
-%vagueness&example\\
-%\midrule
-%crisp 	& 	Choose the square with 5 dots \\
-%vague	&	Choose the square with few dots\\
-%\bottomrule
-%\end{tabular}
-%\end{table}
-%
-%\begin{figure}[tbp]
-%\fitfigure{images/stimuluse1}
-%\caption{An example stimulus from Experiment 1.
-First the referring expression was presented (left panel).
-Then after a keypress, and a fixation cross (not pictured) the squares and dots were presented without repetition of the referring expression (right panel)}
-%\label{stimuluse1}
-%\end{figure}
-%
-%\begin{figure}[htbp]
-%\fitfigure{images/resultse1.pdf}
-%\caption{Experiment 1 results. }
-%\label{resultse1}
-%\end{figure}
-%
-%\begin{table}[htbp]
-%\caption{Table of instructions arranged by condition for the triple (6,15,24). Experiment 2}
-%\label{instructionse2}
-%\begin{tabular}{rlll}
-%\toprule
-%vagueness&instruction format&selection task&example\\
-%\midrule
-%vague	&	numeric 	& matching	&	Choose a square with about 10 dots\\
-%vague	&	verbal	& comparison	&	Choose a square with few dots\\
-%crisp 	& 	numeric	& matching	&	Choose the square with 6 dots \\
-%crisp		&	verbal	& comparison	&	Choose the square with the fewest dots\\
-%\bottomrule
-%\end{tabular}
-%\end{table}
-%
-%\begin{figure}[tbp]
-%\centering
-%\includegraphics[width=.4\textwidth]{images/stimuluse2}
-%\caption{An example stimulus from Experiment 2}
-%\label{stimuluse2}
-%\end{figure}
-%
-%
-%\begin{figure}[htbp]
-%\centering
-%\includegraphics[width=.65\textwidth]{images/response-time-sep-1.pdf}
-%\caption{Response time results for Experiment 2}
-%\label{resultse2}
-%\end{figure}
-%
-%\begin{figure}[htbp]
-%\centering
-%\includegraphics[]{images/histbd4-ggplot-version-1.pdf}
-%\caption{Expt 2: Distribution of Nearest match, borderline case, and furthest match responses}
-%\label{borderline-response-distribution-e2}
-%\end{figure}
-%
-%\begin{figure}[htbp]
-%\centering
-%\includegraphics[width=.5\textwidth]{images/response-time-two-panels-1.pdf}
-%\caption{Results for Experiment 3}
-%\label{resultse3}
-%\end{figure}
-%
-%\begin{table}[htbp]
-%\centering
-%\caption{Experiment 3: Instructions, assuming $6,15,24$ dots as the Item, and showing \emph{fewer} instead of \emph{more}}
-%\label{Instructions for e3}
-%\begin{tabular}{llll}
-%\hline
-%vagueness&instruction format&selection task&instruction\\
-%\hline
-%crisp & numeric&matching & Choose a square with 6 dots \\ 
-%crisp & numeric&comparison & Choose a square with fewer than 20 dots \\
-%vague & numeric&matching & Choose a square with about 10 dots \\ 
-%vague & numeric&comparison & Choose a square with far fewer than 20 dots \\ 
-%\hline
-%\end{tabular}
-%\end{table}%
-%
-%
-%\begin{table*}[htbp]
-%\centering
-%\caption{Experiment 4: Instructions}
-%\label{Instructions for e4}
-%\begin{tabular}{lllp{7cm}}
-%\hline
-%vagueness&instruction format& selection task&instruction\\
-%\hline
-%crisp & verbal&matching & Choose a square with the same number of dots as the target \\ 
-%crisp & verbal&comparison& Choose a square with fewer dots than the target \\
-%vague & verbal&matching & Choose a square with about the same number of dots as the target \\ 
-%vague & verbal&comparison& Choose a square with far fewer dots than the target \\ 
-%\hline
-%\end{tabular}
-%\end{table*}%
-%
-%\begin{figure}[htbp]
-%\centering
-%\includegraphics[width=.5\textwidth]{images/response-time-two-panels-1-e4.pdf}
-%\caption{Results for Experiment 4}
-%\label{resultse4}
-%\end{figure}
-%crisp  numeric matching  Choose a square with 6 dots 
-%
-%crisp  numeric comparison  Choose a square with fewer than 20 dots 
-%
-%vague  numeric matching  Choose a square with about 10 dots 
-%
-%vague  numeric comparison  Choose a square with far fewer than 20 dots 
-%
-%crisp  verbal matching  Choose a square with the same number of dots as the target 
-%
-%crisp  verbal comparison Choose a square with fewer dots than the target 
-%
-%vague  verbal matching  Choose a square with about the same number of dots as the target 
-%
-%vague  verbal comparison Choose a square with far fewer dots than the target
-
-%(H1) A main effect RT advantage for vagueness.
-%
-%(H2) An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
-%
-%(H3) No large main effect of instruction format (numeric versus verbal). 
-%%MG If we are right that vagueness was responsible for the vagueness advantage in e 1, that could also have been a number avoidance advantage, then we would not expect that number avoidance, or, instruction format, plays a role (?)
-%
-%(H4) On the basis of Experiment 1, we would expect faster responses to more discriminable triples (i.e., an effect of triple).
-%
-%(H5) Participants should make more borderline case choices for vague than crisp instructions.
-
-%(H3$_{rt}$) Numerical distance and vagueness interacted significantly for larger numerical distances when modelling RT: essentially there were diminishing returns for vagueness as numerical distance increased.% KvD Is it worth saying how they interacted?
-% MG I think this is answered now?
-%The interactions at the different levels of numerical distance were: level one: ($\beta=.001$, $se=.016$, $t=.03$, $p=.974$); level two ($\beta=-.039$, $se=.009$, $t=-4.460$, $p<.001$); level three ($\beta=-.043$, $se=.006$, $t=-7.0$, $p<.001$).
-%In the crisp conditions RTs started out much slower than in the vague conditions, at the smallest numerical distance, but the two conditions converged to very fast times at the largest numerical distance.
-
-%
-%\emph{Vagueness} was sum coded: \emph{vague} $= -.5$, \emph{crisp} $= .5$; \emph{numerical distance} was Helmert coded, i.e., the variable is ordered and each level is compared against the mean of all the previous levels to that point.
-% MG a reviewer said he was not familiar with Helmert contrasts
-%Level one of this contrast is numerical distance 5 compared with numerical distance 10; level two is the mean of numerical distance 5 and numerical distance 10 versus numerical distance 15; and level three is the mean of numerical distances 5, 10 and 15 versus numerical distance 20.
-%
-%(H1$_{rt}$) RTs were faster for vague instructions than for crisp instructions ($\beta=.109$, $se=.022$, $t=4.9$, $p<.001$).
-
-%(H1$_{er}$) The effect of vagueness on error rates approached significance, with the vague instructions leading to fewer errors ($\beta=.307, se=.173, t=1.8, p=.077$).
-
-%In the present experiment 2 we pitted these alternative interpretations against each other in a factorial design.
-
- %Fig \ref{stimuluse1} shows an example stimulus.
-
- %Table \ref{instructionse1} shows examples from each condition.
-
- % KvD Is this addition defining "erroneous" OK? MG Yes, I moved it though
-
- % MG reviews asked for an explanation of the difference for numeric crisp over items. Also reviews mention ratio differences as something that we did not control.
-
- %indicating that response times differed in some way according to which triple was presented. % However further analysis revealed that there was no consistent smooth trend across different triples.
-
- % Missing bibitems: now added
-% Fine 1975. Kit Fine. Vagueness, Truth and Logic. {\em Synthese} {\bf 30}, pp.265-300
-% Edgington 1996. Dorothy Edgington. Vagueness by Degrees. In Keefe and Smith 1997.
-% Zadeh 1965. Lotfi Zadeh. Fuzzy Sets. {\em Information and Control} 8 (3) 1965, 338-353.
-% Ioannis Konstas and Mirella Lapata. 2013. A global model for concept-to-text generation. J. Artif. Intell. Res. (JAIR), 48:305Ð346.
-
-%The \emph{instruction format} and \emph{vagueness} variables were sum-coded and \emph{Item} was centred.
-%The fixed effects in the model were \emph{instruction format} and \emph{vagueness} and their interaction, and \emph{item}.
-%The random effects in the model were \emph{participant}, and slopes over \emph{participant} for \emph{instruction format}, \emph{vagueness}, and their interaction.
-
-\begin{table}[htbp]
-\centering
-\caption{Experiment 4: Instructions and targets}
-\label{Instructions for e4}
-\begin{tabular}{llll}
-vagueness	&	instruction		&	selection							&	instruction: Choose a square with \ldots\\
-			&	format		&	task		&			\\
-\toprule
-\multirow{ 2}{*}{crisp} 	 	& \multirow{ 2}{*}{verbal} 	&	matching 		& \ldots the same number of dots as the target \\ 
-	& 	&	comparison										& \ldots fewer dots than the target \\
-	\midrule
-\multirow{ 2}{*}{vague} 	& \multirow{ 2}{*}{verbal} 	&	matching 			& \ldots about the same number of dots as the target \\ 
-	& 	&	comparison										& \ldots far fewer dots than the target \\ 
-\bottomrule
-\end{tabular}
-\end{table}%

### 2016-09-27 experiments/e3/writeup/preprocessing/preprocessing.Rmd
pruned some columns from the data for e3 for upload to pure

@@
 # This data set non-impossible trials and renamed/recomputed variables
-write.table(dd, file="../data.txt")
+dd2 <- subset(dd, select=c(Subject, Trial, Item, c_Itm, Vagueness, c_Vag, Selection, c_Sel, Instruction, Order, Quantity, RT, RT_log, switchResponse))
+write.table(dd2, file="../data.txt")
 ```

### 2016-10-01 experiments/e2/writeup/e2-analysis-extra-stuff.Rmd
a messy commit

---
title: "Experiment 2 of 4"
author: "Matt Green"
output:
  html_document:
    toc: yes
    <!--toc_float: true-->
    mathjax: default
    template: default
    <!--code_folding: show-->
    theme: default
---  
<style>
table, th, td {
    border: 1px grey;
    border-collapse: collapse;
    padding:2px 5px 2px 5px
}  
</style>

```{r setup, include=FALSE, cache=FALSE}
library(knitr)
```

```{r loadFonts, message=FALSE, echo=FALSE}
library(fontcm)
library(extrafont)
#font_import()
#loadfonts()
```

```{r loadLibs, message=FALSE, echo=FALSE}
library(beepr)                     #
library(car)                       #
library(ggplot2)                   #
library(grid)                      #
library(gridExtra)                 #
library(languageR)                 #
library(lattice)                   #
library(latticeExtra)              # for doubleYScale
library(lme4)                      #
library(lmerTest)                  #
library(LMERConvenienceFunctions)  # for perSubjectTrim
library(MASS)                      #
library(plyr)                      #
library(reshape2)                  #
library(rms)                       # for varclus
library(Rmisc)                     # for summarySEwithin and friends
library(xtable)                    #
```

```{r preprocessing}
source('e2-preprocessing.R')
if( !file.exists('e2-b-data.txt') ) {
  dat <- gatherData() 
  dat <- declareImpossibleRT(dat)
  dat <- removeImpossibleTrials(dat)
  dat <- subset(dat, select=c(Subject, Trial, Item, 
                              discriminability, 
                              Instruction,
                              nchar_instr,
                              Vagueness, Number, Order, Quantity, 
                              c_Vag, c_Num, c_Ord, c_Qty,
                              NaType, response_category, 
                              isBorderline, RT)
  )
  write.table(dat, file='e2-b-data.txt', sep='\t', quote=FALSE)
}
dd <- read.delim('e2-b-data.txt')
```

```{r showCompression, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='Ratios for different numbers of dots in the arrays: smaller values are more discriminable. Blue is for the ratio between the smallest number in the array and the largest number in an array. Red is for the mean of two ratios, one for the smallest number to the middle number, the other for the middle number to the largest number in the array', echo=FALSE}
#  Show compression of ratios of numbers of dots in the items
discriminability = sort(unique(as.numeric(dd$discriminability)))
plot(y=c(1.5,1.5,1.5,1.5), 
     x=discriminability, 
     ylab='',
     xlab='',
     yaxt='n',
     type='n', 
     xaxt='n',
     axes=F)
lines(y=c(1.5,1.5,1.5,1.5), x=discriminability, type='o', pch=19)
text(y=c(1.15,1.15,1.15,1.15), x=discriminability, labels=round(discriminability,2),  cex=.7)
mtext('mean (ratio min/med,\nratio med/max)', side = 2, line = 1)
mtext('Discriminability index', side=1, line=3)
mtext('Discriminability for each item', side=3, line=1, cex=1.5)
axis(1)
box(which='plot', col='grey')
```

#Lmer model: before outlier removal

```{r}
dd$RT_log <- log(dd$RT)
dd$c_Trl <- as.numeric(scale(dd$Trial))
```

```{r lmer5, tidy=FALSE, cache=TRUE}
vlmerTest <- lmerTest::lmer(data=dd,
                 RT_log ~
                   c_Vag + c_Num + c_Qty + c_Ord +
                   c_Num:c_Vag:c_Qty +
                   discriminability +
                   c_Trl +
                   #RTprev_log +
                   nchar_instr +
                   (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
```

```{r printXtableV5, echo=FALSE, results='asis'}
print(
  xtable(
    summary(vlmerTest)$coef,
    caption = 'lmerTest version',
    digits=c(0, 3,3,1,1,4)),
  type='html',
  caption.placement='top'
)
```

```{r llmer5, tidy=FALSE, cache=TRUE}
vlme4 <- lme4::lmer(data=dd,
                 RT_log ~
                   c_Vag + c_Num + c_Qty + c_Ord +
                   c_Num:c_Vag:c_Qty +
                   discriminability +
                   c_Trl +
                   #RTprev_log +
                   nchar_instr +
                   (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
```

```{r printXtableV521c, echo=FALSE, results='asis'}
print(
  xtable(
    summary(vlme4)$coef,
    caption = 'lme4 version',
    digits=c(0, 3,3,1)),
  type='html',
  caption.placement='top'
)
```

```{r, echo=1:3}
dfr <- data.frame(coef = summary(vlme4)$coef[,1], 
                  ci2.5  = confint(vlme4, method='Wald')[17:25, 1],
                  ci97.5 = confint(vlme4, method='Wald')[17:25, 2] )
dfr <- dfr[-1,]
dfr <- dfr[rev(rownames(dfr)),]
par(mar=c(5,10,1,1))
plot(y=1:nrow(dfr), 
     x=dfr$coef, 
     xlim=range(dfr$ci2.5, dfr$ci97.5), 
     type='n', axes=F, xlab="", ylab="")
abline(v=0, lty=3, col='grey')
abline(h=c(1,2,3,4,5,6,7,8,9), col='grey', lty=3)
arrows(y0=1:nrow(dfr), y1=1:nrow(dfr), 
       x0=dfr$ci2.5, x1=dfr$ci97.5, 
       length=0.035,
       col='black',
       lwd=2, code=3, angle=90)
points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white')
axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE)
axis(1)
#mtext("Coefficient estimates, in units of log(RT),\nwith 95% confidence intervals", side=1, line=4, cex.lab=1, las=1)
#mtext("Model terms", side=2, line=8)
mtext('estimate with 95% CI', side=1, line=3, cex=.75)
box(which='plot', col='grey')
```

### 2016-10-01 experiments/e2/writeup/e2-b-knitr.Rmd
a messy commit

---
title: "Experiment 2 of 4"
author: "Matt Green"
output:
  html_document:
    toc: yes
    <!--toc_float: true-->
    mathjax: default
    template: default
    <!--code_folding: show-->
    theme: default
---  
<style>
table, th, td {
    border: 1px grey;
    border-collapse: collapse;
    padding:2px 5px 2px 5px
}  
</style>

```{setup, include=FALSE, cache=FALSE}
library(knitr)
```

```{loadFonts, message=FALSE, echo=FALSE}
library(fontcm)
library(extrafont)
#font_import()
#loadfonts()
```

```{loadLibs, message=FALSE, echo=FALSE}
library(beepr)                     #
library(car)                       #
library(ggplot2)                   #
library(grid)                      #
library(gridExtra)                 #
library(languageR)                 #
library(lattice)                   #
library(latticeExtra)              # for doubleYScale
library(lme4)                      #
library(lmerTest)                  #
library(LMERConvenienceFunctions)  # for perSubjectTrim
library(MASS)                      #
library(plyr)                      #
library(reshape2)                  #
library(rms)                       # for varclus
library(Rmisc)                     # for summarySEwithin and friends
library(xtable)                    #
```

```{preprocessing}
source('e2-preprocessing.R')
if( !file.exists('e2-b-data.txt') ) {
  dat <- gatherData() 
  dat <- declareImpossibleRT(dat)
  dat <- removeImpossibleTrials(dat)
  dat <- subset(dat, select=c(Subject, Trial, Item, 
                              discriminability, 
                              Instruction,
                              nchar_instr,
                              Vagueness, Number, Order, Quantity, 
                              c_Vag, c_Num, c_Ord, c_Qty,
                              NaType, response_category, 
                              isBorderline, RT)
  )
  write.table(dat, file='e2-b-data.txt', sep='\t', quote=FALSE)
}
dat <- read.delim('e2-b-data.txt')
dd <- read.delim('e2-b-data.txt')
```

```{showCompression, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='Ratios for different numbers of dots in the arrays: smaller values are more discriminable. Blue is for the ratio between the smallest number in the array and the largest number in an array. Red is for the mean of two ratios, one for the smallest number to the middle number, the other for the middle number to the largest number in the array', echo=FALSE}
#  Show compression of ratios of numbers of dots in the items
discriminability = sort(unique(as.numeric(dd$discriminability)))
plot(y=c(1.5,1.5,1.5,1.5), 
     x=discriminability, 
     ylab='',
     xlab='',
     yaxt='n',
     type='n', 
     xaxt='n',
     axes=F)
lines(y=c(1.5,1.5,1.5,1.5), x=discriminability, type='o', pch=19)
text(y=c(1.15,1.15,1.15,1.15), x=discriminability, labels=round(discriminability,2),  cex=.7)
mtext('mean (ratio min/med,\nratio med/max)', side = 2, line = 1)
mtext('Discriminability index', side=1, line=3)
mtext('Discriminability for each item', side=3, line=1, cex=1.5)
axis(1)
box(which='plot', col='grey')
```

#Lmer model: before outlier removal

```{rmme}
load("data_processed.Rda")
```

```{lmer5, tidy=FALSE, cache=TRUE}
v5 <- lme4::lmer(data=dd,
                 RT_log ~
                   c_Vag + c_Num + c_Qty + c_Ord +
                   c_Num:c_Vag:c_Qty +
                   discriminability +
                   s_Trl +
                   RTprev_log +
                   nchar_instr +
                   (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
```

```{summaryV5, echo=FALSE}
summary(v5)
```

```{printXtableV5, echo=FALSE, results='asis'}
print(
  xtable(
    coef(summary(v5)
         ),
  caption = 'xtable v5'
  ),
  table.placement='htbp',
  latex.environments='center',
  size='small'
)
```

```{rSquaredInfo, echo=FALSE}
cat("R^2")
cor(fitted(v5), dd$RT_log)^2
```

```{plotModelCoefsAndCis, fig.width=5.5, fig.height=4, fig.cap='Coefficient estimates and their (Wald) 95 per cent confidence intervals', fig.pos='hbtp', echo=FALSE}
# Plot model coefficients and ci's
dfr <- data.frame(coef = summary(v5)$coef[-1, 1], 
                     ci2.5  = confint(v5, method='Wald')[18:26, 1],
                     ci97.5 = confint(v5, method='Wald')[18:26, 2] )
dfr <- dfr[rev(rownames(dfr)),]
par(mar=c(5,10,1,1))
plot(y=1:nrow(dfr), 
     x=dfr$coef, 
     xlim=range(dfr$ci2.5, dfr$ci97.5), 
     type='n', axes=F, xlab="", ylab="")
abline(v=0, lty=3, col='grey')
abline(h=c(1,2,3,4,5,6,7,8,9), col='grey', lty=3)
arrows(y0=1:nrow(dfr), y1=1:nrow(dfr), 
       x0=dfr$ci2.5, x1=dfr$ci97.5, 
       length=0.035,
       col='black',
       lwd=2, code=3, angle=90)
points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white')
axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE)
axis(1)
mtext("Coefficient estimates, in units of log(RT),\nwith 95% confidence intervals", side=1, line=3.5, cex.lab=1, las=1)
mtext("Model terms", side=2, line=8)
box(which='plot', col='grey')

```

```{plotLMERfnc, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='plotMLERfnc', echo=TRUE}
par(mfrow=c(2,4))
plotLMER.fnc(v5)

```

```{baayenPLots99, fig.width=6, fig.height=4, fig.cap='Baayen Model Criticism Plots', fig.pos='hbtp', echo=FALSE}
# Baayen 4-plot model criticism
par(mfrow=c(1,3), pty='s')
# create scaled residuals
dd$rstand = as.vector(scale(resid(v5)))
# plot scaled residuals density
plot(density(dd$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(dd$rstand, cex=.5)
qqline(dd$rstand)
# plot standardised residuals versus fitted values
plot(dd$rstand ~ fitted(v5), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))

```

#lmerTest Version}

```{lmerTestVersionOfModel, tidy=FALSE, cache=TRUE, echo=TRUE}
v6 <- lmerTest::lmer(data=dd,
                     RT_log ~
                       c_Vag + c_Num + c_Qty + c_Ord +
                       c_Num:c_Vag:c_Qty +
                       discriminability +
                       s_Trl +
                       RTprev_log +
                       nchar_instr +
                       (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
```

```{summaryLmerTest, cache=TRUE}
summary(v6)
```

#Lmer model: after outlier removal
not done yet.

#Borderline responses

```{barplotBorderline, fig.width=6, fig.height=4, echo=FALSE}
dd$response_category <- relevel(dd$response_category, ref = 'expected')
barplot(table(dd$Vagueness,dd$response_category), beside=TRUE, legend=TRUE, space=c(0,.5))

```

```{tabBdl, echo=FALSE, results='asis'}
print(xtable(table(dd$response_category,dd$Vagueness),
             caption='Borderline cases counts'),
      table.placement='htbp',
      latex.environments='center',
      size='small')
```

### 2016-10-01 experiments/e2/writeup/e2-b-knitr.tex
a messy commit

--- 2016-08-12 experiments/e2/writeup/e2-b-knitr.tex
@@
 \newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+\let\hlipl\hlkwb
 
@@
 
-\clearpage
 \section{Get the data}
@@
 \begin{alltt}
-\hlcom{# if the file data_processed.Rda already exists then load it, else do data wrangling}
-\hlkwa{if} \hlstd{(}\hlkwd{file.exists}\hlstd{(}\hlstr{"data_processed.Rda"}\hlstd{)) \{}
-    \hlkwd{load}\hlstd{(}\hlstr{"data_processed.Rda"}\hlstd{)}
-\hlstd{\}} \hlkwa{else} \hlstd{\{}
-    \hlcom{# declare local variables}
-    \hlstd{number_of_valid_subjects} \hlkwb{<-} \hlnum{30}  \hlcom{# = 30}
-    \hlstd{number_of_rows} \hlkwb{<-} \hlnum{7680}  \hlcom{# 7680}
-    \hlstd{number_of_trials_per_subject} \hlkwb{<-} \hlstd{number_of_rows}\hlopt{/}\hlstd{number_of_valid_subjects}  \hlcom{# = 256}
-    \hlcom{# call functions}
-    \hlstd{dat} \hlkwb{<-} \hlkwd{gatherData}\hlstd{(number_of_valid_subjects)}  \hlcom{# = 30}
-    \hlstd{dat} \hlkwb{<-} \hlkwd{classifyResponses}\hlstd{(dat)}  \hlcom{# classify the response as expected, near, or far}
-    \hlstd{dat} \hlkwb{<-} \hlkwd{processData}\hlstd{(dat)}  \hlcom{# remove impossible trials and re-do previous rt measures}
-    \hlstd{dd} \hlkwb{<-} \hlkwd{postProcessData}\hlstd{(dat)}
-\hlstd{\}}  \hlcom{# end else do data wrangling}
+\hlkwd{source}\hlstd{(}\hlstr{"e2-preprocessing.R"}\hlstd{)}
+\hlkwa{if} \hlstd{(}\hlopt{!}\hlkwd{file.exists}\hlstd{(}\hlstr{"e2-b-data.txt"}\hlstd{)) \{}
+    \hlstd{dat} \hlkwb{<-} \hlkwd{gatherData}\hlstd{()}
+    \hlstd{dat} \hlkwb{<-} \hlkwd{declareImpossibleRT}\hlstd{(dat)}
+    \hlstd{dat} \hlkwb{<-} \hlkwd{removeImpossibleTrials}\hlstd{(dat)}
+    \hlstd{dat} \hlkwb{<-} \hlkwd{subset}\hlstd{(dat,} \hlkwc{select} \hlstd{=} \hlkwd{c}\hlstd{(Subject, Trial, Item, discriminability, Instruction, nchar_instr, Vagueness,}
+        \hlstd{Number, Order, Quantity, c_Vag, c_Num, c_Ord, c_Qty, NaType, response_category, isBorderline,}
+        \hlstd{RT))}
+    \hlkwd{write.table}\hlstd{(dat,} \hlkwc{file} \hlstd{=} \hlstr{"e2-b-data.txt"}\hlstd{,} \hlkwc{sep} \hlstd{=} \hlstr{"\textbackslash{}t"}\hlstd{,} \hlkwc{quote} \hlstd{=} \hlnum{FALSE}\hlstd{)}
+\hlstd{\}}
+\hlstd{dat} \hlkwb{<-} \hlkwd{read.delim}\hlstd{(}\hlstr{"e2-b-data.txt"}\hlstd{)}
+\hlstd{dd} \hlkwb{<-} \hlkwd{read.delim}\hlstd{(}\hlstr{"e2-b-data.txt"}\hlstd{)}
 \end{alltt}
@@
 \begin{verbatim}
- [1] "id"                "Subject"           "Trial"             "Condition"        
- [5] "Order"             "Quantity"          "Vagueness"         "Number"           
- [9] "Item"              "discriminability"  "c_Trl"             "s_Trl"            
-[13] "c_Itm"             "c_Vag"             "c_Num"             "f_Cnd"            
-[17] "c_Ord"             "c_Qty"             "RT"                "RT_log"           
-[21] "RT_raw"            "RTprev"            "RTprev_log"        "RTprev_raw"       
-[25] "Exp_Num"           "Bline_Num"         "Extr_Num"          "Exp_side"         
-[29] "Bline_side"        "Extr_side"         "response_num"      "response_side"    
-[33] "response_category" "Left"              "Mid"               "Right"            
-[37] "Instruction"       "nchar_instr"      
+ [1] "Subject"           "Trial"             "Item"              "discriminability" 
+ [5] "Instruction"       "nchar_instr"       "Vagueness"         "Number"           
+ [9] "Order"             "Quantity"          "c_Vag"             "c_Num"            
+[13] "c_Ord"             "c_Qty"             "NaType"            "response_category"
+[17] "isBorderline"      "RT"               
 \end{verbatim}
@@
 \begin{verbatim}
-        id Subject Trial Condition Order Quantity Vagueness  Number     Item discriminability
-1 s01:t001     s01     1         2  RtoL    Small     Crisp Numeric 06:15:24        0.4875000
-2 s01:t002     s01     2         1  LtoR    Large     Vague Numeric 16:25:34        0.3123529
-3 s01:t003     s01     3         4  LtoR    Small     Crisp  Verbal 26:35:44        0.2308442
-4 s01:t004     s01     4         1  RtoL    Large     Vague Numeric 36:45:54        0.1833333
-5 s01:t005     s01     5         1  RtoL    Small     Vague Numeric 06:15:24        0.4875000
-6 s01:t006     s01     6         3  LtoR    Large     Vague  Verbal 16:25:34        0.3123529
-   c_Trl     s_Trl c_Itm c_Vag c_Num f_Cnd c_Ord c_Qty   RT   RT_log RT_raw RTprev RTprev_log
-1 -127.5 -1.725186 -0.75  -0.5  -0.5 Cr:Nm   0.5  -0.5 1517 7.324490   1517   1517   7.324490
-2 -126.5 -1.711655 -0.25   0.5  -0.5 Vg:Nm  -0.5   0.5 1920 7.560080   1920   1517   7.324490
-3 -125.5 -1.698124  0.25  -0.5   0.5 Cr:Vb  -0.5  -0.5 2346 7.760467   2346   1920   7.560080
-4 -124.5 -1.684593  0.75   0.5  -0.5 Vg:Nm   0.5   0.5 1773 7.480428   1773   2346   7.760467
-5 -123.5 -1.671062 -0.75   0.5  -0.5 Vg:Nm   0.5  -0.5 2556 7.846199   2556   1773   7.480428
-6 -122.5 -1.657531 -0.25   0.5   0.5 Vg:Vb  -0.5   0.5 2043 7.622175   2043   2556   7.846199
-  RTprev_raw Exp_Num Bline_Num Extr_Num Exp_side Bline_side Extr_side response_num response_side
-1       1517       6        15       24    right        mid      left            6         right
-2       1517      34        25       16    right        mid      left           25           mid
-3       1920      26        35       44     left        mid     right           26          left
-4       2346      54        45       36     left        mid     right           45           mid
-5       1773       6        15       24    right        mid      left            6         right
-6       2556      34        25       16    right        mid      left           34         right
-  response_category Left Mid Right                            Instruction nchar_instr
-1          expected   24  15     6          Choose the square with 6 dots          29
-2        borderline   16  25    34     Choose a square with about 30 dots          34
-3          expected   26  35    44 Choose the square with the fewest dots          38
-4        borderline   54  45    36     Choose a square with about 50 dots          34
-5          expected   24  15     6     Choose a square with about 10 dots          34
-6          expected   16  25    34         Choose a square with many dots          30
+  Subject Trial     Item discriminability                            Instruction nchar_instr
+1     s01     1 06:15:24        0.4875000          Choose the square with 6 dots          29
+2     s01     2 16:25:34        0.3123529     Choose a square with about 30 dots          34
+3     s01     3 26:35:44        0.2308442 Choose the square with the fewest dots          38
+4     s01     4 36:45:54        0.1833333     Choose a square with about 50 dots          34
+5     s01     5 06:15:24        0.4875000     Choose a square with about 10 dots          34
+6     s01     6 16:25:34        0.3123529         Choose a square with many dots          30
+  Vagueness  Number Order Quantity c_Vag c_Num c_Ord c_Qty      NaType response_category
+1     Crisp Numeric  RtoL    Small  -0.5  -0.5   0.5  -0.5 valid_trial          expected
+2     Vague Numeric  LtoR    Large   0.5  -0.5  -0.5   0.5 valid_trial        borderline
+3     Crisp  Verbal  LtoR    Small  -0.5   0.5  -0.5  -0.5 valid_trial          expected
+4     Vague Numeric  RtoL    Large   0.5  -0.5   0.5   0.5 valid_trial        borderline
+5     Vague Numeric  RtoL    Small   0.5  -0.5   0.5  -0.5 valid_trial          expected
+6     Vague  Verbal  LtoR    Large   0.5   0.5  -0.5   0.5 valid_trial          expected
+  isBorderline   RT
+1        FALSE 1517
+2         TRUE 1920
+3        FALSE 2346
+4         TRUE 1773
+5        FALSE 2556
+6        FALSE 2043
 \end{verbatim}
@@
 \begin{verbatim}
-        id          Subject         Trial         Condition    Order       Quantity    Vagueness   
- s01:t001:   1   s01    : 256   Min.   :  1.0   Min.   :1.0   LtoR:3838   Small:3840   Crisp:3840  
- s01:t002:   1   s02    : 256   1st Qu.: 65.0   1st Qu.:2.0   RtoL:3839   Large:3837   Vague:3837  
- s01:t003:   1   s03    : 256   Median :129.0   Median :3.0                                        
- s01:t004:   1   s04    : 256   Mean   :128.5   Mean   :2.5                                        
- s01:t005:   1   s05    : 256   3rd Qu.:193.0   3rd Qu.:4.0                                        
- s01:t006:   1   s06    : 256   Max.   :256.0   Max.   :4.0                                        
- (Other) :7671   (Other):6141                                                                      
-     Number           Item      discriminability     c_Trl                s_Trl           
- Numeric:3838   06:15:24:1919   Min.   :0.1833   Min.   :-127.50000   Min.   :-1.7251858  
- Verbal :3839   16:25:34:1919   1st Qu.:0.2308   1st Qu.: -63.50000   1st Qu.:-0.8592102  
-                26:35:44:1920   Median :0.2308   Median :   0.50000   Median : 0.0067654  
-                36:45:54:1919   Mean   :0.3035   Mean   :   0.02377   Mean   : 0.0003217  
-                                3rd Qu.:0.3124   3rd Qu.:  64.50000   3rd Qu.: 0.8727411  
-                                Max.   :0.4875   Max.   : 127.50000   Max.   : 1.7251858
+    Subject         Trial             Item      discriminability
+ s01    : 256   Min.   :  1.0   06:15:24:1919   Min.   :0.1833  
+ s02    : 256   1st Qu.: 65.0   16:25:34:1919   1st Qu.:0.2308  
+ s03    : 256   Median :129.0   26:35:44:1920   Median :0.2308  
+ s04    : 256   Mean   :128.5   36:45:54:1919   Mean   :0.3035  
+ s05    : 256   3rd Qu.:193.0                   3rd Qu.:0.3124  
+ s06    : 256   Max.   :256.0                   Max.   :0.4875  
+ (Other):6141                                                   
+                                 Instruction    nchar_instr    Vagueness        Number    
+ Choose a square with few dots         : 960   Min.   :29.00   Crisp:3840   Numeric:3838  
+ Choose the square with the fewest dots: 960   1st Qu.:30.00   Vague:3837   Verbal :3839  
+ Choose the square with the most dots  : 960   Median :30.00                              
+ Choose a square with many dots        : 959   Mean   :32.59                              
+ Choose a square with about 30 dots    : 480   3rd Qu.:36.00                              
+ Choose a square with about 40 dots    : 480   Max.   :38.00                              
+ (Other)                               :2878                                              
+  Order       Quantity        c_Vag                c_Num               c_Ord          
+ LtoR:3838   Large:3837   Min.   :-0.5000000   Min.   :-5.00e-01   Min.   :-5.00e-01  
+ RtoL:3839   Small:3840   1st Qu.:-0.5000000   1st Qu.:-5.00e-01   1st Qu.:-5.00e-01  
+                          Median :-0.5000000   Median : 5.00e-01   Median : 5.00e-01  
+                          Mean   :-0.0001954   Mean   : 6.51e-05   Mean   : 6.51e-05  
+                          3rd Qu.: 0.5000000   3rd Qu.: 5.00e-01   3rd Qu.: 5.00e-01  
+                          Max.   : 0.5000000   Max.   : 5.00e-01   Max.   : 5.00e-01
 
-     c_Itm               c_Vag                c_Num             f_Cnd          c_Ord          
- Min.   :-7.50e-01   Min.   :-0.5000000   Min.   :-5.00e-01   Vg:Nm:1918   Min.   :-5.00e-01  
- 1st Qu.:-2.50e-01   1st Qu.:-0.5000000   1st Qu.:-5.00e-01   Cr:Nm:1920   1st Qu.:-5.00e-01  
- Median : 2.50e-01   Median :-0.5000000   Median : 5.00e-01   Vg:Vb:1919   Median : 5.00e-01  
- Mean   : 3.26e-05   Mean   :-0.0001954   Mean   : 6.51e-05   Cr:Vb:1920   Mean   : 6.51e-05  
- 3rd Qu.: 2.50e-01   3rd Qu.: 0.5000000   3rd Qu.: 5.00e-01                3rd Qu.: 5.00e-01  
- Max.   : 7.50e-01   Max.   : 0.5000000   Max.   : 5.00e-01                Max.   : 5.00e-01
-
-     c_Qty                  RT            RT_log           RT_raw          RTprev     
- Min.   :-0.5000000   Min.   :  445   Min.   : 6.098   Min.   :  445   Min.   :  445  
- 1st Qu.:-0.5000000   1st Qu.: 1240   1st Qu.: 7.123   1st Qu.: 1240   1st Qu.: 1240  
- Median :-0.5000000   Median : 1727   Median : 7.454   Median : 1727   Median : 1727  
- Mean   :-0.0001954   Mean   : 2840   Mean   : 7.595   Mean   : 2840   Mean   : 2835  
- 3rd Qu.: 0.5000000   3rd Qu.: 2699   3rd Qu.: 7.901   3rd Qu.: 2699   3rd Qu.: 2697  
- Max.   : 0.5000000   Max.   :42685   Max.   :10.662   Max.   :42685   Max.   :42685
-
-   RTprev_log       RTprev_raw       Exp_Num     Bline_Num     Extr_Num    Exp_side        
- Min.   : 6.098   Min.   :  445   Min.   : 6   Min.   :15   Min.   : 6   Length:7677       
- 1st Qu.: 7.123   1st Qu.: 1240   1st Qu.:16   1st Qu.:25   1st Qu.:24   Class :character  
- Median : 7.454   Median : 1727   Median :26   Median :35   Median :34   Mode  :character  
- Mean   : 7.594   Mean   : 2835   Mean   :30   Mean   :30   Mean   :30                     
- 3rd Qu.: 7.900   3rd Qu.: 2697   3rd Qu.:36   3rd Qu.:35   3rd Qu.:44                     
- Max.   :10.662   Max.   :42685   Max.   :54   Max.   :45   Max.   :54
-
-  Bline_side         Extr_side          response_num   response_side  response_category
- Length:7677        Length:7677        Min.   : 6.00   left :3215    borderline:1274   
- Class :character   Class :character   1st Qu.:24.00   mid  :1274    expected  :6108   
- Mode  :character   Mode  :character   Median :34.00   right:3188    extreme   : 295   
-                                       Mean   :30.87                                   
-                                       3rd Qu.:44.00                                   
-                                       Max.   :54.00
-
-      Left         Mid         Right                                    Instruction  
- Min.   : 6   Min.   :15   Min.   : 6   Choose a square with few dots         : 960  
- 1st Qu.:24   1st Qu.:25   1st Qu.:24   Choose the square with the fewest dots: 960  
- Median :34   Median :35   Median :26   Choose the square with the most dots  : 960  
- Mean   :30   Mean   :30   Mean   :30   Choose a square with many dots        : 959  
- 3rd Qu.:44   3rd Qu.:35   3rd Qu.:36   Choose a square with about 30 dots    : 480  
- Max.   :54   Max.   :45   Max.   :54   Choose a square with about 40 dots    : 480  
-                                        (Other)                               :2878  
-  nchar_instr   
- Min.   :29.00  
- 1st Qu.:30.00  
- Median :30.00  
- Mean   :32.59  
- 3rd Qu.:36.00  
- Max.   :38.00
+     c_Qty                    NaType      response_category isBorderline          RT       
+ Min.   :-0.5000000   valid_trial:7677   borderline:1274    Mode :logical   Min.   :  445  
+ 1st Qu.:-0.5000000                      expected  :6108    FALSE:6403      1st Qu.: 1240  
+ Median :-0.5000000                      extreme   : 295    TRUE :1274      Median : 1727  
+ Mean   :-0.0001954                                         NA's :0         Mean   : 2840  
+ 3rd Qu.: 0.5000000                                                         3rd Qu.: 2699  
+ Max.   : 0.5000000                                                         Max.   :42685
 
@@
 \begin{verbatim}
-'data.frame':	7677 obs. of  38 variables:
- $ id               : Factor w/ 7680 levels "s01:t001","s01:t002",..: 1 2 3 4 5 6 7 8 9 10 ...
+'data.frame':	7677 obs. of  18 variables:
  $ Subject          : Factor w/ 30 levels "s01","s02","s03",..: 1 1 1 1 1 1 1 1 1 1 ...
  $ Trial            : int  1 2 3 4 5 6 7 8 9 10 ...
- $ Condition        : int  2 1 4 1 1 3 3 4 4 2 ...
- $ Order            : Factor w/ 2 levels "LtoR","RtoL": 2 1 1 2 2 1 2 2 1 2 ...
- $ Quantity         : Factor w/ 2 levels "Small","Large": 1 2 1 2 1 2 1 1 2 2 ...
+ $ Item             : Factor w/ 4 levels "06:15:24","16:25:34",..: 1 2 3 4 1 2 3 4 1 2 ...
+ $ discriminability : num  0.487 0.312 0.231 0.183 0.487 ...
+ $ Instruction      : Factor w/ 17 levels "Choose a square with about 10 dots",..: 15 3 16 5 1 7 6 16 17 11 ...
+ $ nchar_instr      : int  29 34 38 34 34 30 29 38 36 30 ...
  $ Vagueness        : Factor w/ 2 levels "Crisp","Vague": 1 2 1 2 2 2 2 1 1 1 ...
  $ Number           : Factor w/ 2 levels "Numeric","Verbal": 1 1 2 1 1 2 2 2 2 1 ...
- $ Item             : Factor w/ 4 levels "06:15:24","16:25:34",..: 1 2 3 4 1 2 3 4 1 2 ...
- $ discriminability : num  0.487 0.312 0.231 0.183 0.487 ...
- $ c_Trl            : num  -128 -126 -126 -124 -124 ...
- $ s_Trl            : num  -1.73 -1.71 -1.7 -1.68 -1.67 ...
- $ c_Itm            : num  -0.75 -0.25 0.25 0.75 -0.75 -0.25 0.25 0.75 -0.75 -0.25 ...
+ $ Order            : Factor w/ 2 levels "LtoR","RtoL": 2 1 1 2 2 1 2 2 1 2 ...
+ $ Quantity         : Factor w/ 2 levels "Large","Small": 2 1 2 1 2 1 2 2 1 1 ...
  $ c_Vag            : num  -0.5 0.5 -0.5 0.5 0.5 0.5 0.5 -0.5 -0.5 -0.5 ...
  $ c_Num            : num  -0.5 -0.5 0.5 -0.5 -0.5 0.5 0.5 0.5 0.5 -0.5 ...
- $ f_Cnd            : Factor w/ 4 levels "Vg:Nm","Cr:Nm",..: 2 1 4 1 1 3 3 4 4 2 ...
  $ c_Ord            : num  0.5 -0.5 -0.5 0.5 0.5 -0.5 0.5 0.5 -0.5 0.5 ...
  $ c_Qty            : num  -0.5 0.5 -0.5 0.5 -0.5 0.5 -0.5 -0.5 0.5 0.5 ...
+ $ NaType           : Factor w/ 1 level "valid_trial": 1 1 1 1 1 1 1 1 1 1 ...
+ $ response_category: Factor w/ 3 levels "borderline","expected",..: 2 1 2 1 2 2 2 2 2 1 ...
+ $ isBorderline     : logi  FALSE TRUE FALSE TRUE FALSE FALSE ...
  $ RT               : int  1517 1920 2346 1773 2556 2043 2384 3078 1760 2218 ...
- $ RT_log           : num  7.32 7.56 7.76 7.48 7.85 ...
- $ RT_raw           : int  1517 1920 2346 1773 2556 2043 2384 3078 1760 2218 ...
- $ RTprev           : int  1517 1517 1920 2346 1773 2556 2043 2384 3078 1760 ...
- $ RTprev_log       : num  7.32 7.32 7.56 7.76 7.48 ...
- $ RTprev_raw       : int  1517 1517 1920 2346 1773 2556 2043 2384 3078 1760 ...
- $ Exp_Num          : num  6 34 26 54 6 34 26 36 24 34 ...
- $ Bline_Num        : num  15 25 35 45 15 25 35 45 15 25 ...
- $ Extr_Num         : num  24 16 44 36 24 16 44 54 6 16 ...
- $ Exp_side         : chr  "right" "right" "left" "left" ...
- $ Bline_side       : chr  "mid" "mid" "mid" "mid" ...
- $ Extr_side        : chr  "left" "left" "right" "right" ...
- $ response_num     : int  6 25 26 45 6 34 26 36 24 25 ...
- $ response_side    : Factor w/ 3 levels "left","mid","right": 3 2 1 2 3 3 3 3 3 2 ...
- $ response_category: Factor w/ 3 levels "borderline","expected",..: 2 1 2 1 2 2 2 2 2 1 ...
- $ Left             : int  24 16 26 54 24 16 44 54 6 34 ...
- $ Mid              : int  15 25 35 45 15 25 35 45 15 25 ...
- $ Right            : int  6 34 44 36 6 34 26 36 24 16 ...
- $ Instruction      : Factor w/ 17 levels "Choose a square with about 10 dots",..: 15 3 16 5 1 7 6 16 17 11 ...
- $ nchar_instr      : int  29 34 38 34 34 30 29 38 36 30 ...
 \end{verbatim}
@@
 \clearpage
+\subsection{Consider using log RT}
 
 \begin{knitrout}\scriptsize
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 
-{\centering \includegraphics[width=\maxwidth]{figure/graphics-plotDistFreqPolys-1}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `*tmp*`[[j]]: subscript out of bounds}}
 
-}
+{\ttfamily\noindent\color{warningcolor}{Warning in mean.default(dat\_transforms[dat\_transforms\$dfSubset == k \& dat\_transforms\$RTtype == : argument is not numeric or logical: returning NA}}
 
-\caption[Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions]{Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions}\label{fig:plotDistFreqPolys}
-\end{figure}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[<-.data.frame`(`*tmp*`, dat\_transforms\$dfSubset == k \& dat\_transforms\$RTtype == : replacement has length zero}}\end{kframe}
+\end{knitrout}
 
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error: measure variables not found in data: RT\_log, RT\_raw}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in ggplot(temp, aes(time, colour = RTtype)): object 'temp' not found}}\end{kframe}
 \end{knitrout}
@@
 \clearpage
+\subsection{How logging RT affects the distribution}
 
 \begin{knitrout}\scriptsize
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 
-{\centering \includegraphics[width=\maxwidth]{figure/graphics-plotBoxPlots-1}
+{\ttfamily\noindent\color{warningcolor}{Warning in readChar(con, 5L, useBytes = TRUE): cannot open compressed file 'data\_processed.Rda', probable reason 'No such file or directory'}}
 
-}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in readChar(con, 5L, useBytes = TRUE): cannot open the connection}}
 
-\caption[Show how transformations of RT affect distribution of times, and how they affect which times are outliers]{Show how transformations of RT affect distribution of times, and how they affect which times are outliers.}\label{fig:plotBoxPlots}
-\end{figure}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error: measure variables not found in data: RT\_log, RT\_raw}}
 
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in ggplot(subdata): object 'subdata' not found}}\end{kframe}
 \end{knitrout}
@@
 \clearpage
+\subsection{Identify fast and slow subjects and items}
 
 \begin{knitrout}\scriptsize
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 
-{\centering \includegraphics[width=\maxwidth]{figure/graphics-plotFastSlowRTs-1}
+{\ttfamily\noindent\color{warningcolor}{Warning in mean.default(dd\$RT\_log): argument is not numeric or logical: returning NA}}
 
-}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in mean(RT\_log): object 'RT\_log' not found}}
 
-\caption[Show how mean times for individual subjects and items vary with respect to the grand mean Log RT]{Show how mean times for individual subjects and items vary with respect to the grand mean Log RT.}\label{fig:plotFastSlowRTs}
-\end{figure}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in subs\$level = 1:30: object 'subs' not found}}
 
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in factor(subs\$effect, levels = c("{}subject"{}, "{}discriminability"{})): object 'subs' not found}}
+
+{\ttfamily\noindent\color{warningcolor}{Warning in mean.default(dd\$RT\_log): argument is not numeric or logical: returning NA}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in mean(RT\_log): object 'RT\_log' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in factor(itms\$effect, levels = c("{}subject"{}, "{}discriminability"{})): object 'itms' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in rbind(subs, itms): object 'subs' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in factor(si\$effect, levels = c("{}subject"{}, "{}discriminability"{})): object 'si' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in unique(si\$emean): object 'si' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(substitute(groups), data, environment(x)): object 'si' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in inherits(obj1, "{}trellis"{}): object 'my1' not found}}\end{kframe}
 \end{knitrout}
@@
 \clearpage
+\subsection{Plot main effects in both transformations}
 
 \begin{knitrout}\scriptsize
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 
-{\centering \includegraphics[width=\maxwidth]{figure/graphics-plotMainsa-1}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error: measure variables not found in data: RT\_raw}}
 
-}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in reshape2::melt(tempraw, id.vars = c("{}transformation"{}, "{}score"{}), : object 'tempraw' not found}}
 
-\caption[Plot main effects on several transformations]{Plot main effects on several transformations}\label{fig:plotMainsa}
-\end{figure}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error: measure variables not found in data: RT\_log}}
 
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in reshape2::melt(templog, id.vars = c("{}transformation"{}, "{}score"{}), : object 'templog' not found}}\end{kframe}
+\end{knitrout}
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in ggplot(tempraw2, aes(y = score, x = level, group = 1, col = effect)): object 'tempraw2' not found}}\end{kframe}
+\end{knitrout}
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in ggplot(templog2, aes(y = score, x = level, group = 1, col = effect)): object 'templog2' not found}}\end{kframe}
 \end{knitrout}
@@
 \clearpage
+\subsection{Main effects in log RT over discriminability}
 
 \begin{knitrout}\scriptsize
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 
-{\centering \includegraphics[width=\maxwidth]{figure/graphics-mainB-1}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
 
-}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in names(nrts)[5] <- "{}Levels"{}: object 'nrts' not found}}
 
-\caption[Just the main effects]{Just the main effects}\label{fig:mainB}
-\end{figure}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
 
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in names(orts)[5] <- "{}Levels"{}: object 'orts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in names(qrts)[5] <- "{}Levels"{}: object 'qrts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in names(vrts)[5] <- "{}Levels"{}: object 'vrts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in rbind(nrts, orts, qrts, vrts): object 'nrts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(expr, envir, enclos): object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(expr, envir, enclos): object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(expr, envir, enclos): object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in ggplot(data = rts, aes(x = discriminability, y = RT\_log, group = Levels)): object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(expr, envir, enclos): object 'p1' not found}}\end{kframe}
 \end{knitrout}
@@
 \clearpage
+\subsection{2-Way interactions over discriminability}
 
 \begin{knitrout}\scriptsize
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 
-{\centering \includegraphics[width=\maxwidth]{figure/graphics-mainD-1}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
 
-}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in names(rts1)[2] <- "{}F1"{}: object 'rts1' not found}}
 
-\caption[Main effects over item ratios]{Main effects over item ratios}\label{fig:mainD}
-\end{figure}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in names(rts1)[3] <- "{}F2"{}: object 'rts1' not found}}
 
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in rts1\$E1 <- "{}Vagueness"{}: object 'rts1' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in rts1\$E2 <- "{}Number"{}: object 'rts1' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in names(rts2)[2] <- "{}F1"{}: object 'rts2' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in names(rts2)[3] <- "{}F2"{}: object 'rts2' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in rts2\$E1 <- "{}Vagueness"{}: object 'rts2' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in rts2\$E2 <- "{}Quantity"{}: object 'rts2' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in names(rts3)[2] <- "{}F1"{}: object 'rts3' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in names(rts3)[3] <- "{}F2"{}: object 'rts3' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in rts3\$E1 <- "{}Vagueness"{}: object 'rts3' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in rts3\$E2 <- "{}Order"{}: object 'rts3' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in names(rts4)[2] <- "{}F1"{}: object 'rts4' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in names(rts4)[3] <- "{}F2"{}: object 'rts4' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in rts4\$E1 <- "{}Quantity"{}: object 'rts4' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in rts4\$E2 <- "{}Number"{}: object 'rts4' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in names(rts5)[2] <- "{}F1"{}: object 'rts5' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in names(rts5)[3] <- "{}F2"{}: object 'rts5' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in rts5\$E1 <- "{}Order"{}: object 'rts5' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in rts5\$E2 <- "{}Number"{}: object 'rts5' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in names(rts6)[2] <- "{}F1"{}: object 'rts6' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in names(rts6)[3] <- "{}F2"{}: object 'rts6' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in rts6\$E1 <- "{}Quantity"{}: object 'rts6' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in rts6\$E2 <- "{}Order"{}: object 'rts6' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in rbind(rts1, rts2, rts3, rts4, rts5, rts6): object 'rts1' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(expr, envir, enclos): object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(expr, envir, enclos): object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(expr, envir, enclos): object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in paste(rts\$F1, rts\$F2): object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in factor(rts\$E1, levels = c("{}Number"{}, "{}Order"{}, "{}Quantity"{}, "{}Vagueness"{})): object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in factor(rts\$E2, levels = c("{}Number"{}, "{}Order"{}, "{}Quantity"{}, "{}Vagueness"{})): object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in ggplot(rts, aes(y = RT\_log, x = discriminability, group = F3, : object 'rts' not found}}\end{kframe}
 \end{knitrout}
@@
 \clearpage
+\subsection{Vagueness by number interaction over items}
 
 \begin{knitrout}\scriptsize
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 
-{\centering \includegraphics[width=\maxwidth]{figure/graphics-22way-1}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
 
-}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in paste(rts\$Number, rts\$Vagueness, sep = "{} "{}): object 'rts' not found}}
 
-\caption[2-way interactions over item ratios]{2-way interactions over item ratios}\label{fig:22way}
-\end{figure}
-
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in ggplot(rts, aes(y = RT\_log, x = Item, ymin = RT\_log - ci, ymax = RT\_log + : object 'rts' not found}}\end{kframe}
 \end{knitrout}
@@
 \clearpage
+\subsection{Vagueness by number interaction over discriminability}
 
 \begin{knitrout}\scriptsize
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 
-{\centering \includegraphics[width=\maxwidth]{figure/graphics-plotInt1-1}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
 
-}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(expr, envir, enclos): object 'rts' not found}}
 
-\caption[vagueness by number interaction over items]{vagueness by number interaction over items}\label{fig:plotInt1}
-\end{figure}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in paste(rts\$Number, rts\$Vagueness, sep = "{} "{}): object 'rts' not found}}
 
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in ggplot(rts, aes(y = RT\_log, x = discriminability, ymin = RT\_log - : object 'rts' not found}}\end{kframe}
 \end{knitrout}
@@
 \clearpage
+\subsection{3-Way interactions}
 
 \begin{knitrout}\scriptsize
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 
-{\centering \includegraphics[width=\maxwidth]{figure/graphics-plotInt2-1}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
 
-}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(expr, envir, enclos): object 'rts' not found}}
 
-\caption[vagueness by number interaction over item ratios]{vagueness by number interaction over item ratios}\label{fig:plotInt2}
-\end{figure}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(expr, envir, enclos): object 'rts' not found}}
 
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in ggplot(rts, aes(y = RT\_log, x = Number, group = Vagueness, ymin = mins, : object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(expr, envir, enclos): object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(expr, envir, enclos): object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in ggplot(rts, aes(y = RT\_log, x = Number, group = Vagueness, ymin = mins, : object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(expr, envir, enclos): object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(expr, envir, enclos): object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in ggplot(rts, aes(y = RT\_log, x = Number, group = Quantity, ymin = mins, : object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(expr, envir, enclos): object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(expr, envir, enclos): object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in ggplot(rts, aes(y = RT\_log, x = Vagueness, group = Quantity, : object 'rts' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in arrangeGrob(...): object 'p1' not found}}\end{kframe}
 \end{knitrout}
@@
 \clearpage
+\subsection{Vagueness by number by quantity over discriminability}
 
 \begin{knitrout}\scriptsize
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 
-{\centering \includegraphics[width=\maxwidth]{figure/graphics-plot3way-1}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
 
-}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in paste(rts\$Number, rts\$Vagueness, sep = "{} "{}): object 'rts' not found}}
 
-\caption[3 way interactions]{3 way interactions}\label{fig:plot3way}
-\end{figure}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(expr, envir, enclos): object 'rts' not found}}
 
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in ggplot(rts, aes(y = RT\_log, x = discriminability, ymin = RT\_log - : object 'rts' not found}}\end{kframe}
 \end{knitrout}
@@
 \clearpage
+\subsection{4-Way interaction}
 
 \begin{knitrout}\scriptsize
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 
-{\centering \includegraphics[width=\maxwidth]{figure/graphics-vnitems-1}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
 
-}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in paste(rts\$Number, rts\$Vagueness, sep = "{} "{}): object 'rts' not found}}
 
-\caption[Plot vagueness by number by quantity over item ratios]{Plot vagueness by number by quantity over item ratios.}\label{fig:vnitems}
-\end{figure}
-
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in ggplot(rts, aes(y = RT\_log, x = Vagueness, ymin = RT\_log - ci, : object 'rts' not found}}\end{kframe}
 \end{knitrout}
@@
 \clearpage
+\subsection{4-Way interaction split over discriminability}
 
 \begin{knitrout}\scriptsize
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
 
-{\centering \includegraphics[width=\maxwidth]{figure/graphics-4way-1}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in `[.data.frame`(xx, , col): undefined columns selected}}
 
-}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in paste(rts\$Number, rts\$Vagueness, sep = "{} "{}): object 'rts' not found}}
 
-\caption[4 way interaction]{4 way interaction}\label{fig:4way}
-\end{figure}
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in eval(expr, envir, enclos): object 'rts' not found}}
 
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in ggplot(rts, aes(y = RT\_log, x = discriminability, ymin = RT\_log - : object 'rts' not found}}\end{kframe}
 \end{knitrout}
@@
 \clearpage
-
-\begin{knitrout}\scriptsize
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[hbtp]
-
-{\centering \includegraphics[width=\maxwidth]{figure/graphics-4waysplit-1}
-
-}
-
-\caption[4 way interaction split over item discriminability]{4 way interaction split over item discriminability}\label{fig:4waysplit}
-\end{figure}
-
-\end{knitrout}
-
-\clearpage
-\section{Lmer model}
+\section{Lmer model: before outlier removal}
 
@@
 \end{alltt}
-\end{kframe}
+
+{\ttfamily\noindent\color{warningcolor}{Warning in readChar(con, 5L, useBytes = TRUE): cannot open compressed file 'data\_processed.Rda', probable reason 'No such file or directory'}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in readChar(con, 5L, useBytes = TRUE): cannot open the connection}}\end{kframe}
 \end{knitrout}
@@
 
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{verbatim}
+Linear mixed model fit by REML ['lmerMod']
+Formula: RT_log ~ c_Vag + c_Num + c_Qty + c_Ord + c_Num:c_Vag:c_Qty +  
+    discriminability + s_Trl + RTprev_log + nchar_instr + (1 +  
+    c_Vag + c_Num + c_Qty + c_Ord | Subject)
+   Data: dd
+
+REML criterion at convergence: 11474.8
+
+Scaled residuals: 
+    Min      1Q  Median      3Q     Max 
+-4.5470 -0.6351 -0.0955  0.5372  5.0914
+
+Random effects:
+ Groups   Name        Variance Std.Dev. Corr                   
+ Subject  (Intercept) 0.153949 0.39236                         
+          c_Vag       0.001546 0.03932   0.69                  
+          c_Num       0.165314 0.40659  -0.67 -0.64            
+          c_Qty       0.008148 0.09027   0.16  0.26 -0.34      
+          c_Ord       0.001559 0.03949  -0.13  0.02 -0.39 -0.52
+ Residual             0.249734 0.49973                         
+Number of obs: 7677, groups:  Subject, 30
+
+Fixed effects:
+                   Estimate Std. Error t value
+(Intercept)        7.171685   0.122236   58.67
+c_Vag              0.060938   0.013873    4.39
+c_Num             -0.433614   0.075148   -5.77
+c_Qty             -0.069067   0.020047   -3.45
+c_Ord              0.017959   0.013497    1.33
+discriminability  -0.771402   0.049267  -15.66
+s_Trl             -0.106972   0.005807  -18.42
+RTprev_log         0.060469   0.009692    6.24
+nchar_instr        0.006086   0.001944    3.13
+c_Vag:c_Num:c_Qty  0.104949   0.046066    2.28
+
+Correlation of Fixed Effects:
+            (Intr) c_Vag  c_Num  c_Qty  c_Ord  dscrmn s_Trl  RTprv_ nchr_n
+c_Vag        0.083                                                        
+c_Num       -0.369 -0.337                                                 
+c_Qty        0.069  0.117 -0.278                                          
+c_Ord       -0.052  0.007 -0.206 -0.228                                   
+discrmnblty -0.140  0.004 -0.001  0.000  0.001                            
+s_Trl       -0.114  0.003 -0.001 -0.002  0.006  0.017                     
+RTprev_log  -0.607  0.006 -0.004 -0.003  0.019  0.015  0.185              
+nchar_instr -0.524  0.237 -0.034  0.018  0.000  0.017  0.000  0.006       
+c_Vg:c_N:_Q  0.073 -0.033  0.005 -0.003  0.000 -0.002 -0.001 -0.002 -0.138
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
+\clearpage
+
 % latex table generated in R 3.3.1 by xtable 1.8-2 package
-% Fri Aug 12 22:28:46 2016
+% Sat Oct  1 13:19:27 2016
 \begin{table}[htbp]
+\centering
 \begingroup\small
@@
 R^2
-[1] 0.533722
 \end{verbatim}
-\end{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{Error in cor(fitted(v5), dd\$RT\_log): supply both 'x' and 'y' or a matrix-like 'x'}}\end{kframe}
 \end{knitrout}
 
-\clearpage
+%\clearpage
 
@@
 \clearpage
+\section{Lmer model: after outlier removal}
+not done yet.
+\clearpage
 \section{Borderline responses}
@@
 % latex table generated in R 3.3.1 by xtable 1.8-2 package
-% Fri Aug 12 22:28:47 2016
+% Sat Oct  1 13:19:28 2016
 \begin{table}[htbp]
@@
 
+\clearpage
+\appendix
+\section{Functions listing}
+\clearpage
+\subsection{Gather Data}
+
+\begin{knitrout}\scriptsize
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
+\begin{alltt}
+\hlkwd{source}\hlstd{(}\hlstr{"e2-preprocessing.R"}\hlstd{,} \hlkwc{echo} \hlstd{= T)}
+\end{alltt}
+\begin{verbatim}
+
+> gatherData = function() {
++     number_of_valid_subjects = 30
++     number_of_trials_per_subject = 256
++     message("Starting to gather data")
++    .... [TRUNCATED]
+
+> declareImpossibleRT = function(dat) {
++     message("Starting to declare impossible RTs")
++     dat$NaType = "valid_trial"
++     dat$NaType[dat$RT_o .... [TRUNCATED]
+
+> removeImpossibleTrials = function(dat) {
++     message("Starting to remove impossible trials")
++     a = nrow(dat)
++     dat <- dat[complete.cases(d .... [TRUNCATED] 
+\end{verbatim}
+\end{kframe}
+\end{knitrout}
+
 \end{document}

### 2016-10-01 experiments/e2/writeup/e2-preprocessing.R

@@ -1,0 +1,226 @@
+gatherData = function() {
+  number_of_valid_subjects=30
+  number_of_trials_per_subject=256
+  message('Starting to gather data')
+  data_dir <- '../experimentCode/output/'
+  column.headers.df <-  head( read.table(
+    paste(data_dir,'subject01.data',sep=''),
+    header=TRUE),0)
+  gathered.data <- column.headers.df
+  for (subject in 1:number_of_valid_subjects) {
+    current.filename <- paste(data_dir, 'subject', sprintf("%02d",subject), '.data', sep='')
+    current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = TRUE)
+    gathered.data <- rbind(gathered.data, current.data)
+  }
+  dat <- gathered.data
+  dat$Order <- factor(dat$Order, levels=c(1,2), labels=c('LtoR','RtoL'))
+  dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)
+  dat$Quantity <- factor(dat$Quantity, levels=c(1,2), labels=c('Small','Large'))
+  dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)
+  dat[ dat$Condition==1 , 'Vagueness'] <- 'Vague'
+  dat[ dat$Condition==2 , 'Vagueness'] <- 'Crisp'
+  dat[ dat$Condition==3 , 'Vagueness'] <- 'Vague'
+  dat[ dat$Condition==4 , 'Vagueness'] <- 'Crisp'
+  dat$Vagueness <- as.factor(dat$Vagueness)
+  dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)
+  dat[ dat$Condition==1 , 'Number'] <- 'Numeric'
+  dat[ dat$Condition==2 , 'Number'] <- 'Numeric'
+  dat[ dat$Condition==3 , 'Number'] <- 'Verbal'
+  dat[ dat$Condition==4 , 'Number'] <- 'Verbal'
+  dat$Number <- as.factor(dat$Number)
+  dat$c_Num <- ifelse(dat$Number=='Numeric', -.5, .5)
+  dat$Subject <- factor(paste("s",sprintf("%02d",dat$Subject),sep=""))
[…]

### 2016-10-01 experiments/e2/writeup/preprocessing/e2-preprocessing.R

@@ -1,296 +1,0 @@
-## ----gatherDataFunction, echo=FALSE--------------------------------------
-gatherData = function(number_of_valid_subjects) {
-  data_dir <- '../../experimentCode/output/'
-  column.headers.df <-  head( read.table(
-    paste(data_dir,'subject01.data',sep=''),
-    header=TRUE),0)
-  gathered.data <- column.headers.df
-  for (subject in 1:number_of_valid_subjects) {
-    current.filename <- paste(data_dir, 'subject', sprintf("%02d",subject), '.data', sep='')
-    current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = FALSE)
-    gathered.data <- rbind(gathered.data, current.data)
-  }
-  message('Returning gathered data')
-  return(gathered.data)
-} # end of gatherData function
-
-## ----classifyResponseFuntion, echo=FALSE---------------------------------
-classifyResponses = function(dat) {
-  # what were they expected to respond?
-  dat$crossed = as.factor(paste('Con', dat$Condition, ':Quan', dat$Quantity, ':Item', dat$Item, sep=''))
-  dat[dat$crossed=='Con1:Quan1:Item1', 'Exp_Num']       <-  6
-  dat[dat$crossed=='Con1:Quan1:Item1', 'Bline_Num']     <- 15
-  dat[dat$crossed=='Con1:Quan1:Item1', 'Extr_Num']      <- 24
-  dat[dat$crossed=='Con1:Quan1:Item2', 'Exp_Num'] <- 16
-  dat[dat$crossed=='Con1:Quan1:Item2', 'Bline_Num']     <- 25
-  dat[dat$crossed=='Con1:Quan1:Item2', 'Extr_Num']      <- 34
-  dat[dat$crossed=='Con1:Quan1:Item3', 'Exp_Num'] <- 26
-  dat[dat$crossed=='Con1:Quan1:Item3', 'Bline_Num']     <- 35
-  dat[dat$crossed=='Con1:Quan1:Item3', 'Extr_Num']      <- 44
-  dat[dat$crossed=='Con1:Quan1:Item4', 'Exp_Num'] <- 36
[…]

### 2016-10-01 experiments/e2/writeup/runme.R

@@ -1,0 +1,20 @@
+source('e2-preprocessing.R')
+if( file.exists('e2-b-data.txt') ) {
+  dd <- read.delim('e2-b-data.txt')
+} else { 
+  dat <- gatherData() 
+  dat <- declareImpossibleRT(dat)
+  dat <- removeImpossibleTrials(dat)
+  dat <- subset(dat, select=c(Subject, Trial, Item, 
+                              discriminability, 
+                              Instruction,
+                              nchar_instr,
+                              Vagueness, Number, Order, Quantity, 
+                              c_Vag, c_Num, c_Ord, c_Qty,
+                              NaType, response_category, 
+                              isBorderline, RT)
+  )
+  write.table(dat, file='e2-b-data.txt', sep='\t', quote=FALSE)
+  dd <- read.delim('e2-b-data.txt')
+} 
+

### 2016-10-01 experiments/e2/writeup/scratch44.R

@@ -1,0 +1,101 @@
+
+```{r rSquaredInfo, echo=FALSE}
+cat("R^2")
+cor(fitted(v5), dd$RT_log)^2
+```
+
+```{r plotModelCoefsAndCis, fig.width=5.5, fig.height=4, fig.cap='Coefficient estimates and their (Wald) 95 per cent confidence intervals', fig.pos='hbtp', echo=FALSE}
+# Plot model coefficients and ci's
+dfr <- data.frame(coef = summary(v5)$coef[-1, 1], 
+                  ci2.5  = confint(v5, method='Wald')[18:26, 1],
+                  ci97.5 = confint(v5, method='Wald')[18:26, 2] )
+dfr <- dfr[rev(rownames(dfr)),]
+par(mar=c(5,10,1,1))
+plot(y=1:nrow(dfr), 
+     x=dfr$coef, 
+     xlim=range(dfr$ci2.5, dfr$ci97.5), 
+     type='n', axes=F, xlab="", ylab="")
+abline(v=0, lty=3, col='grey')
+abline(h=c(1,2,3,4,5,6,7,8,9), col='grey', lty=3)
+arrows(y0=1:nrow(dfr), y1=1:nrow(dfr), 
+       x0=dfr$ci2.5, x1=dfr$ci97.5, 
+       length=0.035,
+       col='black',
+       lwd=2, code=3, angle=90)
+points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white')
+axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE)
+axis(1)
+mtext("Coefficient estimates, in units of log(RT),\nwith 95% confidence intervals", side=1, line=3.5, cex.lab=1, las=1)
+mtext("Model terms", side=2, line=8)
+box(which='plot', col='grey')
+
+```
+
+
+
+
+```{r plotLMERfnc, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='plotMLERfnc', echo=TRUE}
+par(mfrow=c(2,4))
+plotLMER.fnc(v5)
+
+```
+
+
+
+```{r baayenPLots99, fig.width=6, fig.height=4, fig.cap='Baayen Model Criticism Plots', fig.pos='hbtp', echo=FALSE}
+# Baayen 4-plot model criticism
+par(mfrow=c(1,3), pty='s')
+# create scaled residuals
+dd$rstand = as.vector(scale(resid(v5)))
[…]

### 2016-10-01 experiments/e2/writeup/scratchgather.R

@@ -1,0 +1,197 @@
+number_of_valid_subjects=30
+number_of_trials_per_subject=256
+message('Starting gather data')
+data_dir <- '../experimentCode/output/'
+column.headers.df <-  head( read.table(
+  paste(data_dir,'subject01.data',sep=''),
+  header=TRUE),0)
+gathered.data <- column.headers.df
+for (subject in 1:number_of_valid_subjects) {
+  current.filename <- paste(data_dir, 'subject', sprintf("%02d",subject), '.data', sep='')
+  current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = TRUE)
+  gathered.data <- rbind(gathered.data, current.data)
+}
+dat <- gathered.data
+dat$Order <- factor(dat$Order, levels=c(1,2), labels=c('LtoR','RtoL'))
+dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)
+dat$Quantity <- factor(dat$Quantity, levels=c(1,2), labels=c('Small','Large'))
+dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)
+dat[ dat$Condition==1 , 'Vagueness'] <- 'Vague'
+dat[ dat$Condition==2 , 'Vagueness'] <- 'Crisp'
+dat[ dat$Condition==3 , 'Vagueness'] <- 'Vague'
+dat[ dat$Condition==4 , 'Vagueness'] <- 'Crisp'
+dat$Vagueness <- as.factor(dat$Vagueness)
+dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)
+dat[ dat$Condition==1 , 'Number'] <- 'Numeric'
+dat[ dat$Condition==2 , 'Number'] <- 'Numeric'
+dat[ dat$Condition==3 , 'Number'] <- 'Verbal'
+dat[ dat$Condition==4 , 'Number'] <- 'Verbal'
+dat$Number <- as.factor(dat$Number)
+dat$c_Num <- ifelse(dat$Number=='Numeric', -.5, .5)
+dat$Subject <- factor(paste("s",sprintf("%02d",dat$Subject),sep=""))
[…]

### 2016-10-01 experiments/e2/writeup/e2-analysis-extra-stuff.Rmd
trying to match plot

@@
 
+```{r e2-rtplot, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE, dev='pdf'}
+rtdata=dd
+source('summarySEwithin2.R')
+rtdataplot1 <- summarySEwithin2(rtdata, measurevar="RT_log", withinvars=c("Condition", "Vagueness", "Number", "Item"), idvar="Subject")
+rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness))
+rtplot=ggplot(rtdataplot1, aes(y=RT_logNormed, x=Item, group=Condition, fill=Vagueness)) +
+  geom_line() +
+  geom_errorbar(width=.2, aes(ymin=RT_log-ci, ymax=RT_log+ci)) +
+  geom_point(size=2, col=1, aes(shape=Vagueness)) +
+  scale_fill_grey(name="Vagueness", start=0, end=1) +
+  scale_shape_manual(name="Vagueness", values=c(21,22)) +
+  ggtitle("Response time") +
+  ylab("Respone time (log (ms))") + 
+  xlab(NULL) +
+  theme_bw() +
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        legend.key = element_blank(), legend.position="top",
+        aspect.ratio=1,
+        plot.background=element_rect(fill=NA, color='white'),
+        strip.background=element_blank(),
+        axis.text.x = element_text(angle = 15)) +
+  facet_grid(~Number)
+rtplot
+```
 #Lmer model: before outlier removal

### 2016-10-01 experiments/e2/writeup/e2-analysis-extra-stuff.Rmd
umm

--- 2016-10-01 experiments/e2/writeup/e2-analysis-extra-stuff.Rmd
@@
 
+```{r e2-rtplot, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE, dev='pdf'}
+dd$RT_log=log(dd$RT)
+rtdata=dd
+source('summarySEwithin2.R')
+rtdataplot1 <- summarySEwithin2(rtdata, measurevar="RT_log", 
+                                withinvars=c("Vagueness", "Number", "Item"), idvar="Subject")
+rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness))
+rtplot=ggplot(rtdataplot1, aes(y=RT_logNormed, x=Item, group=Condition, fill=Vagueness)) +
+  geom_line() +
+  geom_errorbar(width=.2, aes(ymin=RT_log-ci, ymax=RT_log+ci)) +
+  geom_point(size=2, col=1, aes(shape=Vagueness)) +
+  scale_fill_grey(name="Vagueness", start=0, end=1) +
+  scale_shape_manual(name="Vagueness", values=c(21,22)) +
+  ggtitle("Response time") +
+  ylab("Respone time (log (ms))") + 
+  xlab(NULL) +
+  theme_bw() +
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        legend.key = element_blank(), legend.position="top",
+        aspect.ratio=1,
+        plot.background=element_rect(fill=NA, color='white'),
+        strip.background=element_blank(),
+        axis.text.x = element_text(angle = 15)) +
+  facet_grid(~Number)
+rtplot
+```
 #Lmer model: before outlier removal

### 2016-10-02 experiments/e2/writeup/e2.Rmd
deleting from e2 some files that dont belong in master

@@
 ```{r loadLibraries, message=FALSE, echo=FALSE, cache=FALSE}
-library(knitr); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions); source('summarySEwithin2.R')
+library(Matrix); library(knitr); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions); source('summarySEwithin2.R'); source('e2-preprocessing.R')
+```
+
+```{r preprocessing}
+source('e2-preprocessing.R')
+if( !file.exists('e2-b-data.txt') ) {
+  dat <- gatherData() 
+  dat <- declareImpossibleRT(dat)
+  #dat <- removeImpossibleTrials(dat)
+  dat <- subset(dat, select=c(Subject, Trial, Item, c_Itm, 
+                              discriminability, 
+                              Instruction,
+                              nchar_instr,
+                              Vagueness, Number, Order, Quantity, 
+                              c_Vag, c_Num, c_Ord, c_Qty,
+                              NaType, response_category, 
+                              isBorderline, RT, RT_log)
+  )
+  write.table(dat, file='e2-b-data.txt', sep='\t', quote=FALSE, row.names = FALSE)
+}
 ```
@@
 ```{r getData, echo=T}
-rawdata <- read.table('data.txt', sep=' ')
-```
-
-```{r}
-tempdata <- perSubjectTrim.fnc(rawdata, response='RT', subject='Subject', trim = 2.5)$data
+rawdata <- read.delim('e2-b-data.txt')
+tempdata <- perSubjectTrim.fnc(rawdata[complete.cases(rawdata),], response='RT', subject='Subject', trim = 2.5)$data
 rtdata <- tempdata
@@
 ```{r computeSummariesForGraphs, cache=TRUE, message=FALSE}
-rtdataplot1 <- summarySEwithin2(rtdata, measurevar="RT_log", withinvars=c("Condition", "Vagueness", "Number", "Item"), idvar="Subject")
+rtdataplot1 <- summarySEwithin2(rtdata, 
+                                measurevar="RT_log", 
+                                withinvars=c("Vagueness", "Number", "Item"), idvar="Subject")
 ```
 
-```{r rtplot, fig.width=5, fig.height=3, echo=F}
+```{r rtplot, echo=F}
 dodge = position_dodge(width=0.2)
@@
         aspect.ratio=1,
-        axis.text.x = element_text(angle = 45, hjust = 1),
+        #axis.text.x = element_text(angle = 45, hjust = 1),
         plot.background=element_rect(fill=NA, color='white')) +

### 2016-10-02 experiments/e2/writeup/runme.R

@@ -1,20 +1,0 @@
-source('e2-preprocessing.R')
-if( file.exists('e2-b-data.txt') ) {
-  dd <- read.delim('e2-b-data.txt')
-} else { 
-  dat <- gatherData() 
-  dat <- declareImpossibleRT(dat)
-  dat <- removeImpossibleTrials(dat)
-  dat <- subset(dat, select=c(Subject, Trial, Item, 
-                              discriminability, 
-                              Instruction,
-                              nchar_instr,
-                              Vagueness, Number, Order, Quantity, 
-                              c_Vag, c_Num, c_Ord, c_Qty,
-                              NaType, response_category, 
-                              isBorderline, RT)
-  )
-  write.table(dat, file='e2-b-data.txt', sep='\t', quote=FALSE)
-  dd <- read.delim('e2-b-data.txt')
-} 
-

### 2016-10-02 experiments/e2/writeup/scratchgather.R

@@ -1,197 +1,0 @@
-number_of_valid_subjects=30
-number_of_trials_per_subject=256
-message('Starting gather data')
-data_dir <- '../experimentCode/output/'
-column.headers.df <-  head( read.table(
-  paste(data_dir,'subject01.data',sep=''),
-  header=TRUE),0)
-gathered.data <- column.headers.df
-for (subject in 1:number_of_valid_subjects) {
-  current.filename <- paste(data_dir, 'subject', sprintf("%02d",subject), '.data', sep='')
-  current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = TRUE)
-  gathered.data <- rbind(gathered.data, current.data)
-}
-dat <- gathered.data
-dat$Order <- factor(dat$Order, levels=c(1,2), labels=c('LtoR','RtoL'))
-dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)
-dat$Quantity <- factor(dat$Quantity, levels=c(1,2), labels=c('Small','Large'))
-dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)
-dat[ dat$Condition==1 , 'Vagueness'] <- 'Vague'
-dat[ dat$Condition==2 , 'Vagueness'] <- 'Crisp'
-dat[ dat$Condition==3 , 'Vagueness'] <- 'Vague'
-dat[ dat$Condition==4 , 'Vagueness'] <- 'Crisp'
-dat$Vagueness <- as.factor(dat$Vagueness)
-dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)
-dat[ dat$Condition==1 , 'Number'] <- 'Numeric'
-dat[ dat$Condition==2 , 'Number'] <- 'Numeric'
-dat[ dat$Condition==3 , 'Number'] <- 'Verbal'
-dat[ dat$Condition==4 , 'Number'] <- 'Verbal'
-dat$Number <- as.factor(dat$Number)
-dat$c_Num <- ifelse(dat$Number=='Numeric', -.5, .5)
-dat$Subject <- factor(paste("s",sprintf("%02d",dat$Subject),sep=""))
[…]

### 2016-10-02 papers/2016-Special-Issue/v-special-issue-2016.tex
master now has the tex for the special issue submission -- but not the R -- for the R you have to git checkout special-issue

@@
 
+% MG change p's to 3 dp
+
 \title{Is Vagueness Beneficial for Hearers? \protect \\ Evidence from Experiments}
 \date{\today}
-\authornote{latex compiled \today ~\currenttime}
+%\authornote{latex compiled \today ~\currenttime}
 \shorttitle{Is Vagueness Beneficial for Hearers?}
@@
 Consequently, it is difficult for the designers of a Natural Language Generation (NLG) system to know when to let the system generate vague expressions.
-The present paper is an attempt to find out what benefits vagueness might have for readers.% , by means of a series of carefully controlled psycholinguistic experiments that focus on the use of vagueness in referring expressions.
+The present paper is an attempt to find out what benefits vagueness might have for readers.
+% , by means of a series of carefully controlled psycholinguistic experiments that focus on the use of vagueness in referring expressions.
 The work is motivated by NLG but has implications for linguistics, philosophical logic and game theory.
 %
-This article reports on a series of experiments that aim to separate the utility of vagueness (as defined by the existence of borderline cases) from the utility of other factors that tend to co-occur with vagueness.
+We report on a series of experiments that aim to separate the utility of vagueness (as defined by the existence of borderline cases) from the utility of other factors that tend to co-occur with vagueness.
 We argue that the evidence so far supports a view where the benefits that vague terms exert are due to other influences, rather than to vagueness itself.
 These factors include: low granularity; the use of evaluative words; the avoidance of overtly numerical words; the existence of comparison strategies; and, lastly and more tentatively, a phenomenon that we call range reduction.
-The paper concludes with a brief discussion of the implications for practical NLG.}
+The paper concludes with a discussion of the implications for practical NLG.}
 
@@
 For example, if the temperature is 27.2 degrees Celsius, this could be expressed as ``27.2 degrees'', ``approximately 27 degrees', ``above 25 degrees", or ``warm'', and the system must decide between these.
-The problem is especially important for NLG systems that take numbers as input, for example in the generation of textual weather reports from numerical weather data such as temperature and wind speed \cite{goldberg1994using,turner2006generating}, and medical decision support on the basis of clinical measurement such as oxygen saturation, heart rhythm, etc.
+The problem is especially important for NLG systems that take numbers as input, for example in the generation of textual weather reports from numerical weather data such as temperature and wind speed \cite{goldberg1994using,turner2006generating}, and medical decision support on the basis of clinical measurements % such as oxygen saturation, heart rhythm, etc.
 \cite{Hripcsak01032009, hunter2008summarising, portet2009automatic}.
@@
 
-Language use may be called vague for various reasons.\footnote{See e.g.
-the entry ``vague'' in \cite{penguin}.} In most academic use though, the word `vagueness' has a specific meaning.
+%Language use may be called vague for various reasons.\footnote{See e.g. the entry ``vague'' in \cite{penguin}.}
+
+In most academic use, the word `vagueness' has a specific meaning.
 Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1997vagueness}, also \citeA{EgreKlinedinst}).
@@
 In this paper, we focus on the first part of Lipman's question.
-Several tentative answers to Lipman's question have been offered  \cite<see>{van2009utility, van2010vagueness}.
+Several tentative answers to Lipman's question have been offered  \cite<see>{van2009utility, vanDeemterBook}.
 Prominent among these answers is the idea that vague expressions are easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite<e.g.,>[]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
@@
 \subsection{General Methodology}\label{method}
-The experiments reported in the present paper put the cost reduction hypothesis to the test. 
-The question that we are trying to answer is whether vague expressions are processed more easily by readers than crisp ones. 
+%The experiments reported in the present paper put the cost reduction hypothesis to the test. The question that we are trying to answer is whether
+
+Are vague expressions processed more easily by readers than crisp ones? 
 Like Lipman, we focus on situations where numerical information is used in order to identify a referent. 
-Reference, in other words, will be the linguistic task on which we focus, partly because of the interest that this topic has recently drawn from the NLG community.  
-In focussing on benefits for the hearer, we will leave aside the question of audience design, leaving this for later research.
+Reference, in other words, will be the communicative task on which we focus, partly because of the interest that this topic has recently drawn from the NLG community.  
+% In focussing on benefits for the hearer, we will leave aside the question of audience design, leaving this for later research.
 
 In using references to quantities to test the cost reduction hypothesis we are only testing one aspect of vagueness in a particular context. 
-This limits the applicability of our results. 
+This limits the direct applicability of our results. 
 However, it has the advantage that it enables us to explore the costs and benefits of vagueness more thoroughly. 
@@
 There was a block of practice trials after which participants could ask any questions, following which the experimenter left the room.
-All $p$ values reported for linear models were calculated using the R package \emph{lmerTest} \cite{lmerTest}.
+All $p$ values reported for linear models were calculated using the R package \emph{lmerTest} \cite{lmerTest}. The complete data and analysis of our first experiment are available from 
+{\tt http://homepages.abdn.ac.uk/k.vdeemter/pages/vagueness-data-analysis/e1.html}, and analogous for experiments 2-4.\\[2ex]
 
@@
 
-When the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the cardinality of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene1996}.
+\noindent When the distance grows between two numbers, they become more easily distinguishable: the \emph{numerical distance effect} has been shown for comparing the cardinality of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene1996}.
 We manipulated the number of dots in each array such that some sets of arrays had smaller numerical distances and others had larger numerical distances.
@@
 We performed a pilot experiment \cite{our-earlier-paper} in which we were able to confirm this finding in the experimental settings on which we are focussing in this paper.
-As one might expect, we found that, when participants were confronted with % MG changed the next line and added the footnote to remove the doubts we had about whether a dot array must have regular layout, or can have irregular layout - I'm still not sure this is a good description though
-a stimulus consisting of two squares containing different numbers of dots\footnote{such a stimulus is referred to hereafter as a stimulus consisting of a set of some number of dot arrays, where the number of dots in each dot array varies and is referred to as its cardinality, and the physical arrangement of dots in each dot array is irregular},
+We found that, when participants were confronted with % MG changed the next line and added the footnote to remove the doubts we had about whether a dot array must have regular layout, or can have irregular layout - I'm still not sure this is a good description though
+a stimulus consisting of two squares containing different numbers of dots\footnote{Such a stimulus is referred to hereafter as consisting of a set of {\em dot arrays}.
+The number of dots in an array is referred to as its cardinality.
+The physical arrangement of dots in each array is irregular.},
 instructions of the form \emph{Choose the square with \emph{n} dots} led to consistently faster response times than instructions of the form \emph{Choose the square with many/few dots} when $2 \leq n \leq 5$; the converse was true for $n>5$.
-We henceforth focussed our studies on non-subitizable numbers, because it is there that vagueness is expected to have benefits.
+Given these findings, we henceforth focussed our studies on non-subitizable numbers, because it is there that vagueness is expected to have benefits.
 
@@
 
-\section{Experiment One}
-
-\subsection{Introduction}
+\section{Experiment 1: initial experiment}
 
@@
 \subsection{Hypothesis}
+\noindent The main aim of this experiment was to test the following hypothesis:
+{\small
 \begin{APAenumerate}
-	\item [(H1)] A main effect advantage for vagueness: vague instructions impose a lower cognitive load for the comprehender than crisp alternatives.
+	\item [(H1)] There exists a main effect advantage for vagueness: vague instructions impose a lower cognitive load for the comprehender than crisp alternatives.
 \end{APAenumerate}
+}
 
-\subsection{Method}
+\subsection{Experimental Method}
 
@@
 \begin{figure}[htbp]
-\fitfigure{images/plotRTandErrorRates-1}
+\fitfigure{images/e1RTandErrorRatesplot-1}
 \caption{Experiment 1 results: response times and error rates.}
@@
 
-\subsubsection{Response times}
-Response times (RTs) for trials with erroneous responses were discarded, leading to the loss of 354 trials from 5120, representing 6.9\% of the trials. 
+%\subsubsection{Response times}
+{\bf Response times} (RTs) for trials with erroneous responses were discarded, leading to the loss of 354 trials from 5120, representing 6.9\% of the trials. 
 The correct response RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of a further 160 trials, or 3.4\% of the remaining correct responses. 
@@
 A linear mixed model of RT was built, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
+Analysis confirms hypothesis H1:
 
+{\small
 \begin{APAenumerate}
@@
 \end{APAenumerate}
+}
 
-\subsubsection{Error rates}
-Error rate data were analysed using a generalized logit mixed model \cite{jaeger2008categorical}, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
+%\subsubsection{Error rates}
+{\bf Error rate} data were analysed using a generalized logit mixed model \cite{jaeger2008categorical}, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
+We found:
 
+{\small
 \begin{APAenumerate}
@@
 \end{APAenumerate}
+}
 
@@
 
-As expected, responses were faster and more accurate for vague than crisp instructions. Post hoc, we formulated some further hypotheses:
+\noindent As expected, responses were faster and more accurate for vague than crisp instructions. Post hoc, we formulated some further hypotheses:
 
+{\small
 \begin{APAenumerate}
@@
 
+{\small
 \begin{APAenumerate}
@@
 \end{APAenumerate}
+}
 
-Looking at error rates, the two post-hoc hypotheses were supported:
+Looking at error rates, the two post-hoc hypotheses were supported as well:
 
+{\small
 \begin{APAenumerate}
@@
 \end{APAenumerate}
+}
 
 So far, these results are in line with the idea of cost reduction.
-Cost reduction explains the vagueness advantage by claiming that the vague referring expressions place less cognitive load on the comprehender than the crisp referring expressions.
+The idea of cost reduction explains the vagueness advantage by claiming that the vague referring expressions place less cognitive load on the comprehender than the crisp referring expressions.
 It explains the diminishing returns for vagueness in more-discriminable stimuli (i.e., the vagueness by numerical distance interaction) by claiming that load is low in both conditions for the easily-discriminable stimuli, and that therefore there is no extra benefit to be had from vagueness in the easily-discriminable stimuli.
@@
 
-\section{Experiment 2}
-
-\subsection{Introduction} % E2
+\section{Experiment 2: separating vagueness from instruction format}
 
@@
 
-To find out what happens when words are used in a context where their potential for vagueness comes to the fore,
-Experiment 2 used three arrays so that the vague quantifiers always had more than one possible referent, and used indefinite articles in the vague instructions to avoid the impression that only one response counted as correct, and was carried out without error feedback.
+To find out what happens when words are used in a context where their potential for vagueness comes to the fore, Experiment 2 used three arrays (rather than two) so that the vague description had more than one possible referent, and used indefinite articles to avoid the impression that only one response counted as correct, and was carried out without error feedback.
 An indication that the potential for vagueness was realised in Experiment 2 is that the borderline response was chosen fairly often: 16\% of the time.
@@
 We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54).
-Each set of arrays had the following properties: it comprised three arrays (instead of two as in Experiment 1); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had more.
+Each set of arrays comprised three arrays (instead of two as in Experiment 1); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had more.
 
-Examples of crisp and vague versions of the numerical and verbal instructions follow: the examples assume the array (6,15,24) and reference to the smaller number of dots, such that 6 was classified as the expected response; 15 was classified as the borderline response; and 24 was classified as the incorrect response.
+Examples of crisp and vague versions of the numerical and verbal instructions follow: the examples assume the array (6,15,24) and reference to the smaller number of dots, such that 6 was classified as the expected response; 15 as the borderline response; and 24 as the incorrect response.
 In the {\em vague numerical} condition we used \emph{Choose a square with about 10 dots}.
@@
 \subsection{Hypotheses} % E2
+\noindent We formulated the following hypotheses for Experiment 2:
+
+{\small
 \begin{APAenumerate}
-	\item [(H1)] A main effect RT advantage for vagueness.
+	\item [(H1)] There exists a main effect RT advantage for vagueness.
 	\item [(H2)] An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
-	\item [(H3)] No large main effect of instruction format (of numeric versus verbal) since we hypothesised that vagueness rather than instruction format drove the effect in experiment 1.
-	\item [(H4)] On the basis of Experiment 1, we would expect faster responses for stimuli with more discriminable arrays.(i.e., an effect of item).
+	\item [(H3)] No large main effect of instruction format (i.e., numeric versus verbal) since we hypothesised that vagueness rather than instruction format drove the effect in Experiment 1.
+	\item [(H4)] On the basis of Experiment 1, we expect faster responses for stimuli with more discriminable arrays.(i.e., an effect of item).
 	\item [(H5)] Participants should make more borderline case choices for vague than crisp instructions.
 \end{APAenumerate}
+}
 
-\subsection{Method} % E2
+\subsection{Experimental Method} % E2
 % Participants were aged between 18 and 45 with a median age of 28. 
-We manipulated as independent variables vagueness and instruction format, yielding four conditions, \emph{vague numeric}; \emph{vague verbal}; \emph{crisp numeric}; \emph{crisp verbal}.
+We manipulated as independent variables vagueness and instruction format (i.e., numeric versus verbal instructions), yielding four conditions, \emph{vague numeric}; \emph{vague verbal}; \emph{crisp numeric}; \emph{crisp verbal}.
 We measured two dependent variables: response time; and the probability of a participant choosing the borderline case.
@@
 
-\subsubsection{Response times} % E2
+%\subsubsection{Response times} % E2
 % MG added the values for these stats on 26 Sept 2016
@@
 A linear mixed model was constructed for the (logged) response times, 
-with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
+with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects. We found:
+{\small
 \begin{APAenumerate}
-	\item [(H1)] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06$, $se=.01$, $t=4.6$, $p<0.0001$)
-	\item [(H2)] In focussed comparisons, vagueness was significantly disadvantageous in the \emph{numeric} conditions ($\beta=0.09$, $se=0.02$, $t=4.5$, $p<0.0001$), and non-significantly disadvantageous in the \emph{verbal} conditions ($\beta=0.02$, $se=0.02$, $t=1.5$, $p=0.1550$), offering partial evidence against hypothesis 2.
+	\item [(H1)] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against Hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06$, $se=.01$, $t=4.6$, $p<0.0001$).
+	\item [(H2)] In focussed comparisons, vagueness was significantly disadvantageous in the \emph{numeric} conditions ($\beta=0.09$, $se=0.02$, $t=4.5$, $p<0.0001$), and non-significantly disadvantageous in the \emph{verbal} conditions ($\beta=0.02$, $se=0.02$, $t=1.5$, $p=0.1550$), offering partial evidence against Hypothesis 2.
 The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to a significant overall interaction effect between vagueness and instruction format ($\beta=-0.13$, $se=0.02$, $t=-6.6$, $p=0.0169$).
-	\item [(H3)] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1, but suggesting that instruction format rather than vagueness drove the effect we observed in experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; ($\beta=-0.36$, $se=0.07$, $t=-5.1$, $p<0.0001$).
+	\item [(H3)] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1, but suggesting that instruction format rather than vagueness drove the effect we observed in Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; ($\beta=-0.36$, $se=0.07$, $t=-5.1$, $p<0.0001$).
 	\item [(H4)] There was a significant main effect of item (i.e., of which triple of numbers of dots were used in the stimulus): ($\beta=0.12$, $se=0.02$, $t=7.1$, $p<0.0001$).
@@
 \end{APAenumerate}
+}
 
-\subsubsection{Borderline responses} % E2
+%\subsubsection{Borderline responses} % E2
 % MG added the values for these stats on 26 Sept 2016
@@
 Participants chose the borderline square on $16.6\%$ of trials overall.
+{\small 
 \begin{APAenumerate}
@@
 \centering
-\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=.49\textwidth]{images/e2rt.pdf}
-\includegraphics[trim = 0mm -23mm 0mm 0mm, clip, width=.49\textwidth]{images/e2bl.pdf}
+\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{images/e2-rtplot-1.pdf}
+\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{images/e2-blBarChart-1}
 \caption{Experiment 2 results: mean response times by condition and item, and counts of borderline case responses by condition.}
@@
 Experiment 2 tested to see whether when borderline cases are present, vague instructions would speed responses as they did in Experiment 1 when there were no borderline squares. 
-We actually found a \emph{disadvantage} of vague instructions: vague instructions slowed people down by 112 ms on average.
+We actually found a small (but statistically significant) \emph{disadvantage} of vague instructions: vague instructions slowed people down by 112 ms on average.
 We also found that the effect of instruction format was significant, with numerical format slowing responses by 689 ms on average, such that the disadvantage of numerical format overwhelmed the contribution of vagueness.
@@
 Observe that, in Experiment 2, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) {\bf selection algorithm}: To see this, consider the task of selecting the dot array that contains ``few dots": to do this, it suffices to {\em compare} the three arrays and select the one that contains the fewest elements.
-Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison ``algorithm" (which is known to be faster than matching).
+To select the dot array that contains "16 dots" seems to require the participant to estimate, and then {\em match}, the cardinality of (at least) one dot array to 16, a process which could plausibly take longer, independently of vagueness.
+Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison ``algorithm".% (which is known to be faster than matching).
 % KvD I think we need to say more about this, citing some appropriate papers to support this idea.
@@
 
-\section{Experiment 3}
+\section{Experiment 3: focus on instructions that contain numerals}
 
-\subsection{Introduction}
-The main aim of experiment 3 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies.
-The main changes from experiment 2 were that the selection task was explicitly controlled, and that all conditions were constrained to mention a number.
-We used the same arrays as in experiment 2.
+The main aim of Experiment 3 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies.
+The main changes from Experiment 2 were that the human selection task was explicitly controlled (i.e., in whether it amounted to matching or comparison), and that all conditions were constrained to mention a number.
+We used the same arrays as in Experiment 2.
 Table \ref{Instructions for e3} shows the instructions for each condition.
+Note the difference between ``fewer than 20" and ``far fewer than 20": whereas the former cannot have borderline cases (i.e., for each number it is clear whether the number is smaller than 20 or not), the latter can.
 
 %KvD I think we should motivate our choice of materials (especially "far ..er")
-\textcolor{red}{{\bf KvD I think we should motivate our choice of materials (especially "far ..er")}}
+%\textcolor{red}{{\bf KvD I think we should motivate our choice of materials (especially "far ..er")}}
+% DONE 28 Sept 2017
 
@@
 \subsection{Hypotheses} % E 3
+\noindent For Experiment 3, we hypothesised:
+{\small
 \begin{APAenumerate}
-	\item [(H1)] Vague instructions are easier for the reader than crisp alternatives (main effect of vagueness)
+	\item [(H1)] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
 	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection task)
-	\item [(H3)] Effects of vagueness are different depending on whether selection is matching or comparison (interaction effect selection x vagueness).
+	\item [(H3)] Effects of vagueness are different depending on whether the selection task is matching or comparison (interaction effect selection x vagueness).
 \end{APAenumerate}
+}
 
-\subsection{Method} % E 3
+\subsection{Experimental Method} % E 3
 
-38 participants were recruited. 
-The design was a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e3}).
-On each trial a referring expression instruction was presented: participants pressed a key to dismiss the instruction, when the dot arrays were presented until the participant responded, and the response time and choice were recorded.
+38 participants were recruited. We used a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e3}).
+On each trial an instruction was presented: participants pressed a key to dismiss the instruction, when the dot arrays were presented until the participant responded, and the response time and choice were recorded.
 
@@
 \centering
-\includegraphics[trim = 0mm 30mm 0mm 10mm, clip, width=.75\textwidth]{images/rtplote3}
+\fitfigure{images/e3-rtplot-1.pdf}
+%\includegraphics[trim = 20mm 0mm 20mm 0mm, clip, width=.75\textwidth]{images/e3-rtplot-1.pdf}
 \caption{Mean response times by condition for Experiment 3 where all instructions were numeric}
@@
 % MG added the values for these stats on 26 Sept 2016
+{\small
 \begin{APAenumerate}
@@
 \end{APAenumerate}
-
+}
 % MG edited this 27 Sept 2016:
 %The cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
-The cost reduction account was wrong to predict significant main effect advantages for vagueness (although there was a non-significant trend in the \emph{direction} predicted by the cost reduction account), and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was significantly advantageous in the comparison task.
+In other words, the cost reduction account was wrong to predict significant main effect advantages for vagueness (although there was a non-significant trend in the \emph{direction} predicted by the cost reduction account), and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was significantly advantageous in the comparison task.
 
@@
 
-\section{Experiment 4} % E 4
+\section{Experiment 4: focus on instructions that do not contain numerals} % E 4
 
-\subsection{Introduction}
-This experiment investigated response times for instructions that did not use a number. 
+This experiment mirrors Experiment 3, but focussing on instructions that did not use a number. 
 We manipulated vagueness and the selection task (comparison and matching). 
@@
 \begin{table}[htp]
-\caption{Instructions and targets by condition for experiment 4}
+\caption{Instructions and targets by condition for Experiment 4}
 \begin{center}
@@
 \subsection{Hypotheses}% E 4
+\noindent For Experiment 4, we hypothesised:
+{\small
 \begin{APAenumerate}
-	\item [(H1)] Vague instructions are easier for the reader than crisp alternatives (main effect of vagueness)
+	\item [(H1)] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
 	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection)
-	\item [(H3)] Effects of vagueness are different depending on whether selection is matching or comparison (interaction effect selection x vagueness, and focussed comparisons at each level of selection).
+	\item [(H3)] Effects of vagueness differ depending on whether selection is matching or comparison (interaction effect selection x vagueness, and focussed comparisons at each level of selection).
 \end{APAenumerate}
+}
 
@@
 The results showed that vagueness was beneficial for comparison but detrimental for matching (the same as Experiment 3) even when no numbers were allowed in the instructions. 
-Figure \ref{resultse4} shows the means by condition.
+Figure \ref{resultse4} shows the means by condition. Our findings were as follows:
 
 % MG added the values for these stats on 27 Sept 2016
+{\small
 \begin{APAenumerate}
@@
 \end{APAenumerate}
+}
 
-The cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
+Once again, the cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
 
@@
 \centering
-\includegraphics[trim = 0mm 30mm 0mm 10mm, clip, width=.75\textwidth]{images/rtplote4.pdf}
+\fitfigure{images/e4-rtplot-1.pdf}
 \caption{Mean response times by condition for Experiment 4}
@@
 
-Experiment 1 showed that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment could not distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format; and mandated a comparison strategy rather than a matching strategy.
+To summarise our findings, Experiment 1 showed that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment could not distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format; and mandated a comparison strategy rather than a matching strategy.
 Experiment 2 showed that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional benefit in either the verbal format instructions or the numerical format instructions.
 
-However, Experiment 2 could not distinguish benefits of number avoidance from benefits of the comparison selection task.
-In Experiments 3 and 4 we manipulated vagueness and the selection task separately at each level of numerical format.
+However, Experiment 2 did not distinguish benefits of number avoidance from benefits of the comparison selection task.
+In Experiments 3 and 4 we manipulated vagueness and the selection task, separately at each level of numerical format.
 Across the two experiments, we found that the comparison-task instructions attracted faster response times than the matching-task instructions.
 Within the two experiments we found that vagueness exerts benefits when the selection task is \emph{comparison}, but not when the task is \emph{matching}.
-
-The benefits of vagueness in the \emph{comparison} task in experiments 3 and 4 could be explained as differences in the number of valid targets for the expression, as follows.
-Taking as an example the stimulus with (6,15,24) dots, it could be argued that the vague comparison instruction (e.g., \emph{a square with far fewer than 20 dots}) has one valid target, the square with 6 dots, while the crisp comparison instruction (e.g., \emph{a square with fewer than 20 dots}) has two valid targets, the squares with 6 and 15 dots.
-In both experiments 3 and 4 we found that people were quicker to identify a square when the instruction only had one valid target.
-This leads us to speculate that the benefit for vagueness here could be due to the vague expression foregrounding a particular valid target while the crisp expression carries with it the additional task of distinguishing between two alternative valid targets, something we propose to call a ``range-reduction'' benefit.
 
@@
 What is one entitled to conclude?
-Given that we were able to identify a class of situations -- namely: situations in which a comparison strategy suffices to identify the intended referent -- in which vague expressions led to faster response times than crisp ones, would it be valid to conclude that we have finally discovered an advantage for vagueness that cannot be ascribed to some other factor?
+Given that we were able to identify a class of situations in which vague expressions led to faster response times than crisp ones, would it be valid to conclude that we have finally discovered an advantage for vagueness that cannot be ascribed to some other factor?
 We believe the answer to this question is negative.
+
 To see why, consider Figures \ref{resultse3} and \ref{resultse4}.
@@
 
-% KvD Something missing here?
-\textcolor{red}{KvD Something missing here?}
+% KvD: I've shortened the text, removing duplication
 
-To see why vagueness thus has opposite effects, depending on whether it is used in matching or comparison situations, compare an instruction like `Choose a square with 6 dots' with its vague counterpart `Choose a square with about 10 dots': by adding the word `about', we broaden the range of squares that the expression might be referring to.
-On the other hand, compare `Choose a square with fewer than 20 dots' with its vague counterpart `Choose a square with far fewer than 20 dots': by adding the word `far', we did not broaden the range of squares denotable by the expression: we narrow it down, because only some of the squares that have fewer dots may have {\em far} fewer dots.
+To see why vagueness has opposite effects, depending on whether it is used in matching or comparison situations, consider the stimulus with (6,15,24) dots.
+Now compare `Choose a square with 6 dots' with its vague counterpart `Choose a square with about 10 dots': by adding the word `about', we broaden the range of squares that the expression might be referring to.
+On the other hand, compare `Choose a square with fewer than 20 dots' with `Choose a square with far fewer than 20 dots': by adding the word `far', we did not broaden the range of squares denotable by the expression: we narrow it down, because only some of the squares that have fewer dots may have {\em far} fewer dots.
+
+The benefits of vagueness in the \emph{comparison} task in experiments 3 and 4 could thus be explained as differences in the number of valid targets for the expression.
+This leads us to speculate that the benefit for vagueness here could be due to the vague expression foregrounding a particular valid target while the crisp expression carries with it the additional task of distinguishing between two alternative valid targets, something we propose to call a ``range-reduction'' benefit.
 
 The observation that conditions with 1 candidate lead to shorter response times than conditions with 2 candidates is consistent with the range reduction hypothesis, but not with the idea that vagueness is beneficial.
-It appears, in other words, that %range reduction causes shorter response times, suggesting that
-shorter response times will only result from a vague expression if this expression leads to range reduction.
-Once again, it is not vagueness itself that has advantages but a phenomenon (namely range reduction) that is an automatic concomitant of vagueness in some types of situations.
-
-Our findings suggest that the observed benefits of vague expressions in certain situations may be due to factors other than vagueness itself: factors like avoiding numbers; permitting comparison tasks; and range reduction.
-The picture that is starting to emerge is subtle: on the one hand, in the situations that we have been studying -- where cooperative speakers refer to an object (e.g., a square) by means of some quantity associated with the object -- vagueness is not intrinsically beneficial.
+It appears, in other words, that shorter response times will only result from a vague expression if this expression leads to range reduction.
+Once again, it is not vagueness itself that has advantages but a phenomenon (namely range reduction) that is an automatic concomitant of vagueness in some types of situations.\\[2ex]
+%
+\noindent Looking at the entire series of experiments, our findings suggest that the observed benefits of vague expressions in certain situations may be due to factors other than vagueness: factors like avoiding numbers; permitting comparison tasks; and range reduction.
+The picture that is starting to emerge is subtle: on the one hand, in the situations that we have been studying
+%-- where cooperative speakers refer to an object (e.g., a square) by means of some quantity associated with the object -- 
+vagueness is not intrinsically beneficial.
 On the other hand, vague expressions frequently possess other features that {\em are} beneficial, and these are what give us the incorrect impression that vagueness itself is beneficial.
 Vagueness may thus have acquired a reputation that it does not deserve.
-To answer to Lipman's question, of why vagueness permeates human language (see our Introduction), may lie in a different direction after all, possibly relating to a combination of pure necessity and benefits for the speaker rather than the hearer \cite<see>[chapter 11, for discussion of some possible directions]{van2010vagueness}.
+The answer to Lipman's question, of why vagueness permeates human language (see our Introduction), may lie in a different direction after all, possibly relating to benefits for the speaker rather than the hearer.
+% \cite<see>[chapter 11, for discussion of some possible directions]{vanDeemterBook}.
 
@@
 %
-{\bf Implications for practical NLG systems.} Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
-The question of how a particular piece of quantitative information is best conveyed through language is an important one for practical NLG.
-Years of research on the logic of vagueness in natural language -- which has given rise to such logical techniques as Partial Logic (e.g, \cite{Fine}, Probabilistic Logic \cite{Edgington}, and Fuzzy Logic \cite{Zadeh} -- have primed the research community to expect that the utility of vagueness is an important part of the answer, but our findings call this expectation into question.
+{\bf Implications for NLG systems.} Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
+The question of how a particular piece of quantitative information is best conveyed through language is an important one for practical NLG; years of research on the logic of vagueness -- giving rise to such techniques as Partial Logic \cite<e.g.,>{Fine}, Probabilistic Logic \cite{Edgington}, and Fuzzy Logic \cite{Zadeh} -- have primed the research community to expect that the utility of vagueness is an important part of the answer, but our findings call this expectation into question.
+
 Although our own studies in this article have focussed on vagueness in descriptive Noun Phrases, it seems plausible that vagueness plays a similar role in other linguistic constructs.
-
-For example, consider air temperature, once again. Given a temperature measurement (or prediction) as input, an NLG system that addresses a non-expert audience might say that it was (or will be)
+For example, consider reports on air temperature (cf., our discussion of NLG at the start of the paper).
+Given a temperature measurement or prediction as input, an NLG system might say that it will be
 %
@@
 If the linguistic literature is to be believed, then options (a) and (c) convey crisp information, whereas (b) and (d) are vague (i.e., they permit borderline cases).
-From our own and previous experiments, there is no evidence that the fact in itself that an expression is vague should confer a benefit on it for hearers.
+From our own experiments and the ones discussed in section 1, there is no evidence that  vagueness should be beneficial for hearers.
 Rather than asking whether a candidate expression is vague, other questions might shed more light on the NLG system's choice, similar to the ones identified in our studies.
@@
 
-First, the experiments by Mishra et al.
-suggest that it is important how much information is conveyed by an expression, and their findings are echoed by our own thoughts about range reduction.
-In the case of (a)--(d) above, it appears that (a) conveys the most detailed information (designating the smallest segment of the temperature scale), followed by (b).
-Expression (b)  appears to be followed by (d), and (d) by (c) (e.g., 40 degrees is above 25, but at 40 Celsius the word ``warm" may no longer be applicable, giving way to words like ``hot"):
+First, the experiments by \citeauthor{Mishra01042011} suggest that it is important how much information is conveyed by an expression, and their findings are echoed by our own thoughts about range reduction (following Experiments 3 and 4).
+In the case of (a)--(d) above, it appears that (a) conveys the most detailed information (designating the smallest segment of the temperature scale), followed by (b), then (d), then (c) (e.g., 40 degrees is above 25, but at 40 Celsius the word ``warm" is likely to give way to ``hot" or ``scorching"):
 %
@@
 %
-If these hunches are correct then an {\em expert} may  to prefer expression (a), because it gives her the most detailed information as a basis for her decisions.
-On the other hand, expression (d) (``warm") is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make ``warm" more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm" (see section 1 and  \cite{peters2009bringing}); the factor of evaluation is especially important if the hearer is unfamiliar with the metric used (e.g., because, being American, they are more used to Fahrenheit), in which case ``warm" is much clearer.
-These considerations suggest that {\em non-experts} might prefer expression (d) over all others.
-If results of this kind were to be confirmed by experiments, they would be applicable in practical NLG, though care would need to be exercised to take the purpose of the communication into account.\footnote{One reason why ``warm" might be less preferred than these considerations suggest is that, in the weather context, some very frequently occurring numbers, such as 20 (degrees), could be relatively easy to process, analogous to the subitizable numbers of \protect\cite{trick1994small}, see our General Methodology section.} In the experiments reported in this paper, the purpose of the utterances investigated was always clear (e.g., instructing people on a weight loss program \cite{Mishra01042011}; informing a patient's choice of hospital \cite{peters2009bringing}; or selecting a dot array (in our own studies).
-% KvD See new footnote above.
+If these hunches are correct then an {\em expert} may  to prefer expression (a), because it gives her the most detailed information on which to base her decisions.
+On the other hand, expression (d) (``warm") is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make (d) more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm" (see our section 1 and  \citeA{peters2009bringing}), which is especially important if the hearer is unfamiliar with the metric used (e.g., because she is more used to Fahrenheit).
+These considerations suggest that {\em non-experts} might prefer expression (d).
+If results of this kind were to be confirmed by experiments, they would be applicable in practical NLG, though care would need to be exercised to take the purpose of the communication into account.\footnote{One reason why ``warm" might be less preferred than these considerations suggest is that, in the weather context, some very frequently occurring numbers, such as 20 (degrees), could be relatively easy to process, analogous to the subitizable numbers of \protect\citeA{trick1994small}, see our section \ref{method}.
+Note that in the experiments reported in this paper, the purpose of the utterances investigated was always clear (e.g., instructing people on a weight loss program \cite{Mishra01042011}; informing a patient's choice of hospital \cite{peters2009bringing}; or selecting a dot array (in our own studies).}
 
-One way to see why vagueness (as defined in our Introduction) may not matter as much for NLG -- and for human communication -- as is often thought, is the following thought experiment.
+One way to see why vagueness (as defined in our Introduction) may not matter much for NLG -- and for human communication -- is the following thought experiment.
 Suppose a group of speakers understand the word ``warm" as vague, agreeing that temperatures above 26 count as warm, and temperatures below 24 do not count as warm, but considering temperatures between 24 and 26 as borderline cases.
@@
 
+% KvD Bracketed out the following for now:
+%A more radical perspective on our findings might say that the distinction between crisp and vague expressions, which is taken for granted in the bulk of the linguistic literature, is mistaken, because expressions such as ``above 25 degrees" (and ``fewer than 20 dots") are not equally felicitous in all situations in which they can be truthfully applied.
+For example, when the temperature is 40 degrees, then a weather report saying that it is ``above 25 degrees" would be correct yet misleading: this description might be felicitous as a description of temperatures between (roughly) 26-28 degrees, borderline felicitous for 29-33 degrees, and infelicitous at even higher temperatures.
+At its most bold, this perspective would assert that the applicability of {\em every} linguistic expression has borderline cases, hence Lipman's question (see section 1) does not arise.
+This radical perspective -- which would have drastic implications for the construction of an NLG system -- can only be indicated here, but not discussed at length.
+
 \setcounter{secnumdepth}{0}

### 2016-10-02 experiments/e2/writeup/e2-analysis.Rmd
renaming e2 to e2-analysis

--- 2016-10-02 experiments/e2/writeup/e2.Rmd
@@
 source('e2-preprocessing.R')
-if( !file.exists('e2-b-data.txt') ) {
+if( !file.exists('e2-data.txt') ) {
   dat <- gatherData() 
@@
   )
-  write.table(dat, file='e2-b-data.txt', sep='\t', quote=FALSE, row.names = FALSE)
+  write.table(dat, file='e2-data.txt', sep='\t', quote=FALSE, row.names = FALSE)
 }

### 2016-10-02 experiments/e2/writeup/e2-analysis.Rmd
e2: simpler file structure; now includes coefficient plots for main analyses; raw data is in writeup

--- 2016-09-27 experiments/e2/writeup/e2.Rmd
@@
 ---
+
 <style>
@@
 
-```{r setup, include=FALSE, cache=FALSE}
-library(knitr)
-knitr::opts_chunk$set(echo = TRUE)
-library(pander)
+In this experiment we did not count any response as an error, instead we coded them as expected, borderline, and extreme.
+
+```{r loadLibraries, message=FALSE, echo=FALSE}
+library(knitr); library(Matrix); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions); source('summarySEwithin2.R'); source('e2-preprocessing.R')
 ```
 
-```{r showInstructions, echo=FALSE, cache=TRUE}
+```{r showInstructions, echo=FALSE}
 v = c("crisp", "crisp", "vague", "vague")
@@
 
-```{r showItems, echo=FALSE, cache=TRUE}
+```{r showItems, echo=FALSE}
 items = c(1,2,3,4,5,6,7,8)
@@
 
-```{r loadLibraries, message=FALSE, echo=FALSE, cache=FALSE}
-library(knitr); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions); source('summarySEwithin2.R')
+```{r preprocessing, echo=FALSE, message=FALSE}
+source('e2-preprocessing.R')
+if( !file.exists('e2-data.txt') ) {
+  dat <- gatherData() 
+  dat <- declareImpossibleRT(dat)
+  #dat <- removeImpossibleTrials(dat)
+  dat <- subset(dat, select=c(Subject, Trial, Item, c_Itm, 
+                              discriminability, 
+                              Instruction,
+                              nchar_instr,
+                              Vagueness, Number, Order, Quantity, 
+                              c_Vag, c_Num, c_Ord, c_Qty,
+                              NaType, response_category, 
+                              isBorderline, RT, RT_log)
+  )
+  write.table(dat, file='e2-data.txt', sep='\t', quote=FALSE, row.names = FALSE)
+}
 ```
 
-In this experiment we did not count any response as an error, instead we coded them as expected, borderline, and extreme.
-
-```{r getData, echo=T}
-rawdata <- read.table('data.txt', sep=' ')
-```
-
-```{r}
-tempdata <- perSubjectTrim.fnc(rawdata, response='RT', subject='Subject', trim = 2.5)$data
+```{r getData, echo=TRUE, tidy=FALSE}
+rawdata <- read.delim('e2-data.txt')
+tempdata <- perSubjectTrim.fnc(data = rawdata[complete.cases(rawdata),], 
+                               response='RT', subject='Subject', trim = 2.5)$data
 rtdata <- tempdata
@@
 
-This code creates the condition means with confidence intervals used for plotting:
-
-```{r computeSummariesForGraphs, cache=TRUE, message=FALSE}
-rtdataplot1 <- summarySEwithin2(rtdata, measurevar="RT_log", withinvars=c("Condition", "Vagueness", "Number", "Item"), idvar="Subject")
-```
-
-```{r rtplot, fig.width=5, fig.height=3, echo=F}
-dodge = position_dodge(width=0.2)
+```{r e2-rtplot, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE, dev='pdf'}
+rtdataplot1 <- summarySEwithin2(rtdata, 
+                                measurevar="RT_log", 
+                                withinvars=c("Vagueness", "Number", "Item"), idvar="Subject")
 rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness))
-rtplot=ggplot(rtdataplot1, aes(y=RT_log, x=Item, ymin=RT_log-ci, ymax=RT_log+ci,
-                group=Condition, fill=Vagueness)) +
-  geom_line(position=dodge) +
-  geom_errorbar(width=.2, position=dodge) +
-  geom_point(pch=21, size=3, position=dodge) +
-  #scale_shape_manual("",values = c(22, 22)) +
-  scale_fill_manual("",values=c("black","white")) +
+rtplot=ggplot(rtdataplot1, aes(y=RT_logNormed, x=Item, group=Condition, fill=Vagueness)) +
+  geom_line() +
+  geom_errorbar(width=.2, aes(ymin=RT_logNormed-ci, ymax=RT_logNormed+ci)) +
+  geom_point(size=2, col=1, aes(shape=Vagueness)) +
+  scale_fill_grey(name="Vagueness", start=0, end=1) +
+  scale_shape_manual(name="Vagueness", values=c(21,22)) +
   ggtitle("Response time") +
   ylab("Respone time (log (ms))") + 
-  xlab("") +
+  xlab(NULL) +
   theme_bw() +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        legend.key = element_blank(), legend.position="top",
+        legend.key = element_blank(), legend.position="right",
         aspect.ratio=1,
-        axis.text.x = element_text(angle = 45, hjust = 1),
-        plot.background=element_rect(fill=NA, color='white')) +
+        plot.background=element_rect(fill=NA, color='white'),
+        strip.background=element_blank(),
+        axis.text.x = element_text(angle = 15)) +
   facet_grid(~Number)
@@
 
-```{r blplot, fig.width=5, fig.height=2.5, echo=F}
-bldata$Condition <- as.factor(paste(sep=':', bldata$Vagueness, bldata$Number))
-bldata$response_category <- relevel(bldata$response_category, ref = 'expected')
-blplot=ggplot(bldata) + 
-  geom_bar(aes(Vagueness, fill=response_category), position = 'dodge') +
-  #geom_errorbar() +
-  facet_wrap(nrow=1,~Number) +
-  scale_fill_grey(name="") +
-  xlab(" ") +
-  ggtitle("Response Type") +
+```{r e2-blBarChart, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE, dev='pdf'}
+bldata$response_category <- relevel(bldata$response_category, ref='expected')
+blBarChart = ggplot(bldata) +
+  geom_bar(aes(response_category, group=Number:Vagueness, fill=Vagueness),
+           position=position_dodge(width=NULL)) +
+  scale_fill_grey() +
+  xlab(NULL) +
+  ggtitle('Borderline response distribution') +
+  facet_grid(~Number) +
   theme_bw() +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        legend.key = element_blank(), legend.position="top", aspect.ratio=1,
-        plot.background=element_rect(fill=NA, color='white'))
-blplot
+        legend.key = element_blank(), legend.position="right", 
+        aspect.ratio=1,
+        plot.background=element_rect(fill=NA, color='white'),
+        strip.background=element_blank(),
+        legend.key.size=unit(4, 'mm'),
+        axis.text.x = element_text(angle = 15))
+blBarChart
 ```
 
-```{r rtlmer, cache=TRUE}
-rtlmer = lmerTest::lmer(data=rtdata,
+Response time models
+
+```{r e2-rtlmerTest1, cache=TRUE, echo=TRUE}
+rtlmerTest1 = lmerTest::lmer(
+  data=rtdata,
   RT_log ~ c_Vag * c_Num + c_Itm +
@@
 
-```{r rtlmersummaryxtable, results='asis', cache=TRUE}
-print(xtable(summary(rtlmer)$coef, digits=c(0,2,2,1,1,4)),type='html')
+```{r e2-rtlmerTest1Visuals, cache=TRUE, echo=FALSE, results='asis', dependson='e2-rtlmerTest1', fig.width=6, fig.height=3, dev='pdf'}
+print(
+  xtable(
+    summary(rtlmerTest1)$coefficients[,-3], 
+    display=c("s","f","f","f","f"),
+    digits=c(0,2,2,1,3),
+    caption="Response time model"),
+  caption.placement="top", 
+  type='html')
+dfr <- data.frame(coef = summary(rtlmerTest1)$coef[,1],
+                  ci2.5 = confint(rtlmerTest1, method='Wald')[,1][!is.na(confint(rtlmerTest1, method='Wald')[,1])],
+                  ci97.5 = confint(rtlmerTest1, method='Wald')[,2][!is.na(confint(rtlmerTest1, method='Wald')[,2])]
+)
+intercept = round(dfr[1,1],2)
+dfr <- dfr[-1,] # lose the intercept for plotting
+dfr <- dfr[rev(rownames(dfr)),]
+par(mar=c(5,7,2,1))
+plot(y=1:nrow(dfr), 
+     x=dfr$coef, 
+     xlim=extendrange(range(dfr$ci2.5, dfr$ci97.5)), 
+     ylim=extendrange(c(1,nrow(dfr))),
+     type='n', axes=F, 
+     xlab=substitute(paste('deviation from an intercept of ', intercept, ' log RT, with 95% CI.'), list(intercept=intercept)),
+     ylab="",
+     main = 'Response time model')
+abline(v=0, lty=1, col='lightgrey')
+arrows(y0=1:nrow(dfr), y1=1:nrow(dfr), 
+       x0=dfr$ci2.5, x1=dfr$ci97.5, 
+       length=0.035,
+       col='black',
+       lwd=2, code=3, angle=90)
+points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white')
+axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE)
+axis(1)
+box(which='plot', col='grey')
 ```
 
-```{r seplmersfornum, results='asis', cache=TRUE}
-s1 <- subset(rtdata, subset=Number=='Numeric')
-s1lm <- lmerTest::lmer(data=s1, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
-print(xtable(summary(s1lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
+```{r e2-seplmerforNum, results='asis', cache=TRUE}
+numOnly <- subset(rtdata, subset=Number=='Numeric')
+numOnlylm <- lmerTest::lmer(data=numOnly, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
+print(
+  xtable(
+    summary(numOnlylm)$coefficients[,-3], 
+    display=c("s","f","f","f","f"),
+    digits=c(0,2,2,1,3),
+    caption="Response time model: Numeric only"),
+  caption.placement="top", 
+  type='html')
 ```
 
-```{r seplmersforvag, results='asis', cache=TRUE}
-s2 <- subset(rtdata, subset=Number=='Verbal')
-s2lm <- lmerTest::lmer(data=s2, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
-print(xtable(summary(s2lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
+```{r e2-seplmerforVerbal, results='asis', cache=TRUE}
+verbalOnly <- subset(rtdata, subset=Number=='Verbal')
+verbalOnlylm <- lmerTest::lmer(data=verbalOnly, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
+print(
+  xtable(
+    summary(verbalOnlylm)$coefficients[,-3], 
+    display=c("s","f","f","f","f"),
+    digits=c(0,2,2,1,3),
+    caption="Response time model: Verbal only"),
+  caption.placement="top", 
+  type='html')
 ```
 
-```{r bllmer, cache=TRUE}
+Borderline response models
+
+```{r e2-bllmer, cache=TRUE, echo=TRUE}
 bllmer = lme4::glmer(data=bldata,
@@
 
-```{r bllmersummaryxtable, results='asis', cache=TRUE}
-print(xtable(summary(bllmer)$coef, digits=c(0,2,2,1,4)),type='html')
+```{r e2-bllmerVisuals, cache=TRUE, echo=FALSE, results='asis', dependson='e2-bllmer', fig.width=6, fig.height=3, dev='pdf'}
+print(
+  xtable(
+    summary(bllmer)$coefficients,
+    digits=c(0, 2, 2, 1, 3),
+    caption="Error rates model"),
+  caption.placement="top", 
+  type='html')
+dfr <- data.frame(coef = summary(bllmer)$coef[,1],
+                  ci2.5 = confint(bllmer, method='Wald')[,1][!is.na(confint(bllmer, method='Wald')[,1])],
+                  ci97.5 = confint(bllmer, method='Wald')[,2][!is.na(confint(bllmer, method='Wald')[,2])]
+)
+intercept = round(dfr[1,1],2)
+dfr <- dfr[-1,] # lose the intercept for plotting
+dfr <- dfr[rev(rownames(dfr)),]
+par(mar=c(5,7,2,1))
+plot(y=1:nrow(dfr), 
+     x=dfr$coef, 
+     xlim=extendrange(range(dfr$ci2.5, dfr$ci97.5)), 
+     ylim=extendrange(c(1,nrow(dfr))),
+     type='n', axes=F, 
+     xlab=substitute(paste('deviation from an intercept of ', intercept, ' with 95% CI.'), list(intercept=intercept)),
+     ylab="",
+     main='Borderline response model')
+abline(v=0, lty=1, col='grey')
+arrows(y0=1:nrow(dfr), y1=1:nrow(dfr), 
+       x0=dfr$ci2.5, x1=dfr$ci97.5, 
+       length=0.035,
+       col='black',
+       lwd=2, code=3, angle=90)
+points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white')
+axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE)
+axis(1)
+box(which='plot', col='grey')
 ```

### 2016-10-02 experiments/e2/writeup/e2-scratch.Rmd
e2: simpler file structure; now includes coefficient plots for main analyses; raw data is in writeup

--- 2016-08-12 experiments/e2/writeup/scratch.Rnw
@@
-<!--html_document:-->
-  <!--toc: true-->
-  <!--toc_float: true-->
-  <!--mathjax: default-->
-  <!--template: default-->
-  <!--code_folding: hide-->
-  <!--theme: default-->
+```{r setup, include=FALSE, cache=FALSE}
+library(knitr)
+```
+
+```{r loadFonts, message=FALSE, echo=FALSE}
+library(fontcm)
+library(extrafont)
+#font_import()
+#loadfonts()
+```
+
+```{r loadLibs, message=FALSE, echo=FALSE}
+library(beepr)                     #
+library(car)                       #
+library(ggplot2)                   #
+library(grid)                      #
+library(gridExtra)                 #
+library(languageR)                 #
+library(lattice)                   #
+library(latticeExtra)              # for doubleYScale
+library(lme4)                      #
+library(lmerTest)                  #
+library(LMERConvenienceFunctions)  # for perSubjectTrim
+library(MASS)                      #
+library(plyr)                      #
+library(reshape2)                  #
+library(rms)                       # for varclus
+library(Rmisc)                     # for summarySEwithin and friends
+library(xtable)                    #
+```
+
+```{r preprocessing}
+source('e2-preprocessing.R')
+if( !file.exists('e2-b-data.txt') ) {
+  dat <- gatherData() 
+  dat <- declareImpossibleRT(dat)
+  dat <- removeImpossibleTrials(dat)
+  dat <- subset(dat, select=c(Subject, Trial, Item, 
+                              discriminability, 
+                              Instruction,
+                              nchar_instr,
+                              Vagueness, Number, Order, Quantity, 
+                              c_Vag, c_Num, c_Ord, c_Qty,
+                              NaType, response_category, 
+                              isBorderline, RT)
+  )
+  write.table(dat, file='e2-b-data.txt', sep='\t', quote=FALSE)
+}
+dd <- read.delim('e2-b-data.txt')
+```
+
+```{r showCompression, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='Ratios for different numbers of dots in the arrays: smaller values are more discriminable.
+Blue is for the ratio between the smallest number in the array and the largest number in an array.
+Red is for the mean of two ratios, one for the smallest number to the middle number, the other for the middle number to the largest number in the array', echo=FALSE}
+#  Show compression of ratios of numbers of dots in the items
+discriminability = sort(unique(as.numeric(dd$discriminability)))
+plot(y=c(1.5,1.5,1.5,1.5), 
+     x=discriminability, 
+     ylab='',
+     xlab='',
+     yaxt='n',
+     type='n', 
+     xaxt='n',
+     axes=F)
+lines(y=c(1.5,1.5,1.5,1.5), x=discriminability, type='o', pch=19)
+text(y=c(1.15,1.15,1.15,1.15), x=discriminability, labels=round(discriminability,2),  cex=.7)
+mtext('mean (ratio min/med,\nratio med/max)', side = 2, line = 1)
+mtext('Discriminability index', side=1, line=3)
+mtext('Discriminability for each item', side=3, line=1, cex=1.5)
+axis(1)
+box(which='plot', col='grey')
+```
+
+```{r e2-rtplot, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE, dev='pdf'}
+dd$RT_log=log(dd$RT)
+rtdata=dd
+source('summarySEwithin2.R')
+rtdataplot1 <- summarySEwithin2(rtdata, measurevar="RT_log", 
+                                withinvars=c("Vagueness", "Number", "Item"), idvar="Subject")
+rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness))
+rtplot=ggplot(rtdataplot1, aes(y=RT_logNormed, x=Item, group=Condition, fill=Vagueness)) +
+  geom_line() +
+  geom_errorbar(width=.2, aes(ymin=RT_log-ci, ymax=RT_log+ci)) +
+  geom_point(size=2, col=1, aes(shape=Vagueness)) +
+  scale_fill_grey(name="Vagueness", start=0, end=1) +
+  scale_shape_manual(name="Vagueness", values=c(21,22)) +
+  ggtitle("Response time") +
+  ylab("Respone time (log (ms))") + 
+  xlab(NULL) +
+  theme_bw() +
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+        legend.key = element_blank(), legend.position="top",
+        aspect.ratio=1,
+        plot.background=element_rect(fill=NA, color='white'),
+        strip.background=element_blank(),
+        axis.text.x = element_text(angle = 15)) +
+  facet_grid(~Number)
+rtplot
+```
+#Lmer model: before outlier removal
+
+```{r}
+dd$RT_log <- log(dd$RT)
+dd$c_Trl <- as.numeric(scale(dd$Trial))
+```
+
+```{r lmer5, tidy=FALSE, cache=TRUE}
+vlmerTest <- lmerTest::lmer(data=dd,
+                 RT_log ~
+                   c_Vag + c_Num + c_Qty + c_Ord +
+                   c_Num:c_Vag:c_Qty +
+                   discriminability +
+                   c_Trl +
+                   #RTprev_log +
+                   nchar_instr +
+                   (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
+```
+
+```{r printXtableV5, echo=FALSE, results='asis'}
+print(
+  xtable(
+    summary(vlmerTest)$coef,
+    caption = 'lmerTest version',
+    digits=c(0, 3,3,1,1,4)),
+  type='html',
+  caption.placement='top'
+)
+```
+
+```{r llmer5, tidy=FALSE, cache=TRUE}
+vlme4 <- lme4::lmer(data=dd,
+                 RT_log ~
+                   c_Vag + c_Num + c_Qty + c_Ord +
+                   c_Num:c_Vag:c_Qty +
+                   discriminability +
+                   c_Trl +
+                   #RTprev_log +
+                   nchar_instr +
+                   (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
+```
+
+```{r printXtableV521c, echo=FALSE, results='asis'}
+print(
+  xtable(
+    summary(vlme4)$coef,
+    caption = 'lme4 version',
+    digits=c(0, 3,3,1)),
+  type='html',
+  caption.placement='top'
+)
+```
 
@@
 
-<<exampleOfFonts, fig.width=3, fig.height=3, fig.cap='font example', echo=FALSE, eval=FALSE>>=
+#FONTS
+
+```{r defineFonts}
+.define.fonts <- function () {
+quartzFonts(avenir =    c('Avenir Book', 'Avenir Black', 'Avenir Book Oblique', 'Avenir Black Oblique'),
+helvetica = c('Helvetica New Light', 'Helvetica New Bold', 'Helvetica New Light Italic', 'Helvetica New Bold Italic'),
+cmodern = c('CMU Serif', 'CMU Serif Extra','CMU Classical Serif','CMU Serif Extra'))
+}
+# .define.fonts()
+```
+
+```{r exampleOfGGplotFonts, fig.width=3, fig.height=3, fig.cap='ggplot special font example', echo=FALSE}
+# ggplot special fonts
 p <- ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point() +
@@
 ggsave("ggplot_CMU_Serif.pdf", p, width=3.5, height=3.5)
-@
+```
 
-<<defineFonts>>=
-# .define.fonts <- function () {
-#     quartzFonts(avenir =    c('Avenir Book', 'Avenir Black', 'Avenir Book Oblique', 'Avenir Black Oblique'),
-#                 helvetica = c('Helvetica New Light', 'Helvetica New Bold', 'Helvetica New Light Italic', 'Helvetica New Bold Italic'),
-#                 cmodern = c('CMU Serif', 'CMU Serif Extra','CMU Classical Serif','CMU Serif Extra'))
-# }
-# .define.fonts()
-@
-
-<<fig.cap='base graphic special font'>>=
-#base graphic special font
+```{r exampleOfBaseGraphicsFonts, fig.width=3, fig.height=3, fig.cap='base graphic special font example'}
+# base graphic special font
 pdf("plot_cm.pdf", family="CM Roman", width=5.5, height=5)
-
 curve(dnorm, from=-3, to=3, main="Normal Distribution")
@@
 e ^ {-frac(x^2, 2)} )))
-
 dev.off()
 embed_fonts("plot_cm.pdf", outfile="plot_cm_embed.pdf")
-@
+```
+
+```{r howManyCellsInDesign}
+# number of higher-level cells in the design (abstracting over Order) =
+# Item(4) * Number(2) * Vagueness(2) * Quantity(2) = 32
+# number of measurements in eaach cell = 8
+# 332 * 8 = 256
+# 256 * 30 = 7680
+dd$measurement=0
+dd$cell=0
+for (s in levels(dd$Subject)) {
+  cellcount=0
+  for (i in levels(dd$Item)) {
+    for (n in levels(dd$Number)) {
+      for (v in levels(dd$Vagueness)) {
+        for (q in levels(dd$Quantity)) {
+          measurementcount=0
+          cellcount=cellcount+1
+          for (t in dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q, "Trial"]) {
+            dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "cell"] <- cellcount
+            measurementcount=measurementcount+1
+            dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "measurement"] = measurementcount
+          }
+        }
+      }
+    }
+  }
+}
+```
+
+```{r rSquaredInfo, echo=FALSE}
+cat("R^2")
+cor(fitted(v5), dd$RT_log)^2
+```
+
+```{r plotModelCoefsAndCis, fig.width=5.5, fig.height=4, fig.cap='Coefficient estimates and their (Wald) 95 per cent confidence intervals', fig.pos='hbtp', echo=FALSE}
+# Plot model coefficients and ci's
+dfr <- data.frame(coef = summary(v5)$coef[-1, 1], 
+                  ci2.5  = confint(v5, method='Wald')[18:26, 1],
+                  ci97.5 = confint(v5, method='Wald')[18:26, 2] )
+dfr <- dfr[rev(rownames(dfr)),]
+par(mar=c(5,10,1,1))
+plot(y=1:nrow(dfr), 
+     x=dfr$coef, 
+     xlim=range(dfr$ci2.5, dfr$ci97.5), 
+     type='n', axes=F, xlab="", ylab="")
+abline(v=0, lty=3, col='grey')
+abline(h=c(1,2,3,4,5,6,7,8,9), col='grey', lty=3)
+arrows(y0=1:nrow(dfr), y1=1:nrow(dfr), 
+       x0=dfr$ci2.5, x1=dfr$ci97.5, 
+       length=0.035,
+       col='black',
+       lwd=2, code=3, angle=90)
+points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white')
+axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE)
+axis(1)
+mtext("Coefficient estimates, in units of log(RT),\nwith 95% confidence intervals", side=1, line=3.5, cex.lab=1, las=1)
+mtext("Model terms", side=2, line=8)
+box(which='plot', col='grey')
+
+```
+
+```{r plotLMERfnc, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='plotMLERfnc', echo=TRUE}
+par(mfrow=c(2,4))
+plotLMER.fnc(v5)
+
+```
+
+```{r baayenPLots99, fig.width=6, fig.height=4, fig.cap='Baayen Model Criticism Plots', fig.pos='hbtp', echo=FALSE}
+# Baayen 4-plot model criticism
+par(mfrow=c(1,3), pty='s')
+# create scaled residuals
+dd$rstand = as.vector(scale(resid(v5)))
+# plot scaled residuals density
+plot(density(dd$rstand))
+# plot sample quantiles versus theoretical quantiles
+qqnorm(dd$rstand, cex=.5)
+qqline(dd$rstand)
+# plot standardised residuals versus fitted values
+plot(dd$rstand ~ fitted(v5), pch='.')
+# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
+abline(h=c(-2.5,2.5))
+
+```
+
+#lmerTest Version
+
+```{r lmerTestVersionOfModel, tidy=FALSE, cache=TRUE, echo=TRUE}
+v6 <- lmerTest::lmer(data=dd,
+                     RT_log ~
+                       c_Vag + c_Num + c_Qty + c_Ord +
+                       c_Num:c_Vag:c_Qty +
+                       discriminability +
+                       s_Trl +
+                       RTprev_log +
+                       nchar_instr +
+                       (1+c_Vag + c_Num + c_Qty + c_Ord|Subject))
+```
+
+```{r summaryLmerTest, cache=TRUE}
+summary(v6)
+```
+
+#Lmer model: after outlier removal
+not done yet.
+
+#Borderline responses
+
+```{r barplotBorderline, fig.width=6, fig.height=4, echo=FALSE}
+dd$response_category <- relevel(dd$response_category, ref = 'expected')
+barplot(table(dd$Vagueness,dd$response_category), beside=TRUE, legend=TRUE, space=c(0,.5))
+
+```
+
+```{r tabBdl, echo=FALSE, results='asis'}
+print(xtable(table(dd$response_category,dd$Vagueness),
+             caption='Borderline cases counts'),
+      table.placement='htbp',
+      latex.environments='center',
+      size='small')
+```

### 2016-10-02 experiments/e2/writeup/scratch44.R

@@ -1,101 +1,0 @@
-
-```{r rSquaredInfo, echo=FALSE}
-cat("R^2")
-cor(fitted(v5), dd$RT_log)^2
-```
-
-```{r plotModelCoefsAndCis, fig.width=5.5, fig.height=4, fig.cap='Coefficient estimates and their (Wald) 95 per cent confidence intervals', fig.pos='hbtp', echo=FALSE}
-# Plot model coefficients and ci's
-dfr <- data.frame(coef = summary(v5)$coef[-1, 1], 
-                  ci2.5  = confint(v5, method='Wald')[18:26, 1],
-                  ci97.5 = confint(v5, method='Wald')[18:26, 2] )
-dfr <- dfr[rev(rownames(dfr)),]
-par(mar=c(5,10,1,1))
-plot(y=1:nrow(dfr), 
-     x=dfr$coef, 
-     xlim=range(dfr$ci2.5, dfr$ci97.5), 
-     type='n', axes=F, xlab="", ylab="")
-abline(v=0, lty=3, col='grey')
-abline(h=c(1,2,3,4,5,6,7,8,9), col='grey', lty=3)
-arrows(y0=1:nrow(dfr), y1=1:nrow(dfr), 
-       x0=dfr$ci2.5, x1=dfr$ci97.5, 
-       length=0.035,
-       col='black',
-       lwd=2, code=3, angle=90)
-points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white')
-axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE)
-axis(1)
-mtext("Coefficient estimates, in units of log(RT),\nwith 95% confidence intervals", side=1, line=3.5, cex.lab=1, las=1)
-mtext("Model terms", side=2, line=8)
-box(which='plot', col='grey')
-
-```
-
-
-
-
-```{r plotLMERfnc, fig.width=6, fig.height=4, fig.pos='hbtp', fig.cap='plotMLERfnc', echo=TRUE}
-par(mfrow=c(2,4))
-plotLMER.fnc(v5)
-
-```
-
-
-
-```{r baayenPLots99, fig.width=6, fig.height=4, fig.cap='Baayen Model Criticism Plots', fig.pos='hbtp', echo=FALSE}
-# Baayen 4-plot model criticism
-par(mfrow=c(1,3), pty='s')
-# create scaled residuals
-dd$rstand = as.vector(scale(resid(v5)))
[…]

### 2017-05-15 experiments/e3/writeup/e3.Rmd
revisited May 2017; the only changes I have made are to get the original code to run

@@
     template: default
-    <!--code_folding: show-->
+    code_folding: show
     theme: default
@@
 ```
+
 ```{r computeSummariesForGraphs, cache=TRUE, message=FALSE}
-rtdataplot1 <- summarySEwithin2(rtdata, measurevar="RT_log", withinvars=c("Condition", "Vagueness", "Selection", "Item"), idvar="Subject")
+rtdataplot1 <- summarySEwithin2(rtdata, measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
 ```
 
-```{r rtplot, fig.width=7, fig.height=3, echo=F}
+```{r rtplot, fig.width=7, fig.height=3}
 dodge = position_dodge(width=0.2)

### 2017-05-16 experiments/e3/writeup/e3.Rmd
ggplot computes summary on the fly using mean_cl_boot

@@
 ```{r loadLibraries, message=FALSE}
-library(knitr); library(MASS); library(ggplot2); library(lme4); library(xtable); library(plyr); library(lattice); library(LMERConvenienceFunctions); library(car); library(grid); library(lmerTest); library(reshape2); library(ggplot2); library(languageR); library(pander); source('summarySEwithin2.R')
+library(knitr); library(MASS); library(ggplot2); library(lme4); library(xtable); library(plyr); library(lattice); library(LMERConvenienceFunctions); library(car); library(grid); library(lmerTest); library(reshape2); library(ggplot2); library(languageR); library(pander); source('summarySEwithin2.R'); library(Hmisc)
 ```
@@
 
-```{r computeSummariesForGraphs, cache=TRUE, message=FALSE}
-rtdataplot1 <- summarySEwithin2(rtdata, measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
-```
-
-```{r rtplot, fig.width=7, fig.height=3}
+```{r rtplot1, fig.width=7, fig.height=3}
+# this plot uses a precomputed summary done by summarySEwithin
+# and data trimmed at 2.5 sd per subject
+rtdataplot1 <- summarySEwithin2(data=perSubjectTrim.fnc(rawdata, response='RT', subject='Subject', trim = 2.5)$data,
+                                measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
 dodge = position_dodge(width=0.2)
 rtdataplot1$Condition <- as.factor(paste(sep=' ', rtdataplot1$Selection, rtdataplot1$Vagueness))
-rtplot=ggplot(rtdataplot1, aes(y=RT_logNormed, x=Item, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci,
+rtplot=ggplot(rtdataplot1,
+              aes(y=RT_logNormed, x=Item, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci,
                 group=Condition, shape=Condition, fill=Condition)) +
@@
 rtplot
+```
+
+```{r rtplotHmisc, fig.width=7, fig.height=3}
+# n.b. rtdata has been trimmed at 2.5 sd: use rawdata if you want untrimmed data
+rawdata$Condition <- as.factor(paste(sep=' ', rawdata$Selection, rawdata$Vagueness))
+ggplot(rawdata, aes(y=RT_log, x=Item, group=Condition, shape=Condition, fill=Condition)) + 
+  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(0.2)) +
+  scale_shape_manual("",values = c(22, 22, 21, 21)) +
+  scale_fill_manual("",values=c("black","white","black","white")) +
+  facet_grid(~Selection)
+
+  mapping = aes(y=RT_log, x=Item, group=Condition, shape=Condition, fill=Condition),
+  stat = stat_summary(fun.data = "mean_cl_boot") )
+
+ggplot(mtcars) +
+  geom_pointrange
++
+  facet_grid(~Selection)
+
+d <- ggplot(mtcars, aes(cyl, mpg)) 
+d + stat_summary(fun.data = "mean_cl_boot", colour = "red", size = 2)
+# +
+  # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
+  #       legend.key = element_blank(), aspect.ratio=1,
+  #       axis.text.x = element_text(angle = 45, hjust = 1),
+  #       plot.background=element_rect(fill=NA, color='white')
 ```

### 2017-05-16 experiments/e3/writeup/package_install_script.R

@@ -1,0 +1,2 @@
+install.packages(c("knitr", "MASS", "ggplot2", "lme4", "xtable", "plyr", "lattice", "LMERConvenienceFunctions", "car", 
+                   "lmerTest", "reshape2", "ggplot2", "languageR", "pander"))

### 2017-05-16 experiments/e3/writeup/e3.Rmd
reworked

@@
   scale_fill_manual("",values=c("black","white","black","white")) +
-  facet_grid(~Selection)
-
-  mapping = aes(y=RT_log, x=Item, group=Condition, shape=Condition, fill=Condition),
-  stat = stat_summary(fun.data = "mean_cl_boot") )
-
-ggplot(mtcars) +
-  geom_pointrange
-+
-  facet_grid(~Selection)
-
-d <- ggplot(mtcars, aes(cyl, mpg)) 
-d + stat_summary(fun.data = "mean_cl_boot", colour = "red", size = 2)
-# +
-  # theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-  #       legend.key = element_blank(), aspect.ratio=1,
-  #       axis.text.x = element_text(angle = 45, hjust = 1),
-  #       plot.background=element_rect(fill=NA, color='white')
+  facet_grid(~Selection) +
+  theme(
+    panel.background=element_rect(fill=NA, color='black'),
+    panel.grid.major=element_line(color='grey95'), 
+        #panel.grid.minor = element_blank(),
+        legend.key=element_blank(), 
+        aspect.ratio=1,
+        axis.text.x=element_text(angle = 45, hjust = 1)
+        )
 ```

### 2017-05-17 experiments/e3/writeup/preprocessing/preprocessing.Rmd
reworked
e3 preprocessing

--- 2016-09-27 experiments/e3/writeup/preprocessing/preprocessing.Rmd
@@
 ```{r readdata}
-dat <- read.delim("../rawdata.txt", stringsAsFactors = FALSE)
+
+# The policy for this file as of May 2017 is not to remove any cases (rows). 
+# That means that one case of RT=0 exists, and one case with a very long time RT=49871 exists.
+# The input is the concatenated individual subject files (data_raw.txt).
+# The output is a file (data_processed.txt) that has the same number of cases but more variables.
+
+dat <- read.delim("data_raw.txt", stringsAsFactors = FALSE)
 
@@
 }
+# now delete the tempoary variable RESPONSE
+dat$RESPONSE <- NULL
+dat$Left <- NULL
+dat$Right <- NULL
+dat$Mid <- NULL
+
 dat$crossed=paste('Con',dat$Condition,':Quan',dat$Quantity,':Item',dat$Item,sep="")
@@
 
+# now delete the temporary variable "crossed"
+dat$crossed <- NULL
+
 dat$isResponseExpected <- dat$choice == dat$ResponseExpected
@@
 dat$isResponseFar <- dat$choice ==      dat$ResponseFar
+# now delete ResponseExpected etc
+dat$ResponseExpected <- NULL
+dat$ResponseNear <- NULL
+dat$ResponseFar <- NULL
+dat$choice <- NULL
 
 for ( row in 1:nrow(dat) ) {
-  dat[row, 'switchResponse'] <- 
+  dat[row, 'response_cat'] <- 
     ifelse(dat[row, 'isResponseExpected']==TRUE, 'Expected', 
@@
 }
+# now delete the temporary variables isResponse*
+dat$isResponseExpected <- NULL
+dat$isResponseNear <- NULL 
+dat$isResponseFar <- NULL
 
-## treat variables
-
-# SUBJECT
 # ensure Subject is a factor 
@@
 
-# TRIAL
 # Trial for subject, 1 to 192
@@
 
-# make a centred Trial for modeling
-dat$c_Trl <-dat$Trial - mean(dat$Trial)
-
-# make a scaled Trial for modelling
-dat$s_Trl <- as.numeric(scale(dat$Trial))
-
-# ID
 # id is a unique identifier for the 7680 row data
@@
 
-# ITEM
-# create a centred numeric item variable for modeling
-dat$c_Itm <- ifelse(dat$Item==1, -.75, 
-                    ifelse(dat$Item==2, -.25,
-                           ifelse(dat$Item==3, .25, .75)))
-
 # make Item be a factor and assign labels
@@
 
-# add ratios for Item
-item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54) 
-# 0.2500000 0.4705882 0.5909091 0.6666667
-item_range_ratio_scaled = as.vector(scale(c(6 / 24, 16 / 34, 26 / 44, 36 / 54)))
-# -1.3441995 -0.1316642  0.5297187  0.9461450
-item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))) 
-# 0.5125000 0.6876471 0.7691558 0.8166667
-item_mean_ratio_scaled = as.vector(scale(c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))))) 
-# -1.37582241 -0.06614191  0.54334858 0.89861574
-
-dat[dat$Item == "06:15:24", "item_range_ratio"] <-  0.2500000
-dat[dat$Item == "16:25:34", "item_range_ratio"] <-  0.4705882
-dat[dat$Item == "26:35:44", "item_range_ratio"] <-  0.5909091
-dat[dat$Item == "36:45:54", "item_range_ratio"] <-  0.6666667
-
-dat[dat$Item == "06:15:24", "item_range_ratio_scaled"] <-  -1.3441995
-dat[dat$Item == "16:25:34", "item_range_ratio_scaled"] <-  -0.1316642
-dat[dat$Item == "26:35:44", "item_range_ratio_scaled"] <-   0.5297187  
-dat[dat$Item == "36:45:54", "item_range_ratio_scaled"] <-   0.9461450
-
-dat[dat$Item == "06:15:24", "item_mean_ratio"] <-  0.5125000 
-dat[dat$Item == "16:25:34", "item_mean_ratio"] <-  0.6876471 
-dat[dat$Item == "26:35:44", "item_mean_ratio"] <-  0.7691558
-dat[dat$Item == "36:45:54", "item_mean_ratio"] <-  0.8166667
-
-dat[dat$Item == "06:15:24", "item_mean_ratio_scaled"] <-   -1.37582241 
-dat[dat$Item == "16:25:34", "item_mean_ratio_scaled"] <-   -0.06614191 
-dat[dat$Item == "26:35:44", "item_mean_ratio_scaled"] <-    0.54334858 
-dat[dat$Item == "36:45:54", "item_mean_ratio_scaled"] <-    0.89861574
-
-# VAGUENESS
 # Create a factor coding for Vagueness
@@
 
-# manually center Vagueness
-dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)
-
 dat$Selection = ""
@@
 dat[dat$Condition %in% c(1,2), "Selection"] <- "Matching"
-dat$c_Sel <- ifelse(dat$Selection=="Comparison", -.5, .5)
+dat$Condition <- NULL
+
 dat$Selection <- as.factor(dat$Selection)
-
-# remove one zero RT by replacing with NA
-dat$RT[dat$RT==0]<-NA
 
@@
 
-# make a manually centred Order
-dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)
-
 # QUANTITY
@@
 
-# make a manually centred Quantity
-dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)
-
 # INSTRUCTION
@@
 dat$nchar_instr = nchar(as.character(dat$Instruction))
-dat$nchar_instr_scaled = as.vector(scale(nchar(as.character(dat$Instruction)), scale=TRUE))
 
@@
 # add transformations of RT
-dat$RT_RecSqd    <- 1/dat$RT^2
-dat$RT_Rec       <- 1/dat$RT
-dat$RT_RecSqt    <- 1/sqrt(dat$RT) 
 dat$RT_log       <- log(dat$RT)
-dat$RT_sqt       <- sqrt(dat$RT)
 dat$RT_raw       <- dat$RT
-dat$RT_sqd       <- dat$RT^2
 
-# The dat data set contains *all* trials including impossible trials 
-# and is mainly for graphs comparing different removals
+# add preceding RT
+my_rt <- dat$RT[-1]
 
-# lose impossible trials
-dd <- dat
-dd$RT[dd$RT == 1 ] <- NA
-dd$RT[dd$RT > 40000 ] <- NA
-dd <- dd[complete.cases(dd),]
-row.names(dd) <- NULL
-
-# add preceding RT: because we removed impossible trials, the value for preceding RT for a trial following an impossible trial is the value of the trial that preceded the impossible trial.
-dd$RTprev <- NA
-for (s in levels(dd$Subject)) {
-  nrows = nrow(dd[dd$Subject==s,])
-  for (i in 1:nrows) {
-    if (i==1) {
-      dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i]
-    }
-    else
-      dd[dd$Subject==s, "RTprev"][i] <- dd[dd$Subject==s, "RT"][i-1]
-  }
-}
+dat$RTprev <- c(NA,dat$RT[1:nrow(dat)-1])
 
 # add transformations of previous RT
-dd$RTprev_RecSqd    <- 1/dd$RTprev^2
-dd$RTprev_Rec       <- 1/dd$RTprev
-dd$RTprev_RecSqt0   <- 1/sqrt(dd$RTprev) 
-dd$RTprev_RecSqt    <- as.vector(scale( 1/sqrt(dd$RTprev)  ))
-dd$RTprev_log       <- log(dd$RTprev)
-dd$RTprev_sqt       <- sqrt(dd$RTprev)
-dd$RTprev_raw       <- dd$RTprev
-dd$RTprev_sqd       <- dd$RTprev^2
+dat$RTprev_log       <- log(dat$RTprev)
+dat$RTprev_raw       <- dat$RTprev
 
-# save dd as data.Rda
+dat$RT <- NULL
+dat$RTprev <- NULL
+
+dat$rt_instr <- dat$Instruction_RT; dat$Instruction_RT <- NULL
+
 # This data set non-impossible trials and renamed/recomputed variables
-write.table(dd, file="../data.txt")
+# dat.out <- subset(dat, select=c(Subject, Trial, Item, c_Itm, Vagueness, c_Vag, Selection, c_Sel, Instruction, Order, Quantity, RT, RT_log, response_catResponse))
+# write.table(dat.out, file="data_processed.txt")
+
+dat <- subset(dat, select=c(Subject,Item,id,Trial,Vagueness,Selection, 
+                             Order, Quantity, Instruction, nchar_instr, RT_raw, RT_log, RTprev_log, RTprev_raw, rt_instr, response_cat ))
+
+write.table(dat, "data_processed.txt")
+summary(dat)
+str(dat)
 ```

### 2017-05-17 experiments/e3/writeup/preprocessing/scratch.R

@@ -1,0 +1,29 @@
+# # add ratios for Item
+# item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54) 
+# # 0.2500000 0.4705882 0.5909091 0.6666667
+# item_range_ratio_scaled = as.vector(scale(c(6 / 24, 16 / 34, 26 / 44, 36 / 54)))
+# # -1.3441995 -0.1316642  0.5297187  0.9461450
+# item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))) 
+# # 0.5125000 0.6876471 0.7691558 0.8166667
+# item_mean_ratio_scaled = as.vector(scale(c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))))) 
+# # -1.37582241 -0.06614191  0.54334858 0.89861574
+
+# dat[dat$Item == "06:15:24", "item_range_ratio"] <-  0.2500000
+# dat[dat$Item == "16:25:34", "item_range_ratio"] <-  0.4705882
+# dat[dat$Item == "26:35:44", "item_range_ratio"] <-  0.5909091
+# dat[dat$Item == "36:45:54", "item_range_ratio"] <-  0.6666667
+# 
+# dat[dat$Item == "06:15:24", "item_range_ratio_scaled"] <-  -1.3441995
+# dat[dat$Item == "16:25:34", "item_range_ratio_scaled"] <-  -0.1316642
+# dat[dat$Item == "26:35:44", "item_range_ratio_scaled"] <-   0.5297187  
+# dat[dat$Item == "36:45:54", "item_range_ratio_scaled"] <-   0.9461450
+# 
+# dat[dat$Item == "06:15:24", "item_mean_ratio"] <-  0.5125000 
+# dat[dat$Item == "16:25:34", "item_mean_ratio"] <-  0.6876471 
+# dat[dat$Item == "26:35:44", "item_mean_ratio"] <-  0.7691558
+# dat[dat$Item == "36:45:54", "item_mean_ratio"] <-  0.8166667
+# 
+# dat[dat$Item == "06:15:24", "item_mean_ratio_scaled"] <-   -1.37582241 
[…]

### 2018-04-18 reviews2018/vagueness-springer-review.tex
Dan Lassiter's review of springer book chapter

\documentclass[10pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{lastpage}
\usepackage{fancyhdr}
\pagestyle{fancy} 
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\begin{document}
\title{The elusive benefits of vagueness}
\author{Dan Lassiter}
\date{%
    Stanford University\\[2ex]%
    April 18, 2018
}
\cfoot{\thepage\ of \pageref{LastPage}}

\maketitle

This is a very interesting paper on an important topic. My main critical comment would be that it is unclear what the inclusion of experiment 1 adds to the paper. The experiment seems to have been problematic, as described in the beginning of the presentation of experiment 2. As a result of this skepticism about the cause of the results I found the attempts to interpret the results theoretically in the discussion to be unconvincing. The results from the second experiment seem to be much more solid and theoretically important. Perhaps the paper could be revised to present and interpret the second experiment only, treating the first as a pilot that revealed some issues that were resolved by the main experiment?

A second point is that the discussion of experiment 2 presents some results as more unequivocal than they appear to be. I have some detailed comments below on this.

A third question is whether the context-dependence of many vague expressions could be considered as a source of utility. E.g., \emph{tall} and \emph{warm} can adjust their meanings in ways that allow them to communicate more information about the location of some object relative to the statistical distribution of objects in a reference class. \emph{1.83 meters tall} and \emph{94C} don't have this ability. I've often thought that this flexibility is one of the main sources of utility for vague expressions with this property, since it allows our lexicon to be fairly expressive without having an enormous number of items.

\section*{Small comments}
\begin{itemize}
\item
While I see the value of presenting Lipman's example as such, the example is fairly strange for the purpose he wants to put it to, since it's not at all obvious that in the context describe \emph{187.9 cm} would indeed be more conversationally useful than \emph{tall} as a description of a person's height. I for one would have no idea how to pick out a person of that exact height from a crowd. Perhaps there is a more convincing modification of the example that would make the point, that we are sometimes inclined to use vague language even though a precise description would be uncontroversially more useful to the hearer?
\item
This sort of comes out in the discussion, but the experiment by Mishra et al. reported in section 2 seem to be of limited relevant since they confuse vagueness with lack of specificity. While \emph{vague} is a typical way to describe an unspecific claim like \emph{I was out with someone} as a response to \emph{Where were you?}, this isn't the way we usually use the term in semantic analysis. So, this is an interesting experiment but I wonder if it is really worth discussing here.
\item
The Peters et al. experiment: the comment \emph{it becomes doubtful whether any borderline cases could be conceived to arise in fact} – indeed it's not obvious but I would be surprised if they didn't. When given natural-sounding descriptions like \emph{poor, fair, good, excellent} alongside hard-to-understand numerical values, I would guess that most participants would simply ignore the numerical values since they make less natural descriptions.
\item
Perhaps the data and analysis could be moved to a repo (e.g., github)? This way the URL remains stable even if someone moves institutions, if the university decides to change web addresses, etc. 
\item
Could you include a sample display for each experiment presented?
\item
Why did the statistical analysis include only random slopes and not random intercepts? It would be standard to use both.
\item
p.13: I wouldn't describe indefinites like \emph{a square with \ldots} as referring expressions
\item
exp 2 questions re presentation: 
\begin{itemize}
\item
why was the central number always presenting in the middle, instead of fully randomizing?
\item
\emph{about 10} seems like a terrible description for both 6 and 15. My guess is this would cause subjects to boggle even if they eventually manage to make a guess.
\item
what was the vague numeric instruction used in stimuli with larger dot arrays, e.g. (16,25,34)? [I assume it wasn't \emph{about 10}]
\item
for the numeric condition, the significant main effect reported at the bottom of p.15 (testing H1) is misleading, since the effect seems to be driven entirely by the much lower response time for crisp \emph{6}. Everything else is nearly indistinguishable. So, it seems like the main thing to explain here is why crisp \emph{6} is so much faster than other numerical expressions – presumably because it's a small number and there's no additional cost of trying to resolve the intention behind the fairly poor description \emph{about 10} for 6?
\item
perhaps the big main effect for instruction format: it seems like the interpretation of this result should be that numbers are in general harder to process than vague verbal expressions. This would seem to account for most of the variance in the response time data of experiment 2.
\item
could the differences in the numeric vs verbal borderline response distribution shown in the right panel of fig. 2 be explained by stimulus design – i.e., the hypothesis I mentioned earlier, that \emph{about 6} is just a poor description of both 6 and 15, with the result that participants were choosing more or less random with a slight bias for the number a little bit closer to 10? This would explain why there is almost no difference between borderline and expected in the vague/numeric condition.
\item
p.17 \emph{we found a small (but statisticslly significant) disadvantage of vague instructions} – again this is misleading since the effect is apparently driven almost entirely by a single condition (6, crisp) that was much faster than expected. There appear to be few differences otherwise \ldots one option would be to exclude the 6:15:24 condition and then re-run the statistical analysis to see if any other effects emerge.
\item
p.18 \emph{allowed participants to use a comparison “algorithm” – the work of Justin Halberda and collaborators on the verification strategies associated with \emph{most} and \emph{more than half} is highly relevant here (and indeed throughout the paper, given that \emph{most} is vague but otherwise fairly similar in meaning to \emph{more than half})
\item
p.19 Testing (H1), we have \emph{Vague instructions were non-significantly easier \ldots} do you mean \emph{not significantly easier}? In any case from the graphs on the next page this seems misleading – what it looks like is a substantial main effect where vague instructions have a main effect of lowering the response time, but this effect is masked by interactions with quantity, most notably (again) the huge speed-up for 6 in the matching/crisp condition. I'd recommend following the method described in Roger Levy}s paper on finding main effects in the presence of interactions, available here: \url{https://arxiv.org/abs/1405.2094}.
\item
p.20 \emph{the cost reduction account was wrong to predict significant main effect advantages for vagueness} – this seems way too strong given the issues I've just described. First, the statistical analysis needs to be redone as I just mentioned. Second, in the kind of statistical analysis being employed here it's not possible to show that something doesn't exist simply by failing to find evidence that it does exist. (APA guidelines: \emph{Never use the unfortunate expression \emph{accept the null hypothesis}}, cf. Wagenmakers, \url{http://www.ejwagenmakers.com/2007/pValueProblems.pdf}). At best, you can say \emph{we failed to find evidence that \ldots}
\item
.p21, table 3: were the matching/crisp and matching/vague conditions unmatched in the target quantity as the table says? Or is this a typo?
\item
p.22 \emph{main interaction effect} - ??
\item
p.22 \emph{In the matching task vagueness slowed response times} – again this is not a good summary since there is a big interaction, and the apparent main efect may be driven by a single condition, here a slowdown for 6:15:24 in the matching/vague condition. Re-running using Levy's method would be advisable here.
\end{itemize}
\item
Typo: Paul Egré's name has an acute accent on the second \emph{e}
\end{itemize}
\end{document}

### 2018-04-22 papers/2018-v-book/v-book.tex
2018: bare bones in Springer inhouse style

\documentclass[graybox,envcountchap,sectrefs%
,footinfo
%,referee
]{svmono}
% Native packages
\usepackage{mathptmx}
\usepackage{helvet}
\usepackage{courier}
\usepackage{type1cm}         
\usepackage{makeidx}		% allows index generation
\usepackage{graphicx}		% standard LaTeX graphics tool
\usepackage{multicol} 		% used for the two-column index
\usepackage[bottom]{footmisc}	% places footnotes at page bottom
% User packages
\usepackage{natbib}
\usepackage{lipsum}
\usepackage{placeins}
\makeindex             			% used for the subject index
                       				% please use the style svind.ist with
                       				% your makeindex program

\begin{document}

\author{Kees van Deemter and Matthew Green}
\title{Vagueness \& Rationality}
\subtitle{The Elusive Benefits of Vagueness: the Evidence So Far}
\maketitle

\frontmatter
\tableofcontents

% BibTeX users please use
\bibliographystyle{spbasic} %plain, harvard, apalike, chicago, astron, authordate, natbib.
\bibliography{bibliography}

\backmatter
\Extrachap{Glossary}
%\runinhead{glossary term} Write here the description of the glossary term. Write here the description of the glossary term. Write here the description of the glossary term.

\runinhead{natural language generation} the process of generating natural language from non-natural language input.

\printindex
\end{document}

### 2018-04-22 papers/2018-v-book/review-dan-lassiter/vagueness-springer-review.tex
add Dan Lassiter review

--- 2018-04-18 reviews2018/vagueness-springer-review.tex
@@
 \item
-p.18 \emph{allowed participants to use a comparison “algorithm” – the work of Justin Halberda and collaborators on the verification strategies associated with \emph{most} and \emph{more than half} is highly relevant here (and indeed throughout the paper, given that \emph{most} is vague but otherwise fairly similar in meaning to \emph{more than half})
+p.18 \emph{allowed participants to use a comparison "algorithm"} – the work of Justin Halberda and collaborators on the verification strategies associated with \emph{most} and \emph{more than half} is highly relevant here (and indeed throughout the paper, given that \emph{most} is vague but otherwise fairly similar in meaning to \emph{more than half})
 \item
@@
 In any case from the graphs on the next page this seems misleading – what it looks like is a substantial main effect where vague instructions have a main effect of lowering the response time, but this effect is masked by interactions with quantity, most notably (again) the huge speed-up for 6 in the matching/crisp condition.
-I'd recommend following the method described in Roger Levy}s paper on finding main effects in the presence of interactions, available here: \url{https://arxiv.org/abs/1405.2094}.
+I'd recommend following the method described in Roger Levy's paper on finding main effects in the presence of interactions, available here: \url{https://arxiv.org/abs/1405.2094}.
 \item

### 2018-04-23 papers/2018-v-book/v-book-2017.tex
starting to fill out

--- 2016-10-02 papers/2016-Special-Issue/v-special-issue-2016.tex
@@
 
-\title{Is Vagueness Beneficial for Hearers? \protect \\ Evidence from Experiments}
+\title{The Elusive Benefits of Vagueness \protect \\ Evidence from Experiments}
 \date{\today}
@@
 
-\abstract{Much of everyday language use is vague, but the causes of this phenomenon are not well understood.
-Consequently, it is difficult for the designers of a Natural Language Generation (NLG) system to know when to let the system generate vague expressions.
-The present paper is an attempt to find out what benefits vagueness might have for readers.
-% , by means of a series of carefully controlled psycholinguistic experiments that focus on the use of vagueness in referring expressions.
-The work is motivated by NLG but has implications for linguistics, philosophical logic and game theory.
-%
-We report on a series of experiments that aim to separate the utility of vagueness (as defined by the existence of borderline cases) from the utility of other factors that tend to co-occur with vagueness.
-We argue that the evidence so far supports a view where the benefits that vague terms exert are due to other influences, rather than to vagueness itself.
-These factors include: low granularity; the use of evaluative words; the avoidance of overtly numerical words; the existence of comparison strategies; and, lastly and more tentatively, a phenomenon that we call range reduction.
-The paper concludes with a discussion of the implications for practical NLG.}
+\abstract{Much of everyday language is vague, yet the benefits of vagueness for hearers and readers are proving to be elusive.
+We discuss a range of earlier controlled experiments with human participants, and we report on a new series of experiments that we conducted in recent years.
+These experiments, which focus on vague expressions that are part of referential noun phrases, aim to separate the utility of vagueness (as defined by the existence of borderline cases) from the utility of other factors that tend to co-occur with vagueness.
+Having presented the evidence, we argue that the evidence supports a view where the benefits that vague terms exert are due to other influences, rather than to vagueness itself.}
 
@@
 \section{Introduction}\label{introduction}
-Vagueness pervades the language that we use on a daily basis, and the challenge of understanding vague language has been a prominent concern in many areas of logic and linguistics, involving both theoretical and applied work, including the area known as Natural Language Generation (NLG).
 
-NLG systems transform data and formulas into language \cite<e.g.,>[]{reiter2000building}.
-NLG systems routinely make decisions between different formulations of the same information.
-For example, if the temperature is 27.2 degrees Celsius, this could be expressed as ``27.2 degrees'', ``approximately 27 degrees', ``above 25 degrees", or ``warm'', and the system must decide between these.
-The problem is especially important for NLG systems that take numbers as input, for example in the generation of textual weather reports from numerical weather data such as temperature and wind speed \cite{goldberg1994using,turner2006generating}, and medical decision support on the basis of clinical measurements % such as oxygen saturation, heart rhythm, etc.
-\cite{Hripcsak01032009, hunter2008summarising, portet2009automatic}.
-Such systems are often forced to make decisions concerning the level of precision in the utterances that they generate on the basis of little more than intuition.
+%Vagueness pervades the language that we use on a daily basis, and the challenge of understanding vague language has been a prominent concern in many areas of logic and linguistics, involving both theoretical and applied work, including the area known as Natural Language Generation (NLG).
+
+%NLG systems can be thought of a small and imperfect models of human language use, and as such they have the ability to make questions about language very tangible, so let us take the perspective of an NLG system for a moment.
+In a nutshell, NLG systems transform data and formulas into language \cite<e.g.,>[]{reiter2000building}, making decisions between different formulations of the same information.
+For example, if the temperature is 27.2 degrees Celsius, the system could express this as ``27.2 degrees'', ``approximately 27 degrees', ``above 25 degrees", or ``warm'', and the system must decide between these.
+The problem is important in the generation of textual weather reports from numerical weather data such as temperature and wind speed \cite{goldberg1994using,turner2006generating}, and medical decision support on the basis of clinical measurements % such as oxygen saturation, heart rhythm, etc.
+%\cite{Hripcsak01032009, hunter2008summarising, portet2009automatic}.
+Such systems often make decisions concerning the level of precision in the utterances that they generate on the basis of little more than intuition.
 Even when NLG systems are designed to mimic human language use \cite<e.g.,>{konstas2013global} there is no guarantee that these decisions taken by these systems benefit readers.
 A better understanding of the benefits (for readers and hearers) of different precision levels would allow these systems to become more useful.
-The present article investigates the benefits,  or de-benefits, of vagueness.
+%The present article investigates the benefits, or de-benefits, of vagueness.
 
@@
 
-\subsection{General Methodology}\label{method}
+\subsection{Our Approach to the Problem}\label{method}
 %The experiments reported in the present paper put the cost reduction hypothesis to the test. The question that we are trying to answer is whether
@@
 Like Lipman, we focus on situations where numerical information is used in order to identify a referent.
-Reference, in other words, will be the communicative task on which we focus, partly because of the interest that this topic has recently drawn from the NLG community.  
-% In focussing on benefits for the hearer, we will leave aside the question of audience design, leaving this for later research.
+Reference, in other words, will be the communicative task on which we focus, partly because of the interest that this topic has recently drawn in various areas of Cognitive Science \cite{vanDeemterCMR}.
+By looking at one specific type of vagueness, we will be able to investigate the costs and benefits of vagueness relatively thoroughly.
+Whether our findings generalise to other uses of vagueness is a question on which we will speculate in the final section of this chapter.
 
-In using references to quantities to test the cost reduction hypothesis we are only testing one aspect of vagueness in a particular context. 
-This limits the direct applicability of our results. 
-However, it has the advantage that it enables us to explore the costs and benefits of vagueness more thoroughly. 
-Since one prevalent view of vagueness is that a vague expression is never preferable to a crisp equivalent, a demonstration of a benefit for vagueness in any context would advance the discussion.
+% KvD Feb 2017 (to address a criticism that came up in the reviews for the special issue):
+We have chosen a narrative strategy in which we address a sequence of four experiments with human reader chronologically, explaining how each experiment helped us refine our research question.
+In order to do justice to our findings, we need to describe these experiments in a fair amount of detail.
 
-In our experiments we used a speeded forced choice task to compare the processing costs of different references to quantities. 
+Let us start by explaining the task that was given to the participants in our experiments.
+We used a {\em speeded forced choice} task to compare the processing costs of different references to quantities.
 In this context, speed and accuracy of responses are the key dimensions on which the different references can be compared.
@@
 The participant was asked to respond as quickly as possible while avoiding errors.
-We manipulated the instructions and the arrays in several ways across a series of four experiments.
+We manipulated the instructions and the arrays in several ways across the four experiments.
 
@@
 
-\section{General Discussion}
+\section{General Discussion}\label{discussion}
 
@@
 %
-{\bf Implications for NLG systems.} Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
-The question of how a particular piece of quantitative information is best conveyed through language is an important one for practical NLG; years of research on the logic of vagueness -- giving rise to such techniques as Partial Logic \cite<e.g.,>{Fine}, Probabilistic Logic \cite{Edgington}, and Fuzzy Logic \cite{Zadeh} -- have primed the research community to expect that the utility of vagueness is an important part of the answer, but our findings call this expectation into question.
+% KvD Feb 2017. I've diminished the role of NLG in the coming paragraph.
+Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
+Years of research on the logic of vagueness -- giving rise to such techniques as Partial Logic \cite<e.g.,>{Fine}, Probabilistic Logic \cite{Edgington}, and Fuzzy Logic \cite{Zadeh} -- have primed the research community to expect that some special utility of vagueness is an important part of the answer, but our findings call this expectation into question.
 
 Although our own studies in this article have focussed on vagueness in descriptive Noun Phrases, it seems plausible that vagueness plays a similar role in other linguistic constructs.
-For example, consider reports on air temperature (cf., our discussion of NLG at the start of the paper).
-Given a temperature measurement or prediction as input, an NLG system might say that it will be
+For example, consider reports on air temperature.
+Given a numerical temperature measurement or prediction, we might word it as
 %
@@
 among other candidate expressions.
+Which of these descriptions is most effective, for example as part of a weather report?
 If the linguistic literature is to be believed, then options (a) and (c) convey crisp information, whereas (b) and (d) are vague (i.e., they permit borderline cases).
 From our own experiments and the ones discussed in section 1, there is no evidence that  vagueness should be beneficial for hearers.
-Rather than asking whether a candidate expression is vague, other questions might shed more light on the NLG system's choice, similar to the ones identified in our studies.
+Rather than asking whether a candidate expression is vague, other questions might shed more light on the choice, similar to the ones identified in our studies.
 These questions might focus on the amount of information that a given expression conveys (i.e., on granularity), on the avoidance of numbers, and on the use of evaluative terms.
@@
 %
-If these hunches are correct then an {\em expert} may  to prefer expression (a), because it gives her the most detailed information on which to base her decisions.
-On the other hand, expression (d) (``warm") is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make (d) more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm" (see our section 1 and  \citeA{peters2009bringing}), which is especially important if the hearer is unfamiliar with the metric used (e.g., because she is more used to Fahrenheit).
+If these hunches are correct then an {\em expert} may prefer to read expression (a), because it gives her the most detailed information on which to base her decisions.
+On the other hand, expression (d) (``warm") is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make (d) more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm" (see our section 1 and  \citeA{peters2009bringing}), which is especially important if the hearer is unfamiliar with the metric used.
 These considerations suggest that {\em non-experts} might prefer expression (d).
-If results of this kind were to be confirmed by experiments, they would be applicable in practical NLG, though care would need to be exercised to take the purpose of the communication into account.\footnote{One reason why ``warm" might be less preferred than these considerations suggest is that, in the weather context, some very frequently occurring numbers, such as 20 (degrees), could be relatively easy to process, analogous to the subitizable numbers of \protect\citeA{trick1994small}, see our section \ref{method}.
+%If results of this kind were to be confirmed by experiments, they would be applicable in practical NLG, though care would need to be exercised to take the purpose of the communication into account.\footnote{One reason why ``warm" might be less preferred than these considerations suggest is that, in the weather context, some very frequently occurring numbers, such as 20 (degrees), could be relatively easy to process, analogous to the subitizable numbers of \protect\citeA{trick1994small}, see our section \ref{method}.
 Note that in the experiments reported in this paper, the purpose of the utterances investigated was always clear (e.g., instructing people on a weight loss program \cite{Mishra01042011}; informing a patient's choice of hospital \cite{peters2009bringing}; or selecting a dot array (in our own studies).}
 
-One way to see why vagueness (as defined in our Introduction) may not matter much for NLG -- and for human communication -- is the following thought experiment.
+One way to see why vagueness (as defined in our Introduction) may not benefit human communication is the following thought experiment.
 Suppose a group of speakers understand the word ``warm" as vague, agreeing that temperatures above 26 count as warm, and temperatures below 24 do not count as warm, but considering temperatures between 24 and 26 as borderline cases.
@@
 For example, when the temperature is 40 degrees, then a weather report saying that it is ``above 25 degrees" would be correct yet misleading: this description might be felicitous as a description of temperatures between (roughly) 26-28 degrees, borderline felicitous for 29-33 degrees, and infelicitous at even higher temperatures.
-At its most bold, this perspective would assert that the applicability of {\em every} linguistic expression has borderline cases, hence Lipman's question (see section 1) does not arise.
-This radical perspective -- which would have drastic implications for the construction of an NLG system -- can only be indicated here, but not discussed at length.
+At its most bold, this perspective would assert that the applicability of {\em every} linguistic expression has borderline cases, hence Lipman's question (see section 1) simply does not arise.
+This radical perspective can only be indicated here, but not discussed at length.
 

### 2018-04-23 papers/2018-v-book/v-book.tex
starting to fill out

@@
 
+\mainmatter%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Your text goes here.
+
+\chapter{Chapter Heading}
+\label{intro1} % Always give a unique label
+% use \chaptermark{}
+% to alter or adjust the chapter heading in the running head
+
+Your text goes here. Separate text sections with the standard \LaTeX\
+sectioning commands.
+
+\section{Section Heading}
+\label{sec:1.1}
+% Always give a unique label
+% and use \ref{<label>} for cross-references
+% and \cite{<label>} for bibliographic references
+% use \sectionmark{}
+% to alter or adjust the section heading in the running head
+Your text goes here. Use the \LaTeX\ automatism for your citations
+\cite{vanDeemterCMR}.
+
+\subsection{Subsection Heading}
+\label{sec:2}
+Your text goes here.
+
+\begin{equation}
+\vec{a}\times\vec{b}=\vec{c}
+\end{equation}
+
+\subsubsection{Subsubsection Heading}
+Your text goes here. Use the \LaTeX\ automatism for cross-references as
+well as for your citations, see Sect.~\ref{sec:1}.
+
+\paragraph{Paragraph Heading} %
+Your text goes here.
+
+\subparagraph{Subparagraph Heading.} Your text goes here.%
+%
+\index{paragraph}
+% Use the \index{} command to code your index words
+%
+% For tables use
+%
+\begin{table}
+\centering
+\caption{Please write your table caption here}
+\label{tab:1}       % Give a unique label
+%
+% For LaTeX tables use
+%
+\begin{tabular}{lll}
+\hline\noalign{\smallskip}
+first & second & third  \\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+number & number & number \\
+number & number & number \\
+\noalign{\smallskip}\hline
+\end{tabular}
+\end{table}
+%
+%
+% For figures use
+%
+\begin{figure}
+\centering
+% Use the relevant command for your figure-insertion program
+% to insert the figure file.
+% For example, with the option graphics use
+%\includegraphics[height=4cm]{figure.eps}
+%
+% If not, use
+%\picplace{5cm}{2cm} % Give the correct figure height and width in cm
+%
+\caption{Please write your figure caption here}
+\label{fig:1}       % Give a unique label
+\end{figure}
+%
+% For built-in environments use
+%
+\begin{theorem}
+Theorem text goes here.
+\end{theorem}
+%
+% or
+%
+\begin{lemma}
+Lemma text goes here.
+\end{lemma}
+%
+%
+% Problems or Exercises should be sorted chapterwise
+\section*{Problems}
+\addcontentsline{toc}{section}{Problems}
+%
+% Use the following environment.
+% Don't forget to label each problem;
+% the label is needed for the solutions' environment
+\begin{prob}
+\label{prob1}
+The problem\footnote{Footnote} is described here. The
+problem is described here. The problem is described here.
+\end{prob}
+
+\begin{prob}
+\label{prob2}
+\textbf{Problem Heading}\\
+(a) The first part of the problem is described here.\\
+(b) The second part of the problem is described here.
+\end{prob}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% our chapter START
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\chapter{The Elusive Benefits of Vagueness: Evidence from Experiments}
+\label{intro} % Always give a unique label
+% use \chaptermark{}
+% to alter or adjust the chapter heading in the running head
+
+Much of everyday language is vague, even in situations where vagueness could have been
+avoided (i.e., where vagueness is used “strategically"). Yet the benefits of vagueness for
+hearers and readers are proving to be elusive. We discuss a range of earlier controlled
+experiments with human participants, and we report on a new series of experiments that
+we conducted in recent years. These experiments, which focus on vague expressions that
+are part of referential noun phrases, aim to separate the utility of vagueness (as defined by
+the existence of borderline cases) from the utility of other factors that tend to co-occur
+with vagueness. Having presented the evidence, we argue that the evidence supports a view
+where the benefits that vague terms exert are due to other influences, and not to vagueness
+itself.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% our chapter END
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\backmatter%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 % BibTeX users please use
 \bibliographystyle{spbasic} %plain, harvard, apalike, chicago, astron, authordate, natbib.
-\bibliography{bibliography}
+\bibliography{v-book}
 
 \backmatter
-\Extrachap{Glossary}
-%\runinhead{glossary term} Write here the description of the glossary term. Write here the description of the glossary term. Write here the description of the glossary term.
-
-\runinhead{natural language generation} the process of generating natural language from non-natural language input.
 

### 2018-07-18 experiments/e0/writeup/e0.Rmd
revisited in July 2018

--- 2016-07-28 experiments/e0/writeup/e0.Rmd
@@
     toc: yes
-    <!--toc_float: true-->
+    toc_float: true
     mathjax: default
     template: default
-    code_folding: hide
+    code_folding: show
     theme: default
+
 ---
@@
 
-Knitr doc options.
+The pilot experiment used subitizable numbers as well as bigger numbers.
+The instructions had the form: _Choose the square with {{few, many},{two, three, four, five, six, seven, eight, nine}} dots._
+The variables were: subitizability; vagueness; target_size; gap_size; side (i.e., which side was the bigger number on?)
 
-```{r docOpts}
+# Libraries and functions
+
+```{r load knitr}
 library(knitr)
-knit_hooks$set(crop=hook_pdfcrop)
 ```
@@
 
-```{r loadLibraries}
-suppressPackageStartupMessages(library(MASS))
-suppressPackageStartupMessages(library(knitr))
-suppressPackageStartupMessages(library(ggplot2))
-suppressPackageStartupMessages(library(lme4))
-suppressPackageStartupMessages(library(xtable))
-suppressPackageStartupMessages(library(plyr))
-suppressPackageStartupMessages(library(reshape))
-suppressPackageStartupMessages(library(lattice))
-suppressPackageStartupMessages(library(LMERConvenienceFunctions))
-suppressPackageStartupMessages(library(car))
-suppressPackageStartupMessages(library(gridExtra))
-suppressPackageStartupMessages(library(grid))
-suppressPackageStartupMessages(library(lmerTest))
+```{r load libraries, message=FALSE, warning=FALSE}
+library(tidyverse)
+library(broom)
+library(dotwhisker)
+library(grid)
+library(gridExtra)
+library(lme4)
+# R CMD INSTALL broom.mixed # https://github.com/bbolker/broom.mixed
 ```
@@
 ```{r sourceCIFunctions}
-source("CIfunctions.R")
+source("CIfunctions.R") # for summarySEwithin etc
 ```
 
-# Preprocessing
+# Data
 
@@
 
-```{r getData}
-if( file.exists('data.txt') ) 
-	{
-	d <- read.table("data.txt", header=FALSE) 
-} else {
-	system("cat ../experimentCode/output/*.dat > data.txt", wait=TRUE)
+```{r "make the data.txt file unless it's already there; read the data from file, add some vars and code the added vars"}
+
+# make the data file from the individual subjects' data unless it's already there in the directory
+if (!file.exists("data.txt")) {
+  system("echo 'gap\ttarget\tside\tvagueness\tpair\tVersion\tLeft\tRight\tText\tCondition\tcorrectAnswer\tsubject\tactualAnswer\tisCorrect\trt' > data.txt")
+  system("cat ../experimentCode/output/*.dat >> data.txt", wait=TRUE)
 }
-mynames <- list("gap", "target", "side", "vagueness", "pair", "Version", "Left",
-								"Right", "Text", "Condition", "correctAnswer", "subject",
-								"actualAnswer", "isCorrect", "rt")
-names(d) <- mynames
-d$obs <- 1:nrow(d)
-d$trial <- 0
-for (s in unique(d$subject)){
-  d[d$subject==s, "trial"] <- 1:nrow(subset(d,subject==s))
+
+# read in the data from file
+din <- read_delim("data.txt", delim="\t")
+
+# make an R object copy of that original data file
+original_data <- din
+
+# add observation number for the whole data set
+din$obs <- 1:nrow(din)
+
+# add trial index for each subject
+din$trial <- 0
+for (s in unique(din$subject)){
+  din[din$subject==s, "trial"] <- 1:nrow(subset(din,subject==s))
 }
-# For some reason there is one rt of -6. Set it to NA
-d[d$rt==-6,"rt"] <- NA
-# make a df
-d <- data.frame(
-  obs = d$obs, 
-  trial = d$trial, 
-  subject = d$subject, 
-  item = factor(d$pair, label = c("2:4", "3:5", "6:8", "7:9", "2:6", "7:3", "8:4", "5:9")),
-  gap = factor(d$gap, levels = c("BigGap", "SmallGap"), labels = c("Big Gap", "Small Gap")), 
-  quan = factor(d$target, levels = c("BigTarget", "SmallTarget"), labels = c("Big Target", "Small Target")), 
-  vag = factor(d$vagueness, levels = c("Precise", "Vague"), labels = c("Crisp", "Vague")), 
-  side = d$side, 
-  err = 1 - d$isCorrect, 
-  rt = d$rt)
-d$subit <- "Not subitizable"
-d$subit[d$item == "2:4"] <- "Subitizable"
-d$subit[d$item == "3:5"] <- "Subitizable"
-d$subit[d$item == "2:6"] <- "Subitizable"
-d$subit[d$item == "7:3"] <- "Subitizable"
-d$subit = as.factor(d$subit)
-d <- subset(d, select = c("obs", "trial", "subject", "item", "gap", "quan", 
-                          "vag", "subit", "side", "err", "rt"))
 ```
 
-Size of the data is \Sexpr{nrow(d)} trials. 1 was removed for being minus 6 rt.
-Remove erroneous response trials. There are \Sexpr{nrow(subset(d,err==1))} error trials.
-
-```{r removeErrorTrials}
-# by setting the RT to NA
-d[d$err==1,"rt"] <- NA
+```{r"make a tibble with factors explicit"}
+# make a df to carry forward
+d <- tibble(
+  obs = din$obs, 
+  trial = din$trial, 
+  subject = as.factor(din$subject), 
+  item = factor(din$pair, label = c("2:4", "3:5", "6:8", "7:9", "2:6", "7:3", "8:4", "5:9")),
+  gap = factor(din$gap, levels = c("BigGap", "SmallGap"), labels = c("Big Gap", "Small Gap")), 
+  quan = factor(din$target, levels = c("BigTarget", "SmallTarget"), labels = c("Big Target", "Small Target")), 
+  vag = factor(din$vagueness, levels = c("Precise", "Vague"), labels = c("Crisp", "Vague")), 
+  subit = factor(ifelse(din$pair %in% c(1,2,5,6), "Subitizable", "Not subitizable")),
+  side = factor(din$side), 
+  text = factor(din$Text),
+  err = 1 - din$isCorrect, 
+  rt_raw = ifelse(din$rt >=0, din$rt, NA),
+  rt_ms = ifelse(1-din$isCorrect==1, NA, ifelse(din$rt >=0, din$rt, NA)), # restricting rt to positive values (rt>=0) converted a score of -6 ms to NA
+  rt_log = log( ifelse(1-din$isCorrect==1, NA, ifelse(din$rt >=0, din$rt, NA)) ),
+  "rt_reciprocal[-1/rt]" = -1/ifelse(1-din$isCorrect==1, NA, ifelse(din$rt >=0, din$rt, NA)), # We multiplied reciprocal scores [1/RT] by -1 [-1/RT] to maintain the direction of effects compatible for the three variants, effectively converting speed [number of responses per ms] into 'rate of slowing' - A linear mixed model analysis of masked repetition priming, Kliegl, Masson & Richter, VISUAL COGNITION,2010,18(5),655--681 (p.~662)
+  rt_reciprocal = -1000/ifelse(1-din$isCorrect==1, NA, ifelse(din$rt >=0, din$rt, NA))
+)
+d$c.vag = ifelse(d$vag=="Crisp", -.5, .5)
 ```
 
-Do some factors and contrasts.  Set subject and item as factors. Sum code vagueness and subitizability.
+Size of the data is `r nrow(d)` trials.
+1 was removed for being impossible ($-6~ms$).
+There are `r nrow(subset(d,err==1))` error trials.
+The variable rt_ms has NA in the case that an error was made.
+The variable rt_raw includes rt for error trials.
 
-```{r postProcessData}
-d$subject <- as.factor(d$subject)
-d$item <- as.factor(d$item)
-d$vag <- relevel(d$vag, ref="Vague")
-contrasts(d$vag) <- contr.sum(2)/2
-d$subit <- relevel(d$subit, ref="Subitizable")
-contrasts(d$subit) <- contr.sum(2)/2
-contrasts(d$quan) <- contr.sum(2)/2
-contrasts(d$gap) <- contr.sum(2)/2
+## Eyeball the RTs
+
+### Uncapped, max of 20,000 ms
+
+```{r fig.show='hide', warning=FALSE, echo=FALSE}
+# all the (correct) RTs
+summary(d$rt_ms)
+
+data_for_plot <- gather(d, key=rt_metric, value=rt_value, rt_ms, rt_log, "rt_reciprocal[-1/rt]") %>%
+  arrange(rt_value)
+
+ga1untrimmed = ggplot(data_for_plot, aes(x=rep(1:1600,times=3), y=rt_value)) + 
+  geom_point(na.rm=TRUE) + 
+  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal[-1/rt]")), scales="free_y") +
+  xlab("observations, ordered by rt_value") + theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))
+
+ga2untrimmed = ggplot(data_for_plot, aes(rt_value)) +
+  geom_density(na.rm=TRUE) +
+  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal[-1/rt]")), scales="free") + theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))
+
+myg1 <- grid.arrange(ga1untrimmed, ga2untrimmed, nrow=2, top = textGrob("Untrimmed",gp=gpar(fontsize=20,font=3)))
 ```
 
-Check the contrasts.
+### Capped at 10,000 ms
 
-```{r checkContrasts}
-contrasts(d$vag)
-contrasts(d$subit)
-contrasts(d$quan)
-contrasts(d$gap)
+```{r fig.show='hide', warning=FALSE, echo=FALSE}
+# now with a simple cap imposed at 10000 ms, half-way through the maximum allotted time of 20000 (2 minute timeout)
+summary(d$rt_ms[d$rt_ms<10000])
+
+data_for_plot <- gather(d %>% filter(rt_ms<10000), key=rt_metric, value=rt_value, rt_ms, rt_log, "rt_reciprocal[-1/rt]") %>%
+  arrange(rt_value)
+
+ga1trimmed = ggplot(data_for_plot, aes(x=rep(1:1556,times=3), y=rt_value)) + 
+  geom_point(na.rm=TRUE) + 
+  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal[-1/rt]")), scales="free_y") +
+  xlab("observations, ordered by rt_value")+ theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))
+
+ga2trimmed = ggplot(data_for_plot, aes(rt_value)) +
+  geom_density(na.rm=TRUE) +
+  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal[-1/rt]")), scales="free") + theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))
+
+myg2 <- grid.arrange(ga1trimmed, ga2trimmed, nrow=2, top = textGrob("Trimmed at 10,1000 ms",gp=gpar(fontsize=20,font=3)))
 ```
 
-Add transformations of rt
-
-```{r addTransform}
-d$logRT <- log(d$rt)
-d$recipRT <- -10000/d$rt
-```
-
-For the untrimmed data, show effects of transforming rt.
-
-```{r plotDistributions, fig.width=4, fig.height=2}
-xs = melt(d,  
-					measure.vars = c("rt","logRT","recipRT"))
-densityplot(
-	~value|variable,
-	data=xs,
-	xlab=NULL,
-	yaxt=F,
-  scales=list(x=list(relation="free", rot=45), 
-   						y=list(relation="free", at=NULL)),
-	cex=.2,
-	plot.points=FALSE,
-	layout=c(3,1)
-	)
-```
-
-Trim the data. This function doesn't work well if there are NA values in the response, so explicitly remove them first.
-
-```{r trimData}
-dummy <- subset(d, !is.na(rt))
-data.trim <- perSubjectTrim.fnc(dummy, response="rt",subject="subject", trim=2.5)$data
-```
-
-Check the distribution of the trimmed data to see if a transformation is suggested.
-
-```{r checkTrim, fig.width=4, fig.height=2}
-xs = melt(data.trim,  
-					measure.vars = c("rt","logRT","recipRT"))
-densityplot(
-	~value|variable,
-	data=xs,
-	xlab=NULL,
-	yaxt=F,
-  scales=list(x=list(relation="free", rot=45), 
-   						y=list(relation="free", at=NULL)),
-	cex=.2,
-	plot.points=FALSE,
-	layout=c(3,1)
-	)
+```{r, fig.width=10,fig.height=10}
+grid.arrange(myg1, myg2, nrow=2)
 ```
@@
 ```{r selectTrim}
-dat <- data.trim
+dat <- d %>% filter(rt_ms<10000)
 ```
@@
 ```{r, lmers}
-lmer2 <- lmer(data=dat, logRT~subit*vag + (subit*vag|subject) + (vag|item))
+lmer2 <- lmer(data=dat, rt_reciprocal~subit*vag + (vag|subject) + (vag|item))
 # lmer3 is closer to the original anova than lmer2
-lmer3 <- lmer(data=dat, logRT~subit*vag+quan+gap + (subit*vag|subject) + (vag|item))
+lmer3 <- tidy(lme4::lmer(data=dat, rt_reciprocal~vag*subit+quan+gap + (vag+subit|subject) + (vag|item)))
+dwplot(lmer3 %>% filter(group=="fixed"), vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2)) + xlab("Coefficient")
 ```
 
-```{r, summaryLMER, results='asis'}
-print(xtable(summary(lmer3)$coefficients, digits=3), type="html")
-```
+<!-- ```{r} -->
+<!-- lmer3b <- lme4::lmer(data=dat, logRT~subit*vag+quan+gap + (subit*vag|subject) + (vag|item))  -->
+<!-- lmer3b_tidy <- tidy(lmer3b) -->
+<!-- lmer3b_tidy$model="foo" -->
+<!-- lmer3b_tidy %>% filter(group=="fixed") %>% dwplot() -->
+<!-- ``` -->
 
-## subitizable
+<!-- ```{r, summaryLMER, results='asis'} -->
+<!-- print(xtable(summary(lmer3)$coefficients, digits=3), type="html") -->
+<!-- ``` -->
 
-```{r, subitRT, results='asis'}
-print(xtable(summary(lmer(data=dat, subset=subit=="Subitizable",logRT~vag+quan+gap +(vag|subject)+(vag|item)))$coefficients, digits=3), type='html')
-```
+<!-- ## subitizable -->
 
-## not subitizable
+<!-- ```{r, subitRT, results='asis'} -->
+<!-- print(xtable(summary(lmer(data=dat, subset=subit=="Subitizable",logRT~vag+quan+gap +(vag|subject)+(vag|item)))$coefficients, digits=3), type='html') -->
+<!-- ``` -->
 
-```{r, notsubitrT, results='asis'}
-print(xtable(summary(lmer(data=dat, subset=subit=="Not subitizable", logRT~vag+quan+gap +(vag|subject)+(vag|item)))$coefficients, digits=3), type='html')
-```
+<!-- ## not subitizable -->
 
-# Error rate analysis
+<!-- ```{r, notsubitrT, results='asis'} -->
+<!-- print(xtable(summary(lmer(data=dat, subset=subit=="Not subitizable", logRT~vag+quan+gap +(vag|subject)+(vag|item)))$coefficients, digits=3), type='html') -->
+<!-- ``` -->
 
-```{r errorrates, results='asis'}
-print(xtable(summary(
-	glmer(data=d, err~vag*subit+quan+gap + (vag|subject) + (1|item), 
-				family=binomial, 
-				control = glmerControl(optimizer = "bobyqa")
-				)
-	)$coefficients,
-	digits=3), 
-	type='html')
-```
+<!-- # Error rate analysis -->
 
-# Plots of significant effects
+<!-- ```{r errorrates, results='asis'} -->
+<!-- print(xtable(summary( -->
+<!-- 	glmer(data=d, err~vag*subit+quan+gap + (vag|subject) + (1|item),  -->
+<!-- 				family=binomial,  -->
+<!-- 				control = glmerControl(optimizer = "bobyqa") -->
+<!-- 				) -->
+<!-- 	)$coefficients, -->
+<!-- 	digits=3),  -->
+<!-- 	type='html') -->
+<!-- ``` -->
 
-```{r plot_sig, fig.width=7.5, fig.height=3}
-a <- summarySEwithin(data=dat, measurevar="logRT", withinvars=c("vag","subit","item"), idvar="subject")
-as <- subset(a,subit=="Subitizable")
-as$item <- as.character(as$item)
-as$item[as$item=="7:3"] <- "3:7"
-as$item <- factor(as$item)
-pd <- position_dodge(.05)
-plot1 = ggplot(data=as, aes(y=logRT, x=item, ymin=logRT-ci, ymax=ci+logRT, fill=vag, group=vag)) + 
-  geom_line(position=pd) +
-  geom_errorbar(width=.2, position=pd)+
-  geom_point(pch=21, size=4, position=pd) +
-  scale_fill_grey(name=element_blank(),start=0,end=1)+
-  theme_bw()+
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key=element_blank()) +
-  ggtitle("Response time")
+<!-- # Plots of significant effects -->
 
-a <- summarySEwithin(data=d, measurevar="err", withinvars=c("vag","subit","item"), idvar="subject")
-as <- subset(a,subit=="Subitizable")
-as$item <- as.character(as$item)
-as$item[as$item=="7:3"] <- "3:7"
-as$item <- factor(as$item)
-pd <- position_dodge(.05)
-plot2 = ggplot(data=as, aes(y=err, x=item, ymin=err-ci, ymax=ci+err, group=vag, fill=vag))+
-  geom_line(position=pd) +
-  geom_errorbar(width=.2, position=pd)+
-  geom_point(pch=21, size=4, position=pd) +
-  scale_fill_grey(name=element_blank(),start=0,end=1)+
-  theme_bw()+
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key=element_blank()) +
-  ggtitle("Error rate")
+<!-- ```{r plot_sig, fig.width=7.5, fig.height=3} -->
+<!-- a <- summarySEwithin(data=dat, measurevar="logRT", withinvars=c("vag","subit","item"), idvar="subject") -->
+<!-- as <- subset(a,subit=="Subitizable") -->
+<!-- as$item <- as.character(as$item) -->
+<!-- as$item[as$item=="7:3"] <- "3:7" -->
+<!-- as$item <- factor(as$item) -->
+<!-- pd <- position_dodge(.05) -->
+<!-- plot1 = ggplot(data=as, aes(y=logRT, x=item, ymin=logRT-ci, ymax=ci+logRT, fill=vag, group=vag)) +  -->
+<!--   geom_line(position=pd) + -->
+<!--   geom_errorbar(width=.2, position=pd)+ -->
+<!--   geom_point(pch=21, size=4, position=pd) + -->
+<!--   scale_fill_grey(name=element_blank(),start=0,end=1)+ -->
+<!--   theme_bw()+ -->
+<!--   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key=element_blank()) + -->
+<!--   ggtitle("Response time") -->
 
-grid.arrange(plot1, plot2, nrow=1, top="Subitizable only")
-```
+<!-- a <- summarySEwithin(data=d, measurevar="err", withinvars=c("vag","subit","item"), idvar="subject") -->
+<!-- as <- subset(a,subit=="Subitizable") -->
+<!-- as$item <- as.character(as$item) -->
+<!-- as$item[as$item=="7:3"] <- "3:7" -->
+<!-- as$item <- factor(as$item) -->
+<!-- pd <- position_dodge(.05) -->
+<!-- plot2 = ggplot(data=as, aes(y=err, x=item, ymin=err-ci, ymax=ci+err, group=vag, fill=vag))+ -->
+<!--   geom_line(position=pd) + -->
+<!--   geom_errorbar(width=.2, position=pd)+ -->
+<!--   geom_point(pch=21, size=4, position=pd) + -->
+<!--   scale_fill_grey(name=element_blank(),start=0,end=1)+ -->
+<!--   theme_bw()+ -->
+<!--   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key=element_blank()) + -->
+<!--   ggtitle("Error rate") -->
 
-```{r barcharts, fig.width=9, fig.height=3}
-bwtheme <- standard.theme("pdf", color=FALSE)
-b1=barchart(data=ddply(dat,.(vag,subit),summarise, logRT=mean(logRT)), 
-						logRT~subit, 
-						group=vag, auto.key=list(space='right', cex=.7), par.settings=bwtheme)
-b2=barchart(data=ddply(d,.(vag,subit),summarise, Error_Rate=mean(err)), 
-						Error_Rate~subit, 
-						group=vag, auto.key=list(space='right', cex=.7), par.settings=bwtheme)
-grid.arrange(b1,b2,ncol=2)
-```
+<!-- grid.arrange(plot1, plot2, nrow=1, top="Subitizable only") -->
+<!-- ``` -->
 
-Subitizability x Vagueness x Quantity
+<!-- ```{r barcharts, fig.width=9, fig.height=3} -->
+<!-- bwtheme <- standard.theme("pdf", color=FALSE) -->
+<!-- b1=barchart(data=ddply(dat,.(vag,subit),summarise, logRT=mean(logRT)),  -->
+<!-- 						logRT~subit,  -->
+<!-- 						group=vag, auto.key=list(space='right', cex=.7), par.settings=bwtheme) -->
+<!-- b2=barchart(data=ddply(d,.(vag,subit),summarise, Error_Rate=mean(err)),  -->
+<!-- 						Error_Rate~subit,  -->
+<!-- 						group=vag, auto.key=list(space='right', cex=.7), par.settings=bwtheme) -->
+<!-- grid.arrange(b1,b2,ncol=2) -->
+<!-- ``` -->
 
-```{r plotMeansa, fig.width=7, fig.height=3}
-a <- summarySEwithin(data=dat, measurevar="logRT", withinvars=c("vag","quan","subit"), idvar="subject")
-pd <- position_dodge(.1)
-	ggplot(
-		a, 
-		aes(y=logRT, x=vag, group=quan, shape=quan, col=quan, ymin=logRT-ci, ymax=ci+logRT)
-	) + 
-	geom_point(size=4, position=pd) + 
-	geom_line(position=pd) + 
-	facet_grid(~subit) + 
-	geom_errorbar(width=.1, position=pd) + 
-	theme_bw() + 
-	theme(panel.grid.major=element_blank()) 
-```
+<!-- Subitizability x Vagueness x Quantity -->
 
-```{r qqplotLMER, fig.width=3, fig.height=3.5}
-qqPlot(residuals(lmer2),cex=.5)
-```
+<!-- ```{r plotMeansa, fig.width=7, fig.height=3} -->
+<!-- a <- summarySEwithin(data=dat, measurevar="logRT", withinvars=c("vag","quan","subit"), idvar="subject") -->
+<!-- pd <- position_dodge(.1) -->
+<!-- 	ggplot( -->
+<!-- 		a,  -->
+<!-- 		aes(y=logRT, x=vag, group=quan, shape=quan, col=quan, ymin=logRT-ci, ymax=ci+logRT) -->
+<!-- 	) +  -->
+<!-- 	geom_point(size=4, position=pd) +  -->
+<!-- 	geom_line(position=pd) +  -->
+<!-- 	facet_grid(~subit) +  -->
+<!-- 	geom_errorbar(width=.1, position=pd) +  -->
+<!-- 	theme_bw() +  -->
+<!-- 	theme(panel.grid.major=element_blank())  -->
+<!-- ``` -->
+
+<!-- ```{r qqplotLMER, fig.width=3, fig.height=3.5} -->
+<!-- qqPlot(residuals(lmer2),cex=.5) -->
+<!-- ``` -->

### 2018-07-18 experiments/e0/writeup/broommm.R

@@ -1,0 +1,23 @@
+library(lme4)
+library(broom)
+
+lmm1 <- lme4::lmer(Reaction ~ Days + (Days | Subject), sleepstudy)
+tidy(lmm1)
+tidy(lmm1, effects = "fixed")
+tidy(lmm1, effects = "fixed", conf.int=TRUE)
+tidy(lmm1, effects = "fixed", conf.int=TRUE, conf.method="profile")
+tidy(lmm1, effects = "ran_modes", conf.int=TRUE)
+head(augment(lmm1, sleepstudy))
+glance(lmm1)
+library(dotwhisker)
+dwplot(lmm1)
+
+
+lmm2 <- lmerTest::lmer(Reaction ~ Days + (Days | Subject), sleepstudy)
+tidy(lmm2)
+tidy(lmm2, effects = "fixed")
+tidy(lmm2, effects = "fixed", conf.int=TRUE)
+tidy(lmm2, effects = "fixed", conf.int=TRUE, conf.method="profile")
+tidy(lmm2, effects = "ran_modes", conf.int=TRUE)
+head(augment(lmm2, sleepstudy))
+glance(lmm2)

### 2018-07-18 experiments/e0/writeup/scratch.R

@@ -1,45 +1,0 @@
-rm(d)
-d <- read.table("data.txt", header=FALSE) # read in the data for all subjects
-mynames <- list("gap_size", 
-                "target_size", 
-                "correct_side", 
-                "vagueness",
-                "pair", 
-                "Version", 
-                "Left",
-                "Right", 
-                "Text", 
-                "Condition", 
-                "correctAnswer",
-                "subject",
-                "actualAnswer",
-                "isCorrect",
-                "rt")
-names(d) <- mynames
-# For some reason there is one rt of -6. Set it to NA
-d[d$rt==-6,"rt"] <- NA
-d$observation <- 1:nrow(d)
-d$trial <- 0
-for (s in unique(d$subject)){
-  d[d$subject==s, "trial"] <- 1:nrow(subset(d,subject==s))
-}
-d$item = ordered(d$pair, label = c("2:4", "3:5", "6:8", "7:9", "2:6", "7:3", "8:4", "5:9"))
-d$subit <- "Not subitizable"
-d$subit[d$item == "2:4"] <- "Subitizable"
-d$subit[d$item == "3:5"] <- "Subitizable"
-d$subit[d$item == "2:6"] <- "Subitizable"
-d$subit[d$item == "7:3"] <- "Subitizable"
-d$dotpair_id=d$item); d$item=NULL
-d$subit = as.factor(d$subit)
-d$gap_size = factor(d$gap_size, levels = c("BigGap", "SmallGap"), labels = c("Big", "Small"))
-d$target_size = factor(d$target_size, levels = c("BigTarget", "SmallTarget"), labels = c("Big", "Small"))
-d$vagueness = factor(d$vagueness, levels = c("Precise", "Vague"), labels = c("Crisp", "Vague")) 
-d$is_error = 1 - d$isCorrect
-d$dots_right=d$Right; d$Right=NULL
-d$dots_left=d$Left; d$Left=NULL
-dd <- subset(d, select=c(
-  observation, subject, trial, gap_size, target_size, correct_side, dots_left, dots_right, dotpair_id, vagueness, Text, is_error, rt))
-
-head(dd,30)
-
-summary(dd)

### 2018-07-20 experiments/e0/writeup/e0.Rmd
e0 is a mess

@@
 
-```{r"make a tibble with factors explicit"}
+```{r "make a tibble with factors explicit"}
 # make a df to carry forward
@@
   err = 1 - din$isCorrect, 
-  rt_raw = ifelse(din$rt >=0, din$rt, NA),
-  rt_ms = ifelse(1-din$isCorrect==1, NA, ifelse(din$rt >=0, din$rt, NA)), # restricting rt to positive values (rt>=0) converted a score of -6 ms to NA
-  rt_log = log( ifelse(1-din$isCorrect==1, NA, ifelse(din$rt >=0, din$rt, NA)) ),
-  "rt_reciprocal[-1/rt]" = -1/ifelse(1-din$isCorrect==1, NA, ifelse(din$rt >=0, din$rt, NA)), # We multiplied reciprocal scores [1/RT] by -1 [-1/RT] to maintain the direction of effects compatible for the three variants, effectively converting speed [number of responses per ms] into 'rate of slowing' - A linear mixed model analysis of masked repetition priming, Kliegl, Masson & Richter, VISUAL COGNITION,2010,18(5),655--681 (p.~662)
-  rt_reciprocal = -1000/ifelse(1-din$isCorrect==1, NA, ifelse(din$rt >=0, din$rt, NA))
+  rt_ms = din$rt,
+  rt_log = log( din$rt ),
+  rt_reciprocal = -1/din$rt # rate of slowing: # We multiplied reciprocal scores [1/RT] by -1 [-1/RT] to maintain the direction of effects compatible for the three variants, effectively converting speed [number of responses per ms] into 'rate of slowing' - A linear mixed model analysis of masked repetition priming, Kliegl, Masson & Richter, VISUAL COGNITION,2010,18(5),655--681 (p.~662)
 )
-d$c.vag = ifelse(d$vag=="Crisp", -.5, .5)
 ```
 
-Size of the data is `r nrow(d)` trials.
-1 was removed for being impossible ($-6~ms$).
-There are `r nrow(subset(d,err==1))` error trials.
-The variable rt_ms has NA in the case that an error was made.
-The variable rt_raw includes rt for error trials.
+```{r "make centred variables for factors"}
+d$c.gap   <- ifelse(d$gap=="Big Gap", -.5, .5)
+d$c.quan  <- ifelse(d$quan=="Big Target", -.5, .5)
+d$c.vag   <- ifelse(d$vag=="Crisp", -.5, .5)
+d$c.subit <- ifelse(d$subit=="Subitizable", -.5, .5)
+d$c.side  <- ifelse(d$side=="Left", -.5, .5)
+```
+
+Size of the data is `r nrow(d)` trials.  There are `r nrow(subset(d,err==1))` error trials. Error trials are removed from analysis of RT
+
+```{r}
+d <- subset(d, err==0)
+```
+
+<!-- 1 was set to NA for being impossible ($-6~ms$). -->
+<!-- The variable rt_ms has NA in the case that an error was made. The variable rt_raw includes rt for error trials. -->
 
@@
 
-### Uncapped, max of 20,000 ms
+### Uncapped
 
@@
 
-data_for_plot <- gather(d, key=rt_metric, value=rt_value, rt_ms, rt_log, "rt_reciprocal[-1/rt]") %>%
+data_for_plot <- d %>% 
+  gather(key=rt_metric, value=rt_value, rt_ms, rt_log, rt_reciprocal) %>%
   arrange(rt_value)
 
-ga1untrimmed = ggplot(data_for_plot, aes(x=rep(1:1600,times=3), y=rt_value)) + 
+ga1untrimmed = ggplot(data_for_plot, aes(x=rep(1:(nrow(data_for_plot)/3),times=3), y=rt_value)) + 
   geom_point(na.rm=TRUE) + 
-  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal[-1/rt]")), scales="free_y") +
+  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal")), scales="free_y") +
   xlab("observations, ordered by rt_value") + theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))
@@
   geom_density(na.rm=TRUE) +
-  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal[-1/rt]")), scales="free") + theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))
+  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal")), scales="free") + theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))
 
@@
 
-### Capped at 10,000 ms
+### Capped at 0 ms and 10,000 ms
 
 ```{r fig.show='hide', warning=FALSE, echo=FALSE}
-# now with a simple cap imposed at 10000 ms, half-way through the maximum allotted time of 20000 (2 minute timeout)
-summary(d$rt_ms[d$rt_ms<10000])
+# now with a simple cap imposed to select RTs between 0 and 10000 ms
+summary(d$rt_ms[d$rt_ms >= 0 & d$rt_ms<10000])
 
-data_for_plot <- gather(d %>% filter(rt_ms<10000), key=rt_metric, value=rt_value, rt_ms, rt_log, "rt_reciprocal[-1/rt]") %>%
+data_for_plot <- d %>% 
+  filter(rt_ms >= 0 & rt_ms <= 10000) %>% 
+  gather(key=rt_metric, value=rt_value, rt_ms, rt_log, rt_reciprocal) %>%
   arrange(rt_value)
 
-ga1trimmed = ggplot(data_for_plot, aes(x=rep(1:1556,times=3), y=rt_value)) + 
+ga1trimmed = ggplot(data_for_plot, aes(x=rep(1:(nrow(data_for_plot)/3), times=3), y=rt_value)) + 
   geom_point(na.rm=TRUE) + 
-  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal[-1/rt]")), scales="free_y") +
+  facet_wrap(~factor(rt_metric, levels=c("rt_ms", "rt_log", "rt_reciprocal")), scales="free_y") +
   xlab("observations, ordered by rt_value")+ theme(aspect.ratio = 1, axis.text.x = element_text(angle = 45, hjust = 1))
@@
 ```{r selectTrim}
-dat <- d %>% filter(rt_ms<10000)
+dat <- d %>% filter(rt_ms >= 0 & rt_ms <= 10000) 
+```
+
+```{r plot rt per trial by subject}
+dat %>% 
+  gather(key=rt_metric, value=rt_value, rt_ms, rt_log, rt_reciprocal) %>%
+  filter(rt_metric=="rt_log") %>%
+  ggplot(aes(x=trial, y=rt_value, col=vag)) +
+  geom_point() +
+  geom_smooth(se=F) +
+  facet_wrap(~subject) +
+  theme(aspect.ratio = 1, panel.background = element_rect(fill="white", colour="black", linetype = "solid"))
+```
+
+```{r}
+dat %>% 
+  gather(key=rt_metric, value=rt_value, rt_ms, rt_log, rt_reciprocal) %>%
+  filter(rt_metric=="rt_log") %>%
+  ggplot(aes(group=vag, y=rt_value)) +
+  geom_boxplot(aes(x=item, col=vag))
 ```
@@
 
-```{r, lmers}
+```{r, lmer1}
+lmer1.ms  <- lmer(data=dat, rt_ms ~         vag + subit + quan + gap + side + (1|subject) + (1|item))
+lmer1.log <- lmer(data=dat, rt_log ~        vag + subit + quan + gap + side + (1|subject) + (1|item))
+lmer1.rcp <- lmer(data=dat, rt_reciprocal ~ vag + subit + quan + gap + side + (1|subject) + (1|item))
+
+models <- 
+rbind(
+  cbind(tidy(lmer1.ms),model="ms"), 
+  cbind(tidy(lmer1.log),model="log"), 
+  cbind(tidy(lmer1.rcp),model="rcp")
+)
+
+dwplot(models, vline = geom_vline(xintercept = 0)) + facet_wrap(~model, scales="free", ncol=1)
+
+```
+
+```{r mler on logrt}
+lmer2.log <- lmer(data=dat, rt_log ~ vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item))
+dwplot(lmer2.log %>% tidy() %>% filter(group=="fixed")) + 
+  ggtitle("rt_log ~ vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item)")
+
 lmer2 <- lmer(data=dat, rt_reciprocal~subit*vag + (vag|subject) + (vag|item))

### 2018-07-20 experiments/e2/writeup/2017_submitted_to_springer/e2_2017_analysis.Rmd
e2 is developing nicely

---
title: "Experiment 2 of 4"
author: "Matt Green"
output:
  html_document:
    toc: yes
    toc_float: true
    mathjax: default
    template: default
    <!--code_folding: show-->
    theme: default
---

<style>
table, th, td {
    border: 1px grey;
    border-collapse: collapse;
    padding:2px 5px 2px 5px
}  
</style>

In this experiment we did not count any response as an error, instead we coded them as expected, borderline, and extreme.

# 2017

```{r "set working directory to 2017_submitted_to_springer"}
root="/home/matt/wd/vagueness/experiments/e2/writeup/2017_submitted_to_springer"
if (getwd() != root) setwd("2017_submitted_to_springer")
```

```{r load Libraries, message=FALSE, echo=FALSE}
library(knitr); library(Matrix); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions) 
```

```{r "source my functions", message=FALSE, echo=FALSE}
source('../functions/summarySEwithin2.R')
source('../functions/e2-preprocessing.R')
```

```{r showInstructions, echo=FALSE}
v = c("crisp", "crisp", "vague", "vague")
n = c('number', 'word', 'number', 'word')
e <- c("Choose the square with 6 dots", "Choose the square with the fewest dots", "Choose a square with about 10 dots", "Choose a square with few dots")
i = data.frame(v, n, e)
names(i) = c("vagueness", "number", "example instruction")
pander(i, justify=c('center', 'left', 'left'))
```

```{r showItems, echo=FALSE}
items = c(1,2,3,4,5,6,7,8)
i = data.frame(Item = c("1","2","3","4"),
               x = c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
names(i) = c('Item', 'Numbers of dots')
pander(i, justify=c('center', 'center'))
```

```{r preprocessing, echo=FALSE, message=FALSE}
if( !file.exists('e2-data.txt') ) {
  dat <- gatherData() 
  dat <- declareImpossibleRT(dat)
  #dat <- removeImpossibleTrials(dat)
  dat <- subset(dat, select=c(Subject, Trial, Item, c_Itm, 
                              discriminability, 
                              Instruction,
                              nchar_instr,
                              Vagueness, Number, Order, Quantity, 
                              c_Vag, c_Num, c_Ord, c_Qty,
                              NaType, response_category, 
                              isBorderline, RT, RT_log)
  )
  write.table(dat, file='e2-data.txt', sep='\t', quote=FALSE, row.names = FALSE)
}
```

```{r getData, echo=TRUE, tidy=FALSE}
rawdata <- read.delim('e2-data.txt')
tempdata <- perSubjectTrim.fnc(data = rawdata[complete.cases(rawdata),], 
                               response='RT', subject='Subject', trim = 2.5)$data
rtdata <- tempdata
bldata <- tempdata
```

```{r e2-rtplot, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE}
rtdataplot1 <- summarySEwithin2(rtdata, 
                                measurevar="RT_log", 
                                withinvars=c("Vagueness", "Number", "Item"), idvar="Subject")
rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness))
rtplot=ggplot(rtdataplot1, aes(y=RT_logNormed, x=Item, group=Condition, fill=Vagueness)) +
  geom_line() +
  geom_errorbar(width=.2, aes(ymin=RT_logNormed-ci, ymax=RT_logNormed+ci)) +
  geom_point(size=2, col=1, aes(shape=Vagueness)) +
  scale_fill_grey(name="Vagueness", start=0, end=1) +
  scale_shape_manual(name="Vagueness", values=c(21,22)) +
  ggtitle("Response time") +
  ylab("Respone time (log (ms))") + 
  xlab(NULL) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.key = element_blank(), legend.position="right",
        aspect.ratio=1,
        plot.background=element_rect(fill=NA, color='white'),
        strip.background=element_blank(),
        axis.text.x = element_text(angle = 15)) +
  facet_grid(~Number)
rtplot
```

```{r e2-blBarChart, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE}
bldata$response_category <- relevel(bldata$response_category, ref='expected')
blBarChart = ggplot(bldata) +
  geom_bar(aes(response_category, group=Number:Vagueness, fill=Vagueness),
           position=position_dodge(width=NULL)) +
  scale_fill_grey() +
  xlab(NULL) +
  ggtitle('Borderline response distribution') +
  facet_grid(~Number) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.key = element_blank(), legend.position="right", 
        aspect.ratio=1,
        plot.background=element_rect(fill=NA, color='white'),
        strip.background=element_blank(),
        legend.key.size=unit(4, 'mm'),
        axis.text.x = element_text(angle = 15))
blBarChart
```

Response time models

```{r e2-rtlmerTest1, cache=TRUE, echo=TRUE}
rtlmerTest1 = lmerTest::lmer(
  data=rtdata,
  RT_log ~ c_Vag * c_Num + c_Itm +
    (1 + c_Vag * c_Num + c_Itm | Subject))
```

```{r e2-rtlmerTest1Visuals, cache=FALSE, echo=FALSE, results='asis', dependson='e2-rtlmerTest1', fig.width=6, fig.height=3}

# print a table of coefficients
print(
  xtable(
    summary(rtlmerTest1)$coefficients[,-3], 
    display=c("s","f","f","f","f"),
    digits=c(0,2,2,1,3),
    caption="Response time model"),
  caption.placement="top", 
  type='html')

# make a data frame for plotting coefficients of rtlmerTest1
dfr <- data.frame(coef = summary(rtlmerTest1)$coef[,1],
                  ci2.5 = confint(rtlmerTest1, method='Wald')[,1][!is.na(confint(rtlmerTest1, method='Wald')[,1])],
                  ci97.5 = confint(rtlmerTest1, method='Wald')[,2][!is.na(confint(rtlmerTest1, method='Wald')[,2])]
)
intercept = round(dfr[1,1],2)
dfr <- dfr[-1,] # lose the intercept for plotting
dfr <- dfr[rev(rownames(dfr)),]

# plotting code for plotting coefficients of rtlmerTest1
par(mar=c(5,7,2,1))
plot(y=1:nrow(dfr), 
     x=dfr$coef, 
     xlim=extendrange(range(dfr$ci2.5, dfr$ci97.5)), 
     ylim=extendrange(c(1,nrow(dfr))),
     type='n', axes=F, 
     xlab=substitute(paste('deviation from an intercept of ', intercept, ' log RT, with 95% CI.'), list(intercept=intercept)),
     ylab="",
     main = 'Response time model')
abline(v=0, lty=1, col='lightgrey')
arrows(y0=1:nrow(dfr), y1=1:nrow(dfr), 
       x0=dfr$ci2.5, x1=dfr$ci97.5, 
       length=0.035,
       col='black',
       lwd=2, code=3, angle=90)
points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white')
axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE)
axis(1)
box(which='plot', col='grey')
```

```{r e2-seplmerforNum, results='asis', cache=TRUE}

# make data for numeric-only model
numOnly <- subset(rtdata, subset=Number=='Numeric')

# run lmer for numeric-only model
numOnlylm <- lmerTest::lmer(data=numOnly, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))

# print coefficients for numeric-only model
print(
  xtable(
    summary(numOnlylm)$coefficients[,-3], 
    display=c("s","f","f","f","f"),
    digits=c(0,2,2,1,3),
    caption="Response time model: Numeric only"),
  caption.placement="top", 
  type='html')
```

```{r e2-seplmerforVerbal, results='asis', cache=TRUE}
# make data for verbal-only model
verbalOnly <- subset(rtdata, subset=Number=='Verbal')

# run lmer for verbal-only model
verbalOnlylm <- lmerTest::lmer(data=verbalOnly, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))

# print coefficients for verbal-only model
print(
  xtable(
    summary(verbalOnlylm)$coefficients[,-3], 
    display=c("s","f","f","f","f"),
    digits=c(0,2,2,1,3),
    caption="Response time model: Verbal only"),
  caption.placement="top", 
  type='html')
```

Borderline response models

```{r e2-bllmer, cache=TRUE, echo=TRUE}
bllmer = lme4::glmer(data=bldata,
                        isBorderline ~ c_Vag * c_Num + c_Itm +
                          (1 + c_Vag * c_Num + c_Itm | Subject),
                        family="binomial", control = glmerControl(optimizer = "bobyqa"))
```

```{r e2-bllmerVisuals, cache=FALSE, echo=FALSE, results='asis', dependson='e2-bllmer', fig.width=6, fig.height=3}
print(
  xtable(
    summary(bllmer)$coefficients,
    digits=c(0, 2, 2, 1, 3),
    caption="Error rates model"),
  caption.placement="top", 
  type='html')
dfr <- data.frame(coef = summary(bllmer)$coef[,1],
                  ci2.5 = confint(bllmer, method='Wald')[,1][!is.na(confint(bllmer, method='Wald')[,1])],
                  ci97.5 = confint(bllmer, method='Wald')[,2][!is.na(confint(bllmer, method='Wald')[,2])]
)
intercept = round(dfr[1,1],2)
dfr <- dfr[-1,] # lose the intercept for plotting
dfr <- dfr[rev(rownames(dfr)),]
par(mar=c(5,7,2,1))
plot(y=1:nrow(dfr), 
     x=dfr$coef, 
     xlim=extendrange(range(dfr$ci2.5, dfr$ci97.5)), 
     ylim=extendrange(c(1,nrow(dfr))),
     type='n', axes=F, 
     xlab=substitute(paste('deviation from an intercept of ', intercept, ' with 95% CI.'), list(intercept=intercept)),
     ylab="",
     main='Borderline response model')
abline(v=0, lty=1, col='grey')
arrows(y0=1:nrow(dfr), y1=1:nrow(dfr), 
       x0=dfr$ci2.5, x1=dfr$ci97.5, 
       length=0.035,
       col='black',
       lwd=2, code=3, angle=90)
points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white')
axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE)
axis(1)
box(which='plot', col='grey')
```

### 2018-07-20 experiments/e2/writeup/2018_tidy_version/e2_2018_analysis.Rmd
e2 is developing nicely

---
title: "Experiment 2 of 4"
author: "Matt Green"
output:
  html_document:
    toc: yes
    toc_float: true
    mathjax: default
    template: default
    <!--code_folding: show-->
    theme: default
---

In this experiment we did not count any response as an error, instead we coded them as expected, borderline, and extreme.

# 2018

```{r "set working directory to 2018_tidy_version"}
root="/home/matt/wd/vagueness/experiments/e2/writeup/2018_tidy_version"
if (getwd() != root) setwd(root)
```

```{r load Libraries, message=FALSE, echo=FALSE}
library(knitr); library(Matrix); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions); library(kableExtra)
```

```{r "source my functions", message=FALSE, echo=FALSE}
source('../functions/summarySEwithin2.R')
source('../functions/e2-preprocessing.R')
```

```{r tidy* libs, message=FALSE, echo=FALSE}
library(tidymodels)
library(tidyverse)
library(broom.mixed)
```

```{r showInstructions, echo=FALSE}
v = c("crisp", "crisp", "vague", "vague")
n = c('number', 'word', 'number', 'word')
e <- c("Choose the square with 6 dots", "Choose the square with the fewest dots", "Choose a square with about 10 dots", "Choose a square with few dots")
i = data.frame(v, n, e)
names(i) = c("vagueness", "number", "example instruction")
pander(i, justify=c('center', 'left', 'left'))
```

```{r showItems, echo=FALSE}
items = c(1,2,3,4,5,6,7,8)
i = data.frame(Item = c("1","2","3","4"),
               x = c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
names(i) = c('Item', 'Numbers of dots')
pander(i, justify=c('center', 'center'))
```

```{r preprocessing, echo=FALSE, message=FALSE}
if( !file.exists('e2-data.txt') ) {
  dat <- gatherData() 
  dat <- declareImpossibleRT(dat)
  #dat <- removeImpossibleTrials(dat)
  dat <- subset(dat, select=c(Subject, Trial, Item, c_Itm, 
                              discriminability, 
                              Instruction,
                              nchar_instr,
                              Vagueness, Number, Order, Quantity, 
                              c_Vag, c_Num, c_Ord, c_Qty,
                              NaType, response_category, 
                              isBorderline, RT, RT_log)
  )
  write.table(dat, file='e2-data.txt', sep='\t', quote=FALSE, row.names = FALSE)
}
```

```{r getData, echo=TRUE, tidy=FALSE}
rawdata <- read.delim('e2-data.txt')
tempdata <- perSubjectTrim.fnc(data = rawdata[complete.cases(rawdata),], 
                               response='RT', subject='Subject', trim = 2.5)$data
rtdata <- tempdata
bldata <- tempdata
```

```{r e2-rtplot, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE}
rtdataplot1 <- summarySEwithin2(rtdata, 
                                measurevar="RT_log", 
                                withinvars=c("Vagueness", "Number", "Item"), idvar="Subject")
rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness))
rtplot=ggplot(rtdataplot1, aes(y=RT_logNormed, x=Item, group=Condition, fill=Vagueness)) +
  geom_line() +
  geom_errorbar(width=.2, aes(ymin=RT_logNormed-ci, ymax=RT_logNormed+ci)) +
  geom_point(size=2, col=1, aes(shape=Vagueness)) +
  scale_fill_grey(name="Vagueness", start=0, end=1) +
  scale_shape_manual(name="Vagueness", values=c(21,22)) +
  ggtitle("Response time") +
  ylab("Respone time (log (ms))") + 
  xlab(NULL) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.key = element_blank(), legend.position="right",
        aspect.ratio=1,
        plot.background=element_rect(fill=NA, color='white'),
        strip.background=element_blank(),
        axis.text.x = element_text(angle = 15)) +
  facet_grid(~Number)
rtplot
```

```{r e2-blBarChart, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE}
bldata$response_category <- relevel(bldata$response_category, ref='expected')
blBarChart = ggplot(bldata) +
  geom_bar(aes(response_category, group=Number:Vagueness, fill=Vagueness),
           position=position_dodge(width=NULL)) +
  scale_fill_grey() +
  xlab(NULL) +
  ggtitle('Borderline response distribution') +
  facet_grid(~Number) +
  theme_bw() +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        legend.key = element_blank(), legend.position="right", 
        aspect.ratio=1,
        plot.background=element_rect(fill=NA, color='white'),
        strip.background=element_blank(),
        legend.key.size=unit(4, 'mm'),
        axis.text.x = element_text(angle = 15))
blBarChart
```

# Response time models

```{r e2-rtlmerTest1, cache=TRUE, echo=TRUE}
rtlmerTest1 = lmerTest::lmer(RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag * c_Num + c_Itm | Subject), rtdata)
```

## Different ways of viewing the tables of coefficients

```{r "scientific penalty"}
#options(scipen = 999)
#options("scipen"=0, "digits"=7)
```

```{r "make a pretty coefficients table for an lmerTest object"}
coef_table <- tidy(rtlmerTest1, effects="fixed")

# set names of the table
names(coef_table) <- c("effect", "term", "coefficient", "std.error", "df", "t_value", "p_value_raw")
coef_table$p_value <- as.character("")
coef_table$star <- as.character("")

# categorise p values and add significance stars
# "\u2007" is a space the same width as a numeric, that is preserved in printing
# "\u0020" is just a space
for (i in 1:nrow(coef_table)) {
  if (!is.na(coef_table$p_value_raw[i])) {
    if (coef_table$p_value_raw[i] >= 0     & coef_table$p_value[i] < 0.001) {coef_table$p_value[i] <-    "<0.001";                        coef_table$star[i] <- paste(sep="","\uFF0A","\uFF0A","\uFF0A")}
    if (coef_table$p_value_raw[i] >= 0.001 & coef_table$p_value[i] < 0.01 ) {coef_table$p_value[i] <-    paste(sep="","<0.01", "\u2007"); coef_table$star[i] <- paste(sep="","\uFF0A","\uFF0A"," ")}
    if (coef_table$p_value_raw[i] >= 0.01  & coef_table$p_value[i] < 0.05 ) {coef_table$p_value[i] <-    paste(sep="","<0.05", "\u2007"); coef_table$star[i] <- paste(sep="","\uFF0A"," "," ")}
    if (coef_table$p_value_raw[i] >= 0.05  ) {coef_table$p_value[i] <-    coef_table$p_value_raw[i]; coef_table$star[i] <-  ""}
  }
}
# un-tibble the table
coef_table <- coef_table %>% as.data.frame()

# convert numerics to formatted chars for pretty printing
coef_table$coefficient <- format(x=coef_table$coefficient, nsmall=2, digits=2)
coef_table$std.error   <- format(x=coef_table$std.error, nsmall=2, digits=2)
coef_table$df          <- format(x=coef_table$df, nsmall=2, digits=1)
coef_table$t_value     <- format(x=coef_table$t_value, nsmall=2, digits=2)
coef_table$p_value_raw <- format(x=coef_table$p_value_raw, nsmall=2, digits=3)
```

```{r}
# print raw df
print(coef_table)

# default kable
kable(coef_table)

# print kable html style with alignment 
kable(coef_table, align='llrrrrrrl')

# no fullwidth
# "\u2007" is a space the same width as a numeric, that is preserved in printing
names(coef_table)[3] <- paste(sep="","\uD835\uDEFD","\u2007")            #'\u03B2' beta
names(coef_table)[4] <- paste(sep="","\u2007","s.e.")
names(coef_table)[5] <- paste(sep="","\u2007","d.f.")                    #
names(coef_table)[6] <- paste(sep="","\uD835\uDC61","\u2007")            # t
names(coef_table)[7] <- paste(sep="","\uD835\uDC5D","\u2007","\u2007")            # p
names(coef_table)[8] <- paste(sep="","Pr(>|t|)")  # Pr(>|t|)
names(coef_table)[9] <- paste(sep="","\u2007","sig.")
coef_table %>%
  select(-effect) %>%
  filter(term!="(Intercept)") %>%
  kable(align='rrllrrrl') %>% 
  kable_styling(full_width = F, position = "left", font_size = 11)

# lmerTest's default summary 
summary(rtlmerTest1)$coefficients

# lmerTest's conversion to F-test style
anova(rtlmerTest1)
```

```{r, results='asis'}
# xtable
print(
  xtable(
    summary(rtlmerTest1)$coefficients[,-3], 
    display=c("s","f","f","f","f"),
    digits=c(0,2,2,1,3),
    caption="Response time model"),
  caption.placement="top", 
  type='html')
```

```{r e2-rtlmerTest1Visuals, cache=FALSE, echo=FALSE, results='asis', dependson='e2-rtlmerTest1', fig.width=6, fig.height=3}
dfr <- data.frame(coef = summary(rtlmerTest1)$coef[,1],
                  ci2.5 = confint(rtlmerTest1, method='Wald')[,1][!is.na(confint(rtlmerTest1, method='Wald')[,1])],
                  ci97.5 = confint(rtlmerTest1, method='Wald')[,2][!is.na(confint(rtlmerTest1, method='Wald')[,2])]
)
intercept = round(dfr[1,1],2)
dfr <- dfr[-1,] # lose the intercept for plotting
dfr <- dfr[rev(rownames(dfr)),]
par(mar=c(5,7,2,1))
plot(y=1:nrow(dfr), 
     x=dfr$coef, 
     xlim=extendrange(range(dfr$ci2.5, dfr$ci97.5)), 
     ylim=extendrange(c(1,nrow(dfr))),
     type='n', axes=F, 
     xlab=substitute(paste('deviation from an intercept of ', intercept, ' log RT, with 95% CI.'), list(intercept=intercept)),
     ylab="",
     main = 'Response time model')
abline(v=0, lty=1, col='lightgrey')
arrows(y0=1:nrow(dfr), y1=1:nrow(dfr), 
       x0=dfr$ci2.5, x1=dfr$ci97.5, 
       length=0.035,
       col='black',
       lwd=2, code=3, angle=90)
points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white')
axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE)
axis(1)
box(which='plot', col='grey')
```

<!-- ```{r e2-seplmerforNum, results='asis', cache=FALSE} -->
<!-- numOnly <- subset(rtdata, subset=Number=='Numeric') -->
<!-- numOnlylm <- lmerTest::lmer(data=numOnly, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject)) -->
<!-- print( -->
<!--   xtable( -->
<!--     summary(numOnlylm)$coefficients[,-3],  -->
<!--     display=c("s","f","f","f","f"), -->
<!--     digits=c(0,2,2,1,3), -->
<!--     caption="Response time model: Numeric only"), -->
<!--   caption.placement="top",  -->
<!--   type='html') -->
<!-- ``` -->

<!-- ```{r e2-seplmerforVerbal, results='asis', cache=FALSE} -->
<!-- verbalOnly <- subset(rtdata, subset=Number=='Verbal') -->
<!-- verbalOnlylm <- lmerTest::lmer(data=verbalOnly, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject)) -->
<!-- print( -->
<!--   xtable( -->
<!--     summary(verbalOnlylm)$coefficients[,-3],  -->
<!--     display=c("s","f","f","f","f"), -->
<!--     digits=c(0,2,2,1,3), -->
<!--     caption="Response time model: Verbal only"), -->
<!--   caption.placement="top",  -->
<!--   type='html') -->
<!-- ``` -->

<!-- # Borderline response models -->

<!-- ```{r e2-bllmer, cache=TRUE, echo=TRUE} -->
<!-- bllmer = lme4::glmer(data=bldata, -->
<!--                         isBorderline ~ c_Vag * c_Num + c_Itm + -->
<!--                           (1 + c_Vag * c_Num + c_Itm | Subject), -->
<!--                         family="binomial", control = glmerControl(optimizer = "bobyqa")) -->
<!-- ``` -->

<!-- ```{r e2-bllmerVisuals, cache=FALSE, echo=FALSE, results='asis', dependson='e2-bllmer', fig.width=6, fig.height=3} -->
<!-- print( -->
<!--   xtable( -->
<!--     summary(bllmer)$coefficients, -->
<!--     digits=c(0, 2, 2, 1, 3), -->
<!--     caption="Error rates model"), -->
<!--   caption.placement="top",  -->
<!--   type='html') -->
<!-- dfr <- data.frame(coef = summary(bllmer)$coef[,1], -->
<!--                   ci2.5 = confint(bllmer, method='Wald')[,1][!is.na(confint(bllmer, method='Wald')[,1])], -->
<!--                   ci97.5 = confint(bllmer, method='Wald')[,2][!is.na(confint(bllmer, method='Wald')[,2])] -->
<!-- ) -->
<!-- intercept = round(dfr[1,1],2) -->
<!-- dfr <- dfr[-1,] # lose the intercept for plotting -->
<!-- dfr <- dfr[rev(rownames(dfr)),] -->
<!-- par(mar=c(5,7,2,1)) -->
<!-- plot(y=1:nrow(dfr),  -->
<!--      x=dfr$coef,  -->
<!--      xlim=extendrange(range(dfr$ci2.5, dfr$ci97.5)),  -->
<!--      ylim=extendrange(c(1,nrow(dfr))), -->
<!--      type='n', axes=F,  -->
<!--      xlab=substitute(paste('deviation from an intercept of ', intercept, ' with 95% CI.'), list(intercept=intercept)), -->
<!--      ylab="", -->
<!--      main='Borderline response model') -->
<!-- abline(v=0, lty=1, col='grey') -->
<!-- arrows(y0=1:nrow(dfr), y1=1:nrow(dfr),  -->
<!--        x0=dfr$ci2.5, x1=dfr$ci97.5,  -->
<!--        length=0.035, -->
<!--        col='black', -->
<!--        lwd=2, code=3, angle=90) -->
<!-- points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white') -->
<!-- axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE) -->
<!-- axis(1) -->
<!-- box(which='plot', col='grey') -->
<!-- ``` -->

### 2018-07-26 experiments/e2/writeup/2018_tidy_version/e2_2018_analysis.Rmd
changed over to tidy all the way

@@
 
-In this experiment we did not count any response as an error, instead we coded them as expected, borderline, and extreme.
-
-# 2018
-
-```{r "set working directory to 2018_tidy_version"}
+```{r "set working directory to 2018_tidy_version", echo=FALSE}
 root="/home/matt/wd/vagueness/experiments/e2/writeup/2018_tidy_version"
@@
 
-```{r load Libraries, message=FALSE, echo=FALSE}
-library(knitr); library(Matrix); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions); library(kableExtra)
-```
-
-```{r "source my functions", message=FALSE, echo=FALSE}
-source('../functions/summarySEwithin2.R')
-source('../functions/e2-preprocessing.R')
-```
-
-```{r tidy* libs, message=FALSE, echo=FALSE}
+```{r "load Libraries and source my functions", message=FALSE, echo=FALSE}
+library(knitr)
+#library(Matrix)
+library(ggplot2)
+library(xtable)
+library(gridExtra)
+library(lme4)
+library(lmerTest)
+library(data.table)
+library(dplyr)
+library(dtplyr)
+#library(pander)
+library(LMERConvenienceFunctions)
+library(kableExtra)
+library(dotwhisker)
 library(tidymodels)
@@
 library(broom.mixed)
+source('../functions/summarySEwithin2.R')
+source('../functions/e2-preprocessing.R')
+source('../functions/pretty_coef_table.R')
+source('../functions/pretty_coef_plot.R')
 ```
 
+In this experiment we did not count any response as an error, instead we coded them as expected, borderline, and extreme.
+
+These are the example instructions:
 ```{r showInstructions, echo=FALSE}
-v = c("crisp", "crisp", "vague", "vague")
-n = c('number', 'word', 'number', 'word')
-e <- c("Choose the square with 6 dots", "Choose the square with the fewest dots", "Choose a square with about 10 dots", "Choose a square with few dots")
-i = data.frame(v, n, e)
-names(i) = c("vagueness", "number", "example instruction")
-pander(i, justify=c('center', 'left', 'left'))
+data.frame(
+  vagueness = c("crisp", "crisp", "vague", "vague"),
+  number = c('number', 'word', 'number', 'word'),
+  example_instruction = c("Choose the square with 6 dots", "Choose the square with the fewest dots", "Choose a square with about 10 dots", "Choose a square with few dots")  
+) %>% kable() %>% kable_styling(full_width = F, position = "left", font_size = 11)
 ```
 
+These are the items:
 ```{r showItems, echo=FALSE}
-items = c(1,2,3,4,5,6,7,8)
-i = data.frame(Item = c("1","2","3","4"),
-               x = c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
-names(i) = c('Item', 'Numbers of dots')
-pander(i, justify=c('center', 'center'))
+data.frame(
+  Item = c("1","2","3","4"),
+  'Numbers of dots' = c("06:15:24", "16:25:34", "26:35:44", "36:45:54")
+) %>% kable() %>% kable_styling(full_width = F, position = "left", font_size = 11)
 ```
 
+Pre-process the raw data.
 ```{r preprocessing, echo=FALSE, message=FALSE}
@@
   #dat <- removeImpossibleTrials(dat)
-  dat <- subset(dat, select=c(Subject, Trial, Item, c_Itm, 
-                              discriminability, 
-                              Instruction,
-                              nchar_instr,
-                              Vagueness, Number, Order, Quantity, 
-                              c_Vag, c_Num, c_Ord, c_Qty,
-                              NaType, response_category, 
-                              isBorderline, RT, RT_log)
-  )
+  dat <- subset(dat, select=c(Subject, Trial, Item, c_Itm, discriminability, Instruction, nchar_instr, Vagueness, Number, Order, Quantity, c_Vag, c_Num, c_Ord, c_Qty, NaType, response_category, isBorderline, RT, RT_log))
   write.table(dat, file='e2-data.txt', sep='\t', quote=FALSE, row.names = FALSE)
@@
 
-```{r getData, echo=TRUE, tidy=FALSE}
-rawdata <- read.delim('e2-data.txt')
-tempdata <- perSubjectTrim.fnc(data = rawdata[complete.cases(rawdata),], 
-                               response='RT', subject='Subject', trim = 2.5)$data
-rtdata <- tempdata
-bldata <- tempdata
+```{r}
+raw_data <- read.delim('e2-data.txt')
+```
+
+```{r}
+trim_data <- perSubjectTrim.fnc(data = raw_data[complete.cases(raw_data),], response='RT', subject='Subject', trim = 2.5)$data
+```
+
+# Response time data
+
+```{r}
+rtdata <- trim_data
 ```
@@
 ```{r e2-rtplot, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE}
-rtdataplot1 <- summarySEwithin2(rtdata, 
-                                measurevar="RT_log", 
-                                withinvars=c("Vagueness", "Number", "Item"), idvar="Subject")
+rtdataplot1 <- summarySEwithin2(rtdata, measurevar="RT_log", withinvars=c("Vagueness", "Number", "Item"), idvar="Subject")
+
 rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness))
-rtplot=ggplot(rtdataplot1, aes(y=RT_logNormed, x=Item, group=Condition, fill=Vagueness)) +
-  geom_line() +
-  geom_errorbar(width=.2, aes(ymin=RT_logNormed-ci, ymax=RT_logNormed+ci)) +
-  geom_point(size=2, col=1, aes(shape=Vagueness)) +
-  scale_fill_grey(name="Vagueness", start=0, end=1) +
-  scale_shape_manual(name="Vagueness", values=c(21,22)) +
-  ggtitle("Response time") +
-  ylab("Respone time (log (ms))") + 
-  xlab(NULL) +
-  theme_bw() +
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        legend.key = element_blank(), legend.position="right",
-        aspect.ratio=1,
-        plot.background=element_rect(fill=NA, color='white'),
-        strip.background=element_blank(),
-        axis.text.x = element_text(angle = 15)) +
-  facet_grid(~Number)
+
+rtplot=ggplot(rtdataplot1, aes(y=RT_logNormed, x=Item, group=Condition, fill=Vagueness)) + geom_line() + geom_errorbar(width=.2, aes(ymin=RT_logNormed-ci, ymax=RT_logNormed+ci)) + geom_point(size=2, col=1, aes(shape=Vagueness)) + scale_fill_grey(name="Vagueness", start=0, end=1) +  scale_shape_manual(name="Vagueness", values=c(21,22)) + ggtitle("Response time") + ylab("Respone time (log (ms))") + xlab(NULL) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), legend.position="right", aspect.ratio=1, plot.background=element_rect(fill=NA, color='white'), strip.background=element_blank(), axis.text.x = element_text(angle = 15)) + facet_grid(~Number)
+
 rtplot
+```
+
+## Response time models
+
+### Full RT model
+
+```{r 'rt-full-model', cache=TRUE, echo=TRUE}
+rtFullModel <- lmerTest::lmer(RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag * c_Num + c_Itm | Subject), rtdata)
+```
+
+```{r 'make a pretty coefficients table for rtFullModel', echo=FALSE, results='asis'}
+pretty_coef_table(rtFullModel, "rtFullModel") %>% print()
+```
+
+```{r, 'make a pretty coefficients plot for rtFullModel', echo=FALSE, fig.width=6, fig.height=3}
+pretty_coef_plot(rtFullModel, "rtFullModel")
+```
+
+### Numeric only RT model
+
+```{r 'rt-numeric-only model', cache=TRUE, echo=TRUE}
+rtNumOnlyModel <- lmerTest::lmer(RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject), subset(rtdata,Number=="Numeric"))
+```
+
+```{r 'make a pretty coefficients table for rtNumOnlyModel', echo=FALSE, results='asis'}
+pretty_coef_table(rtNumOnlyModel, "rtNumOnlyModel") %>% print()
+```
+
+```{r, 'make a pretty coefficients plot for rtNumOnlyModel', echo=FALSE, fig.width=6, fig.height=3}
+pretty_coef_plot(rtNumOnlyModel, "rtNumOnlyModel")
+```
+
+### Verbal only RT model
+
+```{r 'rt-verbal-only model', cache=TRUE, echo=TRUE}
+rtVerbalOnlyModel <- lmerTest::lmer(RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject), subset(rtdata,Number=="Verbal"))
+```
+
+```{r 'make a pretty coefficients table for rtVerbalOnlyModel', echo=FALSE, results='asis'}
+pretty_coef_table(rtVerbalOnlyModel, "rtVerbalOnlyModel") %>% print()
+```
+
+```{r, 'make a pretty coefficients plot for rtVerbalOnlyModel', echo=FALSE, fig.width=6, fig.height=3}
+pretty_coef_plot(rtVerbalOnlyModel, "rtVerbalOnlyModel")
+```
+
+# Borderline response data
+
+```{r}
+bldata <- trim_data
 ```
@@
 bldata$response_category <- relevel(bldata$response_category, ref='expected')
-blBarChart = ggplot(bldata) +
-  geom_bar(aes(response_category, group=Number:Vagueness, fill=Vagueness),
-           position=position_dodge(width=NULL)) +
-  scale_fill_grey() +
-  xlab(NULL) +
-  ggtitle('Borderline response distribution') +
-  facet_grid(~Number) +
-  theme_bw() +
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
-        legend.key = element_blank(), legend.position="right", 
-        aspect.ratio=1,
-        plot.background=element_rect(fill=NA, color='white'),
-        strip.background=element_blank(),
-        legend.key.size=unit(4, 'mm'),
-        axis.text.x = element_text(angle = 15))
+
+blBarChart = ggplot(bldata) + geom_bar(aes(response_category, group=Number:Vagueness, fill=Vagueness), possition=position_dodge(width=NULL)) +   scale_fill_grey() + xlab(NULL) + ggtitle('Borderline response distribution') + facet_grid(~Number) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), legend.position="right", aspect.ratio=1, plot.background = element_rect(fill=NA, color='white'), strip.background=element_blank(), legend.key.size=unit(4, 'mm'), axis.text.x = element_text(angle = 15))
+
 blBarChart
@@
 
-# Response time models
+## Borderline response models
 
-```{r e2-rtlmerTest1, cache=TRUE, echo=TRUE}
-rtlmerTest1 = lmerTest::lmer(RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag * c_Num + c_Itm | Subject), rtdata)
+### Full borderline response model
+
+```{r 'bl-full-model', cache=TRUE, echo=TRUE}
+blFullModel <- lme4::glmer(isBorderline ~ c_Vag * c_Num + c_Itm + (1 + c_Vag * c_Num + c_Itm | Subject), bldata, family="binomial", control = glmerControl(optimizer = "bobyqa"))
 ```
 
-## Different ways of viewing the tables of coefficients
-
-```{r "scientific penalty"}
-#options(scipen = 999)
-#options("scipen"=0, "digits"=7)
+```{r 'make a pretty coefficients table for blFullModel', echo=FALSE, results='asis'}
+pretty_coef_table(blFullModel, "blFullModel") %>% print()
 ```
 
-```{r "make a pretty coefficients table for an lmerTest object"}
-coef_table <- tidy(rtlmerTest1, effects="fixed")
-
-# set names of the table
-names(coef_table) <- c("effect", "term", "coefficient", "std.error", "df", "t_value", "p_value_raw")
-coef_table$p_value <- as.character("")
-coef_table$star <- as.character("")
-
-# categorise p values and add significance stars
-# "\u2007" is a space the same width as a numeric, that is preserved in printing
-# "\u0020" is just a space
-for (i in 1:nrow(coef_table)) {
-  if (!is.na(coef_table$p_value_raw[i])) {
-    if (coef_table$p_value_raw[i] >= 0     & coef_table$p_value[i] < 0.001) {coef_table$p_value[i] <-    "<0.001";                        coef_table$star[i] <- paste(sep="","\uFF0A","\uFF0A","\uFF0A")}
-    if (coef_table$p_value_raw[i] >= 0.001 & coef_table$p_value[i] < 0.01 ) {coef_table$p_value[i] <-    paste(sep="","<0.01", "\u2007"); coef_table$star[i] <- paste(sep="","\uFF0A","\uFF0A"," ")}
-    if (coef_table$p_value_raw[i] >= 0.01  & coef_table$p_value[i] < 0.05 ) {coef_table$p_value[i] <-    paste(sep="","<0.05", "\u2007"); coef_table$star[i] <- paste(sep="","\uFF0A"," "," ")}
-    if (coef_table$p_value_raw[i] >= 0.05  ) {coef_table$p_value[i] <-    coef_table$p_value_raw[i]; coef_table$star[i] <-  ""}
-  }
-}
-# un-tibble the table
-coef_table <- coef_table %>% as.data.frame()
-
-# convert numerics to formatted chars for pretty printing
-coef_table$coefficient <- format(x=coef_table$coefficient, nsmall=2, digits=2)
-coef_table$std.error   <- format(x=coef_table$std.error, nsmall=2, digits=2)
-coef_table$df          <- format(x=coef_table$df, nsmall=2, digits=1)
-coef_table$t_value     <- format(x=coef_table$t_value, nsmall=2, digits=2)
-coef_table$p_value_raw <- format(x=coef_table$p_value_raw, nsmall=2, digits=3)
+```{r, 'make a pretty coefficients plot for blFullModel', echo=FALSE, fig.width=6, fig.height=3}
+pretty_coef_plot(blFullModel, "blFullModel")
 ```
-
-```{r}
-# print raw df
-print(coef_table)
-
-# default kable
-kable(coef_table)
-
-# print kable html style with alignment 
-kable(coef_table, align='llrrrrrrl')
-
-# no fullwidth
-# "\u2007" is a space the same width as a numeric, that is preserved in printing
-names(coef_table)[3] <- paste(sep="","\uD835\uDEFD","\u2007")            #'\u03B2' beta
-names(coef_table)[4] <- paste(sep="","\u2007","s.e.")
-names(coef_table)[5] <- paste(sep="","\u2007","d.f.")                    #
-names(coef_table)[6] <- paste(sep="","\uD835\uDC61","\u2007")            # t
-names(coef_table)[7] <- paste(sep="","\uD835\uDC5D","\u2007","\u2007")            # p
-names(coef_table)[8] <- paste(sep="","Pr(>|t|)")  # Pr(>|t|)
-names(coef_table)[9] <- paste(sep="","\u2007","sig.")
-coef_table %>%
-  select(-effect) %>%
-  filter(term!="(Intercept)") %>%
-  kable(align='rrllrrrl') %>% 
-  kable_styling(full_width = F, position = "left", font_size = 11)
-
-# lmerTest's default summary 
-summary(rtlmerTest1)$coefficients
-
-# lmerTest's conversion to F-test style
-anova(rtlmerTest1)
-```
-
-```{r, results='asis'}
-# xtable
-print(
-  xtable(
-    summary(rtlmerTest1)$coefficients[,-3], 
-    display=c("s","f","f","f","f"),
-    digits=c(0,2,2,1,3),
-    caption="Response time model"),
-  caption.placement="top", 
-  type='html')
-```
-
-```{r e2-rtlmerTest1Visuals, cache=FALSE, echo=FALSE, results='asis', dependson='e2-rtlmerTest1', fig.width=6, fig.height=3}
-dfr <- data.frame(coef = summary(rtlmerTest1)$coef[,1],
-                  ci2.5 = confint(rtlmerTest1, method='Wald')[,1][!is.na(confint(rtlmerTest1, method='Wald')[,1])],
-                  ci97.5 = confint(rtlmerTest1, method='Wald')[,2][!is.na(confint(rtlmerTest1, method='Wald')[,2])]
-)
-intercept = round(dfr[1,1],2)
-dfr <- dfr[-1,] # lose the intercept for plotting
-dfr <- dfr[rev(rownames(dfr)),]
-par(mar=c(5,7,2,1))
-plot(y=1:nrow(dfr), 
-     x=dfr$coef, 
-     xlim=extendrange(range(dfr$ci2.5, dfr$ci97.5)), 
-     ylim=extendrange(c(1,nrow(dfr))),
-     type='n', axes=F, 
-     xlab=substitute(paste('deviation from an intercept of ', intercept, ' log RT, with 95% CI.'), list(intercept=intercept)),
-     ylab="",
-     main = 'Response time model')
-abline(v=0, lty=1, col='lightgrey')
-arrows(y0=1:nrow(dfr), y1=1:nrow(dfr), 
-       x0=dfr$ci2.5, x1=dfr$ci97.5, 
-       length=0.035,
-       col='black',
-       lwd=2, code=3, angle=90)
-points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white')
-axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE)
-axis(1)
-box(which='plot', col='grey')
-```
-
-<!-- ```{r e2-seplmerforNum, results='asis', cache=FALSE} -->
-<!-- numOnly <- subset(rtdata, subset=Number=='Numeric') -->
-<!-- numOnlylm <- lmerTest::lmer(data=numOnly, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject)) -->
-<!-- print( -->
-<!--   xtable( -->
-<!--     summary(numOnlylm)$coefficients[,-3],  -->
-<!--     display=c("s","f","f","f","f"), -->
-<!--     digits=c(0,2,2,1,3), -->
-<!--     caption="Response time model: Numeric only"), -->
-<!--   caption.placement="top",  -->
-<!--   type='html') -->
-<!-- ``` -->
-
-<!-- ```{r e2-seplmerforVerbal, results='asis', cache=FALSE} -->
-<!-- verbalOnly <- subset(rtdata, subset=Number=='Verbal') -->
-<!-- verbalOnlylm <- lmerTest::lmer(data=verbalOnly, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject)) -->
-<!-- print( -->
-<!--   xtable( -->
-<!--     summary(verbalOnlylm)$coefficients[,-3],  -->
-<!--     display=c("s","f","f","f","f"), -->
-<!--     digits=c(0,2,2,1,3), -->
-<!--     caption="Response time model: Verbal only"), -->
-<!--   caption.placement="top",  -->
-<!--   type='html') -->
-<!-- ``` -->
-
-<!-- # Borderline response models -->
-
-<!-- ```{r e2-bllmer, cache=TRUE, echo=TRUE} -->
-<!-- bllmer = lme4::glmer(data=bldata, -->
-<!--                         isBorderline ~ c_Vag * c_Num + c_Itm + -->
-<!--                           (1 + c_Vag * c_Num + c_Itm | Subject), -->
-<!--                         family="binomial", control = glmerControl(optimizer = "bobyqa")) -->
-<!-- ``` -->
-
-<!-- ```{r e2-bllmerVisuals, cache=FALSE, echo=FALSE, results='asis', dependson='e2-bllmer', fig.width=6, fig.height=3} -->
-<!-- print( -->
-<!--   xtable( -->
-<!--     summary(bllmer)$coefficients, -->
-<!--     digits=c(0, 2, 2, 1, 3), -->
-<!--     caption="Error rates model"), -->
-<!--   caption.placement="top",  -->
-<!--   type='html') -->
-<!-- dfr <- data.frame(coef = summary(bllmer)$coef[,1], -->
-<!--                   ci2.5 = confint(bllmer, method='Wald')[,1][!is.na(confint(bllmer, method='Wald')[,1])], -->
-<!--                   ci97.5 = confint(bllmer, method='Wald')[,2][!is.na(confint(bllmer, method='Wald')[,2])] -->
-<!-- ) -->
-<!-- intercept = round(dfr[1,1],2) -->
-<!-- dfr <- dfr[-1,] # lose the intercept for plotting -->
-<!-- dfr <- dfr[rev(rownames(dfr)),] -->
-<!-- par(mar=c(5,7,2,1)) -->
-<!-- plot(y=1:nrow(dfr),  -->
-<!--      x=dfr$coef,  -->
-<!--      xlim=extendrange(range(dfr$ci2.5, dfr$ci97.5)),  -->
-<!--      ylim=extendrange(c(1,nrow(dfr))), -->
-<!--      type='n', axes=F,  -->
-<!--      xlab=substitute(paste('deviation from an intercept of ', intercept, ' with 95% CI.'), list(intercept=intercept)), -->
-<!--      ylab="", -->
-<!--      main='Borderline response model') -->
-<!-- abline(v=0, lty=1, col='grey') -->
-<!-- arrows(y0=1:nrow(dfr), y1=1:nrow(dfr),  -->
-<!--        x0=dfr$ci2.5, x1=dfr$ci97.5,  -->
-<!--        length=0.035, -->
-<!--        col='black', -->
-<!--        lwd=2, code=3, angle=90) -->
-<!-- points(y=1:nrow(dfr), x=dfr$coef, pch=21, bg='white') -->
-<!-- axis(2, labels=row.names(dfr), at=1:nrow(dfr), las=1, tick=FALSE) -->
-<!-- axis(1) -->
-<!-- box(which='plot', col='grey') -->
-<!-- ``` -->

### 2018-07-27 experiments/e2/writeup/2018_tidy_version/e2_2018_analysis.Rmd
typos

@@
 library(knitr)
-#library(Matrix)
-library(ggplot2)
-library(xtable)
-library(gridExtra)
+library(tidyverse)
+library(tidymodels)
+library(broom.mixed)
 library(lme4)
 library(lmerTest)
+library(dotwhisker)
+library(kableExtra)
+library(LMERConvenienceFunctions)
 library(data.table)
-library(dplyr)
-library(dtplyr)
-#library(pander)
-library(LMERConvenienceFunctions)
-library(kableExtra)
-library(dotwhisker)
-library(tidymodels)
-library(tidyverse)
-library(broom.mixed)
 source('../functions/summarySEwithin2.R')
@@
 
-blBarChart = ggplot(bldata) + geom_bar(aes(response_category, group=Number:Vagueness, fill=Vagueness), possition=position_dodge(width=NULL)) +   scale_fill_grey() + xlab(NULL) + ggtitle('Borderline response distribution') + facet_grid(~Number) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), legend.position="right", aspect.ratio=1, plot.background = element_rect(fill=NA, color='white'), strip.background=element_blank(), legend.key.size=unit(4, 'mm'), axis.text.x = element_text(angle = 15))
+blBarChart = ggplot(bldata) + geom_bar(aes(response_category, group=Number:Vagueness, fill=Vagueness), position=position_dodge(width=NULL)) +   scale_fill_grey() + xlab(NULL) + ggtitle('Borderline response distribution') + facet_grid(~Number) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), legend.position="right", aspect.ratio=1, plot.background = element_rect(fill=NA, color='white'), strip.background=element_blank(), legend.key.size=unit(4, 'mm'), axis.text.x = element_text(angle = 15))
 

### 2018-07-27 experiments/e2/writeup/2018_levy-method/2018_levy-method-e1Springer.Rmd
levy start

--- 2018-07-27 experiments/e2/writeup/2018_tidy_version/e2_2018_analysis.Rmd
@@
 ---
-title: "Experiment 2 of 4"
+title: "This is experiment one in the springer 1 to 3 numbering (originaly called e2)"
 author: "Matt Green"
@@
 
-```{r "set working directory to 2018_tidy_version", echo=FALSE}
-root="/home/matt/wd/vagueness/experiments/e2/writeup/2018_tidy_version"
+```{r 'set working directory', echo=FALSE}
+root="/home/matt/wd/vagueness/experiments/e2/writeup/2018_levy-method"
 if (getwd() != root) setwd(root)
@@
 
-```{r "load Libraries and source my functions", message=FALSE, echo=FALSE}
+```{r 'load Libraries and source my functions', message=FALSE, echo=FALSE}
 library(knitr)
@@
 library(data.table)
-source('../functions/summarySEwithin2.R')
-source('../functions/e2-preprocessing.R')
-source('../functions/pretty_coef_table.R')
-source('../functions/pretty_coef_plot.R')
+source('summarySEwithin2.R')
+source('concatenate_raw_data.R')
+source('declare_impossible_rt.R')
+source('remove_impossible_trials.R')
+source('pretty_coef_table.R')
+source('pretty_coef_plot.R')
 ```
@@
 These are the example instructions:
-```{r showInstructions, echo=FALSE}
+```{r 'showInstructions', echo=FALSE}
 data.frame(
@@
 These are the items:
-```{r showItems, echo=FALSE}
+```{r 'showItems', echo=FALSE}
 data.frame(
@@
 
-Pre-process the raw data.
-```{r preprocessing, echo=FALSE, message=FALSE}
-if( !file.exists('e2-data.txt') ) {
-  dat <- gatherData() 
-  dat <- declareImpossibleRT(dat)
-  #dat <- removeImpossibleTrials(dat)
-  dat <- subset(dat, select=c(Subject, Trial, Item, c_Itm, discriminability, Instruction, nchar_instr, Vagueness, Number, Order, Quantity, c_Vag, c_Num, c_Ord, c_Qty, NaType, response_category, isBorderline, RT, RT_log))
-  write.table(dat, file='e2-data.txt', sep='\t', quote=FALSE, row.names = FALSE)
-}
+Concatenate the raw data.
+```{r, 'Concatenate the raw data'}
+dat <- concatenate_raw_data()
+dat %>% select(RT, RESPONSE, response_category) %>% summary() 
 ```
 
-```{r}
+Catch impossible RTs: sticky fingers (RT<=1) and near timeouts (RT>59,998ms)
+```{r, 'declare_impossible_rt'}
+dat <- declare_impossible_rt(dat)
+dat %>% select(RT, RESPONSE, response_category, NaType) %>% summary()
+# show the NA trials
+dat %>% filter(is.na(RT)) %>% select(RT_original, RESPONSE, NaType, RT, response_category) 
+```
+
+```{r, 'write the raw data to file'}
+write.table(dat %>% select(Subject, Trial, Item, c_Itm, discriminability, Instruction, nchar_instr, Vagueness, Number, Order, Quantity, c_Vag, c_Num, c_Ord, c_Qty, NaType, response_category, isBorderline, RT, RT_log), file='e2-data.txt', sep='\t', quote=FALSE, row.names = FALSE)
+```
+
+Get the raw data.
+```{r, 'get the raw data from file'}
 raw_data <- read.delim('e2-data.txt')
@@
 
-```{r}
+Trim the data.
+```{r, 'trim the data'}
 trim_data <- perSubjectTrim.fnc(data = raw_data[complete.cases(raw_data),], response='RT', subject='Subject', trim = 2.5)$data
@@
 
-```{r}
+```{r, 'select the response time data'}
 rtdata <- trim_data
+```
+
+```{r, 'plot the distribution of the trimmed data', fig.width=3, fig.height=3}
+ggplot(data=rtdata, aes(RT_log)) + geom_density() + theme(aspect.ratio = 1)
 ```

### 2018-07-27 experiments/e2/writeup/2018_levy-method/concatenate_raw_data.R

@@ -1,0 +1,192 @@
+concatenate_raw_data = function() {
+  number_of_valid_subjects=30
+  number_of_trials_per_subject=256
+  message('Starting to gather data')
+  data_dir <- '../../experimentCode/output/'
+  column.headers.df <-  head( read.table(
+    paste(data_dir,'subject01.data',sep=''),
+    header=TRUE),0)
+  gathered.data <- column.headers.df
+  for (subject in 1:number_of_valid_subjects) {
+    current.filename <- paste(data_dir, 'subject', sprintf("%02d",subject), '.data', sep='')
+    current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = TRUE)
+    gathered.data <- rbind(gathered.data, current.data)
+  }
+  dat <- gathered.data
+  dat$Order <- factor(dat$Order, levels=c(1,2), labels=c('LtoR','RtoL'))
+  dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)
+  dat$Quantity <- factor(dat$Quantity, levels=c(1,2), labels=c('Small','Large'))
+  dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)
+  dat[ dat$Condition==1 , 'Vagueness'] <- 'Vague'
+  dat[ dat$Condition==2 , 'Vagueness'] <- 'Crisp'
+  dat[ dat$Condition==3 , 'Vagueness'] <- 'Vague'
+  dat[ dat$Condition==4 , 'Vagueness'] <- 'Crisp'
+  dat$Vagueness <- as.factor(dat$Vagueness)
+  dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)
+  dat[ dat$Condition==1 , 'Number'] <- 'Numeric'
+  dat[ dat$Condition==2 , 'Number'] <- 'Numeric'
+  dat[ dat$Condition==3 , 'Number'] <- 'Verbal'
+  dat[ dat$Condition==4 , 'Number'] <- 'Verbal'
+  dat$Number <- as.factor(dat$Number)
+  dat$c_Num <- ifelse(dat$Number=='Numeric', -.5, .5)
[…]

### 2018-07-27 experiments/e2/writeup/2018_levy-method/remove_impossible_trials.R

@@ -1,0 +1,8 @@
+remove_impossible_trials = function(dat){
+  message('Starting to remove impossible trials')
+  a=nrow(dat)
+  dat <- dat[ complete.cases(dat), ]
+  b=nrow(dat)
+  message(paste('removed', a-b, 'impossible trials'))
+  return(dat)
+}

### 2018-08-02 experiments/B_pilot_2/writeup/e1.Rmd
massive restructuring Aug 2018

--- 2016-09-25 experiments/e1/writeup/e1.Rmd
@@
     padding:2px 5px 2px 5px
+}  
+</style>
+
+<!-- no wrap table entries on their white space -->
+
+<style>
+table {
+  white-space: nowrap;
 }
@@
 ```{r loadLibraries, message=FALSE, echo=FALSE, cache=FALSE}
-library(knitr); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions); source('summarySEwithin2.R')
+library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions); source('summarySEwithin2.R')
+```
+
+```{r tidyr, echo=FALSE, message=FALSE}
+library(tidyr)
+```
+
+```{r rawdata, cache=FALSE}
+rawdata <- read.table('data.txt', sep=' ')
+rawdata %>% head() %>% kable()
 ```
@@
 ```{r separateDataFramesForRTandErrorAnalyses, echo=TRUE, cache=FALSE}
-rawdata <- read.table('data.txt', sep=' ')
 rtdata <- subset(rawdata, error==FALSE)
@@
 ```
+
+There were `r nrow(rawdata)` trials; and `r  nrow(subset(rawdata,error==TRUE))` error trials, representing 
+`r round(nrow(subset(rawdata,error==TRUE))/nrow(rawdata) * 100, 1)`% of the raw data.
 
@@
 
-```{r computeSummariesForGraphs, cache=TRUE}
+```{r computeSummariesForGraphs, cache=FALSE}
 rtdataplot1 = summarySEwithin2(rtdata,  measurevar="logrt",  withinvars=c("Vgn","Gap"), idvar="subject")
@@
 
-```{r rtplot, echo = FALSE}
+```{r rtplot, echo=FALSE}
 dodge=.06

### 2018-08-02 experiments/C_exp_1/archive/data_wrangling.Rmd
massive restructuring Aug 2018

--- 2016-07-28 experiments/e3/writeup/experiment3.Rmd
@@
----
-title: "Experiment 3 of 4"
-output:
-  html_document:
-    toc: true
-    <!--toc_float: true-->
-    mathjax: default
-    template: default
-    code_folding: hide
-    theme: default
----  
-# Experiment 3
+Get the data if file dat.Rda does not already exist.
 
-```{r docOpts}
-library(knitr)
-```
+```{r, get_the_data}
+# if the output file dat.Rda already exists then load it, else do data wrangling
+if( file.exists('dat.Rda') ) load('dat.Rda') else {
 
-```{r loadLibraries}
-suppressPackageStartupMessages(library(MASS))
-suppressPackageStartupMessages(library(ggplot2))
-suppressPackageStartupMessages(library(lme4))
-suppressPackageStartupMessages(library(xtable))
-suppressPackageStartupMessages(library(plyr))
-suppressPackageStartupMessages(library(reshape))
-suppressPackageStartupMessages(library(lattice))
-suppressPackageStartupMessages(library(LMERConvenienceFunctions))
-suppressPackageStartupMessages(library(car))
-suppressPackageStartupMessages(library(grid))
-suppressPackageStartupMessages(library(lmerTest))
-suppressPackageStartupMessages(library(reshape2))
-suppressPackageStartupMessages(library(ggplot2))
-suppressPackageStartupMessages(require(languageR))
-```
-
-# Experiment Structure
-Each subject saw 192 trials, being 3 cycles of 64 trials. There were 38 participants. There were 7296 rows in the full data set.
-
-# Trial Procedure
-The procedure for each trial was as follows.
-First an instruction that was a referring expression was presented until participants dismissed it with a key press.
-Then the squares with dots were presented until the participant responded with a key press, with a 60 second timeout if they did not respond.
-
-# Prepare the data
-Source additional R files with functions for computing confidence intervals.
-
-```{r sourceFunctions}
-source("functions/CIfunctions.R")
-```
-
-Get data.
-
-```{r readdata}
-dat <- read.delim("data/allresults.txt", stringsAsFactors = FALSE)
+  # source required data wrangling functions
+  source('functions/classifyResponse.R')
+  source('functions/gatherData.R')
 
   # declare local variables
-number_of_valid_subjects <- 38
-number_of_rows <- 7296
-number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # 192
+  number_of_valid_subjects <- 30
+  number_of_rows <- 7680 # 7680
+  number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # 256
 
-# sort out borderline responses into expected near and far
-# How many squares in the square they chose? 
-dat$RESPONSE <- as.character(dat$RESPONSE)
-for (row in 1 : nrow(dat) ) {
-  switch(dat[row,'RESPONSE'],
-         'LEFT' = {dat[row, 'choice'] <- dat[row, 'Left']},
-         'MIDDLE' = {dat[row, 'choice'] <- dat[row, 'Mid']},
-         'RIGHT' = {dat[row, 'choice'] <- dat[row, 'Right']}         
-  )
-}
-dat$crossed=paste('Con',dat$Condition,':Quan',dat$Quantity,':Item',dat$Item,sep="")
+    ## call user-defined functions
 
-dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseExpected'] <-  6 
-dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseNear']     <- 15  
-dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseFar']      <- 24
+  # collect all individual subject files into one data frame
+  dat <- gatherData(number_of_valid_subjects)
 
-dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseExpected'] <- 16
-dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseFar']      <- 34
-
-dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseExpected'] <- 26
-dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseFar']      <- 44
-
-dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseExpected'] <- 36
-dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseFar']      <- 54
-
-dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseExpected'] <- 24
-dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseNear']     <- 15 
-dat[dat$crossed=='Con1:Quan2:Item1', 'ResponseFar']      <-  6
-
-dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseExpected'] <- 34
-dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con1:Quan2:Item2', 'ResponseFar']      <- 16
-
-dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseExpected'] <- 44
-dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con1:Quan2:Item3', 'ResponseFar']      <- 26
-
-dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseExpected'] <- 54
-dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con1:Quan2:Item4', 'ResponseFar']      <- 36
-
-dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseExpected'] <-  6 
-dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseNear']     <- 15  
-dat[dat$crossed=='Con2:Quan1:Item1', 'ResponseFar']      <- 24
-
-dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseExpected'] <- 16
-dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con2:Quan1:Item2', 'ResponseFar']      <- 34
-
-dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseExpected'] <- 26
-dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con2:Quan1:Item3', 'ResponseFar']      <- 44
-
-dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseExpected'] <- 36
-dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con2:Quan1:Item4', 'ResponseFar']      <- 54
-
-dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseExpected'] <- 24
-dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseNear']     <- 15 
-dat[dat$crossed=='Con2:Quan2:Item1', 'ResponseFar']      <-  6
-
-dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseExpected'] <- 34
-dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con2:Quan2:Item2', 'ResponseFar']      <- 16
-
-dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseExpected'] <- 44
-dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con2:Quan2:Item3', 'ResponseFar']      <- 26
-
-dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseExpected'] <- 54
-dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con2:Quan2:Item4', 'ResponseFar']      <- 36
-
-dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseExpected'] <-  6 
-dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseNear']     <- 15  
-dat[dat$crossed=='Con3:Quan1:Item1', 'ResponseFar']      <- 24
-
-dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseExpected'] <- 16
-dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con3:Quan1:Item2', 'ResponseFar']      <- 34
-
-dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseExpected'] <- 26
-dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con3:Quan1:Item3', 'ResponseFar']      <- 44
-
-dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseExpected'] <- 36
-dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con3:Quan1:Item4', 'ResponseFar']      <- 54
-
-dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseExpected'] <- 24
-dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseNear']     <- 15 
-dat[dat$crossed=='Con3:Quan2:Item1', 'ResponseFar']      <-  6
-
-dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseExpected'] <- 34
-dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con3:Quan2:Item2', 'ResponseFar']      <- 16
-
-dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseExpected'] <- 44
-dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con3:Quan2:Item3', 'ResponseFar']      <- 26
-
-dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseExpected'] <- 54
-dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con3:Quan2:Item4', 'ResponseFar']      <- 36
-
-dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseExpected'] <-  6 
-dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseNear']     <- 15  
-dat[dat$crossed=='Con4:Quan1:Item1', 'ResponseFar']      <- 24
-
-dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseExpected'] <- 16
-dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con4:Quan1:Item2', 'ResponseFar']      <- 34
-
-dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseExpected'] <- 26
-dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con4:Quan1:Item3', 'ResponseFar']      <- 44
-
-dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseExpected'] <- 36
-dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con4:Quan1:Item4', 'ResponseFar']      <- 54
-
-dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseExpected'] <- 24
-dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseNear']     <- 15 
-dat[dat$crossed=='Con4:Quan2:Item1', 'ResponseFar']      <-  6
-
-dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseExpected'] <- 34
-dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseNear']     <- 25  
-dat[dat$crossed=='Con4:Quan2:Item2', 'ResponseFar']      <- 16
-
-dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseExpected'] <- 44
-dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseNear']     <- 35
-dat[dat$crossed=='Con4:Quan2:Item3', 'ResponseFar']      <- 26
-
-dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseExpected'] <- 54
-dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseNear']     <- 45
-dat[dat$crossed=='Con4:Quan2:Item4', 'ResponseFar']      <- 36
-
-dat$isResponseExpected <- dat$choice == dat$ResponseExpected
-dat$isResponseNear <- dat$choice == dat$ResponseNear
-dat$isResponseFar <- dat$choice == dat$ResponseFar
-
-for ( row in 1:nrow(dat) ) {
-  dat[row, 'switchResponse'] <- 
-    ifelse(dat[row, 'isResponseExpected']==TRUE, 'Expected', 
-           ifelse(dat[row, 'isResponseNear']==TRUE, 'Near', 'Far') ) 
-}
+  # classify the response as expected, near, or far
+  dat <- classifyResponse(dat)
 
@@
   # TRIAL
-# Trial for subject, 1 to 192
+  # Trial for subject, 1 to 256
   dat$Trial = rep(x = 1:number_of_trials_per_subject, times = number_of_valid_subjects)
@@
 
-dat$Selection = ""
-dat[dat$Condition %in% c(1,2), "Selection"] <- "matching"
-dat[dat$Condition %in% c(3,4), "Selection"] <- "comparison"
-dat$Selection.sum <- ifelse(dat$Selection=="matching", .5, -.5)
-dat$Selection <- as.factor(dat$Selection)
+  # NUMBER
+  # Create a factor coding for Number use
+  dat[ dat$Condition==1 , 'Number'] <- 'Numeric'
+  dat[ dat$Condition==2 , 'Number'] <- 'Numeric'
+  dat[ dat$Condition==3 , 'Number'] <- 'Verbal'
+  dat[ dat$Condition==4 , 'Number'] <- 'Verbal'
+  dat$Number <- as.factor(dat$Number)
 
-# remove one zero RT by replacing with NA
-dat$RT[dat$RT==0]<-NA
+  # manually center Number
+  dat$c_Num <- ifelse(dat$Number=='Numeric', -.5, .5)
+
+  # CONDITION
+  # make a factor out of Condition, as f_Cnd
+  dat$f_Cnd <- factor(dat$Condition, levels=c(1,2,3,4), labels=c('Vg:Nm', 'Cr:Nm', 'Vg:Vb', 'Cr:Vb'))
 
@@
   # add number of characters in the instruction # 29 30 34 36 38
-dat$nchar_instr = nchar(as.character(dat$Instruction))
-dat$nchar_instr_scaled = as.vector(scale(nchar(as.character(dat$Instruction)), scale=TRUE))
+  dat$nchar_instr = nchar(dat$Instruction)
+  dat$nchar_instr_scaled = as.vector(scale(nchar(dat$Instruction), scale=TRUE))
 
@@
 
+  # print to file a table with information about the design
+  design_info <- unique(subset(dat, select=c(Item, Condition, Vagueness, Number, Quantity, Order, 
+                                             Left, Mid, Right, Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side,Instruction)))
+  design.info <- design_info[order(design_info$Item, design_info$Condition, design_info$Quantity, design_info$Order),]
+  row.names(design_info) <- NULL
+  capture.output(  print.data.frame(design_info, row.names=F, print.gap=3, quote=F, right=F), file="design_info.table")
+
+  # put dat in better column order
+  dat <- subset(dat, 
+                select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number, 
+                           Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,
+                           c_Trl, s_Trl, c_Itm, c_Vag, c_Num, f_Cnd, c_Ord, c_Qty,
+                           RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
+                           Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
+                           Instruction, nchar_instr
+                ))
+
   # save dat as dat.Rda
-# This data set contains *all* trials including impossible trials and is mainly for graphs comparing different removals
+  # This data set contains *all* trials 7680 including impossible trials and is mainly for graphs comparing different removals
   save(dat, file="dat.Rda")
+
+  ####################################################
+  # remove impossible trials from dat in new data dd #
+  ####################################################
+
+  # Throw out RT = 1 and RT = 59998, and RTprev = 1 and RTprev = 59998 
+  # i.e., throw out sticky fingers and timeouts, 
+  # and the trials that followed sticky fingers and timeouts 
+  # since they were likely affected by unusual previous trials.
 
@@
   dd$RT[dd$RT == 1 ] <- NA
-  dd$RT[dd$RT > 40000 ] <- NA
+  dd$RT[dd$RT == 59998 ] <- NA
   dd <- dd[complete.cases(dd),]
@@
 
-# save dd as dd.Rda
-# This data set non-impossible trials and renamed/recomputed variables
+  # number of higher-level cells in the design (abstracting over Order) = Item(4) * Number(2) * Vagueness(2) * Quantity(2) = 32
+  # number of measurements in eaach cell = 8
+  # 332 * 8 = 256
+  # 256 * 30 = 7680
+
+  dd$measurement=0
+  dd$cell=0
+  for (s in levels(dd$Subject)) {
+    cellcount=0
+    for (i in levels(dd$Item)) {
+      for (n in levels(dd$Number)) {
+        for (v in levels(dd$Vagueness)) {
+          for (q in levels(dd$Quantity)) {
+            measurementcount=0
+            cellcount=cellcount+1
+            for (t in dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q, "Trial"]) {
+              dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "cell"] <- cellcount
+              measurementcount=measurementcount+1
+              dd[dd$Subject==s & dd$Item==i & dd$Number==n & dd$Vagueness==v & dd$Quantity == q & dd$Trial==t, "measurement"] = measurementcount
+            }
+          }
+        }
+      }
+    }    
+  }
+
+  # put dd in better column order
+  dd <- subset(dd, 
+               select = c(id, Subject, Trial, Condition, Order, Quantity, Vagueness, Number, 
+                          Item, item_range_ratio, item_range_ratio_scaled, item_mean_ratio, item_mean_ratio_scaled,
+                          c_Trl, s_Trl, c_Itm, c_Vag, c_Num, f_Cnd, c_Ord, c_Qty,
+                          RT, RT_RecSqd, RT_Rec, RT_RecSqt, RT_log, RT_sqt, RT_raw, RT_sqd,
+                          RTprev, RTprev_RecSqd, RTprev_Rec, RTprev_RecSqt, RTprev_log, RTprev_sqt, RTprev_raw, RTprev_sqd,
+                          Exp_Num, Bline_Num, Extr_Num, Exp_side, Bline_side, Extr_side, response_num, response_side, response_category, Left, Mid, Right,
+                          Instruction, nchar_instr,
+                          cell, measurement
+               ))
+
+  # save dd
   save(dd, file="dd.Rda")
+}
 ```
-
-Check the distribution of the trimmed data to see if a transformation is suggested. Log transformation seems to be best so we adopt that for the rest of the analyses.
-
-```{r boxplot subjects and items}
-load("dd.Rda")
-# allow reference to transformation
-subdata = melt(dd,
-               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw"),
-               variable.name="transformation",
-               value.name="value") 
-# boxplots for subjects and items separately for each dependent variable
-ggplot(subdata) + 
-  facet_wrap(~ variable, scales="free", ncol=4) +
-  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
-  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
-  theme_bw() +
-  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
-  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
-  xlab("") + ylab("")
-```
-
-```{r load_dd}
-load("dd.Rda")
-```
-
-```{r do_lmer}
-v5 <- lmer(data=dd, 
-           RT_RecSqt ~ 
-             c_Vag + Selection.sum + c_Qty + c_Ord + 
-             Selection.sum:c_Vag:c_Qty +
-             item_mean_ratio + 
-             s_Trl +
-             RTprev_RecSqt +
-             nchar_instr +
-             (1+c_Vag + Selection.sum + c_Qty + c_Ord|Subject)
-)
-```
-
-```{r summary_lmer}
-summary(v5)
-```
-
-```{r R2}
-cat("R^2")
-cor(fitted(v5), dd$RT_RecSqt)^2
-```
-
-```{r plotLMER}
-par(mfrow=c(2,5))
-plotLMER.fnc(v5)
-```
-
-Plot model coefficients and ci's
-
-```{r coeff}
-par(las=2, mar=c(14,6,1,1))
-y=as.data.frame(summary(v5)$coef[-1,1]) # estimates, without intercept
-q=confint(v5)[-c(1,2,3),]
-y=cbind(y,q)
-plot(y[,1], ylim=range(y[,1],y[,2],y[,3]), xaxt="n", xlab="", ylab="estimate\n",  pch='-', cex=2)
-segments(x0=1:nrow(y), x1=1:nrow(y), y[,2], y[,3])
-abline(h=0)
-axis(1, labels=row.names(y), at=1:nrow(y), srt=45)
-grid()
-```
-
-Model criticism baayen plots
-
-```{r baayenplots, fig.width=9, fig.height=3}
-# Baayen 4-plot model criticism
-
-par(mfrow=c(1,4), pty='s')
-# create scaled residuals
-v5$rstand = as.vector(scale(resid(v5)))
-# plot scaled residuals density
-plot(density(v5$rstand))
-# plot sample quantiles versus theoretical quantiles
-qqnorm(v5$rstand, cex=.5)
-qqline(v5$rstand)
-# plot standardised residuals versus fitted values
-plot(v5$rstand ~ fitted(v5), pch='.')
-# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
-abline(h=c(-2.5,2.5))
-# create dffits
-dffits=abs(resid(v5, "dffits"))
-# plot dffits
-plot(dffits, type='h')
-```

### 2018-08-02 experiments/C_exp_1/archive/gelman_style_plot_regression_lines.Rmd
massive restructuring Aug 2018

## Plot simple regressions on each individual predictor using log

```{r, fig.height=8, eval=TRUE, cache=FALSE}
par(mfrow=c(3,2))
subdata = subset(fulldata, subset=dfSubset== "RT>1&RT<59999")

# trial
fit.trial <- lm(subdata$log ~ subdata$Trial)
with(subdata, plot(log~Trial, cex=.5, col="grey", main="Trial"))
curve (coef(fit.trial)[1] + coef(fit.trial)[2]*x, add=TRUE)

# vagueness
fit.vagueness <- lm(subdata$log ~ subdata$c_Vag)
with(subdata, plot(log~jitter(c_Vag, 1.5), cex=.5, col="grey", xlab="", xaxt="n", main="Vagueness"))
axis(1, at=c(-.5,.5), labels=c("crisp","vague"))
curve (coef(fit.vagueness)[1] + coef(fit.vagueness)[2]*x, add=TRUE)

# number
fit.number <- lm(subdata$log ~ subdata$c_Num)
with(subdata, plot(log~jitter(c_Num, 1.5), cex=.5, col="grey", xlab="", xaxt="n", main="Number"))
axis(1, at=c(-.5,.5), labels=c("verbal","numeric"))
curve (coef(fit.number)[1] + coef(fit.number)[2]*x, add=TRUE)

# item
fit.item <- lm(subdata$log ~ subdata$c_Itm)
with(subdata, plot(log~jitter(c_Itm, 1.5), cex=.5, col="grey", xlab="", xaxt="n", main="Item"))
axis(1, at=c(-1.5, -.5, .5, 1.5), labels=levels(subdata$f_Itm))
curve (coef(fit.item)[1] + coef(fit.item)[2]*x, add=TRUE)

# order
fit.order <- lm(subdata$log ~ subdata$Order)
with(subdata, plot(log~jitter(as.numeric(Order), 1.5), cex=.5, col="grey", xlab="", xaxt="n", main="Order"))
axis(1, at=c(1,2), labels=c("small on left","small on right"))
curve (coef(fit.order)[1] + coef(fit.order)[2]*x, add=TRUE)

# quantity
fit.quantity <- lm(subdata$log ~ subdata$Quantity)
with(subdata, plot(log~jitter(as.numeric(Quantity), 1.5), cex=.5, col="grey", xlab="", xaxt="n", main="Quantity"))
axis(1, at=c(1,2), labels=c("choose small","choose large"))
curve (coef(fit.quantity)[1] + coef(fit.quantity)[2]*x, add=TRUE)
```

### 2018-08-02 experiments/C_exp_1/archive/model_criticism_from_Baayen_Milin_2010_usingRTinv.Rmd
massive restructuring Aug 2018

## Model criticism for logRT model

The residuals are bad for this model.
Baayen says to drop data points for which the absolute standardized residuals of the model exceed 2.5 standard deviations, using code like below

```{r}
# example code from Baayen and Milin, 2010 "Analyzing Reaction Times"
# Baayen, R. Harald, and Petar Milin. "Analyzing reaction times." International Journal of Psychological Research 3.2 (2015): 12-28.
library(languageR) # pulls up lexdec
qqmath(~1/exp(RT) | Subject, data = lexdec, aspect=1, pch='.')

lexdec2A = lexdec2[abs(scale(resid(lexdec2.lmer))) < 2.5, ]
lexdec2A.lmer = lmer(-1000 * RTinv ~ NativeLanguage + Class +
                       + Frequency + Length + (1 | Subject) + (1 | Word), data = lexdec2A)
cor(fitted(lexdec2A.lmer), -1000 * lexdec2A$RTinv)^2
```

```{r, fig.width=4, fig.height=4}
dat2 = dat[abs(scale(resid(lmer.logRT.mm0))) < 2.5, ]
lmer.logRT.mm3 <- lmer(data=dat2,
                       logRT~
                         c_Num * c_Vag * c_Itm + log(Trial.s) +
                         (c_Num * c_Vag + c_Itm | Subject) )
lmer.logRT.mm3.summary <- summary(lmer.logRT.mm3)
cor(fitted(lmer.logRT.mm3), -1000 * dat2$logRT)^2
qqPlot(residuals(lmer.logRT.mm3))
```

### 2018-08-02 experiments/C_exp_1/archive/plot_fast_and_slow_subjects_and_items.Rmd
massive restructuring Aug 2018

Show how mean times for individual subjects and items vary with respect to the grand mean RT.

```{r, fig.width=10, fig.height=5}
# Show how mean times for individual subjects and items vary with respect to the grand mean RT.

suppressMessages(library(plyr))
suppressMessages(library(lattice))
suppressMessages(library(latticeExtra))

subs=ddply(dd, .(level=Subject), summarise, effect="subject", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
itms=ddply(dd, .(level=Item), summarise, effect="item", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
si=rbind(subs,itms)
si$effect <- factor(si$effect, levels=c("subject", "item"))
si$lsign <- factor(si$lsign, levels=c("slow","fast"))

my1 <- xyplot(data=si, lmean~level | effect, 
              aspect=1,
              layout=c(2,1),
              ylim = extendrange(si$lmean, f = 0.10),
              par.settings=list(trellis.par.set(superpose.symbol = list(col = c("blue","red"), pch=19), superpose.line=list(col = c("blue","red")))),
              groups=si$lsign,
              xlab=NULL,
              ylab="RT (in ms)",
              key=list(text=list(levels(si$lsign)), lines=list(lty=c(1,1), col=trellis.par.get('superpose.line')$col, pch=c(19,19), type='o'), divide=1),
              mean=unique(si$emean),
              scales=list(x=list(relation="free", at=list(15.5, 32.5), labels=list("subject number", "item identifier"), cex=1 ), 
                          y=list(at=c(0, 2000, unique(si$emean), 4000, 6000, 8000,  10000), labels=c(0, 2000,  "Mean", 4000,  6000, 8000, 10000), cex=.9) ),
              panel=function(x,y,type,subscripts,groups=groups,...){
                panel.xyplot(x,y,type='p',groups=groups, subscripts=subscripts,...)
                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, groups=groups, subscripts=subscripts, code=3, length=0, col=trellis.par.get('superpose.line')$col[groups][subscripts], ...)
                panel.abline(h=unique(si$emean), col="lightgrey",lty=1)
                panel.text(x, y=(ifelse(y<unique(si$emean),y-300,y+300)), as.numeric(si$level)[subscripts][panel.number()==1], cex=.75)
                panel.text(x, y=(ifelse(y<unique(si$emean),y-300,y+300)), as.character(si$level)[subscripts][panel.number()==2], cex=.75)
              })

my2 <- xyplot(data=si, ldiff~level | effect, 
              aspect=1, type='n',
              ylab="Deviation from mean RT (in ms)",
              ylim=extendrange(si$ldiff,f=0.10))

print(doubleYScale(my1, my2, add.axis=TRUE, add.ylab2=TRUE, use.style=FALSE))
```

### 2018-08-02 experiments/C_exp_1/archive/plot_useful_plotting_code_for_using_later.Rmd
massive restructuring Aug 2018

```{r}
with(rts, xyplot(data=rts, RecSqt ~ F1 | Effect, groups=F2, mins=mins, maxs=maxs,
                 par.settings = standard.theme(color = FALSE),
                 scales=list(x=list(relation="free",at=list( c(1,2),c(1,2),c(1,2),c(3,4),c(3,4),c(5,6) ))),
                 ylim=c(0.970, .984),
                 aspect = 1, layout=c(6,1), xlab="", 
                 panel=function(x,y,groups,subscripts,...){ 
                   panel.arrows(x,mins[subscripts],x,maxs[subscripts],...,lwd=1.5, angle=90, code=3, length=.01)
                   panel.xyplot(x,y,..., groups=groups, subscripts=subscripts, type='l', lty=3)
                   panel.xyplot(x,y,..., groups=groups, subscripts=subscripts, cex=1)
                   panel.abline(h=7.95, col="lightgrey")
                   draw.key(key=list(
                     text = list(text = levels(groups)[levels(groups)%in%groups[subscripts]]), 
                     points = list(pch =  trellis.par.get('superpose.symbol')$pch[1:nlevels(groups)][levels(groups)%in%groups[subscripts]])),
                     draw=TRUE , vp=grid::viewport(x=0.35,y=0.9))}) )
```

### 2018-08-02 experiments/C_exp_1/writeup/2017_submitted_to_springer/2017_submitted_to_springer.Rmd
massive restructuring Aug 2018

--- 2018-07-20 experiments/e2/writeup/2017_submitted_to_springer/e2_2017_analysis.Rmd
@@
 ```{r load Libraries, message=FALSE, echo=FALSE}
-library(knitr); library(Matrix); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions)
+library(knitr); library(Matrix); library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions); library(kableExtra)
 ```
@@
 
-```{r e2-rtplot, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE}
+```{r e2-rtplot, fig.width=7, fig.height=3.5, echo=TRUE, message=FALSE}
 rtdataplot1 <- summarySEwithin2(rtdata, 
@@
 rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness))
+
+rtdataplot1 %>% kable() %>% kable_styling(full_width = F, position = "left", font_size = 11)
+
 rtplot=ggplot(rtdataplot1, aes(y=RT_logNormed, x=Item, group=Condition, fill=Vagueness)) +

### 2018-08-02 experiments/C_exp_1/writeup/2018_tidy_version/e2_2018_analysis.Rmd
massive restructuring Aug 2018

--- 2018-07-27 experiments/e2/writeup/2018_tidy_version/e2_2018_analysis.Rmd
@@
 ```{r}
+ggplot(data=raw_data, aes(RT_log)) + geom_density() +xlim(6,11)+ylim(0,0.8)
+```
+
+```{r}
 trim_data <- perSubjectTrim.fnc(data = raw_data[complete.cases(raw_data),], response='RT', subject='Subject', trim = 2.5)$data
+```
+
+```{r}
+ggplot(data=trim_data, aes(RT_log)) + geom_density()+xlim(6,11)+ylim(0,0.8)
+```
+
+```{r}
+data.frame(obs=1:7680, raw=raw_data$RT_log, trim=c(trim_data$RT_log,rep(NA, times=239))) %>% 
+  gather(key="type", value="value", raw, trim) %>%
+  ggplot(aes(x=value, col=type)) + geom_density() + theme_bw() + theme(panel.grid = element_blank()) 
 ```
@@
 rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness))
+
+rtdataplot1 %>% kable() %>% kable_styling(full_width = F, position = "left", font_size = 11)
 

### 2018-08-02 experiments/D_exp_2/archive/graphs.Rmd
massive restructuring Aug 2018

Vagueness experiment 4 graphs
========================================================

```{r loaddata, echo=FALSE}
rm(list=ls())
load("e4vagueness.RData")
levels(x$vagueness)<-c("Crisp","Vague")
levels(x$selection)<-c("Comparison","Matching")
levels(x$Item)<-c("6:15:24","16:25:34","26:35:44","36:45:54")
source("CIfunctions.R")
```

RT cutoffs: subset=RT>0&RT<49000

```{r loadlibraries, echo=FALSE}
require(ggplot2, quietly=TRUE)
require(reshape2, quietly=TRUE)
require(lattice, quietly=TRUE)
require(gridExtra, quietly=TRUE)
library(plyr)
```

Show distribution of response times.

```{r rtdist1 ,echo=FALSE}
# Untransformed
untransformed = ggplot(x, aes(x=RT)) + 
  geom_histogram(aes(y=..density..),
                 binwidth=500,
                 colour="black", fill="white") +
  geom_density(alpha=.2, fill="blue")
```

```{r logrtdist, echo=FALSE }
# Log transformed
logtransformed = ggplot(x, aes(x=logRT)) + 
  geom_histogram(aes(y=..density..),
                 binwidth=.1,
                 colour="black", fill="white") +
  geom_density(alpha=.2, fill="blue")
```

```{r showtransform, echo=FALSE, fig.width=7, fig.height=3, warning=FALSE}
grid.arrange(untransformed,logtransformed, ncol=2)
```

Show distribution separately for each group.

First way.

```{r density1,echo=FALSE}
# Density plot per group overlaid with means
cdf <- ddply(x, .(Effect), summarise, logRT.mean=mean(logRT))
distributionbygroup = ggplot(x, aes(x=logRT, colour=Effect)) + geom_density() + theme_bw()

distributionbygroupAddMeans = distributionbygroup  + geom_vline(data=cdf, aes(xintercept=logRT.mean, colour=Effect), linetype="dashed", size=.5)  
```

```{r density2,echo=FALSE, fig.width=7, fig.height=3}
print(distributionbygroup)
```

```{r densityfaceted,echo=FALSE}
# Density plot per group faceted
distributionbygroupv2 = ggplot(x, aes(x=logRT)) + 
  geom_histogram(binwidth=.1, colour="black", fill="white")  +
  geom_vline(data=cdf, aes(xintercept=logRT.mean),
             linetype="dashed", size=1, colour="red") +
facet_grid(Effect~.) 
```

Second way

```{r showway2dist,echo=FALSE,fig.width=5,fig.height=5,warning=FALSE}
print(distributionbygroupv2)
```

Boxplot with mean indicator.

```{r boxplot1, fig.width=3,fig.height=3, echo=FALSE}
# Boxplot with mean indicator
boxplotmeanind = ggplot(x, aes(x=vagueness, y=logRT)) + geom_boxplot() +
  stat_summary(fun.y=mean, geom="point", shape=5, size=4) +
  facet_wrap(~selection)
print(boxplotmeanind)
```

```{r, conditionmeans, echo=FALSE}
# condition means - main effect of selection
meanRT <- tapply(x$RT,x$selection,mean)
selection <- factor(rep(levels(x$selection),each=1))
rt1=ggplot(data.frame(selection,meanRT), aes(y=meanRT,x=selection)) +
  geom_bar(position="dodge",stat="identity") + scale_fill_grey()+theme_bw()+
  coord_cartesian(ylim = c(0, 3200))
```

```{r moreconditionmeans,echo=FALSE}
# condition means - main effect of vagueness
meanRT <- tapply(x$RT,x$vagueness,mean)
vagueness <- factor(rep(levels(x$vagueness),each=1))
rt2=ggplot(data.frame(vagueness,meanRT), aes(y=meanRT,x=vagueness)) +
  geom_bar(position="dodge",stat="identity") + scale_fill_grey()+theme_bw()+
  coord_cartesian(ylim = c(0, 3200))
```

Show the main effects of vagueness and selection separately.

```{r showmaineffects, echo=FALSE, fig.width=7, fig.height=3}
grid.arrange(rt1,rt2,ncol=2)
```

```{r showmaineffects2, echo=FALSE}
# condition means but with ggplot2
# This was the code for precogsci2013 bar plot of RT by condition
meanRT <- tapply(x$RT,x$selection:x$vagueness,mean)
vagueness <- factor(rep(levels(x$vagueness),2))
selection <- factor(rep(levels(x$selection),each=2))
rt3=ggplot(data.frame(vagueness,selection,meanRT), 
           aes(y=meanRT,x=selection,fill=vagueness)) +
  geom_bar(position="dodge",stat="identity") + 
  scale_fill_grey() + 
  theme_bw() + coord_cartesian(ylim = c(0, 3200)) +
  geom_segment(aes(x = .75, y = 2000, xend = .75, yend = 2050))+
  geom_segment(aes(x =.75, y=2050, xend=1.25,yend=2050))+
  geom_segment(aes(x =1.25, y=2050, xend=1.25,yend=2000))+
  annotate("text",x=1, y=2100, label="*", size=7)+
  geom_segment(aes(x = 1.75, y = 2600, xend = 1.75, yend = 2650))+
  geom_segment(aes(x = 1.75, y = 2650, xend = 2.25, yend = 2650))+
  geom_segment(aes(x = 2.25, y = 2650, xend = 2.25, yend = 2600))+
  annotate("text",x=2, y=2720, label="NS", size=3)+
  geom_segment(aes(x=1,y=2200,xend=1,yend=3000))+
  geom_segment(aes(x=1,y=3000,xend=2,yend=3000))+
  geom_segment(aes(x=1,y=3000,xend=2,yend=3000))+
  geom_segment(aes(x=2,y=3000,xend=2,yend=2800))+
  annotate("text",x=1.5,y=3050, label="***", size=7)
#+
#              ggtitle(label="Response Time")
```

```{r showmaineffects3,echo=FALSE}
# condition means the other way
rt4=ggplot(data.frame(vagueness,selection,meanRT), aes(y=meanRT,x=vagueness,fill=selection)) + 
  geom_bar(position="dodge",stat="identity") + scale_fill_grey()+theme_bw()+coord_cartesian(ylim = c(0, 3200))+ 
  geom_segment(aes(x=1,y=2600,xend=1,yend=3000))+
  geom_segment(aes(x=1,y=3000,xend=2,yend=3000))+
  geom_segment(aes(x=1,y=3000,xend=2,yend=3000))+
  geom_segment(aes(x=2,y=3000,xend=2,yend=2600))+
  annotate("text",x=1.5,y=3050, label="NS", size=3)
```

Show the Selection x Vagueness interaction.

```{r selxvag, fig.width=8, fig.height=3, echo=FALSE}
grid.arrange(rt3,rt4,ncol=2)
```

```{r PEXmake, echo=FALSE}
# Probability of choosing an extreme response
pEX <- tapply(x$EXTREME,x$selection:x$vagueness,mean)
vagueness <- factor(rep(levels(x$vagueness),2))
selection <- factor(rep(levels(x$selection),each=2))

pex1 = 
  ggplot(data.frame(vagueness,selection,pEX), 
                       aes(y=pEX,x=selection,fill=vagueness)) + 
  geom_bar(position="dodge",stat="identity") + 
  scale_fill_grey() + 
  theme_bw() +
  scale_y_continuous(breaks=c(0,.25,.5,.75,1),
                     limits=c(0,1.23))+
#  ggtitle(label="Probability of choosing the extreme answer") +
  geom_segment(aes(x = .75, y = .9, xend = .75, yend = 1)) + # left upstroke
  geom_segment(aes(x = .75, y = 1,  xend = 1.25, yend = 1)) + # left across
  geom_segment(aes(x =1.25, y=1, xend=1.25,yend=.95)) + # left downstroke
  annotate("text",x=1, y=1.01, label="*", size=7) + # left annotation
  geom_segment(aes(x = 1.75, y = .8, xend = 1.75, yend = .9))+ # right
  geom_segment(aes(x = 1.75, y = .9, xend = 2.25, yend = .9))+
  geom_segment(aes(x = 2.25, y = .9, xend = 2.25, yend = .6))+
  annotate("text",x=2, y=.92, label="*", size=7) +
  geom_segment(aes(x=1,y=1.1,xend=1,yend=1.15)) + # middle
  geom_segment(aes(x=1,y=1.15,xend=2,yend=1.15)) +
  geom_segment(aes(x=1,y=1.15,xend=2,yend=1.15)) +
  geom_segment(aes(x=2,y=1.15,xend=2,yend=1)) +
  annotate("text",x=1.5,y=1.2, label="***", size=7)

pex1.inverted = ggplot(data.frame(vagueness,selection,pEX), 
              aes(y=1-pEX,x=selection,fill=vagueness)) + 
  geom_bar(position="dodge",stat="identity") + 
  scale_fill_grey() + 
  theme_bw() +
  scale_y_continuous(breaks=c(0,.25,.5,.75,1))+
#   coord_cartesian(ylim = c(0, 1.3)) + 
#  ggtitle(label="1 - Probability of choosing the extreme answer") +
  geom_segment(aes(x = .75, y = .2, xend = .75, yend = .3))+
  geom_segment(aes(x =.75, y=.3, xend=1.25,yend=.3))+
  geom_segment(aes(x =1.25, y=.3, xend=1.25,yend=.15))+
  annotate("text",x=1, y=.33, label="*", size=7) +
  geom_segment(aes(x = 1.75, y = .4, xend = 1.75, yend = .6))+
  geom_segment(aes(x = 1.75, y = .6, xend = 2.25, yend = .6))+
  geom_segment(aes(x = 2.25, y = .6, xend = 2.25, yend = .55))+
  annotate("text",x=2, y=.65, label="*", size=7) +
  geom_segment(aes(x=1,y=.4,xend=1,yend=.8))+
  geom_segment(aes(x=1,y=.8,xend=2,yend=.8))+
  geom_segment(aes(x=1,y=.8,xend=2,yend=.8))+
  geom_segment(aes(x=2,y=.8,xend=2,yend=.7))+
  annotate("text",x=1.5,y=.85, label="***", size=7)
```

Probability of an extreme response.

```{r pexshow, fig.width=8,fig.height=3, echo=FALSE}
grid.arrange(pex1,pex1.inverted, ncol=2)
```

Plots for poster preCogSci 2013

```{r makePlotData}
poda=summarySEwithin(x,measurevar="RT",withinvars=c("vagueness","selection", "Item"), idvar="Subject")
poda$mygroup <- paste(poda$selection,poda$vagueness)
```

```{r makeCogSciPlot}
cogsci=
  ggplot(data=poda,
       aes(y=RT,ymin=RT-ci,ymax=RT+ci,x=Item,group=mygroup,col=mygroup,shape=mygroup,fill=mygroup)) +
  geom_errorbar(position=position_dodge(width=.2), width=.2) +
  geom_line(position=position_dodge(width=.2)) +
  geom_point(position=position_dodge(width=.2), size=4) +
  xlab("Dots")+
  ylab("Reaction time (ms)")+
  scale_colour_manual(values=c("red","blue","red","blue")) +
  scale_shape_manual(values=c(21,21,23,23))+
  scale_fill_manual(values=c("red","blue","white","white")) +
  ggtitle(label="Forced choice task") +
  theme(legend.key = element_blank(),
        legend.title = element_blank(), 
        legend.text = element_text(face="bold", size=12),
        legend.key.size = unit(c(1.5,1.5), "lines"),
        axis.title.x = element_text(face="bold", size=12, vjust=-1.5),
        axis.title.y = element_text(face="bold", size=12, angle=90, vjust=-0.01),
        plot.title = element_text(face="bold", size=14, vjust=1.5),
        plot.margin = unit(c(2, -9.0, 2, 2), "lines"),
        panel.background = element_rect(fill='white', colour='black'))
```

Show cogsci plot
```{r ShowCogSciPlot, fig.width=7,fig.height=5}
print(cogsci)
```

```{r savePlot}
pdf(file="figure/precogsci2013e4.pdf",width=7, height=5)
print(cogsci)
dev.off()
```

```{bwhitevxs}
b <- summarySEwithin(x,measurevar="RT",withinvars=c("vagueness","selection"), idvar="Subject")
b$mygroup=paste(b$selection, b$vagueness)
pd=position_dodge(0.1)
ggplot(data=b,aes(y=RT,x=vagueness,group=selection,col=mygroup,shape=mygroup,fill=mygroup,ymax=RT+ci,ymin=RT-ci))+geom_errorbar(position=pd, width=.1)+geom_line(position=pd)+geom_point(position=pd,size=4)+  scale_colour_manual(values = c("black","black", "black", "black")) +   scale_shape_manual(values = c(21, 23, 21, 23)) +    scale_fill_manual(values = c("black", "black", "white","white"))
```

### 2018-08-02 experiments/D_exp_2/archive/models.Rmd
massive restructuring Aug 2018

Vagueness experiment 4 models
========================================================

```{r loaddata, echo=FALSE}
rm(list=ls())
load("e4vagueness.RData")
source("CIfunctions.R")
```

```{r loadlibs}
suppressPackageStartupMessages(library(lme4,quietly=TRUE,warn.conflicts=FALSE))
suppressPackageStartupMessages(library(lmerTest,quietly=TRUE,warn.conflicts=FALSE))
```

Set anova contrasts

```{r setAnovaContrast}
anova.contrast <- matrix(c(-.5,-.5,.5,.5, -.5,.5,-.5,.5, .5,-.5,-.5,.5), 4,3,                      dimnames = list(c("comparison:crisp", "comparison:vague", "matching:crisp", "matching:vague"),  c("selection", "vagueness", "selectionXvagueness")))  
contrasts(x$Effect) <- anova.contrast
```

Do omnibus lmer using custom anova contrast.

```{r maximalLMERrt}
maximal.lmer <- lmer(data=x, logRT ~ Effect + (Effect|Item) + (Effect|Subject))
```

Planned comparisons crisp versus vague at each level of selection.

Comparison vague versus comparison crisp.

```{r comparisonLMERrt}
comparison.lmer <- lmer(data=x, subset=selection=="comparison", logRT ~ vagueness +  (vagueness|Item) + (vagueness|Subject))
```

Matching vague versus matching crisp.

```{r matchingLMERrt}
matching.lmer <- lmer(data=x, subset=selection=="matching", logRT ~ vagueness  + (vagueness|Item) + (vagueness|Subject))
```

```{r}
maximal.lmer
comparison.lmer
matching.lmer
```

PLot the comparison vague versus comparison crisp

```{r}
x2 <- subset(x, subset=selection=="comparison", select=c(Subject,vagueness,RT))
x3 <- summarySEwithin(x2,measurevar="RT",withinvars="vagueness",idvar="Subject")
ggplot(data=x3,aes(y=RT,x=vagueness,ymin=RT-ci,ymax=RT+ci,group=vagueness,stat="identity"))+geom_errorbar(width=.2)+geom_point()

```

### 2018-08-02 experiments/E_exp_3/writeup/scratch.Rmd
massive restructuring Aug 2018

```{r, loadLibraries, echo=FALSE}
# Load libraries.
suppressPackageStartupMessages(library(MASS))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lme4))
suppressPackageStartupMessages(library(xtable))
suppressPackageStartupMessages(library(plyr))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library(lattice))
suppressPackageStartupMessages(library(latticeExtra))
suppressPackageStartupMessages(library(LMERConvenienceFunctions))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(grid))
suppressPackageStartupMessages(library(lmerTest))
```

Compare distributions of the various transformations of RT against random samples from normal distributions with the same mean and sd to see which transformations best approximate normal distributions.

```{r foo,fig.width=20, fig.height=12, message=FALSE}
## plot the comparison of the transformations against their normal equivalents to see distributional properties using full data

# first create a data frame specially for transformations
# this won't be used for analysis, just for this plot
dat_transforms <- dat

# dfSubset says whether any data points were removed
dat_transforms$dfSubset <- "original"

# RTtype says whether the row contains data for an observed RT or 
# a sample from a normal distribution with the same mean and sd
dat_transforms$RTtype <- "observed"

# create extra rows for dfs with removed (outlier) data points
dat1 <- dat_transforms                                       # nrow 7680
dat2 <- subset(dat_transforms, RT<9000)              # nrow 7677
dat2$dfSubset <- "RT<9000"

# bring the dfs back into main df
dat_transforms <- rbind(dat1,dat2)
dat_transforms$dfSubset <- factor(dat_transforms$dfSubset, levels=c("original", "RT<9000"))

# create extra rows for normally distributed transformed variables
dato = dat_transforms
datn = dat_transforms; datn$RTtype <- "normal"
dat_transforms <- rbind(dato,datn)
dat_transforms$RTtype <- factor(dat_transforms$RTtype, levels=c("observed","normal"))

# for safety
dat_transforms[dat_transforms$RTtype=="normal", c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd")] <- NA

# Create the normally distributed transformed variables, replacing the NAs declared above
for (k in levels(dat_transforms$dfSubset)){
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_RecSqd"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_RecSqd"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_RecSqd"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_RecSqd"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_Rec"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_Rec"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_Rec"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_Rec"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_RecSqt"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_RecSqt"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_RecSqt"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_RecSqt"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_log"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_log"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_log"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_log"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_sqt"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_sqt"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_sqt"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_sqt"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_raw"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_raw"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_raw"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_raw"]))
  dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="normal", "RT_sqd"]  <- 
    rnorm(n=length(dat_transforms[dat_transforms$dfSubset==k    & dat_transforms$RTtype=="observed", "RT_sqd"]), 
          mean=mean(dat_transforms[dat_transforms$dfSubset==k & dat_transforms$RTtype=="observed", "RT_sqd"]), 
          sd=sd(dat_transforms[dat_transforms$dfSubset==k     & dat_transforms$RTtype=="observed", "RT_sqd"]))
}

temp <- reshape2::melt(data=dat_transforms, 
             id.vars=c("dfSubset", "RTtype", "Item", "Vagueness", "Selection", "Quantity", "Order"),
             measure.vars=c("RT_RecSqd", "RT_Rec", "RT_RecSqt", "RT_log", "RT_sqt", "RT_raw", "RT_sqd"))

ggplot(temp, 
			 aes(value, colour=RTtype)) + 
	geom_freqpoly() + 
  facet_wrap(dfSubset~variable, scales="free", nrow=2) +
  theme_bw() + theme(aspect.ratio=1) + scale_color_manual(values=c("blue","red"))
```

Show how transformations of RT affect distribution of times, and how they affect which times are outliers.

```{r bar, fig.width=20, fig.height=5}
# boxplot subjects and items
load("data.Rda")
# allow reference to transformation
subdata = melt(dd,
               measure.vars=c("RT_Rec", "RT_RecSqt", "RT_log", "RT_raw")) 
# boxplots for subjects and items separately for each dependent variable
ggplot(subdata) + 
  facet_wrap(~ variable, scales="free", ncol=4) +
  geom_boxplot(aes(y=value, x=Item, col="Items")) + 
  geom_boxplot(aes(y=value, x=Subject, col="Subjects")) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), aspect.ratio=1, panel.grid=element_blank()) + 
  scale_colour_manual("",values=c("Items"="red","Subjects"="blue")) +
  xlab("") + ylab("")
```

Show how mean times for individual subjects and items vary with respect to the grand mean RT.

```{r, bar2, fig.width=10, fig.height=5}
# Show how mean times for individual subjects and items vary with respect to the grand mean RT.
subs=ddply(dd, .(level=Subject), summarise, effect="subject", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
itms=ddply(dd, .(level=Item), summarise, effect="item", emean=mean(dd$RT), lmean=mean(RT), ldiff=mean(RT)-mean(dd$RT), lsign=ifelse(mean(RT)-mean(dd$RT)>0,"slow","fast"))
si=rbind(subs,itms)
si$effect <- factor(si$effect, levels=c("subject", "item"))
si$lsign <- factor(si$lsign, levels=c("slow","fast"))

my1 <- xyplot(data=si, lmean~level | effect, 
              aspect=1,
              layout=c(2,1),
              ylim = extendrange(si$lmean, f = 0.10),
              par.settings=list(trellis.par.set(superpose.symbol = list(col = c("blue","red"), pch=19), superpose.line=list(col = c("blue","red")))),
              groups=si$lsign,
              xlab=NULL,
              ylab="RT (in ms)",
              key=list(text=list(levels(si$lsign)), lines=list(lty=c(1,1), col=trellis.par.get('superpose.line')$col, pch=c(19,19), type='o'), divide=1),
              mean=unique(si$emean),
              scales=list(x=list(relation="free", at=list(15.5, 32.5), labels=list("subject number", "item identifier"), cex=1 ), 
                          y=list(at=c(0, 2000, unique(si$emean), 4000, 6000, 8000,  10000), labels=c(0, 2000,  "Mean", 4000,  6000, 8000, 10000), cex=.9) ),
              panel=function(x,y,type,subscripts,groups=groups,...){
                panel.xyplot(x,y,type='p',groups=groups, subscripts=subscripts,...)
                panel.arrows(x0=x, y0=unique(si$emean), x1=x, y1=y, groups=groups, subscripts=subscripts, code=3, length=0, col=trellis.par.get('superpose.line')$col[groups][subscripts], ...)
                panel.abline(h=unique(si$emean), col="lightgrey",lty=1)
                panel.text(x, y=(ifelse(y<unique(si$emean),y-200,y+200)), as.numeric(si$level)[subscripts][panel.number()==1], cex=.75)
                panel.text(x, y=(ifelse(y<unique(si$emean),y-200,y+200)), as.character(si$level)[subscripts][panel.number()==2], cex=.75)
              })

my2 <- xyplot(data=si, ldiff~level | effect, 
              aspect=1, type='n',
              ylab="Deviation from mean RT (in ms)",
              ylim=extendrange(si$ldiff,f=0.10))

print(doubleYScale(my1, my2, add.axis=TRUE, add.ylab2=TRUE, use.style=FALSE))
```

Show how practice and fatigue effects vary over subjects

```{r barr3, eval=TRUE, fig.width=20, fig.height=10, cache=FALSE, echo=TRUE}
# Show how practice/fatigue effects vary over subjects
ggplot(dd, aes(y=RT_log, x=Trial)) + facet_wrap(~Subject,nrow=3) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial", breaks=NULL) + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1) 
```

Show practice and fatigue effects vary over items.

```{r, var3, eval=TRUE, fig.width=9, fig.height=4, cache=FALSE, echo=TRUE}
# Show how practice/fatigue effects vary over items
ggplot(dd, aes(y=RT_log, x=Trial)) + facet_grid(~Item) + theme(panel.grid=element_blank(), panel.margin = unit(0, "lines"), panel.background= element_rect(fill = 'white', colour='black' ), legend.position="top") + scale_x_continuous("Trial") + geom_point(pch=19, cex=.25, alpha=.5) + stat_smooth(fill='lightblue',col="red",lwd=.5, show.legend=TRUE) + theme(aspect.ratio = 1) 
```

# Show means.

Main effects

```{r, bar4}
# prep main effects rts data frame
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin2(dd, measurevar="RT_log", withinvars=c("Vagueness"), idvar = "Subject"));names(vrts)[3]<-"Levels"
nrts <- cbind(Effect="Selection",x=factor(c(1,2)), summarySEwithin2(dd, measurevar="RT_log", withinvars=c("Selection"), idvar = "Subject"));names(nrts)[3]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin2(dd, measurevar="RT_log", withinvars=c("Quantity"), idvar = "Subject"));names(qrts)[3]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin2(dd, measurevar="RT_log", withinvars=c("Order"), idvar = "Subject"));names(orts)[3]<-"Levels"
irts <- cbind(Effect="Item",x=factor(c(1,2,3,4)), summarySEwithin2(dd, measurevar="RT_log", withinvars=c("Item"), idvar = "Subject"));names(irts)[3]<-"Levels"
rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_log-rts$ci
rts$maxs=rts$RT_log+rts$ci

# plot main effects
ggplot(rts, aes(y=RT_log, x=Levels, group=1, ymin=mins, ymax=maxs)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_grid(~Effect, scale="free_x") + 
  theme_bw() +
  theme(aspect.ratio=1) 
```

Main effects over items

```{r, bar5, message=FALSE, fig.width=10, fig.height=3}
vrts <- cbind(Effect="Vagueness",x=factor(c(1,2)),summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Vagueness")));names(vrts)[4]<-"Levels"
nrts <- cbind(Effect="Selection",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Selection")));names(nrts)[4]<-"Levels"
qrts <- cbind(Effect="Quantity",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Quantity")));names(qrts)[4]<-"Levels"
orts <- cbind(Effect="Order",x=factor(c(1,2)), summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio","Order")));names(orts)[4]<-"Levels"

rts=rbind(vrts,nrts,qrts,orts)
rts$mins=rts$RT_log-rts$ci
rts$maxs=rts$RT_log+rts$ci

rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))

ggplot(rts, aes(y=RT_log, x=item_mean_ratio, group=Levels, ymin=mins, ymax=maxs, shape=Levels, fill=Levels, col=Levels)) +
  scale_shape_manual(values=c(21,21,22,22,23,23,24,24)) +
  scale_fill_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
  scale_color_manual(values=c("black","gray","black","gray","black","gray","black","gray")) +
  geom_line() +
  geom_errorbar(width=.01) +
  geom_point(cex=3) + 
  facet_grid(~Effect) + 
  theme_bw() +
  theme(aspect.ratio=1, 
        panel.grid=element_blank()) +
  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1)) 
```

2 way interactions ignoring item

```{r, bar6}
# prep 2 ways rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Selection")); names(rts1)[1] <- "F1"; names(rts1)[2] <- "F2"
rts1$Effect <- rep("Vagueness:Selection",4)
rts2 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Quantity")); names(rts2)[1] <- "F1"; names(rts2)[2] <- "F2"
rts2$Effect <- rep("Vagueness:Quantity",4)
rts3 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Vagueness","Order")); names(rts3)[1] <- "F1"; names(rts3)[2] <- "F2"
rts3$Effect <- rep("Vagueness:Order",4)
rts5 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection","Quantity")); names(rts5)[1] <- "F1"; names(rts5)[2] <- "F2"
rts5$Effect <- rep("Selection:Quantity",4)
rts6 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Selection","Order")); names(rts6)[1] <- "F1"; names(rts6)[2] <- "F2"
rts6$Effect <- rep("Selection:Order",4)
rts8 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("Quantity","Order")); names(rts8)[1] <- "F1"; names(rts8)[2] <- "F2"
rts8$Effect <- rep("Quantity:Order",4)
rts <- rbind(rts1,rts2,rts3,rts5,rts6,rts8)
rts <- subset(rts, select=c(Effect, F1, F2, RT_log, RT_log_norm, sd, se, ci))
rts$mins=rts$RT_log-rts$ci
rts$maxs=rts$RT_log+rts$ci
```

```{r,bar7,fig.width=10, fig.height=3}
# plot 2 way interactions
ggplot(rts, aes(y=RT_log, x=F2, group=F1, ymin=mins, ymax=maxs, pch=F1)) + 
  geom_line() +
  geom_errorbar(width=.1) +
  geom_point(cex=3) + 
  facet_wrap(~Effect, scale="free_x", nrow=1) + 
  theme_bw() +
  theme(aspect.ratio=1) +
  theme(legend.title=element_blank()) +
  xlab("Levels")
```

Plot 2 way interactions over items

```{r foo2, message=FALSE}
# prep 2 ways over items rts data frame
rts1 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Vagueness","Selection")); names(rts1)[2] <- "F1"; names(rts1)[3] <- "F2"
rts1$E1 <- "Vagueness"
rts1$E2 <- "Selection"
rts2 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Vagueness","Quantity")); names(rts2)[2] <- "F1"; names(rts2)[3] <- "F2"
rts2$E1 <- "Vagueness"
rts2$E2 <- "Quantity"
rts3 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Vagueness","Order")); names(rts3)[2] <- "F1"; names(rts3)[3] <- "F2"
rts3$E1 <- "Vagueness"
rts3$E2 <- "Order"
rts4 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Quantity", "Selection")); names(rts4)[2] <- "F1"; names(rts4)[3] <- "F2"
rts4$E1 <- "Quantity"
rts4$E2 <- "Selection"
rts5 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Order","Selection")); names(rts5)[2] <- "F1"; names(rts5)[3] <- "F2"
rts5$E1 <- "Order"
rts5$E2 <- "Selection"
rts6 <- summarySEwithin(dd, measurevar="RT_log", withinvars=c("item_mean_ratio", "Quantity","Order")); names(rts6)[2] <- "F1"; names(rts6)[3] <- "F2"
rts6$E1 <- "Quantity"
rts6$E2 <- "Order"

rts <- rbind(rts1,rts2,rts3,rts4,rts5,rts6)
rts$mins=rts$RT_log-rts$ci
rts$maxs=rts$RT_log+rts$ci

rts$item_mean_ratio <- as.numeric(as.character(rts$item_mean_ratio))
rts$F3 <- factor(paste(rts$F1,rts$F2))

rts$E1 <- factor(rts$E1, levels=c("Selection","Order","Quantity","Vagueness"))
rts$E2 <- factor(rts$E2, levels=c("Selection","Order","Quantity","Vagueness"))

# The palette with black:
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

ggplot(rts, aes(y=RT_log, x=item_mean_ratio, group=F3, col=F2, pch=F1)) + 
  facet_grid(E2~E1, drop=F) + 
  geom_point(cex=3) + geom_line() + theme_bw() + 
  theme(aspect.ratio=1, panel.grid=element_blank()) +
  scale_color_manual(name="colour",values=cbbPalette) +
  scale_shape_discrete(name="shape") +
  scale_x_continuous(breaks = round(sort(unique(rts$item_mean_ratio)),1))
```

# RT model all items

```{r, foo3, rtmodel-allitems}
# backing off A LOT from maximal model to avoid warning about maxfun
rtm <- lme4::lmer(
  data=dd,
  RT_log ~ 
    c_Vag * Selection.sum + item_mean_ratio + c_Qty + c_Ord + s_Trl +
    RTprev_log  + 
  ( c_Vag * Selection.sum | Subject)
)
rtm.summary <- summary(rtm)
rtm.confint <- confint(rtm, method='Wald')[13:20,]
```

Print a table of coefficients.

```{r, foo4, rtmodeltable, results='asis', echo=FALSE}
print(xtable(
  digits=c(0,2,2,1), 
  rtm.summary$coefficients), type='html')
```

Plot coefficients

```{r, foo5}
par(las=2, mar=c(14,6,1,1))
y=coef(rtm.summary)[-1,1] # omit intercept
n=length(y)
plot(y, xaxt="n", xlab="", ylab="RT_log\n", pch=19, ylim=extendrange(y, f=.20), main="Coefficients")
abline(h=0)
axis(1, labels=names(y) , at=1:n)
grid()
segments(x0=1:n, x1=1:n, y0=rtm.confint[,1], y1=rtm.confint[,2], lwd=2)
```

Plot partial effects

```{r foo6}
par(mfrow=c(2,4))
plotLMER.fnc(rtm)
```

```{r foo7}
suppressPackageStartupMessages(require(languageR))
cat("k is:",
collin.fnc(dd[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])$cnumber)
```

```{r foo8, fig.width=6, fig.height=5}
suppressPackageStartupMessages(require(rms))
par(mar=c(1.1,3.1,1.1,1.1), pty='s')
plot(varclus(as.matrix(dd[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])))
```

Model criticism baayen plots

```{r daz, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

par(mfrow=c(1,4), pty='s')
# create scaled residuals
dd$rstand = as.vector(scale(resid(rtm)))
# plot scaled residuals density
plot(density(dd$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(dd$rstand, cex=.5)
qqline(dd$rstand)
# plot standardised residuals versus fitted values
plot(dd$rstand ~ fitted(rtm), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))

```

Refit model with outliers removed (3.5 sd)

```{r daz2}
dd2 <- subset(dd, abs(rstand)<3.5 )
```

```{r daz3}
# backing off A LOT from maximal model to avoid warning about maxfun
rtm <- lmer(
  data=dd2,
  RT_log ~ 
    c_Vag * Selection.sum + item_mean_ratio + c_Qty + c_Ord + s_Trl +
    RTprev_log  + 
  ( c_Vag * Selection.sum | Subject)
)
rtm.summary <- summary(rtm)
rtm.confint <- confint(rtm, method='Wald')
```

Print a table of coefficients.

```{r daz4,  results='asis', echo=FALSE}
print(xtable(
  digits=c(0,2,2,1,1,3), 
  rtm.summary$coefficients), type='html')
```

Plot coefficients

```{r daz5}
par(las=2, mar=c(14,6,1,1))
y=coef(rtm.summary)[-1,1] # omit intercept
n=length(y)
plot(y, xaxt="n", xlab="", ylab="RT_log\n", pch=19, ylim=extendrange(y, f=.20), main="Coefficients")
abline(h=0)
axis(1, labels=names(y) , at=1:n)
grid()
segments(x0=1:n, x1=1:n, y0=rtm.confint[13:20,1], y1=rtm.confint[13:20,2], lwd=2)
```

Plot partial effects

```{r daz6}
par(mfrow=c(2,4))
plotLMER.fnc(rtm)
```

```{r daz7}
suppressPackageStartupMessages(require(languageR))
cat("k is:",
collin.fnc(dd2[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])$cnumber
)
```

```{r daz9, fig.width=6, fig.height=5}
suppressPackageStartupMessages(require(rms))
par(mar=c(1.1,3.1,1.1,1.1), pty='s')
plot(varclus(as.matrix(dd2[,c("c_Trl", "Selection.sum", "c_Vag", "c_Qty", "c_Ord", "item_mean_ratio", "RTprev_log")])))
```

Model criticism baayen plots

```{r mc1, fig.width=9, fig.height=3}
# Baayen 4-plot model criticism

par(mfrow=c(1,4), pty='s')
# create scaled residuals
dd2$rstand = as.vector(scale(resid(rtm)))
# plot scaled residuals density
plot(density(dd2$rstand))
# plot sample quantiles versus theoretical quantiles
qqnorm(dd2$rstand, cex=.5)
qqline(dd2$rstand)
# plot standardised residuals versus fitted values
plot(dd2$rstand ~ fitted(rtm), pch='.')
# absolute standardised residuals greater than 2.5 are candidates for being outliers, the abline identifies them on the plot
abline(h=c(-2.5,2.5))

```

Separate analyses for vague vs crisp at each level of selection (matching, and comparison)

not done yet.

```{r sep1,  results='asis', echo=FALSE}
print(xtable(
  digits=c(0,2,2,1,1,3), 
  rtm.summary$coefficients), type='html')
```

# Borderline analysis

not done yet.

### 2018-08-02 springer/current/e1.tex
massive restructuring Aug 2018

--- 2018-04-23 papers/2018-v-book/v-book-2017.tex
@@
-\documentclass[%
-%			% choose man or doc immediately below
-%doc,	% doc is single spaced
-man,		% man is double spaced
-floatsintext,%
-apacite%
-]{apa6}
-\usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{color}
-\usepackage{booktabs}
-\usepackage{layout}
-\usepackage{apacite}
-\usepackage{datetime}
-
-%\renewcommand{\baselinestretch}{1.7} % added 21 Sept 2016, copied from CRM book, but see documentclass option 'man' for "manuscript'' above which makes it double spaced
-
-% MG change p's to 3 dp
-
-\title{The Elusive Benefits of Vagueness \protect \\ Evidence from Experiments}
-\date{\today}
-%\authornote{latex compiled \today ~\currenttime}
-\shorttitle{Is Vagueness Beneficial for Hearers?}
-\twoauthors{Matt Green}{Kees van Deemter}
-\twoaffiliations{Computing Science\\ University of Aberdeen}{Computing Science\\ University of Aberdeen}
-
-\abstract{Much of everyday language is vague, yet the benefits of vagueness for hearers and readers are proving to be elusive.
-We discuss a range of earlier controlled experiments with human participants, and we report on a new series of experiments that we conducted in recent years.
-These experiments, which focus on vague expressions that are part of referential noun phrases, aim to separate the utility of vagueness (as defined by the existence of borderline cases) from the utility of other factors that tend to co-occur with vagueness.
-Having presented the evidence, we argue that the evidence supports a view where the benefits that vague terms exert are due to other influences, rather than to vagueness itself.}
-
-\begin{document}
-\maketitle   
-\setcounter{secnumdepth}{3}
-\section{Introduction}\label{introduction}
-
-%Vagueness pervades the language that we use on a daily basis, and the challenge of understanding vague language has been a prominent concern in many areas of logic and linguistics, involving both theoretical and applied work, including the area known as Natural Language Generation (NLG).
-
-%NLG systems can be thought of a small and imperfect models of human language use, and as such they have the ability to make questions about language very tangible, so let us take the perspective of an NLG system for a moment.
-In a nutshell, NLG systems transform data and formulas into language \cite<e.g.,>[]{reiter2000building}, making decisions between different formulations of the same information.
-For example, if the temperature is 27.2 degrees Celsius, the system could express this as ``27.2 degrees'', ``approximately 27 degrees', ``above 25 degrees", or ``warm'', and the system must decide between these.
-The problem is important in the generation of textual weather reports from numerical weather data such as temperature and wind speed \cite{goldberg1994using,turner2006generating}, and medical decision support on the basis of clinical measurements % such as oxygen saturation, heart rhythm, etc.
-%\cite{Hripcsak01032009, hunter2008summarising, portet2009automatic}.
-Such systems often make decisions concerning the level of precision in the utterances that they generate on the basis of little more than intuition.
-Even when NLG systems are designed to mimic human language use \cite<e.g.,>{konstas2013global} there is no guarantee that these decisions taken by these systems benefit readers.
-A better understanding of the benefits (for readers and hearers) of different precision levels would allow these systems to become more useful.
-%The present article investigates the benefits, or de-benefits, of vagueness.
-
-%.
-In the context of an NLG system faced with a practical decision of this kind, Lipman's question becomes ``Under what circumstances should vague terms be produced?''  Relevant applications include weather forecasting
-
-%Language use may be called vague for various reasons.\footnote{See e.g. the entry ``vague'' in \cite{penguin}.}
-
-In most academic use, the word `vagueness' has a specific meaning.
-Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1997vagueness}, also \citeA{EgreKlinedinst}).
-The crucial criterion is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise'' \cite[p.\ 1]{lipmanvague}.
-A typical example is the word ``tall'', as applied to people for example, because here is no precise, known height which separates those who are tall from those who are not.
-The crucial point is that ``tall'' admits borderline cases (i.e., people who may or may not count as tall), which are the hallmark of vagueness as we use the term.
-
-Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions \cite{Lipman:2000fk, lipmanvague}.
-By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss'' \cite<>[p.~1]{lipmanvague}.
-Lipman explains the point by means of a scenario in which a speaker describes a person to a hearer, who needs to identify that person in the arrivals hall of of an airport.
-In such a scenario, a precise description of the person's height (e.g., ``The person's height is 187.96 cm'') would be more useful than a vague one (``The person is tall'').
-Lipman uses this scenario to explain why standard game theory models of communication \cite<e.g.,>{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs.
-%The relevant conditions are, broadly speaking, that both interlocutors know all the relevant facts (e.g., both know the person's height precisely) and that the setting for the communication is co-operative.
-These conditions exclude deliberate deception, and rhetorical situations like political debate and advertising where the intention is to persuade the interlocutor to adopt some point of view, where the persuasion might not be intended to be in the addressee's best interest.
-
-Lipman argued that the efficiency loss resulting from vague expressions would be unlikely to have arisen unless there are advantages as well as disadvantages associated with vague expressions.
-Lipman asked, essentially, what these advantages might be.
-%, and how they might find a place in a game-theoretical explanation.
-In this paper, we focus on the first part of Lipman's question.
-Several tentative answers to Lipman's question have been offered  \cite<see>{van2009utility, vanDeemterBook}.
-Prominent among these answers is the idea that vague expressions are easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite<e.g.,>[]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
-For example, \citeA<>[p.\ 11]{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze''.
-We shall refer to this as the \emph{cost reduction} hypothesis.
-%The idea of vagueness as cost reduction can take various shapes, but the basic hypothesis is that it is easier for people to think in terms of loosely defined categories (such as ``quite a few'', or ``many'') than in terms of crisply defined ones (such as ``thirteen'', or ``237'').
-%This hypothesis predicts that when a vague expression can perform the same communicative task as a precise one, then, the things being equal, it is rational to choose the vague expression.
-The corollary for comprehension is that vague expressions should be understood more readily than precise expressions.
-
-%The cost reduction hypothesis is of direct relevance to psycholinguists interested in language comprehension, and additionally to psycholinguists interested in language production, for example in connection with the question of audience design \cite{Clark1982287}.
-For, to the extent that speakers and writers choose vague expressions over and above crisp ones because the former are easier to process for hearers than the latter, the cost reduction hypothesis suggests that speakers design their utterance for optimal benefit to their hearers -- out of altruism, so to speak.
-
-Charting the utility of vagueness is the attested aim of a small number of studies, but most of these have focussed on vagueness in a different sense, and focussing on different types of benefits for hearers.
-Two recent studies can illustrate both issues.
-
-In a series of studies of behaviour modification, \citeA<>{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss.
-In the weight loss study, participants were told that the study was designed to test the validity of a new (actually fictitious) health index, the HHI (Holistic Health Index).
-They were told that an ideal HHI score lies in the range of 45 to 55.
-In a longitudinal study, participants submitted their weight to a computer each week.
-Participants were told that two algorithms would be used to compute their HHI, and that the two might give different values initially, in which case the true score lay between the two values.
-In one condition, which the authors called the precise condition, the two algorithms gave the same score.
-In the other condition, which the authors called the vague condition, one algorithm added 3\% to the score while the other algorithm subtracted 3\% from the score, yielding a range of values whose midpoint was the same as the two values given in the precise condition.
-
-One group of participants was given HHI scores in the ideal range: for this group their weight loss did not differ depending on whether they were given vague or precise HHI values.
-However for the other group, who were given HHI scores outside the ideal range, their weight loss was significantly greater if they were given vague HHI scores than if they were given precise HHI scores.
-The authors explain the improvement in the vague condition for this group as resulting from the participants' freedom to think of themselves a positioned on one end of the range - the end closest to the ideal HHI scores.
-This ``illusion of proximity'' \cite<>[p.~4]{Mishra01042011} to the goal is argued to allow participants to generate positive expectancies that lead to behaviours that improve performance.
-In contrast, in the precise conditions, participants did not have this freedom of interpretation, and could not distort the information to bring about the beneficial \emph{illusion of proximity}.
-These results are interesting, and of obvious potential practical importance.
-We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \cite{keefe1997vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases.
-In the terminology of \citeA{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}.
-Furthermore, the experiments of \citeA{Mishra01042011} did not explore benefits in terms of processing cost, but in terms of long-term behaviour change.
-
-Similar issues arise from the work of \citeA{peters2009bringing}.
-The authors carried out a series of studies where participants were required to rate hospitals based on various sources of information about quality of care.
-There was a between-subjects manipulation based on numeracy.
-The format of the information was manipulated within subjects: either numbers only were presented, or both numbers and evaluative categories were presented (e.g., \emph{Poor}, \emph{Fair}, \emph{Good}, \emph{Excellent}, with crisp visual boundary lines between the categories).
-Results showed that, for low-numeracy participants, the presence of evaluative categories resulted in a diminished influence of an irrelevant affective state on the ratings.
-For all participants, the presence of evaluative categories resulted in better decisions and in a greater use of the most important and reliable types of information, such as survival rates.
-
-It is, however, questionable whether the ``evaluative categories" manipulation in this study can be considered a manipulation of vagueness.
-Certainly, terms like \emph{Fair} admit the possibility of borderline cases.
-However, given that the boundaries between the categories were marked crisply, and that therefore the categories mapped crisply to numerical values, it becomes doubtful whether any borderline cases could be conceived to arise in fact.
-For example, \emph{Fair} was mapped to 60\% -- 70\% for the variable \emph{percentage of heart attack patients given recommended treatment (ACE inhibitor)}.
-Accordingly, rather than the vagueness of categories such as \emph{Poor}, Peters et al.
-emphasise the evaluative content inherent in these categories, and the affective potential of the evaluative content rather than the vagueness of the terms like {\em Fair}.
-
-\subsection{Our Approach to the Problem}\label{method}
-%The experiments reported in the present paper put the cost reduction hypothesis to the test. The question that we are trying to answer is whether
-
-Are vague expressions processed more easily by readers than crisp ones?
-Like Lipman, we focus on situations where numerical information is used in order to identify a referent.
-Reference, in other words, will be the communicative task on which we focus, partly because of the interest that this topic has recently drawn in various areas of Cognitive Science \cite{vanDeemterCMR}.
-By looking at one specific type of vagueness, we will be able to investigate the costs and benefits of vagueness relatively thoroughly.
-Whether our findings generalise to other uses of vagueness is a question on which we will speculate in the final section of this chapter.
-
-% KvD Feb 2017 (to address a criticism that came up in the reviews for the special issue):
-We have chosen a narrative strategy in which we address a sequence of four experiments with human reader chronologically, explaining how each experiment helped us refine our research question.
-In order to do justice to our findings, we need to describe these experiments in a fair amount of detail.
-
-Let us start by explaining the task that was given to the participants in our experiments.
-We used a {\em speeded forced choice} task to compare the processing costs of different references to quantities.
-In this context, speed and accuracy of responses are the key dimensions on which the different references can be compared.
-Each stimulus in the experiments was a set of dot arrays containing various number of dots, together with a preceding instruction (in the form of a referring expression) to choose one of the arrays with respect to its cardinality.
-The participant was asked to respond as quickly as possible while avoiding errors.
-We manipulated the instructions and the arrays in several ways across the four experiments.
-
-% MG General methodology moved to here: per-experiment method sections only talk about deviations from this general method
-All the experiments shared the following properties: 
-Stimuli were created using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.
-The position of the dots was randomised per-trial.
-The order in which trials were presented was randomised per-participant. 
-There were 256 trials, presented in 4 blocks of 64 each, between which the participant could rest. 
-A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants and recorded responses. 
-Participants were recruited using email lists at the University of Aberdeen, and paid ten pounds for participating.
-All participants self-reported fluency in English, and had normal, or corrected-to-normal vision.
-The experiment was conducted in a quiet room. 
-Participants were asked to respond as quickly as possible while avoiding errors.
-There was a block of practice trials after which participants could ask any questions, following which the experimenter left the room.
-All $p$ values reported for linear models were calculated using the R package \emph{lmerTest} \cite{lmerTest}. The complete data and analysis of our first experiment are available from 
-{\tt http://homepages.abdn.ac.uk/k.vdeemter/pages/vagueness-data-analysis/e1.html}, and analogous for experiments 2-4.\\[2ex]
-
-% KvD I've moved the following two paragraphs towards the end of the section.
-% MG added/moved out of expt 1
-
-\noindent When the distance grows between two numbers, they become more easily distinguishable: the \emph{numerical distance effect} has been shown for comparing the cardinality of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene1996}.
-We manipulated the number of dots in each array such that some sets of arrays had smaller numerical distances and others had larger numerical distances.
-Where a number was mentioned in the instructions, it was always in the form of an Arabic numeral.
-% MG added/moved out of expt 1
-When two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}.
-We controlled which side the smaller number appeared on to avoid systematic influences of this effect.
-
-% KvD added the following from a previous version of the paper:
-There is abundant evidence \cite<e.g.,>{trick1994small} that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities.
-We performed a pilot experiment \cite{our-earlier-paper} in which we were able to confirm this finding in the experimental settings on which we are focussing in this paper.
-We found that, when participants were confronted with % MG changed the next line and added the footnote to remove the doubts we had about whether a dot array must have regular layout, or can have irregular layout - I'm still not sure this is a good description though
-a stimulus consisting of two squares containing different numbers of dots\footnote{Such a stimulus is referred to hereafter as consisting of a set of {\em dot arrays}.
-The number of dots in an array is referred to as its cardinality.
-The physical arrangement of dots in each array is irregular.},
-instructions of the form \emph{Choose the square with \emph{n} dots} led to consistently faster response times than instructions of the form \emph{Choose the square with many/few dots} when $2 \leq n \leq 5$; the converse was true for $n>5$.
-Given these findings, we henceforth focussed our studies on non-subitizable numbers, because it is there that vagueness is expected to have benefits.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 1                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 1: initial experiment}
-
-We used a forced choice task to elicit responses to crisp or vague forms of instructions that required the participant to choose one of two dot arrays by referring to its cardinality.
-The participant was presented with an instruction in the form \emph{Choose the square with \ldots dots}.
-Then a set of two dot arrays was presented, each in the form of a square containing some number of dots.
-The participant was required to identify the dot array that corresponded with the instruction, by pressing the appropriate key, as quickly as possible while avoiding errors. 
-Response time and accuracy were recorded for analysis.
-
-We manipulated how discriminable the dot arrays were by varying the numerical distance between them. 
-One array always contained 25 dots: the other contained either 5, 10, 15, 20, 30, 35, 40, or 45 dots. 
-This gave us numerical distances of 5, 10, 15, and 20, with smaller numerical distances resulting in less discriminable arrays and larger distances resulting in more discriminable arrays.
-
-Our main manipulation was of the vagueness of the instruction, with two levels, \emph{crisp} and \emph{vague}.
-Assuming the dot array [5, 25], and the instruction referring to the smaller cardinality, the \emph{crisp} instruction was \emph{Choose the square with 5 dots} and its \emph{vague} counterpart was \emph{Choose the square with few dots}.
-
-\subsection{Hypothesis}
-\noindent The main aim of this experiment was to test the following hypothesis:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] There exists a main effect advantage for vagueness: vague instructions impose a lower cognitive load for the comprehender than crisp alternatives.
-\end{APAenumerate}
-}
-
-\subsection{Experimental Method}
-
-%Twenty participants were recruited, aged between 18 and 45 with a median age of 26.
-On each trial a participant was presented with an instruction to choose one of two dot arrays on screen by reference to its cardinality. 
-Following a keypress to indicate that the participant had read the instruction there was a central fixation cross for 1000 ms, and a blank screen for 500 ms, followed by the array (without repetition of the referring expression).
-The arrays would stay on screen until the participant responded (there was no timing-out). 
-Response time was measured as the latency between the presentation of the arrays, and the keypress identifying the choice: in this way, the response time was separated from time spent reading the instructions, which is important since we are only interested in the former.
-A response was counted as erroneous if the square with the wrong number of dots was chosen (when the instruction contained a number); 
-if the square with the larger number of dots was selected (when the instruction was \emph{Choose the square with few dots}); 
-or if the square with the smaller number of dots was selected (when the instruction was \emph{Choose the square with many dots}). 
-No feedback was given on correct trials, but there was feedback on error trials in the form of the word ``\textsc{wrong!!}'' which flashed on screen.
-Hypothesis H1 was tested looking both at response times and at error rates.
-
-\subsection{Results}
-
-\begin{figure}[htbp]
-\fitfigure{images/e1RTandErrorRatesplot-1}
-\caption{Experiment 1 results: response times and error rates.}
-\label{resultse1}
-\end{figure}
-
-%\subsubsection{Response times}
-{\bf Response times} (RTs) for trials with erroneous responses were discarded, leading to the loss of 354 trials from 5120, representing 6.9\% of the trials. 
-The correct response RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of a further 160 trials, or 3.4\% of the remaining correct responses. 
-Means for response times and error rates are given in Fig. (\ref{resultse1}).
-A linear mixed model of RT was built, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
-Analysis confirms hypothesis H1:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1$_{rt}$)] RTs were faster for vague instructions than for crisp instructions 
-	($\beta=-0.11$, $se=.02$, $t=4.7$, $p<.001$).  % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-%\subsubsection{Error rates}
-{\bf Error rate} data were analysed using a generalized logit mixed model \cite{jaeger2008categorical}, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
-We found:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1$_{er}$)] There was a statistically non-significant effect of vagueness on error rates: the vague instructions tended to have a lower error rate than the than the crisp instructions 
-	($\beta=-.22$, $se=.18$, $t=-1.2$, $p=.230$).  % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-\subsection{Discussion}
-
-\noindent As expected, responses were faster and more accurate for vague than crisp instructions. Post hoc, we formulated some further hypotheses:
-
-{\small
-\begin{APAenumerate}
-	\item [(H2)] A main effect advantage of increasing numerical distance: the task will become easier as the numerical distance increases, because the two arrays are then more discriminable.
-	\item [(H3)] An interaction between vagueness and numerical distance: i.e., any facilitation for vagueness should be greater at smaller numerical distances than at larger numerical distances.
-\end{APAenumerate}
-}
-
-Looking at response times, these two post-hoc hypotheses were supported:
-
-{\small
-\begin{APAenumerate}
-	\item [(H2$_{rt}$)] RTs increased with numerical distance 
-	($\beta=-0.18$, $se=0.01$, $t=-11.6$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-	Since discriminability of the arrays is easier for larger numerical distances, discriminability probably underlies this effect. 
-	\item [(H3$_{rt}$)] Numerical distance and vagueness interacted significantly: essentially there were diminishing returns for vagueness as numerical distance increased: 
-	($\beta=0.07$, $se=0.01$, $t=5.9$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-Looking at error rates, the two post-hoc hypotheses were supported as well:
-
-{\small
-\begin{APAenumerate}
-	\item [(H2$_{er}$)] Error rates decreased as numerical distance increased: 
-	($\beta=-0.78$, $se=0.08$, $z=-9.6$, $p<0.0005$).% MG added the values for these stats on 25 Sept 2016
-	\item [(H3$_{er}$)] Numerical distance and vagueness interacted significantly: essentially there were diminishing returns for vagueness as numerical distance increased until the biggest numerical distance when vagueness was disadvantageous:
-	($\beta=0.80$, $se=0.15$, $z=5.27$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-So far, these results are in line with the idea of cost reduction.
-The idea of cost reduction explains the vagueness advantage by claiming that the vague referring expressions place less cognitive load on the comprehender than the crisp referring expressions.
-It explains the diminishing returns for vagueness in more-discriminable stimuli (i.e., the vagueness by numerical distance interaction) by claiming that load is low in both conditions for the easily-discriminable stimuli, and that therefore there is no extra benefit to be had from vagueness in the easily-discriminable stimuli.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 2                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 2: separating vagueness from instruction format}
-
-% KvD: I've added some syntactic sugar, in a number of places, to clarify the relation between Ex 1 and Ex 2.
-
-However, the picture painted by these findings might be misleading.
+%However, the picture painted by these findings might be misleading.
 First of all, there is a possibly confounding factor.
 Contrast, from Experiment 1, an expression from the vague condition: `the square with few dots' with an expression from the crisp condition: `the square with 5 dots'.
-One difference is that `few' has the potential for vagueness, whereas `5' is crisp. But
-another difference is that `few' is verbal while `5' is numerical, in the sense that a number is mentioned explicitly. 
+One difference is that `few' has the potential for vagueness, whereas `5' is crisp.
+But another difference is that `few' is verbal while `5' is numerical, in the sense that a number is mentioned explicitly.
 Since these two differences could not be separated in Experiment 1, the vagueness advantage finding is vulnerable to an alternative interpretation, that what we saw as a vagueness advantage was in contrast an advantage for the verbal form of the quantifier.
@@
 
-Another potential problem with Experiment 1 is the following.
+%Another potential problem with Experiment 1 is the following.
 Participants chose one of two squares: therefore the `vague' quantifiers (e.g., `few') uniquely identified one square.
@@
 
-% KvD I've re-instated the table below, because this makes it easier to understand the instruction format/selection algorithm confound.
-
-\begin{table}[htbp]
-\caption{Experiment 1: instructions arranged by condition for the dot triple $(6,15,24)$ and instructions indicating the smaller numbers}
-\label{instre3}
+\begin{table}
+\centering
+\caption{Experiment 2: instructions arranged by condition for the dot triple $(6,15,24)$ and instructions indicating the smaller numbers}
+\label{instructionse2}
 \begin{tabular}{rll}
+\hline\noalign{\smallskip}
 &crisp&vague\\
-\midrule
+\noalign{\smallskip}\hline\noalign{\smallskip}
 numerical	&	Choose the square with 6 dots 			& 	Choose a square with about 10 dots	\\
 verbal		&	Choose the square with the fewest dots	&	Choose a square with few dots		\\
-\midrule
+\noalign{\smallskip}\hline
 \end{tabular}
@@
 
-\subsection{Hypotheses} % E2
-\noindent We formulated the following hypotheses for Experiment 2:
+\subsection{Hypotheses}
 
-{\small
-\begin{APAenumerate}
-	\item [(H1)] There exists a main effect RT advantage for vagueness.
-	\item [(H2)] An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
-	\item [(H3)] No large main effect of instruction format (i.e., numeric versus verbal) since we hypothesised that vagueness rather than instruction format drove the effect in Experiment 1.
-	\item [(H4)] On the basis of Experiment 1, we expect faster responses for stimuli with more discriminable arrays.(i.e., an effect of item).
-	\item [(H5)] Participants should make more borderline case choices for vague than crisp instructions.
-\end{APAenumerate}
-}
+We formulated the following hypotheses for Experiment 2:
 
-\subsection{Experimental Method} % E2
+\begin{description}
+	\item [Hypothesis 1] There exists a main effect RT advantage for vagueness.
+	\item [Hypothesis 2] An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
+	\item [Hypothesis 3] No large main effect of instruction format (i.e., numeric versus verbal) since we hypothesised that vagueness rather than instruction format drove the effect in Experiment 1.
+	\item [Hypothesis 4] On the basis of Experiment 1, we expect faster responses for stimuli with more discriminable arrays.(i.e., an effect of item).
+	\item [Hypothesis 5] Participants should make more borderline case choices for vague than crisp instructions.
+\end{description}
+
+\subsection{Experimental Method}
+
 % Participants were aged between 18 and 45 with a median age of 28. 
@@
 
-\subsection{Results} % E2
+\subsection{Results}
 
-%\subsubsection{Response times} % E2
-% MG added the values for these stats on 26 Sept 2016
 Means for response times and proportion of borderline responses are given in Fig. (\ref{resultse2}). 
@@
 with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects. We found:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against Hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06$, $se=.01$, $t=4.6$, $p<0.0001$).
-	\item [(H2)] In focussed comparisons, vagueness was significantly disadvantageous in the \emph{numeric} conditions ($\beta=0.09$, $se=0.02$, $t=4.5$, $p<0.0001$), and non-significantly disadvantageous in the \emph{verbal} conditions ($\beta=0.02$, $se=0.02$, $t=1.5$, $p=0.1550$), offering partial evidence against Hypothesis 2.
+
+\begin{description}
+	\item [Test of Hypothesis 1] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against Hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06$, $se=.01$, $t=4.6$, $p<0.0001$).
+	\item [Test of Hypothesis 2] In focussed comparisons, vagueness was significantly disadvantageous in the \emph{numeric} conditions ($\beta=0.09$, $se=0.02$, $t=4.5$, $p<0.0001$), and non-significantly disadvantageous in the \emph{verbal} conditions ($\beta=0.02$, $se=0.02$, $t=1.5$, $p=0.1550$), offering partial evidence against Hypothesis 2.
 The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to a significant overall interaction effect between vagueness and instruction format ($\beta=-0.13$, $se=0.02$, $t=-6.6$, $p=0.0169$).
-	\item [(H3)] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1, but suggesting that instruction format rather than vagueness drove the effect we observed in Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; ($\beta=-0.36$, $se=0.07$, $t=-5.1$, $p<0.0001$).
-	\item [(H4)] There was a significant main effect of item (i.e., of which triple of numbers of dots were used in the stimulus): ($\beta=0.12$, $se=0.02$, $t=7.1$, $p<0.0001$).
+	\item [Test of Hypothesis 3] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1, but suggesting that instruction format rather than vagueness drove the effect we observed in Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; ($\beta=-0.36$, $se=0.07$, $t=-5.1$, $p<0.0001$).
+	\item [Test of Hypothesis 4] There was a significant main effect of item (i.e., of which triple of numbers of dots were used in the stimulus): ($\beta=0.12$, $se=0.02$, $t=7.1$, $p<0.0001$).
 This effect seems likely to be due to the very fast responses for stimuli using arrangements of the smallest numbers of dots (6,15,24), which also had the largest difference in ratio of smallest number to largest number in the stimulus, suggesting that stimuli using these numbers of dots may have been particularly discriminable for participants.
-\end{APAenumerate}
-}
+\end{description}
 
-%\subsubsection{Borderline responses} % E2
-% MG added the values for these stats on 26 Sept 2016
-A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
+A generalized linear mixed model \citet{jaeger2008categorical} was fit to the data for selection of the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
 The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{resultse2}.
 Participants chose the borderline square on $16.6\%$ of trials overall.
-{\small 
-\begin{APAenumerate}
-	\item [(H5)] Participants were significantly more likely to choose the borderline option for vague instructions than for crisp instructions (21.9\% vs 11.3\%: $\beta=0.62$, $se=0.22$, $z=2.8$, $p=0.0059$).
+
+\begin{description}
+	\item [Test of Hypothesis 5] Participants were significantly more likely to choose the borderline option for vague instructions than for crisp instructions (21.9\% vs 11.3\%: $\beta=0.62$, $se=0.22$, $z=2.8$, $p=0.0059$).
 Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%: $\beta=-3.35$, $se=0.23$, $z=-14.6$, $p<0.0001$).
-\end{APAenumerate}
-}
+\end{description}
 
@@
 \centering
-\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{images/e2-rtplot-1.pdf}
-\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{images/e2-blBarChart-1}
+\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{figures/e2-rtplot-1.pdf}
+\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{figures/e2-blBarChart-1}
 \caption{Experiment 2 results: mean response times by condition and item, and counts of borderline case responses by condition.}
@@
 
-\subsection{Discussion} % E2
+\subsection{Discussion}
 
@@
 
-% KvD I've expanded the following:
 However, once again there is a possibly confounding factor.
-Observe that, in Experiment 2, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) {\bf selection algorithm}: To see this, consider the task of selecting the dot array that contains ``few dots": to do this, it suffices to {\em compare} the three arrays and select the one that contains the fewest elements.
-To select the dot array that contains "16 dots" seems to require the participant to estimate, and then {\em match}, the cardinality of (at least) one dot array to 16, a process which could plausibly take longer, independently of vagueness.
-Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison ``algorithm".% (which is known to be faster than matching).
-% KvD I think we need to say more about this, citing some appropriate papers to support this idea.
-\textcolor{red}{{\bf--Insert citation}}
+Observe that, in Experiment 2, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) {\bf selection algorithm}: To see this, consider the task of selecting the dot array that contains ``few dots'': to do this, it suffices to {\em compare} the three arrays and select the one that contains the fewest elements.
+To select the dot array that contains ``16 dots'' seems to require the participant to estimate, and then {\em match}, the cardinality of (at least) one dot array to 16, a process which could plausibly take longer, independently of vagueness.
+Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison ``algorithm''.
 
@@
 This allowed us to distinguish between the predictions of the selection task account and the instruction format account.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 3                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 3: focus on instructions that contain numerals}
-
-The main aim of Experiment 3 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies.
-The main changes from Experiment 2 were that the human selection task was explicitly controlled (i.e., in whether it amounted to matching or comparison), and that all conditions were constrained to mention a number.
-We used the same arrays as in Experiment 2.
-Table \ref{Instructions for e3} shows the instructions for each condition.
-Note the difference between ``fewer than 20" and ``far fewer than 20": whereas the former cannot have borderline cases (i.e., for each number it is clear whether the number is smaller than 20 or not), the latter can.
-
-%KvD I think we should motivate our choice of materials (especially "far ..er")
-%\textcolor{red}{{\bf KvD I think we should motivate our choice of materials (especially "far ..er")}}
-% DONE 28 Sept 2017
-
-\begin{table}[htbp]
-\centering
-\caption{Experiment 3: Instructions arranged by condition for stimuli with $(6,15,24)$ dots and instructions indicating the smaller numbers} 
-\label{Instructions for e3}
-\begin{tabular}{llll}
-instruction				&vagueness			&	selection		&	instruction							\\
-format				&					&	task			&									\\
-\toprule
-\multirow{4}{*}{numeric}	&\multirow{2}{*}{crisp} 	&	matching		& Choose a square with 6 dots 				\\ 
-					&					&	comparison 	& Choose a square with fewer than 20 dots 	\\
-					\cline{2-4}
-					&\multirow{2}{*}{vague} 	&	matching 		& Choose a square with about 10 dots 		\\ 
- 					&					&	comparison 	& Choose a square with far fewer than 20 dots 	\\ 
-\bottomrule
-\end{tabular}
-\end{table}%
-
-\subsection{Hypotheses} % E 3
-\noindent For Experiment 3, we hypothesised:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
-	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection task)
-	\item [(H3)] Effects of vagueness are different depending on whether the selection task is matching or comparison (interaction effect selection x vagueness).
-\end{APAenumerate}
-}
-
-\subsection{Experimental Method} % E 3
-
-38 participants were recruited. We used a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e3}).
-On each trial an instruction was presented: participants pressed a key to dismiss the instruction, when the dot arrays were presented until the participant responded, and the response time and choice were recorded.
-
-\subsection{Results} % E 3
-
-Response times were trimmed at 2.5 SD separately for each subject, leading to the loss of 204 trials (2.8\% of the trials).
-Condition means for the remaining (logged) RTs are plotted in Figure \ref{resultse3}.
-A linear mixed model was constructed for the logged response times, 
-with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
-
-\begin{figure}[htbp]
-\centering
-\fitfigure{images/e3-rtplot-1.pdf}
-%\includegraphics[trim = 20mm 0mm 20mm 0mm, clip, width=.75\textwidth]{images/e3-rtplot-1.pdf}
-\caption{Mean response times by condition for Experiment 3 where all instructions were numeric}
-\label{resultse3}
-\end{figure}
-
-% MG added the values for these stats on 26 Sept 2016
-{\small
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions were non-significantly easier for the reader than crisp instructions when vagueness was considered as a main effect ($\beta=-0.01$, $se=0.01$, $t=-0.4$, $p=0.7158$). 
-	\item [(H2)] There was a statistically significant main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=0.16$, $se=0.03$, $t=6.2$, $p<0.0001$).
-	\item [(H3)] Vagueness exerted effects in different directions for the comparison task and for the matching task: there was a significant interaction between vagueness and selection task ($\beta=0.13$, $se=0.03$, $t=4.2$, $p<0.0001$).
-Separate analyses were conducted testing for effects of vagueness at each level of the selection task.
-Within the comparison task vagueness significantly speeded response times compared with crisp controls ($\beta=-0.07$, $se=0.02$, $t=-3.5$, $p<0.0012$). 
-Within the matching task vagueness significantly \emph{slowed} response times compared with crisp controls ($\beta=0.06$, $se=0.02$, $t=2.9$, $p<0.0061$). 
-\end{APAenumerate}
-}
-% MG edited this 27 Sept 2016:
-%The cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
-In other words, the cost reduction account was wrong to predict significant main effect advantages for vagueness (although there was a non-significant trend in the \emph{direction} predicted by the cost reduction account), and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was significantly advantageous in the comparison task.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 4                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 4: focus on instructions that do not contain numerals} % E 4
-
-This experiment mirrors Experiment 3, but focussing on instructions that did not use a number. 
-We manipulated vagueness and the selection task (comparison and matching). 
-In order to implement the experiment without mentioning numbers in the instructions, we changed the sequence of each trial to include a `target' (i.e., a dot array of a particular cardinality) before the instruction, so that we could then refer back to the target's cardinality in the instruction using expressions like \emph{the same number of dots as the target}; \emph{fewer dots than the target}.
-%
-This presentation of a target before the main body of the trial shares some features with \citeA[Experiment 2]{Izard20081221}, although in that experiment participants were told the cardinality of the target (called an \emph{inducer} in that paper) whereas in our experiment we did not tell participants the cardinality of the prime array.
-An item was thus a combination of a target dot array, an instruction that did not contain a number, and a set of dot arrays taking their cardinalities from the same triples used in Experiment 2.
-Table \ref{Instructions for e4} spells out how the instructions were constrained not to mention a numeral and gives examples of targets.
-
-\begin{table}[htp]
-\caption{Instructions and targets by condition for Experiment 4}
-\begin{center}
-\begin{tabular}{llcrl}
-Item					&Selection				&Vagueness	&Target	&Choose a square with\ldots						\\
-\toprule
-\multirow{4}{*}{06:15:24} 	&\multirow{2}{*}{Comparison} 	&Crisp		&20		&\ldots fewer dots than the target					\\
-\cline{3-5}
-					&						&Vague		&20		&\ldots far fewer dots than the target					\\
-					\cline{2-5}
-					&\multirow{2}{*}{Matching}		& Crisp		&6		&\ldots the same number of dots as the target			\\
-					\cline{3-5}
-					& 		 				&Vague		&10 		&\ldots about the same number of dots as the target		\\
-					\bottomrule
-\end{tabular}
-\end{center}
-\label{Instructions for e4}
-\end{table}%
-
-\subsection{Hypotheses}% E 4
-\noindent For Experiment 4, we hypothesised:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
-	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection)
-	\item [(H3)] Effects of vagueness differ depending on whether selection is matching or comparison (interaction effect selection x vagueness, and focussed comparisons at each level of selection).
-\end{APAenumerate}
-}
-
-\subsection{Results}% E 4
-40 volunteers participated.
-The results showed that vagueness was beneficial for comparison but detrimental for matching (the same as Experiment 3) even when no numbers were allowed in the instructions. 
-Figure \ref{resultse4} shows the means by condition. Our findings were as follows:
-
-% MG added the values for these stats on 27 Sept 2016
-{\small
-\begin{APAenumerate}
-	\item [(H1)] There was no significant main effect of vagueness ($\beta=-0.02$, $se=0.01$, $t=-1.5$, $p=0.1296$). 
-	\item [(H2)] There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=0.18$, $se=0.02$, $t=10.4$, $p<0.0001$).
-This effect was in the same direction as Experiment 3.
-	\item[(H3)] Vagueness did exert different effects depending on the selection task (main interaction effect of vagueness by selection $\beta=0.12$, $se=0.02$, $t=5.1$, $p<0.0001$). 
-Separate analyses of the effect of vagueness were conducted for the comparison task and for the matching task using Bonferroni-adjusted significance thresholds. 
-In the comparison task, vagueness resulted in faster response times ($\beta=-0.08$, $se=0.02$, $t=-4.3$, $p<0.0001$). 
-In the matching task vagueness slowed response times ($\beta=0.05$, $se=0.01$, $t=3.7$, $p=0.0004$). 
-These results are in the same direction as Experiment 3.
-\end{APAenumerate}
-}
-
-Once again, the cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
-
-\begin{figure}[htbp]
-\centering
-\fitfigure{images/e4-rtplot-1.pdf}
-\caption{Mean response times by condition for Experiment 4}
-\label{resultse4}
-\end{figure}
-
-\section{Discussion of experiments 3 and 4}
-
-The main aim of these two experiments was to test whether vagueness confers any cognitive benefits over and above those due to differences in the selection task according to whether the instruction mandates a \emph{comparison} selection task or a \emph{matching} selection task, when number-use is held constant.
-The main effect of selection task showed that the assumption that the \emph{comparison} task is easier than the \emph{matching} task is well-founded.
-In both experiments people were reliably quicker to respond in the \emph{comparison} task.
-
-Vagueness, which was the phenomenon on which our investigation focussed, did not exert a significant main effect in response time.
-However when the comparison and selection tasks were analysed separately, there was small significant advantage for vagueness in the \emph{comparison} tasks, but a small significant disadvantage for vagueness in the \emph{matching} tasks.
-
-\section{General Discussion}\label{discussion}
-
-To summarise our findings, Experiment 1 showed that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment could not distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format; and mandated a comparison strategy rather than a matching strategy.
-Experiment 2 showed that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional benefit in either the verbal format instructions or the numerical format instructions.
-
-However, Experiment 2 did not distinguish benefits of number avoidance from benefits of the comparison selection task.
-In Experiments 3 and 4 we manipulated vagueness and the selection task, separately at each level of numerical format.
-Across the two experiments, we found that the comparison-task instructions attracted faster response times than the matching-task instructions.
-Within the two experiments we found that vagueness exerts benefits when the selection task is \emph{comparison}, but not when the task is \emph{matching}.
-
-\begin{table}[htbp]
-\caption{Vagueness as range reduction: a summary of Experiments 3 and 4}
-\label{Vagueness as range reduction}
-\centering
-\begin{tabular}{cccc}
-selection task 				& vagueness		& candidates	& effect of vageueness	\\
-\toprule
-\multirow{ 2}{*}{comparison} 	& crisp 			& 2			& \multirow{ 2}{*}{vagueness advantage}  		\\
-\cline{2-3}
-  						& vague			& 1			&                                               				\\
-\midrule
-\multirow{ 2}{*}{matching} 		& crisp 			& 1 			&\multirow{ 2}{*}{vagueness disadvantage}	\\
-\cline{2-3}
- 						& vague			& 2			&								 	\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-What is one entitled to conclude?
-Given that we were able to identify a class of situations in which vague expressions led to faster response times than crisp ones, would it be valid to conclude that we have finally discovered an advantage for vagueness that cannot be ascribed to some other factor?
-We believe the answer to this question is negative.
-
-To see why, consider Figures \ref{resultse3} and \ref{resultse4}.
-Both figures depict four conditions, depending on whether the expression was crisp or vague, and depending on whether the referent could be identified using a comparison strategy or not.
-Two of the resulting four conditions result in an expression that can denote either of two referents; the other two conditions result in an expression that can only denote one referent, with the other possible referent being a marginal candidate at best.
-
-% KvD: I've shortened the text, removing duplication
-
-To see why vagueness has opposite effects, depending on whether it is used in matching or comparison situations, consider the stimulus with (6,15,24) dots.
-Now compare `Choose a square with 6 dots' with its vague counterpart `Choose a square with about 10 dots': by adding the word `about', we broaden the range of squares that the expression might be referring to.
-On the other hand, compare `Choose a square with fewer than 20 dots' with `Choose a square with far fewer than 20 dots': by adding the word `far', we did not broaden the range of squares denotable by the expression: we narrow it down, because only some of the squares that have fewer dots may have {\em far} fewer dots.
-
-The benefits of vagueness in the \emph{comparison} task in experiments 3 and 4 could thus be explained as differences in the number of valid targets for the expression.
-This leads us to speculate that the benefit for vagueness here could be due to the vague expression foregrounding a particular valid target while the crisp expression carries with it the additional task of distinguishing between two alternative valid targets, something we propose to call a ``range-reduction'' benefit.
-
-The observation that conditions with 1 candidate lead to shorter response times than conditions with 2 candidates is consistent with the range reduction hypothesis, but not with the idea that vagueness is beneficial.
-It appears, in other words, that shorter response times will only result from a vague expression if this expression leads to range reduction.
-Once again, it is not vagueness itself that has advantages but a phenomenon (namely range reduction) that is an automatic concomitant of vagueness in some types of situations.\\[2ex]
-%
-\noindent Looking at the entire series of experiments, our findings suggest that the observed benefits of vague expressions in certain situations may be due to factors other than vagueness: factors like avoiding numbers; permitting comparison tasks; and range reduction.
-The picture that is starting to emerge is subtle: on the one hand, in the situations that we have been studying
-%-- where cooperative speakers refer to an object (e.g., a square) by means of some quantity associated with the object -- 
-vagueness is not intrinsically beneficial.
-On the other hand, vague expressions frequently possess other features that {\em are} beneficial, and these are what give us the incorrect impression that vagueness itself is beneficial.
-Vagueness may thus have acquired a reputation that it does not deserve.
-The answer to Lipman's question, of why vagueness permeates human language (see our Introduction), may lie in a different direction after all, possibly relating to benefits for the speaker rather than the hearer.
-% \cite<see>[chapter 11, for discussion of some possible directions]{vanDeemterBook}.
-
-A comparison may clarify the logic of the situation.
-In recent years a number of studies, focussing on red wine, have suggested that alcohol, consumed in low doses, may have health benefits.
-An alternative explanation, however, asserts that it is not the alcohol in the wine that was beneficial, but antioxidants from grapes.
-If this alternative explanation is correct, then alcohol may not be healthy after all.\\[2ex]
-%
-% KvD Feb 2017. I've diminished the role of NLG in the coming paragraph.
-Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
-Years of research on the logic of vagueness -- giving rise to such techniques as Partial Logic \cite<e.g.,>{Fine}, Probabilistic Logic \cite{Edgington}, and Fuzzy Logic \cite{Zadeh} -- have primed the research community to expect that some special utility of vagueness is an important part of the answer, but our findings call this expectation into question.
-
-Although our own studies in this article have focussed on vagueness in descriptive Noun Phrases, it seems plausible that vagueness plays a similar role in other linguistic constructs.
-For example, consider reports on air temperature.
-Given a numerical temperature measurement or prediction, we might word it as
-%
-\begin{quote}
-(a) {\em 27.2 degrees Celsius}, or\\
-(b) {\em approximately 27 degrees}, or\\
-(c) {\em above 25 degrees}, or\\
-(d) {\em warm}, 
-\end{quote}
-%
-among other candidate expressions.
-Which of these descriptions is most effective, for example as part of a weather report?
-If the linguistic literature is to be believed, then options (a) and (c) convey crisp information, whereas (b) and (d) are vague (i.e., they permit borderline cases).
-From our own experiments and the ones discussed in section 1, there is no evidence that  vagueness should be beneficial for hearers.
-Rather than asking whether a candidate expression is vague, other questions might shed more light on the choice, similar to the ones identified in our studies.
-These questions might focus on the amount of information that a given expression conveys (i.e., on granularity), on the avoidance of numbers, and on the use of evaluative terms.
-Let's see how this might pan out for the examples at hand.
-
-First, the experiments by \citeauthor{Mishra01042011} suggest that it is important how much information is conveyed by an expression, and their findings are echoed by our own thoughts about range reduction (following Experiments 3 and 4).
-In the case of (a)--(d) above, it appears that (a) conveys the most detailed information (designating the smallest segment of the temperature scale), followed by (b), then (d), then (c) (e.g., 40 degrees is above 25, but at 40 Celsius the word ``warm" is likely to give way to ``hot" or ``scorching"):
-%
-\begin{quote}
-$a < b < d < c$
-\end{quote}
-%
-If these hunches are correct then an {\em expert} may prefer to read expression (a), because it gives her the most detailed information on which to base her decisions.
-On the other hand, expression (d) (``warm") is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make (d) more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm" (see our section 1 and  \citeA{peters2009bringing}), which is especially important if the hearer is unfamiliar with the metric used.
-These considerations suggest that {\em non-experts} might prefer expression (d).
-%If results of this kind were to be confirmed by experiments, they would be applicable in practical NLG, though care would need to be exercised to take the purpose of the communication into account.\footnote{One reason why ``warm" might be less preferred than these considerations suggest is that, in the weather context, some very frequently occurring numbers, such as 20 (degrees), could be relatively easy to process, analogous to the subitizable numbers of \protect\citeA{trick1994small}, see our section \ref{method}.
-Note that in the experiments reported in this paper, the purpose of the utterances investigated was always clear (e.g., instructing people on a weight loss program \cite{Mishra01042011}; informing a patient's choice of hospital \cite{peters2009bringing}; or selecting a dot array (in our own studies).}
-
-One way to see why vagueness (as defined in our Introduction) may not benefit human communication is the following thought experiment.
-Suppose a group of speakers understand the word ``warm" as vague, agreeing that temperatures above 26 count as warm, and temperatures below 24 do not count as warm, but considering temperatures between 24 and 26 as borderline cases.
-Now one day these speakers agree to sharpen up their definition deciding that, henceforth, ``warm" means ``$>25$ degrees" (as in (c) above): this decision resolves the borderline cases, while everything else remains the same.
-It seems unlikely that this change in language use, from a vague meaning to a crisp one (i.e., one that has no borderline cases anymore), would lower the utility of the word.
-Our experimental findings are consistent with this idea.
-
-% KvD Bracketed out the following for now:
-%A more radical perspective on our findings might say that the distinction between crisp and vague expressions, which is taken for granted in the bulk of the linguistic literature, is mistaken, because expressions such as ``above 25 degrees" (and ``fewer than 20 dots") are not equally felicitous in all situations in which they can be truthfully applied.
-For example, when the temperature is 40 degrees, then a weather report saying that it is ``above 25 degrees" would be correct yet misleading: this description might be felicitous as a description of temperatures between (roughly) 26-28 degrees, borderline felicitous for 29-33 degrees, and infelicitous at even higher temperatures.
-At its most bold, this perspective would assert that the applicability of {\em every} linguistic expression has borderline cases, hence Lipman's question (see section 1) simply does not arise.
-This radical perspective can only be indicated here, but not discussed at length.
-
-\setcounter{secnumdepth}{0}
-\bibliography{v-special-issue-2016}
-\end{document}

### 2018-08-02 springer/current/e2.tex
massive restructuring Aug 2018

--- 2018-04-23 papers/2018-v-book/v-book-2017.tex
@@
-\documentclass[%
-%			% choose man or doc immediately below
-%doc,	% doc is single spaced
-man,		% man is double spaced
-floatsintext,%
-apacite%
-]{apa6}
-\usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{color}
-\usepackage{booktabs}
-\usepackage{layout}
-\usepackage{apacite}
-\usepackage{datetime}
-
-%\renewcommand{\baselinestretch}{1.7} % added 21 Sept 2016, copied from CRM book, but see documentclass option 'man' for "manuscript'' above which makes it double spaced
-
-% MG change p's to 3 dp
-
-\title{The Elusive Benefits of Vagueness \protect \\ Evidence from Experiments}
-\date{\today}
-%\authornote{latex compiled \today ~\currenttime}
-\shorttitle{Is Vagueness Beneficial for Hearers?}
-\twoauthors{Matt Green}{Kees van Deemter}
-\twoaffiliations{Computing Science\\ University of Aberdeen}{Computing Science\\ University of Aberdeen}
-
-\abstract{Much of everyday language is vague, yet the benefits of vagueness for hearers and readers are proving to be elusive.
-We discuss a range of earlier controlled experiments with human participants, and we report on a new series of experiments that we conducted in recent years.
-These experiments, which focus on vague expressions that are part of referential noun phrases, aim to separate the utility of vagueness (as defined by the existence of borderline cases) from the utility of other factors that tend to co-occur with vagueness.
-Having presented the evidence, we argue that the evidence supports a view where the benefits that vague terms exert are due to other influences, rather than to vagueness itself.}
-
-\begin{document}
-\maketitle   
-\setcounter{secnumdepth}{3}
-\section{Introduction}\label{introduction}
-
-%Vagueness pervades the language that we use on a daily basis, and the challenge of understanding vague language has been a prominent concern in many areas of logic and linguistics, involving both theoretical and applied work, including the area known as Natural Language Generation (NLG).
-
-%NLG systems can be thought of a small and imperfect models of human language use, and as such they have the ability to make questions about language very tangible, so let us take the perspective of an NLG system for a moment.
-In a nutshell, NLG systems transform data and formulas into language \cite<e.g.,>[]{reiter2000building}, making decisions between different formulations of the same information.
-For example, if the temperature is 27.2 degrees Celsius, the system could express this as ``27.2 degrees'', ``approximately 27 degrees', ``above 25 degrees", or ``warm'', and the system must decide between these.
-The problem is important in the generation of textual weather reports from numerical weather data such as temperature and wind speed \cite{goldberg1994using,turner2006generating}, and medical decision support on the basis of clinical measurements % such as oxygen saturation, heart rhythm, etc.
-%\cite{Hripcsak01032009, hunter2008summarising, portet2009automatic}.
-Such systems often make decisions concerning the level of precision in the utterances that they generate on the basis of little more than intuition.
-Even when NLG systems are designed to mimic human language use \cite<e.g.,>{konstas2013global} there is no guarantee that these decisions taken by these systems benefit readers.
-A better understanding of the benefits (for readers and hearers) of different precision levels would allow these systems to become more useful.
-%The present article investigates the benefits, or de-benefits, of vagueness.
-
-%.
-In the context of an NLG system faced with a practical decision of this kind, Lipman's question becomes ``Under what circumstances should vague terms be produced?''  Relevant applications include weather forecasting
-
-%Language use may be called vague for various reasons.\footnote{See e.g. the entry ``vague'' in \cite{penguin}.}
-
-In most academic use, the word `vagueness' has a specific meaning.
-Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1997vagueness}, also \citeA{EgreKlinedinst}).
-The crucial criterion is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise'' \cite[p.\ 1]{lipmanvague}.
-A typical example is the word ``tall'', as applied to people for example, because here is no precise, known height which separates those who are tall from those who are not.
-The crucial point is that ``tall'' admits borderline cases (i.e., people who may or may not count as tall), which are the hallmark of vagueness as we use the term.
-
-Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions \cite{Lipman:2000fk, lipmanvague}.
-By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss'' \cite<>[p.~1]{lipmanvague}.
-Lipman explains the point by means of a scenario in which a speaker describes a person to a hearer, who needs to identify that person in the arrivals hall of of an airport.
-In such a scenario, a precise description of the person's height (e.g., ``The person's height is 187.96 cm'') would be more useful than a vague one (``The person is tall'').
-Lipman uses this scenario to explain why standard game theory models of communication \cite<e.g.,>{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs.
-%The relevant conditions are, broadly speaking, that both interlocutors know all the relevant facts (e.g., both know the person's height precisely) and that the setting for the communication is co-operative.
-These conditions exclude deliberate deception, and rhetorical situations like political debate and advertising where the intention is to persuade the interlocutor to adopt some point of view, where the persuasion might not be intended to be in the addressee's best interest.
-
-Lipman argued that the efficiency loss resulting from vague expressions would be unlikely to have arisen unless there are advantages as well as disadvantages associated with vague expressions.
-Lipman asked, essentially, what these advantages might be.
-%, and how they might find a place in a game-theoretical explanation.
-In this paper, we focus on the first part of Lipman's question.
-Several tentative answers to Lipman's question have been offered  \cite<see>{van2009utility, vanDeemterBook}.
-Prominent among these answers is the idea that vague expressions are easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite<e.g.,>[]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
-For example, \citeA<>[p.\ 11]{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze''.
-We shall refer to this as the \emph{cost reduction} hypothesis.
-%The idea of vagueness as cost reduction can take various shapes, but the basic hypothesis is that it is easier for people to think in terms of loosely defined categories (such as ``quite a few'', or ``many'') than in terms of crisply defined ones (such as ``thirteen'', or ``237'').
-%This hypothesis predicts that when a vague expression can perform the same communicative task as a precise one, then, the things being equal, it is rational to choose the vague expression.
-The corollary for comprehension is that vague expressions should be understood more readily than precise expressions.
-
-%The cost reduction hypothesis is of direct relevance to psycholinguists interested in language comprehension, and additionally to psycholinguists interested in language production, for example in connection with the question of audience design \cite{Clark1982287}.
-For, to the extent that speakers and writers choose vague expressions over and above crisp ones because the former are easier to process for hearers than the latter, the cost reduction hypothesis suggests that speakers design their utterance for optimal benefit to their hearers -- out of altruism, so to speak.
-
-Charting the utility of vagueness is the attested aim of a small number of studies, but most of these have focussed on vagueness in a different sense, and focussing on different types of benefits for hearers.
-Two recent studies can illustrate both issues.
-
-In a series of studies of behaviour modification, \citeA<>{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss.
-In the weight loss study, participants were told that the study was designed to test the validity of a new (actually fictitious) health index, the HHI (Holistic Health Index).
-They were told that an ideal HHI score lies in the range of 45 to 55.
-In a longitudinal study, participants submitted their weight to a computer each week.
-Participants were told that two algorithms would be used to compute their HHI, and that the two might give different values initially, in which case the true score lay between the two values.
-In one condition, which the authors called the precise condition, the two algorithms gave the same score.
-In the other condition, which the authors called the vague condition, one algorithm added 3\% to the score while the other algorithm subtracted 3\% from the score, yielding a range of values whose midpoint was the same as the two values given in the precise condition.
-
-One group of participants was given HHI scores in the ideal range: for this group their weight loss did not differ depending on whether they were given vague or precise HHI values.
-However for the other group, who were given HHI scores outside the ideal range, their weight loss was significantly greater if they were given vague HHI scores than if they were given precise HHI scores.
-The authors explain the improvement in the vague condition for this group as resulting from the participants' freedom to think of themselves a positioned on one end of the range - the end closest to the ideal HHI scores.
-This ``illusion of proximity'' \cite<>[p.~4]{Mishra01042011} to the goal is argued to allow participants to generate positive expectancies that lead to behaviours that improve performance.
-In contrast, in the precise conditions, participants did not have this freedom of interpretation, and could not distort the information to bring about the beneficial \emph{illusion of proximity}.
-These results are interesting, and of obvious potential practical importance.
-We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \cite{keefe1997vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases.
-In the terminology of \citeA{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}.
-Furthermore, the experiments of \citeA{Mishra01042011} did not explore benefits in terms of processing cost, but in terms of long-term behaviour change.
-
-Similar issues arise from the work of \citeA{peters2009bringing}.
-The authors carried out a series of studies where participants were required to rate hospitals based on various sources of information about quality of care.
-There was a between-subjects manipulation based on numeracy.
-The format of the information was manipulated within subjects: either numbers only were presented, or both numbers and evaluative categories were presented (e.g., \emph{Poor}, \emph{Fair}, \emph{Good}, \emph{Excellent}, with crisp visual boundary lines between the categories).
-Results showed that, for low-numeracy participants, the presence of evaluative categories resulted in a diminished influence of an irrelevant affective state on the ratings.
-For all participants, the presence of evaluative categories resulted in better decisions and in a greater use of the most important and reliable types of information, such as survival rates.
-
-It is, however, questionable whether the ``evaluative categories" manipulation in this study can be considered a manipulation of vagueness.
-Certainly, terms like \emph{Fair} admit the possibility of borderline cases.
-However, given that the boundaries between the categories were marked crisply, and that therefore the categories mapped crisply to numerical values, it becomes doubtful whether any borderline cases could be conceived to arise in fact.
-For example, \emph{Fair} was mapped to 60\% -- 70\% for the variable \emph{percentage of heart attack patients given recommended treatment (ACE inhibitor)}.
-Accordingly, rather than the vagueness of categories such as \emph{Poor}, Peters et al.
-emphasise the evaluative content inherent in these categories, and the affective potential of the evaluative content rather than the vagueness of the terms like {\em Fair}.
-
-\subsection{Our Approach to the Problem}\label{method}
-%The experiments reported in the present paper put the cost reduction hypothesis to the test. The question that we are trying to answer is whether
-
-Are vague expressions processed more easily by readers than crisp ones?
-Like Lipman, we focus on situations where numerical information is used in order to identify a referent.
-Reference, in other words, will be the communicative task on which we focus, partly because of the interest that this topic has recently drawn in various areas of Cognitive Science \cite{vanDeemterCMR}.
-By looking at one specific type of vagueness, we will be able to investigate the costs and benefits of vagueness relatively thoroughly.
-Whether our findings generalise to other uses of vagueness is a question on which we will speculate in the final section of this chapter.
-
-% KvD Feb 2017 (to address a criticism that came up in the reviews for the special issue):
-We have chosen a narrative strategy in which we address a sequence of four experiments with human reader chronologically, explaining how each experiment helped us refine our research question.
-In order to do justice to our findings, we need to describe these experiments in a fair amount of detail.
-
-Let us start by explaining the task that was given to the participants in our experiments.
-We used a {\em speeded forced choice} task to compare the processing costs of different references to quantities.
-In this context, speed and accuracy of responses are the key dimensions on which the different references can be compared.
-Each stimulus in the experiments was a set of dot arrays containing various number of dots, together with a preceding instruction (in the form of a referring expression) to choose one of the arrays with respect to its cardinality.
-The participant was asked to respond as quickly as possible while avoiding errors.
-We manipulated the instructions and the arrays in several ways across the four experiments.
-
-% MG General methodology moved to here: per-experiment method sections only talk about deviations from this general method
-All the experiments shared the following properties: 
-Stimuli were created using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.
-The position of the dots was randomised per-trial.
-The order in which trials were presented was randomised per-participant. 
-There were 256 trials, presented in 4 blocks of 64 each, between which the participant could rest. 
-A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants and recorded responses. 
-Participants were recruited using email lists at the University of Aberdeen, and paid ten pounds for participating.
-All participants self-reported fluency in English, and had normal, or corrected-to-normal vision.
-The experiment was conducted in a quiet room. 
-Participants were asked to respond as quickly as possible while avoiding errors.
-There was a block of practice trials after which participants could ask any questions, following which the experimenter left the room.
-All $p$ values reported for linear models were calculated using the R package \emph{lmerTest} \cite{lmerTest}. The complete data and analysis of our first experiment are available from 
-{\tt http://homepages.abdn.ac.uk/k.vdeemter/pages/vagueness-data-analysis/e1.html}, and analogous for experiments 2-4.\\[2ex]
-
-% KvD I've moved the following two paragraphs towards the end of the section.
-% MG added/moved out of expt 1
-
-\noindent When the distance grows between two numbers, they become more easily distinguishable: the \emph{numerical distance effect} has been shown for comparing the cardinality of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene1996}.
-We manipulated the number of dots in each array such that some sets of arrays had smaller numerical distances and others had larger numerical distances.
-Where a number was mentioned in the instructions, it was always in the form of an Arabic numeral.
-% MG added/moved out of expt 1
-When two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}.
-We controlled which side the smaller number appeared on to avoid systematic influences of this effect.
-
-% KvD added the following from a previous version of the paper:
-There is abundant evidence \cite<e.g.,>{trick1994small} that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities.
-We performed a pilot experiment \cite{our-earlier-paper} in which we were able to confirm this finding in the experimental settings on which we are focussing in this paper.
-We found that, when participants were confronted with % MG changed the next line and added the footnote to remove the doubts we had about whether a dot array must have regular layout, or can have irregular layout - I'm still not sure this is a good description though
-a stimulus consisting of two squares containing different numbers of dots\footnote{Such a stimulus is referred to hereafter as consisting of a set of {\em dot arrays}.
-The number of dots in an array is referred to as its cardinality.
-The physical arrangement of dots in each array is irregular.},
-instructions of the form \emph{Choose the square with \emph{n} dots} led to consistently faster response times than instructions of the form \emph{Choose the square with many/few dots} when $2 \leq n \leq 5$; the converse was true for $n>5$.
-Given these findings, we henceforth focussed our studies on non-subitizable numbers, because it is there that vagueness is expected to have benefits.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 1                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 1: initial experiment}
-
-We used a forced choice task to elicit responses to crisp or vague forms of instructions that required the participant to choose one of two dot arrays by referring to its cardinality.
-The participant was presented with an instruction in the form \emph{Choose the square with \ldots dots}.
-Then a set of two dot arrays was presented, each in the form of a square containing some number of dots.
-The participant was required to identify the dot array that corresponded with the instruction, by pressing the appropriate key, as quickly as possible while avoiding errors. 
-Response time and accuracy were recorded for analysis.
-
-We manipulated how discriminable the dot arrays were by varying the numerical distance between them. 
-One array always contained 25 dots: the other contained either 5, 10, 15, 20, 30, 35, 40, or 45 dots. 
-This gave us numerical distances of 5, 10, 15, and 20, with smaller numerical distances resulting in less discriminable arrays and larger distances resulting in more discriminable arrays.
-
-Our main manipulation was of the vagueness of the instruction, with two levels, \emph{crisp} and \emph{vague}.
-Assuming the dot array [5, 25], and the instruction referring to the smaller cardinality, the \emph{crisp} instruction was \emph{Choose the square with 5 dots} and its \emph{vague} counterpart was \emph{Choose the square with few dots}.
-
-\subsection{Hypothesis}
-\noindent The main aim of this experiment was to test the following hypothesis:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] There exists a main effect advantage for vagueness: vague instructions impose a lower cognitive load for the comprehender than crisp alternatives.
-\end{APAenumerate}
-}
-
-\subsection{Experimental Method}
-
-%Twenty participants were recruited, aged between 18 and 45 with a median age of 26.
-On each trial a participant was presented with an instruction to choose one of two dot arrays on screen by reference to its cardinality. 
-Following a keypress to indicate that the participant had read the instruction there was a central fixation cross for 1000 ms, and a blank screen for 500 ms, followed by the array (without repetition of the referring expression).
-The arrays would stay on screen until the participant responded (there was no timing-out). 
-Response time was measured as the latency between the presentation of the arrays, and the keypress identifying the choice: in this way, the response time was separated from time spent reading the instructions, which is important since we are only interested in the former.
-A response was counted as erroneous if the square with the wrong number of dots was chosen (when the instruction contained a number); 
-if the square with the larger number of dots was selected (when the instruction was \emph{Choose the square with few dots}); 
-or if the square with the smaller number of dots was selected (when the instruction was \emph{Choose the square with many dots}). 
-No feedback was given on correct trials, but there was feedback on error trials in the form of the word ``\textsc{wrong!!}'' which flashed on screen.
-Hypothesis H1 was tested looking both at response times and at error rates.
-
-\subsection{Results}
-
-\begin{figure}[htbp]
-\fitfigure{images/e1RTandErrorRatesplot-1}
-\caption{Experiment 1 results: response times and error rates.}
-\label{resultse1}
-\end{figure}
-
-%\subsubsection{Response times}
-{\bf Response times} (RTs) for trials with erroneous responses were discarded, leading to the loss of 354 trials from 5120, representing 6.9\% of the trials. 
-The correct response RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of a further 160 trials, or 3.4\% of the remaining correct responses. 
-Means for response times and error rates are given in Fig. (\ref{resultse1}).
-A linear mixed model of RT was built, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
-Analysis confirms hypothesis H1:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1$_{rt}$)] RTs were faster for vague instructions than for crisp instructions 
-	($\beta=-0.11$, $se=.02$, $t=4.7$, $p<.001$).  % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-%\subsubsection{Error rates}
-{\bf Error rate} data were analysed using a generalized logit mixed model \cite{jaeger2008categorical}, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
-We found:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1$_{er}$)] There was a statistically non-significant effect of vagueness on error rates: the vague instructions tended to have a lower error rate than the than the crisp instructions 
-	($\beta=-.22$, $se=.18$, $t=-1.2$, $p=.230$).  % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-\subsection{Discussion}
-
-\noindent As expected, responses were faster and more accurate for vague than crisp instructions. Post hoc, we formulated some further hypotheses:
-
-{\small
-\begin{APAenumerate}
-	\item [(H2)] A main effect advantage of increasing numerical distance: the task will become easier as the numerical distance increases, because the two arrays are then more discriminable.
-	\item [(H3)] An interaction between vagueness and numerical distance: i.e., any facilitation for vagueness should be greater at smaller numerical distances than at larger numerical distances.
-\end{APAenumerate}
-}
-
-Looking at response times, these two post-hoc hypotheses were supported:
-
-{\small
-\begin{APAenumerate}
-	\item [(H2$_{rt}$)] RTs increased with numerical distance 
-	($\beta=-0.18$, $se=0.01$, $t=-11.6$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-	Since discriminability of the arrays is easier for larger numerical distances, discriminability probably underlies this effect. 
-	\item [(H3$_{rt}$)] Numerical distance and vagueness interacted significantly: essentially there were diminishing returns for vagueness as numerical distance increased: 
-	($\beta=0.07$, $se=0.01$, $t=5.9$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-Looking at error rates, the two post-hoc hypotheses were supported as well:
-
-{\small
-\begin{APAenumerate}
-	\item [(H2$_{er}$)] Error rates decreased as numerical distance increased: 
-	($\beta=-0.78$, $se=0.08$, $z=-9.6$, $p<0.0005$).% MG added the values for these stats on 25 Sept 2016
-	\item [(H3$_{er}$)] Numerical distance and vagueness interacted significantly: essentially there were diminishing returns for vagueness as numerical distance increased until the biggest numerical distance when vagueness was disadvantageous:
-	($\beta=0.80$, $se=0.15$, $z=5.27$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-So far, these results are in line with the idea of cost reduction.
-The idea of cost reduction explains the vagueness advantage by claiming that the vague referring expressions place less cognitive load on the comprehender than the crisp referring expressions.
-It explains the diminishing returns for vagueness in more-discriminable stimuli (i.e., the vagueness by numerical distance interaction) by claiming that load is low in both conditions for the easily-discriminable stimuli, and that therefore there is no extra benefit to be had from vagueness in the easily-discriminable stimuli.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 2                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 2: separating vagueness from instruction format}
-
-% KvD: I've added some syntactic sugar, in a number of places, to clarify the relation between Ex 1 and Ex 2.
-
-However, the picture painted by these findings might be misleading.
-First of all, there is a possibly confounding factor.
-Contrast, from Experiment 1, an expression from the vague condition: `the square with few dots' with an expression from the crisp condition: `the square with 5 dots'.
-One difference is that `few' has the potential for vagueness, whereas `5' is crisp. But
-another difference is that `few' is verbal while `5' is numerical, in the sense that a number is mentioned explicitly. 
-Since these two differences could not be separated in Experiment 1, the vagueness advantage finding is vulnerable to an alternative interpretation, that what we saw as a vagueness advantage was in contrast an advantage for the verbal form of the quantifier.
-In Experiment 2 we therefore created verbal and numeric versions of each of the vague and crisp instructions so that we could compare vague and crisp conditions while taking account of verbal / numeric format.
-
-Another potential problem with Experiment 1 is the following.
-Participants chose one of two squares: therefore the `vague' quantifiers (e.g., `few') uniquely identified one square.
-Recall our definition of vague  -- ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise''.
-In Experiment 1, the quantifiers in the vague conditions did not realise their potential for vagueness.
-This is because there were no borderline cases of the referent that could make the referent set `not well-defined', and perhaps because using definite articles in the instructions implied that only one option was correct.
-Using error feedback in Experiment 1 could have exacerbated this.
-
-To find out what happens when words are used in a context where their potential for vagueness comes to the fore, Experiment 2 used three arrays (rather than two) so that the vague description had more than one possible referent, and used indefinite articles to avoid the impression that only one response counted as correct, and was carried out without error feedback.
-An indication that the potential for vagueness was realised in Experiment 2 is that the borderline response was chosen fairly often: 16\% of the time.
-
-In Experiment 2, an item was a referring expression instruction followed by a set of three dot arrays defined by a triple of numbers, representing the number of dots in the left, middle, and right arrays.
-We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54).
-Each set of arrays comprised three arrays (instead of two as in Experiment 1); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had more.
-
-Examples of crisp and vague versions of the numerical and verbal instructions follow: the examples assume the array (6,15,24) and reference to the smaller number of dots, such that 6 was classified as the expected response; 15 as the borderline response; and 24 as the incorrect response.
-In the {\em vague numerical} condition we used \emph{Choose a square with about 10 dots}.
-None of the squares contained 10 dots.
-10 is slightly closer to 6 than to 15, justifying 6 as the best response and 15 as the borderline response.
-In the {\em vague verbal} condition we used \emph{Choose a square with few dots}.
-In the {\em crisp numerical} condition we used \emph{Choose the square with 6 dots}, and one square always did contain the number mentioned.
-For {\em crisp verbal}, we used \emph{Choose the square with the fewest dots}.
-
-% KvD I've re-instated the table below, because this makes it easier to understand the instruction format/selection algorithm confound.
-
-\begin{table}[htbp]
-\caption{Experiment 1: instructions arranged by condition for the dot triple $(6,15,24)$ and instructions indicating the smaller numbers}
-\label{instre3}
-\begin{tabular}{rll}
-&crisp&vague\\
-\midrule
-numerical	&	Choose the square with 6 dots 			& 	Choose a square with about 10 dots	\\
-verbal	&	Choose the square with the fewest dots	&	Choose a square with few dots		\\
-\midrule
-\end{tabular}
-\end{table}
-
-\subsection{Hypotheses} % E2
-\noindent We formulated the following hypotheses for Experiment 2:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1)] There exists a main effect RT advantage for vagueness.
-	\item [(H2)] An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
-	\item [(H3)] No large main effect of instruction format (i.e., numeric versus verbal) since we hypothesised that vagueness rather than instruction format drove the effect in Experiment 1.
-	\item [(H4)] On the basis of Experiment 1, we expect faster responses for stimuli with more discriminable arrays.(i.e., an effect of item).
-	\item [(H5)] Participants should make more borderline case choices for vague than crisp instructions.
-\end{APAenumerate}
-}
-
-\subsection{Experimental Method} % E2
-% Participants were aged between 18 and 45 with a median age of 28. 
-We manipulated as independent variables vagueness and instruction format (i.e., numeric versus verbal instructions), yielding four conditions, \emph{vague numeric}; \emph{vague verbal}; \emph{crisp numeric}; \emph{crisp verbal}.
-We measured two dependent variables: response time; and the probability of a participant choosing the borderline case.
-On each trial, first the referring expression that constituted the instruction for that trial was displayed. 
-Participants then pressed a key to indicate that they had read the instruction. 
-After 1000 ms, the arrays were presented, while preserving the text of the referring expression. 
-The response time dependent variable was measured from the presentation of the arrays, until the keypress indicating the participant's choice, which was also recorded. 
-The trial would timeout after 60 seconds if there was no response.
-In this experiment, no feedback was given. 
-This was because, in the vague conditions, we did not regard any response as `correct' or `incorrect', but instead as `borderline response', or `not borderline response', and we did not want to draw participants' attention to this distinction explicitly.
-We simply recorded whether the participant chose the best referent, the borderline case or the poorest referent, and how long it took the participant to respond.
-
-\subsection{Results} % E2
-
-%\subsubsection{Response times} % E2
-% MG added the values for these stats on 26 Sept 2016
-Means for response times and proportion of borderline responses are given in Fig. (\ref{resultse2}). 
-Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data. 
-A linear mixed model was constructed for the (logged) response times, 
-with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects. We found:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against Hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06$, $se=.01$, $t=4.6$, $p<0.0001$).
-	\item [(H2)] In focussed comparisons, vagueness was significantly disadvantageous in the \emph{numeric} conditions ($\beta=0.09$, $se=0.02$, $t=4.5$, $p<0.0001$), and non-significantly disadvantageous in the \emph{verbal} conditions ($\beta=0.02$, $se=0.02$, $t=1.5$, $p=0.1550$), offering partial evidence against Hypothesis 2.
-The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to a significant overall interaction effect between vagueness and instruction format ($\beta=-0.13$, $se=0.02$, $t=-6.6$, $p=0.0169$).
-	\item [(H3)] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1, but suggesting that instruction format rather than vagueness drove the effect we observed in Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; ($\beta=-0.36$, $se=0.07$, $t=-5.1$, $p<0.0001$).
-	\item [(H4)] There was a significant main effect of item (i.e., of which triple of numbers of dots were used in the stimulus): ($\beta=0.12$, $se=0.02$, $t=7.1$, $p<0.0001$).
-This effect seems likely to be due to the very fast responses for stimuli using arrangements of the smallest numbers of dots (6,15,24), which also had the largest difference in ratio of smallest number to largest number in the stimulus, suggesting that stimuli using these numbers of dots may have been particularly discriminable for participants.
-\end{APAenumerate}
-}
-
-%\subsubsection{Borderline responses} % E2
-% MG added the values for these stats on 26 Sept 2016
-A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
-The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{resultse2}.
-Participants chose the borderline square on $16.6\%$ of trials overall.
-{\small 
-\begin{APAenumerate}
-	\item [(H5)] Participants were significantly more likely to choose the borderline option for vague instructions than for crisp instructions (21.9\% vs 11.3\%: $\beta=0.62$, $se=0.22$, $z=2.8$, $p=0.0059$).
-Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%: $\beta=-3.35$, $se=0.23$, $z=-14.6$, $p<0.0001$).
-\end{APAenumerate}
-}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{images/e2-rtplot-1.pdf}
-\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{images/e2-blBarChart-1}
-\caption{Experiment 2 results: mean response times by condition and item, and counts of borderline case responses by condition.}
-\label{resultse2}
-\end{figure}
-
-\subsection{Discussion} % E2
-
-Experiment 2 tested to see whether when borderline cases are present, vague instructions would speed responses as they did in Experiment 1 when there were no borderline squares. 
-We actually found a small (but statistically significant) \emph{disadvantage} of vague instructions: vague instructions slowed people down by 112 ms on average.
-We also found that the effect of instruction format was significant, with numerical format slowing responses by 689 ms on average, such that the disadvantage of numerical format overwhelmed the contribution of vagueness.
-The \emph{verbal vague} condition was still responded to faster than the \emph{numerical crisp} condition, so the pattern from Experiment 1 was reproduced, but in the light of the evidence from Experiment 2, in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like an advantage of verbal instruction format.
-
-% KvD I've expanded the following:
-However, once again there is a possibly confounding factor.
-Observe that, in Experiment 2, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) {\bf selection algorithm}: To see this, consider the task of selecting the dot array that contains ``few dots": to do this, it suffices to {\em compare} the three arrays and select the one that contains the fewest elements.
-To select the dot array that contains "16 dots" seems to require the participant to estimate, and then {\em match}, the cardinality of (at least) one dot array to 16, a process which could plausibly take longer, independently of vagueness.
-Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison ``algorithm".% (which is known to be faster than matching).
-% KvD I think we need to say more about this, citing some appropriate papers to support this idea.
-\textcolor{red}{{\bf--Insert citation}}
-
-In the next two experiments we pitted the comparison algorithm and matching algorithm selection tasks against each other while controlling vagueness and instruction format.
-In Experiment 3 we restricted all the instructions to \emph{numeric} quantifiers while factorially manipulating vagueness and selection task.
-In Experiment 4 we ensured that all instructions used \emph{verbal} quantifiers, while also factorially manipulating vagueness and selection task.
-This allowed us to distinguish between the predictions of the selection task account and the instruction format account.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 3                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 3: focus on instructions that contain numerals}
-
 The main aim of Experiment 3 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies.
@@
 Table \ref{Instructions for e3} shows the instructions for each condition.
-Note the difference between ``fewer than 20" and ``far fewer than 20": whereas the former cannot have borderline cases (i.e., for each number it is clear whether the number is smaller than 20 or not), the latter can.
+Note the difference between ``fewer than 20'' and ``far fewer than 20'': whereas the former cannot have borderline cases (i.e., for each number it is clear whether the number is smaller than 20 or not), the latter can.
 
-%KvD I think we should motivate our choice of materials (especially "far ..er")
-%\textcolor{red}{{\bf KvD I think we should motivate our choice of materials (especially "far ..er")}}
-% DONE 28 Sept 2017
-
-\begin{table}[htbp]
+\begin{table}
 \centering
@@
 \label{Instructions for e3}
-\begin{tabular}{llll}
-instruction				&vagueness			&	selection		&	instruction							\\
-format				&					&	task			&									\\
-\toprule
+\begin{tabular}{cccc}
+\hline\noalign{\smallskip}
+Instruction	format          &Vagueness					&	Selection task	& Instruction									\\
+\noalign{\smallskip}\hline\noalign{\smallskip}
 \multirow{4}{*}{numeric}	&\multirow{2}{*}{crisp} 	&	matching		& Choose a square with 6 dots 					\\ 
@@
 							&							&	comparison 		& Choose a square with far fewer than 20 dots 	\\ 
-\bottomrule
+\noalign{\smallskip}\hline
 \end{tabular}
@@
 \noindent For Experiment 3, we hypothesised:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
-	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection task)
-	\item [(H3)] Effects of vagueness are different depending on whether the selection task is matching or comparison (interaction effect selection x vagueness).
-\end{APAenumerate}
-}
+
+\begin{description}
+	\item [Hypothesis 1] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
+	\item [Hypothesis 2] Comparison is easier for the reader than matching (main effect of selection task)
+	\item [Hypothesis 3] Effects of vagueness are different depending on whether the selection task is matching or comparison (interaction effect selection x vagueness).
+\end{description}
 
@@
 \centering
-\fitfigure{images/e3-rtplot-1.pdf}
-%\includegraphics[trim = 20mm 0mm 20mm 0mm, clip, width=.75\textwidth]{images/e3-rtplot-1.pdf}
+\includegraphics[width=\textwidth]{figures/e3-rtplot-1.pdf}
 \caption{Mean response times by condition for Experiment 3 where all instructions were numeric}
@@
 
-% MG added the values for these stats on 26 Sept 2016
-{\small
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions were non-significantly easier for the reader than crisp instructions when vagueness was considered as a main effect ($\beta=-0.01$, $se=0.01$, $t=-0.4$, $p=0.7158$). 
-	\item [(H2)] There was a statistically significant main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=0.16$, $se=0.03$, $t=6.2$, $p<0.0001$).
-	\item [(H3)] Vagueness exerted effects in different directions for the comparison task and for the matching task: there was a significant interaction between vagueness and selection task ($\beta=0.13$, $se=0.03$, $t=4.2$, $p<0.0001$).
+\begin{description}
+	\item [Test of Hypothesis 1] Vague instructions were non-significantly easier for the reader than crisp instructions when vagueness was considered as a main effect ($\beta=-0.01$, $se=0.01$, $t=-0.4$, $p=0.7158$).
+	\item [Test of Hypothesis 2] There was a statistically significant main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=0.16$, $se=0.03$, $t=6.2$, $p<0.0001$).
+	\item [Test of Hypothesis 3] Vagueness exerted effects in different directions for the comparison task and for the matching task: there was a significant interaction between vagueness and selection task ($\beta=0.13$, $se=0.03$, $t=4.2$, $p<0.0001$).
 Separate analyses were conducted testing for effects of vagueness at each level of the selection task.
@@
 Within the matching task vagueness significantly \emph{slowed} response times compared with crisp controls ($\beta=0.06$, $se=0.02$, $t=2.9$, $p<0.0061$). 
-\end{APAenumerate}
-}
-% MG edited this 27 Sept 2016:
-%The cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
+\end{description}
+
 In other words, the cost reduction account was wrong to predict significant main effect advantages for vagueness (although there was a non-significant trend in the \emph{direction} predicted by the cost reduction account), and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was significantly advantageous in the comparison task.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 4                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 4: focus on instructions that do not contain numerals} % E 4
-
-This experiment mirrors Experiment 3, but focussing on instructions that did not use a number. 
-We manipulated vagueness and the selection task (comparison and matching). 
-In order to implement the experiment without mentioning numbers in the instructions, we changed the sequence of each trial to include a `target' (i.e., a dot array of a particular cardinality) before the instruction, so that we could then refer back to the target's cardinality in the instruction using expressions like \emph{the same number of dots as the target}; \emph{fewer dots than the target}.
-%
-This presentation of a target before the main body of the trial shares some features with \citeA[Experiment 2]{Izard20081221}, although in that experiment participants were told the cardinality of the target (called an \emph{inducer} in that paper) whereas in our experiment we did not tell participants the cardinality of the prime array.
-An item was thus a combination of a target dot array, an instruction that did not contain a number, and a set of dot arrays taking their cardinalities from the same triples used in Experiment 2.
-Table \ref{Instructions for e4} spells out how the instructions were constrained not to mention a numeral and gives examples of targets.
-
-\begin{table}[htp]
-\caption{Instructions and targets by condition for Experiment 4}
-\begin{center}
-\begin{tabular}{llcrl}
-Item					&Selection				&Vagueness	&Target	&Choose a square with\ldots						\\
-\toprule
-\multirow{4}{*}{06:15:24} 	&\multirow{2}{*}{Comparison} 	&Crisp		&20		&\ldots fewer dots than the target					\\
-\cline{3-5}
-					&						&Vague		&20		&\ldots far fewer dots than the target					\\
-					\cline{2-5}
-					&\multirow{2}{*}{Matching}		& Crisp		&6		&\ldots the same number of dots as the target			\\
-					\cline{3-5}
-					& 		 				&Vague		&10 		&\ldots about the same number of dots as the target		\\
-					\bottomrule
-\end{tabular}
-\end{center}
-\label{Instructions for e4}
-\end{table}%
-
-\subsection{Hypotheses}% E 4
-\noindent For Experiment 4, we hypothesised:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
-	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection)
-	\item [(H3)] Effects of vagueness differ depending on whether selection is matching or comparison (interaction effect selection x vagueness, and focussed comparisons at each level of selection).
-\end{APAenumerate}
-}
-
-\subsection{Results}% E 4
-40 volunteers participated.
-The results showed that vagueness was beneficial for comparison but detrimental for matching (the same as Experiment 3) even when no numbers were allowed in the instructions. 
-Figure \ref{resultse4} shows the means by condition. Our findings were as follows:
-
-% MG added the values for these stats on 27 Sept 2016
-{\small
-\begin{APAenumerate}
-	\item [(H1)] There was no significant main effect of vagueness ($\beta=-0.02$, $se=0.01$, $t=-1.5$, $p=0.1296$). 
-	\item [(H2)] There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=0.18$, $se=0.02$, $t=10.4$, $p<0.0001$).
-This effect was in the same direction as Experiment 3.
-	\item[(H3)] Vagueness did exert different effects depending on the selection task (main interaction effect of vagueness by selection $\beta=0.12$, $se=0.02$, $t=5.1$, $p<0.0001$). 
-Separate analyses of the effect of vagueness were conducted for the comparison task and for the matching task using Bonferroni-adjusted significance thresholds. 
-In the comparison task, vagueness resulted in faster response times ($\beta=-0.08$, $se=0.02$, $t=-4.3$, $p<0.0001$). 
-In the matching task vagueness slowed response times ($\beta=0.05$, $se=0.01$, $t=3.7$, $p=0.0004$). 
-These results are in the same direction as Experiment 3.
-\end{APAenumerate}
-}
-
-Once again, the cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
-
-\begin{figure}[htbp]
-\centering
-\fitfigure{images/e4-rtplot-1.pdf}
-\caption{Mean response times by condition for Experiment 4}
-\label{resultse4}
-\end{figure}
-
-\section{Discussion of experiments 3 and 4}
-
-The main aim of these two experiments was to test whether vagueness confers any cognitive benefits over and above those due to differences in the selection task according to whether the instruction mandates a \emph{comparison} selection task or a \emph{matching} selection task, when number-use is held constant.
-The main effect of selection task showed that the assumption that the \emph{comparison} task is easier than the \emph{matching} task is well-founded.
-In both experiments people were reliably quicker to respond in the \emph{comparison} task.
-
-Vagueness, which was the phenomenon on which our investigation focussed, did not exert a significant main effect in response time.
-However when the comparison and selection tasks were analysed separately, there was small significant advantage for vagueness in the \emph{comparison} tasks, but a small significant disadvantage for vagueness in the \emph{matching} tasks.
-
-\section{General Discussion}\label{discussion}
-
-To summarise our findings, Experiment 1 showed that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment could not distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format; and mandated a comparison strategy rather than a matching strategy.
-Experiment 2 showed that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional benefit in either the verbal format instructions or the numerical format instructions.
-
-However, Experiment 2 did not distinguish benefits of number avoidance from benefits of the comparison selection task.
-In Experiments 3 and 4 we manipulated vagueness and the selection task, separately at each level of numerical format.
-Across the two experiments, we found that the comparison-task instructions attracted faster response times than the matching-task instructions.
-Within the two experiments we found that vagueness exerts benefits when the selection task is \emph{comparison}, but not when the task is \emph{matching}.
-
-\begin{table}[htbp]
-\caption{Vagueness as range reduction: a summary of Experiments 3 and 4}
-\label{Vagueness as range reduction}
-\centering
-\begin{tabular}{cccc}
-selection task 				& vagueness		& candidates	& effect of vageueness	\\
-\toprule
-\multirow{ 2}{*}{comparison} 	& crisp 			& 2			& \multirow{ 2}{*}{vagueness advantage}  		\\
-\cline{2-3}
-  						& vague			& 1			&                                               				\\
-\midrule
-\multirow{ 2}{*}{matching} 		& crisp 			& 1 			&\multirow{ 2}{*}{vagueness disadvantage}	\\
-\cline{2-3}
- 						& vague			& 2			&								 	\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-What is one entitled to conclude?
-Given that we were able to identify a class of situations in which vague expressions led to faster response times than crisp ones, would it be valid to conclude that we have finally discovered an advantage for vagueness that cannot be ascribed to some other factor?
-We believe the answer to this question is negative.
-
-To see why, consider Figures \ref{resultse3} and \ref{resultse4}.
-Both figures depict four conditions, depending on whether the expression was crisp or vague, and depending on whether the referent could be identified using a comparison strategy or not.
-Two of the resulting four conditions result in an expression that can denote either of two referents; the other two conditions result in an expression that can only denote one referent, with the other possible referent being a marginal candidate at best.
-
-% KvD: I've shortened the text, removing duplication
-
-To see why vagueness has opposite effects, depending on whether it is used in matching or comparison situations, consider the stimulus with (6,15,24) dots.
-Now compare `Choose a square with 6 dots' with its vague counterpart `Choose a square with about 10 dots': by adding the word `about', we broaden the range of squares that the expression might be referring to.
-On the other hand, compare `Choose a square with fewer than 20 dots' with `Choose a square with far fewer than 20 dots': by adding the word `far', we did not broaden the range of squares denotable by the expression: we narrow it down, because only some of the squares that have fewer dots may have {\em far} fewer dots.
-
-The benefits of vagueness in the \emph{comparison} task in experiments 3 and 4 could thus be explained as differences in the number of valid targets for the expression.
-This leads us to speculate that the benefit for vagueness here could be due to the vague expression foregrounding a particular valid target while the crisp expression carries with it the additional task of distinguishing between two alternative valid targets, something we propose to call a ``range-reduction'' benefit.
-
-The observation that conditions with 1 candidate lead to shorter response times than conditions with 2 candidates is consistent with the range reduction hypothesis, but not with the idea that vagueness is beneficial.
-It appears, in other words, that shorter response times will only result from a vague expression if this expression leads to range reduction.
-Once again, it is not vagueness itself that has advantages but a phenomenon (namely range reduction) that is an automatic concomitant of vagueness in some types of situations.\\[2ex]
-%
-\noindent Looking at the entire series of experiments, our findings suggest that the observed benefits of vague expressions in certain situations may be due to factors other than vagueness: factors like avoiding numbers; permitting comparison tasks; and range reduction.
-The picture that is starting to emerge is subtle: on the one hand, in the situations that we have been studying
-%-- where cooperative speakers refer to an object (e.g., a square) by means of some quantity associated with the object -- 
-vagueness is not intrinsically beneficial.
-On the other hand, vague expressions frequently possess other features that {\em are} beneficial, and these are what give us the incorrect impression that vagueness itself is beneficial.
-Vagueness may thus have acquired a reputation that it does not deserve.
-The answer to Lipman's question, of why vagueness permeates human language (see our Introduction), may lie in a different direction after all, possibly relating to benefits for the speaker rather than the hearer.
-% \cite<see>[chapter 11, for discussion of some possible directions]{vanDeemterBook}.
-
-A comparison may clarify the logic of the situation.
-In recent years a number of studies, focussing on red wine, have suggested that alcohol, consumed in low doses, may have health benefits.
-An alternative explanation, however, asserts that it is not the alcohol in the wine that was beneficial, but antioxidants from grapes.
-If this alternative explanation is correct, then alcohol may not be healthy after all.\\[2ex]
-%
-% KvD Feb 2017. I've diminished the role of NLG in the coming paragraph.
-Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
-Years of research on the logic of vagueness -- giving rise to such techniques as Partial Logic \cite<e.g.,>{Fine}, Probabilistic Logic \cite{Edgington}, and Fuzzy Logic \cite{Zadeh} -- have primed the research community to expect that some special utility of vagueness is an important part of the answer, but our findings call this expectation into question.
-
-Although our own studies in this article have focussed on vagueness in descriptive Noun Phrases, it seems plausible that vagueness plays a similar role in other linguistic constructs.
-For example, consider reports on air temperature.
-Given a numerical temperature measurement or prediction, we might word it as
-%
-\begin{quote}
-(a) {\em 27.2 degrees Celsius}, or\\
-(b) {\em approximately 27 degrees}, or\\
-(c) {\em above 25 degrees}, or\\
-(d) {\em warm}, 
-\end{quote}
-%
-among other candidate expressions.
-Which of these descriptions is most effective, for example as part of a weather report?
-If the linguistic literature is to be believed, then options (a) and (c) convey crisp information, whereas (b) and (d) are vague (i.e., they permit borderline cases).
-From our own experiments and the ones discussed in section 1, there is no evidence that  vagueness should be beneficial for hearers.
-Rather than asking whether a candidate expression is vague, other questions might shed more light on the choice, similar to the ones identified in our studies.
-These questions might focus on the amount of information that a given expression conveys (i.e., on granularity), on the avoidance of numbers, and on the use of evaluative terms.
-Let's see how this might pan out for the examples at hand.
-
-First, the experiments by \citeauthor{Mishra01042011} suggest that it is important how much information is conveyed by an expression, and their findings are echoed by our own thoughts about range reduction (following Experiments 3 and 4).
-In the case of (a)--(d) above, it appears that (a) conveys the most detailed information (designating the smallest segment of the temperature scale), followed by (b), then (d), then (c) (e.g., 40 degrees is above 25, but at 40 Celsius the word ``warm" is likely to give way to ``hot" or ``scorching"):
-%
-\begin{quote}
-$a < b < d < c$
-\end{quote}
-%
-If these hunches are correct then an {\em expert} may prefer to read expression (a), because it gives her the most detailed information on which to base her decisions.
-On the other hand, expression (d) (``warm") is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make (d) more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm" (see our section 1 and  \citeA{peters2009bringing}), which is especially important if the hearer is unfamiliar with the metric used.
-These considerations suggest that {\em non-experts} might prefer expression (d).
-%If results of this kind were to be confirmed by experiments, they would be applicable in practical NLG, though care would need to be exercised to take the purpose of the communication into account.\footnote{One reason why ``warm" might be less preferred than these considerations suggest is that, in the weather context, some very frequently occurring numbers, such as 20 (degrees), could be relatively easy to process, analogous to the subitizable numbers of \protect\citeA{trick1994small}, see our section \ref{method}.
-Note that in the experiments reported in this paper, the purpose of the utterances investigated was always clear (e.g., instructing people on a weight loss program \cite{Mishra01042011}; informing a patient's choice of hospital \cite{peters2009bringing}; or selecting a dot array (in our own studies).}
-
-One way to see why vagueness (as defined in our Introduction) may not benefit human communication is the following thought experiment.
-Suppose a group of speakers understand the word ``warm" as vague, agreeing that temperatures above 26 count as warm, and temperatures below 24 do not count as warm, but considering temperatures between 24 and 26 as borderline cases.
-Now one day these speakers agree to sharpen up their definition deciding that, henceforth, ``warm" means ``$>25$ degrees" (as in (c) above): this decision resolves the borderline cases, while everything else remains the same.
-It seems unlikely that this change in language use, from a vague meaning to a crisp one (i.e., one that has no borderline cases anymore), would lower the utility of the word.
-Our experimental findings are consistent with this idea.
-
-% KvD Bracketed out the following for now:
-%A more radical perspective on our findings might say that the distinction between crisp and vague expressions, which is taken for granted in the bulk of the linguistic literature, is mistaken, because expressions such as ``above 25 degrees" (and ``fewer than 20 dots") are not equally felicitous in all situations in which they can be truthfully applied.
-For example, when the temperature is 40 degrees, then a weather report saying that it is ``above 25 degrees" would be correct yet misleading: this description might be felicitous as a description of temperatures between (roughly) 26-28 degrees, borderline felicitous for 29-33 degrees, and infelicitous at even higher temperatures.
-At its most bold, this perspective would assert that the applicability of {\em every} linguistic expression has borderline cases, hence Lipman's question (see section 1) simply does not arise.
-This radical perspective can only be indicated here, but not discussed at length.
-
-\setcounter{secnumdepth}{0}
-\bibliography{v-special-issue-2016}
-\end{document}

### 2018-08-02 springer/current/e3.tex
massive restructuring Aug 2018

--- 2016-09-27 papers/2016-Special-Issue/v-special-issue-2016.tex
@@
-\documentclass[%
-%			% choose man or doc immediately below
-%doc,	% doc is single spaced
-man,		% man is double spaced
-floatsintext,%
-apacite%
-]{apa6}
-\usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{color}
-\usepackage{booktabs}
-\usepackage{layout}
-\usepackage{apacite}
-\usepackage{datetime}
-
-%\renewcommand{\baselinestretch}{1.7} % added 21 Sept 2016, copied from CRM book, but see documentclass option 'man' for "manuscript'' above which makes it double spaced
-
-\title{Is Vagueness Beneficial for Hearers? \protect \\ Evidence from Experiments}
-\date{\today}
-\authornote{latex compiled \today ~\currenttime}
-\shorttitle{Is Vagueness Beneficial for Hearers?}
-\twoauthors{Matt Green}{Kees van Deemter}
-\twoaffiliations{Computing Science\\ University of Aberdeen}{Computing Science\\ University of Aberdeen}
-
-\abstract{Much of everyday language use is vague, but the causes of this phenomenon are not well understood.
-Consequently, it is difficult for the designers of a Natural Language Generation (NLG) system to know when to let the system generate vague expressions.
-The present paper is an attempt to find out what benefits vagueness might have for readers.% , by means of a series of carefully controlled psycholinguistic experiments that focus on the use of vagueness in referring expressions.
-The work is motivated by NLG but has implications for linguistics, philosophical logic and game theory.
-%
-This article reports on a series of experiments that aim to separate the utility of vagueness (as defined by the existence of borderline cases) from the utility of other factors that tend to co-occur with vagueness.
-We argue that the evidence so far supports a view where the benefits that vague terms exert are due to other influences, rather than to vagueness itself.
-These factors include: low granularity; the use of evaluative words; the avoidance of overtly numerical words; the existence of comparison strategies; and, lastly and more tentatively, a phenomenon that we call range reduction.
-The paper concludes with a brief discussion of the implications for practical NLG.}
-
-\begin{document}
-\maketitle   
-\setcounter{secnumdepth}{3}
-\section{Introduction}\label{introduction}
-Vagueness pervades the language that we use on a daily basis, and the challenge of understanding vague language has been a prominent concern in many areas of logic and linguistics, involving both theoretical and applied work, including the area known as Natural Language Generation (NLG).
-
-NLG systems transform data and formulas into language \cite<e.g.,>[]{reiter2000building}.
-NLG systems routinely make decisions between different formulations of the same information.
-For example, if the temperature is 27.2 degrees Celsius, this could be expressed as ``27.2 degrees'', ``approximately 27 degrees', ``above 25 degrees", or ``warm'', and the system must decide between these.
-The problem is especially important for NLG systems that take numbers as input, for example in the generation of textual weather reports from numerical weather data such as temperature and wind speed \cite{goldberg1994using,turner2006generating}, and medical decision support on the basis of clinical measurement such as oxygen saturation, heart rhythm, etc.
-\cite{Hripcsak01032009, hunter2008summarising, portet2009automatic}.
-Such systems are often forced to make decisions concerning the level of precision in the utterances that they generate on the basis of little more than intuition.
-Even when NLG systems are designed to mimic human language use \cite<e.g.,>{konstas2013global} there is no guarantee that these decisions taken by these systems benefit readers.
-A better understanding of the benefits (for readers and hearers) of different precision levels would allow these systems to become more useful.
-The present article investigates the benefits,  or de-benefits, of vagueness.
-
-%.
-In the context of an NLG system faced with a practical decision of this kind, Lipman's question becomes ``Under what circumstances should vague terms be produced?''  Relevant applications include weather forecasting
-
-Language use may be called vague for various reasons.\footnote{See e.g.
-the entry ``vague'' in \cite{penguin}.} In most academic use though, the word `vagueness' has a specific meaning.
-Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1997vagueness}, also \citeA{EgreKlinedinst}).
-The crucial criterion is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise'' \cite[p.\ 1]{lipmanvague}.
-A typical example is the word ``tall'', as applied to people for example, because here is no precise, known height which separates those who are tall from those who are not.
-The crucial point is that ``tall'' admits borderline cases (i.e., people who may or may not count as tall), which are the hallmark of vagueness as we use the term.
-
-Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions \cite{Lipman:2000fk, lipmanvague}.
-By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss'' \cite<>[p.~1]{lipmanvague}.
-Lipman explains the point by means of a scenario in which a speaker describes a person to a hearer, who needs to identify that person in the arrivals hall of of an airport.
-In such a scenario, a precise description of the person's height (e.g., ``The person's height is 187.96 cm'') would be more useful than a vague one (``The person is tall'').
-Lipman uses this scenario to explain why standard game theory models of communication \cite<e.g.,>{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs.
-%The relevant conditions are, broadly speaking, that both interlocutors know all the relevant facts (e.g., both know the person's height precisely) and that the setting for the communication is co-operative.
-These conditions exclude deliberate deception, and rhetorical situations like political debate and advertising where the intention is to persuade the interlocutor to adopt some point of view, where the persuasion might not be intended to be in the addressee's best interest.
-
-Lipman argued that the efficiency loss resulting from vague expressions would be unlikely to have arisen unless there are advantages as well as disadvantages associated with vague expressions.
-Lipman asked, essentially, what these advantages might be.
-%, and how they might find a place in a game-theoretical explanation.
-In this paper, we focus on the first part of Lipman's question.
-Several tentative answers to Lipman's question have been offered  \cite<see>{van2009utility, van2010vagueness}.
-Prominent among these answers is the idea that vague expressions are easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite<e.g.,>[]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
-For example, \citeA<>[p.\ 11]{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze''.
-We shall refer to this as the \emph{cost reduction} hypothesis.
-%The idea of vagueness as cost reduction can take various shapes, but the basic hypothesis is that it is easier for people to think in terms of loosely defined categories (such as ``quite a few'', or ``many'') than in terms of crisply defined ones (such as ``thirteen'', or ``237'').
-%This hypothesis predicts that when a vague expression can perform the same communicative task as a precise one, then, the things being equal, it is rational to choose the vague expression.
-The corollary for comprehension is that vague expressions should be understood more readily than precise expressions.
-
-%The cost reduction hypothesis is of direct relevance to psycholinguists interested in language comprehension, and additionally to psycholinguists interested in language production, for example in connection with the question of audience design \cite{Clark1982287}.
-For, to the extent that speakers and writers choose vague expressions over and above crisp ones because the former are easier to process for hearers than the latter, the cost reduction hypothesis suggests that speakers design their utterance for optimal benefit to their hearers -- out of altruism, so to speak.
-
-Charting the utility of vagueness is the attested aim of a small number of studies, but most of these have focussed on vagueness in a different sense, and focussing on different types of benefits for hearers.
-Two recent studies can illustrate both issues.
-
-In a series of studies of behaviour modification, \citeA<>{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss.
-In the weight loss study, participants were told that the study was designed to test the validity of a new (actually fictitious) health index, the HHI (Holistic Health Index).
-They were told that an ideal HHI score lies in the range of 45 to 55.
-In a longitudinal study, participants submitted their weight to a computer each week.
-Participants were told that two algorithms would be used to compute their HHI, and that the two might give different values initially, in which case the true score lay between the two values.
-In one condition, which the authors called the precise condition, the two algorithms gave the same score.
-In the other condition, which the authors called the vague condition, one algorithm added 3\% to the score while the other algorithm subtracted 3\% from the score, yielding a range of values whose midpoint was the same as the two values given in the precise condition.
-
-One group of participants was given HHI scores in the ideal range: for this group their weight loss did not differ depending on whether they were given vague or precise HHI values.
-However for the other group, who were given HHI scores outside the ideal range, their weight loss was significantly greater if they were given vague HHI scores than if they were given precise HHI scores.
-The authors explain the improvement in the vague condition for this group as resulting from the participants' freedom to think of themselves a positioned on one end of the range - the end closest to the ideal HHI scores.
-This ``illusion of proximity'' \cite<>[p.~4]{Mishra01042011} to the goal is argued to allow participants to generate positive expectancies that lead to behaviours that improve performance.
-In contrast, in the precise conditions, participants did not have this freedom of interpretation, and could not distort the information to bring about the beneficial \emph{illusion of proximity}.
-These results are interesting, and of obvious potential practical importance.
-We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \cite{keefe1997vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases.
-In the terminology of \citeA{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}.
-Furthermore, the experiments of \citeA{Mishra01042011} did not explore benefits in terms of processing cost, but in terms of long-term behaviour change.
-
-Similar issues arise from the work of \citeA{peters2009bringing}.
-The authors carried out a series of studies where participants were required to rate hospitals based on various sources of information about quality of care.
-There was a between-subjects manipulation based on numeracy.
-The format of the information was manipulated within subjects: either numbers only were presented, or both numbers and evaluative categories were presented (e.g., \emph{Poor}, \emph{Fair}, \emph{Good}, \emph{Excellent}, with crisp visual boundary lines between the categories).
-Results showed that, for low-numeracy participants, the presence of evaluative categories resulted in a diminished influence of an irrelevant affective state on the ratings.
-For all participants, the presence of evaluative categories resulted in better decisions and in a greater use of the most important and reliable types of information, such as survival rates.
-
-It is, however, questionable whether the ``evaluative categories" manipulation in this study can be considered a manipulation of vagueness.
-Certainly, terms like \emph{Fair} admit the possibility of borderline cases.
-However, given that the boundaries between the categories were marked crisply, and that therefore the categories mapped crisply to numerical values, it becomes doubtful whether any borderline cases could be conceived to arise in fact.
-For example, \emph{Fair} was mapped to 60\% -- 70\% for the variable \emph{percentage of heart attack patients given recommended treatment (ACE inhibitor)}.
-Accordingly, rather than the vagueness of categories such as \emph{Poor}, Peters et al.
-emphasise the evaluative content inherent in these categories, and the affective potential of the evaluative content rather than the vagueness of the terms like {\em Fair}.
-
-\subsection{General Methodology}\label{method}
-The experiments reported in the present paper put the cost reduction hypothesis to the test. 
-The question that we are trying to answer is whether vague expressions are processed more easily by readers than crisp ones. 
-Like Lipman, we focus on situations where numerical information is used in order to identify a referent. 
-Reference, in other words, will be the linguistic task on which we focus, partly because of the interest that this topic has recently drawn from the NLG community.  
-In focussing on benefits for the hearer, we will leave aside the question of audience design, leaving this for later research.
-
-In using references to quantities to test the cost reduction hypothesis we are only testing one aspect of vagueness in a particular context. 
-This limits the applicability of our results. 
-However, it has the advantage that it enables us to explore the costs and benefits of vagueness more thoroughly. 
-Since one prevalent view of vagueness is that a vague expression is never preferable to a crisp equivalent, a demonstration of a benefit for vagueness in any context would advance the discussion.
-
-In our experiments we used a speeded forced choice task to compare the processing costs of different references to quantities. 
-In this context, speed and accuracy of responses are the key dimensions on which the different references can be compared. 
-Each stimulus in the experiments was a set of dot arrays containing various number of dots, together with a preceding instruction (in the form of a referring expression) to choose one of the arrays with respect to its cardinality.
-The participant was asked to respond as quickly as possible while avoiding errors. 
-We manipulated the instructions and the arrays in several ways across a series of four experiments.
-
-% MG General methodology moved to here: per-experiment method sections only talk about deviations from this general method
-All the experiments shared the following properties: 
-Stimuli were created using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.
-The position of the dots was randomised per-trial.
-The order in which trials were presented was randomised per-participant. 
-There were 256 trials, presented in 4 blocks of 64 each, between which the participant could rest. 
-A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants and recorded responses. 
-Participants were recruited using email lists at the University of Aberdeen, and paid ten pounds for participating.
-All participants self-reported fluency in English, and had normal, or corrected-to-normal vision.
-The experiment was conducted in a quiet room. 
-Participants were asked to respond as quickly as possible while avoiding errors.
-There was a block of practice trials after which participants could ask any questions, following which the experimenter left the room.
-All $p$ values reported for linear models were calculated using the R package \emph{lmerTest} \cite{lmerTest}.
-
-% KvD I've moved the following two paragraphs towards the end of the section.
-% MG added/moved out of expt 1
-
-When the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the cardinality of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene1996}.
-We manipulated the number of dots in each array such that some sets of arrays had smaller numerical distances and others had larger numerical distances.
-Where a number was mentioned in the instructions, it was always in the form of an Arabic numeral.
-% MG added/moved out of expt 1
-When two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}.
-We controlled which side the smaller number appeared on to avoid systematic influences of this effect.
-
-% KvD added the following from a previous version of the paper:
-There is abundant evidence \cite<e.g.,>{trick1994small} that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities.
-We performed a pilot experiment \cite{our-earlier-paper} in which we were able to confirm this finding in the experimental settings on which we are focussing in this paper.
-As one might expect, we found that, when participants were confronted with % MG changed the next line and added the footnote to remove the doubts we had about whether a dot array must have regular layout, or can have irregular layout - I'm still not sure this is a good description though
-a stimulus consisting of two squares containing different numbers of dots\footnote{such a stimulus is referred to hereafter as a stimulus consisting of a set of some number of dot arrays, where the number of dots in each dot array varies and is referred to as its cardinality, and the physical arrangement of dots in each dot array is irregular},
-instructions of the form \emph{Choose the square with \emph{n} dots} led to consistently faster response times than instructions of the form \emph{Choose the square with many/few dots} when $2 \leq n \leq 5$; the converse was true for $n>5$.
-We henceforth focussed our studies on non-subitizable numbers, because it is there that vagueness is expected to have benefits.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 1                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment One}
-
-\subsection{Introduction}
-
-We used a forced choice task to elicit responses to crisp or vague forms of instructions that required the participant to choose one of two dot arrays by referring to its cardinality.
-The participant was presented with an instruction in the form \emph{Choose the square with \ldots dots}.
-Then a set of two dot arrays was presented, each in the form of a square containing some number of dots.
-The participant was required to identify the dot array that corresponded with the instruction, by pressing the appropriate key, as quickly as possible while avoiding errors. 
-Response time and accuracy were recorded for analysis.
-
-We manipulated how discriminable the dot arrays were by varying the numerical distance between them. 
-One array always contained 25 dots: the other contained either 5, 10, 15, 20, 30, 35, 40, or 45 dots. 
-This gave us numerical distances of 5, 10, 15, and 20, with smaller numerical distances resulting in less discriminable arrays and larger distances resulting in more discriminable arrays.
-
-Our main manipulation was of the vagueness of the instruction, with two levels, \emph{crisp} and \emph{vague}.
-Assuming the dot array [5, 25], and the instruction referring to the smaller cardinality, the \emph{crisp} instruction was \emph{Choose the square with 5 dots} and its \emph{vague} counterpart was \emph{Choose the square with few dots}.
-
-\subsection{Hypothesis}
-\begin{APAenumerate}
-	\item [(H1)] A main effect advantage for vagueness: vague instructions impose a lower cognitive load for the comprehender than crisp alternatives.
-\end{APAenumerate}
-
-\subsection{Method}
-
-%Twenty participants were recruited, aged between 18 and 45 with a median age of 26.
-On each trial a participant was presented with an instruction to choose one of two dot arrays on screen by reference to its cardinality. 
-Following a keypress to indicate that the participant had read the instruction there was a central fixation cross for 1000 ms, and a blank screen for 500 ms, followed by the array (without repetition of the referring expression).
-The arrays would stay on screen until the participant responded (there was no timing-out). 
-Response time was measured as the latency between the presentation of the arrays, and the keypress identifying the choice: in this way, the response time was separated from time spent reading the instructions, which is important since we are only interested in the former.
-A response was counted as erroneous if the square with the wrong number of dots was chosen (when the instruction contained a number); 
-if the square with the larger number of dots was selected (when the instruction was \emph{Choose the square with few dots}); 
-or if the square with the smaller number of dots was selected (when the instruction was \emph{Choose the square with many dots}). 
-No feedback was given on correct trials, but there was feedback on error trials in the form of the word ``\textsc{wrong!!}'' which flashed on screen.
-Hypothesis H1 was tested looking both at response times and at error rates.
-
-\subsection{Results}
-
-\begin{figure}[htbp]
-\fitfigure{images/plotRTandErrorRates-1}
-\caption{Experiment 1 results: response times and error rates.}
-\label{resultse1}
-\end{figure}
-
-\subsubsection{Response times}
-Response times (RTs) for trials with erroneous responses were discarded, leading to the loss of 354 trials from 5120, representing 6.9\% of the trials. 
-The correct response RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of a further 160 trials, or 3.4\% of the remaining correct responses. 
-Means for response times and error rates are given in Fig. (\ref{resultse1}).
-A linear mixed model of RT was built, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
-
-\begin{APAenumerate}
-	\item [(H1$_{rt}$)] RTs were faster for vague instructions than for crisp instructions 
-	($\beta=-0.11$, $se=.02$, $t=4.7$, $p<.001$).  % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-
-\subsubsection{Error rates}
-Error rate data were analysed using a generalized logit mixed model \cite{jaeger2008categorical}, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
-
-\begin{APAenumerate}
-	\item [(H1$_{er}$)] There was a statistically non-significant effect of vagueness on error rates: the vague instructions tended to have a lower error rate than the than the crisp instructions 
-	($\beta=-.22$, $se=.18$, $t=-1.2$, $p=.230$).  % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-
-\subsection{Discussion}
-
-As expected, responses were faster and more accurate for vague than crisp instructions. Post hoc, we formulated some further hypotheses:
-
-\begin{APAenumerate}
-	\item [(H2)] A main effect advantage of increasing numerical distance: the task will become easier as the numerical distance increases, because the two arrays are then more discriminable.
-	\item [(H3)] An interaction between vagueness and numerical distance: i.e., any facilitation for vagueness should be greater at smaller numerical distances than at larger numerical distances.
-\end{APAenumerate}
-
-Looking at response times, these two post-hoc hypotheses were supported:
-
-\begin{APAenumerate}
-	\item [(H2$_{rt}$)] RTs increased with numerical distance 
-	($\beta=-0.18$, $se=0.01$, $t=-11.6$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-	Since discriminability of the arrays is easier for larger numerical distances, discriminability probably underlies this effect. 
-	\item [(H3$_{rt}$)] Numerical distance and vagueness interacted significantly: essentially there were diminishing returns for vagueness as numerical distance increased: 
-	($\beta=0.07$, $se=0.01$, $t=5.9$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-
-Looking at error rates, the two post-hoc hypotheses were supported:
-
-\begin{APAenumerate}
-	\item [(H2$_{er}$)] Error rates decreased as numerical distance increased: 
-	($\beta=-0.78$, $se=0.08$, $z=-9.6$, $p<0.0005$).% MG added the values for these stats on 25 Sept 2016
-	\item [(H3$_{er}$)] Numerical distance and vagueness interacted significantly: essentially there were diminishing returns for vagueness as numerical distance increased until the biggest numerical distance when vagueness was disadvantageous:
-	($\beta=0.80$, $se=0.15$, $z=5.27$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-
-So far, these results are in line with the idea of cost reduction.
-Cost reduction explains the vagueness advantage by claiming that the vague referring expressions place less cognitive load on the comprehender than the crisp referring expressions.
-It explains the diminishing returns for vagueness in more-discriminable stimuli (i.e., the vagueness by numerical distance interaction) by claiming that load is low in both conditions for the easily-discriminable stimuli, and that therefore there is no extra benefit to be had from vagueness in the easily-discriminable stimuli.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 2                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 2}
-
-\subsection{Introduction} % E2
-
-% KvD: I've added some syntactic sugar, in a number of places, to clarify the relation between Ex 1 and Ex 2.
-
-However, the picture painted by these findings might be misleading.
-First of all, there is a possibly confounding factor.
-Contrast, from Experiment 1, an expression from the vague condition: `the square with few dots' with an expression from the crisp condition: `the square with 5 dots'.
-One difference is that `few' has the potential for vagueness, whereas `5' is crisp. But
-another difference is that `few' is verbal while `5' is numerical, in the sense that a number is mentioned explicitly. 
-Since these two differences could not be separated in Experiment 1, the vagueness advantage finding is vulnerable to an alternative interpretation, that what we saw as a vagueness advantage was in contrast an advantage for the verbal form of the quantifier.
-In Experiment 2 we therefore created verbal and numeric versions of each of the vague and crisp instructions so that we could compare vague and crisp conditions while taking account of verbal / numeric format.
-
-Another potential problem with Experiment 1 is the following.
-Participants chose one of two squares: therefore the `vague' quantifiers (e.g., `few') uniquely identified one square.
-Recall our definition of vague  -- ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise''.
-In Experiment 1, the quantifiers in the vague conditions did not realise their potential for vagueness.
-This is because there were no borderline cases of the referent that could make the referent set `not well-defined', and perhaps because using definite articles in the instructions implied that only one option was correct.
-Using error feedback in Experiment 1 could have exacerbated this.
-
-To find out what happens when words are used in a context where their potential for vagueness comes to the fore,
-Experiment 2 used three arrays so that the vague quantifiers always had more than one possible referent, and used indefinite articles in the vague instructions to avoid the impression that only one response counted as correct, and was carried out without error feedback.
-An indication that the potential for vagueness was realised in Experiment 2 is that the borderline response was chosen fairly often: 16\% of the time.
-
-In Experiment 2, an item was a referring expression instruction followed by a set of three dot arrays defined by a triple of numbers, representing the number of dots in the left, middle, and right arrays.
-We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54).
-Each set of arrays had the following properties: it comprised three arrays (instead of two as in Experiment 1); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had more.
-
-Examples of crisp and vague versions of the numerical and verbal instructions follow: the examples assume the array (6,15,24) and reference to the smaller number of dots, such that 6 was classified as the expected response; 15 was classified as the borderline response; and 24 was classified as the incorrect response.
-In the {\em vague numerical} condition we used \emph{Choose a square with about 10 dots}.
-None of the squares contained 10 dots.
-10 is slightly closer to 6 than to 15, justifying 6 as the best response and 15 as the borderline response.
-In the {\em vague verbal} condition we used \emph{Choose a square with few dots}.
-In the {\em crisp numerical} condition we used \emph{Choose the square with 6 dots}, and one square always did contain the number mentioned.
-For {\em crisp verbal}, we used \emph{Choose the square with the fewest dots}.
-
-% KvD I've re-instated the table below, because this makes it easier to understand the instruction format/selection algorithm confound.
-
-\begin{table}[htbp]
-\caption{Experiment 1: instructions arranged by condition for the dot triple $(6,15,24)$ and instructions indicating the smaller numbers}
-\label{instre3}
-\begin{tabular}{rll}
-&crisp&vague\\
-\midrule
-numerical	&	Choose the square with 6 dots 			& 	Choose a square with about 10 dots	\\
-verbal	&	Choose the square with the fewest dots	&	Choose a square with few dots		\\
-\midrule
-\end{tabular}
-\end{table}
-
-\subsection{Hypotheses} % E2
-\begin{APAenumerate}
-	\item [(H1)] A main effect RT advantage for vagueness.
-	\item [(H2)] An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
-	\item [(H3)] No large main effect of instruction format (of numeric versus verbal) since we hypothesised that vagueness rather than instruction format drove the effect in experiment 1.
-	\item [(H4)] On the basis of Experiment 1, we would expect faster responses for stimuli with more discriminable arrays.(i.e., an effect of item).
-	\item [(H5)] Participants should make more borderline case choices for vague than crisp instructions.
-\end{APAenumerate}
-
-\subsection{Method} % E2
-% Participants were aged between 18 and 45 with a median age of 28. 
-We manipulated as independent variables vagueness and instruction format, yielding four conditions, \emph{vague numeric}; \emph{vague verbal}; \emph{crisp numeric}; \emph{crisp verbal}.
-We measured two dependent variables: response time; and the probability of a participant choosing the borderline case.
-On each trial, first the referring expression that constituted the instruction for that trial was displayed. 
-Participants then pressed a key to indicate that they had read the instruction. 
-After 1000 ms, the arrays were presented, while preserving the text of the referring expression. 
-The response time dependent variable was measured from the presentation of the arrays, until the keypress indicating the participant's choice, which was also recorded. 
-The trial would timeout after 60 seconds if there was no response.
-In this experiment, no feedback was given. 
-This was because, in the vague conditions, we did not regard any response as `correct' or `incorrect', but instead as `borderline response', or `not borderline response', and we did not want to draw participants' attention to this distinction explicitly.
-We simply recorded whether the participant chose the best referent, the borderline case or the poorest referent, and how long it took the participant to respond.
-
-\subsection{Results} % E2
-
-\subsubsection{Response times} % E2
-% MG added the values for these stats on 26 Sept 2016
-Means for response times and proportion of borderline responses are given in Fig. (\ref{resultse2}). 
-Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data. 
-A linear mixed model was constructed for the (logged) response times, 
-with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
-\begin{APAenumerate}
-	\item [(H1)] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06$, $se=.01$, $t=4.6$, $p<0.0001$)
-	\item [(H2)] In focussed comparisons, vagueness was significantly disadvantageous in the \emph{numeric} conditions ($\beta=0.09$, $se=0.02$, $t=4.5$, $p<0.0001$), and non-significantly disadvantageous in the \emph{verbal} conditions ($\beta=0.02$, $se=0.02$, $t=1.5$, $p=0.1550$), offering partial evidence against hypothesis 2.
-The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to a significant overall interaction effect between vagueness and instruction format ($\beta=-0.13$, $se=0.02$, $t=-6.6$, $p=0.0169$).
-	\item [(H3)] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1, but suggesting that instruction format rather than vagueness drove the effect we observed in experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; ($\beta=-0.36$, $se=0.07$, $t=-5.1$, $p<0.0001$).
-	\item [(H4)] There was a significant main effect of item (i.e., of which triple of numbers of dots were used in the stimulus): ($\beta=0.12$, $se=0.02$, $t=7.1$, $p<0.0001$).
-This effect seems likely to be due to the very fast responses for stimuli using arrangements of the smallest numbers of dots (6,15,24), which also had the largest difference in ratio of smallest number to largest number in the stimulus, suggesting that stimuli using these numbers of dots may have been particularly discriminable for participants.
-\end{APAenumerate}
-
-\subsubsection{Borderline responses} % E2
-% MG added the values for these stats on 26 Sept 2016
-A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
-The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{resultse2}.
-Participants chose the borderline square on $16.6\%$ of trials overall.
-\begin{APAenumerate}
-	\item [(H5)] Participants were significantly more likely to choose the borderline option for vague instructions than for crisp instructions (21.9\% vs 11.3\%: $\beta=0.62$, $se=0.22$, $z=2.8$, $p=0.0059$).
-Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%: $\beta=-3.35$, $se=0.23$, $z=-14.6$, $p<0.0001$).
-\end{APAenumerate}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=.49\textwidth]{images/e2rt.pdf}
-\includegraphics[trim = 0mm -23mm 0mm 0mm, clip, width=.49\textwidth]{images/e2bl.pdf}
-\caption{Experiment 2 results: mean response times by condition and item, and counts of borderline case responses by condition.}
-\label{resultse2}
-\end{figure}
-
-\subsection{Discussion} % E2
-
-Experiment 2 tested to see whether when borderline cases are present, vague instructions would speed responses as they did in Experiment 1 when there were no borderline squares. 
-We actually found a \emph{disadvantage} of vague instructions: vague instructions slowed people down by 112 ms on average.
-We also found that the effect of instruction format was significant, with numerical format slowing responses by 689 ms on average, such that the disadvantage of numerical format overwhelmed the contribution of vagueness.
-The \emph{verbal vague} condition was still responded to faster than the \emph{numerical crisp} condition, so the pattern from Experiment 1 was reproduced, but in the light of the evidence from Experiment 2, in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like an advantage of verbal instruction format.
-
-% KvD I've expanded the following:
-However, once again there is a possibly confounding factor.
-Observe that, in Experiment 2, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) {\bf selection algorithm}: To see this, consider the task of selecting the dot array that contains ``few dots": to do this, it suffices to {\em compare} the three arrays and select the one that contains the fewest elements.
-Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison ``algorithm" (which is known to be faster than matching).
-% KvD I think we need to say more about this, citing some appropriate papers to support this idea.
-\textcolor{red}{{\bf--Insert citation}}
-
-In the next two experiments we pitted the comparison algorithm and matching algorithm selection tasks against each other while controlling vagueness and instruction format.
-In Experiment 3 we restricted all the instructions to \emph{numeric} quantifiers while factorially manipulating vagueness and selection task.
-In Experiment 4 we ensured that all instructions used \emph{verbal} quantifiers, while also factorially manipulating vagueness and selection task.
-This allowed us to distinguish between the predictions of the selection task account and the instruction format account.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 3                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 3}
-
-\subsection{Introduction}
-The main aim of experiment 3 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies.
-The main changes from experiment 2 were that the selection task was explicitly controlled, and that all conditions were constrained to mention a number.
-We used the same arrays as in experiment 2.
-Table \ref{Instructions for e3} shows the instructions for each condition.
-
-%KvD I think we should motivate our choice of materials (especially "far ..er")
-\textcolor{red}{{\bf KvD I think we should motivate our choice of materials (especially "far ..er")}}
-
-\begin{table}[htbp]
-\centering
-\caption{Experiment 3: Instructions arranged by condition for stimuli with $(6,15,24)$ dots and instructions indicating the smaller numbers} 
-\label{Instructions for e3}
-\begin{tabular}{llll}
-instruction				&vagueness			&	selection		&	instruction							\\
-format				&					&	task			&									\\
-\toprule
-\multirow{4}{*}{numeric}	&\multirow{2}{*}{crisp} 	&	matching		& Choose a square with 6 dots 				\\ 
-					&					&	comparison 	& Choose a square with fewer than 20 dots 	\\
-					\cline{2-4}
-					&\multirow{2}{*}{vague} 	&	matching 		& Choose a square with about 10 dots 		\\ 
- 					&					&	comparison 	& Choose a square with far fewer than 20 dots 	\\ 
-\bottomrule
-\end{tabular}
-\end{table}%
-
-\subsection{Hypotheses} % E 3
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions are easier for the reader than crisp alternatives (main effect of vagueness)
-	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection task)
-	\item [(H3)] Effects of vagueness are different depending on whether selection is matching or comparison (interaction effect selection x vagueness).
-\end{APAenumerate}
-
-\subsection{Method} % E 3
-
-38 participants were recruited. 
-The design was a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e3}).
-On each trial a referring expression instruction was presented: participants pressed a key to dismiss the instruction, when the dot arrays were presented until the participant responded, and the response time and choice were recorded.
-
-\subsection{Results} % E 3
-
-Response times were trimmed at 2.5 SD separately for each subject, leading to the loss of 204 trials (2.8\% of the trials).
-Condition means for the remaining (logged) RTs are plotted in Figure \ref{resultse3}.
-A linear mixed model was constructed for the logged response times, 
-with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[trim = 0mm 30mm 0mm 10mm, clip, width=.75\textwidth]{images/rtplote3}
-\caption{Mean response times by condition for Experiment 3 where all instructions were numeric}
-\label{resultse3}
-\end{figure}
-
-% MG added the values for these stats on 26 Sept 2016
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions were non-significantly easier for the reader than crisp instructions when vagueness was considered as a main effect ($\beta=-0.01$, $se=0.01$, $t=-0.4$, $p=0.7158$). 
-	\item [(H2)] There was a statistically significant main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=0.16$, $se=0.03$, $t=6.2$, $p<0.0001$).
-	\item [(H3)] Vagueness exerted effects in different directions for the comparison task and for the matching task: there was a significant interaction between vagueness and selection task ($\beta=0.13$, $se=0.03$, $t=4.2$, $p<0.0001$).
-Separate analyses were conducted testing for effects of vagueness at each level of the selection task.
-Within the comparison task vagueness significantly speeded response times compared with crisp controls ($\beta=-0.07$, $se=0.02$, $t=-3.5$, $p<0.0012$). 
-Within the matching task vagueness significantly \emph{slowed} response times compared with crisp controls ($\beta=0.06$, $se=0.02$, $t=2.9$, $p<0.0061$). 
-\end{APAenumerate}
-
-% MG edited this 27 Sept 2016:
-%The cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
-The cost reduction account was wrong to predict significant main effect advantages for vagueness (although there was a non-significant trend in the \emph{direction} predicted by the cost reduction account), and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was significantly advantageous in the comparison task.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 4                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 4} % E 4
-
-\subsection{Introduction}
-This experiment investigated response times for instructions that did not use a number. 
+This experiment mirrors Experiment 3, but focussing on instructions that did not use a number. 
 We manipulated vagueness and the selection task (comparison and matching). 
 In order to implement the experiment without mentioning numbers in the instructions, we changed the sequence of each trial to include a `target' (i.e., a dot array of a particular cardinality) before the instruction, so that we could then refer back to the target's cardinality in the instruction using expressions like \emph{the same number of dots as the target}; \emph{fewer dots than the target}.
-%
-This presentation of a target before the main body of the trial shares some features with \citeA[Experiment 2]{Izard20081221}, although in that experiment participants were told the cardinality of the target (called an \emph{inducer} in that paper) whereas in our experiment we did not tell participants the cardinality of the prime array.
+This presentation of a target before the main body of the trial shares some features with \citeauthor[Experiment 2]{Izard20081221}, although in that experiment participants were told the cardinality of the target (called an \emph{inducer} in that paper) whereas in our experiment we did not tell participants the cardinality of the prime array.
 An item was thus a combination of a target dot array, an instruction that did not contain a number, and a set of dot arrays taking their cardinalities from the same triples used in Experiment 2.
@@
 \begin{table}[htp]
-\caption{Instructions and targets by condition for experiment 4}
+\caption{Instructions and targets by condition for Experiment 4}
 \begin{center}
 \begin{tabular}{llcrl}
-Item					&Selection				&Vagueness	&Target	&Choose a square with\ldots						\\
-\toprule
-\multirow{4}{*}{06:15:24} 	&\multirow{2}{*}{Comparison} 	&Crisp		&20		&\ldots fewer dots than the target					\\
+\hline\noalign{\smallskip}
+Item					    &Selection task					& Vagueness	& Target		&Choose a square with\ldots							\\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+\multirow{4}{*}{06:15:24} 	&\multirow{2}{*}{Comparison} 	& Crisp		& 20			&\ldots fewer dots than the target					\\
 \cline{3-5}
-					&						&Vague		&20		&\ldots far fewer dots than the target					\\
+							&								& Vague		& 20			&\ldots far fewer dots than the target				\\
 \cline{2-5}
-					&\multirow{2}{*}{Matching}		& Crisp		&6		&\ldots the same number of dots as the target			\\
+							&\multirow{2}{*}{Matching}		& Crisp		& 6			&\ldots the same number of dots as the target		    \\
 \cline{3-5}
-					& 		 				&Vague		&10 		&\ldots about the same number of dots as the target		\\
-					\bottomrule
+							& 		 						& Vague		& 10 		&\ldots about the same number of dots as the target	    \\
+\noalign{\smallskip}\hline
 \end{tabular}
@@
 \subsection{Hypotheses}% E 4
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions are easier for the reader than crisp alternatives (main effect of vagueness)
-	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection)
-	\item [(H3)] Effects of vagueness are different depending on whether selection is matching or comparison (interaction effect selection x vagueness, and focussed comparisons at each level of selection).
-\end{APAenumerate}
+\noindent For Experiment 4, we hypothesised:
+
+\begin{description}
+	\item [Hypothesis 1] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
+	\item [Hypothesis 2] Comparison is easier for the reader than matching (main effect of selection)
+	\item [Hypothesis 3] Effects of vagueness differ depending on whether selection is matching or comparison (interaction effect selection x vagueness, and focussed comparisons at each level of selection).
+\end{description}
 
@@
 The results showed that vagueness was beneficial for comparison but detrimental for matching (the same as Experiment 3) even when no numbers were allowed in the instructions. 
-Figure \ref{resultse4} shows the means by condition.
+Figure \ref{resultse4} shows the means by condition. Our findings were as follows:
 
-% MG added the values for these stats on 27 Sept 2016
-\begin{APAenumerate}
-	\item [(H1)] There was no significant main effect of vagueness ($\beta=-0.02$, $se=0.01$, $t=-1.5$, $p=0.1296$). 
-	\item [(H2)] There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=0.18$, $se=0.02$, $t=10.4$, $p<0.0001$).
+\begin{description}
+	\item [Test of hypothesis 1] There was no significant main effect of vagueness ($\beta=-0.02$, $se=0.01$, $t=-1.5$, $p=0.1296$). 
+	\item [Test of hypothesis 2] There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=0.18$, $se=0.02$, $t=10.4$, $p<0.0001$).
 This effect was in the same direction as Experiment 3.
-	\item[(H3)] Vagueness did exert different effects depending on the selection task (main interaction effect of vagueness by selection $\beta=0.12$, $se=0.02$, $t=5.1$, $p<0.0001$). 
+	\item [Test of hypothesis 3] Vagueness did exert different effects depending on the selection task (main interaction effect of vagueness by selection $\beta=0.12$, $se=0.02$, $t=5.1$, $p<0.0001$). 
 Separate analyses of the effect of vagueness were conducted for the comparison task and for the matching task using Bonferroni-adjusted significance thresholds. 
@@
 These results are in the same direction as Experiment 3.
-\end{APAenumerate}
+\end{description}
 
-The cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
+Once again, the cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
 
@@
 \centering
-\includegraphics[trim = 0mm 30mm 0mm 10mm, clip, width=.75\textwidth]{images/rtplote4.pdf}
+\includegraphics[width=\textwidth]{figures/e4-rtplot-1.pdf}
 \caption{Mean response times by condition for Experiment 4}
@@
 \end{figure}
-
-\section{Discussion of experiments 3 and 4}
-
-The main aim of these two experiments was to test whether vagueness confers any cognitive benefits over and above those due to differences in the selection task according to whether the instruction mandates a \emph{comparison} selection task or a \emph{matching} selection task, when number-use is held constant.
-The main effect of selection task showed that the assumption that the \emph{comparison} task is easier than the \emph{matching} task is well-founded.
-In both experiments people were reliably quicker to respond in the \emph{comparison} task.
-
-Vagueness, which was the phenomenon on which our investigation focussed, did not exert a significant main effect in response time.
-However when the comparison and selection tasks were analysed separately, there was small significant advantage for vagueness in the \emph{comparison} tasks, but a small significant disadvantage for vagueness in the \emph{matching} tasks.
-
-\section{General Discussion}
-
-Experiment 1 showed that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment could not distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format; and mandated a comparison strategy rather than a matching strategy.
-Experiment 2 showed that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional benefit in either the verbal format instructions or the numerical format instructions.
-
-However, Experiment 2 could not distinguish benefits of number avoidance from benefits of the comparison selection task.
-In Experiments 3 and 4 we manipulated vagueness and the selection task separately at each level of numerical format.
-Across the two experiments, we found that the comparison-task instructions attracted faster response times than the matching-task instructions.
-Within the two experiments we found that vagueness exerts benefits when the selection task is \emph{comparison}, but not when the task is \emph{matching}.
-
-The benefits of vagueness in the \emph{comparison} task in experiments 3 and 4 could be explained as differences in the number of valid targets for the expression, as follows.
-Taking as an example the stimulus with (6,15,24) dots, it could be argued that the vague comparison instruction (e.g., \emph{a square with far fewer than 20 dots}) has one valid target, the square with 6 dots, while the crisp comparison instruction (e.g., \emph{a square with fewer than 20 dots}) has two valid targets, the squares with 6 and 15 dots.
-In both experiments 3 and 4 we found that people were quicker to identify a square when the instruction only had one valid target.
-This leads us to speculate that the benefit for vagueness here could be due to the vague expression foregrounding a particular valid target while the crisp expression carries with it the additional task of distinguishing between two alternative valid targets, something we propose to call a ``range-reduction'' benefit.
-
-\begin{table}[htbp]
-\caption{Vagueness as range reduction: a summary of Experiments 3 and 4}
-\label{Vagueness as range reduction}
-\centering
-\begin{tabular}{cccc}
-selection task 				& vagueness		& candidates	& effect of vageueness	\\
-\toprule
-\multirow{ 2}{*}{comparison} 	& crisp 			& 2			& \multirow{ 2}{*}{vagueness advantage}  		\\
-\cline{2-3}
-  						& vague			& 1			&                                               				\\
-\midrule
-\multirow{ 2}{*}{matching} 		& crisp 			& 1 			&\multirow{ 2}{*}{vagueness disadvantage}	\\
-\cline{2-3}
- 						& vague			& 2			&								 	\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-What is one entitled to conclude?
-Given that we were able to identify a class of situations -- namely: situations in which a comparison strategy suffices to identify the intended referent -- in which vague expressions led to faster response times than crisp ones, would it be valid to conclude that we have finally discovered an advantage for vagueness that cannot be ascribed to some other factor?
-We believe the answer to this question is negative.
-To see why, consider Figures \ref{resultse3} and \ref{resultse4}.
-Both figures depict four conditions, depending on whether the expression was crisp or vague, and depending on whether the referent could be identified using a comparison strategy or not.
-Two of the resulting four conditions result in an expression that can denote either of two referents; the other two conditions result in an expression that can only denote one referent, with the other possible referent being a marginal candidate at best.
-
-% KvD Something missing here?
-\textcolor{red}{KvD Something missing here?}
-
-To see why vagueness thus has opposite effects, depending on whether it is used in matching or comparison situations, compare an instruction like `Choose a square with 6 dots' with its vague counterpart `Choose a square with about 10 dots': by adding the word `about', we broaden the range of squares that the expression might be referring to.
-On the other hand, compare `Choose a square with fewer than 20 dots' with its vague counterpart `Choose a square with far fewer than 20 dots': by adding the word `far', we did not broaden the range of squares denotable by the expression: we narrow it down, because only some of the squares that have fewer dots may have {\em far} fewer dots.
-
-The observation that conditions with 1 candidate lead to shorter response times than conditions with 2 candidates is consistent with the range reduction hypothesis, but not with the idea that vagueness is beneficial.
-It appears, in other words, that %range reduction causes shorter response times, suggesting that
-shorter response times will only result from a vague expression if this expression leads to range reduction.
-Once again, it is not vagueness itself that has advantages but a phenomenon (namely range reduction) that is an automatic concomitant of vagueness in some types of situations.
-
-Our findings suggest that the observed benefits of vague expressions in certain situations may be due to factors other than vagueness itself: factors like avoiding numbers; permitting comparison tasks; and range reduction.
-The picture that is starting to emerge is subtle: on the one hand, in the situations that we have been studying -- where cooperative speakers refer to an object (e.g., a square) by means of some quantity associated with the object -- vagueness is not intrinsically beneficial.
-On the other hand, vague expressions frequently possess other features that {\em are} beneficial, and these are what give us the incorrect impression that vagueness itself is beneficial.
-Vagueness may thus have acquired a reputation that it does not deserve.
-To answer to Lipman's question, of why vagueness permeates human language (see our Introduction), may lie in a different direction after all, possibly relating to a combination of pure necessity and benefits for the speaker rather than the hearer \cite<see>[chapter 11, for discussion of some possible directions]{van2010vagueness}.
-
-A comparison may clarify the logic of the situation.
-In recent years a number of studies, focussing on red wine, have suggested that alcohol, consumed in low doses, may have health benefits.
-An alternative explanation, however, asserts that it is not the alcohol in the wine that was beneficial, but antioxidants from grapes.
-If this alternative explanation is correct, then alcohol may not be healthy after all.\\[2ex]
-%
-{\bf Implications for practical NLG systems.} Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
-The question of how a particular piece of quantitative information is best conveyed through language is an important one for practical NLG.
-Years of research on the logic of vagueness in natural language -- which has given rise to such logical techniques as Partial Logic (e.g, \cite{Fine}, Probabilistic Logic \cite{Edgington}, and Fuzzy Logic \cite{Zadeh} -- have primed the research community to expect that the utility of vagueness is an important part of the answer, but our findings call this expectation into question.
-Although our own studies in this article have focussed on vagueness in descriptive Noun Phrases, it seems plausible that vagueness plays a similar role in other linguistic constructs.
-
-For example, consider air temperature, once again. Given a temperature measurement (or prediction) as input, an NLG system that addresses a non-expert audience might say that it was (or will be)
-%
-\begin{quote}
-(a) {\em 27.2 degrees Celsius}, or\\
-(b) {\em approximately 27 degrees}, or\\
-(c) {\em above 25 degrees}, or\\
-(d) {\em warm}, 
-\end{quote}
-%
-among other candidate expressions.
-If the linguistic literature is to be believed, then options (a) and (c) convey crisp information, whereas (b) and (d) are vague (i.e., they permit borderline cases).
-From our own and previous experiments, there is no evidence that the fact in itself that an expression is vague should confer a benefit on it for hearers.
-Rather than asking whether a candidate expression is vague, other questions might shed more light on the NLG system's choice, similar to the ones identified in our studies.
-These questions might focus on the amount of information that a given expression conveys (i.e., on granularity), on the avoidance of numbers, and on the use of evaluative terms.
-Let's see how this might pan out for the examples at hand.
-
-First, the experiments by Mishra et al.
-suggest that it is important how much information is conveyed by an expression, and their findings are echoed by our own thoughts about range reduction.
-In the case of (a)--(d) above, it appears that (a) conveys the most detailed information (designating the smallest segment of the temperature scale), followed by (b).
-Expression (b)  appears to be followed by (d), and (d) by (c) (e.g., 40 degrees is above 25, but at 40 Celsius the word ``warm" may no longer be applicable, giving way to words like ``hot"):
-%
-\begin{quote}
-$a < b < d < c$
-\end{quote}
-%
-If these hunches are correct then an {\em expert} may  to prefer expression (a), because it gives her the most detailed information as a basis for her decisions.
-On the other hand, expression (d) (``warm") is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make ``warm" more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm" (see section 1 and  \cite{peters2009bringing}); the factor of evaluation is especially important if the hearer is unfamiliar with the metric used (e.g., because, being American, they are more used to Fahrenheit), in which case ``warm" is much clearer.
-These considerations suggest that {\em non-experts} might prefer expression (d) over all others.
-If results of this kind were to be confirmed by experiments, they would be applicable in practical NLG, though care would need to be exercised to take the purpose of the communication into account.\footnote{One reason why ``warm" might be less preferred than these considerations suggest is that, in the weather context, some very frequently occurring numbers, such as 20 (degrees), could be relatively easy to process, analogous to the subitizable numbers of \protect\cite{trick1994small}, see our General Methodology section.} In the experiments reported in this paper, the purpose of the utterances investigated was always clear (e.g., instructing people on a weight loss program \cite{Mishra01042011}; informing a patient's choice of hospital \cite{peters2009bringing}; or selecting a dot array (in our own studies).
-% KvD See new footnote above.
-
-One way to see why vagueness (as defined in our Introduction) may not matter as much for NLG -- and for human communication -- as is often thought, is the following thought experiment.
-Suppose a group of speakers understand the word ``warm" as vague, agreeing that temperatures above 26 count as warm, and temperatures below 24 do not count as warm, but considering temperatures between 24 and 26 as borderline cases.
-Now one day these speakers agree to sharpen up their definition deciding that, henceforth, ``warm" means ``$>25$ degrees" (as in (c) above): this decision resolves the borderline cases, while everything else remains the same.
-It seems unlikely that this change in language use, from a vague meaning to a crisp one (i.e., one that has no borderline cases anymore), would lower the utility of the word.
-Our experimental findings are consistent with this idea.
-
-\setcounter{secnumdepth}{0}
-\bibliography{v-special-issue-2016}
-\end{document}

### 2018-08-02 springer/current/general-discussion.tex
massive restructuring Aug 2018

--- 2018-04-23 papers/2018-v-book/v-book-2017.tex
@@
-\documentclass[%
-%			% choose man or doc immediately below
-%doc,	% doc is single spaced
-man,		% man is double spaced
-floatsintext,%
-apacite%
-]{apa6}
-\usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{color}
-\usepackage{booktabs}
-\usepackage{layout}
-\usepackage{apacite}
-\usepackage{datetime}
+To summarise our findings, we asked why strategic vagueness is as frequent as it is and we decided to focus on what we see as the most promising explanation, namely that vague expressions are easy to process by speakers and hearers: the cost hypothesis, as we have called it.
+We decided to test this hypothesis by experimentally investigating whether vague descriptions are resolved by hearers more quickly than crisp ones.
+Although we were able to find some interesting (and statistically highly significant) effects, it appears to us that if our sequence of experiments is assessed as a whole, it cannot be seen as confirmation of the cost hypothesis.
 
-%\renewcommand{\baselinestretch}{1.7} % added 21 Sept 2016, copied from CRM book, but see documentclass option 'man' for "manuscript'' above which makes it double spaced
-
-% MG change p's to 3 dp
-
-\title{The Elusive Benefits of Vagueness \protect \\ Evidence from Experiments}
-\date{\today}
-%\authornote{latex compiled \today ~\currenttime}
-\shorttitle{Is Vagueness Beneficial for Hearers?}
-\twoauthors{Matt Green}{Kees van Deemter}
-\twoaffiliations{Computing Science\\ University of Aberdeen}{Computing Science\\ University of Aberdeen}
-
-\abstract{Much of everyday language is vague, yet the benefits of vagueness for hearers and readers are proving to be elusive.
-We discuss a range of earlier controlled experiments with human participants, and we report on a new series of experiments that we conducted in recent years.
-These experiments, which focus on vague expressions that are part of referential noun phrases, aim to separate the utility of vagueness (as defined by the existence of borderline cases) from the utility of other factors that tend to co-occur with vagueness.
-Having presented the evidence, we argue that the evidence supports a view where the benefits that vague terms exert are due to other influences, rather than to vagueness itself.}
-
-\begin{document}
-\maketitle   
-\setcounter{secnumdepth}{3}
-\section{Introduction}\label{introduction}
-
-%Vagueness pervades the language that we use on a daily basis, and the challenge of understanding vague language has been a prominent concern in many areas of logic and linguistics, involving both theoretical and applied work, including the area known as Natural Language Generation (NLG).
-
-%NLG systems can be thought of a small and imperfect models of human language use, and as such they have the ability to make questions about language very tangible, so let us take the perspective of an NLG system for a moment.
-In a nutshell, NLG systems transform data and formulas into language \cite<e.g.,>[]{reiter2000building}, making decisions between different formulations of the same information.
-For example, if the temperature is 27.2 degrees Celsius, the system could express this as ``27.2 degrees'', ``approximately 27 degrees', ``above 25 degrees", or ``warm'', and the system must decide between these.
-The problem is important in the generation of textual weather reports from numerical weather data such as temperature and wind speed \cite{goldberg1994using,turner2006generating}, and medical decision support on the basis of clinical measurements % such as oxygen saturation, heart rhythm, etc.
-%\cite{Hripcsak01032009, hunter2008summarising, portet2009automatic}.
-Such systems often make decisions concerning the level of precision in the utterances that they generate on the basis of little more than intuition.
-Even when NLG systems are designed to mimic human language use \cite<e.g.,>{konstas2013global} there is no guarantee that these decisions taken by these systems benefit readers.
-A better understanding of the benefits (for readers and hearers) of different precision levels would allow these systems to become more useful.
-%The present article investigates the benefits, or de-benefits, of vagueness.
-
-%.
-In the context of an NLG system faced with a practical decision of this kind, Lipman's question becomes ``Under what circumstances should vague terms be produced?''  Relevant applications include weather forecasting
-
-%Language use may be called vague for various reasons.\footnote{See e.g. the entry ``vague'' in \cite{penguin}.}
-
-In most academic use, the word `vagueness' has a specific meaning.
-Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1997vagueness}, also \citeA{EgreKlinedinst}).
-The crucial criterion is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise'' \cite[p.\ 1]{lipmanvague}.
-A typical example is the word ``tall'', as applied to people for example, because here is no precise, known height which separates those who are tall from those who are not.
-The crucial point is that ``tall'' admits borderline cases (i.e., people who may or may not count as tall), which are the hallmark of vagueness as we use the term.
-
-Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions \cite{Lipman:2000fk, lipmanvague}.
-By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss'' \cite<>[p.~1]{lipmanvague}.
-Lipman explains the point by means of a scenario in which a speaker describes a person to a hearer, who needs to identify that person in the arrivals hall of of an airport.
-In such a scenario, a precise description of the person's height (e.g., ``The person's height is 187.96 cm'') would be more useful than a vague one (``The person is tall'').
-Lipman uses this scenario to explain why standard game theory models of communication \cite<e.g.,>{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs.
-%The relevant conditions are, broadly speaking, that both interlocutors know all the relevant facts (e.g., both know the person's height precisely) and that the setting for the communication is co-operative.
-These conditions exclude deliberate deception, and rhetorical situations like political debate and advertising where the intention is to persuade the interlocutor to adopt some point of view, where the persuasion might not be intended to be in the addressee's best interest.
-
-Lipman argued that the efficiency loss resulting from vague expressions would be unlikely to have arisen unless there are advantages as well as disadvantages associated with vague expressions.
-Lipman asked, essentially, what these advantages might be.
-%, and how they might find a place in a game-theoretical explanation.
-In this paper, we focus on the first part of Lipman's question.
-Several tentative answers to Lipman's question have been offered  \cite<see>{van2009utility, vanDeemterBook}.
-Prominent among these answers is the idea that vague expressions are easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite<e.g.,>[]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
-For example, \citeA<>[p.\ 11]{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze''.
-We shall refer to this as the \emph{cost reduction} hypothesis.
-%The idea of vagueness as cost reduction can take various shapes, but the basic hypothesis is that it is easier for people to think in terms of loosely defined categories (such as ``quite a few'', or ``many'') than in terms of crisply defined ones (such as ``thirteen'', or ``237'').
-%This hypothesis predicts that when a vague expression can perform the same communicative task as a precise one, then, the things being equal, it is rational to choose the vague expression.
-The corollary for comprehension is that vague expressions should be understood more readily than precise expressions.
-
-%The cost reduction hypothesis is of direct relevance to psycholinguists interested in language comprehension, and additionally to psycholinguists interested in language production, for example in connection with the question of audience design \cite{Clark1982287}.
-For, to the extent that speakers and writers choose vague expressions over and above crisp ones because the former are easier to process for hearers than the latter, the cost reduction hypothesis suggests that speakers design their utterance for optimal benefit to their hearers -- out of altruism, so to speak.
-
-Charting the utility of vagueness is the attested aim of a small number of studies, but most of these have focussed on vagueness in a different sense, and focussing on different types of benefits for hearers.
-Two recent studies can illustrate both issues.
-
-In a series of studies of behaviour modification, \citeA<>{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss.
-In the weight loss study, participants were told that the study was designed to test the validity of a new (actually fictitious) health index, the HHI (Holistic Health Index).
-They were told that an ideal HHI score lies in the range of 45 to 55.
-In a longitudinal study, participants submitted their weight to a computer each week.
-Participants were told that two algorithms would be used to compute their HHI, and that the two might give different values initially, in which case the true score lay between the two values.
-In one condition, which the authors called the precise condition, the two algorithms gave the same score.
-In the other condition, which the authors called the vague condition, one algorithm added 3\% to the score while the other algorithm subtracted 3\% from the score, yielding a range of values whose midpoint was the same as the two values given in the precise condition.
-
-One group of participants was given HHI scores in the ideal range: for this group their weight loss did not differ depending on whether they were given vague or precise HHI values.
-However for the other group, who were given HHI scores outside the ideal range, their weight loss was significantly greater if they were given vague HHI scores than if they were given precise HHI scores.
-The authors explain the improvement in the vague condition for this group as resulting from the participants' freedom to think of themselves a positioned on one end of the range - the end closest to the ideal HHI scores.
-This ``illusion of proximity'' \cite<>[p.~4]{Mishra01042011} to the goal is argued to allow participants to generate positive expectancies that lead to behaviours that improve performance.
-In contrast, in the precise conditions, participants did not have this freedom of interpretation, and could not distort the information to bring about the beneficial \emph{illusion of proximity}.
-These results are interesting, and of obvious potential practical importance.
-We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \cite{keefe1997vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases.
-In the terminology of \citeA{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}.
-Furthermore, the experiments of \citeA{Mishra01042011} did not explore benefits in terms of processing cost, but in terms of long-term behaviour change.
-
-Similar issues arise from the work of \citeA{peters2009bringing}.
-The authors carried out a series of studies where participants were required to rate hospitals based on various sources of information about quality of care.
-There was a between-subjects manipulation based on numeracy.
-The format of the information was manipulated within subjects: either numbers only were presented, or both numbers and evaluative categories were presented (e.g., \emph{Poor}, \emph{Fair}, \emph{Good}, \emph{Excellent}, with crisp visual boundary lines between the categories).
-Results showed that, for low-numeracy participants, the presence of evaluative categories resulted in a diminished influence of an irrelevant affective state on the ratings.
-For all participants, the presence of evaluative categories resulted in better decisions and in a greater use of the most important and reliable types of information, such as survival rates.
-
-It is, however, questionable whether the ``evaluative categories" manipulation in this study can be considered a manipulation of vagueness.
-Certainly, terms like \emph{Fair} admit the possibility of borderline cases.
-However, given that the boundaries between the categories were marked crisply, and that therefore the categories mapped crisply to numerical values, it becomes doubtful whether any borderline cases could be conceived to arise in fact.
-For example, \emph{Fair} was mapped to 60\% -- 70\% for the variable \emph{percentage of heart attack patients given recommended treatment (ACE inhibitor)}.
-Accordingly, rather than the vagueness of categories such as \emph{Poor}, Peters et al.
-emphasise the evaluative content inherent in these categories, and the affective potential of the evaluative content rather than the vagueness of the terms like {\em Fair}.
-
-\subsection{Our Approach to the Problem}\label{method}
-%The experiments reported in the present paper put the cost reduction hypothesis to the test. The question that we are trying to answer is whether
-
-Are vague expressions processed more easily by readers than crisp ones?
-Like Lipman, we focus on situations where numerical information is used in order to identify a referent.
-Reference, in other words, will be the communicative task on which we focus, partly because of the interest that this topic has recently drawn in various areas of Cognitive Science \cite{vanDeemterCMR}.
-By looking at one specific type of vagueness, we will be able to investigate the costs and benefits of vagueness relatively thoroughly.
-Whether our findings generalise to other uses of vagueness is a question on which we will speculate in the final section of this chapter.
-
-% KvD Feb 2017 (to address a criticism that came up in the reviews for the special issue):
-We have chosen a narrative strategy in which we address a sequence of four experiments with human reader chronologically, explaining how each experiment helped us refine our research question.
-In order to do justice to our findings, we need to describe these experiments in a fair amount of detail.
-
-Let us start by explaining the task that was given to the participants in our experiments.
-We used a {\em speeded forced choice} task to compare the processing costs of different references to quantities.
-In this context, speed and accuracy of responses are the key dimensions on which the different references can be compared.
-Each stimulus in the experiments was a set of dot arrays containing various number of dots, together with a preceding instruction (in the form of a referring expression) to choose one of the arrays with respect to its cardinality.
-The participant was asked to respond as quickly as possible while avoiding errors.
-We manipulated the instructions and the arrays in several ways across the four experiments.
-
-% MG General methodology moved to here: per-experiment method sections only talk about deviations from this general method
-All the experiments shared the following properties: 
-Stimuli were created using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.
-The position of the dots was randomised per-trial.
-The order in which trials were presented was randomised per-participant. 
-There were 256 trials, presented in 4 blocks of 64 each, between which the participant could rest. 
-A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants and recorded responses. 
-Participants were recruited using email lists at the University of Aberdeen, and paid ten pounds for participating.
-All participants self-reported fluency in English, and had normal, or corrected-to-normal vision.
-The experiment was conducted in a quiet room. 
-Participants were asked to respond as quickly as possible while avoiding errors.
-There was a block of practice trials after which participants could ask any questions, following which the experimenter left the room.
-All $p$ values reported for linear models were calculated using the R package \emph{lmerTest} \cite{lmerTest}. The complete data and analysis of our first experiment are available from 
-{\tt http://homepages.abdn.ac.uk/k.vdeemter/pages/vagueness-data-analysis/e1.html}, and analogous for experiments 2-4.\\[2ex]
-
-% KvD I've moved the following two paragraphs towards the end of the section.
-% MG added/moved out of expt 1
-
-\noindent When the distance grows between two numbers, they become more easily distinguishable: the \emph{numerical distance effect} has been shown for comparing the cardinality of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene1996}.
-We manipulated the number of dots in each array such that some sets of arrays had smaller numerical distances and others had larger numerical distances.
-Where a number was mentioned in the instructions, it was always in the form of an Arabic numeral.
-% MG added/moved out of expt 1
-When two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}.
-We controlled which side the smaller number appeared on to avoid systematic influences of this effect.
-
-% KvD added the following from a previous version of the paper:
-There is abundant evidence \cite<e.g.,>{trick1994small} that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities.
-We performed a pilot experiment \cite{our-earlier-paper} in which we were able to confirm this finding in the experimental settings on which we are focussing in this paper.
-We found that, when participants were confronted with % MG changed the next line and added the footnote to remove the doubts we had about whether a dot array must have regular layout, or can have irregular layout - I'm still not sure this is a good description though
-a stimulus consisting of two squares containing different numbers of dots\footnote{Such a stimulus is referred to hereafter as consisting of a set of {\em dot arrays}.
-The number of dots in an array is referred to as its cardinality.
-The physical arrangement of dots in each array is irregular.},
-instructions of the form \emph{Choose the square with \emph{n} dots} led to consistently faster response times than instructions of the form \emph{Choose the square with many/few dots} when $2 \leq n \leq 5$; the converse was true for $n>5$.
-Given these findings, we henceforth focussed our studies on non-subitizable numbers, because it is there that vagueness is expected to have benefits.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 1                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 1: initial experiment}
-
-We used a forced choice task to elicit responses to crisp or vague forms of instructions that required the participant to choose one of two dot arrays by referring to its cardinality.
-The participant was presented with an instruction in the form \emph{Choose the square with \ldots dots}.
-Then a set of two dot arrays was presented, each in the form of a square containing some number of dots.
-The participant was required to identify the dot array that corresponded with the instruction, by pressing the appropriate key, as quickly as possible while avoiding errors. 
-Response time and accuracy were recorded for analysis.
-
-We manipulated how discriminable the dot arrays were by varying the numerical distance between them. 
-One array always contained 25 dots: the other contained either 5, 10, 15, 20, 30, 35, 40, or 45 dots. 
-This gave us numerical distances of 5, 10, 15, and 20, with smaller numerical distances resulting in less discriminable arrays and larger distances resulting in more discriminable arrays.
-
-Our main manipulation was of the vagueness of the instruction, with two levels, \emph{crisp} and \emph{vague}.
-Assuming the dot array [5, 25], and the instruction referring to the smaller cardinality, the \emph{crisp} instruction was \emph{Choose the square with 5 dots} and its \emph{vague} counterpart was \emph{Choose the square with few dots}.
-
-\subsection{Hypothesis}
-\noindent The main aim of this experiment was to test the following hypothesis:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] There exists a main effect advantage for vagueness: vague instructions impose a lower cognitive load for the comprehender than crisp alternatives.
-\end{APAenumerate}
-}
-
-\subsection{Experimental Method}
-
-%Twenty participants were recruited, aged between 18 and 45 with a median age of 26.
-On each trial a participant was presented with an instruction to choose one of two dot arrays on screen by reference to its cardinality. 
-Following a keypress to indicate that the participant had read the instruction there was a central fixation cross for 1000 ms, and a blank screen for 500 ms, followed by the array (without repetition of the referring expression).
-The arrays would stay on screen until the participant responded (there was no timing-out). 
-Response time was measured as the latency between the presentation of the arrays, and the keypress identifying the choice: in this way, the response time was separated from time spent reading the instructions, which is important since we are only interested in the former.
-A response was counted as erroneous if the square with the wrong number of dots was chosen (when the instruction contained a number); 
-if the square with the larger number of dots was selected (when the instruction was \emph{Choose the square with few dots}); 
-or if the square with the smaller number of dots was selected (when the instruction was \emph{Choose the square with many dots}). 
-No feedback was given on correct trials, but there was feedback on error trials in the form of the word ``\textsc{wrong!!}'' which flashed on screen.
-Hypothesis H1 was tested looking both at response times and at error rates.
-
-\subsection{Results}
-
-\begin{figure}[htbp]
-\fitfigure{images/e1RTandErrorRatesplot-1}
-\caption{Experiment 1 results: response times and error rates.}
-\label{resultse1}
-\end{figure}
-
-%\subsubsection{Response times}
-{\bf Response times} (RTs) for trials with erroneous responses were discarded, leading to the loss of 354 trials from 5120, representing 6.9\% of the trials. 
-The correct response RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of a further 160 trials, or 3.4\% of the remaining correct responses. 
-Means for response times and error rates are given in Fig. (\ref{resultse1}).
-A linear mixed model of RT was built, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
-Analysis confirms hypothesis H1:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1$_{rt}$)] RTs were faster for vague instructions than for crisp instructions 
-	($\beta=-0.11$, $se=.02$, $t=4.7$, $p<.001$).  % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-%\subsubsection{Error rates}
-{\bf Error rate} data were analysed using a generalized logit mixed model \cite{jaeger2008categorical}, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
-We found:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1$_{er}$)] There was a statistically non-significant effect of vagueness on error rates: the vague instructions tended to have a lower error rate than the than the crisp instructions 
-	($\beta=-.22$, $se=.18$, $t=-1.2$, $p=.230$).  % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-\subsection{Discussion}
-
-\noindent As expected, responses were faster and more accurate for vague than crisp instructions. Post hoc, we formulated some further hypotheses:
-
-{\small
-\begin{APAenumerate}
-	\item [(H2)] A main effect advantage of increasing numerical distance: the task will become easier as the numerical distance increases, because the two arrays are then more discriminable.
-	\item [(H3)] An interaction between vagueness and numerical distance: i.e., any facilitation for vagueness should be greater at smaller numerical distances than at larger numerical distances.
-\end{APAenumerate}
-}
-
-Looking at response times, these two post-hoc hypotheses were supported:
-
-{\small
-\begin{APAenumerate}
-	\item [(H2$_{rt}$)] RTs increased with numerical distance 
-	($\beta=-0.18$, $se=0.01$, $t=-11.6$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-	Since discriminability of the arrays is easier for larger numerical distances, discriminability probably underlies this effect. 
-	\item [(H3$_{rt}$)] Numerical distance and vagueness interacted significantly: essentially there were diminishing returns for vagueness as numerical distance increased: 
-	($\beta=0.07$, $se=0.01$, $t=5.9$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-Looking at error rates, the two post-hoc hypotheses were supported as well:
-
-{\small
-\begin{APAenumerate}
-	\item [(H2$_{er}$)] Error rates decreased as numerical distance increased: 
-	($\beta=-0.78$, $se=0.08$, $z=-9.6$, $p<0.0005$).% MG added the values for these stats on 25 Sept 2016
-	\item [(H3$_{er}$)] Numerical distance and vagueness interacted significantly: essentially there were diminishing returns for vagueness as numerical distance increased until the biggest numerical distance when vagueness was disadvantageous:
-	($\beta=0.80$, $se=0.15$, $z=5.27$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-So far, these results are in line with the idea of cost reduction.
-The idea of cost reduction explains the vagueness advantage by claiming that the vague referring expressions place less cognitive load on the comprehender than the crisp referring expressions.
-It explains the diminishing returns for vagueness in more-discriminable stimuli (i.e., the vagueness by numerical distance interaction) by claiming that load is low in both conditions for the easily-discriminable stimuli, and that therefore there is no extra benefit to be had from vagueness in the easily-discriminable stimuli.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 2                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 2: separating vagueness from instruction format}
-
-% KvD: I've added some syntactic sugar, in a number of places, to clarify the relation between Ex 1 and Ex 2.
-
-However, the picture painted by these findings might be misleading.
-First of all, there is a possibly confounding factor.
-Contrast, from Experiment 1, an expression from the vague condition: `the square with few dots' with an expression from the crisp condition: `the square with 5 dots'.
-One difference is that `few' has the potential for vagueness, whereas `5' is crisp. But
-another difference is that `few' is verbal while `5' is numerical, in the sense that a number is mentioned explicitly. 
-Since these two differences could not be separated in Experiment 1, the vagueness advantage finding is vulnerable to an alternative interpretation, that what we saw as a vagueness advantage was in contrast an advantage for the verbal form of the quantifier.
-In Experiment 2 we therefore created verbal and numeric versions of each of the vague and crisp instructions so that we could compare vague and crisp conditions while taking account of verbal / numeric format.
-
-Another potential problem with Experiment 1 is the following.
-Participants chose one of two squares: therefore the `vague' quantifiers (e.g., `few') uniquely identified one square.
-Recall our definition of vague  -- ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise''.
-In Experiment 1, the quantifiers in the vague conditions did not realise their potential for vagueness.
-This is because there were no borderline cases of the referent that could make the referent set `not well-defined', and perhaps because using definite articles in the instructions implied that only one option was correct.
-Using error feedback in Experiment 1 could have exacerbated this.
-
-To find out what happens when words are used in a context where their potential for vagueness comes to the fore, Experiment 2 used three arrays (rather than two) so that the vague description had more than one possible referent, and used indefinite articles to avoid the impression that only one response counted as correct, and was carried out without error feedback.
-An indication that the potential for vagueness was realised in Experiment 2 is that the borderline response was chosen fairly often: 16\% of the time.
-
-In Experiment 2, an item was a referring expression instruction followed by a set of three dot arrays defined by a triple of numbers, representing the number of dots in the left, middle, and right arrays.
-We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54).
-Each set of arrays comprised three arrays (instead of two as in Experiment 1); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had more.
-
-Examples of crisp and vague versions of the numerical and verbal instructions follow: the examples assume the array (6,15,24) and reference to the smaller number of dots, such that 6 was classified as the expected response; 15 as the borderline response; and 24 as the incorrect response.
-In the {\em vague numerical} condition we used \emph{Choose a square with about 10 dots}.
-None of the squares contained 10 dots.
-10 is slightly closer to 6 than to 15, justifying 6 as the best response and 15 as the borderline response.
-In the {\em vague verbal} condition we used \emph{Choose a square with few dots}.
-In the {\em crisp numerical} condition we used \emph{Choose the square with 6 dots}, and one square always did contain the number mentioned.
-For {\em crisp verbal}, we used \emph{Choose the square with the fewest dots}.
-
-% KvD I've re-instated the table below, because this makes it easier to understand the instruction format/selection algorithm confound.
-
-\begin{table}[htbp]
-\caption{Experiment 1: instructions arranged by condition for the dot triple $(6,15,24)$ and instructions indicating the smaller numbers}
-\label{instre3}
-\begin{tabular}{rll}
-&crisp&vague\\
-\midrule
-numerical	&	Choose the square with 6 dots 			& 	Choose a square with about 10 dots	\\
-verbal	&	Choose the square with the fewest dots	&	Choose a square with few dots		\\
-\midrule
-\end{tabular}
-\end{table}
-
-\subsection{Hypotheses} % E2
-\noindent We formulated the following hypotheses for Experiment 2:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1)] There exists a main effect RT advantage for vagueness.
-	\item [(H2)] An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
-	\item [(H3)] No large main effect of instruction format (i.e., numeric versus verbal) since we hypothesised that vagueness rather than instruction format drove the effect in Experiment 1.
-	\item [(H4)] On the basis of Experiment 1, we expect faster responses for stimuli with more discriminable arrays.(i.e., an effect of item).
-	\item [(H5)] Participants should make more borderline case choices for vague than crisp instructions.
-\end{APAenumerate}
-}
-
-\subsection{Experimental Method} % E2
-% Participants were aged between 18 and 45 with a median age of 28. 
-We manipulated as independent variables vagueness and instruction format (i.e., numeric versus verbal instructions), yielding four conditions, \emph{vague numeric}; \emph{vague verbal}; \emph{crisp numeric}; \emph{crisp verbal}.
-We measured two dependent variables: response time; and the probability of a participant choosing the borderline case.
-On each trial, first the referring expression that constituted the instruction for that trial was displayed. 
-Participants then pressed a key to indicate that they had read the instruction. 
-After 1000 ms, the arrays were presented, while preserving the text of the referring expression. 
-The response time dependent variable was measured from the presentation of the arrays, until the keypress indicating the participant's choice, which was also recorded. 
-The trial would timeout after 60 seconds if there was no response.
-In this experiment, no feedback was given. 
-This was because, in the vague conditions, we did not regard any response as `correct' or `incorrect', but instead as `borderline response', or `not borderline response', and we did not want to draw participants' attention to this distinction explicitly.
-We simply recorded whether the participant chose the best referent, the borderline case or the poorest referent, and how long it took the participant to respond.
-
-\subsection{Results} % E2
-
-%\subsubsection{Response times} % E2
-% MG added the values for these stats on 26 Sept 2016
-Means for response times and proportion of borderline responses are given in Fig. (\ref{resultse2}). 
-Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data. 
-A linear mixed model was constructed for the (logged) response times, 
-with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects. We found:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against Hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06$, $se=.01$, $t=4.6$, $p<0.0001$).
-	\item [(H2)] In focussed comparisons, vagueness was significantly disadvantageous in the \emph{numeric} conditions ($\beta=0.09$, $se=0.02$, $t=4.5$, $p<0.0001$), and non-significantly disadvantageous in the \emph{verbal} conditions ($\beta=0.02$, $se=0.02$, $t=1.5$, $p=0.1550$), offering partial evidence against Hypothesis 2.
-The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to a significant overall interaction effect between vagueness and instruction format ($\beta=-0.13$, $se=0.02$, $t=-6.6$, $p=0.0169$).
-	\item [(H3)] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1, but suggesting that instruction format rather than vagueness drove the effect we observed in Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; ($\beta=-0.36$, $se=0.07$, $t=-5.1$, $p<0.0001$).
-	\item [(H4)] There was a significant main effect of item (i.e., of which triple of numbers of dots were used in the stimulus): ($\beta=0.12$, $se=0.02$, $t=7.1$, $p<0.0001$).
-This effect seems likely to be due to the very fast responses for stimuli using arrangements of the smallest numbers of dots (6,15,24), which also had the largest difference in ratio of smallest number to largest number in the stimulus, suggesting that stimuli using these numbers of dots may have been particularly discriminable for participants.
-\end{APAenumerate}
-}
-
-%\subsubsection{Borderline responses} % E2
-% MG added the values for these stats on 26 Sept 2016
-A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
-The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{resultse2}.
-Participants chose the borderline square on $16.6\%$ of trials overall.
-{\small 
-\begin{APAenumerate}
-	\item [(H5)] Participants were significantly more likely to choose the borderline option for vague instructions than for crisp instructions (21.9\% vs 11.3\%: $\beta=0.62$, $se=0.22$, $z=2.8$, $p=0.0059$).
-Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%: $\beta=-3.35$, $se=0.23$, $z=-14.6$, $p<0.0001$).
-\end{APAenumerate}
-}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{images/e2-rtplot-1.pdf}
-\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{images/e2-blBarChart-1}
-\caption{Experiment 2 results: mean response times by condition and item, and counts of borderline case responses by condition.}
-\label{resultse2}
-\end{figure}
-
-\subsection{Discussion} % E2
-
-Experiment 2 tested to see whether when borderline cases are present, vague instructions would speed responses as they did in Experiment 1 when there were no borderline squares. 
-We actually found a small (but statistically significant) \emph{disadvantage} of vague instructions: vague instructions slowed people down by 112 ms on average.
-We also found that the effect of instruction format was significant, with numerical format slowing responses by 689 ms on average, such that the disadvantage of numerical format overwhelmed the contribution of vagueness.
-The \emph{verbal vague} condition was still responded to faster than the \emph{numerical crisp} condition, so the pattern from Experiment 1 was reproduced, but in the light of the evidence from Experiment 2, in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like an advantage of verbal instruction format.
-
-% KvD I've expanded the following:
-However, once again there is a possibly confounding factor.
-Observe that, in Experiment 2, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) {\bf selection algorithm}: To see this, consider the task of selecting the dot array that contains ``few dots": to do this, it suffices to {\em compare} the three arrays and select the one that contains the fewest elements.
-To select the dot array that contains "16 dots" seems to require the participant to estimate, and then {\em match}, the cardinality of (at least) one dot array to 16, a process which could plausibly take longer, independently of vagueness.
-Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison ``algorithm".% (which is known to be faster than matching).
-% KvD I think we need to say more about this, citing some appropriate papers to support this idea.
-\textcolor{red}{{\bf--Insert citation}}
-
-In the next two experiments we pitted the comparison algorithm and matching algorithm selection tasks against each other while controlling vagueness and instruction format.
-In Experiment 3 we restricted all the instructions to \emph{numeric} quantifiers while factorially manipulating vagueness and selection task.
-In Experiment 4 we ensured that all instructions used \emph{verbal} quantifiers, while also factorially manipulating vagueness and selection task.
-This allowed us to distinguish between the predictions of the selection task account and the instruction format account.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 3                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 3: focus on instructions that contain numerals}
-
-The main aim of Experiment 3 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies.
-The main changes from Experiment 2 were that the human selection task was explicitly controlled (i.e., in whether it amounted to matching or comparison), and that all conditions were constrained to mention a number.
-We used the same arrays as in Experiment 2.
-Table \ref{Instructions for e3} shows the instructions for each condition.
-Note the difference between ``fewer than 20" and ``far fewer than 20": whereas the former cannot have borderline cases (i.e., for each number it is clear whether the number is smaller than 20 or not), the latter can.
-
-%KvD I think we should motivate our choice of materials (especially "far ..er")
-%\textcolor{red}{{\bf KvD I think we should motivate our choice of materials (especially "far ..er")}}
-% DONE 28 Sept 2017
-
-\begin{table}[htbp]
-\centering
-\caption{Experiment 3: Instructions arranged by condition for stimuli with $(6,15,24)$ dots and instructions indicating the smaller numbers} 
-\label{Instructions for e3}
-\begin{tabular}{llll}
-instruction				&vagueness			&	selection		&	instruction							\\
-format				&					&	task			&									\\
-\toprule
-\multirow{4}{*}{numeric}	&\multirow{2}{*}{crisp} 	&	matching		& Choose a square with 6 dots 				\\ 
-					&					&	comparison 	& Choose a square with fewer than 20 dots 	\\
-					\cline{2-4}
-					&\multirow{2}{*}{vague} 	&	matching 		& Choose a square with about 10 dots 		\\ 
- 					&					&	comparison 	& Choose a square with far fewer than 20 dots 	\\ 
-\bottomrule
-\end{tabular}
-\end{table}%
-
-\subsection{Hypotheses} % E 3
-\noindent For Experiment 3, we hypothesised:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
-	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection task)
-	\item [(H3)] Effects of vagueness are different depending on whether the selection task is matching or comparison (interaction effect selection x vagueness).
-\end{APAenumerate}
-}
-
-\subsection{Experimental Method} % E 3
-
-38 participants were recruited. We used a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e3}).
-On each trial an instruction was presented: participants pressed a key to dismiss the instruction, when the dot arrays were presented until the participant responded, and the response time and choice were recorded.
-
-\subsection{Results} % E 3
-
-Response times were trimmed at 2.5 SD separately for each subject, leading to the loss of 204 trials (2.8\% of the trials).
-Condition means for the remaining (logged) RTs are plotted in Figure \ref{resultse3}.
-A linear mixed model was constructed for the logged response times, 
-with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
-
-\begin{figure}[htbp]
-\centering
-\fitfigure{images/e3-rtplot-1.pdf}
-%\includegraphics[trim = 20mm 0mm 20mm 0mm, clip, width=.75\textwidth]{images/e3-rtplot-1.pdf}
-\caption{Mean response times by condition for Experiment 3 where all instructions were numeric}
-\label{resultse3}
-\end{figure}
-
-% MG added the values for these stats on 26 Sept 2016
-{\small
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions were non-significantly easier for the reader than crisp instructions when vagueness was considered as a main effect ($\beta=-0.01$, $se=0.01$, $t=-0.4$, $p=0.7158$). 
-	\item [(H2)] There was a statistically significant main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=0.16$, $se=0.03$, $t=6.2$, $p<0.0001$).
-	\item [(H3)] Vagueness exerted effects in different directions for the comparison task and for the matching task: there was a significant interaction between vagueness and selection task ($\beta=0.13$, $se=0.03$, $t=4.2$, $p<0.0001$).
-Separate analyses were conducted testing for effects of vagueness at each level of the selection task.
-Within the comparison task vagueness significantly speeded response times compared with crisp controls ($\beta=-0.07$, $se=0.02$, $t=-3.5$, $p<0.0012$). 
-Within the matching task vagueness significantly \emph{slowed} response times compared with crisp controls ($\beta=0.06$, $se=0.02$, $t=2.9$, $p<0.0061$). 
-\end{APAenumerate}
-}
-% MG edited this 27 Sept 2016:
-%The cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
-In other words, the cost reduction account was wrong to predict significant main effect advantages for vagueness (although there was a non-significant trend in the \emph{direction} predicted by the cost reduction account), and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was significantly advantageous in the comparison task.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 4                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 4: focus on instructions that do not contain numerals} % E 4
-
-This experiment mirrors Experiment 3, but focussing on instructions that did not use a number. 
-We manipulated vagueness and the selection task (comparison and matching). 
-In order to implement the experiment without mentioning numbers in the instructions, we changed the sequence of each trial to include a `target' (i.e., a dot array of a particular cardinality) before the instruction, so that we could then refer back to the target's cardinality in the instruction using expressions like \emph{the same number of dots as the target}; \emph{fewer dots than the target}.
-%
-This presentation of a target before the main body of the trial shares some features with \citeA[Experiment 2]{Izard20081221}, although in that experiment participants were told the cardinality of the target (called an \emph{inducer} in that paper) whereas in our experiment we did not tell participants the cardinality of the prime array.
-An item was thus a combination of a target dot array, an instruction that did not contain a number, and a set of dot arrays taking their cardinalities from the same triples used in Experiment 2.
-Table \ref{Instructions for e4} spells out how the instructions were constrained not to mention a numeral and gives examples of targets.
-
-\begin{table}[htp]
-\caption{Instructions and targets by condition for Experiment 4}
-\begin{center}
-\begin{tabular}{llcrl}
-Item					&Selection				&Vagueness	&Target	&Choose a square with\ldots						\\
-\toprule
-\multirow{4}{*}{06:15:24} 	&\multirow{2}{*}{Comparison} 	&Crisp		&20		&\ldots fewer dots than the target					\\
-\cline{3-5}
-					&						&Vague		&20		&\ldots far fewer dots than the target					\\
-					\cline{2-5}
-					&\multirow{2}{*}{Matching}		& Crisp		&6		&\ldots the same number of dots as the target			\\
-					\cline{3-5}
-					& 		 				&Vague		&10 		&\ldots about the same number of dots as the target		\\
-					\bottomrule
-\end{tabular}
-\end{center}
-\label{Instructions for e4}
-\end{table}%
-
-\subsection{Hypotheses}% E 4
-\noindent For Experiment 4, we hypothesised:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
-	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection)
-	\item [(H3)] Effects of vagueness differ depending on whether selection is matching or comparison (interaction effect selection x vagueness, and focussed comparisons at each level of selection).
-\end{APAenumerate}
-}
-
-\subsection{Results}% E 4
-40 volunteers participated.
-The results showed that vagueness was beneficial for comparison but detrimental for matching (the same as Experiment 3) even when no numbers were allowed in the instructions. 
-Figure \ref{resultse4} shows the means by condition. Our findings were as follows:
-
-% MG added the values for these stats on 27 Sept 2016
-{\small
-\begin{APAenumerate}
-	\item [(H1)] There was no significant main effect of vagueness ($\beta=-0.02$, $se=0.01$, $t=-1.5$, $p=0.1296$). 
-	\item [(H2)] There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=0.18$, $se=0.02$, $t=10.4$, $p<0.0001$).
-This effect was in the same direction as Experiment 3.
-	\item[(H3)] Vagueness did exert different effects depending on the selection task (main interaction effect of vagueness by selection $\beta=0.12$, $se=0.02$, $t=5.1$, $p<0.0001$). 
-Separate analyses of the effect of vagueness were conducted for the comparison task and for the matching task using Bonferroni-adjusted significance thresholds. 
-In the comparison task, vagueness resulted in faster response times ($\beta=-0.08$, $se=0.02$, $t=-4.3$, $p<0.0001$). 
-In the matching task vagueness slowed response times ($\beta=0.05$, $se=0.01$, $t=3.7$, $p=0.0004$). 
-These results are in the same direction as Experiment 3.
-\end{APAenumerate}
-}
-
-Once again, the cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
-
-\begin{figure}[htbp]
-\centering
-\fitfigure{images/e4-rtplot-1.pdf}
-\caption{Mean response times by condition for Experiment 4}
-\label{resultse4}
-\end{figure}
-
-\section{Discussion of experiments 3 and 4}
-
-The main aim of these two experiments was to test whether vagueness confers any cognitive benefits over and above those due to differences in the selection task according to whether the instruction mandates a \emph{comparison} selection task or a \emph{matching} selection task, when number-use is held constant.
-The main effect of selection task showed that the assumption that the \emph{comparison} task is easier than the \emph{matching} task is well-founded.
-In both experiments people were reliably quicker to respond in the \emph{comparison} task.
-
-Vagueness, which was the phenomenon on which our investigation focussed, did not exert a significant main effect in response time.
-However when the comparison and selection tasks were analysed separately, there was small significant advantage for vagueness in the \emph{comparison} tasks, but a small significant disadvantage for vagueness in the \emph{matching} tasks.
-
-\section{General Discussion}\label{discussion}
-
-To summarise our findings, Experiment 1 showed that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment could not distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format; and mandated a comparison strategy rather than a matching strategy.
+To explain why this is, let us summarise our findings so far: Experiment 1 showed that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment was unable to distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format, and they mandated a comparison strategy rather than a matching strategy.
 Experiment 2 showed that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional benefit in either the verbal format instructions or the numerical format instructions.
-
 However, Experiment 2 did not distinguish benefits of number avoidance from benefits of the comparison selection task.
@@
 Within the two experiments we found that vagueness exerts benefits when the selection task is \emph{comparison}, but not when the task is \emph{matching}.
+
+What is one entitled to conclude?
+Given that we were able to identify a class of situations in which vague expressions led to faster response times than crisp ones, would it be valid to conclude that we have finally discovered an advantage for vagueness that cannot be ascribed to some other factor?
+We believe the answer to this question is negative.
 
@@
 \begin{tabular}{cccc}
-selection task 				& vagueness		& candidates	& effect of vageueness	\\
-\toprule
+\hline\noalign{\smallskip}
+selection task 					& vagueness		& candidates	& effect of vagueness						\\
+\noalign{\smallskip}\hline\noalign{\smallskip}
 \multirow{ 2}{*}{comparison} 	& crisp 		& 2				& \multirow{ 2}{*}{vagueness advantage}  	\\
@@
 								& vague			& 1				&                                           \\
-\midrule
-\multirow{ 2}{*}{matching} 		& crisp 			& 1 			&\multirow{ 2}{*}{vagueness disadvantage}	\\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+\multirow{ 2}{*}{matching} 		& crisp 		& 1 			& \multirow{ 2}{*}{vagueness disadvantage}	\\
 \cline{2-3}
 								& vague			& 2				&								 			\\
-\bottomrule
+\noalign{\smallskip}\hline
 \end{tabular}
@@
 
-What is one entitled to conclude?
-Given that we were able to identify a class of situations in which vague expressions led to faster response times than crisp ones, would it be valid to conclude that we have finally discovered an advantage for vagueness that cannot be ascribed to some other factor?
-We believe the answer to this question is negative.
-
-To see why, consider Figures \ref{resultse3} and \ref{resultse4}.
+To see why, consider Figures 3 and 4.
 Both figures depict four conditions, depending on whether the expression was crisp or vague, and depending on whether the referent could be identified using a comparison strategy or not.
 Two of the resulting four conditions result in an expression that can denote either of two referents; the other two conditions result in an expression that can only denote one referent, with the other possible referent being a marginal candidate at best.
-
-% KvD: I've shortened the text, removing duplication
 
@@
 %
-\noindent Looking at the entire series of experiments, our findings suggest that the observed benefits of vague expressions in certain situations may be due to factors other than vagueness: factors like avoiding numbers; permitting comparison tasks; and range reduction.
-The picture that is starting to emerge is subtle: on the one hand, in the situations that we have been studying
-%-- where cooperative speakers refer to an object (e.g., a square) by means of some quantity associated with the object -- 
-vagueness is not intrinsically beneficial.
-On the other hand, vague expressions frequently possess other features that {\em are} beneficial, and these are what give us the incorrect impression that vagueness itself is beneficial.
+\noindent Looking at the entire series of experiments, our findings suggest that the observed benefits of vague expressions may be due to factors other than vagueness: factors like avoiding numbers; permitting comparison tasks; and range reduction.
+The picture that is starting to emerge is subtle: on the one hand, in the situations that we have been studying vagueness is not intrinsically beneficial.
+On the other hand, vague expressions frequently possess other features that \emph{are} beneficial, and these are what give us the incorrect impression that vagueness itself is beneficial.
 Vagueness may thus have acquired a reputation that it does not deserve.
 The answer to Lipman's question, of why vagueness permeates human language (see our Introduction), may lie in a different direction after all, possibly relating to benefits for the speaker rather than the hearer.
-% \cite<see>[chapter 11, for discussion of some possible directions]{vanDeemterBook}.
 
@@
 Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
-Years of research on the logic of vagueness -- giving rise to such techniques as Partial Logic \cite<e.g.,>{Fine}, Probabilistic Logic \cite{Edgington}, and Fuzzy Logic \cite{Zadeh} -- have primed the research community to expect that some special utility of vagueness is an important part of the answer, but our findings call this expectation into question.
+Years of research on the logic of vagueness -- giving rise to such techniques as Partial Logic \citet[e.g.,]{Fine}, Probabilistic Logic \citet{Edgington}, and Fuzzy Logic \citet{Zadeh} -- have primed the research community to expect that some special utility of vagueness is an important part of the answer, but our findings call this expectation into question.
 
-Although our own studies in this article have focussed on vagueness in descriptive Noun Phrases, it seems plausible that vagueness plays a similar role in other linguistic constructs.
-For example, consider reports on air temperature.
-Given a numerical temperature measurement or prediction, we might word it as
+Although our own studies in this article have focussed on vagueness in descriptive Noun Phrases only, it seems to us plausible that vagueness plays a similar role in other
+linguistic constructs. For example, consider reports on air temperature. Given a numerical
+temperature measurement or prediction, we might word it as 
 %
@@
 If the linguistic literature is to be believed, then options (a) and (c) convey crisp information, whereas (b) and (d) are vague (i.e., they permit borderline cases).
-From our own experiments and the ones discussed in section 1, there is no evidence that  vagueness should be beneficial for hearers.
+In the situations studied in our own experiments and the ones discussed in section 1, we found no evidence that vagueness is beneficial for hearers.
 Rather than asking whether a candidate expression is vague, other questions might shed more light on the choice, similar to the ones identified in our studies.
 These questions might focus on the amount of information that a given expression conveys (i.e., on granularity), on the avoidance of numbers, and on the use of evaluative terms.
-Let's see how this might pan out for the examples at hand.
+Let's see how this might pan out for the above examples from the weather domain.
 
 First, the experiments by \citeauthor{Mishra01042011} suggest that it is important how much information is conveyed by an expression, and their findings are echoed by our own thoughts about range reduction (following Experiments 3 and 4).
-In the case of (a)--(d) above, it appears that (a) conveys the most detailed information (designating the smallest segment of the temperature scale), followed by (b), then (d), then (c) (e.g., 40 degrees is above 25, but at 40 Celsius the word ``warm" is likely to give way to ``hot" or ``scorching"):
+In the case of (a)--(d) above, it appears that (a) conveys the most detailed information (designating the smallest segment of the temperature scale), followed by (b), then (d), then (c) (e.g., 40 degrees is above 25, but at 40 Celsius the word ``warm'' is likely to give way to ``hot'' or ``scorching''):
 %
@@
 %
-If these hunches are correct then an {\em expert} may prefer to read expression (a), because it gives her the most detailed information on which to base her decisions.
-On the other hand, expression (d) (``warm") is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make (d) more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm" (see our section 1 and  \citeA{peters2009bringing}), which is especially important if the hearer is unfamiliar with the metric used.
+If these hunches are correct, then it seems to us that it is relatively unimportant whether a given expression is vague or crisp.
+Other factors seem more important; moreover, it may depend on the task and the audience which of a-d is preferred.
+For example, an \emph{expert} may prefer to read expression (a), because it gives her the most detailed information on which to base her decisions.
+On the other hand, expression (d) (``warm'') is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make (d) more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm'' (see our section 1 and  \citeauthor{peters2009bringing}), which is especially important if the hearer is unfamiliar with the metric used.
 These considerations suggest that {\em non-experts} might prefer expression (d).
-%If results of this kind were to be confirmed by experiments, they would be applicable in practical NLG, though care would need to be exercised to take the purpose of the communication into account.\footnote{One reason why ``warm" might be less preferred than these considerations suggest is that, in the weather context, some very frequently occurring numbers, such as 20 (degrees), could be relatively easy to process, analogous to the subitizable numbers of \protect\citeA{trick1994small}, see our section \ref{method}.
-Note that in the experiments reported in this paper, the purpose of the utterances investigated was always clear (e.g., instructing people on a weight loss program \cite{Mishra01042011}; informing a patient's choice of hospital \cite{peters2009bringing}; or selecting a dot array (in our own studies).}
 
 One way to see why vagueness (as defined in our Introduction) may not benefit human communication is the following thought experiment.
-Suppose a group of speakers understand the word ``warm" as vague, agreeing that temperatures above 26 count as warm, and temperatures below 24 do not count as warm, but considering temperatures between 24 and 26 as borderline cases.
-Now one day these speakers agree to sharpen up their definition deciding that, henceforth, ``warm" means ``$>25$ degrees" (as in (c) above): this decision resolves the borderline cases, while everything else remains the same.
+Suppose a group of speakers understand the word ``warm'' as vague, agreeing that temperatures above 26 count as warm, and temperatures below 24 do not count as warm, but considering temperatures between 24 and 26 as borderline cases.
+Now one day these speakers agree to sharpen up their definition deciding that, henceforth, ``warm'' means ``$>25$ degrees'' (as in (c) above): this decision resolves the borderline cases, while everything else remains the same.
 It seems unlikely that this change in language use, from a vague meaning to a crisp one (i.e., one that has no borderline cases anymore), would lower the utility of the word.
-Our experimental findings are consistent with this idea.
-
-% KvD Bracketed out the following for now:
-%A more radical perspective on our findings might say that the distinction between crisp and vague expressions, which is taken for granted in the bulk of the linguistic literature, is mistaken, because expressions such as ``above 25 degrees" (and ``fewer than 20 dots") are not equally felicitous in all situations in which they can be truthfully applied.
-For example, when the temperature is 40 degrees, then a weather report saying that it is ``above 25 degrees" would be correct yet misleading: this description might be felicitous as a description of temperatures between (roughly) 26-28 degrees, borderline felicitous for 29-33 degrees, and infelicitous at even higher temperatures.
-At its most bold, this perspective would assert that the applicability of {\em every} linguistic expression has borderline cases, hence Lipman's question (see section 1) simply does not arise.
-This radical perspective can only be indicated here, but not discussed at length.
-
-\setcounter{secnumdepth}{0}
-\bibliography{v-special-issue-2016}
-\end{document}
+Our experimental findings,
+and the conclusions that we draw from them, are consistent with this idea.

### 2018-08-02 springer/current/our-approach-to-the-problem.tex
massive restructuring Aug 2018

--- 2018-04-23 papers/2018-v-book/v-book-2017.tex
@@
-\documentclass[%
-%			% choose man or doc immediately below
-%doc,	% doc is single spaced
-man,		% man is double spaced
-floatsintext,%
-apacite%
-]{apa6}
-\usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{color}
-\usepackage{booktabs}
-\usepackage{layout}
-\usepackage{apacite}
-\usepackage{datetime}
-
-%\renewcommand{\baselinestretch}{1.7} % added 21 Sept 2016, copied from CRM book, but see documentclass option 'man' for "manuscript'' above which makes it double spaced
-
-% MG change p's to 3 dp
-
-\title{The Elusive Benefits of Vagueness \protect \\ Evidence from Experiments}
-\date{\today}
-%\authornote{latex compiled \today ~\currenttime}
-\shorttitle{Is Vagueness Beneficial for Hearers?}
-\twoauthors{Matt Green}{Kees van Deemter}
-\twoaffiliations{Computing Science\\ University of Aberdeen}{Computing Science\\ University of Aberdeen}
-
-\abstract{Much of everyday language is vague, yet the benefits of vagueness for hearers and readers are proving to be elusive.
-We discuss a range of earlier controlled experiments with human participants, and we report on a new series of experiments that we conducted in recent years.
-These experiments, which focus on vague expressions that are part of referential noun phrases, aim to separate the utility of vagueness (as defined by the existence of borderline cases) from the utility of other factors that tend to co-occur with vagueness.
-Having presented the evidence, we argue that the evidence supports a view where the benefits that vague terms exert are due to other influences, rather than to vagueness itself.}
-
-\begin{document}
-\maketitle   
-\setcounter{secnumdepth}{3}
-\section{Introduction}\label{introduction}
-
-%Vagueness pervades the language that we use on a daily basis, and the challenge of understanding vague language has been a prominent concern in many areas of logic and linguistics, involving both theoretical and applied work, including the area known as Natural Language Generation (NLG).
-
-%NLG systems can be thought of a small and imperfect models of human language use, and as such they have the ability to make questions about language very tangible, so let us take the perspective of an NLG system for a moment.
-In a nutshell, NLG systems transform data and formulas into language \cite<e.g.,>[]{reiter2000building}, making decisions between different formulations of the same information.
-For example, if the temperature is 27.2 degrees Celsius, the system could express this as ``27.2 degrees'', ``approximately 27 degrees', ``above 25 degrees", or ``warm'', and the system must decide between these.
-The problem is important in the generation of textual weather reports from numerical weather data such as temperature and wind speed \cite{goldberg1994using,turner2006generating}, and medical decision support on the basis of clinical measurements % such as oxygen saturation, heart rhythm, etc.
-%\cite{Hripcsak01032009, hunter2008summarising, portet2009automatic}.
-Such systems often make decisions concerning the level of precision in the utterances that they generate on the basis of little more than intuition.
-Even when NLG systems are designed to mimic human language use \cite<e.g.,>{konstas2013global} there is no guarantee that these decisions taken by these systems benefit readers.
-A better understanding of the benefits (for readers and hearers) of different precision levels would allow these systems to become more useful.
-%The present article investigates the benefits, or de-benefits, of vagueness.
-
-%.
-In the context of an NLG system faced with a practical decision of this kind, Lipman's question becomes ``Under what circumstances should vague terms be produced?''  Relevant applications include weather forecasting
-
-%Language use may be called vague for various reasons.\footnote{See e.g. the entry ``vague'' in \cite{penguin}.}
-
-In most academic use, the word `vagueness' has a specific meaning.
-Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1997vagueness}, also \citeA{EgreKlinedinst}).
-The crucial criterion is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise'' \cite[p.\ 1]{lipmanvague}.
-A typical example is the word ``tall'', as applied to people for example, because here is no precise, known height which separates those who are tall from those who are not.
-The crucial point is that ``tall'' admits borderline cases (i.e., people who may or may not count as tall), which are the hallmark of vagueness as we use the term.
-
-Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions \cite{Lipman:2000fk, lipmanvague}.
-By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss'' \cite<>[p.~1]{lipmanvague}.
-Lipman explains the point by means of a scenario in which a speaker describes a person to a hearer, who needs to identify that person in the arrivals hall of of an airport.
-In such a scenario, a precise description of the person's height (e.g., ``The person's height is 187.96 cm'') would be more useful than a vague one (``The person is tall'').
-Lipman uses this scenario to explain why standard game theory models of communication \cite<e.g.,>{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs.
-%The relevant conditions are, broadly speaking, that both interlocutors know all the relevant facts (e.g., both know the person's height precisely) and that the setting for the communication is co-operative.
-These conditions exclude deliberate deception, and rhetorical situations like political debate and advertising where the intention is to persuade the interlocutor to adopt some point of view, where the persuasion might not be intended to be in the addressee's best interest.
-
-Lipman argued that the efficiency loss resulting from vague expressions would be unlikely to have arisen unless there are advantages as well as disadvantages associated with vague expressions.
-Lipman asked, essentially, what these advantages might be.
-%, and how they might find a place in a game-theoretical explanation.
-In this paper, we focus on the first part of Lipman's question.
-Several tentative answers to Lipman's question have been offered  \cite<see>{van2009utility, vanDeemterBook}.
-Prominent among these answers is the idea that vague expressions are easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite<e.g.,>[]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
-For example, \citeA<>[p.\ 11]{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze''.
-We shall refer to this as the \emph{cost reduction} hypothesis.
-%The idea of vagueness as cost reduction can take various shapes, but the basic hypothesis is that it is easier for people to think in terms of loosely defined categories (such as ``quite a few'', or ``many'') than in terms of crisply defined ones (such as ``thirteen'', or ``237'').
-%This hypothesis predicts that when a vague expression can perform the same communicative task as a precise one, then, the things being equal, it is rational to choose the vague expression.
-The corollary for comprehension is that vague expressions should be understood more readily than precise expressions.
-
-%The cost reduction hypothesis is of direct relevance to psycholinguists interested in language comprehension, and additionally to psycholinguists interested in language production, for example in connection with the question of audience design \cite{Clark1982287}.
-For, to the extent that speakers and writers choose vague expressions over and above crisp ones because the former are easier to process for hearers than the latter, the cost reduction hypothesis suggests that speakers design their utterance for optimal benefit to their hearers -- out of altruism, so to speak.
-
-Charting the utility of vagueness is the attested aim of a small number of studies, but most of these have focussed on vagueness in a different sense, and focussing on different types of benefits for hearers.
-Two recent studies can illustrate both issues.
-
-In a series of studies of behaviour modification, \citeA<>{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss.
-In the weight loss study, participants were told that the study was designed to test the validity of a new (actually fictitious) health index, the HHI (Holistic Health Index).
-They were told that an ideal HHI score lies in the range of 45 to 55.
-In a longitudinal study, participants submitted their weight to a computer each week.
-Participants were told that two algorithms would be used to compute their HHI, and that the two might give different values initially, in which case the true score lay between the two values.
-In one condition, which the authors called the precise condition, the two algorithms gave the same score.
-In the other condition, which the authors called the vague condition, one algorithm added 3\% to the score while the other algorithm subtracted 3\% from the score, yielding a range of values whose midpoint was the same as the two values given in the precise condition.
-
-One group of participants was given HHI scores in the ideal range: for this group their weight loss did not differ depending on whether they were given vague or precise HHI values.
-However for the other group, who were given HHI scores outside the ideal range, their weight loss was significantly greater if they were given vague HHI scores than if they were given precise HHI scores.
-The authors explain the improvement in the vague condition for this group as resulting from the participants' freedom to think of themselves a positioned on one end of the range - the end closest to the ideal HHI scores.
-This ``illusion of proximity'' \cite<>[p.~4]{Mishra01042011} to the goal is argued to allow participants to generate positive expectancies that lead to behaviours that improve performance.
-In contrast, in the precise conditions, participants did not have this freedom of interpretation, and could not distort the information to bring about the beneficial \emph{illusion of proximity}.
-These results are interesting, and of obvious potential practical importance.
-We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \cite{keefe1997vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases.
-In the terminology of \citeA{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}.
-Furthermore, the experiments of \citeA{Mishra01042011} did not explore benefits in terms of processing cost, but in terms of long-term behaviour change.
-
-Similar issues arise from the work of \citeA{peters2009bringing}.
-The authors carried out a series of studies where participants were required to rate hospitals based on various sources of information about quality of care.
-There was a between-subjects manipulation based on numeracy.
-The format of the information was manipulated within subjects: either numbers only were presented, or both numbers and evaluative categories were presented (e.g., \emph{Poor}, \emph{Fair}, \emph{Good}, \emph{Excellent}, with crisp visual boundary lines between the categories).
-Results showed that, for low-numeracy participants, the presence of evaluative categories resulted in a diminished influence of an irrelevant affective state on the ratings.
-For all participants, the presence of evaluative categories resulted in better decisions and in a greater use of the most important and reliable types of information, such as survival rates.
-
-It is, however, questionable whether the ``evaluative categories" manipulation in this study can be considered a manipulation of vagueness.
-Certainly, terms like \emph{Fair} admit the possibility of borderline cases.
-However, given that the boundaries between the categories were marked crisply, and that therefore the categories mapped crisply to numerical values, it becomes doubtful whether any borderline cases could be conceived to arise in fact.
-For example, \emph{Fair} was mapped to 60\% -- 70\% for the variable \emph{percentage of heart attack patients given recommended treatment (ACE inhibitor)}.
-Accordingly, rather than the vagueness of categories such as \emph{Poor}, Peters et al.
-emphasise the evaluative content inherent in these categories, and the affective potential of the evaluative content rather than the vagueness of the terms like {\em Fair}.
-
-\subsection{Our Approach to the Problem}\label{method}
-%The experiments reported in the present paper put the cost reduction hypothesis to the test. The question that we are trying to answer is whether
-
 Are vague expressions processed more easily by readers than crisp ones?
 Like Lipman, we focus on situations where numerical information is used in order to identify a referent.
-Reference, in other words, will be the communicative task on which we focus, partly because of the interest that this topic has recently drawn in various areas of Cognitive Science \cite{vanDeemterCMR}.
+Reference, in other words, will be the communicative task on which we focus, partly because of the interest that this topic has recently drawn in various areas of Cognitive Science \citep{vanDeemterCMR}.
 By looking at one specific type of vagueness, we will be able to investigate the costs and benefits of vagueness relatively thoroughly.
@@
 
-% KvD Feb 2017 (to address a criticism that came up in the reviews for the special issue):
 We have chosen a narrative strategy in which we address a sequence of four experiments with human reader chronologically, explaining how each experiment helped us refine our research question.
@@
 
-% MG General methodology moved to here: per-experiment method sections only talk about deviations from this general method
-All the experiments shared the following properties: 
-Stimuli were created using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.
+All the experiments shared the following properties: Stimuli were created using the language GNU Octave \citep{eaton:2002} and the Psychophysics Toolbox extensions \citep{ptbx1, ptbx2}.
 The position of the dots was randomised per-trial.
@@
 There was a block of practice trials after which participants could ask any questions, following which the experimenter left the room.
-All $p$ values reported for linear models were calculated using the R package \emph{lmerTest} \cite{lmerTest}. The complete data and analysis of our first experiment are available from 
-{\tt http://homepages.abdn.ac.uk/k.vdeemter/pages/vagueness-data-analysis/e1.html}, and analogous for experiments 2-4.\\[2ex]
+All $p$ values reported for linear models were calculated using the R package \emph{lmerTest} \citep{lmerTest}.
+The complete data and analysis for this experiment and the others in this paper are available at \texttt{https://github.com/mjgreen/vagueness}.
+%\texttt{\url{http://homepages.abdn.ac.uk/k.vdeemter/pages/vagueness-data-analysis/e1.html}
 
-% KvD I've moved the following two paragraphs towards the end of the section.
-% MG added/moved out of expt 1
-
-\noindent When the distance grows between two numbers, they become more easily distinguishable: the \emph{numerical distance effect} has been shown for comparing the cardinality of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene1996}.
+When the distance grows between two numbers, they become more easily distinguishable: the \emph{numerical distance effect} has been shown for comparing the cardinality of two sets of dots \citep{van123} and for processing Arabic numerals and number words \citep{Dehaene1996}.
 We manipulated the number of dots in each array such that some sets of arrays had smaller numerical distances and others had larger numerical distances.
 Where a number was mentioned in the instructions, it was always in the form of an Arabic numeral.
-% MG added/moved out of expt 1
-When two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}.
+When two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \citep{dehaene1993mental, gevers2006automatic}.
 We controlled which side the smaller number appeared on to avoid systematic influences of this effect.
 
-% KvD added the following from a previous version of the paper:
-There is abundant evidence \cite<e.g.,>{trick1994small} that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities.
-We performed a pilot experiment \cite{our-earlier-paper} in which we were able to confirm this finding in the experimental settings on which we are focussing in this paper.
-We found that, when participants were confronted with % MG changed the next line and added the footnote to remove the doubts we had about whether a dot array must have regular layout, or can have irregular layout - I'm still not sure this is a good description though
-a stimulus consisting of two squares containing different numbers of dots\footnote{Such a stimulus is referred to hereafter as consisting of a set of {\em dot arrays}.
+There is abundant evidence \citep[e.g.,][]{trick1994small} that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities.
+We performed a pilot experiment \citep{green2011costreduction} in which we were able to confirm this finding in the experimental settings on which we are focussing in this paper.
+We found that, when participants were confronted with a stimulus consisting of two squares containing different numbers of dots\footnote{Such a stimulus is referred to hereafter as consisting of a set of {\em dot arrays}.
 The number of dots in an array is referred to as its cardinality.
-The physical arrangement of dots in each array is irregular.},
-instructions of the form \emph{Choose the square with \emph{n} dots} led to consistently faster response times than instructions of the form \emph{Choose the square with many/few dots} when $2 \leq n \leq 5$; the converse was true for $n>5$.
+The physical arrangement of dots in each array is irregular.}, instructions of the form \emph{Choose the square with \emph{n} dots} led to consistently faster response times than instructions of the form \emph{Choose the square with many/few dots} when $2 \leq n \leq 5$; the converse was true for $n>5$.
 Given these findings, we henceforth focussed our studies on non-subitizable numbers, because it is there that vagueness is expected to have benefits.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 1                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 1: initial experiment}
-
-We used a forced choice task to elicit responses to crisp or vague forms of instructions that required the participant to choose one of two dot arrays by referring to its cardinality.
-The participant was presented with an instruction in the form \emph{Choose the square with \ldots dots}.
-Then a set of two dot arrays was presented, each in the form of a square containing some number of dots.
-The participant was required to identify the dot array that corresponded with the instruction, by pressing the appropriate key, as quickly as possible while avoiding errors. 
-Response time and accuracy were recorded for analysis.
-
-We manipulated how discriminable the dot arrays were by varying the numerical distance between them. 
-One array always contained 25 dots: the other contained either 5, 10, 15, 20, 30, 35, 40, or 45 dots. 
-This gave us numerical distances of 5, 10, 15, and 20, with smaller numerical distances resulting in less discriminable arrays and larger distances resulting in more discriminable arrays.
-
-Our main manipulation was of the vagueness of the instruction, with two levels, \emph{crisp} and \emph{vague}.
-Assuming the dot array [5, 25], and the instruction referring to the smaller cardinality, the \emph{crisp} instruction was \emph{Choose the square with 5 dots} and its \emph{vague} counterpart was \emph{Choose the square with few dots}.
-
-\subsection{Hypothesis}
-\noindent The main aim of this experiment was to test the following hypothesis:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] There exists a main effect advantage for vagueness: vague instructions impose a lower cognitive load for the comprehender than crisp alternatives.
-\end{APAenumerate}
-}
-
-\subsection{Experimental Method}
-
-%Twenty participants were recruited, aged between 18 and 45 with a median age of 26.
-On each trial a participant was presented with an instruction to choose one of two dot arrays on screen by reference to its cardinality. 
-Following a keypress to indicate that the participant had read the instruction there was a central fixation cross for 1000 ms, and a blank screen for 500 ms, followed by the array (without repetition of the referring expression).
-The arrays would stay on screen until the participant responded (there was no timing-out). 
-Response time was measured as the latency between the presentation of the arrays, and the keypress identifying the choice: in this way, the response time was separated from time spent reading the instructions, which is important since we are only interested in the former.
-A response was counted as erroneous if the square with the wrong number of dots was chosen (when the instruction contained a number); 
-if the square with the larger number of dots was selected (when the instruction was \emph{Choose the square with few dots}); 
-or if the square with the smaller number of dots was selected (when the instruction was \emph{Choose the square with many dots}). 
-No feedback was given on correct trials, but there was feedback on error trials in the form of the word ``\textsc{wrong!!}'' which flashed on screen.
-Hypothesis H1 was tested looking both at response times and at error rates.
-
-\subsection{Results}
-
-\begin{figure}[htbp]
-\fitfigure{images/e1RTandErrorRatesplot-1}
-\caption{Experiment 1 results: response times and error rates.}
-\label{resultse1}
-\end{figure}
-
-%\subsubsection{Response times}
-{\bf Response times} (RTs) for trials with erroneous responses were discarded, leading to the loss of 354 trials from 5120, representing 6.9\% of the trials. 
-The correct response RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of a further 160 trials, or 3.4\% of the remaining correct responses. 
-Means for response times and error rates are given in Fig. (\ref{resultse1}).
-A linear mixed model of RT was built, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
-Analysis confirms hypothesis H1:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1$_{rt}$)] RTs were faster for vague instructions than for crisp instructions 
-	($\beta=-0.11$, $se=.02$, $t=4.7$, $p<.001$).  % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-%\subsubsection{Error rates}
-{\bf Error rate} data were analysed using a generalized logit mixed model \cite{jaeger2008categorical}, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
-We found:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1$_{er}$)] There was a statistically non-significant effect of vagueness on error rates: the vague instructions tended to have a lower error rate than the than the crisp instructions 
-	($\beta=-.22$, $se=.18$, $t=-1.2$, $p=.230$).  % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-\subsection{Discussion}
-
-\noindent As expected, responses were faster and more accurate for vague than crisp instructions. Post hoc, we formulated some further hypotheses:
-
-{\small
-\begin{APAenumerate}
-	\item [(H2)] A main effect advantage of increasing numerical distance: the task will become easier as the numerical distance increases, because the two arrays are then more discriminable.
-	\item [(H3)] An interaction between vagueness and numerical distance: i.e., any facilitation for vagueness should be greater at smaller numerical distances than at larger numerical distances.
-\end{APAenumerate}
-}
-
-Looking at response times, these two post-hoc hypotheses were supported:
-
-{\small
-\begin{APAenumerate}
-	\item [(H2$_{rt}$)] RTs increased with numerical distance 
-	($\beta=-0.18$, $se=0.01$, $t=-11.6$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-	Since discriminability of the arrays is easier for larger numerical distances, discriminability probably underlies this effect. 
-	\item [(H3$_{rt}$)] Numerical distance and vagueness interacted significantly: essentially there were diminishing returns for vagueness as numerical distance increased: 
-	($\beta=0.07$, $se=0.01$, $t=5.9$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-Looking at error rates, the two post-hoc hypotheses were supported as well:
-
-{\small
-\begin{APAenumerate}
-	\item [(H2$_{er}$)] Error rates decreased as numerical distance increased: 
-	($\beta=-0.78$, $se=0.08$, $z=-9.6$, $p<0.0005$).% MG added the values for these stats on 25 Sept 2016
-	\item [(H3$_{er}$)] Numerical distance and vagueness interacted significantly: essentially there were diminishing returns for vagueness as numerical distance increased until the biggest numerical distance when vagueness was disadvantageous:
-	($\beta=0.80$, $se=0.15$, $z=5.27$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-So far, these results are in line with the idea of cost reduction.
-The idea of cost reduction explains the vagueness advantage by claiming that the vague referring expressions place less cognitive load on the comprehender than the crisp referring expressions.
-It explains the diminishing returns for vagueness in more-discriminable stimuli (i.e., the vagueness by numerical distance interaction) by claiming that load is low in both conditions for the easily-discriminable stimuli, and that therefore there is no extra benefit to be had from vagueness in the easily-discriminable stimuli.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 2                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 2: separating vagueness from instruction format}
-
-% KvD: I've added some syntactic sugar, in a number of places, to clarify the relation between Ex 1 and Ex 2.
-
-However, the picture painted by these findings might be misleading.
-First of all, there is a possibly confounding factor.
-Contrast, from Experiment 1, an expression from the vague condition: `the square with few dots' with an expression from the crisp condition: `the square with 5 dots'.
-One difference is that `few' has the potential for vagueness, whereas `5' is crisp. But
-another difference is that `few' is verbal while `5' is numerical, in the sense that a number is mentioned explicitly. 
-Since these two differences could not be separated in Experiment 1, the vagueness advantage finding is vulnerable to an alternative interpretation, that what we saw as a vagueness advantage was in contrast an advantage for the verbal form of the quantifier.
-In Experiment 2 we therefore created verbal and numeric versions of each of the vague and crisp instructions so that we could compare vague and crisp conditions while taking account of verbal / numeric format.
-
-Another potential problem with Experiment 1 is the following.
-Participants chose one of two squares: therefore the `vague' quantifiers (e.g., `few') uniquely identified one square.
-Recall our definition of vague  -- ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise''.
-In Experiment 1, the quantifiers in the vague conditions did not realise their potential for vagueness.
-This is because there were no borderline cases of the referent that could make the referent set `not well-defined', and perhaps because using definite articles in the instructions implied that only one option was correct.
-Using error feedback in Experiment 1 could have exacerbated this.
-
-To find out what happens when words are used in a context where their potential for vagueness comes to the fore, Experiment 2 used three arrays (rather than two) so that the vague description had more than one possible referent, and used indefinite articles to avoid the impression that only one response counted as correct, and was carried out without error feedback.
-An indication that the potential for vagueness was realised in Experiment 2 is that the borderline response was chosen fairly often: 16\% of the time.
-
-In Experiment 2, an item was a referring expression instruction followed by a set of three dot arrays defined by a triple of numbers, representing the number of dots in the left, middle, and right arrays.
-We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54).
-Each set of arrays comprised three arrays (instead of two as in Experiment 1); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had more.
-
-Examples of crisp and vague versions of the numerical and verbal instructions follow: the examples assume the array (6,15,24) and reference to the smaller number of dots, such that 6 was classified as the expected response; 15 as the borderline response; and 24 as the incorrect response.
-In the {\em vague numerical} condition we used \emph{Choose a square with about 10 dots}.
-None of the squares contained 10 dots.
-10 is slightly closer to 6 than to 15, justifying 6 as the best response and 15 as the borderline response.
-In the {\em vague verbal} condition we used \emph{Choose a square with few dots}.
-In the {\em crisp numerical} condition we used \emph{Choose the square with 6 dots}, and one square always did contain the number mentioned.
-For {\em crisp verbal}, we used \emph{Choose the square with the fewest dots}.
-
-% KvD I've re-instated the table below, because this makes it easier to understand the instruction format/selection algorithm confound.
-
-\begin{table}[htbp]
-\caption{Experiment 1: instructions arranged by condition for the dot triple $(6,15,24)$ and instructions indicating the smaller numbers}
-\label{instre3}
-\begin{tabular}{rll}
-&crisp&vague\\
-\midrule
-numerical	&	Choose the square with 6 dots 			& 	Choose a square with about 10 dots	\\
-verbal	&	Choose the square with the fewest dots	&	Choose a square with few dots		\\
-\midrule
-\end{tabular}
-\end{table}
-
-\subsection{Hypotheses} % E2
-\noindent We formulated the following hypotheses for Experiment 2:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1)] There exists a main effect RT advantage for vagueness.
-	\item [(H2)] An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
-	\item [(H3)] No large main effect of instruction format (i.e., numeric versus verbal) since we hypothesised that vagueness rather than instruction format drove the effect in Experiment 1.
-	\item [(H4)] On the basis of Experiment 1, we expect faster responses for stimuli with more discriminable arrays.(i.e., an effect of item).
-	\item [(H5)] Participants should make more borderline case choices for vague than crisp instructions.
-\end{APAenumerate}
-}
-
-\subsection{Experimental Method} % E2
-% Participants were aged between 18 and 45 with a median age of 28. 
-We manipulated as independent variables vagueness and instruction format (i.e., numeric versus verbal instructions), yielding four conditions, \emph{vague numeric}; \emph{vague verbal}; \emph{crisp numeric}; \emph{crisp verbal}.
-We measured two dependent variables: response time; and the probability of a participant choosing the borderline case.
-On each trial, first the referring expression that constituted the instruction for that trial was displayed. 
-Participants then pressed a key to indicate that they had read the instruction. 
-After 1000 ms, the arrays were presented, while preserving the text of the referring expression. 
-The response time dependent variable was measured from the presentation of the arrays, until the keypress indicating the participant's choice, which was also recorded. 
-The trial would timeout after 60 seconds if there was no response.
-In this experiment, no feedback was given. 
-This was because, in the vague conditions, we did not regard any response as `correct' or `incorrect', but instead as `borderline response', or `not borderline response', and we did not want to draw participants' attention to this distinction explicitly.
-We simply recorded whether the participant chose the best referent, the borderline case or the poorest referent, and how long it took the participant to respond.
-
-\subsection{Results} % E2
-
-%\subsubsection{Response times} % E2
-% MG added the values for these stats on 26 Sept 2016
-Means for response times and proportion of borderline responses are given in Fig. (\ref{resultse2}). 
-Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data. 
-A linear mixed model was constructed for the (logged) response times, 
-with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects. We found:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against Hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06$, $se=.01$, $t=4.6$, $p<0.0001$).
-	\item [(H2)] In focussed comparisons, vagueness was significantly disadvantageous in the \emph{numeric} conditions ($\beta=0.09$, $se=0.02$, $t=4.5$, $p<0.0001$), and non-significantly disadvantageous in the \emph{verbal} conditions ($\beta=0.02$, $se=0.02$, $t=1.5$, $p=0.1550$), offering partial evidence against Hypothesis 2.
-The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to a significant overall interaction effect between vagueness and instruction format ($\beta=-0.13$, $se=0.02$, $t=-6.6$, $p=0.0169$).
-	\item [(H3)] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1, but suggesting that instruction format rather than vagueness drove the effect we observed in Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; ($\beta=-0.36$, $se=0.07$, $t=-5.1$, $p<0.0001$).
-	\item [(H4)] There was a significant main effect of item (i.e., of which triple of numbers of dots were used in the stimulus): ($\beta=0.12$, $se=0.02$, $t=7.1$, $p<0.0001$).
-This effect seems likely to be due to the very fast responses for stimuli using arrangements of the smallest numbers of dots (6,15,24), which also had the largest difference in ratio of smallest number to largest number in the stimulus, suggesting that stimuli using these numbers of dots may have been particularly discriminable for participants.
-\end{APAenumerate}
-}
-
-%\subsubsection{Borderline responses} % E2
-% MG added the values for these stats on 26 Sept 2016
-A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
-The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{resultse2}.
-Participants chose the borderline square on $16.6\%$ of trials overall.
-{\small 
-\begin{APAenumerate}
-	\item [(H5)] Participants were significantly more likely to choose the borderline option for vague instructions than for crisp instructions (21.9\% vs 11.3\%: $\beta=0.62$, $se=0.22$, $z=2.8$, $p=0.0059$).
-Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%: $\beta=-3.35$, $se=0.23$, $z=-14.6$, $p<0.0001$).
-\end{APAenumerate}
-}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{images/e2-rtplot-1.pdf}
-\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{images/e2-blBarChart-1}
-\caption{Experiment 2 results: mean response times by condition and item, and counts of borderline case responses by condition.}
-\label{resultse2}
-\end{figure}
-
-\subsection{Discussion} % E2
-
-Experiment 2 tested to see whether when borderline cases are present, vague instructions would speed responses as they did in Experiment 1 when there were no borderline squares. 
-We actually found a small (but statistically significant) \emph{disadvantage} of vague instructions: vague instructions slowed people down by 112 ms on average.
-We also found that the effect of instruction format was significant, with numerical format slowing responses by 689 ms on average, such that the disadvantage of numerical format overwhelmed the contribution of vagueness.
-The \emph{verbal vague} condition was still responded to faster than the \emph{numerical crisp} condition, so the pattern from Experiment 1 was reproduced, but in the light of the evidence from Experiment 2, in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like an advantage of verbal instruction format.
-
-% KvD I've expanded the following:
-However, once again there is a possibly confounding factor.
-Observe that, in Experiment 2, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) {\bf selection algorithm}: To see this, consider the task of selecting the dot array that contains ``few dots": to do this, it suffices to {\em compare} the three arrays and select the one that contains the fewest elements.
-To select the dot array that contains "16 dots" seems to require the participant to estimate, and then {\em match}, the cardinality of (at least) one dot array to 16, a process which could plausibly take longer, independently of vagueness.
-Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison ``algorithm".% (which is known to be faster than matching).
-% KvD I think we need to say more about this, citing some appropriate papers to support this idea.
-\textcolor{red}{{\bf--Insert citation}}
-
-In the next two experiments we pitted the comparison algorithm and matching algorithm selection tasks against each other while controlling vagueness and instruction format.
-In Experiment 3 we restricted all the instructions to \emph{numeric} quantifiers while factorially manipulating vagueness and selection task.
-In Experiment 4 we ensured that all instructions used \emph{verbal} quantifiers, while also factorially manipulating vagueness and selection task.
-This allowed us to distinguish between the predictions of the selection task account and the instruction format account.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 3                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 3: focus on instructions that contain numerals}
-
-The main aim of Experiment 3 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies.
-The main changes from Experiment 2 were that the human selection task was explicitly controlled (i.e., in whether it amounted to matching or comparison), and that all conditions were constrained to mention a number.
-We used the same arrays as in Experiment 2.
-Table \ref{Instructions for e3} shows the instructions for each condition.
-Note the difference between ``fewer than 20" and ``far fewer than 20": whereas the former cannot have borderline cases (i.e., for each number it is clear whether the number is smaller than 20 or not), the latter can.
-
-%KvD I think we should motivate our choice of materials (especially "far ..er")
-%\textcolor{red}{{\bf KvD I think we should motivate our choice of materials (especially "far ..er")}}
-% DONE 28 Sept 2017
-
-\begin{table}[htbp]
-\centering
-\caption{Experiment 3: Instructions arranged by condition for stimuli with $(6,15,24)$ dots and instructions indicating the smaller numbers} 
-\label{Instructions for e3}
-\begin{tabular}{llll}
-instruction				&vagueness			&	selection		&	instruction							\\
-format				&					&	task			&									\\
-\toprule
-\multirow{4}{*}{numeric}	&\multirow{2}{*}{crisp} 	&	matching		& Choose a square with 6 dots 				\\ 
-					&					&	comparison 	& Choose a square with fewer than 20 dots 	\\
-					\cline{2-4}
-					&\multirow{2}{*}{vague} 	&	matching 		& Choose a square with about 10 dots 		\\ 
- 					&					&	comparison 	& Choose a square with far fewer than 20 dots 	\\ 
-\bottomrule
-\end{tabular}
-\end{table}%
-
-\subsection{Hypotheses} % E 3
-\noindent For Experiment 3, we hypothesised:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
-	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection task)
-	\item [(H3)] Effects of vagueness are different depending on whether the selection task is matching or comparison (interaction effect selection x vagueness).
-\end{APAenumerate}
-}
-
-\subsection{Experimental Method} % E 3
-
-38 participants were recruited. We used a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e3}).
-On each trial an instruction was presented: participants pressed a key to dismiss the instruction, when the dot arrays were presented until the participant responded, and the response time and choice were recorded.
-
-\subsection{Results} % E 3
-
-Response times were trimmed at 2.5 SD separately for each subject, leading to the loss of 204 trials (2.8\% of the trials).
-Condition means for the remaining (logged) RTs are plotted in Figure \ref{resultse3}.
-A linear mixed model was constructed for the logged response times, 
-with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
-
-\begin{figure}[htbp]
-\centering
-\fitfigure{images/e3-rtplot-1.pdf}
-%\includegraphics[trim = 20mm 0mm 20mm 0mm, clip, width=.75\textwidth]{images/e3-rtplot-1.pdf}
-\caption{Mean response times by condition for Experiment 3 where all instructions were numeric}
-\label{resultse3}
-\end{figure}
-
-% MG added the values for these stats on 26 Sept 2016
-{\small
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions were non-significantly easier for the reader than crisp instructions when vagueness was considered as a main effect ($\beta=-0.01$, $se=0.01$, $t=-0.4$, $p=0.7158$). 
-	\item [(H2)] There was a statistically significant main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=0.16$, $se=0.03$, $t=6.2$, $p<0.0001$).
-	\item [(H3)] Vagueness exerted effects in different directions for the comparison task and for the matching task: there was a significant interaction between vagueness and selection task ($\beta=0.13$, $se=0.03$, $t=4.2$, $p<0.0001$).
-Separate analyses were conducted testing for effects of vagueness at each level of the selection task.
-Within the comparison task vagueness significantly speeded response times compared with crisp controls ($\beta=-0.07$, $se=0.02$, $t=-3.5$, $p<0.0012$). 
-Within the matching task vagueness significantly \emph{slowed} response times compared with crisp controls ($\beta=0.06$, $se=0.02$, $t=2.9$, $p<0.0061$). 
-\end{APAenumerate}
-}
-% MG edited this 27 Sept 2016:
-%The cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
-In other words, the cost reduction account was wrong to predict significant main effect advantages for vagueness (although there was a non-significant trend in the \emph{direction} predicted by the cost reduction account), and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was significantly advantageous in the comparison task.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 4                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 4: focus on instructions that do not contain numerals} % E 4
-
-This experiment mirrors Experiment 3, but focussing on instructions that did not use a number. 
-We manipulated vagueness and the selection task (comparison and matching). 
-In order to implement the experiment without mentioning numbers in the instructions, we changed the sequence of each trial to include a `target' (i.e., a dot array of a particular cardinality) before the instruction, so that we could then refer back to the target's cardinality in the instruction using expressions like \emph{the same number of dots as the target}; \emph{fewer dots than the target}.
-%
-This presentation of a target before the main body of the trial shares some features with \citeA[Experiment 2]{Izard20081221}, although in that experiment participants were told the cardinality of the target (called an \emph{inducer} in that paper) whereas in our experiment we did not tell participants the cardinality of the prime array.
-An item was thus a combination of a target dot array, an instruction that did not contain a number, and a set of dot arrays taking their cardinalities from the same triples used in Experiment 2.
-Table \ref{Instructions for e4} spells out how the instructions were constrained not to mention a numeral and gives examples of targets.
-
-\begin{table}[htp]
-\caption{Instructions and targets by condition for Experiment 4}
-\begin{center}
-\begin{tabular}{llcrl}
-Item					&Selection				&Vagueness	&Target	&Choose a square with\ldots						\\
-\toprule
-\multirow{4}{*}{06:15:24} 	&\multirow{2}{*}{Comparison} 	&Crisp		&20		&\ldots fewer dots than the target					\\
-\cline{3-5}
-					&						&Vague		&20		&\ldots far fewer dots than the target					\\
-					\cline{2-5}
-					&\multirow{2}{*}{Matching}		& Crisp		&6		&\ldots the same number of dots as the target			\\
-					\cline{3-5}
-					& 		 				&Vague		&10 		&\ldots about the same number of dots as the target		\\
-					\bottomrule
-\end{tabular}
-\end{center}
-\label{Instructions for e4}
-\end{table}%
-
-\subsection{Hypotheses}% E 4
-\noindent For Experiment 4, we hypothesised:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
-	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection)
-	\item [(H3)] Effects of vagueness differ depending on whether selection is matching or comparison (interaction effect selection x vagueness, and focussed comparisons at each level of selection).
-\end{APAenumerate}
-}
-
-\subsection{Results}% E 4
-40 volunteers participated.
-The results showed that vagueness was beneficial for comparison but detrimental for matching (the same as Experiment 3) even when no numbers were allowed in the instructions. 
-Figure \ref{resultse4} shows the means by condition. Our findings were as follows:
-
-% MG added the values for these stats on 27 Sept 2016
-{\small
-\begin{APAenumerate}
-	\item [(H1)] There was no significant main effect of vagueness ($\beta=-0.02$, $se=0.01$, $t=-1.5$, $p=0.1296$). 
-	\item [(H2)] There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=0.18$, $se=0.02$, $t=10.4$, $p<0.0001$).
-This effect was in the same direction as Experiment 3.
-	\item[(H3)] Vagueness did exert different effects depending on the selection task (main interaction effect of vagueness by selection $\beta=0.12$, $se=0.02$, $t=5.1$, $p<0.0001$). 
-Separate analyses of the effect of vagueness were conducted for the comparison task and for the matching task using Bonferroni-adjusted significance thresholds. 
-In the comparison task, vagueness resulted in faster response times ($\beta=-0.08$, $se=0.02$, $t=-4.3$, $p<0.0001$). 
-In the matching task vagueness slowed response times ($\beta=0.05$, $se=0.01$, $t=3.7$, $p=0.0004$). 
-These results are in the same direction as Experiment 3.
-\end{APAenumerate}
-}
-
-Once again, the cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
-
-\begin{figure}[htbp]
-\centering
-\fitfigure{images/e4-rtplot-1.pdf}
-\caption{Mean response times by condition for Experiment 4}
-\label{resultse4}
-\end{figure}
-
-\section{Discussion of experiments 3 and 4}
-
-The main aim of these two experiments was to test whether vagueness confers any cognitive benefits over and above those due to differences in the selection task according to whether the instruction mandates a \emph{comparison} selection task or a \emph{matching} selection task, when number-use is held constant.
-The main effect of selection task showed that the assumption that the \emph{comparison} task is easier than the \emph{matching} task is well-founded.
-In both experiments people were reliably quicker to respond in the \emph{comparison} task.
-
-Vagueness, which was the phenomenon on which our investigation focussed, did not exert a significant main effect in response time.
-However when the comparison and selection tasks were analysed separately, there was small significant advantage for vagueness in the \emph{comparison} tasks, but a small significant disadvantage for vagueness in the \emph{matching} tasks.
-
-\section{General Discussion}\label{discussion}
-
-To summarise our findings, Experiment 1 showed that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment could not distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format; and mandated a comparison strategy rather than a matching strategy.
-Experiment 2 showed that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional benefit in either the verbal format instructions or the numerical format instructions.
-
-However, Experiment 2 did not distinguish benefits of number avoidance from benefits of the comparison selection task.
-In Experiments 3 and 4 we manipulated vagueness and the selection task, separately at each level of numerical format.
-Across the two experiments, we found that the comparison-task instructions attracted faster response times than the matching-task instructions.
-Within the two experiments we found that vagueness exerts benefits when the selection task is \emph{comparison}, but not when the task is \emph{matching}.
-
-\begin{table}[htbp]
-\caption{Vagueness as range reduction: a summary of Experiments 3 and 4}
-\label{Vagueness as range reduction}
-\centering
-\begin{tabular}{cccc}
-selection task 				& vagueness		& candidates	& effect of vageueness	\\
-\toprule
-\multirow{ 2}{*}{comparison} 	& crisp 			& 2			& \multirow{ 2}{*}{vagueness advantage}  		\\
-\cline{2-3}
-  						& vague			& 1			&                                               				\\
-\midrule
-\multirow{ 2}{*}{matching} 		& crisp 			& 1 			&\multirow{ 2}{*}{vagueness disadvantage}	\\
-\cline{2-3}
- 						& vague			& 2			&								 	\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-What is one entitled to conclude?
-Given that we were able to identify a class of situations in which vague expressions led to faster response times than crisp ones, would it be valid to conclude that we have finally discovered an advantage for vagueness that cannot be ascribed to some other factor?
-We believe the answer to this question is negative.
-
-To see why, consider Figures \ref{resultse3} and \ref{resultse4}.
-Both figures depict four conditions, depending on whether the expression was crisp or vague, and depending on whether the referent could be identified using a comparison strategy or not.
-Two of the resulting four conditions result in an expression that can denote either of two referents; the other two conditions result in an expression that can only denote one referent, with the other possible referent being a marginal candidate at best.
-
-% KvD: I've shortened the text, removing duplication
-
-To see why vagueness has opposite effects, depending on whether it is used in matching or comparison situations, consider the stimulus with (6,15,24) dots.
-Now compare `Choose a square with 6 dots' with its vague counterpart `Choose a square with about 10 dots': by adding the word `about', we broaden the range of squares that the expression might be referring to.
-On the other hand, compare `Choose a square with fewer than 20 dots' with `Choose a square with far fewer than 20 dots': by adding the word `far', we did not broaden the range of squares denotable by the expression: we narrow it down, because only some of the squares that have fewer dots may have {\em far} fewer dots.
-
-The benefits of vagueness in the \emph{comparison} task in experiments 3 and 4 could thus be explained as differences in the number of valid targets for the expression.
-This leads us to speculate that the benefit for vagueness here could be due to the vague expression foregrounding a particular valid target while the crisp expression carries with it the additional task of distinguishing between two alternative valid targets, something we propose to call a ``range-reduction'' benefit.
-
-The observation that conditions with 1 candidate lead to shorter response times than conditions with 2 candidates is consistent with the range reduction hypothesis, but not with the idea that vagueness is beneficial.
-It appears, in other words, that shorter response times will only result from a vague expression if this expression leads to range reduction.
-Once again, it is not vagueness itself that has advantages but a phenomenon (namely range reduction) that is an automatic concomitant of vagueness in some types of situations.\\[2ex]
-%
-\noindent Looking at the entire series of experiments, our findings suggest that the observed benefits of vague expressions in certain situations may be due to factors other than vagueness: factors like avoiding numbers; permitting comparison tasks; and range reduction.
-The picture that is starting to emerge is subtle: on the one hand, in the situations that we have been studying
-%-- where cooperative speakers refer to an object (e.g., a square) by means of some quantity associated with the object -- 
-vagueness is not intrinsically beneficial.
-On the other hand, vague expressions frequently possess other features that {\em are} beneficial, and these are what give us the incorrect impression that vagueness itself is beneficial.
-Vagueness may thus have acquired a reputation that it does not deserve.
-The answer to Lipman's question, of why vagueness permeates human language (see our Introduction), may lie in a different direction after all, possibly relating to benefits for the speaker rather than the hearer.
-% \cite<see>[chapter 11, for discussion of some possible directions]{vanDeemterBook}.
-
-A comparison may clarify the logic of the situation.
-In recent years a number of studies, focussing on red wine, have suggested that alcohol, consumed in low doses, may have health benefits.
-An alternative explanation, however, asserts that it is not the alcohol in the wine that was beneficial, but antioxidants from grapes.
-If this alternative explanation is correct, then alcohol may not be healthy after all.\\[2ex]
-%
-% KvD Feb 2017. I've diminished the role of NLG in the coming paragraph.
-Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
-Years of research on the logic of vagueness -- giving rise to such techniques as Partial Logic \cite<e.g.,>{Fine}, Probabilistic Logic \cite{Edgington}, and Fuzzy Logic \cite{Zadeh} -- have primed the research community to expect that some special utility of vagueness is an important part of the answer, but our findings call this expectation into question.
-
-Although our own studies in this article have focussed on vagueness in descriptive Noun Phrases, it seems plausible that vagueness plays a similar role in other linguistic constructs.
-For example, consider reports on air temperature.
-Given a numerical temperature measurement or prediction, we might word it as
-%
-\begin{quote}
-(a) {\em 27.2 degrees Celsius}, or\\
-(b) {\em approximately 27 degrees}, or\\
-(c) {\em above 25 degrees}, or\\
-(d) {\em warm}, 
-\end{quote}
-%
-among other candidate expressions.
-Which of these descriptions is most effective, for example as part of a weather report?
-If the linguistic literature is to be believed, then options (a) and (c) convey crisp information, whereas (b) and (d) are vague (i.e., they permit borderline cases).
-From our own experiments and the ones discussed in section 1, there is no evidence that  vagueness should be beneficial for hearers.
-Rather than asking whether a candidate expression is vague, other questions might shed more light on the choice, similar to the ones identified in our studies.
-These questions might focus on the amount of information that a given expression conveys (i.e., on granularity), on the avoidance of numbers, and on the use of evaluative terms.
-Let's see how this might pan out for the examples at hand.
-
-First, the experiments by \citeauthor{Mishra01042011} suggest that it is important how much information is conveyed by an expression, and their findings are echoed by our own thoughts about range reduction (following Experiments 3 and 4).
-In the case of (a)--(d) above, it appears that (a) conveys the most detailed information (designating the smallest segment of the temperature scale), followed by (b), then (d), then (c) (e.g., 40 degrees is above 25, but at 40 Celsius the word ``warm" is likely to give way to ``hot" or ``scorching"):
-%
-\begin{quote}
-$a < b < d < c$
-\end{quote}
-%
-If these hunches are correct then an {\em expert} may prefer to read expression (a), because it gives her the most detailed information on which to base her decisions.
-On the other hand, expression (d) (``warm") is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make (d) more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm" (see our section 1 and  \citeA{peters2009bringing}), which is especially important if the hearer is unfamiliar with the metric used.
-These considerations suggest that {\em non-experts} might prefer expression (d).
-%If results of this kind were to be confirmed by experiments, they would be applicable in practical NLG, though care would need to be exercised to take the purpose of the communication into account.\footnote{One reason why ``warm" might be less preferred than these considerations suggest is that, in the weather context, some very frequently occurring numbers, such as 20 (degrees), could be relatively easy to process, analogous to the subitizable numbers of \protect\citeA{trick1994small}, see our section \ref{method}.
-Note that in the experiments reported in this paper, the purpose of the utterances investigated was always clear (e.g., instructing people on a weight loss program \cite{Mishra01042011}; informing a patient's choice of hospital \cite{peters2009bringing}; or selecting a dot array (in our own studies).}
-
-One way to see why vagueness (as defined in our Introduction) may not benefit human communication is the following thought experiment.
-Suppose a group of speakers understand the word ``warm" as vague, agreeing that temperatures above 26 count as warm, and temperatures below 24 do not count as warm, but considering temperatures between 24 and 26 as borderline cases.
-Now one day these speakers agree to sharpen up their definition deciding that, henceforth, ``warm" means ``$>25$ degrees" (as in (c) above): this decision resolves the borderline cases, while everything else remains the same.
-It seems unlikely that this change in language use, from a vague meaning to a crisp one (i.e., one that has no borderline cases anymore), would lower the utility of the word.
-Our experimental findings are consistent with this idea.
-
-% KvD Bracketed out the following for now:
-%A more radical perspective on our findings might say that the distinction between crisp and vague expressions, which is taken for granted in the bulk of the linguistic literature, is mistaken, because expressions such as ``above 25 degrees" (and ``fewer than 20 dots") are not equally felicitous in all situations in which they can be truthfully applied.
-For example, when the temperature is 40 degrees, then a weather report saying that it is ``above 25 degrees" would be correct yet misleading: this description might be felicitous as a description of temperatures between (roughly) 26-28 degrees, borderline felicitous for 29-33 degrees, and infelicitous at even higher temperatures.
-At its most bold, this perspective would assert that the applicability of {\em every} linguistic expression has borderline cases, hence Lipman's question (see section 1) simply does not arise.
-This radical perspective can only be indicated here, but not discussed at length.
-
-\setcounter{secnumdepth}{0}
-\bibliography{v-special-issue-2016}
-\end{document}

### 2018-08-02 springer/current/related-work.tex
massive restructuring Aug 2018

--- 2016-09-27 papers/2016-Special-Issue/v-special-issue-2016.tex
@@
-\documentclass[%
-%			% choose man or doc immediately below
-%doc,	% doc is single spaced
-man,		% man is double spaced
-floatsintext,%
-apacite%
-]{apa6}
-\usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{color}
-\usepackage{booktabs}
-\usepackage{layout}
-\usepackage{apacite}
-\usepackage{datetime}
-
-%\renewcommand{\baselinestretch}{1.7} % added 21 Sept 2016, copied from CRM book, but see documentclass option 'man' for "manuscript'' above which makes it double spaced
-
-\title{Is Vagueness Beneficial for Hearers? \protect \\ Evidence from Experiments}
-\date{\today}
-\authornote{latex compiled \today ~\currenttime}
-\shorttitle{Is Vagueness Beneficial for Hearers?}
-\twoauthors{Matt Green}{Kees van Deemter}
-\twoaffiliations{Computing Science\\ University of Aberdeen}{Computing Science\\ University of Aberdeen}
-
-\abstract{Much of everyday language use is vague, but the causes of this phenomenon are not well understood.
-Consequently, it is difficult for the designers of a Natural Language Generation (NLG) system to know when to let the system generate vague expressions.
-The present paper is an attempt to find out what benefits vagueness might have for readers.% , by means of a series of carefully controlled psycholinguistic experiments that focus on the use of vagueness in referring expressions.
-The work is motivated by NLG but has implications for linguistics, philosophical logic and game theory.
-%
-This article reports on a series of experiments that aim to separate the utility of vagueness (as defined by the existence of borderline cases) from the utility of other factors that tend to co-occur with vagueness.
-We argue that the evidence so far supports a view where the benefits that vague terms exert are due to other influences, rather than to vagueness itself.
-These factors include: low granularity; the use of evaluative words; the avoidance of overtly numerical words; the existence of comparison strategies; and, lastly and more tentatively, a phenomenon that we call range reduction.
-The paper concludes with a brief discussion of the implications for practical NLG.}
-
-\begin{document}
-\maketitle   
-\setcounter{secnumdepth}{3}
-\section{Introduction}\label{introduction}
-Vagueness pervades the language that we use on a daily basis, and the challenge of understanding vague language has been a prominent concern in many areas of logic and linguistics, involving both theoretical and applied work, including the area known as Natural Language Generation (NLG).
-
-NLG systems transform data and formulas into language \cite<e.g.,>[]{reiter2000building}.
-NLG systems routinely make decisions between different formulations of the same information.
-For example, if the temperature is 27.2 degrees Celsius, this could be expressed as ``27.2 degrees'', ``approximately 27 degrees', ``above 25 degrees", or ``warm'', and the system must decide between these.
-The problem is especially important for NLG systems that take numbers as input, for example in the generation of textual weather reports from numerical weather data such as temperature and wind speed \cite{goldberg1994using,turner2006generating}, and medical decision support on the basis of clinical measurement such as oxygen saturation, heart rhythm, etc.
-\cite{Hripcsak01032009, hunter2008summarising, portet2009automatic}.
-Such systems are often forced to make decisions concerning the level of precision in the utterances that they generate on the basis of little more than intuition.
-Even when NLG systems are designed to mimic human language use \cite<e.g.,>{konstas2013global} there is no guarantee that these decisions taken by these systems benefit readers.
-A better understanding of the benefits (for readers and hearers) of different precision levels would allow these systems to become more useful.
-The present article investigates the benefits,  or de-benefits, of vagueness.
-
-%.
-In the context of an NLG system faced with a practical decision of this kind, Lipman's question becomes ``Under what circumstances should vague terms be produced?''  Relevant applications include weather forecasting
-
-Language use may be called vague for various reasons.\footnote{See e.g.
-the entry ``vague'' in \cite{penguin}.} In most academic use though, the word `vagueness' has a specific meaning.
-Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1997vagueness}, also \citeA{EgreKlinedinst}).
-The crucial criterion is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise'' \cite[p.\ 1]{lipmanvague}.
-A typical example is the word ``tall'', as applied to people for example, because here is no precise, known height which separates those who are tall from those who are not.
-The crucial point is that ``tall'' admits borderline cases (i.e., people who may or may not count as tall), which are the hallmark of vagueness as we use the term.
-
-Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions \cite{Lipman:2000fk, lipmanvague}.
-By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss'' \cite<>[p.~1]{lipmanvague}.
-Lipman explains the point by means of a scenario in which a speaker describes a person to a hearer, who needs to identify that person in the arrivals hall of of an airport.
-In such a scenario, a precise description of the person's height (e.g., ``The person's height is 187.96 cm'') would be more useful than a vague one (``The person is tall'').
-Lipman uses this scenario to explain why standard game theory models of communication \cite<e.g.,>{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs.
-%The relevant conditions are, broadly speaking, that both interlocutors know all the relevant facts (e.g., both know the person's height precisely) and that the setting for the communication is co-operative.
-These conditions exclude deliberate deception, and rhetorical situations like political debate and advertising where the intention is to persuade the interlocutor to adopt some point of view, where the persuasion might not be intended to be in the addressee's best interest.
-
-Lipman argued that the efficiency loss resulting from vague expressions would be unlikely to have arisen unless there are advantages as well as disadvantages associated with vague expressions.
-Lipman asked, essentially, what these advantages might be.
-%, and how they might find a place in a game-theoretical explanation.
-In this paper, we focus on the first part of Lipman's question.
-Several tentative answers to Lipman's question have been offered  \cite<see>{van2009utility, van2010vagueness}.
-Prominent among these answers is the idea that vague expressions are easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite<e.g.,>[]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
-For example, \citeA<>[p.\ 11]{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze''.
-We shall refer to this as the \emph{cost reduction} hypothesis.
-%The idea of vagueness as cost reduction can take various shapes, but the basic hypothesis is that it is easier for people to think in terms of loosely defined categories (such as ``quite a few'', or ``many'') than in terms of crisply defined ones (such as ``thirteen'', or ``237'').
-%This hypothesis predicts that when a vague expression can perform the same communicative task as a precise one, then, the things being equal, it is rational to choose the vague expression.
-The corollary for comprehension is that vague expressions should be understood more readily than precise expressions.
-
-%The cost reduction hypothesis is of direct relevance to psycholinguists interested in language comprehension, and additionally to psycholinguists interested in language production, for example in connection with the question of audience design \cite{Clark1982287}.
-For, to the extent that speakers and writers choose vague expressions over and above crisp ones because the former are easier to process for hearers than the latter, the cost reduction hypothesis suggests that speakers design their utterance for optimal benefit to their hearers -- out of altruism, so to speak.
-
 Charting the utility of vagueness is the attested aim of a small number of studies, but most of these have focussed on vagueness in a different sense, and focussing on different types of benefits for hearers.
@@
 
-In a series of studies of behaviour modification, \citeA<>{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss.
+In a series of studies of behaviour modification, \citeauthor{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss.
 In the weight loss study, participants were told that the study was designed to test the validity of a new (actually fictitious) health index, the HHI (Holistic Health Index).
@@
 The authors explain the improvement in the vague condition for this group as resulting from the participants' freedom to think of themselves a positioned on one end of the range - the end closest to the ideal HHI scores.
-This ``illusion of proximity'' \cite<>[p.~4]{Mishra01042011} to the goal is argued to allow participants to generate positive expectancies that lead to behaviours that improve performance.
+This ``illusion of proximity'' \citep[][p.~4]{Mishra01042011} to the goal is argued to allow participants to generate positive expectancies that lead to behaviours that improve performance.
 In contrast, in the precise conditions, participants did not have this freedom of interpretation, and could not distort the information to bring about the beneficial \emph{illusion of proximity}.
 These results are interesting, and of obvious potential practical importance.
-We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \cite{keefe1997vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases.
-In the terminology of \citeA{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}.
-Furthermore, the experiments of \citeA{Mishra01042011} did not explore benefits in terms of processing cost, but in terms of long-term behaviour change.
+We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \citep{keefe1997vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases.
+In the terminology of \citeauthor{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}.
+Furthermore, the experiments of \citeauthor{Mishra01042011} did not explore benefits in terms of processing cost, but in terms of long-term behaviour change.
 
-Similar issues arise from the work of \citeA{peters2009bringing}.
+Similar issues arise from the work of \citet{peters2009bringing}.
 The authors carried out a series of studies where participants were required to rate hospitals based on various sources of information about quality of care.
@@
 
-It is, however, questionable whether the ``evaluative categories" manipulation in this study can be considered a manipulation of vagueness.
+It is, however, questionable whether the ``evaluative categories'' manipulation in this study can be considered a manipulation of vagueness.
 Certainly, terms like \emph{Fair} admit the possibility of borderline cases.
@@
 emphasise the evaluative content inherent in these categories, and the affective potential of the evaluative content rather than the vagueness of the terms like {\em Fair}.
-
-\subsection{General Methodology}\label{method}
-The experiments reported in the present paper put the cost reduction hypothesis to the test. 
-The question that we are trying to answer is whether vague expressions are processed more easily by readers than crisp ones. 
-Like Lipman, we focus on situations where numerical information is used in order to identify a referent. 
-Reference, in other words, will be the linguistic task on which we focus, partly because of the interest that this topic has recently drawn from the NLG community.  
-In focussing on benefits for the hearer, we will leave aside the question of audience design, leaving this for later research.
-
-In using references to quantities to test the cost reduction hypothesis we are only testing one aspect of vagueness in a particular context. 
-This limits the applicability of our results. 
-However, it has the advantage that it enables us to explore the costs and benefits of vagueness more thoroughly. 
-Since one prevalent view of vagueness is that a vague expression is never preferable to a crisp equivalent, a demonstration of a benefit for vagueness in any context would advance the discussion.
-
-In our experiments we used a speeded forced choice task to compare the processing costs of different references to quantities. 
-In this context, speed and accuracy of responses are the key dimensions on which the different references can be compared. 
-Each stimulus in the experiments was a set of dot arrays containing various number of dots, together with a preceding instruction (in the form of a referring expression) to choose one of the arrays with respect to its cardinality.
-The participant was asked to respond as quickly as possible while avoiding errors. 
-We manipulated the instructions and the arrays in several ways across a series of four experiments.
-
-% MG General methodology moved to here: per-experiment method sections only talk about deviations from this general method
-All the experiments shared the following properties: 
-Stimuli were created using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.
-The position of the dots was randomised per-trial.
-The order in which trials were presented was randomised per-participant. 
-There were 256 trials, presented in 4 blocks of 64 each, between which the participant could rest. 
-A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants and recorded responses. 
-Participants were recruited using email lists at the University of Aberdeen, and paid ten pounds for participating.
-All participants self-reported fluency in English, and had normal, or corrected-to-normal vision.
-The experiment was conducted in a quiet room. 
-Participants were asked to respond as quickly as possible while avoiding errors.
-There was a block of practice trials after which participants could ask any questions, following which the experimenter left the room.
-All $p$ values reported for linear models were calculated using the R package \emph{lmerTest} \cite{lmerTest}.
-
-% KvD I've moved the following two paragraphs towards the end of the section.
-% MG added/moved out of expt 1
-
-When the distance grows between two numbers, they become more easily distinguishable from each other: the \emph{numerical distance effect}, which has been shown for comparing the cardinality of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene1996}.
-We manipulated the number of dots in each array such that some sets of arrays had smaller numerical distances and others had larger numerical distances.
-Where a number was mentioned in the instructions, it was always in the form of an Arabic numeral.
-% MG added/moved out of expt 1
-When two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}.
-We controlled which side the smaller number appeared on to avoid systematic influences of this effect.
-
-% KvD added the following from a previous version of the paper:
-There is abundant evidence \cite<e.g.,>{trick1994small} that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities.
-We performed a pilot experiment \cite{our-earlier-paper} in which we were able to confirm this finding in the experimental settings on which we are focussing in this paper.
-As one might expect, we found that, when participants were confronted with % MG changed the next line and added the footnote to remove the doubts we had about whether a dot array must have regular layout, or can have irregular layout - I'm still not sure this is a good description though
-a stimulus consisting of two squares containing different numbers of dots\footnote{such a stimulus is referred to hereafter as a stimulus consisting of a set of some number of dot arrays, where the number of dots in each dot array varies and is referred to as its cardinality, and the physical arrangement of dots in each dot array is irregular},
-instructions of the form \emph{Choose the square with \emph{n} dots} led to consistently faster response times than instructions of the form \emph{Choose the square with many/few dots} when $2 \leq n \leq 5$; the converse was true for $n>5$.
-We henceforth focussed our studies on non-subitizable numbers, because it is there that vagueness is expected to have benefits.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 1                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment One}
-
-\subsection{Introduction}
-
-We used a forced choice task to elicit responses to crisp or vague forms of instructions that required the participant to choose one of two dot arrays by referring to its cardinality.
-The participant was presented with an instruction in the form \emph{Choose the square with \ldots dots}.
-Then a set of two dot arrays was presented, each in the form of a square containing some number of dots.
-The participant was required to identify the dot array that corresponded with the instruction, by pressing the appropriate key, as quickly as possible while avoiding errors. 
-Response time and accuracy were recorded for analysis.
-
-We manipulated how discriminable the dot arrays were by varying the numerical distance between them. 
-One array always contained 25 dots: the other contained either 5, 10, 15, 20, 30, 35, 40, or 45 dots. 
-This gave us numerical distances of 5, 10, 15, and 20, with smaller numerical distances resulting in less discriminable arrays and larger distances resulting in more discriminable arrays.
-
-Our main manipulation was of the vagueness of the instruction, with two levels, \emph{crisp} and \emph{vague}.
-Assuming the dot array [5, 25], and the instruction referring to the smaller cardinality, the \emph{crisp} instruction was \emph{Choose the square with 5 dots} and its \emph{vague} counterpart was \emph{Choose the square with few dots}.
-
-\subsection{Hypothesis}
-\begin{APAenumerate}
-	\item [(H1)] A main effect advantage for vagueness: vague instructions impose a lower cognitive load for the comprehender than crisp alternatives.
-\end{APAenumerate}
-
-\subsection{Method}
-
-%Twenty participants were recruited, aged between 18 and 45 with a median age of 26.
-On each trial a participant was presented with an instruction to choose one of two dot arrays on screen by reference to its cardinality. 
-Following a keypress to indicate that the participant had read the instruction there was a central fixation cross for 1000 ms, and a blank screen for 500 ms, followed by the array (without repetition of the referring expression).
-The arrays would stay on screen until the participant responded (there was no timing-out). 
-Response time was measured as the latency between the presentation of the arrays, and the keypress identifying the choice: in this way, the response time was separated from time spent reading the instructions, which is important since we are only interested in the former.
-A response was counted as erroneous if the square with the wrong number of dots was chosen (when the instruction contained a number); 
-if the square with the larger number of dots was selected (when the instruction was \emph{Choose the square with few dots}); 
-or if the square with the smaller number of dots was selected (when the instruction was \emph{Choose the square with many dots}). 
-No feedback was given on correct trials, but there was feedback on error trials in the form of the word ``\textsc{wrong!!}'' which flashed on screen.
-Hypothesis H1 was tested looking both at response times and at error rates.
-
-\subsection{Results}
-
-\begin{figure}[htbp]
-\fitfigure{images/plotRTandErrorRates-1}
-\caption{Experiment 1 results: response times and error rates.}
-\label{resultse1}
-\end{figure}
-
-\subsubsection{Response times}
-Response times (RTs) for trials with erroneous responses were discarded, leading to the loss of 354 trials from 5120, representing 6.9\% of the trials. 
-The correct response RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of a further 160 trials, or 3.4\% of the remaining correct responses. 
-Means for response times and error rates are given in Fig. (\ref{resultse1}).
-A linear mixed model of RT was built, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
-
-\begin{APAenumerate}
-	\item [(H1$_{rt}$)] RTs were faster for vague instructions than for crisp instructions 
-	($\beta=-0.11$, $se=.02$, $t=4.7$, $p<.001$).  % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-
-\subsubsection{Error rates}
-Error rate data were analysed using a generalized logit mixed model \cite{jaeger2008categorical}, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
-
-\begin{APAenumerate}
-	\item [(H1$_{er}$)] There was a statistically non-significant effect of vagueness on error rates: the vague instructions tended to have a lower error rate than the than the crisp instructions 
-	($\beta=-.22$, $se=.18$, $t=-1.2$, $p=.230$).  % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-
-\subsection{Discussion}
-
-As expected, responses were faster and more accurate for vague than crisp instructions. Post hoc, we formulated some further hypotheses:
-
-\begin{APAenumerate}
-	\item [(H2)] A main effect advantage of increasing numerical distance: the task will become easier as the numerical distance increases, because the two arrays are then more discriminable.
-	\item [(H3)] An interaction between vagueness and numerical distance: i.e., any facilitation for vagueness should be greater at smaller numerical distances than at larger numerical distances.
-\end{APAenumerate}
-
-Looking at response times, these two post-hoc hypotheses were supported:
-
-\begin{APAenumerate}
-	\item [(H2$_{rt}$)] RTs increased with numerical distance 
-	($\beta=-0.18$, $se=0.01$, $t=-11.6$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-	Since discriminability of the arrays is easier for larger numerical distances, discriminability probably underlies this effect. 
-	\item [(H3$_{rt}$)] Numerical distance and vagueness interacted significantly: essentially there were diminishing returns for vagueness as numerical distance increased: 
-	($\beta=0.07$, $se=0.01$, $t=5.9$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-
-Looking at error rates, the two post-hoc hypotheses were supported:
-
-\begin{APAenumerate}
-	\item [(H2$_{er}$)] Error rates decreased as numerical distance increased: 
-	($\beta=-0.78$, $se=0.08$, $z=-9.6$, $p<0.0005$).% MG added the values for these stats on 25 Sept 2016
-	\item [(H3$_{er}$)] Numerical distance and vagueness interacted significantly: essentially there were diminishing returns for vagueness as numerical distance increased until the biggest numerical distance when vagueness was disadvantageous:
-	($\beta=0.80$, $se=0.15$, $z=5.27$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-
-So far, these results are in line with the idea of cost reduction.
-Cost reduction explains the vagueness advantage by claiming that the vague referring expressions place less cognitive load on the comprehender than the crisp referring expressions.
-It explains the diminishing returns for vagueness in more-discriminable stimuli (i.e., the vagueness by numerical distance interaction) by claiming that load is low in both conditions for the easily-discriminable stimuli, and that therefore there is no extra benefit to be had from vagueness in the easily-discriminable stimuli.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 2                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 2}
-
-\subsection{Introduction} % E2
-
-% KvD: I've added some syntactic sugar, in a number of places, to clarify the relation between Ex 1 and Ex 2.
-
-However, the picture painted by these findings might be misleading.
-First of all, there is a possibly confounding factor.
-Contrast, from Experiment 1, an expression from the vague condition: `the square with few dots' with an expression from the crisp condition: `the square with 5 dots'.
-One difference is that `few' has the potential for vagueness, whereas `5' is crisp. But
-another difference is that `few' is verbal while `5' is numerical, in the sense that a number is mentioned explicitly. 
-Since these two differences could not be separated in Experiment 1, the vagueness advantage finding is vulnerable to an alternative interpretation, that what we saw as a vagueness advantage was in contrast an advantage for the verbal form of the quantifier.
-In Experiment 2 we therefore created verbal and numeric versions of each of the vague and crisp instructions so that we could compare vague and crisp conditions while taking account of verbal / numeric format.
-
-Another potential problem with Experiment 1 is the following.
-Participants chose one of two squares: therefore the `vague' quantifiers (e.g., `few') uniquely identified one square.
-Recall our definition of vague  -- ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise''.
-In Experiment 1, the quantifiers in the vague conditions did not realise their potential for vagueness.
-This is because there were no borderline cases of the referent that could make the referent set `not well-defined', and perhaps because using definite articles in the instructions implied that only one option was correct.
-Using error feedback in Experiment 1 could have exacerbated this.
-
-To find out what happens when words are used in a context where their potential for vagueness comes to the fore,
-Experiment 2 used three arrays so that the vague quantifiers always had more than one possible referent, and used indefinite articles in the vague instructions to avoid the impression that only one response counted as correct, and was carried out without error feedback.
-An indication that the potential for vagueness was realised in Experiment 2 is that the borderline response was chosen fairly often: 16\% of the time.
-
-In Experiment 2, an item was a referring expression instruction followed by a set of three dot arrays defined by a triple of numbers, representing the number of dots in the left, middle, and right arrays.
-We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54).
-Each set of arrays had the following properties: it comprised three arrays (instead of two as in Experiment 1); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had more.
-
-Examples of crisp and vague versions of the numerical and verbal instructions follow: the examples assume the array (6,15,24) and reference to the smaller number of dots, such that 6 was classified as the expected response; 15 was classified as the borderline response; and 24 was classified as the incorrect response.
-In the {\em vague numerical} condition we used \emph{Choose a square with about 10 dots}.
-None of the squares contained 10 dots.
-10 is slightly closer to 6 than to 15, justifying 6 as the best response and 15 as the borderline response.
-In the {\em vague verbal} condition we used \emph{Choose a square with few dots}.
-In the {\em crisp numerical} condition we used \emph{Choose the square with 6 dots}, and one square always did contain the number mentioned.
-For {\em crisp verbal}, we used \emph{Choose the square with the fewest dots}.
-
-% KvD I've re-instated the table below, because this makes it easier to understand the instruction format/selection algorithm confound.
-
-\begin{table}[htbp]
-\caption{Experiment 1: instructions arranged by condition for the dot triple $(6,15,24)$ and instructions indicating the smaller numbers}
-\label{instre3}
-\begin{tabular}{rll}
-&crisp&vague\\
-\midrule
-numerical	&	Choose the square with 6 dots 			& 	Choose a square with about 10 dots	\\
-verbal	&	Choose the square with the fewest dots	&	Choose a square with few dots		\\
-\midrule
-\end{tabular}
-\end{table}
-
-\subsection{Hypotheses} % E2
-\begin{APAenumerate}
-	\item [(H1)] A main effect RT advantage for vagueness.
-	\item [(H2)] An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
-	\item [(H3)] No large main effect of instruction format (of numeric versus verbal) since we hypothesised that vagueness rather than instruction format drove the effect in experiment 1.
-	\item [(H4)] On the basis of Experiment 1, we would expect faster responses for stimuli with more discriminable arrays.(i.e., an effect of item).
-	\item [(H5)] Participants should make more borderline case choices for vague than crisp instructions.
-\end{APAenumerate}
-
-\subsection{Method} % E2
-% Participants were aged between 18 and 45 with a median age of 28. 
-We manipulated as independent variables vagueness and instruction format, yielding four conditions, \emph{vague numeric}; \emph{vague verbal}; \emph{crisp numeric}; \emph{crisp verbal}.
-We measured two dependent variables: response time; and the probability of a participant choosing the borderline case.
-On each trial, first the referring expression that constituted the instruction for that trial was displayed. 
-Participants then pressed a key to indicate that they had read the instruction. 
-After 1000 ms, the arrays were presented, while preserving the text of the referring expression. 
-The response time dependent variable was measured from the presentation of the arrays, until the keypress indicating the participant's choice, which was also recorded. 
-The trial would timeout after 60 seconds if there was no response.
-In this experiment, no feedback was given. 
-This was because, in the vague conditions, we did not regard any response as `correct' or `incorrect', but instead as `borderline response', or `not borderline response', and we did not want to draw participants' attention to this distinction explicitly.
-We simply recorded whether the participant chose the best referent, the borderline case or the poorest referent, and how long it took the participant to respond.
-
-\subsection{Results} % E2
-
-\subsubsection{Response times} % E2
-% MG added the values for these stats on 26 Sept 2016
-Means for response times and proportion of borderline responses are given in Fig. (\ref{resultse2}). 
-Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data. 
-A linear mixed model was constructed for the (logged) response times, 
-with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
-\begin{APAenumerate}
-	\item [(H1)] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06$, $se=.01$, $t=4.6$, $p<0.0001$)
-	\item [(H2)] In focussed comparisons, vagueness was significantly disadvantageous in the \emph{numeric} conditions ($\beta=0.09$, $se=0.02$, $t=4.5$, $p<0.0001$), and non-significantly disadvantageous in the \emph{verbal} conditions ($\beta=0.02$, $se=0.02$, $t=1.5$, $p=0.1550$), offering partial evidence against hypothesis 2.
-The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to a significant overall interaction effect between vagueness and instruction format ($\beta=-0.13$, $se=0.02$, $t=-6.6$, $p=0.0169$).
-	\item [(H3)] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1, but suggesting that instruction format rather than vagueness drove the effect we observed in experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; ($\beta=-0.36$, $se=0.07$, $t=-5.1$, $p<0.0001$).
-	\item [(H4)] There was a significant main effect of item (i.e., of which triple of numbers of dots were used in the stimulus): ($\beta=0.12$, $se=0.02$, $t=7.1$, $p<0.0001$).
-This effect seems likely to be due to the very fast responses for stimuli using arrangements of the smallest numbers of dots (6,15,24), which also had the largest difference in ratio of smallest number to largest number in the stimulus, suggesting that stimuli using these numbers of dots may have been particularly discriminable for participants.
-\end{APAenumerate}
-
-\subsubsection{Borderline responses} % E2
-% MG added the values for these stats on 26 Sept 2016
-A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
-The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{resultse2}.
-Participants chose the borderline square on $16.6\%$ of trials overall.
-\begin{APAenumerate}
-	\item [(H5)] Participants were significantly more likely to choose the borderline option for vague instructions than for crisp instructions (21.9\% vs 11.3\%: $\beta=0.62$, $se=0.22$, $z=2.8$, $p=0.0059$).
-Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%: $\beta=-3.35$, $se=0.23$, $z=-14.6$, $p<0.0001$).
-\end{APAenumerate}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[trim = 0mm 0mm 0mm 0mm, clip, width=.49\textwidth]{images/e2rt.pdf}
-\includegraphics[trim = 0mm -23mm 0mm 0mm, clip, width=.49\textwidth]{images/e2bl.pdf}
-\caption{Experiment 2 results: mean response times by condition and item, and counts of borderline case responses by condition.}
-\label{resultse2}
-\end{figure}
-
-\subsection{Discussion} % E2
-
-Experiment 2 tested to see whether when borderline cases are present, vague instructions would speed responses as they did in Experiment 1 when there were no borderline squares. 
-We actually found a \emph{disadvantage} of vague instructions: vague instructions slowed people down by 112 ms on average.
-We also found that the effect of instruction format was significant, with numerical format slowing responses by 689 ms on average, such that the disadvantage of numerical format overwhelmed the contribution of vagueness.
-The \emph{verbal vague} condition was still responded to faster than the \emph{numerical crisp} condition, so the pattern from Experiment 1 was reproduced, but in the light of the evidence from Experiment 2, in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like an advantage of verbal instruction format.
-
-% KvD I've expanded the following:
-However, once again there is a possibly confounding factor.
-Observe that, in Experiment 2, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) {\bf selection algorithm}: To see this, consider the task of selecting the dot array that contains ``few dots": to do this, it suffices to {\em compare} the three arrays and select the one that contains the fewest elements.
-Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison ``algorithm" (which is known to be faster than matching).
-% KvD I think we need to say more about this, citing some appropriate papers to support this idea.
-\textcolor{red}{{\bf--Insert citation}}
-
-In the next two experiments we pitted the comparison algorithm and matching algorithm selection tasks against each other while controlling vagueness and instruction format.
-In Experiment 3 we restricted all the instructions to \emph{numeric} quantifiers while factorially manipulating vagueness and selection task.
-In Experiment 4 we ensured that all instructions used \emph{verbal} quantifiers, while also factorially manipulating vagueness and selection task.
-This allowed us to distinguish between the predictions of the selection task account and the instruction format account.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 3                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 3}
-
-\subsection{Introduction}
-The main aim of experiment 3 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies.
-The main changes from experiment 2 were that the selection task was explicitly controlled, and that all conditions were constrained to mention a number.
-We used the same arrays as in experiment 2.
-Table \ref{Instructions for e3} shows the instructions for each condition.
-
-%KvD I think we should motivate our choice of materials (especially "far ..er")
-\textcolor{red}{{\bf KvD I think we should motivate our choice of materials (especially "far ..er")}}
-
-\begin{table}[htbp]
-\centering
-\caption{Experiment 3: Instructions arranged by condition for stimuli with $(6,15,24)$ dots and instructions indicating the smaller numbers} 
-\label{Instructions for e3}
-\begin{tabular}{llll}
-instruction				&vagueness			&	selection		&	instruction							\\
-format				&					&	task			&									\\
-\toprule
-\multirow{4}{*}{numeric}	&\multirow{2}{*}{crisp} 	&	matching		& Choose a square with 6 dots 				\\ 
-					&					&	comparison 	& Choose a square with fewer than 20 dots 	\\
-					\cline{2-4}
-					&\multirow{2}{*}{vague} 	&	matching 		& Choose a square with about 10 dots 		\\ 
- 					&					&	comparison 	& Choose a square with far fewer than 20 dots 	\\ 
-\bottomrule
-\end{tabular}
-\end{table}%
-
-\subsection{Hypotheses} % E 3
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions are easier for the reader than crisp alternatives (main effect of vagueness)
-	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection task)
-	\item [(H3)] Effects of vagueness are different depending on whether selection is matching or comparison (interaction effect selection x vagueness).
-\end{APAenumerate}
-
-\subsection{Method} % E 3
-
-38 participants were recruited. 
-The design was a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e3}).
-On each trial a referring expression instruction was presented: participants pressed a key to dismiss the instruction, when the dot arrays were presented until the participant responded, and the response time and choice were recorded.
-
-\subsection{Results} % E 3
-
-Response times were trimmed at 2.5 SD separately for each subject, leading to the loss of 204 trials (2.8\% of the trials).
-Condition means for the remaining (logged) RTs are plotted in Figure \ref{resultse3}.
-A linear mixed model was constructed for the logged response times, 
-with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[trim = 0mm 30mm 0mm 10mm, clip, width=.75\textwidth]{images/rtplote3}
-\caption{Mean response times by condition for Experiment 3 where all instructions were numeric}
-\label{resultse3}
-\end{figure}
-
-% MG added the values for these stats on 26 Sept 2016
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions were non-significantly easier for the reader than crisp instructions when vagueness was considered as a main effect ($\beta=-0.01$, $se=0.01$, $t=-0.4$, $p=0.7158$). 
-	\item [(H2)] There was a statistically significant main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=0.16$, $se=0.03$, $t=6.2$, $p<0.0001$).
-	\item [(H3)] Vagueness exerted effects in different directions for the comparison task and for the matching task: there was a significant interaction between vagueness and selection task ($\beta=0.13$, $se=0.03$, $t=4.2$, $p<0.0001$).
-Separate analyses were conducted testing for effects of vagueness at each level of the selection task.
-Within the comparison task vagueness significantly speeded response times compared with crisp controls ($\beta=-0.07$, $se=0.02$, $t=-3.5$, $p<0.0012$). 
-Within the matching task vagueness significantly \emph{slowed} response times compared with crisp controls ($\beta=0.06$, $se=0.02$, $t=2.9$, $p<0.0061$). 
-\end{APAenumerate}
-
-% MG edited this 27 Sept 2016:
-%The cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
-The cost reduction account was wrong to predict significant main effect advantages for vagueness (although there was a non-significant trend in the \emph{direction} predicted by the cost reduction account), and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was significantly advantageous in the comparison task.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 4                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 4} % E 4
-
-\subsection{Introduction}
-This experiment investigated response times for instructions that did not use a number. 
-We manipulated vagueness and the selection task (comparison and matching). 
-In order to implement the experiment without mentioning numbers in the instructions, we changed the sequence of each trial to include a `target' (i.e., a dot array of a particular cardinality) before the instruction, so that we could then refer back to the target's cardinality in the instruction using expressions like \emph{the same number of dots as the target}; \emph{fewer dots than the target}.
-%
-This presentation of a target before the main body of the trial shares some features with \citeA[Experiment 2]{Izard20081221}, although in that experiment participants were told the cardinality of the target (called an \emph{inducer} in that paper) whereas in our experiment we did not tell participants the cardinality of the prime array.
-An item was thus a combination of a target dot array, an instruction that did not contain a number, and a set of dot arrays taking their cardinalities from the same triples used in Experiment 2.
-Table \ref{Instructions for e4} spells out how the instructions were constrained not to mention a numeral and gives examples of targets.
-
-\begin{table}[htp]
-\caption{Instructions and targets by condition for experiment 4}
-\begin{center}
-\begin{tabular}{llcrl}
-Item					&Selection				&Vagueness	&Target	&Choose a square with\ldots						\\
-\toprule
-\multirow{4}{*}{06:15:24} 	&\multirow{2}{*}{Comparison} 	&Crisp		&20		&\ldots fewer dots than the target					\\
-\cline{3-5}
-					&						&Vague		&20		&\ldots far fewer dots than the target					\\
-					\cline{2-5}
-					&\multirow{2}{*}{Matching}		& Crisp		&6		&\ldots the same number of dots as the target			\\
-					\cline{3-5}
-					& 		 				&Vague		&10 		&\ldots about the same number of dots as the target		\\
-					\bottomrule
-\end{tabular}
-\end{center}
-\label{Instructions for e4}
-\end{table}%
-
-\subsection{Hypotheses}% E 4
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions are easier for the reader than crisp alternatives (main effect of vagueness)
-	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection)
-	\item [(H3)] Effects of vagueness are different depending on whether selection is matching or comparison (interaction effect selection x vagueness, and focussed comparisons at each level of selection).
-\end{APAenumerate}
-
-\subsection{Results}% E 4
-40 volunteers participated.
-The results showed that vagueness was beneficial for comparison but detrimental for matching (the same as Experiment 3) even when no numbers were allowed in the instructions. 
-Figure \ref{resultse4} shows the means by condition.
-
-% MG added the values for these stats on 27 Sept 2016
-\begin{APAenumerate}
-	\item [(H1)] There was no significant main effect of vagueness ($\beta=-0.02$, $se=0.01$, $t=-1.5$, $p=0.1296$). 
-	\item [(H2)] There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=0.18$, $se=0.02$, $t=10.4$, $p<0.0001$).
-This effect was in the same direction as Experiment 3.
-	\item[(H3)] Vagueness did exert different effects depending on the selection task (main interaction effect of vagueness by selection $\beta=0.12$, $se=0.02$, $t=5.1$, $p<0.0001$). 
-Separate analyses of the effect of vagueness were conducted for the comparison task and for the matching task using Bonferroni-adjusted significance thresholds. 
-In the comparison task, vagueness resulted in faster response times ($\beta=-0.08$, $se=0.02$, $t=-4.3$, $p<0.0001$). 
-In the matching task vagueness slowed response times ($\beta=0.05$, $se=0.01$, $t=3.7$, $p=0.0004$). 
-These results are in the same direction as Experiment 3.
-\end{APAenumerate}
-
-The cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[trim = 0mm 30mm 0mm 10mm, clip, width=.75\textwidth]{images/rtplote4.pdf}
-\caption{Mean response times by condition for Experiment 4}
-\label{resultse4}
-\end{figure}
-
-\section{Discussion of experiments 3 and 4}
-
-The main aim of these two experiments was to test whether vagueness confers any cognitive benefits over and above those due to differences in the selection task according to whether the instruction mandates a \emph{comparison} selection task or a \emph{matching} selection task, when number-use is held constant.
-The main effect of selection task showed that the assumption that the \emph{comparison} task is easier than the \emph{matching} task is well-founded.
-In both experiments people were reliably quicker to respond in the \emph{comparison} task.
-
-Vagueness, which was the phenomenon on which our investigation focussed, did not exert a significant main effect in response time.
-However when the comparison and selection tasks were analysed separately, there was small significant advantage for vagueness in the \emph{comparison} tasks, but a small significant disadvantage for vagueness in the \emph{matching} tasks.
-
-\section{General Discussion}
-
-Experiment 1 showed that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment could not distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format; and mandated a comparison strategy rather than a matching strategy.
-Experiment 2 showed that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional benefit in either the verbal format instructions or the numerical format instructions.
-
-However, Experiment 2 could not distinguish benefits of number avoidance from benefits of the comparison selection task.
-In Experiments 3 and 4 we manipulated vagueness and the selection task separately at each level of numerical format.
-Across the two experiments, we found that the comparison-task instructions attracted faster response times than the matching-task instructions.
-Within the two experiments we found that vagueness exerts benefits when the selection task is \emph{comparison}, but not when the task is \emph{matching}.
-
-The benefits of vagueness in the \emph{comparison} task in experiments 3 and 4 could be explained as differences in the number of valid targets for the expression, as follows.
-Taking as an example the stimulus with (6,15,24) dots, it could be argued that the vague comparison instruction (e.g., \emph{a square with far fewer than 20 dots}) has one valid target, the square with 6 dots, while the crisp comparison instruction (e.g., \emph{a square with fewer than 20 dots}) has two valid targets, the squares with 6 and 15 dots.
-In both experiments 3 and 4 we found that people were quicker to identify a square when the instruction only had one valid target.
-This leads us to speculate that the benefit for vagueness here could be due to the vague expression foregrounding a particular valid target while the crisp expression carries with it the additional task of distinguishing between two alternative valid targets, something we propose to call a ``range-reduction'' benefit.
-
-\begin{table}[htbp]
-\caption{Vagueness as range reduction: a summary of Experiments 3 and 4}
-\label{Vagueness as range reduction}
-\centering
-\begin{tabular}{cccc}
-selection task 				& vagueness		& candidates	& effect of vageueness	\\
-\toprule
-\multirow{ 2}{*}{comparison} 	& crisp 			& 2			& \multirow{ 2}{*}{vagueness advantage}  		\\
-\cline{2-3}
-  						& vague			& 1			&                                               				\\
-\midrule
-\multirow{ 2}{*}{matching} 		& crisp 			& 1 			&\multirow{ 2}{*}{vagueness disadvantage}	\\
-\cline{2-3}
- 						& vague			& 2			&								 	\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-What is one entitled to conclude?
-Given that we were able to identify a class of situations -- namely: situations in which a comparison strategy suffices to identify the intended referent -- in which vague expressions led to faster response times than crisp ones, would it be valid to conclude that we have finally discovered an advantage for vagueness that cannot be ascribed to some other factor?
-We believe the answer to this question is negative.
-To see why, consider Figures \ref{resultse3} and \ref{resultse4}.
-Both figures depict four conditions, depending on whether the expression was crisp or vague, and depending on whether the referent could be identified using a comparison strategy or not.
-Two of the resulting four conditions result in an expression that can denote either of two referents; the other two conditions result in an expression that can only denote one referent, with the other possible referent being a marginal candidate at best.
-
-% KvD Something missing here?
-\textcolor{red}{KvD Something missing here?}
-
-To see why vagueness thus has opposite effects, depending on whether it is used in matching or comparison situations, compare an instruction like `Choose a square with 6 dots' with its vague counterpart `Choose a square with about 10 dots': by adding the word `about', we broaden the range of squares that the expression might be referring to.
-On the other hand, compare `Choose a square with fewer than 20 dots' with its vague counterpart `Choose a square with far fewer than 20 dots': by adding the word `far', we did not broaden the range of squares denotable by the expression: we narrow it down, because only some of the squares that have fewer dots may have {\em far} fewer dots.
-
-The observation that conditions with 1 candidate lead to shorter response times than conditions with 2 candidates is consistent with the range reduction hypothesis, but not with the idea that vagueness is beneficial.
-It appears, in other words, that %range reduction causes shorter response times, suggesting that
-shorter response times will only result from a vague expression if this expression leads to range reduction.
-Once again, it is not vagueness itself that has advantages but a phenomenon (namely range reduction) that is an automatic concomitant of vagueness in some types of situations.
-
-Our findings suggest that the observed benefits of vague expressions in certain situations may be due to factors other than vagueness itself: factors like avoiding numbers; permitting comparison tasks; and range reduction.
-The picture that is starting to emerge is subtle: on the one hand, in the situations that we have been studying -- where cooperative speakers refer to an object (e.g., a square) by means of some quantity associated with the object -- vagueness is not intrinsically beneficial.
-On the other hand, vague expressions frequently possess other features that {\em are} beneficial, and these are what give us the incorrect impression that vagueness itself is beneficial.
-Vagueness may thus have acquired a reputation that it does not deserve.
-To answer to Lipman's question, of why vagueness permeates human language (see our Introduction), may lie in a different direction after all, possibly relating to a combination of pure necessity and benefits for the speaker rather than the hearer \cite<see>[chapter 11, for discussion of some possible directions]{van2010vagueness}.
-
-A comparison may clarify the logic of the situation.
-In recent years a number of studies, focussing on red wine, have suggested that alcohol, consumed in low doses, may have health benefits.
-An alternative explanation, however, asserts that it is not the alcohol in the wine that was beneficial, but antioxidants from grapes.
-If this alternative explanation is correct, then alcohol may not be healthy after all.\\[2ex]
-%
-{\bf Implications for practical NLG systems.} Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
-The question of how a particular piece of quantitative information is best conveyed through language is an important one for practical NLG.
-Years of research on the logic of vagueness in natural language -- which has given rise to such logical techniques as Partial Logic (e.g, \cite{Fine}, Probabilistic Logic \cite{Edgington}, and Fuzzy Logic \cite{Zadeh} -- have primed the research community to expect that the utility of vagueness is an important part of the answer, but our findings call this expectation into question.
-Although our own studies in this article have focussed on vagueness in descriptive Noun Phrases, it seems plausible that vagueness plays a similar role in other linguistic constructs.
-
-For example, consider air temperature, once again. Given a temperature measurement (or prediction) as input, an NLG system that addresses a non-expert audience might say that it was (or will be)
-%
-\begin{quote}
-(a) {\em 27.2 degrees Celsius}, or\\
-(b) {\em approximately 27 degrees}, or\\
-(c) {\em above 25 degrees}, or\\
-(d) {\em warm}, 
-\end{quote}
-%
-among other candidate expressions.
-If the linguistic literature is to be believed, then options (a) and (c) convey crisp information, whereas (b) and (d) are vague (i.e., they permit borderline cases).
-From our own and previous experiments, there is no evidence that the fact in itself that an expression is vague should confer a benefit on it for hearers.
-Rather than asking whether a candidate expression is vague, other questions might shed more light on the NLG system's choice, similar to the ones identified in our studies.
-These questions might focus on the amount of information that a given expression conveys (i.e., on granularity), on the avoidance of numbers, and on the use of evaluative terms.
-Let's see how this might pan out for the examples at hand.
-
-First, the experiments by Mishra et al.
-suggest that it is important how much information is conveyed by an expression, and their findings are echoed by our own thoughts about range reduction.
-In the case of (a)--(d) above, it appears that (a) conveys the most detailed information (designating the smallest segment of the temperature scale), followed by (b).
-Expression (b)  appears to be followed by (d), and (d) by (c) (e.g., 40 degrees is above 25, but at 40 Celsius the word ``warm" may no longer be applicable, giving way to words like ``hot"):
-%
-\begin{quote}
-$a < b < d < c$
-\end{quote}
-%
-If these hunches are correct then an {\em expert} may  to prefer expression (a), because it gives her the most detailed information as a basis for her decisions.
-On the other hand, expression (d) (``warm") is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make ``warm" more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm" (see section 1 and  \cite{peters2009bringing}); the factor of evaluation is especially important if the hearer is unfamiliar with the metric used (e.g., because, being American, they are more used to Fahrenheit), in which case ``warm" is much clearer.
-These considerations suggest that {\em non-experts} might prefer expression (d) over all others.
-If results of this kind were to be confirmed by experiments, they would be applicable in practical NLG, though care would need to be exercised to take the purpose of the communication into account.\footnote{One reason why ``warm" might be less preferred than these considerations suggest is that, in the weather context, some very frequently occurring numbers, such as 20 (degrees), could be relatively easy to process, analogous to the subitizable numbers of \protect\cite{trick1994small}, see our General Methodology section.} In the experiments reported in this paper, the purpose of the utterances investigated was always clear (e.g., instructing people on a weight loss program \cite{Mishra01042011}; informing a patient's choice of hospital \cite{peters2009bringing}; or selecting a dot array (in our own studies).
-% KvD See new footnote above.
-
-One way to see why vagueness (as defined in our Introduction) may not matter as much for NLG -- and for human communication -- as is often thought, is the following thought experiment.
-Suppose a group of speakers understand the word ``warm" as vague, agreeing that temperatures above 26 count as warm, and temperatures below 24 do not count as warm, but considering temperatures between 24 and 26 as borderline cases.
-Now one day these speakers agree to sharpen up their definition deciding that, henceforth, ``warm" means ``$>25$ degrees" (as in (c) above): this decision resolves the borderline cases, while everything else remains the same.
-It seems unlikely that this change in language use, from a vague meaning to a crisp one (i.e., one that has no borderline cases anymore), would lower the utility of the word.
-Our experimental findings are consistent with this idea.
-
-\setcounter{secnumdepth}{0}
-\bibliography{v-special-issue-2016}
-\end{document}

### 2018-08-02 springer/current/v-book-2018.tex
massive restructuring Aug 2018

\documentclass[ %
  graybox       %
 ,envcountchap  %
 ,sectrefs      %
 ,footinfo      %
 ,graphics      %
%,referee
]{svmono}
% Native packages
\usepackage{mathptmx}
\usepackage{helvet}
\usepackage{courier}
\usepackage{type1cm}
\usepackage{makeidx}		    % allows index generation
\usepackage{graphicx}		    % standard LaTeX graphics tool
\usepackage{multicol} 		    % used for the two-column index
\usepackage[bottom]{footmisc}	% places footnotes at page bottom
% User packages
\usepackage{natbib}
\usepackage{lipsum}
\usepackage{placeins}
\usepackage{multirow}			% added by Matt
\usepackage{marginnote}			% added by Matt
%\makeindex             		% used for the subject index: please use the style svind.ist with your makeindex program
\begin{document}
\author{Kees van Deemter and Matthew Green}
\title{Vagueness \& Rationality}
\subtitle{The Elusive Benefits of Vagueness: the Evidence So Far}
\input{authoring-notes}\newpage~\newpage
\maketitle
\frontmatter
\tableofcontents
\mainmatter
\chapter{The Elusive Benefits of Vagueness: Evidence from Experiments}\label{chapterlabel}
\input{abstract}
\section{Introduction}\label{introduction}\input{introduction}
\section{Related work}\label{related-work}\input{related-work}
\section{Our Approach to the Problem}\label{our-approach-to-the-problem}\input{our-approach-to-the-problem}
%\section{Experiment 0: the pilot experiment} % This was never in the submitted version
%\section{Experiment 1: initial experiment}\label{e1}\input{e1}% The reviewer wants this experiment excised.
\marginnote{the former E1 was excised here. The new numbering is the former E2,3,4 map to the new E1,2,3.}
\section{Experiment 1: separating vagueness from instruction format}\label{e1}\input{e1}
\section{Experiment 2: focus on instructions that contain numerals}\label{e2}\input{e2}
\section{Experiment 3: focus on instructions that do not contain numerals}\label{e3}\input{e3}
\section{Discussion of experiments 2 and 3}\label{discussion-of-e2-and-e3}\input{discussion-of-e2-and-e3}
\section{General Discussion}\label{general-discussion}\input{general-discussion}
\backmatter
\bibliographystyle{spbasic}
\bibliography{v-book-2018}
\backmatter
%\printindex
\end{document}

### 2018-08-02 springer/review/vagueness-springer-review.tex
massive restructuring Aug 2018

--- 2018-04-18 reviews2018/vagueness-springer-review.tex
@@
-\documentclass[10pt,a4paper]{article}
-\usepackage[utf8]{inputenc}
-\usepackage{lastpage}
-\usepackage{fancyhdr}
-\pagestyle{fancy} 
-\usepackage{amsmath}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
+\documentclass{tufte-handout}
+\title{Review of "The elusive benefits of vagueness: the evidence so far"}
+\date{Manuscript submitted May 2017; Dan Lassiter's review recieved April 2018}
+
+%\geometry{showframe} % display margins for debugging page layout
+
+\usepackage{graphicx} % allow embedded images
+  \setkeys{Gin}{width=\linewidth,totalheight=\textheight,keepaspectratio}
+  \graphicspath{{graphics/}} % set of paths to search for images
+\usepackage{amsmath}  % extended mathematics
+\usepackage{booktabs} % book-quality tables
+\usepackage{units}    % non-stacked fractions and better unit spacing
+\usepackage{multicol} % multiple column layout facilities
+\usepackage{lipsum}   % filler text
+\usepackage{fancyvrb} % extended verbatim environments
+  \fvset{fontsize=\normalsize}% default font size for fancy-verbatim environments
+
+% Standardize command font styles and environments
+\newcommand{\doccmd}[1]{\texttt{\textbackslash#1}}% command name -- adds backslash automatically
+\newcommand{\docopt}[1]{\ensuremath{\langle}\textrm{\textit{#1}}\ensuremath{\rangle}}% optional command argument
+\newcommand{\docarg}[1]{\textrm{\textit{#1}}}% (required) command argument
+\newcommand{\docenv}[1]{\textsf{#1}}% environment name
+\newcommand{\docpkg}[1]{\texttt{#1}}% package name
+\newcommand{\doccls}[1]{\texttt{#1}}% document class name
+\newcommand{\docclsopt}[1]{\texttt{#1}}% document class option name
+\newenvironment{docspec}{\begin{quote}\noindent}{\end{quote}}% command specification environment
+
+\usepackage{textcomp}
+
 \begin{document}
-\title{The elusive benefits of vagueness}
-\author{Dan Lassiter}
-\date{%
-    Stanford University\\[2ex]%
-    April 18, 2018
-}
-\cfoot{\thepage\ of \pageref{LastPage}}
-
 \maketitle
@@
 The results from the second experiment seem to be much more solid and theoretically important.
-Perhaps the paper could be revised to present and interpret the second experiment only, treating the first as a pilot that revealed some issues that were resolved by the main experiment?
+\begin{itemize}
+\item Perhaps the paper could be revised to present and interpret\marginnote{I agree, I think we should remove this experiment and present three experiments only} the second experiment only, treating the first as a pilot that revealed some issues that were resolved by the main experiment?
+\end{itemize}
 
@@
 E.g., \emph{tall} and \emph{warm} can adjust their meanings in ways that allow them to communicate more information about the location of some object relative to the statistical distribution of objects in a reference class.
-\emph{1.83 meters tall} and \emph{94C} don't have this ability.
+"$1.83$ meters tall" and "$94$\textdegree C" don't have this ability.
 I've often thought that this flexibility is one of the main sources of utility for vague expressions with this property, since it allows our lexicon to be fairly expressive without having an enormous number of items.
@@
 \section*{Small comments}
-\begin{itemize}
+\begin{enumerate}
 \item
-While I see the value of presenting Lipman's example as such, the example is fairly strange for the purpose he wants to put it to, since it's not at all obvious that in the context describe \emph{187.9 cm} would indeed be more conversationally useful than \emph{tall} as a description of a person's height.
+While I see the value of presenting Lipman's example as such, the example is fairly strange for the purpose he wants to put it to, since it's not at all obvious that in the context described \emph{187.9 cm} would indeed be more conversationally useful than \emph{tall} as a description of a person's height.
 I for one would have no idea how to pick out a person of that exact height from a crowd.
@@
 The Peters et al.
-experiment: the comment \emph{it becomes doubtful whether any borderline cases could be conceived to arise in fact} – indeed it's not obvious but I would be surprised if they didn't.
+experiment: the comment \emph{it becomes doubtful whether any borderline cases could be conceived to arise in fact} -- indeed it's not obvious but I would be surprised if they didn't.
 When given natural-sounding descriptions like \emph{poor, fair, good, excellent} alongside hard-to-understand numerical values, I would guess that most participants would simply ignore the numerical values since they make less natural descriptions.
@@
 \begin{itemize}
-\item
+\item [(a)]
 why was the central number always presenting in the middle, instead of fully randomizing?
-\item
+\item [(b)]
 \emph{about 10} seems like a terrible description for both 6 and 15. My guess is this would cause subjects to boggle even if they eventually manage to make a guess.
-\item
+\item [(c)]
 what was the vague numeric instruction used in stimuli with larger dot arrays, e.g. (16,25,34)? [I assume it wasn't \emph{about 10}]
+\end{itemize}
 \item
@@
 Everything else is nearly indistinguishable.
-So, it seems like the main thing to explain here is why crisp \emph{6} is so much faster than other numerical expressions – presumably because it's a small number and there's no additional cost of trying to resolve the intention behind the fairly poor description \emph{about 10} for 6?
+So, it seems like the main thing to explain here is why crisp \emph{6} is so much faster than other numerical expressions -- presumably because it's a small number and there's no additional cost of trying to resolve the intention behind the fairly poor description \emph{about 10} for 6?
 \item
@@
 could the differences in the numeric vs verbal borderline response distribution shown in the right panel of fig.
-2 be explained by stimulus design – i.e., the hypothesis I mentioned earlier, that \emph{about 6} is just a poor description of both 6 and 15, with the result that participants were choosing more or less random with a slight bias for the number a little bit closer to 10?
+2 be explained by stimulus design -- i.e., the hypothesis I mentioned earlier, that \emph{about 6} is just a poor description of both 6 and 15, with the result that participants were choosing more or less at random with a slight bias for the number a little bit closer to 10?
 This would explain why there is almost no difference between borderline and expected in the vague/numeric condition.
 \item
-p.17 \emph{we found a small (but statisticslly significant) disadvantage of vague instructions} – again this is misleading since the effect is apparently driven almost entirely by a single condition (6, crisp) that was much faster than expected.
+p.17 \emph{we found a small (but statistically significant) disadvantage of vague instructions} -- again this is misleading since the effect is apparently driven almost entirely by a single condition (6, crisp) that was much faster than expected.
 There appear to be few differences otherwise \ldots one option would be to exclude the 6:15:24 condition and then re-run the statistical analysis to see if any other effects emerge.
 \item
-p.18 \emph{allowed participants to use a comparison “algorithm” – the work of Justin Halberda and collaborators on the verification strategies associated with \emph{most} and \emph{more than half} is highly relevant here (and indeed throughout the paper, given that \emph{most} is vague but otherwise fairly similar in meaning to \emph{more than half})
+p.18 \emph{allowed participants to use a comparison "algorithm"} -- the work of Justin Halberda and collaborators on the verification strategies associated with \emph{most} and \emph{more than half} is highly relevant here (and indeed throughout the paper, given that \emph{most} is vague but otherwise fairly similar in meaning to \emph{more than half})
 \item
 p.19 Testing (H1), we have \emph{Vague instructions were non-significantly easier \ldots} do you mean \emph{not significantly easier}?
-In any case from the graphs on the next page this seems misleading – what it looks like is a substantial main effect where vague instructions have a main effect of lowering the response time, but this effect is masked by interactions with quantity, most notably (again) the huge speed-up for 6 in the matching/crisp condition.
-I'd recommend following the method described in Roger Levy}s paper on finding main effects in the presence of interactions, available here: \url{https://arxiv.org/abs/1405.2094}.
+In any case from the graphs on the next page this seems misleading -- what it looks like is a substantial main effect where vague instructions have a main effect of lowering the response time, but this effect is masked by interactions with quantity, most notably (again) the huge speed-up for 6 in the matching/crisp condition.
+I'd recommend following the method described in Roger Levy's paper on finding main effects in the presence of interactions (\url{https://arxiv.org/abs/1405.2094}).
 \item
-p.20 \emph{the cost reduction account was wrong to predict significant main effect advantages for vagueness} – this seems way too strong given the issues I've just described.
+p.20 \emph{the cost reduction account was wrong to predict significant main effect advantages for vagueness} -- this seems way too strong given the issues I've just described.
 First, the statistical analysis needs to be redone as I just mentioned.
 Second, in the kind of statistical analysis being employed here it's not possible to show that something doesn't exist simply by failing to find evidence that it does exist.
-(APA guidelines: \emph{Never use the unfortunate expression \emph{accept the null hypothesis}}, cf.
-Wagenmakers, \url{http://www.ejwagenmakers.com/2007/pValueProblems.pdf}).
+(APA guidelines: \emph{Never use the unfortunate expression "accept the null hypothesis"}, cf.
+Wagenmakers (\url{http://www.ejwagenmakers.com/2007/pValueProblems.pdf}).
 At best, you can say \emph{we failed to find evidence that \ldots}
 \item
-.p21, table 3: were the matching/crisp and matching/vague conditions unmatched in the target quantity as the table says? Or is this a typo?
+p.21, table 3: were the matching/crisp and matching/vague conditions unmatched in the target quantity as the table says? Or is this a typo?
 \item
@@
 \item
-p.22 \emph{In the matching task vagueness slowed response times} – again this is not a good summary since there is a big interaction, and the apparent main efect may be driven by a single condition, here a slowdown for 6:15:24 in the matching/vague condition.
+p.22 \emph{In the matching task vagueness slowed response times} -- again this is not a good summary since there is a big interaction, and the apparent main efect may be driven by a single condition, here a slowdown for 6:15:24 in the matching/vague condition.
 Re-running using Levy's method would be advisable here.
-\end{itemize}
 \item
-Typo: Paul Egré's name has an acute accent on the second \emph{e}
-\end{itemize}
+Typo: Paul Egr\a'e's name has an acute accent on the second \emph{e}
+\marginnote{corrected in the .bib}
+\end{enumerate}
 \end{document}

### 2018-08-02 springer/submission/v-book-2017.tex
massive restructuring Aug 2018

--- 2018-04-23 papers/2018-v-book/v-book-2017.tex
@@
 apacite%
-]{apa6}
+]{apa6} % I'm using a local (as in, in this folder) apa6.cls which edits footnotesep from 16pt which is the real apa6 in 2018 back to 6pt in order to make it line up with the pdf generated for the 2017 submission
 \usepackage{graphicx}
@@
 
-%\renewcommand{\baselinestretch}{1.7} % added 21 Sept 2016, copied from CRM book, but see documentclass option 'man' for "manuscript'' above which makes it double spaced
-
-% MG change p's to 3 dp
-
 \title{The Elusive Benefits of Vagueness \protect \\ Evidence from Experiments}
 \date{\today}
-%\authornote{latex compiled \today ~\currenttime}
+\authornote{This is a layout-identical version of the document submitted for review in May 2017. LaTeX compiled \today~at \currenttime}
 \shorttitle{Is Vagueness Beneficial for Hearers?}
@@
 
-\abstract{Much of everyday language is vague, yet the benefits of vagueness for hearers and readers are proving to be elusive.
+\abstract{Much of everyday language is vague, even in situations where vagueness could have been avoided (i.e., where vagueness is used "strategically").
+Yet the benefits of vagueness for hearers and readers are proving to be elusive.
 We discuss a range of earlier controlled experiments with human participants, and we report on a new series of experiments that we conducted in recent years.
 These experiments, which focus on vague expressions that are part of referential noun phrases, aim to separate the utility of vagueness (as defined by the existence of borderline cases) from the utility of other factors that tend to co-occur with vagueness.
-Having presented the evidence, we argue that the evidence supports a view where the benefits that vague terms exert are due to other influences, rather than to vagueness itself.}
+Having presented the evidence, we argue that the evidence supports a view where the benefits that vague terms exert are due to other influences, and not to vagueness
+itself.}
 
@@
 \section{Introduction}\label{introduction}
-
-%Vagueness pervades the language that we use on a daily basis, and the challenge of understanding vague language has been a prominent concern in many areas of logic and linguistics, involving both theoretical and applied work, including the area known as Natural Language Generation (NLG).
-
-%NLG systems can be thought of a small and imperfect models of human language use, and as such they have the ability to make questions about language very tangible, so let us take the perspective of an NLG system for a moment.
-In a nutshell, NLG systems transform data and formulas into language \cite<e.g.,>[]{reiter2000building}, making decisions between different formulations of the same information.
-For example, if the temperature is 27.2 degrees Celsius, the system could express this as ``27.2 degrees'', ``approximately 27 degrees', ``above 25 degrees", or ``warm'', and the system must decide between these.
-The problem is important in the generation of textual weather reports from numerical weather data such as temperature and wind speed \cite{goldberg1994using,turner2006generating}, and medical decision support on the basis of clinical measurements % such as oxygen saturation, heart rhythm, etc.
-%\cite{Hripcsak01032009, hunter2008summarising, portet2009automatic}.
-Such systems often make decisions concerning the level of precision in the utterances that they generate on the basis of little more than intuition.
-Even when NLG systems are designed to mimic human language use \cite<e.g.,>{konstas2013global} there is no guarantee that these decisions taken by these systems benefit readers.
-A better understanding of the benefits (for readers and hearers) of different precision levels would allow these systems to become more useful.
-%The present article investigates the benefits, or de-benefits, of vagueness.
-
-%.
-In the context of an NLG system faced with a practical decision of this kind, Lipman's question becomes ``Under what circumstances should vague terms be produced?''  Relevant applications include weather forecasting
-
-%Language use may be called vague for various reasons.\footnote{See e.g. the entry ``vague'' in \cite{penguin}.}
 
@@
 
-Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions \cite{Lipman:2000fk, lipmanvague}.
+Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions \cite{Lipman:2000fk, lipmanvague}, which are
+often used even in situations where the speaker could have used an expression that is not vague (i.e., crisp); in these situations we say that vagueness is used \emph{strategically}.
 By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss'' \cite<>[p.~1]{lipmanvague}.
@@
 Lipman uses this scenario to explain why standard game theory models of communication \cite<e.g.,>{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs.
-%The relevant conditions are, broadly speaking, that both interlocutors know all the relevant facts (e.g., both know the person's height precisely) and that the setting for the communication is co-operative.
-These conditions exclude deliberate deception, and rhetorical situations like political debate and advertising where the intention is to persuade the interlocutor to adopt some point of view, where the persuasion might not be intended to be in the addressee's best interest.
 
@@
 Lipman asked, essentially, what these advantages might be.
-%, and how they might find a place in a game-theoretical explanation.
-In this paper, we focus on the first part of Lipman's question.
 Several tentative answers to Lipman's question have been offered  \cite<see>{van2009utility, vanDeemterBook}.
-Prominent among these answers is the idea that vague expressions are easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite<e.g.,>[]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
+One of the most promising answers appears to be the idea that vague expressions are easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite<e.g.,>[]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
 For example, \citeA<>[p.\ 11]{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze''.
 We shall refer to this as the \emph{cost reduction} hypothesis.
-%The idea of vagueness as cost reduction can take various shapes, but the basic hypothesis is that it is easier for people to think in terms of loosely defined categories (such as ``quite a few'', or ``many'') than in terms of crisply defined ones (such as ``thirteen'', or ``237'').
-%This hypothesis predicts that when a vague expression can perform the same communicative task as a precise one, then, the things being equal, it is rational to choose the vague expression.
-The corollary for comprehension is that vague expressions should be understood more readily than precise expressions.
 
-%The cost reduction hypothesis is of direct relevance to psycholinguists interested in language comprehension, and additionally to psycholinguists interested in language production, for example in connection with the question of audience design \cite{Clark1982287}.
-For, to the extent that speakers and writers choose vague expressions over and above crisp ones because the former are easier to process for hearers than the latter, the cost reduction hypothesis suggests that speakers design their utterance for optimal benefit to their hearers -- out of altruism, so to speak.
+This article brings an experimental approach to these issues, focussing on vagueness in descriptions (e.g., "the square with few dots") and its effect on the hearer's ability to act on a given description, as measured by the time that it takes hearers to click on the referent of a description\footnote{Other metrics could have been chosen, such as hearers' ability to remember information, for example, or error rates.
+Although error rates play a minor role in the present paper, for reasons that will become clear, we focus on response times in particular.}.
+We find that, although hearer benefits from vague descriptions are straightforward to demonstrate in many cases, a closer experimental analysis militates against the conclusion that vagueness itself -- as defined above, in terms of the existence of borderline cases -- lies at the heart of these results.
+Instead, it is other factors, such as the presence of an overt numerical expression in the description, that proved to be decisive.
+We believe that, despite the fact that our experiments are unavoidably focussed only on a specific class of vague expressions (since any experiment can only deal with a limited number of different stimuli), these findings are potentially important, because they call into question whether "strategic" vagueness (i.e., vagueness where the speaker had a choice, because she could have been produced a crisp expression instead) has any advantages at all.
+In other words, returning to Lipman's question, it is possible that vagueness has evolved partly as a necessary evil (e.g., because of the limits of observation and prediction) and partly as a side effect of other factors.
+
+\section{Related work}
 
@@
 \subsection{Our Approach to the Problem}\label{method}
-%The experiments reported in the present paper put the cost reduction hypothesis to the test. The question that we are trying to answer is whether
 
@@
 
-% MG General methodology moved to here: per-experiment method sections only talk about deviations from this general method
 All the experiments shared the following properties: 
@@
 
-% KvD I've moved the following two paragraphs towards the end of the section.
-% MG added/moved out of expt 1
-
 \noindent When the distance grows between two numbers, they become more easily distinguishable: the \emph{numerical distance effect} has been shown for comparing the cardinality of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene1996}.
@@
 Where a number was mentioned in the instructions, it was always in the form of an Arabic numeral.
-% MG added/moved out of expt 1
 When two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}.
@@
 
-% KvD added the following from a previous version of the paper:
 There is abundant evidence \cite<e.g.,>{trick1994small} that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities.
 We performed a pilot experiment \cite{our-earlier-paper} in which we were able to confirm this finding in the experimental settings on which we are focussing in this paper.
-We found that, when participants were confronted with % MG changed the next line and added the footnote to remove the doubts we had about whether a dot array must have regular layout, or can have irregular layout - I'm still not sure this is a good description though
-a stimulus consisting of two squares containing different numbers of dots\footnote{Such a stimulus is referred to hereafter as consisting of a set of {\em dot arrays}.
+We found that, when participants were confronted with a stimulus consisting of two squares containing different numbers of dots\footnote{Such a stimulus is referred to hereafter as consisting of a set of {\em dot arrays}.
 The number of dots in an array is referred to as its cardinality.
-The physical arrangement of dots in each array is irregular.},
-instructions of the form \emph{Choose the square with \emph{n} dots} led to consistently faster response times than instructions of the form \emph{Choose the square with many/few dots} when $2 \leq n \leq 5$; the converse was true for $n>5$.
+The physical arrangement of dots in each array is irregular.}, instructions of the form \emph{Choose the square with \emph{n} dots} led to consistently faster response times than instructions of the form \emph{Choose the square with many/few dots} when $2 \leq n \leq 5$; the converse was true for $n>5$.
 Given these findings, we henceforth focussed our studies on non-subitizable numbers, because it is there that vagueness is expected to have benefits.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 1                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-%\subsubsection{Response times}
 {\bf Response times} (RTs) for trials with erroneous responses were discarded, leading to the loss of 354 trials from 5120, representing 6.9\% of the trials. 
@@
 
-% KvD: I've added some syntactic sugar, in a number of places, to clarify the relation between Ex 1 and Ex 2.
-
 However, the picture painted by these findings might be misleading.
@@
 Contrast, from Experiment 1, an expression from the vague condition: `the square with few dots' with an expression from the crisp condition: `the square with 5 dots'.
-One difference is that `few' has the potential for vagueness, whereas `5' is crisp. But
-another difference is that `few' is verbal while `5' is numerical, in the sense that a number is mentioned explicitly. 
+One difference is that `few' has the potential for vagueness, whereas `5' is crisp.
+But another difference is that `few' is verbal while `5' is numerical, in the sense that a number is mentioned explicitly.
 Since these two differences could not be separated in Experiment 1, the vagueness advantage finding is vulnerable to an alternative interpretation, that what we saw as a vagueness advantage was in contrast an advantage for the verbal form of the quantifier.
@@
 For {\em crisp verbal}, we used \emph{Choose the square with the fewest dots}.
-
-% KvD I've re-instated the table below, because this makes it easier to understand the instruction format/selection algorithm confound.
 
@@
 %\subsubsection{Response times} % E2
-% MG added the values for these stats on 26 Sept 2016
 Means for response times and proportion of borderline responses are given in Fig. (\ref{resultse2}). 
@@
 
-%\subsubsection{Borderline responses} % E2
-% MG added the values for these stats on 26 Sept 2016
 A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
@@
 
-% KvD I've expanded the following:
 However, once again there is a possibly confounding factor.
@@
 To select the dot array that contains "16 dots" seems to require the participant to estimate, and then {\em match}, the cardinality of (at least) one dot array to 16, a process which could plausibly take longer, independently of vagueness.
-Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison ``algorithm".% (which is known to be faster than matching).
-% KvD I think we need to say more about this, citing some appropriate papers to support this idea.
-\textcolor{red}{{\bf--Insert citation}}
+Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison ``algorithm".
 
@@
 Note the difference between ``fewer than 20" and ``far fewer than 20": whereas the former cannot have borderline cases (i.e., for each number it is clear whether the number is smaller than 20 or not), the latter can.
-
-%KvD I think we should motivate our choice of materials (especially "far ..er")
-%\textcolor{red}{{\bf KvD I think we should motivate our choice of materials (especially "far ..er")}}
-% DONE 28 Sept 2017
 
@@
 \fitfigure{images/e3-rtplot-1.pdf}
-%\includegraphics[trim = 20mm 0mm 20mm 0mm, clip, width=.75\textwidth]{images/e3-rtplot-1.pdf}
 \caption{Mean response times by condition for Experiment 3 where all instructions were numeric}
@@
 
-% MG added the values for these stats on 26 Sept 2016
 {\small
@@
 }
-% MG edited this 27 Sept 2016:
-%The cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
 In other words, the cost reduction account was wrong to predict significant main effect advantages for vagueness (although there was a non-significant trend in the \emph{direction} predicted by the cost reduction account), and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was significantly advantageous in the comparison task.
@@
 In order to implement the experiment without mentioning numbers in the instructions, we changed the sequence of each trial to include a `target' (i.e., a dot array of a particular cardinality) before the instruction, so that we could then refer back to the target's cardinality in the instruction using expressions like \emph{the same number of dots as the target}; \emph{fewer dots than the target}.
-%
 This presentation of a target before the main body of the trial shares some features with \citeA[Experiment 2]{Izard20081221}, although in that experiment participants were told the cardinality of the target (called an \emph{inducer} in that paper) whereas in our experiment we did not tell participants the cardinality of the prime array.
@@
 
-% MG added the values for these stats on 27 Sept 2016
 {\small
@@
 
-To summarise our findings, Experiment 1 showed that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment could not distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format; and mandated a comparison strategy rather than a matching strategy.
+To summarise our findings, we asked why strategic vagueness is as frequent as it is and we decided to focus on what we see as the most promising explanation, namely that vague expressions are easy to process by speakers and hearers: the cost hypothesis, as we have called it.
+We decided to test this hypothesis by experimentally investigating whether vague descriptions are resolved by hearers more quickly than crisp ones.
+Although we were able to find some interesting (and statistically highly significant) effects, it appears to us that if our sequence of experiments is assessed as a whole, it cannot be seen as confirmation of the cost hypothesis.
+
+To explain why this is, let us summarise our findings so far: Experiment 1 showed that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment was unable to distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format, and they mandated a comparison strategy rather than a matching strategy.
 Experiment 2 showed that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional benefit in either the verbal format instructions or the numerical format instructions.
-
 However, Experiment 2 did not distinguish benefits of number avoidance from benefits of the comparison selection task.
@@
 \midrule
-\multirow{ 2}{*}{matching} 		& crisp 			& 1 			&\multirow{ 2}{*}{vagueness disadvantage}	\\
+\multirow{ 2}{*}{matching} 		& crisp 		& 1 			& \multirow{ 2}{*}{vagueness disadvantage}	\\
 \cline{2-3}
@@
 
-Although our own studies in this article have focussed on vagueness in descriptive Noun Phrases, it seems plausible that vagueness plays a similar role in other linguistic constructs.
-For example, consider reports on air temperature.
-Given a numerical temperature measurement or prediction, we might word it as
+Although our own studies in this article have focussed on vagueness in descriptive Noun Phrases only, it seems to us plausible that vagueness plays a similar role in other
+linguistic constructs. For example, consider reports on air temperature. Given a numerical
+temperature measurement or prediction, we might word it as 
 %
@@
 If the linguistic literature is to be believed, then options (a) and (c) convey crisp information, whereas (b) and (d) are vague (i.e., they permit borderline cases).
-From our own experiments and the ones discussed in section 1, there is no evidence that  vagueness should be beneficial for hearers.
+In the situations studied in our own experiments and the ones discussed in section 1, we found no evidence that vagueness is beneficial for hearers.
 Rather than asking whether a candidate expression is vague, other questions might shed more light on the choice, similar to the ones identified in our studies.
 These questions might focus on the amount of information that a given expression conveys (i.e., on granularity), on the avoidance of numbers, and on the use of evaluative terms.
-Let's see how this might pan out for the examples at hand.
+Let's see how this might pan out for the above examples from the weather domain.
 
@@
 %
-If these hunches are correct then an {\em expert} may prefer to read expression (a), because it gives her the most detailed information on which to base her decisions.
+If these hunches are correct, then it seems to us that it is relatively unimportant whether a given expression is vague or crisp.
+Other factors seem more important; moreover, it may depend on the task and the audience which of a-d is preferred.
+For example, an \emph{expert} may prefer to read expression (a), because it gives her the most detailed information on which to base her decisions.
 On the other hand, expression (d) (``warm") is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make (d) more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm" (see our section 1 and  \citeA{peters2009bringing}), which is especially important if the hearer is unfamiliar with the metric used.
 These considerations suggest that {\em non-experts} might prefer expression (d).
-%If results of this kind were to be confirmed by experiments, they would be applicable in practical NLG, though care would need to be exercised to take the purpose of the communication into account.\footnote{One reason why ``warm" might be less preferred than these considerations suggest is that, in the weather context, some very frequently occurring numbers, such as 20 (degrees), could be relatively easy to process, analogous to the subitizable numbers of \protect\citeA{trick1994small}, see our section \ref{method}.
-Note that in the experiments reported in this paper, the purpose of the utterances investigated was always clear (e.g., instructing people on a weight loss program \cite{Mishra01042011}; informing a patient's choice of hospital \cite{peters2009bringing}; or selecting a dot array (in our own studies).}
 
@@
 It seems unlikely that this change in language use, from a vague meaning to a crisp one (i.e., one that has no borderline cases anymore), would lower the utility of the word.
-Our experimental findings are consistent with this idea.
-
-% KvD Bracketed out the following for now:
-%A more radical perspective on our findings might say that the distinction between crisp and vague expressions, which is taken for granted in the bulk of the linguistic literature, is mistaken, because expressions such as ``above 25 degrees" (and ``fewer than 20 dots") are not equally felicitous in all situations in which they can be truthfully applied.
-For example, when the temperature is 40 degrees, then a weather report saying that it is ``above 25 degrees" would be correct yet misleading: this description might be felicitous as a description of temperatures between (roughly) 26-28 degrees, borderline felicitous for 29-33 degrees, and infelicitous at even higher temperatures.
-At its most bold, this perspective would assert that the applicability of {\em every} linguistic expression has borderline cases, hence Lipman's question (see section 1) simply does not arise.
-This radical perspective can only be indicated here, but not discussed at length.
+Our experimental findings,
+and the conclusions that we draw from them, are consistent with this idea.
 
 \setcounter{secnumdepth}{0}
-\bibliography{v-special-issue-2016}
+\bibliography{v-book-2017}
 \end{document}

### 2018-08-02 experiments/C_exp_1/writeup/2018_levy-method/annotate_the_raw_data.R

@@ -1,0 +1,41 @@
+annotate_the_raw_data = function (dat) {
+  
+  # annotate vagueness
+  dat[ dat$Condition==1 , 'Vagueness'] <- 'Vague'
+  dat[ dat$Condition==2 , 'Vagueness'] <- 'Crisp'
+  dat[ dat$Condition==3 , 'Vagueness'] <- 'Vague'
+  dat[ dat$Condition==4 , 'Vagueness'] <- 'Crisp'
+  dat$Vagueness <- as.factor(dat$Vagueness)
+  
+  # annotate use of a number / word in the instructions
+  dat[ dat$Condition==1 , 'Number'] <- 'Numeric'
+  dat[ dat$Condition==2 , 'Number'] <- 'Numeric'
+  dat[ dat$Condition==3 , 'Number'] <- 'Verbal'
+  dat[ dat$Condition==4 , 'Number'] <- 'Verbal'
+  dat$Number <- as.factor(dat$Number)
+  
+  # annotate Item
+  dat$Item <- factor(dat$Item, levels=c(1,2,3,4), labels=c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
+  
+  # annotate Order
+  dat$Order <- factor(dat$Order, levels=c(1,2), labels=c('LtoR','RtoL'))
+  
+  # annotate Quantity
+  dat$Quantity <- factor(dat$Quantity, levels=c(1,2), labels=c('Small','Large'))
+  
+  # annotate Subject
+  dat$Subject <- factor(paste("s",sprintf("%02d",dat$Subject),sep=""))
+  
+  # annotate Trial index for each participant (1:256 trials for 30 subjects)
+  dat$Trial = rep(x = 1:256, times = 30)
+  
+  # add an Observation index, over the whole data set (1:7680)
+  dat$Obs <- 1:7680
+  
+  # remove variables that don't matter for purposes of RT analysis
+  dat <- select(dat, Obs, Trial, Instruction, Subject, Item, Quantity, Order, Vagueness, Number, RT)
+  
+  return(dat)
+}  
+
+

### 2018-08-02 experiments/C_exp_1/writeup/2018_levy-method/concatenate_raw_data.R

@@ -1,0 +1,21 @@
+concatenate_raw_data = function() {
+  
+  number_of_valid_subjects=30
+  number_of_trials_per_subject=256
+  data_dir <- 'per_subject_raw_data_files/'
+  
+  column.headers.df <-  head( read.table(
+    paste(data_dir,'subject01.data',sep=''),
+    header=TRUE),0)
+
+  gathered.data <- column.headers.df
+  
+  for (subject in 1:number_of_valid_subjects) {
+    current.filename <- paste(data_dir, 'subject', sprintf("%02d",subject), '.data', sep='')
+    current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = TRUE)
+    gathered.data <- rbind(gathered.data, current.data)
+  }
+  
+  return(gathered.data)
+  
+}

### 2018-08-02 experiments/C_exp_1/writeup/2018_levy-method/declare_impossible_rt.R

@@ -1,0 +1,29 @@
+declare_impossible_rt = function(dat) {
+  message('Starting to declare impossible RTs')
+  dat$NaType = 'valid_trial'
+  # declare impossible trials as NA
+  # impossible trials are sticky fingers and timeouts
+  # sticky fingers
+  dat$NaType[dat$RT_original <= 1] <- 'sticky_fingers'
+  dat$RT[dat$RT_original <= 1] <- NA
+  dat$isBorderline[dat$RT_original <= 1] <- NA
+  dat$response_category[dat$RT_original <= 1] <- NA
+  # timeout
+  dat$NaType[dat$RT_original >= 59998] <- 'near_timeout'
+  dat$RT[dat$RT_original >= 59998 ] <- NA
+  dat$isBorderline[dat$RT_original >= 59998] <- NA
+  dat$response_category[dat$RT_original >= 59998] <- NA
+  
+  #dat$NaType[is.na(dat$isBorderline_original)] <- 'timeout'
+ # dat$isBorderline[dat$isBorderline_original=='timeout'] <- NA
+ # dat$response_category[dat$isBorderline_original=='timeout'] <- NA
+  
+  dat$response_category <- as.factor(dat$response_category)
+  dat$NaType <- as.factor(dat$NaType)
+  #dat$RT_original <- NULL
+  #dat$isBorderline_original <- NULL
+  dat$RT_log <- log(dat$RT)
+  dat$RT_rcp <- -1/dat$RT
+  message('Returning data with impossible RTs declared as NA in RT and isBorderline')
+  return(dat)
+}

### 2018-08-02 experiments/e2/writeup/2018_levy-method/declare_impossible_rt.R

@@ -1,28 +1,0 @@
-declare_impossible_rt = function(dat) {
-  message('Starting to declare impossible RTs')
-  dat$NaType = 'valid_trial'
-  # declare impossible trials as NA
-  # impossible trials are sticky fingers and timeouts
-  # sticky fingers
-  dat$NaType[dat$RT_original <= 1] <- 'sticky_fingers'
-  dat$RT[dat$RT_original <= 1] <- NA
-  dat$isBorderline[dat$RT_original <= 1] <- NA
-  dat$response_category[dat$RT_original <= 1] <- NA
-  # timeout
-  dat$NaType[dat$RT_original >= 59998] <- 'near_timeout'
-  dat$RT[dat$RT_original >= 59998 ] <- NA
-  dat$isBorderline[dat$RT_original >= 59998] <- NA
-  dat$response_category[dat$RT_original >= 59998] <- NA
-  
-  #dat$NaType[is.na(dat$isBorderline_original)] <- 'timeout'
- # dat$isBorderline[dat$isBorderline_original=='timeout'] <- NA
- # dat$response_category[dat$isBorderline_original=='timeout'] <- NA
-  
-  dat$response_category <- as.factor(dat$response_category)
-  dat$NaType <- as.factor(dat$NaType)
-  #dat$RT_original <- NULL
-  #dat$isBorderline_original <- NULL
-  dat$RT_log <- log(dat$RT)
-  message('Returning data with impossible RTs declared as NA in RT and isBorderline')
-  return(dat)
-}

### 2018-08-02 experiments/C_exp_1/writeup/C_exp_1.Rmd
more massive restructuring

---
title: "This is experiment one in the springer 1 to 3 numbering (originaly called e2)"
author: "Matt Green"
output:
  html_document:
    toc: yes
    toc_float: true
    mathjax: default
    template: default
    <!--code_folding: show-->
    theme: default
---

```{r 'set working directory', echo=FALSE}
root="/home/matt/wd/vagueness/experiments/e2/writeup/2018_levy-method"
if (getwd() != root) setwd(root)
```

```{r 'load Libraries and source my functions', message=FALSE, echo=FALSE}
library(knitr)
library(tidyverse)
library(tidymodels)
library(broom.mixed)
library(lme4)
library(lmerTest)
library(dotwhisker)
library(kableExtra)
library(LMERConvenienceFunctions)
library(data.table)
library(gridExtra)
source('summarySEwithin2.R')
source('concatenate_raw_data.R')
source('annotate_the_raw_data.R')
source('declare_impossible_rt.R')
source('remove_impossible_trials.R')
source('pretty_coef_table.R')
source('pretty_coef_plot.R')
```

```{r "read in the data", echo=F}
dat <- concatenate_raw_data()
dat <- annotate_the_raw_data(dat)
# at this point RT has min=1 (resp_type="sticky") and max=59,999 (resp_type="timeout")
dat <- subset(dat, RT>1 & RT<59999)
# at this point RT has min=445, max=42,685 (42 seconds) and nrow(dat) is 7677 down from 7680
# Now trim the data
dat <- perSubjectTrim.fnc(data = dat, response='RT', subject='Subject', trim = 2.5)$data %>% select(-SD, -Mean, -Scaled)
# Now log the remaining RT
dat$RT_log <- log(dat$RT)
```

# Analysis including 6:15:24

```{r "make means table", echo=F, message=F, warning=F, results='hide'}
dat_plot <- summarySEwithin2(dat, measurevar="RT_log", withinvars=c("Vagueness", "Number", "Item"), idvar="Subject")
```

```{r "plotmeans", echo=F}
# correction from Morey (2008) applied to condition means
mywidth=0
pdodge=position_dodge(width=mywidth)
#
ggplot(dat_plot, aes(x=Item, y=RT_logNormed, group=Vagueness, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci, shape=Vagueness, fill=Vagueness)) +
  facet_wrap(~Number) +
  scale_fill_grey(name="Vagueness", start=0, end=1) +  
  scale_shape_manual(name="Vagueness", values=c(21,22)) +
  theme(aspect.ratio = 1, panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background=element_rect(fill="white", colour="black")) +
  ylab("RT log(ms)") +
  #
  geom_errorbar(position=pdodge, width=0.25) +
  geom_line(position=pdodge) +
  geom_point(position=pdodge, size=2) 
```

Original full model
```{r "original full RT model", echo=F}
dat_model <- dat
dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5)
dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
rtFullModel <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag * c_Num + c_Itm | Subject), dat_model)
pretty_coef_table(rtFullModel, "rtFullModel")
```

The reviewer wants us to re-run this analysis without 6:15:24 but just removing that level results in a model that doesn't converge.
I had to drop c_Vag * c_Itm and c_Itm slopes over subjects to get it to converge.

```{r "original full Model after dropping 6:15:24", echo=F}
dat_model <- droplevels(subset(dat, Item!="06:15:24"))
dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5)
dat_model$c_Itm <- ifelse(dat_model$Item=="16:25:34", -.3333, ifelse(dat_model$Item=="26:35:44", .0000, .3333))

# now trying dropping c_Vag * Item slopes for subject, but the model didn't converge.
#rtFullModelMinus <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag + c_Num + c_Itm | Subject), dat_model)

# now tring droppping c_Vag * Item slopes for subject, and c_Itm slopes for subject
rtFullModelMinus <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag + c_Num  | Subject), dat_model)
pretty_coef_table(rtFullModelMinus, "rtFullModelMinus 6:15:24")
```

So, after dropping the 6:15:24 level, there is still a disadvantage for Vagueness, but it is not significant (p=.24). There is still a significant disadvantage for Numeric instruction format (p<.001). There is still a disadvantage for increasing the number of dots but it is not significant (p=.323). The interaction between Vagueness and Instruction format changes sign, and is now non-significant (beta=-.022, p=.374).

However the full model and the reduced model are not really equivalent because the terms are different in order for the reduced model to converge. What we really want is a full model structure that converges with and without the 6:15:24 item.

If we use the structure of the reduced model for the structure of the full model we get this for the full model:

```{r "full RT model", echo=F}
dat_model <- dat
dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5)
dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
rtFullModel2 <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag + c_Num  | Subject), dat_model)
pretty_coef_table(rtFullModel2, "rtFullModel2")
```

and this for the reduced model:

```{r}
pretty_coef_table(rtFullModelMinus, "rtFullModelMinus 6:15:24")
```

fullmodel with Helmert contrasts on Item
```{r "full RT model Helmert", echo=F}
dat_model <- dat
dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5)
contrasts(dat_model$Item) <- contr.helmert(4)
# had to drop c_Vag * Item slopes for subject
rtFullModel_helmert <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + Item + (1 + c_Vag + c_Num + Item | Subject), dat_model)
pretty_coef_table(rtFullModel_helmert, "c_Vag * c_Num + Item + (1 + c_Vag + c_Num + Item | Subject)")
```

reducedmodel with Helmert contrasts on Item
```{r "reduced RT model Helmert", echo=F}
dat_model <- droplevels(subset(dat, Item!="06:15:24"))
dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5)
contrasts(dat_model$Item) <- contr.helmert(3)
# had to drop c_Vag * Item slopes for subject
rtReducedModel_helmert <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + Item + (1 + c_Vag + c_Num + Item | Subject), dat_model)
pretty_coef_table(rtReducedModel_helmert, "c_Vag * c_Num + Item + (1 + c_Vag + c_Num + Item | Subject)")
```

However the plot of means suggests that the main variables are c_Num and Item so it might be better to have a model structure that includes a c_Num*Item term.

```{r}
dat_model <- dat
dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5)
contrasts(dat_model$Item) <- contr.helmert(4)
myMod1 <- lmerTest::lmer(1 + RT_log ~ c_Vag + c_Num + Item + c_Num:Item +(1|Subject), dat_model)
pretty_coef_table(myMod1, "myMod1")
```

```{r}
dat_model <- droplevels(subset(dat, Item!="06:15:24"))
dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5)
contrasts(dat_model$Item) <- contr.helmert(3)
myMod2 <- lmerTest::lmer(1 + RT_log ~ c_Vag + c_Num + Item + c_Num:Item +(1|Subject), dat_model)
pretty_coef_table(myMod2, "myMod2:no6:15:24")
```

<!-- Show which RTs were removed by trimming -->
<!-- ```{r} -->
<!-- ggplot(data=dat, aes(x=Trial, y=RT)) + -->
<!--   geom_point(aes(colour=was_RT_removed_by_trimming), size=.5) +  -->
<!--   facet_wrap(~Subject)+ -->
<!--   theme(aspect.ratio = 1, panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background=element_rect(fill="white", colour="black")) -->
<!-- ``` -->

<!-- Show which Rts were removed by trimming, shown as log RT -->
<!-- ```{r} -->
<!-- ggplot(data=dat, aes(x=Trial, y=log(RT))) + -->
<!--   geom_point(aes(colour=was_RT_removed_by_trimming), size=.5) +  -->
<!--   facet_wrap(~Subject)+ -->
<!--   theme(aspect.ratio = 1, panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background=element_rect(fill="white", colour="black")) -->
<!-- ``` -->

<!-- Show how distributions of Rt change according to trimming, in log RT -->
<!-- ```{r} -->
<!-- dat %>% gather(key=whether_trimmed, value=time, RT, RT_trimmed) %>% -->
<!--   ggplot(aes(log(time))) + -->
<!--   geom_density(aes(group=whether_trimmed, colour=whether_trimmed)) +  -->
<!--   facet_wrap(~Subject, scales="free") + -->
<!--   theme(aspect.ratio = 1, panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background=element_rect(fill="white", colour="black"))  -->
<!-- ``` -->

<!-- Make RT log and reciprocal, from the trimmed RT in ms. -->
<!-- ```{r} -->
<!-- dat$RT_trimmed_log <- log(dat$RT_trimmed) -->
<!-- dat$RT_trimmed_rcp <- -1/(dat$RT_trimmed) -->
<!-- ``` -->

<!-- plot transformations of rt in the trimmed data -->

<!-- ```{r, warning=FALSE, echo=F} -->
<!-- p.dens <- -->
<!-- dat %>% -->
<!--   gather(key=metric, value=value, RT_trimmed, RT_trimmed_log, RT_trimmed_rcp) %>% -->
<!--     ggplot(aes(value)) + -->
<!--     geom_density() + -->
<!--     facet_wrap(~metric, scales="free") + -->
<!--     theme(aspect.ratio = 1, panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background=element_rect(fill="white", colour="black")) -->
<!-- ``` -->

<!-- ```{r, warning=FALSE, echo=F} -->
<!-- p.point <- -->
<!-- dat %>% -->
<!--   gather(key=metric, value=value, RT_trimmed, RT_trimmed_log, RT_trimmed_rcp) %>% -->
<!--     ggplot(aes(x=Obs, y=value)) + -->
<!--     geom_point(size=.1) + -->
<!--     facet_wrap(~metric, scales="free") + -->
<!--     theme(aspect.ratio = 1, panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background=element_rect(fill="white", colour="black")) -->
<!-- ``` -->

<!-- ```{r, warning=FALSE, echo=F} -->
<!-- grid.arrange(p.dens,p.point) -->
<!-- ``` -->

<!-- Compute the ci's from untransformed data that were trimmed using summarysewithin2 and plot them on a log scale to reproduce the plot in the Springer submission. -->

<!-- ```{re2-rtplot, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE} -->
<!-- rtdataplot1 <- summarySEwithin2(dat, measurevar="RT_trimmed_log", withinvars=c("Vagueness", "Number", "Item"), idvar="Subject") -->

<!-- rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness)) -->

<!-- ggplot(rtdataplot1, aes(x=Item, group=Condition, fill=Vagueness)) +  -->
<!--   geom_line(aes(x=Item, y=log(RT_trimmedNormed))) + -->
<!--   facet_grid(~Number) -->

<!-- +  -->
<!--   geom_errorbar(width=.2, aes(ymin=RT_trimmedNormed-ci, ymax=RT_trimmedNormed+ci)) +  -->
<!--   geom_point(size=2, col=1, aes(shape=Vagueness)) +  -->
<!--   scale_fill_grey(name="Vagueness", start=0, end=1) +   -->
<!--   scale_shape_manual(name="Vagueness", values=c(21,22)) +  -->
<!--   ggtitle("Response time") +  -->
<!--   ylab("Respone time (log (ms))") +  -->
<!--   xlab(NULL) +  -->
<!--   theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), legend.position="right", aspect.ratio=1, plot.background=element_rect(fill=NA, color='white'), strip.background=element_blank(), axis.text.x = element_text(angle = 15))  -->

<!-- ``` -->

<!-- ```{r e2-rtplot, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE} -->
<!-- rtdataplot1 <- summarySEwithin2(dat, measurevar="RT_trimmed_log", withinvars=c("Vagueness", "Number", "Item"), idvar="Subject") -->

<!-- rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness)) -->

<!-- rtplot <-  -->
<!--   ggplot(rtdataplot1, aes(y=RT_trimmed_logNormed, x=Item, group=Condition, fill=Vagueness)) +  -->
<!--   geom_line() +  -->
<!--   geom_errorbar(width=.2, aes(ymin=RT_trimmed_logNormed-ci, ymax=RT_trimmed_logNormed+ci)) +  -->
<!--   geom_point(size=2, col=1, aes(shape=Vagueness)) +  -->
<!--   scale_fill_grey(name="Vagueness", start=0, end=1) +   -->
<!--   scale_shape_manual(name="Vagueness", values=c(21,22)) +  -->
<!--   ggtitle("Response time") +  -->
<!--   ylab("Respone time (log (ms))") +  -->
<!--   xlab(NULL) +  -->
<!--   theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), legend.position="right", aspect.ratio=1, plot.background=element_rect(fill=NA, color='white'), strip.background=element_blank(), axis.text.x = element_text(angle = 15)) + facet_grid(~Number) -->

<!-- rtplot -->
<!-- ``` -->

<!-- ## Response time models  -->

<!-- ### Full RT model -->

<!-- ```{r 'rt-full-model', cache=TRUE, echo=TRUE} -->
<!-- rtFullModel <- lmerTest::lmer(RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag * c_Num + c_Itm | Subject), rtdata) -->
<!-- ``` -->

<!-- ```{r 'make a pretty coefficients table for rtFullModel', echo=FALSE, results='asis'} -->
<!-- pretty_coef_table(rtFullModel, "rtFullModel") %>% print() -->
<!-- ``` -->

<!-- ```{r, 'make a pretty coefficients plot for rtFullModel', echo=FALSE, fig.width=6, fig.height=3} -->
<!-- pretty_coef_plot(rtFullModel, "rtFullModel") -->
<!-- ``` -->

<!-- ### Numeric only RT model -->

<!-- ```{r 'rt-numeric-only model', cache=TRUE, echo=TRUE} -->
<!-- rtNumOnlyModel <- lmerTest::lmer(RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject), subset(rtdata,Number=="Numeric")) -->
<!-- ``` -->

<!-- ```{r 'make a pretty coefficients table for rtNumOnlyModel', echo=FALSE, results='asis'} -->
<!-- pretty_coef_table(rtNumOnlyModel, "rtNumOnlyModel") %>% print() -->
<!-- ``` -->

<!-- ```{r, 'make a pretty coefficients plot for rtNumOnlyModel', echo=FALSE, fig.width=6, fig.height=3} -->
<!-- pretty_coef_plot(rtNumOnlyModel, "rtNumOnlyModel") -->
<!-- ``` -->

<!-- ### Verbal only RT model -->

<!-- ```{r 'rt-verbal-only model', cache=TRUE, echo=TRUE} -->
<!-- rtVerbalOnlyModel <- lmerTest::lmer(RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject), subset(rtdata,Number=="Verbal")) -->
<!-- ``` -->

<!-- ```{r 'make a pretty coefficients table for rtVerbalOnlyModel', echo=FALSE, results='asis'} -->
<!-- pretty_coef_table(rtVerbalOnlyModel, "rtVerbalOnlyModel") %>% print() -->
<!-- ``` -->

<!-- ```{r, 'make a pretty coefficients plot for rtVerbalOnlyModel', echo=FALSE, fig.width=6, fig.height=3} -->
<!-- pretty_coef_plot(rtVerbalOnlyModel, "rtVerbalOnlyModel") -->
<!-- ``` -->

<!-- # Borderline response data -->

<!-- ```{r} -->
<!-- bldata <- trim_data -->
<!-- ``` -->

<!-- ```{r e2-blBarChart, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE} -->
<!-- bldata$response_category <- relevel(bldata$response_category, ref='expected') -->

<!-- blBarChart = ggplot(bldata) + geom_bar(aes(response_category, group=Number:Vagueness, fill=Vagueness), position=position_dodge(width=NULL)) +   scale_fill_grey() + xlab(NULL) + ggtitle('Borderline response distribution') + facet_grid(~Number) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), legend.position="right", aspect.ratio=1, plot.background = element_rect(fill=NA, color='white'), strip.background=element_blank(), legend.key.size=unit(4, 'mm'), axis.text.x = element_text(angle = 15)) -->

<!-- blBarChart -->
<!-- ``` -->

<!-- ## Borderline response models -->

<!-- ### Full borderline response model -->

<!-- ```{r 'bl-full-model', cache=TRUE, echo=TRUE} -->
<!-- blFullModel <- lme4::glmer(isBorderline ~ c_Vag * c_Num + c_Itm + (1 + c_Vag * c_Num + c_Itm | Subject), bldata, family="binomial", control = glmerControl(optimizer = "bobyqa")) -->
<!-- ``` -->

<!-- ```{r 'make a pretty coefficients table for blFullModel', echo=FALSE, results='asis'} -->
<!-- pretty_coef_table(blFullModel, "blFullModel") %>% print() -->
<!-- ``` -->

<!-- ```{r, 'make a pretty coefficients plot for blFullModel', echo=FALSE, fig.width=6, fig.height=3} -->
<!-- pretty_coef_plot(blFullModel, "blFullModel") -->
<!-- ``` -->

<!-- Original fullmodel with Helmert contrasts on Item -->
<!-- ```{r "original full RT model Helmert", echo=4:5} -->
<!-- dat_model <- dat -->
<!-- dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5) -->
<!-- dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5) -->
<!-- contrasts(dat_model$Item) <- contr.helmert(4) -->
<!-- # had to drop c_Vag * Item slopes for subject -->
<!-- rtFullModel_helmert <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + Item + (1 + c_Vag + c_Num + Item | Subject), dat_model) -->
<!-- pretty_coef_table(rtFullModel_helmert, "rtFullModel Helmert") -->
<!-- ``` -->

<!-- # Re-run without 6:15:24 -->
<!-- ```{r, echo=F, message=F, warning=F, results='hide'} -->
<!-- dat_reduced <- droplevels(subset(dat, Item!="06:15:24")) -->
<!-- dat_plot <- summarySEwithin2(dat_reduced, measurevar="RT_log", withinvars=c("Vagueness", "Number", "Item"), idvar="Subject") -->
<!-- ``` -->

<!-- ```{r, echo=F, results='hide'} -->
<!-- # correction from Morey (2008) applied to condition means -->
<!-- mywidth=0 -->
<!-- pdodge=position_dodge(width=mywidth) -->
<!-- # -->
<!-- ggplot(dat_plot, aes(x=Item, y=RT_logNormed, group=Vagueness, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci, shape=Vagueness, fill=Vagueness)) + -->
<!--   facet_wrap(~Number) + -->
<!--   scale_fill_grey(name="Vagueness", start=0, end=1) +   -->
<!--   scale_shape_manual(name="Vagueness", values=c(21,22)) + -->
<!--   theme(aspect.ratio = 1, panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background=element_rect(fill="white", colour="black")) + -->
<!--   ylab("RT log(ms)") + -->
<!--   # -->
<!--   geom_errorbar(position=pdodge, width=0.25) + -->
<!--   geom_line(position=pdodge) + -->
<!--   geom_point(position=pdodge, size=2)  -->
<!-- ``` -->

### 2018-08-02 experiments/C_exp_1/writeup/C_exp_1.Rmd
Don't know what to call this commit

@@
 ---
-title: "This is experiment one in the springer 1 to 3 numbering (originaly called e2)"
+title: "C_exp_1"
 author: "Matt Green"
@@
     toc: yes
-    toc_float: true
-    mathjax: default
-    template: default
-    <!--code_folding: show-->
-    theme: default
+    toc_float: false
 ---
-
-```{r 'set working directory', echo=FALSE}
-root="/home/matt/wd/vagueness/experiments/e2/writeup/2018_levy-method"
-if (getwd() != root) setwd(root)
-```
 
@@
 
-```{r "read in the data", echo=F}
+```{r, echo=FALSE}
 dat <- concatenate_raw_data()
 dat <- annotate_the_raw_data(dat)
+```
+
+# Info and Procedure (Method)
+
+To find out what happens when words are used in a context where their potential for vagueness comes to the fore, Experiment 1 used three arrays (rather than two arrays as in pilot experiment B) so that the vague description had more than one possible referent; it used indefinite articles in the vague instructions to avoid the impression that only one response counted as correct; and it was carried out without error feedback.
+
+An indication that the potential for vagueness was realised in Experiment 1 is that the borderline response was chosen fairly often: 16\% of the time.
+
+In Experiment 1, an item was a referring expression instruction followed by a set of three dot arrays defined by a triple of numbers, representing the number of dots in the left, middle, and right arrays.
+We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54).
+Each set of arrays comprised three arrays (instead of two as in pilot experiment B); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had more, and these flanking arrays appeared equally often on the left and right of the central array.
+
+The way in which borderline responses were construed is as follows, using as an example the array (6:15:24) and instructions that identified the smaller flanking array (6).
+6 was classified as the expected response.
+15 was classified as the borderline response.
+24 was classified as the extreme response.
+
+* In the "vague numerical" condition the instruction was "Choose a square with about 10 dots" -- none of the arrays contained exactly 10 dots, but 10 is closer to 6 than it is to 15, making 6 a better response to that instruction, 15 a borderline response, and 24 an extreme response.
+
+* In the vague verbal condition we used "Choose a square with few dots". We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
+
+* In the crisp numerical condition we used "Choose the square with 6 dots".
+The smaller flanking array always contained exactly the specified number of dots.
+We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
+
+* For crisp verbal, we used Choose the square with the fewest dots. We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
+
+On each trial, first the referring expression that constituted the instruction for that trial was displayed. 
+Participants then pressed a key to indicate that they had read the instruction. 
+After 1000 ms, the arrays were presented, while preserving the text of the referring expression. 
+The response time dependent variable was measured from the presentation of the arrays, until the keypress indicating the participant's choice, which was also recorded. 
+The trial would timeout after 60 seconds if there was no response.
+In this experiment, no feedback was given.
+This was because, in the vague conditions, we did not regard any response as "correct" or "incorrect", but instead as "expected response"; "borderline response"; and "extreme response", and we did not want to draw participants' attention to this distinction explicitly.
+We simply recorded whether the participant chose the best referent, the borderline case or the poorest referent, and how long it took the participant to respond.
+
+# Sample display
+
+Remember to do a screenshot for the sample display.
+
+# Full table of instructions
+
+```{r, echo=FALSE}
+instructions_table <- 
+  dat %>%
+    select(Item, Quantity, Vagueness, Number, Instruction) %>%
+    unique() %>%
+    arrange(Item, Quantity, Vagueness, Number) %>%
+    spread(key=Vagueness, value=Instruction)
+```
+
+```{r, echo=FALSE}
+instructions_table %>% 
+  kable(align='cccll', caption="Full table of instructions") %>% 
+      kable_styling(full_width = F, position = "left", font_size = 11) 
+```
+
+```{r, echo=FALSE, eval=FALSE}
+kable(instructions_table, format="latex", booktabs = TRUE)
+```
+
+# Means plots
+
+```{r, echo=F, message=F, warning=F, results='hide'}
 # at this point RT has min=1 (resp_type="sticky") and max=59,999 (resp_type="timeout")
@@
 dat <- perSubjectTrim.fnc(data = dat, response='RT', subject='Subject', trim = 2.5)$data %>% select(-SD, -Mean, -Scaled)
-# Now log the remaining RT
+# Now add log(RT) for the remaining RTs
 dat$RT_log <- log(dat$RT)
@@
 
-# Analysis including 6:15:24
-
-```{r "make means table", echo=F, message=F, warning=F, results='hide'}
+```{r, echo=F}
 dat_plot <- summarySEwithin2(dat, measurevar="RT_log", withinvars=c("Vagueness", "Number", "Item"), idvar="Subject")
@@
 
-```{r "plotmeans", echo=F}
+```{r, echo=F}
 # correction from Morey (2008) applied to condition means
@@
 
-Original full model
-```{r "original full RT model", echo=F}
+# Hypotheses
+
+We formulated the following hypotheses for Experiment 1:
+
+* [Hypothesis 1] Crisp/Vague RT:
+    * Vague instructions should result in faster responses than crisp instructions; and this pattern should hold when the model is restricted to numeric-only data and when it is restricted to verbal-only data.
+* [Hypothesis 2] Numeric/Verbal RT: 
+    * There should be no real difference between responses to Numeric instructions and Verbal instructions (based on our interpretation of pilot experiment B, where we thought that vague instructions alone were driving the advantage for instructions that were both vague, and also in verbal format).
+* [Hypothesis 3] Item RT: 
+    * Responses should take longer as the number of dots in the display grows larger (i.e., as the levels of Item increase).
+* [Hypothesis 4] Crisp/Vague Response Type: 
+    * Vague instructions should lead to more borderline responses than crisp instructions.
+
+# Results
+
+## Full data
+
+Original full model: RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag * c_Num + c_Itm | Subject)
+```{r "original full RT model", echo=F, cache=TRUE}
 dat_model <- dat
@@
 
-The reviewer wants us to re-run this analysis without 6:15:24 but just removing that level results in a model that doesn't converge.
-I had to drop c_Vag * c_Itm and c_Itm slopes over subjects to get it to converge.
+Numeric only
+```{r, echo=FALSE}
+dat_model <- droplevels(subset(dat, Number=="Numeric"))
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
+rtFullModel_num <- lmerTest::lmer(1 + RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject), dat_model)
+pretty_coef_table(rtFullModel_num, "rtFullModel_num")
+```
 
-```{r "original full Model after dropping 6:15:24", echo=F}
+Verbal only
+```{r, echo=FALSE}
+dat_model <- droplevels(subset(dat, Number=="Verbal"))
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
+rtFullModel_verb <- lmerTest::lmer(1 + RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject), dat_model)
+pretty_coef_table(rtFullModel_verb, "rtFullModel_verb")
+```
+
+On the basis of the initial full model:
+
+* [Hypothesis 1] Crisp/Vague RT:
+    * Vague instructions actually led to significantly slower responses than crisp instructions, against Hypothesis 1.
+    * When the model was restricted to numeric-only instructions Vague instructions still led to significantly slower responses than crisp instructions
+    * When the model was restricted to verbal-only instructions Vague instructions tended to slow responses, but not significantly.
+* [Hypothesis 2] Numeric/Verbal RT: 
+    * There was actually a significant difference between numeric and verbal instructions, with numeric instructions leading to longer responses than verbal instructions, against Hypothesis 2
+* [Hypothesis 3] Item RT: 
+    * Responses took longer as the levels of Item increased, supporting Hypothesis 3
+* [Hypothesis 4] Response Type:
+    * Participants were significantly more likely to choose the borderline option for vague instructions than for crisp instructions (Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format).
+* _Interaction comment goes here_
+
+## Data less 6:15:24
+
+However, given that the plot shows that responses to 6:15:24 in the "crisp numeric" instructions condition were extremely fast relative to the "vague numeric" instructions to 6:15:24, the effects in the model of the full dataset could be driven by this difference.
+
+A clearer picture of the effects of interest might be obtained by removing the 6:15:24 level of Item from the data set, and fitting the model to this restricted data.
+Doing this results in the effects tabled below.
+
+With less data available, the model formula had to be simplified in order to converge -- specifically the following terms were dropped: per-subject slopes for the Vagueness by Item interaction, and per-subject slopes for the effect of Item.
+
+Model of the data after the 6:15:24 level of Item is removed: RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag + c_Num  | Subject)
+```{r, echo=F, cache=TRUE}
 dat_model <- droplevels(subset(dat, Item!="06:15:24"))
@@
 dat_model$c_Itm <- ifelse(dat_model$Item=="16:25:34", -.3333, ifelse(dat_model$Item=="26:35:44", .0000, .3333))
-
-# now trying dropping c_Vag * Item slopes for subject, but the model didn't converge.
-#rtFullModelMinus <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag + c_Num + c_Itm | Subject), dat_model)
-
-# now tring droppping c_Vag * Item slopes for subject, and c_Itm slopes for subject
-rtFullModelMinus <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag + c_Num  | Subject), dat_model)
-pretty_coef_table(rtFullModelMinus, "rtFullModelMinus 6:15:24")
+rtRestrictedModel <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag + c_Num  | Subject), dat_model)
+pretty_coef_table(rtRestrictedModel, "rtRestrictedModel")
 ```
 
-So, after dropping the 6:15:24 level, there is still a disadvantage for Vagueness, but it is not significant (p=.24).
-There is still a significant disadvantage for Numeric instruction format (p<.001).
-There is still a disadvantage for increasing the number of dots but it is not significant (p=.323).
-The interaction between Vagueness and Instruction format changes sign, and is now non-significant (beta=-.022, p=.374).
-
-However the full model and the reduced model are not really equivalent because the terms are different in order for the reduced model to converge.
-What we really want is a full model structure that converges with and without the 6:15:24 item.
-
-If we use the structure of the reduced model for the structure of the full model we get this for the full model:
-
-```{r "full RT model", echo=F}
-dat_model <- dat
+Numeric only after the 6:15:24 level of Item is removed
+```{r, echo=F, cache=TRUE}
+dat_model <- droplevels(subset(dat, Item!="06:15:24" & Number=="Numeric"))
 dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
-dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5)
-dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
-rtFullModel2 <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag + c_Num  | Subject), dat_model)
-pretty_coef_table(rtFullModel2, "rtFullModel2")
+dat_model$c_Itm <- ifelse(dat_model$Item=="16:25:34", -.3333, ifelse(dat_model$Item=="26:35:44", .0000, .3333))
+rtRestrictedModel_num <- lmerTest::lmer(1 + RT_log ~ c_Vag + c_Itm + (1 + c_Vag | Subject), dat_model)
+pretty_coef_table(rtRestrictedModel_num, "rtRestrictedModel_num")
 ```
 
-and this for the reduced model:
-
-```{r}
-pretty_coef_table(rtFullModelMinus, "rtFullModelMinus 6:15:24")
+Verbal-only after the 6:15:24 level of Item is removed
+```{r, echo=F, cache=TRUE}
+dat_model <- droplevels(subset(dat, Item!="06:15:24" & Number=="Verbal"))
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="16:25:34", -.3333, ifelse(dat_model$Item=="26:35:44", .0000, .3333))
+rtRestrictedModel_verb <- lmerTest::lmer(1 + RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject), dat_model)
+pretty_coef_table(rtRestrictedModel_verb, "rtRestrictedModel_verb")
 ```
 
-fullmodel with Helmert contrasts on Item
-```{r "full RT model Helmert", echo=F}
-dat_model <- dat
-dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
-dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5)
-contrasts(dat_model$Item) <- contr.helmert(4)
-# had to drop c_Vag * Item slopes for subject
-rtFullModel_helmert <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + Item + (1 + c_Vag + c_Num + Item | Subject), dat_model)
-pretty_coef_table(rtFullModel_helmert, "c_Vag * c_Num + Item + (1 + c_Vag + c_Num + Item | Subject)")
-```
+On the basis of the restricted model:
 
-reducedmodel with Helmert contrasts on Item
-```{r "reduced RT model Helmert", echo=F}
-dat_model <- droplevels(subset(dat, Item!="06:15:24"))
-dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
-dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5)
-contrasts(dat_model$Item) <- contr.helmert(3)
-# had to drop c_Vag * Item slopes for subject
-rtReducedModel_helmert <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + Item + (1 + c_Vag + c_Num + Item | Subject), dat_model)
-pretty_coef_table(rtReducedModel_helmert, "c_Vag * c_Num + Item + (1 + c_Vag + c_Num + Item | Subject)")
-```
+* [Hypothesis 1] Crisp/Vague RT:
+    * after dropping the 6:15:24 level, there is still a disadvantage for Vagueness, but it is not significant (p=.24).
+    * When the model was restricted to numeric-only instructions the disadvantage for vagueness was very small and not significant.
+    * When the model was restricted to verbal-only instructions there was a small and non-significant disadvantage for vagueness.
+* [Hypothesis 2] Numeric/Verbal RT: 
+    * There is still a significant disadvantage for Numeric instruction format (p<.001).
+* [Hypothesis 3] Item RT: 
+    * There is still a disadvantage for increasing the number of dots but it is not significant (p=.323).
+* _Interaction comment goes here_ The interaction between Vagueness and Instruction format changes sign, and is now non-significant (beta=-.022, p=.374).
 
-However the plot of means suggests that the main variables are c_Num and Item so it might be better to have a model structure that includes a c_Num*Item term.
+# Discussion
 
-```{r}
-dat_model <- dat
-dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
-dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5)
-contrasts(dat_model$Item) <- contr.helmert(4)
-myMod1 <- lmerTest::lmer(1 + RT_log ~ c_Vag + c_Num + Item + c_Num:Item +(1|Subject), dat_model)
-pretty_coef_table(myMod1, "myMod1")
-```
+<!-- So,     -->
 
-```{r}
-dat_model <- droplevels(subset(dat, Item!="06:15:24"))
-dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
-dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5)
-contrasts(dat_model$Item) <- contr.helmert(3)
-myMod2 <- lmerTest::lmer(1 + RT_log ~ c_Vag + c_Num + Item + c_Num:Item +(1|Subject), dat_model)
-pretty_coef_table(myMod2, "myMod2:no6:15:24")
-```
+<!-- However the full model and the reduced model are not really equivalent because the terms are different in order for the reduced model to converge.
+What we really want is a full model structure that converges with and without the 6:15:24 item.
+-->
+
+<!-- If we use the structure of the reduced model for the structure of the full model we get this for the full model: -->
+
+<!-- ```{r "full RT model", echo=F} -->
+<!-- dat_model <- dat -->
+<!-- dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5) -->
+<!-- dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5) -->
+<!-- dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75))) -->
+<!-- rtFullModel2 <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag + c_Num  | Subject), dat_model) -->
+<!-- pretty_coef_table(rtFullModel2, "rtFullModel2") -->
+<!-- ``` -->
+
+<!-- and this for the reduced model: -->
+
+<!-- ```{r} -->
+<!-- pretty_coef_table(rtFullModelMinus, "rtFullModelMinus 6:15:24") -->
+<!-- ``` -->
+
+<!-- fullmodel with Helmert contrasts on Item -->
+<!-- ```{r "full RT model Helmert", echo=F} -->
+<!-- dat_model <- dat -->
+<!-- dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5) -->
+<!-- dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5) -->
+<!-- contrasts(dat_model$Item) <- contr.helmert(4) -->
+<!-- # had to drop c_Vag * Item slopes for subject -->
+<!-- rtFullModel_helmert <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + Item + (1 + c_Vag + c_Num + Item | Subject), dat_model) -->
+<!-- pretty_coef_table(rtFullModel_helmert, "c_Vag * c_Num + Item + (1 + c_Vag + c_Num + Item | Subject)") -->
+<!-- ``` -->
+
+<!-- reducedmodel with Helmert contrasts on Item -->
+<!-- ```{r "reduced RT model Helmert", echo=F} -->
+<!-- dat_model <- droplevels(subset(dat, Item!="06:15:24")) -->
+<!-- dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5) -->
+<!-- dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5) -->
+<!-- contrasts(dat_model$Item) <- contr.helmert(3) -->
+<!-- # had to drop c_Vag * Item slopes for subject -->
+<!-- rtReducedModel_helmert <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + Item + (1 + c_Vag + c_Num + Item | Subject), dat_model) -->
+<!-- pretty_coef_table(rtReducedModel_helmert, "c_Vag * c_Num + Item + (1 + c_Vag + c_Num + Item | Subject)") -->
+<!-- ``` -->
+
+<!-- However the plot of means suggests that the main variables are c_Num and Item so it might be better to have a model structure that includes a c_Num*Item term. -->
+
+<!-- ```{r} -->
+<!-- dat_model <- dat -->
+<!-- dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5) -->
+<!-- dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5) -->
+<!-- contrasts(dat_model$Item) <- contr.helmert(4) -->
+<!-- myMod1 <- lmerTest::lmer(1 + RT_log ~ c_Vag + c_Num + Item + c_Num:Item +(1|Subject), dat_model) -->
+<!-- pretty_coef_table(myMod1, "myMod1") -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- dat_model <- droplevels(subset(dat, Item!="06:15:24")) -->
+<!-- dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5) -->
+<!-- dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5) -->
+<!-- contrasts(dat_model$Item) <- contr.helmert(3) -->
+<!-- myMod2 <- lmerTest::lmer(1 + RT_log ~ c_Vag + c_Num + Item + c_Num:Item +(1|Subject), dat_model) -->
+<!-- pretty_coef_table(myMod2, "myMod2:no6:15:24") -->
+<!-- ``` -->
+
+<!-- ################################################################################################## -->
 

### 2018-08-02 springer/current/e1.tex
Don't know what to call this commit

@@
 \centering
-\caption{Experiment 2: instructions arranged by condition for the dot triple $(6,15,24)$ and instructions indicating the smaller numbers}
+\caption{Experiment 1 instructions arranged by condition. The instructions given in the table started with ``Choose \ldots"}
 \label{instructionse2}
-\begin{tabular}{rll}
+\begin{tabular}{cccll}
 \hline\noalign{\smallskip}
-&crisp&vague\\
+Item & Quantity & Number & Crisp & Vague\\
 \noalign{\smallskip}\hline\noalign{\smallskip}
-numerical	&	Choose the square with 6 dots 			& 	Choose a square with about 10 dots	\\
-verbal		&	Choose the square with the fewest dots	&	Choose a square with few dots		\\
+06:15:24 & Small & Numeric & the square with 6 dots & a square with about 10 dots\\
+06:15:24 & Small & Verbal & the square with the fewest dots & a square with few dots\\
+06:15:24 & Large & Numeric & the square with 24 dots & a square with about 20 dots\\
+06:15:24 & Large & Verbal & the square with the most dots & a square with many dots\\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+16:25:34 & Small & Numeric & the square with 16 dots & a square with about 20 dots\\
+16:25:34 & Small & Verbal & the square with the fewest dots & a square with few dots\\
+16:25:34 & Large & Numeric & the square with 34 dots & a square with about 30 dots\\
+16:25:34 & Large & Verbal & the square with the most dots & a square with many dots\\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+26:35:44 & Small & Numeric & the square with 26 dots & a square with about 30 dots\\
+26:35:44 & Small & Verbal & the square with the fewest dots & a square with few dots\\
+26:35:44 & Large & Numeric & the square with 44 dots & a square with about 40 dots\\
+26:35:44 & Large & Verbal & the square with the most dots & a square with many dots\\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+36:45:54 & Small & Numeric & the square with 36 dots & a square with about 40 dots\\
+36:45:54 & Small & Verbal & the square with the fewest dots & a square with few dots\\
+36:45:54 & Large & Numeric & the square with 54 dots & a square with about 50 dots\\
+36:45:54 & Large & Verbal & the square with the most dots & a square with many dots\\
 \noalign{\smallskip}\hline

### 2018-08-03 experiments/C_exp_1/writeup/C_exp_1.Rmd
minor
added discussion

@@
 library(tidyverse)
-library(tidymodels)
+#library(tidymodels)
 library(broom.mixed)
@@
     * When the model was restricted to numeric-only instructions the disadvantage for vagueness was very small and not significant.
-    * When the model was restricted to verbal-only instructions there was a small and non-significant disadvantage for vagueness.
+    * When the model was restricted to verbal-only instructions the disadvantage for vagueness was small and not significant.
 * [Hypothesis 2] Numeric/Verbal RT: 
@@
 # Discussion
+
+This experiment tested whether vague instructions would result in faster responses than crisp instructions, when borderline cases were present.
+Faster responses for vague instructions were found in pilot experiment B, but there were no borderline cases in that experiment.
+
+In this experiment we found in contrast that vague instructions resulted in slower responses than crisp instructions: a difference that was significant when considering the full data (112ms), but which was not significant after removing the smallest arrays from the analysis, which had a pattern opposite to the main trends in the rest of the data.
+
+We also found that the effect of instruction format was significant, with numerical format slowing responses by 689 ms on average, such that the disadvantage of numerical format overwhelmed the contribution of vagueness.
+The verbal vague condition still yielded faster responses than the numerical crisp condition, so the pattern from pilot experiment B was reproduced, but in the light of the evidence from this experiment (Experiment 1), in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like an advantage of verbal instruction format.
+
+However, once again there is a possibly confounding factor.
+Observe that, in Experiment 1, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) "selection algorithm": To see this, consider the task of selecting the dot array that contains "few dots":" to do this, it suffices to _compare_ the three arrays and select the one that contains the fewest elements.
+To select the dot array that contains "16 dots" seems to require the participant to estimate, and then _match_, the cardinality of (at least) one dot array to 16, a process which could plausibly take longer, independently of vagueness.
+Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use _comparison_ rather than having to resort to _matching_.
+
+In the next two experiments we pitted the comparison algorithm and matching algorithm selection tasks against each other while controlling vagueness and instruction format.
+In Experiment 2 we restricted all the instructions to numeric quantifiers while factorially manipulating vagueness and selection task.
+In Experiment 3 we ensured that all instructions used verbal quantifiers, while also factorially manipulating vagueness and selection task.
+This allowed us to distinguish between the predictions of the selection task account and the instruction format account.
 

### 2018-08-03 experiments/C_exp_1/writeup/C_exp_1.Rmd
added borderline analysis

@@
 library(tidyverse)
-#library(tidymodels)
+library(tidymodels)
 library(broom.mixed)
@@
 source('concatenate_raw_data.R')
+source('add_borderline_vars.R')
 source('annotate_the_raw_data.R')
@@
 
-```{r, echo=FALSE}
-dat <- concatenate_raw_data()
-dat <- annotate_the_raw_data(dat)
+```{r, "get data for RT analysis", echo=F, message=F, warning=F, results='hide'}
+dat <- 
+  concatenate_raw_data() %>%
+  annotate_the_raw_data() %>% # at this point RT has min=1 (resp_type="sticky") and max=59,999 (resp_type="timeout")
+  filter(RT>1 & RT<59999) # at this point RT has min=445, max=42,685 (42 seconds) and nrow(dat) is 7677 down from 7680
+dat <- perSubjectTrim.fnc(dat, response='RT', subject='Subject', trim = 2.5)$data %>% select(-SD, -Mean, -Scaled)
+dat$RT_log <- log(dat$RT)
+```
+
+```{r, "get data for borderline analysis", echo=F}
+dat_borderline <- 
+  concatenate_raw_data() %>%
+  add_borderline_vars() %>%
+  annotate_the_raw_data() %>% # at this point RT has min=1 (resp_type="sticky") and max=59,999 (resp_type="timeout")
+  filter(RT>1 & RT<59999) # at this point RT has min=445, max=42,685 (42 seconds) and nrow(dat) is 7677 down from 7680
 ```
@@
 
-```{r, echo=F, message=F, warning=F, results='hide'}
-# at this point RT has min=1 (resp_type="sticky") and max=59,999 (resp_type="timeout")
-dat <- subset(dat, RT>1 & RT<59999)
-# at this point RT has min=445, max=42,685 (42 seconds) and nrow(dat) is 7677 down from 7680
-# Now trim the data
-dat <- perSubjectTrim.fnc(data = dat, response='RT', subject='Subject', trim = 2.5)$data %>% select(-SD, -Mean, -Scaled)
-# Now add log(RT) for the remaining RTs
-dat$RT_log <- log(dat$RT)
-```
-
 ```{r, echo=F}
@@
 
-```{r, echo=F}
+```{r, "RT condition means plot", fig.width=7, fig.height=3.5, echo=FALSE, results='hide'}
 # correction from Morey (2008) applied to condition means
@@
 ggplot(dat_plot, aes(x=Item, y=RT_logNormed, group=Vagueness, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci, shape=Vagueness, fill=Vagueness)) +
+  ggtitle("Response time condition means") +
   facet_wrap(~Number) +
@@
   scale_shape_manual(name="Vagueness", values=c(21,22)) +
-  theme(aspect.ratio = 1, panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background=element_rect(fill="white", colour="black")) +
-  ylab("RT log(ms)") +
+  theme(aspect.ratio = 1, panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_rect(fill="white", colour="black"), strip.background=element_blank(), legend.key = element_blank(), legend.key.size=unit(4, 'mm'), legend.position="top", axis.text.x = element_text(angle = 15)) +
+  ylab("RT log(ms)") + xlab(NULL) +
   #
@@
   geom_point(position=pdodge, size=2)
+
+```
+
+```{r, "Borderline response distribution plot", fig.width=7, fig.height=3.5, echo=FALSE, results='hide'}
+dat_borderline$response_category <- relevel(dat_borderline$response_category, ref='expected')
+
+ggplot(dat_borderline) + 
+  geom_bar(aes(response_category, group=Number:Vagueness, fill=Vagueness), position=position_dodge(width=NULL)) +   
+  scale_fill_grey() + 
+  xlab(NULL) + 
+  ggtitle('Borderline response distribution') + 
+  facet_grid(~Number) +
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), legend.position="top", aspect.ratio=1, plot.background = element_rect(fill=NA, color='white'), strip.background=element_blank(), legend.key.size=unit(4, 'mm'), axis.text.x = element_text(angle = 15))
+
 ```
@@
 
-Original full model: RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag * c_Num + c_Itm | Subject)
+Original full model
 ```{r "original full RT model", echo=F, cache=TRUE}
@@
 
+Borderline model
+```{r 'bl-full-model', cache=TRUE, echo=FALSE}
+dat_model <- dat_borderline
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
+blFullModel <- lme4::glmer(isBorderline ~ c_Vag * c_Num + c_Itm + (1 + c_Vag * c_Num + c_Itm | Subject), dat_model, family="binomial", control = glmerControl(optimizer = "bobyqa"))
+pretty_coef_table(blFullModel, "blFullModel")
+```
+
 On the basis of the initial full model:
@@
 
-Model of the data after the 6:15:24 level of Item is removed: RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag + c_Num  | Subject)
+Model of the data after the 6:15:24 level of Item is removed
 ```{r, echo=F, cache=TRUE}
@@
     * There is still a disadvantage for increasing the number of dots but it is not significant (p=.323).
-* _Interaction comment goes here_ The interaction between Vagueness and Instruction format changes sign, and is now non-significant (beta=-.022, p=.374).
+* The interaction between Vagueness and Instruction format changes sign, and is now non-significant (beta=-.022, p=.374).
 
@@
 However, once again there is a possibly confounding factor.
-Observe that, in Experiment 1, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) "selection algorithm": To see this, consider the task of selecting the dot array that contains "few dots":" to do this, it suffices to _compare_ the three arrays and select the one that contains the fewest elements.
+Observe that, in Experiment 1, instruction format (i.e., the difference between numeric and verbal) went hand in hand with might be called the (human) "selection algorithm": To see this, consider the task of selecting the dot array that contains "few dots":" to do this, it suffices to _compare_ the three arrays and select the one that contains the fewest elements.
 To select the dot array that contains "16 dots" seems to require the participant to estimate, and then _match_, the cardinality of (at least) one dot array to 16, a process which could plausibly take longer, independently of vagueness.
@@
 This allowed us to distinguish between the predictions of the selection task account and the instruction format account.
-
-<!-- So,     -->
-
-<!-- However the full model and the reduced model are not really equivalent because the terms are different in order for the reduced model to converge.
-What we really want is a full model structure that converges with and without the 6:15:24 item.
--->
-
-<!-- If we use the structure of the reduced model for the structure of the full model we get this for the full model: -->
-
-<!-- ```{r "full RT model", echo=F} -->
-<!-- dat_model <- dat -->
-<!-- dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5) -->
-<!-- dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5) -->
-<!-- dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75))) -->
-<!-- rtFullModel2 <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag + c_Num  | Subject), dat_model) -->
-<!-- pretty_coef_table(rtFullModel2, "rtFullModel2") -->
-<!-- ``` -->
-
-<!-- and this for the reduced model: -->
-
-<!-- ```{r} -->
-<!-- pretty_coef_table(rtFullModelMinus, "rtFullModelMinus 6:15:24") -->
-<!-- ``` -->
-
-<!-- fullmodel with Helmert contrasts on Item -->
-<!-- ```{r "full RT model Helmert", echo=F} -->
-<!-- dat_model <- dat -->
-<!-- dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5) -->
-<!-- dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5) -->
-<!-- contrasts(dat_model$Item) <- contr.helmert(4) -->
-<!-- # had to drop c_Vag * Item slopes for subject -->
-<!-- rtFullModel_helmert <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + Item + (1 + c_Vag + c_Num + Item | Subject), dat_model) -->
-<!-- pretty_coef_table(rtFullModel_helmert, "c_Vag * c_Num + Item + (1 + c_Vag + c_Num + Item | Subject)") -->
-<!-- ``` -->
-
-<!-- reducedmodel with Helmert contrasts on Item -->
-<!-- ```{r "reduced RT model Helmert", echo=F} -->
-<!-- dat_model <- droplevels(subset(dat, Item!="06:15:24")) -->
-<!-- dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5) -->
-<!-- dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5) -->
-<!-- contrasts(dat_model$Item) <- contr.helmert(3) -->
-<!-- # had to drop c_Vag * Item slopes for subject -->
-<!-- rtReducedModel_helmert <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + Item + (1 + c_Vag + c_Num + Item | Subject), dat_model) -->
-<!-- pretty_coef_table(rtReducedModel_helmert, "c_Vag * c_Num + Item + (1 + c_Vag + c_Num + Item | Subject)") -->
-<!-- ``` -->
-
-<!-- However the plot of means suggests that the main variables are c_Num and Item so it might be better to have a model structure that includes a c_Num*Item term. -->
-
-<!-- ```{r} -->
-<!-- dat_model <- dat -->
-<!-- dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5) -->
-<!-- dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5) -->
-<!-- contrasts(dat_model$Item) <- contr.helmert(4) -->
-<!-- myMod1 <- lmerTest::lmer(1 + RT_log ~ c_Vag + c_Num + Item + c_Num:Item +(1|Subject), dat_model) -->
-<!-- pretty_coef_table(myMod1, "myMod1") -->
-<!-- ``` -->
-
-<!-- ```{r} -->
-<!-- dat_model <- droplevels(subset(dat, Item!="06:15:24")) -->
-<!-- dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5) -->
-<!-- dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5) -->
-<!-- contrasts(dat_model$Item) <- contr.helmert(3) -->
-<!-- myMod2 <- lmerTest::lmer(1 + RT_log ~ c_Vag + c_Num + Item + c_Num:Item +(1|Subject), dat_model) -->
-<!-- pretty_coef_table(myMod2, "myMod2:no6:15:24") -->
-<!-- ``` -->
-
-<!-- ################################################################################################## -->
-
-<!-- Show which RTs were removed by trimming -->
-<!-- ```{r} -->
-<!-- ggplot(data=dat, aes(x=Trial, y=RT)) + -->
-<!--   geom_point(aes(colour=was_RT_removed_by_trimming), size=.5) +  -->
-<!--   facet_wrap(~Subject)+ -->
-<!--   theme(aspect.ratio = 1, panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background=element_rect(fill="white", colour="black")) -->
-<!-- ``` -->
-
-<!-- Show which Rts were removed by trimming, shown as log RT -->
-<!-- ```{r} -->
-<!-- ggplot(data=dat, aes(x=Trial, y=log(RT))) + -->
-<!--   geom_point(aes(colour=was_RT_removed_by_trimming), size=.5) +  -->
-<!--   facet_wrap(~Subject)+ -->
-<!--   theme(aspect.ratio = 1, panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background=element_rect(fill="white", colour="black")) -->
-<!-- ``` -->
-
-<!-- Show how distributions of Rt change according to trimming, in log RT -->
-<!-- ```{r} -->
-<!-- dat %>% gather(key=whether_trimmed, value=time, RT, RT_trimmed) %>% -->
-<!--   ggplot(aes(log(time))) + -->
-<!--   geom_density(aes(group=whether_trimmed, colour=whether_trimmed)) +  -->
-<!--   facet_wrap(~Subject, scales="free") + -->
-<!--   theme(aspect.ratio = 1, panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background=element_rect(fill="white", colour="black"))  -->
-<!-- ``` -->
-
-<!-- Make RT log and reciprocal, from the trimmed RT in ms. -->
-<!-- ```{r} -->
-<!-- dat$RT_trimmed_log <- log(dat$RT_trimmed) -->
-<!-- dat$RT_trimmed_rcp <- -1/(dat$RT_trimmed) -->
-<!-- ``` -->
-
-<!-- plot transformations of rt in the trimmed data -->
-
-<!-- ```{r, warning=FALSE, echo=F} -->
-<!-- p.dens <- -->
-<!-- dat %>% -->
-<!--   gather(key=metric, value=value, RT_trimmed, RT_trimmed_log, RT_trimmed_rcp) %>% -->
-<!--     ggplot(aes(value)) + -->
-<!--     geom_density() + -->
-<!--     facet_wrap(~metric, scales="free") + -->
-<!--     theme(aspect.ratio = 1, panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background=element_rect(fill="white", colour="black")) -->
-<!-- ``` -->
-
-<!-- ```{r, warning=FALSE, echo=F} -->
-<!-- p.point <- -->
-<!-- dat %>% -->
-<!--   gather(key=metric, value=value, RT_trimmed, RT_trimmed_log, RT_trimmed_rcp) %>% -->
-<!--     ggplot(aes(x=Obs, y=value)) + -->
-<!--     geom_point(size=.1) + -->
-<!--     facet_wrap(~metric, scales="free") + -->
-<!--     theme(aspect.ratio = 1, panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background=element_rect(fill="white", colour="black")) -->
-<!-- ``` -->
-
-<!-- ```{r, warning=FALSE, echo=F} -->
-<!-- grid.arrange(p.dens,p.point) -->
-<!-- ``` -->
-
-<!-- Compute the ci's from untransformed data that were trimmed using summarysewithin2 and plot them on a log scale to reproduce the plot in the Springer submission. -->
-
-<!-- ```{re2-rtplot, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE} -->
-<!-- rtdataplot1 <- summarySEwithin2(dat, measurevar="RT_trimmed_log", withinvars=c("Vagueness", "Number", "Item"), idvar="Subject") -->
-
-<!-- rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness)) -->
-
-<!-- ggplot(rtdataplot1, aes(x=Item, group=Condition, fill=Vagueness)) +  -->
-<!--   geom_line(aes(x=Item, y=log(RT_trimmedNormed))) + -->
-<!--   facet_grid(~Number) -->
-
-<!-- +  -->
-<!--   geom_errorbar(width=.2, aes(ymin=RT_trimmedNormed-ci, ymax=RT_trimmedNormed+ci)) +  -->
-<!--   geom_point(size=2, col=1, aes(shape=Vagueness)) +  -->
-<!--   scale_fill_grey(name="Vagueness", start=0, end=1) +   -->
-<!--   scale_shape_manual(name="Vagueness", values=c(21,22)) +  -->
-<!--   ggtitle("Response time") +  -->
-<!--   ylab("Respone time (log (ms))") +  -->
-<!--   xlab(NULL) +  -->
-<!--   theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), legend.position="right", aspect.ratio=1, plot.background=element_rect(fill=NA, color='white'), strip.background=element_blank(), axis.text.x = element_text(angle = 15))  -->
-
-<!-- ``` -->
-
-<!-- ```{r e2-rtplot, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE} -->
-<!-- rtdataplot1 <- summarySEwithin2(dat, measurevar="RT_trimmed_log", withinvars=c("Vagueness", "Number", "Item"), idvar="Subject") -->
-
-<!-- rtdataplot1$Condition <- as.factor(paste(sep=':', rtdataplot1$Number, rtdataplot1$Vagueness)) -->
-
-<!-- rtplot <-  -->
-<!--   ggplot(rtdataplot1, aes(y=RT_trimmed_logNormed, x=Item, group=Condition, fill=Vagueness)) +  -->
-<!--   geom_line() +  -->
-<!--   geom_errorbar(width=.2, aes(ymin=RT_trimmed_logNormed-ci, ymax=RT_trimmed_logNormed+ci)) +  -->
-<!--   geom_point(size=2, col=1, aes(shape=Vagueness)) +  -->
-<!--   scale_fill_grey(name="Vagueness", start=0, end=1) +   -->
-<!--   scale_shape_manual(name="Vagueness", values=c(21,22)) +  -->
-<!--   ggtitle("Response time") +  -->
-<!--   ylab("Respone time (log (ms))") +  -->
-<!--   xlab(NULL) +  -->
-<!--   theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), legend.position="right", aspect.ratio=1, plot.background=element_rect(fill=NA, color='white'), strip.background=element_blank(), axis.text.x = element_text(angle = 15)) + facet_grid(~Number) -->
-
-<!-- rtplot -->
-<!-- ``` -->
-
-<!-- ## Response time models  -->
-
-<!-- ### Full RT model -->
-
-<!-- ```{r 'rt-full-model', cache=TRUE, echo=TRUE} -->
-<!-- rtFullModel <- lmerTest::lmer(RT_log ~ c_Vag * c_Num + c_Itm + (1 + c_Vag * c_Num + c_Itm | Subject), rtdata) -->
-<!-- ``` -->
-
-<!-- ```{r 'make a pretty coefficients table for rtFullModel', echo=FALSE, results='asis'} -->
-<!-- pretty_coef_table(rtFullModel, "rtFullModel") %>% print() -->
-<!-- ``` -->
-
-<!-- ```{r, 'make a pretty coefficients plot for rtFullModel', echo=FALSE, fig.width=6, fig.height=3} -->
-<!-- pretty_coef_plot(rtFullModel, "rtFullModel") -->
-<!-- ``` -->
-
-<!-- ### Numeric only RT model -->
-
-<!-- ```{r 'rt-numeric-only model', cache=TRUE, echo=TRUE} -->
-<!-- rtNumOnlyModel <- lmerTest::lmer(RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject), subset(rtdata,Number=="Numeric")) -->
-<!-- ``` -->
-
-<!-- ```{r 'make a pretty coefficients table for rtNumOnlyModel', echo=FALSE, results='asis'} -->
-<!-- pretty_coef_table(rtNumOnlyModel, "rtNumOnlyModel") %>% print() -->
-<!-- ``` -->
-
-<!-- ```{r, 'make a pretty coefficients plot for rtNumOnlyModel', echo=FALSE, fig.width=6, fig.height=3} -->
-<!-- pretty_coef_plot(rtNumOnlyModel, "rtNumOnlyModel") -->
-<!-- ``` -->
-
-<!-- ### Verbal only RT model -->
-
-<!-- ```{r 'rt-verbal-only model', cache=TRUE, echo=TRUE} -->
-<!-- rtVerbalOnlyModel <- lmerTest::lmer(RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject), subset(rtdata,Number=="Verbal")) -->
-<!-- ``` -->
-
-<!-- ```{r 'make a pretty coefficients table for rtVerbalOnlyModel', echo=FALSE, results='asis'} -->
-<!-- pretty_coef_table(rtVerbalOnlyModel, "rtVerbalOnlyModel") %>% print() -->
-<!-- ``` -->
-
-<!-- ```{r, 'make a pretty coefficients plot for rtVerbalOnlyModel', echo=FALSE, fig.width=6, fig.height=3} -->
-<!-- pretty_coef_plot(rtVerbalOnlyModel, "rtVerbalOnlyModel") -->
-<!-- ``` -->
-
-<!-- # Borderline response data -->
-
-<!-- ```{r} -->
-<!-- bldata <- trim_data -->
-<!-- ``` -->
-
-<!-- ```{r e2-blBarChart, fig.width=7, fig.height=3.5, echo=FALSE, message=FALSE} -->
-<!-- bldata$response_category <- relevel(bldata$response_category, ref='expected') -->
-
-<!-- blBarChart = ggplot(bldata) + geom_bar(aes(response_category, group=Number:Vagueness, fill=Vagueness), position=position_dodge(width=NULL)) +   scale_fill_grey() + xlab(NULL) + ggtitle('Borderline response distribution') + facet_grid(~Number) + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), legend.position="right", aspect.ratio=1, plot.background = element_rect(fill=NA, color='white'), strip.background=element_blank(), legend.key.size=unit(4, 'mm'), axis.text.x = element_text(angle = 15)) -->
-
-<!-- blBarChart -->
-<!-- ``` -->
-
-<!-- ## Borderline response models -->
-
-<!-- ### Full borderline response model -->
-
-<!-- ```{r 'bl-full-model', cache=TRUE, echo=TRUE} -->
-<!-- blFullModel <- lme4::glmer(isBorderline ~ c_Vag * c_Num + c_Itm + (1 + c_Vag * c_Num + c_Itm | Subject), bldata, family="binomial", control = glmerControl(optimizer = "bobyqa")) -->
-<!-- ``` -->
-
-<!-- ```{r 'make a pretty coefficients table for blFullModel', echo=FALSE, results='asis'} -->
-<!-- pretty_coef_table(blFullModel, "blFullModel") %>% print() -->
-<!-- ``` -->
-
-<!-- ```{r, 'make a pretty coefficients plot for blFullModel', echo=FALSE, fig.width=6, fig.height=3} -->
-<!-- pretty_coef_plot(blFullModel, "blFullModel") -->
-<!-- ``` -->
-
-<!-- Original fullmodel with Helmert contrasts on Item -->
-<!-- ```{r "original full RT model Helmert", echo=4:5} -->
-<!-- dat_model <- dat -->
-<!-- dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5) -->
-<!-- dat_model$c_Num <- ifelse(dat_model$Number=="Verbal", -0.5, 0.5) -->
-<!-- contrasts(dat_model$Item) <- contr.helmert(4) -->
-<!-- # had to drop c_Vag * Item slopes for subject -->
-<!-- rtFullModel_helmert <- lmerTest::lmer(1 + RT_log ~ c_Vag * c_Num + Item + (1 + c_Vag + c_Num + Item | Subject), dat_model) -->
-<!-- pretty_coef_table(rtFullModel_helmert, "rtFullModel Helmert") -->
-<!-- ``` -->
-
-<!-- # Re-run without 6:15:24 -->
-<!-- ```{r, echo=F, message=F, warning=F, results='hide'} -->
-<!-- dat_reduced <- droplevels(subset(dat, Item!="06:15:24")) -->
-<!-- dat_plot <- summarySEwithin2(dat_reduced, measurevar="RT_log", withinvars=c("Vagueness", "Number", "Item"), idvar="Subject") -->
-<!-- ``` -->
-
-<!-- ```{r, echo=F, results='hide'} -->
-<!-- # correction from Morey (2008) applied to condition means -->
-<!-- mywidth=0 -->
-<!-- pdodge=position_dodge(width=mywidth) -->
-<!-- # -->
-<!-- ggplot(dat_plot, aes(x=Item, y=RT_logNormed, group=Vagueness, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci, shape=Vagueness, fill=Vagueness)) + -->
-<!--   facet_wrap(~Number) + -->
-<!--   scale_fill_grey(name="Vagueness", start=0, end=1) +   -->
-<!--   scale_shape_manual(name="Vagueness", values=c(21,22)) + -->
-<!--   theme(aspect.ratio = 1, panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background=element_rect(fill="white", colour="black")) + -->
-<!--   ylab("RT log(ms)") + -->
-<!--   # -->
-<!--   geom_errorbar(position=pdodge, width=0.25) + -->
-<!--   geom_line(position=pdodge) + -->
-<!--   geom_point(position=pdodge, size=2)  -->
-<!-- ``` -->

### 2018-08-03 experiments/C_exp_1/writeup/annotate_the_raw_data.R

@@ -33,7 +33,7 @@ annotate_the_raw_data = function (dat) {
   dat$Obs <- 1:7680
   
   # remove variables that don't matter for purposes of RT analysis
-  dat <- select(dat, Obs, Trial, Instruction, Subject, Item, Quantity, Order, Vagueness, Number, RT)
+  #dat <- select(dat, Obs, Trial, Instruction, Subject, Item, Quantity, Order, Vagueness, Number, response_category, isBorderline, RT)
   
   return(dat)
 }  

### 2018-08-11 experiments/D_exp_2/writeup/D_exp_2.Rmd
D_exp_2
only edits that made the code run.

--- 2017-05-16 experiments/e3/writeup/e3.Rmd
@@
 ```{r getData, echo=T}
-rawdata <- read.table('data.txt', sep=' ')
+# data_processed.txt has RT subset=RT>0 & RT<49871
+rawdata <- read.table('preprocessing/data_processed.txt', sep=' ')
 ```
@@
 ```{r rtlmer, cache=TRUE}
-rtlmer = lmerTest::lmer(data=rtdata,
+dat_model <- rtdata
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Sel <- ifelse(dat_model$Selection=="Comparison", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
+rtlmer = lmerTest::lmer(data=dat_model,
                         RT_log ~ c_Vag * c_Sel + c_Itm +
@@
 ```{r seplmersforcomparison, results='asis', cache=TRUE, echo=-3}
-s1 <- subset(rtdata, subset=Selection=='Comparison')
-s1lm <- lmerTest::lmer(data=s1, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
+dat_model <- subset(rtdata, subset=Selection=='Comparison')
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
+s1lm <- lmerTest::lmer(data=dat_model, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
 print(xtable(summary(s1lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
@@
 ```{r seplmersformatching, results='asis', cache=TRUE, echo=-3}
-s2 <- subset(rtdata, subset=Selection=='Matching')
-s2lm <- lmerTest::lmer(data=s2, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
+dat_model <- subset(rtdata, subset=Selection=='Matching')
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
+s2lm <- lmerTest::lmer(data=dat_model, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
 print(xtable(summary(s2lm)$coef, digits=c(0,2,2,1,1,4)), type='html')

### 2018-08-11 experiments/D_exp_2/writeup/preprocessing/preprocessing.Rmd
D_exp_2
only edits that made the code run.

--- 2017-05-17 experiments/e3/writeup/preprocessing/preprocessing.Rmd
@@
 ```{r readdata}
+
+# The policy as of Aug 2018 is that:
+# RT=0 is stripped, and one case with a very long time RT=49871 is stripped.
 
@@
 
+# Create a factor coding for Selection
 dat$Selection = ""
@@
 dat$Condition <- NULL
-
 dat$Selection <- as.factor(dat$Selection)
@@
 # add preceding RT
-my_rt <- dat$RT[-1]
-
 dat$RTprev <- c(NA,dat$RT[1:nrow(dat)-1])
@@
 
-dat$RT <- NULL
 dat$RTprev <- NULL
@@
 
-# This data set non-impossible trials and renamed/recomputed variables
-# dat.out <- subset(dat, select=c(Subject, Trial, Item, c_Itm, Vagueness, c_Vag, Selection, c_Sel, Instruction, Order, Quantity, RT, RT_log, response_catResponse))
-# write.table(dat.out, file="data_processed.txt")
-
-dat <- subset(dat, select=c(Subject,Item,id,Trial,Vagueness,Selection, 
-                             Order, Quantity, Instruction, nchar_instr, RT_raw, RT_log, RTprev_log, RTprev_raw, rt_instr, response_cat ))
+dat <- subset(dat, 
+              subset=RT>0 & RT<49871,
+              select=c(Subject,Item,id,Trial,Vagueness,Selection, Order, Quantity, Instruction, nchar_instr, RT, RT_raw, RT_log, RTprev_log, RTprev_raw, rt_instr, response_cat ))
 

### 2018-08-11 experiments/E_exp_3/writeup/E_exp_3.Rmd
D_exp_3
Only added an Rproj and a jpg and html

--- 2016-09-27 experiments/e4/writeup/e4.Rmd
@@
 <figure>
-<img src="./flow-of-experiment4.pdf" alt="Mountain View" style="width:480px;height:300px;">
-<figcaption>Flow of experiment 4.</figcaption>
+<img src="./E_exp_3-flow.jpg" alt="flow-of-D_exp_3" style="width:480px;height:300px;">
+<figcaption>Flow of experiment E_exp_3.</figcaption>
 </figure>

### 2018-08-11 experiments/D_exp_2/writeup/D_exp_2.Rmd
pretty coef D

--- 2017-05-16 experiments/e3/writeup/e3.Rmd
@@
 library(knitr); library(MASS); library(ggplot2); library(lme4); library(xtable); library(plyr); library(lattice); library(LMERConvenienceFunctions); library(car); library(grid); library(lmerTest); library(reshape2); library(ggplot2); library(languageR); library(pander); source('summarySEwithin2.R'); library(Hmisc)
+library(broom)
+library(broom.mixed)
+source("add_stars.R")
+library(kableExtra)
+source("pretty_coef_table.R")
 ```
@@
 
-```{r rtlmer, cache=TRUE}
-rtlmer = lmerTest::lmer(data=rtdata,
+```{r rtlmer, cache=TRUE, results='asis'}
+dat_model <- rtdata
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Sel <- ifelse(dat_model$Selection=="Comparison", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
+rtlmer = lmerTest::lmer(data=dat_model,
                         RT_log ~ c_Vag * c_Sel + c_Itm +
                           (1 + c_Vag * c_Sel + c_Itm | Subject))
-```
-
-```{r rtlmersummaryxtable, results='asis', cache=TRUE, echo=F}
-print(xtable(summary(rtlmer)$coef, digits=c(0,2,2,1,1,4)),type='html')
+pretty_coef_table(rtlmer, "rtlmer")
 ```
@@
 ```{r seplmersforcomparison, results='asis', cache=TRUE, echo=-3}
-s1 <- subset(rtdata, subset=Selection=='Comparison')
-s1lm <- lmerTest::lmer(data=s1, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
-print(xtable(summary(s1lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
+dat_model <- subset(rtdata, subset=Selection=='Comparison')
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
+s1lm <- lmerTest::lmer(data=dat_model, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
+pretty_coef_table(s1lm, "s1lm")
 ```
@@
 ```{r seplmersformatching, results='asis', cache=TRUE, echo=-3}
-s2 <- subset(rtdata, subset=Selection=='Matching')
-s2lm <- lmerTest::lmer(data=s2, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
-print(xtable(summary(s2lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
+dat_model <- subset(rtdata, subset=Selection=='Matching')
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
+s2lm <- lmerTest::lmer(data=dat_model, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
+pretty_coef_table(s2lm, "s2lm")
 ```

### 2018-08-11 experiments/E_exp_3/writeup/E_exp_3.Rmd
E minor edits

@@
 library(knitr); library(MASS); library(ggplot2); library(lme4); library(xtable); library(plyr); library(lattice); library(LMERConvenienceFunctions); library(car); library(grid); library(lmerTest); library(reshape2); library(ggplot2); library(languageR); library(pander); source('summarySEwithin2.R')
+library(kableExtra)
+library(broom)
+library(broom.mixed)
+source("pretty_coef_table.R")
 ```
@@
 
-```{r rtlmer, cache=TRUE}
-rtlmer = lmerTest::lmer(data=rtdata,
-                        RT_log ~ c_Vag * c_Sel + c_Itm +
-                          (1 + c_Vag * c_Sel + c_Itm | Subject))
-```
-
-```{r rtlmersummaryxtable, results='asis', cache=TRUE, echo=F}
-print(xtable(summary(rtlmer)$coef, digits=c(0,2,2,1,1,4)),type='html')
+```{r rtlmer, results='asis', cache=TRUE, echo=-2}
+rtlmer = lmerTest::lmer(data=rtdata, RT_log ~ c_Vag * c_Sel + c_Itm + (1 + c_Vag * c_Sel + c_Itm | Subject))
+pretty_coef_table(rtlmer, "rtlmer")
 ```
@@
 s1lm <- lmerTest::lmer(data=s1, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
-print(xtable(summary(s1lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
+pretty_coef_table(s1lm, "s1lm")
 ```
@@
 s2lm <- lmerTest::lmer(data=s2, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
-print(xtable(summary(s2lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
+pretty_coef_table(s2lm, "s2lm")
 ```

### 2018-08-11 experiments/C_exp_1/writeup/C_exp_1.Rmd
restructure or standardise C

@@
 library(tidyverse)
-library(tidymodels)
+library(broom)
 library(broom.mixed)
@@
 library(gridExtra)
-source('summarySEwithin2.R')
-source('concatenate_raw_data.R')
-source('add_borderline_vars.R')
-source('annotate_the_raw_data.R')
-source('declare_impossible_rt.R')
-source('remove_impossible_trials.R')
-source('pretty_coef_table.R')
-source('pretty_coef_plot.R')
+source("C_exp_1_functions/concatenate_raw_data.R")
+source("C_exp_1_functions/annotate_the_raw_data.R")
+source("C_exp_1_functions/add_borderline_vars.R")
+source("C_exp_1_functions/pretty_coef_table.R")
+source("C_exp_1_functions/summarySEwithin2.R")
+root_dir <- getwd()
+data_dir <- "C_exp_1_data"
 ```
@@
 dat <- 
-  concatenate_raw_data() %>%
+  concatenate_raw_data(root_dir, data_dir) %>%
   annotate_the_raw_data() %>% # at this point RT has min=1 (resp_type="sticky") and max=59,999 (resp_type="timeout")
@@
 dat_borderline <- 
-  concatenate_raw_data() %>%
+  concatenate_raw_data(root_dir, data_dir) %>%
   add_borderline_vars() %>%
@@
     spread(key=Vagueness, value=Instruction)
-```
-
-```{r, echo=FALSE}
+# kable(instructions_table, format="latex", booktabs = TRUE)
 instructions_table %>% 
@@
       kable_styling(full_width = F, position = "left", font_size = 11) 
-```
-
-```{r, echo=FALSE, eval=FALSE}
-kable(instructions_table, format="latex", booktabs = TRUE)
 ```

### 2018-08-11 experiments/C_exp_1/writeup/add_stars.R

@@ -1,75 +1,0 @@
-add_stars = function(modl) {
-  
-  coef_table <- as.data.frame(tidy(modl, effects="fixed"))
-  
-  # be sensitive to whether the model is an lmer or a glmer 
-  
-  if (class(modl) == "lmerModLmerTest") {
-    
-    # sort the names out
-    names(coef_table) <- c("effect", "term", "coefficient", "std.error", "df", "t_value", "p_value_raw")
-    coef_table$p_value <- as.character("")
-    coef_table$star <- as.character("")
-    
-    # categorise p values and add significance stars
-    for (i in 1:nrow(coef_table)) {
-      if (!is.na(coef_table$p_value_raw[i])) {
-        if (coef_table$p_value_raw[i] >= 0     & coef_table$p_value[i] < 0.001) {coef_table$p_value[i] <-    "<0.001"; coef_table$star[i] <- "***"}
-        if (coef_table$p_value_raw[i] >= 0.001 & coef_table$p_value[i] < 0.01 ) {coef_table$p_value[i] <-    "<0.01"; coef_table$star[i] <- "**"}
-        if (coef_table$p_value_raw[i] >= 0.01  & coef_table$p_value[i] < 0.05 ) {coef_table$p_value[i] <-    "<0.05"; coef_table$star[i] <- "*"}
-        if (coef_table$p_value_raw[i] >= 0.05  ) {coef_table$p_value[i] <-    round(coef_table$p_value_raw[i],3); coef_table$star[i] <-  ""}
-      }
-    }
-
-    # convert numerics to formatted chars for pretty printing
-    coef_table$coefficient <- format(x=coef_table$coefficient, nsmall=2, digits=2)
-    coef_table$std.error   <- format(x=coef_table$std.error, nsmall=2, digits=2)
-    coef_table$df          <- format(x=coef_table$df, nsmall=2, digits=1)
-    coef_table$t_value     <- format(x=coef_table$t_value, nsmall=2, digits=2)
-    coef_table$p_value_raw <- format(x=coef_table$p_value_raw, nsmall=2, digits=3)
-    
-    # select rows and cols
-    coef_table <- coef_table %>%
-      select(-effect) %>%
-      filter(term!="(Intercept)") 
-    
-    return(coef_table)
-  
-  }
-  
-  if (class(modl) == "glmerMod") {
-    
-    # sort the names out
[…]

### 2018-08-11 experiments/C_exp_1/writeup/concatenate_raw_data.R

@@ -1,21 +1,0 @@
-concatenate_raw_data = function() {
-  
-  number_of_valid_subjects=30
-  number_of_trials_per_subject=256
-  data_dir <- 'per_subject_raw_data_files/'
-  
-  column.headers.df <-  head( read.table(
-    paste(data_dir,'subject01.data',sep=''),
-    header=TRUE),0)
-
-  gathered.data <- column.headers.df
-  
-  for (subject in 1:number_of_valid_subjects) {
-    current.filename <- paste(data_dir, 'subject', sprintf("%02d",subject), '.data', sep='')
-    current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = TRUE)
-    gathered.data <- rbind(gathered.data, current.data)
-  }
-  
-  return(gathered.data)
-  
-}

### 2018-08-11 experiments/C_exp_1/writeup/declare_impossible_rt.R

@@ -1,29 +1,0 @@
-declare_impossible_rt = function(dat) {
-  message('Starting to declare impossible RTs')
-  dat$NaType = 'valid_trial'
-  # declare impossible trials as NA
-  # impossible trials are sticky fingers and timeouts
-  # sticky fingers
-  dat$NaType[dat$RT_original <= 1] <- 'sticky_fingers'
-  dat$RT[dat$RT_original <= 1] <- NA
-  dat$isBorderline[dat$RT_original <= 1] <- NA
-  dat$response_category[dat$RT_original <= 1] <- NA
-  # timeout
-  dat$NaType[dat$RT_original >= 59998] <- 'near_timeout'
-  dat$RT[dat$RT_original >= 59998 ] <- NA
-  dat$isBorderline[dat$RT_original >= 59998] <- NA
-  dat$response_category[dat$RT_original >= 59998] <- NA
-  
-  #dat$NaType[is.na(dat$isBorderline_original)] <- 'timeout'
- # dat$isBorderline[dat$isBorderline_original=='timeout'] <- NA
- # dat$response_category[dat$isBorderline_original=='timeout'] <- NA
-  
-  dat$response_category <- as.factor(dat$response_category)
-  dat$NaType <- as.factor(dat$NaType)
-  #dat$RT_original <- NULL
-  #dat$isBorderline_original <- NULL
-  dat$RT_log <- log(dat$RT)
-  dat$RT_rcp <- -1/dat$RT
-  message('Returning data with impossible RTs declared as NA in RT and isBorderline')
-  return(dat)
-}

### 2018-08-11 experiments/C_exp_1/writeup/preprocessing.R

@@ -1,227 +1,0 @@
-gatherData = function() {
-  number_of_valid_subjects=30
-  number_of_trials_per_subject=256
-  message('Starting to gather data')
-  data_dir <- '../../experimentCode/output/'
-  column.headers.df <-  head( read.table(
-    paste(data_dir,'subject01.data',sep=''),
-    header=TRUE),0)
-  gathered.data <- column.headers.df
-  for (subject in 1:number_of_valid_subjects) {
-    current.filename <- paste(data_dir, 'subject', sprintf("%02d",subject), '.data', sep='')
-    current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = TRUE)
-    gathered.data <- rbind(gathered.data, current.data)
-  }
-  dat <- gathered.data
-  dat$Order <- factor(dat$Order, levels=c(1,2), labels=c('LtoR','RtoL'))
-  dat$c_Ord <- ifelse(dat$Order=="LtoR",-.5,.5)
-  dat$Quantity <- factor(dat$Quantity, levels=c(1,2), labels=c('Small','Large'))
-  dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)
-  dat[ dat$Condition==1 , 'Vagueness'] <- 'Vague'
-  dat[ dat$Condition==2 , 'Vagueness'] <- 'Crisp'
-  dat[ dat$Condition==3 , 'Vagueness'] <- 'Vague'
-  dat[ dat$Condition==4 , 'Vagueness'] <- 'Crisp'
-  dat$Vagueness <- as.factor(dat$Vagueness)
-  dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)
-  dat[ dat$Condition==1 , 'Number'] <- 'Numeric'
-  dat[ dat$Condition==2 , 'Number'] <- 'Numeric'
-  dat[ dat$Condition==3 , 'Number'] <- 'Verbal'
-  dat[ dat$Condition==4 , 'Number'] <- 'Verbal'
-  dat$Number <- as.factor(dat$Number)
-  dat$c_Num <- ifelse(dat$Number=='Numeric', -.5, .5)
-  dat$Subject <- factor(paste("s",sprintf("%02d",dat$Subject),sep=""))
[…]

### 2018-08-11 experiments/C_exp_1/writeup/remove_impossible_trials.R

@@ -1,8 +1,0 @@
-remove_impossible_trials = function(dat){
-  message('Starting to remove impossible trials')
-  a=nrow(dat)
-  dat <- dat[ complete.cases(dat), ]
-  b=nrow(dat)
-  message(paste('removed', a-b, 'impossible trials'))
-  return(dat)
-}

### 2018-08-11 experiments/C_exp_1/writeup/scratch_borderline.R

@@ -1,173 +1,0 @@
-# In this experiment we did not count any response as an error, instead we coded them as expected, borderline, and extreme
-
-dat$c_Qty <- ifelse(dat$Quantity=="Small",-.5,.5)
-
-dat$c_Vag <- ifelse(dat$Vagueness=='Crisp', -.5, .5)
-
-dat$c_Num <- ifelse(dat$Number=='Numeric', -.5, .5)
-
-
-
-dat$c_Itm <- ifelse(dat$Item==1, -.75, ifelse(dat$Item==2, -.25, ifelse(dat$Item==3, .25, .75)))
-
-dat$c_Trl <-dat$Trial - mean(dat$Trial)
-dat$c_Itm <- ifelse(dat$Item==1, -.75, ifelse(dat$Item==2, -.25, ifelse(dat$Item==3, .25, .75)))
-dat$nchar_instr = nchar(as.character(dat$Instruction))
-discriminability = c(0.4875000, 0.3123529, 0.2308442, 0.1833333)
-dat[dat$Item_fac == "06:15:24", "discriminability"] <- discriminability[1]#0.4875000
-dat[dat$Item_fac == "16:25:34", "discriminability"] <- discriminability[2]#0.3123529
-dat[dat$Item_fac == "26:35:44", "discriminability"] <- discriminability[3]#0.2308442
-dat[dat$Item_fac == "36:45:54", "discriminability"] <- discriminability[4]#0.1833333
-# what were they expected to respond?
-message('figuring out borderline response')
-dat$crossed = as.factor(paste('Con', dat$Condition, ':Quan', as.numeric(dat$Quantity), ':Item', dat$Item, sep=''))
-dat[dat$crossed=='Con1:Quan1:Item1', 'Exp_Num']       <-  6
-dat[dat$crossed=='Con1:Quan1:Item1', 'Bline_Num']     <- 15
-dat[dat$crossed=='Con1:Quan1:Item1', 'Extr_Num']      <- 24
-dat[dat$crossed=='Con1:Quan1:Item2', 'Exp_Num'] <- 16
-dat[dat$crossed=='Con1:Quan1:Item2', 'Bline_Num']     <- 25
[…]

### 2018-08-11 experiments/D_exp_2/writeup/D_exp_2.Rmd
making D be like C

--- 2017-05-16 experiments/e3/writeup/e3.Rmd
@@
 ---
-title: "Experiment 3 of 4"
+title: "D_exp_2"
 author: "Matt Green"
@@
     toc: true
-    <!--toc_float: true-->
-    mathjax: default
-    template: default
-    code_folding: show
-    theme: default
+    toc_float: false
 ---
-<style>
-table, th, td {
-    border: 1px grey;
-    border-collapse: collapse;
-    padding:2px 5px 2px 5px
-}  
-</style>
 
-# Experiment 3
+```{r, echo=FALSE, message=FALSE}
+library(knitr)
+library(tidyverse)
+library(broom)
+library(broom.mixed)
+library(lme4)
+library(lmerTest)
+library(dotwhisker)
+library(kableExtra)
+library(LMERConvenienceFunctions)
+library(data.table)
+library(gridExtra)
+source("D_exp_2_functions/preprocessing.R")
+source("D_exp_2_functions/pretty_coef_table.R")
+source("D_exp_2_functions/summarySEwithin2.R")
+root_dir <- getwd()
+data_dir <- "D_exp_2_data"
+```
 
-```{r loadLibraries, message=FALSE}
-library(knitr); library(MASS); library(ggplot2); library(lme4); library(xtable); library(plyr); library(lattice); library(LMERConvenienceFunctions); library(car); library(grid); library(lmerTest); library(reshape2); library(ggplot2); library(languageR); library(pander); source('summarySEwithin2.R'); library(Hmisc)
+```{r}
+dat <- preprocessing(root_dir, data_dir)
+dat <- perSubjectTrim.fnc(dat, response='RT', subject='Subject', trim = 2.5)$data
+dat$RT_log <- log(dat$RT)
 ```
+
+# Info and Procedure (Method)
 
@@
 
-```{r showInstructions, echo=FALSE, cache=TRUE}
-v = c("crisp", "crisp", "vague", "vague")
-n = c('number', 'number', 'number', 'number')
-t = c("matching", "comparison", "matching", "comparison")
-e = c("Choose a square with 6 dots", "Choose a square with fewer than 20 dots", "Choose a square with about 10 dots", "Choose a square with far fewer than 20 dots")
-i = data.frame(v, n, t, e)
-names(i) = c("vagueness", "instruction format", "selection task", "example instruction")
-pander(i, justify=c('center', 'left', 'left', 'left'))
+# Sample display
+
+Remember to do a screenshot for the sample display.
+
+# Full table of instructions
+
+```{r, echo=FALSE}
+instructions_table <- 
+  dat %>%
+    select(Item, Quantity, Vagueness, Selection, Instruction) %>%
+    unique() %>%
+    arrange(Item, Quantity, Vagueness, Selection) %>%
+    spread(key=Vagueness, value=Instruction)
+# kable(instructions_table, format="latex", booktabs = TRUE)
+instructions_table %>% 
+  kable(align='cccll', caption="Full table of instructions") %>% 
+      kable_styling(full_width = F, position = "left", font_size = 11) 
 ```
 
-```{r showItems, echo=FALSE, cache=TRUE}
-items = c(1,2,3,4,5,6,7,8)
-i = data.frame(Item = c("1","2","3","4"),
-               x = c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
-names(i) = c('Item', 'Numbers of dots')
-pander(i, justify=c('center', 'center'))
-```
-
-```{r getData, echo=T}
-rawdata <- read.table('data.txt', sep=' ')
-```
+# Means plots
 
 ```{r trimat25sd}
-trimdata <- perSubjectTrim.fnc(rawdata, response='RT', subject='Subject', trim = 2.5)$data
-rtdata <- trimdata
+dat_plot <- summarySEwithin2(dat, measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
 ```
@@
 ```{r rtplot1, fig.width=7, fig.height=3}
-# this plot uses a precomputed summary done by summarySEwithin
-# and data trimmed at 2.5 sd per subject
-rtdataplot1 <- summarySEwithin2(data=perSubjectTrim.fnc(rawdata, response='RT', subject='Subject', trim = 2.5)$data,
-                                measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
 dodge = position_dodge(width=0.2)
-rtdataplot1$Condition <- as.factor(paste(sep=' ', rtdataplot1$Selection, rtdataplot1$Vagueness))
-rtplot=ggplot(rtdataplot1,
+dat_plot$Condition <- as.factor(paste(sep=' ', dat_plot$Selection, dat_plot$Vagueness))
+rtplot=ggplot(dat_plot,
               aes(y=RT_logNormed, x=Item, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci,
@@
 
-```{r rtplotHmisc, fig.width=7, fig.height=3}
-# n.b. rtdata has been trimmed at 2.5 sd: use rawdata if you want untrimmed data
-rawdata$Condition <- as.factor(paste(sep=' ', rawdata$Selection, rawdata$Vagueness))
-ggplot(rawdata, aes(y=RT_log, x=Item, group=Condition, shape=Condition, fill=Condition)) + 
-  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(0.2)) +
-  scale_shape_manual("",values = c(22, 22, 21, 21)) +
-  scale_fill_manual("",values=c("black","white","black","white")) +
-  facet_grid(~Selection) +
-  theme(
-    panel.background=element_rect(fill=NA, color='black'),
-    panel.grid.major=element_line(color='grey95'), 
-        #panel.grid.minor = element_blank(),
-        legend.key=element_blank(), 
-        aspect.ratio=1,
-        axis.text.x=element_text(angle = 45, hjust = 1)
-        )
+```{r}
+dat_borderline <- dat
 ```
 
-```{r rtlmer, cache=TRUE}
-rtlmer = lmerTest::lmer(data=rtdata,
+```{r, "Borderline response distribution plot", fig.width=7, fig.height=3.5, echo=FALSE, results='hide'}
+dat_borderline$response_cat <- factor(dat_borderline$response_cat, levels= c("Near", "Expected", "Far"))
+
+ggplot(dat_borderline) + 
+  geom_bar(aes(response_cat, group=Selection:Vagueness, fill=Vagueness), position=position_dodge(width=NULL)) +   
+  scale_fill_grey() + 
+  xlab(NULL) + 
+  ggtitle('Borderline response distribution') + 
+  facet_grid(~Selection) +
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), legend.position="top", aspect.ratio=1, plot.background = element_rect(fill=NA, color='white'), strip.background=element_blank(), legend.key.size=unit(4, 'mm'), axis.text.x = element_text(angle = 15))
+
+```
+
+# Hypotheses
+
+# Results
+
+```{r rtlmer, cache=TRUE, results='asis'}
+dat_model <- dat
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Sel <- ifelse(dat_model$Selection=="Comparison", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
+rtlmer = lmerTest::lmer(data=dat_model,
                         RT_log ~ c_Vag * c_Sel + c_Itm +
                           (1 + c_Vag * c_Sel + c_Itm | Subject))
+pretty_coef_table(rtlmer, "rtlmer")
 ```
 
-```{r rtlmersummaryxtable, results='asis', cache=TRUE, echo=F}
-print(xtable(summary(rtlmer)$coef, digits=c(0,2,2,1,1,4)),type='html')
+```{r seplmersforcomparison, results='asis', cache=TRUE, echo=-5}
+dat_model <- subset(dat, subset=Selection=='Comparison')
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
+s1lm <- lmerTest::lmer(data=dat_model, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
+pretty_coef_table(s1lm, "s1lm")
 ```
 
-```{r seplmersforcomparison, results='asis', cache=TRUE, echo=-3}
-s1 <- subset(rtdata, subset=Selection=='Comparison')
-s1lm <- lmerTest::lmer(data=s1, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
-print(xtable(summary(s1lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
+```{r seplmersformatching, results='asis', cache=TRUE, echo=-5}
+dat_model <- subset(dat, subset=Selection=='Matching')
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
+s2lm <- lmerTest::lmer(data=dat_model, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
+pretty_coef_table(s2lm, "s2lm")
 ```
 
-```{r seplmersformatching, results='asis', cache=TRUE, echo=-3}
-s2 <- subset(rtdata, subset=Selection=='Matching')
-s2lm <- lmerTest::lmer(data=s2, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
-print(xtable(summary(s2lm)$coef, digits=c(0,2,2,1,1,4)), type='html')
-```
+# Discussion

### 2018-08-11 experiments/D_exp_2/writeup/D_exp_2_functions/preprocessing.R

@@ -1,0 +1,230 @@
+preprocessing = function(root_dir, data_dir) {
+  
+  # The policy as of Aug 2018 is that:
+  # RT=0 is stripped, and one case with a very long time RT=49871 is stripped.
+  
+  dat <- read.delim(file.path(root_dir, data_dir, "data_raw.txt"), stringsAsFactors = FALSE)
+  
+  # declare local variables
+  number_of_valid_subjects <- 38
+  number_of_rows <- 7296
+  number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # 192
+  
+  # sort out borderline responses into expected near and far
+  dat$RESPONSE <- as.character(dat$RESPONSE)
+  for (row in 1 : nrow(dat) ) {
+    switch(dat[row,'RESPONSE'],
+           'LEFT' = {dat[row, 'choice'] <- dat[row, 'Left']},
+           'MIDDLE' = {dat[row, 'choice'] <- dat[row, 'Mid']},
+           'RIGHT' = {dat[row, 'choice'] <- dat[row, 'Right']}         
+    )
+  }
+  # now delete the tempoary variable RESPONSE
+  dat$RESPONSE <- NULL
+  dat$Left <- NULL
+  dat$Right <- NULL
+  dat$Mid <- NULL
+  
+  dat$crossed=paste('Con',dat$Condition,':Quan',dat$Quantity,':Item',dat$Item,sep="")
+  
+  dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseExpected'] <-  6 
+  dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseNear']     <- 15  
+  dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseFar']      <- 24
+  
+  dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseExpected'] <- 16
+  dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseNear']     <- 25  
+  dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseFar']      <- 34
+  
+  dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseExpected'] <- 26
+  dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseNear']     <- 35
[…]

### 2018-08-11 experiments/D_exp_2/writeup/add_stars.R

@@ -1,77 +1,0 @@
-add_stars = function(modl) {
-  
-  require(broom.mixed)
-  
-  coef_table <- as.data.frame(tidy(modl, effects="fixed"))
-  
-  # be sensitive to whether the model is an lmer or a glmer 
-  
-  if (class(modl) == "lmerModLmerTest") {
-    
-    # sort the names out
-    names(coef_table) <- c("effect", "term", "coefficient", "std.error", "df", "t_value", "p_value_raw")
-    coef_table$p_value <- as.character("")
-    coef_table$star <- as.character("")
-    
-    # categorise p values and add significance stars
-    for (i in 1:nrow(coef_table)) {
-      if (!is.na(coef_table$p_value_raw[i])) {
-        if (coef_table$p_value_raw[i] >= 0     & coef_table$p_value[i] < 0.001) {coef_table$p_value[i] <-    "<0.001"; coef_table$star[i] <- "***"}
-        if (coef_table$p_value_raw[i] >= 0.001 & coef_table$p_value[i] < 0.01 ) {coef_table$p_value[i] <-    "<0.01"; coef_table$star[i] <- "**"}
-        if (coef_table$p_value_raw[i] >= 0.01  & coef_table$p_value[i] < 0.05 ) {coef_table$p_value[i] <-    "<0.05"; coef_table$star[i] <- "*"}
-        if (coef_table$p_value_raw[i] >= 0.05  ) {coef_table$p_value[i] <-    round(coef_table$p_value_raw[i],3); coef_table$star[i] <-  ""}
-      }
-    }
-
-    # convert numerics to formatted chars for pretty printing
-    coef_table$coefficient <- format(x=coef_table$coefficient, nsmall=2, digits=2)
-    coef_table$std.error   <- format(x=coef_table$std.error, nsmall=2, digits=2)
-    coef_table$df          <- format(x=coef_table$df, nsmall=2, digits=1)
-    coef_table$t_value     <- format(x=coef_table$t_value, nsmall=2, digits=2)
-    coef_table$p_value_raw <- format(x=coef_table$p_value_raw, nsmall=2, digits=3)
-    
-    # select rows and cols
-    coef_table <- coef_table %>%
-      select(-effect) %>%
-      filter(term!="(Intercept)") 
-    
-    return(coef_table)
-  
-  }
-  
-  if (class(modl) == "glmerMod") {
-    
-    # sort the names out
[…]

### 2018-08-11 experiments/D_exp_2/writeup/package_install_script.R

@@ -1,2 +1,0 @@
-install.packages(c("knitr", "MASS", "ggplot2", "lme4", "xtable", "plyr", "lattice", "LMERConvenienceFunctions", "car", 
-                   "lmerTest", "reshape2", "ggplot2", "languageR", "pander"))

### 2018-08-11 experiments/D_exp_2/writeup/preprocessing/scratch.R

@@ -1,29 +1,0 @@
-# # add ratios for Item
-# item_range_ratio = c(6 / 24, 16 / 34, 26 / 44, 36 / 54) 
-# # 0.2500000 0.4705882 0.5909091 0.6666667
-# item_range_ratio_scaled = as.vector(scale(c(6 / 24, 16 / 34, 26 / 44, 36 / 54)))
-# # -1.3441995 -0.1316642  0.5297187  0.9461450
-# item_mean_ratio = c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))) 
-# # 0.5125000 0.6876471 0.7691558 0.8166667
-# item_mean_ratio_scaled = as.vector(scale(c(mean(c(6 / 15, 15 / 24)), mean(c(16 / 25, 25 / 34)), mean(c(26 /35, 35 / 44)), mean(c(36 / 45, 45 / 54))))) 
-# # -1.37582241 -0.06614191  0.54334858 0.89861574
-
-# dat[dat$Item == "06:15:24", "item_range_ratio"] <-  0.2500000
-# dat[dat$Item == "16:25:34", "item_range_ratio"] <-  0.4705882
-# dat[dat$Item == "26:35:44", "item_range_ratio"] <-  0.5909091
-# dat[dat$Item == "36:45:54", "item_range_ratio"] <-  0.6666667
-# 
-# dat[dat$Item == "06:15:24", "item_range_ratio_scaled"] <-  -1.3441995
-# dat[dat$Item == "16:25:34", "item_range_ratio_scaled"] <-  -0.1316642
-# dat[dat$Item == "26:35:44", "item_range_ratio_scaled"] <-   0.5297187  
-# dat[dat$Item == "36:45:54", "item_range_ratio_scaled"] <-   0.9461450
-# 
-# dat[dat$Item == "06:15:24", "item_mean_ratio"] <-  0.5125000 
-# dat[dat$Item == "16:25:34", "item_mean_ratio"] <-  0.6876471 
-# dat[dat$Item == "26:35:44", "item_mean_ratio"] <-  0.7691558
-# dat[dat$Item == "36:45:54", "item_mean_ratio"] <-  0.8166667
-# 
-# dat[dat$Item == "06:15:24", "item_mean_ratio_scaled"] <-   -1.37582241 
[…]

### 2018-08-12 experiments/C_exp_1/writeup/C_exp_1.Rmd
consistent C, D, and E formats in RMD

@@
 
-```{r, echo=FALSE}
+```{r, "full table of instructions", echo=FALSE}
 instructions_table <- 
@@
 
-```{r, echo=F}
+```{r, "make summary data for rt plot", echo=FALSE}
 dat_plot <- summarySEwithin2(dat, measurevar="RT_log", withinvars=c("Vagueness", "Number", "Item"), idvar="Subject")
@@
 
-```{r, "RT condition means plot", fig.width=7, fig.height=3.5, echo=FALSE, results='hide'}
-# correction from Morey (2008) applied to condition means
+```{r, "EXP_C_RT_condition_means_plot", fig.width=7, fig.height=3.5, echo=FALSE}
 mywidth=0
 pdodge=position_dodge(width=mywidth)
-#
 ggplot(dat_plot, aes(x=Item, y=RT_logNormed, group=Vagueness, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci, shape=Vagueness, fill=Vagueness)) +
@@
   ylab("RT log(ms)") + xlab(NULL) +
-  #
   geom_errorbar(position=pdodge, width=0.25) +
@@
   geom_point(position=pdodge, size=2)
-
 ```
 
-```{r, "Borderline response distribution plot", fig.width=7, fig.height=3.5, echo=FALSE, results='hide'}
+```{r, "EXP_C_Borderline_response_distribution_plot", fig.width=7, fig.height=3.5, echo=FALSE}
 dat_borderline$response_category <- relevel(dat_borderline$response_category, ref='expected')
@@
 Original full model
-```{r "original full RT model", echo=F, cache=TRUE}
+```{r "original full RT model", cache=TRUE}
 dat_model <- dat
@@
 Numeric only
-```{r, echo=FALSE}
+```{r "numeric only RT model", cache=TRUE}
 dat_model <- droplevels(subset(dat, Number=="Numeric"))
@@
 Verbal only
-```{r, echo=FALSE}
+```{r "verbal only RT model", cache=TRUE}
 dat_model <- droplevels(subset(dat, Number=="Verbal"))
@@
 Borderline model
-```{r 'bl-full-model', cache=TRUE, echo=FALSE}
+```{r "borderline full model", cache=TRUE}
 dat_model <- dat_borderline
@@
 Model of the data after the 6:15:24 level of Item is removed
-```{r, echo=F, cache=TRUE}
+```{r, "full RT model after drop item", cache=TRUE}
 dat_model <- droplevels(subset(dat, Item!="06:15:24"))
@@
 Numeric only after the 6:15:24 level of Item is removed
-```{r, echo=F, cache=TRUE}
+```{r, "num only RT model after drop item", cache=TRUE}
 dat_model <- droplevels(subset(dat, Item!="06:15:24" & Number=="Numeric"))
@@
 Verbal-only after the 6:15:24 level of Item is removed
-```{r, echo=F, cache=TRUE}
+```{r, "verbal only RT model after drop item", cache=TRUE}
 dat_model <- droplevels(subset(dat, Item!="06:15:24" & Number=="Verbal"))

### 2018-08-12 experiments/D_exp_2/writeup/D_exp_2.Rmd
consistent C, D, and E formats in RMD

@@
 ```{r}
-dat <- preprocessing(root_dir, data_dir)
+dat_full <- preprocessing(root_dir, data_dir)
+dat <- subset(dat_full, subset=RT>0 & RT<49871)
+dat_borderline <- dat
 dat <- perSubjectTrim.fnc(dat, response='RT', subject='Subject', trim = 2.5)$data
@@
 
-```{r trimat25sd}
+```{r}
 dat_plot <- summarySEwithin2(dat, measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
@@
 
-```{r rtplot1, fig.width=7, fig.height=3}
+```{r, "EXP_D_RT_means_plot", fig.width=7, fig.height=3, echo=FALSE}
 dodge = position_dodge(width=0.2)
 dat_plot$Condition <- as.factor(paste(sep=' ', dat_plot$Selection, dat_plot$Vagueness))
-rtplot=ggplot(dat_plot,
-              aes(y=RT_logNormed, x=Item, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci,
-                group=Condition, shape=Condition, fill=Condition)) +
+ggplot(dat_plot, aes(y=RT_logNormed, x=Item, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci, group=Condition, shape=Condition, fill=Condition)) +
   geom_line(position=dodge) +
@@
   facet_grid(~Selection) 
-rtplot
 ```
 
-```{r}
-dat_borderline <- dat
-```
-
-```{r, "Borderline response distribution plot", fig.width=7, fig.height=3.5, echo=FALSE, results='hide'}
+```{r, "EXP_D_Borderline_response_distribution_plot", fig.width=7, fig.height=3.5, echo=FALSE}
 dat_borderline$response_cat <- factor(dat_borderline$response_cat, levels= c("Near", "Expected", "Far"))
-
 ggplot(dat_borderline) + 
@@
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), legend.position="top", aspect.ratio=1, plot.background = element_rect(fill=NA, color='white'), strip.background=element_blank(), legend.key.size=unit(4, 'mm'), axis.text.x = element_text(angle = 15))
-
 ```
@@
 
-```{r rtlmer, cache=TRUE, results='asis'}
+```{r, "rt_lmer_full_model_all_items", cache=TRUE}
 dat_model <- dat
@@
 dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
-rtlmer = lmerTest::lmer(data=dat_model,
-                        RT_log ~ c_Vag * c_Sel + c_Itm +
-                          (1 + c_Vag * c_Sel + c_Itm | Subject))
-pretty_coef_table(rtlmer, "rtlmer")
+rtlmer = lmerTest::lmer(data=dat_model, RT_log ~ c_Vag * c_Sel + c_Itm + (1 + c_Vag * c_Sel + c_Itm | Subject))
+pretty_coef_table(rtlmer, "rt_lmer_full_model_all_items")
 ```
 
-```{r seplmersforcomparison, results='asis', cache=TRUE, echo=-5}
+```{r, "rt_lmer_for_comparison_all_items", cache=TRUE}
 dat_model <- subset(dat, subset=Selection=='Comparison')
@@
 dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
-s1lm <- lmerTest::lmer(data=dat_model, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
-pretty_coef_table(s1lm, "s1lm")
+comp_lmer <- lmerTest::lmer(data=dat_model, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
+pretty_coef_table(comp_lmer, "rt_lmer_for_comparison_all_items")
 ```
 
-```{r seplmersformatching, results='asis', cache=TRUE, echo=-5}
+```{r, "rt_lmer_for_matching_all_items", cache=TRUE}
 dat_model <- subset(dat, subset=Selection=='Matching')
@@
 dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
-s2lm <- lmerTest::lmer(data=dat_model, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
-pretty_coef_table(s2lm, "s2lm")
+match_lmer <- lmerTest::lmer(data=dat_model, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
+pretty_coef_table(match_lmer, "rt_lmer_for_matching_all_items")
 ```

### 2018-08-12 experiments/E_exp_3/writeup/E_exp_3.Rmd
consistent C, D, and E formats in RMD

@@
 ---
-title: "Experiment 4 of 4"
+title: "E exp 3"
 author: "Matt Green"
@@
     toc: true
-    <!--toc_float: true-->
-    mathjax: default
-    template: default
-    <!--code_folding: hide-->
-    theme: default
+    toc_float: false
 ---
-<style>
-table, th, td {
-    border: 1px grey;
-    border-collapse: collapse;
-    padding:2px 5px 2px 5px
-}  
-</style>
 
-```{r libs, message=FALSE}
-library(knitr); library(MASS); library(ggplot2); library(lme4); library(xtable); library(plyr); library(lattice); library(LMERConvenienceFunctions); library(car); library(grid); library(lmerTest); library(reshape2); library(ggplot2); library(languageR); library(pander); source('summarySEwithin2.R')
-library(kableExtra)
+```{r, "libs and functions", echo=FALSE, message=FALSE}
+library(knitr)
+library(tidyverse)
 library(broom)
 library(broom.mixed)
-source("pretty_coef_table.R")
+library(lme4)
+library(lmerTest)
+library(dotwhisker)
+library(kableExtra)
+library(LMERConvenienceFunctions)
+library(data.table)
+library(gridExtra)
+source("E_exp_3_functions/preprocessing.R")
+source("E_exp_3_functions/pretty_coef_table.R")
+source("E_exp_3_functions/summarySEwithin2.R")
+root_dir <- getwd()
+data_dir <- "E_exp_3_data"
 ```
 
-# Experiment 4
+```{r, get_data}
+dat_full <- preprocessing(root_dir, data_dir)
+dat <- subset(dat_full, RT>0 & RT<9500)
+dat_borderline <- dat
+dat <- perSubjectTrim.fnc(dat, response='RT', subject='Subject', trim = 2.5)$data
+dat$RT_log <- log(dat$RT)
+```
+
+# Info and Procedure (Method)
+
+The distinctive thing in this experiment is that in order to avoid mentioning a number, we had to show a "Target" (similar to a prime) that presented an array with a certain number of dots which we could later refer to as "the target" in expressions like "the same number of dots as the target".
+
+RTs were overwhelmingly faster in this experiment than in experiments where numbers were used in the instructions.
 
@@
 
-# Trial Procedure
 First an instruction was presented until subjects dismissed it with a key press.
@@
 
+# Sample display
+
 <figure>
-<img src="./E_exp_3-flow.jpg" alt="flow-of-D_exp_3" style="width:480px;height:300px;">
+<img src="./E_exp_3_figures_manual/E_exp_3_flow.jpg" alt="flow-of-D_exp_3" style="width:480px;height:300px;">
 <figcaption>Flow of experiment E_exp_3.</figcaption>
@@
 
-# Describe the data
+# Full table of instructions
 
-Here is a description of each column in the raw data.
-
-* **Sub** identifier for subject. There are 40 subjects.
-* **Trl** identifier for the trial. There are 64 Trials for each subject.
-* **Itm** identifier for item.
-There are 4 items.
-An item is a set of Selections for the dots in the squares, e.g., Item 1 is (6,15,24) in any order.
-This should be included as a (possibly ordered) factor because the task will probably take longer for more dots.
-Include this in the final model if it does increase rt.
-Table below shows the dots for each item.
-* **Cnd** just a single vector for the combination of vagueness and selection. Useful for some plotting routines that don't accept pairs of grouping variables. 
-* **Ord** whether the order was for example (6,15,24) or (24,14,6).
-This matters because when the smaller quantity is on the left people are quicker than when the smaller quantity is on the right (i.e, SNARC effect).
-It is a control variable rather than a direct predictor.
-It could be left out of the main model assuming it is non-significant.
-* **Qty** whether the instruction identifies bigger or smaller quantities.
-This is a factor with 2 levels sum coded.
-An example of bigger is _"Choose a square with more dots than the target"_.
-An example of smaller is _"Choose a square with fewer dots than the target"_.
-* **Prm** how many dots there were in the prime (prime is _target_ in the instructions to the participants). Primes were 6 10 16 20 24 26 30 34 36 40 44 50 54.
-* **Lft** how many dots there were in the left square
-* **Mid** how many dots there were in the middle square
-* **Rgt** how many dots there were in the right-hand side square
-* **key** which key the subject pressed as a response on a trial.
-* **rt** raw response time in milliseconds
-* **Ins** a character array for the instruction that was presented to the subject on that trial. Table below shows an example of an instruction in each condition.
-* **V** factor for vagueness with 2 levels sum coded
-* **S** factor for selection task with 2 levels sum coded
-
-Here is a table of instructions.
-
-<!-- <style type="text/css"> -->
-<!-- .table {width: 50%;} -->
-<!-- </style> -->
-
-vagueness | selection | instruction
-----------|-----------|--------------------------------------------------------------------
-crisp     | match     | Choose a square with the same number of dots as the target  
-crisp     | compare   | Choose a square with fewer dots than the target 
-vague     | match     | Choose a square with about the same number of dots as the target 
-vague     | compare   | Choose a square with far fewer dots than the target
-
-Item   | dots
--------|-----------
-Item 1 | (6,15,24)
-Item 2 | (16,25,34)
-Item 3 | (26,35,44)
-Item 4 | (36,45,54)
-
-```{r getdata}
-rawdata <- read.table('data.txt')
+```{r, "Full table of instructions", echo=FALSE}
+instructions_table <- 
+  dat %>%
+    select(Item, Target, Quantity, Vagueness, Selection, Instruction) %>%
+    unique() %>%
+    arrange(Item, Target, Quantity, Vagueness, Selection) %>%
+    spread(key=Vagueness, value=Instruction)
+# kable(instructions_table, format="latex", booktabs = TRUE)
+instructions_table %>% 
+  kable(align='ccccll', caption="Full table of instructions") %>% 
+      kable_styling(full_width = F, position = "left", font_size = 11) 
 ```
 
-```{r targets}
-target <- subset(rawdata, subset= Item=="06:15:24" & Quantity=="Small")
-target2 <- subset(target, select=c(Item, Condition, Ins, Target))
-target3 <- unique(target2[order(target$Condition),] )
-row.names(target3) <- NULL
-pander(target3)
+# Means plots
+
+```{r, "make data for rt plot"}
+dat_plot <- summarySEwithin2(dat, measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
 ```
 
-```{r trimat25sd}
-trimdata <- perSubjectTrim.fnc(rawdata, response='RT', subject='Subject', trim = 2.5)$data
-rtdata <- trimdata
-```
-
-```{r computeSummariesForGraphs, cache=TRUE, message=FALSE}
-rtdataplot1 <- summarySEwithin2(rtdata, measurevar="RT_log", withinvars=c("Condition", "Vagueness", "Selection", "Item"), idvar="Subject")
-```
-
-```{r rtplot, fig.width=7, fig.height=3, echo=F}
+```{r, "EXP_E_rt_means_plot", fig.width=7, fig.height=3, echo=FALSE}
 dodge = position_dodge(width=0.2)
-rtdataplot1$Condition <- as.factor(paste(sep=' ', rtdataplot1$Selection, rtdataplot1$Vagueness))
-rtplot=ggplot(rtdataplot1, aes(y=RT_logNormed, x=Item, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci,
-                group=Condition, shape=Condition, fill=Condition)) +
+dat_plot$Condition <- as.factor(paste(sep=' ', dat_plot$Selection, dat_plot$Vagueness))
+ggplot(dat_plot, aes(y=RT_logNormed, x=Item, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci, group=Condition, shape=Condition, fill=Condition)) +
   geom_line(position=dodge) +
@@
   facet_grid(~Selection) 
-rtplot
 ```
 
-```{r rtlmer, results='asis', cache=TRUE, echo=-2}
-rtlmer = lmerTest::lmer(data=rtdata, RT_log ~ c_Vag * c_Sel + c_Itm + (1 + c_Vag * c_Sel + c_Itm | Subject))
-pretty_coef_table(rtlmer, "rtlmer")
+```{r, "EXP_E_Borderline_response_distribution_plot", fig.width=7, fig.height=3, echo=FALSE}
+dat_borderline$response_cat <- factor(dat_borderline$response_cat, levels= c("Near", "Expected", "Far"))
+ggplot(dat_borderline) + 
+  geom_bar(aes(response_cat, group=Selection:Vagueness, fill=Vagueness), position=position_dodge(width=NULL)) +   
+  scale_fill_grey() + 
+  xlab(NULL) + 
+  ggtitle('Borderline response distribution') + 
+  facet_grid(~Selection) +
+  theme_bw() + 
+  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank(), legend.position="top", aspect.ratio=1, plot.background = element_rect(fill=NA, color='white'), strip.background=element_blank(), legend.key.size=unit(4, 'mm'), axis.text.x = element_text(angle = 15))
 ```
 
-```{r seplmersforcomparison, results='asis', cache=TRUE, echo=-3}
-s1 <- subset(rtdata, subset=Selection=='Comparison')
-s1lm <- lmerTest::lmer(data=s1, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
-pretty_coef_table(s1lm, "s1lm")
+# Hypotheses
+
+# Results
+
+```{r, "rt_lmer_full_model_all_items", cache=TRUE}
+dat_model <- dat
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Sel <- ifelse(dat_model$Selection=="Comparison", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
+rtlmer = lmerTest::lmer(data=dat_model, RT_log ~ c_Vag * c_Sel + c_Itm + (1 + c_Vag * c_Sel + c_Itm | Subject))
+pretty_coef_table(rtlmer, "rt_lmer_full_model_all_items")
 ```
 
-```{r seplmersformatching, results='asis', cache=TRUE, echo=-3}
-s2 <- subset(rtdata, subset=Selection=='Matching')
-s2lm <- lmerTest::lmer(data=s2, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
-pretty_coef_table(s2lm, "s2lm")
+```{r, "rt_lmer_for_comparison_all_items", cache=TRUE}
+dat_model <- subset(dat, subset=Selection=='Comparison')
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
+comp_lmer <- lmerTest::lmer(data=dat_model, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
+pretty_coef_table(comp_lmer, "rt_lmer_for_comparison_all_items")
 ```
+
+```{r "rt_lmer_for_matching_all_items", cache=TRUE}
+dat_model <- subset(dat, subset=Selection=='Matching')
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
+match_lmer <- lmerTest::lmer(data=dat_model, RT_log ~ c_Vag + c_Itm + (1 + c_Vag + c_Itm | Subject))
+pretty_coef_table(match_lmer, "rt_lmer_for_matching_all_items")
+```
+
+# Discussion

### 2018-08-12 experiments/C_exp_1/writeup/C_exp_1_functions/annotate_the_raw_data.R

@@ -32,9 +32,6 @@ annotate_the_raw_data = function (dat) {
   # add an Observation index, over the whole data set (1:7680)
   dat$Obs <- 1:7680
   
-  # remove variables that don't matter for purposes of RT analysis
-  #dat <- select(dat, Obs, Trial, Instruction, Subject, Item, Quantity, Order, Vagueness, Number, response_category, isBorderline, RT)
-  
   return(dat)
 }  
 

### 2018-08-12 experiments/C_exp_1/writeup/C_exp_1.Rmd
removed archives and added a screenshot to C

@@
 
-On each trial, first the referring expression that constituted the instruction for that trial was displayed. 
+On each trial, first the referring expression that constituted the instruction for that trial was displayed (e.g., "Choose a square with about 10 dots").
 Participants then pressed a key to indicate that they had read the instruction.
-After 1000 ms, the arrays were presented, while preserving the text of the referring expression. 
-The response time dependent variable was measured from the presentation of the arrays, until the keypress indicating the participant's choice, which was also recorded. 
-The trial would timeout after 60 seconds if there was no response.
+
+The instruction remained on screen, and after 1000 ms, the arrays appeared (see Figure below).
+<figure>
+<img src="C_exp_1_screenshots/EXP_C_example_screenshot.bmp" style="width:480px;height:300px;">
+</figure>
+
+Response time was measured from the presentation of the arrays until the keypress indicating the participant's choice. The trial would timeout after 60 seconds if there was no response.
+
 In this experiment, no feedback was given.
 This was because, in the vague conditions, we did not regard any response as "correct" or "incorrect", but instead as "expected response"; "borderline response"; and "extreme response", and we did not want to draw participants' attention to this distinction explicitly.
-We simply recorded whether the participant chose the best referent, the borderline case or the poorest referent, and how long it took the participant to respond.
-
-# Sample display
-
-Remember to do a screenshot for the sample display.
+Which choice the participant made was recorded for analysis.
 

### 2018-08-12 experiments/B_pilot_2/trashme/CIfunctions.R

@@ -1,140 +1,0 @@
-## Summarizes data.
-## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
-##   data: a data frame.
-##   measurevar: the name of a column that contains the variable to be summariezed
-##   groupvars: a vector containing names of columns that contain grouping variables
-##   na.rm: a boolean that indicates whether to ignore NA's
-##   conf.interval: the percent range of the confidence interval (default is 95%)
-summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
-                      conf.interval=.95, .drop=TRUE) {
-  require(plyr)
-  
-  # New version of length which can handle NA's: if na.rm==T, don't count them
-  length2 <- function (x, na.rm=FALSE) {
-    if (na.rm) sum(!is.na(x))
-    else       length(x)
-  }
-  
-  # This is does the summary; it's not easy to understand...
-  datac <- ddply(data, groupvars, .drop=.drop,
-                 .fun= function(xx, col, na.rm) {
-                   c( N    = length2(xx[,col], na.rm=na.rm),
-                      mean = mean   (xx[,col], na.rm=na.rm),
-                      sd   = sd     (xx[,col], na.rm=na.rm)
-                   )
-                 },
-                 measurevar,
-                 na.rm
-  )
-  
-  # Rename the "mean" column    
-  datac <- rename(datac, c("mean"=measurevar))
-  
-  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
-  
-  # Confidence interval multiplier for standard error
-  # Calculate t-statistic for confidence interval: 
-  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
-  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
-  datac$ci <- datac$se * ciMult
-  
-  return(datac)
-}
-
-
-## Summarizes data, handling within-subjects variables by removing inter-subject variability.
-## It will still work if there are no within-S variables.
-## Gives count, un-normed mean, normed mean (with same between-group mean),
-##   standard deviation, standard error of the mean, and confidence interval.
-## If there are within-subject variables, calculate adjusted values using method from Morey (2008).
-##   data: a data frame.
-##   measurevar: the name of a column that contains the variable to be summariezed
[…]

### 2018-08-12 experiments/B_pilot_2/trashme/e1.R

@@ -1,97 +1,0 @@
-## ----loadLibs------------------------------------------------------------
-suppressPackageStartupMessages(library(knitr))
-suppressPackageStartupMessages(library(ggplot2))
-suppressPackageStartupMessages(library(xtable))
-suppressPackageStartupMessages(library(gridExtra))
-suppressPackageStartupMessages(library(lme4))
-suppressPackageStartupMessages(library(lmerTest)) 
-suppressPackageStartupMessages(library(data.table)) 
-suppressPackageStartupMessages(library(dplyr))
-suppressPackageStartupMessages(library(dtplyr))  
-source('summarySEwithin2.R') 
-
-## ----separateDataFramesForRTandErrorAnalyses, echo=T---------------------
-dat <- read.table('data.txt', sep=' ')
-str(dat)
-erdata <- dat
-rtdata <- subset(dat, error==FALSE)
-
-## ----rtplot, message=FALSE-----------------------------------------------
-(rtdataplot1 = summarySEwithin2(rtdata,  measurevar="logrt",  withinvars=c("Vgn","Gap"), idvar="subject"))
-dodge=.06
-rtplot = ggplot(rtdataplot1, aes(y=logrtNormed, x=Gap, group=Vgn, fill=Vgn, shape=Vgn, ymin=logrtNormed-ci, ymax=logrtNormed+ci)) + 
-  geom_line(position=position_dodge(dodge)) + 
-  geom_errorbar(width=.3, lty=1, position=position_dodge(dodge)) + 
-  geom_point(pch=21, size=3, position=position_dodge(dodge)) +
-  ggtitle("Response time") + 
-  ylab("log RT") + 
-  xlab("Gap Size") +
-  scale_fill_grey(name=element_blank(), start=0, end=1) +
-  theme_bw() +
-  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), legend.key = element_blank() ) +
-  theme(aspect.ratio=1) + xlab("")
-rtplot
-
-## ----tableRTmodelSimple, results='asis'----------------------------------
-print(
-  xtable(
-    digits=3, 
-    summary(
-      rt.lmer <- lmer(data=rtdata, logrt ~ c_Vgn * c_Gap + (c_Vgn * c_Gap|subject))
-    )$coefficients,
-    caption="a response time model", 
-    caption.placement="top"), type='html')
-
-## ----tableRTmodelMaximal, results='asis'---------------------------------
-print(
-  xtable(
-    digits=3, 
-    summary(
[…]

### 2018-08-12 experiments/C_exp_1/archive/commentblock.R

@@ -1,3 +1,0 @@
-##################################
-##                              ##
-##################################

### 2018-08-12 experiments/C_exp_1/archive/which_subjects_violate_normality.R

@@ -1,12 +1,0 @@
-# which subjects violate normality
-f = function(dfr) return(shapiro.test(dfr$logRT)$p.value)
-p = as.vector(by(dat, dat$Subject, f))
-names(p) = levels(dat$Subject)
-names(p[p < 0.05])
-length(names(p[p < 0.05]))
-qqmath(~logRT | Subject, data = dat,
-       prepanel = prepanel.qqmathline,
-       panel = function(x, ...) {
-         panel.qqmathline(x, ...)
-         panel.qqmath(x, ...)
-       })

### 2018-08-12 experiments/C_exp_1/experimentCode/do_c.m

@@ -1,0 +1,778 @@
+function do_c(SUBJECT)
+% This third study has the following structure:
+% Practice, Rest, Block1 Block 2 Block 3, Block 4
+% There are 256 trials
+% There are 64 unique trials
+% There are 4 Blocks
+% Each trial consists of 3 squares each with a number of dots.
+%
+% A trial consists of:
+% the instruction on screen while participants read it for as long as they like;
+% a button press on SPACEBAR to bring up
+% a screen with the instruction and the dots presented
+%
+% Participant id > 100 is used for testing and can be overwritten
+
+howLongToWaitForAResponse = 60; % seconds. Set low for test runs
+testing = 1; % 0 for real runs, 1 for testing
+takescreenshot=1;
+SUBJECT=1001;
+%takescreenshot=0;
+
+% Clear Matlab/Octave window:
+clc;
+
+% check for Opengl compatibility, abort otherwise:
+AssertOpenGL;
+
+% Reseed the random-number generator for each expt.
+rand('state',sum(100*clock));
+
+% Make sure keyboard mapping is the same on all supported operating systems
+% Apple MacOS/X, MS-Windows and GNU/Linux:
+KbName('UnifyKeyNames');
+
+% Define keys by ID
+leftresp=KbName('c'); % is 6
+middleresp=KbName('b'); % is 5
+rightresp=KbName('m'); % is 16
+escapekey=KbName('ESCAPE'); % is 41
+spacebar=KbName('space'); % is 44
+returnkey=KbName('return'); % is 40
+
+%%%%%%%%%%%%%%%%%%%%%%
+% define stimuli
+%%%%%%%%%%%%%%%%%%%%%%
+
+% order in headers gives the answer to "Which square had the small quantity"? and does not give the correct response
+
+%Condition 1 is numerical vague
+%Condition 2 is numerical precise
+%Condition 3 is visual vague
+%Condition 4 is visual precise
+%
+%Order 1 is small number on left
+%Order 2 is small number on right
+%
+%Quantity 1 is small
+%Quantity 2 is big
+
+
+exp3headers={
+	"Item","Condition","Order","Quantity","Left","Mid","Right","Instruction"};
+
+exp3practicestimuli = {
+	0,0,0,0,6,15,24,"Choose a square with about 10 dots";
+	0,0,0,0,34,25,16,"Choose the square with 16 dots";
[…]

### 2018-08-12 experiments/E_exp_3/archive/scratch-eg-diff-and-error-chunks.R

@@ -1,309 +1,0 @@
-
-<<lmer-rt-1, echo=FALSE, cache=FALSE>>=
-  # rt model at word level
-  lmer.rt <- lmer(data=d, rt~aggregation*similarity*nconj + wlen + (1|subject) + (1|item))
-@
-
-<<lmer-msch-1>>=
-  # msch model at word level
-  lmer.msch <- lmer(data=d, msch~aggregation*similarity*nconj + wlen + (1|subject) + (1|item))
-@
-
-<<lmer-rat-1>>=
-  # rating ratings model at sentence level
-  dr <- unique(subset(d, select=c(subject,id,item,aggregation,similarity,nconj,rating)))
-lmer.DIF <- glmer(data=dr, rating ~ nconj*aggregation*similarity + (1|subject) + (1|item), family=poisson)
-@
-
-<<lmer-error-1>>=
-  # error rates model at sentence level
-  derr <- unique(subset(d, select=c(subject,id,item,aggregation,similarity,nconj,error)))
-lmer.ERROR <- glmer(data=derr, error~nconj+aggregation+similarity + (1|subject) + (1|item), family=binomial)
-@
-
-<<overviewallmeasures, echo=FALSE, fig.path="e1aggreg-writeup-figs/", fig.height=3, fig.width=9, fig.show='hide'>>=
-  d5 <- melt(d, measure.vars=c("rt","rating","error"))
-grandmeans <- dcast(d5, similarity+aggregation+nconj~variable, function(x)mean(x,na.rm=TRUE))
-g.data <- melt(grandmeans, measure.vars=c("rt","rating","error"))
-ggplot(data=g.data, aes(y=value, x=nconj, shape=similarity, linetype=aggregation, group=similarity:aggregation))  +   facet_wrap(~variable, scales="free_y") +   geom_point(aes(shape=similarity),size=3) +   geom_line() +  theme(legend.key = element_blank()) 
-@
-
-<<plotCoefsWithCisrt, echo=FALSE, fig.path="e1aggreg-writeup-figs/", fig.height=5, fig.width=7, fig.show='hide'>>=
-  xvals=as.character(names((summary(lmer.rt)$coefficients)[,3]))
-yvals=as.numeric((summary(lmer.rt)$coefficients)[,3])
-coef_values <- as.numeric(summary(lmer.rt)$coefficients[,1])
-st_err <- as.numeric(summary(lmer.rt)$coefficients[,2])
-upperCI <-  coef_values + 1.96*st_err
[…]

### 2018-08-12 experiments/D_exp_2/writeup/D_exp_2.Rmd
D has hypotheses and results for first set of models but not for the levy 2nd set

@@
 dat_full <- preprocessing(root_dir, data_dir)
-dat <- subset(dat_full, subset=RT>0 & RT<49871)
+dat <- subset(dat_full, subset=RT>0 & RT<25000) # loses 2 RTs that were judged to be outliers 25978 49871
 dat_borderline <- dat
@@
 
-Experiment Structure: 
-Each subject saw 192 trials, being 3 cycles of 64 trials. There were 38 participants. There were 7296 rows in the full data set.
+The main aim of this experiment was to see whether vague instructions would confer advantages over crisp alternatives when all instructions used numerals, and when there were vague and crisp versions of each of the comparison and matching strategies.
 
-Trial Procedure: 
-First an instruction that was a referring expression was presented until participants dismissed it with a key press.
-Then the squares with dots were presented until the participant responded with a key press, with a 60 second timeout if they did not respond.
+On each trial, first the referring expression that constituted the instruction for that trial was displayed (e.g., "Choose a square with about 10 dots").
+Participants then pressed a key to indicate that they had read the instruction.
 
-# Sample display
+The instruction remained on screen, and after 1000 ms, the arrays appeared (see Figure below).
+<figure>
+<img src="D_exp_2_screenshots/EXP_D_example_screenshot.bmp" style="width:480px;height:300px;">
+</figure>
 
-Remember to do a screenshot for the sample display.
+Response time was measured from the presentation of the arrays until the keypress indicating the participant's choice. The trial would timeout after 60 seconds if there was no response.
+
+In this experiment, no feedback was given.
+This was because, in the vague conditions, we did not regard any response as "correct" or "incorrect", but instead as "expected response"; "borderline response"; and "extreme response", and we did not want to draw participants' attention to this distinction explicitly.
+Which choice the participant made was recorded for analysis.
 
 # Full table of instructions
+
+The instructions for this experiment were similar to C_exp_2, except that we changed "fewer than" to "far fewer than": this was because "far fewer than" can have borderline cases whereas "fewer than" cannot have borderline cases.
 
@@
 
+We formulated the following hypotheses:
+
+* [Hypothesis 1] Crisp/Vague RT:
+    * Vague instructions should result in faster responses than crisp instructions.
+* [Hypothesis 2] Comparison/Matching RT:
+    * Instructions that allow comparison should result in faster responses than instructions that necessitate matching.
+* [Hypothesis 3] Vagueness effect differs between comparison and matching: 
+    * The effects of vagueness should differ when the selection task is comparison versus when it is matching.
+
 # Results
+
+Response times were trimmed at 25,000 ms, resulting in the loss of 2 trials that were judged to have outlier RTs.
+The remaining data were trimmed separately for each participant at 2.5 standard deviations below and above that participant's mean RT, leading to the loss of 204 trials (2.8% of the remaining trials).
+RTs in milliseconds were log-transformed so that they better approximated a normal distribution for the purposes of analysis.
+
+## The first set of models
+
+The first set of models considered included main effects of Vagueness and Item as well as their interaction.
 
@@
 
+Considering the first set of models with Vagueness and Item as main effects and the Vagueness:Item interaction in the same model:
+
+* [Hypothesis 1] Crisp/Vague RT:
+    * _Vague instructions should result in faster responses than crisp instructions_ : Vague instructions resulted in faster responses than crisp instructions but the difference was not significant (p=0.716).
+* [Hypothesis 2] Comparison/Matching RT:
+    * _Instructions that allow comparison should result in faster responses than instructions that necessitate matching_ : comparison instructions resulted in faster responses than matching instructions, and the difference was significant (p<0.001).
+* [Hypothesis 3] Vagueness effect differs between comparison and matching
+    * _The effects of vagueness should differ when the selection task is comparison versus when it is matching_ : Vagueness exerted effects in different directions for the two selection tasks (p<0.001).
+Separate analyses at each level of selection provided evidence that vague instructions resulted in faster responses than crisp instructions for the comparison condition (p<0.01); and slower responses than crisp instructions in the matching condition (p<0.01).
+
+## The models conducted to test for main effects _in the presence of interactions involving those main effects_ (after Levy, 2018)
+
 # Discussion

### 2018-08-12 experiments/D_exp_2/experimentCode/e3.m

@@ -1,317 +1,0 @@
-function e4(SUBJECT)
-ListenChar(2); % disable keyboard input to octave while running experiment (no echo to terminal)
-Screen('Preference', 'SkipSyncTests', 0); % 0 = do run sync tests; 1 = do not run sync tests
-% variables
-howLongToWaitForAResponse = 60; % seconds. Set low for test runs
-% Clear Matlab/Octave window:
-clc;
-% check for Opengl compatibility, abort otherwise:
-AssertOpenGL;
-% Reseed the random-number generator for each expt.
-rand('state',sum(100*clock));
-% Make sure keyboard mapping is the same on all supported operating systems
-KbName('UnifyKeyNames');
-% Define keys by ID
-leftresp=KbName('c'); % is 6
-middleresp=KbName('b'); % is 5
-rightresp=KbName('m'); % is 16
-escapekey=KbName('ESCAPE'); % is 41
-spacebar=KbName('space'); % is 44
-returnkey=KbName('return'); % is 40
-% read separate files containing dotlocations, stimuli, and strings 
-source('e4stimuli.m');
-source('e4strings.m');
-source('e4dotlocations.m');
-
-% set up output files
-% practice data output file to write to
-practicedatafilename = strcat('output/practice/subject',...
-    num2str(SUBJECT),'Practice','.data');
-% experimental data output file to write to
-datafilename = strcat('output/subject',...
-    num2str(SUBJECT),'.data');
-% check for existing result file to prevent accidentally overwriting
-% files from a previous subject/session (except for subject  numbers >
-% 99, which are reserved for testing)
-if SUBJECT<99 && fopen(datafilename, 'rt')~=-1
-    fclose('all');
-    error(strcat('Result data file already exists!',...
-        'Choose a different subject number.'));
-else
-    outputFilePointer = fopen(datafilename,'wt');
-        fprintf(outputFilePointer,...
-            '%s\t%s\t%s\t%s\t%s\t%s\t%s\t%s\t%s\t%s\t%s\t%s\n',...
-            'Item','Condition','Order','Quantity','Left','Mid','Right','Instruction','Subject','Instruction_RT','RESPONSE','RT')
-    outputFilePointerForPracticeTrials = ...
-        fopen(practicedatafilename,'wt');
-        fprintf(outputFilePointerForPracticeTrials,...
-            '%s\t%s\t%s\t%s\t%s\t%s\t%s\t%s\t%s\t%s\t%s\t%s\n',...
[…]

### 2018-08-12 experiments/D_exp_2/experimentCode/e3dotlocations.m

@@ -1,21 +1,0 @@
-    %%%%%%%%%%%%%%%%%%%
-    % set up the dots %
-    %%%%%%%%%%%%%%%%%%%
-
-    % describe allowable dot locations
-    % In this experiment we need up to 74 dots to be displayed at any one
-    % time in any one square. The setup allows 100 locations.
-
-    x=0;y=0;
-    dots = [...
-        [x-90;y-90] [x-70;y-90] [x-50;y-90] [x-30;y-90] [x-10;y-90] [x+10;y-90] [x+30;y-90] [x+50;y-90] [x+70;y-90] [x+90;y-90]...
-        [x-90;y-70] [x-70;y-70] [x-50;y-70] [x-30;y-70] [x-10;y-70] [x+10;y-70] [x+30;y-70] [x+50;y-70] [x+70;y-70] [x+90;y-70]...
-        [x-90;y-50] [x-70;y-50] [x-50;y-50] [x-30;y-50] [x-10;y-50] [x+10;y-50] [x+30;y-50] [x+50;y-50] [x+70;y-50] [x+90;y-50]...
-        [x-90;y-30] [x-70;y-30] [x-50;y-30] [x-30;y-30] [x-10;y-30] [x+10;y-30] [x+30;y-30] [x+50;y-30] [x+70;y-30] [x+90;y-30]...
-        [x-90;y-10] [x-70;y-10] [x-50;y-10] [x-30;y-10] [x-10;y-10] [x+10;y-10] [x+30;y-10] [x+50;y-10] [x+70;y-10] [x+90;y-10]...
-        [x-90;y+10] [x-70;y+10] [x-50;y+10] [x-30;y+10] [x-10;y+10] [x+10;y+10] [x+30;y+10] [x+50;y+10] [x+70;y+10] [x+90;y+10]...
-        [x-90;y+30] [x-70;y+30] [x-50;y+30] [x-30;y+30] [x-10;y+30] [x+10;y+30] [x+30;y+30] [x+50;y+30] [x+70;y+30] [x+90;y+30]...
-        [x-90;y+50] [x-70;y+50] [x-50;y+50] [x-30;y+50] [x-10;y+50] [x+10;y+50] [x+30;y+50] [x+50;y+50] [x+70;y+50] [x+90;y+50]...
[…]

### 2018-08-12 experiments/D_exp_2/experimentCode/expDstimuli.m

@@ -1,0 +1,99 @@
+%%%%%%%%%%%%%%%%%%%%%%
+% define stimuli
+%%%%%%%%%%%%%%%%%%%%%%
+
+% a column named _order_ in headers gives a answer to 
+% 'Which square had a small quantity'? and does not give a correct response
+
+%Condition 1 is numerical vague matching
+%Condition 2 is numerical crisp matching
+%Condition 3 is numerical vague comparison
+%Condition 4 is numerical crisp comparison
+%
+%Order 1 is small number on left
+%Order 2 is small number on right
+%
+%Quantity 1 is small
+%Quantity 2 is big
+
+e4headers={
+	'Item','Condition','Order','Quantity','Left','Mid','Right','Instruction'};
+
+e4practicestimuli = {
+	0,0,0,0,6,15,24,'Choose a square with about 10 dots';               % numerical vague matching
+	0,0,0,0,34,25,16,'Choose a square with 16 dots';                    % numerical crisp matching
+	0,0,0,0,44,35,26,'Choose a square with far fewer than 40 dots';     % numerical vague comparison
+	0,0,0,0,36,45,54,'Choose a square with more than 40 dots'} ;        % numerical crisp comparison
+
+
+e4stimuli={
+% item 1 
+1,1,1,1,6,15,24,'Choose a square with about 10 dots'; 		% numerical vague matching, small q on left, identify small
+1,1,1,2,6,15,24,'Choose a square with about 20 dots'; 		% numerical vague matching, small q on left, identify large
+1,1,2,1,24,15,6,'Choose a square with about 10 dots'; 		% numerical vague matching, small q on right, identify small
+1,1,2,2,24,15,6,'Choose a square with about 20 dots'; 		% numerical vague matching, small q on right, identify large
+1,2,1,1,6,15,24,'Choose a square with 6 dots';            	% numerical crisp matching, small q on left, identify small
+1,2,1,2,6,15,24,'Choose a square with 24 dots';           	% numerical crisp matching, small q on left, identify large
+1,2,2,1,24,15,6,'Choose a square with 6 dots';            	% numerical crisp matching, small q on right, identify small
+1,2,2,2,24,15,6,'Choose a square with 24 dots';           	% numerical crisp matching, small q on right, identify large
[…]

### 2018-08-12 experiments/D_exp_2/writeup/D_exp_2.Rmd
D and E have hypotheses and results for the first set of models but not the levy models

@@
 
-The first set of models considered included main effects of Vagueness and Item as well as their interaction.
-
 ```{r, "rt_lmer_full_model_all_items", cache=TRUE}
@@
 
-Considering the first set of models with Vagueness and Item as main effects and the Vagueness:Item interaction in the same model:
+Considering the first set of models:
 
@@
 
-## The models conducted to test for main effects _in the presence of interactions involving those main effects_ (after Levy, 2018)
+## The second set of models
+
+The second set of models considered were conducted to test for main effects _in the presence of interactions involving those main effects_ (after Levy, 2018)
 

### 2018-08-12 experiments/E_exp_3/writeup/E_exp_3.Rmd
D and E have hypotheses and results for the first set of models but not the levy models

@@
 
+We formulated the following hypotheses:
+
+* [Hypothesis 1] Crisp/Vague RT:
+    * Vague instructions should result in faster responses than crisp instructions.
+* [Hypothesis 2] Comparison/Matching RT:
+    * Instructions that allow comparison should result in faster responses than instructions that necessitate matching.
+* [Hypothesis 3] Vagueness effect differs between comparison and matching: 
+    * The effects of vagueness should differ when the selection task is comparison versus when it is matching.
+
 # Results
+
+An initial observation is that response times were much faster than in our previous experiments.
+It seems that when there are no numbers in the instructions, and the instructions refer to a previously-seen trial-initial array, participants are able to resolve those instructions an order of magnitude more quickly than when the instructions use numbers, and make no reference to any previously-seen trial-initial array.
+
+Response times were trimmed at 9,500 ms, resulting in the loss of 9 trials that were judged to have outlier RTs.
+The remaining data were trimmed separately for each participant at 2.5 standard deviations below and above that participant's mean RT, leading to the loss of 211 trials (2.8% of the remaining trials).
+RTs in milliseconds were log-transformed so that they better approximated a normal distribution for the purposes of analysis.
+
+## The first set of models
 
@@
 
+Considering the first set of models:
+
+* [Hypothesis 1] Crisp/Vague RT:
+    * _Vague instructions should result in faster responses than crisp instructions_ : Vague instructions resulted in faster responses than crisp instructions, but this effect was very small and not significant (p=0.13)
+* [Hypothesis 2] Comparison/Matching RT:
+    * _Instructions that allow comparison should result in faster responses than instructions that necessitate matching_ : Comparison instructions resulted in faster responses than matching and this difference was significant (p<0.001)
+* [Hypothesis 3] Vagueness effect differs between comparison and matching:
+    * _The effects of vagueness should differ when the selection task is comparison versus when it is matching_ : Vague instructions did have different effects depending on whether the selection task is comprison or matching (p<0.001).
+In the comparison condition vague instructions resulted in significantly faster responses (p<0.001).
+In the selection condition vague instructions slowed down responses significantly (p<0.001)
+
+## The second set of models
+
+The second set of models considered were conducted to test for main effects _in the presence of interactions involving those main effects_ (after Levy, 2018)
+
 # Discussion

### 2018-08-17 experiments/D_exp_2/writeup/D_exp_2.Rmd
trying roger in D_exp_2

@@
 
+```{r}
+dat$Condition <- as.factor(paste(dat$Item, dat$Vagueness))
+ggplot(data=dat, mapping=aes(x=Item, y=RT_log, group=Vagueness)) +
+  theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), aspect.ratio = 1) +
+  facet_wrap(~Selection) +
+  stat_summary(mapping=aes(y=RT_log, x=Item, group=Vagueness, colour=Vagueness), fun.y=mean, geom="line", lwd=2) +
+  geom_smooth(mapping=aes(y=RT_log, x=Item, group=Vagueness, colour=Vagueness), method="loess")
+```
+
 ```{r, "EXP_D_Borderline_response_distribution_plot", fig.width=7, fig.height=3.5, echo=FALSE}
@@
 
-The second set of models considered were conducted to test for main effects _in the presence of interactions involving those main effects_ (after Levy, 2018)
+The second set of models considered were conducted to test for main effects _in the presence of interactions involving those main effects_ (after Levy, 2018).
+
+Testing for a main effect of X when X is a two-level factor and the model also includes a main effect of a three-level factor  and the interaction between X and Y.
+
+Testing for a main effect of Vagueness when Vagueness is a two-level factor and the model also includes a main effect of a four-level factor Item and the interaction between Vagueness and Item.
+
+```{r}
+Item.numeric <- sapply(dat$Item, function(i) contr.sum(4)[i,])
+dat$Item1 <- Item.numeric[1,]
+dat$Item2 <- Item.numeric[2,]
+dat$Item3 <- Item.numeric[3,]
+m0 <- lm(RT_log ~ 
+           Item1 + Vagueness:Item1 + 
+           Item2 + Vagueness:Item2 + 
+           Item3 + Vagueness:Item3 +
+           Selection +
+           Selection:Vagueness, 
+         dat)
+contrasts(dat$Item) <- contr.sum(4)
+m.test <- lm(RT_log ~ 
+               Item +
+           Vagueness:Item1 + 
+           Vagueness:Item2 + 
+           Vagueness:Item3 +
+           Selection +
+           Selection:Vagueness, 
+         dat)
+m1 <- lm(RT_log ~ 
+           Item1 + Vagueness:Item1 + 
+           Item2 + Vagueness:Item2 + 
+           Item3 + Vagueness:Item3 +
+           Selection +
+           Selection:Vagueness +
+           Vagueness,
+         dat)
+m100 <- lm(RT_log ~
+             Item +
+             Vagueness +
+             Item:Vagueness+
+             Selection+
+             Selection:Vagueness,
+           dat)
+#m1 and m100 are the same model
+anova(m0,m1,test="F")
+```
+
+```{r}
+Item.numeric <- sapply(dat$Item, function(i) contr.sum(4)[i,])
+dat$Item1 <- Item.numeric[1,]
+dat$Item2 <- Item.numeric[2,]
+dat$Item3 <- Item.numeric[3,]
+m3 <- lm(RT_log ~ 
+           Item1 + 
+           Item2 +  
+           Item3 , 
+         dat)
+contrasts(dat$Item) <- contr.sum(4)
+m4 <- lm(RT_log ~
+           Item,
+         dat)
+# m3 and m4 are the same model
+```
+
+```{r}
+Item.numeric <- sapply(dat$Item, function(i) contr.sum(4)[i,])
+dat$Item1 <- Item.numeric[1,]
+dat$Item2 <- Item.numeric[2,]
+dat$Item3 <- Item.numeric[3,]
+
+m.full <- lm(RT_log ~
+             Item +
+             Vagueness +
+             Selection + 
+             Item:Vagueness +
+             Item:Selection + 
+             Vagueness:Selection,
+           dat)
+m.reduced <- lm(RT_log ~
+                Item1 + 
+                Item2 + 
+                Item3 + 
+                Vagueness:Item1 + 
+                Vagueness:Item2 +
+                Vagueness:Item3 +
+                Vagueness:Selection +
+                Selection +
+                Item:Selection
+                ,dat  )
+anova(m.reduced, m.full, test="F")
+```
 

### 2018-08-18 springer/review/vagueness-springer-review.tex
numbering off review points

@@
 This is a very interesting paper on an important topic.
-My main critical comment would be that it is unclear what the inclusion of experiment 1 adds to the paper.
+
+\section{Main points}
+
+\begin{enumerate}
+
+\item My main critical comment would be that it is unclear what the inclusion of experiment 1 adds to the paper.
 The experiment seems to have been problematic, as described in the beginning of the presentation of experiment 2.
@@
 The results from the second experiment seem to be much more solid and theoretically important.
-\begin{itemize}
-\item Perhaps the paper could be revised to present and interpret\marginnote{I agree, I think we should remove this experiment and present three experiments only} the second experiment only, treating the first as a pilot that revealed some issues that were resolved by the main experiment?
-\end{itemize}
+Perhaps the paper could be revised to present and interpret the second experiment only, treating the first as a pilot that revealed some issues that were resolved by the main experiment?
+\marginnote[-3cm]{1: I agree, I think we should remove this experiment and present three experiments only}
 
-A second point is that the discussion of experiment 2 presents some results as more unequivocal than they appear to be. I have some detailed comments below on this.
+\item A second point is that the discussion of experiment 2 presents some results as more unequivocal than they appear to be. I have some detailed comments below on this.
+\marginnote[]{Addressed in points x y and z below}
 
-A third question is whether the context-dependence of many vague expressions could be considered as a source of utility.
-E.g., \emph{tall} and \emph{warm} can adjust their meanings in ways that allow them to communicate more information about the location of some object relative to the statistical distribution of objects in a reference class.
+\item A third question is whether the context-dependence of many vague expressions could be considered as a source of utility; e.g., \emph{tall} and \emph{warm} can adjust their meanings in ways that allow them to communicate more information about the location of some object relative to the statistical distribution of objects in a reference class.
 "$1.83$ meters tall" and "$94$\textdegree C" don't have this ability.
@@
 \section*{Small comments}
-\begin{enumerate}
-\item
-While I see the value of presenting Lipman's example as such, the example is fairly strange for the purpose he wants to put it to, since it's not at all obvious that in the context described \emph{187.9 cm} would indeed be more conversationally useful than \emph{tall} as a description of a person's height.
+
+\item While I see the value of presenting Lipman's example as such, the example is fairly strange for the purpose he wants to put it to, since it's not at all obvious that in the context described \emph{187.9 cm} would indeed be more conversationally useful than \emph{tall} as a description of a person's height.
 I for one would have no idea how to pick out a person of that exact height from a crowd.
 Perhaps there is a more convincing modification of the example that would make the point, that we are sometimes inclined to use vague language even though a precise description would be uncontroversially more useful to the hearer?
-\item
-This sort of comes out in the discussion, but the experiment by Mishra et al.
+
+\item This sort of comes out in the discussion, but the experiment by Mishra et al.
 reported in section 2 seem to be of limited relevant since they confuse vagueness with lack of specificity.
@@
 So, this is an interesting experiment but I wonder if it is really worth discussing here.
-\item
-The Peters et al.
+
+\item The Peters et al.
 experiment: the comment \emph{it becomes doubtful whether any borderline cases could be conceived to arise in fact} -- indeed it's not obvious but I would be surprised if they didn't.
 When given natural-sounding descriptions like \emph{poor, fair, good, excellent} alongside hard-to-understand numerical values, I would guess that most participants would simply ignore the numerical values since they make less natural descriptions.
-\item
-Perhaps the data and analysis could be moved to a repo (e.g., github)? This way the URL remains stable even if someone moves institutions, if the university decides to change web addresses, etc. 
-\item
-Could you include a sample display for each experiment presented?
-\item
-Why did the statistical analysis include only random slopes and not random intercepts? It would be standard to use both.
-\item
-p.13: I wouldn't describe indefinites like \emph{a square with \ldots} as referring expressions
-\item
-exp 2 questions re presentation: 
-\begin{itemize}
-\item [(a)]
-why was the central number always presenting in the middle, instead of fully randomizing?
-\item [(b)]
-\emph{about 10} seems like a terrible description for both 6 and 15. My guess is this would cause subjects to boggle even if they eventually manage to make a guess.
-\item [(c)]
-what was the vague numeric instruction used in stimuli with larger dot arrays, e.g. (16,25,34)? [I assume it wasn't \emph{about 10}]
-\end{itemize}
-\item
-for the numeric condition, the significant main effect reported at the bottom of p.15 (testing H1) is misleading, since the effect seems to be driven entirely by the much lower response time for crisp \emph{6}.
+
+\item Perhaps the data and analysis could be moved to a repo (e.g., github)? This way the URL remains stable even if someone moves institutions, if the university decides to change web addresses, etc.
+
+\item Could you include a sample display for each experiment presented?
+
+\item Why did the statistical analysis include only random slopes and not random intercepts? It would be standard to use both.
+
+\item p.13 in the submitted version (in paragraph 4 of the intro to C-exp-1): I wouldn't describe indefinites like \emph{a square with \ldots} as referring expressions
+\marginnote[-1cm]{10: Kees?}
+
+\item re: exp 2 presentation 1/3: why was the central number always presenting in the middle, instead of fully randomizing?
+
+\item re: exp 2 presentation 2/3: \emph{about 10} seems like a terrible description for both 6 and 15. My guess is this would cause subjects to boggle even if they eventually manage to make a guess.
+
+\item re: exp 2 presentation 3/3: what was the vague numeric instruction used in stimuli with larger dot arrays, e.g. (16,25,34)? [I assume it wasn't \emph{about 10}]
+
+\item for the numeric condition, the significant main effect reported at the bottom of p.15 (testing H1) is misleading, since the effect seems to be driven entirely by the much lower response time for crisp \emph{6}.
 Everything else is nearly indistinguishable.
 So, it seems like the main thing to explain here is why crisp \emph{6} is so much faster than other numerical expressions -- presumably because it's a small number and there's no additional cost of trying to resolve the intention behind the fairly poor description \emph{about 10} for 6?
-\item
-perhaps the big main effect for instruction format: it seems like the interpretation of this result should be that numbers are in general harder to process than vague verbal expressions.
+
+\item perhaps the big main effect for instruction format: it seems like the interpretation of this result should be that numbers are in general harder to process than vague verbal expressions.
 This would seem to account for most of the variance in the response time data of experiment 2.
-\item
-could the differences in the numeric vs verbal borderline response distribution shown in the right panel of fig.
+
+\item could the differences in the numeric vs verbal borderline response distribution shown in the right panel of fig.
 2 be explained by stimulus design -- i.e., the hypothesis I mentioned earlier, that \emph{about 6} is just a poor description of both 6 and 15, with the result that participants were choosing more or less at random with a slight bias for the number a little bit closer to 10?
 This would explain why there is almost no difference between borderline and expected in the vague/numeric condition.
-\item
-p.17 \emph{we found a small (but statistically significant) disadvantage of vague instructions} -- again this is misleading since the effect is apparently driven almost entirely by a single condition (6, crisp) that was much faster than expected.
+
+\item p.17 \emph{we found a small (but statistically significant) disadvantage of vague instructions} -- again this is misleading since the effect is apparently driven almost entirely by a single condition (6, crisp) that was much faster than expected.
 There appear to be few differences otherwise \ldots one option would be to exclude the 6:15:24 condition and then re-run the statistical analysis to see if any other effects emerge.
-\item
-p.18 \emph{allowed participants to use a comparison "algorithm"} -- the work of Justin Halberda and collaborators on the verification strategies associated with \emph{most} and \emph{more than half} is highly relevant here (and indeed throughout the paper, given that \emph{most} is vague but otherwise fairly similar in meaning to \emph{more than half})
-\item
-p.19 Testing (H1), we have \emph{Vague instructions were non-significantly easier \ldots} do you mean \emph{not significantly easier}?
+
+\item p.18 \emph{allowed participants to use a comparison "algorithm"} -- the work of Justin Halberda and collaborators on the verification strategies associated with \emph{most} and \emph{more than half} is highly relevant here (and indeed throughout the paper, given that \emph{most} is vague but otherwise fairly similar in meaning to \emph{more than half})
+
+\item p.19 Testing (H1), we have \emph{Vague instructions were non-significantly easier \ldots} do you mean \emph{not significantly easier}?
 In any case from the graphs on the next page this seems misleading -- what it looks like is a substantial main effect where vague instructions have a main effect of lowering the response time, but this effect is masked by interactions with quantity, most notably (again) the huge speed-up for 6 in the matching/crisp condition.
 I'd recommend following the method described in Roger Levy's paper on finding main effects in the presence of interactions (\url{https://arxiv.org/abs/1405.2094}).
-\item
-p.20 \emph{the cost reduction account was wrong to predict significant main effect advantages for vagueness} -- this seems way too strong given the issues I've just described.
+
+\item p.20 \emph{the cost reduction account was wrong to predict significant main effect advantages for vagueness} -- this seems way too strong given the issues I've just described.
 First, the statistical analysis needs to be redone as I just mentioned.
@@
 At best, you can say \emph{we failed to find evidence that \ldots}
-\item
-p.21, table 3: were the matching/crisp and matching/vague conditions unmatched in the target quantity as the table says? Or is this a typo?
-\item
-p.22 \emph{main interaction effect} - ??
-\item
-p.22 \emph{In the matching task vagueness slowed response times} -- again this is not a good summary since there is a big interaction, and the apparent main efect may be driven by a single condition, here a slowdown for 6:15:24 in the matching/vague condition.
+
+\item p.21, table 3: were the matching/crisp and matching/vague conditions unmatched in the target quantity as the table says? Or is this a typo?
+
+\item p.22 \emph{main interaction effect} - ??
+
+\item p.22 \emph{In the matching task vagueness slowed response times} -- again this is not a good summary since there is a big interaction, and the apparent main efect may be driven by a single condition, here a slowdown for 6:15:24 in the matching/vague condition.
 Re-running using Levy's method would be advisable here.
-\item
-Typo: Paul Egr\a'e's name has an acute accent on the second \emph{e}
+
+\item Typo: Paul Egr\a'e's name has an acute accent on the second \emph{e}
 \marginnote{corrected in the .bib}
+
 \end{enumerate}
+
 \end{document}

### 2018-08-18 springer/current/e1.tex
added a full table of instructiosn for D-Exp-2

@@
 
+\marginnote{C\_Exp\_1}
+
 To find out what happens when words are used in a context where their potential for vagueness comes to the fore, Experiment 2 used three arrays (rather than two) so that the vague description had more than one possible referent, and used indefinite articles to avoid the impression that only one response counted as correct, and was carried out without error feedback.
@@
 
-Examples of crisp and vague versions of the numerical and verbal instructions follow: the examples assume the array (6,15,24) and reference to the smaller number of dots, such that 6 was classified as the expected response; 15 as the borderline response; and 24 as the incorrect response.
+Examples of crisp and vague versions of the numerical and verbal instructions are given next: the examples assume the array (6,15,24) and reference to the smaller number of dots, such that 6 was classified as the expected response; 15 as the borderline response; and 24 as the incorrect response.
 In the {\em vague numerical} condition we used \emph{Choose a square with about 10 dots}.

### 2018-08-18 springer/current/e2.tex
added a placeholder citation for Levy article on main effects and interactions
added a full table of instructiosn for D-Exp-2

@@
-The main aim of Experiment 3 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies.
+\marginnote{D\_Exp\_2}
+
+The main aim of Experiment 2 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies.
 The main changes from Experiment 2 were that the human selection task was explicitly controlled (i.e., in whether it amounted to matching or comparison), and that all conditions were constrained to mention a number.
 We used the same arrays as in Experiment 2.
-Table \ref{Instructions for e3} shows the instructions for each condition.
+Table \ref{Instructions for e2} shows the instructions for each condition.
 Note the difference between ``fewer than 20'' and ``far fewer than 20'': whereas the former cannot have borderline cases (i.e., for each number it is clear whether the number is smaller than 20 or not), the latter can.
@@
 \centering
-\caption{Experiment 3: Instructions arranged by condition for stimuli with $(6,15,24)$ dots and instructions indicating the smaller numbers} 
-\label{Instructions for e3}
-\begin{tabular}{cccc}
+\caption{Experiment 2: Instructions arranged by condition. The instructions given in the table started with ``Choose a square with \ldots"} 
+\label{Instructions for e2}
+\begin{tabular}{cccll}
 \hline\noalign{\smallskip}
-Instruction	format          &Vagueness					&	Selection task	& Instruction									\\
+Item & Quantity & Selection & Crisp & Vague \\ 
 \noalign{\smallskip}\hline\noalign{\smallskip}
-\multirow{4}{*}{numeric}	&\multirow{2}{*}{crisp} 	&	matching		& Choose a square with 6 dots 					\\ 
-							&							&	comparison 		& Choose a square with fewer than 20 dots 		\\
-                             \cline{2-4}
-							&\multirow{2}{*}{vague} 	&	matching 		& Choose a square with about 10 dots 			\\ 
-							&							&	comparison 		& Choose a square with far fewer than 20 dots 	\\ 
+06:15:24 & Small & Comparison & fewer than 20 dots & far fewer than 20 dots \\ 
+06:15:24 & Small & Matching & 6 dots & about 10 dots \\ 
+06:15:24 & Large & Comparison & more than 10 dots & far more than 10 dots \\ 
+06:15:24 & Large & Matching & 24 dots & about 20 dots \\ 
+\noalign{\smallskip}\hline\noalign{\smallskip}
+16:25:34 & Small & Comparison & fewer than 30 dots & far fewer than 30 dots \\ 
+16:25:34 & Small & Matching & 16 dots & about 20 dots \\ 
+16:25:34 & Large & Comparison & more than 20 dots & far more than 20 dots \\ 
+16:25:34 & Large & Matching & 34 dots & about 30 dots \\ 
+\noalign{\smallskip}\hline\noalign{\smallskip}
+26:35:44 & Small & Comparison & fewer than 40 dots & far fewer than 40 dots \\ 
+26:35:44 & Small & Matching & 26 dots & about 30 dots \\ 
+26:35:44 & Large & Comparison & more than 30 dots & far more than 30 dots \\ 
+26:35:44 & Large & Matching & 44 dots & about 40 dots \\ 
+\noalign{\smallskip}\hline\noalign{\smallskip}
+36:45:54 & Small & Comparison & fewer than 50 dots & far fewer than 50 dots \\ 
+36:45:54 & Small & Matching & 36 dots & about 40 dots \\ 
+36:45:54 & Large & Comparison & more than 40 dots & far more than 40 dots \\ 
+36:45:54 & Large & Matching & 54 dots & about 50 dots \\ 
 \noalign{\smallskip}\hline
 \end{tabular}
-\end{table}%
+\end{table}
 
-\subsection{Hypotheses} % E 3
-\noindent For Experiment 3, we hypothesised:
+\subsection{Hypotheses} % D_Exp_2
+\noindent For Experiment 2, we hypothesised:
 
@@
 
-\subsection{Experimental Method} % E 3
+\subsection{Experimental Method} % D_Exp_2
 
-38 participants were recruited. We used a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e3}).
+38 participants were recruited. We used a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e2}).
 On each trial an instruction was presented: participants pressed a key to dismiss the instruction, when the dot arrays were presented until the participant responded, and the response time and choice were recorded.
 
-\subsection{Results} % E 3
+\subsection{Results} % D_Exp_2
+
+We used the method proposed by \citet{Levy:MainEffectsInteractions} to test for the main effect of Vagueness in the presence of its higher-order interactions.
 
@@
 \includegraphics[width=\textwidth]{figures/e3-rtplot-1.pdf}
-\caption{Mean response times by condition for Experiment 3 where all instructions were numeric}
+\caption{Mean response times by condition for Experiment 2 where all instructions were numeric}
 \label{resultse3}

### 2018-08-18 experiments/D_exp_2/writeup/D_exp_2.Rmd
recompiled D-exp-2

@@
     spread(key=Vagueness, value=Instruction)
-# kable(instructions_table, format="latex", booktabs = TRUE)
 instructions_table %>% 

### 2018-08-19 springer/current/e1.tex
finished renumbering

@@
-%However, the picture painted by these findings might be misleading.
-First of all, there is a possibly confounding factor.
-Contrast, from Experiment 1, an expression from the vague condition: `the square with few dots' with an expression from the crisp condition: `the square with 5 dots'.
-One difference is that `few' has the potential for vagueness, whereas `5' is crisp.
-But another difference is that `few' is verbal while `5' is numerical, in the sense that a number is mentioned explicitly.
-Since these two differences could not be separated in Experiment 1, the vagueness advantage finding is vulnerable to an alternative interpretation, that what we saw as a vagueness advantage was in contrast an advantage for the verbal form of the quantifier.
-In Experiment 2 we therefore created verbal and numeric versions of each of the vague and crisp instructions so that we could compare vague and crisp conditions while taking account of verbal / numeric format.
+\marginnote{C-Exp-1}
 
-%Another potential problem with Experiment 1 is the following.
-Participants chose one of two squares: therefore the `vague' quantifiers (e.g., `few') uniquely identified one square.
-Recall our definition of vague  -- ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise''.
-In Experiment 1, the quantifiers in the vague conditions did not realise their potential for vagueness.
-This is because there were no borderline cases of the referent that could make the referent set `not well-defined', and perhaps because using definite articles in the instructions implied that only one option was correct.
-Using error feedback in Experiment 1 could have exacerbated this.
+To find out what happens when words are used in a context where their potential for vagueness comes to the fore, Experiment 1 used three arrays (rather than two) so that the vague description had more than one possible referent, and used indefinite articles to avoid the impression that only one response counted as correct, and was carried out without error feedback.
+An indication that the potential for vagueness was indeed realised in Experiment 1 is that the borderline response was chosen fairly often: 16\% of the time.
 
-\marginnote{C\_Exp\_1}
-
-To find out what happens when words are used in a context where their potential for vagueness comes to the fore, Experiment 2 used three arrays (rather than two) so that the vague description had more than one possible referent, and used indefinite articles to avoid the impression that only one response counted as correct, and was carried out without error feedback.
-An indication that the potential for vagueness was realised in Experiment 2 is that the borderline response was chosen fairly often: 16\% of the time.
-
-In Experiment 2, an item was a referring expression instruction followed by a set of three dot arrays defined by a triple of numbers, representing the number of dots in the left, middle, and right arrays.
+In Experiment 1, an item was a referring expression instruction followed by a set of three dot arrays defined by a triple of numbers, representing the number of dots in the left, middle, and right arrays.
 We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54).
 Each set of arrays comprised three arrays (instead of two as in Experiment 1); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had more.
+An example stimulus is given in Figure \ref{Experiment1and2examplestimulus}.
 
@@
 \caption{Experiment 1 instructions arranged by condition. The instructions given in the table started with ``Choose \ldots"}
-\label{instructionse2}
+\label{instructionsC-exp-1}
 \begin{tabular}{cccll}
@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.75\textwidth]{figures/Ce1-example-screenshot}
+\caption{Experiment 1 and 2 example stimulus}
+\label{Experiment1and2examplestimulus}
+\end{figure}
+
 \subsection{Hypotheses}
 
-We formulated the following hypotheses for Experiment 2:
+We formulated the following hypotheses for Experiment 1:
 
@@
 
-Means for response times and proportion of borderline responses are given in Fig. (\ref{resultse2}). 
+Means for response times and proportion of borderline responses are given in Fig.
+(\ref{resultsC-exp-1}).
 Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data.
-A linear mixed model was constructed for the (logged) response times, 
-with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects. We found:
+A linear mixed model was constructed for the (logged) response times, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
+We found:
 
@@
 A generalized linear mixed model \citet{jaeger2008categorical} was fit to the data for selection of the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
-The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{resultse2}.
+The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{resultsC-exp-1}.
 Participants chose the borderline square on $16.6\%$ of trials overall.
@@
 \centering
-\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{figures/e2-rtplot-1.pdf}
-\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{figures/e2-blBarChart-1}
-\caption{Experiment 2 results: mean response times by condition and item, and counts of borderline case responses by condition.}
-\label{resultse2}
+\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{figures/Ce1-rtplot-1.pdf}
+\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{figures/Ce1-blBarChart-1}
+\caption{Experiment 1 results: mean response times by condition and item, and counts of borderline case responses by condition.}
+\label{resultsC-exp-1}
 \end{figure}
@@
 
-Experiment 2 tested to see whether when borderline cases are present, vague instructions would speed responses as they did in Experiment 1 when there were no borderline squares. 
+Experiment 1 tested to see whether when borderline cases are present, vague instructions would speed responses as they did in Experiment pilot exp 2 when there were no borderline squares.
+
 We actually found a small (but statistically significant) \emph{disadvantage} of vague instructions: vague instructions slowed people down by 112 ms on average.
 We also found that the effect of instruction format was significant, with numerical format slowing responses by 689 ms on average, such that the disadvantage of numerical format overwhelmed the contribution of vagueness.
-The \emph{verbal vague} condition was still responded to faster than the \emph{numerical crisp} condition, so the pattern from Experiment 1 was reproduced, but in the light of the evidence from Experiment 2, in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like an advantage of verbal instruction format.
+The \emph{verbal vague} condition was still responded to faster than the \emph{numerical crisp} condition, so the pattern from Experiment pilot exp 2 was reproduced, but in the light of the evidence from Experiment 1, in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like an advantage of verbal instruction format.
 
 However, once again there is a possibly confounding factor.
-Observe that, in Experiment 2, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) {\bf selection algorithm}: To see this, consider the task of selecting the dot array that contains ``few dots'': to do this, it suffices to {\em compare} the three arrays and select the one that contains the fewest elements.
+Observe that, in Experiment 1, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) {\bf selection algorithm}: To see this, consider the task of selecting the dot array that contains ``few dots'': to do this, it suffices to {\em compare} the three arrays and select the one that contains the fewest elements.
 To select the dot array that contains ``16 dots'' seems to require the participant to estimate, and then {\em match}, the cardinality of (at least) one dot array to 16, a process which could plausibly take longer, independently of vagueness.
@@
 In the next two experiments we pitted the comparison algorithm and matching algorithm selection tasks against each other while controlling vagueness and instruction format.
-In Experiment 3 we restricted all the instructions to \emph{numeric} quantifiers while factorially manipulating vagueness and selection task.
-In Experiment 4 we ensured that all instructions used \emph{verbal} quantifiers, while also factorially manipulating vagueness and selection task.
+In Experiment 2 we restricted all the instructions to \emph{numeric} quantifiers while factorially manipulating vagueness and selection task.
+In Experiment 3 we ensured that all instructions used \emph{verbal} quantifiers, while also factorially manipulating vagueness and selection task.
 This allowed us to distinguish between the predictions of the selection task account and the instruction format account.

### 2018-08-19 springer/current/e2.tex
finished renumbering

@@
-\marginnote{D\_Exp\_2}
+\marginnote{D-Exp-2}
 
 The main aim of Experiment 2 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies.
-The main changes from Experiment 2 were that the human selection task was explicitly controlled (i.e., in whether it amounted to matching or comparison), and that all conditions were constrained to mention a number.
-We used the same arrays as in Experiment 2.
+The main changes from Experiment 1 were that the human selection task was explicitly controlled (i.e., in whether it amounted to matching or comparison), and that all conditions were constrained to mention a number.
+We used the same arrays as in Experiment 1 (an example stimulus is given in Figure \ref{Experiment1and2examplestimulus}).
 Table \ref{Instructions for e2} shows the instructions for each condition.
@@
 
-\subsection{Hypotheses} % D_Exp_2
-\noindent For Experiment 2, we hypothesised:
+\subsection{Hypotheses}
+
+For Experiment 2, we hypothesised:
 
@@
 
-\subsection{Experimental Method} % D_Exp_2
+\subsection{Experimental Method}
 
-38 participants were recruited. We used a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e2}).
-On each trial an instruction was presented: participants pressed a key to dismiss the instruction, when the dot arrays were presented until the participant responded, and the response time and choice were recorded.
+38 participants were recruited.
+We used a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e2}).
+On each trial an instruction was presented: participants pressed a key to dismiss the instruction, at which time the dot arrays were presented until the participant responded, and the response time and choice were recorded.
 
-\subsection{Results} % D_Exp_2
+\subsection{Results}
 
@@
 Response times were trimmed at 2.5 SD separately for each subject, leading to the loss of 204 trials (2.8\% of the trials).
-Condition means for the remaining (logged) RTs are plotted in Figure \ref{resultse3}.
-A linear mixed model was constructed for the logged response times, 
-with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
+Condition means for the remaining (logged) RTs are plotted in Figure \ref{resultsD-exp-2}.
+A linear mixed model was constructed for the logged response times, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
 
@@
 \centering
-\includegraphics[width=\textwidth]{figures/e3-rtplot-1.pdf}
+\includegraphics[width=\textwidth]{figures/De2-rtplot-1.pdf}
 \caption{Mean response times by condition for Experiment 2 where all instructions were numeric}
-\label{resultse3}
+\label{resultsD-exp-2}
 \end{figure}

### 2018-08-19 springer/current/e3.tex
added a full table of instructiosn for D-Exp-2
finished renumbering

@@
-This experiment mirrors Experiment 3, but focussing on instructions that did not use a number. 
+\marginnote{E-Exp-3}
+
+This experiment mirrors Experiment 2, but focussing on instructions that did not use a number.
 We manipulated vagueness and the selection task (comparison and matching).
-In order to implement the experiment without mentioning numbers in the instructions, we changed the sequence of each trial to include a `target' (i.e., a dot array of a particular cardinality) before the instruction, so that we could then refer back to the target's cardinality in the instruction using expressions like \emph{the same number of dots as the target}; \emph{fewer dots than the target}.
+In order to implement the experiment without mentioning numbers in the instructions, we changed the sequence of each trial to include a `target' (i.e., a dot array of a particular cardinality) before the array, so that we could then refer to the target's cardinality in the instruction using expressions like \emph{the same number of dots as the target}; \emph{fewer dots than the target}.
+An example of this sequence is given in Figure \ref{Experiment3examplestimulus}.
 This presentation of a target before the main body of the trial shares some features with \citeauthor[Experiment 2]{Izard20081221}, although in that experiment participants were told the cardinality of the target (called an \emph{inducer} in that paper) whereas in our experiment we did not tell participants the cardinality of the prime array.
-An item was thus a combination of a target dot array, an instruction that did not contain a number, and a set of dot arrays taking their cardinalities from the same triples used in Experiment 2.
-Table \ref{Instructions for e4} spells out how the instructions were constrained not to mention a numeral and gives examples of targets.
+An item was thus a combination of a target dot array, an instruction that did not contain a number, and a set of dot arrays taking their cardinalities from the same triples used in Experiments 1 and 2.
+Table \ref{instructionsE-exp-3} spells out how the instructions were constrained not to mention a numeral and gives examples of targets.
 
-\begin{table}[htp]
-\caption{Instructions and targets by condition for Experiment 4}
-\begin{center}
-\begin{tabular}{llcrl}
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.75\textwidth]{figures/Ee3-flow.jpg}
+\caption{Experiment 3 example stimulus}
+\label{Experiment3examplestimulus}
+\end{figure}
+
+\begin{table}
+\centering
+\caption{Experiment 3 instructions arranged by condition. The instructions given in the table started with ``Choose a square with \ldots"}
+\label{instructionsE-exp-3}
+\resizebox{\textwidth}{!}{\begin{tabular}{ccccll}
 \hline\noalign{\smallskip}
-Item					    &Selection task					& Vagueness	& Target		&Choose a square with\ldots							\\
+Item & Target & Quantity & Selection & Crisp & Vague \\ 
 \noalign{\smallskip}\hline\noalign{\smallskip}
-\multirow{4}{*}{06:15:24} 	&\multirow{2}{*}{Comparison} 	& Crisp		& 20			&\ldots fewer dots than the target					\\
-\cline{3-5}
-							&								& Vague		& 20			&\ldots far fewer dots than the target				\\
-\cline{2-5}
-							&\multirow{2}{*}{Matching}		& Crisp		& 6			&\ldots the same number of dots as the target		    \\
-\cline{3-5}
-							& 		 						& Vague		& 10 		&\ldots about the same number of dots as the target	    \\
+06:15:24 &   6 & Small & Matching 	& the same number of dots as the target & NA \\ 
+06:15:24 &  10 & Small & Matching 	& NA & about the same number of dots as the target \\ 
+06:15:24 &  10 & Large & Comparison & more dots than the target & far more dots than the target \\ 
+06:15:24 &  20 & Small & Comparison & fewer dots than the target & far fewer dots than the target \\ 
+06:15:24 &  20 & Large & Matching 	& NA & about the same number of dots as the target \\ 
+06:15:24 &  24 & Large & Matching 	& the same number of dots as the target & NA \\ 
+\noalign{\smallskip}\hline\noalign{\smallskip}
+16:25:34 &  16 & Small & Matching 	& the same number of dots as the target & NA \\ 
+16:25:34 &  20 & Small & Matching 	& NA & about the same number of dots as the target \\ 
+16:25:34 &  20 & Large & Comparison & more dots than the target & far more dots than the target \\ 
+16:25:34 &  30 & Small & Comparison & fewer dots than the target & far fewer dots than the target \\ 
+16:25:34 &  30 & Large & Matching 	& NA & about the same number of dots as the target \\ 
+16:25:34 &  34 & Large & Matching 	& the same number of dots as the target & NA \\ 
+\noalign{\smallskip}\hline\noalign{\smallskip}
+26:35:44 &  26 & Small & Matching 	& the same number of dots as the target & NA \\ 
+26:35:44 &  30 & Small & Matching 	& NA & about the same number of dots as the target \\ 
+26:35:44 &  30 & Large & Comparison & more dots than the target & far more dots than the target \\ 
+26:35:44 &  40 & Small & Comparison & fewer dots than the target & far fewer dots than the target \\ 
+26:35:44 &  40 & Large & Matching 	& NA & about the same number of dots as the target \\ 
+26:35:44 &  44 & Large & Matching 	& the same number of dots as the target & NA \\ 
+\noalign{\smallskip}\hline\noalign{\smallskip}
+36:45:54 &  36 & Small & Matching 	& the same number of dots as the target & NA \\ 
+36:45:54 &  40 & Small & Matching 	& NA & about the same number of dots as the target \\ 
+36:45:54 &  40 & Large & Comparison & more dots than the target & far more dots than the target \\ 
+36:45:54 &  50 & Small & Comparison & fewer dots than the target & far fewer dots than the target \\ 
+36:45:54 &  50 & Large & Matching 	& NA & about the same number of dots as the target \\ 
+36:45:54 &  54 & Large & Matching 	& the same number of dots as the target & NA \\ 
 \noalign{\smallskip}\hline
-\end{tabular}
-\end{center}
-\label{Instructions for e4}
-\end{table}%
+\end{tabular}}
+\end{table}
 
-\subsection{Hypotheses}% E 4
-\noindent For Experiment 4, we hypothesised:
+\subsection{Hypotheses}
+
+For Experiment 3, we hypothesised:
 
@@
 
-\subsection{Results}% E 4
+\subsection{Results}
 40 volunteers participated.
-The results showed that vagueness was beneficial for comparison but detrimental for matching (the same as Experiment 3) even when no numbers were allowed in the instructions. 
-Figure \ref{resultse4} shows the means by condition. Our findings were as follows:
+The results showed that vagueness was beneficial for comparison but detrimental for matching (the same as Experiment 2) even when no numbers were allowed in the instructions.
+Figure \ref{resultsE-exp-3} shows the means by condition.
+Our findings were as follows:
 
@@
 \centering
-\includegraphics[width=\textwidth]{figures/e4-rtplot-1.pdf}
-\caption{Mean response times by condition for Experiment 4}
-\label{resultse4}
+\includegraphics[width=\textwidth]{figures/Ee3-rtplot-1.pdf}
+\caption{Mean response times by condition for Experiment 3}
+\label{resultsE-exp-3}
 \end{figure}

### 2018-08-20 springer/review/vagueness-springer-review.tex
review is now a canvas for replies
added experiment breaks in the review

@@
 Perhaps the paper could be revised to present and interpret the second experiment only, treating the first as a pilot that revealed some issues that were resolved by the main experiment?
-\marginnote[-3cm]{1: I agree, I think we should remove this experiment and present three experiments only}
+\marginnote[-3cm]{1: Done \ldots}
+\begin{quote}What was E1 has been cut out and replaced by a couple of paragraphs on our 2013 cogsci paper\end{quote}
 
 \item A second point is that the discussion of experiment 2 presents some results as more unequivocal than they appear to be. I have some detailed comments below on this.
-\marginnote[]{Addressed in points x y and z below}
+\marginnote[-.5cm]{2: Addressed in points x y and z below}
+\begin{quote}\ldots\end{quote}
 
@@
 I've often thought that this flexibility is one of the main sources of utility for vague expressions with this property, since it allows our lexicon to be fairly expressive without having an enormous number of items.
+\marginnote[-3cm]{3: \ldots}
+\begin{quote}\ldots\end{quote}
 
-\section*{Small comments}
+\end{enumerate}
+
+\section{Small comments}
+
+\begin{enumerate}\setcounter{enumi}{3}
 
@@
 Perhaps there is a more convincing modification of the example that would make the point, that we are sometimes inclined to use vague language even though a precise description would be uncontroversially more useful to the hearer?
+\marginnote[-1.6cm]{4: \ldots}
+\begin{quote}\ldots\end{quote}
 
@@
 So, this is an interesting experiment but I wonder if it is really worth discussing here.
+\marginnote[-1.5cm]{5: \ldots}
+\begin{quote}\ldots\end{quote}
 
@@
 When given natural-sounding descriptions like \emph{poor, fair, good, excellent} alongside hard-to-understand numerical values, I would guess that most participants would simply ignore the numerical values since they make less natural descriptions.
+\marginnote[-1.5cm]{6: \ldots}
+\begin{quote}\ldots\end{quote}
 
-\item Perhaps the data and analysis could be moved to a repo (e.g., github)? This way the URL remains stable even if someone moves institutions, if the university decides to change web addresses, etc.
+\item Perhaps the data and analysis could be moved to a repo (e.g., github)?
+This way the URL remains stable even if someone moves institutions, if the university decides to change web addresses, etc.
+\marginnote[-.5cm]{7: Done}
+\begin{quote}https://github.com/mjgreen/vagueness\end{quote}
 
 \item Could you include a sample display for each experiment presented?
+\marginnote[-.5cm]{8: \ldots}
+\begin{quote}\ldots\end{quote}
+
+\section{C-exp-1}
 
 \item Why did the statistical analysis include only random slopes and not random intercepts? It would be standard to use both.
+\marginnote[-.5cm]{9: \ldots}
+\begin{quote}\ldots\end{quote}
 
@@
 \marginnote[-1cm]{10: Kees?}
+\begin{quote}\ldots\end{quote}
 
 \item re: exp 2 presentation 1/3: why was the central number always presenting in the middle, instead of fully randomizing?
+\marginnote[-0.25cm]{11: \ldots}
+\begin{quote}\ldots\end{quote}
 
 \item re: exp 2 presentation 2/3: \emph{about 10} seems like a terrible description for both 6 and 15. My guess is this would cause subjects to boggle even if they eventually manage to make a guess.
+\marginnote[-0.5cm]{12: \ldots}
+\begin{quote}\ldots\end{quote}
 
 \item re: exp 2 presentation 3/3: what was the vague numeric instruction used in stimuli with larger dot arrays, e.g. (16,25,34)? [I assume it wasn't \emph{about 10}]
+\marginnote[-0.5cm]{13: \ldots}
+\begin{quote}\ldots\end{quote}
 
@@
 So, it seems like the main thing to explain here is why crisp \emph{6} is so much faster than other numerical expressions -- presumably because it's a small number and there's no additional cost of trying to resolve the intention behind the fairly poor description \emph{about 10} for 6?
+\marginnote[-1.5cm]{14: \ldots}
+\begin{quote}\ldots\end{quote}
 
@@
 This would seem to account for most of the variance in the response time data of experiment 2.
+\marginnote[-1.5cm]{15: \ldots}
+\begin{quote}\ldots\end{quote}
 
@@
 This would explain why there is almost no difference between borderline and expected in the vague/numeric condition.
+\marginnote[-1.5cm]{16: \ldots}
+\begin{quote}\ldots\end{quote}
 
@@
 There appear to be few differences otherwise \ldots one option would be to exclude the 6:15:24 condition and then re-run the statistical analysis to see if any other effects emerge.
+\marginnote[-1.5cm]{17: \ldots}
+\begin{quote}\ldots\end{quote}
 
 \item p.18 \emph{allowed participants to use a comparison "algorithm"} -- the work of Justin Halberda and collaborators on the verification strategies associated with \emph{most} and \emph{more than half} is highly relevant here (and indeed throughout the paper, given that \emph{most} is vague but otherwise fairly similar in meaning to \emph{more than half})
+\marginnote[-1.5cm]{18: \ldots}
+\begin{quote}\ldots\end{quote}
+
+\section{D-exp-2}
 
@@
 I'd recommend following the method described in Roger Levy's paper on finding main effects in the presence of interactions (\url{https://arxiv.org/abs/1405.2094}).
+\marginnote[-1.5cm]{19: \ldots}
+\begin{quote}\ldots\end{quote}
 
@@
 At best, you can say \emph{we failed to find evidence that \ldots}
+\marginnote[-1.5cm]{20: \ldots}
+\begin{quote}\ldots\end{quote}
+
+\section{D-exp-3}
 
 \item p.21, table 3: were the matching/crisp and matching/vague conditions unmatched in the target quantity as the table says? Or is this a typo?
+\marginnote[-1cm]{21: \ldots}
+\begin{quote}\ldots\end{quote}
 
 \item p.22 \emph{main interaction effect} - ??
+\marginnote[-0cm]{22: \ldots}
+\begin{quote}\ldots\end{quote}
 
@@
 Re-running using Levy's method would be advisable here.
+\marginnote[-1.5cm]{23: \ldots}
+\begin{quote}\ldots\end{quote}
 
 \item Typo: Paul Egr\a'e's name has an acute accent on the second \emph{e}
-\marginnote{corrected in the .bib}
+\marginnote{24: corrected in the .bib}
+\begin{quote}\ldots\end{quote}
 

### 2018-08-20 experiments/D_exp_2/writeup/D_exp_2.Rmd
dexp2 now has functioning levy i think

--- 2018-08-12 experiments/D_exp_2/writeup/D_exp_2.Rmd
@@
 
-The instructions for this experiment were similar to C_exp_2, except that we changed "fewer than" to "far fewer than": this was because "far fewer than" can have borderline cases whereas "fewer than" cannot have borderline cases.
+We changed "fewer than" (crisp) to "far fewer than" (in vague): this was because "far fewer than" can have borderline cases whereas "fewer than" cannot have borderline cases.
 
@@
 
-```{r}
+```{r, cho=FALSE}
 dat_plot <- summarySEwithin2(dat, measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
@@
 ```
-
 ```{r, "EXP_D_Borderline_response_distribution_plot", fig.width=7, fig.height=3.5, echo=FALSE}
@@
 
-The second set of models considered were conducted to test for main effects _in the presence of interactions involving those main effects_ (after Levy, 2018)
+The second set of models considered were conducted to test for main effects _in the presence of interactions involving those main effects_ (after Levy, 2018).
 
-# Discussion
+The variables are:
+Vagueness 2 levels (X) sumcoded -.5, +.5 
+Item 4 levels (Y) 
+Selection 2 levels (Z)
+
+Note that the null-hypothsesis model _includes_ random slopes for the main effect of X; Y; and Z; as well as for their interactions -- including a seperate random slope for all combiations of X, Y, and Z achieves this.
+
+```{r}
+dat$XYZ <- with(dat, factor(paste(Vagueness,Item,Selection)))
+```
+
+Convert a K level factor Y into a sum-coding representation.
+
+```{r}
+Item.numeric <- sapply(dat$Item,function(i) contr.sum(4)[i,])
+dat$Item1 <- Item.numeric[1,]
+dat$Item2 <- Item.numeric[2,]
+dat$Item3 <- Item.numeric[3,]
+```
+
+```{r}
+dat_model <- dat
+Item.numeric <- sapply(dat_model$Item,function(i) contr.sum(4)[i,])
+dat_model$Item1 <- Item.numeric[1,]
+dat_model$Item2 <- Item.numeric[2,]
+dat_model$Item3 <- Item.numeric[3,]
+dat_model$c_Sel <- ifelse(dat_model$Selection=="Comparison", -0.5, 0.5)
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Sel <- ifelse(dat_model$Selection=="Comparison", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
+
+null.model <- lmer(RT_log ~ 
+                     Item1 + c_Vag:Item1 +
+                     Item2 + c_Vag:Item2 + 
+                     Item3 + c_Vag:Item3 + 
+                     c_Sel + c_Vag:c_Sel + 
+                     (1 + c_Vag * c_Itm * c_Sel|Subject), dat_model, REML=FALSE)
+pretty_coef_table(null.model, "null model")
+
+full.model <- lmer(RT_log ~ 
+                     Item1 + c_Vag:Item1 +
+                     Item2 + c_Vag:Item2 + 
+                     Item3 + c_Vag:Item3 + 
+                     c_Sel + c_Vag:c_Sel + c_Vag +
+                     (1 + c_Vag * c_Itm * c_Sel|Subject), dat_model, REML=FALSE)
+pretty_coef_table(full.model, "full model")
+
+anova(null.model, full.model)
+```

### 2018-08-20 experiments/D_exp_2/writeup/D_exp_2.Rmd
foo

@@
 The variables are:
-Vagueness 2 levels (X) sumcoded -.5, +.5 
+Vagueness 2 levels (X) 
 Item 4 levels (Y) 
@@
 ```{r}
-dat$XYZ <- with(dat, factor(paste(Vagueness,Item,Selection)))
-```
-
-Convert a K level factor Y into a sum-coding representation.
-
-```{r}
-Item.numeric <- sapply(dat$Item,function(i) contr.sum(4)[i,])
-dat$Item1 <- Item.numeric[1,]
-dat$Item2 <- Item.numeric[2,]
-dat$Item3 <- Item.numeric[3,]
+#dat$XYZ <- with(dat, factor(paste(Vagueness,Item,Selection)))
 ```
@@
 ```{r}
-dat_model <- dat
+dat_model <- subset(dat, select=c(Subject, Item, Vagueness, Selection, RT_log))
+
 Item.numeric <- sapply(dat_model$Item,function(i) contr.sum(4)[i,])
@@
 dat_model$Item3 <- Item.numeric[3,]
-dat_model$c_Sel <- ifelse(dat_model$Selection=="Comparison", -0.5, 0.5)
-dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
-dat_model$c_Sel <- ifelse(dat_model$Selection=="Comparison", -0.5, 0.5)
+
+Selection.numeric <- sapply(dat_model$Selection,function(i) contr.sum(2)[i,])
+dat_model$Selection1 <- Selection.numeric[]
+
+Vagueness.numeric <- sapply(dat_model$Vagueness,function(i) contr.sum(2)[i,])
+dat_model$Vagueness1 <- Vagueness.numeric[]
+
 dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
 
+head(dat_model)
+```
+
+```{r, "null model 0"}
+# make the null model
 null.model <- lmer(RT_log ~ 
-                     Item1 + c_Vag:Item1 +
-                     Item2 + c_Vag:Item2 + 
-                     Item3 + c_Vag:Item3 + 
-                     c_Sel + c_Vag:c_Sel + 
-                     (1 + c_Vag * c_Itm * c_Sel|Subject), dat_model, REML=FALSE)
+                     Item1 + Vagueness1:Item1 +
+                     Item2 + Vagueness1:Item2 + 
+                     Item3 + Vagueness1:Item3 + 
+                     Selection1 + Vagueness1:Selection1 + 
+                     (Vagueness1 * Selection1 * c_Itm |Subject), dat_model, REML=FALSE)
 pretty_coef_table(null.model, "null model")
+```
 
+```{r, "full model0"}
+# make the full model
 full.model <- lmer(RT_log ~ 
-                     Item1 + c_Vag:Item1 +
-                     Item2 + c_Vag:Item2 + 
-                     Item3 + c_Vag:Item3 + 
-                     c_Sel + c_Vag:c_Sel + c_Vag +
-                     (1 + c_Vag * c_Itm * c_Sel|Subject), dat_model, REML=FALSE)
+                     Item1 + Vagueness1:Item1 +
+                     Item2 + Vagueness1:Item2 + 
+                     Item3 + Vagueness1:Item3 + 
+                     Selection1 + Vagueness1:Selection1 + Vagueness1 +
+                     (Vagueness1 * Selection1 * c_Itm |Subject), dat_model, REML=FALSE)
 pretty_coef_table(full.model, "full model")
+```
 
+```{r}
+# do model comparison between the null and full models
 anova(null.model, full.model)

### 2018-08-25 springer/review/vagueness-springer-review.tex
added some responses

@@
 \item Could you include a sample display for each experiment presented?
-\marginnote[-.5cm]{8: \ldots}
-\begin{quote}\ldots\end{quote}
+\marginnote[-.5cm]{8: done}
+\begin{quote}We added sample displays for each experiment presented.\end{quote}
 
@@
 \item re: exp 2 presentation 3/3: what was the vague numeric instruction used in stimuli with larger dot arrays, e.g. (16,25,34)? [I assume it wasn't \emph{about 10}]
-\marginnote[-0.5cm]{13: \ldots}
-\begin{quote}\ldots\end{quote}
+\marginnote[-0.5cm]{13: done\ldots}
+\begin{quote}We have provided full tables of instructions.\end{quote}
 
@@
 There appear to be few differences otherwise \ldots one option would be to exclude the 6:15:24 condition and then re-run the statistical analysis to see if any other effects emerge.
-\marginnote[-1.5cm]{17: \ldots}
-\begin{quote}\ldots\end{quote}
+\marginnote[-1.5cm]{17: done in the online supplement but not yet added to the document}
+\begin{quote}done in the online supplement but not yet added to the document\end{quote}
 
@@
 I'd recommend following the method described in Roger Levy's paper on finding main effects in the presence of interactions (\url{https://arxiv.org/abs/1405.2094}).
-\marginnote[-1.5cm]{19: \ldots}
-\begin{quote}\ldots\end{quote}
+\marginnote[-1.5cm]{19: done}
+\begin{quote}This is done for D exp 2. The effect was not significant when using levy's method\end{quote}
 
@@
 \marginnote[-1.5cm]{20: \ldots}
-\begin{quote}\ldots\end{quote}
+\begin{quote}I have toned the language down here (after also re-doing the statitical analysis in the requested way\end{quote}
 
@@
 Re-running using Levy's method would be advisable here.
-\marginnote[-1.5cm]{23: \ldots}
-\begin{quote}\ldots\end{quote}
+\marginnote[-1.5cm]{23: done\ldots}
+\begin{quote}This is done for E exp 3. The effect was not significant when using levy's method\end{quote}
 

### 2018-08-25 springer/current/e1.tex
recompiling

@@
 
-\begin{table}
+\begin{table}[htbp]
 \centering
@@
 
-\subsection{Hypotheses}
+\subsection{Hypotheses (Experiment 1)}
 
@@
 	\item [Hypothesis 3] No large main effect of instruction format (i.e., numeric versus verbal) since we hypothesised that vagueness rather than instruction format drove the effect in Experiment 1.
-	\item [Hypothesis 4] On the basis of Experiment 1, we expect faster responses for stimuli with more discriminable arrays.(i.e., an effect of item).
+	\item [Hypothesis 4] On the basis of our experiment in \citet{green2013utility}, we expect faster responses for stimuli with more discriminable arrays.(i.e., an effect of item).
 	\item [Hypothesis 5] Participants should make more borderline case choices for vague than crisp instructions.
@@
 
-\subsection{Experimental Method}
+\subsection{Experimental Method (Experiment 1)}
 
@@
 
-\subsection{Results}
+\subsection{Results (Experiment 1)}
 
@@
 
+\begin{figure}[htbp]
+\centering
+\includegraphics[trim = 20mm 0mm 30mm 0mm, clip, width=.49\textwidth]{figures/Ce1-rtplot-1.pdf}
+\includegraphics[trim = 20mm 0mm 25mm 0mm, clip, width=.49\textwidth]{figures/Ce1-blBarChart-1}
+\caption{Experiment 1 results: mean response times by condition and item, and counts of borderline case responses by condition.}
+\label{resultsC-exp-1}
+\end{figure}
+
 \begin{description}
@@
 The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to a significant overall interaction effect between vagueness and instruction format ($\beta=-0.13$, $se=0.02$, $t=-6.6$, $p=0.0169$).
-	\item [Test of Hypothesis 3] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1, but suggesting that instruction format rather than vagueness drove the effect we observed in Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; ($\beta=-0.36$, $se=0.07$, $t=-5.1$, $p<0.0001$).
+	\item [Test of Hypothesis 3] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1, but suggesting that instruction format rather than vagueness drove the effect we observed in \citet{green2013utility} (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; ($\beta=-0.36$, $se=0.07$, $t=-5.1$, $p<0.0001$).
 	\item [Test of Hypothesis 4] There was a significant main effect of item (i.e., of which triple of numbers of dots were used in the stimulus): ($\beta=0.12$, $se=0.02$, $t=7.1$, $p<0.0001$).
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{figures/Ce1-rtplot-1.pdf}
-\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{figures/Ce1-blBarChart-1}
-\caption{Experiment 1 results: mean response times by condition and item, and counts of borderline case responses by condition.}
-\label{resultsC-exp-1}
-\end{figure}
+\subsection{Discussion (Experiment 1)}
 
-\subsection{Discussion}
-
-Experiment 1 tested to see whether when borderline cases are present, vague instructions would speed responses as they did in Experiment pilot exp 2 when there were no borderline squares.
+Experiment 1 tested to see whether when borderline cases are present, vague instructions would speed responses as they did in our experiment in \citet{green2013utility} when there were no borderline squares.
 
@@
 We also found that the effect of instruction format was significant, with numerical format slowing responses by 689 ms on average, such that the disadvantage of numerical format overwhelmed the contribution of vagueness.
-The \emph{verbal vague} condition was still responded to faster than the \emph{numerical crisp} condition, so the pattern from Experiment pilot exp 2 was reproduced, but in the light of the evidence from Experiment 1, in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like an advantage of verbal instruction format.
+The \emph{verbal vague} condition was still responded to faster than the \emph{numerical crisp} condition, so the pattern from \citet{green2013utility} was reproduced, but in the light of the evidence from Experiment 1, in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like an advantage of verbal instruction format.
 

### 2018-08-25 springer/current/e2.tex
recompiling

@@
 
-\subsection{Hypotheses}
+\subsection{Hypotheses (Experiment 2)}
 
-For Experiment 2, we hypothesised:
+For Experiment 2, we formulated the following hypotheses:
 
 \begin{description}
-	\item [Hypothesis 1] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
-	\item [Hypothesis 2] Comparison is easier for the reader than matching (main effect of selection task)
-	\item [Hypothesis 3] Effects of vagueness are different depending on whether the selection task is matching or comparison (interaction effect selection x vagueness).
+\item [Hypothesis 1] Vague instructions should result in faster responses than crisp instructions (main effect of vagueness).
+\item [Hypothesis 2] Instructions that allow comparison should result in faster responses than instructions that necessitate matching (main effect of selection task).
+\item [Hypothesis 3] Vagueness effect should differ between comparison and matching: The effects of vagueness should differ when the selection task is comparison versus when it is matching (interaction effect selection x vagueness).
 \end{description}
 
-\subsection{Experimental Method}
+\subsection{Experimental Method (Experiment 2)}
 
@@
 
-\subsection{Results}
+\subsection{Results (Experiment 2)}
 
-We used the method proposed by \citet{Levy:MainEffectsInteractions} to test for the main effect of Vagueness in the presence of its higher-order interactions.
+%We used the method proposed by \citet{Levy:MainEffectsInteractions} to test for the main effect of Vagueness in the presence of its higher-order interactions.
 
@@
 
-\begin{description}
-	\item [Test of Hypothesis 1] Vague instructions were non-significantly easier for the reader than crisp instructions when vagueness was considered as a main effect ($\beta=-0.01$, $se=0.01$, $t=-0.4$, $p=0.7158$).
-	\item [Test of Hypothesis 2] There was a statistically significant main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=0.16$, $se=0.03$, $t=6.2$, $p<0.0001$).
-	\item [Test of Hypothesis 3] Vagueness exerted effects in different directions for the comparison task and for the matching task: there was a significant interaction between vagueness and selection task ($\beta=0.13$, $se=0.03$, $t=4.2$, $p<0.0001$).
-Separate analyses were conducted testing for effects of vagueness at each level of the selection task.
-Within the comparison task vagueness significantly speeded response times compared with crisp controls ($\beta=-0.07$, $se=0.02$, $t=-3.5$, $p<0.0012$). 
-Within the matching task vagueness significantly \emph{slowed} response times compared with crisp controls ($\beta=0.06$, $se=0.02$, $t=2.9$, $p<0.0061$). 
+\begin{description} %
+	\item [Test of Hypothesis 1] Vague instructions resulted in faster responses than crisp instructions on average.
+However this difference was not significant in the full model ($\beta=-0.0057$, $se=0.0137$, $t=-0.42$, $p=0.678$).
+Using Levy's method \citep{Levy:MainEffectsInteractions} to test for main effects in the presence of higher-order interactions, by doing model comparison between a null model that included all interaction terms involving Vagueness but leaving out a term for the main effect of Vagueness, against a full model that differed only by including Vagueness as a main effect, showed that the full model was no better than the reduced model (df=1, p=0.6764) - consituting more evidence that Vagueness did not exert a significant main effect on response times.
+	\item [Test of Hypothesis 2] comparison instructions resulted in faster responses than matching instructions, and the difference was significant ($\beta=0.1618$, $se=0.0255$, $t=6.34$, $p<0.001$).
+	\item [Test of Hypothesis 3] Although Vagueness did not exert a significant main effect, Vagueness did exert effects in interactions with some other variables: the interaction between Vagueness and Selection task  was significant ($\beta=0.1306$, $se=0.0205$, $t=6.38$, $p<0.001$), suggesting that Vagueness speeded RTs in the comparison condition but slowed them down in the matching task.
+	Separate analyses at each level of selection \ldots
+	%provided evidence that vague instructions resulted in faster responses than crisp instructions for the comparison condition (p<0.01); and slower responses than crisp instructions in the matching condition (p<0.01).
+%
+%Separate analyses were conducted testing for effects of vagueness at each level of the selection task.
+%Within the comparison task vagueness significantly speeded response times compared with crisp controls ($\beta=-0.07$, $se=0.02$, $t=-3.5$, $p<0.0012$). 
+%Within the matching task vagueness significantly \emph{slowed} response times compared with crisp controls ($\beta=0.06$, $se=0.02$, $t=2.9$, $p<0.0061$). 
 \end{description}
 
-In other words, the cost reduction account was wrong to predict significant main effect advantages for vagueness (although there was a non-significant trend in the \emph{direction} predicted by the cost reduction account), and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was significantly advantageous in the comparison task.
+\subsection{Discussion (Experiment 2)}
+The cost reduction account predicted that there should be a significant main effect of Vagueness such that responses would be faster for Vague intructions than for crisp instructions.
+We found that although there was a very small effect in that direction, the effect was not statistically significant.
+Models that differed only in the presence of Vagueness as a main effect were shown not to differ signigicantly in their explanatory value.
+However we did find that Vagueness exerted effects on other variables: Vgueness speeded RTs in the comparison task and slowed RTs in the matching task.
+
+%In other words, the cost reduction account was wrong to predict significant main effect advantages for vagueness (although there was a non-significant trend in the \emph{direction} predicted by the cost reduction account), and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was significantly advantageous in the comparison task.

### 2018-08-25 springer/current/e3.tex
recompiling

@@
 \centering
-\includegraphics[width=.75\textwidth]{figures/Ee3-flow.jpg}
+\includegraphics[width=.5\textwidth]{figures/Ee3-flow.jpg}
 \caption{Experiment 3 example stimulus}
@@
 
-\subsection{Hypotheses}
+\subsection{Hypotheses (Experiment 3)}
 
@@
 
-\subsection{Results}
+\subsection{Results (Experiment 3)}
 40 volunteers participated.
@@
 
-\begin{description}
-	\item [Test of hypothesis 1] There was no significant main effect of vagueness ($\beta=-0.02$, $se=0.01$, $t=-1.5$, $p=0.1296$). 
-	\item [Test of hypothesis 2] There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=0.18$, $se=0.02$, $t=10.4$, $p<0.0001$).
-This effect was in the same direction as Experiment 3.
-	\item [Test of hypothesis 3] Vagueness did exert different effects depending on the selection task (main interaction effect of vagueness by selection $\beta=0.12$, $se=0.02$, $t=5.1$, $p<0.0001$). 
-Separate analyses of the effect of vagueness were conducted for the comparison task and for the matching task using Bonferroni-adjusted significance thresholds. 
-In the comparison task, vagueness resulted in faster response times ($\beta=-0.08$, $se=0.02$, $t=-4.3$, $p<0.0001$). 
-In the matching task vagueness slowed response times ($\beta=0.05$, $se=0.01$, $t=3.7$, $p=0.0004$). 
-These results are in the same direction as Experiment 3.
-\end{description}
-
-Once again, the cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
-
 \begin{figure}[htbp]
@@
 \includegraphics[width=\textwidth]{figures/Ee3-rtplot-1.pdf}
-\caption{Mean response times by condition for Experiment 3}
+\caption{Mean response times by condition for Experiment 3 where all instructions were verbal}
 \label{resultsE-exp-3}
 \end{figure}
+
+\begin{description}
+	\item [Test of hypothesis 1] Vague instructions resulted in faster responses than crisp instructions on average.
+However this difference was not significant in the full model ($\beta=-0.015$, $se=0.0097$, $t=-1.58$, $p=0.119$).
+Using Levy's method \citep{Levy:MainEffectsInteractions} to test for main effects in the presence of higher-order interactions, by doing model comparison between a null model that included all interaction terms involving Vagueness but leaving out a term for the main effect of Vagueness, against a full model that differed only by including Vagueness as a main effect, showed that the full model was no better than the reduced model (df=1, p=0.118) - consituting more evidence that Vagueness did not exert a significant main effect on response times.
+	\item [Test of hypothesis 2] comparison instructions resulted in faster responses than matching instructions, and the difference was significant ($\beta=0.176$, $se=0.0168$, $t=10.51$, $p<0.001$).
+	\item [Test of hypothesis 3] Although Vagueness did not exert a significant main effect, Vagueness did exert effects in interactions with some other variables: the interaction between Vagueness and Selection task  was significant ($\beta=0.123$, $se=0.0166$, $t=7.42$, $p<0.001$), suggesting that Vagueness speeded RTs in the comparison condition but slowed them down in the matching task.
+	Separate analyses at each level of selection \ldots
+	%provided evidence that vague instructions resulted in faster responses than crisp instructions for the comparison condition (p<0.01); and slower responses than crisp instructions in the matching condition (p<0.01).
+	%
+%Separate analyses of the effect of vagueness were conducted for the comparison task and for the matching task using Bonferroni-adjusted significance thresholds. 
+%In the comparison task, vagueness resulted in faster response times ($\beta=-0.08$, $se=0.02$, $t=-4.3$, $p<0.0001$). 
+%In the matching task vagueness slowed response times ($\beta=0.05$, $se=0.01$, $t=3.7$, $p=0.0004$). 
+%These results are in the same direction as Experiment 2.
+\end{description}
+
+\subsection{Discussion (Experiment 3)}
+Once again, the cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.

### 2018-08-25 springer/current/general-discussion.tex
finished renumbering
recompiling

@@
 
-To explain why this is, let us summarise our findings so far: Experiment 1 showed that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment was unable to distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format, and they mandated a comparison strategy rather than a matching strategy.
-Experiment 2 showed that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional benefit in either the verbal format instructions or the numerical format instructions.
-However, Experiment 2 did not distinguish benefits of number avoidance from benefits of the comparison selection task.
-In Experiments 3 and 4 we manipulated vagueness and the selection task, separately at each level of numerical format.
+To explain why this is, let us summarise our findings so far: %
+%\citet{green2013utility} showed that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment was unable to distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format, and they mandated a comparison strategy rather than a matching strategy.
+%
+Experiment 1 showed that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional benefit in either the verbal format instructions or the numerical format instructions.
+However, Experiment 1 did not distinguish benefits of number avoidance from benefits of the comparison selection task.
+In Experiments 2 and 3 we manipulated vagueness and the selection task, separately at each level of numerical format.
 Across the two experiments, we found that the comparison-task instructions attracted faster response times than the matching-task instructions.
@@
 \begin{table}[htbp]
-\caption{Vagueness as range reduction: a summary of Experiments 3 and 4}
+\caption{Vagueness as range reduction: a summary of Experiments 2 and 3}
 \label{Vagueness as range reduction}
@@
 \multirow{ 2}{*}{comparison} 	& crisp 		& 2				& \multirow{ 2}{*}{vagueness advantage}  	\\
-\cline{2-3}
+\noalign{\smallskip}\cline{2-3}\noalign{\smallskip}
 								& vague			& 1				&                                           \\
@@
 \multirow{ 2}{*}{matching} 		& crisp 		& 1 			& \multirow{ 2}{*}{vagueness disadvantage}	\\
-\cline{2-3}
+\noalign{\smallskip}\cline{2-3}\noalign{\smallskip}
 								& vague			& 2				&								 			\\
@@
 
-To see why, consider Figures 3 and 4.
+To see why, consider Figures \ref{resultsD-exp-2} and \ref{resultsE-exp-3}.
 Both figures depict four conditions, depending on whether the expression was crisp or vague, and depending on whether the referent could be identified using a comparison strategy or not.
@@
 
-First, the experiments by \citeauthor{Mishra01042011} suggest that it is important how much information is conveyed by an expression, and their findings are echoed by our own thoughts about range reduction (following Experiments 3 and 4).
+First, the experiments by \citeauthor{Mishra01042011} suggest that it is important how much information is conveyed by an expression, and their findings are echoed by our own thoughts about range reduction (following Experiments 2 and 3).
 In the case of (a)--(d) above, it appears that (a) conveys the most detailed information (designating the smallest segment of the temperature scale), followed by (b), then (d), then (c) (e.g., 40 degrees is above 25, but at 40 Celsius the word ``warm'' is likely to give way to ``hot'' or ``scorching''):

### 2018-08-25 experiments/D_exp_2/writeup/D_exp_2.Rmd
think i forgot to commit some old changes

--- 2018-08-18 experiments/D_exp_2/writeup/D_exp_2.Rmd
@@
   facet_grid(~Selection) 
-```
-
-```{r}
-dat$Condition <- as.factor(paste(dat$Item, dat$Vagueness))
-ggplot(data=dat, mapping=aes(x=Item, y=RT_log, group=Vagueness)) +
-  theme_bw() +  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), aspect.ratio = 1) +
-  facet_wrap(~Selection) +
-  stat_summary(mapping=aes(y=RT_log, x=Item, group=Vagueness, colour=Vagueness), fun.y=mean, geom="line", lwd=2) +
-  geom_smooth(mapping=aes(y=RT_log, x=Item, group=Vagueness, colour=Vagueness), method="loess")
 ```
@@
 
-Testing for a main effect of X when X is a two-level factor and the model also includes a main effect of a three-level factor  and the interaction between X and Y.
+The variables are:
+Vagueness 2 levels (X) 
+Item 4 levels (Y) 
+Selection 2 levels (Z)
 
-Testing for a main effect of Vagueness when Vagueness is a two-level factor and the model also includes a main effect of a four-level factor Item and the interaction between Vagueness and Item.
+Note that the null-hypothsesis model _includes_ random slopes for the main effect of X; Y; and Z; as well as for their interactions -- including a seperate random slope for all combinations of X, Y, and Z achieves this.
 
 ```{r}
-Item.numeric <- sapply(dat$Item, function(i) contr.sum(4)[i,])
-dat$Item1 <- Item.numeric[1,]
-dat$Item2 <- Item.numeric[2,]
-dat$Item3 <- Item.numeric[3,]
-m0 <- lm(RT_log ~ 
-           Item1 + Vagueness:Item1 + 
-           Item2 + Vagueness:Item2 + 
-           Item3 + Vagueness:Item3 +
-           Selection +
-           Selection:Vagueness, 
-         dat)
-contrasts(dat$Item) <- contr.sum(4)
-m.test <- lm(RT_log ~ 
-               Item +
-           Vagueness:Item1 + 
-           Vagueness:Item2 + 
-           Vagueness:Item3 +
-           Selection +
-           Selection:Vagueness, 
-         dat)
-m1 <- lm(RT_log ~ 
-           Item1 + Vagueness:Item1 + 
-           Item2 + Vagueness:Item2 + 
-           Item3 + Vagueness:Item3 +
-           Selection +
-           Selection:Vagueness +
-           Vagueness,
-         dat)
-m100 <- lm(RT_log ~
-             Item +
-             Vagueness +
-             Item:Vagueness+
-             Selection+
-             Selection:Vagueness,
-           dat)
-#m1 and m100 are the same model
-anova(m0,m1,test="F")
+dat_model <- dat
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Sel <- ifelse(dat_model$Selection=="Comparison", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
+
+Item.numeric <- sapply(dat_model$Item,function(i) contr.sum(4)[i,])
+dat_model$Item1 <- Item.numeric[1,]
+dat_model$Item2 <- Item.numeric[2,]
+dat_model$Item3 <- Item.numeric[3,]
 ```
@@
 ```{r}
-Item.numeric <- sapply(dat$Item, function(i) contr.sum(4)[i,])
-dat$Item1 <- Item.numeric[1,]
-dat$Item2 <- Item.numeric[2,]
-dat$Item3 <- Item.numeric[3,]
-m3 <- lm(RT_log ~ 
-           Item1 + 
-           Item2 +  
-           Item3 , 
-         dat)
-contrasts(dat$Item) <- contr.sum(4)
-m4 <- lm(RT_log ~
-           Item,
-         dat)
-# m3 and m4 are the same model
+null.model2 <- lmer(RT_log ~ 
+                     c_Sel + c_Vag:c_Sel + # main effect of selection and the interaction with vagueness
+                     Item1 + c_Vag:Item1 +         # item effect and interaction with vagueness
+                     Item2 + c_Vag:Item2 +         # item effect and interaction with vagueness
+                     Item3 + c_Vag:Item3 +         # item effect and interaction with vagueness
+                     (c_Sel + c_Vag + c_Itm| Subject), # per-subject
+                   dat_model, REML=FALSE) 
 ```
@@
 ```{r}
-Item.numeric <- sapply(dat$Item, function(i) contr.sum(4)[i,])
-dat$Item1 <- Item.numeric[1,]
-dat$Item2 <- Item.numeric[2,]
-dat$Item3 <- Item.numeric[3,]
-
-m.full <- lm(RT_log ~
-             Item +
-             Vagueness +
-             Selection + 
-             Item:Vagueness +
-             Item:Selection + 
-             Vagueness:Selection,
-           dat)
-m.reduced <- lm(RT_log ~
-                Item1 + 
-                Item2 + 
-                Item3 + 
-                Vagueness:Item1 + 
-                Vagueness:Item2 +
-                Vagueness:Item3 +
-                Vagueness:Selection +
-                Selection +
-                Item:Selection
-                ,dat  )
-anova(m.reduced, m.full, test="F")
+pretty_coef_table(null.model2, "null.model2")
 ```
 
-# Discussion
+```{r}
+full.model2 <- lmer(RT_log ~ 
+                     c_Vag + 
+                     c_Sel + c_Vag:c_Sel + # main effect of selection and the interaction with vagueness
+                     Item1 + c_Vag:Item1 +         # item effect and interaction with vagueness
+                     Item2 + c_Vag:Item2 +         # item effect and interaction with vagueness
+                     Item3 + c_Vag:Item3 +         # item effect and interaction with vagueness
+                     (c_Sel + c_Vag + c_Itm| Subject), # per-subject
+                   dat_model, REML=FALSE) 
+```
+
+```{r}
+pretty_coef_table(full.model2, "full.model2")
+```
+
+```{r}
+anova(null.model2, full.model2)
+```
+
+* [Hypothesis 1] Crisp/Vague RT:
+    * _Vague instructions should result in faster responses than crisp instructions_ : Vague instructions resulted in faster responses than crisp instructions on average (beta = -0.0057 units of log(ms)).
+However this difference was not significant in the full model (p=0.678).
+Using Levy's method to test for main effects in the presence of higher-order interactions, by doing model comparison between a null model that included all interaction terms involving Vagueness but leaving out a term for the main effect of Vagueness, against a full model that differed only by including Vagueness as a main effect, showed that the full model was no better than the reduced model (df=1, p=0.6764) - consituting more evidence that Vagueness did not exert a significant main effect on response times.
+* [Hypothesis 2] Comparison/Matching RT:
+    * _Instructions that allow comparison should result in faster responses than instructions that necessitate matching_ : comparison instructions resulted in faster responses than matching instructions, and the difference was significant (beta=0.1618, p<0.001).
+* [Hypothesis 3] Vagueness effect differs between comparison and matching
+    * _The effects of vagueness should differ when the selection task is comparison versus when it is matching_ : Although Vagueness did not exert a significant main effect, Vagueness did exert effects in interactions with some other variables: the interaction between Vagueness and Selection task  was significant (beta=0.1206, p<0.001), suggesting that Vagueness speeded RTs in the comparison condition but slowed them down in the matching task.
+Separate analyses at each level of selection provided evidence that vague instructions resulted in faster responses than crisp instructions for the comparison condition (p<0.01); and slower responses than crisp instructions in the matching condition (p<0.01).

### 2018-08-25 experiments/E_exp_3/writeup/E_exp_3.Rmd
think i forgot to commit some old changes

@@
 
-```{r, "make data for rt plot"}
+```{r, "make data for rt plot", echo=FALSE}
 dat_plot <- summarySEwithin2(dat, measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
@@
 
-The second set of models considered were conducted to test for main effects _in the presence of interactions involving those main effects_ (after Levy, 2018)
+The second set of models considered were conducted to test for main effects _in the presence of interactions involving those main effects_ (after Levy, 2018).
 
-# Discussion
+The variables are:
+Vagueness 2 levels (X) 
+Item 4 levels (Y) 
+Selection 2 levels (Z)
+
+Note that the null-hypothsesis model _includes_ random slopes for the main effect of X; Y; and Z; as well as for their interactions -- including a seperate random slope for all combinations of X, Y, and Z achieves this.
+
+```{r}
+dat_model <- dat
+dat_model$c_Vag <- ifelse(dat_model$Vagueness=="Crisp", -0.5, 0.5)
+dat_model$c_Sel <- ifelse(dat_model$Selection=="Comparison", -0.5, 0.5)
+dat_model$c_Itm <- ifelse(dat_model$Item=="06:15:24", -.75, ifelse(dat_model$Item=="16:25:34", -.25, ifelse(dat_model$Item=="26:35:44", .25, .75)))
+
+Item.numeric <- sapply(dat_model$Item,function(i) contr.sum(4)[i,])
+dat_model$Item1 <- Item.numeric[1,]
+dat_model$Item2 <- Item.numeric[2,]
+dat_model$Item3 <- Item.numeric[3,]
+```
+
+```{r}
+null.model2 <- lmer(RT_log ~ 
+                     c_Sel + c_Vag:c_Sel + # main effect of selection and the interaction with vagueness
+                     Item1 + c_Vag:Item1 +         # item effect and interaction with vagueness
+                     Item2 + c_Vag:Item2 +         # item effect and interaction with vagueness
+                     Item3 + c_Vag:Item3 +         # item effect and interaction with vagueness
+                     (c_Sel + c_Vag + c_Itm| Subject), # per-subject
+                   dat_model, REML=FALSE) 
+```
+
+```{r}
+pretty_coef_table(null.model2, "null.model2")
+```
+
+```{r}
+full.model2 <- lmer(RT_log ~ 
+                     c_Vag + 
+                     c_Sel + c_Vag:c_Sel + # main effect of selection and the interaction with vagueness
+                     Item1 + c_Vag:Item1 +         # item effect and interaction with vagueness
+                     Item2 + c_Vag:Item2 +         # item effect and interaction with vagueness
+                     Item3 + c_Vag:Item3 +         # item effect and interaction with vagueness
+                     (c_Sel + c_Vag + c_Itm| Subject), # per-subject
+                   dat_model, REML=FALSE) 
+```
+
+```{r}
+pretty_coef_table(full.model2, "full.model2")
+```
+
+```{r}
+anova(null.model2, full.model2)
+```
+
+* [Hypothesis 1] Crisp/Vague RT:
+    * _Vague instructions should result in faster responses than crisp instructions_ : Vague instructions resulted in faster responses than crisp instructions on average (beta = -0.015 units of log(ms)).
+However this difference was not significant in the full model (p=0.119).
+Using Levy's method to test for main effects in the presence of higher-order interactions, by doing model comparison between a null model that included all interaction terms involving Vagueness but leaving out a term for the main effect of Vagueness, against a full model that differed only by including Vagueness as a main effect, showed that the full model was no better than the reduced model (df=1, p=0.118) - consituting more evidence that Vagueness did not exert a significant main effect on response times.
+
+* [Hypothesis 2] Comparison/Matching RT:
+    * _Instructions that allow comparison should result in faster responses than instructions that necessitate matching_ : comparison instructions resulted in faster responses than matching instructions, and the difference was significant (beta=0.176, p<0.001).
+
+* [Hypothesis 3] Vagueness effect differs between comparison and matching:
+    * _The effects of vagueness should differ when the selection task is comparison versus when it is matching_ : 
+    Vagueness did not exert a significant main effect, Vagueness did exert effects in interactions with some other variables: the interaction between Vagueness and Selection task  was significant (beta=0.123, p<0.001), suggesting that Vagueness speeded RTs in the comparison condition but slowed them down in the matching task.
+Separate analyses at each level of selection provided evidence that vague instructions resulted in faster responses than crisp instructions for the comparison condition (p<0.01); and slower responses than crisp instructions in the matching condition (p<0.01).

### 2018-08-27 springer/current/e1.tex
did a bit more on e1, still needs stats numbers imported from Rmds

@@
 We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54).
-Each set of arrays comprised three arrays (instead of two as in Experiment 1); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had more.
+Each set of arrays comprised three arrays (instead of two as in Experiment 1); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had more, and these flanking arrays appeared equally often on the left and right of the central array.
+
+Table \ref{instructionsC-exp-1} gives the full set of stimuli and associated instructions. 
+%
+The way in which borderline responses were construed is as follows, using as an example the array (6:15:24) and instructions that identified the smaller flanking array (6).
+6 was classified as the expected response.
+15 was classified as the borderline response.
+24 was classified as the extreme response.
+
+\begin{itemize}
+\item In the vague numerical condition the instruction was "Choose a square with about 10 dots" -- none of the arrays contained exactly 10 dots, but 10 is closer to 6 than it is to 15, making 6 a better response to that instruction, 15 a borderline response, and 24 an extreme response.
+\item In the vague verbal condition we used "Choose a square with few dots". We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
+\item In the crisp numerical condition we used "Choose the square with 6 dots".
+The smaller flanking array always contained exactly the specified number of dots.
+We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
+\item For crisp verbal, we used "Choose the square with the fewest dots". We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
+\end{itemize}
+
+On each trial, first the referring expression that constituted the instruction for that trial was displayed (e.g., "Choose a square with about 10 dots").
+Participants then pressed a key to indicate that they had read the instruction.
+The instruction remained on screen, and after 1000 ms, the arrays appeared.
 An example stimulus is given in Figure \ref{Experiment1and2examplestimulus}.
-
-Examples of crisp and vague versions of the numerical and verbal instructions are given next: the examples assume the array (6,15,24) and reference to the smaller number of dots, such that 6 was classified as the expected response; 15 as the borderline response; and 24 as the incorrect response.
-In the {\em vague numerical} condition we used \emph{Choose a square with about 10 dots}.
-None of the squares contained 10 dots.
-10 is slightly closer to 6 than to 15, justifying 6 as the best response and 15 as the borderline response.
-In the {\em vague verbal} condition we used \emph{Choose a square with few dots}.
-In the {\em crisp numerical} condition we used \emph{Choose the square with 6 dots}, and one square always did contain the number mentioned.
-For {\em crisp verbal}, we used \emph{Choose the square with the fewest dots}.
+Response time was measured from the presentation of the arrays until the keypress indicating the participant's choice.
+The trial would timeout after 60 seconds if there was no response.
+%
+In this experiment, no feedback was given.
+This was because, in the vague conditions, we did not regard any response as "correct" or "incorrect", but instead as "expected response"; "borderline response"; and "extreme response", and we did not want to draw participants' attention to this distinction explicitly.
+Which choice the participant made was recorded for analysis.
 
@@
 \begin{description}
-	\item [Hypothesis 1] There exists a main effect RT advantage for vagueness.
-	\item [Hypothesis 2] An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
-	\item [Hypothesis 3] No large main effect of instruction format (i.e., numeric versus verbal) since we hypothesised that vagueness rather than instruction format drove the effect in Experiment 1.
-	\item [Hypothesis 4] On the basis of our experiment in \citet{green2013utility}, we expect faster responses for stimuli with more discriminable arrays.(i.e., an effect of item).
-	\item [Hypothesis 5] Participants should make more borderline case choices for vague than crisp instructions.
+	\item [Hypothesis 1 (Crisp/Vague RT)] Vague instructions should result in faster responses than crisp instructions; and this pattern should hold when the model is restricted to numeric-only data and when it is restricted to verbal-only data.
+	\item [Hypothesis 2 (Numeric/Verbal RT)] There should be no real difference between responses to Numeric instructions and Verbal instructions (based on our interpretation of the experiment in \citet{green2013utility}, where we thought that vague instructions alone were driving the advantage for instructions that were both vague, and also in verbal format).
+	\item [Hypothesis 3 (Item RT)] Responses should take longer as the number of dots in the display grows larger (i.e., as the levels of Item increase).
+	\item [Hypothesis 4 (Response Type)] Vague instructions should lead to more borderline responses than crisp instructions.
 \end{description}
 
-\subsection{Experimental Method (Experiment 1)}
-
-% Participants were aged between 18 and 45 with a median age of 28. 
-We manipulated as independent variables vagueness and instruction format (i.e., numeric versus verbal instructions), yielding four conditions, \emph{vague numeric}; \emph{vague verbal}; \emph{crisp numeric}; \emph{crisp verbal}.
-We measured two dependent variables: response time; and the probability of a participant choosing the borderline case.
-On each trial, first the referring expression that constituted the instruction for that trial was displayed.
-Participants then pressed a key to indicate that they had read the instruction.
-After 1000 ms, the arrays were presented, while preserving the text of the referring expression.
-The response time dependent variable was measured from the presentation of the arrays, until the keypress indicating the participant's choice, which was also recorded.
-The trial would timeout after 60 seconds if there was no response.
-In this experiment, no feedback was given.
-This was because, in the vague conditions, we did not regard any response as `correct' or `incorrect', but instead as `borderline response', or `not borderline response', and we did not want to draw participants' attention to this distinction explicitly.
-We simply recorded whether the participant chose the best referent, the borderline case or the poorest referent, and how long it took the participant to respond.
-
 \subsection{Results (Experiment 1)} 
+\subsubsection{Full model}
+\marginnote{needs numbers for the stats}
 
@@
 Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data.
-A linear mixed model was constructed for the (logged) response times, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
+A linear mixed model was constructed for the (logged) response times, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and per-participant intercepts and slopes for sum-coded vagueness, instruction format, (and their interaction) as random effects.
 We found:
@@
 \begin{description}
-	\item [Test of Hypothesis 1] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against Hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06$, $se=.01$, $t=4.6$, $p<0.0001$).
-	\item [Test of Hypothesis 2] In focussed comparisons, vagueness was significantly disadvantageous in the \emph{numeric} conditions ($\beta=0.09$, $se=0.02$, $t=4.5$, $p<0.0001$), and non-significantly disadvantageous in the \emph{verbal} conditions ($\beta=0.02$, $se=0.02$, $t=1.5$, $p=0.1550$), offering partial evidence against Hypothesis 2.
-The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to a significant overall interaction effect between vagueness and instruction format ($\beta=-0.13$, $se=0.02$, $t=-6.6$, $p=0.0169$).
-	\item [Test of Hypothesis 3] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1, but suggesting that instruction format rather than vagueness drove the effect we observed in \citet{green2013utility} (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; ($\beta=-0.36$, $se=0.07$, $t=-5.1$, $p<0.0001$).
-	\item [Test of Hypothesis 4] There was a significant main effect of item (i.e., of which triple of numbers of dots were used in the stimulus): ($\beta=0.12$, $se=0.02$, $t=7.1$, $p<0.0001$).
-This effect seems likely to be due to the very fast responses for stimuli using arrangements of the smallest numbers of dots (6,15,24), which also had the largest difference in ratio of smallest number to largest number in the stimulus, suggesting that stimuli using these numbers of dots may have been particularly discriminable for participants.
+\item [Hypothesis 1 (Crisp/Vague RT)] Vague instructions actually led to significantly slower responses than crisp instructions, against Hypothesis 1: $\beta=$, $se=$, $t=$, $p$.
+When the model was restricted to numeric-only instructions Vague instructions still led to significantly slower responses than crisp instructions $\beta=$, $se=$, $t=$, $p$.
+When the model was restricted to verbal-only instructions Vague instructions tended to slow responses, but not significantly: $\beta=$, $se=$, $t=$, $p$.
+\item [Hypothesis 2 (Numeric/Verbal RT)] There was actually a significant difference between numeric and verbal instructions, with numeric instructions leading to longer responses than verbal instructions, against Hypothesis 2: $\beta=$, $se=$, $t=$, $p$.
+\item [Hypothesis 3 (Item RT)] Responses took longer as the levels of Item increased, supporting Hypothesis 3: $\beta=$, $se=$, $t=$, $p$.
 \end{description}
 
-A generalized linear mixed model \citet{jaeger2008categorical} was fit to the data for selection of the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
+A generalized linear mixed model \citet{jaeger2008categorical} was fit to the data for selection of the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant, as well as per-participant intercepts as random effects.
 The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{resultsC-exp-1}.
@@
 \begin{description}
-	\item [Test of Hypothesis 5] Participants were significantly more likely to choose the borderline option for vague instructions than for crisp instructions (21.9\% vs 11.3\%: $\beta=0.62$, $se=0.22$, $z=2.8$, $p=0.0059$).
+\item [Hypothesis 4 (Response Type)] Participants were significantly more likely to choose the borderline option for vague instructions than for crisp instructions (21.9\% vs 11.3\%: $\beta=0.62$, $se=0.22$, $z=2.8$, $p=0.0059$).
 Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%: $\beta=-3.35$, $se=0.23$, $z=-14.6$, $p<0.0001$).
+\end{description}
+
+\subsubsection{Restricted model}
+
+However, given that the plot shows that responses to 6:15:24 in the "crisp numeric" instructions condition were extremely fast relative to the "vague numeric" instructions to 6:15:24, the effects in the model of the full dataset could be driven by this difference.
+
+A clearer picture of the effects of interest might be obtained by removing the 6:15:24 level of Item from the data set, and fitting the model to this restricted data.
+Doing this results in the effects tabled below.
+
+With less data available, the model formula had to be simplified in order to converge -- specifically the following terms were dropped: per-subject slopes for the Vagueness by Item interaction, and per-subject slopes for the effect of Item.
+
+On the basis of the restricted model:
+
+\begin{description}
+\item [Hypothesis 1 (Crisp/Vague RT)] After dropping the 6:15:24 level, there is still a disadvantage for Vagueness, but it is not significant (p=.24) $\beta=$, $se=$, $t=$, $p$.
+When the model was restricted to numeric-only instructions the disadvantage for vagueness was very small and not significant $\beta=$, $se=$, $t=$, $p$.
+When the model was restricted to verbal-only instructions the disadvantage for vagueness was small and not significant $\beta=$, $se=$, $t=$, $p$.
+\item [Hypothesis 2 (Numeric/Verbal RT)] There is still a significant disadvantage for Numeric instruction format (p<.001).
+\item [Hypothesis 3 (Item RT)] There is still a disadvantage for increasing the number of dots but it is not significant (p=.323). 
+%\item The interaction between Vagueness and Instruction format changes sign, and is now non-significant (beta=-.022, p=.374). 
 \end{description}
@@
 
-Experiment 1 tested to see whether when borderline cases are present, vague instructions would speed responses as they did in our experiment in \citet{green2013utility} when there were no borderline squares.
+This experiment tested whether vague instructions would result in faster responses than crisp instructions, when borderline cases were present.
+Faster responses for vague instructions were found in pilot experiment B, but there were no borderline cases in that experiment.
 
-We actually found a small (but statistically significant) \emph{disadvantage} of vague instructions: vague instructions slowed people down by 112 ms on average.
+In this experiment we found in contrast that vague instructions resulted in slower responses than crisp instructions: a difference that was significant when considering the full data (112ms), but which was not significant after removing the smallest arrays from the analysis, which had a pattern opposite to the main trends in the rest of the data.
+
 We also found that the effect of instruction format was significant, with numerical format slowing responses by 689 ms on average, such that the disadvantage of numerical format overwhelmed the contribution of vagueness.
-The \emph{verbal vague} condition was still responded to faster than the \emph{numerical crisp} condition, so the pattern from \citet{green2013utility} was reproduced, but in the light of the evidence from Experiment 1, in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like an advantage of verbal instruction format.
+The verbal vague condition still yielded faster responses than the numerical crisp condition, so the pattern from pilot experiment B was reproduced, but in the light of the evidence from this experiment (Experiment 1), in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like an advantage of verbal instruction format.
 
 However, once again there is a possibly confounding factor.
-Observe that, in Experiment 1, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) {\bf selection algorithm}: To see this, consider the task of selecting the dot array that contains ``few dots'': to do this, it suffices to {\em compare} the three arrays and select the one that contains the fewest elements.
-To select the dot array that contains ``16 dots'' seems to require the participant to estimate, and then {\em match}, the cardinality of (at least) one dot array to 16, a process which could plausibly take longer, independently of vagueness.
-Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison ``algorithm''.
+Observe that, in Experiment 1, instruction format (i.e., the difference between numeric and verbal) went hand in hand with might be called the (human) "selection algorithm": To see this, consider the task of selecting the dot array that contains "few dots":" to do this, it suffices to \emph{compare} the three arrays and select the one that contains the fewest elements.
+To select the dot array that contains "16 dots" seems to require the participant to estimate, and then \emph{match}, the cardinality of (at least) one dot array to 16, a process which could plausibly take longer, independently of vagueness.
+Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use \emph{comparison} rather than having to resort to \emph{matching}.
 
 In the next two experiments we pitted the comparison algorithm and matching algorithm selection tasks against each other while controlling vagueness and instruction format.
-In Experiment 2 we restricted all the instructions to \emph{numeric} quantifiers while factorially manipulating vagueness and selection task.
-In Experiment 3 we ensured that all instructions used \emph{verbal} quantifiers, while also factorially manipulating vagueness and selection task.
+In Experiment 2 we restricted all the instructions to numeric quantifiers while factorially manipulating vagueness and selection task.
+In Experiment 3 we ensured that all instructions used verbal quantifiers, while also factorially manipulating vagueness and selection task.
 This allowed us to distinguish between the predictions of the selection task account and the instruction format account.

### 2018-08-29 experiments/B_pilot_2/writeup/B_pilot_2.Rmd
renaming for B_pilot_2

--- 2018-08-02 experiments/B_pilot_2/writeup/e1.Rmd
@@
 ---
-title: "Experiment 1 of 4"
+title: "B_pilot_2"
 author: "Matt Green"
@@
     toc: yes
-    <!--toc_float: true-->
-    mathjax: default
-    template: default
-    <!--code_folding: show-->
-    theme: default
 ---
+
 <style>

### 2018-08-29 experiments/A_pilot_1/writeup/broommm.R
renaming to A_pilot_1

@@ -1,23 +1,0 @@
-library(lme4)
-library(broom)
-
-lmm1 <- lme4::lmer(Reaction ~ Days + (Days | Subject), sleepstudy)
-tidy(lmm1)
-tidy(lmm1, effects = "fixed")
-tidy(lmm1, effects = "fixed", conf.int=TRUE)
-tidy(lmm1, effects = "fixed", conf.int=TRUE, conf.method="profile")
-tidy(lmm1, effects = "ran_modes", conf.int=TRUE)
-head(augment(lmm1, sleepstudy))
-glance(lmm1)
-library(dotwhisker)
-dwplot(lmm1)
-
-
-lmm2 <- lmerTest::lmer(Reaction ~ Days + (Days | Subject), sleepstudy)
-tidy(lmm2)
-tidy(lmm2, effects = "fixed")
-tidy(lmm2, effects = "fixed", conf.int=TRUE)
-tidy(lmm2, effects = "fixed", conf.int=TRUE, conf.method="profile")
-tidy(lmm2, effects = "ran_modes", conf.int=TRUE)
-head(augment(lmm2, sleepstudy))
-glance(lmm2)

### 2018-08-29 experiments/A_pilot_1/writeup/A_pilot_1.Rmd
renaming to A_pilot_1
tidying A

--- 2018-07-20 experiments/e0/writeup/e0.Rmd
@@
 ---
-title: "Experiment 0 pilot"
+title: "A_pilot_1"
 author: "Matt Green"
@@
     toc: yes
-    toc_float: true
-    mathjax: default
-    template: default
-    code_folding: show
-    theme: default
-
 ---
@@
 
-```{r load knitr}
-library(knitr)
-```
-
 Load libraries.
@@
 ```{r load libraries, message=FALSE, warning=FALSE}
+library(knitr)
 library(tidyverse)
@@
 
-```{r "make the data.txt file unless it's already there; read the data from file, add some vars and code the added vars"}
+```{r "make the data.txt file unless it's already there; read the data from file, add some vars and code the added vars", message=FALSE}
 
@@
 
-```{r}
+```{r, "remove error trials"}
 d <- subset(d, err==0)
@@
 
-### Uncapped
+### Untrimmed
 
-```{r fig.show='hide', warning=FALSE, echo=FALSE}
+Notice min is -6; max in 20258.
+
+```{r, "make untrimmed plot", fig.show='hide', warning=FALSE, echo=FALSE}
 # all the (correct) RTs
@@
 
-myg1 <- grid.arrange(ga1untrimmed, ga2untrimmed, nrow=2, top = textGrob("Untrimmed",gp=gpar(fontsize=20,font=3)))
+myg1 <- grid.arrange(ga1untrimmed, ga2untrimmed, nrow=2, top = textGrob("Untrimmed", gp = gpar(fontsize=20, font=3)))
 ```
 
-### Capped at 0 ms and 10,000 ms
+### Trimmed at 0 ms and 10,000 ms
 
-```{r fig.show='hide', warning=FALSE, echo=FALSE}
+```{r, "make trimmed plot", fig.show='hide', warning=FALSE, echo=FALSE}
 # now with a simple cap imposed to select RTs between 0 and 10000 ms
@@
 
-myg2 <- grid.arrange(ga1trimmed, ga2trimmed, nrow=2, top = textGrob("Trimmed at 10,1000 ms",gp=gpar(fontsize=20,font=3)))
+myg2 <- grid.arrange(ga1trimmed, ga2trimmed, nrow=2, top = textGrob("Trimmed at 0 and 10,1000 ms", gp = gpar(fontsize=20, font=3)))
 ```
 
-```{r, fig.width=10,fig.height=10}
+```{r, "put trimmed-untrimmed plot", fig.width=10,fig.height=10, echo=FALSE}
 grid.arrange(myg1, myg2, nrow=2)
@@
 
-```{r}
+```{r boxplot crisp and vague, fig.width=3, fig.height=3}
 dat %>% 
@@
   ggplot(aes(group=vag, y=rt_value)) +
-  geom_boxplot(aes(x=item, col=vag))
+  geom_boxplot(aes(x=vag, col=vag)) + 
+  theme(aspect.ratio=1)
 ```
@@
 
-```{r, lmer1}
+```{r, "lmer 1 compare different rt metrics using the same model structure", warning=FALSE, cache=TRUE}
 lmer1.ms  <- lmer(data=dat, rt_ms ~         vag + subit + quan + gap + side + (1|subject) + (1|item))
@@
 dwplot(models, vline = geom_vline(xintercept = 0)) + facet_wrap(~model, scales="free", ncol=1)
-
 ```
 
-```{r mler on logrt}
+```{r lmer on logrt, warning=FALSE, cache=TRUE}
 lmer2.log <- lmer(data=dat, rt_log ~ vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item))
-dwplot(lmer2.log %>% tidy() %>% filter(group=="fixed")) + 
-  ggtitle("rt_log ~ vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item)")
+dwplot(lmer2.log %>% tidy() %>% filter(group=="fixed")) + ggtitle("rt_log ~ vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item)")
 
-lmer2 <- lmer(data=dat, rt_reciprocal~subit*vag + (vag|subject) + (vag|item))
+lmer2.rcp <- lmer(data=dat, rt_reciprocal~subit*vag + (vag|subject) + (vag|item))
+
 # lmer3 is closer to the original anova than lmer2
-lmer3 <- tidy(lme4::lmer(data=dat, rt_reciprocal~vag*subit+quan+gap + (vag+subit|subject) + (vag|item)))
+lmer3 <- tidy(lmer(data=dat, rt_reciprocal~vag*subit+quan+gap + (vag+subit|subject) + (vag|item)))
 dwplot(lmer3 %>% filter(group=="fixed"), vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2)) + xlab("Coefficient")

### 2018-08-30 experiments/B_pilot_2/writeup/B_pilot_2.Rmd
reompile for print

@@
 
-<style>
-table, th, td {
-    border: 1px grey;
-    border-collapse: collapse;
-    padding:2px 5px 2px 5px
-}  
-</style>
+<!-- <style> -->
+<!-- table, th, td { -->
+<!--     border: 1px grey; -->
+<!--     border-collapse: collapse; -->
+<!--     padding:2px 5px 2px 5px -->
+<!-- }   -->
+<!-- </style> -->
 
@@
 
-<style>
-table {
-  white-space: nowrap;
-}
-</style>
+<!-- <style> -->
+<!-- table { -->
+<!--   white-space: nowrap; -->
+<!-- } -->
+<!-- </style> -->
 
-```{r pander, echo=FALSE, cache=FALSE}
-library(pander)
+```{r pander, echo=FALSE, cache=FALSE, message=FALSE}
 library(knitr)
+library(tidyverse)
+library(broom)
+library(broom.mixed)
+library(lme4)
+library(lmerTest)
+library(dotwhisker)
+library(kableExtra)
+library(LMERConvenienceFunctions)
+library(data.table)
+library(gridExtra)
+source('summarySEwithin2.R')
+source("B_pilot_2_functions/pretty_coef_table.R")
 ```
@@
 e <- c("Choose the the square with many dots", "Choose the square with 20 dots")
-i = data.frame(v, e)
-names(i) = c("vagueness", "example instruction")
-pander(i, justify=c('center', 'left'))
+i = data.frame(Vagueness = v, Instruction = e)
+kable(i) %>% kable_styling(full_width = F, position = "left", font_size = 11) 
 ```
@@
 names(i) = c('Gap Size', 'Small Target', 'Large Target')
-pander(i, justify=c('center', 'center', 'center'))
-```
-
-```{r loadLibraries, message=FALSE, echo=FALSE, cache=FALSE}
-library(ggplot2); library(xtable); library(gridExtra); library(lme4); library(lmerTest);  library(data.table); library(dplyr); library(dtplyr); library(pander); library(LMERConvenienceFunctions); source('summarySEwithin2.R')
-```
-
-```{r tidyr, echo=FALSE, message=FALSE}
-library(tidyr)
+kable(i) %>% kable_styling(full_width = F, position = "left", font_size = 11) 
 ```
@@
 rawdata <- read.table('data.txt', sep=' ')
-rawdata %>% head() %>% kable()
+rawdata %>% head() 
 ```
@@
 ```{r, results='asis', echo=F}
-print(
-  xtable(rtdataplot1,
-         caption='Condition means'),
-  include.rownames=FALSE,
-  caption.placement='top',
-  type='html')
+kable(rtdataplot1, caption="condition means") %>% kable_styling(full_width = F, position = "left", font_size = 11) 
 ```
@@
 ```{r tableRTmodelSimple, results='asis', echo=FALSE, cache=TRUE}
-x = as.data.frame(format(summary(rt.lmer.1)$coef, scientific=FALSE), stringsAsFactors = FALSE)
-
-for (i in 1:(ncol(x)-1)) {
-  for (j in 1:nrow(x)) {
-    x[j,i] <-  substr(x[j,i], start=1, stop=7)
-  }
-}
-
-for (i in 1:(ncol(x)-1)) {
-  x[,i] <- as.numeric(x[,i])
-}
-
-for (i in ncol(x)) {
-  for (j in 1:nrow(x)) {
-    x[j,i] <-  substr(x[j,i], start=1, stop=7)
-  }
-}
-print(
-  xtable(
-    align='lrrrrr', 
-    digits=c(0, 2, 2, 1, 1, 3),
-    x,
-    caption="Response time model"), 
-    caption.placement="top", 
-  type='html')
+pretty_coef_table(rt.lmer.1, "rt.lmer.1")
 ```
@@
 ```{r erdataplot1, results='asis', echo=F}
-print(
-  xtable(erdataplot1,
-         caption='Condition means'),
-  include.rownames=FALSE,
-  caption.placement='top',
-  type='html')
+kable(erdataplot1, caption="condition means") %>% kable_styling(full_width = F, position = "left", font_size = 11) 
 ```
@@
 ```{r tableErrorModelsimple, results='asis', cache=TRUE, echo=FALSE}
-print(
-  xtable(
-    summary(
-    ac.lmer.1
-    )$coefficients,
-    digits=c(0, 2, 2, 2, 7), 
-    caption="Error rates model"),
-  caption.placement="top", type='html')
+pretty_coef_table(ac.lmer.1, "ac.lmer.1")
 ```

### 2018-08-30 experiments/A_pilot_1/writeup/A_pilot_1.Rmd
applied custom .css to A

@@
     toc: yes
+    theme: default
+    css: styles.css
 ---
@@
   geom_smooth(se=F) +
-  facet_wrap(~subject) +
+  facet_wrap(~subject, ncol = 7) +
   theme(aspect.ratio = 1, panel.background = element_rect(fill="white", colour="black", linetype = "solid"))
@@
 
-```{r lmer on logrt, warning=FALSE, cache=TRUE}
+```{r lmer on logrt, warning=FALSE, cache=TRUE, fig.width=6, fig.height=2}
 lmer2.log <- lmer(data=dat, rt_log ~ vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item))
-dwplot(lmer2.log %>% tidy() %>% filter(group=="fixed")) + ggtitle("rt_log ~ vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item)")
+
+dwplot(lmer2.log %>% tidy() %>% filter(group=="fixed"), vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 1)) +
+  xlab("Coefficient") +
+  ggtitle("rt_log ~ vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item)") + 
+  theme(plot.title = element_text(size = 8), axis.title.x = element_text(size=8), axis.text.x = element_text(size=8))
 
@@
 # lmer3 is closer to the original anova than lmer2
-lmer3 <- tidy(lmer(data=dat, rt_reciprocal~vag*subit+quan+gap + (vag+subit|subject) + (vag|item)))
-dwplot(lmer3 %>% filter(group=="fixed"), vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 2)) + xlab("Coefficient")
+lmer3 <- lmer(data=dat, rt_reciprocal~vag*subit+quan+gap + (vag+subit|subject) + (vag|item))
+
+dwplot(lmer3 %>% tidy() %>% filter(group=="fixed"), vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 1)) +
+  xlab("Coefficient") + 
+  ggtitle("rt_reciprocal ~ vag * subit + quan + gap + (vag+subit|subject) + (vag|item)") + 
+  theme(plot.title = element_text(size = 8), axis.title.x = element_text(size=8), axis.text.x = element_text(size=8))
 ```

### 2018-08-31 experiments/A_pilot_1/writeup/A_pilot_1.Rmd
consistent YAML

@@
 
-This document aims to be a complete and up to date record of the analysis for vagueness 0, the pilot experiment.
-
-The pilot experiment used subitizable numbers as well as bigger numbers.
-The instructions had the form: _Choose the square with {{few, many},{two, three, four, five, six, seven, eight, nine}} dots._
-The variables were: subitizability; vagueness; target_size; gap_size; side (i.e., which side was the bigger number on?)
-
-# Libraries and functions
-
-Load libraries.
-
 ```{r load libraries, message=FALSE, warning=FALSE}
@@
 library(broom)
+library(broom.mixed)
+library(lme4)
+library(lmerTest)
 library(dotwhisker)
-library(grid)
+library(kableExtra)
+library(LMERConvenienceFunctions)
+library(data.table)
 library(gridExtra)
-library(lme4)
-# R CMD INSTALL broom.mixed # https://github.com/bbolker/broom.mixed
+source('summarySEwithin2.R')
+source("A_pilot_1_functions/pretty_coef_table.R")
 ```
 
-Source additional R files with functions for computing confidence intervals.
+The experiment used subitizable numbers as well as bigger numbers.
+The instructions had the form: _Choose the square with {{few, many},{two, three, four, five, six, seven, eight, nine}} dots._ The variables were: subitizability; vagueness; target_size; gap_size; side (i.e., which side was the bigger number on?)
 
-```{r sourceCIFunctions}
-source("CIfunctions.R") # for summarySEwithin etc
-```
-
-# Data
-
-Get the data.
-
-```{r "make the data.txt file unless it's already there; read the data from file, add some vars and code the added vars", message=FALSE}
-
-# make the data file from the individual subjects' data unless it's already there in the directory
-if (!file.exists("data.txt")) {
-  system("echo 'gap\ttarget\tside\tvagueness\tpair\tVersion\tLeft\tRight\tText\tCondition\tcorrectAnswer\tsubject\tactualAnswer\tisCorrect\trt' > data.txt")
-  system("cat ../experimentCode/output/*.dat >> data.txt", wait=TRUE)
-}
-
+```{r "read the data from file, add some vars", message=FALSE}
 # read in the data from file
@@
   rt_log = log( din$rt ),
-  rt_reciprocal = -1/din$rt # rate of slowing: # We multiplied reciprocal scores [1/RT] by -1 [-1/RT] to maintain the direction of effects compatible for the three variants, effectively converting speed [number of responses per ms] into 'rate of slowing' - A linear mixed model analysis of masked repetition priming, Kliegl, Masson & Richter, VISUAL COGNITION,2010,18(5),655--681 (p.~662)
-)
+  rt_reciprocal = -1/din$rt)
+# reciprocal is rate of slowing: "We multiplied reciprocal scores [1/RT] by -1 [-1/RT] to maintain the direction of effects compatible for the three variants, effectively converting speed [number of responses per ms] into 'rate of slowing' " - A linear mixed model analysis of masked repetition priming, Kliegl, Masson & Richter, VISUAL COGNITION,2010,18(5),655--681 (p.~662)
 ```

### 2018-08-31 experiments/B_pilot_2/writeup/B_pilot_2.Rmd
consistent YAML

@@
     toc: yes
+    theme: default
+    css: styles.css
 ---
-
-<!-- <style> -->
-<!-- table, th, td { -->
-<!--     border: 1px grey; -->
-<!--     border-collapse: collapse; -->
-<!--     padding:2px 5px 2px 5px -->
-<!-- }   -->
-<!-- </style> -->
-
-<!-- no wrap table entries on their white space -->
-
-<!-- <style> -->
-<!-- table { -->
-<!--   white-space: nowrap; -->
-<!-- } -->
-<!-- </style> -->
 

### 2018-08-31 experiments/C_exp_1/writeup/C_exp_1.Rmd
did a bit more on e1, still needs stats numbers imported from Rmds
consistent YAML

@@
     toc: yes
-    toc_float: false
+    theme: default
+    css: styles.css    
 ---
@@
 
-* For crisp verbal, we used Choose the square with the fewest dots. We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
+* For crisp verbal, we used "Choose the square with the fewest dots". We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
 

### 2018-08-31 experiments/A_pilot_1/writeup/A_pilot_1.Rmd
consistent YAML

@@
 library(data.table)
+library(grid)
 library(gridExtra)

### 2018-08-31 experiments/E_exp_3/writeup/E_exp_3.Rmd
consistent YAML

@@
     toc: true
-    toc_float: false
+    theme: default
+    css: styles.css
 ---

### 2018-08-31 experiments/A_pilot_1/writeup/A_pilot_1.Rmd
A update Rmd and project str

--- 2018-08-31 experiments/A_pilot_1/writeup/A_pilot_1.Rmd
@@
 library(data.table)
+library(grid)
 library(gridExtra)
-source('summarySEwithin2.R')
 source("A_pilot_1_functions/pretty_coef_table.R")
+source("A_pilot_1_functions/summarySEwithin2.R")
 ```
@@
 
-```{r, "lmer 1 compare different rt metrics using the same model structure", warning=FALSE, cache=TRUE}
+```{r, "lmer 1 compare different rt metrics using the same model structure", warning=FALSE}
 lmer1.ms  <- lmer(data=dat, rt_ms ~         vag + subit + quan + gap + side + (1|subject) + (1|item))
@@
 lmer1.rcp <- lmer(data=dat, rt_reciprocal ~ vag + subit + quan + gap + side + (1|subject) + (1|item))
+```
 
+```{r}
 models <- 
@@
 
-```{r lmer on logrt, warning=FALSE, cache=TRUE, fig.width=6, fig.height=2}
+```{r lmer on logrt, warning=FALSE}
 lmer2.log <- lmer(data=dat, rt_log ~ vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item))
+```
 
-dwplot(lmer2.log %>% tidy() %>% filter(group=="fixed"), vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 1)) +
+```{r, fig.width=6, fig.height=2}
+dwplot(lmer2.log %>% tidy() %>% filter(effect=="fixed"), vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 1)) +
   xlab("Coefficient") +
@@
   theme(plot.title = element_text(size = 8), axis.title.x = element_text(size=8), axis.text.x = element_text(size=8))
+```
 
+```{r}
 lmer2.rcp <- lmer(data=dat, rt_reciprocal~subit*vag + (vag|subject) + (vag|item))
+```
 
+```{r}
 # lmer3 is closer to the original anova than lmer2
 lmer3 <- lmer(data=dat, rt_reciprocal~vag*subit+quan+gap + (vag+subit|subject) + (vag|item))
+```
 
-dwplot(lmer3 %>% tidy() %>% filter(group=="fixed"), vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 1)) +
+```{r, fig.width=6, fig.height=2}
+dwplot(lmer3 %>% tidy() %>% filter(effect=="fixed"), vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 1)) +
   xlab("Coefficient") + 

### 2018-08-31 experiments/A_pilot_1/writeup/CIfunctions.R

@@ -1,144 +1,0 @@
-# This file contains functions that implement Morey's correction to Cousineau within subjects Cis 
-
-
-
-## Summarizes data.
-## Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
-##   data: a data frame.
-##   measurevar: the name of a column that contains the variable to be summariezed
-##   groupvars: a vector containing names of columns that contain grouping variables
-##   na.rm: a boolean that indicates whether to ignore NA's
-##   conf.interval: the percent range of the confidence interval (default is 95%)
-summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
-                      conf.interval=.95, .drop=TRUE) {
-  require(plyr)
-  
-  # New version of length which can handle NA's: if na.rm==T, don't count them
-  length2 <- function (x, na.rm=FALSE) {
-    if (na.rm) sum(!is.na(x))
-    else       length(x)
-  }
-  
-  # This is does the summary; it's not easy to understand...
-  datac <- ddply(data, groupvars, .drop=.drop,
-                 .fun= function(xx, col, na.rm) {
-                   c( N    = length2(xx[,col], na.rm=na.rm),
-                      mean = mean   (xx[,col], na.rm=na.rm),
-                      sd   = sd     (xx[,col], na.rm=na.rm)
-                   )
-                 },
-                 measurevar,
-                 na.rm
-  )
-  
-  # Rename the "mean" column    
-  datac <- rename(datac, c("mean"=measurevar))
-  
-  datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
-  
-  # Confidence interval multiplier for standard error
-  # Calculate t-statistic for confidence interval: 
-  # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
-  ciMult <- qt(conf.interval/2 + .5, datac$N-1)
-  datac$ci <- datac$se * ciMult
-  
-  return(datac)
-}
-
-
-## Summarizes data, handling within-subjects variables by removing inter-subject variability.
-## It will still work if there are no within-S variables.
-## Gives count, un-normed mean, normed mean (with same between-group mean),
-##   standard deviation, standard error of the mean, and confidence interval.
-## If there are within-subject variables, calculate adjusted values using method from Morey (2008).
-##   data: a data frame.
[…]

### 2018-08-31 experiments/B_pilot_2/B_pilot_2.Rmd
mvd writeup contents up

--- 2018-08-31 experiments/B_pilot_2/writeup/B_pilot_2.Rmd
@@
 library(data.table)
+library(grid)
 library(gridExtra)
-source('summarySEwithin2.R')
+source("B_pilot_2_functions/summarySEwithin2.R")
 source("B_pilot_2_functions/pretty_coef_table.R")
@@
 
-```{r rawdata, cache=FALSE}
+```{r rawdata, cache=FALSE, comment=NA}
 rawdata <- read.table('data.txt', sep=' ')

### 2018-08-31 experiments/B_pilot_2/preprocessing/preprocessing.Rmd
mvd writeup contents up

# Make data frame

```{r makeDataFrame}
dat <- read.delim("data_raw.txt", header=FALSE)
names(dat) <- c("condition", "leftN", "rightN", "vgn", "side", "string", "subject", "rt", "error")
dat$condition <- as.factor(dat$condition)
# item_ratio
for (c in unique(dat$condition)) {
  x=subset(dat, condition==c)
  x$item_ratio=unique(min(x$leftN,x$rightN)/max(x$leftN,x$rightN))
  dat[dat$condition==c, 'item_ratio']=x$item_ratio
}
# trial
for (s in unique(dat$subject)) {
  x=subset(dat, subject==s)
  x$trial=1:nrow(x)
  dat[dat$subject==s, 'trial']=x$trial
}
dat$subject <- as.factor(paste("sub",sep="",sprintf("%02d",dat$subject,sep="")))
dat$error <- as.logical(1-dat$error)
dat$Ord <- as.factor(ifelse(dat$leftN < dat$rightN, 'LtoR', 'RtoL'))
dat$c_Ord <- ifelse(dat$Ord=='LtoR', -.5, .5)
dat$bigside <- ifelse(dat$leftN > dat$rightN, 'L', 'R')
dat$Qty <- as.factor(ifelse(dat$side==dat$bigside, "Large", "Small"))
dat$bigside=NULL
dat$c_Qty <- ifelse(dat$Qty=='Small', -.5, .5)
dat$Gap <- as.factor(as.character(paste("gap", sep=" ", sprintf("%02d",abs(dat$leftN-dat$rightN)))))
dat$c_Gap[dat$Gap=="gap 05"] <- -1.5
dat$c_Gap[dat$Gap=="gap 10"] <- -.5
dat$c_Gap[dat$Gap=="gap 15"] <- .5
dat$c_Gap[dat$Gap=="gap 20"] <- 1.5
dat$Vgn <- as.factor(ifelse(dat$vgn=="P", "Crisp", "Vague"))
dat$vgn = NULL
dat$c_Vgn <- ifelse(dat$Vgn=='Crisp', -.5, .5)
dat$logrt <- log(dat$rt)
dat <- subset(dat, select=c(subject, trial, condition, leftN, rightN, item_ratio, side, string, Vgn, c_Vgn, Ord, c_Ord, Qty, c_Qty, Gap, c_Gap, error, rt, logrt))
row.names(dat) <- NULL
write.table(x=dat, "../data.txt")
```

### 2018-08-31 experiments/B_pilot_2/experimentCode/experimentpresentation.m

@@ -1,690 +1,0 @@
-function experimentpresentation(SUBJECT)
-% This second study has the following structure: practice block followed by:4 blocks with a rest in
-% between. Each block has 2 presentations of 32 trials. The 32 trial are
-% composed of 8 pairs of quantities presented once with vague instructions
-% and once with precise instructions, with then element of the pair that is
-% being picked out presented on the left and on the right (8 pairs * 2
-% sides * 2 instructions types * sides = 32 trials, presented twice per
-% block = 64 trials; each block presented 4 times = 256 trials).
-%
-% A trial consists of the instruction on screen while participants read it
-% for as long as they like; a button press on SPACEBAR to dismiss the
-% instruction; a blank screen of 100ms duration; a fixation cross of 100ms
-% duration; a blank of 100ms duration; a screen with the pair presented
-% (currently for unlimited duration terminated by a key press from a
-% restricted set (left cmd, right cmd) ); 1000ms ISI.
-%
-% Participant id > 100 is used for testing and can be overwritten
-
-howLongToWaitForAResponse = 60; % seconds. Set low for test runs
-testing = 0; % 0 for real runs
-
-% Clear Matlab/Octave window:
-clc;
-
-% check for Opengl compatibility, abort otherwise:
-AssertOpenGL;
-
-% Reseed the random-number generator for each expt.
-rand('state',sum(100*clock));
-
-% Make sure keyboard mapping is the same on all supported operating systems
-% Apple MacOS/X, MS-Windows and GNU/Linux:
-KbName('UnifyKeyNames');
-
-%     % Turn off the annoying echo to the command line
-%     ListenChar(2);
-
-% Define allowable keys
-leftresp=KbName('q'); %227
-rightresp=KbName('p'); %231
-spacebar=KbName('space'); %44
-%    RestrictKeysForKbCheck([leftresp,rightresp,spacebar]);
-
-%%%%%%%%%%%%%%%%%%%%%%
-% file handling
-%%%%%%%%%%%%%%%%%%%%%%
-
-%practice stimuli to read from
-practiceStimuliFilename = 'practiceStimuli.txt';
-
-%practice data output file to write to
-practicedatafilename = strcat('dummy/subject',...
-    num2str(SUBJECT),'Practice','.dat');
-
-% experimental stimuli to read from
-experimentalStimuliFilename = 'experimentalStimuli.txt';
-
[…]

### 2018-08-31 experiments/E_exp_3/experimentCode/e4.m

@@ -1,387 +1,0 @@
-function e5(SUBJECT)
-
-diary on;
-more off;
-newLevel=3; % adjust verbosity
-oldLevel = Screen('Preference', 'Verbosity', [newLevel]);
-ListenChar(2); % disable keyboard input to octave while running experiment (no echo to terminal)
-% If we are in testing mode then do skip synch tests, If we are in real mode don't skip synch tests
-skipTestValue=0
-if SUBJECT >99
-	skipTestValue=1
-end
-
-
-Screen('Preference', 'SkipSyncTests', skipTestValue); % 0 = don't skip sync tests; 1 = do skip run sync tests
-
-% VARIABLES
-howLongToWaitForAResponse = 10; % in seconds. Set low for test runs
-timeOutTime = 2; % in seconds.
-maskDuration = 1; % in seconds
-interTrialDelay = 1; % in seconds
-leftBox  = [170 300 370 500];
-midBox   = [540 300 740 500];
-rightBox = [910 300 1110 500];
-leftCenter  = [270 400];
-midCenter   = [640 400];
-rightCenter = [1010 400];
-messagey = 150; 
-instructionY = 400;
-instructionX = 100;
-gray=127.5;
-white=225;
-black=0;
-startrt=0;   % initialise start rt variable
-endrt=0;     % initialise end rt variable
-instruction_on_time=0;
-instruction_off_time=0;
-numberOfBlocks = 4; % 1 is practice, 2 is block1, 3 is block2, 4 is block3
-numberOfColumns = 9; % there are always 9 columns in the input
-rand('state',sum(100*clock));
-
-% COMMANDS
-% Clear Matlab/Octave window:
-clc;
-% check for Opengl compatibility, abort otherwise:
-AssertOpenGL;
-% Make sure keyboard mapping is the same on all supported operating systems
-KbName('UnifyKeyNames');
-% Define keys by ID
-leftresp=KbName('c'); 
-middleresp=KbName('b'); 
-rightresp=KbName('m'); 
-escapekey=KbName('ESCAPE'); 
-spacebar=KbName('space'); 
-returnkey=KbName('return'); 
-% read separate files containing dotlocations, stimuli, and strings 
-source('e5stimuli.m');
-source('e5strings.m');
-source('e5dotlocations.m');
-
-% set up output files
-% practice data output file to write to
-practicedatafilename = strcat('output/practice/subject',...
-    num2str(SUBJECT),'Practice','.data');
-% experimental data output file to write to
[…]

### 2018-08-31 experiments/E_exp_3/experimentCode/e4strings.m

@@ -1,10 +1,0 @@
-    %%%%%%%%%%%%%%%%%%%%%%%
-    % define text strings %
-    %%%%%%%%%%%%%%%%%%%%%%%
-
-    instructionText = sprintf(strcat('Press Enter to start the practice trials.'));
-
-    oopsText = sprintf(strcat(...
-        'Ooops... Something went wrong.\n\n',...
-        'Please tell the experimenter.'));
-

### 2018-08-31 experiments/E_exp_3/experimentCode/touchpadoff.sh

@@ -1,2 +1,0 @@
-synclient TouchpadOff=1
-

### 2018-09-01 springer/current/e1.tex
e1 first pass done completely in springer latex

@@
-\marginnote{C-Exp-1}
+% C_exp_1
 
@@
 
-In Experiment 1, an item was a referring expression instruction followed by a set of three dot arrays defined by a triple of numbers, representing the number of dots in the left, middle, and right arrays.
+In Experiment 1, an item was an instruction followed by a set of three dot arrays defined by a triple of numbers, representing the number of dots in the left, middle, and right arrays.
 We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54).
-Each set of arrays comprised three arrays (instead of two as in Experiment 1); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had more, and these flanking arrays appeared equally often on the left and right of the central array.
+Each set of arrays comprised three arrays (instead of two as in \citet{green2013utility}); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had more, and these flanking arrays appeared equally often on the left and right of the central array.
 
 Table \ref{instructionsC-exp-1} gives the full set of stimuli and associated instructions.
-%
 The way in which borderline responses were construed is as follows, using as an example the array (6:15:24) and instructions that identified the smaller flanking array (6).
@@
 The trial would timeout after 60 seconds if there was no response.
-%
 In this experiment, no feedback was given.
@@
 \subsection{Results (Experiment 1)}
-\subsubsection{Full model}
-\marginnote{needs numbers for the stats}
 
-Means for response times and proportion of borderline responses are given in Fig.
-(\ref{resultsC-exp-1}).
 Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data.
-A linear mixed model was constructed for the (logged) response times, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and per-participant intercepts and slopes for sum-coded vagueness, instruction format, (and their interaction) as random effects.
-We found:
+The distribution of remaining response times was skewed with many long responses.
+These remaining response times were log-transformed which reduced this skew so that their distribution more closely approximated a normal distribution.
+Means for response times and proportion of borderline responses are given in Figure \ref{resultsC-exp-1}.
 
@@
 
+A linear mixed model was constructed for the (log-transformed) response times, with sum-coded vagueness, instruction format, and their interaction, and item, as fixed effects, and per-participant intercepts and slopes for sum-coded vagueness, instruction format, and their interaction as random effects.
+
 \begin{description}
-\item [Hypothesis 1 (Crisp/Vague RT)] Vague instructions actually led to significantly slower responses than crisp instructions, against Hypothesis 1: $\beta=$, $se=$, $t=$, $p$.
-When the model was restricted to numeric-only instructions Vague instructions still led to significantly slower responses than crisp instructions $\beta=$, $se=$, $t=$, $p$.
-When the model was restricted to verbal-only instructions Vague instructions tended to slow responses, but not significantly: $\beta=$, $se=$, $t=$, $p$.
-\item [Hypothesis 2 (Numeric/Verbal RT)] There was actually a significant difference between numeric and verbal instructions, with numeric instructions leading to longer responses than verbal instructions, against Hypothesis 2: $\beta=$, $se=$, $t=$, $p$.
-\item [Hypothesis 3 (Item RT)] Responses took longer as the levels of Item increased, supporting Hypothesis 3: $\beta=$, $se=$, $t=$, $p$.
+\item [Hypothesis 1 (Crisp/Vague RT)] Vague instructions actually led to significantly slower responses than crisp instructions, against Hypothesis 1: $\beta=0.058$, $se=0.013$, $t=4.55$, $p<0.001$.
+When the model was restricted to numeric-only instructions Vague instructions still led to significantly slower responses than crisp instructions $\beta=0.093$, $se=0.021$, $t=4.51$, $p<0.001$.
+When the model was restricted to verbal-only instructions Vague instructions tended to slow responses, but not significantly: $\beta=0.024$, $se=0.016$, $t=1.46$, $p=0.155$.
+\item [Hypothesis 2 (Numeric/Verbal RT)] There was actually a significant difference between numeric and verbal instructions, with numeric instructions leading to longer responses than verbal instructions, against Hypothesis 2: $\beta0.265=$, $se=0.072$, $t=5.08$, $p<0.001$.
+\item [Hypothesis 3 (Item RT)] Responses took longer as the levels of Item increased, supporting Hypothesis 3: $\beta=0.120$, $se=0.017$, $t=7.11$, $p<0.001$.
 \end{description}
 
+However, given that the response time plot in the left panel of Figure \ref{resultsC-exp-1} shows that responses to 6:15:24 in the "crisp numeric" instructions condition were extremely fast relative to the "vague numeric" instructions to 6:15:24, the effects in the model of the full dataset could be driven by this difference.
+A clearer picture of the effects of interest might be obtained by removing the 6:15:24 level of Item from the data set, and fitting the model to this restricted data.
+Doing this did not affect the direction of the effects in the full dataset, but whereas the effects were significant in the full dataset, they were not significant in the restricted dataset.
+Full details of the analysis of the restricted dataset are available in the online materials.
+
 A generalized linear mixed model \citet{jaeger2008categorical} was fit to the data for selection of the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant, as well as per-participant intercepts as random effects.
-The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{resultsC-exp-1}.
+The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Figure \ref{resultsC-exp-1}.
 Participants chose the borderline square on $16.6\%$ of trials overall.
@@
 Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%: $\beta=-3.35$, $se=0.23$, $z=-14.6$, $p<0.0001$).
-\end{description}
-
-\subsubsection{Restricted model}
-
-However, given that the plot shows that responses to 6:15:24 in the "crisp numeric" instructions condition were extremely fast relative to the "vague numeric" instructions to 6:15:24, the effects in the model of the full dataset could be driven by this difference.
-
-A clearer picture of the effects of interest might be obtained by removing the 6:15:24 level of Item from the data set, and fitting the model to this restricted data.
-Doing this results in the effects tabled below.
-
-With less data available, the model formula had to be simplified in order to converge -- specifically the following terms were dropped: per-subject slopes for the Vagueness by Item interaction, and per-subject slopes for the effect of Item.
-
-On the basis of the restricted model:
-
-\begin{description}
-\item [Hypothesis 1 (Crisp/Vague RT)] After dropping the 6:15:24 level, there is still a disadvantage for Vagueness, but it is not significant (p=.24) $\beta=$, $se=$, $t=$, $p$.
-When the model was restricted to numeric-only instructions the disadvantage for vagueness was very small and not significant $\beta=$, $se=$, $t=$, $p$.
-When the model was restricted to verbal-only instructions the disadvantage for vagueness was small and not significant $\beta=$, $se=$, $t=$, $p$.
-\item [Hypothesis 2 (Numeric/Verbal RT)] There is still a significant disadvantage for Numeric instruction format (p<.001).
-\item [Hypothesis 3 (Item RT)] There is still a disadvantage for increasing the number of dots but it is not significant (p=.323). 
-%\item The interaction between Vagueness and Instruction format changes sign, and is now non-significant (beta=-.022, p=.374). 
 \end{description}

### 2018-09-01 springer/current/e1.tex
some consistency edits

@@
 
+30 participants were recruited.
 Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data.
@@
 These remaining response times were log-transformed which reduced this skew so that their distribution more closely approximated a normal distribution.
-Means for response times and proportion of borderline responses are given in Figure \ref{resultsC-exp-1}.
+Condition means for response times and proportion of borderline responses are given in Figure \ref{resultsC-exp-1}.
 
@@
 \begin{description}
-\item [Hypothesis 1 (Crisp/Vague RT)] Vague instructions actually led to significantly slower responses than crisp instructions, against Hypothesis 1: $\beta=0.058$, $se=0.013$, $t=4.55$, $p<0.001$.
+	\item [Test of Hypothesis 1 (Crisp/Vague RT)] Vague instructions actually led to significantly slower responses than crisp instructions, against Hypothesis 1: $\beta=0.058$, $se=0.013$, $t=4.55$, $p<0.001$.
 When the model was restricted to numeric-only instructions Vague instructions still led to significantly slower responses than crisp instructions $\beta=0.093$, $se=0.021$, $t=4.51$, $p<0.001$.
 When the model was restricted to verbal-only instructions Vague instructions tended to slow responses, but not significantly: $\beta=0.024$, $se=0.016$, $t=1.46$, $p=0.155$.
-\item [Hypothesis 2 (Numeric/Verbal RT)] There was actually a significant difference between numeric and verbal instructions, with numeric instructions leading to longer responses than verbal instructions, against Hypothesis 2: $\beta0.265=$, $se=0.072$, $t=5.08$, $p<0.001$.
-\item [Hypothesis 3 (Item RT)] Responses took longer as the levels of Item increased, supporting Hypothesis 3: $\beta=0.120$, $se=0.017$, $t=7.11$, $p<0.001$.
+	\item [Test of Hypothesis 2 (Numeric/Verbal RT)] There was actually a significant difference between numeric and verbal instructions, with numeric instructions leading to longer responses than verbal instructions, against Hypothesis 2: $\beta0.265=$, $se=0.072$, $t=5.08$, $p<0.001$.
+	\item [Test of Hypothesis 3 (Item RT)] Responses took longer as the levels of Item increased, supporting Hypothesis 3: $\beta=0.120$, $se=0.017$, $t=7.11$, $p<0.001$.
 \end{description}

### 2018-09-01 springer/current/e2.tex
did a bit more on e1, still needs stats numbers imported from Rmds
some consistency edits

--- 2018-08-19 springer/current/e2.tex
@@
-\marginnote{D-Exp-2}
+% D_Exp_2
 
 The main aim of Experiment 2 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies.
-The main changes from Experiment 1 were that the human selection task was explicitly controlled (i.e., in whether it amounted to matching or comparison), and that all conditions were constrained to mention a number.
+The main changes from Experiment 1 were that the human selection task was explicitly controlled (i.e., whether the task amounted to matching or comparison), and that all conditions were constrained to mention a number.
 We used the same arrays as in Experiment 1 (an example stimulus is given in Figure \ref{Experiment1and2examplestimulus}).
+We used a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e2}).
+On each trial an instruction was presented: participants pressed a key to dismiss the instruction, at which time the dot arrays were presented until the participant responded, and the response time and choice were recorded.
 Table \ref{Instructions for e2} shows the instructions for each condition.
@@
 
-\subsection{Hypotheses}
+\subsection{Hypotheses (Experiment 2)}
 
-For Experiment 2, we hypothesised:
+For Experiment 2, we formulated the following hypotheses:
 
 \begin{description}
-	\item [Hypothesis 1] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
-	\item [Hypothesis 2] Comparison is easier for the reader than matching (main effect of selection task)
-	\item [Hypothesis 3] Effects of vagueness are different depending on whether the selection task is matching or comparison (interaction effect selection x vagueness).
+	\item [Hypothesis 1 (Crisp/Vague RT)] Vague instructions should result in faster responses than crisp instructions.
+	\item [Hypothesis 2 (Comparison/Matching RT)] Instructions that allow comparison should result in faster responses than instructions that necessitate matching.
+	\item [Hypothesis 3 (Interaction)] The vagueness effect should differ according to whether the selection task is comparison or matching.
 \end{description}
 
-\subsection{Experimental Method}
+\subsection{Results (Experiment 2)}
 
 38 participants were recruited.
-We used a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e2}).
-On each trial an instruction was presented: participants pressed a key to dismiss the instruction, at which time the dot arrays were presented until the participant responded, and the response time and choice were recorded.
-
-\subsection{Results}
-
-We used the method proposed by \citet{Levy:MainEffectsInteractions} to test for the main effect of Vagueness in the presence of its higher-order interactions.
-
-Response times were trimmed at 2.5 SD separately for each subject, leading to the loss of 204 trials (2.8\% of the trials).
-Condition means for the remaining (logged) RTs are plotted in Figure \ref{resultsD-exp-2}.
-A linear mixed model was constructed for the logged response times, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
+Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 204 trials (2.8\% of the trials).
+The distribution of remaining response times was skewed with many long responses.
+These remaining response times were log-transformed which reduced this skew so that their distribution more closely approximated a normal distribution.
+Condition means for response times are plotted in Figure \ref{resultsD-exp-2}.
 
@@
 
+A linear mixed model was constructed for the logged response times, with sum-coded vagueness, selection task, and their interaction, and item as fixed effects, and per-participant intercepts and slopes for sum-coded vagueness, selection task, and their interaction, and item for random effects.
+
 \begin{description} 
-	\item [Test of Hypothesis 1] Vague instructions were non-significantly easier for the reader than crisp instructions when vagueness was considered as a main effect ($\beta=-0.01$, $se=0.01$, $t=-0.4$, $p=0.7158$).
-	\item [Test of Hypothesis 2] There was a statistically significant main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=0.16$, $se=0.03$, $t=6.2$, $p<0.0001$).
-	\item [Test of Hypothesis 3] Vagueness exerted effects in different directions for the comparison task and for the matching task: there was a significant interaction between vagueness and selection task ($\beta=0.13$, $se=0.03$, $t=4.2$, $p<0.0001$).
-Separate analyses were conducted testing for effects of vagueness at each level of the selection task.
-Within the comparison task vagueness significantly speeded response times compared with crisp controls ($\beta=-0.07$, $se=0.02$, $t=-3.5$, $p<0.0012$). 
-Within the matching task vagueness significantly \emph{slowed} response times compared with crisp controls ($\beta=0.06$, $se=0.02$, $t=2.9$, $p<0.0061$). 
+	\item [Test of Hypothesis 1 (Crisp/Vague RT)] Vague instructions resulted in faster responses than crisp instructions on average.
+However this difference was not significant in the full model ($\beta=-0.0057$, $se=0.0137$, $t=-0.42$, $p=0.678$).
+Using Levy's method \citep{Levy:MainEffectsInteractions} to test for main effects in the presence of higher-order interactions, by doing model comparison between a null model that included all interaction terms involving Vagueness but leaving out a term for the main effect of Vagueness, against a full model that differed only by including Vagueness as a main effect, showed that the full model was no better than the reduced model ($df=1$, $p=0.676$), consituting more evidence that Vagueness did not exert a significant main effect on response times.
+	\item [Test of Hypothesis 2 (Comparison/Matching RT)] comparison instructions resulted in faster responses than matching instructions, and the difference was significant ($\beta=0.1618$, $se=0.0255$, $t=6.34$, $p<0.001$).
+	\item [Test of Hypothesis 3 (Interaction)] Although Vagueness did not exert a significant main effect, Vagueness did exert effects in interactions with some other variables: the interaction between Vagueness and Selection task  was significant ($\beta=0.1306$, $se=0.0205$, $t=6.38$, $p<0.001$), suggesting that Vagueness speeded RTs in the comparison condition but slowed them down in the matching task.
 \end{description}
 
-In other words, the cost reduction account was wrong to predict significant main effect advantages for vagueness (although there was a non-significant trend in the \emph{direction} predicted by the cost reduction account), and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was significantly advantageous in the comparison task.
+\subsection{Discussion (Experiment 2)}
+
+The cost reduction account predicted that there should be a significant main effect of Vagueness such that responses would be faster for Vague intructions than for crisp instructions.
+We found that although there was a very small effect in that direction, the effect was not statistically significant.
+Models that differed only in the presence of Vagueness as a main effect were shown not to differ significantly in their explanatory value.
+However we did find that Vagueness exerted effects on other variables: Vagueness speeded RTs in the comparison task and slowed RTs in the matching task.

### 2018-09-01 springer/current/e3.tex
some consistency edits

--- 2018-08-19 springer/current/e3.tex
@@
-\marginnote{E-Exp-3}
+%E_Exp_3
 
@@
 \label{instructionsE-exp-3}
-\resizebox{\textwidth}{!}{\begin{tabular}{ccccll}
+\resizebox{\textwidth}{!}{\begin{tabular}{ccccl}
 \hline\noalign{\smallskip}
-Item & Target & Quantity & Selection & Crisp & Vague \\ 
+Item & Target & Selection & Vagueness & Instruction \\ 
 \noalign{\smallskip}\hline\noalign{\smallskip}
-06:15:24 &   6 & Small & Matching 	& the same number of dots as the target & NA \\ 
-06:15:24 &  10 & Small & Matching 	& NA & about the same number of dots as the target \\ 
-06:15:24 &  10 & Large & Comparison & more dots than the target & far more dots than the target \\ 
-06:15:24 &  20 & Small & Comparison & fewer dots than the target & far fewer dots than the target \\ 
-06:15:24 &  20 & Large & Matching 	& NA & about the same number of dots as the target \\ 
-06:15:24 &  24 & Large & Matching 	& the same number of dots as the target & NA \\ 
-\noalign{\smallskip}\hline\noalign{\smallskip}
-16:25:34 &  16 & Small & Matching 	& the same number of dots as the target & NA \\ 
-16:25:34 &  20 & Small & Matching 	& NA & about the same number of dots as the target \\ 
-16:25:34 &  20 & Large & Comparison & more dots than the target & far more dots than the target \\ 
-16:25:34 &  30 & Small & Comparison & fewer dots than the target & far fewer dots than the target \\ 
-16:25:34 &  30 & Large & Matching 	& NA & about the same number of dots as the target \\ 
-16:25:34 &  34 & Large & Matching 	& the same number of dots as the target & NA \\ 
-\noalign{\smallskip}\hline\noalign{\smallskip}
-26:35:44 &  26 & Small & Matching 	& the same number of dots as the target & NA \\ 
-26:35:44 &  30 & Small & Matching 	& NA & about the same number of dots as the target \\ 
-26:35:44 &  30 & Large & Comparison & more dots than the target & far more dots than the target \\ 
-26:35:44 &  40 & Small & Comparison & fewer dots than the target & far fewer dots than the target \\ 
-26:35:44 &  40 & Large & Matching 	& NA & about the same number of dots as the target \\ 
-26:35:44 &  44 & Large & Matching 	& the same number of dots as the target & NA \\ 
-\noalign{\smallskip}\hline\noalign{\smallskip}
-36:45:54 &  36 & Small & Matching 	& the same number of dots as the target & NA \\ 
-36:45:54 &  40 & Small & Matching 	& NA & about the same number of dots as the target \\ 
-36:45:54 &  40 & Large & Comparison & more dots than the target & far more dots than the target \\ 
-36:45:54 &  50 & Small & Comparison & fewer dots than the target & far fewer dots than the target \\ 
-36:45:54 &  50 & Large & Matching 	& NA & about the same number of dots as the target \\ 
-36:45:54 &  54 & Large & Matching 	& the same number of dots as the target & NA \\ 
+06:15:24 &   6 & Matching 	& Crisp 	& the same number of dots as the target 	    \\ 
+06:15:24 &  10 & Matching	& Vague		& about the same number of dots as the target	\\
+06:15:24 &  20 & Comparison & Crisp 	& fewer dots than the target             	    \\ 
+06:15:24 &  20 & Comparison	& Vague		& far fewer dots than the target				\\
 \noalign{\smallskip}\hline
@@
 \begin{description}
-	\item [Hypothesis 1] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
-	\item [Hypothesis 2] Comparison is easier for the reader than matching (main effect of selection)
-	\item [Hypothesis 3] Effects of vagueness differ depending on whether selection is matching or comparison (interaction effect selection x vagueness, and focussed comparisons at each level of selection).
+	\item [Hypothesis 1 (Crisp/Vague RT)] Vague instructions are easier for the reader than crisp ones.
+	\item [Hypothesis 2 (Comparison/Matching RT)] Comparison is easier for the reader than matching.
+	\item [Hypothesis 3 (Interaction)] The effect of vagueness differs depending on whether the selection task mandated by the instructions is matching or comparison.
 \end{description}
 
-\subsection{Results}
+\subsection{Results (Experiment 3)}
+
 40 volunteers participated.
-The results showed that vagueness was beneficial for comparison but detrimental for matching (the same as Experiment 2) even when no numbers were allowed in the instructions.
-Figure \ref{resultsE-exp-3} shows the means by condition.
-Our findings were as follows:
-
-\begin{description}
-	\item [Test of hypothesis 1] There was no significant main effect of vagueness ($\beta=-0.02$, $se=0.01$, $t=-1.5$, $p=0.1296$). 
-	\item [Test of hypothesis 2] There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=0.18$, $se=0.02$, $t=10.4$, $p<0.0001$).
-This effect was in the same direction as Experiment 3.
-	\item [Test of hypothesis 3] Vagueness did exert different effects depending on the selection task (main interaction effect of vagueness by selection $\beta=0.12$, $se=0.02$, $t=5.1$, $p<0.0001$). 
-Separate analyses of the effect of vagueness were conducted for the comparison task and for the matching task using Bonferroni-adjusted significance thresholds. 
-In the comparison task, vagueness resulted in faster response times ($\beta=-0.08$, $se=0.02$, $t=-4.3$, $p<0.0001$). 
-In the matching task vagueness slowed response times ($\beta=0.05$, $se=0.01$, $t=3.7$, $p=0.0004$). 
-These results are in the same direction as Experiment 3.
-\end{description}
-
-Once again, the cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
+Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 211 trials (2.8\% of the trials).
+The distribution of remaining response times was skewed with many long responses.
+These remaining response times were log-transformed which reduced this skew so that their distribution more closely approximated a normal distribution.
+Condition means for response times are plotted in Figure \ref{resultsE-exp-3}.
 
@@
 \includegraphics[width=\textwidth]{figures/Ee3-rtplot-1.pdf}
-\caption{Mean response times by condition for Experiment 3}
+\caption{Mean response times by condition for Experiment 3 where all instructions were verbal}
 \label{resultsE-exp-3}
 \end{figure}
+
+A linear mixed model was constructed for the logged response times, with sum-coded vagueness, selection task, and their interaction, and item as fixed effects, and per-participant intercepts and slopes for sum-coded vagueness, selection task, and their interaction, and item for random effects.
+
+\begin{description}
+	\item [Test of Hypothesis 1 (Crisp/Vague RT)] Vague instructions resulted in faster responses than crisp instructions on average.
+However this difference was not significant in the full model ($\beta=-0.015$, $se=0.0097$, $t=-1.58$, $p=0.119$).
+Using Levy's method \citep{Levy:MainEffectsInteractions} to test for main effects in the presence of higher-order interactions, by doing model comparison between a null model that included all interaction terms involving Vagueness but leaving out a term for the main effect of Vagueness, against a full model that differed only by including Vagueness as a main effect, showed that the full model was no better than the reduced model ($df=1$, $p=0.118$), consituting more evidence that Vagueness did not exert a significant main effect on response times.
+	\item [Test of Hypothesis 2 (Comparison/Matching RT)] Comparison instructions resulted in significantly faster responses than matching instructions ($\beta=0.176$, $se=0.0168$, $t=10.51$, $p<0.001$).
+	\item [Test of Hypothesis 3 (Interaction)] Although Vagueness did not exert a significant main effect, Vagueness did exert effects in its interaction with Selection task ($\beta=0.123$, $se=0.0166$, $t=7.42$, $p<0.001$), suggesting that Vagueness speeded response times in the comparison condition but slowed response times in the matching task.
+\end{description}
+
+\subsection{Discussion (Experiment 3)}
+
+The cost reduction account predicted that there should be a significant main effect of Vagueness such that responses would be faster for Vague intructions than for crisp instructions.
+We found that although there was a very small effect in that direction, the effect was not statistically significant.
+Models that differed only in the presence of Vagueness as a main effect were shown not to differ significantly in their explanatory value.
+
+However, the results also showed that vagueness did exert an influence on reaction times according to whether the task was comparison or matching: vagueness was beneficial for comparison and detrimental for matching (the same as Experiment 2) even when no numbers were allowed in the instructions.

### 2018-09-02 springer/current/e1.tex
Added separate analyses for V at each level of selection task for D and E.

@@
 
+\paragraph{\textbf{Response times}}
+
 30 participants were recruited.
@@
 These remaining response times were log-transformed which reduced this skew so that their distribution more closely approximated a normal distribution.
-Condition means for response times and proportion of borderline responses are given in Figure \ref{resultsC-exp-1}.
+Condition means for response times are given in Figure \ref{resultsC-exp-1-RT}.
 
@@
 \centering
-\includegraphics[trim = 20mm 0mm 30mm 0mm, clip, width=.49\textwidth]{figures/Ce1-rtplot-1.pdf}
-\includegraphics[trim = 20mm 0mm 25mm 0mm, clip, width=.49\textwidth]{figures/Ce1-blBarChart-1}
-\caption{Experiment 1 results: mean response times by condition and item, and counts of borderline case responses by condition.}
-\label{resultsC-exp-1}
+\includegraphics[width=\textwidth]{figures/Ce1-rtplot-1.pdf}
+\caption{Experiment 1 results: mean response times by condition and item}
+\label{resultsC-exp-1-RT}
 \end{figure}
@@
 
-However, given that the response time plot in the left panel of Figure \ref{resultsC-exp-1} shows that responses to 6:15:24 in the "crisp numeric" instructions condition were extremely fast relative to the "vague numeric" instructions to 6:15:24, the effects in the model of the full dataset could be driven by this difference.
+However, given that the response time plot in Figure \ref{resultsC-exp-1-RT} shows that responses to 6:15:24 in the "crisp numeric" instructions condition were extremely fast relative to the "vague numeric" instructions to 6:15:24, the effects in the model of the full dataset could be driven by this difference.
 A clearer picture of the effects of interest might be obtained by removing the 6:15:24 level of Item from the data set, and fitting the model to this restricted data.
@@
 
-A generalized linear mixed model \citet{jaeger2008categorical} was fit to the data for selection of the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant, as well as per-participant intercepts as random effects.
-The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Figure \ref{resultsC-exp-1}.
+\paragraph{\textbf{Borderline cases}}
+
+A generalized linear mixed model \citet{jaeger2008categorical} was fit to the data for the distribution of responses indicating the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant, as well as per-participant intercepts as random effects.
+The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Figure \ref{resultsC-exp-1-BL}.
 Participants chose the borderline square on $16.6\%$ of trials overall.
 
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=\textwidth]{figures/Ce1-blBarChart-1}
+\caption{Experiment 1 results: counts of borderline case responses by condition.}
+\label{resultsC-exp-1-BL}
+\end{figure}
+
 \begin{description}
-	\item [Hypothesis 4 (Response Type)] Participants were significantly more likely to choose the borderline option for vague instructions than for crisp instructions (21.9\% vs 11.3\%: $\beta=0.62$, $se=0.22$, $z=2.8$, $p=0.0059$).
+	\item [Test of Hypothesis 4 (Response Type)] Participants were significantly more likely to choose the borderline option for vague instructions than for crisp instructions (21.9\% vs 11.3\%: $\beta=0.62$, $se=0.22$, $z=2.8$, $p=0.0059$).
 Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%: $\beta=-3.35$, $se=0.23$, $z=-14.6$, $p<0.0001$).

### 2018-09-02 springer/current/e2.tex
Added separate analyses for V at each level of selection task for D and E.

@@
 Using Levy's method \citep{Levy:MainEffectsInteractions} to test for main effects in the presence of higher-order interactions, by doing model comparison between a null model that included all interaction terms involving Vagueness but leaving out a term for the main effect of Vagueness, against a full model that differed only by including Vagueness as a main effect, showed that the full model was no better than the reduced model ($df=1$, $p=0.676$), consituting more evidence that Vagueness did not exert a significant main effect on response times.
-	\item [Test of Hypothesis 2 (Comparison/Matching RT)] comparison instructions resulted in faster responses than matching instructions, and the difference was significant ($\beta=0.1618$, $se=0.0255$, $t=6.34$, $p<0.001$).
-	\item [Test of Hypothesis 3 (Interaction)] Although Vagueness did not exert a significant main effect, Vagueness did exert effects in interactions with some other variables: the interaction between Vagueness and Selection task  was significant ($\beta=0.1306$, $se=0.0205$, $t=6.38$, $p<0.001$), suggesting that Vagueness speeded RTs in the comparison condition but slowed them down in the matching task.
+	\item [Test of Hypothesis 2 (Comparison/Matching RT)] Instructions involving comparison resulted in faster responses than instructions involving matching, and the difference was significant ($\beta=0.1618$, $se=0.0255$, $t=6.34$, $p<0.001$).
+	\item [Test of Hypothesis 3 (Interaction)] The interaction between Vagueness and Selection task was significant ($\beta=0.1306$, $se=0.0205$, $t=6.38$, $p<0.001$), showing that vagueness exerted different effects on response times according to whether the instructions involved comparison or matching.
+When separate analyses were carried out testing for the effect of vagueness in comparison-only and in matching-only conditions, vague instructions led to significantly \emph{faster} responses than crisp instructions in the comparison-only conditions ($\beta=-0.071$, $se=0.020$, $t=-3.52$, $p<0.01$); and in the matching-only conditions vague instructions led to significantly \emph{slower} responses than crisp instructions ($\beta=0.062$, $se=0.021$, $t=2.91$, $p<0.01$).
 \end{description}

### 2018-09-02 springer/current/e3.tex
Added separate analyses for V at each level of selection task for D and E.

@@
 \includegraphics[width=.5\textwidth]{figures/Ee3-flow.jpg}
-\caption{Experiment 3 example stimulus}
+\caption{Experiment 3 example stimulus, showing the sequence of presentation, with the instruction screen at the top, followed by the presentation of the "target" array, followed by the stimulus that participants responded to.}
 \label{Experiment3examplestimulus}
@@
 \centering
-\caption{Experiment 3 instructions arranged by condition. The instructions given in the table started with ``Choose a square with \ldots"}
+\caption{Experiment 3 instructions for the smallest array, arranged by condition. The instructions given in the table started with ``Choose a square with \ldots"}
 \label{instructionsE-exp-3}
@@
 	\item [Test of Hypothesis 2 (Comparison/Matching RT)] Comparison instructions resulted in significantly faster responses than matching instructions ($\beta=0.176$, $se=0.0168$, $t=10.51$, $p<0.001$).
-	\item [Test of Hypothesis 3 (Interaction)] Although Vagueness did not exert a significant main effect, Vagueness did exert effects in its interaction with Selection task ($\beta=0.123$, $se=0.0166$, $t=7.42$, $p<0.001$), suggesting that Vagueness speeded response times in the comparison condition but slowed response times in the matching task.
+	\item [Test of Hypothesis 3 (Interaction)] The interaction between Vagueness and Selection task was significant ($\beta=0.123$, $se=0.0166$, $t=7.42$, $p<0.001$), showing that Vagueness had a different influence on response times when the instructions mandated comparison versus when the instructions mandated matching.
+When separate analyses were carried out testing for the effect of vagueness in comparison-only and in matching-only conditions, vague instructions led to significantly \emph{faster} responses than crisp instructions in the comparison-only conditions ($\beta=-0.077$, $se=0.018$, $t=-4.31$, $p<0.001$); and in the matching-only conditions vague instructions led to significantly \emph{slower} responses than crisp instructions ($\beta=0.047$, $se=0.013$, $t=3.71$, $p<0.001$).
 \end{description}

### 2018-09-02 springer/current/general-discussion.tex
Added separate analyses for V at each level of selection task for D and E.

@@
 Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
-Years of research on the logic of vagueness -- giving rise to such techniques as Partial Logic \citet[e.g.,]{Fine}, Probabilistic Logic \citet{Edgington}, and Fuzzy Logic \citet{Zadeh} -- have primed the research community to expect that some special utility of vagueness is an important part of the answer, but our findings call this expectation into question.
+Years of research on the logic of vagueness -- giving rise to such techniques as Partial Logic \citep[e.g.,]{Fine}, Probabilistic Logic \citep{Edgington}, and Fuzzy Logic \citep{Zadeh} -- have primed the research community to expect that some special utility of vagueness is an important part of the answer, but our findings call this expectation into question.
 

### 2018-09-03 springer/current/related-work.tex
Update related-work.tex

@@
-Charting the utility of vagueness is the attested aim of a small number of studies, but most of these have focussed on vagueness in a different sense, and focussing on different types of benefits for hearers.
+A number of answers to Litman’s question have been proposed, but as we have argued elsewhere [REF KvD JPL and REF book], some of the most often cited answers do not hold up to scrutiny. For example,
+%
+\begin{itemize}
+\item It has been proposed that the existence of vague words makes a language more efficient [REF Barwise Cooper].
+The idea is that vagueness allows the use of one and the same word (e.g., ``big”) in different situations, which require different standards.
+For instance, the vague word ``’big” can denote a mouse a well as an elephant.
+— The idea is plausible, but the problem is that ``big" is efficient not because it is vague (i.e., not because it allows borderline cases) but because it is context dependent.
+To see this, it suffices to realise that superlatives are not vague, but they {\em are} context dependent: to be ``the biggest $x$" ascribes different sizes to the referent depending on the word $x$.
+It is for this reason that we can use the word ``biggest” to talk about the biggest mouse, the biggest elephant, and so on.
+Context-dependence, however, does not imply vagueness.
+%
+\item It has been proposed that the ability to combine a statement of quantity (along some dimension) with an evaluative statement [REF Veltman].
+The idea is that when we say “The patient has a fever”, we do not merely assert that her temperature is above a certain point in the temperature dimension, we also imply that the deviation is large enough to be clinically significant (i.e., something’s not enirely right with the patient).
+— The problem with this equally plausible idea is that evaluation does not imply vagueness.
+For example, the medical term “obese” is evaluative (i.e., it’s not healthy to be obese), yet its standard medical definition in terms of Body Mass Index is perfectly crisp.
+\end{itemize}
+%
+The findings on which we are reporting in the present paper will follow a similar pattern: we will explore a possible answer to Litman’s question, only to conclude that it does not stand up to experimental scrutiny.
+
+Charting the utility of vagueness is also the attested aim of a small number of experimental studies.
+However, to the best of our knowledge, few of these studies have truly focussed on vagueness in the sense on which we are focussing (i.e., they did not address Litman's challenge).
 Two recent studies can illustrate both issues.

### 2018-09-06 docs/D_exp_2.Rmd
github please build from /docs/

--- 2018-08-25 experiments/D_exp_2/writeup/D_exp_2.Rmd
@@
     toc: true
-    toc_float: false
+    theme: default
+    css: styles.css    
 ---

### 2018-09-06 docs/D_exp_2_functions/preprocessing.R

@@ -1,0 +1,207 @@
+preprocessing = function(root_dir, data_dir) {
+  
+  # The policy as of Aug 2018 is that:
+  # RT=0 is stripped, and one case with a very long time RT=49871 is stripped.
+  
+  dat <- read.delim(file.path(root_dir, data_dir, "data_raw.txt"), stringsAsFactors = FALSE)
+  
+  # declare local variables
+  number_of_valid_subjects <- 38
+  number_of_rows <- 7296
+  number_of_trials_per_subject <- number_of_rows / number_of_valid_subjects # 192
+  
+  # sort out borderline responses into expected near and far
+  dat$RESPONSE <- as.character(dat$RESPONSE)
+  for (row in 1 : nrow(dat) ) {
+    switch(dat[row,'RESPONSE'],
+           'LEFT' = {dat[row, 'choice'] <- dat[row, 'Left']},
+           'MIDDLE' = {dat[row, 'choice'] <- dat[row, 'Mid']},
+           'RIGHT' = {dat[row, 'choice'] <- dat[row, 'Right']}         
+    )
+  }
+
+  dat$crossed=paste('Con',dat$Condition,':Quan',dat$Quantity,':Item',dat$Item,sep="")
+  
+  dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseExpected'] <-  6 
+  dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseNear']     <- 15  
+  dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseFar']      <- 24
+  
+  dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseExpected'] <- 16
+  dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseNear']     <- 25  
+  dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseFar']      <- 34
+  
+  dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseExpected'] <- 26
+  dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseNear']     <- 35
+  dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseFar']      <- 44
+  
+  dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseExpected'] <- 36
[…]

### 2018-09-06 experiments/C_exp_1/C_exp_1_functions/annotate_the_raw_data.R

@@ -1,38 +1,0 @@
-annotate_the_raw_data = function (dat) {
-  
-  # annotate vagueness
-  dat[ dat$Condition==1 , 'Vagueness'] <- 'Vague'
-  dat[ dat$Condition==2 , 'Vagueness'] <- 'Crisp'
-  dat[ dat$Condition==3 , 'Vagueness'] <- 'Vague'
-  dat[ dat$Condition==4 , 'Vagueness'] <- 'Crisp'
-  dat$Vagueness <- as.factor(dat$Vagueness)
-  
-  # annotate use of a number / word in the instructions
-  dat[ dat$Condition==1 , 'Number'] <- 'Numeric'
-  dat[ dat$Condition==2 , 'Number'] <- 'Numeric'
-  dat[ dat$Condition==3 , 'Number'] <- 'Verbal'
-  dat[ dat$Condition==4 , 'Number'] <- 'Verbal'
-  dat$Number <- as.factor(dat$Number)
-  
-  # annotate Item
-  dat$Item <- factor(dat$Item, levels=c(1,2,3,4), labels=c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
-  
-  # annotate Order
-  dat$Order <- factor(dat$Order, levels=c(1,2), labels=c('LtoR','RtoL'))
-  
-  # annotate Quantity
-  dat$Quantity <- factor(dat$Quantity, levels=c(1,2), labels=c('Small','Large'))
-  
-  # annotate Subject
-  dat$Subject <- factor(paste("s",sprintf("%02d",dat$Subject),sep=""))
-  
-  # annotate Trial index for each participant (1:256 trials for 30 subjects)
-  dat$Trial = rep(x = 1:256, times = 30)
-  
-  # add an Observation index, over the whole data set (1:7680)
-  dat$Obs <- 1:7680
-  
-  return(dat)
-}  
-
-

### 2018-09-06 experiments/C_exp_1/C_exp_1_functions/pretty_coef_plot.R

@@ -1,6 +1,0 @@
-pretty_coef_plot = function(modl, mycaption) {
-  dwplot(tidy(modl, effects="fixed")) + 
-    theme(panel.background = element_rect(colour="black", fill="white"), panel.grid = element_blank()) + 
-    geom_vline(xintercept=0, alpha=0.2) + 
-    ggtitle(paste(mycaption, 'coefficient estimates'), subtitle = paste('deviation from an intercept of', round(summary(modl)$coefficients[1,1],2), 'log RT, with 95% CI.'))
-}

### 2018-09-06 experiments/C_exp_1/C_exp_1_functions/pretty_coef_table.R

@@ -1,101 +1,0 @@
-pretty_coef_table = function(modl, mycaption){
-  
-  coef_table <- tidy(modl, effects="fixed") 
-
-  # be sensitive to whether the model is an lmer or a glmer 
-  
-  if (class(modl) == "lmerModLmerTest") {
-    
-    # sort the names out
-    names(coef_table) <- c("effect", "term", "coefficient", "std.error", "df", "t_value", "p_value_raw")
-    coef_table$p_value <- as.character("")
-    coef_table$star <- as.character("")
-    
-    # categorise p values and add significance stars
-    # "\u2007" is a space the same width as a numeric, that is preserved in printing
-    # "\u0020" is just a space
-    for (i in 1:nrow(coef_table)) {
-      if (!is.na(coef_table$p_value_raw[i])) {
-        if (coef_table$p_value_raw[i] >= 0     & coef_table$p_value[i] < 0.001) {coef_table$p_value[i] <-    "<0.001";                        coef_table$star[i] <- paste(sep="","\uFF0A","\uFF0A","\uFF0A")}
-        if (coef_table$p_value_raw[i] >= 0.001 & coef_table$p_value[i] < 0.01 ) {coef_table$p_value[i] <-    paste(sep="","<0.01", "\u2007"); coef_table$star[i] <- paste(sep="","\uFF0A","\uFF0A"," ")}
-        if (coef_table$p_value_raw[i] >= 0.01  & coef_table$p_value[i] < 0.05 ) {coef_table$p_value[i] <-    paste(sep="","<0.05", "\u2007"); coef_table$star[i] <- paste(sep="","\uFF0A"," "," ")}
-        if (coef_table$p_value_raw[i] >= 0.05  ) {coef_table$p_value[i] <-    round(coef_table$p_value_raw[i],3); coef_table$star[i] <-  ""}
-      }
-    }
-    # un-tibble the table
-    coef_table <- coef_table %>% as.data.frame()
-    
-    # convert numerics to formatted chars for pretty printing
-    coef_table$coefficient <- format(x=coef_table$coefficient, nsmall=2, digits=2)
-    coef_table$std.error   <- format(x=coef_table$std.error, nsmall=2, digits=2)
-    coef_table$df          <- format(x=coef_table$df, nsmall=2, digits=1)
-    coef_table$t_value     <- format(x=coef_table$t_value, nsmall=2, digits=2)
[…]

### 2018-09-08 docs/D_exp_2.R
recompiled with new titles; purled all

@@ -1,0 +1,131 @@
+## ---- echo=FALSE, message=FALSE------------------------------------------
+library(knitr)
+library(tidyverse)
+library(broom)
+library(broom.mixed)
+library(lme4)
+library(lmerTest)
+library(dotwhisker)
+library(kableExtra)
+library(LMERConvenienceFunctions)
+library(data.table)
+library(grid)
+library(gridExtra)
+source("D_exp_2_functions/preprocessing.R")
+source("D_exp_2_functions/pretty_coef_table.R")
+source("D_exp_2_functions/summarySEwithin2.R")
+root_dir <- getwd()
+data_dir <- "D_exp_2_data"
+
+## ------------------------------------------------------------------------
+dat_full <- preprocessing(root_dir, data_dir)
+dat <- subset(dat_full, subset=RT>0 & RT<25000) # loses 2 RTs that were judged to be outliers 25978 49871
+dat_borderline <- dat
+dat <- perSubjectTrim.fnc(dat, response='RT', subject='Subject', trim = 2.5)$data
+dat$RT_log <- log(dat$RT)
+
+## ---- echo=FALSE---------------------------------------------------------
+instructions_table <- 
+  dat %>%
+    select(Item, Quantity, Vagueness, Selection, Instruction) %>%
+    unique() %>%
+    arrange(Item, Quantity, Vagueness, Selection) %>%
+    spread(key=Vagueness, value=Instruction)
+instructions_table %>% 
+  kable(align='cccll', caption="Full table of instructions") %>% 
+      kable_styling(full_width = F, position = "left", font_size = 11) 
+
+## ---- cho=FALSE----------------------------------------------------------
+dat_plot <- summarySEwithin2(dat, measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
+
+## ---- "EXP_D_RT_means_plot", fig.width=7, fig.height=3, echo=FALSE-------
+dodge = position_dodge(width=0.2)
+dat_plot$Condition <- as.factor(paste(sep=' ', dat_plot$Selection, dat_plot$Vagueness))
+ggplot(dat_plot, aes(y=RT_logNormed, x=Item, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci, group=Condition, shape=Condition, fill=Condition)) +
+  geom_line(position=dodge) +
+  geom_errorbar(width=.2, position=dodge) +
+  geom_point(size=2, position=dodge) +
[…]

### 2018-09-08 docs/E_exp_3.R

@@ -1,0 +1,131 @@
+## ---- "libs and functions", echo=FALSE, message=FALSE--------------------
+library(knitr)
+library(tidyverse)
+library(broom)
+library(broom.mixed)
+library(lme4)
+library(lmerTest)
+library(dotwhisker)
+library(kableExtra)
+library(LMERConvenienceFunctions)
+library(data.table)
+library(grid)
+library(gridExtra)
+source("E_exp_3_functions/preprocessing.R")
+source("E_exp_3_functions/pretty_coef_table.R")
+source("E_exp_3_functions/summarySEwithin2.R")
+root_dir <- getwd()
+data_dir <- "E_exp_3_data"
+
+## ---- get_data-----------------------------------------------------------
+dat_full <- preprocessing(root_dir, data_dir)
+dat <- subset(dat_full, RT>0 & RT<9500)
+dat_borderline <- dat
+dat <- perSubjectTrim.fnc(dat, response='RT', subject='Subject', trim = 2.5)$data
+dat$RT_log <- log(dat$RT)
+
+## ---- "Full table of instructions", echo=FALSE---------------------------
+instructions_table <- 
+  dat %>%
+    select(Item, Target, Quantity, Vagueness, Selection, Instruction) %>%
+    unique() %>%
+    arrange(Item, Target, Quantity, Vagueness, Selection) %>%
+    spread(key=Vagueness, value=Instruction)
+instructions_table %>% 
+  kable(align='ccccll', caption="Full table of instructions") %>% 
+      kable_styling(full_width = F, position = "left", font_size = 11) 
+
+## ---- "make data for rt plot", echo=FALSE--------------------------------
+dat_plot <- summarySEwithin2(dat, measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
+
+## ---- "EXP_E_rt_means_plot", fig.width=7, fig.height=3, echo=FALSE-------
+dodge = position_dodge(width=0.2)
+dat_plot$Condition <- as.factor(paste(sep=' ', dat_plot$Selection, dat_plot$Vagueness))
+ggplot(dat_plot, aes(y=RT_logNormed, x=Item, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci, group=Condition, shape=Condition, fill=Condition)) +
+  geom_line(position=dodge) +
+  geom_errorbar(width=.2, position=dodge) +
+  geom_point(size=2, position=dodge) +
[…]

### 2018-09-12 chapter-latex/abstract.tex
added the springer chapter latex --post-review --post-response-to-review --pre-send-to-editor

--- 2018-04-22 papers/2018-v-book/v-book.tex
@@
-\documentclass[graybox,envcountchap,sectrefs%
-,footinfo
-%,referee
-]{svmono}
-% Native packages
-\usepackage{mathptmx}
-\usepackage{helvet}
-\usepackage{courier}
-\usepackage{type1cm}         
-\usepackage{makeidx}		% allows index generation
-\usepackage{graphicx}		% standard LaTeX graphics tool
-\usepackage{multicol} 		% used for the two-column index
-\usepackage[bottom]{footmisc}	% places footnotes at page bottom
-% User packages
-\usepackage{natbib}
-\usepackage{lipsum}
-\usepackage{placeins}
-\makeindex             			% used for the subject index
-                       				% please use the style svind.ist with
-                       				% your makeindex program
-
-\begin{document}
-
-\author{Kees van Deemter and Matthew Green}
-\title{Vagueness \& Rationality}
-\subtitle{The Elusive Benefits of Vagueness: the Evidence So Far}
-\maketitle
-
-\frontmatter
-\tableofcontents
-
-% BibTeX users please use
-\bibliographystyle{spbasic} %plain, harvard, apalike, chicago, astron, authordate, natbib.
-\bibliography{bibliography}
-
-\backmatter
-\Extrachap{Glossary}
-%\runinhead{glossary term} Write here the description of the glossary term. Write here the description of the glossary term. Write here the description of the glossary term.
-
-\runinhead{natural language generation} the process of generating natural language from non-natural language input.
-
-\printindex
-\end{document}
+Much of everyday language is vague, even in situations where vagueness could have been avoided (i.e., where vagueness is used ``strategically'').
+Yet the benefits of vagueness for hearers and readers are proving to be elusive.
+We discuss a range of earlier controlled experiments with human participants, and we report on a new series of experiments that we conducted in recent years.
+These experiments, which focus on vague expressions that are part of referential noun phrases, aim to separate the utility of vagueness (as defined by the existence of borderline cases) from the utility of other factors that tend to co-occur with vagueness.
+Having presented the evidence, we argue that the evidence supports a view where the benefits that vague terms exert are due to other influences, and not to vagueness itself.

### 2018-09-12 chapter-latex/e1.tex
added the springer chapter latex --post-review --post-response-to-review --pre-send-to-editor

--- 2018-09-02 springer/current/e1.tex
@@
 \begin{itemize}
-	\item In the vague numerical condition the instruction was "Choose a square with about 10 dots" -- none of the arrays contained exactly 10 dots, but 10 is closer to 6 than it is to 15, making 6 a better response to that instruction, 15 a borderline response, and 24 an extreme response.
-	\item In the vague verbal condition we used "Choose a square with few dots". We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
-	\item In the crisp numerical condition we used "Choose the square with 6 dots".
+	\item In the vague numerical condition the instruction was ``Choose a square with about 10 dots" -- none of the arrays contained exactly 10 dots, but 10 is closer to 6 than it is to 15, making 6 a better response to that instruction, 15 a borderline response, and 24 an extreme response.
+	\item In the vague verbal condition we used ``Choose a square with few dots". We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
+	\item In the crisp numerical condition we used ``Choose the square with 6 dots".
 The smaller flanking array always contained exactly the specified number of dots.
 We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
-	\item For crisp verbal, we used "Choose the square with the fewest dots". We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
+	\item For crisp verbal, we used ``Choose the square with the fewest dots". We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
 \end{itemize}
 
-On each trial, first the referring expression that constituted the instruction for that trial was displayed (e.g., "Choose a square with about 10 dots").
+On each trial, first the referring expression that constituted the instruction for that trial was displayed (e.g., ``Choose a square with about 10 dots").
 Participants then pressed a key to indicate that they had read the instruction.
@@
 In this experiment, no feedback was given.
-This was because, in the vague conditions, we did not regard any response as "correct" or "incorrect", but instead as "expected response"; "borderline response"; and "extreme response", and we did not want to draw participants' attention to this distinction explicitly.
+This was because, in the vague conditions, we did not regard any response as ``correct" or ``incorrect", but instead as ``expected response"; ``borderline response"; and ``extreme response", and we did not want to draw participants' attention to this distinction explicitly.
 Which choice the participant made was recorded for analysis.

### 2018-09-12 chapter-latex/general-discussion.tex
added the springer chapter latex --post-review --post-response-to-review --pre-send-to-editor

--- 2018-09-02 springer/current/general-discussion.tex
@@
 Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
-Years of research on the logic of vagueness -- giving rise to such techniques as Partial Logic \citep[e.g.,]{Fine}, Probabilistic Logic \citep{Edgington}, and Fuzzy Logic \citep{Zadeh} -- have primed the research community to expect that some special utility of vagueness is an important part of the answer, but our findings call this expectation into question.
+Years of research on the logic of vagueness -- giving rise to such techniques as Partial Logic \citep[e.g.,][]{Fine}, Probabilistic Logic \citep{Edgington}, and Fuzzy Logic \citep{Zadeh} -- have primed the research community to expect that some special utility of vagueness is an important part of the answer, but our findings call this expectation into question.
 
@@
 For example, an \emph{expert} may prefer to read expression (a), because it gives her the most detailed information on which to base her decisions.
-On the other hand, expression (d) (``warm'') is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make (d) more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm'' (see our section 1 and  \citeauthor{peters2009bringing}), which is especially important if the hearer is unfamiliar with the metric used.
+On the other hand, expression (d) (``warm'') is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make (d) more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm'' (recall our section 1.2), which is especially important if the hearer is unfamiliar with the metric used.
 These considerations suggest that {\em non-experts} might prefer expression (d).

### 2018-09-12 chapter-latex/introduction.tex
added the springer chapter latex --post-review --post-response-to-review --pre-send-to-editor

--- 2018-08-02 springer/submission/v-book-2017.tex
@@
-\documentclass[%
-%			% choose man or doc immediately below
-%doc,	    % doc is single spaced
-man,		% man is double spaced
-floatsintext,%
-apacite%
-]{apa6} % I'm using a local (as in, in this folder) apa6.cls which edits footnotesep from 16pt which is the real apa6 in 2018 back to 6pt in order to make it line up with the pdf generated for the 2017 submission
-\usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{color}
-\usepackage{booktabs}
-\usepackage{layout}
-\usepackage{apacite}
-\usepackage{datetime}
-
-\title{The Elusive Benefits of Vagueness \protect \\ Evidence from Experiments}
-\date{\today}
-\authornote{This is a layout-identical version of the document submitted for review in May 2017. LaTeX compiled \today~at \currenttime}
-\shorttitle{Is Vagueness Beneficial for Hearers?}
-\twoauthors{Matt Green}{Kees van Deemter}
-\twoaffiliations{Computing Science\\ University of Aberdeen}{Computing Science\\ University of Aberdeen}
-
-\abstract{Much of everyday language is vague, even in situations where vagueness could have been avoided (i.e., where vagueness is used "strategically").
-Yet the benefits of vagueness for hearers and readers are proving to be elusive.
-We discuss a range of earlier controlled experiments with human participants, and we report on a new series of experiments that we conducted in recent years.
-These experiments, which focus on vague expressions that are part of referential noun phrases, aim to separate the utility of vagueness (as defined by the existence of borderline cases) from the utility of other factors that tend to co-occur with vagueness.
-Having presented the evidence, we argue that the evidence supports a view where the benefits that vague terms exert are due to other influences, and not to vagueness
-itself.}
-
-\begin{document}
-\maketitle   
-\setcounter{secnumdepth}{3}
-\section{Introduction}\label{introduction}
-
 In most academic use, the word `vagueness' has a specific meaning.
-Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1997vagueness}, also \citeA{EgreKlinedinst}).
+Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \citep[p.~4]{keefe1997vagueness}, also \citet{EgreKlinedinst}.
 The crucial criterion is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise'' \cite[p.\ 1]{lipmanvague}.
+By contrast, a word is vague if it is not precise'' \citet[p.~1]{lipmanvague}.
 A typical example is the word ``tall'', as applied to people for example, because here is no precise, known height which separates those who are tall from those who are not.
@@
 
-Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions \cite{Lipman:2000fk, lipmanvague}, which are
+Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions \citet{Lipman:2000fk, lipmanvague}, which are
 often used even in situations where the speaker could have used an expression that is not vague (i.e., crisp); in these situations we say that vagueness is used \emph{strategically}.
-By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss'' \cite<>[p.~1]{lipmanvague}.
+By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss'' \citet[][p.~1]{lipmanvague}.
 Lipman explains the point by means of a scenario in which a speaker describes a person to a hearer, who needs to identify that person in the arrivals hall of of an airport.
 In such a scenario, a precise description of the person's height (e.g., ``The person's height is 187.96 cm'') would be more useful than a vague one (``The person is tall'').
-Lipman uses this scenario to explain why standard game theory models of communication \cite<e.g.,>{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs.
+Lipman uses this scenario to explain why standard game theory models of communication \citep[e.g.,][]{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs.
 
@@
 Lipman asked, essentially, what these advantages might be.
-Several tentative answers to Lipman's question have been offered  \cite<see>{van2009utility, vanDeemterBook}.
-One of the most promising answers appears to be the idea that vague expressions are easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite<e.g.,>[]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
-For example, \citeA<>[p.\ 11]{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze''.
+Several tentative answers to Lipman's question have been offered \citep[see][]{van2009utility, vanDeemterBook}.
+One of the most promising answers appears to be the idea that vague expressions are easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \citep[e.g.,][]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
+For example, \citeauthor{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze'' (\citeyear[][p.~11]{lipmanvague}).
 We shall refer to this as the \emph{cost reduction} hypothesis.
 
-This article brings an experimental approach to these issues, focussing on vagueness in descriptions (e.g., "the square with few dots") and its effect on the hearer's ability to act on a given description, as measured by the time that it takes hearers to click on the referent of a description\footnote{Other metrics could have been chosen, such as hearers' ability to remember information, for example, or error rates.
+This article brings an experimental approach to these issues, focussing on vagueness in descriptions (e.g., ``the square with few dots'') and its effect on the hearer's ability to act on a given description, as measured by the time that it takes hearers to click on the referent of a description\footnote{Other metrics could have been chosen, such as hearers' ability to remember information, for example, or error rates.
 Although error rates play a minor role in the present paper, for reasons that will become clear, we focus on response times in particular.}.
@@
 Instead, it is other factors, such as the presence of an overt numerical expression in the description, that proved to be decisive.
-We believe that, despite the fact that our experiments are unavoidably focussed only on a specific class of vague expressions (since any experiment can only deal with a limited number of different stimuli), these findings are potentially important, because they call into question whether "strategic" vagueness (i.e., vagueness where the speaker had a choice, because she could have been produced a crisp expression instead) has any advantages at all.
+We believe that, despite the fact that our experiments are unavoidably focussed only on a specific class of vague expressions (since any experiment can only deal with a limited number of different stimuli), these findings are potentially important, because they call into question whether ``strategic'' vagueness (i.e., vagueness where the speaker had a choice, because she could have been produced a crisp expression instead) has any advantages at all.
 In other words, returning to Lipman's question, it is possible that vagueness has evolved partly as a necessary evil (e.g., because of the limits of observation and prediction) and partly as a side effect of other factors.
-
-\section{Related work}
-
-Charting the utility of vagueness is the attested aim of a small number of studies, but most of these have focussed on vagueness in a different sense, and focussing on different types of benefits for hearers.
-Two recent studies can illustrate both issues.
-
-In a series of studies of behaviour modification, \citeA<>{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss.
-In the weight loss study, participants were told that the study was designed to test the validity of a new (actually fictitious) health index, the HHI (Holistic Health Index).
-They were told that an ideal HHI score lies in the range of 45 to 55.
-In a longitudinal study, participants submitted their weight to a computer each week.
-Participants were told that two algorithms would be used to compute their HHI, and that the two might give different values initially, in which case the true score lay between the two values.
-In one condition, which the authors called the precise condition, the two algorithms gave the same score.
-In the other condition, which the authors called the vague condition, one algorithm added 3\% to the score while the other algorithm subtracted 3\% from the score, yielding a range of values whose midpoint was the same as the two values given in the precise condition.
-
-One group of participants was given HHI scores in the ideal range: for this group their weight loss did not differ depending on whether they were given vague or precise HHI values.
-However for the other group, who were given HHI scores outside the ideal range, their weight loss was significantly greater if they were given vague HHI scores than if they were given precise HHI scores.
-The authors explain the improvement in the vague condition for this group as resulting from the participants' freedom to think of themselves a positioned on one end of the range - the end closest to the ideal HHI scores.
-This ``illusion of proximity'' \cite<>[p.~4]{Mishra01042011} to the goal is argued to allow participants to generate positive expectancies that lead to behaviours that improve performance.
-In contrast, in the precise conditions, participants did not have this freedom of interpretation, and could not distort the information to bring about the beneficial \emph{illusion of proximity}.
-These results are interesting, and of obvious potential practical importance.
-We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \cite{keefe1997vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases.
-In the terminology of \citeA{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}.
-Furthermore, the experiments of \citeA{Mishra01042011} did not explore benefits in terms of processing cost, but in terms of long-term behaviour change.
-
-Similar issues arise from the work of \citeA{peters2009bringing}.
-The authors carried out a series of studies where participants were required to rate hospitals based on various sources of information about quality of care.
-There was a between-subjects manipulation based on numeracy.
-The format of the information was manipulated within subjects: either numbers only were presented, or both numbers and evaluative categories were presented (e.g., \emph{Poor}, \emph{Fair}, \emph{Good}, \emph{Excellent}, with crisp visual boundary lines between the categories).
-Results showed that, for low-numeracy participants, the presence of evaluative categories resulted in a diminished influence of an irrelevant affective state on the ratings.
-For all participants, the presence of evaluative categories resulted in better decisions and in a greater use of the most important and reliable types of information, such as survival rates.
-
-It is, however, questionable whether the ``evaluative categories" manipulation in this study can be considered a manipulation of vagueness.
-Certainly, terms like \emph{Fair} admit the possibility of borderline cases.
-However, given that the boundaries between the categories were marked crisply, and that therefore the categories mapped crisply to numerical values, it becomes doubtful whether any borderline cases could be conceived to arise in fact.
-For example, \emph{Fair} was mapped to 60\% -- 70\% for the variable \emph{percentage of heart attack patients given recommended treatment (ACE inhibitor)}.
-Accordingly, rather than the vagueness of categories such as \emph{Poor}, Peters et al.
-emphasise the evaluative content inherent in these categories, and the affective potential of the evaluative content rather than the vagueness of the terms like {\em Fair}.
-
-\subsection{Our Approach to the Problem}\label{method}
-
-Are vague expressions processed more easily by readers than crisp ones?
-Like Lipman, we focus on situations where numerical information is used in order to identify a referent.
-Reference, in other words, will be the communicative task on which we focus, partly because of the interest that this topic has recently drawn in various areas of Cognitive Science \cite{vanDeemterCMR}.
-By looking at one specific type of vagueness, we will be able to investigate the costs and benefits of vagueness relatively thoroughly.
-Whether our findings generalise to other uses of vagueness is a question on which we will speculate in the final section of this chapter.
-
-We have chosen a narrative strategy in which we address a sequence of four experiments with human reader chronologically, explaining how each experiment helped us refine our research question.
-In order to do justice to our findings, we need to describe these experiments in a fair amount of detail.
-
-Let us start by explaining the task that was given to the participants in our experiments.
-We used a {\em speeded forced choice} task to compare the processing costs of different references to quantities.
-In this context, speed and accuracy of responses are the key dimensions on which the different references can be compared.
-Each stimulus in the experiments was a set of dot arrays containing various number of dots, together with a preceding instruction (in the form of a referring expression) to choose one of the arrays with respect to its cardinality.
-The participant was asked to respond as quickly as possible while avoiding errors.
-We manipulated the instructions and the arrays in several ways across the four experiments.
-
-All the experiments shared the following properties: 
-Stimuli were created using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.
-The position of the dots was randomised per-trial.
-The order in which trials were presented was randomised per-participant. 
-There were 256 trials, presented in 4 blocks of 64 each, between which the participant could rest. 
-A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants and recorded responses. 
-Participants were recruited using email lists at the University of Aberdeen, and paid ten pounds for participating.
-All participants self-reported fluency in English, and had normal, or corrected-to-normal vision.
-The experiment was conducted in a quiet room. 
-Participants were asked to respond as quickly as possible while avoiding errors.
-There was a block of practice trials after which participants could ask any questions, following which the experimenter left the room.
-All $p$ values reported for linear models were calculated using the R package \emph{lmerTest} \cite{lmerTest}. The complete data and analysis of our first experiment are available from 
-{\tt http://homepages.abdn.ac.uk/k.vdeemter/pages/vagueness-data-analysis/e1.html}, and analogous for experiments 2-4.\\[2ex]
-
-\noindent When the distance grows between two numbers, they become more easily distinguishable: the \emph{numerical distance effect} has been shown for comparing the cardinality of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene1996}.
-We manipulated the number of dots in each array such that some sets of arrays had smaller numerical distances and others had larger numerical distances.
-Where a number was mentioned in the instructions, it was always in the form of an Arabic numeral.
-When two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}.
-We controlled which side the smaller number appeared on to avoid systematic influences of this effect.
-
-There is abundant evidence \cite<e.g.,>{trick1994small} that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities.
-We performed a pilot experiment \cite{our-earlier-paper} in which we were able to confirm this finding in the experimental settings on which we are focussing in this paper.
-We found that, when participants were confronted with a stimulus consisting of two squares containing different numbers of dots\footnote{Such a stimulus is referred to hereafter as consisting of a set of {\em dot arrays}.
-The number of dots in an array is referred to as its cardinality.
-The physical arrangement of dots in each array is irregular.}, instructions of the form \emph{Choose the square with \emph{n} dots} led to consistently faster response times than instructions of the form \emph{Choose the square with many/few dots} when $2 \leq n \leq 5$; the converse was true for $n>5$.
-Given these findings, we henceforth focussed our studies on non-subitizable numbers, because it is there that vagueness is expected to have benefits.
-
-\section{Experiment 1: initial experiment}
-
-We used a forced choice task to elicit responses to crisp or vague forms of instructions that required the participant to choose one of two dot arrays by referring to its cardinality.
-The participant was presented with an instruction in the form \emph{Choose the square with \ldots dots}.
-Then a set of two dot arrays was presented, each in the form of a square containing some number of dots.
-The participant was required to identify the dot array that corresponded with the instruction, by pressing the appropriate key, as quickly as possible while avoiding errors. 
-Response time and accuracy were recorded for analysis.
-
-We manipulated how discriminable the dot arrays were by varying the numerical distance between them. 
-One array always contained 25 dots: the other contained either 5, 10, 15, 20, 30, 35, 40, or 45 dots. 
-This gave us numerical distances of 5, 10, 15, and 20, with smaller numerical distances resulting in less discriminable arrays and larger distances resulting in more discriminable arrays.
-
-Our main manipulation was of the vagueness of the instruction, with two levels, \emph{crisp} and \emph{vague}.
-Assuming the dot array [5, 25], and the instruction referring to the smaller cardinality, the \emph{crisp} instruction was \emph{Choose the square with 5 dots} and its \emph{vague} counterpart was \emph{Choose the square with few dots}.
-
-\subsection{Hypothesis}
-\noindent The main aim of this experiment was to test the following hypothesis:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] There exists a main effect advantage for vagueness: vague instructions impose a lower cognitive load for the comprehender than crisp alternatives.
-\end{APAenumerate}
-}
-
-\subsection{Experimental Method}
-
-%Twenty participants were recruited, aged between 18 and 45 with a median age of 26.
-On each trial a participant was presented with an instruction to choose one of two dot arrays on screen by reference to its cardinality. 
-Following a keypress to indicate that the participant had read the instruction there was a central fixation cross for 1000 ms, and a blank screen for 500 ms, followed by the array (without repetition of the referring expression).
-The arrays would stay on screen until the participant responded (there was no timing-out). 
-Response time was measured as the latency between the presentation of the arrays, and the keypress identifying the choice: in this way, the response time was separated from time spent reading the instructions, which is important since we are only interested in the former.
-A response was counted as erroneous if the square with the wrong number of dots was chosen (when the instruction contained a number); 
-if the square with the larger number of dots was selected (when the instruction was \emph{Choose the square with few dots}); 
-or if the square with the smaller number of dots was selected (when the instruction was \emph{Choose the square with many dots}). 
-No feedback was given on correct trials, but there was feedback on error trials in the form of the word ``\textsc{wrong!!}'' which flashed on screen.
-Hypothesis H1 was tested looking both at response times and at error rates.
-
-\subsection{Results}
-
-\begin{figure}[htbp]
-\fitfigure{images/e1RTandErrorRatesplot-1}
-\caption{Experiment 1 results: response times and error rates.}
-\label{resultse1}
-\end{figure}
-
-{\bf Response times} (RTs) for trials with erroneous responses were discarded, leading to the loss of 354 trials from 5120, representing 6.9\% of the trials. 
-The correct response RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of a further 160 trials, or 3.4\% of the remaining correct responses. 
-Means for response times and error rates are given in Fig. (\ref{resultse1}).
-A linear mixed model of RT was built, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
-Analysis confirms hypothesis H1:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1$_{rt}$)] RTs were faster for vague instructions than for crisp instructions 
-	($\beta=-0.11$, $se=.02$, $t=4.7$, $p<.001$).  % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-%\subsubsection{Error rates}
-{\bf Error rate} data were analysed using a generalized logit mixed model \cite{jaeger2008categorical}, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
-We found:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1$_{er}$)] There was a statistically non-significant effect of vagueness on error rates: the vague instructions tended to have a lower error rate than the than the crisp instructions 
-	($\beta=-.22$, $se=.18$, $t=-1.2$, $p=.230$).  % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-\subsection{Discussion}
-
-\noindent As expected, responses were faster and more accurate for vague than crisp instructions. Post hoc, we formulated some further hypotheses:
-
-{\small
-\begin{APAenumerate}
-	\item [(H2)] A main effect advantage of increasing numerical distance: the task will become easier as the numerical distance increases, because the two arrays are then more discriminable.
-	\item [(H3)] An interaction between vagueness and numerical distance: i.e., any facilitation for vagueness should be greater at smaller numerical distances than at larger numerical distances.
-\end{APAenumerate}
-}
-
-Looking at response times, these two post-hoc hypotheses were supported:
-
-{\small
-\begin{APAenumerate}
-	\item [(H2$_{rt}$)] RTs increased with numerical distance 
-	($\beta=-0.18$, $se=0.01$, $t=-11.6$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-	Since discriminability of the arrays is easier for larger numerical distances, discriminability probably underlies this effect. 
-	\item [(H3$_{rt}$)] Numerical distance and vagueness interacted significantly: essentially there were diminishing returns for vagueness as numerical distance increased: 
-	($\beta=0.07$, $se=0.01$, $t=5.9$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-Looking at error rates, the two post-hoc hypotheses were supported as well:
-
-{\small
-\begin{APAenumerate}
-	\item [(H2$_{er}$)] Error rates decreased as numerical distance increased: 
-	($\beta=-0.78$, $se=0.08$, $z=-9.6$, $p<0.0005$).% MG added the values for these stats on 25 Sept 2016
-	\item [(H3$_{er}$)] Numerical distance and vagueness interacted significantly: essentially there were diminishing returns for vagueness as numerical distance increased until the biggest numerical distance when vagueness was disadvantageous:
-	($\beta=0.80$, $se=0.15$, $z=5.27$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-So far, these results are in line with the idea of cost reduction.
-The idea of cost reduction explains the vagueness advantage by claiming that the vague referring expressions place less cognitive load on the comprehender than the crisp referring expressions.
-It explains the diminishing returns for vagueness in more-discriminable stimuli (i.e., the vagueness by numerical distance interaction) by claiming that load is low in both conditions for the easily-discriminable stimuli, and that therefore there is no extra benefit to be had from vagueness in the easily-discriminable stimuli.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 2                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 2: separating vagueness from instruction format}
-
-However, the picture painted by these findings might be misleading.
-First of all, there is a possibly confounding factor.
-Contrast, from Experiment 1, an expression from the vague condition: `the square with few dots' with an expression from the crisp condition: `the square with 5 dots'.
-One difference is that `few' has the potential for vagueness, whereas `5' is crisp.
-But another difference is that `few' is verbal while `5' is numerical, in the sense that a number is mentioned explicitly.
-Since these two differences could not be separated in Experiment 1, the vagueness advantage finding is vulnerable to an alternative interpretation, that what we saw as a vagueness advantage was in contrast an advantage for the verbal form of the quantifier.
-In Experiment 2 we therefore created verbal and numeric versions of each of the vague and crisp instructions so that we could compare vague and crisp conditions while taking account of verbal / numeric format.
-
-Another potential problem with Experiment 1 is the following.
-Participants chose one of two squares: therefore the `vague' quantifiers (e.g., `few') uniquely identified one square.
-Recall our definition of vague  -- ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise''.
-In Experiment 1, the quantifiers in the vague conditions did not realise their potential for vagueness.
-This is because there were no borderline cases of the referent that could make the referent set `not well-defined', and perhaps because using definite articles in the instructions implied that only one option was correct.
-Using error feedback in Experiment 1 could have exacerbated this.
-
-To find out what happens when words are used in a context where their potential for vagueness comes to the fore, Experiment 2 used three arrays (rather than two) so that the vague description had more than one possible referent, and used indefinite articles to avoid the impression that only one response counted as correct, and was carried out without error feedback.
-An indication that the potential for vagueness was realised in Experiment 2 is that the borderline response was chosen fairly often: 16\% of the time.
-
-In Experiment 2, an item was a referring expression instruction followed by a set of three dot arrays defined by a triple of numbers, representing the number of dots in the left, middle, and right arrays.
-We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54).
-Each set of arrays comprised three arrays (instead of two as in Experiment 1); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had more.
-
-Examples of crisp and vague versions of the numerical and verbal instructions follow: the examples assume the array (6,15,24) and reference to the smaller number of dots, such that 6 was classified as the expected response; 15 as the borderline response; and 24 as the incorrect response.
-In the {\em vague numerical} condition we used \emph{Choose a square with about 10 dots}.
-None of the squares contained 10 dots.
-10 is slightly closer to 6 than to 15, justifying 6 as the best response and 15 as the borderline response.
-In the {\em vague verbal} condition we used \emph{Choose a square with few dots}.
-In the {\em crisp numerical} condition we used \emph{Choose the square with 6 dots}, and one square always did contain the number mentioned.
-For {\em crisp verbal}, we used \emph{Choose the square with the fewest dots}.
-
-\begin{table}[htbp]
-\caption{Experiment 1: instructions arranged by condition for the dot triple $(6,15,24)$ and instructions indicating the smaller numbers}
-\label{instre3}
-\begin{tabular}{rll}
-&crisp&vague\\
-\midrule
-numerical	&	Choose the square with 6 dots 			& 	Choose a square with about 10 dots	\\
-verbal	&	Choose the square with the fewest dots	&	Choose a square with few dots		\\
-\midrule
-\end{tabular}
-\end{table}
-
-\subsection{Hypotheses} % E2
-\noindent We formulated the following hypotheses for Experiment 2:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1)] There exists a main effect RT advantage for vagueness.
-	\item [(H2)] An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
-	\item [(H3)] No large main effect of instruction format (i.e., numeric versus verbal) since we hypothesised that vagueness rather than instruction format drove the effect in Experiment 1.
-	\item [(H4)] On the basis of Experiment 1, we expect faster responses for stimuli with more discriminable arrays.(i.e., an effect of item).
-	\item [(H5)] Participants should make more borderline case choices for vague than crisp instructions.
-\end{APAenumerate}
-}
-
-\subsection{Experimental Method} % E2
-% Participants were aged between 18 and 45 with a median age of 28. 
-We manipulated as independent variables vagueness and instruction format (i.e., numeric versus verbal instructions), yielding four conditions, \emph{vague numeric}; \emph{vague verbal}; \emph{crisp numeric}; \emph{crisp verbal}.
-We measured two dependent variables: response time; and the probability of a participant choosing the borderline case.
-On each trial, first the referring expression that constituted the instruction for that trial was displayed. 
-Participants then pressed a key to indicate that they had read the instruction. 
-After 1000 ms, the arrays were presented, while preserving the text of the referring expression. 
-The response time dependent variable was measured from the presentation of the arrays, until the keypress indicating the participant's choice, which was also recorded. 
-The trial would timeout after 60 seconds if there was no response.
-In this experiment, no feedback was given. 
-This was because, in the vague conditions, we did not regard any response as `correct' or `incorrect', but instead as `borderline response', or `not borderline response', and we did not want to draw participants' attention to this distinction explicitly.
-We simply recorded whether the participant chose the best referent, the borderline case or the poorest referent, and how long it took the participant to respond.
-
-\subsection{Results} % E2
-
-%\subsubsection{Response times} % E2
-Means for response times and proportion of borderline responses are given in Fig. (\ref{resultse2}). 
-Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data. 
-A linear mixed model was constructed for the (logged) response times, 
-with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects. We found:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against Hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06$, $se=.01$, $t=4.6$, $p<0.0001$).
-	\item [(H2)] In focussed comparisons, vagueness was significantly disadvantageous in the \emph{numeric} conditions ($\beta=0.09$, $se=0.02$, $t=4.5$, $p<0.0001$), and non-significantly disadvantageous in the \emph{verbal} conditions ($\beta=0.02$, $se=0.02$, $t=1.5$, $p=0.1550$), offering partial evidence against Hypothesis 2.
-The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to a significant overall interaction effect between vagueness and instruction format ($\beta=-0.13$, $se=0.02$, $t=-6.6$, $p=0.0169$).
-	\item [(H3)] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1, but suggesting that instruction format rather than vagueness drove the effect we observed in Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; ($\beta=-0.36$, $se=0.07$, $t=-5.1$, $p<0.0001$).
-	\item [(H4)] There was a significant main effect of item (i.e., of which triple of numbers of dots were used in the stimulus): ($\beta=0.12$, $se=0.02$, $t=7.1$, $p<0.0001$).
-This effect seems likely to be due to the very fast responses for stimuli using arrangements of the smallest numbers of dots (6,15,24), which also had the largest difference in ratio of smallest number to largest number in the stimulus, suggesting that stimuli using these numbers of dots may have been particularly discriminable for participants.
-\end{APAenumerate}
-}
-
-A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
-The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{resultse2}.
-Participants chose the borderline square on $16.6\%$ of trials overall.
-{\small 
-\begin{APAenumerate}
-	\item [(H5)] Participants were significantly more likely to choose the borderline option for vague instructions than for crisp instructions (21.9\% vs 11.3\%: $\beta=0.62$, $se=0.22$, $z=2.8$, $p=0.0059$).
-Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%: $\beta=-3.35$, $se=0.23$, $z=-14.6$, $p<0.0001$).
-\end{APAenumerate}
-}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{images/e2-rtplot-1.pdf}
-\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{images/e2-blBarChart-1}
-\caption{Experiment 2 results: mean response times by condition and item, and counts of borderline case responses by condition.}
-\label{resultse2}
-\end{figure}
-
-\subsection{Discussion} % E2
-
-Experiment 2 tested to see whether when borderline cases are present, vague instructions would speed responses as they did in Experiment 1 when there were no borderline squares. 
-We actually found a small (but statistically significant) \emph{disadvantage} of vague instructions: vague instructions slowed people down by 112 ms on average.
-We also found that the effect of instruction format was significant, with numerical format slowing responses by 689 ms on average, such that the disadvantage of numerical format overwhelmed the contribution of vagueness.
-The \emph{verbal vague} condition was still responded to faster than the \emph{numerical crisp} condition, so the pattern from Experiment 1 was reproduced, but in the light of the evidence from Experiment 2, in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like an advantage of verbal instruction format.
-
-However, once again there is a possibly confounding factor.
-Observe that, in Experiment 2, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) {\bf selection algorithm}: To see this, consider the task of selecting the dot array that contains ``few dots": to do this, it suffices to {\em compare} the three arrays and select the one that contains the fewest elements.
-To select the dot array that contains "16 dots" seems to require the participant to estimate, and then {\em match}, the cardinality of (at least) one dot array to 16, a process which could plausibly take longer, independently of vagueness.
-Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison ``algorithm".
-
-In the next two experiments we pitted the comparison algorithm and matching algorithm selection tasks against each other while controlling vagueness and instruction format.
-In Experiment 3 we restricted all the instructions to \emph{numeric} quantifiers while factorially manipulating vagueness and selection task.
-In Experiment 4 we ensured that all instructions used \emph{verbal} quantifiers, while also factorially manipulating vagueness and selection task.
-This allowed us to distinguish between the predictions of the selection task account and the instruction format account.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 3                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 3: focus on instructions that contain numerals}
-
-The main aim of Experiment 3 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies.
-The main changes from Experiment 2 were that the human selection task was explicitly controlled (i.e., in whether it amounted to matching or comparison), and that all conditions were constrained to mention a number.
-We used the same arrays as in Experiment 2.
-Table \ref{Instructions for e3} shows the instructions for each condition.
-Note the difference between ``fewer than 20" and ``far fewer than 20": whereas the former cannot have borderline cases (i.e., for each number it is clear whether the number is smaller than 20 or not), the latter can.
-
-\begin{table}[htbp]
-\centering
-\caption{Experiment 3: Instructions arranged by condition for stimuli with $(6,15,24)$ dots and instructions indicating the smaller numbers} 
-\label{Instructions for e3}
-\begin{tabular}{llll}
-instruction					&vagueness					&	selection		&	instruction									\\
-format						&							&	task			&												\\
-\toprule
-\multirow{4}{*}{numeric}	&\multirow{2}{*}{crisp} 	&	matching		& Choose a square with 6 dots 					\\ 
-							&							&	comparison 		& Choose a square with fewer than 20 dots 		\\
-\cline{2-4}
-							&\multirow{2}{*}{vague} 	&	matching 		& Choose a square with about 10 dots 			\\ 
-							&							&	comparison 		& Choose a square with far fewer than 20 dots 	\\ 
-\bottomrule
-\end{tabular}
-\end{table}%
-
-\subsection{Hypotheses} % E 3
-\noindent For Experiment 3, we hypothesised:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
-	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection task)
-	\item [(H3)] Effects of vagueness are different depending on whether the selection task is matching or comparison (interaction effect selection x vagueness).
-\end{APAenumerate}
-}
-
-\subsection{Experimental Method} % E 3
-
-38 participants were recruited. We used a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e3}).
-On each trial an instruction was presented: participants pressed a key to dismiss the instruction, when the dot arrays were presented until the participant responded, and the response time and choice were recorded.
-
-\subsection{Results} % E 3
-
-Response times were trimmed at 2.5 SD separately for each subject, leading to the loss of 204 trials (2.8\% of the trials).
-Condition means for the remaining (logged) RTs are plotted in Figure \ref{resultse3}.
-A linear mixed model was constructed for the logged response times, 
-with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
-
-\begin{figure}[htbp]
-\centering
-\fitfigure{images/e3-rtplot-1.pdf}
-\caption{Mean response times by condition for Experiment 3 where all instructions were numeric}
-\label{resultse3}
-\end{figure}
-
-{\small
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions were non-significantly easier for the reader than crisp instructions when vagueness was considered as a main effect ($\beta=-0.01$, $se=0.01$, $t=-0.4$, $p=0.7158$). 
-	\item [(H2)] There was a statistically significant main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=0.16$, $se=0.03$, $t=6.2$, $p<0.0001$).
-	\item [(H3)] Vagueness exerted effects in different directions for the comparison task and for the matching task: there was a significant interaction between vagueness and selection task ($\beta=0.13$, $se=0.03$, $t=4.2$, $p<0.0001$).
-Separate analyses were conducted testing for effects of vagueness at each level of the selection task.
-Within the comparison task vagueness significantly speeded response times compared with crisp controls ($\beta=-0.07$, $se=0.02$, $t=-3.5$, $p<0.0012$). 
-Within the matching task vagueness significantly \emph{slowed} response times compared with crisp controls ($\beta=0.06$, $se=0.02$, $t=2.9$, $p<0.0061$). 
-\end{APAenumerate}
-}
-In other words, the cost reduction account was wrong to predict significant main effect advantages for vagueness (although there was a non-significant trend in the \emph{direction} predicted by the cost reduction account), and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was significantly advantageous in the comparison task.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 4                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 4: focus on instructions that do not contain numerals} % E 4
-
-This experiment mirrors Experiment 3, but focussing on instructions that did not use a number. 
-We manipulated vagueness and the selection task (comparison and matching). 
-In order to implement the experiment without mentioning numbers in the instructions, we changed the sequence of each trial to include a `target' (i.e., a dot array of a particular cardinality) before the instruction, so that we could then refer back to the target's cardinality in the instruction using expressions like \emph{the same number of dots as the target}; \emph{fewer dots than the target}.
-This presentation of a target before the main body of the trial shares some features with \citeA[Experiment 2]{Izard20081221}, although in that experiment participants were told the cardinality of the target (called an \emph{inducer} in that paper) whereas in our experiment we did not tell participants the cardinality of the prime array.
-An item was thus a combination of a target dot array, an instruction that did not contain a number, and a set of dot arrays taking their cardinalities from the same triples used in Experiment 2.
-Table \ref{Instructions for e4} spells out how the instructions were constrained not to mention a numeral and gives examples of targets.
-
-\begin{table}[htp]
-\caption{Instructions and targets by condition for Experiment 4}
-\begin{center}
-\begin{tabular}{llcrl}
-Item					    &Selection						&Vagueness	&Target		&Choose a square with\ldots							\\
-\toprule
-\multirow{4}{*}{06:15:24} 	&\multirow{2}{*}{Comparison} 	&Crisp		&20			&\ldots fewer dots than the target					\\
-\cline{3-5}
-							&								&Vague		&20			&\ldots far fewer dots than the target				\\
-\cline{2-5}
-							&\multirow{2}{*}{Matching}		& Crisp		&6			&\ldots the same number of dots as the target		\\
-\cline{3-5}
-							& 		 						&Vague		&10 		&\ldots about the same number of dots as the target	\\
-\bottomrule
-\end{tabular}
-\end{center}
-\label{Instructions for e4}
-\end{table}%
-
-\subsection{Hypotheses}% E 4
-\noindent For Experiment 4, we hypothesised:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
-	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection)
-	\item [(H3)] Effects of vagueness differ depending on whether selection is matching or comparison (interaction effect selection x vagueness, and focussed comparisons at each level of selection).
-\end{APAenumerate}
-}
-
-\subsection{Results}% E 4
-40 volunteers participated.
-The results showed that vagueness was beneficial for comparison but detrimental for matching (the same as Experiment 3) even when no numbers were allowed in the instructions. 
-Figure \ref{resultse4} shows the means by condition. Our findings were as follows:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1)] There was no significant main effect of vagueness ($\beta=-0.02$, $se=0.01$, $t=-1.5$, $p=0.1296$). 
-	\item [(H2)] There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=0.18$, $se=0.02$, $t=10.4$, $p<0.0001$).
-This effect was in the same direction as Experiment 3.
-	\item[(H3)] Vagueness did exert different effects depending on the selection task (main interaction effect of vagueness by selection $\beta=0.12$, $se=0.02$, $t=5.1$, $p<0.0001$). 
-Separate analyses of the effect of vagueness were conducted for the comparison task and for the matching task using Bonferroni-adjusted significance thresholds. 
-In the comparison task, vagueness resulted in faster response times ($\beta=-0.08$, $se=0.02$, $t=-4.3$, $p<0.0001$). 
-In the matching task vagueness slowed response times ($\beta=0.05$, $se=0.01$, $t=3.7$, $p=0.0004$). 
-These results are in the same direction as Experiment 3.
-\end{APAenumerate}
-}
-
-Once again, the cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
-
-\begin{figure}[htbp]
-\centering
-\fitfigure{images/e4-rtplot-1.pdf}
-\caption{Mean response times by condition for Experiment 4}
-\label{resultse4}
-\end{figure}
-
-\section{Discussion of experiments 3 and 4}
-
-The main aim of these two experiments was to test whether vagueness confers any cognitive benefits over and above those due to differences in the selection task according to whether the instruction mandates a \emph{comparison} selection task or a \emph{matching} selection task, when number-use is held constant.
-The main effect of selection task showed that the assumption that the \emph{comparison} task is easier than the \emph{matching} task is well-founded.
-In both experiments people were reliably quicker to respond in the \emph{comparison} task.
-
-Vagueness, which was the phenomenon on which our investigation focussed, did not exert a significant main effect in response time.
-However when the comparison and selection tasks were analysed separately, there was small significant advantage for vagueness in the \emph{comparison} tasks, but a small significant disadvantage for vagueness in the \emph{matching} tasks.
-
-\section{General Discussion}\label{discussion}
-
-To summarise our findings, we asked why strategic vagueness is as frequent as it is and we decided to focus on what we see as the most promising explanation, namely that vague expressions are easy to process by speakers and hearers: the cost hypothesis, as we have called it.
-We decided to test this hypothesis by experimentally investigating whether vague descriptions are resolved by hearers more quickly than crisp ones.
-Although we were able to find some interesting (and statistically highly significant) effects, it appears to us that if our sequence of experiments is assessed as a whole, it cannot be seen as confirmation of the cost hypothesis.
-
-To explain why this is, let us summarise our findings so far: Experiment 1 showed that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment was unable to distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format, and they mandated a comparison strategy rather than a matching strategy.
-Experiment 2 showed that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional benefit in either the verbal format instructions or the numerical format instructions.
-However, Experiment 2 did not distinguish benefits of number avoidance from benefits of the comparison selection task.
-In Experiments 3 and 4 we manipulated vagueness and the selection task, separately at each level of numerical format.
-Across the two experiments, we found that the comparison-task instructions attracted faster response times than the matching-task instructions.
-Within the two experiments we found that vagueness exerts benefits when the selection task is \emph{comparison}, but not when the task is \emph{matching}.
-
-What is one entitled to conclude?
-Given that we were able to identify a class of situations in which vague expressions led to faster response times than crisp ones, would it be valid to conclude that we have finally discovered an advantage for vagueness that cannot be ascribed to some other factor?
-We believe the answer to this question is negative.
-
-\begin{table}[htbp]
-\caption{Vagueness as range reduction: a summary of Experiments 3 and 4}
-\label{Vagueness as range reduction}
-\centering
-\begin{tabular}{cccc}
-selection task 					& vagueness		& candidates	& effect of vageueness						\\
-\toprule
-\multirow{ 2}{*}{comparison} 	& crisp 		& 2				& \multirow{ 2}{*}{vagueness advantage}  	\\
-\cline{2-3}
-								& vague			& 1				&                                           \\
-\midrule
-\multirow{ 2}{*}{matching} 		& crisp 		& 1 			& \multirow{ 2}{*}{vagueness disadvantage}	\\
-\cline{2-3}
-								& vague			& 2				&								 			\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-To see why, consider Figures 3 and 4.
-Both figures depict four conditions, depending on whether the expression was crisp or vague, and depending on whether the referent could be identified using a comparison strategy or not.
-Two of the resulting four conditions result in an expression that can denote either of two referents; the other two conditions result in an expression that can only denote one referent, with the other possible referent being a marginal candidate at best.
-
-To see why vagueness has opposite effects, depending on whether it is used in matching or comparison situations, consider the stimulus with (6,15,24) dots.
-Now compare `Choose a square with 6 dots' with its vague counterpart `Choose a square with about 10 dots': by adding the word `about', we broaden the range of squares that the expression might be referring to.
-On the other hand, compare `Choose a square with fewer than 20 dots' with `Choose a square with far fewer than 20 dots': by adding the word `far', we did not broaden the range of squares denotable by the expression: we narrow it down, because only some of the squares that have fewer dots may have {\em far} fewer dots.
-
-The benefits of vagueness in the \emph{comparison} task in experiments 3 and 4 could thus be explained as differences in the number of valid targets for the expression.
-This leads us to speculate that the benefit for vagueness here could be due to the vague expression foregrounding a particular valid target while the crisp expression carries with it the additional task of distinguishing between two alternative valid targets, something we propose to call a ``range-reduction'' benefit.
-
-The observation that conditions with 1 candidate lead to shorter response times than conditions with 2 candidates is consistent with the range reduction hypothesis, but not with the idea that vagueness is beneficial.
-It appears, in other words, that shorter response times will only result from a vague expression if this expression leads to range reduction.
-Once again, it is not vagueness itself that has advantages but a phenomenon (namely range reduction) that is an automatic concomitant of vagueness in some types of situations.\\[2ex]
-%
-\noindent Looking at the entire series of experiments, our findings suggest that the observed benefits of vague expressions may be due to factors other than vagueness: factors like avoiding numbers; permitting comparison tasks; and range reduction.
-The picture that is starting to emerge is subtle: on the one hand, in the situations that we have been studying vagueness is not intrinsically beneficial.
-On the other hand, vague expressions frequently possess other features that \emph{are} beneficial, and these are what give us the incorrect impression that vagueness itself is beneficial.
-Vagueness may thus have acquired a reputation that it does not deserve.
-The answer to Lipman's question, of why vagueness permeates human language (see our Introduction), may lie in a different direction after all, possibly relating to benefits for the speaker rather than the hearer.
-
-A comparison may clarify the logic of the situation.
-In recent years a number of studies, focussing on red wine, have suggested that alcohol, consumed in low doses, may have health benefits.
-An alternative explanation, however, asserts that it is not the alcohol in the wine that was beneficial, but antioxidants from grapes.
-If this alternative explanation is correct, then alcohol may not be healthy after all.\\[2ex]
-%
-% KvD Feb 2017. I've diminished the role of NLG in the coming paragraph.
-Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
-Years of research on the logic of vagueness -- giving rise to such techniques as Partial Logic \cite<e.g.,>{Fine}, Probabilistic Logic \cite{Edgington}, and Fuzzy Logic \cite{Zadeh} -- have primed the research community to expect that some special utility of vagueness is an important part of the answer, but our findings call this expectation into question.
-
-Although our own studies in this article have focussed on vagueness in descriptive Noun Phrases only, it seems to us plausible that vagueness plays a similar role in other
-linguistic constructs. For example, consider reports on air temperature. Given a numerical
-temperature measurement or prediction, we might word it as 
-%
-\begin{quote}
-(a) {\em 27.2 degrees Celsius}, or\\
-(b) {\em approximately 27 degrees}, or\\
-(c) {\em above 25 degrees}, or\\
-(d) {\em warm}, 
-\end{quote}
-%
-among other candidate expressions.
-Which of these descriptions is most effective, for example as part of a weather report?
-If the linguistic literature is to be believed, then options (a) and (c) convey crisp information, whereas (b) and (d) are vague (i.e., they permit borderline cases).
-In the situations studied in our own experiments and the ones discussed in section 1, we found no evidence that vagueness is beneficial for hearers.
-Rather than asking whether a candidate expression is vague, other questions might shed more light on the choice, similar to the ones identified in our studies.
-These questions might focus on the amount of information that a given expression conveys (i.e., on granularity), on the avoidance of numbers, and on the use of evaluative terms.
-Let's see how this might pan out for the above examples from the weather domain.
-
-First, the experiments by \citeauthor{Mishra01042011} suggest that it is important how much information is conveyed by an expression, and their findings are echoed by our own thoughts about range reduction (following Experiments 3 and 4).
-In the case of (a)--(d) above, it appears that (a) conveys the most detailed information (designating the smallest segment of the temperature scale), followed by (b), then (d), then (c) (e.g., 40 degrees is above 25, but at 40 Celsius the word ``warm" is likely to give way to ``hot" or ``scorching"):
-%
-\begin{quote}
-$a < b < d < c$
-\end{quote}
-%
-If these hunches are correct, then it seems to us that it is relatively unimportant whether a given expression is vague or crisp.
-Other factors seem more important; moreover, it may depend on the task and the audience which of a-d is preferred.
-For example, an \emph{expert} may prefer to read expression (a), because it gives her the most detailed information on which to base her decisions.
-On the other hand, expression (d) (``warm") is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make (d) more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm" (see our section 1 and  \citeA{peters2009bringing}), which is especially important if the hearer is unfamiliar with the metric used.
-These considerations suggest that {\em non-experts} might prefer expression (d).
-
-One way to see why vagueness (as defined in our Introduction) may not benefit human communication is the following thought experiment.
-Suppose a group of speakers understand the word ``warm" as vague, agreeing that temperatures above 26 count as warm, and temperatures below 24 do not count as warm, but considering temperatures between 24 and 26 as borderline cases.
-Now one day these speakers agree to sharpen up their definition deciding that, henceforth, ``warm" means ``$>25$ degrees" (as in (c) above): this decision resolves the borderline cases, while everything else remains the same.
-It seems unlikely that this change in language use, from a vague meaning to a crisp one (i.e., one that has no borderline cases anymore), would lower the utility of the word.
-Our experimental findings,
-and the conclusions that we draw from them, are consistent with this idea.
-
-\setcounter{secnumdepth}{0}
-\bibliography{v-book-2017}
-\end{document}

### 2018-09-12 chapter-latex/our-approach-to-the-problem.tex
added the springer chapter latex --post-review --post-response-to-review --pre-send-to-editor

--- 2018-08-02 springer/current/our-approach-to-the-problem.tex
@@
 Given these findings, we henceforth focussed our studies on non-subitizable numbers, because it is there that vagueness is expected to have benefits.
+
+In a second pilot experiment \citep{green2013utility}, we again presented two dot arrays with the instruction in the form \emph{Choose the square with \ldots dots}.
+The arrays contained larger numbers of dots than in the first pilot: one array always contained 25 dots, and the other contained either 5, 10, 15, 20, 30, 35, 40, or 45 dots.
+Each stimulus can therefore be seen in terms of the numerical difference between the number of dots in one array and the number of dots in the other: giving numerical distances of 5, 10, 15, and 20, with smaller numerical distances resulting in less discriminable arrays and larger distances resulting in more discriminable arrays.
+Our main manipulation was of the vagueness of the instruction, with two levels, \emph{crisp} and \emph{vague}.
+Assuming the dot array [5, 25], and the instruction referring to the smaller cardinality, the \emph{crisp} instruction was \emph{Choose the square with 5 dots} and its \emph{vague} counterpart was \emph{Choose the square with few dots}.
+We found, as expected, that responses were faster and more accurate for vague instructions than for crisp instructions.
+We also found an interaction between vagueness and numerical distance such that there were diminishing returns for vagueness as numerical distance increased, until, at the biggest numerical distance, there was no real difference any more between crisp and vague instructions.
+We interpreted this pattern as showing that cognitive load is relatively easy in both conditions when the stimuli are most discriminable, and so vagueness confers no additional advantage in those most easily discriminable stimuli.
+
+However, the picture painted by these findings from the second pilot experiment might be misleading.
+First of all, there is a possibly confounding factor.
+Contrast an expression from the vague condition: `the square with few dots' with an expression from the crisp condition: `the square with 5 dots'.
+One difference is that `few' has the potential for vagueness, whereas `5' is crisp.
+But another difference is that `few' is verbal while `5' is numerical, in the sense that a number is mentioned explicitly.
+Since these two differences could not be separated in that experiment, the finding of a vagueness advantage is vulnerable to an alternative interpretation, that the crisp-vague difference was really an advantage for the verbal form of the quantifier.
+In our next experiment (reported below as Experiment 1) we therefore created verbal and numeric versions of each of the vague and crisp instructions so that we could compare the crisp-vague difference and the numeric-verbal difference in the same experiment.
+
+In Experiment 1, reported below, we also addressed another issue with the second pilot experiment: participants chose one of two dot arrays - therefore the `vague' quantifiers (few and many) uniquely identified one square.
+Recall our definition of vague  -- ``a word is precise if it describes a well-defined set of objects.
+By contrast, a word is vague if it is not precise''.
+The quantifiers few and many might not have realised their potential for vagueness in the context of a choice between (only) two alternatives - where there is no borderline alternative that could result in few and many being "not well-defined".
+Furthermore using definite articles in the instructions may have reinforced the participant's impression that only one choice counted as correct, and this impression could have been reinforced by the use of error feedback.
+
+In the experiments we report below (Experiments 1, 2 and 3) we used what we learned from the pilot experiments to design situations that we believe did address the difference between crisp and vague instructions while taking account of the most important alternative explanations of this crisp-vague difference.

### 2018-09-12 docs/A_pilot_1.Rmd
added the pilot experiments R stuff
typo fix

--- 2018-08-31 experiments/A_pilot_1/writeup/A_pilot_1.Rmd
@@
 The experiment used subitizable numbers as well as bigger numbers.
-The instructions had the form: _Choose the square with {{few, many},{two, three, four, five, six, seven, eight, nine}} dots._ The variables were: subitizability; vagueness; target_size; gap_size; side (i.e., which side was the bigger number on?)
+The instructions had the form: _Choose the square with (few, many), (two, three, four, five, six, seven, eight, nine) dots._ The variables were: subitizability; vagueness; target_size; gap_size; side (i.e., which side was the bigger number on?)
 
@@
 # read in the data from file
-din <- read_delim("data.txt", delim="\t")
+din <- read_delim("A_pilot_1_data/data.txt", delim="\t")
 

### 2018-09-12 docs/B_pilot_2_functions/summarySEwithin2.R
ln -s in docs

@@ -1,95 +1,0 @@
-# https://gist.github.com/hauselin/a83b6d2f05b0c90c0428017455f73744#file-summarysewithin2-r
-summarySE2 <- function (data = NULL, measurevar, groupvars = NULL, na.rm = TRUE, conf.interval = 0.95) {
-  library(data.table)
-  data <- data.table(data)
-  
-  length2 <- function(x, na.rm = FALSE) {
-    if (na.rm) 
-      sum(!is.na(x))
-    else length(x)
-  }
-  
-  datac <- data[, .(unlist(lapply(.SD, length2, na.rm = na.rm)), 
-                    unlist(lapply(.SD, mean, na.rm = na.rm)),
-                    unlist(lapply(.SD, sd, na.rm = na.rm))),
-                by = groupvars, .SDcols = measurevar]
-  names(datac) <- c(groupvars, "N", measurevar, "sd")
-  setkeyv(datac, groupvars)
-  
-  datac[, se := unlist(sd) / sqrt(unlist(N))] #compute standard error
-  
-  ciMult <- qt(conf.interval / 2 + 0.5, unlist(datac$N) - 1)
-  datac[, ci := se * ciMult]
-  datac <- data.frame(datac)
-  return(datac)
-}
-
-normDataWithin2 <- function (data = NULL, idvar, measurevar, betweenvars = NULL, 
-                             na.rm = TRUE) {
-  library(data.table); library(dplyr)
-  data <- data.table(data)
-  setkeyv(data, idvar)
-  
-  data.subjMean <- data[, .(unlist(lapply(.SD, mean, na.rm = na.rm))), by = idvar, .SDcols = measurevar]
-  names(data.subjMean) <- c(idvar, 'subjMean')
-  data <- merge(data, data.subjMean)
-  setkeyv(data, c(idvar, betweenvars))
-  
-  measureNormedVar <- paste(measurevar, "Normed", sep = "")
-  data <- data.frame(data)
-  
-  data[, measureNormedVar] <- data[, measurevar] - unlist(data[, "subjMean"]) + mean(data[, measurevar], na.rm = na.rm)
-  return(data)
-}
-
-
-#normed and un-normed versions
-summarySEwithin2 <- function (data = NULL, measurevar, betweenvars = NULL, withinvars = NULL, 
-                              idvar = NULL, na.rm = TRUE, conf.interval = 0.95) {
-  
-  # Ensure that the betweenvars and withinvars are factors
[…]

### 2018-09-12 docs/C_exp_1.R

@@ -1,130 +1,1 @@
-## ----'load Libraries and source my functions', message=FALSE, echo=FALSE----
-library(knitr)
-library(tidyverse)
-library(broom)
-library(broom.mixed)
-library(lme4)
-library(lmerTest)
-library(dotwhisker)
-library(kableExtra)
-library(LMERConvenienceFunctions)
-library(data.table)
-library(grid)
-library(gridExtra)
-source("C_exp_1_functions/concatenate_raw_data.R")
-source("C_exp_1_functions/annotate_the_raw_data.R")
-source("C_exp_1_functions/add_borderline_vars.R")
-source("C_exp_1_functions/pretty_coef_table.R")
-source("C_exp_1_functions/summarySEwithin2.R")
-root_dir <- getwd()
-data_dir <- "C_exp_1_data"
-
-## ---- "get data for RT analysis", echo=F, message=F, warning=F, results='hide'----
-dat <- 
-  concatenate_raw_data(root_dir, data_dir) %>%
-  annotate_the_raw_data() %>% # at this point RT has min=1 (resp_type="sticky") and max=59,999 (resp_type="timeout")
-  filter(RT>1 & RT<59999) # at this point RT has min=445, max=42,685 (42 seconds) and nrow(dat) is 7677 down from 7680
-dat <- perSubjectTrim.fnc(dat, response='RT', subject='Subject', trim = 2.5)$data %>% select(-SD, -Mean, -Scaled)
-dat$RT_log <- log(dat$RT)
-
-## ---- "get data for borderline analysis", echo=F-------------------------
-dat_borderline <- 
-  concatenate_raw_data(root_dir, data_dir) %>%
-  add_borderline_vars() %>%
-  annotate_the_raw_data() %>% # at this point RT has min=1 (resp_type="sticky") and max=59,999 (resp_type="timeout")
-  filter(RT>1 & RT<59999) # at this point RT has min=445, max=42,685 (42 seconds) and nrow(dat) is 7677 down from 7680
-
-## ---- "full table of instructions", echo=FALSE---------------------------
-instructions_table <- 
-  dat %>%
-    select(Item, Quantity, Vagueness, Number, Instruction) %>%
-    unique() %>%
-    arrange(Item, Quantity, Vagueness, Number) %>%
[…]

### 2018-09-12 docs/D_exp_2.R

@@ -1,131 +1,1 @@
-## ---- echo=FALSE, message=FALSE------------------------------------------
-library(knitr)
-library(tidyverse)
-library(broom)
-library(broom.mixed)
-library(lme4)
-library(lmerTest)
-library(dotwhisker)
-library(kableExtra)
-library(LMERConvenienceFunctions)
-library(data.table)
-library(grid)
-library(gridExtra)
-source("D_exp_2_functions/preprocessing.R")
-source("D_exp_2_functions/pretty_coef_table.R")
-source("D_exp_2_functions/summarySEwithin2.R")
-root_dir <- getwd()
-data_dir <- "D_exp_2_data"
-
-## ------------------------------------------------------------------------
-dat_full <- preprocessing(root_dir, data_dir)
-dat <- subset(dat_full, subset=RT>0 & RT<25000) # loses 2 RTs that were judged to be outliers 25978 49871
-dat_borderline <- dat
-dat <- perSubjectTrim.fnc(dat, response='RT', subject='Subject', trim = 2.5)$data
-dat$RT_log <- log(dat$RT)
-
-## ---- echo=FALSE---------------------------------------------------------
-instructions_table <- 
-  dat %>%
-    select(Item, Quantity, Vagueness, Selection, Instruction) %>%
-    unique() %>%
-    arrange(Item, Quantity, Vagueness, Selection) %>%
-    spread(key=Vagueness, value=Instruction)
-instructions_table %>% 
-  kable(align='cccll', caption="Full table of instructions") %>% 
-      kable_styling(full_width = F, position = "left", font_size = 11) 
-
-## ---- cho=FALSE----------------------------------------------------------
-dat_plot <- summarySEwithin2(dat, measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
-
-## ---- "EXP_D_RT_means_plot", fig.width=7, fig.height=3, echo=FALSE-------
-dodge = position_dodge(width=0.2)
-dat_plot$Condition <- as.factor(paste(sep=' ', dat_plot$Selection, dat_plot$Vagueness))
-ggplot(dat_plot, aes(y=RT_logNormed, x=Item, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci, group=Condition, shape=Condition, fill=Condition)) +
-  geom_line(position=dodge) +
-  geom_errorbar(width=.2, position=dodge) +
-  geom_point(size=2, position=dodge) +
[…]

### 2018-09-12 docs/E_exp_3.R

@@ -1,131 +1,1 @@
-## ---- "libs and functions", echo=FALSE, message=FALSE--------------------
-library(knitr)
-library(tidyverse)
-library(broom)
-library(broom.mixed)
-library(lme4)
-library(lmerTest)
-library(dotwhisker)
-library(kableExtra)
-library(LMERConvenienceFunctions)
-library(data.table)
-library(grid)
-library(gridExtra)
-source("E_exp_3_functions/preprocessing.R")
-source("E_exp_3_functions/pretty_coef_table.R")
-source("E_exp_3_functions/summarySEwithin2.R")
-root_dir <- getwd()
-data_dir <- "E_exp_3_data"
-
-## ---- get_data-----------------------------------------------------------
-dat_full <- preprocessing(root_dir, data_dir)
-dat <- subset(dat_full, RT>0 & RT<9500)
-dat_borderline <- dat
-dat <- perSubjectTrim.fnc(dat, response='RT', subject='Subject', trim = 2.5)$data
-dat$RT_log <- log(dat$RT)
-
-## ---- "Full table of instructions", echo=FALSE---------------------------
-instructions_table <- 
-  dat %>%
-    select(Item, Target, Quantity, Vagueness, Selection, Instruction) %>%
-    unique() %>%
-    arrange(Item, Target, Quantity, Vagueness, Selection) %>%
-    spread(key=Vagueness, value=Instruction)
-instructions_table %>% 
-  kable(align='ccccll', caption="Full table of instructions") %>% 
-      kable_styling(full_width = F, position = "left", font_size = 11) 
-
-## ---- "make data for rt plot", echo=FALSE--------------------------------
-dat_plot <- summarySEwithin2(dat, measurevar="RT_log", withinvars=c("Vagueness", "Selection", "Item"), idvar="Subject")
-
-## ---- "EXP_E_rt_means_plot", fig.width=7, fig.height=3, echo=FALSE-------
-dodge = position_dodge(width=0.2)
-dat_plot$Condition <- as.factor(paste(sep=' ', dat_plot$Selection, dat_plot$Vagueness))
-ggplot(dat_plot, aes(y=RT_logNormed, x=Item, ymin=RT_logNormed-ci, ymax=RT_logNormed+ci, group=Condition, shape=Condition, fill=Condition)) +
-  geom_line(position=dodge) +
-  geom_errorbar(width=.2, position=dodge) +
-  geom_point(size=2, position=dodge) +
[…]

### 2018-09-12 experiments/C_exp_1_functions/add_borderline_vars.R

@@ -1,0 +1,151 @@
+add_borderline_vars = function (dat) {
+  
+  dat$crossed = as.factor(paste('Con', dat$Condition, ':Quan', as.numeric(dat$Quantity), ':Item', dat$Item, sep=''))
+  
+  dat[dat$crossed=='Con1:Quan1:Item1', 'Exp_Num']       <-  6
+  dat[dat$crossed=='Con1:Quan1:Item1', 'Bline_Num']     <- 15
+  dat[dat$crossed=='Con1:Quan1:Item1', 'Extr_Num']      <- 24
+  dat[dat$crossed=='Con1:Quan1:Item2', 'Exp_Num'] <- 16
+  dat[dat$crossed=='Con1:Quan1:Item2', 'Bline_Num']     <- 25
+  dat[dat$crossed=='Con1:Quan1:Item2', 'Extr_Num']      <- 34
+  dat[dat$crossed=='Con1:Quan1:Item3', 'Exp_Num'] <- 26
+  dat[dat$crossed=='Con1:Quan1:Item3', 'Bline_Num']     <- 35
+  dat[dat$crossed=='Con1:Quan1:Item3', 'Extr_Num']      <- 44
+  dat[dat$crossed=='Con1:Quan1:Item4', 'Exp_Num'] <- 36
+  dat[dat$crossed=='Con1:Quan1:Item4', 'Bline_Num']     <- 45
+  dat[dat$crossed=='Con1:Quan1:Item4', 'Extr_Num']      <- 54
+  dat[dat$crossed=='Con1:Quan2:Item1', 'Exp_Num'] <- 24
+  dat[dat$crossed=='Con1:Quan2:Item1', 'Bline_Num']     <- 15
+  dat[dat$crossed=='Con1:Quan2:Item1', 'Extr_Num']      <-  6
+  dat[dat$crossed=='Con1:Quan2:Item2', 'Exp_Num'] <- 34
+  dat[dat$crossed=='Con1:Quan2:Item2', 'Bline_Num']     <- 25
+  dat[dat$crossed=='Con1:Quan2:Item2', 'Extr_Num']      <- 16
+  dat[dat$crossed=='Con1:Quan2:Item3', 'Exp_Num'] <- 44
+  dat[dat$crossed=='Con1:Quan2:Item3', 'Bline_Num']     <- 35
+  dat[dat$crossed=='Con1:Quan2:Item3', 'Extr_Num']      <- 26
[…]

### 2018-09-12 experiments/C_exp_1_functions/pretty_coef_table.R

@@ -1,0 +1,101 @@
+pretty_coef_table = function(modl, mycaption){
+  
+  coef_table <- tidy(modl, effects="fixed") 
+
+  # be sensitive to whether the model is an lmer or a glmer 
+  
+  if (class(modl) == "lmerModLmerTest") {
+    
+    # sort the names out
+    names(coef_table) <- c("effect", "term", "coefficient", "std.error", "df", "t_value", "p_value_raw")
+    coef_table$p_value <- as.character("")
+    coef_table$star <- as.character("")
+    
+    # categorise p values and add significance stars
+    # "\u2007" is a space the same width as a numeric, that is preserved in printing
+    # "\u0020" is just a space
+    for (i in 1:nrow(coef_table)) {
+      if (!is.na(coef_table$p_value_raw[i])) {
+        if (coef_table$p_value_raw[i] >= 0     & coef_table$p_value[i] < 0.001) {coef_table$p_value[i] <-    "<0.001";                        coef_table$star[i] <- paste(sep="","\uFF0A","\uFF0A","\uFF0A")}
+        if (coef_table$p_value_raw[i] >= 0.001 & coef_table$p_value[i] < 0.01 ) {coef_table$p_value[i] <-    paste(sep="","<0.01", "\u2007"); coef_table$star[i] <- paste(sep="","\uFF0A","\uFF0A"," ")}
+        if (coef_table$p_value_raw[i] >= 0.01  & coef_table$p_value[i] < 0.05 ) {coef_table$p_value[i] <-    paste(sep="","<0.05", "\u2007"); coef_table$star[i] <- paste(sep="","\uFF0A"," "," ")}
+        if (coef_table$p_value_raw[i] >= 0.05  ) {coef_table$p_value[i] <-    round(coef_table$p_value_raw[i],3); coef_table$star[i] <-  ""}
+      }
+    }
+    # un-tibble the table
+    coef_table <- coef_table %>% as.data.frame()
+    
+    # convert numerics to formatted chars for pretty printing
+    coef_table$coefficient <- format(x=coef_table$coefficient, nsmall=2, digits=2)
+    coef_table$std.error   <- format(x=coef_table$std.error, nsmall=2, digits=2)
+    coef_table$df          <- format(x=coef_table$df, nsmall=2, digits=1)
+    coef_table$t_value     <- format(x=coef_table$t_value, nsmall=2, digits=2)
[…]

### 2018-09-12 experiments/E_exp_3_functions/preprocessing.R

@@ -1,0 +1,210 @@
+preprocessing = function(root_dir, data_dir) {
+  
+  dat <- read.delim(file.path(root_dir, data_dir, "data_raw.txt"), stringsAsFactors = FALSE)
+  
+  # declare local variables
+  Number_of_valid_subjects <- 40
+  Number_of_rows <- 7680
+  Number_of_trials_per_subject <- Number_of_rows / Number_of_valid_subjects # 192
+  
+  # sort out borderline responses into expected near and far
+  # How many squares in the square they chose? 
+  dat$RESPONSE <- as.character(dat$key)
+  for (row in 1 : nrow(dat) ) {
+    switch(dat[row,'RESPONSE'],
+           'L' = {dat[row, 'choice'] <- dat[row, 'Lft']},
+           'M' = {dat[row, 'choice'] <- dat[row, 'Mid']},
+           'R' = {dat[row, 'choice'] <- dat[row, 'Rgt']}         
+    )
+  }
+  dat$Item <- dat$Itm
+  dat$Itm <- NULL
+  dat$crossed=paste('Con',dat$Cnd,':Quan',dat$Qty,':Item',dat$Item,sep="")
+  
+  dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseExpected'] <-  6 
+  dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseNear']     <- 15  
+  dat[dat$crossed=='Con1:Quan1:Item1', 'ResponseFar']      <- 24
+  
+  dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseExpected'] <- 16
+  dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseNear']     <- 25  
+  dat[dat$crossed=='Con1:Quan1:Item2', 'ResponseFar']      <- 34
+  
+  dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseExpected'] <- 26
+  dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseNear']     <- 35
+  dat[dat$crossed=='Con1:Quan1:Item3', 'ResponseFar']      <- 44
+  
+  dat[dat$crossed=='Con1:Quan1:Item4', 'ResponseExpected'] <- 36
[…]

### 2018-09-12 chapter-latex/discussion-of-e2-and-e3.tex
add all files

--- 2018-08-02 springer/submission/v-book-2017.tex
@@
-\documentclass[%
-%			% choose man or doc immediately below
-%doc,	    % doc is single spaced
-man,		% man is double spaced
-floatsintext,%
-apacite%
-]{apa6} % I'm using a local (as in, in this folder) apa6.cls which edits footnotesep from 16pt which is the real apa6 in 2018 back to 6pt in order to make it line up with the pdf generated for the 2017 submission
-\usepackage{graphicx}
-\usepackage{multirow}
-\usepackage{color}
-\usepackage{booktabs}
-\usepackage{layout}
-\usepackage{apacite}
-\usepackage{datetime}
-
-\title{The Elusive Benefits of Vagueness \protect \\ Evidence from Experiments}
-\date{\today}
-\authornote{This is a layout-identical version of the document submitted for review in May 2017. LaTeX compiled \today~at \currenttime}
-\shorttitle{Is Vagueness Beneficial for Hearers?}
-\twoauthors{Matt Green}{Kees van Deemter}
-\twoaffiliations{Computing Science\\ University of Aberdeen}{Computing Science\\ University of Aberdeen}
-
-\abstract{Much of everyday language is vague, even in situations where vagueness could have been avoided (i.e., where vagueness is used "strategically").
-Yet the benefits of vagueness for hearers and readers are proving to be elusive.
-We discuss a range of earlier controlled experiments with human participants, and we report on a new series of experiments that we conducted in recent years.
-These experiments, which focus on vague expressions that are part of referential noun phrases, aim to separate the utility of vagueness (as defined by the existence of borderline cases) from the utility of other factors that tend to co-occur with vagueness.
-Having presented the evidence, we argue that the evidence supports a view where the benefits that vague terms exert are due to other influences, and not to vagueness
-itself.}
-
-\begin{document}
-\maketitle   
-\setcounter{secnumdepth}{3}
-\section{Introduction}\label{introduction}
-
-In most academic use, the word `vagueness' has a specific meaning.
-Keefe and Smith, for example, state ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \cite[p.\ 4]{keefe1997vagueness}, also \citeA{EgreKlinedinst}).
-The crucial criterion is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise'' \cite[p.\ 1]{lipmanvague}.
-A typical example is the word ``tall'', as applied to people for example, because here is no precise, known height which separates those who are tall from those who are not.
-The crucial point is that ``tall'' admits borderline cases (i.e., people who may or may not count as tall), which are the hallmark of vagueness as we use the term.
-
-Linguists, philosophers of language, and more recently game theorists, have asked why natural languages contain so many vague expressions \cite{Lipman:2000fk, lipmanvague}, which are
-often used even in situations where the speaker could have used an expression that is not vague (i.e., crisp); in these situations we say that vagueness is used \emph{strategically}.
-By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss'' \cite<>[p.~1]{lipmanvague}.
-Lipman explains the point by means of a scenario in which a speaker describes a person to a hearer, who needs to identify that person in the arrivals hall of of an airport.
-In such a scenario, a precise description of the person's height (e.g., ``The person's height is 187.96 cm'') would be more useful than a vague one (``The person is tall'').
-Lipman uses this scenario to explain why standard game theory models of communication \cite<e.g.,>{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs.
-
-Lipman argued that the efficiency loss resulting from vague expressions would be unlikely to have arisen unless there are advantages as well as disadvantages associated with vague expressions.
-Lipman asked, essentially, what these advantages might be.
-Several tentative answers to Lipman's question have been offered  \cite<see>{van2009utility, vanDeemterBook}.
-One of the most promising answers appears to be the idea that vague expressions are easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \cite<e.g.,>[]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
-For example, \citeA<>[p.\ 11]{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze''.
-We shall refer to this as the \emph{cost reduction} hypothesis.
-
-This article brings an experimental approach to these issues, focussing on vagueness in descriptions (e.g., "the square with few dots") and its effect on the hearer's ability to act on a given description, as measured by the time that it takes hearers to click on the referent of a description\footnote{Other metrics could have been chosen, such as hearers' ability to remember information, for example, or error rates.
-Although error rates play a minor role in the present paper, for reasons that will become clear, we focus on response times in particular.}.
-We find that, although hearer benefits from vague descriptions are straightforward to demonstrate in many cases, a closer experimental analysis militates against the conclusion that vagueness itself -- as defined above, in terms of the existence of borderline cases -- lies at the heart of these results.
-Instead, it is other factors, such as the presence of an overt numerical expression in the description, that proved to be decisive.
-We believe that, despite the fact that our experiments are unavoidably focussed only on a specific class of vague expressions (since any experiment can only deal with a limited number of different stimuli), these findings are potentially important, because they call into question whether "strategic" vagueness (i.e., vagueness where the speaker had a choice, because she could have been produced a crisp expression instead) has any advantages at all.
-In other words, returning to Lipman's question, it is possible that vagueness has evolved partly as a necessary evil (e.g., because of the limits of observation and prediction) and partly as a side effect of other factors.
-
-\section{Related work}
-
-Charting the utility of vagueness is the attested aim of a small number of studies, but most of these have focussed on vagueness in a different sense, and focussing on different types of benefits for hearers.
-Two recent studies can illustrate both issues.
-
-In a series of studies of behaviour modification, \citeA<>{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss.
-In the weight loss study, participants were told that the study was designed to test the validity of a new (actually fictitious) health index, the HHI (Holistic Health Index).
-They were told that an ideal HHI score lies in the range of 45 to 55.
-In a longitudinal study, participants submitted their weight to a computer each week.
-Participants were told that two algorithms would be used to compute their HHI, and that the two might give different values initially, in which case the true score lay between the two values.
-In one condition, which the authors called the precise condition, the two algorithms gave the same score.
-In the other condition, which the authors called the vague condition, one algorithm added 3\% to the score while the other algorithm subtracted 3\% from the score, yielding a range of values whose midpoint was the same as the two values given in the precise condition.
-
-One group of participants was given HHI scores in the ideal range: for this group their weight loss did not differ depending on whether they were given vague or precise HHI values.
-However for the other group, who were given HHI scores outside the ideal range, their weight loss was significantly greater if they were given vague HHI scores than if they were given precise HHI scores.
-The authors explain the improvement in the vague condition for this group as resulting from the participants' freedom to think of themselves a positioned on one end of the range - the end closest to the ideal HHI scores.
-This ``illusion of proximity'' \cite<>[p.~4]{Mishra01042011} to the goal is argued to allow participants to generate positive expectancies that lead to behaviours that improve performance.
-In contrast, in the precise conditions, participants did not have this freedom of interpretation, and could not distort the information to bring about the beneficial \emph{illusion of proximity}.
-These results are interesting, and of obvious potential practical importance.
-We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \cite{keefe1997vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases.
-In the terminology of \citeA{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}.
-Furthermore, the experiments of \citeA{Mishra01042011} did not explore benefits in terms of processing cost, but in terms of long-term behaviour change.
-
-Similar issues arise from the work of \citeA{peters2009bringing}.
-The authors carried out a series of studies where participants were required to rate hospitals based on various sources of information about quality of care.
-There was a between-subjects manipulation based on numeracy.
-The format of the information was manipulated within subjects: either numbers only were presented, or both numbers and evaluative categories were presented (e.g., \emph{Poor}, \emph{Fair}, \emph{Good}, \emph{Excellent}, with crisp visual boundary lines between the categories).
-Results showed that, for low-numeracy participants, the presence of evaluative categories resulted in a diminished influence of an irrelevant affective state on the ratings.
-For all participants, the presence of evaluative categories resulted in better decisions and in a greater use of the most important and reliable types of information, such as survival rates.
-
-It is, however, questionable whether the ``evaluative categories" manipulation in this study can be considered a manipulation of vagueness.
-Certainly, terms like \emph{Fair} admit the possibility of borderline cases.
-However, given that the boundaries between the categories were marked crisply, and that therefore the categories mapped crisply to numerical values, it becomes doubtful whether any borderline cases could be conceived to arise in fact.
-For example, \emph{Fair} was mapped to 60\% -- 70\% for the variable \emph{percentage of heart attack patients given recommended treatment (ACE inhibitor)}.
-Accordingly, rather than the vagueness of categories such as \emph{Poor}, Peters et al.
-emphasise the evaluative content inherent in these categories, and the affective potential of the evaluative content rather than the vagueness of the terms like {\em Fair}.
-
-\subsection{Our Approach to the Problem}\label{method}
-
-Are vague expressions processed more easily by readers than crisp ones?
-Like Lipman, we focus on situations where numerical information is used in order to identify a referent.
-Reference, in other words, will be the communicative task on which we focus, partly because of the interest that this topic has recently drawn in various areas of Cognitive Science \cite{vanDeemterCMR}.
-By looking at one specific type of vagueness, we will be able to investigate the costs and benefits of vagueness relatively thoroughly.
-Whether our findings generalise to other uses of vagueness is a question on which we will speculate in the final section of this chapter.
-
-We have chosen a narrative strategy in which we address a sequence of four experiments with human reader chronologically, explaining how each experiment helped us refine our research question.
-In order to do justice to our findings, we need to describe these experiments in a fair amount of detail.
-
-Let us start by explaining the task that was given to the participants in our experiments.
-We used a {\em speeded forced choice} task to compare the processing costs of different references to quantities.
-In this context, speed and accuracy of responses are the key dimensions on which the different references can be compared.
-Each stimulus in the experiments was a set of dot arrays containing various number of dots, together with a preceding instruction (in the form of a referring expression) to choose one of the arrays with respect to its cardinality.
-The participant was asked to respond as quickly as possible while avoiding errors.
-We manipulated the instructions and the arrays in several ways across the four experiments.
-
-All the experiments shared the following properties: 
-Stimuli were created using the language GNU Octave \cite{eaton:2002} and the Psychophysics Toolbox extensions \cite{ptbx1, ptbx2}.
-The position of the dots was randomised per-trial.
-The order in which trials were presented was randomised per-participant. 
-There were 256 trials, presented in 4 blocks of 64 each, between which the participant could rest. 
-A MacBook Pro laptop computer with a 13 inch screen presented the stimuli to the participants and recorded responses. 
-Participants were recruited using email lists at the University of Aberdeen, and paid ten pounds for participating.
-All participants self-reported fluency in English, and had normal, or corrected-to-normal vision.
-The experiment was conducted in a quiet room. 
-Participants were asked to respond as quickly as possible while avoiding errors.
-There was a block of practice trials after which participants could ask any questions, following which the experimenter left the room.
-All $p$ values reported for linear models were calculated using the R package \emph{lmerTest} \cite{lmerTest}. The complete data and analysis of our first experiment are available from 
-{\tt http://homepages.abdn.ac.uk/k.vdeemter/pages/vagueness-data-analysis/e1.html}, and analogous for experiments 2-4.\\[2ex]
-
-\noindent When the distance grows between two numbers, they become more easily distinguishable: the \emph{numerical distance effect} has been shown for comparing the cardinality of two sets of dots \cite{van123} and for processing Arabic numerals and number words \cite{Dehaene1996}.
-We manipulated the number of dots in each array such that some sets of arrays had smaller numerical distances and others had larger numerical distances.
-Where a number was mentioned in the instructions, it was always in the form of an Arabic numeral.
-When two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \cite{dehaene1993mental, gevers2006automatic}.
-We controlled which side the smaller number appeared on to avoid systematic influences of this effect.
-
-There is abundant evidence \cite<e.g.,>{trick1994small} that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities.
-We performed a pilot experiment \cite{our-earlier-paper} in which we were able to confirm this finding in the experimental settings on which we are focussing in this paper.
-We found that, when participants were confronted with a stimulus consisting of two squares containing different numbers of dots\footnote{Such a stimulus is referred to hereafter as consisting of a set of {\em dot arrays}.
-The number of dots in an array is referred to as its cardinality.
-The physical arrangement of dots in each array is irregular.}, instructions of the form \emph{Choose the square with \emph{n} dots} led to consistently faster response times than instructions of the form \emph{Choose the square with many/few dots} when $2 \leq n \leq 5$; the converse was true for $n>5$.
-Given these findings, we henceforth focussed our studies on non-subitizable numbers, because it is there that vagueness is expected to have benefits.
-
-\section{Experiment 1: initial experiment}
-
-We used a forced choice task to elicit responses to crisp or vague forms of instructions that required the participant to choose one of two dot arrays by referring to its cardinality.
-The participant was presented with an instruction in the form \emph{Choose the square with \ldots dots}.
-Then a set of two dot arrays was presented, each in the form of a square containing some number of dots.
-The participant was required to identify the dot array that corresponded with the instruction, by pressing the appropriate key, as quickly as possible while avoiding errors. 
-Response time and accuracy were recorded for analysis.
-
-We manipulated how discriminable the dot arrays were by varying the numerical distance between them. 
-One array always contained 25 dots: the other contained either 5, 10, 15, 20, 30, 35, 40, or 45 dots. 
-This gave us numerical distances of 5, 10, 15, and 20, with smaller numerical distances resulting in less discriminable arrays and larger distances resulting in more discriminable arrays.
-
-Our main manipulation was of the vagueness of the instruction, with two levels, \emph{crisp} and \emph{vague}.
-Assuming the dot array [5, 25], and the instruction referring to the smaller cardinality, the \emph{crisp} instruction was \emph{Choose the square with 5 dots} and its \emph{vague} counterpart was \emph{Choose the square with few dots}.
-
-\subsection{Hypothesis}
-\noindent The main aim of this experiment was to test the following hypothesis:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] There exists a main effect advantage for vagueness: vague instructions impose a lower cognitive load for the comprehender than crisp alternatives.
-\end{APAenumerate}
-}
-
-\subsection{Experimental Method}
-
-%Twenty participants were recruited, aged between 18 and 45 with a median age of 26.
-On each trial a participant was presented with an instruction to choose one of two dot arrays on screen by reference to its cardinality. 
-Following a keypress to indicate that the participant had read the instruction there was a central fixation cross for 1000 ms, and a blank screen for 500 ms, followed by the array (without repetition of the referring expression).
-The arrays would stay on screen until the participant responded (there was no timing-out). 
-Response time was measured as the latency between the presentation of the arrays, and the keypress identifying the choice: in this way, the response time was separated from time spent reading the instructions, which is important since we are only interested in the former.
-A response was counted as erroneous if the square with the wrong number of dots was chosen (when the instruction contained a number); 
-if the square with the larger number of dots was selected (when the instruction was \emph{Choose the square with few dots}); 
-or if the square with the smaller number of dots was selected (when the instruction was \emph{Choose the square with many dots}). 
-No feedback was given on correct trials, but there was feedback on error trials in the form of the word ``\textsc{wrong!!}'' which flashed on screen.
-Hypothesis H1 was tested looking both at response times and at error rates.
-
-\subsection{Results}
-
-\begin{figure}[htbp]
-\fitfigure{images/e1RTandErrorRatesplot-1}
-\caption{Experiment 1 results: response times and error rates.}
-\label{resultse1}
-\end{figure}
-
-{\bf Response times} (RTs) for trials with erroneous responses were discarded, leading to the loss of 354 trials from 5120, representing 6.9\% of the trials. 
-The correct response RTs were trimmed at 2.5 standard deviations for each subject, leading to the loss of a further 160 trials, or 3.4\% of the remaining correct responses. 
-Means for response times and error rates are given in Fig. (\ref{resultse1}).
-A linear mixed model of RT was built, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
-Analysis confirms hypothesis H1:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1$_{rt}$)] RTs were faster for vague instructions than for crisp instructions 
-	($\beta=-0.11$, $se=.02$, $t=4.7$, $p<.001$).  % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-%\subsubsection{Error rates}
-{\bf Error rate} data were analysed using a generalized logit mixed model \cite{jaeger2008categorical}, with vagueness and numerical distance and their interaction as independent variables, and with random slopes for vagueness and numerical distance over participants.
-We found:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1$_{er}$)] There was a statistically non-significant effect of vagueness on error rates: the vague instructions tended to have a lower error rate than the than the crisp instructions 
-	($\beta=-.22$, $se=.18$, $t=-1.2$, $p=.230$).  % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-\subsection{Discussion}
-
-\noindent As expected, responses were faster and more accurate for vague than crisp instructions. Post hoc, we formulated some further hypotheses:
-
-{\small
-\begin{APAenumerate}
-	\item [(H2)] A main effect advantage of increasing numerical distance: the task will become easier as the numerical distance increases, because the two arrays are then more discriminable.
-	\item [(H3)] An interaction between vagueness and numerical distance: i.e., any facilitation for vagueness should be greater at smaller numerical distances than at larger numerical distances.
-\end{APAenumerate}
-}
-
-Looking at response times, these two post-hoc hypotheses were supported:
-
-{\small
-\begin{APAenumerate}
-	\item [(H2$_{rt}$)] RTs increased with numerical distance 
-	($\beta=-0.18$, $se=0.01$, $t=-11.6$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-	Since discriminability of the arrays is easier for larger numerical distances, discriminability probably underlies this effect. 
-	\item [(H3$_{rt}$)] Numerical distance and vagueness interacted significantly: essentially there were diminishing returns for vagueness as numerical distance increased: 
-	($\beta=0.07$, $se=0.01$, $t=5.9$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-Looking at error rates, the two post-hoc hypotheses were supported as well:
-
-{\small
-\begin{APAenumerate}
-	\item [(H2$_{er}$)] Error rates decreased as numerical distance increased: 
-	($\beta=-0.78$, $se=0.08$, $z=-9.6$, $p<0.0005$).% MG added the values for these stats on 25 Sept 2016
-	\item [(H3$_{er}$)] Numerical distance and vagueness interacted significantly: essentially there were diminishing returns for vagueness as numerical distance increased until the biggest numerical distance when vagueness was disadvantageous:
-	($\beta=0.80$, $se=0.15$, $z=5.27$, $p<0.0005$). % MG added the values for these stats on 25 Sept 2016
-\end{APAenumerate}
-}
-
-So far, these results are in line with the idea of cost reduction.
-The idea of cost reduction explains the vagueness advantage by claiming that the vague referring expressions place less cognitive load on the comprehender than the crisp referring expressions.
-It explains the diminishing returns for vagueness in more-discriminable stimuli (i.e., the vagueness by numerical distance interaction) by claiming that load is low in both conditions for the easily-discriminable stimuli, and that therefore there is no extra benefit to be had from vagueness in the easily-discriminable stimuli.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 2                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 2: separating vagueness from instruction format}
-
-However, the picture painted by these findings might be misleading.
-First of all, there is a possibly confounding factor.
-Contrast, from Experiment 1, an expression from the vague condition: `the square with few dots' with an expression from the crisp condition: `the square with 5 dots'.
-One difference is that `few' has the potential for vagueness, whereas `5' is crisp.
-But another difference is that `few' is verbal while `5' is numerical, in the sense that a number is mentioned explicitly.
-Since these two differences could not be separated in Experiment 1, the vagueness advantage finding is vulnerable to an alternative interpretation, that what we saw as a vagueness advantage was in contrast an advantage for the verbal form of the quantifier.
-In Experiment 2 we therefore created verbal and numeric versions of each of the vague and crisp instructions so that we could compare vague and crisp conditions while taking account of verbal / numeric format.
-
-Another potential problem with Experiment 1 is the following.
-Participants chose one of two squares: therefore the `vague' quantifiers (e.g., `few') uniquely identified one square.
-Recall our definition of vague  -- ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise''.
-In Experiment 1, the quantifiers in the vague conditions did not realise their potential for vagueness.
-This is because there were no borderline cases of the referent that could make the referent set `not well-defined', and perhaps because using definite articles in the instructions implied that only one option was correct.
-Using error feedback in Experiment 1 could have exacerbated this.
-
-To find out what happens when words are used in a context where their potential for vagueness comes to the fore, Experiment 2 used three arrays (rather than two) so that the vague description had more than one possible referent, and used indefinite articles to avoid the impression that only one response counted as correct, and was carried out without error feedback.
-An indication that the potential for vagueness was realised in Experiment 2 is that the borderline response was chosen fairly often: 16\% of the time.
-
-In Experiment 2, an item was a referring expression instruction followed by a set of three dot arrays defined by a triple of numbers, representing the number of dots in the left, middle, and right arrays.
-We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54).
-Each set of arrays comprised three arrays (instead of two as in Experiment 1); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had more.
-
-Examples of crisp and vague versions of the numerical and verbal instructions follow: the examples assume the array (6,15,24) and reference to the smaller number of dots, such that 6 was classified as the expected response; 15 as the borderline response; and 24 as the incorrect response.
-In the {\em vague numerical} condition we used \emph{Choose a square with about 10 dots}.
-None of the squares contained 10 dots.
-10 is slightly closer to 6 than to 15, justifying 6 as the best response and 15 as the borderline response.
-In the {\em vague verbal} condition we used \emph{Choose a square with few dots}.
-In the {\em crisp numerical} condition we used \emph{Choose the square with 6 dots}, and one square always did contain the number mentioned.
-For {\em crisp verbal}, we used \emph{Choose the square with the fewest dots}.
-
-\begin{table}[htbp]
-\caption{Experiment 1: instructions arranged by condition for the dot triple $(6,15,24)$ and instructions indicating the smaller numbers}
-\label{instre3}
-\begin{tabular}{rll}
-&crisp&vague\\
-\midrule
-numerical	&	Choose the square with 6 dots 			& 	Choose a square with about 10 dots	\\
-verbal	&	Choose the square with the fewest dots	&	Choose a square with few dots		\\
-\midrule
-\end{tabular}
-\end{table}
-
-\subsection{Hypotheses} % E2
-\noindent We formulated the following hypotheses for Experiment 2:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1)] There exists a main effect RT advantage for vagueness.
-	\item [(H2)] An RT advantage for vagueness both in \emph{numeric} and in \emph{verbal} instructions.
-	\item [(H3)] No large main effect of instruction format (i.e., numeric versus verbal) since we hypothesised that vagueness rather than instruction format drove the effect in Experiment 1.
-	\item [(H4)] On the basis of Experiment 1, we expect faster responses for stimuli with more discriminable arrays.(i.e., an effect of item).
-	\item [(H5)] Participants should make more borderline case choices for vague than crisp instructions.
-\end{APAenumerate}
-}
-
-\subsection{Experimental Method} % E2
-% Participants were aged between 18 and 45 with a median age of 28. 
-We manipulated as independent variables vagueness and instruction format (i.e., numeric versus verbal instructions), yielding four conditions, \emph{vague numeric}; \emph{vague verbal}; \emph{crisp numeric}; \emph{crisp verbal}.
-We measured two dependent variables: response time; and the probability of a participant choosing the borderline case.
-On each trial, first the referring expression that constituted the instruction for that trial was displayed. 
-Participants then pressed a key to indicate that they had read the instruction. 
-After 1000 ms, the arrays were presented, while preserving the text of the referring expression. 
-The response time dependent variable was measured from the presentation of the arrays, until the keypress indicating the participant's choice, which was also recorded. 
-The trial would timeout after 60 seconds if there was no response.
-In this experiment, no feedback was given. 
-This was because, in the vague conditions, we did not regard any response as `correct' or `incorrect', but instead as `borderline response', or `not borderline response', and we did not want to draw participants' attention to this distinction explicitly.
-We simply recorded whether the participant chose the best referent, the borderline case or the poorest referent, and how long it took the participant to respond.
-
-\subsection{Results} % E2
-
-%\subsubsection{Response times} % E2
-Means for response times and proportion of borderline responses are given in Fig. (\ref{resultse2}). 
-Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data. 
-A linear mixed model was constructed for the (logged) response times, 
-with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects. We found:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] The main effect of \emph{vagueness} was to slow responses down, in contrast with Experiment 1, and offering evidence against Hypothesis 1 (vague: 2668 ms; crisp: 2450 ms; a difference of 218 ms; $\beta=.06$, $se=.01$, $t=4.6$, $p<0.0001$).
-	\item [(H2)] In focussed comparisons, vagueness was significantly disadvantageous in the \emph{numeric} conditions ($\beta=0.09$, $se=0.02$, $t=4.5$, $p<0.0001$), and non-significantly disadvantageous in the \emph{verbal} conditions ($\beta=0.02$, $se=0.02$, $t=1.5$, $p=0.1550$), offering partial evidence against Hypothesis 2.
-The disadvantage for vagueness was greater in the numerical than in the verbal conditions, leading to a significant overall interaction effect between vagueness and instruction format ($\beta=-0.13$, $se=0.02$, $t=-6.6$, $p=0.0169$).
-	\item [(H3)] There was a significant effect of \emph{instruction format} with numerical conditions attracting longer responses than the verbal conditions: consistent with Experiment 1, but suggesting that instruction format rather than vagueness drove the effect we observed in Experiment 1 (numeric: 3284 ms; verbal 1866 ms; a difference of 1418 ms; ($\beta=-0.36$, $se=0.07$, $t=-5.1$, $p<0.0001$).
-	\item [(H4)] There was a significant main effect of item (i.e., of which triple of numbers of dots were used in the stimulus): ($\beta=0.12$, $se=0.02$, $t=7.1$, $p<0.0001$).
-This effect seems likely to be due to the very fast responses for stimuli using arrangements of the smallest numbers of dots (6,15,24), which also had the largest difference in ratio of smallest number to largest number in the stimulus, suggesting that stimuli using these numbers of dots may have been particularly discriminable for participants.
-\end{APAenumerate}
-}
-
-A generalized linear mixed model \cite{jaeger2008categorical} was fit to the data for selection of the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
-The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Fig, \ref{resultse2}.
-Participants chose the borderline square on $16.6\%$ of trials overall.
-{\small 
-\begin{APAenumerate}
-	\item [(H5)] Participants were significantly more likely to choose the borderline option for vague instructions than for crisp instructions (21.9\% vs 11.3\%: $\beta=0.62$, $se=0.22$, $z=2.8$, $p=0.0059$).
-Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%: $\beta=-3.35$, $se=0.23$, $z=-14.6$, $p<0.0001$).
-\end{APAenumerate}
-}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{images/e2-rtplot-1.pdf}
-\includegraphics[trim = 20mm 0mm 35mm 0mm, clip, width=.49\textwidth]{images/e2-blBarChart-1}
-\caption{Experiment 2 results: mean response times by condition and item, and counts of borderline case responses by condition.}
-\label{resultse2}
-\end{figure}
-
-\subsection{Discussion} % E2
-
-Experiment 2 tested to see whether when borderline cases are present, vague instructions would speed responses as they did in Experiment 1 when there were no borderline squares. 
-We actually found a small (but statistically significant) \emph{disadvantage} of vague instructions: vague instructions slowed people down by 112 ms on average.
-We also found that the effect of instruction format was significant, with numerical format slowing responses by 689 ms on average, such that the disadvantage of numerical format overwhelmed the contribution of vagueness.
-The \emph{verbal vague} condition was still responded to faster than the \emph{numerical crisp} condition, so the pattern from Experiment 1 was reproduced, but in the light of the evidence from Experiment 2, in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like an advantage of verbal instruction format.
-
-However, once again there is a possibly confounding factor.
-Observe that, in Experiment 2, instruction format (i.e., the choice between numeric and verbal) went hand in hand with might be called the (human) {\bf selection algorithm}: To see this, consider the task of selecting the dot array that contains ``few dots": to do this, it suffices to {\em compare} the three arrays and select the one that contains the fewest elements.
-To select the dot array that contains "16 dots" seems to require the participant to estimate, and then {\em match}, the cardinality of (at least) one dot array to 16, a process which could plausibly take longer, independently of vagueness.
-Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use a comparison ``algorithm".
-
-In the next two experiments we pitted the comparison algorithm and matching algorithm selection tasks against each other while controlling vagueness and instruction format.
-In Experiment 3 we restricted all the instructions to \emph{numeric} quantifiers while factorially manipulating vagueness and selection task.
-In Experiment 4 we ensured that all instructions used \emph{verbal} quantifiers, while also factorially manipulating vagueness and selection task.
-This allowed us to distinguish between the predictions of the selection task account and the instruction format account.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 3                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 3: focus on instructions that contain numerals}
-
-The main aim of Experiment 3 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies.
-The main changes from Experiment 2 were that the human selection task was explicitly controlled (i.e., in whether it amounted to matching or comparison), and that all conditions were constrained to mention a number.
-We used the same arrays as in Experiment 2.
-Table \ref{Instructions for e3} shows the instructions for each condition.
-Note the difference between ``fewer than 20" and ``far fewer than 20": whereas the former cannot have borderline cases (i.e., for each number it is clear whether the number is smaller than 20 or not), the latter can.
-
-\begin{table}[htbp]
-\centering
-\caption{Experiment 3: Instructions arranged by condition for stimuli with $(6,15,24)$ dots and instructions indicating the smaller numbers} 
-\label{Instructions for e3}
-\begin{tabular}{llll}
-instruction					&vagueness					&	selection		&	instruction									\\
-format						&							&	task			&												\\
-\toprule
-\multirow{4}{*}{numeric}	&\multirow{2}{*}{crisp} 	&	matching		& Choose a square with 6 dots 					\\ 
-							&							&	comparison 		& Choose a square with fewer than 20 dots 		\\
-\cline{2-4}
-							&\multirow{2}{*}{vague} 	&	matching 		& Choose a square with about 10 dots 			\\ 
-							&							&	comparison 		& Choose a square with far fewer than 20 dots 	\\ 
-\bottomrule
-\end{tabular}
-\end{table}%
-
-\subsection{Hypotheses} % E 3
-\noindent For Experiment 3, we hypothesised:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
-	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection task)
-	\item [(H3)] Effects of vagueness are different depending on whether the selection task is matching or comparison (interaction effect selection x vagueness).
-\end{APAenumerate}
-}
-
-\subsection{Experimental Method} % E 3
-
-38 participants were recruited. We used a 2 x 2 factorial manipulation of vagueness and selection task (see Table \ref{Instructions for e3}).
-On each trial an instruction was presented: participants pressed a key to dismiss the instruction, when the dot arrays were presented until the participant responded, and the response time and choice were recorded.
-
-\subsection{Results} % E 3
-
-Response times were trimmed at 2.5 SD separately for each subject, leading to the loss of 204 trials (2.8\% of the trials).
-Condition means for the remaining (logged) RTs are plotted in Figure \ref{resultse3}.
-A linear mixed model was constructed for the logged response times, 
-with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant for random effects.
-
-\begin{figure}[htbp]
-\centering
-\fitfigure{images/e3-rtplot-1.pdf}
-\caption{Mean response times by condition for Experiment 3 where all instructions were numeric}
-\label{resultse3}
-\end{figure}
-
-{\small
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions were non-significantly easier for the reader than crisp instructions when vagueness was considered as a main effect ($\beta=-0.01$, $se=0.01$, $t=-0.4$, $p=0.7158$). 
-	\item [(H2)] There was a statistically significant main effect of selection task, with the comparison task speeding responses compared to the matching task ($\beta=0.16$, $se=0.03$, $t=6.2$, $p<0.0001$).
-	\item [(H3)] Vagueness exerted effects in different directions for the comparison task and for the matching task: there was a significant interaction between vagueness and selection task ($\beta=0.13$, $se=0.03$, $t=4.2$, $p<0.0001$).
-Separate analyses were conducted testing for effects of vagueness at each level of the selection task.
-Within the comparison task vagueness significantly speeded response times compared with crisp controls ($\beta=-0.07$, $se=0.02$, $t=-3.5$, $p<0.0012$). 
-Within the matching task vagueness significantly \emph{slowed} response times compared with crisp controls ($\beta=0.06$, $se=0.02$, $t=2.9$, $p<0.0061$). 
-\end{APAenumerate}
-}
-In other words, the cost reduction account was wrong to predict significant main effect advantages for vagueness (although there was a non-significant trend in the \emph{direction} predicted by the cost reduction account), and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was significantly advantageous in the comparison task.
-
-%%%%%%%%%%%%%%%%%%%%%%%%
-%                            E 4                                     %
-%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Experiment 4: focus on instructions that do not contain numerals} % E 4
-
-This experiment mirrors Experiment 3, but focussing on instructions that did not use a number. 
-We manipulated vagueness and the selection task (comparison and matching). 
-In order to implement the experiment without mentioning numbers in the instructions, we changed the sequence of each trial to include a `target' (i.e., a dot array of a particular cardinality) before the instruction, so that we could then refer back to the target's cardinality in the instruction using expressions like \emph{the same number of dots as the target}; \emph{fewer dots than the target}.
-This presentation of a target before the main body of the trial shares some features with \citeA[Experiment 2]{Izard20081221}, although in that experiment participants were told the cardinality of the target (called an \emph{inducer} in that paper) whereas in our experiment we did not tell participants the cardinality of the prime array.
-An item was thus a combination of a target dot array, an instruction that did not contain a number, and a set of dot arrays taking their cardinalities from the same triples used in Experiment 2.
-Table \ref{Instructions for e4} spells out how the instructions were constrained not to mention a numeral and gives examples of targets.
-
-\begin{table}[htp]
-\caption{Instructions and targets by condition for Experiment 4}
-\begin{center}
-\begin{tabular}{llcrl}
-Item					    &Selection						&Vagueness	&Target		&Choose a square with\ldots							\\
-\toprule
-\multirow{4}{*}{06:15:24} 	&\multirow{2}{*}{Comparison} 	&Crisp		&20			&\ldots fewer dots than the target					\\
-\cline{3-5}
-							&								&Vague		&20			&\ldots far fewer dots than the target				\\
-\cline{2-5}
-							&\multirow{2}{*}{Matching}		& Crisp		&6			&\ldots the same number of dots as the target		\\
-\cline{3-5}
-							& 		 						&Vague		&10 		&\ldots about the same number of dots as the target	\\
-\bottomrule
-\end{tabular}
-\end{center}
-\label{Instructions for e4}
-\end{table}%
-
-\subsection{Hypotheses}% E 4
-\noindent For Experiment 4, we hypothesised:
-{\small
-\begin{APAenumerate}
-	\item [(H1)] Vague instructions are easier for the reader than crisp ones (main effect of vagueness)
-	\item [(H2)] Comparison is easier for the reader than matching (main effect of selection)
-	\item [(H3)] Effects of vagueness differ depending on whether selection is matching or comparison (interaction effect selection x vagueness, and focussed comparisons at each level of selection).
-\end{APAenumerate}
-}
-
-\subsection{Results}% E 4
-40 volunteers participated.
-The results showed that vagueness was beneficial for comparison but detrimental for matching (the same as Experiment 3) even when no numbers were allowed in the instructions. 
-Figure \ref{resultse4} shows the means by condition. Our findings were as follows:
-
-{\small
-\begin{APAenumerate}
-	\item [(H1)] There was no significant main effect of vagueness ($\beta=-0.02$, $se=0.01$, $t=-1.5$, $p=0.1296$). 
-	\item [(H2)] There was a main effect of selection, with comparison task instructions leading to faster responses than the matching task instructions ($\beta=0.18$, $se=0.02$, $t=10.4$, $p<0.0001$).
-This effect was in the same direction as Experiment 3.
-	\item[(H3)] Vagueness did exert different effects depending on the selection task (main interaction effect of vagueness by selection $\beta=0.12$, $se=0.02$, $t=5.1$, $p<0.0001$). 
-Separate analyses of the effect of vagueness were conducted for the comparison task and for the matching task using Bonferroni-adjusted significance thresholds. 
-In the comparison task, vagueness resulted in faster response times ($\beta=-0.08$, $se=0.02$, $t=-4.3$, $p<0.0001$). 
-In the matching task vagueness slowed response times ($\beta=0.05$, $se=0.01$, $t=3.7$, $p=0.0004$). 
-These results are in the same direction as Experiment 3.
-\end{APAenumerate}
-}
-
-Once again, the cost reduction account was wrong to predict main effect advantages for vagueness, and wrong to predict that vagueness should be beneficial at each level of the selection task: however vagueness was advantageous in the comparison task.
-
-\begin{figure}[htbp]
-\centering
-\fitfigure{images/e4-rtplot-1.pdf}
-\caption{Mean response times by condition for Experiment 4}
-\label{resultse4}
-\end{figure}
-
-\section{Discussion of experiments 3 and 4}
-
 The main aim of these two experiments was to test whether vagueness confers any cognitive benefits over and above those due to differences in the selection task according to whether the instruction mandates a \emph{comparison} selection task or a \emph{matching} selection task, when number-use is held constant.
@@
 However when the comparison and selection tasks were analysed separately, there was small significant advantage for vagueness in the \emph{comparison} tasks, but a small significant disadvantage for vagueness in the \emph{matching} tasks.
-
-\section{General Discussion}\label{discussion}
-
-To summarise our findings, we asked why strategic vagueness is as frequent as it is and we decided to focus on what we see as the most promising explanation, namely that vague expressions are easy to process by speakers and hearers: the cost hypothesis, as we have called it.
-We decided to test this hypothesis by experimentally investigating whether vague descriptions are resolved by hearers more quickly than crisp ones.
-Although we were able to find some interesting (and statistically highly significant) effects, it appears to us that if our sequence of experiments is assessed as a whole, it cannot be seen as confirmation of the cost hypothesis.
-
-To explain why this is, let us summarise our findings so far: Experiment 1 showed that responses were faster and more accurate when the instructions were vague than when they were crisp, but the experiment was unable to distinguish effects of vagueness from those of number-avoidance or selection task: the vague conditions were also in verbal rather than numerical format, and they mandated a comparison strategy rather than a matching strategy.
-Experiment 2 showed that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional benefit in either the verbal format instructions or the numerical format instructions.
-However, Experiment 2 did not distinguish benefits of number avoidance from benefits of the comparison selection task.
-In Experiments 3 and 4 we manipulated vagueness and the selection task, separately at each level of numerical format.
-Across the two experiments, we found that the comparison-task instructions attracted faster response times than the matching-task instructions.
-Within the two experiments we found that vagueness exerts benefits when the selection task is \emph{comparison}, but not when the task is \emph{matching}.
-
-What is one entitled to conclude?
-Given that we were able to identify a class of situations in which vague expressions led to faster response times than crisp ones, would it be valid to conclude that we have finally discovered an advantage for vagueness that cannot be ascribed to some other factor?
-We believe the answer to this question is negative.
-
-\begin{table}[htbp]
-\caption{Vagueness as range reduction: a summary of Experiments 3 and 4}
-\label{Vagueness as range reduction}
-\centering
-\begin{tabular}{cccc}
-selection task 					& vagueness		& candidates	& effect of vageueness						\\
-\toprule
-\multirow{ 2}{*}{comparison} 	& crisp 		& 2				& \multirow{ 2}{*}{vagueness advantage}  	\\
-\cline{2-3}
-								& vague			& 1				&                                           \\
-\midrule
-\multirow{ 2}{*}{matching} 		& crisp 		& 1 			& \multirow{ 2}{*}{vagueness disadvantage}	\\
-\cline{2-3}
-								& vague			& 2				&								 			\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-To see why, consider Figures 3 and 4.
-Both figures depict four conditions, depending on whether the expression was crisp or vague, and depending on whether the referent could be identified using a comparison strategy or not.
-Two of the resulting four conditions result in an expression that can denote either of two referents; the other two conditions result in an expression that can only denote one referent, with the other possible referent being a marginal candidate at best.
-
-To see why vagueness has opposite effects, depending on whether it is used in matching or comparison situations, consider the stimulus with (6,15,24) dots.
-Now compare `Choose a square with 6 dots' with its vague counterpart `Choose a square with about 10 dots': by adding the word `about', we broaden the range of squares that the expression might be referring to.
-On the other hand, compare `Choose a square with fewer than 20 dots' with `Choose a square with far fewer than 20 dots': by adding the word `far', we did not broaden the range of squares denotable by the expression: we narrow it down, because only some of the squares that have fewer dots may have {\em far} fewer dots.
-
-The benefits of vagueness in the \emph{comparison} task in experiments 3 and 4 could thus be explained as differences in the number of valid targets for the expression.
-This leads us to speculate that the benefit for vagueness here could be due to the vague expression foregrounding a particular valid target while the crisp expression carries with it the additional task of distinguishing between two alternative valid targets, something we propose to call a ``range-reduction'' benefit.
-
-The observation that conditions with 1 candidate lead to shorter response times than conditions with 2 candidates is consistent with the range reduction hypothesis, but not with the idea that vagueness is beneficial.
-It appears, in other words, that shorter response times will only result from a vague expression if this expression leads to range reduction.
-Once again, it is not vagueness itself that has advantages but a phenomenon (namely range reduction) that is an automatic concomitant of vagueness in some types of situations.\\[2ex]
-%
-\noindent Looking at the entire series of experiments, our findings suggest that the observed benefits of vague expressions may be due to factors other than vagueness: factors like avoiding numbers; permitting comparison tasks; and range reduction.
-The picture that is starting to emerge is subtle: on the one hand, in the situations that we have been studying vagueness is not intrinsically beneficial.
-On the other hand, vague expressions frequently possess other features that \emph{are} beneficial, and these are what give us the incorrect impression that vagueness itself is beneficial.
-Vagueness may thus have acquired a reputation that it does not deserve.
-The answer to Lipman's question, of why vagueness permeates human language (see our Introduction), may lie in a different direction after all, possibly relating to benefits for the speaker rather than the hearer.
-
-A comparison may clarify the logic of the situation.
-In recent years a number of studies, focussing on red wine, have suggested that alcohol, consumed in low doses, may have health benefits.
-An alternative explanation, however, asserts that it is not the alcohol in the wine that was beneficial, but antioxidants from grapes.
-If this alternative explanation is correct, then alcohol may not be healthy after all.\\[2ex]
-%
-% KvD Feb 2017. I've diminished the role of NLG in the coming paragraph.
-Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
-Years of research on the logic of vagueness -- giving rise to such techniques as Partial Logic \cite<e.g.,>{Fine}, Probabilistic Logic \cite{Edgington}, and Fuzzy Logic \cite{Zadeh} -- have primed the research community to expect that some special utility of vagueness is an important part of the answer, but our findings call this expectation into question.
-
-Although our own studies in this article have focussed on vagueness in descriptive Noun Phrases only, it seems to us plausible that vagueness plays a similar role in other
-linguistic constructs. For example, consider reports on air temperature. Given a numerical
-temperature measurement or prediction, we might word it as 
-%
-\begin{quote}
-(a) {\em 27.2 degrees Celsius}, or\\
-(b) {\em approximately 27 degrees}, or\\
-(c) {\em above 25 degrees}, or\\
-(d) {\em warm}, 
-\end{quote}
-%
-among other candidate expressions.
-Which of these descriptions is most effective, for example as part of a weather report?
-If the linguistic literature is to be believed, then options (a) and (c) convey crisp information, whereas (b) and (d) are vague (i.e., they permit borderline cases).
-In the situations studied in our own experiments and the ones discussed in section 1, we found no evidence that vagueness is beneficial for hearers.
-Rather than asking whether a candidate expression is vague, other questions might shed more light on the choice, similar to the ones identified in our studies.
-These questions might focus on the amount of information that a given expression conveys (i.e., on granularity), on the avoidance of numbers, and on the use of evaluative terms.
-Let's see how this might pan out for the above examples from the weather domain.
-
-First, the experiments by \citeauthor{Mishra01042011} suggest that it is important how much information is conveyed by an expression, and their findings are echoed by our own thoughts about range reduction (following Experiments 3 and 4).
-In the case of (a)--(d) above, it appears that (a) conveys the most detailed information (designating the smallest segment of the temperature scale), followed by (b), then (d), then (c) (e.g., 40 degrees is above 25, but at 40 Celsius the word ``warm" is likely to give way to ``hot" or ``scorching"):
-%
-\begin{quote}
-$a < b < d < c$
-\end{quote}
-%
-If these hunches are correct, then it seems to us that it is relatively unimportant whether a given expression is vague or crisp.
-Other factors seem more important; moreover, it may depend on the task and the audience which of a-d is preferred.
-For example, an \emph{expert} may prefer to read expression (a), because it gives her the most detailed information on which to base her decisions.
-On the other hand, expression (d) (``warm") is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make (d) more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm" (see our section 1 and  \citeA{peters2009bringing}), which is especially important if the hearer is unfamiliar with the metric used.
-These considerations suggest that {\em non-experts} might prefer expression (d).
-
-One way to see why vagueness (as defined in our Introduction) may not benefit human communication is the following thought experiment.
-Suppose a group of speakers understand the word ``warm" as vague, agreeing that temperatures above 26 count as warm, and temperatures below 24 do not count as warm, but considering temperatures between 24 and 26 as borderline cases.
-Now one day these speakers agree to sharpen up their definition deciding that, henceforth, ``warm" means ``$>25$ degrees" (as in (c) above): this decision resolves the borderline cases, while everything else remains the same.
-It seems unlikely that this change in language use, from a vague meaning to a crisp one (i.e., one that has no borderline cases anymore), would lower the utility of the word.
-Our experimental findings,
-and the conclusions that we draw from them, are consistent with this idea.
-
-\setcounter{secnumdepth}{0}
-\bibliography{v-book-2017}
-\end{document}

### 2018-09-12 chapter-latex/e1.tex
add all files

@@
 \begin{description}
-	\item [Test of Hypothesis 1 (Crisp/Vague RT)] Vague instructions actually led to significantly slower responses than crisp instructions, against Hypothesis 1: $\beta=0.058$, $se=0.013$, $t=4.55$, $p<0.001$.
-When the model was restricted to numeric-only instructions Vague instructions still led to significantly slower responses than crisp instructions $\beta=0.093$, $se=0.021$, $t=4.51$, $p<0.001$.
-When the model was restricted to verbal-only instructions Vague instructions tended to slow responses, but not significantly: $\beta=0.024$, $se=0.016$, $t=1.46$, $p=0.155$.
-	\item [Test of Hypothesis 2 (Numeric/Verbal RT)] There was actually a significant difference between numeric and verbal instructions, with numeric instructions leading to longer responses than verbal instructions, against Hypothesis 2: $\beta0.265=$, $se=0.072$, $t=5.08$, $p<0.001$.
-	\item [Test of Hypothesis 3 (Item RT)] Responses took longer as the levels of Item increased, supporting Hypothesis 3: $\beta=0.120$, $se=0.017$, $t=7.11$, $p<0.001$.
+	\item [Test of Hypothesis 1 (Crisp/Vague RT)] Vague instructions actually led to significantly slower responses than crisp instructions, against Hypothesis 1 ($\beta=0.058$, $se=0.013$, $t=4.55$, $p<0.001$).
+When the model was restricted to numeric-only instructions Vague instructions still led to significantly slower responses than crisp instructions ($\beta=0.093$, $se=0.021$, $t=4.51$, $p<0.001$).
+When the model was restricted to verbal-only instructions Vague instructions tended to slow responses, but not significantly ($\beta=0.024$, $se=0.016$, $t=1.46$, $p=0.155$).
+	\item [Test of Hypothesis 2 (Numeric/Verbal RT)] There was actually a significant difference between numeric and verbal instructions, with numeric instructions leading to longer responses than verbal instructions, against Hypothesis 2 ($\beta0.265=$, $se=0.072$, $t=5.08$, $p<0.001$).
+	\item [Test of Hypothesis 3 (Item RT)] Responses took longer as the levels of Item increased, supporting Hypothesis 3 ($\beta=0.120$, $se=0.017$, $t=7.11$, $p<0.001$).
 \end{description}

### 2018-09-12 chapter-latex/related-work.tex
add all files

--- 2018-08-02 springer/current/related-work.tex
@@
-Charting the utility of vagueness is the attested aim of a small number of studies, but most of these have focussed on vagueness in a different sense, and focussing on different types of benefits for hearers.
-Two recent studies can illustrate both issues.
+A number of answers to Lipman's question have been proposed, but as we have argued elsewhere, it appears that some of these do not to hold up to scrutiny \citep{van2009utility,vanDeemterBook}.
+For example,
+%
+\begin{itemize}
+\item It has been proposed that the existence of vague words makes a language more efficient \citep{barwiseperry}.
+The idea is that vagueness allows the use of one and the same word (e.g., ``big'') in different situations, which require different standards.
+For instance, the vague word ``big'' can denote a mouse a well as an elephant.
+The idea is plausible, but the problem is that ``big" is efficient not because it is vague (i.e., not because it allows borderline cases) but because it is context dependent.
+To see this, note for example that superlatives are not vague, but they {\em are} context dependent: ``the biggest $x$" ascribes different sizes to the referent depending on $x$.
+It is for this reason that we can use the word ``biggest''  to talk about the biggest mouse, the biggest elephant, and so on.
+Context-dependence, however, does not imply vagueness.
+%
+\item It has been pointed out that vague words are capable of combining a statement of quantity (along some dimension) with an evaluative statement \citep{veltman}.
+The idea is that when we say ``The patient has a fever'', we do not merely assert that her temperature lies above a certain value, we also imply that the deviation is large enough to be clinically significant (i.e., something is not right with the patient).
+The problem with this plausible idea is that evaluation does not imply vagueness.
+For example, the medical term ``obese'' is evaluative (i.e., it is not healthy to be obese), yet its standard medical definition in terms of Body Mass Index is perfectly crisp, with BMI values above 30 counting as obese.
+\end{itemize}
+%
+Analogous observations can be made about a number of other purported benefits of vagueness; one of the few hypotheses left standing at the moment is the {\em cost reduction} hypothesis, which we mentioned in the previous section.
+The findings on which we are reporting in the present paper will follow a familiar pattern: we will explore the cost reduction hypothesis, only to conclude that it does not stand up to scrutiny.
+Our approach, this time, will be experimental: we conduct controlled experiments to investigate the effect that vague expressions have on a hearer.
 
-In a series of studies of behaviour modification, \citeauthor{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss.
+Charting the utility of vagueness is also the attested aim of a small number of previous experimental studies.
+But, as it happens, few of these studies have truly focussed on vagueness in the sense on which we are focussing (i.e., they did not address Lipman's challenge).
+Two recent studies illustrate this issue; let's discuss them briefly.
+
+In a series of studies of behaviour modification, \citet{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss.
 In the weight loss study, participants were told that the study was designed to test the validity of a new (actually fictitious) health index, the HHI (Holistic Health Index).

### 2018-09-12 docs/C_exp_1.R

@@ -1,0 +1,1 @@
+../experiments/C_exp_1.R

### 2018-09-12 experiments/C_exp_1_functions/annotate_the_raw_data.R

@@ -1,0 +1,38 @@
+annotate_the_raw_data = function (dat) {
+  
+  # annotate vagueness
+  dat[ dat$Condition==1 , 'Vagueness'] <- 'Vague'
+  dat[ dat$Condition==2 , 'Vagueness'] <- 'Crisp'
+  dat[ dat$Condition==3 , 'Vagueness'] <- 'Vague'
+  dat[ dat$Condition==4 , 'Vagueness'] <- 'Crisp'
+  dat$Vagueness <- as.factor(dat$Vagueness)
+  
+  # annotate use of a number / word in the instructions
+  dat[ dat$Condition==1 , 'Number'] <- 'Numeric'
+  dat[ dat$Condition==2 , 'Number'] <- 'Numeric'
+  dat[ dat$Condition==3 , 'Number'] <- 'Verbal'
+  dat[ dat$Condition==4 , 'Number'] <- 'Verbal'
+  dat$Number <- as.factor(dat$Number)
+  
+  # annotate Item
+  dat$Item <- factor(dat$Item, levels=c(1,2,3,4), labels=c("06:15:24", "16:25:34", "26:35:44", "36:45:54"))
+  
+  # annotate Order
+  dat$Order <- factor(dat$Order, levels=c(1,2), labels=c('LtoR','RtoL'))
+  
+  # annotate Quantity
+  dat$Quantity <- factor(dat$Quantity, levels=c(1,2), labels=c('Small','Large'))
+  
+  # annotate Subject
+  dat$Subject <- factor(paste("s",sprintf("%02d",dat$Subject),sep=""))
+  
+  # annotate Trial index for each participant (1:256 trials for 30 subjects)
+  dat$Trial = rep(x = 1:256, times = 30)
+  
+  # add an Observation index, over the whole data set (1:7680)
+  dat$Obs <- 1:7680
+  
+  return(dat)
+}  
+
+

### 2018-09-12 chapter-latex/related-work.tex
chapter is the version Kees sent to the editor

@@
 The idea is plausible, but the problem is that ``big" is efficient not because it is vague (i.e., not because it allows borderline cases) but because it is context dependent.
-To see this, note for example that superlatives are not vague, but they {\em are} context dependent: ``the biggest $x$" ascribes different sizes to the referent depending on $x$.
+To see this, note for example that superlatives are not vague (i.e., they do not permit borderline cases), but they {\em are} context dependent: ``the biggest $x$" ascribes different sizes to the referent depending on $x$.
 It is for this reason that we can use the word ``biggest''  to talk about the biggest mouse, the biggest elephant, and so on.
@@
 %
-\item It has been pointed out that vague words are capable of combining a statement of quantity (along some dimension) with an evaluative statement \citep{veltman}.
-The idea is that when we say ``The patient has a fever'', we do not merely assert that her temperature lies above a certain value, we also imply that the deviation is large enough to be clinically significant (i.e., something is not right with the patient).
+\item Vague words are capable of combining a statement of quantity with an evaluative statement \citep{veltman}.
+The idea is that when we say ``The patient has a fever'', we do not merely assert that her temperature lies above a certain value, we also imply that the deviation is clinically significant (i.e., something is not right with the patient).
 The problem with this plausible idea is that evaluation does not imply vagueness.
-For example, the medical term ``obese'' is evaluative (i.e., it is not healthy to be obese), yet its standard medical definition in terms of Body Mass Index is perfectly crisp, with BMI values above 30 counting as obese.
+For example, the medical term ``obese'' is evaluative (i.e., it is not healthy to be obese), yet its standard medical definition in terms of Body Mass Index is perfectly crisp (i.e., without borderline cases), with all and only BMI values above 30 counting as obese.
 \end{itemize}
@@
 Analogous observations can be made about a number of other purported benefits of vagueness; one of the few hypotheses left standing at the moment is the {\em cost reduction} hypothesis, which we mentioned in the previous section.
-The findings on which we are reporting in the present paper will follow a familiar pattern: we will explore the cost reduction hypothesis, only to conclude that it does not stand up to scrutiny.
+The findings on which we are reporting in the present paper will follow a familiar pattern: we will explore the cost reduction hypothesis, only to conclude that it may not stand up to scrutiny.
 Our approach, this time, will be experimental: we conduct controlled experiments to investigate the effect that vague expressions have on a hearer.

### 2018-10-15 experiment_data_and_analyses/C_exp_1.R
was experiments

@@ -1,0 +1,130 @@
+## ----'load Libraries and source my functions', message=FALSE, echo=FALSE----
+library(knitr)
+library(tidyverse)
+library(broom)
+library(broom.mixed)
+library(lme4)
+library(lmerTest)
+library(dotwhisker)
+library(kableExtra)
+library(LMERConvenienceFunctions)
+library(data.table)
+library(grid)
+library(gridExtra)
+source("C_exp_1_functions/concatenate_raw_data.R")
+source("C_exp_1_functions/annotate_the_raw_data.R")
+source("C_exp_1_functions/add_borderline_vars.R")
+source("C_exp_1_functions/pretty_coef_table.R")
+source("C_exp_1_functions/summarySEwithin2.R")
+root_dir <- getwd()
+data_dir <- "C_exp_1_data"
+
+## ---- "get data for RT analysis", echo=F, message=F, warning=F, results='hide'----
+dat <- 
+  concatenate_raw_data(root_dir, data_dir) %>%
+  annotate_the_raw_data() %>% # at this point RT has min=1 (resp_type="sticky") and max=59,999 (resp_type="timeout")
+  filter(RT>1 & RT<59999) # at this point RT has min=445, max=42,685 (42 seconds) and nrow(dat) is 7677 down from 7680
+dat <- perSubjectTrim.fnc(dat, response='RT', subject='Subject', trim = 2.5)$data %>% select(-SD, -Mean, -Scaled)
+dat$RT_log <- log(dat$RT)
+
+## ---- "get data for borderline analysis", echo=F-------------------------
+dat_borderline <- 
+  concatenate_raw_data(root_dir, data_dir) %>%
+  add_borderline_vars() %>%
+  annotate_the_raw_data() %>% # at this point RT has min=1 (resp_type="sticky") and max=59,999 (resp_type="timeout")
+  filter(RT>1 & RT<59999) # at this point RT has min=445, max=42,685 (42 seconds) and nrow(dat) is 7677 down from 7680
+
+## ---- "full table of instructions", echo=FALSE---------------------------
+instructions_table <- 
+  dat %>%
+    select(Item, Quantity, Vagueness, Number, Instruction) %>%
+    unique() %>%
+    arrange(Item, Quantity, Vagueness, Number) %>%
[…]

### 2018-10-15 experiment_data_and_analyses/C_exp_1_functions/concatenate_raw_data.R

@@ -1,0 +1,18 @@
+concatenate_raw_data = function(root_dir, data_dir) {
+  
+  number_of_valid_subjects=30
+  number_of_trials_per_subject=256
+
+  column.headers.df <-  head(read.table(file.path(root_dir, data_dir, "subject01.data"), header=TRUE),0)
+
+  gathered.data <- column.headers.df
+  
+  for (subject_number in 1:number_of_valid_subjects) {
+    current.filename <- file.path(root_dir, data_dir, paste(sep="", "subject", sprintf("%02d", subject_number), ".data"))
+    current.data <- read.table(file=current.filename, header=TRUE, stringsAsFactors = TRUE)
+    gathered.data <- rbind(gathered.data, current.data)
+  }
+  
+  return(gathered.data)
+  
+}

### 2018-10-27 manuscript/e1.tex
removed undirected quote marks and replaced with aostrophe-style quotes

--- 2018-09-12 chapter-latex/e1.tex
@@
 \begin{itemize}
-	\item In the vague numerical condition the instruction was ``Choose a square with about 10 dots" -- none of the arrays contained exactly 10 dots, but 10 is closer to 6 than it is to 15, making 6 a better response to that instruction, 15 a borderline response, and 24 an extreme response.
-	\item In the vague verbal condition we used ``Choose a square with few dots". We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
-	\item In the crisp numerical condition we used ``Choose the square with 6 dots".
+	\item In the vague numerical condition the instruction was ``Choose a square with about 10 dots'' -- none of the arrays contained exactly 10 dots, but 10 is closer to 6 than it is to 15, making 6 a better response to that instruction, 15 a borderline response, and 24 an extreme response.
+	\item In the vague verbal condition we used ``Choose a square with few dot''. We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
+	\item In the crisp numerical condition we used ``Choose the square with 6 dots''.
 The smaller flanking array always contained exactly the specified number of dots.
 We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
-	\item For crisp verbal, we used ``Choose the square with the fewest dots". We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
+	\item For crisp verbal, we used ``Choose the square with the fewest dots''. We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
 \end{itemize}
 
-On each trial, first the referring expression that constituted the instruction for that trial was displayed (e.g., ``Choose a square with about 10 dots").
+On each trial, first the referring expression that constituted the instruction for that trial was displayed (e.g., ``Choose a square with about 10 dots'').
 Participants then pressed a key to indicate that they had read the instruction.
@@
 In this experiment, no feedback was given.
-This was because, in the vague conditions, we did not regard any response as ``correct" or ``incorrect", but instead as ``expected response"; ``borderline response"; and ``extreme response", and we did not want to draw participants' attention to this distinction explicitly.
+This was because, in the vague conditions, we did not regard any response as ``correct'' or ``incorrect'', but instead as ``expected response''; ``borderline response''; and ``extreme response'', and we did not want to draw participants' attention to this distinction explicitly.
 Which choice the participant made was recorded for analysis.
@@
 \centering
-\caption{Experiment 1 instructions arranged by condition. The instructions given in the table started with ``Choose \ldots"}
+\caption{Experiment 1 instructions arranged by condition. The instructions given in the table started with ``Choose \ldots''}
 \label{instructionsC-exp-1}
@@
 \centering
-\includegraphics[width=.75\textwidth]{figures/Ce1-example-screenshot}
+\includegraphics[width=.75\textwidth]{figures/Ce1-example-screenshot-border}
 \caption{Experiment 1 and 2 example stimulus}
@@
 
-However, given that the response time plot in Figure \ref{resultsC-exp-1-RT} shows that responses to 6:15:24 in the "crisp numeric" instructions condition were extremely fast relative to the "vague numeric" instructions to 6:15:24, the effects in the model of the full dataset could be driven by this difference.
+However, given that the response time plot in Figure \ref{resultsC-exp-1-RT} shows that responses to 6:15:24 in the `crisp numeric' instructions condition were extremely fast relative to the `vague numeric' instructions to 6:15:24, the effects in the model of the full dataset could be driven by this difference.
 A clearer picture of the effects of interest might be obtained by removing the 6:15:24 level of Item from the data set, and fitting the model to this restricted data.
@@
 However, once again there is a possibly confounding factor.
-Observe that, in Experiment 1, instruction format (i.e., the difference between numeric and verbal) went hand in hand with might be called the (human) "selection algorithm": To see this, consider the task of selecting the dot array that contains "few dots":" to do this, it suffices to \emph{compare} the three arrays and select the one that contains the fewest elements.
-To select the dot array that contains "16 dots" seems to require the participant to estimate, and then \emph{match}, the cardinality of (at least) one dot array to 16, a process which could plausibly take longer, independently of vagueness.
+Observe that, in Experiment 1, instruction format (i.e., the difference between numeric and verbal) went hand in hand with might be called the (human) `selection algorithm': To see this, consider the task of selecting the dot array that contains `few dots': to do this, it suffices to \emph{compare} the three arrays and select the one that contains the fewest elements.
+To select the dot array that contains `16 dots' seems to require the participant to estimate, and then \emph{match}, the cardinality of (at least) one dot array to 16, a process which could plausibly take longer, independently of vagueness.
 Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use \emph{comparison} rather than having to resort to \emph{matching}.

### 2018-10-27 manuscript/e3.tex
removed undirected quote marks and replaced with aostrophe-style quotes

--- 2018-09-02 springer/current/e3.tex
@@
 \includegraphics[width=.5\textwidth]{figures/Ee3-flow.jpg}
-\caption{Experiment 3 example stimulus, showing the sequence of presentation, with the instruction screen at the top, followed by the presentation of the "target" array, followed by the stimulus that participants responded to.}
+\caption{Experiment 3 example stimulus, showing the sequence of presentation, with the instruction screen at the top, followed by the presentation of the ``target'' array, followed by the stimulus that participants responded to.}
 \label{Experiment3examplestimulus}
@@
 \centering
-\caption{Experiment 3 instructions for the smallest array, arranged by condition. The instructions given in the table started with ``Choose a square with \ldots"}
+\caption{Experiment 3 instructions for the smallest array, arranged by condition. The instructions given in the table started with ``Choose a square with \ldots''}
 \label{instructionsE-exp-3}

### 2018-10-27 manuscript/our-approach-to-the-problem.tex
removed undirected quote marks and replaced with aostrophe-style quotes

--- 2018-09-12 chapter-latex/our-approach-to-the-problem.tex
@@
 All $p$ values reported for linear models were calculated using the R package \emph{lmerTest} \citep{lmerTest}.
-The complete data and analysis for this experiment and the others in this paper are available at \texttt{https://github.com/mjgreen/vagueness}.
-%\texttt{\url{http://homepages.abdn.ac.uk/k.vdeemter/pages/vagueness-data-analysis/e1.html}
 
@@
 By contrast, a word is vague if it is not precise''.
-The quantifiers few and many might not have realised their potential for vagueness in the context of a choice between (only) two alternatives - where there is no borderline alternative that could result in few and many being "not well-defined".
+The quantifiers few and many might not have realised their potential for vagueness in the context of a choice between (only) two alternatives - where there is no borderline alternative that could result in few and many being ``not well-defined''.
 Furthermore using definite articles in the instructions may have reinforced the participant's impression that only one choice counted as correct, and this impression could have been reinforced by the use of error feedback.
@@
 In the experiments we report below (Experiments 1, 2 and 3) we used what we learned from the pilot experiments to design situations that we believe did address the difference between crisp and vague instructions while taking account of the most important alternative explanations of this crisp-vague difference.
+The complete data and analysis for the experiments in this paper are available at \texttt{https://mjgreen.github.io/vagueness}.

### 2018-10-27 manuscript/related-work.tex
removed undirected quote marks and replaced with aostrophe-style quotes

--- 2018-09-12 chapter-latex/related-work.tex
@@
 For instance, the vague word ``big'' can denote a mouse a well as an elephant.
-The idea is plausible, but the problem is that ``big" is efficient not because it is vague (i.e., not because it allows borderline cases) but because it is context dependent.
-To see this, note for example that superlatives are not vague (i.e., they do not permit borderline cases), but they {\em are} context dependent: ``the biggest $x$" ascribes different sizes to the referent depending on $x$.
+The idea is plausible, but the problem is that ``big'' is efficient not because it is vague (i.e., not because it allows borderline cases) but because it is context dependent.
+To see this, note for example that superlatives are not vague (i.e., they do not permit borderline cases), but they {\em are} context dependent: ``the biggest $x$'' ascribes different sizes to the referent depending on $x$.
 It is for this reason that we can use the word ``biggest''  to talk about the biggest mouse, the biggest elephant, and so on.

### 2018-10-27 manuscript/v-book-2018.tex
was chapter_latex
removed undirected quote marks and replaced with aostrophe-style quotes
merged separate tex into v-book-2018.tex

--- 2018-08-02 springer/current/v-book-2018.tex
@@
  ,sectrefs      %
- ,footinfo      %
- ,graphics      %
-%,referee
+% ,footinfo     %
+% ,referee
 ]{svmono}
-% Native packages
 \usepackage{mathptmx}
@@
 \usepackage{type1cm}
-\usepackage{makeidx}		    % allows index generation
-\usepackage{graphicx}		    % standard LaTeX graphics tool
-\usepackage{multicol} 		    % used for the two-column index
-\usepackage[bottom]{footmisc}	% places footnotes at page bottom
-% User packages
+\usepackage{graphicx}
+\usepackage{multicol}
+\usepackage[bottom]{footmisc}
 \usepackage{natbib}
@@
 \usepackage{placeins}
-\usepackage{multirow}			% added by Matt
-\usepackage{marginnote}			% added by Matt
-%\makeindex             		% used for the subject index: please use the style svind.ist with your makeindex program
+\usepackage{multirow}
 \begin{document}
-\author{Kees van Deemter and Matthew Green}
+\author{Matthew James Green and Kees van Deemter}
 \title{Vagueness \& Rationality}
-\subtitle{The Elusive Benefits of Vagueness: the Evidence So Far}
-\input{authoring-notes}\newpage~\newpage
+\subtitle{The Elusive Benefits of Vagueness: The Evidence So Far}
+\date{}
 \maketitle
@@
 \mainmatter
-\chapter{The Elusive Benefits of Vagueness: Evidence from Experiments}\label{chapterlabel}
-\input{abstract}
-\section{Introduction}\label{introduction}\input{introduction}
-\section{Related work}\label{related-work}\input{related-work}
-\section{Our Approach to the Problem}\label{our-approach-to-the-problem}\input{our-approach-to-the-problem}
-%\section{Experiment 0: the pilot experiment} % This was never in the submitted version
-%\section{Experiment 1: initial experiment}\label{e1}\input{e1}% The reviewer wants this experiment excised.
-\marginnote{the former E1 was excised here. The new numbering is the former E2,3,4 map to the new E1,2,3.}
-\section{Experiment 1: separating vagueness from instruction format}\label{e1}\input{e1}
-\section{Experiment 2: focus on instructions that contain numerals}\label{e2}\input{e2}
-\section{Experiment 3: focus on instructions that do not contain numerals}\label{e3}\input{e3}
-\section{Discussion of experiments 2 and 3}\label{discussion-of-e2-and-e3}\input{discussion-of-e2-and-e3}
-\section{General Discussion}\label{general-discussion}\input{general-discussion}
+
+\chapter{The Elusive Benefits of Vagueness: Evidence from Experiments}
+Much of everyday language is vague, even in situations where vagueness could have been avoided (i.e., where vagueness is used ``strategically'').
+Yet the benefits of vagueness for hearers and readers are proving to be elusive.
+We discuss a range of earlier controlled experiments with human participants, and we report on a new series of experiments that we ourselves have conducted in recent years.
+These experiments, which focus on vague expressions that are part of referential noun phrases, aim to separate the utility of vagueness (as defined by the existence of borderline cases) from the utility of other factors that tend to co-occur with vagueness.
+After presenting the evidence, we argue that it supports a view where the benefits that vague terms exert are due to other influences, and not to vagueness itself.
+
+\section{Introduction}
+In academic use, the word `vagueness' has a specific meaning.
+Keefe and Smith, for example, state that ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \citep[p.~4]{keefe1997vagueness}, as do \citet{EgreKlinedinst}.
+The crucial criterion is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects.
+By contrast, a word is vague if it is not precise'' \citet[p.~1]{lipmanvague}.
+A typical example is the word ``tall'', as applied to people for example, because there is no precise, known height which separates those who are tall from those who are not.
+The crucial point is that ``tall'' admits borderline cases (i.e., people who may or may not count as tall), which are the hallmark of vagueness as we use the term.
+
+Linguists, philosophers of language, and more recently game theorists have asked why natural languages contain so many vague expressions \citet{Lipman:2000fk, lipmanvague}, which are
+often used even in situations where the speaker could have used an expression that is not vague (i.e., crisp); in these situations we say that vagueness is used \emph{strategically}.
+By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss'' \citet[][p.~1]{lipmanvague}.
+Lipman explains the point by means of a scenario in which a speaker describes a person to a hearer, who needs to identify that person in the arrivals hall of of an airport.
+In such a scenario, a precise description of the person's height (e.g., ``The person's height is 187.96 cm'') would be more useful than a vague one (``The person is tall'').
+Lipman uses this scenario to explain why standard game theory models of communication \citep[e.g.,][]{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs.
+
+Lipman argues that the efficiency loss resulting from vague expressions would be unlikely to have arisen unless there were advantages as well as disadvantages associated with vague expressions.
+Lipman asks, essentially, what these advantages might be.
+Several tentative answers to Lipman's question have been offered \citep[see][]{van2009utility, vanDeemterBook}.
+One of the most promising answers appears to be the idea that vague expressions are easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \citep[e.g.,][]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
+For example, \citeauthor{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze'' (\citeyear[][p.~11]{lipmanvague}).
+We shall refer to this as the \emph{cost reduction} hypothesis.
+
+This article brings an experimental approach to these issues, focussing on vagueness in descriptions (e.g., ``the square with few dots'') and its effect on the hearer's ability to act on a given description, as measured by the time that it takes hearers to click on the referent of a description.\footnote{Other metrics could have been chosen, such as hearers' ability to remember information, for example, or error rates.
+Although error rates play a minor role in the present paper, for reasons that will become clear, we focus on response times in particular.}
+We find that, although hearer benefits from vague descriptions are straightforward to demonstrate in many cases, a closer experimental analysis militates against the conclusion that vagueness itself --- as defined above, in terms of the existence of borderline cases --- lies at the heart of these results.
+Instead, it is other factors, such as the presence of an overt numerical expression in the description, that proved to be decisive.
+We believe that, despite the fact that our experiments are unavoidably focussed on a specific class of vague expressions (since any experiment can only deal with a limited number of different stimuli), these findings are potentially important, because they call into question whether ``strategic'' vagueness (i.e., vagueness where the speaker had a choice, because she could have produced a crisp expression instead) has any advantages at all.
+In other words, returning to Lipman's question, it is possible that vagueness has evolved partly as a necessary evil (e.g., because of the limits of observation and prediction) and partly as a side effect of other factors.
+
+\section{Related work}\label{related-work}
+A number of answers to Lipman's question have been proposed, but as we have argued elsewhere, it appears that some of these do not to hold up to scrutiny \citep{van2009utility,vanDeemterBook}.
+For example,
+\begin{itemize}
+\item It has been proposed that the existence of vague words makes a language more efficient \citep{barwiseperry}.
+The idea is that vagueness allows the use of one and the same word (e.g., ``big'') in different situations, which require different standards.
+For instance, the vague word ``big'' can denote a mouse as well as an elephant.
+The idea is plausible, but the problem is that ``big'' is efficient not because it is vague (i.e., not because it allows borderline cases) but because it is context dependent.
+To see this, note for example that superlatives are not vague (i.e., they do not permit borderline cases), but they {\em are} context dependent: ``the biggest $x$'' ascribes different sizes to the referent depending on $x$.
+It is for this reason that we can use the word ``biggest''  to talk about the biggest mouse, the biggest elephant, and so on.
+Context-dependence, however, does not imply vagueness.
+\item Vague words are capable of combining a statement of quantity with an evaluative statement \citep{veltman}.
+The idea is that when we say ``The patient has a fever,'' we do not merely assert that her temperature lies above a certain value, we also imply that the deviation is clinically significant (i.e., something is not right with the patient).
+The problem with this plausible idea is that evaluation does not imply vagueness.
+For example, the medical term ``obese'' is evaluative (i.e., it is not healthy to be obese), yet its standard medical definition in terms of Body Mass Index is perfectly crisp (i.e., without borderline cases), with all and only BMI values above 30 counting as obese.
+\end{itemize}
+Analogous observations can be made about a number of other purported benefits of vagueness; one of the few hypotheses left standing at the moment is the {\em cost reduction} hypothesis, which we mentioned in the previous section.
+The findings on which we are reporting in the present paper will follow a familiar pattern: we will explore the cost reduction hypothesis, only to conclude that it may not stand up to scrutiny.
+Our approach, this time, will be experimental: we conduct controlled experiments to investigate the effect that vague expressions have on a hearer.
+
+Charting the utility of vagueness is also the attested aim of a small number of previous experimental studies.
+But, as it happens, few of these studies have truly focussed on vagueness in the sense in which we do here (i.e., they do not address Lipman's challenge).
+Two recent studies illustrate this issue; let's discuss them briefly.
+
+In a series of studies of behaviour modification, \citet{Mishra01042011} manipulated the presentation format of information about quantities in the domains of mental acuity, physical strength, and weight loss.
+In the weight loss study, participants were told that the study was designed to test the validity of a new (actually fictitious) health index, the HHI (Holistic Health Index).
+They were told that an ideal HHI score lies in the range of 45 to 55.
+In a longitudinal study, participants submitted their weight to a computer each week.
+Participants were told that two algorithms would be used to compute their HHI, and that the two might give different values initially, in which case the true score lay between the two values.
+In one condition, which the authors called the precise condition, the two algorithms gave the same score.
+In the other condition, which the authors called the vague condition, one algorithm added 3\% to the score while the other algorithm subtracted 3\% from the score, yielding a range of values whose midpoint was the same as the two values given in the precise condition.
+
+One group of participants was given HHI scores in the ideal range: for this group their weight loss did not differ depending on whether they were given vague or precise HHI values.
+However for the other group, who were given HHI scores outside the ideal range, their weight loss was significantly greater if they were given vague HHI scores than if they were given precise HHI scores.
+The authors explain the improvement in the vague condition for this group as resulting from the participants' freedom to think of themselves a positioned on one end of the range --- the end closest to the ideal HHI scores.
+This ``illusion of proximity'' \citep[][p.~4]{Mishra01042011} to the goal is argued to have allowed participants to generate positive expectancies that lead to behaviours that improved performance.
+In contrast, in the precise conditions, participants did not have this freedom of interpretation, and could not distort the information to bring about the beneficial \emph{illusion of proximity}.
+These results are interesting, and of obvious potential practical importance.
+We note, however, that information presented as an exact range of values does not conform with the standard definition of vagueness \citep{keefe1997vagueness, EgreKlinedinst}, since an exact range does not admit borderline cases.
+In the terminology of \citeauthor{Hobbs85granularity}, the difference between a range and a single midpoint value is a difference of \emph{granularity}.
+Furthermore, the experiments of \citeauthor{Mishra01042011} did not explore benefits in terms of processing cost, but in terms of long-term behaviour change.
+
+Similar issues arise from the work of \citet{peters2009bringing}.
+The authors carried out a series of studies where participants were required to rate hospitals based on various sources of information about quality of care.
+There was a between-subjects manipulation based on numeracy.
+The format of the information was manipulated within subjects: either numbers only were presented, or both numbers and evaluative categories were presented (e.g., \emph{Poor}, \emph{Fair}, \emph{Good}, \emph{Excellent}, with crisp visual boundary lines between the categories).
+Results showed that, for low-numeracy participants, the presence of evaluative categories resulted in a diminished influence of an irrelevant affective state on the ratings.
+For all participants, the presence of evaluative categories resulted in better decisions and in a greater use of the most important and reliable types of information, such as survival rates.
+
+It is, however, questionable whether the ``evaluative categories'' manipulation in this study can be considered a manipulation of vagueness.
+Certainly, terms like \emph{Fair} admit the possibility of borderline cases.
+However, given that the boundaries between the categories were marked crisply, and that therefore the categories mapped crisply to numerical values, it becomes doubtful whether any borderline cases could be conceived to arise in fact.
+For example, \emph{Fair} was mapped to 60--70\% for the variable \emph{percentage of heart attack patients given recommended treatment (ACE inhibitor)}.
+Accordingly, rather than the vagueness of categories such as \emph{Poor}, Peters et al.
+emphasise the evaluative content inherent in these categories, and the affective potential of the evaluative content rather than the vagueness of the terms like {\em Fair}.
+
+\section{Our Approach to the Problem}
+Are vague expressions processed more easily by readers than crisp ones?
+Like Lipman, we focus on situations where numerical information is used in order to identify a referent.
+Reference, in other words, will be the communicative task on which we focus, partly because of the interest that this topic has recently drawn in various areas of Cognitive Science \citep{vanDeemterCMR}.
+By looking at one specific type of vagueness, we will be able to investigate the costs and benefits of vagueness relatively thoroughly.
+Whether our findings generalise to other uses of vagueness is a question on which we will speculate in the final section of this chapter.
+
+We have chosen a narrative strategy in which we address a sequence of four experiments with human readers chronologically, explaining how each experiment helped us refine our research question.
+In order to do justice to our findings, we need to describe these experiments in a fair amount of detail.
+
+Let us start by explaining the task that was given to the participants in our experiments.
+We used a {\em speeded forced choice} task to compare the processing costs of different references to quantities.
+In this context, speed and accuracy of responses are the key dimensions on which the different references can be compared.
+Each stimulus in the experiments was a set of dot arrays containing various number of dots, together with a preceding instruction (in the form of a referring expression) to choose one of the arrays with respect to its cardinality.
+The participant was asked to respond as quickly as possible while avoiding errors.
+We manipulated the instructions and the arrays in several ways across the four experiments.
+
+All the experiments shared the following properties: stimuli were created using the language GNU Octave \citep{eaton:2002} and the Psychophysics Toolbox extensions \citep{ptbx1, ptbx2}.
+The position of the dots was randomised per-trial.
+The order in which trials were presented was randomised per-participant.
+There were 256 trials, presented in 4 blocks of 64 each, between which the participant could rest.
+A MacBook Pro laptop computer with a 13-inch screen presented the stimuli to the participants and recorded responses.
+Participants were recruited using email lists at the University of Aberdeen, and paid ten pounds for participating.
+All participants self-reported fluency in English, and had normal, or corrected-to-normal vision.
+The experiment was conducted in a quiet room.
+Participants were asked to respond as quickly as possible while avoiding errors.
+There was a block of practice trials after which participants could ask questions, following which the experimenter left the room.
+All $p$ values reported for linear models were calculated using the R package \emph{lmerTest} \citep{lmerTest}.
+
+When the distance grows between two numbers, they become more easily distinguishable: the \emph{numerical distance effect} has been shown for comparing the cardinality of two sets of dots \citep{van123} and for processing Arabic numerals and number words \citep{Dehaene1996}.
+We manipulated the number of dots in each array such that some sets of arrays had smaller numerical distances and others had larger numerical distances.
+Where a number was mentioned in the instructions, it was always in the form of an Arabic numeral.
+When two numbers are presented with the smaller on the left, this left-side presentation facilitates responses indicating the smaller number: the \emph{Spatial-Numerical Association of Response Codes (SNARC)} effect \citep{dehaene1993mental, gevers2006automatic}.
+We controlled which side the smaller number appeared on to avoid systematic influences of this effect.
+
+There is abundant evidence \citep[e.g.,][]{trick1994small} that very small (i.e., \emph{subitizable}) quantities are recognised and processed by a distinct psychological mechanism that differs from that used to process larger quantities.
+We performed a pilot experiment \citep{green2011costreduction} in which we were able to confirm this finding in the experimental settings on which we are focussing in this paper.
+We found that, when participants were confronted with a stimulus consisting of two squares containing different numbers of dots.\footnote{Such a stimulus is referred to hereafter as consisting of a set of {\em dot arrays} The number of dots in an array is referred to as its cardinality.
+The physical arrangement of dots in each array is irregular.}, instructions of the form \emph{Choose the square with \emph{n} dots} led to consistently faster response times than instructions of the form \emph{Choose the square with many/few dots} when $2 \leq n \leq 5$; the converse was true for $n>5$.
+Given these findings, we henceforth focussed our studies on non-subitizable numbers, because it is there that vagueness is expected to have benefits.
+
+In a second pilot experiment \citep{green2013utility}, we again presented two dot arrays with the instruction in the form \emph{Choose the square with \ldots dots}.
+The arrays contained larger numbers of dots than in the first pilot: one array always contained 25 dots, and the other contained either 5, 10, 15, 20, 30, 35, 40, or 45 dots.
+Each stimulus can therefore be seen in terms of the numerical difference between the number of dots in one array and the number of dots in the other: giving numerical distances of 5, 10, 15, and 20, with smaller numerical distances resulting in less discriminable arrays and larger distances resulting in more discriminable arrays.
+Our main manipulation was of the vagueness of the instruction, with two levels, \emph{crisp} and \emph{vague}.
+Assuming the dot array (5, 25), and the instruction referring to the smaller cardinality, the \emph{crisp} instruction was \emph{Choose the square with 5 dots} and its \emph{vague} counterpart was \emph{Choose the square with few dots}.
+We found, as expected, that responses were faster and more accurate for vague instructions than for crisp instructions.
+We also found an interaction between vagueness and numerical distance such that there were diminishing returns for vagueness as numerical distance increased, until, at the biggest numerical distance, there was no real difference any more between crisp and vague instructions.
+We interpreted this pattern as showing that cognitive load is relatively easy in both conditions when the stimuli are most discriminable, and so vagueness confers no additional advantage in those most easily discriminable stimuli.
+
+However, the picture painted by these findings from the second pilot experiment might be misleading.
+First of all, there is a possibly confounding factor.
+Contrast an expression from the vague condition: `the square with few dots' with an expression from the crisp condition: `the square with 5 dots'.
+One difference is that `few' has the potential for vagueness, whereas `5' is crisp.
+But another difference is that `few' is verbal while `5' is numerical, in the sense that a number is mentioned explicitly.
+Since these two differences could not be separated in the experiment, the finding of a vagueness advantage is vulnerable to an alternative interpretation, namely that the crisp--vague difference was really an advantage for the verbal form of the quantifier.
+In our next experiment (reported below as Experiment 1) we therefore created verbal and numeric versions of each of the vague and crisp instructions so that we could compare the crisp--vague difference and the numeric--verbal difference in the same experiment.
+
+In Experiment 1, reported below, we also addressed another issue with the second pilot experiment: participants chose one of two dot arrays --- therefore the `vague' quantifiers (few and many) uniquely identified one square.
+Recall our definition of vague --- ``a word is precise if it describes a well-defined set of objects.
+By contrast, a word is vague if it is not precise''.
+The quantifiers `few' and `many' might not have realised their potential for vagueness in the context of a choice between (only) two alternatives --- where there is no borderline alternative that could result in few and many being ``not well-defined''.
+Furthermore, using definite articles in the instructions may have reinforced the participant's impression that only one choice counted as correct, and this impression could have been reinforced by the use of error feedback.
+
+In the experiments we report below (Experiments 1, 2, and 3) we used what we learned from the pilot experiments to design situations that we believe did address the difference between crisp and vague instructions while taking account of the most important alternative explanations of this crisp--vague difference.
+The complete data and analysis for the experiments in this paper are available at \texttt{https://mjgreen.github.io/vagueness}.
+
+\section{Experiment 1: Separating vagueness from instruction format}
+To find out what happens when words are used in a context where their potential for vagueness comes to the fore, Experiment 1 used three arrays (rather than two) so that the vague description had more than one possible referent, used indefinite articles to avoid the impression that only one response counted as correct, and was carried out without error feedback.
+An indication that the potential for vagueness was indeed realised in Experiment 1 is that the borderline response was chosen fairly often: 16\% of the time.
+
+In Experiment 1, an item was an instruction followed by a set of three dot arrays defined by a triple of numbers, representing the number of dots in the left, middle, and right arrays.
+We used four different triples of numbers: (6,15,24); (16,25,34); (26,35,44); (36,45,54).
+Each set of arrays comprised three arrays (instead of two as in \citet{green2013utility}); the array representing the central number was always presented in the middle of the three; there were two flanking arrays where one had fewer dots than the central array and the other had more, and these flanking arrays appeared equally often on the left and right of the central array.
+
+Table~\ref{instructionsC-exp-1} gives the full set of stimuli and associated instructions.
+The way in which borderline responses were construed is as follows, using as an example the array (6:15:24) and instructions that identified the smaller flanking array (6).
+6 was classified as the expected response.
+15 was classified as the borderline response.
+24 was classified as the extreme response.
+
+\begin{itemize}
+	\item In the vague numerical condition the instruction was ``Choose a square with about 10 dots'' --- none of the arrays contained exactly 10 dots, but 10 is closer to 6 than it is to 15, making 6 a better response to that instruction, 15 a borderline response, and 24 an extreme response.
+	\item In the vague verbal condition we used ``Choose a square with few dot''. We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
+	\item In the crisp numerical condition we used ``Choose the square with 6 dots''.
+The smaller flanking array always contained exactly the specified number of dots.
+We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
+	\item For crisp verbal, we used ``Choose the square with the fewest dots''. We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
+\end{itemize}
+
+On each trial, first the referring expression that constituted the instruction for that trial was displayed (e.g., ``Choose a square with about 10 dots'').
+Participants then pressed a key to indicate that they had read the instruction.
+The instruction remained on screen, and after 1000 ms, the arrays appeared.
+An example stimulus is given in Figure~\ref{Experiment1and2examplestimulus}.
+Response time was measured from the presentation of the arrays until the keypress indicating the participant's choice.
+The trial would timeout after 60 seconds if there was no response.
+In this experiment, no feedback was given.
+This was because, in the vague conditions, we did not regard any response as ``correct'' or ``incorrect'', but instead as ``expected response''; ``borderline response''; and ``extreme response'', and we did not want to draw participants' attention to this distinction explicitly.
+The participant's choice was recorded for analysis.
+
+\begin{table}[htbp]
+\centering
+\caption{Experiment 1 instructions arranged by condition. The instructions given in the table started with ``Choose \ldots''}
+\label{instructionsC-exp-1}
+\begin{tabular}{cccll}
+\hline\noalign{\smallskip}
+Item & Quantity & Number & Crisp & Vague\\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+06:15:24 & Small & Numeric & the square with 6 dots & a square with about 10 dots\\
+06:15:24 & Small & Verbal & the square with the fewest dots & a square with few dots\\
+06:15:24 & Large & Numeric & the square with 24 dots & a square with about 20 dots\\
+06:15:24 & Large & Verbal & the square with the most dots & a square with many dots\\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+16:25:34 & Small & Numeric & the square with 16 dots & a square with about 20 dots\\
+16:25:34 & Small & Verbal & the square with the fewest dots & a square with few dots\\
+16:25:34 & Large & Numeric & the square with 34 dots & a square with about 30 dots\\
+16:25:34 & Large & Verbal & the square with the most dots & a square with many dots\\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+26:35:44 & Small & Numeric & the square with 26 dots & a square with about 30 dots\\
+26:35:44 & Small & Verbal & the square with the fewest dots & a square with few dots\\
+26:35:44 & Large & Numeric & the square with 44 dots & a square with about 40 dots\\
+26:35:44 & Large & Verbal & the square with the most dots & a square with many dots\\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+36:45:54 & Small & Numeric & the square with 36 dots & a square with about 40 dots\\
+36:45:54 & Small & Verbal & the square with the fewest dots & a square with few dots\\
+36:45:54 & Large & Numeric & the square with 54 dots & a square with about 50 dots\\
+36:45:54 & Large & Verbal & the square with the most dots & a square with many dots\\
+\noalign{\smallskip}\hline
+\end{tabular}
+\end{table}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.75\textwidth]{figures/Ce1-example-screenshot-border}
+\caption{Experiment 1 and 2 example stimulus}
+\label{Experiment1and2examplestimulus}
+\end{figure}
+
+\subsection{Hypotheses (Experiment 1)}
+We formulated the following hypotheses for Experiment 1:
+\begin{description}
+	\item [Hypothesis 1 (Crisp/Vague RT)] Vague instructions should result in faster responses than crisp instructions; and this pattern should hold when the model is restricted to numeric-only data and when it is restricted to verbal-only data.
+	\item [Hypothesis 2 (Numeric/Verbal RT)] There should be no real difference between responses to numeric instructions and verbal instructions (based on our interpretation of the experiment in \citet{green2013utility}, where we thought that vague instructions alone were driving the advantage for instructions that were both vague, and also in verbal format).
+	\item [Hypothesis 3 (Item RT)] Responses should take longer as the number of dots in the display grows larger (i.e., as the levels of \emph{Item} increase).
+	\item [Hypothesis 4 (Response Type)] Vague instructions should lead to more borderline responses than crisp instructions.
+\end{description}
+
+\subsection{Results (Experiment 1)}
+\paragraph{\textbf{Response times}}
+30 participants were recruited.
+Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 236 trials, 3.1\% of the data.
+The distribution of remaining response times was skewed with many long responses.
+These remaining response times were log-transformed, which reduced this skew so that their distribution more closely approximated a normal distribution.
+Condition means for response times are given in Figure~\ref{resultsC-exp-1-RT}.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=\textwidth]{figures/Ce1-rtplot-1.pdf}
+\caption{Experiment 1 results: Mean response times by condition and item}
+\label{resultsC-exp-1-RT}
+\end{figure}
+
+A linear mixed model was constructed for the (log-transformed) response times, with sum-coded vagueness, instruction format, and their interaction, and item, as fixed effects, and per-participant intercepts and slopes for sum-coded vagueness, instruction format, and their interaction as random effects.
+
+\begin{description}
+	\item [Test of Hypothesis 1 (Crisp/Vague RT)] Vague instructions actually led to significantly slower responses than crisp instructions, against Hypothesis 1 ($\beta=0.058$, $se=0.013$, $t=4.55$, $p<0.001$).
+When the model was restricted to numeric-only instructions, vague instructions still led to significantly slower responses than crisp instructions ($\beta=0.093$, $se=0.021$, $t=4.51$, $p<0.001$).
+When the model was restricted to verbal-only instructions, vague instructions tended to slow responses, but not significantly ($\beta=0.024$, $se=0.016$, $t=1.46$, $p=0.155$).
+	\item [Test of Hypothesis 2 (Numeric/Verbal RT)] There was actually a significant difference between numeric and verbal instructions, with numeric instructions leading to longer responses than verbal instructions, against Hypothesis 2 ($\beta0.265=$, $se=0.072$, $t=5.08$, $p<0.001$).
+	\item [Test of Hypothesis 3 (Item RT)] Responses took longer as the levels of \emph{Item} increased, supporting Hypothesis 3 ($\beta=0.120$, $se=0.017$, $t=7.11$, $p<0.001$).
+\end{description}
+
+However, given that the response time plot in Figure~\ref{resultsC-exp-1-RT} shows that responses to 6:15:24 in the `crisp numeric' instructions condition were extremely fast relative to the `vague numeric' instructions to 6:15:24, the effects in the model of the full dataset could be driven by this difference.
+A clearer picture of the effects of interest might be obtained by removing the 6:15:24 level of Item from the data set, and fitting the model to this restricted data.
+Doing this did not affect the direction of the effects in the full dataset, but whereas the effects were significant in the full dataset, they were not significant in the restricted dataset.
+Full details of the analysis of the restricted dataset are available in the online materials.
+
+\paragraph{\textbf{Borderline cases}}
+
+A generalized linear mixed model \citep{jaeger2008categorical} was fit to the data for the distribution of responses indicating the borderline response, with sum-coded vagueness, instruction format, (and their interaction), and item as fixed effects, and the same effects as slopes over participant, as well as per-participant intercepts as random effects.
+The distribution of responses over the nearest match square, the borderline square, and the furthest match square are given in Figure~\ref{resultsC-exp-1-BL}.
+Participants chose the borderline square on $16.6\%$ of trials overall.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=\textwidth]{figures/Ce1-blBarChart-1}
+\caption{Experiment 1 results: Counts of borderline case responses by condition.}
+\label{resultsC-exp-1-BL}
+\end{figure}
+
+\begin{description}
+	\item [Test of Hypothesis 4 (Response Type)] Participants were significantly more likely to choose the borderline option for vague instructions than for crisp instructions (21.9\% vs 11.3\%: $\beta=0.62$, $se=0.22$, $z=2.8$, $p=0.0059$).
+Participants were also significantly more likely to choose the borderline square when the instruction used the numerical format rather than the verbal format (30.1\% vs 3.0\%: $\beta=-3.35$, $se=0.23$, $z=-14.6$, $p<0.0001$).
+\end{description}
+
+\subsection{Discussion (Experiment 1)}
+
+This experiment tested whether vague instructions would result in faster responses than crisp instructions, when borderline cases were present.
+Faster responses for vague instructions were found in pilot experiment B, but there were no borderline cases in that experiment.
+
+In this experiment we found in contrast that vague instructions resulted in slower responses than crisp instructions: a difference that was significant when considering the full data (112ms), but which was not significant after removing the smallest arrays from the analysis, which had a pattern opposite to the main trends in the rest of the data.
+
+We also found that the effect of instruction format was significant, with numerical format slowing responses by 689 ms on average, such that the disadvantage of numerical format overwhelmed the contribution of vagueness.
+The verbal vague condition still yielded faster responses than the numerical crisp condition, so the pattern from pilot experiment B was reproduced, but in the light of the evidence from this experiment (Experiment 1), in the presence of borderline cases, the advantage that was ascribed to vagueness before now looks more like an advantage of verbal instruction format.
+
+However, once again there is a possibly confounding factor.
+Observe that, in Experiment 1, instruction format (i.e., the difference between numeric and verbal) went hand in hand with what might be called the (human) `selection algorithm': To see this, consider the task of selecting the dot array that contains `few dots': to do this, it suffices to \emph{compare} the three arrays and select the one that contains the fewest elements.
+To select the dot array that contains `16 dots' seems to require the participant to estimate, and then \emph{match}, the cardinality of (at least) one dot array to 16, a process which could plausibly take longer, independently of vagueness.
+Therefore, our results so far permit the interpretation that what made the instructions in the verbal condition fast is not the fact that they were worded verbally, but that they allowed participants to use \emph{comparison} rather than having to resort to \emph{matching}.
+
+In the next two experiments we pitted the comparison algorithm and matching algorithm selection tasks against each other while controlling vagueness and instruction format.
+In Experiment 2 we restricted all the instructions to numeric quantifiers while factorially manipulating vagueness and selection task.
+In Experiment 3 we ensured that all instructions used verbal quantifiers, while also factorially manipulating `vagueness' and `selection task'.
+This allowed us to distinguish between the predictions of the selection task account and the instruction format account.
+
+\section{Experiment 2: Focus on instructions that contain numerals}
+The main aim of Experiment 2 was to see whether vagueness would exert beneficial effects when all conditions used numerals in the instructions, and when there were vague and crisp versions of the instructions for both comparison and matching strategies.
+The main changes from Experiment 1 were that the human selection task was explicitly controlled (i.e., whether the task amounted to matching or comparison), and that all conditions were constrained to mention a number.
+We used the same arrays as in Experiment 1 (an example stimulus is given in Figure~\ref{Experiment1and2examplestimulus}).
+We used a 2 x 2 factorial manipulation of vagueness and selection task (see Table~\ref{Instructions for e2}).
+On each trial an instruction was presented: participants pressed a key to dismiss the instruction, at which time the dot arrays were presented until the participant responded, and the response time and choice were recorded.
+Table~\ref{Instructions for e2} shows the instructions for each condition.
+Note the difference between ``fewer than 20'' and ``far fewer than 20'': whereas the former cannot have borderline cases (i.e., for each number it is clear whether the number is smaller than 20 or not), the latter can.
+
+\begin{table}
+\centering
+\caption{Experiment 2: Instructions arranged by condition. The instructions given in the table started with ``Choose a square with \ldots''}
+\label{Instructions for e2}
+\begin{tabular}{cccll}
+\hline\noalign{\smallskip}
+Item & Quantity & Selection & Crisp & Vague \\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+06:15:24 & Small & Comparison & fewer than 20 dots & far fewer than 20 dots \\
+06:15:24 & Small & Matching & 6 dots & about 10 dots \\
+06:15:24 & Large & Comparison & more than 10 dots & far more than 10 dots \\
+06:15:24 & Large & Matching & 24 dots & about 20 dots \\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+16:25:34 & Small & Comparison & fewer than 30 dots & far fewer than 30 dots \\
+16:25:34 & Small & Matching & 16 dots & about 20 dots \\
+16:25:34 & Large & Comparison & more than 20 dots & far more than 20 dots \\
+16:25:34 & Large & Matching & 34 dots & about 30 dots \\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+26:35:44 & Small & Comparison & fewer than 40 dots & far fewer than 40 dots \\
+26:35:44 & Small & Matching & 26 dots & about 30 dots \\
+26:35:44 & Large & Comparison & more than 30 dots & far more than 30 dots \\
+26:35:44 & Large & Matching & 44 dots & about 40 dots \\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+36:45:54 & Small & Comparison & fewer than 50 dots & far fewer than 50 dots \\
+36:45:54 & Small & Matching & 36 dots & about 40 dots \\
+36:45:54 & Large & Comparison & more than 40 dots & far more than 40 dots \\
+36:45:54 & Large & Matching & 54 dots & about 50 dots \\
+\noalign{\smallskip}\hline
+\end{tabular}
+\end{table}
+
+\subsection{Hypotheses (Experiment 2)}
+
+For Experiment 2, we formulated the following hypotheses:
+
+\begin{description}
+	\item [Hypothesis 1 (Crisp/Vague RT)] Vague instructions should result in faster responses than crisp instructions.
+	\item [Hypothesis 2 (Comparison/Matching RT)] Instructions that allow comparison should result in faster responses than instructions that necessitate matching.
+	\item [Hypothesis 3 (Interaction)] The vagueness effect should differ according to whether the selection task is comparison or matching.
+\end{description}
+
+\subsection{Results (Experiment 2)}
+
+38 participants were recruited.
+Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 204 trials (2.8\% of the trials).
+The distribution of remaining response times was skewed with many long responses.
+These remaining response times were log-transformed, which reduced this skew so that their distribution more closely approximated a normal distribution.
+Condition means for response times are plotted in Figure~\ref{resultsD-exp-2}.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=\textwidth]{figures/De2-rtplot-1.pdf}
+\caption{Mean response times by condition for Experiment 2 where all instructions were numeric}
+\label{resultsD-exp-2}
+\end{figure}
+
+A linear mixed model was constructed for the logged response times, with sum-coded vagueness, selection task, and their interaction, and item as fixed effects, and per-participant intercepts and slopes for sum-coded vagueness, selection task, and their interaction, and item for random effects.
+
+\begin{description}
+	\item [Test of Hypothesis 1 (Crisp/Vague RT)] Vague instructions resulted in faster responses than crisp instructions on average.
+However this difference was not significant in the full model ($\beta=-0.0057$, $se=0.0137$, $t=-0.42$, $p=0.678$).
+Using Levy's method \citep{Levy:MainEffectsInteractions} to test for main effects in the presence of higher-order interactions, by doing model comparison between a null model that included all interaction terms involving vagueness but leaving out a term for the main effect of vagueness, against a full model that differed only by including vagueness as a main effect, showed that the full model was no better than the reduced model ($df=1$, $p=0.676$), consituting more evidence that vagueness did not exert a significant main effect on response times.
+	\item [Test of Hypothesis 2 (Comparison/Matching RT)] Instructions involving comparison resulted in faster responses than instructions involving matching, and the difference was significant ($\beta=0.1618$, $se=0.0255$, $t=6.34$, $p<0.001$).
+	\item [Test of Hypothesis 3 (Interaction)] The interaction between vagueness and selection task was significant ($\beta=0.1306$, $se=0.0205$, $t=6.38$, $p<0.001$), showing that vagueness exerted different effects on response times according to whether the instructions involved comparison or matching.
+When separate analyses were carried out testing for the effect of vagueness in comparison-only and in matching-only conditions, vague instructions led to significantly \emph{faster} responses than crisp instructions in the comparison-only conditions ($\beta=-0.071$, $se=0.020$, $t=-3.52$, $p<0.01$); and in the matching-only conditions vague instructions led to significantly \emph{slower} responses than crisp instructions ($\beta=0.062$, $se=0.021$, $t=2.91$, $p<0.01$).
+\end{description}
+
+\subsection{Discussion (Experiment 2)}
+
+The cost reduction account predicted that there should be a significant main effect of vagueness such that responses would be faster for vague intructions than for crisp instructions.
+We found that although there was a very small effect in that direction, the effect was not statistically significant.
+Models that differed only in the presence of vagueness as a main effect were shown not to differ significantly in their explanatory value.
+However we did find that vagueness exerted effects on other variables: vagueness speeded RTs in the comparison task and slowed RTs in the matching task.
+
+\section{Experiment 3: Focus on instructions that do not contain numerals}
+This experiment mirrors Experiment 2, but focuses on instructions that do not use a number.
+We manipulated vagueness and the selection task (comparison and matching).
+In order to implement the experiment without mentioning numbers in the instructions, we changed the sequence of each trial to include a `target' (i.e., a dot array of a particular cardinality) before the array, so that we could then refer to the target's cardinality in the instruction using expressions like \emph{the same number of dots as the target}; \emph{fewer dots than the target}.
+An example of this sequence is given in Figure~\ref{Experiment3examplestimulus}.
+This presentation of a target before the main body of the trial shares some features with \citeauthor[Experiment 2]{Izard20081221}, although in that experiment participants were told the cardinality of the target (called an \emph{inducer} in that paper), whereas in our experiment we did not tell participants the cardinality of the prime array.
+An item was thus a combination of a target dot array, an instruction that did not contain a number, and a set of dot arrays taking their cardinalities from the same triples used in Experiments 1 and 2.
+Table~\ref{instructionsE-exp-3} spells out how the instructions were constrained so as not to mention a numeral and gives examples of targets.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.5\textwidth]{figures/Ee3-flow.jpg}
+\caption{Experiment 3 example stimulus, showing the sequence of presentation, with the instruction screen at the top, followed by the presentation of the ``target'' array, followed by the stimulus to which participants responded.}
+\label{Experiment3examplestimulus}
+\end{figure}
+
+\begin{table}
+\centering
+\caption{Experiment 3 instructions for the smallest array, arranged by condition. The instructions given in the table started with ``Choose a square with \ldots''}
+\label{instructionsE-exp-3}
+\resizebox{\textwidth}{!}{\begin{tabular}{ccccl}
+\hline\noalign{\smallskip}
+Item & Target & Selection & Vagueness & Instruction \\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+06:15:24 &   6 & Matching 	& Crisp 	& the same number of dots as the target 	    \\
+06:15:24 &  10 & Matching	& Vague		& about the same number of dots as the target	\\
+06:15:24 &  20 & Comparison & Crisp 	& fewer dots than the target             	    \\
+06:15:24 &  20 & Comparison	& Vague		& far fewer dots than the target				\\
+\noalign{\smallskip}\hline
+\end{tabular}}
+\end{table}
+
+\subsection{Hypotheses (Experiment 3)}
+
+For Experiment 3, we hypothesised:
+
+\begin{description}
+	\item [Hypothesis 1 (Crisp/Vague RT)] Vague instructions are easier for the reader than crisp ones.
+	\item [Hypothesis 2 (Comparison/Matching RT)] Comparison is easier for the reader than matching.
+	\item [Hypothesis 3 (Interaction)] The effect of vagueness differs depending on whether the selection task mandated by the instructions is matching or comparison.
+\end{description}
+
+\subsection{Results (Experiment 3)}
+
+40 volunteers participated.
+Response times from all trials were trimmed at 2.5 standard deviations for each subject, leading to the loss of 211 trials (2.8\% of the trials).
+The distribution of remaining response times was skewed with many long responses.
+These remaining response times were log-transformed, which reduced this skew so that their distribution more closely approximated a normal distribution.
+Condition means for response times are plotted in Figure~\ref{resultsE-exp-3}.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=\textwidth]{figures/Ee3-rtplot-1.pdf}
+\caption{Mean response times by condition for Experiment 3, where all instructions were verbal}
+\label{resultsE-exp-3}
+\end{figure}
+
+A linear mixed model was constructed for the logged response times, with sum-coded vagueness, selection task, and their interaction, and item as fixed effects, and per-participant intercepts and slopes for sum-coded vagueness, selection task, and their interaction, and item for random effects.
+
+\begin{description}
+	\item [Test of Hypothesis 1 (Crisp/Vague RT)] Vague instructions resulted in faster responses than crisp instructions on average.
+However this difference was not significant in the full model ($\beta=-0.015$, $se=0.0097$, $t=-1.58$, $p=0.119$).
+Using Levy's method \citep{Levy:MainEffectsInteractions} to test for main effects in the presence of higher-order interactions, by doing model comparison between a null model that included all interaction terms involving Vagueness but leaving out a term for the main effect of vagueness, against a full model that differed only by including vagueness as a main effect, showed that the full model was no better than the reduced model ($df=1$, $p=0.118$), consituting more evidence that vagueness did not exert a significant main effect on response times.
+	\item [Test of Hypothesis 2 (Comparison/Matching RT)] Comparison instructions resulted in significantly faster responses than matching instructions ($\beta=0.176$, $se=0.0168$, $t=10.51$, $p<0.001$).
+	\item [Test of Hypothesis 3 (Interaction)] The interaction between vagueness and selection task was significant ($\beta=0.123$, $se=0.0166$, $t=7.42$, $p<0.001$), showing that vagueness had a different influence on response times when the instructions mandated comparison versus when the instructions mandated matching.
+When separate analyses were carried out testing for the effect of vagueness in comparison-only and in matching-only conditions, vague instructions led to significantly \emph{faster} responses than crisp instructions in the comparison-only conditions ($\beta=-0.077$, $se=0.018$, $t=-4.31$, $p<0.001$); and in the matching-only conditions vague instructions led to significantly \emph{slower} responses than crisp instructions ($\beta=0.047$, $se=0.013$, $t=3.71$, $p<0.001$).
+\end{description}
+
+\subsection{Discussion (Experiment 3)}
+
+The cost reduction account predicted that there should be a significant main effect of vagueness such that responses would be faster for vague intructions than for crisp instructions.
+We found that although there was a very small effect in that direction, the effect was not statistically significant.
+Models that differed only in the presence of vagueness as a main effect were shown not to differ significantly in their explanatory value.
+
+However, the results also showed that vagueness did exert an influence on reaction times according to whether the task was comparison or matching: vagueness was beneficial for comparison and detrimental for matching (the same as Experiment 2) even when no numbers were allowed in the instructions.
+
+\section{Discussion of Experiments 2 and 3}
+The main aim of these two experiments was to test whether vagueness confers any cognitive benefits over and above those that are due to differences in the selection task according to whether the instruction mandates a \emph{comparison} selection task or a \emph{matching} selection task, when number-use is held constant.
+The main effect of selection task showed that the assumption that the \emph{comparison} task is easier than the \emph{matching} task is well-founded.
+In both experiments people were reliably quicker to respond in the \emph{comparison} task.
+
+Vagueness, which was the phenomenon on which our investigation focussed, did not exert a significant main effect in response time.
+However when the comparison and selection tasks were analysed separately, there was small significant advantage for vagueness in the \emph{comparison} tasks, but a small significant disadvantage for vagueness in the \emph{matching} tasks.
+
+\section{General Discussion}
+To summarise our findings, we asked why strategic vagueness is as frequent as it is and we decided to focus on what we see as the most promising explanation, namely that vague expressions are easy to process for speakers and hearers: the cost hypothesis, as we have called it.
+We decided to test this hypothesis by experimentally investigating whether vague descriptions are resolved by hearers more quickly than crisp ones.
+Although we were able to find some interesting (and statistically highly significant) effects, it appears to us that if our sequence of experiments is assessed as a whole, it cannot be seen as confirmation of the cost hypothesis.
+
+To explain why this is, let us summarise our findings so far: Experiment 1 showed that number avoidance in the verbal format instructions is an important factor driving the faster response times in the task, and that vagueness does not have any additional benefit in either the verbal format instructions or the numerical format instructions.
+However, Experiment 1 did not distinguish the benefits of number avoidance from the benefits of the comparison selection task.
+In Experiments 2 and 3 we manipulated vagueness and the selection task, separately at each level of numerical format.
+Across the two experiments, we found that the comparison-task instructions attracted faster response times than the matching-task instructions.
+Within the two experiments we found that vagueness exerts benefits when the selection task is \emph{comparison}, but not when the task is \emph{matching}.
+
+What is one entitled to conclude?
+Given that we were able to identify a class of situations in which vague expressions led to faster response times than crisp ones, would it be valid to conclude that we have finally discovered an advantage for vagueness that cannot be ascribed to some other factor?
+We believe the answer to this question is negative.
+
+\begin{table}[htbp]
+\caption{Vagueness as range reduction: A summary of Experiments 2 and 3}
+\label{Vagueness as range reduction}
+\centering
+\begin{tabular}{cccc}
+\hline\noalign{\smallskip}
+selection task 					& vagueness		& candidates	& effect of vagueness						\\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+\multirow{ 2}{*}{comparison} 	& crisp 		& 2				& \multirow{ 2}{*}{vagueness advantage}  	\\
+\noalign{\smallskip}\cline{2-3}\noalign{\smallskip}
+								& vague			& 1				&                                           \\
+\noalign{\smallskip}\hline\noalign{\smallskip}
+\multirow{ 2}{*}{matching} 		& crisp 		& 1 			& \multirow{ 2}{*}{vagueness disadvantage}	\\
+\noalign{\smallskip}\cline{2-3}\noalign{\smallskip}
+								& vague			& 2				&								 			\\
+\noalign{\smallskip}\hline
+\end{tabular}
+\end{table}
+
+To see why, consider Figures~\ref{resultsD-exp-2} and~\ref{resultsE-exp-3}.
+Both figures depict four conditions, depending on whether the expression was crisp or vague, and depending on whether the referent could be identified using a comparison strategy or not.
+Two of the resulting four conditions result in an expression that can denote either of two referents; the other two conditions result in an expression that can only denote one referent, with the other possible referent being a marginal candidate at best.
+
+To see why vagueness has opposite effects depending on whether it is used in matching or comparison situations, consider the stimulus with (6,15,24) dots.
+Now compare `Choose a square with 6 dots' with its vague counterpart `Choose a square with about 10 dots': by adding the word `about,' we broaden the range of squares to which the expression might be referring.
+On the other hand, compare `Choose a square with fewer than 20 dots' with `Choose a square with far fewer than 20 dots': by adding the word `far,' we did not broaden the range of squares denotable by the expression: we narrow it down, because only some of the squares that have fewer dots may have {\em far} fewer dots.
+
+The benefits of vagueness in the \emph{comparison} task in Experiments 3 and 4 could thus be explained as differences in the number of valid targets for the expression.
+This leads us to speculate that the benefit for vagueness here could be due to the vague expression foregrounding a particular valid target, while the crisp expression carries with it the additional task of distinguishing between two alternative valid targets, something we propose to call a ``range-reduction'' benefit.
+
+The observation that conditions with 1 candidate lead to shorter response times than conditions with 2 candidates is consistent with the range reduction hypothesis, but not with the idea that vagueness is beneficial.
+It appears, in other words, that shorter response times will only result from a vague expression if this expression leads to range reduction.
+Once again, it is not vagueness itself that has advantages but a phenomenon (namely range reduction) that is an automatic concomitant of vagueness in some types of situations.\\[2ex]
+%
+Looking at the entire series of experiments, our findings suggest that the observed benefits of vague expressions may be due to factors other than vagueness: factors like avoiding numbers, permitting comparison tasks, and range reduction.
+The picture that is starting to emerge is subtle: on the one hand, in the situations that we have been studying, vagueness is not intrinsically beneficial.
+On the other hand, vague expressions frequently possess other features that \emph{are} beneficial, and these are what give us the incorrect impression that vagueness itself is beneficial.
+Vagueness may thus have acquired a reputation that it does not deserve.
+The answer to Lipman's question, of why vagueness permeates human language (see our Introduction), may lie in a different direction after all, possibly relating to benefits for the speaker rather than the hearer.
+
+A comparison may clarify the logic of the situation.
+In recent years a number of studies, focussing on red wine, have suggested that alcohol, consumed in low doses, may have health benefits.
+An alternative explanation, however, asserts that it is not the alcohol in the wine that is beneficial, but antioxidants from grapes.
+If this alternative explanation is correct, then alcohol may not be healthy after all.\\[2ex]
+%
+Our findings suggest a re-think of the questions on which much research on the utility of vagueness rests.
+Years of research on the logic of vagueness --- giving rise to such techniques as Partial Logic \citep[e.g.,][]{Fine}, Probabilistic Logic \citep{Edgington}, and Fuzzy Logic \citep{Zadeh} --- have primed the research community to expect that some special utility of vagueness is an important part of the answer; but our findings call this expectation into question.
+
+Although our own studies in this article have focussed on vagueness in descriptive noun phrases only, it seems plausible that vagueness plays a similar role in other linguistic constructs.
+For example, consider reports on air temperature. Given a numerical temperature measurement or prediction, we might word it as
+%
+\begin{quote}
+(a) {\em 27.2 degrees Celsius}, or\\
+(b) {\em approximately 27 degrees}, or\\
+(c) {\em above 25 degrees}, or\\
+(d) {\em warm},
+\end{quote}
+%
+among other candidate expressions.
+Which of these descriptions is most effective, for example as part of a weather report?
+If the linguistic literature is to be believed, then options (a) and (c) convey crisp information, whereas (b) and (d) are vague (i.e., they permit borderline cases).
+In the situations studied in our own experiments and the ones discussed in section 1, we found no evidence that vagueness is beneficial for hearers.
+Rather than asking whether a candidate expression is vague, other questions might shed more light on the choice, similar to the ones identified in our studies.
+These questions might focus on the amount of information that a given expression conveys (i.e., on granularity), on the avoidance of numbers, and on the use of evaluative terms.
+Let's see how this might pan out for the above examples from the weather domain.
+
+First, the experiments by \citeauthor{Mishra01042011} suggest that it is important how much information is conveyed by an expression, and their findings are echoed by our own thoughts about range reduction (following Experiments 2 and 3).
+In the case of (a)--(d) above, it appears that (a) conveys the most detailed information (designating the smallest segment of the temperature scale), followed by (b), then (d), then (c) (e.g., 40 degrees is above 25, but at 40 Celsius the word ``warm'' is likely to give way to ``hot'' or ``scorching''):
+%
+\begin{quote}
+$a < b < d < c$
+\end{quote}
+%
+If these hunches are correct, then it seems to us that it is relatively unimportant whether a given expression is vague or crisp.
+Other factors seem more important; moreover, it may depend on the task and the audience which of a--d is preferred.
+For example, an \emph{expert} may prefer to read expression (a), because it gives her the most detailed information on which to base her decisions.
+On the other hand, expression (d) (``warm'') is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make (d) more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm'' (recall section~\ref{related-work}), which is especially important if the hearer is unfamiliar with the metric used.
+These considerations suggest that {\em non-experts} might prefer expression (d).
+
+One way to see why vagueness (as defined in our Introduction) may not benefit human communication is the following thought experiment.
+Suppose a group of speakers understand the word ``warm'' as vague, agreeing that temperatures above 26 count as warm, and temperatures below 24 do not count as warm, but considering temperatures between 24 and 26 as borderline cases.
+Now one day these speakers agree to sharpen up their definition, deciding that, henceforth, ``warm'' means ``$>25$ degrees'' (as in (c) above): this decision resolves the borderline cases, while everything else remains the same.
+It seems unlikely that this change in language use, from a vague meaning to a crisp one (i.e., one that has no borderline cases anymore), would lower the utility of the word.
+Our experimental findings, and the conclusions that we draw from them, are consistent with this idea.
+
 \backmatter
@@
 \backmatter
-%\printindex
 \end{document}

### 2018-10-27 manuscript/v-book-2018.tex
replaced double quotes with single quotes throughout

@@
 \chapter{The Elusive Benefits of Vagueness: Evidence from Experiments}
-Much of everyday language is vague, even in situations where vagueness could have been avoided (i.e., where vagueness is used ``strategically'').
+Much of everyday language is vague, even in situations where vagueness could have been avoided (i.e., where vagueness is used `strategically').
 Yet the benefits of vagueness for hearers and readers are proving to be elusive.
@@
 In academic use, the word `vagueness' has a specific meaning.
-Keefe and Smith, for example, state that ``vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes'' \citep[p.~4]{keefe1997vagueness}, as do \citet{EgreKlinedinst}.
-The crucial criterion is the existence of borderline cases: ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise'' \citet[p.~1]{lipmanvague}.
-A typical example is the word ``tall'', as applied to people for example, because there is no precise, known height which separates those who are tall from those who are not.
-The crucial point is that ``tall'' admits borderline cases (i.e., people who may or may not count as tall), which are the hallmark of vagueness as we use the term.
+Keefe and Smith, for example, state that `vague predicates have borderline cases, have fuzzy boundaries, and are susceptible to sorites paradoxes' \citep[p.~4]{keefe1997vagueness}, as do \citet{EgreKlinedinst}.
+The crucial criterion is the existence of borderline cases: `a word is precise if it describes a well-defined set of objects.
+By contrast, a word is vague if it is not precise' \citet[p.~1]{lipmanvague}.
+A typical example is the word `tall', as applied to people for example, because there is no precise, known height which separates those who are tall from those who are not.
+The crucial point is that `tall' admits borderline cases (i.e., people who may or may not count as tall), which are the hallmark of vagueness as we use the term.
 
@@
 often used even in situations where the speaker could have used an expression that is not vague (i.e., crisp); in these situations we say that vagueness is used \emph{strategically}.
-By introducing borderline cases, these expressions create potential misunderstandings, thereby creating ``a worldwide several-thousand year efficiency loss'' \citet[][p.~1]{lipmanvague}.
+By introducing borderline cases, these expressions create potential misunderstandings, thereby creating `a worldwide several-thousand year efficiency loss' \citet[][p.~1]{lipmanvague}.
 Lipman explains the point by means of a scenario in which a speaker describes a person to a hearer, who needs to identify that person in the arrivals hall of of an airport.
-In such a scenario, a precise description of the person's height (e.g., ``The person's height is 187.96 cm'') would be more useful than a vague one (``The person is tall'').
+In such a scenario, a precise description of the person's height (e.g., `The person's height is 187.96 cm') would be more useful than a vague one (`The person is tall').
 Lipman uses this scenario to explain why standard game theory models of communication \citep[e.g.,][]{Crawford:1982lr} predict that, under certain conditions, a crisp act of communication will always have more utility than a vague act that communicates the same state of affairs.
@@
 One of the most promising answers appears to be the idea that vague expressions are easier to process, by a speaker and/or a hearer, than expressions that are not vague (i.e., crisp) \citep[e.g.,][]{lipmanvague,De-Jaegher:2003lr,vanrooij2003lr}.
-For example, \citeauthor{lipmanvague} writes: ``For the listener, information which is too specific may require more effort to analyze'' (\citeyear[][p.~11]{lipmanvague}).
+For example, \citeauthor{lipmanvague} writes: `For the listener, information which is too specific may require more effort to analyze' (\citeyear[][p.~11]{lipmanvague}).
 We shall refer to this as the \emph{cost reduction} hypothesis.
 
-This article brings an experimental approach to these issues, focussing on vagueness in descriptions (e.g., ``the square with few dots'') and its effect on the hearer's ability to act on a given description, as measured by the time that it takes hearers to click on the referent of a description.\footnote{Other metrics could have been chosen, such as hearers' ability to remember information, for example, or error rates.
+This article brings an experimental approach to these issues, focussing on vagueness in descriptions (e.g., `the square with few dots') and its effect on the hearer's ability to act on a given description, as measured by the time that it takes hearers to click on the referent of a description.\footnote{Other metrics could have been chosen, such as hearers' ability to remember information, for example, or error rates.
 Although error rates play a minor role in the present paper, for reasons that will become clear, we focus on response times in particular.}
@@
 Instead, it is other factors, such as the presence of an overt numerical expression in the description, that proved to be decisive.
-We believe that, despite the fact that our experiments are unavoidably focussed on a specific class of vague expressions (since any experiment can only deal with a limited number of different stimuli), these findings are potentially important, because they call into question whether ``strategic'' vagueness (i.e., vagueness where the speaker had a choice, because she could have produced a crisp expression instead) has any advantages at all.
+We believe that, despite the fact that our experiments are unavoidably focussed on a specific class of vague expressions (since any experiment can only deal with a limited number of different stimuli), these findings are potentially important, because they call into question whether `strategic' vagueness (i.e., vagueness where the speaker had a choice, because she could have produced a crisp expression instead) has any advantages at all.
 In other words, returning to Lipman's question, it is possible that vagueness has evolved partly as a necessary evil (e.g., because of the limits of observation and prediction) and partly as a side effect of other factors.
@@
 \item It has been proposed that the existence of vague words makes a language more efficient \citep{barwiseperry}.
-The idea is that vagueness allows the use of one and the same word (e.g., ``big'') in different situations, which require different standards.
-For instance, the vague word ``big'' can denote a mouse as well as an elephant.
-The idea is plausible, but the problem is that ``big'' is efficient not because it is vague (i.e., not because it allows borderline cases) but because it is context dependent.
-To see this, note for example that superlatives are not vague (i.e., they do not permit borderline cases), but they {\em are} context dependent: ``the biggest $x$'' ascribes different sizes to the referent depending on $x$.
-It is for this reason that we can use the word ``biggest''  to talk about the biggest mouse, the biggest elephant, and so on.
+The idea is that vagueness allows the use of one and the same word (e.g., `big') in different situations, which require different standards.
+For instance, the vague word `big' can denote a mouse as well as an elephant.
+The idea is plausible, but the problem is that `big' is efficient not because it is vague (i.e., not because it allows borderline cases) but because it is context dependent.
+To see this, note for example that superlatives are not vague (i.e., they do not permit borderline cases), but they {\em are} context dependent: `the biggest $x$' ascribes different sizes to the referent depending on $x$.
+It is for this reason that we can use the word `biggest'  to talk about the biggest mouse, the biggest elephant, and so on.
 Context-dependence, however, does not imply vagueness.
 \item Vague words are capable of combining a statement of quantity with an evaluative statement \citep{veltman}.
-The idea is that when we say ``The patient has a fever,'' we do not merely assert that her temperature lies above a certain value, we also imply that the deviation is clinically significant (i.e., something is not right with the patient).
+The idea is that when we say `The patient has a fever,' we do not merely assert that her temperature lies above a certain value, we also imply that the deviation is clinically significant (i.e., something is not right with the patient).
 The problem with this plausible idea is that evaluation does not imply vagueness.
-For example, the medical term ``obese'' is evaluative (i.e., it is not healthy to be obese), yet its standard medical definition in terms of Body Mass Index is perfectly crisp (i.e., without borderline cases), with all and only BMI values above 30 counting as obese.
+For example, the medical term `obese' is evaluative (i.e., it is not healthy to be obese), yet its standard medical definition in terms of Body Mass Index is perfectly crisp (i.e., without borderline cases), with all and only BMI values above 30 counting as obese.
 \end{itemize}
@@
 The authors explain the improvement in the vague condition for this group as resulting from the participants' freedom to think of themselves a positioned on one end of the range --- the end closest to the ideal HHI scores.
-This ``illusion of proximity'' \citep[][p.~4]{Mishra01042011} to the goal is argued to have allowed participants to generate positive expectancies that lead to behaviours that improved performance.
+This `illusion of proximity' \citep[][p.~4]{Mishra01042011} to the goal is argued to have allowed participants to generate positive expectancies that lead to behaviours that improved performance.
 In contrast, in the precise conditions, participants did not have this freedom of interpretation, and could not distort the information to bring about the beneficial \emph{illusion of proximity}.
@@
 
-It is, however, questionable whether the ``evaluative categories'' manipulation in this study can be considered a manipulation of vagueness.
+It is, however, questionable whether the `evaluative categories' manipulation in this study can be considered a manipulation of vagueness.
 Certainly, terms like \emph{Fair} admit the possibility of borderline cases.
@@
 In Experiment 1, reported below, we also addressed another issue with the second pilot experiment: participants chose one of two dot arrays --- therefore the `vague' quantifiers (few and many) uniquely identified one square.
-Recall our definition of vague --- ``a word is precise if it describes a well-defined set of objects.
-By contrast, a word is vague if it is not precise''.
-The quantifiers `few' and `many' might not have realised their potential for vagueness in the context of a choice between (only) two alternatives --- where there is no borderline alternative that could result in few and many being ``not well-defined''.
+Recall our definition of vague --- `a word is precise if it describes a well-defined set of objects.
+By contrast, a word is vague if it is not precise'.
+The quantifiers `few' and `many' might not have realised their potential for vagueness in the context of a choice between (only) two alternatives --- where there is no borderline alternative that could result in few and many being `not well-defined'.
 Furthermore, using definite articles in the instructions may have reinforced the participant's impression that only one choice counted as correct, and this impression could have been reinforced by the use of error feedback.
@@
 \begin{itemize}
-	\item In the vague numerical condition the instruction was ``Choose a square with about 10 dots'' --- none of the arrays contained exactly 10 dots, but 10 is closer to 6 than it is to 15, making 6 a better response to that instruction, 15 a borderline response, and 24 an extreme response.
-	\item In the vague verbal condition we used ``Choose a square with few dot''. We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
-	\item In the crisp numerical condition we used ``Choose the square with 6 dots''.
+	\item In the vague numerical condition the instruction was `Choose a square with about 10 dots' --- none of the arrays contained exactly 10 dots, but 10 is closer to 6 than it is to 15, making 6 a better response to that instruction, 15 a borderline response, and 24 an extreme response.
+	\item In the vague verbal condition we used `Choose a square with few dot'. We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
+	\item In the crisp numerical condition we used `Choose the square with 6 dots'.
 The smaller flanking array always contained exactly the specified number of dots.
 We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
-	\item For crisp verbal, we used ``Choose the square with the fewest dots''. We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
+	\item For crisp verbal, we used `Choose the square with the fewest dots'. We considered this to be equivalent in terms of which responses were expected (6), borderline (15) and extreme (24).
 \end{itemize}
 
-On each trial, first the referring expression that constituted the instruction for that trial was displayed (e.g., ``Choose a square with about 10 dots'').
+On each trial, first the referring expression that constituted the instruction for that trial was displayed (e.g., `Choose a square with about 10 dots').
 Participants then pressed a key to indicate that they had read the instruction.
@@
 In this experiment, no feedback was given.
-This was because, in the vague conditions, we did not regard any response as ``correct'' or ``incorrect'', but instead as ``expected response''; ``borderline response''; and ``extreme response'', and we did not want to draw participants' attention to this distinction explicitly.
+This was because, in the vague conditions, we did not regard any response as `correct' or `incorrect', but instead as `expected response'; `borderline response'; and `extreme response', and we did not want to draw participants' attention to this distinction explicitly.
 The participant's choice was recorded for analysis.
@@
 \centering
-\caption{Experiment 1 instructions arranged by condition. The instructions given in the table started with ``Choose \ldots''}
+\caption{Experiment 1 instructions arranged by condition. The instructions given in the table started with `Choose \ldots'}
 \label{instructionsC-exp-1}
@@
 Table~\ref{Instructions for e2} shows the instructions for each condition.
-Note the difference between ``fewer than 20'' and ``far fewer than 20'': whereas the former cannot have borderline cases (i.e., for each number it is clear whether the number is smaller than 20 or not), the latter can.
+Note the difference between `fewer than 20' and `far fewer than 20': whereas the former cannot have borderline cases (i.e., for each number it is clear whether the number is smaller than 20 or not), the latter can.
 
@@
 \centering
-\caption{Experiment 2: Instructions arranged by condition. The instructions given in the table started with ``Choose a square with \ldots''}
+\caption{Experiment 2: Instructions arranged by condition. The instructions given in the table started with `Choose a square with \ldots'}
 \label{Instructions for e2}
@@
 \includegraphics[width=.5\textwidth]{figures/Ee3-flow.jpg}
-\caption{Experiment 3 example stimulus, showing the sequence of presentation, with the instruction screen at the top, followed by the presentation of the ``target'' array, followed by the stimulus to which participants responded.}
+\caption{Experiment 3 example stimulus, showing the sequence of presentation, with the instruction screen at the top, followed by the presentation of the `target' array, followed by the stimulus to which participants responded.}
 \label{Experiment3examplestimulus}
@@
 \centering
-\caption{Experiment 3 instructions for the smallest array, arranged by condition. The instructions given in the table started with ``Choose a square with \ldots''}
+\caption{Experiment 3 instructions for the smallest array, arranged by condition. The instructions given in the table started with `Choose a square with \ldots'}
 \label{instructionsE-exp-3}
@@
 The benefits of vagueness in the \emph{comparison} task in Experiments 3 and 4 could thus be explained as differences in the number of valid targets for the expression.
-This leads us to speculate that the benefit for vagueness here could be due to the vague expression foregrounding a particular valid target, while the crisp expression carries with it the additional task of distinguishing between two alternative valid targets, something we propose to call a ``range-reduction'' benefit.
+This leads us to speculate that the benefit for vagueness here could be due to the vague expression foregrounding a particular valid target, while the crisp expression carries with it the additional task of distinguishing between two alternative valid targets, something we propose to call a `range-reduction' benefit.
 
@@
 First, the experiments by \citeauthor{Mishra01042011} suggest that it is important how much information is conveyed by an expression, and their findings are echoed by our own thoughts about range reduction (following Experiments 2 and 3).
-In the case of (a)--(d) above, it appears that (a) conveys the most detailed information (designating the smallest segment of the temperature scale), followed by (b), then (d), then (c) (e.g., 40 degrees is above 25, but at 40 Celsius the word ``warm'' is likely to give way to ``hot'' or ``scorching''):
+In the case of (a)--(d) above, it appears that (a) conveys the most detailed information (designating the smallest segment of the temperature scale), followed by (b), then (d), then (c) (e.g., 40 degrees is above 25, but at 40 Celsius the word `warm' is likely to give way to `hot' or `scorching'):
 %
@@
 For example, an \emph{expert} may prefer to read expression (a), because it gives her the most detailed information on which to base her decisions.
-On the other hand, expression (d) (``warm'') is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make (d) more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of ``warm'' (recall section~\ref{related-work}), which is especially important if the hearer is unfamiliar with the metric used.
+On the other hand, expression (d) (`warm') is shorter than the other three and avoids the use of numbers; our experiments suggest that this may make (d) more rapidly understood than its competitors; earlier experiments point in the same direction, given the evaluative nature of `warm' (recall section~\ref{related-work}), which is especially important if the hearer is unfamiliar with the metric used.
 These considerations suggest that {\em non-experts} might prefer expression (d).
@@
 One way to see why vagueness (as defined in our Introduction) may not benefit human communication is the following thought experiment.
-Suppose a group of speakers understand the word ``warm'' as vague, agreeing that temperatures above 26 count as warm, and temperatures below 24 do not count as warm, but considering temperatures between 24 and 26 as borderline cases.
-Now one day these speakers agree to sharpen up their definition, deciding that, henceforth, ``warm'' means ``$>25$ degrees'' (as in (c) above): this decision resolves the borderline cases, while everything else remains the same.
+Suppose a group of speakers understand the word `warm' as vague, agreeing that temperatures above 26 count as warm, and temperatures below 24 do not count as warm, but considering temperatures between 24 and 26 as borderline cases.
+Now one day these speakers agree to sharpen up their definition, deciding that, henceforth, `warm' means `$>25$ degrees' (as in (c) above): this decision resolves the borderline cases, while everything else remains the same.
 It seems unlikely that this change in language use, from a vague meaning to a crisp one (i.e., one that has no borderline cases anymore), would lower the utility of the word.

### 2019-04-15 experiment_data_and_analyses/A_pilot_1.Rmd
pilot 1 formatting and plotting and tables

--- 2018-09-12 docs/A_pilot_1.Rmd
@@
     theme: default
-    css: styles.css
+    code_folding: show
 ---
+
+<style>
+.main-container{width: 1500px;max-width:2800px;}
+body{font-size: 13pt;}
+.r{font-size: 12px;}
+pre{font-size: 12px}
+</style>
 
@@
 ```{r, "lmer 1 compare different rt metrics using the same model structure", warning=FALSE}
+# no interaction
 lmer1.ms  <- lmer(data=dat, rt_ms ~         vag + subit + quan + gap + side + (1|subject) + (1|item))
 lmer1.log <- lmer(data=dat, rt_log ~        vag + subit + quan + gap + side + (1|subject) + (1|item))
-lmer1.rcp <- lmer(data=dat, rt_reciprocal ~ vag + subit + quan + gap + side + (1|subject) + (1|item))
+# with interaction
+lmer2.ms <- lmer(data=dat, rt_ms ~          vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item))
+lmer2.log <- lmer(data=dat, rt_log ~        vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item))
 ```
@@
   rbind(
-  cbind(tidy(lmer1.ms),model="ms"), 
-  cbind(tidy(lmer1.log),model="log"), 
-  cbind(tidy(lmer1.rcp),model="rcp")
-)
-
-dwplot(models, vline = geom_vline(xintercept = 0)) + facet_wrap(~model, scales="free", ncol=1)
+    cbind(tidy(lmer1.ms), model="ms no interaction"), 
+    cbind(tidy(lmer1.log), model="log no interaction"),
+    cbind(tidy(lmer2.ms), model="ms with interaction"), 
+    cbind(tidy(lmer2.log), model="log with interaction")
+  ) %>%
+  filter(effect == 'fixed', term != "(Intercept)") %>%
+  select(-group)
 ```
 
-```{r lmer on logrt, warning=FALSE}
-lmer2.log <- lmer(data=dat, rt_log ~ vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item))
-```
-
-```{r, fig.width=6, fig.height=2}
-dwplot(lmer2.log %>% tidy() %>% filter(effect=="fixed"), vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 1)) +
-  xlab("Coefficient") +
-  ggtitle("rt_log ~ vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item)") + 
-  theme(plot.title = element_text(size = 8), axis.title.x = element_text(size=8), axis.text.x = element_text(size=8))
+```{r, fig.height=12}
+dwplot(models, vline = geom_vline(xintercept = 0)) + facet_wrap(vars(model), scales="free", ncol=1, drop=TRUE)
 ```
@@
 ```{r}
-lmer2.rcp <- lmer(data=dat, rt_reciprocal~subit*vag + (vag|subject) + (vag|item))
+summary(lmer2.log)$coefficients %>% kable() %>% kable_styling(full_width = FALSE, position='left')
 ```
 
-```{r}
-# lmer3 is closer to the original anova than lmer2
-lmer3 <- lmer(data=dat, rt_reciprocal~vag*subit+quan+gap + (vag+subit|subject) + (vag|item))
-```
+<!-- Lmer2 includes the interaction of vagueness and subitizability. -->
 
-```{r, fig.width=6, fig.height=2}
-dwplot(lmer3 %>% tidy() %>% filter(effect=="fixed"), vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 1)) +
-  xlab("Coefficient") + 
-  ggtitle("rt_reciprocal ~ vag * subit + quan + gap + (vag+subit|subject) + (vag|item)") + 
-  theme(plot.title = element_text(size = 8), axis.title.x = element_text(size=8), axis.text.x = element_text(size=8))
-```
+<!-- ```{r lmer on logrt, warning=FALSE} -->
+<!-- lmer2.log <- lmer(data=dat, rt_log ~ vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item)) -->
+<!-- ``` -->
+
+<!-- ```{r, fig.width=6, fig.height=2} -->
+<!-- dwplot(lmer2.log %>% tidy() %>% filter(effect=="fixed"), vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 1)) + -->
+<!--   xlab("Coefficient") + -->
+<!--   ggtitle("rt_log ~ vag * subit + quan + gap + side + (1+vag|subject) + (1+vag|item)") +  -->
+<!--   theme(plot.title = element_text(size = 8), axis.title.x = element_text(size=8), axis.text.x = element_text(size=8)) -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- lmer2.rcp <- lmer(data=dat, rt_reciprocal~subit*vag + (vag|subject) + (vag|item)) -->
+<!-- ``` -->
+
+<!-- ```{r} -->
+<!-- # lmer3 is closer to the original anova than lmer2 -->
+<!-- lmer3 <- lmer(data=dat, rt_reciprocal~vag*subit+quan+gap + (vag+subit|subject) + (vag|item)) -->
+<!-- ``` -->
+
+<!-- ```{r, fig.width=6, fig.height=2} -->
+<!-- dwplot(lmer3 %>% tidy() %>% filter(effect=="fixed"), vline = geom_vline(xintercept = 0, colour = "grey50", linetype = 1)) + -->
+<!--   xlab("Coefficient") +  -->
+<!--   ggtitle("rt_reciprocal ~ vag * subit + quan + gap + (vag+subit|subject) + (vag|item)") +  -->
+<!--   theme(plot.title = element_text(size = 8), axis.title.x = element_text(size=8), axis.text.x = element_text(size=8)) -->
+<!-- ``` -->
 
